### 2009-12-11 trunk/deliverables/D2_2_Ontology/classes.tex
D2_2 and D2_4 added
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@385 d8b0c59b-6544-0410-94d9-83a5658a058e

\subsubsection{Class: \texttt{fast:Resource}}
\label{subs:Resource}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & Resource \\ \hline
\textcolor{white}{\textbf{description}} & Anything that is part of a gadget (or the gadget itself). Tentatively 
    anything that can be `touched' and moved around in the FAST IDE. \\ \hline
\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fast:Screen\_Flow}}{subs:ScreenFlow}, \htmlref{\texttt{fast:Screen}}{subs:Screen}, \htmlref{\texttt{fast:Backend\_Service}}{subs:BackendService}, \htmlref{\texttt{fast:Flow\_Control\_Element}}{subs:FlowControlElement}, \htmlref{\texttt{fast:Operator}}{subs:Operator}, \htmlref{\texttt{fast:Form\_Element}}{subs:FormElement}, \htmlref{\texttt{fast:Condition}}{subs:Condition}, \htmlref{\texttt{fast:WithConditions}}{subs:WithConditions} \\ \hline
\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fast:contains}}{subs:contains}, \htmlref{\texttt{fast:hasIcon}}{subs:hasIcon}, \htmlref{\texttt{fast:hasScreenshot}}{subs:hasScreenshot} \\ \hline
\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fast:contains}}{subs:contains} \\ \hline
\end{tabular}
\subsubsection{Class: \texttt{fast:Screen\_Flow}}
\label{subs:ScreenFlow}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & Screen Flow \\ \hline
\textcolor{white}{\textbf{description}} & The complete gadget, a set of screens. \\ \hline
\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
\end{tabular}
\subsubsection{Class: \texttt{fast:Screen}}
\label{subs:Screen}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & Screen \\ \hline
\textcolor{white}{\textbf{description}} & An individual screen. \\ \hline
\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
\end{tabular}
\subsubsection{Class: \texttt{fast:Backend\_Service}}
\label{subs:BackendService}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & Backend Service \\ \hline
\textcolor{white}{\textbf{description}} & A Web service which provides data and/or functionality to a screen. 
    A backend service will often be external to FAST, and will probably have to be 
    wrapped by the screen. \\ \hline
\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
\end{tabular}
\subsubsection{Class: \texttt{fast:Flow\_Control\_Element}}
\label{subs:FlowControlElement}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & Flow Control Element \\ \hline
\textcolor{white}{\textbf{description}} & Any kind of component which can restrict the default flow of screens 
    in a gadget. \\ \hline
\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fast:Screen\_Flow\_Start}}{subs:ScreenFlowStart}, \htmlref{\texttt{fast:Screen\_Flow\_End}}{subs:ScreenFlowEnd}, \htmlref{\texttt{fast:Connector}}{subs:Connector} \\ \hline
\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
\end{tabular}
\subsubsection{Class: \texttt{fast:Screen\_Flow\_Start}}
\label{subs:ScreenFlowStart}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & Screen Flow Start \\ \hline
\textcolor{white}{\textbf{description}} & The entry point to a wigdet; the first screen. \\ \hline
\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Flow\_Control\_Element}}{subs:FlowControlElement} \\ \hline
\end{tabular}
\subsubsection{Class: \texttt{fast:Screen\_Flow\_End}}
\label{subs:ScreenFlowEnd}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & Screen Flow End \\ \hline
\textcolor{white}{\textbf{description}} & A screen that ends the workflow of the gadget. \\ \hline
\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Flow\_Control\_Element}}{subs:FlowControlElement} \\ \hline
\end{tabular}
\subsubsection{Class: \texttt{fast:Connector}}
\label{subs:Connector}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & Connector \\ \hline
\textcolor{white}{\textbf{description}} & An explicit connection between two screens. \\ \hline
\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Flow\_Control\_Element}}{subs:FlowControlElement} \\ \hline
\end{tabular}
\subsubsection{Class: \texttt{fast:Operator}}
\label{subs:Operator}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & Operator \\ \hline
\textcolor{white}{\textbf{description}} & Any kind of component that is used to connect backend services to 
    form elements. Examples are simple pipes, aggregators or various kinds of 
    filters. \\ \hline
\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
\end{tabular}
\subsubsection{Class: \texttt{fast:Form\_Element}}
\label{subs:FormElement}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & Form Element \\ \hline
\textcolor{white}{\textbf{description}} & Form elements are UI elements in a particular screen. \\ \hline
\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
\end{tabular}
\subsubsection{Class: \texttt{fast:Condition}}
\label{subs:Condition}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & Condition \\ \hline
\textcolor{white}{\textbf{description}} & The pre- or post-condition of either a screen or a screenflow. In 
    the latter case, each target platform will use these conditions in its own way, 
    or may also ignore them. E.g., in EzWeb pre- and post-conditions correspond to 
    the concepts of slot and event. \\ \hline
\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fast:hasPattern}}{subs:hasPattern}, \htmlref{\texttt{fast:isPositive}}{subs:isPositive} \\ \hline
\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fast:hasPreCondition}}{subs:hasPreCondition}, \htmlref{\texttt{fast:hasPostCondition}}{subs:hasPostCondition} \\ \hline
\end{tabular}
\subsubsection{Class: \texttt{fast:WithConditions}}
\label{subs:WithConditions}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & With-condition \\ \hline
\textcolor{white}{\textbf{description}} & Those kinds of resource which can have pre- or post-conditions 
    (i.e., screens and screen flows). \\ \hline
\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fast:hasPreCondition}}{subs:hasPreCondition}, \htmlref{\texttt{fast:hasPostCondition}}{subs:hasPostCondition} \\ \hline
\textcolor{white}{\textbf{unionOf}} & \htmlref{\texttt{fast:Screen}}{subs:Screen}, \htmlref{\texttt{fast:Screen\_Flow}}{subs:ScreenFlow} \\ \hline
\end{tabular}

### 2009-12-11 trunk/deliverables/D2_2_Ontology/fast_d2_2_ontology.tex
D2_2 and D2_4 added
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@385 d8b0c59b-6544-0410-94d9-83a5658a058e

\documentclass{fast_latex}
% yay! We can use any kind of funky diacritic:
\usepackage[utf8]{inputenc}

\usepackage{lastpage}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage[pdfborder={0 0 0}]{hyperref}		% turn on when latex is used (not miktec)
\usepackage{url} % LEO: urls \url{}
\usepackage{verbatim} % code and comment
\usepackage{longtable}
\usepackage{xspace}   % whitespace after a macro if no punctuation after the macro
\usepackage{multirow}
\usepackage{colortbl}
\usepackage{longtable}
\usepackage{array}
\usepackage{amssymb}
\parindent0pt

\newcommand\deliverableNumber{D2.2.1}
\newcommand\deliverableTitle{Ontology and Conceptual Model for the Semantic Characterisation of Complex Gadgets}
\newcommand\deliverableTitleShort{Ontology and Conceptual Model}
\newcommand\workpackageNumber{2}
\newcommand\workpackageTitle{Definition of Conceptual Model}
\newcommand\authorOne{Knud M\"oller, NUIG}
\newcommand\authorTwo{Ismael Rivera, NUIG}
\newcommand\authorThree{Author 3}
\newcommand\authorFour{Author 4}
\newtheorem{example}{\emph{Example}}

\begin{document}
% explicit hyphenations
\hyphenation{RDF-Re-po-si-to-ry}
\hyphenation{name-space}
\hyphenation{So-cket-A-dap-ter}

%\fontfamily{tahoma}\selectfont
\def\note#1{\marginpar{\footnotesize#1}} % use this to show the notes in the document
%\def\note#1{} % use this to hide the notes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TITLE PAGES 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\thispagestyle{empty}

\pagenumbering{roman}

\begin{flushright}
	\includegraphics[width=3cm]{images/FP7_logo}
\end{flushright}

\vspace{1cm}

%\begin{minipage}[p]{15cm}
	\begin{center}
		\includegraphics{images/FAST_logo}\\
		\vspace{1cm}
		{\LARGE{\sffamily \emph{FAST AND ADVANCED STORYBOARD TOOLS}}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{FP7-ICT-2007-1-216048}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{http://fast.morfeo-project.eu}}\\
		\vspace{4cm}
		{\LARGE \sffamily \textbf{Deliverable \deliverableNumber}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \textbf{\deliverableTitle}}\\
		\vspace{2cm}
		{\large \sffamily \authorOne}\\
		{\large \sffamily \authorTwo}\\
		\vspace{0.5cm}
		\vfill
		{\large \sffamily Date: 27/02/2009}\\
		\vspace{1cm}
		{\sffamily FAST is partially funded by the E.C. (grant code: FP7-ICT-2007-1-216048).}

	\end{center}
%\end{minipage}

\clearpage
%%%%%%%%%%%%%%
% NEXT PAGES %
%%%%%%%%%%%%%%
\pagestyle{scrheadings}

\lohead{\includegraphics[width=4cm]{images/FAST_logo_transparent}}
%\cohead{\small\textcolor{fast@lightgrey}{\deliverableTitle}}
%\rohead{\small{\today}}
%\lofoot{\small\textcolor{fast@lightgrey}{Task Force Ontologies}}
\cofoot{\small{FAST 216048 --- \deliverableTitleShort}}
\rofoot{\small{\thepage}}

\newpage
%
%\section*{Authors}
%\authorOne \\
%%\authorTwo \\
%%\authorThree\\
%%\authorFour\\
%
%
%\vfill
%%\section*{Project Co-ordinator}
%%\hspace*{-2,5cm}\begin{minipage}[p]{14cm}
%%Dr. Ansgar Bernardi \\
%%German Research Center for Artificial Intelligence (DFKI) GmbH \\
%%Trippstadter Strasse 122 \\
%%D 67663 Kaiserslautern \\
%%Germany \\
%%Email: bernardi@dfki.uni-kl.de, phone: +49 631 205 3582, fax: +49 631 205 4910 \\
%%\end{minipage}
%
%
%\section*{Partners}
%Telef\'onica Investigaci\'on y Desarrollo (Telef\'onica I+D), Spain \\
%National University of Ireland, Galway (NUIG), Republic of Ireland \\
%Kassel University, Germany \\
%Cyntelix Corporation, Republic of Ireland \\
%SAP AG, Germany \\
%Universidad Polit\'ecnica de Madrid (UPM), Spain
%
%
%\vspace{0.3cm}
%\begin{footnotesize}
%Copyright: FAST Consortium 2009\\
%\end{footnotesize}

%\clearpage

\section*{Version History}

\begin{small}
\begin{tabular}{|l|l|l|p{7.5cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Rev. No.}} &
                            \textcolor{white}{\textbf{Date}} &
                            \textcolor{white}{\textbf{Author (Partner)}} &
							\textcolor{white}{\textbf{Change description}}\\ \hline
%0.1 & 15.12.2008 & Knud M\"oller (NUIG) & template modelled \\ \hline
%0.2 & 15.02.2009 & Knud M\"oller (NUIG) & content moved from wiki after internal review \\ \hline
1.0 & 27.02.2009 & Knud M\"oller (NUIG) & final version ready for external review \\ \hline
\end{tabular}
\end{small}

\color{black}

\vfill
%{\bf Explanations of abbreviations on front page}\\
%\\
%%Nature \\
%R: Report \\
%P: Prototype \\
%R/P: Report and Prototype \\
%O: Other \\
% \\
%Dissemination level \\
%PU: Public \\
%PP: Restricted to other FP6 participants \\
%RE: Restricted to specified group \\
%CO: Confidential, only for NEPOMUK partners \\

\newpage

%%%%%%%%%%%%%%%%%%%%%
% Executive Summary %
%%%%%%%%%%%%%%%%%%%%%

\clearpage

\section*{Executive Summary}
\doublespacing

This deliverable defines the ontology and conceptual model for the semantic characterisation of complex gadgets in the FAST project. As such, it feeds directly into the various implementation efforts within the project. In the document, we present the ontology development methodology (Methontology) and design principles we have applied. Following Methontology, our design and development process results in a number of different living documents, which both function as a development tool, as well as documentation for the ontology itself. The major part of the deliverable is made up of these documents, resulting in the final \emph{implementation document} which formally defines the ontology in OWL DL semantics.

\newpage

%%%%%%%%%%%%%%%%%%%%%
% Document Summary %
%%%%%%%%%%%%%%%%%%%%%

\clearpage

\section*{Document Summary}
% double spacing from here on:
\singlespacing
\begin{small}

\begin{tabular}
	%{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{6cm}|p{2cm}|p{2cm}|}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{6cm}|p{2cm}|p{2cm}|}
	\hline
	\textcolor{white}{\textbf{Code}} & {FP7-ICT-2007-1-216048} & {\textbf{Acronym}} & {FAST}\\ \hline
	\textcolor{white}{\textbf{Full title}} & \multicolumn{3}{l|}{Fast and Advanced Storyboard Tools}\\ \hline
	\textcolor{white}{\textbf{URL}} & \multicolumn{3}{l|}{\url{http://fast.morfeo-project.eu}}\\ \hline
	\textcolor{white}{\textbf{Project officer}} & \multicolumn{3}{l|}{Annalisa Bogliolo}\\ \hline
\end{tabular}

\vspace{0.5cm}

\begin{tabular}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.25cm}|p{1cm}|p{1cm}|p{6.32cm}|}
	\hline
	\textcolor{white}{\textbf{Deliverable}} & {\textbf{Number}} & {\deliverableNumber} & {\textbf{Name}} & {\deliverableTitle}\\ \hline
	\textcolor{white}{\textbf{Work package}} & {\textbf{Number}} & {\workpackageNumber} & {\textbf{Name}} & {\workpackageTitle}\\ \hline
\end{tabular}

\vspace{0.5cm}

\begin{tabular}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
	\hline
	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2009} & {\textbf{Submitted}} & {27/02/2009}\\ \hline
	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.2.1 is the first iteration of the FAST Gadget ontology. The ontology provides a formal description of the conceptual model of FAST in general, and of the screen/screenflow architecture in particular. The document comprises a discussion of the ontology development methodology used, a number of intermediate representations which document the development process, and the final product in the form of an OWL ontology.}\\ \hline
	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne, \authorTwo}\\
%	{} & \multicolumn{4}{l|}{}\\ 
%	{} & \multicolumn{4}{l|}{}\\ 
%	{} & \multicolumn{4}{l|}{}\\
	\hline
	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {knud.moeller@deri.org} \\ \cline{4-5}
	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495086} \\ \hline
	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, conceptual model, ontology, OWL, RDFS}\\ \hline
\end{tabular}
\end{small}
\newpage

%%%%%%%%%%%%%%%%%%%%%
% TABLE OF CONTENTS %
%%%%%%%%%%%%%%%%%%%%%
\doublespacing
\setcounter{tocdepth}{3}
\tableofcontents
\cleardoublepage
\pagenumbering{arabic}

%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGINNING OF SECTIONS %
%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\rofoot{\small{Page \thepage\ of \pageref{LastPage}}}

\section{Introduction} % (fold)
\label{sec:introduction}

\subsection{Goal and Scope} % (fold)
\label{sub:goal_and_scope}

As the corner stone of the theoretical work undertaken in the FAST project, the definition of a conceptual model and ontology for the characterisation of complex gadgets feeds directly into the project's three main development activities: the development environment for complex gadgets (GVS), the development of gadget components and the semantic catalogue which represents the backend of the architecture. For all these different aspects of FAST, the ontology developed and presented in this deliverable provides a structured definition of their common domain of discourse. The ontology formally defines the parts which a complex gadget is made of, how the parts interrelate, what kind of metadata is available for each part, how users of the system are represented, etc.

While following an ontology design methodology which is agnostic to particular data models or implementation languages (see below), we have chosen to represent the final ontology specification using the Resource Description Framework (RDF) and OWL semantics.

In FAST, we have adopted the term \emph{gadget} for the small, Web-based software components which the project focusses on. There are small differences in meaning between this term and the term \emph{widget}, which is also widely used. However, for the purpose of this document, it should be noted that whenever we use the term \emph{gadget}, we also refer to what \emph{widget} at the same time.

% subsection goal_and_scope (end)

\subsection{Structure of the Document} % (fold)
\label{sub:structure_of_the_document}

This deliverable is structured as follows: as part of the introduction, we will present the methodology we chose to adopt for the development of the FAST gadget ontology, by discussing  each of its proposed stages in detail. Following, we will conclude the introduction by elaborating more on some of the general design decisions we are following. After briefly addressing the topic of related work in Sect.~\ref{sec:related_work} --- which is mainly covered in a different deliverable from this work package, \cite{urmetzer2009fast_state_of_the_art} --- the structure of the remainder of this document follows directly from our adopted methodology: we will perform a domain analysis in Sect.~\ref{sec:domain_analysis}, covering the stages of specification, conceptualisation and integration of existing vocabularies. This is then followed by the implementation step in Sect.~\ref{sec:ontology}, which covers the ontology in its final form (at this moment of the project). Finishing the deliverable, we present our conclusions and outlook on years two and three of the project in Sect.~\ref{sec:conclusions}.

% subsection structure_of_the_document (end)

\subsection{\emph{Methontology}: A Methodology for Ontology Development} % (fold)
\label{sub:methontology_a_methodology_for_ontology_development}

In order to put the design of the ontology on a stable footing, we chose to adopt a tried and tested design methodology, rather than using a purely ad-hoc approach. We decided to use the \emph{Methontology} methodology, which was first introduced in \cite{gomez_perez1996conceptualise_ontologies} and \cite{fernandez1997methontology}. Methontology provides a good balance of formalisation and streamlining of the design process on the one hand, and a looseness in its requirements on how strictly one needs to follow the individual phases and steps. The authors specifically advocate an ``evolving prototype life cycle'', which allows the ontology to ``grow depending on its needs'' \cite{fernandez1997methontology}.

Methontology consists of seven different stages, which will be briefly described in the following sections. For a more detailed discussion, we point the reader to the original papers proposing this methodology. The phases are \emph{specification}, \emph{knowledge acquisistion}, \emph{conceptualisation}, \emph{integration}, \emph{implementation}, \emph{evaluation} and \emph{documentation}. It is important to note that the phases do not have to be visited slavishly in the order presented here. Instead, a much more likely scenario is that the ontology designers will visit each phase roughly in this order, but are free to revisit each phase at any time to apply changes.

\subsubsection{Specification} % (fold)
\label{ssub:specification}

The specification phase sets the stage for the rest of the ontology development process. The central activity in this phase is the setting up of an \emph{ontology requirement specification document}, which will act as a guideline for most of the other phases. The level of formality of the specification document is left open and can range from pure natural language, over a set of intermediate representations to using competency questions. Most likely, it will be a mix of those formats. Regardless of the chosen format of the specification document, it must answer questions regarding the \emph{purpose} of the ontology (intended use and users, scenarios, etc.), its \emph{scope} (what are the things that need to be covered), its \emph{level of formality} (e.g. highly informal, semi-formal or rigorously formal \cite{uschold1996ontologies}) and the \emph{sources of knowledge} used when researching the ontology. It should be noted that the scope as defined in the specification document does not need to be complete. Rather, it should provide a good impression of the kind of terms that need to be covered by the ontology.

The \emph{Domain Analysis} section of this deliverable implements an ontology requirement specification document for the FAST gadget ontology (which is not to be confused with the FAST requirements deliverable \cite{villoslada2009fast_requirements}!).
% subsubsection specification (end)

\subsubsection{Knowledge Acquisition} % (fold)
\label{ssub:knowledge_acquisition}

Knowledge Acquisition in the context of Methontology is a loose collective term to all activities which are aimed at gathering background knowledge to clearly determine purpose and scope of the ontology. As such, it feeds directly into the specification phase, but can just as well be relevant at later stages of the design and development process. Similarly to evaluation and documentation, knowledge acquisition is a supporting activity, which takes place throughout the whole development process.

Typical ways of knowledge acquisition are referring to handbooks, encyclopaedias or other written material (through formal or informal text analysis), performing interviews with domain experts (structured or non-structured) or evaluating previous conceptualisation work done for the domain in question.

The primary sources used for knowledge acquisition for the the FAST gadget ontology are referred to in Sect.~\ref{sec:related_work}.
% subsubsection knowledge_acquisition (end)

\subsubsection{Conceptualisation} % (fold)
\label{ssub:conceptualisation}

The conceptualisation phase stands at the core of the ontology development process. The ontology is still in a language-/format-independent state, but is becoming increasingly formal. Based on the specification document, the ontology designers now create a number of increasingly detailed and fine-grained tables and dictionaries which cover all terms to be included in the ontology (at the current stage - an ontology in the Semantic Web sense is never complete). In the first iteration, a so-called \emph{glossary of terms} (GT) is built, which includes all concepts (or classes), instances and properties (or verbs). The GT does not have to be built from scratch, but can instead be understood as an extension of the scope definition in the specification phase.  In contrast to the specification document, the GT is a living document that should always reflect the latest and most complete list of terms. Once completed (in a first iteration), the terms in the GT will be grouped according to relatedness. For each group, the concepts are arranged in a classification tree. Instances, constants and attributes are grouped in corresponding tables, while properties are captured in verb diagrams. If necessary, tables with formulas and rules are set up at the end of the conceptualisation phase.
% subsubsection conceptualisation (end)

\subsubsection{Integration} % (fold)
\label{ssub:integration}

The integration phase offers an opportunity for the ontology designer to ensure that their ontology is well integrated with other, existing ontologies. This can either mean connecting the emerging ontology to upper or meta ontologies (in the case that there are appropriate super classes or properties in those ontologies which can act as connection joints), or the reuse of terms from other ontologies (in the case that other ontologies have already defined classes or properties which fall in the scope of the emerging ontology). A typical example for a meta-ontology is OpenCyc, while the FOAF vocabulary \cite{brickley2004foaf} is a example of an often-reused ontology to describe people. As the output of this phase, an \emph{integration document} is suggested, which gives detailed information about which terms were taken from which external ontology.
% subsubsection integration (end)

\subsubsection{Implementation} % (fold)
\label{ssub:implementation}

The implementation phase is what is often (wrongly so) considered to be the main activity in developing an ontology: materialising the various ontology terms in a concrete language, which can be RDFS or OWL for the Semantic Web, or any other formal language in other contexts (even a programming language such as C++). The authors of Methontology do not go into a lot of details regarding the carrying out of this phase, other than that the outcome will be the formalisation of the ontology in the ontology language of choice (the \emph{implementation document}).
% subsubsection implementation (end)

\subsubsection{Evaluation} % (fold)
\label{ssub:evaluation}

According to [Fernández et  al., 1997], evaluation ``means to carry out a technical judgment of the ontologies, their software environment and documentation with respect to a frame of reference'' (in this case the requirement specification document). Evaluation can take place at any time during the ontology development process, and comprises activities such as \emph{validation} and \emph{verification}.

For the FAST Gadget Ontology, most of the evaluation process will take place as part of the work carried out in WP6 (Experimentation and Evaluation) in years two and three of the project.
% subsubsection evaluation (end)

\subsubsection{Documentation} % (fold)
\label{ssub:documentation}

In Methontology, documentation is an activity which is automatically carried out throughout the development process, in the form of the various documents which form the output of the individual phases. Since this deliverable contains the current versions of all those documents, it is at the same time the ontology itself as well as its documentation.
% subsubsection documentation (end)

% subsection methontology_a_methodology_for_ontology_development (end)

\subsection{General Design Decisions} % (fold)
\label{sub:general_design_decisions}

\subsubsection{Reuse of Existing Ontologies} % (fold)
\label{ssub:reuse_of_existing_ontologies}

An important aspect to consider when designing any ontology which is aimed to be used not only within a restricted institution or community is to reuse terms from existing ontologies and vocabularies as much as possible. Reuse serves the purpose of lowering the entry barrier for third parties to adopt the ontology --- if they can use familiar concepts and terms, they will be more inclined to try out something new. Equally important is the fact that reuse of terms facilitates data integration, which, after all, is one of the major goals of the Semantic Web at large.
% subsubsection reuse_of_existing_ontologies (end)

\subsubsection{Modularisation} % (fold)
\label{ssub:modularisation}

Apart from reusing existing terms, another basic design principle which helps to improve accessibility and understanding of the ontology is modularisation. While all terms will (for the time being) reside in the same namespace, classes and properties will nevertheless be grouped according to the specific sub-task which they belong to. E.g., terms will be grouped into defining \emph{resources}, \emph{pre- and post-conditions}, \emph{general annotation} or \emph{users and user profiles}.
% subsubsection modularisation (end)

% subsection general_design_decisions (end)

% section introduction (end)

\section{Related Work} % (fold)
\label{sec:related_work}

We are not aware of any other project which has already developed or planned to develop an ontology of the gadget domain. However, there is a range of related material from which we take inspiration regarding the description of the gadget domain. This material comes from a number of different sources and directions. It includes work done under the hood of the W3C which may eventually lead to an official specification of various technical aspects of the gadget domain, as well as a number of different gadget APIs. While both are not strictly speaking ontologies, they nevertheless provide a very good insight in how to conceptualise the domain. Additionally, we consider work done in the area of Semantic Web Services, firstly because the gadgets designed in FAST will connect to Semantic and conventional Web services, and secondly because there are a number of similarities between the way SWS are often formalised and the way we envision the interaction of the different components of a gadget in the FAST IDE.

All of these references are discussed in \cite{urmetzer2009fast_state_of_the_art}, and we refer the reader to this deliverable for more detail.
% section related_work (end)

\section{Domain Analysis} % (fold)
\label{sec:domain_analysis}

In this section, we will apply a number of the phases proposed in Methontology to the process of developing the FAST gadget ontology. The names of the following sections reflect the names of the phases in Methontology.

\subsection{Specification} % (fold)
\label{sub:specification}

The specification phase of Methontology requires setting up an ontology requirements specification document. We have compiled such a document for the FAST Gadget Ontology, as in Tab~\ref{tab:ontology_requirements_spec}.

\singlespacing
\begin{small}
\begin{longtable}[t]{|>{\columncolor{fast@lightgrey}}p{3.25cm}|p{11cm}|}
\caption{\label{tab:ontology_requirements_spec}FAST Ontology Requirements Specification Document}\\
\hline
\textcolor{white}{\textbf{Name}} & \textbf{FAST Gadget Ontology} \\ \hline
\textcolor{white}{\textbf{Domain}} & Intelligent Gadgets \\ \hline
\textcolor{white}{\textbf{Purpose}} & The FAST gadget ontology conceptualises the domain of intelligent gadgets as defined in the FAST development platform. Gadgets consist of several interrelated parts, all of which are covered in the ontology. In FAST, a description of each gadget component and resource is available to the IDE. Form this description, the IDE can construct its interface, determine which components can be connected to which other components, or make suggestions to the user in order to aid in the gadget development process.

Furthermore, FAST gadgets are capable of connecting to a variety of backend Web services, which can either be Semantic Web services, or conventional, non-semantic Web services. Therefore, the FAST Gadget Ontology must facilitate the description of those backend services as well. Where a semantic description of the service already exists, it may be necessary to perform ontology mediation between the ontology of the backend service description and the FAST ontology.

Finally, the FAST IDE will support individual user profiles to allow personalisation of the gadget development process. This means that the FAST gadget ontology will also have to cover the description of users and user profiles.

In summary, the ontology is supposed to facilitate support for the user in the design and generation of FAST gadgets, as well as in searching and browsing those gadgets.\\ \hline 
\textcolor{white}{\textbf{Level of Formality}} & formal (OWL ontology) \\ \hline
\textcolor{white}{\textbf{Scope}} & \textbf{Concepts:} \emph{Component}, \emph{Resource}, \emph{Screen}, \emph{Backend Service}, \emph{Flow Control Element}, \emph{Operator}, \emph{Screenflow Start}, \emph{Screenflow End}, \emph{Connector}, \emph{Form Element}, \emph{Pre-condition}, \emph{Post-condition}, \emph{Query}, \emph{Label}, \emph{Icon}, \emph{User}, \emph{User Profile}, \emph{Tag}

\textbf{Properties:} \emph{containsScreen}, \emph{hasLabel}, \emph{hasIcon}, \emph{hasPreCondition}, \emph{hasPostCondition}, \emph{hasTag}, \emph{hasProfile}\\ \hline
\textcolor{white}{\textbf{Sources of Knowledge}} & FAST Architecture deliverable~\cite{solero2009fast_architecture}, FAST Requirements Specification~\cite{villoslada2009fast_requirements}, FAST State-of-the-art Document~\cite{urmetzer2009fast_state_of_the_art}, EzWeb documentation~\cite{lizcano2008ezweb} \\ \hline
\end{longtable}
\end{small}
\doublespacing

% subsection specification (end)

\subsection{Conceptualisation} % (fold)
\label{sub:conceptualisation}

In the conceptualisation phase, we first produce a \emph{glossary of terms} which lists all classes and properties of our ontology. The glossary is seeded from the specification document (see previous section), but groups the terms according to relatedness. Also, new terms are added to this document, whereas the specification document stays the same.

\subsubsection{Glossary of Terms} % (fold)
\label{ssub:glossary_of_terms}

\paragraph{Classes} % (fold)
\label{par:classes}

``Classes'' here should be read as ``types of things'' in general. I.e., if a term is listed as a \emph{class} here, this has no direct implications on the concrete implementation of this term. For example, while \emph{label} is a type of thing that is important in FAST, it will not necessarily be represented as an \texttt{rdfs:Class} in the implementation step. It could just as likely be represented as a property.

\emph{Components/Resources}

\begin{itemize}
	\item \texttt{Component} (or \texttt{Resource}) - Anything that is part of a gadget (or the gadget itself). Tentatively anything that can be ``touched'' and moved around in the FAST IDE.
	\begin{itemize}
		\item \texttt{Screen\_Flow} - A set of screens from which a gadget for a given target platform can be generated.
		\item \texttt{Screen} - An individual screen.
		\item \texttt{Backend\_Service} - A Web service which provides data and/or functionality to a screen. A backend service will often be external to FAST, and will probably have to be wrapped by the screen.
		\item \texttt{Flow\_Control\_Element} - Any kind of component which can restrict the default flow of screens in a gadget.
		\begin{itemize}
			\item \texttt{Screen\_Flow\_Start} - The entry point to a wigdet; the first screen.
			\item \texttt{Screen\_Flow\_End} - A screen that ends the workflow of the gadget.
			\item \texttt{Connector} - An explicit connection between two screens.
		\end{itemize}
		\item \texttt{Operator} - Any kind of component that is used to connect backend services to form elements. Examples are simple pipes, aggregators or various kinds of filters. Operators concern the inner mechanics of a screen and will be covered in years 2 and 3 of the project.
		\item \texttt{Form\_Element} - Form elements are UI elements in a particular screen.
	\end{itemize}
\end{itemize}

\emph{Pre- and Post-conditions}

\begin{itemize}
	\item \texttt{Condition} - The pre- or post-condition of either a screen or a screenflow. In the latter case, each target platform will use these conditions in its own way, or may also ignore them. E.g., in EzWeb pre- and post-conditions correspond to the concepts of \emph{slot} and \emph{event}.
	\item \texttt{Pattern} - A set of facts which formally defines a condition. A pattern may contain one or more variables.
	\item \texttt{Fact} - ``the basic information unit of a FAST gadget''~\cite{solero2009fast_architecture} In terms of RDF, a fact will probably be one statement consisting of \emph{(S, P, O)}.
\end{itemize}

\emph{Annotation}

\begin{itemize}
	\item \texttt{String} - A catch-all class of objects that serves for representing short labels, longer descriptions, dates, patterns, etc. If meant for human consumption, strings can have multiple representations for different languages.
	\item \texttt{Image} - Images are any kind of graphical object, such as icons, screenshots, etc.
\end{itemize}

\emph{Users}

\begin{itemize}
	\item \texttt{User} - A human user of the FAST IDE.
	\item \texttt{User\_Profile} - The settings and data known about a particular user.
\end{itemize}

% paragraph classes (end)

\paragraph{Properties} % (fold)
\label{par:properties}

All properties are listed in Tab.~\ref{tab:glossary_of_terms_properties} (many of the properties covered in this table will also have an inverse).

\singlespacing
\begin{small}
\begin{longtable}{|p{3.25cm}|p{11cm}|}
\caption{\label{tab:glossary_of_terms_properties}FAST Glossary of Terms, Properties}\\
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Name}} & 
	\textcolor{white}{\textbf{Description}} \\ \hline
\endfirsthead
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Name}} & 
	\textcolor{white}{\textbf{Description}} \\ \hline
\endhead
\multicolumn{2}{|c|}{\textbf{Components/Resources}} \\ \hline
\texttt{contains} & Many kinds of components in FAST can contain other components: screenflows contain screens, screens contain forms or form elements, etc. \\ \hline
\texttt{hasPreCondition} & This property links a screen or screenflow to its pre-condition, i.e., the facts that need to be fulfilled in order for this screen or screenflow to be reachable. \\ \hline
\texttt{hasPostCondition} & This property links a screen or screenflow to its post-condition, i.e., the facts that are produced once the screen or screenflow has been executed. \\ \hline
\multicolumn{2}{|c|}{\textbf{Pre- and Post-conditions}} \\ \hline
\texttt{hasPattern} & This property links a condition resource to the pattern which formally defines it. \\ \hline
\texttt{isPositive} & Conditions can be positive or negative, depending on whether they must be fulfilled or must not be fulfilled (in the case of pre-conditions), or whether their facts will be added to the canvas or removed (in the case of post-conditions). \\ \hline
\multicolumn{2}{|c|}{\textbf{Annotation}} \\ \hline
\texttt{hasLabel} & A string attached to any FAST component or sub-component, which represents a short ($\sim$1-2 word) human-readable description of the component. \\ \hline
\texttt{hasDescription} & A string attached to any FAST component or sub-component, which represents a longer, more detailed human-readable description of the component in natural language. \\ \hline
\texttt{hasImage} & An abstract property defining the relation between a thing in FAST and an image which somehow represents it. \\ \hline
\texttt{hasIcon} & A small graphical representation of any FAST component or sub-component. \\ \hline
\texttt{hasScreenshot} & An image which shows a particular screen or screenflow in action, to aid users in deciding which screen or screenflow to choose out of many. \\ \hline
\texttt{hasRights} & A human-readable description of license information or similar for a screen, screenflow or possibly other components. \\ \hline
\texttt{hasCreator} & Every resource generated within FAST will contain metainformation about the person who created it. \\ \hline
\texttt{hasVersion} & A string representing the version number of a resource in FAST. \\ \hline
\texttt{hasCreationDate} & A date-formatted string representing the date when a resource in FAST was created first. \\ \hline
\texttt{hasHomepage} & Links a resource in FAST to a main Webpage with information about it. We expect screenflows/gadgets to have homepages, as well as users. Other resources will most likely not have homepages. \\ \hline
\texttt{hasDomainContext} & A catch-all property to annotate a resource in FAST with information about the domain for which it is relevant. The domain context will be utilised for searching, matching, etc. Domain contexts can be expressed either as tags or structured objects. \\ \hline
\multicolumn{2}{|c|}{\textbf{User and User Profiles}} \\ \hline
\texttt{hasName} & The full name of a FAST user. \\ \hline
\texttt{hasUserName} & The user name of a FAST user, which will often be a short form of the full name, or a nick name. The user name must be a unique ID within an instance of FAST. \\ \hline
\texttt{hasEmail} & The e-mail address of a FAST user. \\ \hline
\texttt{hasProfile} & Connecting a user with their profile. \\ \hline
\texttt{hasInterests} & Interests of a user as specified in their profile. Interests could be matched with a component's domain context to allow the FAST catalogue (and thereby the GVS) to suggest screens and other components to the user. \\ \hline
\end{longtable}
\end{small}
\doublespacing

\subsection{Integration} % (fold)
\label{sub:integration}

In this section we will investigate out a number of established ontologies which cover part of the requirements laid out in the specification document, as well as the glossary of terms from the conceptualisation stage. In particular, we will look at the Friend of a Friend (FOAF) ontology, Semantically Interlinked Online Communities (SIOC) and the Dublin Core metadata specification. The outcome of this process will be a living \emph{integration document} at \ref{ssub:integration_document}.

\subsubsection{FOAF} % (fold)
\label{ssub:foaf}

The Friend of a Friend vocabulary is a simple ontology which main purpose to describe people --- represented as instances of the \texttt{foaf:Person} class --- in terms of their contact information, interests, or Web presence. More importantly, however, is the fact that FOAF provides some simple terms for specifying who knows who (using the propery \texttt{foaf:knows}), thereby effectively allowing to build a social network of people. There is no such thing as a centralised FOAF service to which people have to sign up. Instead, each person can maintain and host their own FOAF description (often in the form of a \emph{FOAF file}), or use one of many external services which provide this functionality, such as LiveJournal, TypePad, Vox, Hi5, etc., thus making the FOAF network a truly decentralised social network. Figure ~\ref{fig:foaf_example} depicts an excerpt of some typical FOAF files, describing two people called ``Knud M\"oller'' and ``Andreas Harth'', including the fact that Knud knows Andreas.

\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{images/foaf_example.pdf}
    \caption{Example of FOAF data}
    \label{fig:foaf_example}
  \end{center}
\end{figure}

FOAF has been \emph{``evolving gradually since its creation in mid-2000''}~\cite{brickley2004foaf}. Over the years, it attracted a lot of attention and has become one of the major successes of the Semantic Web. Compared to most other ontologies or vocabularies, it has received a much higher uptake, with the number of FOAF files on the internet probably numbering tens of millions now. Its popularity in the community and beyond is also underlined by the fact that many other ontologies are integrating FOAF in order to represent basic information regarding people, or they are using it as a point of reference and extend the basic FOAF classes and properties for their own needs. Examples for this kind of integration are the increasingly popular SIOC (Semantically-Interlinked Online Communities) ontology~\cite{breslin2005communitiesEswc}, which uses FOAF description to unify identities in different online communities, or the Semantic Web Conference ontology~\cite{moeller2007dogfood}, where conference attendees or authors are represented as FOAF person instances.

The specification document of the FAST ontology defines that part of the purpose of the ontology is ``the description of users and user profiles''. The most obvious solution for implementing this requirement is to integrate the FAST ontology with FOAF. This means that each user of FAST would be modelled as a \texttt{foaf:Person}. A lot of the basic necessities in terms of user profile information is covered by properties defined in FOAF (names, contact details, interests, user pictures, etc.) and could therefore be used as is. In other cases, rather generic FOAF terms might be good starting points for extension towards more specific needs that occur in FAST. The icons of various components in the GVS may serve as an example here: the relation that holds between an icon and the thing X of which it is an icon could be described as ``the icon depicts X''. FOAF provides the property \texttt{depiction} to express this general relation. However, \texttt{hasIcon} (our property for saying that something is the icon of a thing) can be considered a specialisation of this, and so we will say that \texttt{fast:hasIcon} is a specialisation of \texttt{foaf:depiction}. In terms of RDFS, we will say that \texttt{<fast:hasIcon> <rdfs:subPropertyOf> <foaf:depiction>}.
% subsubsection foaf (end)

\subsubsection{SIOC} % (fold)
\label{ssub:sioc}

Another vocabulary that is gaining a lot of uptake and support on the Semantic Web is SIOC, which received additional weight when it became a W3C member submission in 2007. The basic idea behind SIOC is that there is an abstract model behind all online community sites which contain an aspect of discussion between members, be it forum sites, discussion boards, blogs, wikis, content management systems, mailing lists, etc. For each of those \emph{sites}, it can be said they contain discussion threads or \emph{fora}, which in turn contain \emph{posts}, which in turn can have \emph{comments}. Furthermore, each post or comment can have a \emph{topic} and a \emph{creator}, which can be a \emph{user} of the site. The authors of SIOC make a point of integrating the vocabulary with FOAF, e.g., by suggesting that a user of a site is in fact held by an instance of \texttt{foaf:Person} (the SIOC specification states that \texttt{sioc:User} is a sub-class of \texttt{foaf:OnlineAccount} \cite{sioc_spec2009}). An overview of the different classes and properties of SIOC is given in Fig.~\ref{fig:sioc_overview}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{images/sioc_overview}
    \caption{Overview of the SIOC ontology}
    \label{fig:sioc_overview}
  \end{center}
\end{figure}

From an implementation point of view, a particular site will use SIOC by exposing its internal data on request with the help of a SIOC exporter. Such components have been provided by various members in the SIOC developer community for different popular platforms, such as the WordPress blogging software, the Drupal CMS or the Twitter micro-blogging platform. The benefit that sites are creating for end users by using SIOC is that a common representation format and reference points such as authors and topics allow data from different sites to be integrated and thus browsed and searched together.

From the point of view of the FAST gadget ontology, the most interesting features of the vocabulary are the classes and properties it provides for modelling users of a site. We will use SIOC in combination with FOAF to model users and their profiles, as well as user groups, as specified in the requirements document. The discussion aspects of SIOC will not have an immediate use within the FAST platform. However, it is conceivable that collaborative features such as support for fora will be added to FAST at some point, which will lead to an extension of the requirements document. These additional requirements could then be fulfilled by the integration of further terms from SIOC.
% subsubsection sioc (end)

\subsubsection{Dublin Core} % (fold)
\label{ssub:dublin_core}

Dublin Core (DC) is a set of metadata elements for describing online resources in order to support indexing, searching and finding them. Typical examples of terms from DC are \emph{title}, \emph{creator}, \emph{subject}, \emph{description}, \emph{date} or \emph{rights}. In principle, DC can be applied to a wide range of resources, but is typically used in the context of describing media (textual documents, video, sound or image). The initial work on DC was done at an invitational workshop in Dublin, Ohio, US (hence the name) in 1995. Since then, several revisions and extensions have been applied to the vocabulary.

While it is possible to express DC in plain HTML, the more widely used language is probably RDF. The original 15 terms were all properties defined in the DC elements namespace (\url{http://purl.org/dc/elements/1.1/}, short \texttt{dc}). The properties were not formally defined with respect to their range and domain, or whether they are object or datatype properties. However, the assumption was that the values for each property would always be literals, as opposed to complex values represented by a URI. More recently, the DC elements namespace has been declared legacy and is now superseded by the DC terms namespace (\url{http://purl.org/dc/terms/}, short \texttt{dcterms})~\cite{dcterms2008}, which includes more precise re-definitions of all 15 original terms, as well as a number of new terms (resulting in a total of 55 terms). This new iteration of DC now provides the facility to use structured values for properties and specifies what type these values will have. E.g., the property \texttt{dcterms:creator} now supersedes the original \texttt{dc:creator} and specifies that its value has the type \texttt{dcterms:Agent}.

In order to reconcile with the simplicity of the original, flat DC elements and the newer, more structured DC terms, DC also includes the so-called ``DumbDown'' principle, which basically says that structured values of DC terms should always carry a literal description as well, in order to cater for ``dumb'' agents processing the data. These literals should be expressed using properties such as \emph{rdfs:label} or \emph{rdf:value}.

Because of the open nature of the RDF model and the open-world assumption usually applied for RDFS and OWL, it is possible to integrate DC with other ontologies such as FOAF. E.g., it is possible and good practice to make a statement saying that the \texttt{dcterms:creator} of a document is an instance of \texttt{foaf:Person}. By inference following from the DC terms specification, this person would then also be a \texttt{dcterms:Agent}. This scenario is illustrated in Fig.~\ref{fig:dcterms_example}. A similar, but slightly out-dated example is given in \url{http://dublincore.org/documents/dcq-rdf-xml/}.

\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{images/dcterms_example.pdf}
    \caption{Example of Dublin Core data}
    \label{fig:dcterms_example}
  \end{center}
\end{figure}

Looking at the requirements specification document, DC can be integrated into the FAST gadget ontology in many ways. A lot of the metadata needs for screens, screenflows and their components are identical to those of the media resources originally intended by DC. For this reason, many properties such as \texttt{title}, \texttt{creator}, \texttt{rights} or various \texttt{date} properties can be used directly (more details in the integration document). 
% subsubsection dublin_core (end)

\subsubsection{Integration Document} % (fold)
\label{ssub:integration_document}
%
This living document specifies how the terms from the glossary of terms are expressed using terms from the various ontologies presented in the previous sections.

\singlespacing
\begin{small}
\begin{longtable}{|p{2.6cm}|p{1.5cm}|p{3cm}|p{7cm}|}
\caption{\label{tab:integration_document}FAST Ontology integration document}\\
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{FAST Term}} & 
\textcolor{white}{\textbf{Integrated Ontology}} & 
\textcolor{white}{\textbf{Integrated Term}} & 
\textcolor{white}{\textbf{Comment}} \\ \hline
\endfirsthead
\multicolumn{4}{|c|}{\emph{\textbf{Classes}}} \\ \hline
\texttt{User} & FOAF & \texttt{foaf:Person} & A user in FAST will be modelled using a combination of FOAF and SIOC, so that an instance of \texttt{foaf:Person} will have an account on the FAST platform, represented by an instance of \texttt{sioc:User}, which is a subclass of \texttt{foaf:OnlineAccount}. The relationship is expressed using \texttt{foaf:holdsAccount}. This way of modelling directly follows the suggestions made by~\cite{sioc_related2007}.\\ \hline
\texttt{User} & SIOC & \texttt{sioc:User} & s.a.\\ \hline
\texttt{Image} & FOAF & \texttt{foaf:Image} & --- \\ \hline
\multicolumn{4}{|c|}{\emph{\textbf{Properties}}} \\ \hline
\texttt{hasLabel} & DC & \texttt{dcterms:title} & ---  \\ \hline
\texttt{hasCreator} & DC & \texttt{dcterms:creator} & DC defines the range of \texttt{dcterms:creator} to be \texttt{dcterms:Agent}. By inference, this obviously also holds within FAST. However, we will explicitly only use \texttt{foaf:Person} (also see Fig.~\ref{fig:dcterms_example}). \\ \hline
\texttt{hasDescription} & DC & \texttt{dcterms:description} & --- \\ \hline
\texttt{hasCreationDate} & DC & \texttt{created} & Dates should be formatted according to ISO 8601. \\ \hline
\texttt{hasRights} & DC & \texttt{rights} & DC terms specify a number of classes to model the rights of resources. \texttt{dcterms:rights} links a resource to a \texttt{dcterms:RightsStatement}. The holder of the rights is specified by pointing from the resource to a \texttt{dcterms:Agent}, using the \texttt{dcterms:rightsHolder} property. \\ \hline
\texttt{hasDomainContext} & DC & \texttt{subject} & DC specifies that subject should be non-literals. To allow simple tagging, we should represent tags as resources (TODO). \\ \hline
\texttt{hasImage} & FOAF & \texttt{depiction} & The inverse \texttt{foaf:depicts} could also be used. \\ \hline
\texttt{hasName} & FOAF & \texttt{name} & Apart from \texttt{foaf:name}, which is used for the full name, more fine-grained properties such as \texttt{foaf:firstName}, \texttt{foaf:surname} or \texttt{foaf:title} and \texttt{foaf:nick} can be used as well. \\ \hline
\texttt{hasEmail} & FOAF & \texttt{mbox} & While \texttt{foaf:mbox} will be used for the plain e-mail address, \texttt{foaf:mbox\_sha1sum} will be used for an obfuscated version of the address. \\ \hline
\texttt{hasInterests} & FOAF & \texttt{interest} & --- \\ \hline
\texttt{hasUserName} & FOAF & \texttt{accountName} & --- \\ \hline
\end{longtable}
\end{small}
\doublespacing

%% subsubsection integration_document (end)
%
%% subsection integration (end)
%
%% section domain_analysis (end)
%
\section{The FAST Gadget Ontology} % (fold)
\label{sec:ontology}

In the following sections we will define the FAST gadget ontology.

\subsection{Defining Pre- and Post-conditions} % (fold)
\label{sub:defining_pre_and_post_conditions}

The facts which define the pre- and post-conditions of screens and screenflows in FAST will be modelled as individual RDF triples. In effect, this means that conditions, which are sets of facts, can be modelled as graph patterns using SPARQL notation (see \cite{sparql2008spec}). E.g., a simple pre-condition such as \emph{``There has to be a user''} could be expressed as in example~\ref{ex:simple_precondition}. Literally, this very simple pattern means \emph{``a variable ?user is of type sioc:User''}. In logical terms, it means something along the lines of \emph{``there exists a sioc:User''}. For the pre-condition to be fulfilled, it needs to be executed against the RDF graph in question (most likely the facts on the current canvas).

\begin{example}
\begin{verbatim}

?user a sioc:User .
\end{verbatim}
\label{ex:simple_precondition}
\end{example}

Post-conditions will be expressed in a similar fashion. At design time, post-condition patterns will have variable in the same way as pre-condition patterns. When a screen is added to the canvas, the pattern needs to be materialised. To do this, each variable will be replaced with a randomly generated URI or blank node. At runtime, variables will be replaced with actual values once the screen for which the post-condition holds has been executed. As an example, consider the post-condition of a login screen. In natural language, it could be \emph{``Once the login process has finished, there will be a user object''}. Using the same notation as before, this could be expressed as shown in example~\ref{ex:postcondition}, extended with some additional facts (\emph{``There is a user object which has an account name. There is also a person which has a name, and which has the user object as an online account.''}).

\begin{example}
\begin{verbatim}

?user a sioc:User ;
      foaf:accountName ?account_name .
?person a foaf:Person ;
      foaf:holdsAccount ?user ;
      foaf:name ?person_name .
\end{verbatim}
\label{ex:postcondition}
\end{example}

Once added to the canvas in the GVS (``instantiated at design time''), each variable would be replaced with a blank node identifier, as shown in example~\ref{ex:postcondition_canvas}. The pre-condition in example~\ref{ex:simple_precondition} could now be executed successfully as a SPARQL query against the canvas graph.

\begin{example}
\begin{verbatim}

_:b0 a sioc:User ;
      foaf:accountName _:b1 .
_:b2 a foaf:Person ;
      foaf:holdsAccount _:b0 ;
      foaf:name _:b3 .
\end{verbatim}
\label{ex:postcondition_canvas}
\end{example}

Finally, during runtime, the post-condition's variables would be replaced with the actual values of the user which completed the login screen (``instantiated during runtime''). E.g., this could result in a graph such as in example~\ref{ex:postcondition_runtime}.

\begin{example}
\begin{verbatim}

<http://fast.org/gvs/knud> a sioc:User ;
      foaf:accountName "dunk" .
<http://kantenwerk.org/metadata/foaf.rdf#me> a foaf:Person ;
      foaf:holdsAccount <http://fast.org/gvs/knud> ;
      foaf:name "Knud Möller" .
\end{verbatim}
\label{ex:postcondition_runtime}
\end{example}

Since graph patterns with variable cannot directly be expressed in RDF (short of using blank nodes, which cannot have labels), the graph patterns are expressed as RDF literals and linked \texttt{fast:Condition} instances using the \texttt{fast:hasPattern} property.
Each such condition will be linked to a screen or screenflow using either the \texttt{fast:hasPreCondition} or \texttt{fast:hasPostCondition} property.

Since SPARQL has no direct support for negation, and since there is a need to allow screens to remove facts from the canvas graph as well as adding them, both pre- and post-conditions can be modelled to be either \emph{positive} or \emph{negative} using the \texttt{fast:isPositive} property. For pre-conditions, the interpretation of \emph{positive} is that the condition must hold, while \emph{negative} means that the negation of the condition must hold. For post-conditions, \emph{positive} leads to the instantiated condition being added to the canvas, while \emph{negative} will remove the instantiated condition. These interpretations are summarised in Tab.~\ref{tab:is_positive}.

\begin{table}[h]
\caption{Different semantics for \texttt{isPositive}}
\label{tab:is_positive}
\begin{center}
\begin{tabular}{|l|c|c|}
\hline
& pre-condition & post-condition \\ \hline
\emph{isPositive ``YES''} & condition must be fulfilled & condition instantiated\\ \hline
\emph{isPositive ``NO''} & $\neg$condition must be fulfilled & instantiated condition removed \\ \hline
\end{tabular}
\end{center}
\end{table}

\subsection{Classes} % (fold)
\label{sub:classes}

This section contains a definition of all classes defined in the \texttt{fast} namespace. For classes integrated from other ontologies, see the integration document~\ref{ssub:integration_document}.

\singlespacing
\begin{small}
\input{classes}
\end{small}
\doublespacing

%subsection classes (end)

\subsection{Properties} % (fold)
\label{sub:properties}

This section contains a definition of all properties defined in the \texttt{fast} namespace. For properties integrated from other ontologies, see the integration document~\ref{ssub:integration_document}.

\singlespacing
\begin{small}
\input{properties}
\end{small}
\doublespacing

% subsection properties (end)

\subsection{Extension towards Specific Domains} % (fold)
\label{sub:extension_towards_specific_domains}

At this point in the development of FAST, there are three main entry points at which external ontologies come into play:

\begin{itemize}
	\item \emph{Domain Contexts:} By providing a domain context for a screen (or any other FAST resource), it is possible to express what domain this resource is relevant for. Examples for domain contexts are ``medicine'', ``EU projects'', ``catering'', ``human resources'', ``books'', etc. As defined in the Integration Document in Sect.~\ref{ssub:integration_document}, domain contexts are given using the \emph{dcterms:subject} property. One option for finding domain identifiers to be used with this property is the use of DBpedia~\cite{Auer07dbpedia} identifiers such as \url{http://dbpedia.org/resource/Medicine} (DBpedia is a large RDF corpus automatically extracted from the info boxes of Wikipedia entries). This is now a widely used and recommended practice in the Semantic Web community (similar corpora could also be used). As an alternative or supplement to DBpedia identifiers, it is possible to use structured tagging approaches such as MOAT\footnote{\url{http://moat-project.org/}}.
	\item \emph{User Interests:} In the same way as domain contexts, a user's interests can be specified using DBpedia identifiers or structured tagging. Both can then be used in combination to suggest resources to the user.
	\item \emph{Pre- and Post-conditions:} In specifying pre- and post-conditions, terminology for an open-ended number of domains is necessary. While a number of classes will be reused quite often, very specific screens will require very specific vocabulary. We cannot, as part of the FAST project, provide vocabulary for all conceivable domains of discourse, and while also here DBpedia identifiers may be an option, often more precise terms will be needed.
\end{itemize}

It should be added that for any given screenflow, the concrete definition of both domain context and could be derived automatically from any semantic description which the backend services might have. While this might happen directly, without any transformation or mapping, this approach would be impractical in practice. The reason for this is that each backend service can potentially (and very probably) express its semantics in different ontologies or vocabularies, which would render the pre- and post-condition mechanism in FAST useless. Therefore, we assume that a mapping or mediation layer will ensure that the different semantics at the backend service level will be mapped to a common set of terms within the FAST platform. This problem is the topic of deliverable 2.4 in FAST~\cite{ambrus2009fast_mediation}.
% subsection extension_towards_specific_domains (end)

% section the_fast_gadget_ontology (end)

\section{Conclusions and Outlook} % (fold)
\label{sec:conclusions}

The current version of deliverable 2.2 provides the foundation for enabling the semantic definition of gadgets within the context of the FAST platform. We have specified what the individual components of a gadget are, and how they interrelate. In developing our ontology, we have followed  well-known methodology for ontology development. During the course of this process, a number of living documents --- the requirements specification document, the glossary of terms, the integration document and finally the implementation document --- were produced. Those documents serve as a tool for development and as documentation at the same time, and will be extended and updated in the coming two years of the project and beyond. The driving force behind these changes will be testing and evaluation of the ontology within the FAST project, and possibly in other contexts as well.

A solution for representing pre- and post-conditions of screens and screen flows --- a central aspect of FAST technology --- has been proposed, which will now have to be implemented and tested both within the FAST catalogue and the FAST GVS.

Finally, to contribute to a lasting legacy of the FAST project, the gadget ontology will serve as input to standardisation efforts for the definition of semantic gadgets, which will go beyond the scope of the project itself.
% section conclusions (end)

\clearpage
\bibliographystyle{apalike}
\addcontentsline{toc}{section}{References}
\bibliography{fast_d2_2_bib}

\clearpage
\section*{Appendix A (Ontology Code)}
\addcontentsline{toc}{section}{Appendix A (Ontology Code)}

Below is the complete code of the FAST gadget ontology in Turtle syntax. The latest version of this file can always be accessed at \url{http://purl.oclc.org/fast/ontology/gadget}.

\singlespacing
\begin{small}
\verbatiminput{ontology/fast_gadget.owl}
\end{small}

\clearpage
\doublespacing
\section*{Appendix B (Lists of Tables and Figures)}
\addcontentsline{toc}{section}{Appendix B (Lists of Tables and Figures)}

\listoftables

\listoffigures

\end{document}

### 2009-12-11 trunk/deliverables/D2_2_Ontology/properties.tex
D2_2 and D2_4 added
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@385 d8b0c59b-6544-0410-94d9-83a5658a058e

\subsubsection{Property: \texttt{fast:contains}}
\label{subs:contains}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & contains \\ \hline
\textcolor{white}{\textbf{description}} & Many kinds of components in FAST can contain other components: 
    screenflows contain screens, screens contain forms or form elements, etc. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
\end{tabular}
\subsubsection{Property: \texttt{fast:hasPreCondition}}
\label{subs:hasPreCondition}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has pre-condition \\ \hline
\textcolor{white}{\textbf{description}} & This property links a screen or screenflow to its pre-condition, i.e., 
    the facts that need to be fulfilled in order for this screen or screenflow to be 
    reachable. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fast:WithConditions}}{subs:WithConditions} \\ \hline
\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fast:Condition}}{subs:Condition} \\ \hline
\end{tabular}
\subsubsection{Property: \texttt{fast:hasPostCondition}}
\label{subs:hasPostCondition}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has post-condition \\ \hline
\textcolor{white}{\textbf{description}} & This property links a screen or screenflow to its post-condition, 
    i.e., the facts that are produced once the screen or screenflow has been 
    executed. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fast:WithConditions}}{subs:WithConditions} \\ \hline
\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fast:Condition}}{subs:Condition} \\ \hline
\end{tabular}
\subsubsection{Property: \texttt{fast:hasIcon}}
\label{subs:hasIcon}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has icon \\ \hline
\textcolor{white}{\textbf{description}} & A small graphical representation of any FAST component or sub-component. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
\textcolor{white}{\textbf{range}} & \texttt{foaf:Image} \\ \hline
\end{tabular}
\subsubsection{Property: \texttt{fast:hasScreenshot}}
\label{subs:hasScreenshot}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has screenshot \\ \hline
\textcolor{white}{\textbf{description}} & An image which shows a particular screen or screenflow in action, to 
    aid users in deciding which screen or screenflow to choose out of many. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
\textcolor{white}{\textbf{range}} & \texttt{foaf:Image} \\ \hline
\end{tabular}
\subsubsection{Property: \texttt{fast:integratesTerm}}
\label{subs:integratesTerm}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & integratesTerm \\ \hline
\textcolor{white}{\textbf{description}} & A way to explicitly say that an ontology uses terms from another namespace. 
    Maybe a bit redundant, but why not. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
\end{tabular}
\subsubsection{Property: \texttt{fast:hasPattern}}
\label{subs:hasPattern}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & has pattern \\ \hline
\textcolor{white}{\textbf{description}} & This property links a screen or screenflow to its pre-condition, 
    i.e., the facts that need to be fulfilled in order for this screen or screenflow 
    to be reachable. \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fast:Condition}}{subs:Condition} \\ \hline
\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
\end{tabular}
\subsubsection{Property: \texttt{fast:isPositive}}
\label{subs:isPositive}
\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
\hline
\textcolor{white}{\textbf{label}} & is positive \\ \hline
\textcolor{white}{\textbf{description}} & Conditions can be positive or negative, depending on whether they 
    must be fulfilled or must not be fulfilled (in the case of pre-conditions), or 
    whether their facts will be added to the canvas or removed (in the case of 
    post-conditions). \\ \hline
\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fast:Condition}}{subs:Condition} \\ \hline
\textcolor{white}{\textbf{range}} & \texttt{xsd:boolean} \\ \hline
\end{tabular}

### 2009-12-11 trunk/deliverables/D2_4_Mediation/fast_d2_4_mediation.tex
D2_2 and D2_4 added
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@385 d8b0c59b-6544-0410-94d9-83a5658a058e

\documentclass{fast_latex}
% yay! We can use any kind of funky diacritic:
\usepackage[utf8]{inputenc}

\usepackage{lastpage}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage[pdfborder={0 0 0}]{hyperref}		% turn on when latex is used (not miktec)
\usepackage{url} % LEO: urls \url{}
\usepackage{verbatim} % code and comment
\usepackage{longtable}
\usepackage{xspace}   % whitespace after a macro if no punctuation after the macro
\usepackage{multirow}
\usepackage{colortbl}
\usepackage{longtable}
\usepackage{array}
\usepackage{textcomp} %vertical apostrophe
\usepackage{listings} %source code listings
\usepackage{amssymb} %math symbols
\usepackage{graphicx} %image placement
\usepackage{supertabular}
\parindent0pt

\newcommand\deliverableNumber{D2.4.1}
\newcommand\deliverableTitle{Mediation amongst ontologies: Application to the FAST ontology}
\newcommand\deliverableTitleShort{Mediation amongst ontologies}
\newcommand\workpackageNumber{2}
\newcommand\workpackageTitle{Definition of Conceptual Model}
\newcommand\authorOne{Oszkar Ambrus, NUIG}
\newcommand\authorTwo{Author 2}
\newcommand\authorThree{Author 3}
\newcommand\authorFour{Author 4}
\newtheorem{example}{\emph{Example}}

\begin{document}
% explicit hyphenations
\hyphenation{RDF-Re-po-si-to-ry}
\hyphenation{name-space}
\hyphenation{So-cket-A-dap-ter}

%\fontfamily{tahoma}\selectfont
\def\note#1{\marginpar{\footnotesize#1}} % use this to show the notes in the document
%\def\note#1{} % use this to hide the notes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TITLE PAGES 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\thispagestyle{empty}

\pagenumbering{roman}

\begin{flushright}
	\includegraphics[width=3cm]{images/FP7_logo.png}
\end{flushright}

\vspace{1cm}

%\begin{minipage}[p]{15cm}
	\begin{center}
		\includegraphics{images/FAST_logo}\\
		\vspace{1cm}
		{\LARGE{\sffamily \emph{FAST AND ADVANCED STORYBOARD TOOLS}}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{FP7-ICT-2007-1-216048}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{http://fast.morfeo-project.eu}}\\
		\vspace{4cm}
		{\LARGE \sffamily \textbf{Deliverable \deliverableNumber}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \textbf{\deliverableTitle}}\\
		\vspace{2cm}
		{\large \sffamily \authorOne}\\
		\vspace{0.5cm}
		\vfill
		{\large \sffamily Date: 27/02/2009}\\
		\vspace{1cm}
		{\sffamily FAST is partially funded by the E.C. (grant code: FP7-ICT-2007-1-216048).}

	\end{center}
%\end{minipage}

\clearpage
%%%%%%%%%%%%%%
% NEXT PAGES %
%%%%%%%%%%%%%%
\pagestyle{scrheadings}

\lohead{\includegraphics[width=4cm]{images/FAST_logo_transparent}}
%\cohead{\small\textcolor{fast@lightgrey}{\deliverableTitle}}
%\rohead{\small{\today}}
%\lofoot{\small\textcolor{fast@lightgrey}{Task Force Ontologies}}
\cofoot{\small{FAST 216048 --- \deliverableTitleShort}}
\rofoot{\small{\thepage}}

\newpage
%
%\section*{Authors}
%\authorOne \\
%%\authorTwo \\
%%\authorThree\\
%%\authorFour\\
%
%
%\vfill
%%\section*{Project Co-ordinator}
%%\hspace*{-2,5cm}\begin{minipage}[p]{14cm}
%%Dr. Ansgar Bernardi \\
%%German Research Center for Artificial Intelligence (DFKI) GmbH \\
%%Trippstadter Strasse 122 \\
%%D 67663 Kaiserslautern \\
%%Germany \\
%%Email: bernardi@dfki.uni-kl.de, phone: +49 631 205 3582, fax: +49 631 205 4910 \\
%%\end{minipage}
%
%
%\section*{Partners}
%Telef\'onica Investigaci\'on y Desarrollo (Telef\'onica I+D), Spain \\
%National University of Ireland, Galway (NUIG), Republic of Ireland \\
%Kassel University, Germany \\
%Cyntelix Corporation, Republic of Ireland \\
%SAP AG, Germany \\
%Universidad Polit\'ecnica de Madrid (UPM), Spain
%
%
%\vspace{0.3cm}
%\begin{footnotesize}
%Copyright: FAST Consortium 2009\\
%\end{footnotesize}

%\clearpage

\section*{Version History}

\begin{small}
\begin{tabular}{|l|l|l|p{7.5cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Rev. No.}} &
                            \textcolor{white}{\textbf{Date}} &
                            \textcolor{white}{\textbf{Author (Partner)}} &
							\textcolor{white}{\textbf{Change description}}\\ \hline
1.0 & 27.02.2009 & Oszkar Ambrus (NUIG) & final version ready for external review \\ \hline
\end{tabular}
\end{small}

\color{black}

\vfill
%{\bf Explanations of abbreviations on front page}\\
%\\
%%Nature \\
%R: Report \\
%P: Prototype \\
%R/P: Report and Prototype \\
%O: Other \\
% \\
%Dissemination level \\
%PU: Public \\
%PP: Restricted to other FP6 participants \\
%RE: Restricted to specified group \\
%CO: Confidential, only for NEPOMUK partners \\

\newpage

%%%%%%%%%%%%%%%%%%%%%
% Executive Summary %
%%%%%%%%%%%%%%%%%%%%%

\clearpage

\section*{Executive Summary}
\doublespacing
Ontology mediation is a critical problem in heterogeneous environments,
since inter-operating services and agents often use different
ontologies, and the differences between them need to be determined and
overcome to allow for a seamless data exchange between these parties.
This is a difficult ask, with which web-based agents must cope.

We review current work and literature, discussing the main areas of the
topic, that is the core problem of ontology mediation, the types of
mismatches that occur between ontologies, and the three main ontology
mediation approaches, namely (1) the representation of correspondences
between ontologies, (2) the (semi-)automatic process of discovering
these correspondences and (3) creation of a new ontology from the
source ontologies.

By the overview of current research in the field of ontology mediation,
we lay the basis of the implementation of ontology mediation in FAST.

\newpage

%%%%%%%%%%%%%%%%%%%%%
% Document Summary %
%%%%%%%%%%%%%%%%%%%%%

\clearpage

\section*{Document Summary}
% double spacing from here on:
\singlespacing
\begin{small}

\begin{tabular}
	%{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{6cm}|p{2cm}|p{2cm}|}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{6cm}|p{2cm}|p{2cm}|}
	\hline
	\textcolor{white}{\textbf{Code}} & {FP7-ICT-2007-1-216048} & {\textbf{Acronym}} & {FAST}\\ \hline
	\textcolor{white}{\textbf{Full title}} & \multicolumn{3}{l|}{Fast and Advanced Storyboard Tools}\\ \hline
	\textcolor{white}{\textbf{URL}} & \multicolumn{3}{l|}{\url{http://fast.morfeo-project.eu}}\\ \hline
	\textcolor{white}{\textbf{Project officer}} & \multicolumn{3}{l|}{Annalisa Bogliolo}\\ \hline
\end{tabular}

\vspace{0.5cm}

\begin{tabular}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.25cm}|p{1cm}|p{1cm}|p{6.32cm}|}
	\hline
	\textcolor{white}{\textbf{Deliverable}} & {\textbf{Number}} & {\deliverableNumber} & {\textbf{Name}} & {\deliverableTitle}\\ \hline
	\textcolor{white}{\textbf{Work package}} & {\textbf{Number}} & {\workpackageNumber} & {\textbf{Name}} & {\workpackageTitle}\\ \hline
\end{tabular}

\vspace{0.5cm}

\begin{tabular}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
	\hline
	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2009} & {\textbf{Submitted}} & {27/02/2009}\\ \hline
	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.4.1 is the first iteration of the FAST ontology mediation approach. This document contains a general overview of current literature regarding ontology mediation, paving the way for futher analysis of the scenarios arising in FAST that require ontology mediation. Based on this general overview, the implementation of ontology mediation within FAST can be researched, which will constitute the topic of the next iteration of this document.}\\ \hline
	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
%	{} & \multicolumn{4}{l|}{}\\ 
%	{} & \multicolumn{4}{l|}{}\\ 
%	{} & \multicolumn{4}{l|}{}\\
	\hline
	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {oszkar.ambrus@deri.org} \\ \cline{4-5}
	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495112} \\ \hline
	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, ontology mediation, ontology matching, ontology alignment, ontology merging}\\ \hline
\end{tabular}
\end{small}
\newpage

%%%%%%%%%%%%%%%%%%%%%
% TABLE OF CONTENTS %
%%%%%%%%%%%%%%%%%%%%%
\doublespacing
\setcounter{tocdepth}{3}
\tableofcontents
\cleardoublepage
\pagenumbering{arabic}

%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGINNING OF SECTIONS %
%%%%%%%%%%%%%%%%%%%%%%%%%
\clearpage
\rofoot{\small{Page \thepage\ of \pageref{LastPage}}}

\section{Introduction} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:introduction}
\paragraph{Goal and Scope.}
The central driver for Enterprise Mashups is the lightweight resource
composition style to build enterprise applications with reusable
building blocks, called Gadgets. The architecture used for this
front-end layer of Gadgets specifies a back-end layer of Services,
designed to support machine-to-machine interaction over a network (e.g.
Web Services, Databases, etc.). Having this case, the communication
problems need to be addressed, that arise from disparate interfaces,
different invocation mechanisms and even different conceptualisations
of the data transmitted between the front-end and the back-end. This
document deals with the third issue, namely the differences between the
conceptualisations of the exchanged data.

As in all fields employing semantic web paradigms, data is annotated
using ontologies. Ontologies conceptualise the data used, describe
their context and their semantic background that should be known to all
agents using that data. Because we do not the have the situation where
the same ontologies are used by all sides, having different ontologies
and even different languages to describe them, we need to integrate the
ontologies present in the back-end with the ones present in the
front-end, and to find a way to mediate between the the different
ontologies used in the two layers.

We try to find a solution to these differences which range from simple
naming differences to more complex data heterogeneity, behavioural
differences, discrepant paradigms and other disparities. The
reconciliation of these differences is called ontology mediation (often
mentioned in the specialised literature as ontology matching, or
ontology alignment, thus we will sometimes use these terms as well). 
\paragraph{Structure of the Document.}
We discuss ontologies in general in Section \ref{sec:ontologies}, present the ontology
mediation problem in Section \ref{sec:mediationproblem} , discuss the types of mismatches between
ontologies in Section \ref{sec:ontologymismatches} and detail the main approaches to ontology
mediation in Section \ref{sec:mediationapproaches}.
\paragraph{Overall Approach for WPX and relation to other Work Packages and Deliverables.}
Ontology mediation will play an important role for the integration of 
external ontologies into the FAST ontology infrastructure.
By giving a detailed overview of the state-of-the-art in ontology
mediation, we lay the groundwork for further research in T2.4 [Task
2.4, see Project Description] with this first iteration of D2.4. Future
versions of the deliverable (i.e., D2.4.2 and D2.4.3) will implement
ontology mediation within the FAST platform.

\section{Ontologies}
\label{sec:ontologies}
This section provides a brief introduction to the concept of ontologies.

\subsection{Definitions}
As specified in \cite{gruber2008ontology} {\textquotedblleft}an ontology defines a
set of representational primitives with which to model a domain of
knowledge or discourse{\textquotedblright}. These primitives are
typically classes, attributes and relations among classes or class
instances (and others, such as function terms, rules, restrictions,
axioms and events). Their definitions provide information about their
semantics (meaning) and constraints on their logically consistent
usage.

Ontologies are typically specified in languages that allow to abstract
away from implementation details, and focus on the so-called semantic
level. This way ontologies provide a representation independent from
data models and implementation, thus providing a means to enable
interoperability among disparate systems and specifying interfaces to
independent, knowledge-based systems. In the technology stack of the
Semantic Web, ontologies define a separate layer, and have a set of
languages and tools (both commercial and open-source) to work with
them.

We would also mention a widely-known and used classical definition of
ontologies, presented in \cite{gruber93towards}: {\textquotedblright}[An
ontology is] an explicit, formal specification of a shared
conceptualisation{\textquotedblright}. \cite{studer1998knowledge_engineering_principles} explains
this definition, by elaborating on the four main terms of the
definition:

\begin{enumerate}
\item it is a conceptualisation, because it models and categorises
relevant concepts from the real world, 
\item it is explicit, because the model explicitly states the types of
the concepts, the relationships between them and the constraints of
their use, 
\item it is formal, because the ontology has to be machine-readable and
consistent, and 
\item it is shared, because an ontology is consensual, i.e. accepted by
a group of people and it is used in a shared environment to allow a
common conceptualisation of data. 
\end{enumerate}

\subsection{General Description}
An ontology is used to define the vocabulary that agents use to exchange
queries and assertions. If an agent commits to a common ontology, it
guarantees the consistency (but not completeness) of the queries and
assertions using the vocabulary defined in the ontology \cite{gruber93towards}.
An agent supporting the interface defined in an ontology is not
required to use the terms of the ontology as an internal encoding of
its knowledge. Nonetheless, the definitions and formal constraints of
the ontology do restrict what can be meaningfully stated in the
language defined.

Two analogies to ontologies are given in \cite{gruber93towards}. Ontologies are
similar to global type declarations in a conventional software library,
and ontological commitments are similar to type restrictions over the
inputs and outputs of program modules. Formal argument restrictions can
be checked mechanically by compilers, to make sure that calling
procedures pass legal data to called procedures. Similarly, sentences
in an exchange between agents can be checked for logical consistency
using the definitions in the ontologies. Just as the formal argument
list hides the internal structure of a procedure from its calling
environment, a common ontology allows one to interact with a
knowledge-based program without committing to or being aware of its
internal encoding of knowledge.

Ontologies are also like conceptual schemata in database systems. Such a
schema provides a logical description of the data being shared,
allowing applications to interact with it without having any knowledge
of internal data structures. So while a conceptual schema defines
relations and constraints on data, an ontology defines terms with which
to represent knowledge, and constraints on their relationships.
\cite{gruber2008ontology} also points out that the requirement of inputs and
outputs to be logically consistent with the definitions and constraints
of an ontology is analogous to the requirement that the rows of a
database table (or insert statements in SQL) must be consistent with
the integrity constraints imposed on the table, stated separately and
independently of the internal data formats.

\cite{gruber2008ontology} enumerates the key applications of ontologies, by saying
that they {\textquotedblleft}are part of the W3C standards stack for
the Semantic Web, in which they are used to specify standard conceptual
vocabularies in which to exchange data among systems, provide services
for answering queries, publish reusable knowledge bases, and offer
services to facilitate interoperability across multiple, heterogeneous
systems and databases.{\textquotedblright} The key role of ontologies
in database systems is to provide a data modeling layer above specified
database designs, so that data can be used across independently
developed systems and services. This made it possible to achieve
database interoperability, cross database search and web service
interrogation.

There are many issues not addressed by common ontologies, related to
knowledge sharing. One important question is how a group of people can
reach a consensus on a common ontology. The discussion of ontology
mediation aims to address this problem by attempting a solution to the
case of several ontologies conceptualising a similar domain of the real
world in different ways.

\section{The Mediation Problem} %%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:mediationproblem}
Ontologies are employed as explicit descriptions of the information
source semantics. According to \cite{wache2001ontology_information_integration}, integrating these
semantics along multiple systems can be done using three main
approaches. (1) The single ontology approach, using a single ontology
as a global shared vocabulary. Unfortunately this can not provide
different views on a domain, and using it for multiple similar
catalogues is unfeasible due to the vast amount of information needed
to be captured, and the susceptibility of the conceptualisation to
change. (2) The multiple ontology approach has each information source
describing its own ontology. This solves problems related to the single
ontology approach, but introduces the problem of differences in
conceptualising the same concepts. Ontology mediation can be used to
map different parts of the ontologies to each other. (3) The hybrid
approach uses multiple ontologies, but uses a global ontology to make
all the ontologies comparable to each other. This is closely related to
ontology merging, where overlapping entities are merged, and the rest
is simply composed into one unified ontology, allowing for the
definition of the concepts of both ontologies. We suggest that the FAST
Gadget Ontology (as defined in D2.1) could act as a global ontology in
the context of FAST, although further research needs to be done in this
direction, since the Gadget Ontology has a slightly different
epistemological commitment.

In the context of FAST, the problem is that of reconciling the
ontologies in the back-end with the ontologies in the front-end, to
allow for interoperability between the two layers.

In the following we present a motivating scenario for ontology mediation
in Section \ref{subsec:mediationproblem_motivatingexample}, and a formal description of the alignment problem and
process, respectively, in Section \ref{subsec:mediationproblem_problemstatement}. After that we show some examples
of ontology mediation in Section \ref{subsec:mediationproblem_examples}.

\subsection{Motivating Example}
\label{subsec:mediationproblem_motivatingexample}
This section presents a brief example, illustrating a problem that
presents the need for ontology mediation.

%\begin{verbatim}
\lstset{frame=single, basicstyle=\ttfamily, caption=Ontology \texttt{O1}.}
\lstset{emph={vsmlVariant, namespace, ontology, concept, ofType, instance, nonFunctionalProperties, endNonFunctionalProperties, memberOf, hasValue}}
\lstset{emphstyle=\textbf\textit\underbar}
\begin{lstlisting}
vsmlVariant -"http://www.smo.org/wsml/wsml-syntax/wsml-flight"

namespace { -"http://see.deri.org/adrian/thesis/ontologies/O1#" ,
       wsml - "http://www.wsmo.org/wsml/wsml-syntax#" ,
         dc - "http://purl.org/dc/elements/1.1/" }

ontology - "http://see.deri.org/thesis/ontologies/O1"
    nonFunctionalProperties
    dc#description hasValue "A simple ontology modeling
                             the concept of a person"
    endNonFunctionalProperties

concept person
    name ofType _string
    age ofType _integer
    hasGender ofType gender
    hasChild ofType person
    marriedTo ofType person

concept gender
    value ofType _string

instance male memberOf gender
    value hasValue "male"
instance female memberOf gender
    value hasValue "female"
\end{lstlisting}

Consider the two ontologies \texttt{O1} and
\texttt{O2} \cite{mocan2008mediation} presented in Listing 1 and
Listing 2 respectively, in the Human Readable Syntax of the Web Service
Modeling Language (WSML) \cite{debruin2005wsml}. Elements from
ontology \texttt{O1} need to be mapped to the elements
described in Ontology \texttt{O2}.

Ontology \texttt{O1} describes the concept
\texttt{person} as one having five attributes, each of
them having a type that is either a concept or a value from a given
data type. The concept gender has two instances defined, that have
attributes pointing to the corresponding values.

\lstset{frame=single, caption=Ontology \texttt{O2}.}
%\lstset{emph={vsmlVariant, namespace, ontology, concept, ofType, instance, nonFunctionalProperties, endNonFunctionalProperties, memberOf, hasValue}}
%\lstset{emphstyle=\textbf}
\begin{lstlisting}
vsmlVariant -"http://www.smo.org/wsml/wsml-syntax/wsml-flight"

namespace { -"http://see.deri.org/adrian/thesis/ontologies/O2#" ,
       wsml - "http://www.wsmo.org/wsml/wsml-syntax#" ,
         dc - "http://purl.org/dc/elements/1.1/" }

ontology - "http://see.deri.org/thesis/ontologies/O2"
    nonFunctionalProperties
    dc#description hasValue "A simple ontology modeling
                             the concept of a human"
    endNonFunctionalProperties

concept human
    name ofType _string
    age ofType _integer
    noOfChildren ofType _integer

concept man subConceptOf human

concept woman subConceptOf human

concept marriage
    hasParticipant ofType human
    date ofType _date
\end{lstlisting}

%\newline
Ontology \texttt{O2} describes the concept
\texttt{human} as one having three attributes. The
concepts \texttt{man} and \texttt{woman} are
subclasses of the concept \texttt{human}.

Additionally a fourth concept called \texttt{marriage} is
defined having two attributes. (For simplicity we do not deal with the
exceptions of polygamy and same-sex marriage).

Suppose that a service (or agent) using ontology O1 needs to interact
with a service accepting and providing data that is semantically
defined in ontology O2.

The first step is to identify candidates to be mapped or to have
taxonomic relationships under an integrated schema \cite{shvaiko2005schema_based}. It is obvious that \texttt{person} and
\texttt{human} are candidates to be mapped, as well as the
class \texttt{man} with the class
\texttt{person} having the attribute
\texttt{gender} equal to
{\textquotedblleft}\texttt{male}{\textquotedblright} and
the class \texttt{woman} with the class
\texttt{person} having the attribute
\texttt{gender} equal to
{\textquotedblleft}\texttt{female}{\textquotedblright}.

After having identified these correspondences, the next step is to
generate query expressions and assertions that automatically translate
data instances of one of these ontologies into the other one, or that
automatically handle their semantic relations under an integrated
ontology.

\subsection{Problem Statement}
\label{subsec:mediationproblem_problemstatement}
Given two ontologies \texttt{o} and
\texttt{o{\textquotesingle}} that need to be mapped to
each other, according to \cite{shvaiko2005schema_based} we can define a
mapping element as a 5-tuple \texttt{{\textless}id, e,
e{\textquotesingle}, n, R{\textgreater}}, where

\begin{itemize}
\item \texttt{id} is a unique identifier, identifying the
mapping element, 
\item \texttt{e} and
\texttt{e{\textquotesingle}} entities (classes,
attributes, relations) of the first and second ontology, respectively, 
\item \texttt{n} is a confidence measure in some
mathematical structure holding the correspondence value between e and
e{\textquotesingle}, 
\item \texttt{R} is the correspondence relation holding
between e and e{\textquotesingle} (e.g.,
\texttt{equivalence (=); more general($\sqsupseteq$); disjointness($\perp$); overlapping($\sqcap$)}.
\end{itemize}

\begin{figure}
    \centering
        \includegraphics[width=72mm, height=45mm]{images/Alignment_process.png}%[bb=0 0 201 126]
        \caption{The alignment process}
    \label{fig:Alignment}
\end{figure}

A mapping is a set of the above defined mapping elements. The alignment
operation determines the mapping
(M\texttt{{\textquotesingle}}) for the pair of ontologies
(\texttt{o} and
\texttt{o{\textquotesingle}}). There are some parameters
than can extend the definition of the alignment process, namely:

\begin{enumerate}
\item an input mapping, \texttt{M}, which is to be used and
completed by the process, 
\item the mapping parameters, \texttt{p} (such as weights
and thresholds), that configure the alignment process, 
\item other external resources used by the alignment,
\texttt{r} (such as dictionaries, thesauri, etc.); See Figure \ref{fig:Alignment}. 
\end{enumerate}
For example, suppose that based on some algorithm (using linguistic and
structural analysis) the confidence measure for the equivalence
relation between the classes \texttt{person} in
\texttt{O1} and \texttt{human} in
\texttt{O2} is 0.62. Suppose that a mapping parameter
specifies a threshold of 0.55 for this algorithm, that is all pairs of
entities with a confidence measure higher that 0.55 will be considered
correct mapping elements. Thus our algorithms would return the
following mapping element: \texttt{{\textless} id11,
person, human, 0.62, = {\textgreater}.}Now, another algorithm
determines the two concepts to mean exactly the same thing, thus it
either does not compute a confidence measure (returning
\texttt{n/a}), or it returns the upper limit of the
confidence measure interval (e.g. 1). Thus the mapping element would
be: \texttt{{\textless} id11, person, human, n/a, =
{\textgreater}.}

It should be noted, that the entities mapped between two ontologies can
be constructed using patterns, i.e. more than just mapping concepts to
concepts, and so on. For example, we can map a concept to a concept
having a certain attribute, like in the case of the concept
\texttt{person} having the attribute
\texttt{hasGender} equal to \texttt{female}
from \texttt{O1} being mapped to the concept
\texttt{woman} from \texttt{O2}.

\subsection{Mediation Examples}
\label{subsec:mediationproblem_examples}
We present two further examples to better illustrate the use of ontology
mediation.

The first example is taken from \cite{giunchiglia2008evaluation}. Based on the
idea that we view ontologies as graph-like structures, we exemplify
ontology matching with the help of fragments of two tree-like
structures, such as Google and Looksmart. Notice that in the general
case the relation holding between nodes is not specialisation, but
classification (parent-child relation).

\begin{figure}[h]
    \centering
        \includegraphics[width=101mm, height=67mm]{images/F2GoogleLooksmart.png}%[bb=0 0 254 167]
        \caption{Fragments of Google and Looksmart directories and some
correspondences. Double arrowheads denote equivalence, while single
arrowheads stand for the more general relation. Image taken from
\cite{giunchiglia2008evaluation}.}
    \label{fig:GoogleLooksmart}
\end{figure}

Suppose we want to merge the merge the two structures. Such situations
can arise, for example when an e-commerce company acquires another one.
One probable mapping is given in Figure \ref{fig:GoogleLooksmart}. We have found, for example,
that \texttt{O1:Basketball} is equivalent to
\texttt{O2:Basketball}, and that
\texttt{O2:Games} is more general than
\texttt{O1:Board\_Games.}

The second example given in \cite{madhavan2001schema_matching} illustrates the
need to use several techniques in matching two ontologies. Figure \ref{fig:POSchemas}
shows two purchase order schemas that need to be matched.

\begin{figure}[h]
    \centering
        \includegraphics [width=101mm, height=42mm]{images/F1POSchemas.png}%[bb=0 0 257 105]
        \caption{Matching purchase order schemas. Image taken from \cite{madhavan2001schema_matching}.}
    \label{fig:POSchemas}
\end{figure}

As stated above in Section 3.2 we look for similarity coefficients
between the two schemas and deduce a mapping from those coefficients.
Suppose, as a first phase we use linguistic matching, using thesauri to
help match names, acronyms, synonyms, etc. resulting in a set of
candidate pairs (such as \texttt{PO} and
\texttt{PurchaseOrder}, or \texttt{Bill} and
\texttt{Invoice}) and language similarity coefficients
(\texttt{lsim}). In the second phase we would perform a
structural matching of the two schemas. For example,
\texttt{Line} would be mapped to
\texttt{ItemNumber}, because their parents and their other
two children match, \texttt{City} under
\texttt{POBillTo} will be mapped to
\texttt{City} under \texttt{InvoiceTo},
because bill is a synonym to invoice. This process results in
structural similarity coefficients (\texttt{ssim}). By
using a weighted similarity, we can compute the final mappings (e.g. by
using the formula \texttt{wsim=ws*wsim+(1-ws)*lsim}, where
\texttt{ws} is the interval [0,1]).

\section{Ontology Mismatches}
\label{sec:ontologymismatches}
Ontology mismatches are the main obstacles in the combined use of
independently developed ontologies. We will now give an overview of the
main type of ontology mismatches as presented in \cite{klein2001combiningOntologies}, which
is the main source of input to this state-of-the-art report. We will
classify ontology mismatches into given types and will find their
relationships.

Firstly, two levels of mismatches can be distinguished. The first level
is the \textit{language} or meta-model level. These are differences in
the primitives that make up the language used to specify the ontology.
They produce mismatches in the mechanics to define classes, relations,
attributes and so on. The second level is the \textit{ontology} or
model level. This is were the actual ontology is located. These are
differences in the way the domain is modeled.

In this section we discuss the overlap and mismatches between concepts,
relations, attributes and instances. We will detail the two levels of
mismatches and the different types that occur at each level.

\subsection{Language Level Mismatches}
\label{subsec:ontologymismatches_languagelevel}
We define four types of mismatches, although they often coincide. These
occur when ontologies written in different languages are combined. We
note that these type of mismatches will be less relevant for FAST
because we assume that standards for ontology languages are now so well
established that in practice we will not very often find ontologies
written in different languages anymore, but we discuss them briefly to
give a better overview of the meditation problem.

\subsubsection{Syntax}
Different languages obviously use different syntax. Therefore the same
concept can (and will be) often modeled in a completely different way.
For example, the concept of chairs would be defined in RDF Schema as
\texttt{{\textless}rdfs:Class
ID={\textquotedblright}Chair{\textquotedblright}{\textgreater}}, while
in LOOM it would be \texttt{(defconcept Chair)}.

This is the simplest possible mismatch, although it often comes coupled
with other mismatches as well. The simplest form is when a language has
several syntactical forms.

The solution is to simply to have a rewrite mechanism that returns the
corresponding translated form.

\subsubsection{Logical representation}
A slightly more complicated mismatch is when a logical notion can not be
represented in a given language. For example, in certain languages it
is possible to assert that two classes are disjoint, say using
\texttt{(disjoint A B),} whereas in other languages it is
necessary to express this using simpler statements, such as the
negation of set-inclusion statements, like \texttt{A
subclass-of (NOT B), B subclass-of (NOT A).}

The actual problem is not the question if something can be expressed,
but finding the language constructs to express a logical notion (not a
concept).

Solving this mismatch is relatively-easy: it consists in finding
translation rules to transform one representation into another.

\subsubsection{Semantics of primitives}
This difference is a more subtle one, and occurs at the meta-model of
the language. It sometimes occurs that the semantics of an identical
name differs from language to language. For example, the expression
\texttt{A equalsTo B} can have many different
interpretations.

Even when two ontologies use the same syntax, the semantics may differ.
For example, the OIL RDF syntax interprets several
\texttt{{\textless}rdfs:domain{\textgreater}} statements
as the intersection of the arguments, whereas RDF Schema interprets
them as the union of the arguments.

\subsubsection{Language expressiveness}
This mismatch, at the meta-model level of the languages, is the one
having the greatest impact on the attempt to combine and reconcile
ontologies.

The difference implies that some languages can express notions that
others can not. For example, some languages have no constructs to
express negation, whereas other do. Other typical examples are support
for sets, lists, etc.

These type of mismatches have the greatest impact, and are sometimes
mentioned as the {\textquotedblleft}fundamental
differences{\textquotedblright} between knowledge models.

\subsection{Ontology Level Mismatches}
\label{subsec:ontologymismatches_ontologylevel}
Mismatches at the model level happen both in the case of having the two
ontologies written in the same language and syntax and in the case of
different languages. We can distinguish between several types of
ontology level mismatches.

A conceptualisation mismatch is a difference in how the domain is
modeled (conceptualised), therefore we will have different concepts and
different relations between them. Explication mismatches are
differences in the way the concepts are specified. These occur as
mismatches in definitions, or terms, or the combination of the two.
Terminological mismatches are naming differences or conflicts. Finally,
encoding mismatches are differences in the way values are specified,
for example currencies for prices.

These kinds of mismatches are likely to occur in the context of FAST and
are therefore important factors in the implementation of ontology
mediation in FAST.

\subsubsection{Conceptualisation Mismatches}
These are differences in the conceptualisation of the domain, that is
not only the specification differs in the two ontologies, but the model
as well.

Sometimes these mismatches can not be reconciled automatically, and need
a domain expert. In this case, a good solution would be to use both
ontologies together. In that case, the overlapping part needs to be
aligned, and the rest can be merged.

\paragraph{Scope.}

Two classes that seemingly represent the same concept, but do not have
the same instances, although some of these are identical. For example
the concept of {\textquotedblleft}employee{\textquotedblright} can be
perceived and specified in different ways by different organisations.

\paragraph{Model coverage and granularity.}
This mismatch is related to the part of the domain that is covered by
the ontologies, and the level of detail respectively. For example, an
ontology could model cars, but no trucks, whereas another would have a
few categories for trucks, and another third ontology could make very
detailed distinction between all the types of trucks and their
properties.

\subsubsection{Explication Mismatches}
The first two types of explication mismatches (\textit{paradigm} and
\textit{concept description}) result from the \textbf{style of
modeling.} The next two types (\textit{synonym terms} and
\textit{homonym terms}) can be classified as \textbf{terminological
mismatches}. Finally, there is one trivial difference, cause by
\textit{encoding} mismatches.

\paragraph{Paradigm.}
Different paradigms can be used to represent different concepts, or
different top-level ontologies can be used, which all result in this
kind of mismatch. For example, one could model time based on interval
logic, while another could use discrete time points.

\paragraph{Concept Description.}
These types of mismatches are called modeling conventions. Several
choices can be made in different modeling questions. One choice, the
categorisation of concepts can be done based on an attribute, or using
subclasses. For example, as presented in Section 3.1, the concepts
{\textquotedblleft}man{\textquotedblright} and
{\textquotedblleft}woman{\textquotedblright} can be modeled in several
ways. They could be instances of the concept, and differ only in an
attribute (e.g. \texttt{gender}), or they can be separate
subclasses of a top-level class representing
{\textquotedblleft}humans{\textquotedblright} in general.

Another choice is the way in which a hierarchy is built - distinction
between features can be made higher or lower in the hierarchy. For
example, consider the place where the difference between scientific and
non-scientific publications is made: a dissertation can be modeled as
\texttt{dissertation }\texttt{{\textless}
book {\textless} scientific publication {\textless} publication}, or as
\texttt{dissertation {\textless} scientific book
{\textless} book {\textless} publication}, or as a subclass of both
\texttt{scientific publication} and
\texttt{book}.

\paragraph{Synonym Terms.}
Concepts are represented by different names. For example, the term
{\textquotedblleft}car{\textquotedblright} from an ontology can be
represented as {\textquotedblleft}automobile{\textquotedblright} in
another.

Although the solution looks simple in using thesauri, in reality much
human effort is needed to solve these problems, and care must be taken
not to overlap with scope mismatches (see above).

\paragraph{Homonym Terms.}
The meaning of a term is different in another context. For example, the
term {\textquotedblleft}conductor{\textquotedblright} has one meaning
in a musical context, and a different meaning in an electric
engineering domain.

These differences are much harder to solve than synonyms, and always
need human contribution.

\paragraph{Encoding.}
The values of attributes in ontologies are encoded in different formats.
For example, date can be encoded in a
\texttt{{\textquotedblleft}yyyy-mm-dd{\textquotedblright}}
form, or a
\texttt{{\textquotedblleft}dd/mm/yyyy{\textquotedblright}},
distance might be specified in miles or kilometers, money amount can be
given in different currencies, and so on.

These kind of mismatches can very easily be solved, by using a wrapper,
or a single transformation step.

\section{Ontology Mediation Approaches}
\label{sec:mediationapproaches}
In this section we give an overview of some of the major approaches to
ontology mediation, following the structure of Section 2: Approaches in
Ontology Mediation of \cite{debruin2005wsml}. Firstly, in Section
\ref{subsec:mediationapproaches_ontologymapping} we present some representative approaches to the specification of
mappings. Then in Section \ref{subsec:mediationapproaches_ontologyalignment} we overview some of the main approaches
and algorithms to overcome mismatches and find mappings between
ontologies, that is we discuss the topic of ontology alignment. Finally
in Section \ref{subsec:mediationapproaches_ontologymerging} we discuss the process of merging several ontologies.
Further research is needed to decide on the type of ontology mediation
to be used in the context of FAST, the algorithms and approaches
applied, and whether to use alignment or merging.

\subsection{Ontology Mapping}
\label{subsec:mediationapproaches_ontologymapping}
An ontology mapping is a declarative specification of the semantic
overlap between two ontologies \cite{debruin2005wsml}. It is the
result of the ontology alignment process (see Figure 3.). This mapping
is usually expressed as a set of axioms in a certain mapping language.

There are three main phases of the mapping process: (1) discovering the
mapping, (2) representing the mapping and (3)
using/exploiting/executing the mapping. This section is focused on (2),
that is we survey a number of approaches in representing ontology
mappings.

A common tendency in ontology mapping is to create an ontology of
mappings, which represents the vocabulary for the representation of
mappings.

\paragraph{MAFRA.}
MAFRA \cite{maedche2002mafra}, or MApping FRAmework for distributed
ontologies supports the interactive, incremental and dynamic process of
ontology mapping, which transforms instances of a source ontology into
instances of a target ontology.

\subparagraph{Framework Components.}
The framework consists of five main phases (called horizontal
dimensions) and for components that run along the entire process,
called vertical components. Within the horizontal dimension, the
authors identified five modules:

\begin{itemize}
\item \textit{Lift \& Normalisation}. Both ontologies must be normalised
to the uniform representation of RDFS. As a result both ontologies are
represented in RDF Schema with their instances in RDF. Thus syntactical
end lexical differences are eliminated and semantic differences become
more apparent. 
\item \textit{Similarity.} This module establishes similarities between
entities using different approaches (\textit{lexical similarity},
\textit{property similarity} based on properties of concepts,
\textit{bottom-up similarity} propagating the similarity from lower
parts of the taxonomy to the upper part and \textit{top-down
similarity} that assumes special relevance when top level concepts have
a higher or lower similarity). 
\item \textit{Semantic Bridging}. Based on the similarities computed in
the previous phase, correspondences are identified between entities,
using so-called semantic bridges, which define the actual mapping. 
\item \textit{Execution}. Transforms instances between the two
ontologies using the semantic bridges. 
\item \textit{Post-processing.} Revisiting the mapping for improvements.

\end{itemize}
The vertical dimension contains modules like evolution (synchronising
changes between in the ontologies and the bridges), the GUI (extensive
graphical support must be given for human assistance of the process),
etc.

\subparagraph{Semantic Bridges.}
Semantic bridges are captured in the Semantic Bridging Ontology (SBO).
Five dimensions have been identified for the semantic bridges, as
presented in the following:

\begin{itemize}
\item \textit{Entity dimension.} Semantic bridges may relate to the
following entities: (i) concepts modeling object classes from the real
world, (ii) relations modeling relationships of objects, (iii)
attributes modeling simple properties and (iv) extensional patterns
modeling the content of instances. 
\item \textit{Cardinality dimension.} Defines the number of entities on
both sides of the semantic bridge (mostly 1:1, 1:n and m:1; m:n is
rarely encountered, and can be usually decomposed onto m:1:n). 
\item \textit{Structural dimension.} Reflects on how elementary bridges
may be combined into more complex bridges. Relations that hold among
bridges are \textit{specialisation} (reuse a bridge and provide
additional information), \textit{abstraction} (should only be used as a
super-class), \textit{composition} (a bridge is composed of other
bridges) and \textit{alternatives} (mutually exclusive bridges). 
\item \textit{Constraint dimension.} Permits to control the execution of
the semantic bridge. Constraints must hold on the source ontology in
order for the transformation procedures to be applied. 
\item \textit{Transformation dimension.} Reflects on how instances of
the source ontology are transformed during the mapping process. 
\end{itemize}
The Semantic Bridging Ontology (SBO) contains a specification of all
semantic bridges, organised in a taxonomy. To actually relate the
source and the target ontology the mapping process creates an instance
of the SBO containing semantic bridge instances along with all the
necessary information to transform instances of entities from the
source to the target ontology.

The five semantic bridge dimensions are employed as follows:

\begin{itemize}
\item There are three basic entity types: \textit{Concepts},
\textit{Relations} and \textit{Attributes}. 
\item The class \textit{SemanticBridge} is the most generic bridge type.
It defines the relations to the source and target entities. It is
specialised according to entity type and cardinality of the relation. 
\item The class \textit{Service} can reference resources that are
responsible to connect to, or to describe transformations. 
\item \textit{Rule} is the general class for constraints and
transformation-relevant information. 
\item The class \textit{Transformation} uses the \textit{inService}
relation to link to the transformation procedure. 
\item The class \textit{Condition} represents the conditions that need
to be satisfied in order to execute the semantic bridge. 
\item The Composition primitive can be accomplished using the
\textit{hasBridge} relation of \textit{SemanticBridge}, without any
cardinality or type constraint. 
\item The Alternative primitive is supported by the
\textit{SemanticBridgeAlt} class. It groups several mutually exclusive
bridges, of which the first consistent bridge (with satisfied
conditions) is executed. 
\end{itemize}
For an extensive and relevant example please refer to Section 3.3. of
\cite{maedche2002mafra}.

\paragraph{RDFT.}
RDFT \cite{omelayenko2002rdft} is a small language (meta-ontology) defined for
mapping XML DTDs to/and RDF Schemas specially targeted for business
integration tasks. It is built on top of RDF Schema and it is used to
map RDF Schemas and concepts like events, messages, vocabularies to
XML-specific parts of the conceptual models that occur in the
integration tasks.

The model is derived from WSDL \cite{christensen2001wsdl} and provides
an RDF Schema for RDF annotations of WSDL documents, and is extended
with the temporal ontology PSL. The business integration task in this
context is seen as a service integration task, where each enterprise is
represented as a Web service specified in WSDL. WSDL annotations made
according to the defined meta-ontology allow performing inference over
WSDL descriptions to validate the links established between the
enterprises.

\subparagraph{PSL.}
To be able reason about the inputs and outputs of the companies being
integrated, the authors developed a conceptual model of WSDL, which is
directly derived from the WSDL specification. WSDL defines the
following basic elements of services, on top of which PSL is built:

\begin{itemize}
\item \texttt{Types} that provide links to the XML Schemas
of the messages exchanged; 
\item abstract definitions of \texttt{Messages} in
accordance with \texttt{Types}; 
\item \texttt{Port} \texttt{Types} that
specify input and output messages; 
\item \texttt{Bindings} that specify concrete protocols and
formats for messages according to \texttt{Types}. 
\end{itemize}
These elements can describe services but do not represent any temporal
knowledge about the messages requiring integration. The Process
Specification Language or PSL temporal ontology includes classes to
capture temporal knowledge:

\begin{itemize}
\item \texttt{activity} , which is performed during a
certain time interval; 
\item \texttt{activity} \texttt{occurrence} ,
contains the snapshot of an activity at a moment in time; 
\item \texttt{timepoint} , marks the time interval of the
activity; 
\item \texttt{objects} , things that do not possess
temporal properties, but may participate in certain activities at given
timepoints. 
\end{itemize}
\subparagraph{RDFT.}
The basic RDFT class is \texttt{Bridge }that connects two
concepts. It describes common properties of bridges, which enables one
to specify correspondences between an entity and a set of entities.
allowing for one-to-many and many-to-one connections.

The bridges contain the \texttt{Relation} property,
pointing to one of the subclasses of
\texttt{BridgeRelation}:

\begin{itemize}
\item \texttt{EquivalenceRelation}, stating that the source
element is equivalent with the target set of elements, or the source
set of elements is equivalent to the target element respectively,
depending on the relation cardinality; 
\item \texttt{VersionRelation}, specifying that the target
set of elements forms a later version of the source set of elements.
Unlike equivalence bridges, assumes that both concepts belong to the
same domain. 
\end{itemize}
Several types of bridges are defined in RDFT:

\begin{itemize}
\item \texttt{Event2Event} bridges link different events,
specify temporal event generation conditions, and link the events to
the messages transmitted with them. They connect instances of the
meta-class \texttt{mediator:Event}. 
\item two kinds of \texttt{RDFBridges}:
\texttt{Class2Class} and
\texttt{Property2Property} bridges between RDF Schema
classes and properties. They contain \texttt{rdfs:Class
}and \texttt{rdfs:Property} instances, respectively. 
\item four kinds of \texttt{XMLBridges}:
\texttt{Tag2Class}, \texttt{Tag2Property},
\texttt{Class2Tag}, \texttt{Class2Property},
which link XML DTD tags and RDF Schema classes and properties. 
\end{itemize}
The bridges are grouped into \texttt{Maps}, which are
collections of bridges serving a single purpose. The maps are
identified by their names and form minimal reusable modules of RDFT
bridges. Each map can include other maps and serves as a container for
\texttt{Bridges}. Described in this manner, as a set of
bridges, mappings are said to be \textit{declarative}, while
\textit{procedural} mappings can be defined as Xpath \cite{clark1999xpath} expressions, transforming instance data.

Connecting two services with RDFT consists of connecting their events
(instantiating \texttt{EventMap}) consisting of
\texttt{Event2Event} bridges. Each of the bridges points
to a \texttt{DocumentMap} aligning the documents attached
to the bridges and, in turn, consisting of
\texttt{RDFBridges}, \texttt{XMLBridges} and
\texttt{VocabularyMaps}.

\paragraph{C-OWL.}
C-OWL \cite{bouquet2004cowl} gives another perspective on ontology
alignment, extending the widely-known OWL \cite{dean2004owl}
language.

\subparagraph{The Vision.}
In their vision, similar to the one discussed at the beginning of
Section 3, two types of domain conceptualisations exist:

\begin{itemize}
\item \textit{ontologies}, that encode a view common to a set of
different parties. They make it easy to exchange information, although
they have the requirement of common consensus and the problem of hard
maintenance. 
\item \textit{contexts}, that encode the view of a party, and are local
conceptualisations, that are not shared. These are easy to maintain,
and no consensus is required, but need explicit mapping among the
contexts of different parties to be able to exchange information. 
\end{itemize}
In this vision, an ontology is contextualised, i.e. it is a
\textit{contextual} \textit{ontology}, when its contents are kept local
(therefore not shared) and can be mapped to the contents of other
ontologies via explicit mappings, allowing for a controlled form of
global visibility. This is opposed to the OWL importing mechanism,
where a set of local models is globalised in a unique shared model, by
importing complete models and using the imported elements by direct
reference.

\subparagraph{Context OWL.}
Context OWL or C-OWL is a language that extends OWL both syntactically
and semantically to support the concept of contextual ontologies.

Therefore a contextual ontology defined in C-OWL is a pair consisting
of:

\begin{enumerate}
\item an OWL ontology, and 
\item mapping between contexts, namely a set of bridge rules with the
same target ontology. 
\end{enumerate}
A C-OWL mapping therefore is a 4-tuple having the following form:

\begin{enumerate}
\item a mapping identifier (URI), 
\item a source context containing an OWL ontology (URI of the ontology),

\item a target context containing an OWL ontology (URI of the ontology),

\item a set of bridge rules form the local language of the source
ontology to the local language of the target ontology. Each mapping is
composed of three elements:

\begin{enumerate}
\item a source element (concept, role or individual) of the source
ontology; 
\item a target element, which must be of the same type as the source
element; 
\item the type of mapping (subsuming($\sqsupseteq$), equivalence(=), disjointness($\perp$) and
overlapping($\sqcap$)). 
\end{enumerate}
\end{enumerate}
Thus C-OWL has the complete representational power of OWL, augmented
with appropriate local model semantics for mapping between contexts
(local ontologies). A nice property of C-OWL is that the two components
(the OWL ontology and the mapping) are orthogonal, thus one can use the
ontology or the contextual component in an independent manner.

The local model semantics defined in C-OWL, as opposed to OWL, considers
that each context has a local set of models and a local domain of
interpretation. Thus, it is possible to have contradicting axioms or
unsatisfiable ontologies in the local models, without having the entire
context space unsatisfiable.

\subsection{Ontology Alignment}
\label{subsec:mediationapproaches_ontologyalignment}
\subsubsection{The Match Operator}
Ontology alignment is the process of discovering similarities between
two source ontologies \cite{debruin2005wsml}. This process can be
described as the application of the so-called \textit{Match} operator
described in \cite{rahm01survey}. The result of the matching
process is a specification of the similarities between the two
ontologies. The input of the Match operator consists mainly of two
ontologies and some other inputs (initial mapping that will be
extended, mapping parameters that configure the process and external
resources used by the matching process; see Section 3.2).

The generic implementation of the Match operator should contain a
uniform internal representation of the schemas to be matched, in order
to significantly reduce complexity. This, of course, requires a
semantics-preserving schema importer and exporter, respectively.

In general it is not possible to fully automate the matching between two
ontologies, primarily because they often have semantics that affects
the matching criteria, but is not formally expressed or often even
documented. Therefore the implementation of Match should determine
match candidates, which the user can accept, reject, or change.
Furthermore the user should be able to specify matches for elements for
which the system was unable to find satisfactory match candidates.

\subsubsection{The Alignment Process}
The alignment process proposed in \cite{debruin2005wsml} relieves the user of
some of the burdens in creating the mappings. The input of the process
consists of two ontologies which are to be aligned. The output is a set
of mappings.

The steps of the alignment are as follows:

\begin{enumerate}
\item \textit{Feature engineering}. Selects only parts of an ontology
definition in order to describe a specific entity. A feature may be as
simple as a label or it may include super- and sub-concepts, relations
or extensional descriptions.\newline
For example, suppose a fragment of an ontology describes the instance
\texttt{Daimler}. In this case we should consider the
generic ontology feature called type which has the values
\texttt{luxury} and \texttt{automobile},
respectively. 
\item \textit{Selection of Next Search Steps}. This step chooses the
next set of candidate pairs. It may choose to compute a restricted
subset of candidate concept pairs of the two ontologies and ignore the
others. 
\item \textit{Similarity Assessment.} Determines the similarity degree
of candidate pairs. Heuristics are used, that is similarity functions
such as on strings, object sets, related concepts, and so on.\newline
For example, one similarity function can check whether the parents of
the compared two concepts are identical. 
\item \textit{Similarity Aggregation}. For a candidate pair we can (and
often do) use several similarity functions. The results of these
functions need to be aggregated to compute the final similarity measure
between the two entities. This can be done by either a simple
averaging, or a complex aggregation function using weighting schemes. 
\item \textit{Interpretation}. This step uses the aggregated similarity
values to align entities. It can use thresholds for similarity, perform
relaxation labelling, or combine structural and similarity
criteria.\newline
For example, suppose \texttt{simil(o1:Person,
o2:Human)=0.62 ${\geq}$ 0.5}, thus it is added to the result as a
mapping element. 
\item \textit{Iteration}. Several algorithms perform more than one
iteration in order to bootstrap the amount of structural knowledge.
Iteration may stop when no new alignments are proposed, or when a given
number of mappings have been discovered. Note that certain steps can be
omitted in later iterations, since we can use previously computed
values (such as selecting features, computing a given similarity,
etc.). 
\end{enumerate}
The output of the process is a set of correspondences between entities
of the two ontologies. We cannot expect all correspondences to be
discovered, therefore the results of the alignment process can be
considered as the input to a manual refinement process.

\subsubsection{Classification of Approaches}
An implementation of Match may use multiple matching algorithms or
\textit{matchers. }This allows us to select matchers according to the
application domain or schema types.

Given that we want two use multiple matchers we distinguish two
subproblems. First, there is the need to realise individual matchers,
each of which uses a single matching criterion to find mappings.
Second, we want to combine different individual matchers, either by
using multiple matching criteria within an integrated \textit{hybrid}
\textit{matcher}, or by combining multiple match results provided by
different match algorithms within a \textit{composite}
\textit{matcher}.

For individual matchers, we consider the following mostly-orthogonal
classification criteria (of which the first one is the most important
distinctive criterion):

\begin{itemize}
\item \textit{instance-based/schema-based} matching: A schema-based
matcher uses different approaches to determine correspondence between
ontologies based on concepts and relations \cite{debruin2005wsml}. 
An instance-based matcher takes instances
(i.e. instance data) belonging to the different concepts in the
ontologies and uses these to discover similarities between the concept.

\item \textit{element-level/structure-level} matching: an element-level
matcher takes individual schema elements, properties of the particular
concept or relation, and uses these to find similarities, whereas a
structure level matcher compares the combination of elements, the
structure (e.g. the concept hierarchy) of the ontologies to find
similarities. 
\item \textit{language/constraint}: a matcher can use a linguistic
approach (e.g. based on names and textual description of schema
elements) or a constraint-based approach (e.g. based on keys and
relationships). 
\item \textit{matching cardinality:} an overall matching result can
relate one or more elements of an ontology to one or more elements of
the other, giving four cases: 1:1, 1:n, m:1 and n:m. In addition each
mapping element can may interrelate one ore more elements of the given
ontologies. Furthermore, there may be different cardinalities at the
instance level. 
\item \textit{auxiliary information}: most matchers rely on external
information besides the two given ontologies, such as global schemas,
dictionaries, previous matching decisions and user input. 
\item \textit{accepted input:} this dimension concerns the type of input
on which the algorithms operate, classified depending on the
data/conceptual models they accept. 
\item \textit{produced output}: this pertains to the way the mapping is
described, whether it is a one-to-one correspondence or a more general
one, whether it is a graded answer (98\% or 4/5), or an all-or-nothing
approach, whether the result focuses only on equivalence between
entities, or can it give a more expressive result, and so on. 
\end{itemize}

\paragraph{Schema-based matching.}
A schema-based matcher \cite{rahm01survey} takes different aspects
of the concepts and relations in ontologies and uses some similarity
measure to determine correspondence. The available information includes
the usual properties of ontology elements, such as name, description,
data types, relationship types (part-of, is-a, etc.), constraints and
ontology structures. In the following we discuss some of the main
aspects of schema-based matching.

\begin{itemize}
\item \textit{Granularity of match}. We distinguish two main
alternatives, element-level and structure-level matching. For each
element of the fist ontology, an \textit{element-level matcher}
determines the matching element of the second ontology at the same
granularity level (usually at the atomic level). For instance, in two
ontology fragments describing addresses we could find that
\texttt{{\textquotedblleft}Address.ZIP}{\textquotedblright}
from one ontology matches
{\textquotedblleft}\texttt{CustomerAddress.PostalCode{\textquotedblright}}
from the other one. \textit{Structure level matching}on the other hand
refers to matching combinations of elements that appear together in a
structure. This can be a complete, or a partial match. Known
equivalence patterns can be kept in a library. For example (this
conforms to a common equivalence pattern), one ontology could have a
class \texttt{ParttimeEmployee}, which is the subclass of
\texttt{Employee}, while another could have a class
\texttt{Employee} having the attribute
\texttt{IsParttime}. A structure-level matcher should
identify these as being identical, when the
\texttt{IsParttime} property equals to true. 
\end{itemize}
\begin{itemize}
\item \textit{Match cardinality}. An element can participate in zero,
one or many mapping elements of a match result. Moreover, within a
mapping element, one or more elements from one ontology can match one
or more elements from the other one. Thus we have the usual
relationship cardinalities of 1:1, n:1, 1:m and n:m between matching
elements, both with respect to different mapping elements
(\textit{global cardinality}) and with respect to an individual mapping
element (\textit{local cardinality}).\newline
Most existing matchers map each element of one schema to one element of
the other schema. This results in local 1:1 mappings and global 1:1 or
1:n mappings. 
\end{itemize}
\begin{itemize}
\item \textit{Linguistic approaches}. Linguistic matchers use name and
text (i.e. words or sentences) to find semantically similar elements.
\textit{Name matching} matches elements with identical or similar names
(using synonyms, common substrings, user-provided name matching, etc.).
\textit{Description matching} uses natural language comments of
ontology elements. For example we can have
{\textquotedblleft}\texttt{empn //employee
name}{\textquotedblright} in an ontology matching
{\textquotedblleft}\texttt{name //name of
employee}{\textquotedblright} in the other. This can be done by simple
keyword extraction, or complex natural language processing. 
\end{itemize}
\begin{itemize}
\item \textit{Constraint-based approaches.} These approaches use the
fact that ontologies often contain constraints on data types and
ranges, relationships, optionality, uniqueness, cardinalities, etc.
This information can be used to determine similarities between ontology
elements. For example, in a candidate pair having the fields
\texttt{Born} and \texttt{BirthDate}
respectively, both having the type \texttt{Date}, a
constraint-based matcher could find these to be similar. 
\end{itemize}
\begin{itemize}
\item \textit{Reusing schema and mapping information}. In addition to
already mentioned auxiliary information, matchers could well reuse
common ontology components and previously determined mappings. Some
mappings occur very often in the same domains, and reusing these
increases effectiveness. For example, in e-commerce domains
substructures often repeat with different message formats (like
address, name, etc.), so we could keep a library of previously found
results on these, so when we find two elements labelled
\texttt{ProductOrder} and \texttt{Product} in
an ontology \texttt{O1}, and \texttt{Porder}
and \texttt{Article} in an ontology
\texttt{O2}, the matcher would already know that these
match (\texttt{ProductOrder} matches
\texttt{POrder} and \texttt{Product} matches
\texttt{Article}). 
\end{itemize}

\paragraph{Instance-based matching.}
An instance-based matcher \cite{rahm01survey} takes instances
(i.e. instance data) belonging to the different concepts in the
ontologies and uses these to discover similarities between the
concepts. It can be useful when the data is semistructured or
particularly useful to uncover incorrect interpretations of schema
information.

The approaches we discuss for instance-level matching primarily work in
finding element-level matches, because of the complexity of comparing
large numbers of combinations of instances.

Most of the approaches of schema-level matchers can be applied here, but
some are especially applicable:

\begin{itemize}
\item For text elements, a linguistic approach based on information
retrieval techniques is the preferred approach. For example, if we have
a field called \texttt{Dept} in an ontology, and two
fields \texttt{DeptName} and \texttt{EmpName}
in another, instance information may help to deduce that
\texttt{DeptName} is the primary match candidate for
\texttt{Dept}. 
\item For more structured data, such as numerical or string elements, we
can apply a constraint-based characterisation, such as numerical value
ranges or character patterns. For instance, this may help recognising
phone numbers, postal codes, birth dates, ISBNs, money-related entries
(e.g. based on currency symbols), etc. 
\end{itemize}
The results of instance-based matchers can be used to enhance
schema-level matchers as a first approach. For instance, a
constraint-based schema-level matcher can more accurately determine
data types and ranges for an element using the results of
instance-based matching. A second approach is to perform instance-level
matching on its own. Matching elements from the first ontology to the
second one, and then vice-versa, we can obtain schema-level matches.
Instance-level matching can also be performed by using auxiliary
information. This is especially helpful for matching text elements by
providing match candidates for individual keywords. For example, a
previous analysis may have revealed that the keyword
\texttt{{\textquotedblleft}HP{\textquotedblright}}
frequently occurs for ontology elements
\texttt{{\textquotedblleft}CompanyName{\textquotedblright}},
\texttt{{\textquotedblleft}Manufacturer{\textquotedblright}},
etc. For a new match task, if an \texttt{O2} ontology
element \texttt{X} frequently contains the term
\texttt{{\textquotedblleft}HP{\textquotedblright}}, this
can be used to generate
\texttt{{\textquotedblleft}CompanyName{\textquotedblright}}
in \texttt{O1} as a match candidate for
\texttt{X}, even if
\texttt{{\textquotedblleft}HP{\textquotedblright}} does
not often occur in instances of \texttt{O1}.

\subsubsection{Overview of Approaches}
In this section we give a brief overview of some of the main approaches
to ontology alignment.

\paragraph{Anchor-PROMPT.}
Anchor-PROMPT \cite{noy2000anchor_prompt} uses a set of heuristics to analyse
non-local context, as opposed to algorithms like PROMPT \cite{noy2000prompt} 
(See Section \ref{subsubsec:mediationapproaches_ontologymerging_PROMPT}) and Chimaera, that only analyse local context.
It doesn{\textquotesingle}t provide a complete solution, but augments
existing methods.

The algorithms takes as input two pairs of related terms (anchors) from
the source ontologies. These anchors are either given by the user, or
generated automatically with string-based techniques or another matcher
using linguistic similarity.

From this set of anchors, the algorithm produces a set of new pairs of
semantically close terms. To do that, Anchor-PROMPT traverses the paths
between the anchors in the corresponding ontologies. A path follows the
links between the classes, defined by the hierarchical relations or by
slots and their domain and range. Anchor-PROMPT then compares these
terms along the paths to find similar terms. Terms with a high
similarity score are presented to the user to improve the set of
possible suggestions, for example, a merging process in PROMPT.

For example, suppose we have ontology \texttt{O1} having
the classes \texttt{A-C-E-H} in a hierarchical chain, and
\texttt{O2} having the classes
\texttt{B-D-F-G} in its hierarchy. Then, suppose we
identify the anchors as \texttt{A-B} and
\texttt{G-H}. We then traverse each path in parallel,
incrementing the similarity score of the classes we encounter at each
step. We repeat the the process for all existing paths between the
anchor points, cumulatively aggregating the similarity score.

The central observation behind Anchor-PROMPT is that if two pairs of
terms are similar and there are paths connecting the terms, the
elements along those paths are often similar as well.

\paragraph{GLUE.}
GLUE \cite{doan2004ontology_matching} applies machine learning techniques to
semi-automatically create semantic mappings between heterogeneous
ontologies based on instance data. It regards ontologies as being
taxonomies of concepts, and focuses on finding 1-to-1 correspondences
between the concepts of the ontologies: for each concept node in one
taxonomy, find the most similar concept node from the other taxonomy.

Instead of committing to a certain similarity definition, it computes
the \textit{joint probability distribution} of the concepts involved,
and lets the application use the joint distribution to compute any
suitable similarity measure. Specifically, for any two concepts
\texttt{A} and \texttt{B} the joint
probability distribution consists of \texttt{P(A,B),
P(A,{\textlnot}B), P({\textlnot}A,{\textlnot}B)}, and
\texttt{P({\textlnot}A,B),} where a term such as
\texttt{P(A,{\textlnot}B)}, is the probability that an
instance in the domain belongs to concept \texttt{A}, but
not to concept \texttt{B}.

Computing this joint distribution is based on the sets of instances that
overlap between the two concepts. A term such as
\texttt{P(A,B)} can be approximated as the fraction of
instances that belong to both A and B, that is we need to decide for
each instance whether it belongs to \texttt{A${\Pi}$B}.
GLUE addresses this by using the instances of \texttt{A}
to learn a classifier for \texttt{A}. Then classifies
instances of \texttt{B} according to that classifier, and
vice-versa.

\paragraph{Semantic Matching.}
Semantic Matching \cite{giunchiglia2004semantic_matching} is an implementation of
the Match operator, that takes two graph-like structures (like database
schemas or ontologies) and produces a mapping between elements of the
two graphs that correspond semantically to each other. The authors
argue that most previous approaches used syntax driven techniques, like
matching labels(substrings, abbreviations, similar soundex) or
syntactical structure and calculating a similarity measure, all being
different variations of syntactic matching, that searches for semantic
correspondences based on syntactic features.

Semantic matching has the following main features:

\begin{itemize}
\item searching is done by mapping meanings (concepts), not labels. It
is not sufficient to consider the meaning of labels in the nodes, but
also the positions the nodes have in the graph. 
\item semantic similarity relations are used between elements (concepts)
instead of syntactical similarity relations. In particular, those
relations are considered, which relate the extensions of the concepts
under consideration. 
\end{itemize}
The semantics of a node is given by the concept attached to that node
(the concept denoted by the label of the node), by the position of the
given node in the graph, and the semantics of all the nodes which are
higher in the hierarchy.

The possible returned relations between elements are equality,
overlapping, mismatch, and more general/specific.

\paragraph{QOM.}
QOM \cite{ehrig2004qom} was designed to create an efficient matching
tool for on-the-fly creation of mappings between ontologies.

Given that the a major ingredient of run-time complexity is the number
of mapping pairs which have to be compared to actually find the best
mappings, QOM uses heuristics to lower the number of candidate
mappings, thus not comparing all entities of the first ontology to all
entities of the second ontology. It uses the ontological structures to
classify candidate mappings into promising and less promising pairs,
using a dynamic programming approach to refine the set of candidate
mappings in each iteration.

The similarity measure is computed using a wide range of similarity
functions, such as string similarity. Several of such similarity
measures are computed, which are all input to the similarity
aggregation function, which computes the actual similarity measure. QOM
then applies a sigmoid function, which emphasises high individual
similarities and deemphasises low individual similarities. The actual
correspondences are then extracted using a threshold to the aggregated
similarity measure.

The output of one iteration of QOM can be used as part of the input of a
subsequent iteration in order to refine the result. After a number of
iterations the actual mapping between the ontologies is obtained. 
\subsubsection{Alignment Examples}
This section provides a few examples of the matching
process{\textquotesingle} results.
\begin{center}
\tablecaption{Constraint-based matching example. {\newline}Taken from \cite{rahm01survey}.}
\label{tab:ConstraintMatching}
\tablehead{}
%begin{supertabular}{|r@{\hspace{6.5mm}}|r@{\hspace{5.5mm}}|r|r|}
\begin{supertabular}{|m{0.4\textwidth}|m{0.4\textwidth}|}
\hline
S1 elements &
S2 elements\\\hline
Employee

\ \ \ \ EmpNo -- int, primary key

\ \ \ \ EmpName -- varchar (50)

\ \ \ \ DeptNo -- int, references

\ \ \ \ Salary -- dec (15,2)

Department

\ \ \ \ Birthdate -- date

\ \ \ \ DeptNo -- int, primary key

\ \ \ \ DeptName -- varchar (40) &
Personnel

\ \ \ \ Pno - int, unique

\ \ \ \ Pname -- string

\ \ \ \ Dept -- string

\ \ \ \ Born -- date\\\hline
\end{supertabular}
\end{center}
The two schemas (corresponding to ontologies) shown in Table \ref{tab:ConstraintMatching} present
an example of constraint-based schema-level matching \cite{rahm01survey}. Type and key information suggest that
\texttt{Born} matches \texttt{Birthdate} and
\texttt{Pno} matches either \texttt{EmpNo} or
\texttt{DeptNo} (this could be refined using
instance-based matching, which could reveal that
\texttt{Pno} matches \texttt{EmpNo}). By
using a matcher that detects more than just atomic-level similarities,
we match \texttt{S2.Personnel }to
\texttt{S1.Employee} joined with
\texttt{S1.Department.} This can be detected automatically
by observing that elements of \texttt{S2.Personnel} match
elements of \texttt{S1.Employee} and
\texttt{S1.Department} and that
\texttt{S1.Employee} and
\texttt{S1.Departmen}t are connected by the foreign key
\texttt{DepNo}. This allows us to determine the correct
SQL-like n:m mapping:
\begin{verbatim}
    S2.Personnel (Pno, Pname, Dept, born) ~
    SELECT S1.Employee.EmpNo,
           S1.Employee.EmpName,
           S1.Department.DeptName,
           S1.Employee.Birthdate
    FROM S1.Employee, S1.Department
    WHERE (S1.Employee.DeptNo = S1.Department.DeptNo)
\end{verbatim}

Table \ref{tab:CardinMatching} \cite{rahm01survey} illustrates match cardinalities for
two given schemas S1 and S2. When matching multiple elements, we can
see that expressions are used to combine them.

\begin{center}
\tablecaption{Matching cardinalities example. {\newline}Taken from \cite{rahm01survey}.}
\label{tab:CardinMatching}
\tablehead{}
\begin{supertabular}{|p{0.20\textwidth}|p{0.20\textwidth}|p{0.20\textwidth}|p{0.20\textwidth}|}
\hline
\textbf{Local match cardinalities} & \textbf{S1 element(s)} & \textbf{S2 element(s)} & \textbf{Matching expression} \\\hline 
1:1, element level & Price & Amount & Amount = Price \\\hline 
n:1, element-level & Price, Tax & Cost & Cost ={\newline}Price*(1+Tax/100) \\\hline 
1:n, element-level & Name & FirstName,\newline
LastName & FirstName,{\newline}LastName =\newline
Extract (Name, \ldots) \\\hline 
n:1 structure-level
(n:m element-level) & B.Title,\newline
B.PuNo,\newline
P.PuNo,\newline
P.Name & A.Book,
A.Publisher & A.Book, A.Publisher =\newline
SELECT B.Title, P.Name\newline
FROM B, P\newline
WHERE B.PuNo=P.PuNo
\\\hline
\end{supertabular}
\end{center}
In the first row, the
match is 1:1, which is the most simple to determine. Row 2 shows how
\texttt{Cost} can be matched to a formula based on
\texttt{Price} and \texttt{Tax}. Row 3
explains how \texttt{FirstName} and
\texttt{LastName} are extracted from
\texttt{Name}. Row 4 uses a SQL expression combining
attributes from two tables. It corresponds to an n:m relationship at
the attribute level (four \texttt{S1} attributes
correspond to two \texttt{S2} attributes) and to an n:1
relationship at the structure level (two tables from
\texttt{S1} match one table from
\texttt{S2}).

The example in Table \ref{tab:MappingPattern} \cite{debruin2005wsml} shows a mapping element from the
motivating example from Section \ref{subsec:mediationproblem_motivatingexample}. Assume we have two ontologies
\texttt{O1} and \texttt{O2} which both
describe humans and their gender. Ontology \texttt{O1}
describes the concept \texttt{Person} with an attribute
\texttt{hasGender}, which has the two possible values
{\textquotedblleft}\texttt{male}{\textquotedblright} and
{\textquotedblleft}\texttt{female}{\textquotedblright}.
Ontology \texttt{O2} describes the concept
\texttt{Human}, and its subclasses
\texttt{Man} and \texttt{Woman} to
distinguish the gender.

\begin{center}
\tablecaption{Class by attribute mapping pattern and example. {\newline}Taken from \cite{debruin2005wsml}.}
\label{tab:MappingPattern}
\tablehead{}
\begin{supertabular}{|m{0.8\linewidth}|}
\hline
\textbf{Name}: Class by Attribute Mapping \\\hline
\textbf{Problem}: The extension of a class in one ontology
corresponds to the extension of a class in another ontology, provided
that all individuals in the extension have a particular attribute
value. \\\hline
\textbf{Solution:}

\textit{Solution description}: A mapping is established
between a class/attribute/attribute value combination in one ontology
and a class in another ontology.

\textit{Mapping syntax:}

\texttt{mapping ::= classMapping(direction A
B attributeValueCondition(P o))} \\\hline
\textbf{Example:}

\texttt{classMapping(Human Female
attributeValueCondition(hasGender
}{\textquotedblleft}female{\textquotedblright})) \\\hline
\end{supertabular}
\end{center}

Notice that this is a typical case of a
conceptual mismatch, namely a mismatch in the style of modeling. The
solution illustrates an elementary mapping pattern. The pattern is
described in terms of its name,the problem addressed, the solution of
the problem, both in natural-language description and in terms of the
actual mapping language, and an example, namely a mapping between a
class \texttt{Human} from ontology
\texttt{O1} to the class \texttt{Woman} from
ontology \texttt{O2}, but only for humans having the
gender
{\textquotedblleft}\texttt{female}{\textquotedblright}.

\subsection{Ontology Merging}
\label{subsec:mediationapproaches_ontologymerging}
Ontology merging is the creation of a new ontology from one or more
source ontologies. The new ontology will unify and in general replace
the original source ontologies.

We distinguish between two main approaches to ontology merging. In the
first approach the input of the process is a collection of ontologies
and the output is a new, merged ontology which captures the original
ontologies. A prominent example of this is PROMPT \cite{noy2000prompt}. 
In the second approach the ontologies are not replaced, but
rather a {\textquotesingle}view{\textquotesingle}, called
\textit{bridge ontology} is created which imports the original
ontologies and specifies the correspondences using bridge axioms.
OntoMerge \cite{dou2002ontology_translation} is an example of this approach, creating a
bridge ontology that imports the original ontologies and relates the
overlapping concepts in these ontologies using bridge axioms. We
describe PROMPT and OntoMerge in more detail below.

\subsubsection{PROMPT}
\label{subsubsec:mediationapproaches_ontologymerging_PROMPT}
PROMPT \cite{noy2000prompt} is an algorithm and interactive tool for
ontology merging and alignment. It was one of the first ontology
merging tools, comparing every pair in the two ontologies and looking
for syntactic similarity between labels. It has been implemented as a
Proteg\'e-2000 extension.

The algorithm of PROMPT defines a number of steps for the interactive
merging process:

\begin{enumerate}
\item An initial list of merging candidates is created, based on class
name similarities. The list is presented to the user as potential
merging operations. Then the following cycle happens: 
\item The user chooses the next operation: selects an option suggested
by the algorithm, or edits the ontology to specify the desired
operation directly. 
\item The third step is composed of the following three sub-steps:

\begin{enumerate}
\item The system performs the requested action and automatically
executes additional changes derived from the operation. 
\item A new list of suggested operations is presented to the user, based
on the structure of the ontology around the arguments of the last
operation. 
\item The system determines the conflicts introduced by the last
operation and finds possible solutions to them. 
\end{enumerate}
\end{enumerate}
PROMPT defines the following set of ontology-merging operations: merge
classes, merge slots, merge bindings between a slot and a class,
perform a deep copy of the class (copy all parents up to the root),
perform a shallow copy of the class.

The conflicts that may appear as a result of these operations are: name
conflicts, dangling references, redundancy in the class hierarchy,
slot-value restrictions that violate class inheritance.

The result of the algorithm is a new ontology which replaces the
original ones.

\subsubsection{OntoMerge}
OntoMerge \cite{dou2002ontology_translation} is an online ontology translation system,
which translates a dataset to a new dataset which captures the same
information in a different ontology.

The approach divides the process into two main parts, namely separating
the syntactic translation from the semantic translation:

\begin{enumerate}
\item Converting the datasets into a uniform internal representation (a
language called Web-PDDL) in order to clear away syntactic differences.

\item Perform the semantic translation from the internal representation
of a dataset in the source ontology to the internal representation of
the a dataset in the target ontology. 
\end{enumerate}
The result of the process is not an ontology replacing the original ones
as in PROMPT, but a bridge ontology that imports the source ontologies
and contains a set of Bridging Axioms, which are translation rules used
to connect the overlapping parts of the source ontologies. The two
source ontologies, together with the set of bridging axioms, are then
treated as a single theory by a theorem prover (optimised for
operations like dataset translation, ontology extension generation
based on the extension of a related ontology, and query rewriting).

\section{Summary}
\label{sec:summary}
With the Semantic Web winning more and more ground both in research and
industry, the mismatches between ontologies used by the inter-operating
systems present an increasing problem. Ontology mediation deals with
identifying these mismatches, and overcoming them.

After presenting the main problem, and the types of mismatches between
ontologies, we have discussed - based on the most prominent research
results in these areas - the three main approaches to ontology
mediation, that is \textit{ontology mapping} (representing
correspondences between entities), \textit{ontology alignment} (the
process of finding the correspondences between ontologies) and
\textit{ontology merging }(the process of creating a new ontology based
on the source ontologies that need to be reconciled).

We will test the various kinds of ontology mediation on FAST, choose a
suitable candidate, adopt it to our needs and implement it in the
platform. The choice of mediation approach will be influenced by an
evaluation of the kinds of ontologies that are used in real-world web
services.

\clearpage
\bibliographystyle{apalike}
\addcontentsline{toc}{section}{References}
\bibliography{fast_d2_4_mediation_bib}

\clearpage
\doublespacing
\section*{Appendix A (Lists of Tables and Figures)}
\addcontentsline{toc}{section}{Appendix A (Lists of Tables and Figures)}

\listoftables

\listoffigures

\end{document}

### 2009-12-14 trunk/deliverables/D2_2_Ontology/classes.tex
latest ontology changes integrated
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@389 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
-\subsubsection{Class: \texttt{fast:Resource}}
+\subsubsection{Class: \texttt{fgo:Resource}}
 \label{subs:Resource}
@@
     anything that can be `touched' and moved around in the FAST IDE. \\ \hline
-\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fast:Screen\_Flow}}{subs:ScreenFlow}, \htmlref{\texttt{fast:Screen}}{subs:Screen}, \htmlref{\texttt{fast:Backend\_Service}}{subs:BackendService}, \htmlref{\texttt{fast:Flow\_Control\_Element}}{subs:FlowControlElement}, \htmlref{\texttt{fast:Operator}}{subs:Operator}, \htmlref{\texttt{fast:Form\_Element}}{subs:FormElement}, \htmlref{\texttt{fast:Condition}}{subs:Condition}, \htmlref{\texttt{fast:WithConditions}}{subs:WithConditions} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fast:contains}}{subs:contains}, \htmlref{\texttt{fast:hasIcon}}{subs:hasIcon}, \htmlref{\texttt{fast:hasScreenshot}}{subs:hasScreenshot} \\ \hline
-\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fast:contains}}{subs:contains} \\ \hline
+\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fgo:ScreenFlow}}{subs:ScreenFlow}, \htmlref{\texttt{fgo:Screen}}{subs:Screen}, \htmlref{\texttt{fgo:FlowControlElement}}{subs:FlowControlElement}, \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent}, \htmlref{\texttt{fgo:Condition}}{subs:Condition}, \htmlref{\texttt{fgo:Fact}}{subs:Fact}, \htmlref{\texttt{fgo:WithPreConditions}}{subs:WithPreConditions}, \htmlref{\texttt{fgo:WithPostConditions}}{subs:WithPostConditions}, \htmlref{\texttt{fgo:WithConditions}}{subs:WithConditions}, \htmlref{\texttt{fgo:Precondition}}{subs:Precondition}, \htmlref{\texttt{fgo:Postcondition}}{subs:Postcondition}, \htmlref{\texttt{fgo:Pipe}}{subs:Pipe}, \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:contains}}{subs:contains}, \htmlref{\texttt{fgo:hasIcon}}{subs:hasIcon}, \htmlref{\texttt{fgo:hasScreenshot}}{subs:hasScreenshot}, \htmlref{\texttt{fgo:hasTag}}{subs:hasTag}, \htmlref{\texttt{fgo:hasVersion}}{subs:hasVersion}, \htmlref{\texttt{fgo:hasId}}{subs:hasId}, \htmlref{\texttt{fgo:hasName}}{subs:hasName}, \htmlref{\texttt{fgo:hasType}}{subs:hasType}, \htmlref{\texttt{fgo:hasDefinition}}{subs:hasDefinition} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fast:Screen\_Flow}}
+\subsubsection{Class: \texttt{fgo:ScreenFlow}}
 \label{subs:ScreenFlow}
@@
 \textcolor{white}{\textbf{description}} & The complete gadget, a set of screens. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fast:Screen}}
+\subsubsection{Class: \texttt{fgo:Screen}}
 \label{subs:Screen}
@@
 \textcolor{white}{\textbf{description}} & An individual screen. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasCode}}{subs:hasCode} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fast:Backend\_Service}}
+\subsubsection{Class: \texttt{fgo:FlowControlElement}}
+\label{subs:FlowControlElement}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Flow Control Element \\ \hline
+\textcolor{white}{\textbf{description}} & Any kind of component which can restrict the default flow of screens 
+    in a gadget. \\ \hline
+\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fgo:ScreenFlowStart}}{subs:ScreenFlowStart}, \htmlref{\texttt{fgo:ScreenFlowEnd}}{subs:ScreenFlowEnd}, \htmlref{\texttt{fgo:Connector}}{subs:Connector} \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:ScreenFlowStart}}
+\label{subs:ScreenFlowStart}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Screen Flow Start \\ \hline
+\textcolor{white}{\textbf{description}} & The entry point to a wigdet; the first screen. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:FlowControlElement}}{subs:FlowControlElement} \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:ScreenFlowEnd}}
+\label{subs:ScreenFlowEnd}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Screen Flow End \\ \hline
+\textcolor{white}{\textbf{description}} & A screen that ends the workflow of the gadget. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:FlowControlElement}}{subs:FlowControlElement} \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:Connector}}
+\label{subs:Connector}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Connector \\ \hline
+\textcolor{white}{\textbf{description}} & An explicit connection between two screens. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:FlowControlElement}}{subs:FlowControlElement} \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:ScreenComponent}}
+\label{subs:ScreenComponent}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Screen Component \\ \hline
+\textcolor{white}{\textbf{description}} & A screen component is any resource which is part of a particular screen. \\ \hline
+\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fgo:Operator}}{subs:Operator}, \htmlref{\texttt{fgo:FormElement}}{subs:FormElement}, \htmlref{\texttt{fgo:BackendService}}{subs:BackendService} \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasAction}}{subs:hasAction}, \htmlref{\texttt{fgo:hasTrigger}}{subs:hasTrigger}, \htmlref{\texttt{fgo:hasLibrary}}{subs:hasLibrary} \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:Operator}}
+\label{subs:Operator}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Operator \\ \hline
+\textcolor{white}{\textbf{description}} & Any kind of component that is used to connect backend services to 
+    form elements. Examples are simple pipes, aggregators or various kinds of 
+    filters. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:FormElement}}
+\label{subs:FormElement}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Form Element \\ \hline
+\textcolor{white}{\textbf{description}} & Form elements are UI elements in a particular screen. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:BackendService}}
 \label{subs:BackendService}
@@
     wrapped by the screen. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fast:Flow\_Control\_Element}}
-\label{subs:FlowControlElement}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Flow Control Element \\ \hline
-\textcolor{white}{\textbf{description}} & Any kind of component which can restrict the default flow of screens 
-    in a gadget. \\ \hline
-\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fast:Screen\_Flow\_Start}}{subs:ScreenFlowStart}, \htmlref{\texttt{fast:Screen\_Flow\_End}}{subs:ScreenFlowEnd}, \htmlref{\texttt{fast:Connector}}{subs:Connector} \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
-\end{tabular}
-\subsubsection{Class: \texttt{fast:Screen\_Flow\_Start}}
-\label{subs:ScreenFlowStart}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Screen Flow Start \\ \hline
-\textcolor{white}{\textbf{description}} & The entry point to a wigdet; the first screen. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Flow\_Control\_Element}}{subs:FlowControlElement} \\ \hline
-\end{tabular}
-\subsubsection{Class: \texttt{fast:Screen\_Flow\_End}}
-\label{subs:ScreenFlowEnd}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Screen Flow End \\ \hline
-\textcolor{white}{\textbf{description}} & A screen that ends the workflow of the gadget. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Flow\_Control\_Element}}{subs:FlowControlElement} \\ \hline
-\end{tabular}
-\subsubsection{Class: \texttt{fast:Connector}}
-\label{subs:Connector}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Connector \\ \hline
-\textcolor{white}{\textbf{description}} & An explicit connection between two screens. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Flow\_Control\_Element}}{subs:FlowControlElement} \\ \hline
-\end{tabular}
-\subsubsection{Class: \texttt{fast:Operator}}
-\label{subs:Operator}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Operator \\ \hline
-\textcolor{white}{\textbf{description}} & Any kind of component that is used to connect backend services to 
-    form elements. Examples are simple pipes, aggregators or various kinds of 
-    filters. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
-\end{tabular}
-\subsubsection{Class: \texttt{fast:Form\_Element}}
-\label{subs:FormElement}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Form Element \\ \hline
-\textcolor{white}{\textbf{description}} & Form elements are UI elements in a particular screen. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
-\end{tabular}
-\subsubsection{Class: \texttt{fast:Condition}}
+\subsubsection{Class: \texttt{fgo:Condition}}
 \label{subs:Condition}
@@
     or may also ignore them. E.g., in EzWeb pre- and post-conditions correspond to 
-    the concepts of slot and event. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fast:hasPattern}}{subs:hasPattern}, \htmlref{\texttt{fast:isPositive}}{subs:isPositive} \\ \hline
-\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fast:hasPreCondition}}{subs:hasPreCondition}, \htmlref{\texttt{fast:hasPostCondition}}{subs:hasPostCondition} \\ \hline
+    the concepts of slot and event.
+	A condition can be seen as a RDF bag of facts, where every fact has to be true
+	for the condition be true as well. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasPreCondition}}{subs:hasPreCondition}, \htmlref{\texttt{fgo:hasPostCondition}}{subs:hasPostCondition}, \htmlref{\texttt{fgo:hasCondition}}{subs:hasCondition} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fast:WithConditions}}
+\subsubsection{Class: \texttt{fgo:Fact}}
+\label{subs:Fact}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Fact \\ \hline
+\textcolor{white}{\textbf{description}} & A fact is the atomic formal representation of a part of a condition.
+	Therefore, several facts compose a condition. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasPattern}}{subs:hasPattern}, \htmlref{\texttt{fgo:hasPatternString}}{subs:hasPatternString}, \htmlref{\texttt{fgo:isPositive}}{subs:isPositive} \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:WithPreConditions}}
+\label{subs:WithPreConditions}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & With-precondition \\ \hline
+\textcolor{white}{\textbf{description}} & Those kinds of resource which can have pre-conditions 
+    (i.e., screens and screen flows). \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasPreCondition}}{subs:hasPreCondition} \\ \hline
+\textcolor{white}{\textbf{unionOf}} & \htmlref{\texttt{fgo:ScreenFlow}}{subs:ScreenFlow}, \htmlref{\texttt{fgo:Screen}}{subs:Screen}, \htmlref{\texttt{fgo:Action}}{subs:Action} \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:WithPostConditions}}
+\label{subs:WithPostConditions}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & With-precondition \\ \hline
+\textcolor{white}{\textbf{description}} & Those kinds of resource which can have pre-conditions 
+    (i.e., screens and screen flows). \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasPostCondition}}{subs:hasPostCondition} \\ \hline
+\textcolor{white}{\textbf{unionOf}} & \htmlref{\texttt{fgo:ScreenFlow}}{subs:ScreenFlow}, \htmlref{\texttt{fgo:Screen}}{subs:Screen}, \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:WithConditions}}
 \label{subs:WithConditions}
@@
 \textcolor{white}{\textbf{label}} & With-condition \\ \hline
-\textcolor{white}{\textbf{description}} & Those kinds of resource which can have pre- or post-conditions 
-    (i.e., screens and screen flows). \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fast:hasPreCondition}}{subs:hasPreCondition}, \htmlref{\texttt{fast:hasPostCondition}}{subs:hasPostCondition} \\ \hline
-\textcolor{white}{\textbf{unionOf}} & \htmlref{\texttt{fast:Screen}}{subs:Screen}, \htmlref{\texttt{fast:Screen\_Flow}}{subs:ScreenFlow} \\ \hline
+\textcolor{white}{\textbf{description}} & Those kinds of resource which can have both pre- or post-conditions. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{unionOf}} & \htmlref{\texttt{fgo:WithPreConditions}}{subs:WithPreConditions}, \htmlref{\texttt{fgo:WithPostConditions}}{subs:WithPostConditions} \\ \hline
 \end{tabular}
+\subsubsection{Class: \texttt{fgo:Action}}
+\label{subs:Action}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Action \\ \hline
+\textcolor{white}{\textbf{description}} & An action represents a specific routine which will be performed when a certain
+	condition is fulfilled within a certain screen component (i.e., the action `showTable' will
+	be performed when data from a service is received). \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasUse}}{subs:hasUse} \\ \hline
+\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasAction}}{subs:hasAction} \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:Library}}
+\label{subs:Library}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Action \\ \hline
+\textcolor{white}{\textbf{description}} & An action represents a specific routine which will be performed when a certain
+	condition is fulfilled within a certain screen component (i.e., the action `showTable' will
+	be performed when data from a service is received). \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasLanguage}}{subs:hasLanguage}, \htmlref{\texttt{fgo:hasSource}}{subs:hasSource} \\ \hline
+\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasLibrary}}{subs:hasLibrary} \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:Precondition}}
+\label{subs:Precondition}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Precondition \\ \hline
+\textcolor{white}{\textbf{description}} & A precondition is a satisfied condition within a screenflow. It can be seen
+    as an input of the screenflow. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasCondition}}{subs:hasCondition} \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:Postcondition}}
+\label{subs:Postcondition}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Postcondition \\ \hline
+\textcolor{white}{\textbf{description}} & A postcondition is a result condition within a screenflow. It can be seen
+    as an output of the screenflow. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasCondition}}{subs:hasCondition} \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:Definition}}
+\label{subs:Definition}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Resource definition \\ \hline
+\textcolor{white}{\textbf{description}} & Structural and behaviour definition of a resource. \\ \hline
+\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fgo:ScreenDefinition}}{subs:ScreenDefinition} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:contains}}{subs:contains} \\ \hline
+\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasDefinition}}{subs:hasDefinition} \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:ResourceReference}}
+\label{subs:ResourceReference}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Resource reference \\ \hline
+\textcolor{white}{\textbf{description}} & It`s a reference to a certain resource. Needs a 'id' for internal
+	identification for the resource which is referencing it, and the `uri' of the
+	resource. \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasId}}{subs:hasId}, \htmlref{\texttt{fgo:hasUri}}{subs:hasUri} \\ \hline
+\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:contains}}{subs:contains}, \htmlref{\texttt{fgo:hasUse}}{subs:hasUse} \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:ScreenDefinition}}
+\label{subs:ScreenDefinition}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Screen definition \\ \hline
+\textcolor{white}{\textbf{description}} & Behaviour definition of a screen. This will contain which form, operators and 
+	backend services the screen is composed, and how they are connected. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Definition}}{subs:Definition} \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:Pipe}}
+\label{subs:Pipe}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & pipe or connector \\ \hline
+\textcolor{white}{\textbf{description}} & Define a pipe or connector between two resources. The connection is made
+	specifying the conditions which will be connected. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasIdBBFrom}}{subs:hasIdBBFrom}, \htmlref{\texttt{fgo:hasIdConditionFrom}}{subs:hasIdConditionFrom}, \htmlref{\texttt{fgo:hasIdBBTo}}{subs:hasIdBBTo}, \htmlref{\texttt{fgo:hasIdConditionTo}}{subs:hasIdConditionTo}, \htmlref{\texttt{fgo:hasIdActionTo}}{subs:hasIdActionTo} \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:Trigger}}
+\label{subs:Trigger}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Trigger \\ \hline
+\textcolor{white}{\textbf{description}} & Define a... \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasIdBBFrom}}{subs:hasIdBBFrom}, \htmlref{\texttt{fgo:hasIdBBTo}}{subs:hasIdBBTo}, \htmlref{\texttt{fgo:hasIdActionTo}}{subs:hasIdActionTo}, \htmlref{\texttt{fgo:hasNameFrom}}{subs:hasNameFrom} \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:FormDefinition}}
+\label{subs:FormDefinition}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Form Definition \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:OperatorDefinition}}
+\label{subs:OperatorDefinition}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Operator Definition \\ \hline
+\end{tabular}
+\subsubsection{Class: \texttt{fgo:BackendServiceDefinition}}
+\label{subs:BackendServiceDefinition}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Backend Service Definition \\ \hline
+\end{tabular}

### 2009-12-14 trunk/deliverables/D2_2_Ontology/properties.tex
latest ontology changes integrated
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@389 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
-\subsubsection{Property: \texttt{fast:contains}}
+\subsubsection{Property: \texttt{fgo:contains}}
 \label{subs:contains}
@@
 \textcolor{white}{\textbf{description}} & Many kinds of components in FAST can contain other components: 
-    screenflows contain screens, screens contain forms or form elements, etc. \\ \hline
+    screenflows contain screens, screens contain forms or form elements, etc.
+    It points to a resource reference to give a unique `id' to each component
+    within the resource. \\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource}, \htmlref{\texttt{fgo:Definition}}{subs:Definition} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:ResourceReference}}{subs:ResourceReference} \\ \hline
 \end{tabular}
-\subsubsection{Property: \texttt{fast:hasPreCondition}}
+\subsubsection{Property: \texttt{fgo:hasPreCondition}}
 \label{subs:hasPreCondition}
@@
 \textcolor{white}{\textbf{label}} & has pre-condition \\ \hline
-\textcolor{white}{\textbf{description}} & This property links a screen or screenflow to its pre-condition, i.e., 
+\textcolor{white}{\textbf{description}} & This property links certain type of resources to its pre-condition, i.e., 
     the facts that need to be fulfilled in order for this screen or screenflow to be 
@@
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fast:WithConditions}}{subs:WithConditions} \\ \hline
-\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fast:Condition}}{subs:Condition} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:WithPreConditions}}{subs:WithPreConditions} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition}, \texttt{rdf:Bag} \\ \hline
 \end{tabular}
-\subsubsection{Property: \texttt{fast:hasPostCondition}}
+\subsubsection{Property: \texttt{fgo:hasPostCondition}}
 \label{subs:hasPostCondition}
@@
 \textcolor{white}{\textbf{label}} & has post-condition \\ \hline
-\textcolor{white}{\textbf{description}} & This property links a screen or screenflow to its post-condition, 
+\textcolor{white}{\textbf{description}} & This property links certain type of resources to its post-condition, 
     i.e., the facts that are produced once the screen or screenflow has been 
@@
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fast:WithConditions}}{subs:WithConditions} \\ \hline
-\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fast:Condition}}{subs:Condition} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:WithPostConditions}}{subs:WithPostConditions} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition}, \texttt{rdf:Bag} \\ \hline
 \end{tabular}
-\subsubsection{Property: \texttt{fast:hasIcon}}
+\subsubsection{Property: \texttt{fgo:hasPattern}}
+\label{subs:hasPattern}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has pattern \\ \hline
+\textcolor{white}{\textbf{description}} & This property links a screen or screenflow to its pre-condition, 
+    i.e., the facts that need to be fulfilled in order for this screen or screenflow 
+    to be reachable. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Fact}}{subs:Fact} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{rdfs:Resource} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasIcon}}
 \label{subs:hasIcon}
@@
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
 \textcolor{white}{\textbf{range}} & \texttt{foaf:Image} \\ \hline
 \end{tabular}
-\subsubsection{Property: \texttt{fast:hasScreenshot}}
+\subsubsection{Property: \texttt{fgo:hasScreenshot}}
 \label{subs:hasScreenshot}
@@
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
 \textcolor{white}{\textbf{range}} & \texttt{foaf:Image} \\ \hline
 \end{tabular}
-\subsubsection{Property: \texttt{fast:integratesTerm}}
+\subsubsection{Property: \texttt{fgo:hasCode}}
+\label{subs:hasCode}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has code \\ \hline
+\textcolor{white}{\textbf{description}} & URL of the executable code of a particular screen. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Screen}}{subs:Screen} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{foaf:Document} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasCondition}}
+\label{subs:hasCondition}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has condition \\ \hline
+\textcolor{white}{\textbf{description}} & This property links a precondition or postcondition to a certain condition. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Precondition}}{subs:Precondition}, \htmlref{\texttt{fgo:Postcondition}}{subs:Postcondition} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition}, \texttt{rdf:Bag} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasAction}}
+\label{subs:hasAction}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has action \\ \hline
+\textcolor{white}{\textbf{description}} & This property indicates which actions are asociated and can be perfomed within 
+	a screen component. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Action}}{subs:Action} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasUse}}
+\label{subs:hasUse}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has use string \\ \hline
+\textcolor{white}{\textbf{description}} & This property indicates concepts used within a building block, without being a 
+	pre/postcondition. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Action}}{subs:Action} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:ResourceReference}}{subs:ResourceReference} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasLibrary}}
+\label{subs:hasLibrary}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has library \\ \hline
+\textcolor{white}{\textbf{description}} & This property indicates which libraries are used by a screen component. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Library}}{subs:Library} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasSource}}
+\label{subs:hasSource}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has source \\ \hline
+\textcolor{white}{\textbf{description}} & URL of the source code of a particular library. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Library}}{subs:Library} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{foaf:Document} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasUri}}
+\label{subs:hasUri}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has uri \\ \hline
+\textcolor{white}{\textbf{description}} & URI of a particular resource pointing from a reference within
+	another resource. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:ResourceReference}}{subs:ResourceReference} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{foaf:Document} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasDefinition}}
+\label{subs:hasDefinition}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has definition \\ \hline
+\textcolor{white}{\textbf{description}} & The structure and behaviour of a resource. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Definition}}{subs:Definition} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:integratesTerm}}
 \label{subs:integratesTerm}
@@
 \end{tabular}
-\subsubsection{Property: \texttt{fast:hasPattern}}
-\label{subs:hasPattern}
+\subsubsection{Property: \texttt{fgo:hasPatternString}}
+\label{subs:hasPatternString}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & has pattern \\ \hline
-\textcolor{white}{\textbf{description}} & This property links a screen or screenflow to its pre-condition, 
-    i.e., the facts that need to be fulfilled in order for this screen or screenflow 
-    to be reachable. \\ \hline
+\textcolor{white}{\textbf{label}} & has pattern string \\ \hline
+\textcolor{white}{\textbf{description}} & This property is the textual representation of a condition. \\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fast:Condition}}{subs:Condition} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Fact}}{subs:Fact} \\ \hline
 \textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
 \end{tabular}
-\subsubsection{Property: \texttt{fast:isPositive}}
+\subsubsection{Property: \texttt{fgo:isPositive}}
 \label{subs:isPositive}
@@
 \textcolor{white}{\textbf{label}} & is positive \\ \hline
-\textcolor{white}{\textbf{description}} & Conditions can be positive or negative, depending on whether they 
-    must be fulfilled or must not be fulfilled (in the case of pre-conditions), or 
-    whether their facts will be added to the canvas or removed (in the case of 
-    post-conditions). \\ \hline
+\textcolor{white}{\textbf{description}} & Facts can be set to a specific scope: design time, execution time,
+	or both of them. This will define when they have to be taken into account by the
+	inference engine or reasoner. \\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fast:Condition}}{subs:Condition} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Fact}}{subs:Fact} \\ \hline
 \textcolor{white}{\textbf{range}} & \texttt{xsd:boolean} \\ \hline
 \end{tabular}
+\subsubsection{Property: \texttt{fgo:hasTag}}
+\label{subs:hasTag}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has tag \\ \hline
+\textcolor{white}{\textbf{description}} & A tag is used to annotate keywords of a particular resource. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasVersion}}
+\label{subs:hasVersion}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has version \\ \hline
+\textcolor{white}{\textbf{description}} & The version of a particular screen or screenflow, such as 1.0, 0.2beta, etc. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasTrigger}}
+\label{subs:hasTrigger}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has trigger string \\ \hline
+\textcolor{white}{\textbf{description}} & This property represents the events fired within a building block. e.g. a 
+	button to clear the shopping cart will fired a "clear cart" trigger. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasLanguage}}
+\label{subs:hasLanguage}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has language string \\ \hline
+\textcolor{white}{\textbf{description}} & This property is the language the library is written in. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Library}}{subs:Library} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasId}}
+\label{subs:hasId}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has id \\ \hline
+\textcolor{white}{\textbf{description}} & The internal ID of the resource within the container resource. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource}, \htmlref{\texttt{fgo:ResourceReference}}{subs:ResourceReference} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasName}}
+\label{subs:hasName}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has name \\ \hline
+\textcolor{white}{\textbf{description}} & The name the user gives to the resource. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasType}}
+\label{subs:hasType}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has type \\ \hline
+\textcolor{white}{\textbf{description}} & The type of the resource. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasIdBBFrom}}
+\label{subs:hasIdBBFrom}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has id building block from \\ \hline
+\textcolor{white}{\textbf{description}} & building block id of the `from' point of the pipe or trigger. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe}, \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasIdConditionFrom}}
+\label{subs:hasIdConditionFrom}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has id condition from \\ \hline
+\textcolor{white}{\textbf{description}} & condition id of the `from' point of the pipe. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasIdBBTo}}
+\label{subs:hasIdBBTo}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has id building block to \\ \hline
+\textcolor{white}{\textbf{description}} & building block id of the `to' point of the pipe or trigger. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe}, \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasIdConditionTo}}
+\label{subs:hasIdConditionTo}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has id condition to \\ \hline
+\textcolor{white}{\textbf{description}} & condition id of the `to' point of the pipe. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasIdActionTo}}
+\label{subs:hasIdActionTo}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has id action to \\ \hline
+\textcolor{white}{\textbf{description}} & action id of the `to' point of the pipe or trigger. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe}, \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection{Property: \texttt{fgo:hasNameFrom}}
+\label{subs:hasNameFrom}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has name from \\ \hline
+\textcolor{white}{\textbf{description}} & name of the trigger. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}

### 2010-01-14 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_mediation.tex
First initial working draft of D2.4.2
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@419 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2009-12-11 trunk/deliverables/D2_4_Mediation/fast_d2_4_mediation.tex
@@
 \usepackage{xspace}   % whitespace after a macro if no punctuation after the macro
+\usepackage{lipsum}
 \usepackage{multirow}
@@
 \usepackage{supertabular}
+
+\usepackage{amsmath}
+\usepackage{pgf} %graph plot
+\usepackage{tikz}
+\usepackage{paralist}
+\usetikzlibrary{plotmarks}
+
 \parindent0pt
 
-\newcommand\deliverableNumber{D2.4.1}
-\newcommand\deliverableTitle{Mediation amongst ontologies: Application to the FAST ontology}
+\newcommand\deliverableNumber{D2.4.2}
+\newcommand\deliverableTitle{Mediation amongst ontologies: Application to the FAST platform}
 \newcommand\deliverableTitleShort{Mediation amongst ontologies}
@@
         \vfill
-		{\large \sffamily Date: 27/02/2009}\\
+        {\large \sffamily Date: 20/02/2010}\\
         \vspace{1cm}
@@
                             \textcolor{white}{\textbf{Change description}}\\ \hline
-1.0 & 27.02.2009 & Oszkar Ambrus (NUIG) & final version ready for external review \\ \hline
+0.1 & 07.01.2010 & Oszkar Ambrus (NUIG) & initial draft \\ \hline
+%0.2 & 19.02.2009 & Oszkar Ambrus (NUIG) & content moved from wiki after internal review \\ \hline
+%0.1 & 07.01.2009 & Oszkar Ambrus (NUIG) & final version ready for external review \\ \hline
 \end{tabular}
@@
 \doublespacing
-Ontology mediation is a critical problem in heterogeneous environments,
-since inter-operating services and agents often use different
-ontologies, and the differences between them need to be determined and
-overcome to allow for a seamless data exchange between these parties.
+Ontology mediation is a critical problem in heterogeneous environments, since inter-operating services and agents often use different
+ontologies, and the differences between them need to be determined and overcome to allow for a seamless data exchange between these parties.
 This is a difficult ask, with which web-based agents must cope.
 
-We review current work and literature, discussing the main areas of the
-topic, that is the core problem of ontology mediation, the types of
-mismatches that occur between ontologies, and the three main ontology
-mediation approaches, namely (1) the representation of correspondences
-between ontologies, (2) the (semi-)automatic process of discovering
-these correspondences and (3) creation of a new ontology from the
-source ontologies.
+In D2.4.1 we reviewed current work and literature, discussing the main areas of the topic, that is the core problem of ontology mediation, the types of mismatches that occur between ontologies, and the three main ontology mediation approaches, namely (1) the representation of correspondences between ontologies, (2) the (semi-)automatic process of discovering these correspondences and (3) creation of a new ontology from the source ontologies.
 
-By the overview of current research in the field of ontology mediation,
-we lay the basis of the implementation of ontology mediation in FAST.
+Based on the theoretical foundations laid down in D2.4.1 we present our approach to integrating ontology matching into the FAST platform.
+We present the roles related to ontology matching in FAST, and the way ontology matching is achieved within each one of those roles.
 
@@
     \textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
-	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.4.1 is the first iteration of the FAST ontology mediation approach.
-This document contains a general overview of current literature regarding ontology mediation, paving the way for futher analysis of the scenarios arising in FAST that require ontology mediation.
-Based on this general overview, the implementation of ontology mediation within FAST can be researched, which will constitute the topic of the next iteration of this document.}\\ \hline
+    \textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.4.2 is the second iteration of the FAST ontology matching approach.
+The firsty iteration contained a general overview of the literature regarding ontology mediation, paving the way for futher analysis of the scenarios arising in FAST that require ontology mediation.
+Based on that general overview, this current interation presents the approach to implement ontology matching within FAST.
+We present the different roles within the platform, tasks and challenges related to them, and the way ontology matching is integrated within each role.}\\ \hline
     \textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
@@
     \textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495112} \\ \hline
-	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, ontology mediation, ontology matching, ontology alignment, ontology merging}\\ \hline
+    \textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, ontology mediation, ontology matching, ontology alignment}\\ \hline
 \end{tabular}
@@
 \paragraph{Goal and Scope.}
-The central driver for Enterprise Mashups is the lightweight resource
-composition style to build enterprise applications with reusable
-building blocks, called Gadgets. The architecture used for this
-front-end layer of Gadgets specifies a back-end layer of Services,
-designed to support machine-to-machine interaction over a network (e.g.
-Web Services, Databases, etc.). Having this case, the communication
-problems need to be addressed, that arise from disparate interfaces,
-different invocation mechanisms and even different conceptualisations
-of the data transmitted between the front-end and the back-end. This
-document deals with the third issue, namely the differences between the
-conceptualisations of the exchanged data.
+FAST is a visual programming platform allowing business users to build enterprise-class mashups, employing various underlying services and generating new ones.
+Resources in FAST are described semantically using different ontologies and vocabularies, and can therefore be combined to what we can call ``intelligent gadgets''.
+In the services created within the FAST environment, the ontologies come from different parties involved in the creation of gadgets and building blocks.
+These ontologies, though different, will sometimes cover the same domain, making ontology matching necessary, as the means to reconcile differences in the various conceptualisations.
 
-As in all fields employing semantic web paradigms, data is annotated
-using ontologies. Ontologies conceptualise the data used, describe
-their context and their semantic background that should be known to all
-agents using that data. Because we do not the have the situation where
-the same ontologies are used by all sides, having different ontologies
-and even different languages to describe them, we need to integrate the
-ontologies present in the back-end with the ones present in the
-front-end, and to find a way to mediate between the the different
-ontologies used in the two layers.
+We aim to integrate ontology matching into the FAST platform to allow for discrepant systems to intercommunicate.
+We try to find a solution to these differences which range from simple naming differences to more complex data heterogeneity, behavioural differences, discrepant paradigms and other disparities.
+The reconciliation of these differences is called ontology matching (often mentioned in the specialised literature as ontology mediation, or ontology alignment, thus we will sometimes use these terms as well).
 
-We try to find a solution to these differences which range from simple
-naming differences to more complex data heterogeneity, behavioural
-differences, discrepant paradigms and other disparities. The
-reconciliation of these differences is called ontology mediation (often
-mentioned in the specialised literature as ontology matching, or
-ontology alignment, thus we will sometimes use these terms as well). 
 \paragraph{Structure of the Document.}
-We discuss ontologies in general in Section \ref{sec:ontologies}, present the ontology
-mediation problem in Section \ref{sec:mediationproblem} , discuss the types of mismatches between
-ontologies in Section \ref{sec:ontologymismatches} and detail the main approaches to ontology
-mediation in Section \ref{sec:mediationapproaches}.
+Part \ref{part:d2.4.2} describes the approach of integrating ontology matching into the FAST environment. 
+%Describe the rest.
+Part \ref{part:d2.4.1} contains the first iteration of this document, D2.4.1, laying down the theoretical foundations of ontology matching.
+We discuss ontologies in general in Section \ref{sec:ontologies}, present the ontology mediation problem in Section \ref{sec:mediationproblem} , discuss the types  of mismatches between ontologies in Section \ref{sec:ontologymismatches} and detail the main approaches to ontology mediation in Section \ref{sec:mediationapproaches}.
+
 \paragraph{Overall Approach for WPX and relation to other Work Packages and Deliverables.}
-Ontology mediation will play an important role for the integration of 
-external ontologies into the FAST ontology infrastructure.
-By giving a detailed overview of the state-of-the-art in ontology
-mediation, we lay the groundwork for further research in T2.4 [Task
-2.4, see Project Description] with this first iteration of D2.4. Future
-versions of the deliverable (i.e., D2.4.2 and D2.4.3) will implement
-ontology mediation within the FAST platform.
+Ontology matching will play an important role for the integration of different underlying resources in the user-built gadgets. 
+We present our initial research in the implementation of ontology matching in this second iteration of D2.4, after having laid the theoretical ground work of ontology matching in the first iteration.
+The final version of the deliverable (i.e., D2.4.3) will finalize the implementation of ontology mediation within the FAST platform.
 
-\section{Ontologies}
-\label{sec:ontologies}
-This section provides a brief introduction to the concept of ontologies.
+\part{Implementing Ontology Matching within the FAST Platform}
+\label{part:d2.4.2}
 
-\subsection{Definitions}
-As specified in \cite{gruber2008ontology} {\textquotedblleft}an ontology defines a
-set of representational primitives with which to model a domain of
-knowledge or discourse{\textquotedblright}. These primitives are
-typically classes, attributes and relations among classes or class
-instances (and others, such as function terms, rules, restrictions,
-axioms and events). Their definitions provide information about their
-semantics (meaning) and constraints on their logically consistent
-usage.
+\include{fast_d2_4_2_part1}
 
-Ontologies are typically specified in languages that allow to abstract
-away from implementation details, and focus on the so-called semantic
-level. This way ontologies provide a representation independent from
-data models and implementation, thus providing a means to enable
-interoperability among disparate systems and specifying interfaces to
-independent, knowledge-based systems. In the technology stack of the
-Semantic Web, ontologies define a separate layer, and have a set of
-languages and tools (both commercial and open-source) to work with
-them.
-
-We would also mention a widely-known and used classical definition of
-ontologies, presented in \cite{gruber93towards}: {\textquotedblright}[An
-ontology is] an explicit, formal specification of a shared
-conceptualisation{\textquotedblright}. \cite{studer1998knowledge_engineering_principles} explains
-this definition, by elaborating on the four main terms of the
-definition:
-
-\begin{enumerate}
-\item it is a conceptualisation, because it models and categorises
-relevant concepts from the real world, 
-\item it is explicit, because the model explicitly states the types of
-the concepts, the relationships between them and the constraints of
-their use, 
-\item it is formal, because the ontology has to be machine-readable and
-consistent, and 
-\item it is shared, because an ontology is consensual, i.e. accepted by
-a group of people and it is used in a shared environment to allow a
-common conceptualisation of data. 
-\end{enumerate}
-
-\subsection{General Description}
-An ontology is used to define the vocabulary that agents use to exchange
-queries and assertions. If an agent commits to a common ontology, it
-guarantees the consistency (but not completeness) of the queries and
-assertions using the vocabulary defined in the ontology \cite{gruber93towards}.
-An agent supporting the interface defined in an ontology is not
-required to use the terms of the ontology as an internal encoding of
-its knowledge. Nonetheless, the definitions and formal constraints of
-the ontology do restrict what can be meaningfully stated in the
-language defined.
-
-Two analogies to ontologies are given in \cite{gruber93towards}. Ontologies are
-similar to global type declarations in a conventional software library,
-and ontological commitments are similar to type restrictions over the
-inputs and outputs of program modules. Formal argument restrictions can
-be checked mechanically by compilers, to make sure that calling
-procedures pass legal data to called procedures. Similarly, sentences
-in an exchange between agents can be checked for logical consistency
-using the definitions in the ontologies. Just as the formal argument
-list hides the internal structure of a procedure from its calling
-environment, a common ontology allows one to interact with a
-knowledge-based program without committing to or being aware of its
-internal encoding of knowledge.
-
-Ontologies are also like conceptual schemata in database systems. Such a
-schema provides a logical description of the data being shared,
-allowing applications to interact with it without having any knowledge
-of internal data structures. So while a conceptual schema defines
-relations and constraints on data, an ontology defines terms with which
-to represent knowledge, and constraints on their relationships.
-\cite{gruber2008ontology} also points out that the requirement of inputs and
-outputs to be logically consistent with the definitions and constraints
-of an ontology is analogous to the requirement that the rows of a
-database table (or insert statements in SQL) must be consistent with
-the integrity constraints imposed on the table, stated separately and
-independently of the internal data formats.
-
-\cite{gruber2008ontology} enumerates the key applications of ontologies, by saying
-that they {\textquotedblleft}are part of the W3C standards stack for
-the Semantic Web, in which they are used to specify standard conceptual
-vocabularies in which to exchange data among systems, provide services
-for answering queries, publish reusable knowledge bases, and offer
-services to facilitate interoperability across multiple, heterogeneous
-systems and databases.{\textquotedblright} The key role of ontologies
-in database systems is to provide a data modeling layer above specified
-database designs, so that data can be used across independently
-developed systems and services. This made it possible to achieve
-database interoperability, cross database search and web service
-interrogation.
-
-There are many issues not addressed by common ontologies, related to
-knowledge sharing. One important question is how a group of people can
-reach a consensus on a common ontology. The discussion of ontology
-mediation aims to address this problem by attempting a solution to the
-case of several ontologies conceptualising a similar domain of the real
-world in different ways.
-
-\section{The Mediation Problem} %%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\label{sec:mediationproblem}
-Ontologies are employed as explicit descriptions of the information
-source semantics. According to \cite{wache2001ontology_information_integration}, integrating these
-semantics along multiple systems can be done using three main
-approaches. (1) The single ontology approach, using a single ontology
-as a global shared vocabulary. Unfortunately this can not provide
-different views on a domain, and using it for multiple similar
-catalogues is unfeasible due to the vast amount of information needed
-to be captured, and the susceptibility of the conceptualisation to
-change. (2) The multiple ontology approach has each information source
-describing its own ontology. This solves problems related to the single
-ontology approach, but introduces the problem of differences in
-conceptualising the same concepts. Ontology mediation can be used to
-map different parts of the ontologies to each other. (3) The hybrid
-approach uses multiple ontologies, but uses a global ontology to make
-all the ontologies comparable to each other. This is closely related to
-ontology merging, where overlapping entities are merged, and the rest
-is simply composed into one unified ontology, allowing for the
-definition of the concepts of both ontologies. We suggest that the FAST
-Gadget Ontology (as defined in D2.1) could act as a global ontology in
-the context of FAST, although further research needs to be done in this
-direction, since the Gadget Ontology has a slightly different
-epistemological commitment.
-
-In the context of FAST, the problem is that of reconciling the
-ontologies in the back-end with the ontologies in the front-end, to
-allow for interoperability between the two layers.
-
-In the following we present a motivating scenario for ontology mediation
-in Section \ref{subsec:mediationproblem_motivatingexample}, and a formal description of the alignment problem and
-process, respectively, in Section \ref{subsec:mediationproblem_problemstatement}. After that we show some examples
-of ontology mediation in Section \ref{subsec:mediationproblem_examples}.
-
-\subsection{Motivating Example}
-\label{subsec:mediationproblem_motivatingexample}
-This section presents a brief example, illustrating a problem that
-presents the need for ontology mediation.
-
-%\begin{verbatim}
-\lstset{frame=single, basicstyle=\ttfamily, caption=Ontology \texttt{O1}.}
-\lstset{emph={vsmlVariant, namespace, ontology, concept, ofType, instance, nonFunctionalProperties, endNonFunctionalProperties, memberOf, hasValue}}
-\lstset{emphstyle=\textbf\textit\underbar}
-\begin{lstlisting}
-vsmlVariant -"http://www.smo.org/wsml/wsml-syntax/wsml-flight"
-
-namespace { -"http://see.deri.org/adrian/thesis/ontologies/O1#" ,
-       wsml - "http://www.wsmo.org/wsml/wsml-syntax#" ,
-         dc - "http://purl.org/dc/elements/1.1/" }
-
-ontology - "http://see.deri.org/thesis/ontologies/O1"
-    nonFunctionalProperties
-    dc#description hasValue "A simple ontology modeling
-                             the concept of a person"
-    endNonFunctionalProperties
-
-concept person
-    name ofType _string
-    age ofType _integer
-    hasGender ofType gender
-    hasChild ofType person
-    marriedTo ofType person
-
-concept gender
-    value ofType _string
-
-instance male memberOf gender
-    value hasValue "male"
-instance female memberOf gender
-    value hasValue "female"
-\end{lstlisting}
-
-Consider the two ontologies \texttt{O1} and
-\texttt{O2} \cite{mocan2008mediation} presented in Listing 1 and
-Listing 2 respectively, in the Human Readable Syntax of the Web Service
-Modeling Language (WSML) \cite{debruin2005wsml}. Elements from
-ontology \texttt{O1} need to be mapped to the elements
-described in Ontology \texttt{O2}.
-
-Ontology \texttt{O1} describes the concept
-\texttt{person} as one having five attributes, each of
-them having a type that is either a concept or a value from a given
-data type. The concept gender has two instances defined, that have
-attributes pointing to the corresponding values.
-
-\lstset{frame=single, caption=Ontology \texttt{O2}.}
-%\lstset{emph={vsmlVariant, namespace, ontology, concept, ofType, instance, nonFunctionalProperties, endNonFunctionalProperties, memberOf, hasValue}}
-%\lstset{emphstyle=\textbf}
-\begin{lstlisting}
-vsmlVariant -"http://www.smo.org/wsml/wsml-syntax/wsml-flight"
-
-namespace { -"http://see.deri.org/adrian/thesis/ontologies/O2#" ,
-       wsml - "http://www.wsmo.org/wsml/wsml-syntax#" ,
-         dc - "http://purl.org/dc/elements/1.1/" }
-
-ontology - "http://see.deri.org/thesis/ontologies/O2"
-    nonFunctionalProperties
-    dc#description hasValue "A simple ontology modeling
-                             the concept of a human"
-    endNonFunctionalProperties
-
-concept human
-    name ofType _string
-    age ofType _integer
-    noOfChildren ofType _integer
-
-concept man subConceptOf human
-
-concept woman subConceptOf human
-
-concept marriage
-    hasParticipant ofType human
-    date ofType _date
-\end{lstlisting}
-
-%\newline
-Ontology \texttt{O2} describes the concept
-\texttt{human} as one having three attributes. The
-concepts \texttt{man} and \texttt{woman} are
-subclasses of the concept \texttt{human}.
-
-Additionally a fourth concept called \texttt{marriage} is
-defined having two attributes. (For simplicity we do not deal with the
-exceptions of polygamy and same-sex marriage).
-
-Suppose that a service (or agent) using ontology O1 needs to interact
-with a service accepting and providing data that is semantically
-defined in ontology O2.
-
-The first step is to identify candidates to be mapped or to have
-taxonomic relationships under an integrated schema \cite{shvaiko2005schema_based}. It is obvious that \texttt{person} and
-\texttt{human} are candidates to be mapped, as well as the
-class \texttt{man} with the class
-\texttt{person} having the attribute
-\texttt{gender} equal to
-{\textquotedblleft}\texttt{male}{\textquotedblright} and
-the class \texttt{woman} with the class
-\texttt{person} having the attribute
-\texttt{gender} equal to
-{\textquotedblleft}\texttt{female}{\textquotedblright}.
-
-After having identified these correspondences, the next step is to
-generate query expressions and assertions that automatically translate
-data instances of one of these ontologies into the other one, or that
-automatically handle their semantic relations under an integrated
-ontology.
-
-\subsection{Problem Statement}
-\label{subsec:mediationproblem_problemstatement}
-Given two ontologies \texttt{o} and
-\texttt{o{\textquotesingle}} that need to be mapped to
-each other, according to \cite{shvaiko2005schema_based} we can define a
-mapping element as a 5-tuple \texttt{{\textless}id, e,
-e{\textquotesingle}, n, R{\textgreater}}, where
-
-\begin{itemize}
-\item \texttt{id} is a unique identifier, identifying the
-mapping element, 
-\item \texttt{e} and
-\texttt{e{\textquotesingle}} entities (classes,
-attributes, relations) of the first and second ontology, respectively, 
-\item \texttt{n} is a confidence measure in some
-mathematical structure holding the correspondence value between e and
-e{\textquotesingle}, 
-\item \texttt{R} is the correspondence relation holding
-between e and e{\textquotesingle} (e.g.,
-\texttt{equivalence (=); more general($\sqsupseteq$); disjointness($\perp$); overlapping($\sqcap$)}.
-\end{itemize}
-
-\begin{figure}
-    \centering
-        \includegraphics[width=72mm, height=45mm]{images/Alignment_process.png}%[bb=0 0 201 126]
-        \caption{The alignment process}
-    \label{fig:Alignment}
-\end{figure}
-
-A mapping is a set of the above defined mapping elements. The alignment
-operation determines the mapping
-(M\texttt{{\textquotesingle}}) for the pair of ontologies
-(\texttt{o} and
-\texttt{o{\textquotesingle}}). There are some parameters
-than can extend the definition of the alignment process, namely:
-
-\begin{enumerate}
-\item an input mapping, \texttt{M}, which is to be used and
-completed by the process, 
-\item the mapping parameters, \texttt{p} (such as weights
-and thresholds), that configure the alignment process, 
-\item other external resources used by the alignment,
-\texttt{r} (such as dictionaries, thesauri, etc.); See Figure \ref{fig:Alignment}. 
-\end{enumerate}
-For example, suppose that based on some algorithm (using linguistic and
-structural analysis) the confidence measure for the equivalence
-relation between the classes \texttt{person} in
-\texttt{O1} and \texttt{human} in
-\texttt{O2} is 0.62. Suppose that a mapping parameter
-specifies a threshold of 0.55 for this algorithm, that is all pairs of
-entities with a confidence measure higher that 0.55 will be considered
-correct mapping elements. Thus our algorithms would return the
-following mapping element: \texttt{{\textless} id11,
-person, human, 0.62, = {\textgreater}.}Now, another algorithm
-determines the two concepts to mean exactly the same thing, thus it
-either does not compute a confidence measure (returning
-\texttt{n/a}), or it returns the upper limit of the
-confidence measure interval (e.g. 1). Thus the mapping element would
-be: \texttt{{\textless} id11, person, human, n/a, =
-{\textgreater}.}
-
-It should be noted, that the entities mapped between two ontologies can
-be constructed using patterns, i.e. more than just mapping concepts to
-concepts, and so on. For example, we can map a concept to a concept
-having a certain attribute, like in the case of the concept
-\texttt{person} having the attribute
-\texttt{hasGender} equal to \texttt{female}
-from \texttt{O1} being mapped to the concept
-\texttt{woman} from \texttt{O2}.
-
-\subsection{Mediation Examples}
-\label{subsec:mediationproblem_examples}
-We present two further examples to better illustrate the use of ontology
-mediation.
-
-The first example is taken from \cite{giunchiglia2008evaluation}. Based on the
-idea that we view ontologies as graph-like structures, we exemplify
-ontology matching with the help of fragments of two tree-like
-structures, such as Google and Looksmart. Notice that in the general
-case the relation holding between nodes is not specialisation, but
-classification (parent-child relation).
-
-\begin{figure}[h]
-    \centering
-        \includegraphics[width=101mm, height=67mm]{images/F2GoogleLooksmart.png}%[bb=0 0 254 167]
-        \caption{Fragments of Google and Looksmart directories and some
-correspondences. Double arrowheads denote equivalence, while single
-arrowheads stand for the more general relation. Image taken from
-\cite{giunchiglia2008evaluation}.}
-    \label{fig:GoogleLooksmart}
-\end{figure}
-
-Suppose we want to merge the merge the two structures. Such situations
-can arise, for example when an e-commerce company acquires another one.
-One probable mapping is given in Figure \ref{fig:GoogleLooksmart}. We have found, for example,
-that \texttt{O1:Basketball} is equivalent to
-\texttt{O2:Basketball}, and that
-\texttt{O2:Games} is more general than
-\texttt{O1:Board\_Games.}
-
-The second example given in \cite{madhavan2001schema_matching} illustrates the
-need to use several techniques in matching two ontologies. Figure \ref{fig:POSchemas}
-shows two purchase order schemas that need to be matched.
-
-\begin{figure}[h]
-    \centering
-        \includegraphics [width=101mm, height=42mm]{images/F1POSchemas.png}%[bb=0 0 257 105]
-        \caption{Matching purchase order schemas. Image taken from \cite{madhavan2001schema_matching}.}
-    \label{fig:POSchemas}
-\end{figure}
-
-As stated above in Section 3.2 we look for similarity coefficients
-between the two schemas and deduce a mapping from those coefficients.
-Suppose, as a first phase we use linguistic matching, using thesauri to
-help match names, acronyms, synonyms, etc. resulting in a set of
-candidate pairs (such as \texttt{PO} and
-\texttt{PurchaseOrder}, or \texttt{Bill} and
-\texttt{Invoice}) and language similarity coefficients
-(\texttt{lsim}). In the second phase we would perform a
-structural matching of the two schemas. For example,
-\texttt{Line} would be mapped to
-\texttt{ItemNumber}, because their parents and their other
-two children match, \texttt{City} under
-\texttt{POBillTo} will be mapped to
-\texttt{City} under \texttt{InvoiceTo},
-because bill is a synonym to invoice. This process results in
-structural similarity coefficients (\texttt{ssim}). By
-using a weighted similarity, we can compute the final mappings (e.g. by
-using the formula \texttt{wsim=ws*wsim+(1-ws)*lsim}, where
-\texttt{ws} is the interval [0,1]).
-
-\section{Ontology Mismatches}
-\label{sec:ontologymismatches}
-Ontology mismatches are the main obstacles in the combined use of
-independently developed ontologies. We will now give an overview of the
-main type of ontology mismatches as presented in \cite{klein2001combiningOntologies}, which
-is the main source of input to this state-of-the-art report. We will
-classify ontology mismatches into given types and will find their
-relationships.
-
-Firstly, two levels of mismatches can be distinguished. The first level
-is the \textit{language} or meta-model level. These are differences in
-the primitives that make up the language used to specify the ontology.
-They produce mismatches in the mechanics to define classes, relations,
-attributes and so on. The second level is the \textit{ontology} or
-model level. This is were the actual ontology is located. These are
-differences in the way the domain is modeled.
-
-In this section we discuss the overlap and mismatches between concepts,
-relations, attributes and instances. We will detail the two levels of
-mismatches and the different types that occur at each level.
-
-\subsection{Language Level Mismatches}
-\label{subsec:ontologymismatches_languagelevel}
-We define four types of mismatches, although they often coincide. These
-occur when ontologies written in different languages are combined. We
-note that these type of mismatches will be less relevant for FAST
-because we assume that standards for ontology languages are now so well
-established that in practice we will not very often find ontologies
-written in different languages anymore, but we discuss them briefly to
-give a better overview of the meditation problem.
-
-\subsubsection{Syntax}
-Different languages obviously use different syntax. Therefore the same
-concept can (and will be) often modeled in a completely different way.
-For example, the concept of chairs would be defined in RDF Schema as
-\texttt{{\textless}rdfs:Class
-ID={\textquotedblright}Chair{\textquotedblright}{\textgreater}}, while
-in LOOM it would be \texttt{(defconcept Chair)}.
-
-This is the simplest possible mismatch, although it often comes coupled
-with other mismatches as well. The simplest form is when a language has
-several syntactical forms.
-
-The solution is to simply to have a rewrite mechanism that returns the
-corresponding translated form.
-
-\subsubsection{Logical representation}
-A slightly more complicated mismatch is when a logical notion can not be
-represented in a given language. For example, in certain languages it
-is possible to assert that two classes are disjoint, say using
-\texttt{(disjoint A B),} whereas in other languages it is
-necessary to express this using simpler statements, such as the
-negation of set-inclusion statements, like \texttt{A
-subclass-of (NOT B), B subclass-of (NOT A).}
-
-The actual problem is not the question if something can be expressed,
-but finding the language constructs to express a logical notion (not a
-concept).
-
-Solving this mismatch is relatively-easy: it consists in finding
-translation rules to transform one representation into another.
-
-\subsubsection{Semantics of primitives}
-This difference is a more subtle one, and occurs at the meta-model of
-the language. It sometimes occurs that the semantics of an identical
-name differs from language to language. For example, the expression
-\texttt{A equalsTo B} can have many different
-interpretations.
-
-Even when two ontologies use the same syntax, the semantics may differ.
-For example, the OIL RDF syntax interprets several
-\texttt{{\textless}rdfs:domain{\textgreater}} statements
-as the intersection of the arguments, whereas RDF Schema interprets
-them as the union of the arguments.
-
-\subsubsection{Language expressiveness}
-This mismatch, at the meta-model level of the languages, is the one
-having the greatest impact on the attempt to combine and reconcile
-ontologies.
-
-The difference implies that some languages can express notions that
-others can not. For example, some languages have no constructs to
-express negation, whereas other do. Other typical examples are support
-for sets, lists, etc.
-
-These type of mismatches have the greatest impact, and are sometimes
-mentioned as the {\textquotedblleft}fundamental
-differences{\textquotedblright} between knowledge models.
-
-\subsection{Ontology Level Mismatches}
-\label{subsec:ontologymismatches_ontologylevel}
-Mismatches at the model level happen both in the case of having the two
-ontologies written in the same language and syntax and in the case of
-different languages. We can distinguish between several types of
-ontology level mismatches.
-
-A conceptualisation mismatch is a difference in how the domain is
-modeled (conceptualised), therefore we will have different concepts and
-different relations between them. Explication mismatches are
-differences in the way the concepts are specified. These occur as
-mismatches in definitions, or terms, or the combination of the two.
-Terminological mismatches are naming differences or conflicts. Finally,
-encoding mismatches are differences in the way values are specified,
-for example currencies for prices.
-
-These kinds of mismatches are likely to occur in the context of FAST and
-are therefore important factors in the implementation of ontology
-mediation in FAST.
-
-\subsubsection{Conceptualisation Mismatches}
-These are differences in the conceptualisation of the domain, that is
-not only the specification differs in the two ontologies, but the model
-as well.
-
-Sometimes these mismatches can not be reconciled automatically, and need
-a domain expert. In this case, a good solution would be to use both
-ontologies together. In that case, the overlapping part needs to be
-aligned, and the rest can be merged.
-
-\paragraph{Scope.}
-
-Two classes that seemingly represent the same concept, but do not have
-the same instances, although some of these are identical. For example
-the concept of {\textquotedblleft}employee{\textquotedblright} can be
-perceived and specified in different ways by different organisations.
-
-\paragraph{Model coverage and granularity.}
-This mismatch is related to the part of the domain that is covered by
-the ontologies, and the level of detail respectively. For example, an
-ontology could model cars, but no trucks, whereas another would have a
-few categories for trucks, and another third ontology could make very
-detailed distinction between all the types of trucks and their
-properties.
-
-\subsubsection{Explication Mismatches}
-The first two types of explication mismatches (\textit{paradigm} and
-\textit{concept description}) result from the \textbf{style of
-modeling.} The next two types (\textit{synonym terms} and
-\textit{homonym terms}) can be classified as \textbf{terminological
-mismatches}. Finally, there is one trivial difference, cause by
-\textit{encoding} mismatches.
-
-\paragraph{Paradigm.}
-Different paradigms can be used to represent different concepts, or
-different top-level ontologies can be used, which all result in this
-kind of mismatch. For example, one could model time based on interval
-logic, while another could use discrete time points.
-
-\paragraph{Concept Description.}
-These types of mismatches are called modeling conventions. Several
-choices can be made in different modeling questions. One choice, the
-categorisation of concepts can be done based on an attribute, or using
-subclasses. For example, as presented in Section 3.1, the concepts
-{\textquotedblleft}man{\textquotedblright} and
-{\textquotedblleft}woman{\textquotedblright} can be modeled in several
-ways. They could be instances of the concept, and differ only in an
-attribute (e.g. \texttt{gender}), or they can be separate
-subclasses of a top-level class representing
-{\textquotedblleft}humans{\textquotedblright} in general.
-
-Another choice is the way in which a hierarchy is built - distinction
-between features can be made higher or lower in the hierarchy. For
-example, consider the place where the difference between scientific and
-non-scientific publications is made: a dissertation can be modeled as
-\texttt{dissertation }\texttt{{\textless}
-book {\textless} scientific publication {\textless} publication}, or as
-\texttt{dissertation {\textless} scientific book
-{\textless} book {\textless} publication}, or as a subclass of both
-\texttt{scientific publication} and
-\texttt{book}.
-
-\paragraph{Synonym Terms.}
-Concepts are represented by different names. For example, the term
-{\textquotedblleft}car{\textquotedblright} from an ontology can be
-represented as {\textquotedblleft}automobile{\textquotedblright} in
-another.
-
-Although the solution looks simple in using thesauri, in reality much
-human effort is needed to solve these problems, and care must be taken
-not to overlap with scope mismatches (see above).
-
-\paragraph{Homonym Terms.}
-The meaning of a term is different in another context. For example, the
-term {\textquotedblleft}conductor{\textquotedblright} has one meaning
-in a musical context, and a different meaning in an electric
-engineering domain.
-
-These differences are much harder to solve than synonyms, and always
-need human contribution.
-
-\paragraph{Encoding.}
-The values of attributes in ontologies are encoded in different formats.
-For example, date can be encoded in a
-\texttt{{\textquotedblleft}yyyy-mm-dd{\textquotedblright}}
-form, or a
-\texttt{{\textquotedblleft}dd/mm/yyyy{\textquotedblright}},
-distance might be specified in miles or kilometers, money amount can be
-given in different currencies, and so on.
-
-These kind of mismatches can very easily be solved, by using a wrapper,
-or a single transformation step.
-
-\section{Ontology Mediation Approaches}
-\label{sec:mediationapproaches}
-In this section we give an overview of some of the major approaches to
-ontology mediation, following the structure of Section 2: Approaches in
-Ontology Mediation of \cite{debruin2005wsml}. Firstly, in Section
-\ref{subsec:mediationapproaches_ontologymapping} we present some representative approaches to the specification of
-mappings. Then in Section \ref{subsec:mediationapproaches_ontologyalignment} we overview some of the main approaches
-and algorithms to overcome mismatches and find mappings between
-ontologies, that is we discuss the topic of ontology alignment. Finally
-in Section \ref{subsec:mediationapproaches_ontologymerging} we discuss the process of merging several ontologies.
-Further research is needed to decide on the type of ontology mediation
-to be used in the context of FAST, the algorithms and approaches
-applied, and whether to use alignment or merging.
-
-\subsection{Ontology Mapping}
-\label{subsec:mediationapproaches_ontologymapping}
-An ontology mapping is a declarative specification of the semantic
-overlap between two ontologies \cite{debruin2005wsml}. It is the
-result of the ontology alignment process (see Figure 3.). This mapping
-is usually expressed as a set of axioms in a certain mapping language.
-
-There are three main phases of the mapping process: (1) discovering the
-mapping, (2) representing the mapping and (3)
-using/exploiting/executing the mapping. This section is focused on (2),
-that is we survey a number of approaches in representing ontology
-mappings.
-
-A common tendency in ontology mapping is to create an ontology of
-mappings, which represents the vocabulary for the representation of
-mappings.
-
-\paragraph{MAFRA.}
-MAFRA \cite{maedche2002mafra}, or MApping FRAmework for distributed
-ontologies supports the interactive, incremental and dynamic process of
-ontology mapping, which transforms instances of a source ontology into
-instances of a target ontology.
-
-\subparagraph{Framework Components.}
-The framework consists of five main phases (called horizontal
-dimensions) and for components that run along the entire process,
-called vertical components. Within the horizontal dimension, the
-authors identified five modules:
-
-\begin{itemize}
-\item \textit{Lift \& Normalisation}. Both ontologies must be normalised
-to the uniform representation of RDFS. As a result both ontologies are
-represented in RDF Schema with their instances in RDF. Thus syntactical
-end lexical differences are eliminated and semantic differences become
-more apparent. 
-\item \textit{Similarity.} This module establishes similarities between
-entities using different approaches (\textit{lexical similarity},
-\textit{property similarity} based on properties of concepts,
-\textit{bottom-up similarity} propagating the similarity from lower
-parts of the taxonomy to the upper part and \textit{top-down
-similarity} that assumes special relevance when top level concepts have
-a higher or lower similarity). 
-\item \textit{Semantic Bridging}. Based on the similarities computed in
-the previous phase, correspondences are identified between entities,
-using so-called semantic bridges, which define the actual mapping. 
-\item \textit{Execution}. Transforms instances between the two
-ontologies using the semantic bridges. 
-\item \textit{Post-processing.} Revisiting the mapping for improvements.
-
-\end{itemize}
-The vertical dimension contains modules like evolution (synchronising
-changes between in the ontologies and the bridges), the GUI (extensive
-graphical support must be given for human assistance of the process),
-etc.
-
-\subparagraph{Semantic Bridges.}
-Semantic bridges are captured in the Semantic Bridging Ontology (SBO).
-Five dimensions have been identified for the semantic bridges, as
-presented in the following:
-
-\begin{itemize}
-\item \textit{Entity dimension.} Semantic bridges may relate to the
-following entities: (i) concepts modeling object classes from the real
-world, (ii) relations modeling relationships of objects, (iii)
-attributes modeling simple properties and (iv) extensional patterns
-modeling the content of instances. 
-\item \textit{Cardinality dimension.} Defines the number of entities on
-both sides of the semantic bridge (mostly 1:1, 1:n and m:1; m:n is
-rarely encountered, and can be usually decomposed onto m:1:n). 
-\item \textit{Structural dimension.} Reflects on how elementary bridges
-may be combined into more complex bridges. Relations that hold among
-bridges are \textit{specialisation} (reuse a bridge and provide
-additional information), \textit{abstraction} (should only be used as a
-super-class), \textit{composition} (a bridge is composed of other
-bridges) and \textit{alternatives} (mutually exclusive bridges). 
-\item \textit{Constraint dimension.} Permits to control the execution of
-the semantic bridge. Constraints must hold on the source ontology in
-order for the transformation procedures to be applied. 
-\item \textit{Transformation dimension.} Reflects on how instances of
-the source ontology are transformed during the mapping process. 
-\end{itemize}
-The Semantic Bridging Ontology (SBO) contains a specification of all
-semantic bridges, organised in a taxonomy. To actually relate the
-source and the target ontology the mapping process creates an instance
-of the SBO containing semantic bridge instances along with all the
-necessary information to transform instances of entities from the
-source to the target ontology.
-
-The five semantic bridge dimensions are employed as follows:
-
-\begin{itemize}
-\item There are three basic entity types: \textit{Concepts},
-\textit{Relations} and \textit{Attributes}. 
-\item The class \textit{SemanticBridge} is the most generic bridge type.
-It defines the relations to the source and target entities. It is
-specialised according to entity type and cardinality of the relation. 
-\item The class \textit{Service} can reference resources that are
-responsible to connect to, or to describe transformations. 
-\item \textit{Rule} is the general class for constraints and
-transformation-relevant information. 
-\item The class \textit{Transformation} uses the \textit{inService}
-relation to link to the transformation procedure. 
-\item The class \textit{Condition} represents the conditions that need
-to be satisfied in order to execute the semantic bridge. 
-\item The Composition primitive can be accomplished using the
-\textit{hasBridge} relation of \textit{SemanticBridge}, without any
-cardinality or type constraint. 
-\item The Alternative primitive is supported by the
-\textit{SemanticBridgeAlt} class. It groups several mutually exclusive
-bridges, of which the first consistent bridge (with satisfied
-conditions) is executed. 
-\end{itemize}
-For an extensive and relevant example please refer to Section 3.3. of
-\cite{maedche2002mafra}.
-
-\paragraph{RDFT.}
-RDFT \cite{omelayenko2002rdft} is a small language (meta-ontology) defined for
-mapping XML DTDs to/and RDF Schemas specially targeted for business
-integration tasks. It is built on top of RDF Schema and it is used to
-map RDF Schemas and concepts like events, messages, vocabularies to
-XML-specific parts of the conceptual models that occur in the
-integration tasks.
-
-The model is derived from WSDL \cite{christensen2001wsdl} and provides
-an RDF Schema for RDF annotations of WSDL documents, and is extended
-with the temporal ontology PSL. The business integration task in this
-context is seen as a service integration task, where each enterprise is
-represented as a Web service specified in WSDL. WSDL annotations made
-according to the defined meta-ontology allow performing inference over
-WSDL descriptions to validate the links established between the
-enterprises.
-
-\subparagraph{PSL.}
-To be able reason about the inputs and outputs of the companies being
-integrated, the authors developed a conceptual model of WSDL, which is
-directly derived from the WSDL specification. WSDL defines the
-following basic elements of services, on top of which PSL is built:
-
-\begin{itemize}
-\item \texttt{Types} that provide links to the XML Schemas
-of the messages exchanged; 
-\item abstract definitions of \texttt{Messages} in
-accordance with \texttt{Types}; 
-\item \texttt{Port} \texttt{Types} that
-specify input and output messages; 
-\item \texttt{Bindings} that specify concrete protocols and
-formats for messages according to \texttt{Types}. 
-\end{itemize}
-These elements can describe services but do not represent any temporal
-knowledge about the messages requiring integration. The Process
-Specification Language or PSL temporal ontology includes classes to
-capture temporal knowledge:
-
-\begin{itemize}
-\item \texttt{activity} , which is performed during a
-certain time interval; 
-\item \texttt{activity} \texttt{occurrence} ,
-contains the snapshot of an activity at a moment in time; 
-\item \texttt{timepoint} , marks the time interval of the
-activity; 
-\item \texttt{objects} , things that do not possess
-temporal properties, but may participate in certain activities at given
-timepoints. 
-\end{itemize}
-\subparagraph{RDFT.}
-The basic RDFT class is \texttt{Bridge }that connects two
-concepts. It describes common properties of bridges, which enables one
-to specify correspondences between an entity and a set of entities.
-allowing for one-to-many and many-to-one connections.
-
-The bridges contain the \texttt{Relation} property,
-pointing to one of the subclasses of
-\texttt{BridgeRelation}:
-
-\begin{itemize}
-\item \texttt{EquivalenceRelation}, stating that the source
-element is equivalent with the target set of elements, or the source
-set of elements is equivalent to the target element respectively,
-depending on the relation cardinality; 
-\item \texttt{VersionRelation}, specifying that the target
-set of elements forms a later version of the source set of elements.
-Unlike equivalence bridges, assumes that both concepts belong to the
-same domain. 
-\end{itemize}
-Several types of bridges are defined in RDFT:
-
-\begin{itemize}
-\item \texttt{Event2Event} bridges link different events,
-specify temporal event generation conditions, and link the events to
-the messages transmitted with them. They connect instances of the
-meta-class \texttt{mediator:Event}. 
-\item two kinds of \texttt{RDFBridges}:
-\texttt{Class2Class} and
-\texttt{Property2Property} bridges between RDF Schema
-classes and properties. They contain \texttt{rdfs:Class
-}and \texttt{rdfs:Property} instances, respectively. 
-\item four kinds of \texttt{XMLBridges}:
-\texttt{Tag2Class}, \texttt{Tag2Property},
-\texttt{Class2Tag}, \texttt{Class2Property},
-which link XML DTD tags and RDF Schema classes and properties. 
-\end{itemize}
-The bridges are grouped into \texttt{Maps}, which are
-collections of bridges serving a single purpose. The maps are
-identified by their names and form minimal reusable modules of RDFT
-bridges. Each map can include other maps and serves as a container for
-\texttt{Bridges}. Described in this manner, as a set of
-bridges, mappings are said to be \textit{declarative}, while
-\textit{procedural} mappings can be defined as Xpath \cite{clark1999xpath} expressions, transforming instance data.
-
-Connecting two services with RDFT consists of connecting their events
-(instantiating \texttt{EventMap}) consisting of
-\texttt{Event2Event} bridges. Each of the bridges points
-to a \texttt{DocumentMap} aligning the documents attached
-to the bridges and, in turn, consisting of
-\texttt{RDFBridges}, \texttt{XMLBridges} and
-\texttt{VocabularyMaps}.
-
-\paragraph{C-OWL.}
-C-OWL \cite{bouquet2004cowl} gives another perspective on ontology
-alignment, extending the widely-known OWL \cite{dean2004owl}
-language.
-
-\subparagraph{The Vision.}
-In their vision, similar to the one discussed at the beginning of
-Section 3, two types of domain conceptualisations exist:
-
-\begin{itemize}
-\item \textit{ontologies}, that encode a view common to a set of
-different parties. They make it easy to exchange information, although
-they have the requirement of common consensus and the problem of hard
-maintenance. 
-\item \textit{contexts}, that encode the view of a party, and are local
-conceptualisations, that are not shared. These are easy to maintain,
-and no consensus is required, but need explicit mapping among the
-contexts of different parties to be able to exchange information. 
-\end{itemize}
-In this vision, an ontology is contextualised, i.e. it is a
-\textit{contextual} \textit{ontology}, when its contents are kept local
-(therefore not shared) and can be mapped to the contents of other
-ontologies via explicit mappings, allowing for a controlled form of
-global visibility. This is opposed to the OWL importing mechanism,
-where a set of local models is globalised in a unique shared model, by
-importing complete models and using the imported elements by direct
-reference.
-
-\subparagraph{Context OWL.}
-Context OWL or C-OWL is a language that extends OWL both syntactically
-and semantically to support the concept of contextual ontologies.
-
-Therefore a contextual ontology defined in C-OWL is a pair consisting
-of:
-
-\begin{enumerate}
-\item an OWL ontology, and 
-\item mapping between contexts, namely a set of bridge rules with the
-same target ontology. 
-\end{enumerate}
-A C-OWL mapping therefore is a 4-tuple having the following form:
-
-\begin{enumerate}
-\item a mapping identifier (URI), 
-\item a source context containing an OWL ontology (URI of the ontology),
-
-\item a target context containing an OWL ontology (URI of the ontology),
-
-\item a set of bridge rules form the local language of the source
-ontology to the local language of the target ontology. Each mapping is
-composed of three elements:
-
-\begin{enumerate}
-\item a source element (concept, role or individual) of the source
-ontology; 
-\item a target element, which must be of the same type as the source
-element; 
-\item the type of mapping (subsuming($\sqsupseteq$), equivalence(=), disjointness($\perp$) and
-overlapping($\sqcap$)). 
-\end{enumerate}
-\end{enumerate}
-Thus C-OWL has the complete representational power of OWL, augmented
-with appropriate local model semantics for mapping between contexts
-(local ontologies). A nice property of C-OWL is that the two components
-(the OWL ontology and the mapping) are orthogonal, thus one can use the
-ontology or the contextual component in an independent manner.
-
-The local model semantics defined in C-OWL, as opposed to OWL, considers
-that each context has a local set of models and a local domain of
-interpretation. Thus, it is possible to have contradicting axioms or
-unsatisfiable ontologies in the local models, without having the entire
-context space unsatisfiable.
-
-\subsection{Ontology Alignment}
-\label{subsec:mediationapproaches_ontologyalignment}
-\subsubsection{The Match Operator}
-Ontology alignment is the process of discovering similarities between
-two source ontologies \cite{debruin2005wsml}. This process can be
-described as the application of the so-called \textit{Match} operator
-described in \cite{rahm01survey}. The result of the matching
-process is a specification of the similarities between the two
-ontologies. The input of the Match operator consists mainly of two
-ontologies and some other inputs (initial mapping that will be
-extended, mapping parameters that configure the process and external
-resources used by the matching process; see Section 3.2).
-
-The generic implementation of the Match operator should contain a
-uniform internal representation of the schemas to be matched, in order
-to significantly reduce complexity. This, of course, requires a
-semantics-preserving schema importer and exporter, respectively.
-
-In general it is not possible to fully automate the matching between two
-ontologies, primarily because they often have semantics that affects
-the matching criteria, but is not formally expressed or often even
-documented. Therefore the implementation of Match should determine
-match candidates, which the user can accept, reject, or change.
-Furthermore the user should be able to specify matches for elements for
-which the system was unable to find satisfactory match candidates.
-
-\subsubsection{The Alignment Process}
-The alignment process proposed in \cite{debruin2005wsml} relieves the user of
-some of the burdens in creating the mappings. The input of the process
-consists of two ontologies which are to be aligned. The output is a set
-of mappings.
-
-The steps of the alignment are as follows:
-
-\begin{enumerate}
-\item \textit{Feature engineering}. Selects only parts of an ontology
-definition in order to describe a specific entity. A feature may be as
-simple as a label or it may include super- and sub-concepts, relations
-or extensional descriptions.\newline
-For example, suppose a fragment of an ontology describes the instance
-\texttt{Daimler}. In this case we should consider the
-generic ontology feature called type which has the values
-\texttt{luxury} and \texttt{automobile},
-respectively. 
-\item \textit{Selection of Next Search Steps}. This step chooses the
-next set of candidate pairs. It may choose to compute a restricted
-subset of candidate concept pairs of the two ontologies and ignore the
-others. 
-\item \textit{Similarity Assessment.} Determines the similarity degree
-of candidate pairs. Heuristics are used, that is similarity functions
-such as on strings, object sets, related concepts, and so on.\newline
-For example, one similarity function can check whether the parents of
-the compared two concepts are identical. 
-\item \textit{Similarity Aggregation}. For a candidate pair we can (and
-often do) use several similarity functions. The results of these
-functions need to be aggregated to compute the final similarity measure
-between the two entities. This can be done by either a simple
-averaging, or a complex aggregation function using weighting schemes. 
-\item \textit{Interpretation}. This step uses the aggregated similarity
-values to align entities. It can use thresholds for similarity, perform
-relaxation labelling, or combine structural and similarity
-criteria.\newline
-For example, suppose \texttt{simil(o1:Person,
-o2:Human)=0.62 ${\geq}$ 0.5}, thus it is added to the result as a
-mapping element. 
-\item \textit{Iteration}. Several algorithms perform more than one
-iteration in order to bootstrap the amount of structural knowledge.
-Iteration may stop when no new alignments are proposed, or when a given
-number of mappings have been discovered. Note that certain steps can be
-omitted in later iterations, since we can use previously computed
-values (such as selecting features, computing a given similarity,
-etc.). 
-\end{enumerate}
-The output of the process is a set of correspondences between entities
-of the two ontologies. We cannot expect all correspondences to be
-discovered, therefore the results of the alignment process can be
-considered as the input to a manual refinement process.
-
-\subsubsection{Classification of Approaches}
-An implementation of Match may use multiple matching algorithms or
-\textit{matchers. }This allows us to select matchers according to the
-application domain or schema types.
-
-Given that we want two use multiple matchers we distinguish two
-subproblems. First, there is the need to realise individual matchers,
-each of which uses a single matching criterion to find mappings.
-Second, we want to combine different individual matchers, either by
-using multiple matching criteria within an integrated \textit{hybrid}
-\textit{matcher}, or by combining multiple match results provided by
-different match algorithms within a \textit{composite}
-\textit{matcher}.
-
-For individual matchers, we consider the following mostly-orthogonal
-classification criteria (of which the first one is the most important
-distinctive criterion):
-
-\begin{itemize}
-\item \textit{instance-based/schema-based} matching: A schema-based
-matcher uses different approaches to determine correspondence between
-ontologies based on concepts and relations \cite{debruin2005wsml}. 
-An instance-based matcher takes instances
-(i.e. instance data) belonging to the different concepts in the
-ontologies and uses these to discover similarities between the concept.
-
-\item \textit{element-level/structure-level} matching: an element-level
-matcher takes individual schema elements, properties of the particular
-concept or relation, and uses these to find similarities, whereas a
-structure level matcher compares the combination of elements, the
-structure (e.g. the concept hierarchy) of the ontologies to find
-similarities. 
-\item \textit{language/constraint}: a matcher can use a linguistic
-approach (e.g. based on names and textual description of schema
-elements) or a constraint-based approach (e.g. based on keys and
-relationships). 
-\item \textit{matching cardinality:} an overall matching result can
-relate one or more elements of an ontology to one or more elements of
-the other, giving four cases: 1:1, 1:n, m:1 and n:m. In addition each
-mapping element can may interrelate one ore more elements of the given
-ontologies. Furthermore, there may be different cardinalities at the
-instance level. 
-\item \textit{auxiliary information}: most matchers rely on external
-information besides the two given ontologies, such as global schemas,
-dictionaries, previous matching decisions and user input. 
-\item \textit{accepted input:} this dimension concerns the type of input
-on which the algorithms operate, classified depending on the
-data/conceptual models they accept. 
-\item \textit{produced output}: this pertains to the way the mapping is
-described, whether it is a one-to-one correspondence or a more general
-one, whether it is a graded answer (98\% or 4/5), or an all-or-nothing
-approach, whether the result focuses only on equivalence between
-entities, or can it give a more expressive result, and so on. 
-\end{itemize}
-
-\paragraph{Schema-based matching.}
-A schema-based matcher \cite{rahm01survey} takes different aspects
-of the concepts and relations in ontologies and uses some similarity
-measure to determine correspondence. The available information includes
-the usual properties of ontology elements, such as name, description,
-data types, relationship types (part-of, is-a, etc.), constraints and
-ontology structures. In the following we discuss some of the main
-aspects of schema-based matching.
-
-\begin{itemize}
-\item \textit{Granularity of match}. We distinguish two main
-alternatives, element-level and structure-level matching. For each
-element of the fist ontology, an \textit{element-level matcher}
-determines the matching element of the second ontology at the same
-granularity level (usually at the atomic level). For instance, in two
-ontology fragments describing addresses we could find that
-\texttt{{\textquotedblleft}Address.ZIP}{\textquotedblright}
-from one ontology matches
-{\textquotedblleft}\texttt{CustomerAddress.PostalCode{\textquotedblright}}
-from the other one. \textit{Structure level matching}on the other hand
-refers to matching combinations of elements that appear together in a
-structure. This can be a complete, or a partial match. Known
-equivalence patterns can be kept in a library. For example (this
-conforms to a common equivalence pattern), one ontology could have a
-class \texttt{ParttimeEmployee}, which is the subclass of
-\texttt{Employee}, while another could have a class
-\texttt{Employee} having the attribute
-\texttt{IsParttime}. A structure-level matcher should
-identify these as being identical, when the
-\texttt{IsParttime} property equals to true. 
-\end{itemize}
-\begin{itemize}
-\item \textit{Match cardinality}. An element can participate in zero,
-one or many mapping elements of a match result. Moreover, within a
-mapping element, one or more elements from one ontology can match one
-or more elements from the other one. Thus we have the usual
-relationship cardinalities of 1:1, n:1, 1:m and n:m between matching
-elements, both with respect to different mapping elements
-(\textit{global cardinality}) and with respect to an individual mapping
-element (\textit{local cardinality}).\newline
-Most existing matchers map each element of one schema to one element of
-the other schema. This results in local 1:1 mappings and global 1:1 or
-1:n mappings. 
-\end{itemize}
-\begin{itemize}
-\item \textit{Linguistic approaches}. Linguistic matchers use name and
-text (i.e. words or sentences) to find semantically similar elements.
-\textit{Name matching} matches elements with identical or similar names
-(using synonyms, common substrings, user-provided name matching, etc.).
-\textit{Description matching} uses natural language comments of
-ontology elements. For example we can have
-{\textquotedblleft}\texttt{empn //employee
-name}{\textquotedblright} in an ontology matching
-{\textquotedblleft}\texttt{name //name of
-employee}{\textquotedblright} in the other. This can be done by simple
-keyword extraction, or complex natural language processing. 
-\end{itemize}
-\begin{itemize}
-\item \textit{Constraint-based approaches.} These approaches use the
-fact that ontologies often contain constraints on data types and
-ranges, relationships, optionality, uniqueness, cardinalities, etc.
-This information can be used to determine similarities between ontology
-elements. For example, in a candidate pair having the fields
-\texttt{Born} and \texttt{BirthDate}
-respectively, both having the type \texttt{Date}, a
-constraint-based matcher could find these to be similar. 
-\end{itemize}
-\begin{itemize}
-\item \textit{Reusing schema and mapping information}. In addition to
-already mentioned auxiliary information, matchers could well reuse
-common ontology components and previously determined mappings. Some
-mappings occur very often in the same domains, and reusing these
-increases effectiveness. For example, in e-commerce domains
-substructures often repeat with different message formats (like
-address, name, etc.), so we could keep a library of previously found
-results on these, so when we find two elements labelled
-\texttt{ProductOrder} and \texttt{Product} in
-an ontology \texttt{O1}, and \texttt{Porder}
-and \texttt{Article} in an ontology
-\texttt{O2}, the matcher would already know that these
-match (\texttt{ProductOrder} matches
-\texttt{POrder} and \texttt{Product} matches
-\texttt{Article}). 
-\end{itemize}
-
-\paragraph{Instance-based matching.}
-An instance-based matcher \cite{rahm01survey} takes instances
-(i.e. instance data) belonging to the different concepts in the
-ontologies and uses these to discover similarities between the
-concepts. It can be useful when the data is semistructured or
-particularly useful to uncover incorrect interpretations of schema
-information.
-
-The approaches we discuss for instance-level matching primarily work in
-finding element-level matches, because of the complexity of comparing
-large numbers of combinations of instances.
-
-Most of the approaches of schema-level matchers can be applied here, but
-some are especially applicable:
-
-\begin{itemize}
-\item For text elements, a linguistic approach based on information
-retrieval techniques is the preferred approach. For example, if we have
-a field called \texttt{Dept} in an ontology, and two
-fields \texttt{DeptName} and \texttt{EmpName}
-in another, instance information may help to deduce that
-\texttt{DeptName} is the primary match candidate for
-\texttt{Dept}. 
-\item For more structured data, such as numerical or string elements, we
-can apply a constraint-based characterisation, such as numerical value
-ranges or character patterns. For instance, this may help recognising
-phone numbers, postal codes, birth dates, ISBNs, money-related entries
-(e.g. based on currency symbols), etc. 
-\end{itemize}
-The results of instance-based matchers can be used to enhance
-schema-level matchers as a first approach. For instance, a
-constraint-based schema-level matcher can more accurately determine
-data types and ranges for an element using the results of
-instance-based matching. A second approach is to perform instance-level
-matching on its own. Matching elements from the first ontology to the
-second one, and then vice-versa, we can obtain schema-level matches.
-Instance-level matching can also be performed by using auxiliary
-information. This is especially helpful for matching text elements by
-providing match candidates for individual keywords. For example, a
-previous analysis may have revealed that the keyword
-\texttt{{\textquotedblleft}HP{\textquotedblright}}
-frequently occurs for ontology elements
-\texttt{{\textquotedblleft}CompanyName{\textquotedblright}},
-\texttt{{\textquotedblleft}Manufacturer{\textquotedblright}},
-etc. For a new match task, if an \texttt{O2} ontology
-element \texttt{X} frequently contains the term
-\texttt{{\textquotedblleft}HP{\textquotedblright}}, this
-can be used to generate
-\texttt{{\textquotedblleft}CompanyName{\textquotedblright}}
-in \texttt{O1} as a match candidate for
-\texttt{X}, even if
-\texttt{{\textquotedblleft}HP{\textquotedblright}} does
-not often occur in instances of \texttt{O1}.
-
-\subsubsection{Overview of Approaches}
-In this section we give a brief overview of some of the main approaches
-to ontology alignment.
-
-\paragraph{Anchor-PROMPT.}
-Anchor-PROMPT \cite{noy2000anchor_prompt} uses a set of heuristics to analyse
-non-local context, as opposed to algorithms like PROMPT \cite{noy2000prompt} 
-(See Section \ref{subsubsec:mediationapproaches_ontologymerging_PROMPT}) and Chimaera, that only analyse local context.
-It doesn{\textquotesingle}t provide a complete solution, but augments
-existing methods.
-
-The algorithms takes as input two pairs of related terms (anchors) from
-the source ontologies. These anchors are either given by the user, or
-generated automatically with string-based techniques or another matcher
-using linguistic similarity.
-
-From this set of anchors, the algorithm produces a set of new pairs of
-semantically close terms. To do that, Anchor-PROMPT traverses the paths
-between the anchors in the corresponding ontologies. A path follows the
-links between the classes, defined by the hierarchical relations or by
-slots and their domain and range. Anchor-PROMPT then compares these
-terms along the paths to find similar terms. Terms with a high
-similarity score are presented to the user to improve the set of
-possible suggestions, for example, a merging process in PROMPT.
-
-For example, suppose we have ontology \texttt{O1} having
-the classes \texttt{A-C-E-H} in a hierarchical chain, and
-\texttt{O2} having the classes
-\texttt{B-D-F-G} in its hierarchy. Then, suppose we
-identify the anchors as \texttt{A-B} and
-\texttt{G-H}. We then traverse each path in parallel,
-incrementing the similarity score of the classes we encounter at each
-step. We repeat the the process for all existing paths between the
-anchor points, cumulatively aggregating the similarity score.
-
-The central observation behind Anchor-PROMPT is that if two pairs of
-terms are similar and there are paths connecting the terms, the
-elements along those paths are often similar as well.
-
-\paragraph{GLUE.}
-GLUE \cite{doan2004ontology_matching} applies machine learning techniques to
-semi-automatically create semantic mappings between heterogeneous
-ontologies based on instance data. It regards ontologies as being
-taxonomies of concepts, and focuses on finding 1-to-1 correspondences
-between the concepts of the ontologies: for each concept node in one
-taxonomy, find the most similar concept node from the other taxonomy.
-
-Instead of committing to a certain similarity definition, it computes
-the \textit{joint probability distribution} of the concepts involved,
-and lets the application use the joint distribution to compute any
-suitable similarity measure. Specifically, for any two concepts
-\texttt{A} and \texttt{B} the joint
-probability distribution consists of \texttt{P(A,B),
-P(A,{\textlnot}B), P({\textlnot}A,{\textlnot}B)}, and
-\texttt{P({\textlnot}A,B),} where a term such as
-\texttt{P(A,{\textlnot}B)}, is the probability that an
-instance in the domain belongs to concept \texttt{A}, but
-not to concept \texttt{B}.
-
-Computing this joint distribution is based on the sets of instances that
-overlap between the two concepts. A term such as
-\texttt{P(A,B)} can be approximated as the fraction of
-instances that belong to both A and B, that is we need to decide for
-each instance whether it belongs to \texttt{A${\Pi}$B}.
-GLUE addresses this by using the instances of \texttt{A}
-to learn a classifier for \texttt{A}. Then classifies
-instances of \texttt{B} according to that classifier, and
-vice-versa.
-
-\paragraph{Semantic Matching.}
-Semantic Matching \cite{giunchiglia2004semantic_matching} is an implementation of
-the Match operator, that takes two graph-like structures (like database
-schemas or ontologies) and produces a mapping between elements of the
-two graphs that correspond semantically to each other. The authors
-argue that most previous approaches used syntax driven techniques, like
-matching labels(substrings, abbreviations, similar soundex) or
-syntactical structure and calculating a similarity measure, all being
-different variations of syntactic matching, that searches for semantic
-correspondences based on syntactic features.
-
-Semantic matching has the following main features:
-
-\begin{itemize}
-\item searching is done by mapping meanings (concepts), not labels. It
-is not sufficient to consider the meaning of labels in the nodes, but
-also the positions the nodes have in the graph. 
-\item semantic similarity relations are used between elements (concepts)
-instead of syntactical similarity relations. In particular, those
-relations are considered, which relate the extensions of the concepts
-under consideration. 
-\end{itemize}
-The semantics of a node is given by the concept attached to that node
-(the concept denoted by the label of the node), by the position of the
-given node in the graph, and the semantics of all the nodes which are
-higher in the hierarchy.
-
-The possible returned relations between elements are equality,
-overlapping, mismatch, and more general/specific.
-
-\paragraph{QOM.}
-QOM \cite{ehrig2004qom} was designed to create an efficient matching
-tool for on-the-fly creation of mappings between ontologies.
-
-Given that the a major ingredient of run-time complexity is the number
-of mapping pairs which have to be compared to actually find the best
-mappings, QOM uses heuristics to lower the number of candidate
-mappings, thus not comparing all entities of the first ontology to all
-entities of the second ontology. It uses the ontological structures to
-classify candidate mappings into promising and less promising pairs,
-using a dynamic programming approach to refine the set of candidate
-mappings in each iteration.
-
-The similarity measure is computed using a wide range of similarity
-functions, such as string similarity. Several of such similarity
-measures are computed, which are all input to the similarity
-aggregation function, which computes the actual similarity measure. QOM
-then applies a sigmoid function, which emphasises high individual
-similarities and deemphasises low individual similarities. The actual
-correspondences are then extracted using a threshold to the aggregated
-similarity measure.
-
-The output of one iteration of QOM can be used as part of the input of a
-subsequent iteration in order to refine the result. After a number of
-iterations the actual mapping between the ontologies is obtained. 
-\subsubsection{Alignment Examples}
-This section provides a few examples of the matching
-process{\textquotesingle} results.
-\begin{center}
-\tablecaption{Constraint-based matching example. {\newline}Taken from \cite{rahm01survey}.}
-\label{tab:ConstraintMatching}
-\tablehead{}
-%begin{supertabular}{|r@{\hspace{6.5mm}}|r@{\hspace{5.5mm}}|r|r|}
-\begin{supertabular}{|m{0.4\textwidth}|m{0.4\textwidth}|}
-\hline
-S1 elements &
-S2 elements\\\hline
-Employee
-
-\ \ \ \ EmpNo -- int, primary key
-
-\ \ \ \ EmpName -- varchar (50)
-
-\ \ \ \ DeptNo -- int, references
-
-\ \ \ \ Salary -- dec (15,2)
-
-Department
-
-\ \ \ \ Birthdate -- date
-
-\ \ \ \ DeptNo -- int, primary key
-
-\ \ \ \ DeptName -- varchar (40) &
-Personnel
-
-\ \ \ \ Pno - int, unique
-
-\ \ \ \ Pname -- string
-
-\ \ \ \ Dept -- string
-
-\ \ \ \ Born -- date\\\hline
-\end{supertabular}
-\end{center}
-The two schemas (corresponding to ontologies) shown in Table \ref{tab:ConstraintMatching} present
-an example of constraint-based schema-level matching \cite{rahm01survey}. Type and key information suggest that
-\texttt{Born} matches \texttt{Birthdate} and
-\texttt{Pno} matches either \texttt{EmpNo} or
-\texttt{DeptNo} (this could be refined using
-instance-based matching, which could reveal that
-\texttt{Pno} matches \texttt{EmpNo}). By
-using a matcher that detects more than just atomic-level similarities,
-we match \texttt{S2.Personnel }to
-\texttt{S1.Employee} joined with
-\texttt{S1.Department.} This can be detected automatically
-by observing that elements of \texttt{S2.Personnel} match
-elements of \texttt{S1.Employee} and
-\texttt{S1.Department} and that
-\texttt{S1.Employee} and
-\texttt{S1.Departmen}t are connected by the foreign key
-\texttt{DepNo}. This allows us to determine the correct
-SQL-like n:m mapping:
-\begin{verbatim}
-    S2.Personnel (Pno, Pname, Dept, born) ~
-    SELECT S1.Employee.EmpNo,
-           S1.Employee.EmpName,
-           S1.Department.DeptName,
-           S1.Employee.Birthdate
-    FROM S1.Employee, S1.Department
-    WHERE (S1.Employee.DeptNo = S1.Department.DeptNo)
-\end{verbatim}
-
-Table \ref{tab:CardinMatching} \cite{rahm01survey} illustrates match cardinalities for
-two given schemas S1 and S2. When matching multiple elements, we can
-see that expressions are used to combine them.
-
-\begin{center}
-\tablecaption{Matching cardinalities example. {\newline}Taken from \cite{rahm01survey}.}
-\label{tab:CardinMatching}
-\tablehead{}
-\begin{supertabular}{|p{0.20\textwidth}|p{0.20\textwidth}|p{0.20\textwidth}|p{0.20\textwidth}|}
-\hline
-\textbf{Local match cardinalities} & \textbf{S1 element(s)} & \textbf{S2 element(s)} & \textbf{Matching expression} \\\hline 
-1:1, element level & Price & Amount & Amount = Price \\\hline 
-n:1, element-level & Price, Tax & Cost & Cost ={\newline}Price*(1+Tax/100) \\\hline 
-1:n, element-level & Name & FirstName,\newline
-LastName & FirstName,{\newline}LastName =\newline
-Extract (Name, \ldots) \\\hline 
-n:1 structure-level
-(n:m element-level) & B.Title,\newline
-B.PuNo,\newline
-P.PuNo,\newline
-P.Name & A.Book,
-A.Publisher & A.Book, A.Publisher =\newline
-SELECT B.Title, P.Name\newline
-FROM B, P\newline
-WHERE B.PuNo=P.PuNo
-\\\hline
-\end{supertabular}
-\end{center}
-In the first row, the
-match is 1:1, which is the most simple to determine. Row 2 shows how
-\texttt{Cost} can be matched to a formula based on
-\texttt{Price} and \texttt{Tax}. Row 3
-explains how \texttt{FirstName} and
-\texttt{LastName} are extracted from
-\texttt{Name}. Row 4 uses a SQL expression combining
-attributes from two tables. It corresponds to an n:m relationship at
-the attribute level (four \texttt{S1} attributes
-correspond to two \texttt{S2} attributes) and to an n:1
-relationship at the structure level (two tables from
-\texttt{S1} match one table from
-\texttt{S2}).
-
-The example in Table \ref{tab:MappingPattern} \cite{debruin2005wsml} shows a mapping element from the
-motivating example from Section \ref{subsec:mediationproblem_motivatingexample}. Assume we have two ontologies
-\texttt{O1} and \texttt{O2} which both
-describe humans and their gender. Ontology \texttt{O1}
-describes the concept \texttt{Person} with an attribute
-\texttt{hasGender}, which has the two possible values
-{\textquotedblleft}\texttt{male}{\textquotedblright} and
-{\textquotedblleft}\texttt{female}{\textquotedblright}.
-Ontology \texttt{O2} describes the concept
-\texttt{Human}, and its subclasses
-\texttt{Man} and \texttt{Woman} to
-distinguish the gender.
-
-\begin{center}
-\tablecaption{Class by attribute mapping pattern and example. {\newline}Taken from \cite{debruin2005wsml}.}
-\label{tab:MappingPattern}
-\tablehead{}
-\begin{supertabular}{|m{0.8\linewidth}|}
-\hline
-\textbf{Name}: Class by Attribute Mapping \\\hline
-\textbf{Problem}: The extension of a class in one ontology
-corresponds to the extension of a class in another ontology, provided
-that all individuals in the extension have a particular attribute
-value. \\\hline
-\textbf{Solution:}
-
-\textit{Solution description}: A mapping is established
-between a class/attribute/attribute value combination in one ontology
-and a class in another ontology.
-
-\textit{Mapping syntax:}
-
-\texttt{mapping ::= classMapping(direction A
-B attributeValueCondition(P o))} \\\hline
-\textbf{Example:}
-
-\texttt{classMapping(Human Female
-attributeValueCondition(hasGender
-}{\textquotedblleft}female{\textquotedblright})) \\\hline
-\end{supertabular}
-\end{center}
-
-Notice that this is a typical case of a
-conceptual mismatch, namely a mismatch in the style of modeling. The
-solution illustrates an elementary mapping pattern. The pattern is
-described in terms of its name,the problem addressed, the solution of
-the problem, both in natural-language description and in terms of the
-actual mapping language, and an example, namely a mapping between a
-class \texttt{Human} from ontology
-\texttt{O1} to the class \texttt{Woman} from
-ontology \texttt{O2}, but only for humans having the
-gender
-{\textquotedblleft}\texttt{female}{\textquotedblright}.
-
-\subsection{Ontology Merging}
-\label{subsec:mediationapproaches_ontologymerging}
-Ontology merging is the creation of a new ontology from one or more
-source ontologies. The new ontology will unify and in general replace
-the original source ontologies.
-
-We distinguish between two main approaches to ontology merging. In the
-first approach the input of the process is a collection of ontologies
-and the output is a new, merged ontology which captures the original
-ontologies. A prominent example of this is PROMPT \cite{noy2000prompt}. 
-In the second approach the ontologies are not replaced, but
-rather a {\textquotesingle}view{\textquotesingle}, called
-\textit{bridge ontology} is created which imports the original
-ontologies and specifies the correspondences using bridge axioms.
-OntoMerge \cite{dou2002ontology_translation} is an example of this approach, creating a
-bridge ontology that imports the original ontologies and relates the
-overlapping concepts in these ontologies using bridge axioms. We
-describe PROMPT and OntoMerge in more detail below.
-
-\subsubsection{PROMPT}
-\label{subsubsec:mediationapproaches_ontologymerging_PROMPT}
-PROMPT \cite{noy2000prompt} is an algorithm and interactive tool for
-ontology merging and alignment. It was one of the first ontology
-merging tools, comparing every pair in the two ontologies and looking
-for syntactic similarity between labels. It has been implemented as a
-Proteg\'e-2000 extension.
-
-The algorithm of PROMPT defines a number of steps for the interactive
-merging process:
-
-\begin{enumerate}
-\item An initial list of merging candidates is created, based on class
-name similarities. The list is presented to the user as potential
-merging operations. Then the following cycle happens: 
-\item The user chooses the next operation: selects an option suggested
-by the algorithm, or edits the ontology to specify the desired
-operation directly. 
-\item The third step is composed of the following three sub-steps:
-
-\begin{enumerate}
-\item The system performs the requested action and automatically
-executes additional changes derived from the operation. 
-\item A new list of suggested operations is presented to the user, based
-on the structure of the ontology around the arguments of the last
-operation. 
-\item The system determines the conflicts introduced by the last
-operation and finds possible solutions to them. 
-\end{enumerate}
-\end{enumerate}
-PROMPT defines the following set of ontology-merging operations: merge
-classes, merge slots, merge bindings between a slot and a class,
-perform a deep copy of the class (copy all parents up to the root),
-perform a shallow copy of the class.
-
-The conflicts that may appear as a result of these operations are: name
-conflicts, dangling references, redundancy in the class hierarchy,
-slot-value restrictions that violate class inheritance.
-
-The result of the algorithm is a new ontology which replaces the
-original ones.
-
-\subsubsection{OntoMerge}
-OntoMerge \cite{dou2002ontology_translation} is an online ontology translation system,
-which translates a dataset to a new dataset which captures the same
-information in a different ontology.
-
-The approach divides the process into two main parts, namely separating
-the syntactic translation from the semantic translation:
-
-\begin{enumerate}
-\item Converting the datasets into a uniform internal representation (a
-language called Web-PDDL) in order to clear away syntactic differences.
-
-\item Perform the semantic translation from the internal representation
-of a dataset in the source ontology to the internal representation of
-the a dataset in the target ontology. 
-\end{enumerate}
-The result of the process is not an ontology replacing the original ones
-as in PROMPT, but a bridge ontology that imports the source ontologies
-and contains a set of Bridging Axioms, which are translation rules used
-to connect the overlapping parts of the source ontologies. The two
-source ontologies, together with the set of bridging axioms, are then
-treated as a single theory by a theorem prover (optimised for
-operations like dataset translation, ontology extension generation
-based on the extension of a related ontology, and query rewriting).
+\part{D2.4.1: Theoretical Foundations for Ontology Mediation}
+\label{part:d2.4.1}
+%\include{fast_d2_4_2_part2}
 
@@
 \label{sec:summary}
-With the Semantic Web winning more and more ground both in research and
-industry, the mismatches between ontologies used by the inter-operating
-systems present an increasing problem. Ontology mediation deals with
-identifying these mismatches, and overcoming them.
+With the Semantic Web winning more and more ground both in research and industry, the mismatches between ontologies used by the inter-operating systems present an increasing problem.
+Ontology matching deals with identifying these mismatches, and overcoming them.
 
-After presenting the main problem, and the types of mismatches between
-ontologies, we have discussed - based on the most prominent research
-results in these areas - the three main approaches to ontology
-mediation, that is \textit{ontology mapping} (representing
-correspondences between entities), \textit{ontology alignment} (the
-process of finding the correspondences between ontologies) and
-\textit{ontology merging }(the process of creating a new ontology based
+In the first iteration (described in Part \ref{part:d2.4.1} we laid the theoretical foundations of ontology matching.
+After presenting the main problem, and the types of mismatches between ontologies, we have discussed - based on the most prominent research results in these areas - the three main approaches to ontology mediation, that is \textit{ontology mapping} (representing  correspondences between entities), \textit{ontology alignment} (the process of finding the correspondences between ontologies) and \textit{ontology merging }(the process of creating a new ontology based
 on the source ontologies that need to be reconciled).
 
-We will test the various kinds of ontology mediation on FAST, choose a
-suitable candidate, adopt it to our needs and implement it in the
-platform. The choice of mediation approach will be influenced by an
-evaluation of the kinds of ontologies that are used in real-world web
-services.
+In this second iteration we established the approach to integrate ontology matching into FAST, and have shown the implementation details as well as test scenarios.
+We have presented a tool for finding the alignments between ontologies, as well as a powerful language to represent them, beyond the capabilities of the tool.
+We have shown how this matching is done at the different phases of the gadget life cycle and the approach taken at every level.
+
+We plan to evaluate the feasibility of ontology matching in real-life tests and finalize it for the completed version of the FAST platform.
 

### 2010-01-14 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_part1.tex
First initial working draft of D2.4.2
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@419 d8b0c59b-6544-0410-94d9-83a5658a058e

% \begin{abstract}
% The FAST environment allows users to graphically compose intelligent, i.e. semantically annotated gadgets from predefined building blocks and deploy them on various mashup platforms, thus enabling the interconnection of different systems and services. Since different parties involved in the creation of gadgets and building blocks conceptualise knowledge in different ontologies, using ontology matching as a means to reconcile differences in the various conceptualisations is a crucial issue.
% This paper discusses first steps in our effort to integrate ontology matching in an end-user-oriented environment such as FAST, using an e-commerce-related scenario. We evaluate a number of tools and approaches for solving different levels of complexity in ontology matching and 
% %, which we address in this paper. We present a scenario of interconnecting e-commerce web services within a FAST gadget, the ontologies created to represent those services and the alignment tool and language used to generate and represent the ontology mappings. The testing aims firstly, to find out whether and how state-of-the-art ontology matching makes it possible to mash-up different systems and secondly, to 
% define the direction of integrating ontology matching into the FAST environment.
% 
% \keywords{Ontology Matching, User-oriented, Mashups, E-commerce}
% \end{abstract}

\section{Overview}
\subsection{Ontology Matching}
\label{mediation}

FAST uses ontology matching to conceptualise the underlying resources used by the different components.
Ontologies embody the fundamental vehicle for conceptualising data on semantic systems; they describe the context and semantic background of data that should be known to all agents using it. An ontology is used to define the vocabulary that agents use to exchange  queries and assertions. If an agent commits to a common ontology, it guarantees the consistency of the queries and assertions using the defined vocabulary \cite{gruber93towards}.
However, different ontologies are often used in the same scenario. This is also true for FAST, where gadget building blocks can originate from different providers, who might use different ontologies to describe them.  The task of ontology matching is therefore critical in FAST.

Given two ontologies $O$ and $O'$ that need to be mapped to each other, we adopt the definition given in \cite{shvaiko2005schema_based}: an ontology mapping element is a 5-tuple $<id, e, e', n, R>$, where 
%\begin{inparaenum}[(i)]
$id$ is a unique identifier, identifying the
mapping element, 
$e$ and $e'$ are entities (formulas, terms, classes, individuals) of the first and second ontology, respectively, 
$n$ is a confidence measure holding the correspondence value between $e$ and
$e'$, 
$R$ is the correspondence relation holding between $e$ and $e'$ (e.g., \texttt{equivalence (=)}, \texttt{more general($\sqsupseteq$)} or \texttt{disjointness($\perp$)}).
%; \texttt{overlapping($\sqcap$)}. 
%\end{inparaenum}
The alignment operation determines the mapping $M'$ for a pair of ontologies $O$ and $O'$. The alignment process can be extended by parameters, such as an input mapping, weights and thresholds and other external resources (dictionaries, thesauri, etc.). Different levels of mappings are defined:

A \textit{level 0} mapping~\cite{euzenat2004api} is a set of the above mapping elements, when the entities are discreet (defined by URIs). E.g., consider the ontology $O1$ with a class \texttt{Person}, and another ontology $O2$ with a class \texttt{Human}. For this case a matching algorithm  could return the mapping element $< id_{11}, Person, Human, 0.67, = >$, meaning that the \texttt{Person} class from the first ontology is found to be equivalent to the \texttt{Human} class in the second one with a confidence measure of \texttt{0.67}.

A \textit{level 1} mapping is a slight refinement of level 0, replacing pairs of elements with pairs of sets of elements.

A \textit{level 2} mapping can be more complex and defines correspondences in first order logic. It uses the ontology mapping language described in \cite{scharffe2005language}. It can describe complex correspondences, such as the one detailed in Section \ref{gr} for matching the Amazon and GoodRelations ontologies.
% :
% 
% \begin{equation*} \label{eq}
% \forall{x,z} grandparent(x,z) \Longrightarrow \exists{y}; parent(x,y) \wedge parent(y,z)
% \end{equation*}

\subsection{Ontology Matching in FAST}
\label{ominfast}

The gadget life cycle in FAST has several phases and roles associated, as detailed in \cite{hoyer-fast}. Here, we list the ones relevant for the ontology matching tasks, in decreasing order of the measure in which knowledge about ontologies is required.
\begin{inparaenum}[(i)]
    \item The \textit{ontology engineer} creates the ontologies used to annotate services and data. This role also includes the process of ontology matching, either automated or manually, determining if the alignment is feasible and creating so-called \emph{matching operator} building blocks, which are basic elements of the FAST screen building. We also mention the \textit{resource developer}, (who, in some cases, can be the same person as the one fulfilling the ontology engineer role), who uses these ontologies to annotate the the resources created in FAST.
    \item Ontology matching is needed by the \textit{screen developer} at the design-time of a screen (a visual building block of a gadget). Screen developers have a dedicated UI component for building screens, in which they can use the matching operators to combine components annotated with different ontologies. No actual matching is performed in this phase, but rather the simple possibility of matching.
%the FAST user interface determines if two resources can be combined or not (see Figure \ref{fig:screens} for a conceptual example of resources that can not be combined).
    \item The \emph{gadget developer} combines screens to screen-flows and gadgets, and only uses ontology matching implicitly.
    \item The \textit{end-user} uses the final deployed gadget, that combines the different services, but is unaware of the underlying resources and ontologies or the matching process. Thus we also need ontology matching at the gadget's run-time, where the mapping will have to be integrated into the actual JavaScript code of the running FAST gadget. %In this paper, we mainly consider the first two cases (i.e. ontology engineering and screen development), but portability of the alignment code to JavaScript must also be kept in mind.
\end{inparaenum}

\section{Ontology Engineering Phase}

\subsection{Alignment Tool}
\label{alignmenttool}

An ontology mapping (or alignment) is a declarative specification of the semantic overlap between two ontologies~\cite{debruin2005wsml}. It is the result of the ontology alignment process. This mapping is represented as a set of axioms in a mapping language. The mapping process has three main phases:
\begin{inparaenum}[(1)]
    \item discovering the mapping (alignment phase), 
    \item representing the mapping and 
    \item exploiting the mapping.
\end{inparaenum}

To accomplish this we need a tool to assist the ontology engineer in the ontology mapping process.
Based on the description given in Sect.~\ref{ominfast} we identify the following requirements for an alignment tool, that we will take as the basis for an ontology matching component in FAST:
\begin{inparaenum}[(i)]
\item All three phases of the process need to be accessible.
\item Matching of OWL and RDFS ontologies should be supported, since we only consider these two languages.
\item The tool should be as independent as possible, performing the alignment process with little or no user interference. This is an important requirement, since FAST is end-user oriented.
\item It should have an open source license, allowing its free use and modification, because it needs to be integrated into the free and open FAST platform.
\item The code should be suitable for porting to other languages (in particular JavaScript), allowing it to be integrated into the FAST environment.
\item It should be well documented.
\end{inparaenum}

We have considered three tools, two of which we have already detailed in \cite{ambrus2009mediation}.
\paragraph{MAFRA} \cite{maedche2002mafra} supports an interactive and incremental process of ontology mapping. It provides an explicit notion of semantic bridges. This representation is serialisable, portable and independent from the mapped languages. The bridges, however, have been designed to be used within the MAFRA system, and the alignment process needs to be done through the provided GUI.

\paragraph{RDFT}
\cite{omelayenko2002rdft} is a small language originally designed to map between XML and RDF. The results are mappings represented in DAML+OIL, that can be executed in a transformation process. No hints are given to add alignment methods or extending the format and, as a major downside, we have not been able to locate the tool for download and testing.

\paragraph{Alignment API} \cite{euzenat2004api} is the tool best matching the requirements, satisfying all the desired conditions. It also being widely used at the annual ontology matching evaluation campaigns and is under active development. It provides an API and its implementation, is open source (GPLv2 or above) and written in Java, providing an easy way to embed it into other programs. Alignment API can be extended by other representations and matching algorithms, it can be invoked through the command line interface (thus working without user interference) or one of the two available GUI implementations, or it can be exposed as an HTTP server. The tool allows for testing different alignment methods and can generate evaluation results based on a reference alignment. One of the major advantages is that it can generate the mapping result in XSLT, making it possible with the use of a simple tool, or a JavaScript package to transform XML data from one ontology into the other, independently from the alignment system.

\subsection{Scenario Description}
\label{scenario}

The scenario we want to test in FAST is taken from the e-commerce domain, based on the often-met case where a commercial agent aggregates data from other commercial websites. We consider the situation where a user needs a gadget which combines major e-commerce services, allowing to aggregate item lists from all of them in a combined interface.
As examples in our scenario. we consider the two most popular online shopping websites, Amazon and eBay\footnote{Based on the Alexa ranking from http://alexa.com/topsites/category/Top/Shopping}, along with BestBuy, which is very close behind the former two. BestBuy is an interesting case, because it exposes its data in RDF using the GoodRelations (GR) ontology~\cite{hepp-goodrelations}, which has recently gained a lot of popularity. It is therefore one of the first major e-commerce sites to provide semantic metadata.

\begin{figure}
    \centering
        \includegraphics[width=8cm]{images/screens_diagram.png}%[bb=0 0 201 126]
        \caption{The scenario: three components retrieving incompatible data}
    \label{fig:screens}
\end{figure}

Figure~\ref{fig:screens} illustrates our scenario. There are three retrieval components that wrap the different e-commerce sites and provide data according to three different ontologies: the GR ontology, the Amazon ontology and the eBay ontology. Another component displays GR items for display to the user. Out of the box, only the BestBuy retrieval and the display component can be combined. However, if the gadget designer wants to aggregate data from all three services in the display, there will have to be a mapping present between the Amazon and eBay on the one hand, and the GR ontology on the other.

%, say we have a component that accesses the BestBuy service and retrieves RDF data that uses the GoodRelations ontology. In other words, the output of this component is GoodRelations data. Another component takes this data and displays them in a list, i.e. this component only accepts GoodRelations data for display. Since both components use the same ontology, they can be combined. However, now suppose we have two additional components, which access the Amazon and eBay services and retrieve items from there, which are then represented using the Amazon and eBay ontologies, respecitvely. The user wants to list items from all three sources in the same display component. However, since all three ontologies are incompatible, this is not possible out of the box. What is necessary is a mapping between the Amazon and GoodRelations, as well as eBay and GoodRelations ontologies.

%We have thus a case, where we want to aggregate data using three different ontologies, so aligning them to a common ontology is the only possible solution.

\subsection{Ontologies}
\label{ontologies}
We have used three ontologies to annotate underlying resources in FAST, the first of which, namely GoodRelations, is an extensive ontology for e-commerce. The other two, i.e., the Amazon and the eBay ontologies were developed by us as simplified versions of what would be used in the real-life scenarios. Nevertheless, they are sufficient for testing the feasibility of the scenario and the options presented by current ontology alignment methods.

\paragraph{GoodRelations:}
This ontology is aimed at annotating so-called ``offerings'' on the Web, which can be products or services. The ontology is  available under the Creative Commons Attribution 3.0 license and features support for ranges of units, measurements, currencies,  shipping and payments, common business functions (sell, lease, repair, etc.) and international standards (ISO 4217 or UNSPSC) and codes (e.g., EAN or UPC) in the field.

\begin{center}
\lstset{captionpos=b, breaklines=true}
\lstset{frame=single, basicstyle=\scriptsize}
\lstset{caption=Basic GoodRelations data in N3 notation, label=listing_gr}
\lstset{language=XML}
\begin{lstlisting}
:Offering_8794691     a gr:Offering;
  gr:hasPriceSpecification 
      :UnitPriceSpecification_8794691_1;
  gr:includesObject :TypeAndQuantityNode_8794691_1;

:UnitPriceSpecification_8794691_1     a gr:UnitPriceSpecification;
  gr:hasCurrency "USD"^^xsd:string;
  gr:hasCurrencyValue "749.99"^^xsd:float;
  gr:hasUnitOfMeasurement "C62"^^xsd:string;

:TypeAndQuantityNode_8794691_1     a gr:TypeAndQuantityNode;
  gr:amountOfThisGood "1.0"^^xsd:float;
  gr:hasUnitOfMeasurement "C62"^^xsd:string;
  gr:typeOfGood :ProductOrServicesSomeInstancesPlaceholder_8794691 .

:ProductOrServicesSomeInstancesPlaceholder_8794691 a gr:ProductOrServicesSomeInstancesPlaceholder;
  gr:hasEAN_UCC-13 "0013803096095"^^xsd:string;
  gr:hasMakeAndModel :PoSM_8794691;
  rdfs:comment "With Auto Optimization and [...]"@en;

:PoSM_8794691     a gr:ProductOrServiceModel;
  gr:hasManufacturer <bbuy:Manufacturer_Canon>;
  rdfs:label "Canon EOS Digital Rebel [...]"@en;
  rdfs:comment "With Auto Optimization and [...]"@en;
\end{lstlisting}
\end{center}

The ontology (see List.~\ref{listing_gr} for an example of GR data)  is centred on the \texttt{Offering} class,
which represents an announcement by a \texttt{BusinessEntity} to provide a \texttt{ProductOrService} with a given \texttt{BusinessFunction}. It may be constrained in terms of eligible business partner, countries, quantities, and other properties. It is also described by a given \texttt{PriceSpecification}. The super-class for all classes describing products or service types is \texttt{ProductOrService}. This top-level concept has sub-classes representing actual product instances, product models and dummy product placeholders. A product is described by its title and description, manufacturer, make and model, etc. While GoodRelations offers terminology to allow a much higher expressivity, we will restrict the discussion to those terms relevant for our scenario.

\paragraph{Amazon Ontology:}
We have created a small Amazon ontology based on the datatypes supported by the web service exposed by Amazon to third-party agents to use its services. It is a simplified version of what would be needed in a real-world application, but it suffices for the scope of this work. The ontology describes \texttt{Items} based on the \texttt{ItemAttributes} description given in the Amazon Product Advertising API documentation\footnote{http://docs.amazonwebservices.com/AWSECommerceService/latest/DG/}.

\begin{center}
\lstset{captionpos=b, breaklines=true}
\lstset{frame=single, basicstyle=\scriptsize}
\lstset{caption=Simplified Amazon ontology data in N3 notation, label=listing_amazon}
%\caption{Minimal Amazon ontology snippet.}
\lstset{language=XML}
%\lstset{emph={vsmlVariant, namespace, ontology, concept, ofType, instance, nonFunctionalProperties, endNonFunctionalProperties, memberOf, hasValue}}
%\lstset{emphstyle=\textbf\textit\underbar}
\begin{lstlisting}
:Item_7590645     a amzn:Item;
    amzn:hasASIN "B0012YA85A";
    amzn:hasManufacturer :Manufacturer_Canon;
    amzn:hasModel "XSI Kit";
    amzn:hasPrice :Price_7590645_1;
    amzn:hasProductGroup "Electronics";
    amzn:hasTitle "Canon Digital Rebel XSi [...]" .

:Manufacturer_Canon     a amzn:Company;
    amzn:hasLegalName "Canon" .

:Price_7590645_1     a amzn:ListPrice;
    amzn:hasAmount "575.55";
    amzn:hasCurrencyCode "GBP" .
\end{lstlisting}
\end{center}

The ontology features three classes for describing a product. Example instance data is given in List.~\ref{listing_amazon}.
\begin{inparaenum}[(1)]
    \item \texttt{Item} represents an Amazon item, defined by a title, a manufacturer, a product group (DVD, Book, etc.), an international EAN code, an ASIN (unique Amazon id), an author (for books) and a \texttt{ListPrice}. 
    \item \texttt{Company}, described by a  legal name, is used for representing the manufacturer of an \texttt{Item}.
    \item \texttt{ListPrice} has two properties: \texttt{hasCurrencyCode}, representing an ISO 4217 currency code (e.g. GBP or EUR), and \texttt{hasAmount} representing the price in the given currency.
\end{inparaenum}

\paragraph{eBay Ontology:}
The eBay ontology was created based on the eBay Shopping API \footnote{http://developer.ebay.com/DevZone/shopping/docs/CallRef/index.html} and is supposed to annotate data retrieved through the web service described by the given API. Like the Amazon ontology, it is a simplified case of what would be used in a real-world implementation.

\begin{center}
\lstset{captionpos=b, breaklines=true}
\lstset{frame=single, basicstyle=\small}
\lstset{caption=Simplified eBay ontology data in N3 notation, label=listing_ebay}
\lstset{language=XML}
\begin{lstlisting}

:SimpleItem_1320648     a ebay:SimpleItem;
  ebay:hasItemID "E012Y090912";
  ebay:hasBidCount :"4";
  ebay:hasCountry "UK";
  ebay:hasCurrentPrice :Price_1320648_1;
  ebay:hasPrimaryCategoryName "Electronics";
  ebay:hasTitle "Canon Digital Rebel XSi [...]" .
  ebay:hasDescription "Camera is practically unused. It's like new."

:Price_1320648_1     a ebay:CurrentPrice;
  ebay:hasAmount "575.55";
  ebay:hasAmountType "GBP" 
\end{lstlisting}
\end{center}

The ontology features three basic classes, (see Listing \ref{listing_ebay} for an example of ebay data)
\begin{inparaenum}[(1)]
    \item \texttt{SimpleItem} represents an eBay \texttt{Item}, that is sold by a \texttt{SimpleUser}. It is described by a title, a \texttt{CurrentPrice} (specifying the highest bid, or the selling price of fix-priced items), primary category name, manufacturer, model, EAN code, item ID (a unique eBay ID), bid count, end time of bid, country where the item is located, and a product ID (which supports major international product codes --- this property is from the Finding API).
    \item The \texttt{CurrentPrice} features a \texttt{hasAmountType} property, specifying the currency code, and a \texttt{hasAmount} property, which is the amount of money for a price per unit.
    \item \texttt{SimpleUser} contains information about eBay users. Users are described by a user ID, about me URL and the seller's positive feedback score. This class will not be used for capturing information on goods for our scenario, but is an essential component of the eBay system, which was the reason for its inclusion in the ontology.
\end{inparaenum}

\subsection{Testing and Results}
\label{testing}

We present the approach an ontology engineer has to take to discover and represent ontology mappings, and a means to exploit them after the they have been discovered and appropriately represented.

There is a major paradigm difference between the GoodRelations ontology and the other two ontologies (see Sect.~\ref{results} for details). After studying the approach of the GR ontology, and examining the matching results between the Amazon and GR ontologies, we have concluded that automated matching between the two is unfeasible using the string-based methods employed using the Alignment API tool in this testing. Therefore, we have have created a manual mapping for this case, and have tested the automated matching on the Amazon--eBay pair, which closely resemble each other in their approach to capturing product details.

\subsubsection{The Matcher}
The matching tool called Alignment API \cite{euzenat2004api} has been developed to support several kinds of algorithms and methods, and provide an efficient environment for testing alignment systems.

\paragraph{The Algorithm:}
We have used a simple string distance-based algorithm provided by the tool, which computes the string distance between the names of the entities to find correspondences between them. Four methods have been used for computing the distance:
\begin{inparaenum}[(1)]
    \item equality, which tests whether the names are identical, 
    \item Levenshtein distance (number of character operations needed), 
    \item SMOA distance (which is a specialised distance for matching ontology identifiers) and
    \item a Wordnet-based~\cite{fellbaum1998wordnet} distance using the JWNL library with Wordnet.
\end{inparaenum}

\paragraph{The Alignment Description:} 
The alignment description is given based on a simple vocabulary, containing a pair of ontologies and a set of correspondences, which express relations holding among entities of the two ontologies.

We used the level 0 mapping representation for representing simple mappings, which map discrete entities of the two ontologies. Thus the representation of the correspondences is given with the five elements described (with the \texttt{id} being optional), as shown in List.~\ref{listing_correspondence}. Level 2 mappings were also used for more complex, manually-created representations, as detailed in Sect.~\ref{results}.

\begin{center}
\lstset{frame=single, basicstyle=\small}
\lstset{caption=Level 0 mapping element example, label=listing_correspondence}
\lstset{captionpos=b}
\lstset{language=XML}
%\lstset{emph={vsmlVariant, namespace, ontology, concept, ofType, instance, onFunctionalProperties, endNonFunctionalProperties, memberOf, hasValue}}
%\lstset{emphstyle=\textbf\textit\underbar}
\begin{lstlisting}
[ a :Cell;
  :entity1 amzn:hasCurrencyCode;
  :entity2 ebay:hasAmountType;
  :measure "1.0"^^xsd:float;
  :relation "=" ]
\end{lstlisting}
\end{center}

\paragraph{Using the Tool:}
The Alignment API tool can be used through a GUI, as a server or from the command-line interface, of which we have chosen the last one. The tool reads two RDF/OWL ontologies, computes the alignment between them, performs some thresholding and displays the results. It can render the output in a number of formats, including HTML and XSLT.

An additional feature of the tool is its ability to evaluate results based on a reference alignment, and output the evaluation results in a table, or plot them as \LaTeX{} graphs.

\subsubsection{Testing Procedure}
We set up the ontologies according to Sect.~\ref{ontologies}. For the Amazon--eBay pair we set up a reference alignment, against which the results are evaluated.
%For the reason of excluding the Amazon - GoodRelations pair from this procedure please refer to Section \ref{results}.

\begin{center}
\lstset{frame=single, basicstyle=\small}
\lstset{caption=Executing the alignment tool on a pair of ontologies, label=listing_cli}
\lstset{language=bash}
\lstset{captionpos=b}
\begin{lstlisting}
 $ java -jar lib/procalign.jar amazon.owl ebay.owl 
     -o results/equal.rdf
\end{lstlisting}
\end{center}

Using the tool's command line interface, we have performed the following steps:
\begin{itemize}
\item generate the RDF alignment output for the four methods: (1) equality, (2) Levenshtein distance with a confidence threshold of \texttt{0.33} (meaning that any correspondence having a smaller confidence measure will be excluded), (3) SMOA distance with a threshold of \texttt{0.5} and (4) Wordnet distance using a threshold of \texttt{0.5}.
\item render the XSLT file for the SMOA distance function;
\item transform an example dataset using the XSLT transformation rules;
\item generate the tables and graphs displaying different parameters of the evaluation.
\end{itemize}

\subsubsection{Results}
\label{results}

\paragraph{GoodRelations:}
\label{gr}

The GoodRelations ontology employs a unique paradigm, different from the paradigms of Amazon and eBay. In GR everything is centred around an instance of \texttt{Offering} and a graph of other instances attached to it, whereas for Amazon (and similarly for eBay), the main class is \texttt{Item}, which holds all relevant properties. In principle, \texttt{Item} would correspond to \texttt{ProductOrService} in GoodRelations, but the properties of the \texttt{Item} class are reflected as properties of many different classes in GR.

Though the infeasibility of automating this alignment became obvious, we have represented the alignment in the mapping language supported by the tool, as a level 2 mapping (described in Sect. \ref{mediation}) (the mapping has been created between Amazon and GoodRelations, since the eBay and Amazon ontologies are very similar in approach, and mapping eBay to GoodRelations would be an indentical task). This mapping description can later be used by the run-time JavaScript code. List.~\ref{listing_gr_a} shows an example mapping between two properties of the two ontologies, specifying that the relationship is \texttt{Equivalence} with a certainty degree of \texttt{1.0}. This fragment does not show, but assumes the equivalence correspondence between the classes \texttt{Item} and \texttt{Offering}, which is a trivial level 0 mapping. This mapping specifies the relation

\begin{equation*} \label{eq}
\begin{split}
\forall{v,z}; hasEAN(v,z) & \Longrightarrow \exists{x, y}; includesObject(v,x) \wedge \\
&typeOfGood(x,y) \wedge hasEAN\_UCC\_13(y,z),
\end{split}
\end{equation*}

meaning that the \texttt{hasEAN} property of \texttt{v} in the Amazon ontology corresponds to the \texttt{hasEAN\_UCC\_13} property of the \texttt{typeOfGood} of the \texttt{includesObject} of \texttt{v} in GoodRelations. The domains and ranges of the properties are inferred, thus it is deduced, that in \texttt{Amazon v} is of type \texttt{Item} and \texttt{z} is \texttt{int}, and in \texttt{GoodRelations v, x, y} and \texttt{z} are instances of the classes \texttt{Offering}, \texttt{TypeAndQuantityNode}, \texttt{ProductOrService} and \texttt{int}, respectively.

\begin{center}
\lstset{frame=single, basicstyle=\small}
\lstset{caption=Fragment of the Amazon--GoodRelations mapping, label=listing_gr_a}
\lstset{language=bash}
\lstset{captionpos=b, breaklines=true}
\begin{lstlisting}
<Cell rdf:about="MappingRule_01">
 <entity1><omwg:Property rdf:about="&amzn;hasEAN"/></entity1>
 <entity2>
  <omwg:Property>
    <first><Relation rdf:about="&gr;includesObject"/></first>
    <next><Relation rdf:about="&gr;typeOfGood"/></next>
    <next><Property rdf:about="&gr;hasEAN_UCC_13"/></next>
  </omwg:Property>
 </entity2>
 <measure rdf:datatype="&xsd;float">1.0</measure>
 <relation>Equivalence</relation>
</Cell>
\end{lstlisting}
\end{center}

Using this representation, complex correspondences can be modelled, using first order logic constructs. We will further investigate the possibilities of converting the level 2 mapping to XSLT, so it can automatically be used in the final JavaScript implementation. In case this is not possible, we will represent the mapping elements in SPARQL.

\paragraph{Alignment Results:}
The results of automatically aligning the Amazon and eBay ontologies were quite favourable. As shown in Tab.~\ref{table_results}, we captured the four main parameters used in information retrieval, as described in \cite{olson2008advanced}. These four parameters are used for evaluating the performance of the alignment methods: 
\begin{inparaenum}[(1)]
    \item \textit{Precision}, the fraction of results that are correct --- the higher, the better, 
    \item \textit{Recall}, the ratio of the correct results to the total number of correct correspondences --- the higher, the better, 
    \item \emph{Fallout}, the fraction of incorrect results - the lower the better, and 
    \item \emph{F-measure}, which measures the overall effectiveness of the retrieval by a harmonic mean of precision and recall --- the higher, the better.
\end{inparaenum}

\begin{center}
\begin{table}
\caption{Alignment results: Precision, Recall, Fallout and F-Measure}
\label{table_results}

\begin{center}
\begin{tabular}[c]{|llll|llll|llll|} \hline
\multicolumn{4}{|l|}{Reference} & \multicolumn{4}{|l|}{Equal} & 
\multicolumn{4}{|l|}{SMOA}\\ \hline
Prec & Rec & Fall & FMeas & Prec & Rec & Fall & FMeas & Prec & Rec & 
Fall & FMeas  \\ \hline
1.00& 1.00& 0.00& 1.00& 1.00& 0.38& 0.00& 0.55& 0.43& 0.75& 0.57& 0.55 \\ \hline
\end{tabular}
\end{center}

\begin{center}
\begin{tabular}{|llll|llll|} \hline
\multicolumn{4}{|l|}{Levenshtein} & \multicolumn{4}{|l|}{JWNL} \\ \hline
 Prec & Rec & Fall & FMeas & Prec & Rec & Fall & FMeas \\ \hline
0.4& 0.75& 0.6& 0.52& 0.67& 0.75& 0.33& 0.71 \\ \hline
\end{tabular}
\end{center}

\end{table}
\end{center}

The first column (Reference) shows the reference alignment, which, naturally, has both perfect precision and recall. We can observe what intuition has predicted, namely that pure string equality (Equal) is far too simple and irrelevant, by only taking identical labels. By using string distances and giving certain thresholds (Levenshtein and SMOA), we can see that the results are much less precise, but have a better recall, since this allows for entities having similar names to be discovered, at the expense of having quite a few incorrect results (lower precision); the thresholds allow for low-scored cases to be eliminated, although this results in the exclusion of some correct correspondences. The last column (JWNL) contains the results of the Wordnet-enabled method, which shows quite an improvement, due to the lexical analysis, which performs a much more relevant comparison of strings, giving a high number of correct results. We can clearly see how it is quite close to the reference alignment, having a correctness of 0.67 and a recall of 0.75. The precision of the JWNL alignment shows only a tiny drop below the recall value, meaning that the number of incorrect correspondences discovered is small, and the main source of error is from the number of correspondences not discovered.

% \begin{center}
% \begin{figure}
% %% Plot generated by GenPlot of alignapi
% \begin{center}
% \begin{tikzpicture}[cap=round, scale=0.75]
% % Draw grid
% \draw[step=1cm,very thin,color=gray] (-0.2,-0.2) grid (10.0,9.0);
% \draw[|-|] (-0,0) -- (10,0);
% %\draw[dashed,very thin] (0,0) -- (5,8.66) -- (10,0);
% \draw[dashed,very thin] (10,0) arc (0:60:10cm);
% \draw[dashed,very thin] (0,0) arc (180:120:10cm);
% \draw (0,-0.3) node {$precision$};
% \draw (10,-0.3) node {$recall$};
% % Plots
% \draw plot[mark=+,] coordinates {(5.0,8.660254037844386)};
% \draw (5.01,8.370254037844385) node[anchor=south west] {Reference};
% \draw plot[mark=+,] coordinates {(9.296875,3.6834922606644636)};
% \draw (9.306875,3.3934922606644633) node[anchor=south west] {Equal};
% \draw plot[mark=+,] coordinates {(3.1058673469387754,2.9531229168449795)};
% \draw (3.115867346938775,2.6631229168449793) node[anchor=south west] {SMOA0.5};
% \draw plot[mark=+,] coordinates {(2.9875000000000003,2.6598578439458005)};
% \draw (2.9975,2.3698578439458003) node[anchor=south west] {Levenshtein0.33};
% \draw plot[mark=+,] coordinates {(4.409722222222222,4.99987943527481)};
% \draw (4.419722222222222,4.70987943527481) node[anchor=south west] {JWNL};
% 
% \end{tikzpicture}
% \end{center}
% \caption{Comparison of the four methods, against the reference alignment,\newline
% showing the distances from 0 precision and 0 recall}
% \label{fig:figure_alignment}
% \end{figure}
% \end{center}

We can deduce that the results provided are satisfactory, even though the methods used were simple, string-based ones, and the process was completely automated without any user input.
This means that through some user assistance or an initial input alignment the tool can achieve 100\% correct results.

\paragraph{XSLT:}
One of the appealing features of the Alignment API is that it can render the results in XSLT, thus specifying a set of rules which make it possible to transform an  XML file from one format to the other. We have generated such an XML transformation description, and with the simple Linux application \texttt{xsltproc} (run from the command line) we have managed to transform data (limited, of course by the incompleteness of the alignment) from the Amazon ontology into data annotated by the eBay vocabulary. 
%Please refer to Appendix A for a listing of an example transformation.

\section{Gadget Building Phase}
Every time ontology matching is performed and an alignment is created between two underlying resources, the corresponding alignment rules are stored and an operator is created in the catalogue. At the gadget building phase, whenever two screens are combined, the catalogue will provide the possible matching operators for the two screens. This way an operator can be placed between two screens that would otherwise be incompatible.

\section{Runtime} 
Employing alignment rules in the runtime phase of the gadget.

### 2010-01-14 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_part2.tex
First initial working draft of D2.4.2
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@419 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2009-12-11 trunk/deliverables/D2_4_Mediation/fast_d2_4_mediation.tex
@@
-\documentclass{fast_latex}
-% yay! We can use any kind of funky diacritic:
-\usepackage[utf8]{inputenc}
-
-\usepackage{lastpage}
-\usepackage{setspace}
-\usepackage{graphicx}
-\usepackage[pdfborder={0 0 0}]{hyperref}		% turn on when latex is used (not miktec)
-\usepackage{url} % LEO: urls \url{}
-\usepackage{verbatim} % code and comment
-\usepackage{longtable}
-\usepackage{xspace}   % whitespace after a macro if no punctuation after the macro
-\usepackage{multirow}
-\usepackage{colortbl}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{textcomp} %vertical apostrophe
-\usepackage{listings} %source code listings
-\usepackage{amssymb} %math symbols
-\usepackage{graphicx} %image placement
-\usepackage{supertabular}
-\parindent0pt
-
-\newcommand\deliverableNumber{D2.4.1}
-\newcommand\deliverableTitle{Mediation amongst ontologies: Application to the FAST ontology}
-\newcommand\deliverableTitleShort{Mediation amongst ontologies}
-\newcommand\workpackageNumber{2}
-\newcommand\workpackageTitle{Definition of Conceptual Model}
-\newcommand\authorOne{Oszkar Ambrus, NUIG}
-\newcommand\authorTwo{Author 2}
-\newcommand\authorThree{Author 3}
-\newcommand\authorFour{Author 4}
-\newtheorem{example}{\emph{Example}}
-
-\begin{document}
-% explicit hyphenations
-\hyphenation{RDF-Re-po-si-to-ry}
-\hyphenation{name-space}
-\hyphenation{So-cket-A-dap-ter}
-
-%\fontfamily{tahoma}\selectfont
-\def\note#1{\marginpar{\footnotesize#1}} % use this to show the notes in the document
-%\def\note#1{} % use this to hide the notes
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% TITLE PAGES 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\thispagestyle{empty}
-
-\pagenumbering{roman}
-
-\begin{flushright}
-	\includegraphics[width=3cm]{images/FP7_logo.png}
-\end{flushright}
-
-\vspace{1cm}
-
-%\begin{minipage}[p]{15cm}
-	\begin{center}
-		\includegraphics{images/FAST_logo}\\
-		\vspace{1cm}
-		{\LARGE{\sffamily \emph{FAST AND ADVANCED STORYBOARD TOOLS}}}\\
-		\vspace{0.5cm}
-		{\LARGE \sffamily \emph{FP7-ICT-2007-1-216048}}\\
-		\vspace{0.5cm}
-		{\LARGE \sffamily \emph{http://fast.morfeo-project.eu}}\\
-		\vspace{4cm}
-		{\LARGE \sffamily \textbf{Deliverable \deliverableNumber}}\\
-		\vspace{0.5cm}
-		{\LARGE \sffamily \textbf{\deliverableTitle}}\\
-		\vspace{2cm}
-		{\large \sffamily \authorOne}\\
-		\vspace{0.5cm}
-		\vfill
-		{\large \sffamily Date: 27/02/2009}\\
-		\vspace{1cm}
-		{\sffamily FAST is partially funded by the E.C. (grant code: FP7-ICT-2007-1-216048).}
-
-	\end{center}
-%\end{minipage}
-
-\clearpage
-%%%%%%%%%%%%%%
-% NEXT PAGES %
-%%%%%%%%%%%%%%
-\pagestyle{scrheadings}
-
-\lohead{\includegraphics[width=4cm]{images/FAST_logo_transparent}}
-%\cohead{\small\textcolor{fast@lightgrey}{\deliverableTitle}}
-%\rohead{\small{\today}}
-%\lofoot{\small\textcolor{fast@lightgrey}{Task Force Ontologies}}
-\cofoot{\small{FAST 216048 --- \deliverableTitleShort}}
-\rofoot{\small{\thepage}}
-
-\newpage
-%
-%\section*{Authors}
-%\authorOne \\
-%%\authorTwo \\
-%%\authorThree\\
-%%\authorFour\\
-%
-%
-%\vfill
-%%\section*{Project Co-ordinator}
-%%\hspace*{-2,5cm}\begin{minipage}[p]{14cm}
-%%Dr. Ansgar Bernardi \\
-%%German Research Center for Artificial Intelligence (DFKI) GmbH \\
-%%Trippstadter Strasse 122 \\
-%%D 67663 Kaiserslautern \\
-%%Germany \\
-%%Email: bernardi@dfki.uni-kl.de, phone: +49 631 205 3582, fax: +49 631 205 4910 \\
-%%\end{minipage}
-%
-%
-%\section*{Partners}
-%Telef\'onica Investigaci\'on y Desarrollo (Telef\'onica I+D), Spain \\
-%National University of Ireland, Galway (NUIG), Republic of Ireland \\
-%Kassel University, Germany \\
-%Cyntelix Corporation, Republic of Ireland \\
-%SAP AG, Germany \\
-%Universidad Polit\'ecnica de Madrid (UPM), Spain
-%
-%
-%\vspace{0.3cm}
-%\begin{footnotesize}
-%Copyright: FAST Consortium 2009\\
-%\end{footnotesize}
-
-%\clearpage
-
-\section*{Version History}
-
-\begin{small}
-\begin{tabular}{|l|l|l|p{7.5cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Rev. No.}} &
-                            \textcolor{white}{\textbf{Date}} &
-                            \textcolor{white}{\textbf{Author (Partner)}} &
-							\textcolor{white}{\textbf{Change description}}\\ \hline
-1.0 & 27.02.2009 & Oszkar Ambrus (NUIG) & final version ready for external review \\ \hline
-\end{tabular}
-\end{small}
-
-\color{black}
-
-\vfill
-%{\bf Explanations of abbreviations on front page}\\
-%\\
-%%Nature \\
-%R: Report \\
-%P: Prototype \\
-%R/P: Report and Prototype \\
-%O: Other \\
-% \\
-%Dissemination level \\
-%PU: Public \\
-%PP: Restricted to other FP6 participants \\
-%RE: Restricted to specified group \\
-%CO: Confidential, only for NEPOMUK partners \\
-
-\newpage
-
-%%%%%%%%%%%%%%%%%%%%%
-% Executive Summary %
-%%%%%%%%%%%%%%%%%%%%%
-
-\clearpage
-
-\section*{Executive Summary}
-\doublespacing
-Ontology mediation is a critical problem in heterogeneous environments,
-since inter-operating services and agents often use different
-ontologies, and the differences between them need to be determined and
-overcome to allow for a seamless data exchange between these parties.
-This is a difficult ask, with which web-based agents must cope.
-
-We review current work and literature, discussing the main areas of the
-topic, that is the core problem of ontology mediation, the types of
-mismatches that occur between ontologies, and the three main ontology
-mediation approaches, namely (1) the representation of correspondences
-between ontologies, (2) the (semi-)automatic process of discovering
-these correspondences and (3) creation of a new ontology from the
-source ontologies.
-
-By the overview of current research in the field of ontology mediation,
-we lay the basis of the implementation of ontology mediation in FAST.
-
-\newpage
-
-%%%%%%%%%%%%%%%%%%%%%
-% Document Summary %
-%%%%%%%%%%%%%%%%%%%%%
-
-\clearpage
-
-\section*{Document Summary}
-% double spacing from here on:
-\singlespacing
-\begin{small}
-
-\begin{tabular}
-	%{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{6cm}|p{2cm}|p{2cm}|}
-	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{6cm}|p{2cm}|p{2cm}|}
-	\hline
-	\textcolor{white}{\textbf{Code}} & {FP7-ICT-2007-1-216048} & {\textbf{Acronym}} & {FAST}\\ \hline
-	\textcolor{white}{\textbf{Full title}} & \multicolumn{3}{l|}{Fast and Advanced Storyboard Tools}\\ \hline
-	\textcolor{white}{\textbf{URL}} & \multicolumn{3}{l|}{\url{http://fast.morfeo-project.eu}}\\ \hline
-	\textcolor{white}{\textbf{Project officer}} & \multicolumn{3}{l|}{Annalisa Bogliolo}\\ \hline
-\end{tabular}
-
-\vspace{0.5cm}
-
-\begin{tabular}
-	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.25cm}|p{1cm}|p{1cm}|p{6.32cm}|}
-	\hline
-	\textcolor{white}{\textbf{Deliverable}} & {\textbf{Number}} & {\deliverableNumber} & {\textbf{Name}} & {\deliverableTitle}\\ \hline
-	\textcolor{white}{\textbf{Work package}} & {\textbf{Number}} & {\workpackageNumber} & {\textbf{Name}} & {\workpackageTitle}\\ \hline
-\end{tabular}
-
-\vspace{0.5cm}
-
-\begin{tabular}
-	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
-	\hline
-	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2009} & {\textbf{Submitted}} & {27/02/2009}\\ \hline
-	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
-	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
-	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.4.1 is the first iteration of the FAST ontology mediation approach.
-This document contains a general overview of current literature regarding ontology mediation, paving the way for futher analysis of the scenarios arising in FAST that require ontology mediation.
-Based on this general overview, the implementation of ontology mediation within FAST can be researched, which will constitute the topic of the next iteration of this document.}\\ \hline
-	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
-%	{} & \multicolumn{4}{l|}{}\\ 
-%	{} & \multicolumn{4}{l|}{}\\ 
-%	{} & \multicolumn{4}{l|}{}\\
-	\hline
-	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {oszkar.ambrus@deri.org} \\ \cline{4-5}
-	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495112} \\ \hline
-	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, ontology mediation, ontology matching, ontology alignment, ontology merging}\\ \hline
-\end{tabular}
-\end{small}
-\newpage
-
-%%%%%%%%%%%%%%%%%%%%%
-% TABLE OF CONTENTS %
-%%%%%%%%%%%%%%%%%%%%%
-\doublespacing
-\setcounter{tocdepth}{3}
-\tableofcontents
-\cleardoublepage
-\pagenumbering{arabic}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-% BEGINNING OF SECTIONS %
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\clearpage
-\rofoot{\small{Page \thepage\ of \pageref{LastPage}}}
-
-\section{Introduction} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\label{sec:introduction}
-\paragraph{Goal and Scope.}
-The central driver for Enterprise Mashups is the lightweight resource
-composition style to build enterprise applications with reusable
-building blocks, called Gadgets. The architecture used for this
-front-end layer of Gadgets specifies a back-end layer of Services,
-designed to support machine-to-machine interaction over a network (e.g.
-Web Services, Databases, etc.). Having this case, the communication
-problems need to be addressed, that arise from disparate interfaces,
-different invocation mechanisms and even different conceptualisations
-of the data transmitted between the front-end and the back-end. This
-document deals with the third issue, namely the differences between the
-conceptualisations of the exchanged data.
-
-As in all fields employing semantic web paradigms, data is annotated
-using ontologies. Ontologies conceptualise the data used, describe
-their context and their semantic background that should be known to all
-agents using that data. Because we do not the have the situation where
-the same ontologies are used by all sides, having different ontologies
-and even different languages to describe them, we need to integrate the
-ontologies present in the back-end with the ones present in the
-front-end, and to find a way to mediate between the the different
-ontologies used in the two layers.
-
-We try to find a solution to these differences which range from simple
-naming differences to more complex data heterogeneity, behavioural
-differences, discrepant paradigms and other disparities. The
-reconciliation of these differences is called ontology mediation (often
-mentioned in the specialised literature as ontology matching, or
-ontology alignment, thus we will sometimes use these terms as well). 
-\paragraph{Structure of the Document.}
-We discuss ontologies in general in Section \ref{sec:ontologies}, present the ontology
-mediation problem in Section \ref{sec:mediationproblem} , discuss the types of mismatches between
-ontologies in Section \ref{sec:ontologymismatches} and detail the main approaches to ontology
-mediation in Section \ref{sec:mediationapproaches}.
-\paragraph{Overall Approach for WPX and relation to other Work Packages and Deliverables.}
-Ontology mediation will play an important role for the integration of 
-external ontologies into the FAST ontology infrastructure.
-By giving a detailed overview of the state-of-the-art in ontology
-mediation, we lay the groundwork for further research in T2.4 [Task
-2.4, see Project Description] with this first iteration of D2.4. Future
-versions of the deliverable (i.e., D2.4.2 and D2.4.3) will implement
-ontology mediation within the FAST platform.
-
 \section{Ontologies}
@@
 based on the extension of a related ontology, and query rewriting).
-
-\section{Summary}
-\label{sec:summary}
-With the Semantic Web winning more and more ground both in research and
-industry, the mismatches between ontologies used by the inter-operating
-systems present an increasing problem. Ontology mediation deals with
-identifying these mismatches, and overcoming them.
-
-After presenting the main problem, and the types of mismatches between
-ontologies, we have discussed - based on the most prominent research
-results in these areas - the three main approaches to ontology
-mediation, that is \textit{ontology mapping} (representing
-correspondences between entities), \textit{ontology alignment} (the
-process of finding the correspondences between ontologies) and
-\textit{ontology merging }(the process of creating a new ontology based
-on the source ontologies that need to be reconciled).
-
-We will test the various kinds of ontology mediation on FAST, choose a
-suitable candidate, adopt it to our needs and implement it in the
-platform. The choice of mediation approach will be influenced by an
-evaluation of the kinds of ontologies that are used in real-world web
-services.
-
-\clearpage
-\bibliographystyle{apalike}
-\addcontentsline{toc}{section}{References}
-\bibliography{fast_d2_4_mediation_bib}
-
-\clearpage
-\doublespacing
-\section*{Appendix A (Lists of Tables and Figures)}
-\addcontentsline{toc}{section}{Appendix A (Lists of Tables and Figures)}
-
-\listoftables
-
-\listoffigures
-
-\end{document}

### 2010-01-21 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_part1.tex
Added newest version of USGS article contents
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@442 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
+\lstset{
+captionpos=b, 
+extendedchars=true,
+basicstyle=\scriptsize \sffamily, 
+stringstyle=\bfseries,
+frame=single,
+frameround=tttt,
+showstringspaces=false,
+breaklines=true,
+}
+
+\lstdefinelanguage{turtle} 
+{morekeywords={@prefix, a}, 
+sensitive=false, 
+morecomment=[l]{\#}, 
+morestring=[b]",
+}
+
 % \begin{abstract}
@@
 
-FAST uses ontology matching to conceptualise the underlying resources used by the different components.
-Ontologies embody the fundamental vehicle for conceptualising data on semantic systems; they describe the context and semantic background of data that should be known to all agents using it.
-An ontology is used to define the vocabulary that agents use to exchange  queries and assertions.
-If an agent commits to a common ontology, it guarantees the consistency of the queries and assertions using the defined vocabulary \cite{gruber93towards}.
-However, different ontologies are often used in the same scenario.
+FAST uses ontologies to conceptualise the underlying resources used by the different components.
+Ontologies embody the fundamental vehicle for conceptualising data on semantic systems; they describe the context and semantic background of data that should be known to all agents using it~\cite{gruber93towards}.
+However, different ontologies are often used to describe the same domain or cover the same scenario.
 This is also true for FAST, where gadget building blocks can originate from different providers, who might use different ontologies to describe them.
-The task of ontology matching is therefore critical in FAST.
+To ensure interoperability, the task of ontology matching is therefore critical in FAST.
 
-Given two ontologies $O$ and $O'$ that need to be mapped to each other, we adopt the definition given in \cite{shvaiko2005schema_based}: an ontology mapping element is a 5-tuple $<id, e, e', n, R>$, where
+Given two ontologies $O$ and $O'$ that need to be mapped to each other, we adopt the definition given in~\cite{shvaiko2005schema_based}: an ontology mapping element is a 5-tuple $<id, e, e', n, R>$, where
 %\begin{inparaenum}[(i)]
@@
 
-A \textit{level 0} mapping~\cite{euzenat2004api} is a set of the above mapping elements, when the entities are discreet (defined by URIs).
+\begin{inparaenum}[(a)]
+    \item A \textit{level 0} mapping~\cite{euzenat2004api} is a set of the above mapping elements, when the entities are discreet (defined by URIs).
 E.g., consider the ontology $O1$ with a class \texttt{Person}, and another ontology $O2$ with a class \texttt{Human}.
 For this case a matching algorithm  could return the mapping element $< id_{11}, Person, Human, 0.67, = >$, meaning that the \texttt{Person} class from the first ontology is found to be equivalent to the \texttt{Human} class in the second one with a confidence measure of \texttt{0.67}.
-
-A \textit{level 1} mapping is a slight refinement of level 0, replacing pairs of elements with pairs of sets of elements.
-
-A \textit{level 2} mapping can be more complex and defines correspondences in first order logic.
-It uses the ontology mapping language described in \cite{scharffe2005language}.
-It can describe complex correspondences, such as the one detailed in Section \ref{gr} for matching the Amazon and GoodRelations ontologies.
+    \item A \textit{level 1} mapping is a slight refinement of level 0, replacing pairs of elements with pairs of sets of elements.
+    \item A \textit{level 2} mapping can be more complex and defines correspondences in first order logic.
+It uses the ontology mapping language described in~\cite{scharffe2005language}.
+It can describe complex correspondences, such as the one detailed in Section \ref{sub:manual_mapping}.
+\end{inparaenum}
 % :
@@
 Here, we list the ones relevant for the ontology matching tasks, in decreasing order of the measure in which knowledge about ontologies is required.
+Note that several roles can be played by the same actor.
 \begin{inparaenum}[(i)]
@@
 This role also includes the process of ontology matching, either automated or manually, determining if the alignment is feasible and creating so-called \emph{matching operator} building blocks, which are basic elements of the FAST screen building.
-We also mention the \textit{resource developer}, (who, in some cases, can be the same person as the one fulfilling the ontology engineer role), who uses these ontologies to annotate the the resources created in FAST.
+The \textit{resource developer} then uses these ontologies to annotate resources created in FAST.
     \item Ontology matching is needed by the \textit{screen developer} at the design-time of a screen (a visual building block of a gadget).
 Screen developers have a dedicated UI component for building screens, in which they can use the matching operators to combine components annotated with different ontologies.
-No actual matching is performed in this phase, but rather the simple possibility of matching.
+No actual matching needs to be performed in this phase, but rather the possibility of matching needs to be determined (i.e., \emph{can} two screens A and B be combined?).
 %the FAST user interface determines if two resources can be combined or not (see Figure \ref{fig:screens} for a conceptual example of resources that can not be combined).
     \item The \emph{gadget developer} combines screens to screen-flows and gadgets, and only uses ontology matching implicitly.
-    \item The \textit{end-user} uses the final deployed gadget, that combines the different services, but is unaware of the underlying resources and ontologies or the matching process.
-Thus we also need ontology matching at the gadget's run-time, where the mapping will have to be integrated into the actual JavaScript code of the running FAST gadget.
-%In this paper, we mainly consider the first two cases (i.e.
-ontology engineering and screen development), but portability of the alignment code to JavaScript must also be kept in mind.
+    \item The \textit{end-user} uses the final deployed gadget at run-time, but is unaware of the underlying resources and ontologies or the matching process.
+Only at run-time the actual mapping of instance data has to be performed.
+In this paper, we mainly consider the first two cases (i.e.
+ontology engineering and screen development).
 \end{inparaenum}
@@
 
-An ontology mapping (or alignment) is a declarative specification of the semantic overlap between two ontologies~\cite{debruin2005wsml}.
-It is the result of the ontology alignment process.
-This mapping is represented as a set of axioms in a mapping language.
+An \emph{ontology mapping} is a declarative specification of the semantic overlap between two ontologies~\cite{debruin2005wsml}.
+It is the result of the \emph{ontology alignment process}.
+This mapping is represented as a set of axioms in a \emph{mapping language}.
 The mapping process has three main phases:
@@
 \item All three phases of the process need to be accessible.
-\item Matching of OWL and RDFS ontologies should be supported, since we only consider these two languages.
+\item Matching of OWL and RDFS ontologies must be supported in the FAST project.
 \item The tool should be as independent as possible, performing the alignment process with little or no user interference. This is an important requirement, since FAST is end-user oriented.
-\item It should have an open source license, allowing its free use and modification, because it needs to be integrated into the free and open FAST platform.
-\item The code should be suitable for porting to other languages (in particular JavaScript), allowing it to be integrated into the FAST environment.
+\item The tool needs to be open source, allowing it to be integrated into the free and open FAST platform.
+\item The code should be suitable for porting to other languages (in particular JavaScript), allowing it to be integrated into the FAST gadget run-time.
 \item It should be well documented.
@@
 
-We have considered three tools, two of which we have already detailed in \cite{ambrus2009mediation}.
-\paragraph{MAFRA} \cite{maedche2002mafra} supports an interactive and incremental process of ontology mapping.
+Based on these requirements, we compared three different tools.
+\emph{MAFRA}~\cite{maedche2002mafra} supports an interactive and incremental process of ontology mapping.
 It provides an explicit notion of semantic bridges.
@@
 The bridges, however, have been designed to be used within the MAFRA system, and the alignment process needs to be done through the provided GUI.
-
-\paragraph{RDFT}
-\cite{omelayenko2002rdft} is a small language originally designed to map between XML and RDF.
+\emph{RDFT}~\cite{omelayenko2002rdft} is a small language originally designed to map between XML and RDF.
 The results are mappings represented in DAML+OIL, that can be executed in a transformation process.
-No hints are given to add alignment methods or extending the format and, as a major downside, we have not been able to locate the tool for download and testing.
-
-\paragraph{Alignment API} \cite{euzenat2004api} is the tool best matching the requirements, satisfying all the desired conditions.
-It also being widely used at the annual ontology matching evaluation campaigns and is under active development.
-It provides an API and its implementation, is open source (GPLv2 or above) and written in Java, providing an easy way to embed it into other programs.
+No hints are given to add alignment methods or extending the format and the tool does not longer seem to be available.
+\emph{Alignment API}~\cite{euzenat2004api} is the tool best matching our requirements, satisfying all the desired conditions.
+It is still under active development, provides an API and its implementation, is open source (GPLv2 or above) and written in Java, providing an easy way to embed it into other programs.
 Alignment API can be extended by other representations and matching algorithms, it can be invoked through the command line interface (thus working without user interference) or one of the two available GUI implementations, or it can be exposed as an HTTP server.
 The tool allows for testing different alignment methods and can generate evaluation results based on a reference alignment.
-One of the major advantages is that it can generate the mapping result in XSLT, making it possible with the use of a simple tool, or a JavaScript package to transform XML data from one ontology into the other, independently from the alignment system.
+Alignment API can generate the mapping results in XSLT, therefore providing an easy way to integrate them into other systems.
 
@@
 
-The scenario we want to test in FAST is taken from the e-commerce domain, based on the often-met case where a commercial agent aggregates data from other commercial websites.
-We consider the situation where a user needs a gadget which combines major e-commerce services, allowing to aggregate item lists from all of them in a combined interface.
-As examples in our scenario.
-we consider the two most popular online shopping websites, Amazon and eBay\footnote{Based on the Alexa ranking from http://alexa.com/topsites/category/Top/Shopping}, along with BestBuy, which is very close behind the former two.
-BestBuy is an interesting case, because it exposes its data in RDF using the GoodRelations (GR) ontology~\cite{hepp-goodrelations}, which has recently gained a lot of popularity.
+In our evaluation scenario, which is taken from the e-commerce domain, a user needs to build a gadget which combines data from major e-commerce services, allowing to aggregate item lists from all of them in a combined interface.
+As examples in our scenario, we consider the two most popular online shopping websites\footnote{\url{http://alexa.com/topsites/category/Top/Shopping}, checked 01/11/2009}, Amazon and eBay, along with the BestBuy site.
+The latter is an interesting case, because it exposes its data in RDF using the GoodRelations (GR) ontology~\cite{hepp-goodrelations}, which has recently gained a lot of popularity.
 It is therefore one of the first major e-commerce sites to provide semantic metadata.
@@
         \includegraphics[width=8cm]{images/screens_diagram.png}%[bb=0 0 201 126]
-        \caption{The scenario: three components retrieving incompatible data}
+        \caption{Three gadget components retrieving incompatible data}
     \label{fig:screens}
@@
 Figure~\ref{fig:screens} illustrates our scenario.
-There are three retrieval components that wrap the different e-commerce sites and provide data according to three different ontologies: the GR ontology, the Amazon ontology and the eBay ontology.
-Another component displays GR items for display to the user.
-Out of the box, only the BestBuy retrieval and the display component can be combined.
-However, if the gadget designer wants to aggregate data from all three services in the display, there will have to be a mapping present between the Amazon and eBay on the one hand, and the GR ontology on the other.
+There are three retrieval components that wrap the different e-commerce sites and provide data according to three different ontologies: the GR ontology, the Amazon ontology (A) and the eBay ontology (E).
+Another component displays GR items for display to the user, but not A or E items.
+If the gadget designer wants to aggregate data from all three services in the display, there will therefore have to be a mapping present between A and E on the one hand, and GR on the other.
 
@@
 \label{ontologies}
-We have used three ontologies to annotate underlying resources in FAST, the first of which, namely GoodRelations, is an extensive ontology for e-commerce.
-The other two, i.e., the Amazon and the eBay ontologies were developed by us as simplified versions of what would be used in the real-life scenarios.
-Nevertheless, they are sufficient for testing the feasibility of the scenario and the options presented by current ontology alignment methods.
+Of the three ontologies used in our evaluation, only GoodRelations is a real-world, extensive ontology for e-commerce.
+The other two, i.e., the Amazon and the eBay ontologies were developed for simulation purposes as simplified versions of what would be used in the real-life scenarios.
+They were designed to showcase particular features of ontology mapping in our scenario.
 
@@
 This ontology is aimed at annotating so-called ``offerings'' on the Web, which can be products or services.
-The ontology is  available under the Creative Commons Attribution 3.0 license and features support for ranges of units, measurements, currencies,  shipping and payments, common business functions (sell, lease, repair, etc.) and international standards (ISO 4217 or UNSPSC) and codes (e.g., EAN or UPC) in the field.
-
-\begin{center}
-\lstset{captionpos=b, breaklines=true}
-\lstset{frame=single, basicstyle=\scriptsize}
-\lstset{caption=Basic GoodRelations data in N3 notation, label=listing_gr}
-\lstset{language=XML}
-\begin{lstlisting}
-:Offering_8794691     a gr:Offering;
-  gr:hasPriceSpecification 
-      :UnitPriceSpecification_8794691_1;
-  gr:includesObject :TypeAndQuantityNode_8794691_1;
-
-:UnitPriceSpecification_8794691_1     a gr:UnitPriceSpecification;
-  gr:hasCurrency "USD"^^xsd:string;
-  gr:hasCurrencyValue "749.99"^^xsd:float;
-  gr:hasUnitOfMeasurement "C62"^^xsd:string;
-
-:TypeAndQuantityNode_8794691_1     a gr:TypeAndQuantityNode;
-  gr:amountOfThisGood "1.0"^^xsd:float;
-  gr:hasUnitOfMeasurement "C62"^^xsd:string;
-  gr:typeOfGood :ProductOrServicesSomeInstancesPlaceholder_8794691 .
-
-:ProductOrServicesSomeInstancesPlaceholder_8794691 a gr:ProductOrServicesSomeInstancesPlaceholder;
-  gr:hasEAN_UCC-13 "0013803096095"^^xsd:string;
-  gr:hasMakeAndModel :PoSM_8794691;
-  rdfs:comment "With Auto Optimization and [...]"@en;
-
-:PoSM_8794691     a gr:ProductOrServiceModel;
-  gr:hasManufacturer <bbuy:Manufacturer_Canon>;
-  rdfs:label "Canon EOS Digital Rebel [...]"@en;
-  rdfs:comment "With Auto Optimization and [...]"@en;
-\end{lstlisting}
-\end{center}
-
-The ontology (see List.~\ref{listing_gr} for an example of GR data)  is centred on the \texttt{Offering} class,
+The ontology features support for ranges of units, measurements, currencies,  shipping and payments, common business functions (sell, lease, repair, etc.) and international standards (ISO 4217 or UNSPSC) and codes (e.g., EAN or UPC) in the field.
+The main class is \texttt{Offering},
 which represents an announcement by a \texttt{BusinessEntity} to provide a \texttt{ProductOrService} with a given \texttt{BusinessFunction}.
@@
 A product is described by its title and description, manufacturer, make and model, etc.
-While GoodRelations offers terminology to allow a much higher expressivity, we will restrict the discussion to those terms relevant for our scenario.
+% While GoodRelations offers terminology to allow a much higher expressivity, we will restrict the discussion to those terms relevant for our scenario.
+
+% \begin{center}
+% \lstset{captionpos=b, breaklines=true}
+% \lstset{frame=single, basicstyle=\scriptsize}
+% \lstset{caption=Basic GoodRelations data in N3 notation, label=listing_gr}
+% \lstset{language=XML}
+% \begin{lstlisting}
+% :Offering_8794691     a gr:Offering;
+%   gr:hasPriceSpecification 
+%       :UnitPriceSpecification_8794691_1;
+%   gr:includesObject :TypeAndQuantityNode_8794691_1;
+% 
+% :UnitPriceSpecification_8794691_1     a gr:UnitPriceSpecification;
+%   gr:hasCurrency "USD"^^xsd:string;
+%   gr:hasCurrencyValue "749.99"^^xsd:float;
+%   gr:hasUnitOfMeasurement "C62"^^xsd:string;
+% 
+% :TypeAndQuantityNode_8794691_1     a gr:TypeAndQuantityNode;
+%   gr:amountOfThisGood "1.0"^^xsd:float;
+%   gr:hasUnitOfMeasurement "C62"^^xsd:string;
+%   gr:typeOfGood :ProductOrServicesSomeInstancesPlaceholder_8794691 .
+% 
+% :ProductOrServicesSomeInstancesPlaceholder_8794691 a gr:ProductOrServicesSomeInstancesPlaceholder;
+%   gr:hasEAN_UCC-13 "0013803096095"^^xsd:string;
+%   gr:hasMakeAndModel :PoSM_8794691;
+%   rdfs:comment "With Auto Optimization and [...]"@en;
+% 
+% :PoSM_8794691     a gr:ProductOrServiceModel;
+%   gr:hasManufacturer <bbuy:Manufacturer_Canon>;
+%   rdfs:label "Canon EOS Digital Rebel [...]"@en;
+%   rdfs:comment "With Auto Optimization and [...]"@en;
+% \end{lstlisting}
+% \end{center}
 
 \paragraph{Amazon Ontology:}
-We have created a small Amazon ontology based on the datatypes supported by the web service exposed by Amazon to third-party agents to use its services.
-It is a simplified version of what would be needed in a real-world application, but it suffices for the scope of this work.
-The ontology describes \texttt{Items} based on the \texttt{ItemAttributes} description given in the Amazon Product Advertising API documentation\footnote{http://docs.amazonwebservices.com/AWSECommerceService/latest/DG/}.
+We have created a small Amazon ontology based on a subset of the datatypes supported by the web service exposed by Amazon to third-party agents.
+The ontology describes \texttt{Items} based on the \texttt{ItemAttributes} description given in the Amazon Product Advertising API documentation\footnote{\url{http://docs.amazonwebservices.com/AWSECommerceService/latest/DG/}}.
+The ontology features three classes for describing a product. Example instance data is given in List.~\ref{listing_amazon}.
+\begin{inparaenum}[(1)]
+    \item \texttt{Item} represents an Amazon item, defined by a title, a manufacturer, a product group (DVD, Book, etc.), an international EAN code, an ASIN (unique Amazon id), an author (for books) and a \texttt{ListPrice}.
+    \item \texttt{Company}, described by a  legal name, is used for representing the manufacturer of an \texttt{Item}.
+    \item \texttt{ListPrice} has two properties: \texttt{hasCurrencyCode}, representing an ISO 4217 currency code (e.g.
+GBP or EUR), and \texttt{hasAmount} representing the price in the given currency.
+\end{inparaenum}
 
-\begin{center}
-\lstset{captionpos=b, breaklines=true}
-\lstset{frame=single, basicstyle=\scriptsize}
+% \begin{figure}[ht]
 \lstset{caption=Simplified Amazon ontology data in N3 notation, label=listing_amazon}
-%\caption{Minimal Amazon ontology snippet.}
-\lstset{language=XML}
-%\lstset{emph={vsmlVariant, namespace, ontology, concept, ofType, instance, nonFunctionalProperties, endNonFunctionalProperties, memberOf, hasValue}}
-%\lstset{emphstyle=\textbf\textit\underbar}
+\lstset{language=turtle}
 \begin{lstlisting}
@@
    amzn:hasTitle "Canon Digital Rebel XSi [...]" .
-
 :Manufacturer_Canon a amzn:Company;
    amzn:hasLegalName "Canon" .
-
 :Price_7590645_1 a amzn:ListPrice;
@@
 \end{lstlisting}
-\end{center}
-
-The ontology features three classes for describing a product. Example instance data is given in List.~\ref{listing_amazon}.
-\begin{inparaenum}[(1)]
-    \item \texttt{Item} represents an Amazon item, defined by a title, a manufacturer, a product group (DVD, Book, etc.), an international EAN code, an ASIN (unique Amazon id), an author (for books) and a \texttt{ListPrice}.
-    \item \texttt{Company}, described by a  legal name, is used for representing the manufacturer of an \texttt{Item}.
-    \item \texttt{ListPrice} has two properties: \texttt{hasCurrencyCode}, representing an ISO 4217 currency code (e.g.
-GBP or EUR), and \texttt{hasAmount} representing the price in the given currency.
-\end{inparaenum}
+% \end{figure}
 
 \paragraph{eBay Ontology:}
-The eBay ontology was created based on the eBay Shopping API \footnote{http://developer.ebay.com/DevZone/shopping/docs/CallRef/index.html} and is supposed to annotate data retrieved through the web service described by the given API.
-Like the Amazon ontology, it is a simplified case of what would be used in a real-world implementation.
-
-\begin{center}
-\lstset{captionpos=b, breaklines=true}
-\lstset{frame=single, basicstyle=\small}
-\lstset{caption=Simplified eBay ontology data in N3 notation, label=listing_ebay}
-\lstset{language=XML}
-\begin{lstlisting}
-
-:SimpleItem_1320648     a ebay:SimpleItem;
-  ebay:hasItemID "E012Y090912";
-  ebay:hasBidCount :"4";
-  ebay:hasCountry "UK";
-  ebay:hasCurrentPrice :Price_1320648_1;
-  ebay:hasPrimaryCategoryName "Electronics";
-  ebay:hasTitle "Canon Digital Rebel XSi [...]" .
-  ebay:hasDescription "Camera is practically unused. It's like new."
-
-:Price_1320648_1     a ebay:CurrentPrice;
-  ebay:hasAmount "575.55";
-  ebay:hasAmountType "GBP" 
-\end{lstlisting}
-\end{center}
-
-The ontology features three basic classes, (see Listing \ref{listing_ebay} for an example of ebay data)
+The eBay ontology was created based on the eBay Shopping API\footnote{\url{http://developer.ebay.com/DevZone/shopping/docs/CallRef/index.html}} and is supposed to annotate data retrieved through the web service described by the API.
+% \begin{center}
+% \lstset{captionpos=b, breaklines=true}
+% \lstset{frame=single, basicstyle=\small}
+% \lstset{caption=Simplified eBay ontology data in N3 notation, label=listing_ebay}
+% \lstset{language=XML}
+% \begin{lstlisting}
+% 
+% :SimpleItem_1320648     a ebay:SimpleItem;
+%   ebay:hasItemID "E012Y090912";
+%   ebay:hasBidCount :"4";
+%   ebay:hasCountry "UK";
+%   ebay:hasCurrentPrice :Price_1320648_1;
+%   ebay:hasPrimaryCategoryName "Electronics";
+%   ebay:hasTitle "Canon Digital Rebel XSi [...]" .
+%   ebay:hasDescription "Camera is practically unused. It's like new."
+% 
+% :Price_1320648_1     a ebay:CurrentPrice;
+%   ebay:hasAmount "575.55";
+%   ebay:hasAmountType "GBP" 
+% \end{lstlisting}
+% \end{center}
+The ontology features three basic classes,
+%  (see Listing \ref{listing_ebay} for an example of ebay data)
 \begin{inparaenum}[(1)]
@@
 
-There is a major paradigm difference between the GoodRelations ontology and the other two ontologies (see Sect.~\ref{results} for details).
-After studying the approach of the GR ontology, and examining the matching results between the Amazon and GR ontologies, we have concluded that automated matching between the two is unfeasible using the string-based methods employed using the Alignment API tool in this testing.
-Therefore, we have have created a manual mapping for this case, and have tested the automated matching on the Amazon--eBay pair, which closely resemble each other in their approach to capturing product details.
+There is a major paradigm difference between the GoodRelations ontology and the other two ontologies (see Sect.~\ref{sub:manual_mapping} for details).
+After some initial testing, we concluded that automatic mapping from GR to A/E using the string-based methods employed by the Alignment API tool was not feasible.
+Therefore, the following sections report on \emph{automatic mapping} for the A--E pair --- which are similar enough to be suitable for level 0 mapping ---, and \emph{manual mapping} for the GR--A/E pairs.
 
-\subsubsection{The Matcher}
-The matching tool called Alignment API \cite{euzenat2004api} has been developed to support several kinds of algorithms and methods, and provide an efficient environment for testing alignment systems.
+\subsubsection{Automatic Mapping}
 
-\paragraph{The Algorithm:}
-We have used a simple string distance-based algorithm provided by the tool, which computes the string distance between the names of the entities to find correspondences between them.
+% \paragraph{The Matcher} % (fold)
+% \label{par:the_matcher}
+
+% paragraph the_matcher (end)
+For automatic mapping of level 0 mappings, we used a simple string distance-based algorithm provided by Alignment API~\cite{euzenat2004api}, which computes the string distance between the names of the entities to find correspondences between them.
 Four methods have been used for computing the distance:
@@
 
-\paragraph{The Alignment Description:} 
-The alignment description is given based on a simple vocabulary, containing a pair of ontologies and a set of correspondences, which express relations holding among entities of the two ontologies.
-
+The alignment description derived from these methods is given based on a simple vocabulary, containing a pair of ontologies and a set of correspondences, which express relations between entities of the two ontologies.
 We used the level 0 mapping representation for representing simple mappings, which map discrete entities of the two ontologies.
 Thus the representation of the correspondences is given with the five elements described (with the \texttt{id} being optional), as shown in List.~\ref{listing_correspondence}.
-Level 2 mappings were also used for more complex, manually-created representations, as detailed in Sect.~\ref{results}.
+Similar mappings were were also used for more complex, manually-created representations (level 2), as detailed in Sect.~\ref{sub:manual_mapping}.
 
-\begin{center}
-\lstset{frame=single, basicstyle=\small}
+% \begin{figure}[ht]
+% \begin{minipage}{\linewidth}
 \lstset{caption=Level 0 mapping element example, label=listing_correspondence}
-\lstset{captionpos=b}
-\lstset{language=XML}
-%\lstset{emph={vsmlVariant, namespace, ontology, concept, ofType, instance, onFunctionalProperties, endNonFunctionalProperties, memberOf, hasValue}}
-%\lstset{emphstyle=\textbf\textit\underbar}
+\lstset{language=turtle}
 \begin{lstlisting}
-[ a :Cell;
-  :entity1 amzn:hasCurrencyCode;
-  :entity2 ebay:hasAmountType;
-  :measure "1.0"^^xsd:float;
-  :relation "=" ]
+<level_0_mapping> a align:Cell;
+  align:entity1 amzn:hasCurrencyCode;
+  align:entity2 ebay:hasAmountType;
+  align:measure "1.0"^^xsd:float;
+  align:relation "=" .
 \end{lstlisting}
-\end{center}
+% \end{minipage}
+% \end{figure}
 
-\paragraph{Using the Tool:}
-The Alignment API tool can be used through a GUI, as a server or from the command-line interface, of which we have chosen the last one.
-The tool reads two RDF/OWL ontologies, computes the alignment between them, performs some thresholding and displays the results.
-It can render the output in a number of formats, including HTML and XSLT.
+% \subsubsection{Using the Tool.}
+% The Alignment API tool can be used through a GUI, as a server or from the
+% command-line interface, of which we have chosen the last one.
+% The tool reads two RDF/OWL ontologies, computes the alignment between them, 
+% performs some thresholding and displays the results. It can render the output
+% in a number of formats, including HTML and XSLT.
+% 
+% An additional feature of the tool is its ability to evaluate results based on
+% a reference alignment, and output the evaluation results in a table,
+% or plot them as \LaTeX{} graphs.
 
-An additional feature of the tool is its ability to evaluate results based on a reference alignment, and output the evaluation results in a table, or plot them as \LaTeX{} graphs.
+\paragraph{Testing Procedure}
+For the Amazon--eBay pair we set up a reference alignment, against which the results are evaluated.
+We then ran the matching process for all for methods:
+\begin{inparaenum}[(1)]
+    \item equality,
+    \item Levenshtein distance with a confidence threshold of \texttt{0.33} (meaning that any correspondence having a smaller confidence measure will be excluded),
+    \item SMOA distance with a threshold of \texttt{0.5} and
+    \item Wordnet distance using a threshold of \texttt{0.5}.
+\end{inparaenum}
+To apply the results, we rendered an XSLT template to transform an example dataset.
 
-\subsubsection{Testing Procedure}
-We set up the ontologies according to Sect.~\ref{ontologies}. For the Amazon--eBay pair we set up a reference alignment, against which the results are evaluated.
-%For the reason of excluding the Amazon - GoodRelations pair from this procedure please refer to Section \ref{results}.
+\paragraph{Results}
 
-\begin{center}
-\lstset{frame=single, basicstyle=\small}
-\lstset{caption=Executing the alignment tool on a pair of ontologies, label=listing_cli}
-\lstset{language=bash}
-\lstset{captionpos=b}
-\begin{lstlisting}
- $ java -jar lib/procalign.jar amazon.owl ebay.owl 
-     -o results/equal.rdf
-\end{lstlisting}
-\end{center}
-
-Using the tool's command line interface, we have performed the following steps:
-\begin{itemize}
-\item generate the RDF alignment output for the four methods: (1) equality, (2) Levenshtein distance with a confidence threshold of \texttt{0.33} (meaning that any correspondence having a smaller confidence measure will be excluded), (3) SMOA distance with a threshold of \texttt{0.5} and (4) Wordnet distance using a threshold of \texttt{0.5}.
-\item render the XSLT file for the SMOA distance function;
-\item transform an example dataset using the XSLT transformation rules;
-\item generate the tables and graphs displaying different parameters of the evaluation.
-\end{itemize}
-
-\subsubsection{Results}
-\label{results}
-
-\paragraph{GoodRelations:}
-\label{gr}
-
-The GoodRelations ontology employs a unique paradigm, different from the paradigms of Amazon and eBay.
-In GR everything is centred around an instance of \texttt{Offering} and a graph of other instances attached to it, whereas for Amazon (and similarly for eBay), the main class is \texttt{Item}, which holds all relevant properties.
-In principle, \texttt{Item} would correspond to \texttt{ProductOrService} in GoodRelations, but the properties of the \texttt{Item} class are reflected as properties of many different classes in GR.
-
-Though the infeasibility of automating this alignment became obvious, we have represented the alignment in the mapping language supported by the tool, as a level 2 mapping (described in Sect.
-\ref{mediation}) (the mapping has been created between Amazon and GoodRelations, since the eBay and Amazon ontologies are very similar in approach, and mapping eBay to GoodRelations would be an indentical task).
-This mapping description can later be used by the run-time JavaScript code.
-List.~\ref{listing_gr_a} shows an example mapping between two properties of the two ontologies, specifying that the relationship is \texttt{Equivalence} with a certainty degree of \texttt{1.0}.
-This fragment does not show, but assumes the equivalence correspondence between the classes \texttt{Item} and \texttt{Offering}, which is a trivial level 0 mapping.
-This mapping specifies the relation
-
-\begin{equation*} \label{eq}
-\begin{split}
-\forall{v,z}; hasEAN(v,z) & \Longrightarrow \exists{x, y}; includesObject(v,x) \wedge \\
-&typeOfGood(x,y) \wedge hasEAN\_UCC\_13(y,z),
-\end{split}
-\end{equation*}
-
-meaning that the \texttt{hasEAN} property of \texttt{v} in the Amazon ontology corresponds to the \texttt{hasEAN\_UCC\_13} property of the \texttt{typeOfGood} of the \texttt{includesObject} of \texttt{v} in GoodRelations.
-The domains and ranges of the properties are inferred, thus it is deduced, that in \texttt{Amazon v} is of type \texttt{Item} and \texttt{z} is \texttt{int}, and in \texttt{GoodRelations v, x, y} and \texttt{z} are instances of the classes \texttt{Offering}, \texttt{TypeAndQuantityNode}, \texttt{ProductOrService} and \texttt{int}, respectively.
-
-\begin{center}
-\lstset{frame=single, basicstyle=\small}
-\lstset{caption=Fragment of the Amazon--GoodRelations mapping, label=listing_gr_a}
-\lstset{language=bash}
-\lstset{captionpos=b, breaklines=true}
-\begin{lstlisting}
-<Cell rdf:about="MappingRule_01">
- <entity1><omwg:Property rdf:about="&amzn;hasEAN"/></entity1>
- <entity2>
-  <omwg:Property>
-    <first><Relation rdf:about="&gr;includesObject"/></first>
-    <next><Relation rdf:about="&gr;typeOfGood"/></next>
-    <next><Property rdf:about="&gr;hasEAN_UCC_13"/></next>
-  </omwg:Property>
- </entity2>
- <measure rdf:datatype="&xsd;float">1.0</measure>
- <relation>Equivalence</relation>
-</Cell>
-\end{lstlisting}
-\end{center}
-
-Using this representation, complex correspondences can be modelled, using first order logic constructs.
-We will further investigate the possibilities of converting the level 2 mapping to XSLT, so it can automatically be used in the final JavaScript implementation.
-In case this is not possible, we will represent the mapping elements in SPARQL.
-
-\paragraph{Alignment Results:}
 The results of automatically aligning the Amazon and eBay ontologies were quite favourable.
@@
 
-\begin{center}
 \begin{table}
+    \centering
     \caption{Alignment results: Precision, Recall, Fallout and F-Measure}
@@
 
-\begin{center}
-\begin{tabular}[c]{|llll|llll|llll|} \hline
-\multicolumn{4}{|l|}{Reference} & \multicolumn{4}{|l|}{Equal} & 
-\multicolumn{4}{|l|}{SMOA}\\ \hline
-Prec & Rec & Fall & FMeas & Prec & Rec & Fall & FMeas & Prec & Rec & 
-Fall & FMeas  \\ \hline
-1.00& 1.00& 0.00& 1.00& 1.00& 0.38& 0.00& 0.55& 0.43& 0.75& 0.57& 0.55 \\ \hline
+    \begin{tabular}{rrrrr}
+        \toprule
+        & \textbf{precision} & \textbf{recall} & \textbf{fallout} & \textbf{f-measure}\\
+        \midrule
+        \textbf{reference} & 1.00 & 1.00 & 0.00 & 1.00\\
+        \textbf{equality} & 1.00 & 0.38 & 0.00 & 0.55\\
+        \textbf{SMOA} & 0.43 & 0.75 & 0.57 & 0.55\\
+        \textbf{Levenshtein} & 0.40 & 0.75 & 0.60 & 0.52\\
+        \textbf{JWNL} & 0.67 & 0.75 & 0.33 & 0.71\\
+        \bottomrule
     \end{tabular}
-\end{center}
+\end{table}
 
-\begin{center}
-\begin{tabular}{|llll|llll|} \hline
-\multicolumn{4}{|l|}{Levenshtein} & \multicolumn{4}{|l|}{JWNL} \\ \hline
- Prec & Rec & Fall & FMeas & Prec & Rec & Fall & FMeas \\ \hline
-0.4& 0.75& 0.6& 0.52& 0.67& 0.75& 0.33& 0.71 \\ \hline
-\end{tabular}
-\end{center}
-
-\end{table}
-\end{center}
-
-The first column (Reference) shows the reference alignment, which, naturally, has both perfect precision and recall.
-We can observe what intuition has predicted, namely that pure string equality (Equal) is far too simple and irrelevant, by only taking identical labels.
+The first row (reference) shows the reference alignment, which, naturally, has both perfect precision and recall.
+We can observe what intuition has predicted, namely that pure string equality (equality) is far too simple and irrelevant, by only taking identical labels.
 By using string distances and giving certain thresholds (Levenshtein and SMOA), we can see that the results are much less precise, but have a better recall, since this allows for entities having similar names to be discovered, at the expense of having quite a few incorrect results (lower precision); the thresholds allow for low-scored cases to be eliminated, although this results in the exclusion of some correct correspondences.
-The last column (JWNL) contains the results of the Wordnet-enabled method, which shows quite an improvement, due to the lexical analysis, which performs a much more relevant comparison of strings, giving a high number of correct results.
-We can clearly see how it is quite close to the reference alignment, having a correctness of 0.67 and a recall of 0.75.
+The last column (JWNL) contains the results of the Wordnet-enabled method, which shows quite an improvement (precision of 0.67 and a recall of 0.75), due to the lexical analysis, which performs a much more relevant comparison of strings, giving a high number of correct results.
 The precision of the JWNL alignment shows only a tiny drop below the recall value, meaning that the number of incorrect correspondences discovered is small, and the main source of error is from the number of correspondences not discovered.
@@
 We can deduce that the results provided are satisfactory, even though the methods used were simple, string-based ones, and the process was completely automated without any user input.
-This means that through some user assistance or an initial input alignment the tool can achieve 100\% correct results.
+We are therefore confident that through some user assistance or an initial input alignment the tool can achieve 100\% correct results.
 
-\paragraph{XSLT:}
-One of the appealing features of the Alignment API is that it can render the results in XSLT, thus specifying a set of rules which make it possible to transform an  XML file from one format to the other.
+% \paragraph{XSLT:}
+% One of the appealing features of the Alignment API is that it can render the results in XSLT, thus specifying a set of rules which make it possible to transform an  XML file from one format to the other.
 We have generated such an XML transformation description, and with the simple Linux application \texttt{xsltproc} (run from the command line) we have managed to transform data (limited, of course by the incompleteness of the alignment) from the Amazon ontology into data annotated by the eBay vocabulary.
 %Please refer to Appendix A for a listing of an example transformation.
+
+\subsubsection{Manual Mapping} % (fold)
+\label{sub:manual_mapping}
+
+The GoodRelations ontology employs a unique paradigm, different from the paradigms of Amazon and eBay.
+In GR everything is centred around an instance of \texttt{Offering} and a graph of other instances attached to it, whereas for Amazon (and similarly for eBay), the main class is \texttt{Item}, which holds all relevant properties.
+In principle, \texttt{Item} would correspond to \texttt{ProductOrService} in GoodRelations, but the properties of the \texttt{Item} class are reflected as properties of many different classes in GR.
+
+Though the infeasibility of automating this alignment became obvious, we have represented the alignment in the mapping language supported by the tool, as a level 2 mapping (described in Sect.
+\ref{mediation}).
+ % (the mapping has been created between Amazon and GoodRelations, since the eBay and Amazon ontologies are very similar in approach, and mapping eBay to GoodRelations would be an indentical task). 
+This mapping description can later be used by the run-time gadget code.
+List.~\ref{listing_gr_a} shows an example mapping between two properties of the two ontologies, specifying that the relationship is \texttt{Equivalence} with a certainty degree of \texttt{1.0}.
+This fragment does not show, but assumes the equivalence correspondence between the classes \texttt{Item} and \texttt{Offering}, which is a trivial level 0 mapping.
+This mapping specifies the relation
+
+\begin{equation*} \label{eq}
+\begin{split}
+\forall{v,z}; hasEAN(v,z) & \Longrightarrow \exists{x, y}; includesObject(v,x) \wedge \\
+&typeOfGood(x,y) \wedge hasEAN\_UCC\_13(y,z),
+\end{split}
+\end{equation*}
+
+meaning that the \texttt{hasEAN} property of \texttt{v} in the Amazon ontology corresponds to the \texttt{hasEAN\_UCC\_13} property of the \texttt{typeOfGood} of the \texttt{includesObject} of \texttt{v} in GoodRelations.
+The domains and ranges of the properties are inferred, thus it is deduced, that in \texttt{Amazon v} is of type \texttt{Item} and \texttt{z} is \texttt{int}, and in \texttt{GoodRelations v, x, y} and \texttt{z} are instances of the classes \texttt{Offering}, \texttt{TypeAndQuantityNode}, \texttt{ProductOrService} and \texttt{int}, respectively.
+
+\begin{figure}[ht]
+\lstset{caption=Fragment of the Amazon--GoodRelations mapping, label=listing_gr_a}
+\lstset{language=turtle}
+\begin{lstlisting}
+<level_2_mapping> a align:Cell;
+   align:entity1 amzn:hasEAN;
+   align:entity2 
+      [ a align:Property;
+        align:first gr:includesObject;
+        align:next gr:hasEAN_UCC_13,
+           gr:typeOfGood ];
+     align:measure "1.0"^^xsd:float;
+     align:relation "Equivalence" .
+
+amzn:hasEAN a align:Property .
+gr:hasEAN_UCC_13 a align:Property .
+gr:includesObject a align:Relation .
+gr:typeOfGood a align:Relation .
+\end{lstlisting}
+\end{figure}
+
+% \begin{center}
+% \lstset{caption=Fragment of the Amazon--GoodRelations mapping, label=listing_gr_a}
+% \lstset{language=xml}
+% \begin{lstlisting}
+% <Cell rdf:about="MappingRule_01">
+%  <entity1><omwg:Property rdf:about="&amzn;hasEAN"/></entity1>
+%  <entity2>
+%   <omwg:Property>
+%     <first><Relation rdf:about="&gr;includesObject"/></first>
+%     <next><Relation rdf:about="&gr;typeOfGood"/></next>
+%     <next><Property rdf:about="&gr;hasEAN_UCC_13"/></next>
+%   </omwg:Property>
+%  </entity2>
+%  <measure rdf:datatype="&xsd;float">1.0</measure>
+%  <relation>Equivalence</relation>
+% </Cell>
+% \end{lstlisting}
+% \end{center}
+
+Using this representation, complex correspondences can be modelled, using first order logic constructs.
 

### 2010-01-21 trunk/deliverables/D4_3_2/fast_d4_3_2.tex
first draft
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@443 d8b0c59b-6544-0410-94d9-83a5658a058e

\documentclass{fast_latex}
% yay! We can use any kind of funky diacritic:
\usepackage[utf8]{inputenc}

\usepackage{lastpage}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage[pdfborder={0 0 0}]{hyperref}		% turn on when latex is used (not miktec)
\usepackage{url} % LEO: urls \url{}
\usepackage{verbatim} % code and comment
\usepackage{longtable}
\usepackage{xspace}   % whitespace after a macro if no punctuation after the macro
\usepackage{multirow}
\usepackage{colortbl}
\usepackage{longtable}
\usepackage{array}
\usepackage{amssymb} 
\parindent0pt

\newcommand\deliverableNumber{D4.3.2}
\newcommand\deliverableTitle{Mechanisms for Gadget-Service Connections and Gadget Functionality}
\newcommand\deliverableTitleShort{Mechanisms for Gadget-Service Connections}
\newcommand\workpackageNumber{4}
\newcommand\workpackageTitle{Visual composition of screen-flow resources and interoperability with back-end Web Services}
\newcommand\authorOne{Ismael Rivera, NUIG}
\newcommand\authorTwo{Author 2}
\newcommand\authorThree{Author 3}
\newcommand\authorFour{Author 4}
\newtheorem{example}{\emph{Example}}

\begin{document}
% explicit hyphenations
%\hyphenation{RDF-Re-po-si-to-ry}
%\hyphenation{name-space}

%\fontfamily{tahoma}\selectfont
\def\note#1{\marginpar{\footnotesize#1}} % use this to show the notes in the document
%\def\note#1{} % use this to hide the notes

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TITLE PAGES 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\thispagestyle{empty}

\pagenumbering{roman}

\begin{flushright}
	\includegraphics[width=3cm]{images/FP7_logo}
\end{flushright}

\vspace{1cm}

%\begin{minipage}[p]{15cm}
	\begin{center}
		\includegraphics{images/FAST_logo}\\
		\vspace{1cm}
		{\LARGE{\sffamily \emph{FAST AND ADVANCED STORYBOARD TOOLS}}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{FP7-ICT-2007-1-216048}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{http://fast.morfeo-project.eu}}\\
		\vspace{4cm}
		{\LARGE \sffamily \textbf{Deliverable \deliverableNumber}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \textbf{\deliverableTitle}}\\
		\vspace{2cm}
		{\large \sffamily \authorOne}\\
%		{\large \sffamily \authorTwo}\\
		\vspace{0.5cm}
		\vfill
		{\large \sffamily Date: 26/02/2010}\\
		\vspace{1cm}
		{\sffamily FAST is partially funded by the E.C. (grant code: FP7-ICT-2007-1-216048).}

	\end{center}
%\end{minipage}

\clearpage
%%%%%%%%%%%%%%
% NEXT PAGES %
%%%%%%%%%%%%%%
\pagestyle{scrheadings}

\lohead{\includegraphics[width=4cm]{images/FAST_logo_transparent}}
%\cohead{\small\textcolor{fast@lightgrey}{\deliverableTitle}}
%\rohead{\small{\today}}
%\lofoot{\small\textcolor{fast@lightgrey}{Task Force Ontologies}}
\cofoot{\small{FAST 216048 --- \deliverableTitleShort}}
\rofoot{\small{\thepage}}

\newpage
%
%\section*{Authors}
%\authorOne \\
%%\authorTwo \\
%%\authorThree\\
%%\authorFour\\
%
%
%\vfill
%%\section*{Project Co-ordinator}
%%\hspace*{-2,5cm}\begin{minipage}[p]{14cm}
%%Dr. Ansgar Bernardi \\
%%German Research Center for Artificial Intelligence (DFKI) GmbH \\
%%Trippstadter Strasse 122 \\
%%D 67663 Kaiserslautern \\
%%Germany \\
%%Email: bernardi@dfki.uni-kl.de, phone: +49 631 205 3582, fax: +49 631 205 4910 \\
%%\end{minipage}
%
%
%\section*{Partners}
%Telef\'onica Investigaci\'on y Desarrollo (Telef\'onica I+D), Spain \\
%National University of Ireland, Galway (NUIG), Republic of Ireland \\
%Kassel University, Germany \\
%Cyntelix Corporation, Republic of Ireland \\
%SAP AG, Germany \\
%Universidad Polit\'ecnica de Madrid (UPM), Spain
%
%
%\vspace{0.3cm}
%\begin{footnotesize}
%Copyright: FAST Consortium 2009\\
%\end{footnotesize}

%\clearpage

\section*{Version History}

\begin{small}
\begin{tabular}{|l|l|l|p{7.5cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Rev. No.}} &
                            \textcolor{white}{\textbf{Date}} &
                            \textcolor{white}{\textbf{Author (Partner)}} &
							\textcolor{white}{\textbf{Change description}}\\ \hline
%0.1 & 15.12.2008 & Knud M\"oller (NUIG) & template modelled \\ \hline
%0.2 & 15.02.2009 & Knud M\"oller (NUIG) & content moved from wiki after internal review \\ \hline
1.0 & 26.02.2010 & Ismael Rivera (NUIG) & Final version ready for external review \\ \hline
\end{tabular}
\end{small}

\color{black}

\vfill
%{\bf Explanations of abbreviations on front page}\\
%\\
%%Nature \\
%R: Report \\
%P: Prototype \\
%R/P: Report and Prototype \\
%O: Other \\
% \\
%Dissemination level \\
%PU: Public \\
%PP: Restricted to other FP6 participants \\
%RE: Restricted to specified group \\
%CO: Confidential, only for NEPOMUK partners \\

\newpage

%%%%%%%%%%%%%%%%%%%%%
% Executive Summary %
%%%%%%%%%%%%%%%%%%%%%

\clearpage

\section*{Executive Summary}
\doublespacing

This deliverable exposes several mechanisms which will allow the connection and interaction between end-user's interfaces to third-party back-end services.

These back-end services cannot be directly used; hence they need to be encapsulated in what in the FAST platform is called Resource Adapters. The application of semantics to the back-end, through the corresponding Resource Adapters, and front-end building blocks assures a powerful instrument in the task of building new gadgets, improving the search, and enhancing the connection among the different building blocks which compose a gadget.

Therefore, the focus of this deliverable is to define how these wrappers will be constructed to allow the FAST platform exploiting web services \cite{GustavoAlonso2003} such as RESTful web services, SOAP-based web services and semantic web services through WSMO, and define mechanisms to connect them within the gadgets.

\newpage

%%%%%%%%%%%%%%%%%%%%%
% Document Summary %
%%%%%%%%%%%%%%%%%%%%%

\clearpage

\section*{Document Summary}
% double spacing from here on:
\singlespacing
\begin{small}

\begin{tabular}
	%{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{6cm}|p{2cm}|p{2cm}|}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{6cm}|p{2cm}|p{2cm}|}
	\hline
	\textcolor{white}{\textbf{Code}} & {FP7-ICT-2007-1-216048} & {\textbf{Acronym}} & {FAST}\\ \hline
	\textcolor{white}{\textbf{Full title}} & \multicolumn{3}{l|}{Fast and Advanced Storyboard Tools}\\ \hline
	\textcolor{white}{\textbf{URL}} & \multicolumn{3}{l|}{\url{http://fast.morfeo-project.eu}}\\ \hline
	\textcolor{white}{\textbf{Project officer}} & \multicolumn{3}{l|}{Annalisa Bogliolo}\\ \hline
\end{tabular}

\vspace{0.5cm}

\begin{tabular}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.25cm}|p{1cm}|p{1cm}|p{6.32cm}|}
	\hline
	\textcolor{white}{\textbf{Deliverable}} & {\textbf{Number}} & {\deliverableNumber} & {\textbf{Name}} & {\deliverableTitle}\\ \hline
	\textcolor{white}{\textbf{Work package}} & {\textbf{Number}} & {\workpackageNumber} & {\textbf{Name}} & {\workpackageTitle}\\ \hline
\end{tabular}

\vspace{0.5cm}

\begin{tabular}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
	\hline
	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2009} & {\textbf{Submitted}} & {27/02/2009}\\ \hline
	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D4.3.2 is...}\\ \hline
	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
%	{} & \multicolumn{4}{l|}{}\\ 
%	{} & \multicolumn{4}{l|}{}\\ 
%	{} & \multicolumn{4}{l|}{}\\
	\hline
	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {ismael.rivera@deri.org} \\ \cline{4-5}
	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 ******} \\ \hline
	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, web services, WSDL, REST, SOAP, WSMO}\\ \hline
\end{tabular}
\end{small}
\newpage

%%%%%%%%%%%%%%%%%%%%%
% TABLE OF CONTENTS %
%%%%%%%%%%%%%%%%%%%%%
\doublespacing
\setcounter{tocdepth}{3}
\tableofcontents
\cleardoublepage
\pagenumbering{arabic}

%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGINNING OF SECTIONS %
%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage
\rofoot{\small{Page \thepage\ of \pageref{LastPage}}}

\section{Introduction} % (fold)
\label{sec:introduction}

\subsection{Goal and Scope} % (fold)
\label{sub:goal_and_scope}

The objective of this deliverable is the analysis and development of mechanisms to facilitate the connection between screen-flow gadgets and underlying Web services, based on front-end user requirements and on the semantic descriptions of the service wrappers, and to develop these service wrappers from of the Web services' APIs and formal descriptions.

% subsection goal_and_scope (end)

\subsection{Structure of the Document} % (fold)
\label{sub:structure_of_the_document}

This deliverable is structured as follows. Section ~\ref{sec:introduction} states the goal, scope and structure of the document. Section ~\ref{sec:web_services_wrapping} describes what a service wrapper is, how the platform is able to interact with web services, how they these wrappers are build and then discovered to be reused in any gadget. Finally, Section ~\ref{sec:restful_web_services_wrapper_tool} defines the service wrapper tool built to facilitate FAST users to create Resource Adapters of RESTful web services.

% section introduction (end)

\clearpage
\section{Web Services Wrapping} % (fold)
\label{sec:web_services_wrapping}

In the context of FAST philosophy, a service is a software element or system, often deployed within enterprise boundaries, designed to support interoperable Machine-to-Machine interaction over a network (e.g. Web Services, Databases, CORBA or RPC interfaces...). These services have to be used by the gadgets, but there are several complications to be resolved in order to communicate the front-end (i.e. the gadget user interface and logic) with those services, such as the disparity of interfaces and invocation mechanisms. Another issue is the complexity for an end-user to interact with those services. Hence, this interaction will be shift to a user-interaction paradigm through a user-friendly graphical interface (service front-end), allowing both humans and machines to interact with the services through a uniform fashion.

Therefore, a complex gadget in terms of FAST is aimed to provide a functional access from a graphical user interface to a set of services (SOAP or REST-based Web services) and data sources (Atom/RSS feeds). These services and data sources need to be modelled and encapsulated within the platform in order to allow the discovery and use of them by other components (i.e. forms and operators). This is called Resource Adapter in the FAST platform. Figure ~\ref{fig:complex_gadget_architecture} illustrates their position and interaction with the rest of the building blocks within a gadget.

\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{images/ComplexGadgetArchitecture.pdf}
    \caption{Complex Gadget Architecture}
    \label{fig:complex_gadget_architecture}
  \end{center}
\end{figure}

Once a brief background has been shown, the following sections explain how a user, in this case a resource developer, can build Resource Adapters, and how another user, a screen developer for instance, will discover and connect them while creating a screen. Hence, these three phases are called: building, discovery and connection.

\subsection{Building} % (fold)
\label{sub:building}

\subsubsection{REST-based Web Services} % (fold)
\label{ssub:rest_based_web_services}

REST is a term to describe an architecture style, not a standard, of networked systems. The acronym REST stands for Representational State Transfer. REST-based or RESTful web services \cite{Fielding2000} are created identifying all of the conceptual entities or resources which want to be exposed as services. Those entities or resources should be nouns, not verbs (orders, tickets, etc.). Then, the interaction with those resources is made by convention using HTTP verbs such as GET, POST, PUT or DELETE, in order to retrieve, create, modify or delete them.

REST is lightweight (not a lot of extra xml markup, human readable results), but unlike SOAP-based web services, which have a standard vocabulary, and commonly used, to describe the web service interface through WSDL, RESTful web services are currently not formally described most of the times. For a service consumer to understand the context and content of the data that must be sent to and received from the service, both the service consumer and service producer must have an out-of-band agreement. This takes the form of documentation, sample code, and an API that the service provider publishes for developers to use. For example, the many web-based services available from Google, Yahoo, Flickr, Amazon, and so on have accompanying artefacts describing how to consume the services. This style of documenting REST-based web services is fine for use by developers, but it averts tools from programmatically consuming such services and generating artefacts specific to programming languages, as a web service described using WSDL allows. Nevertheless, Web Application Description Language (WADL) attempts to resolve some of these issues by providing a means to describe services in terms of schemas, HTTP methods, and the request or response structures exchanged, but this language is not widely adopted yet by RESTful web service developers.

To permit a high number of REST-based web services to be integrated to the FAST platform, the approach taken is from a manual development perspective. There is no need of a formal document such as WADL defining the web service, for this reason, building service wrappers for these services involve the correct understanding of the service by a human being and a tool to facilitate this task. This tool is explained in detail in Section ~\ref{sec:restful_web_services_wrapper_tool}.

% subsubsection rest_based_web_services (end)

\subsubsection{SOAP Web Services} % (fold)
\label{ssub:soap_web_services}

SOAP-based web Services or "Heavyweight Web Services" use Extensible Markup Language (XML) \cite{XML1.1} messages that follow the Simple Object Access Protocol (SOAP) standard \cite{SOAP1.2} and have been popular with traditional enterprise, usually relying on HTTP for message negotiation and transmission. In such services, there is often a machine-readable description of the operations offered by the service written in a Web Services Description Language (WSDL) document. Hence, the advantage of using WSDL is that it can be programmatically processed.

Shortly, a WSDL definition of a service, regarding the WSDL 2.0 specification \cite{WSDL2.0}, will contain the following information:
\begin{description}
	\item[Interfaces] A set of Interface components describing sequences of messages that a service sends and/or receives.
	\item[Bindings] A set of Binding components describing concrete message formats and transmission protocols which may be used to define the endpoints.
	\item[Services] A set of Service components describing a set of endpoints at which a particular deployed implementation of the service is provided.
	\item[Element declarations] A set of Element Declaration components defining the name and content model of the element information items such as that defined by an XML Schema global element declaration.
	\item[Type definitions] A set of Type Definition components defining the content model of the element information items such as that defined by an XML Schema global type definition.
\end{description}

From a specific WSDL definition, a set of methods or operations can be easily extracted which encloses a set of inputs and outputs. These operations would be transformed into actions, and the inputs and outputs would be used to define Resource Adapter pre/postconditions. With the use of this information along with the XML Schema defining the types, a Resource Adapter can be semantically defined.

Moreover, there are several frameworks to facilitate the programmatic use of these formal definitions of web services. These frameworks shall be use to transform a WSDL definition into an executable Resource Adapter ready to use inside a gadget.

% subsubsection soap_web_services (end)

\subsubsection{WSMO Web Services} % (fold)
\label{ssub:wsmo_web_services}

Semantic web services bring a number of advantages in the creation of the resources adapters over classic web services. Formal and semantic descriptions of web services allow mechanisms to raise its exploitation in a more automatic way. In this section, some general notions of the Web Service Modelling Ontology (WSMO) will be explained and how FAST can make use of it.

In a few words, WSMO provides means to describe all relevant aspects of semantic web services in a unified manner. A web service description in WSMO consists of five sub-components: non-functional properties, imported ontologies, used mediators, a capability and interfaces. However, we will focus on the capabilities and the interfaces since they are components which will make possible the integration of these services in FAST.

Capabilities and interfaces are the two types of Web Service description in WSMO. The capabilities describe the different functions of WSMO, while the interfaces specify:
\begin{enumerate}
	\item How to communicate with a web service in order to avail of its functionality. This is called Choreography.
	\item How the functionality of a web service is enabled by interacting with other Web Services. This is called Orchestration.
\end{enumerate}

A web service in WSMO defines one and only one capability. The capability of a web service defines its functionality in terms of pre/postconditions, assumptions and effects. A web service capability is defined by specifying the following elements: non-functional properties, imported ontologies, used mediators, shared variables, precondition, postcondition, assumption, and effect.

Basically, the web service will offers to a client a postcondition when some conditions are met in the information space (precondition). Effects and assumptions can stay out of this study, since they will not be taken into account by any of the building blocks of a gadget.

NEED TO BE FINISHED

% subsection building (end)

\subsection{Discovering} % (fold)
\label{sub:discovering}

In the building phase, every Resource Adapter is created sharing a common structure for screen components, such as forms, operators and resources. Hence, it will have a set of actions which will contain a set of preconditions to satisfy in order to be executed, and after its execution, it may produce any of the conditions of their postconditions.

For this reason, a screen developer during the screen development phase is able to reuse these screen components which are already stored in the catalogue, establishing which the screen's pre/postconditions are, and then the system will suggest several screen components which may satisfy these pre/postconditions. In a first step, the system will suggest screen components which have any of the preconditions which can be connected to any of the screen preconditions (the connection will be represented by an internal pipe), and the same would happen for the postconditions. Once a screen component is inserted into the screen, the system will take its definition into consideration to recommend new screen components which can be connected to the ones the screen is composed of.

In these bases, the mechanism is relatively similar to the screen recommender, used to find the best screens to create a screen flow. Once more, it is based on the pre/postconditions of a set of building blocks.

% subsection discovery (end)

% section web_services_wrapping (end)

\clearpage
\section{RESTful Web Services Wrapper Tool} % (fold)
\label{sec:restful_web_services_wrapper_tool}

In FAST, the wrapper tool will be in charge of the Resource Adapters' building phase. At the moment, just supports RESTful web services. The building phase for this type of web services is done in two steps: a first step will be in charge of the construction of a service request and a second step will analyse the response got from the execution of the service, allowing the extraction of facts contained in that response and mapping them to domain-specific concepts from the ontologies used within FAST by any building block.

\subsection{Constructing service requests} % (fold)
\label{sub:constructing_service_requests}

As a first approach, the interaction with these services will be limited to retrieve information to feed the gadgets using simply GET requests. A service request would be assembled using a certain URL and a set of parameters. As an example, Ebay Shopping web service will be studied. The following URL is invoked to retrieve a list of items corresponding to certain search keywords:

\url{http://open.api.sandbox.ebay.com/shopping?appid=KasselUn-efea-4b93-9505-5dc2ef1ceecd&version=517&callname=FindItems&ItemSort=EndTime&QueryKeywords=USB&responseencoding=XML}

As you may see, the URL invoked is http://open.api.sandbox.ebay.com/shopping and the parameters used in the example are:
\begin{description}
	\item[appid] this is the application ID obtained to use the API.
	\item[version] the API version.
	\item[callname] in this case FindItems to search through all items in Ebay.
	\item[itemsort] sorting method for the list of items.
	\item[querykeywords] list of keywords.
	\item[responseencoding] format of the response message obtained by the invocation of the request.
\end{description}

If needed, a detailed specification of the Ebay Shopping API can be found at \cite{eBayShoppingAPIs}.

The above URL for searching items in Ebay is followed by the query parameters, which take the form \textit{argument=value}, where the arguments and values are URL encoded, and are separated by an ampersand (\&). For instance, the only relevant parameter the user would need to specify is \emph{querykeywords}, thus somehow the service has to receive an input value for it, while the other parameters can be set to a default value. However, in order to develop a generic wrapper for the any service, all parameters might be set by the user.

To achieve this, service wrappers will be handled as pseudo-screens with preconditions and postconditions (inputs and outputs). Therefore, these precondition ports might be used to determine values to parameters like query keywords coming from a building block inside the screen the service is placed or even an external screen as a screen precondition. The service wrapper tool composes the request using a template string which will contain placeholders for precondition values. Before sending the request to the service, the placeholders are replaced with their corresponding values and the resulting URL is then ready to be sent the service request.

Figure ~\ref{fig:complex_gadget_architecture} shows the screen to construct the service requests.

\begin{figure}
  \begin{center}
    %\includegraphics[width=\linewidth]{images/ComplexGadgetArchitecture.pdf}
    \caption{...}
    \label{fig:construct_service_request}
  \end{center}
\end{figure}

In the upper part of Figure ~\ref{fig:complex_gadget_architecture}, the user establishes the name of the Resource Adapter and which parameters or input ports the wrapper is composed of. For each input port, the user provides a name, a type, and an example value. The example value is just to simulate the behaviour of the wrapper. While constructing the request template, these placeholders have to be surrounded by <>. As you see in the above example, the request contains <query\_keywords> in the position its value needs to be placed.

\subsubsection{Limitations} % (fold)
\label{ssub:limitations}

It is worth pointing out that currently the wrapper tool is able to construct input ports for the wrappers using just basic types. However, every building block accepts any (complex) concept as a pre/postcondition, so the wrapper tool needs to be adapted to permit any concept as input or precondition, extracting the required text to construct the request.

% subsubsection limitations (end)

% subsection constructing_service_requests (end)

\subsection{Interpreting service responses} % (fold)
\label{sub:interpreting_service_responses}

Once the service request is constructed and sent to the service provider, it will send back a response. This response message could be serialize in any format, though the most common formats used nowadays are XML or JSON among others. To continue the example it was started in the previous section, the response of the Ebay Shopping service will be in XML as specified in the request, which is the format supported by the wrapper tool.

\subsubsection{Translation XML into Facts} % (fold)
\label{ssub:translation_xml_into_facts}

Figure ~\ref{fig:response_service_execution} shows the response obtained after the execution of the service request.

\begin{figure}
  \begin{center}
    %\includegraphics[width=\linewidth]{images/ComplexGadgetArchitecture.pdf}
    \caption{...}
    \label{fig:response_service_execution}
  \end{center}
\end{figure}

Once the service response, in XML format, has been retrieved, it is displayed both in plain text (debugging/test purposes) and as an interactive object. To construct this interactive object, the XML document has to been parsed into a DOM, and a simplified tree representation of that DOM (see lower left part in Figure ~\ref{fig:response_service_execution}) is built up. This tree representation of the XML data is will be used as an input to construct transformation rules.

A transformation rule is used to analyse the XML data and to generate domain-specific facts from concepts from the ontologies used by the pre/postconditions of the different building blocks. A transformation rule is composed of three elements. First, from indicates the origin of the rule, which will be the tagname of a DOM element from the XML document. Second, the type of the rule will be set, taking one of the following values: \emph{createObject}, \emph{fillAttributes} or \emph{dummy}. And third, the target of the rule specifies a certain concept or attribute, to be created or filled. A detailed explanation of the type of actions to be trigger from the transformation rules is:
\begin{description}
	\item[action \emph{createObject}] specifies the creation of a new fact object. The type of that new fact is provided in the third compartment. In the example is being explained, the root rule searches for XML elements with tagname \emph{FindItemsResponse} and for each such element a \emph{List} fact is created. The resulting facts are shown in a facts tree in the lower right of Figure ~\ref{fig:response_service_execution}.
	\item[action \emph{fillAttributes}] does not create a new fact but it fills the value of the attribute provided as third part of such rules. In our example, the third transformation rule searches for XML elements with tagname \emph{Title}. Note, the rule is a sub-rule of the second rule, which generates \emph{Product} facts. Thus, the sub-rule searches for \emph{Title} tags only in the subtree of the XML data that has been identified by an application of the parent rule before. For example, the \emph{Item} rule may just have been applied to the first \emph{Item} element of the XML data. Then, the \emph{Title} rule is applied only to the first \emph{Item} sub-tree of the XML data and thus it will find only one \emph{Title} element in that sub-tree (not visible in Figure ~\ref{fig:response_service_execution} ). The value of that \emph{Title} element is then transfered to the \emph{productName} attribute of the corresponding \emph{Product} fact.
	\item[action \emph{dummy}] does not create or modify any facts but such rules are just used to narrow the search space for their sub-rules. For example, in the Amazon case, the XML data for an item contains sections for \emph{minimum price}, \emph{maximum price}, and \emph{average price}. Each such section contains the \emph{plain price} and the \emph{formatted price}. Thus, in the Amazon case, a rule that searches for \emph{formatted price} elements within an \emph{Item} element would retrieve three matches. Using a dummy rule, we may first search for \emph{minimum price} elements and then search for \emph{formatted price} elements within that sub-tree.
\end{description}

Since FAST is storyboard oriented, the service wrapper designer follows the storyboard idea as well. Any time, a change to a transformation rule is done, the transformation process is triggered and the resulted facts tree is directly shown. This process helps the user to deal with the slightly complex semantics of the transformation rules avoiding errors or mistakes. In addition, FAST is semantic-driven, therefore, the service wrapper designer shall retrieve the domain-specific types from a FAST ontology server together with the structure of each type, i.e. together with a description of the attributes of each fact. Thus, the transformation rule editor is able to provide selection boxes for the target element of the rules. For a \emph{createObject} rule, this selection box shows the fact types available for that domain. For the \emph{fillAttributes} rules, the selection box shows the attributes of the fact type chosen in the parent rule. In addition, we may provide some analysis tool, which will help to guarantee that the facts generated by the transformation rules conform to the fact types defined in the corresponding FAST ontology. This helps to ensure that the facts generated by the designed pseudo screen will be compatible for precondition ports of subsequent filter steps and or screens.

% subsubsection translation_xml_into_facts (end)

% subsection interpreting_service_responses (end)

\subsection{Generating a Resource Adapter} % (fold)
\label{sub:generating_a_resource_adapter}

Once the wrapping of a service has been defined and tested in the service wrapper tool, it shall generate an implementation of the desired Resource Adapter in XML, HTML, and JavaScript, ready to be deployed and executed inside a gadget. This service wrapper implementation is compliant with the formats required by the Gadget Visual Storyboard Tool (GVS) and it shall be stored inside the FAST catalogue, in order to be found and used by any user.

% subsection generating_a_resource_adapter (end)

\subsection{Limitations} % (fold)
\label{sub:limitations}

The rule driven approach presented above is somewhat limited. It is deliberately restricted to such a simple rule mechanism in order to keep things simple enough for end-users. Still, the selected approach suffices for most practical and real world cases. However, there are some transformations that these rules cannot perform. For example, the XML data for a person may provide two different tags for the first and the last name of a person. Contrarily, a person fact which conforms to a certain ontology for that domain may provide only one \emph{fullname} attribute that shall be filled by a concatenation of the first and the last name. To cover such cases directly requires rather complex extensions of our transformation rules. To avoid this, intermediate fact formats can be used which would allow generating person facts with separated first and last name attributes. Then, additional filters may be used allowing the combination of the first and last name in a single \emph{fullname} attribute. However, this can be the role of the operators, since some generic operators can act as filters for aggregation and conversions of facts from multiple sources. Then, service wrappers in combination with these filter operators will allow covering these complex cases.

% subsection limitations (end)

% section restful_web_services_wrapper_tool (end)

\clearpage
\section{Related Work} % (fold)
\label{sec:related_work}

Write about some related work such as:

Wrapping WSDL-Described Web Services as Moby Services (using SAWSDL)
http://biomoby.open-bio.org/CVS_CONTENT/moby-live/Java/docs/sawsdlServlet.html

% section related_work (end)

\clearpage
\bibliographystyle{apalike}
\addcontentsline{toc}{section}{References}
\bibliography{fast_d4_3_2}

\clearpage
\doublespacing
\section*{Appendix A (Lists of Tables and Figures)}
\addcontentsline{toc}{section}{Appendix A (Lists of Tables and Figures)}

\listoftables

\listoffigures

\end{document}

### 2010-01-22 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_part1.tex
Added outlines of Kassel's contribution to come
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@445 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 Note that several roles can be played by the same actor.
-\begin{inparaenum}[(i)]
+\begin{enumerate}[(i)]
     \item The \textit{ontology engineer} creates the ontologies used to annotate services and data.
 This role also includes the process of ontology matching, either automated or manually, determining if the alignment is feasible and creating so-called \emph{matching operator} building blocks, which are basic elements of the FAST screen building.
-The \textit{resource developer} then uses these ontologies to annotate resources created in FAST.
-    \item Ontology matching is needed by the \textit{screen developer} at the design-time of a screen (a visual building block of a gadget).
-Screen developers have a dedicated UI component for building screens, in which they can use the matching operators to combine components annotated with different ontologies.
+The \textit{resource developer} then uses these ontologies to annotate resources created in FAST, which resources will be used by the \textit{screen developer}, who will also specify the resource type at the pre- and postconditions.
+    \item Ontology matching is needed by the \textit{gadget developer} at the design-time of a screen-flow (gadget).
+Gadget developers use the GVS (Gadget Visual Storyboard) for building gadgets, in which they can use the matching operators to combine screens having input and output resources annotated with different ontologies.
 No actual matching needs to be performed in this phase, but rather the possibility of matching needs to be determined (i.e., \emph{can} two screens A and B be combined?).
 %the FAST user interface determines if two resources can be combined or not (see Figure \ref{fig:screens} for a conceptual example of resources that can not be combined).
-    \item The \emph{gadget developer} combines screens to screen-flows and gadgets, and only uses ontology matching implicitly.
+    %\item The \emph{gadget developer} combines screens to screen-flows and gadgets, and only uses ontology matching implicitly.
     \item The \textit{end-user} uses the final deployed gadget at run-time, but is unaware of the underlying resources and ontologies or the matching process.
@@
 ontology engineering and screen development).
-\end{inparaenum}
+\end{enumerate}
 
@@
 
+\subsection{Service Wrapping Tool}
+\textbf{\textit{Kassel will describe the service wrapping tool they've developed, that creates mapping instances between ontologies. 
+A description of the front-end and the creation of the mapping rules will be given here.}}
+
+\subsection{Creating the Operator}
+
+\textbf{\textit{Kassel will describe the creation of the matching operator in the catalogue and the storage of the mapping rules.}}
+
 \section{Gadget Building Phase}
 Every time ontology matching is performed and an alignment is created between two underlying resources, the corresponding alignment rules are stored and an operator is created in the catalogue.
+
 At the gadget building phase, whenever two screens are combined, the catalogue will provide the possible matching operators for the two screens.
-This way an operator can be placed between two screens that would otherwise be incompatible.
+This way an operator can be placed between two screens that would otherwise be incompatible, making interoperation between the two possible.
 
 \section{Runtime} 
-Employing alignment rules in the runtime phase of the gadget.
+At runtime, the matching operator's rules are tranformed into JSON data used by the JavaScript gadget.
+Thus, when data come from the two disparate resources, the gadget will automatically transform data instances of the ontology belonging to the first resource, into instances of the ontology belonging to the second resource.
+
+\textbf{\textit{Kassel will describe the way the mapping rules are transformed into JSON and how they are used within the JavaScript gadget.}}

### 2010-01-22 trunk/GVS/media/js/gvs/dialogs/ManageScreensDialog.js
Updates in the gallery css
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@446 d8b0c59b-6544-0410-94d9-83a5658a058e

@@ -98,7 +98,7 @@ var ManageScreensDialog = Class.create(GalleryDialog /** @lends ManageScreensDia
                                 '<span class="bold">Version: </span>' +
                                      screen.version,
                                 '<span class="bold">Tags: </span>' +
-                                    data.domainContext.tags.join(","),
+                                    data.domainContext.tags.join(", "),
                                 '<span class="bold">Description </span><br />'+
                                     data.description['en-gb'],
                                 '<span class=' + (screen.uri ? '"shared"': '"unshared"') +

### 2010-01-22 trunk/GVS/media/js/gvs/documents/ScreenDocument.js
Allow resizing of screen document areas
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@447 d8b0c59b-6544-0410-94d9-83a5658a058e

@@ -35,12 +35,12 @@ var ScreenDocument = Class.create(PaletteDocument,
                                 getBuildingBlockFactory(Constants.BuildingBlock.DOMAIN_CONCEPT));
         
         // Dropping areas
-        var formArea = new Area('form', $A([Constants.BuildingBlock.FORM]), this._drop.bind(this));
-        var operatorArea = new Area('operator', $A([Constants.BuildingBlock.OPERATOR]), this._drop.bind(this));
-        var resourceArea = new Area('resource', $A([Constants.BuildingBlock.RESOURCE]), this._drop.bind(this));
-        var preArea = new Area('pre', $A([Constants.BuildingBlock.DOMAIN_CONCEPT]), this._drop.bind(this));
-        var postArea = new Area('post', $A([Constants.BuildingBlock.DOMAIN_CONCEPT]), this._drop.bind(this));
-        
+        var formArea = new Area('form', $A([Constants.BuildingBlock.FORM]), this._drop.bind(this), {splitter: true, region: 'top', minHeight:300});
+        var operatorArea = new Area('operator', $A([Constants.BuildingBlock.OPERATOR]), this._drop.bind(this), {splitter: true, region: 'center', minHeight:200, minWidth:200});
+        var resourceArea = new Area('resource', $A([Constants.BuildingBlock.RESOURCE]), this._drop.bind(this), {splitter: true, region: 'bottom', minHeight:100});
+        var preArea = new Area('pre', $A([Constants.BuildingBlock.DOMAIN_CONCEPT]), this._drop.bind(this), {splitter: true, region: 'left', minWidth:100});
+        var postArea = new Area('post', $A([Constants.BuildingBlock.DOMAIN_CONCEPT]), this._drop.bind(this), {splitter: true, region: 'right', minWidth:100});
+
         var areas = $A([formArea, operatorArea, resourceArea, preArea, postArea]);
         
         $super(title, $A([formSet, operatorSet, resourceSet, domainConceptSet]),
@@ -49,8 +49,19 @@ var ScreenDocument = Class.create(PaletteDocument,
         
         // Adding the dropping areas to the document
         areas.each(function(area) {
-            this._tabContent.appendChild(area.getNode());     
+            var contentPane = area.getWidget();
+            this._screenDesignContainer.addChild(contentPane);
+
+            var contentNode = area.getNode();
+
+            contentNode.observe('click', function() {
+                this._onClick();
+            }.bind(this));
[…]

### 2010-01-22 trunk/GVS/media/js/gvs/dragndrop/Area.js

@@ -5,8 +5,12 @@ var Area = Class.create( /** @lends Area.prototype */ {
      * @param Function onDropHandler(DropZone zone, ComponentInstance droppedInstance)
      * @constructs
      */ 
-    initialize: function(/** String */ areaClass, /** Array */ acceptedElements, /** Function */ onDropHandler) {
-        
+    initialize: function(/** String */ areaClass, /** Array */ acceptedElements, /** Function */ onDropHandler, /** Object */ _options) {
+        var options = Utils.variableOrDefault(_options, {});
+        options = Object.extend ({
+            style : ""
+        }, options);
+
         /**
          * List of valid elements to be dropped in the area
          * @type Array
@@ -23,7 +27,23 @@ var Area = Class.create( /** @lends Area.prototype */ {
          * @private
          */
         this._onDropHandler = onDropHandler;
-        
+
+        if (options.region != 'center') {
+            if (options['minHeight'] != null) {
+                options.style += "height: " + options['minHeight'] + 'px;';
+            }
+            if (options['minWidth'] != null) {
+                options.style += "width: " + options['minWidth'] + 'px;';
+            }
+        }
+
+        /**
+         * ContentPane of the area
+         * @type dijit.layout.ContentPane
+         * @private
+         */
+        this._contentPane = new dijit.layout.ContentPane(options);
+
         /**
          * DOM Node of the area
          * @type DOMNode
@@ -32,7 +52,7 @@ var Area = Class.create( /** @lends Area.prototype */ {
         this._node = new Element('div', {
             'class': 'dropArea ' + areaClass
         });
-        
+        this._contentPane.setContent(this._node);
     },
     
 
@@ -43,10 +63,17 @@ var Area = Class.create( /** @lends Area.prototype */ {
      * @type DOMNode
      */
     getNode: function() {
-        return this._node;    
+        return this._node;
     },
     
-    
+    /**
+     * Returns the  widget
+     * @type dijit.*
+     */
+    getWidget: function () {
+        return this._contentPane;
+    },
+
     /**
      * Implementing DropZone Interface: drop
      * @type DOMNode

### 2010-01-22 trunk/GVS/settings.py

@@ -19,6 +19,7 @@ ADMINS = (
 MANAGERS = ADMINS
 
 DATABASE_ENGINE = 'mysql'           # 'postgresql_psycopg2', 'postgresql', 'mysql', 'sqlite3' or 'oracle'.
+DATABASE_OPTIONS = {"init_command": "SET storage_engine=InnoDB"}
 DATABASE_NAME = 'fast'             # Or path to database file if using sqlite3.
 DATABASE_USER = 'fast'             # Not used with sqlite3.
 DATABASE_PASSWORD = 'fast'         # Not used with sqlite3.
@@ -124,7 +125,7 @@ AUTH_PROFILE_MODULE = 'user.UserProfile'
 NOT_PROXY_FOR = ['localhost', '127.0.0.1']
 
 # Url to the server that holds the semantic catalogue
-CATALOGUE_URL = 'http://localhost:8080/catalogue/'
+CATALOGUE_URL = 'http://localhost:8080/Catalogue-0.1-dev/'
 
 # Url to the server that holds the storage service (empty if local storage in GVS)
 STORAGE_URL = None

### 2010-01-22 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_mediation.tex
Added newest version of USGS article contents
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@442 d8b0c59b-6544-0410-94d9-83a5658a058e
Final rough version before Albert's contribution (still needs a final read-through)
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@448 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \usetikzlibrary{plotmarks}
+
+\usepackage{listings} %source code listings
+\usepackage{paralist}
+\usepackage{booktabs}
 
@@
 \section{Introduction} %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\label{sec:introduction}
+\label{sec:intro}
 \paragraph{Goal and Scope.}
@@
 Resources in FAST are described semantically using different ontologies and vocabularies, and can therefore be combined to what we can call ``intelligent gadgets''.
-In the services created within the FAST environment, the ontologies come from different parties involved in the creation of gadgets and building blocks.
+In the services used within the FAST environment, the ontologies involved in the creation of gadgets and building blocks come from different parties.
 These ontologies, though different, will sometimes cover the same domain, making ontology matching necessary, as the means to reconcile differences in the various conceptualisations.
@@
 Part \ref{part:d2.4.2} describes the approach of integrating ontology matching into the FAST environment.
+Section \ref{sec:overview} gives the big picture of ontology matching in fast, which the remaining three sections expand.
+Thus, Section \ref{sec:ontologyengineering} describes a tool to automatically discover alignments, a scenario and example ontologies to demonstrate the capabilites of the tool, and the testing procedure with the results.
+The remaing part of the section deals with a tool to define the discovered mapping rules in a way that is consistent with other similar tasks within the FAST platform, and describes the way in which matching operators are stored within the Semantic Catalogue.
+Section \ref{sec:gadgetbuilding} shortly describes the way the matching operator is used when composing screenflows and Section \ref{sec:runtime} describes how the matching operator and the defined rules are converted and used in the final, deployed gadget.
+
 %Describe the rest.
@@
 \label{part:d2.4.1}
-%\include{fast_d2_4_2_part2}
+\include{fast_d2_4_2_part2}
 
@@
 
-In this second iteration we established the approach to integrate ontology matching into FAST, and have shown the implementation details as well as test scenarios.
-We have presented a tool for finding the alignments between ontologies, as well as a powerful language to represent them, beyond the capabilities of the tool.
-We have shown how this matching is done at the different phases of the gadget life cycle and the approach taken at every level.
+In this second iteration we established the approach to integrate ontology matching into FAST, and we have shown how this matching is done at the different phases of the gadget life cycle and the approach taken at every level.
+Firstly, have shown the implementation details as well as test scenarios, we have presented a tool for finding the alignments between ontologies as well as a powerful language to represent them, we have described a tool to compose mapping rules in a consistent manner with other FAST components and the means to save them in the Semantic Catalogue.
+Secondly, we described the way in which mapping operators are used in the GVS to combine discrepant screens.
+Lastly, we presented the way the matching operator and mapping rules are deployed and exploited in the JavaScript gadget.
 
-We plan to evaluate the feasibility of ontology matching in real-life tests and finalize it for the completed version of the FAST platform.
+For the last and final iteration of deliverable D2.4 we plan to evaluate real-life scenarios and finalize the implementation of ontology matching for the FAST platform.
 

### 2010-01-22 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_part1.tex
Final rough version before Albert's contribution (still needs a final read-through)
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@448 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \section{Overview}
+\label{sec:overview}
+
 \subsection{Ontology Matching}
@@
 \section{Ontology Engineering Phase}
+\label{sec:ontologyengineering}
 
@@
 \section{Gadget Building Phase}
+\label{sec:gadgetbuilding}
 Every time ontology matching is performed and an alignment is created between two underlying resources, the corresponding alignment rules are stored and an operator is created in the catalogue.
@@
 \section{Runtime} 
+\label{sec:runtime}
 At runtime, the matching operator's rules are tranformed into JSON data used by the JavaScript gadget.

### 2010-01-22 trunk/GVS/media/js/gvs/documents/ScreenDocument.js
Solved a bug when repainting pipes
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@449 d8b0c59b-6544-0410-94d9-83a5658a058e

@@ -35,11 +35,26 @@ var ScreenDocument = Class.create(PaletteDocument,
                                 getBuildingBlockFactory(Constants.BuildingBlock.DOMAIN_CONCEPT));
         
         // Dropping areas
-        var formArea = new Area('form', $A([Constants.BuildingBlock.FORM]), this._drop.bind(this), {splitter: true, region: 'top', minHeight:300});
-        var operatorArea = new Area('operator', $A([Constants.BuildingBlock.OPERATOR]), this._drop.bind(this), {splitter: true, region: 'center', minHeight:200, minWidth:200});
-        var resourceArea = new Area('resource', $A([Constants.BuildingBlock.RESOURCE]), this._drop.bind(this), {splitter: true, region: 'bottom', minHeight:100});
-        var preArea = new Area('pre', $A([Constants.BuildingBlock.DOMAIN_CONCEPT]), this._drop.bind(this), {splitter: true, region: 'left', minWidth:100});
-        var postArea = new Area('post', $A([Constants.BuildingBlock.DOMAIN_CONCEPT]), this._drop.bind(this), {splitter: true, region: 'right', minWidth:100});
+        var formArea = new Area('form',
+                                $A([Constants.BuildingBlock.FORM]),
+                                this._drop.bind(this),
+                                {splitter: true, region: 'top', minHeight:300});
+        var operatorArea = new Area('operator',
+                                    $A([Constants.BuildingBlock.OPERATOR]),
+                                    this._drop.bind(this),
+                                    {splitter: true, region: 'center', minHeight:200, minWidth:200});
+        var resourceArea = new Area('resource',
+                                    $A([Constants.BuildingBlock.RESOURCE]),
+                                    this._drop.bind(this),
+                                    {splitter: true, region: 'bottom', minHeight:100});
+        var preArea = new Area('pre',
+                                $A([Constants.BuildingBlock.DOMAIN_CONCEPT]),
+                                this._drop.bind(this),
+                                {splitter: true, region: 'left', minWidth:100});
+        var postArea = new Area('post',
+                                $A([Constants.BuildingBlock.DOMAIN_CONCEPT]),
+                                this._drop.bind(this),
+                                {splitter: true, region: 'right', minWidth:100});
 
         var areas = $A([formArea, operatorArea, resourceArea, preArea, postArea]);
         
@@ -462,8 +477,11 @@ var ScreenDocument = Class.create(PaletteDocument,
      * @private
      */
     _repaint: function() {
[…]

### 2010-01-22 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_mediation.tex
Final version before Albert's input
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@451 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 Part \ref{part:d2.4.2} describes the approach of integrating ontology matching into the FAST environment.
-Section \ref{sec:overview} gives the big picture of ontology matching in fast, which the remaining three sections expand.
-Thus, Section \ref{sec:ontologyengineering} describes a tool to automatically discover alignments, a scenario and example ontologies to demonstrate the capabilites of the tool, and the testing procedure with the results.
+Sect.~\ref{sec:overview} gives the big picture of ontology matching in fast, which the remaining three sections expand.
+Thus, Sect.~\ref{sec:ontologyengineering} describes a tool to automatically discover alignments, a scenario and example ontologies to demonstrate the capabilites of the tool, and the testing procedure with the results.
 The remaing part of the section deals with a tool to define the discovered mapping rules in a way that is consistent with other similar tasks within the FAST platform, and describes the way in which matching operators are stored within the Semantic Catalogue.
-Section \ref{sec:gadgetbuilding} shortly describes the way the matching operator is used when composing screenflows and Section \ref{sec:runtime} describes how the matching operator and the defined rules are converted and used in the final, deployed gadget.
+Sect.~\ref{sec:gadgetbuilding} shortly describes the way the matching operator is used when composing screenflows and Sect.~\ref{sec:runtime} describes how the matching operator and the defined rules are converted and used in the final, deployed gadget.
 
@@
 Part \ref{part:d2.4.1} contains the first iteration of this document, D2.4.1, laying down the theoretical foundations of ontology matching.
-We discuss ontologies in general in Section \ref{sec:ontologies}, present the ontology mediation problem in Section \ref{sec:mediationproblem} , discuss the types  of mismatches between ontologies in Section \ref{sec:ontologymismatches} and detail the main approaches to ontology mediation in Section \ref{sec:mediationapproaches}.
+We discuss ontologies in general in Sect.~\ref{sec:ontologies}, present the ontology mediation problem in Sect.~\ref{sec:mediationproblem} , discuss the types  of mismatches between ontologies in Sect.~\ref{sec:ontologymismatches} and detail the main approaches to ontology mediation in Sect.~\ref{sec:mediationapproaches}.
 

### 2010-01-22 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_part1.tex
Final version before Albert's input
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@451 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 It uses the ontology mapping language described in~\cite{scharffe2005language}.
-It can describe complex correspondences, such as the one detailed in Section \ref{sub:manual_mapping}.
+It can describe complex correspondences, such as the one detailed in Sect.~\ref{sub:manual_mapping}.
 \end{inparaenum}
@@
     \item The \textit{ontology engineer} creates the ontologies used to annotate services and data.
-This role also includes the process of ontology matching, either automated or manually, determining if the alignment is feasible and creating so-called \emph{matching operator} building blocks, which are basic elements of the FAST screen building.
-The \textit{resource developer} then uses these ontologies to annotate resources created in FAST, which resources will be used by the \textit{screen developer}, who will also specify the resource type at the pre- and postconditions.
+This role also includes the process of ontology matching, either automated or manually, determining if the alignment is feasible and creating so-called \emph{matching operator} building blocks, which are basic elements of the FAST gadget building.
+The \textit{resource developer} then uses these ontologies to annotate resources created in FAST, which resources will be used by the \textit{screen developer}, who will also specify the resource type at the pre- and postconditions of a screen.
     \item Ontology matching is needed by the \textit{gadget developer} at the design-time of a screen-flow (gadget).
-Gadget developers use the GVS (Gadget Visual Storyboard) for building gadgets, in which they can use the matching operators to combine screens having input and output resources annotated with different ontologies.
+Gadget developers use the GVS (Gadget Visual Storyboard) for building gadgets, where they can use the matching operators to combine screens having input and output conditions specifying resources annotated with different ontologies.
 No actual matching needs to be performed in this phase, but rather the possibility of matching needs to be determined (i.e., \emph{can} two screens A and B be combined?).
@@
 Only at run-time the actual mapping of instance data has to be performed.
-In this paper, we mainly consider the first two cases (i.e.
-ontology engineering and screen development).
 \end{enumerate}
@@
 Based on these requirements, we compared three different tools. 
-\emph{MAFRA}~\cite{maedche2002mafra} supports an interactive and incremental process of ontology mapping.
+\paragraph{MAFRA} \emph{MAFRA}~\cite{maedche2002mafra} supports an interactive and incremental process of ontology mapping.
 It provides an explicit notion of semantic bridges.
@@
 The bridges, however, have been designed to be used within the MAFRA system, and the alignment process needs to be done through the provided GUI.
-\emph{RDFT}~\cite{omelayenko2002rdft} is a small language originally designed to map between XML and RDF.
+
+\paragraph{RDFT} \emph{RDFT}~\cite{omelayenko2002rdft} is a small language originally designed to map between XML and RDF.
 The results are mappings represented in DAML+OIL, that can be executed in a transformation process.
 No hints are given to add alignment methods or extending the format and the tool does not longer seem to be available.
-\emph{Alignment API}~\cite{euzenat2004api} is the tool best matching our requirements, satisfying all the desired conditions.
+\paragraph{Alignment API} \emph{Alignment API}~\cite{euzenat2004api} is the tool best matching our requirements, satisfying all the desired conditions.
 It is still under active development, provides an API and its implementation, is open source (GPLv2 or above) and written in Java, providing an easy way to embed it into other programs.
@@
 
-Though the infeasibility of automating this alignment became obvious, we have represented the alignment in the mapping language supported by the tool, as a level 2 mapping (described in Sect.
-\ref{mediation}).
+Though the infeasibility of automating this alignment became obvious, we have represented the alignment in the mapping language supported by the tool, as a level 2 mapping (described in Sect.~\ref{mediation}).
  % (the mapping has been created between Amazon and GoodRelations, since the eBay and Amazon ontologies are very similar in approach, and mapping eBay to GoodRelations would be an indentical task). 

### 2010-01-25 trunk/deliverables/D2_2_Ontology/fast_d2_2_2_ontology.tex
common tag added
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@463 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2009-12-11 trunk/deliverables/D2_2_Ontology/fast_d2_2_ontology.tex
@@
 \usepackage{amssymb}
+\usepackage{listings}
+
+\lstdefinelanguage{turtle} 
+{morekeywords={@prefix, a}, 
+sensitive=false, 
+morecomment=[l]{\#}, 
+morestring=[b]",
+}
+
+\lstdefinelanguage{sparql} 
+{morekeywords={PREFIX, a, SELECT, CONSTRUCT, WHERE, UNION, OPTIONAL, FILTER, DISTINCT, ORDER, BY, GRAPH}, 
+sensitive=false, 
+morecomment=[l]{\#}, 
+morestring=[b]",
+}
+
+\lstset{
+	captionpos=b, 
+	extendedchars=true,
+	basicstyle=\footnotesize \sffamily, 
+	stringstyle=\bfseries,
+	frame=single,
+	frameround=tttt,
+	showstringspaces=false,
+	breaklines=true,
+	% inputencoding=utf8,
+}
+
 \parindent0pt
 
-\newcommand\deliverableNumber{D2.2.1}
+\newcommand\deliverableNumber{D2.2.2}
 \newcommand\deliverableTitle{Ontology and Conceptual Model for the Semantic Characterisation of Complex Gadgets}
@@
 		\vfill
-		{\large \sffamily Date: 27/02/2009}\\
+		{\large \sffamily Date: 27/02/2010}\\
 		\vspace{1cm}
@@
 %0.2 & 15.02.2009 & Knud M\"oller (NUIG) & content moved from wiki after internal review \\ \hline
-1.0 & 27.02.2009 & Knud M\"oller (NUIG) & final version ready for external review \\ \hline
+1.0 & 27.02.2009 & Knud M\"oller (NUIG) & final version (D2.2.1) ready for external review \\ \hline
+2.0 & 27.02.2010 & Knud M\"oller (NUIG) & final version (D2.2.2) ready for external review \\ \hline
 \end{tabular}
@@
 	\hline
-	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2009} & {\textbf{Submitted}} & {27/02/2009}\\ \hline
+	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2010} & {\textbf{Submitted}} & {27/02/2010}\\ \hline
 	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
 	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
-	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.2.1 is the first iteration of the FAST Gadget ontology.
+	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{\deliverableNumber{} is the second iteration of the FAST Gadget ontology.
 The ontology provides a formal description of the conceptual model of FAST in general, and of the screen/screenflow architecture in particular.
@@
 
-While following an ontology design methodology which is agnostic to particular data models or implementation languages (see below), we have chosen to represent the final ontology specification using the Resource Description Framework (RDF) and OWL semantics.
+While following an ontology design methodology which is agnostic to particular data models or implementation languages (see Appendix A), we have chosen to represent the final ontology specification using the Resource Description Framework (RDF) and OWL semantics.
 
@@
 Following, we will conclude the introduction by elaborating more on some of the general design decisions we are following.
-After briefly addressing the topic of related work in Sect.~\ref{sec:related_work} --- which is mainly covered in a different deliverable from this work package, \cite{urmetzer2009fast_state_of_the_art} --- the structure of the remainder of this document follows directly from our adopted methodology: we will perform a domain analysis in Sect.~\ref{sec:domain_analysis}, covering the stages of specification, conceptualisation and integration of existing vocabularies.
+After briefly addressing the topic of related work in Sect.~\ref{sec:related_work} --- which is mainly covered in a different deliverable from this work package, \cite{urmetzer2010fast_state_of_the_art} --- the structure of the remainder of this document follows directly from our adopted methodology: we will perform a domain analysis in Sect.~\ref{sec:domain_analysis}, covering the stages of specification, conceptualisation and integration of existing vocabularies.
 This is then followed by the implementation step in Sect.~\ref{sec:ontology}, which covers the ontology in its final form (at this moment of the project).
@@
 
-\subsection{\emph{Methontology}: A Methodology for Ontology Development} % (fold)
-\label{sub:methontology_a_methodology_for_ontology_development}
+\subsection{Changes from Previous Version} % (fold)
+\label{sub:changes_from_previous_version}
 
-In order to put the design of the ontology on a stable footing, we chose to adopt a tried and tested design methodology, rather than using a purely ad-hoc approach.
-We decided to use the \emph{Methontology} methodology, which was first introduced in \cite{gomez_perez1996conceptualise_ontologies} and \cite{fernandez1997methontology}.
-Methontology provides a good balance of formalisation and streamlining of the design process on the one hand, and a looseness in its requirements on how strictly one needs to follow the individual phases and steps.
-The authors specifically advocate an ``evolving prototype life cycle'', which allows the ontology to ``grow depending on its needs'' \cite{fernandez1997methontology}.
+The following changes have been made with respect to the previous version of this deliverable, D2.2.1.~\cite{moeller2009fast_ontology}:
 
-Methontology consists of seven different stages, which will be briefly described in the following sections.
-For a more detailed discussion, we point the reader to the original papers proposing this methodology.
-The phases are \emph{specification}, \emph{knowledge acquisistion}, \emph{conceptualisation}, \emph{integration}, \emph{implementation}, \emph{evaluation} and \emph{documentation}.
-It is important to note that the phases do not have to be visited slavishly in the order presented here.
-Instead, a much more likely scenario is that the ontology designers will visit each phase roughly in this order, but are free to revisit each phase at any time to apply changes.
+\begin{itemize}
+	\item Sect.~\ref{sub:goal_and_scope} has been revised to better define the scope of this deliverable with respect to other deliverables.
+	\item Sect.~1.3 (Methontology) has been moved to the end of the document (Sect.~\ref{sub:methontology_a_methodology_for_ontology_development}), but otherwise did not change.
+	\item The section on design principles (Sect.~\ref{sub:general_design_decisions}) has been extended to reflect research done by the authors within the past year.
+	\item In the integration section (Sect.~\ref{sub:integration}) of the ontology definition, the Common Tag vocabulary has been added. It is briefly introduced, and the integrated terms are specified.
+\end{itemize}
 
-\subsubsection{Specification} % (fold)
-\label{ssub:specification}
+% subsection changes_from_previous_version (end)
 
-The specification phase sets the stage for the rest of the ontology development process.
-The central activity in this phase is the setting up of an \emph{ontology requirement specification document}, which will act as a guideline for most of the other phases.
-The level of formality of the specification document is left open and can range from pure natural language, over a set of intermediate representations to using competency questions.
-Most likely, it will be a mix of those formats.
-Regardless of the chosen format of the specification document, it must answer questions regarding the \emph{purpose} of the ontology (intended use and users, scenarios, etc.), its \emph{scope} (what are the things that need to be covered), its \emph{level of formality} (e.g.
-highly informal, semi-formal or rigorously formal \cite{uschold1996ontologies}) and the \emph{sources of knowledge} used when researching the ontology.
-It should be noted that the scope as defined in the specification document does not need to be complete.
-Rather, it should provide a good impression of the kind of terms that need to be covered by the ontology.
+\subsection{Usability as a General Design Principles} % (fold)
+\label{sub:general_design_decisions}
 
-The \emph{Domain Analysis} section of this deliverable implements an ontology requirement specification document for the FAST gadget ontology (which is not to be confused with the FAST requirements deliverable \cite{villoslada2009fast_requirements}!).
-% subsubsection specification (end)
-
-\subsubsection{Knowledge Acquisition} % (fold)
-\label{ssub:knowledge_acquisition}
-
-Knowledge Acquisition in the context of Methontology is a loose collective term to all activities which are aimed at gathering background knowledge to clearly determine purpose and scope of the ontology.
-As such, it feeds directly into the specification phase, but can just as well be relevant at later stages of the design and development process.
-Similarly to evaluation and documentation, knowledge acquisition is a supporting activity, which takes place throughout the whole development process.
-
-Typical ways of knowledge acquisition are referring to handbooks, encyclopaedias or other written material (through formal or informal text analysis), performing interviews with domain experts (structured or non-structured) or evaluating previous conceptualisation work done for the domain in question.
-
-The primary sources used for knowledge acquisition for the the FAST gadget ontology are referred to in Sect.~\ref{sec:related_work}.
-% subsubsection knowledge_acquisition (end)
-
-\subsubsection{Conceptualisation} % (fold)
-\label{ssub:conceptualisation}
-
-The conceptualisation phase stands at the core of the ontology development process.
-The ontology is still in a language-/format-independent state, but is becoming increasingly formal.
-Based on the specification document, the ontology designers now create a number of increasingly detailed and fine-grained tables and dictionaries which cover all terms to be included in the ontology (at the current stage - an ontology in the Semantic Web sense is never complete).
-In the first iteration, a so-called \emph{glossary of terms} (GT) is built, which includes all concepts (or classes), instances and properties (or verbs).
-The GT does not have to be built from scratch, but can instead be understood as an extension of the scope definition in the specification phase.
-In contrast to the specification document, the GT is a living document that should always reflect the latest and most complete list of terms.
-Once completed (in a first iteration), the terms in the GT will be grouped according to relatedness.
-For each group, the concepts are arranged in a classification tree.
-Instances, constants and attributes are grouped in corresponding tables, while properties are captured in verb diagrams.
-If necessary, tables with formulas and rules are set up at the end of the conceptualisation phase.
-% subsubsection conceptualisation (end)
-
-\subsubsection{Integration} % (fold)
-\label{ssub:integration}
-
-The integration phase offers an opportunity for the ontology designer to ensure that their ontology is well integrated with other, existing ontologies.
-This can either mean connecting the emerging ontology to upper or meta ontologies (in the case that there are appropriate super classes or properties in those ontologies which can act as connection joints), or the reuse of terms from other ontologies (in the case that other ontologies have already defined classes or properties which fall in the scope of the emerging ontology).
-A typical example for a meta-ontology is OpenCyc, while the FOAF vocabulary \cite{brickley2004foaf} is a example of an often-reused ontology to describe people.
-As the output of this phase, an \emph{integration document} is suggested, which gives detailed information about which terms were taken from which external ontology.
-% subsubsection integration (end)
-
-\subsubsection{Implementation} % (fold)
-\label{ssub:implementation}
-
-The implementation phase is what is often (wrongly so) considered to be the main activity in developing an ontology: materialising the various ontology terms in a concrete language, which can be RDFS or OWL for the Semantic Web, or any other formal language in other contexts (even a programming language such as C++).
-The authors of Methontology do not go into a lot of details regarding the carrying out of this phase, other than that the outcome will be the formalisation of the ontology in the ontology language of choice (the \emph{implementation document}).
-% subsubsection implementation (end)
-
-\subsubsection{Evaluation} % (fold)
-\label{ssub:evaluation}
-
-According to [Fernández et  al., 1997], evaluation ``means to carry out a technical judgment of the ontologies, their software environment and documentation with respect to a frame of reference'' (in this case the requirement specification document).
-Evaluation can take place at any time during the ontology development process, and comprises activities such as \emph{validation} and \emph{verification}.
-
-For the FAST Gadget Ontology, most of the evaluation process will take place as part of the work carried out in WP6 (Experimentation and Evaluation) in years two and three of the project.
-% subsubsection evaluation (end)
-
-\subsubsection{Documentation} % (fold)
-\label{ssub:documentation}
-
-In Methontology, documentation is an activity which is automatically carried out throughout the development process, in the form of the various documents which form the output of the individual phases.
-Since this deliverable contains the current versions of all those documents, it is at the same time the ontology itself as well as its documentation.
-% subsubsection documentation (end)
-
-% subsection methontology_a_methodology_for_ontology_development (end)
-
-\subsection{General Design Decisions} % (fold)
-\label{sub:general_design_decisions}
+Apart from ``hard'' metrics for the evaluation of ontologies, so-called ``soft'' metrics for ontologies are relevant for facilitating better understanding by users and implementers, and therefore ultimately greater uptake.
 
@@
 
-All of these references are discussed in \cite{urmetzer2009fast_state_of_the_art}, and we refer the reader to this deliverable for more detail.
+All of these references are discussed in \cite{urmetzer2010fast_state_of_the_art}, and we refer the reader to this deliverable for more detail.
 % section related_work (end)
@@
 \textbf{Properties:} \emph{containsScreen}, \emph{hasLabel}, \emph{hasIcon}, \emph{hasPreCondition}, \emph{hasPostCondition}, \emph{hasTag}, \emph{hasProfile}\\ \hline
-\textcolor{white}{\textbf{Sources of Knowledge}} & FAST Architecture deliverable~\cite{solero2009fast_architecture}, FAST Requirements Specification~\cite{villoslada2009fast_requirements}, FAST State-of-the-art Document~\cite{urmetzer2009fast_state_of_the_art}, EzWeb documentation~\cite{lizcano2008ezweb} \\ \hline
+\textcolor{white}{\textbf{Sources of Knowledge}} & FAST Architecture deliverable~\cite{solero2009fast_architecture}, FAST Requirements Specification~\cite{villoslada2009fast_requirements}, FAST State-of-the-art Document~\cite{urmetzer2010fast_state_of_the_art}, EzWeb documentation~\cite{lizcano2008ezweb} \\ \hline
 \end{longtable}
@@
 	\item \texttt{Image} - Images are any kind of graphical object, such as icons, screenshots, etc.
+	\item \texttt{Tag} - A short (usually one word) description of a resource.
 \end{itemize}
@@
 
+\subsubsection{Common Tag} % (fold)
+\label{ssub:common_tag}
+
+Common Tag\footnote{\url{http://www.commontag.org}, 24/01/2010} is an open format and initiative for conceptual tagging, which is developed and backed by a number of relevant players in the Web community, including Yahoo!
+and DERI.
+The basic idea of conceptual tagging is to address the problem of ambiguity in the meaning of tags (e.g., does ``jaguar'' mean the car, the animal or the operating system?) as it occurs in free-text tagging.
+Rather than relying on methods such as clustering, applied to a whole corpus of tagged resources, conceptual tagging intends to disambiguate a tag right from the moment it is used, by linking it to an unambiguous concept from vocabularies such as DBpedia (representing Wikipedia as linked data) or Freebase.
+E.g., in order to disambiguate ``cocoa'' as meaning the API, not the drink or plant, it would be represented as a resource (rather than a plain literal) and linked to the DBpedia resource \url{http://dbpedia.org/resource/Cocoa_%28API%29}.
+
+Apart from facilitating the disambiguation of tags, Common Tag also allows to make further assertions about the nature of a tag, such as who applied it or when it was applied.
+As an example, List.~\ref{list:common_tag} shows how a blog post has been tagged on 20/01/2010 with ``cocoa'', meaning the API, not the drink or plant.
+
+\singlespacing
+\lstset{
+	caption={Common Tag example --- disambiguating the tag ``cocoa''}, 
+	label=list:common_tag,
+	language=turtle
+}
+\begin{figure}[ht]
+	\lstinputlisting{exampleCode/common_tag_example.ttl}
+\end{figure}
+\doublespacing
+
+Within FAST, Common Tag is used to cover the tagging needs as specified in the ontology requirements specification, and in particular in order to define the domain context as specified in the annotation part of the glossary of terms (\texttt{Tag} and \texttt{hasDomainContext}).
+
+% subsubsection common_tag (end)
+
 \subsubsection{Integration Document} % (fold)
@@
 \texttt{Image} & FOAF & \texttt{foaf:Image} & --- \\ \hline
+\texttt{Tag} & Common Tag & \texttt{ctag:Tag} & Complex or conceptual tags are being used e.g. to model the domain context of resources.	 \\ \hline
 \multicolumn{4}{|c|}{\emph{\textbf{Properties}}} \\ \hline
@@
 \\ \hline
-\texttt{hasInterests} & FOAF & \texttt{interest} & --- \\ \hline
-\texttt{hasUserName} & FOAF & \texttt{accountName} & --- \\ \hline
+\texttt{hasInterests} & FOAF & \texttt{foaf:interest} & --- \\ \hline
+\texttt{hasUserName} & FOAF & \texttt{foaf:accountName} & --- \\ \hline
+\texttt{hasDomainContext} & Common Tag & \texttt{ctag:tagged} & The domain context of a resource in FAST can be expressed as a complex tag.
+Further Common Tag properties can then be applied to make assertions about the tag itself.
+\\ \hline
 \end{longtable}
@@
 
+\subsection{Instance Examples} % (fold)
+\label{sub:instance_examples}
+
+\subsubsection{Complete Screen with Opaque Code} % (fold)
+\label{ssub:complete_screen}
+
+\singlespacing
+\lstset{
+	caption={Complete Screen with Opaque Code}, 
+	label=list:screen_01
+}
+% \setlength\parindent{0in}
+% \begin{minipage}[t]{\textwidth}
+\lstinputlisting{exampleCode/01_screenExample.ttl}
+% \end{minipage}
+% \setlength\parindent{0.21in}
+\doublespacing
+
+% subsubsection complete_screen (end)
+
+\subsubsection{Complete Screen with Declarative Definition} % (fold)
+\label{ssub:complete_screen_with_declarative_definition}
+
+\singlespacing
+\lstset{
+	caption={Complete Screen with Declarative Definition}, 
+	label=list:screen_01
+}
+% \setlength\parindent{0in}
+% \begin{minipage}[t]{\textwidth}
+\lstinputlisting{exampleCode/02_screenExample.ttl}
+% \end{minipage}
+% \setlength\parindent{0.21in}
+\doublespacing
+% subsubsection complete_screen_with_declarative_definition (end)
+
+% subsection instance_examples (end)
+
 \subsection{Classes} % (fold)
@@
 
-\clearpage
-\bibliographystyle{apalike}
-\addcontentsline{toc}{section}{References}
-\bibliography{fast_d2_2_bib}
+% \clearpage
+% \phantomsection
+% \section*{Appendix A (Instance Examples)}
+% \addcontentsline{toc}{section}{Appendix A (Instance Examples)}
+%
 
 \clearpage
-\section*{Appendix A (Ontology Code)}
-\addcontentsline{toc}{section}{Appendix A (Ontology Code)}
+\phantomsection
+\appendix
+
+\section{Development Methodology}
+% \addcontentsline{toc}{section}{Development Methodology}
+
+\subsection{\emph{Methontology}: A Methodology for Ontology Development} % (fold)
+\label{sub:methontology_a_methodology_for_ontology_development}
+
+In order to put the design of the ontology on a stable footing, we chose to adopt a tried and tested design methodology, rather than using a purely ad-hoc approach.
+We decided to use the \emph{Methontology} methodology, which was first introduced in \cite{gomez_perez1996conceptualise_ontologies} and \cite{fernandez1997methontology}.
+Methontology provides a good balance of formalisation and streamlining of the design process on the one hand, and a looseness in its requirements on how strictly one needs to follow the individual phases and steps.
+The authors specifically advocate an ``evolving prototype life cycle'', which allows the ontology to ``grow depending on its needs'' \cite{fernandez1997methontology}.
+
+Methontology consists of seven different stages, which will be briefly described in the following sections.
+For a more detailed discussion, we point the reader to the original papers proposing this methodology.
+The phases are \emph{specification}, \emph{knowledge acquisistion}, \emph{conceptualisation}, \emph{integration}, \emph{implementation}, \emph{evaluation} and \emph{documentation}.
+It is important to note that the phases do not have to be visited slavishly in the order presented here.
+Instead, a much more likely scenario is that the ontology designers will visit each phase roughly in this order, but are free to revisit each phase at any time to apply changes.
+
+\subsubsection{Specification} % (fold)
+\label{ssub:specification}
+
+The specification phase sets the stage for the rest of the ontology development process.
+The central activity in this phase is the setting up of an \emph{ontology requirement specification document}, which will act as a guideline for most of the other phases.
+The level of formality of the specification document is left open and can range from pure natural language, over a set of intermediate representations to using competency questions.
+Most likely, it will be a mix of those formats.
+Regardless of the chosen format of the specification document, it must answer questions regarding the \emph{purpose} of the ontology (intended use and users, scenarios, etc.), its \emph{scope} (what are the things that need to be covered), its \emph{level of formality} (e.g.
+highly informal, semi-formal or rigorously formal \cite{uschold1996ontologies}) and the \emph{sources of knowledge} used when researching the ontology.
+It should be noted that the scope as defined in the specification document does not need to be complete.
+Rather, it should provide a good impression of the kind of terms that need to be covered by the ontology.
+
+The \emph{Domain Analysis} section of this deliverable implements an ontology requirement specification document for the FAST gadget ontology (which is not to be confused with the FAST requirements deliverable \cite{villoslada2009fast_requirements}!).
+% subsubsection specification (end)
+
+\subsubsection{Knowledge Acquisition} % (fold)
+\label{ssub:knowledge_acquisition}
+
+Knowledge Acquisition in the context of Methontology is a loose collective term to all activities which are aimed at gathering background knowledge to clearly determine purpose and scope of the ontology.
+As such, it feeds directly into the specification phase, but can just as well be relevant at later stages of the design and development process.
+Similarly to evaluation and documentation, knowledge acquisition is a supporting activity, which takes place throughout the whole development process.
+
+Typical ways of knowledge acquisition are referring to handbooks, encyclopaedias or other written material (through formal or informal text analysis), performing interviews with domain experts (structured or non-structured) or evaluating previous conceptualisation work done for the domain in question.
+
+The primary sources used for knowledge acquisition for the the FAST gadget ontology are referred to in Sect.~\ref{sec:related_work}.
+% subsubsection knowledge_acquisition (end)
+
+\subsubsection{Conceptualisation} % (fold)
+\label{ssub:conceptualisation}
+
+The conceptualisation phase stands at the core of the ontology development process.
+The ontology is still in a language-/format-independent state, but is becoming increasingly formal.
+Based on the specification document, the ontology designers now create a number of increasingly detailed and fine-grained tables and dictionaries which cover all terms to be included in the ontology (at the current stage - an ontology in the Semantic Web sense is never complete).
+In the first iteration, a so-called \emph{glossary of terms} (GT) is built, which includes all concepts (or classes), instances and properties (or verbs).
+The GT does not have to be built from scratch, but can instead be understood as an extension of the scope definition in the specification phase.
+In contrast to the specification document, the GT is a living document that should always reflect the latest and most complete list of terms.
+Once completed (in a first iteration), the terms in the GT will be grouped according to relatedness.
+For each group, the concepts are arranged in a classification tree.
+Instances, constants and attributes are grouped in corresponding tables, while properties are captured in verb diagrams.
+If necessary, tables with formulas and rules are set up at the end of the conceptualisation phase.
+% subsubsection conceptualisation (end)
+
+\subsubsection{Integration} % (fold)
+\label{ssub:integration}
+
+The integration phase offers an opportunity for the ontology designer to ensure that their ontology is well integrated with other, existing ontologies.
+This can either mean connecting the emerging ontology to upper or meta ontologies (in the case that there are appropriate super classes or properties in those ontologies which can act as connection joints), or the reuse of terms from other ontologies (in the case that other ontologies have already defined classes or properties which fall in the scope of the emerging ontology).
+A typical example for a meta-ontology is OpenCyc, while the FOAF vocabulary \cite{brickley2004foaf} is a example of an often-reused ontology to describe people.
+As the output of this phase, an \emph{integration document} is suggested, which gives detailed information about which terms were taken from which external ontology.
+% subsubsection integration (end)
+
+\subsubsection{Implementation} % (fold)
+\label{ssub:implementation}
+
+The implementation phase is what is often (wrongly so) considered to be the main activity in developing an ontology: materialising the various ontology terms in a concrete language, which can be RDFS or OWL for the Semantic Web, or any other formal language in other contexts (even a programming language such as C++).
+The authors of Methontology do not go into a lot of details regarding the carrying out of this phase, other than that the outcome will be the formalisation of the ontology in the ontology language of choice (the \emph{implementation document}).
+% subsubsection implementation (end)
+
+\subsubsection{Evaluation} % (fold)
+\label{ssub:evaluation}
+
+According to [Fernández et  al., 1997], evaluation ``means to carry out a technical judgment of the ontologies, their software environment and documentation with respect to a frame of reference'' (in this case the requirement specification document).
+Evaluation can take place at any time during the ontology development process, and comprises activities such as \emph{validation} and \emph{verification}.
+
+For the FAST Gadget Ontology, most of the evaluation process will take place as part of the work carried out in WP6 (Experimentation and Evaluation) in years two and three of the project.
+% subsubsection evaluation (end)
+
+\subsubsection{Documentation} % (fold)
+\label{ssub:documentation}
+
+In Methontology, documentation is an activity which is automatically carried out throughout the development process, in the form of the various documents which form the output of the individual phases.
+Since this deliverable contains the current versions of all those documents, it is at the same time the ontology itself as well as its documentation.
+% subsubsection documentation (end)
+
+% subsection methontology_a_methodology_for_ontology_development (end)
+
+\clearpage
+\phantomsection
+\section{Ontology Overview}
+% \addcontentsline{toc}{section}{Ontology Overview}
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=.82\linewidth]{images/fgo_uml_diagram}
+    \caption{FAST Gadget Ontology Class Hierarchy}
+    \label{fig:ontology_uml}
+  \end{center}
+\end{figure}
+
+\clearpage
+\phantomsection
+\section{Ontology Code}
+% \addcontentsline{toc}{section}{Ontology Code}
 
@@
 \begin{small}
-\verbatiminput{ontology/fast_gadget.owl}
+\verbatiminput{fast_gadget.owl}
 \end{small}
@@
 \doublespacing
-\section*{Appendix B (Lists of Tables and Figures)}
-\addcontentsline{toc}{section}{Appendix B (Lists of Tables and Figures)}
+\phantomsection
+\section*{Lists of Tables and Figures}
+\addcontentsline{toc}{section}{Lists of Tables and Figures}
 
 \listoftables
-
 \listoffigures
 
+\clearpage
+\phantomsection
+\bibliographystyle{apalike}
+\addcontentsline{toc}{section}{References}
+\bibliography{fast_d2_2_bib}
+
 \end{document}

### 2010-01-25 trunk/catalogue_service/src/eu/morfeoproject/fast/catalogue/ontologies/DefaultOntologies.java
minor changes
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@464 d8b0c59b-6544-0410-94d9-83a5658a058e

@@ -45,22 +45,22 @@ public class DefaultOntologies {
     	new Ontology(new URIImpl("http://commontag.org/ns#"), "ctag.owl", Syntax.RdfXml, true);
 
     /**-- The DBPedia ontology (no datasets included) --**/
-    public static PublicOntology DBPEDIA = 
-    	new PublicOntology(new URIImpl("http://dbpedia.org/ontology/#"),
-    			"http://downloads.dbpedia.org/3.2/en/dbpedia-ontology.owl",
-    			Syntax.RdfXml,
-    			true);
+//    public static PublicOntology DBPEDIA = 
+//    	new PublicOntology(new URIImpl("http://dbpedia.org/ontology/#"),
+//    			"http://downloads.dbpedia.org/3.2/en/dbpedia-ontology.owl",
+//    			Syntax.RdfXml,
+//    			true);
 
     /**-- Ontologies for test purposes --**/
     public static Ontology AMAZON_MOCKUP =
     	new Ontology(new URIImpl("http://aws.amazon.com/AWSECommerceService#"), "amazon-mockup.rdf", Syntax.RdfXml, true);
     public static Ontology DEMO =
     	new Ontology(new URIImpl("http://www.morfeoproject.eu/fast/demo#"), "demo.rdf", Syntax.RdfXml, true);
-    public static PublicOntology GR =
-    	new PublicOntology(new URIImpl("http://purl.org/goodrelations/v1#"),
-    			"http://www.heppnetz.de/ontologies/goodrelations/v1.owl",
-    			Syntax.RdfXml,
-    			true);
+//    public static PublicOntology GR =
+//    	new PublicOntology(new URIImpl("http://purl.org/goodrelations/v1#"),
+//    			"http://www.heppnetz.de/ontologies/goodrelations/v1.owl",
+//    			Syntax.RdfXml,
+//    			true);
     
     
     /**

### 2010-01-25 trunk/catalogue_service/src/eu/morfeoproject/fast/vocabulary/FGO.java

@@ -4,7 +4,7 @@ import org.ontoware.rdf2go.model.node.URI;
 import org.ontoware.rdf2go.model.node.impl.URIImpl;
 
 /**
- * Vocabulary File. Created by org.ontoware.rdf2go.util.VocabularyWriter on Wed Dec 16 16:01:15 GMT 2009
+ * Vocabulary File. Created by org.ontoware.rdf2go.util.VocabularyWriter on Fri Jan 22 13:22:40 GMT 2010
  * input file: src/eu/morfeoproject/fast/catalogue/ontologies/fgo.rdf
  * namespace: http://purl.oclc.org/fast/ontology/gadget#
  */
@@ -456,6 +456,14 @@ public interface FGO {
      * Range: http://xmlns.com/foaf/0.1/Document 
      */
     public static final URI hasUri = new URIImpl("http://purl.oclc.org/fast/ontology/gadget#hasUri", false);
+
+    /**
+     * Label: tag creator@en 
+     * Comment: The creator of the tag.@en 
+     * Comment: http://commontag.org/ns#Tag 
+     * Range: http://xmlns.com/foaf/0.1/Agent 
+     */
+    public static final URI tagCreator = new URIImpl("http://purl.oclc.org/fast/ontology/gadget#tagCreator", false);
 
     /**
      * Label: has pre-condition@en 

### 2010-01-25 trunk/GVS/media/js/gvs/catalogue/factories/BuildingBlockFactory.js
First steps on opening screens
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@466 d8b0c59b-6544-0410-94d9-83a5658a058e

@@ -36,9 +36,19 @@ var BuildingBlockFactory = Class.create(
     /**
      * Gets building block descriptions by URI
      * @type {BuildingBlockDescription[]}
+     * @abstract
      */
     getBuildingBlocks: function (/*...*/) {
         throw "Abstract method invocation. BuildingBlockFactory::getBuildingBlockDescriptions";
+    },
+
+    /**
+     * Gets a new instance of the type of the factory
+     * @abstract
+     * @type ComponentInstance
+     */
+    getInstance: function(/** BuildingBlockDescription */description, /** InferenceEngine */ engine) {
+        throw "Abstract method invocation. BuildingBlockFactory::getInstance";
     }
 
 });

### 2010-01-25 trunk/GVS/media/js/gvs/catalogue/factories/FormFactory.js

@@ -75,6 +75,15 @@ var FormFactory = Class.create(BuildingBlockFactory,
             callback();
         }       
     },
+
+    /**
+     * Gets a new instance of the type of the factory
+     * @override
+     * @type FormInstance
+     */
+    getInstance: function(/** BuildingBlockDescription */description, /** InferenceEngine */ engine) {
+        return new FormInstance(description, engine);
+    },
     
     // **************** PRIVATE METHODS **************** //
     

### 2010-01-25 trunk/GVS/media/js/gvs/catalogue/factories/OperatorFactory.js

@@ -75,6 +75,15 @@ var OperatorFactory = Class.create(BuildingBlockFactory,
             callback();
         }       
     },
+
+    /**
+     * Gets a new instance of the type of the factory
+     * @override
+     * @type OperatorInstance
+     */
+    getInstance: function(/** BuildingBlockDescription */description, /** InferenceEngine */ engine) {
+        return new OperatorInstance(description, engine);
+    },
     
     // **************** PRIVATE METHODS **************** //
     

### 2010-01-25 trunk/GVS/media/js/gvs/catalogue/factories/ResourceFactory.js

@@ -47,6 +47,15 @@ var ResourceFactory = Class.create(BuildingBlockFactory,
         }.bind(this));
         return result;
     },
+
+    /**
+     * Gets a new instance of the type of the factory
+     * @override
+     * @type ResourceInstance
+     */
+    getInstance: function(/** BuildingBlockDescription */description, /** InferenceEngine */ engine) {
+        return new ResourceInstance(description, engine);
+    },
     
     /**
      * This function retrieves the pending elements from the serverside

### 2010-01-25 trunk/GVS/media/js/gvs/catalogue/factories/ScreenFactory.js

@@ -75,6 +75,15 @@ var ScreenFactory = Class.create(BuildingBlockFactory,
             callback();
         }       
     },
+
+    /**
+     * Gets a new instance of the type of the factory
+     * @override
+     * @type ScreenInstance
+     */
+    getInstance: function(/** BuildingBlockDescription */description, /** InferenceEngine */ engine) {
+        return new ScreenInstance(description, engine);
+    },
     
     // **************** PRIVATE METHODS **************** //
     

### 2010-01-25 trunk/GVS/media/js/gvs/catalogue/inferenceEngine/InferenceEngine.js

@@ -29,11 +29,11 @@ var InferenceEngine = Class.create( /** @lends InferenceEngine.prototype */ {
      * callback upon completion.
      */
     findCheck: function (/**Array*/ canvas, /** Array*/ elements,
-                         /** Array */ domainContext, 
+                         /** Array */ tags,
                          /** String*/ criteria,
                          /** Function */ callback) {
         
-        var body = this._constructBody(canvas, elements, domainContext, criteria);
+        var body = this._constructBody(canvas, elements, tags, criteria);
         var persistenceEngine = PersistenceEngineFactory.getInstance();
         var context = {
             'callback': callback,
@@ -47,10 +47,10 @@ var InferenceEngine = Class.create( /** @lends InferenceEngine.prototype */ {
      * This function calls the check operation in the catalogue
      */
     check: function (/**Array*/ canvas, /** Object */ elements,
-                    /** Array */ domainContext, 
+                    /** Array */ tags,
                     /** String*/ criteria, 
                     /** Function */ callback) {
-        var body = this._constructBody(canvas, elements, domainContext, criteria, "check");
+        var body = this._constructBody(canvas, elements, tags, criteria, "check");
         var persistenceEngine = PersistenceEngineFactory.getInstance();
         var context = {
             'callback': callback,
@@ -189,6 +189,10 @@ var InferenceEngine = Class.create( /** @lends InferenceEngine.prototype */ {
      * @private
      */
     _onError: function(transport, e){
+        Utils.showMessage("Error loading building blocks. Maybe is it not working?", {
+            'error': true,
+            'hide': true
+        });
         Logger.serverError(transport,e);
     },
     

### 2010-01-25 trunk/GVS/media/js/gvs/catalogue/inferenceEngine/ScreenInferenceEngine.js

@@ -23,16 +23,16 @@ var ScreenInferenceEngine = Class.create( /** @lends ScreenInferenceEngine.proto
      * @type String
      */
     _constructBody: function(/**Array*/ canvas, /** Object */ elements,
-                    /** Array */ domainContext, 
+                    /** Array */ tags,
                     /** String*/ criteria, /** String(Optional) */ _method) {
         var method = Utils.variableOrDefault(_method, "");
-        var domain = {
-            'tags': domainContext,
+        var domainContext = {
+            'tags': tags,
             'user': null /* TODO: add user here */
         };
         var body = {
             'canvas': canvas,
-            'domainContext': domain,
+            'domainContext': domainContext,
             'criterion': criteria
         };
         if (method == "check") {

### 2010-01-25 trunk/GVS/media/js/gvs/catalogue/inferenceEngine/ScreenflowInferenceEngine.js

@@ -48,16 +48,16 @@ var ScreenflowInferenceEngine = Class.create(InferenceEngine /** @lends Screenfl
      * @type String
      */
     _constructBody: function(/**Array*/ canvas, /** Array*/ elements,
-                    /** Array */ domainContext, 
+                    /** Array */ tags,
                     /** String*/ criteria) {
-        var domain = {
-            'tags': domainContext,
+        var domainContext = {
+            'tags': tags,
             'user': null /* TODO: add user here */
         };
         var body = {
             'canvas': canvas,
             'elements': elements,
-            'domainContext': domain,
+            'domainContext': domainContext,
             'criterion': criteria
         };
         return Object.toJSON(body);

### 2010-01-25 trunk/GVS/media/js/gvs/dialogs/GalleryDialog.js

@@ -30,6 +30,7 @@ var GalleryDialog = Class.create(FormDialog, /** @lends GalleryDialog.prototype 
         // Assigning the passed parameters, or defaults
         this._properties.set('showTitleRow', (properties.showTitleRow || false));
         this._properties.set('elementsPerPage', (properties.elementsPerPage || 10));
+        this._properties.set('onDblClick', (properties.onDblClick || null));
 
         /**
          * Table fields
@@ -147,9 +148,12 @@ var GalleryDialog = Class.create(FormDialog, /** @lends GalleryDialog.prototype 
                 element.addClassName("selected");
                 
                 this._selectedRow = row;
-                // TODO
-                // this._showButtons();
             }.bind(this));
+            if (this._properties.get('onDblClick')) {
+                rowNode.observe('dblclick', function() {
+                    this._properties.get('onDblClick')(row.key);
+                }.bind(this));
+            }
             content.appendChild(rowNode);
         }.bind(this));
         if (loadAll && this._rows.size() > 0) {

### 2010-01-25 trunk/GVS/media/js/gvs/dialogs/ManageScreensDialog.js

@@ -6,7 +6,7 @@ var ManageScreensDialog = Class.create(GalleryDialog /** @lends ManageScreensDia
      * @extends GalleryDialog
      */ 
     initialize: function($super) {  
-        $super("Screen browser");
+        $super("Screen browser", {'onDblClick': this._openScreen.bind(this)});
 
         /**
          * List of screens

### 2010-01-25 trunk/GVS/media/js/gvs/dialogs/NewScreenDialog.js

@@ -64,11 +64,11 @@ var NewScreenDialog = Class.create(ConfirmDialog /** @lends NewScreenDialog.prot
     _onOk: function($super){
         if (this._getFormWidget().validate()) {
             var name = $F(this._getForm().name);
-            var domainContext = $F(this._getForm().domaincontext).split(/[\s,]+/).without("");
+            var tags = $F(this._getForm().domaincontext).split(/[\s,]+/).without("");
             var version = $F(this._getForm().version);
             
             var documentController = GVSSingleton.getInstance().getDocumentController();
-            documentController.createScreen(name, $A(domainContext), version);
+            documentController.createScreen(name, $A(tags), version);
             
             $super();
         }       

### 2010-01-25 trunk/GVS/media/js/gvs/dialogs/PropertiesDialog.js

@@ -6,7 +6,8 @@ var PropertiesDialog = Class.create(ConfirmDialog /** @lends PropertiesDialog.pr
      * @extends ConfirmDialog
      */ 
     initialize: function($super, /** String */ type,
-                        /** BuildingBlockDescription */ description) {
+                        /** BuildingBlockDescription */ description,
+                        /** Function */ onChangeHandler) {
         
         /**
          * Title of the dialog
@@ -25,9 +26,16 @@ var PropertiesDialog = Class.create(ConfirmDialog /** @lends PropertiesDialog.pr
         /**
          * Handler to be called when the dialog finishes
          * @private
-         * @type String
+         * @type Function
          */
-        this._handler = null;
+        this._onFinishHandler = null;
+
+        /**
+         * Handler to be called when the dialog finishes with changes
+         * @private
+         * @type Function
+         */
+        this._onChangeHandler = onChangeHandler;
 
         $super(this._title);
         
@@ -41,7 +49,7 @@ var PropertiesDialog = Class.create(ConfirmDialog /** @lends PropertiesDialog.pr
      */
     show: function($super, /** Function(Optional) */ _handler) {
         var handler = Utils.variableOrDefault(_handler, null);      
-        this._handler = handler;
+        this._onFinishHandler = handler;
         $super();
     },
 
@@ -118,11 +126,21 @@ var PropertiesDialog = Class.create(ConfirmDialog /** @lends PropertiesDialog.pr
             var elements = [form.name, form.version, form.creator, form.rights,
                 form.icon, form.screenshot, form.homepage];
             Object.extend(description, Form.serializeElements(elements, {'hash': true}));
+            var changed = false;
+            $H(description).each(function(pair) {
+                if (this._description[pair.key] != pair.value) {
+                    changed = true;
+                }
+            }.bind(this));
+
             this._description.addProperties(description);
             $super();
-            if (this._handler && this._handler instanceof Function) {
-                this._handler();
-                this._handler = null;
+            if (changed) {
+                this._onChangeHandler();
+            }
+            if (this._onFinishHandler && this._onFinishHandler instanceof Function) {
+                this._onFinishHandler();
+                this._onFinishHandler = null;
             }
         }
     },

### 2010-01-25 trunk/GVS/media/js/gvs/documents/AbstractDocument.js

@@ -122,6 +122,15 @@ var AbstractDocument = Class.create(ToolbarModel, /** @lends AbstractDocument.pr
         var gvs = GVSSingleton.getInstance();
         gvs.getDocumentController().closeDocument(this._tabId);
         return true;
+    },
+
+    /**
+     * Sets the document title
+     * @private
+     */
+    _setTitle: function(/** String */ title) {
+        this._title = title;
+        this._tab.attr("title", this._title);
     }
 });
 

### 2010-01-25 trunk/GVS/media/js/gvs/documents/DocumentController.js

@@ -82,11 +82,14 @@ var DocumentController = Class.create(
     /**
      * Creates a new screen document
      */
-    createScreen: function(/** String */ name, /** String */ domainContext,
+    createScreen: function(/** String */ name, /** Array */ tags,
                             /** String */ version){
         var screen = new ScreenDocument({
                 'name': name,
-                'domainContext': domainContext,
+                'domainContext': {
+                    'user': null,
+                    'tags': tags
+                },
                 'version': version
             });
         this.addDocument(screen);
@@ -96,7 +99,9 @@ var DocumentController = Class.create(
      * Opens an existing screen by its id
      */
     loadScreen: function(/** String */ id) {
-        
+        var persistenceEngine = PersistenceEngineFactory.getInstance();
+        var uri = URIs.buildingblock + id;
+        persistenceEngine.sendGet(uri, this, this._onScreenLoadSuccess, this._onLoadError);
     },
     
     /**
@@ -233,6 +238,24 @@ var DocumentController = Class.create(
      */
     _onKeyPressed: function(/** String */ key) {
         this._currentDocument.onKeyPressed(key);
+    },
+
+    /**
+     * On screen load success
+     * @private
+     */
+    _onScreenLoadSuccess: function (/** XMLHttpRequest */ transport) {
+        var screenData = JSON.parse(transport.responseText);
+        var screen = new ScreenDocument(screenData);
+        this.addDocument(screen);
+        screen.loadInstances();
+    },
+
+    _onLoadError: function() {
+        Utils.showMessage("Can not open the selected element", {
+            'error': true,
+            'hide': true
+        });
     }
 });
 

### 2010-01-25 trunk/GVS/media/js/gvs/documents/PaletteDocument.js

@@ -12,14 +12,14 @@ var PaletteDocument = Class.create(AbstractDocument, /** @lends PaletteDocument.
             /** String */ title, 
             /** Array */ buildingBlockSets,
             /** Array */ dropZones,
-            /** Array */ domainContext, /** InferenceEngine */ inferenceEngine) {
+            /** Array */ tags, /** InferenceEngine */ inferenceEngine) {
         $super(title);
         
         /**
          * List of tags
          * @type Array
          */
-        this._domainContext = domainContext;
+        this._tags = tags;
         
         /**
          * Main border container

### 2010-01-25 trunk/GVS/media/js/gvs/documents/ScreenCanvasCache.js

@@ -1,0 +1,154 @@
+var ScreenCanvasCache = Class.create( /** @lends ScreenCanvasCache.prototype */ {
+    /**
+     * 
+     * @constructs
+     */ 
+    initialize: function (/** Object */ properties) {
+        /**
+         * Buildingblocks of the screen
+         * @type Array
+         * @private
+         */
+        this._buildingblocks = properties.definition.buildingblocks;
+
+        /**
+         * Pipes of the screen
+         * @type Array
+         * @private
+         */
+        this._pipes = properties.definition.pipes;
+
+        /**
+         * Triggers of the screen
+         * @type Array
+         * @private
+         */
+        this._triggers = properties.definition.triggers;
+
+        /**
+         * Preconditions of the screen
+         * @type Array
+         * @private
+         */
+        this._preconditions = properties.preconditions;
+
+        /**
+         * Postconditions of the screen
+         * @type Array
+         * @private
+         */
+        this._postconditions = properties.definition.buildingblocks;
+
+        /**
+         * Array of already loaded building block types
+         * @type Array
+         * @private
+         */
+        this._elementsLoaded = new Array();
+    },
+  
+    
+    // **************** PRIVATE METHODS **************** //
+    /**
+     * Gets the uri of the form (if any) in form of Array, to
+     * get compatibility with factory method
+     * @type Array
+     */
+    getFormURI: function () {
+        var form = this._buildingblocks.detect(function(element) {
+            return (element.uri.search(/FormElement/i) != -1);
+        });
+        if (form) {
+            return [form.uri];
+        } else {
+            return [];
+        }
+        
+    },
+
+    /**
+     * Returns the list of operator URIs
+     * @type Array
+     */
+    getOperatorURIs: function() {
+        var elements = this._buildingblocks.findAll(function(element) {
+            return (element.uri.search(/Operator/i) != -1);
+        });
+        var result = new Array();
+        elements.each(function(element) {
+            if (result.indexOf(element.uri) == -1) {
+                result.push(element.uri);
+            }
+        });
+        return result;
+    },
+
+    /**
+     * Returns the list of resource URIs
+     * @type Array
+     */
+    getResourceURIs: function() {
+        var elements = this._buildingblocks.findAll(function(element) {
+            return (element.uri.search(/BackendService/i) != -1);
+        });
+        var result = new Array();
+        elements.each(function(element) {
[…]

### 2010-01-25 trunk/GVS/media/js/gvs/documents/ScreenDocument.js

@@ -7,7 +7,7 @@ var ScreenDocument = Class.create(PaletteDocument,
      * @extends PaletteDocument
      */
     initialize: function($super, /** Object */ properties) {
-        var domainContext = properties.domainContext;
+        var tags = properties.domainContext.tags;
         var name = properties.name;
         var version = properties.version;
 
@@ -29,13 +29,13 @@ var ScreenDocument = Class.create(PaletteDocument,
         var catalogue = CatalogueSingleton.getInstance();
 
         // Palette sets
-        var formSet = new BuildingBlockSet(domainContext, catalogue.
+        var formSet = new BuildingBlockSet(tags, catalogue.
                             getBuildingBlockFactory(Constants.BuildingBlock.FORM));
-        var operatorSet = new BuildingBlockSet(domainContext, catalogue.
+        var operatorSet = new BuildingBlockSet(tags, catalogue.
                                 getBuildingBlockFactory(Constants.BuildingBlock.OPERATOR));
-        var resourceSet = new BuildingBlockSet(domainContext, catalogue.
+        var resourceSet = new BuildingBlockSet(tags, catalogue.
                                 getBuildingBlockFactory(Constants.BuildingBlock.RESOURCE));
-        var domainConceptSet = new DomainConceptSet(domainContext, catalogue.
+        var domainConceptSet = new DomainConceptSet(tags, catalogue.
                                 getBuildingBlockFactory(Constants.BuildingBlock.DOMAIN_CONCEPT));
         
         // Dropping areas
@@ -60,14 +60,25 @@ var ScreenDocument = Class.create(PaletteDocument,
                                 this._drop.bind(this),
                                 {splitter: true, region: 'right', minWidth:100});
 
-        var areas = $A([formArea, operatorArea, resourceArea, preArea, postArea]);
+        /**
+         * Areas of the canvas
+         * @type Hash
+         * @private
+         */
+        this._areas = $H({
+            'form': formArea,
+            'operator': operatorArea,
+            'resource': resourceArea,
+            'pre': preArea,
+            'post': postArea
+        });
         
         $super(name, $A([formSet, operatorSet, resourceSet, domainConceptSet]),
-                areas,
-                domainContext,  new ScreenInferenceEngine());
+                this._areas.values(),
+                tags,  new ScreenInferenceEngine());
         
         // Adding the dropping areas to the document
-        areas.each(function(area) {
+        this._areas.values().each(function(area) {
             var contentPane = area.getWidget();
[…]

### 2010-01-25 trunk/GVS/media/js/gvs/documents/ScreenflowDocument.js

@@ -6,7 +6,7 @@ var ScreenflowDocument = Class.create(PaletteDocument,
      * @constructs
      * @extends PaletteDocument
      */
-    initialize: function($super, /** String */ title, /** Array */ domainContext, /** String */ version) {
+    initialize: function($super, /** String */ title, /** Array */ tags, /** String */ version) {
         this._inspectorArea = this._createInspectorArea(); 
                
         /**
@@ -24,9 +24,9 @@ var ScreenflowDocument = Class.create(PaletteDocument,
         
         var catalogue = CatalogueSingleton.getInstance();
         
-        var screenSet = new BuildingBlockSet(domainContext, catalogue.
+        var screenSet = new BuildingBlockSet(tags, catalogue.
                             getBuildingBlockFactory(Constants.BuildingBlock.SCREEN));
-        var domainConceptSet = new DomainConceptSet(domainContext, catalogue.
+        var domainConceptSet = new DomainConceptSet(tags, catalogue.
                             getBuildingBlockFactory(Constants.BuildingBlock.DOMAIN_CONCEPT));
         
         /**
@@ -36,7 +36,7 @@ var ScreenflowDocument = Class.create(PaletteDocument,
          */
         this._planPanel = new PlanPanel();
         
-        $super(title, [screenSet, domainConceptSet], [this], domainContext, new ScreenflowInferenceEngine());
+        $super(title, [screenSet, domainConceptSet], [this], tags, new ScreenflowInferenceEngine());
         
         this._planPanel.setDropZone(this);
         this._planPanel.setInferenceEngine(this._inferenceEngine);
@@ -60,7 +60,8 @@ var ScreenflowDocument = Class.create(PaletteDocument,
             'name': title,
             'version': version,
             'domainContext': {
-                'tags': domainContext
+                'tags': tags,
+                'user': null
             }
         });
 
@@ -78,7 +79,7 @@ var ScreenflowDocument = Class.create(PaletteDocument,
         this._inferenceEngine.findCheck(
                 this._getCanvas(),
                 /* palette elements */ [],
-                this._domainContext,
+                this._tags,
                 'reachability',
                 this._findCheckCallback.bind(this)
         );
@@ -386,10 +387,10 @@ var ScreenflowDocument = Class.create(PaletteDocument,
         var palette = this._paletteController.getComponentUris(Constants.BuildingBlock.SCREEN);
         
         if (URIs.catalogueFlow =='check') {
-            this._inferenceEngine.check(canvas, palette, this._domainContext, 
[…]

### 2010-01-25 trunk/GVS/media/js/gvs/dragndrop/DragHandler.js

@@ -174,7 +174,7 @@ var DragHandler = Class.create(
         if (this._isChangingZone()) {
             var dropZone = this._inWhichDropZone();          
             if (dropZone) {
-                dropPosition = Geometry.adaptDropPosition(dropZone, draggableNode);
+                dropPosition = Geometry.adaptDropPosition(dropZone.getNode(), draggableNode);
                 this._initialArea.removeChild(draggableNode);
                 accepted = dropZone.drop(this._draggedObject, dropPosition);
             } else {

### 2010-01-25 trunk/GVS/media/js/gvs/instances/ComponentInstance.js

@@ -23,7 +23,7 @@ var ComponentInstance = Class.create(DragSource,
          * @type String
          * @private
          */
-        this._id = "";
+        this._id = null;
 
         /**
          * BuildingBlock description graphical representation

### 2010-01-25 trunk/GVS/media/js/gvs/persistenceEngine/PersistenceEngine.js

@@ -41,8 +41,8 @@ var PersistenceEngineFactory = function () {
                 method: 'get',
                 parameters: arguments[4],
                 onSuccess: successHandler.bind(context),
-                onFailure: errorHandler.bind(context),
-                onException: errorHandler.bind(context),
+                //onFailure: errorHandler.bind(context),
+                //onException: errorHandler.bind(context),
 				requestHeaders: requestHeaders
             });
         }

### 2010-01-25 trunk/GVS/media/js/gvs/utils/Geometry.js

@@ -114,9 +114,8 @@ Object.extend(Geometry, {
         }
     },
 
-    adaptDropPosition: function(/** DropZone */ dropZone, /** Element */ node) {
+    adaptDropPosition: function(/** Element */ containerElement, /** Element */ node) {
         var element = this.getRectangle(node);
-        var containerElement = dropZone.getNode();
         var containerBounds = this.getClientRectangle(containerElement);
 
         var width = element.right - element.left;
@@ -139,5 +138,41 @@ Object.extend(Geometry, {
             'top': element.top - containerBounds.top
         }
 
+    },
+
+    adaptInitialPosition: function(/** Element */ containerElement, /** Element */ node,
+                                    /** Object */ position) {
+        var containerBounds = {
+            'top': 0,
+            'left': 0,
+            'right': containerElement.clientWidth,
+            'bottom': containerElement.clientHeight
+        }
+        
+        var elementBounds = {
+            'top': position.top,
+            'left': position.left,
+            'right': position.left + node.offsetWidth,
+            'bottom': position.top + node.offsetHeight      
+        }
+
+        var result = {
+            'top': position.top,
+            'left': position.left
+        };
+
+        if (elementBounds.right > containerBounds.right)
+            result.left = containerBounds.right - node.offsetWidth;
+
+        if (elementBounds.left < containerBounds.left)
+            result.left = containerBounds.left;
+
+        if (elementBounds.bottom > containerBounds.bottom)
+            result.top = containerBounds.bottom - node.offsetHeight;
+
+        if (elementBounds.top < containerBounds.top)
+            result.top = containerBounds.top;
+
+        return result;
     }
 }); 

### 2010-01-25 trunk/GVS/media/js/gvs/utils/UIDGenerator.js

@@ -42,7 +42,8 @@ var UIDGeneratorSingleton = function () {
          * @type String
          */
         generate: function (/** String */ element) {
-            var sanitized = element.replace(new RegExp('\\s', 'g'), "");
+            var sanitized = element.replace(new RegExp('\\s', 'g'), "")
+                                    .replace("_","");
             var nextId = this._nextIds[sanitized];
             
             if (!nextId){
@@ -52,6 +53,18 @@ var UIDGeneratorSingleton = function () {
             this._nextIds[sanitized] = nextId + 1;
             
             return sanitized + "_" + nextId;
+        },
+
+        /**
+         * Sets the initial id for a given name
+         */
+        setStartId: function(/** String */ id) {
+            var pieces = id.split("_");
+            var name = pieces[0];
+            var lastId = parseInt(pieces[1]);
+            if (!this._nextIds[name] || this._nextIds[name] <= lastId) {
+                this._nextIds[name] = lastId + 1;
+            }
         }
     });
         

### 2010-01-25 trunk/deliverables/D2_2_Ontology/fast_d2_2_2_ontology.tex
started on evaluation section
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@467 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 	\item Sect.~\ref{sub:goal_and_scope} has been revised to better define the scope of this deliverable with respect to other deliverables.
-	\item Sect.~1.3 (Methontology) has been moved to the end of the document (Sect.~\ref{sub:methontology_a_methodology_for_ontology_development}), but otherwise did not change.
-	\item The section on design principles (Sect.~\ref{sub:general_design_decisions}) has been extended to reflect research done by the authors within the past year.
+	\item The section on our adopted development methodology (previously Sect.~1.3) has been moved to the end of the document (Sect.~\ref{sub:methontology_a_methodology_for_ontology_development}), but otherwise did not change.
+	\item An evaluation section has been added (Sect.~\ref{sec:evaluation}).
+	\item The section on design principles (previously Sect.~1.4, now Sect.~\ref{sub:general_design_decisions}) has been moved to the general evaluation section and extended to reflect research done by the authors within the past year.
 	\item In the integration section (Sect.~\ref{sub:integration}) of the ontology definition, the Common Tag vocabulary has been added. It is briefly introduced, and the integrated terms are specified.
@@
 % subsection changes_from_previous_version (end)
-
-\subsection{Usability as a General Design Principles} % (fold)
-\label{sub:general_design_decisions}
-
-Apart from ``hard'' metrics for the evaluation of ontologies, so-called ``soft'' metrics for ontologies are relevant for facilitating better understanding by users and implementers, and therefore ultimately greater uptake.
-
-\subsubsection{Reuse of Existing Ontologies} % (fold)
-\label{ssub:reuse_of_existing_ontologies}
-
-An important aspect to consider when designing any ontology which is aimed to be used not only within a restricted institution or community is to reuse terms from existing ontologies and vocabularies as much as possible.
-Reuse serves the purpose of lowering the entry barrier for third parties to adopt the ontology --- if they can use familiar concepts and terms, they will be more inclined to try out something new.
-Equally important is the fact that reuse of terms facilitates data integration, which, after all, is one of the major goals of the Semantic Web at large.
-% subsubsection reuse_of_existing_ontologies (end)
-
-\subsubsection{Modularisation} % (fold)
-\label{ssub:modularisation}
-
-Apart from reusing existing terms, another basic design principle which helps to improve accessibility and understanding of the ontology is modularisation.
-While all terms will (for the time being) reside in the same namespace, classes and properties will nevertheless be grouped according to the specific sub-task which they belong to.
-E.g., terms will be grouped into defining \emph{resources}, \emph{pre- and post-conditions}, \emph{general annotation} or \emph{users and user profiles}.
-% subsubsection modularisation (end)
-
-% subsection general_design_decisions (end)
 
@@
 The facts which define the pre- and post-conditions of screens and screenflows in FAST will be modelled as individual RDF triples.
-In effect, this means that conditions, which are sets of facts, can be modelled as graph patterns using SPARQL notation (see \cite{sparql2008spec}).
-E.g., a simple pre-condition such as \emph{``There has to be a user''} could be expressed as in example~\ref{ex:simple_precondition}.
+In effect, this means that conditions, which are sets of facts, can be modelled as graph patterns using SPARQL notation \cite{sparql2008spec}.
+E.g., a simple pre-condition such as \emph{``There has to be a user''} could be expressed as in List.~\ref{ex:simple_precondition}.
 Literally, this very simple pattern means \emph{``a variable ?user is of type sioc:User''}.
@@
 
-\begin{example}
-\begin{verbatim}
-
+\singlespacing
+\lstset{
+	caption={Simple pre-condition}, 
+	label=ex:simple_precondition,
+	language=sparql
+}
+\begin{figure}[ht]
+\begin{lstlisting}
 	?user a sioc:User .
-\end{verbatim}
-\label{ex:simple_precondition}
-\end{example}
+\end{lstlisting}
+\end{figure}
+\doublespacing
 
@@
 In natural language, it could be \emph{``Once the login process has finished, there will be a user object''}.
-Using the same notation as before, this could be expressed as shown in example~\ref{ex:postcondition}, extended with some additional facts (\emph{``There is a user object which has an account name.
+Using the same notation as before, this could be expressed as shown in List.~\ref{ex:postcondition}, extended with some additional facts (\emph{``There is a user object which has an account name.
 There is also a person which has a name, and which has the user object as an online account.''}).
 
-\begin{example}
-\begin{verbatim}
-
+\singlespacing
+\lstset{
+	caption={Simple post-condition}, 
+	label=ex:postcondition,
+	language=sparql
+}
+\begin{figure}[ht]
+\begin{lstlisting}
 	?user a sioc:User ;
@@
 	      foaf:name ?person_name .
-\end{verbatim}
-\label{ex:postcondition}
-\end{example}
+\end{lstlisting}
+\end{figure}
+\doublespacing
 
@@
 
-\begin{example}
-\begin{verbatim}
-
+\singlespacing
+\lstset{
+	caption={Post-condition on the FAST GVS canvas}, 
+	label=ex:postcondition_canvas,
+	language=sparql
+}
+\begin{figure}[ht]
+\begin{lstlisting}
 	_:b0 a sioc:User ;
@@
 	      foaf:name _:b3 .
-\end{verbatim}
-\label{ex:postcondition_canvas}
-\end{example}
+\end{lstlisting}
+\end{figure}
+\doublespacing
 
@@
 
-\begin{example}
-\begin{verbatim}
-
+\singlespacing
+\lstset{
+	caption={Post-condition during runtime}, 
+	label=ex:postcondition_runtime,
+	language=sparql
+}
+\begin{figure}[ht]
+\begin{lstlisting}
 	<http://fast.org/gvs/knud> a sioc:User ;
@@
 	      foaf:holdsAccount <http://fast.org/gvs/knud> ;
-      foaf:name "Knud Möller" .
-\end{verbatim}
-\label{ex:postcondition_runtime}
-\end{example}
+	      foaf:name "Knud Moeller" .
+\end{lstlisting}
+\end{figure}
+\doublespacing
 
@@
 Therefore, we assume that a mapping or mediation layer will ensure that the different semantics at the backend service level will be mapped to a common set of terms within the FAST platform.
-This problem is the topic of deliverable 2.4 in FAST~\cite{ambrus2009fast_mediation}.
+This problem is the topic of deliverable 2.4 in FAST~\cite{ambrus2010fast_mediation} and has also been discussed in \cite{Ambrus:2009it}.
 % subsection extension_towards_specific_domains (end)
@@
 % section the_fast_gadget_ontology (end)
+
+\section{Evaluation} % (fold)
+\label{sec:evaluation}
+
+\subsection{Usability as a General Design Principles} % (fold)
+\label{sub:general_design_decisions}
+
+Apart from ``hard'' metrics for the evaluation of ontologies, so-called ``soft'' factors for ontologies are relevant for facilitating better understanding by users and implementers, and therefore ultimately greater usability and uptake.
+\cite{moeller2009ontology_soft_skills} define five such soft factors as guidelines for ontology developers:
+
+\singlespacing
+\begin{itemize}
+	\item \textbf{reuse external ontologies} to facilitate integration and compatibility,
+	\item \textbf{avoid over-engineering} and keep the number of ontology terms down to reduce the cognitive load on implementers,
+	\item \textbf{be a good Web citizen} by following the principles of linked data,
+	\item \textbf{be well documented} to make it easy for implementers to understand and use the ontology correctly and
+	\item \textbf{provide tool support} to ease the work-load involved in generating instance data.
+\end{itemize}
+\doublespacing
+
+In the following, we will look at each of those guidelines and discuss in how far the FAST gadget ontology fulfils them.
+
+\subsubsection{Reuse of Existing Ontologies} % (fold)
+\label{ssub:reuse_of_existing_ontologies}
+
+An important aspect to consider when designing any ontology which is aimed to be used not only within a restricted institution or community is to reuse terms from existing ontologies and vocabularies as much as possible.
+Reuse serves the purpose of lowering the entry barrier for third parties to adopt the ontology --- if they can use familiar concepts and terms, they will be more inclined to try out something new.
+Equally important is the fact that reuse of terms facilitates data integration, which, after all, is one of the major goals of the Semantic Web at large.
+
+Rather than reinventing the wheel in the various terms that have been defined in its requirement specification, the FAST gadget ontology integrates a number of well-established external ontologies.
+In particular, FOAF and SIOC are used to model tool users and developers, Dublin Core is applied for its general purpose annotation vocabulary and Common Tag is integrated to allow the modelling of a FAST resource's domain context as complex conceptual tags.
+Details are given in Sect.~\ref{sub:integration}.
+% subsubsection reuse_of_existing_ontologies (end)
+
+\subsubsection{Avoid of Over-engineering} % (fold)
+\label{ssub:avoid_of_over_engineering}
+
+% subsubsection avoid_of_over_engineering (end)
+
+\subsubsection{Be a good Web Citizen} % (fold)
+\label{ssub:be_a_good_web_citizen}
+
+% subsubsection be_a_good_web_citizen (end)
+
+\subsubsection{Be Well Documented} % (fold)
+\label{ssub:be_well_documented}
+
+% subsubsection be_well_documented (end)
+
+\subsubsection{Provide Tool Support} % (fold)
+\label{ssub:provide_tool_support}
+
+% subsubsection provide_tool_support (end)
+
+\subsubsection{Modularisation} % (fold)
+\label{ssub:modularisation}
+
+Apart from reusing existing terms, another basic design principle which helps to improve accessibility and understanding of the ontology is modularisation.
+While all terms will (for the time being) reside in the same namespace, classes and properties will nevertheless be grouped according to the specific sub-task which they belong to.
+E.g., terms will be grouped into defining \emph{resources}, \emph{pre- and post-conditions}, \emph{general annotation} or \emph{users and user profiles}.
+% subsubsection modularisation (end)
+
+% subsection general_design_decisions (end)
+
+% section evaluation (end)
 

### 2010-01-25 trunk/catalogue_service/src/eu/morfeoproject/fast/server/ScreenComponentFindCheckServlet.java
fixes bug while checking pipe satisfaction
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@468 d8b0c59b-6544-0410-94d9-83a5658a058e

@@ -253,7 +253,7 @@ public class ScreenComponentFindCheckServlet extends GenericServlet {
 			conTo = getConditionById(postconditions, pipe.getIdConditionTo());
 		} else {
 			ScreenComponent sc = CatalogueAccessPoint.getCatalogue().getScreenComponent(new URIImpl(pipe.getIdBBTo()));
-			conTo = getPreconditionById(sc, pipe.getIdConditionFrom());
+			conTo = getPreconditionById(sc, pipe.getIdConditionTo());
 		}
 		satisfied = isConditionCompatible(conFrom, conTo);
 		return satisfied;

### 2010-01-25 trunk/deliverables/D5_2_1_appendix/D5_2_1.tex
deliverable 5.2.1 (M12)
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@470 d8b0c59b-6544-0410-94d9-83a5658a058e

\documentclass{fast_latex}

\usepackage{inputenc}
\usepackage{lastpage}
\usepackage{setspace}
\usepackage{graphicx}
\usepackage[pdfborder={0 0 0}]{hyperref}		% turn on when latex is used (not miktec)
\usepackage{url} % LEO: urls \url{}
\usepackage{verbatim} % code and comment
\usepackage{longtable}
\usepackage{xspace}   % whitespace after a macro if no punctuation after the macro
\usepackage{multirow}
\usepackage{colortbl}
\usepackage{longtable}
\usepackage{array}
\usepackage{amssymb}
\parindent0pt

\newcommand\deliverableNumber{D5.2.1}
\newcommand\deliverableTitle{Prototype Semantic Catalogue for Screen-Flow Gadgets and Back-End Services}
\newcommand\deliverableTitleShort{Prototype Semantic Catalogue}
\newcommand\workpackageNumber{5}
\newcommand\workpackageTitle{Semantic catalogue of screen-flow resources and back-end Web Services}
\newcommand\authorOne{Ismael Rivera, NUIG}
\newcommand\authorTwo{Author 2}
\newcommand\authorThree{Author 3}
\newcommand\authorFour{Author 4}

\begin{document}
% explicit hyphenations
\hyphenation{RDF-Re-po-si-to-ry}
\hyphenation{name-space}

%\fontfamily{tahoma}\selectfont
\def\note#1{\marginpar{\footnotesize#1}} % use this to show the notes in the document
%\def\note#1{} % use this to hide the notes

\newenvironment{definition}{}{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TITLE PAGES 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\thispagestyle{empty}

\pagenumbering{roman}

\begin{flushright}
	\includegraphics[width=3cm]{images/FP7_logo}
\end{flushright}

\vspace{1cm}

%\begin{minipage}[p]{15cm}
	\begin{center}
		\includegraphics{images/FAST_logo}\\
		\vspace{1cm}
		{\LARGE{\sffamily \emph{FAST AND ADVANCED STORYBOARD TOOLS}}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{FP7-ICT-2007-1-216048}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \emph{http://fast.morfeo-project.eu}}\\
		\vspace{4cm}
		{\LARGE \sffamily \textbf{Deliverable \deliverableNumber}}\\
		\vspace{0.5cm}
		{\LARGE \sffamily \textbf{\deliverableTitle}}\\
		\vspace{2cm}
		{\large \sffamily \authorOne}\\
		\vspace{0.5cm}
		\vfill
		{\large \sffamily Date: 20/02/2009}\\
		\vspace{1cm}
		{\sffamily FAST is partially funded by the E.C. (grant code: FP7-ICT-2007-1-216048).}

	\end{center}
%\end{minipage}

\clearpage
%%%%%%%%%%%%%%
% NEXT PAGES %
%%%%%%%%%%%%%%
\pagestyle{scrheadings}

\lohead{\includegraphics[width=4cm]{images/FAST_logo_transparent}}
%\cohead{\small\textcolor{fast@lightgrey}{\deliverableTitle}}
\rohead{\small{\today}}
%\lofoot{\small\textcolor{fast@lightgrey}{Task Force Ontologies}}
\cofoot{\small{FAST 216048 --- \deliverableTitleShort}}
\rofoot{\small{\thepage}}

\newpage

\clearpage

\section*{Version History}

\begin{small}
\begin{tabular}{|l|l|l|p{7.5cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Rev. No.}} &
                         \textcolor{white}{\textbf{Date}} &
                         \textcolor{white}{\textbf{Author (Partner)}} &
                         \textcolor{white}{\textbf{Change description}}\\ \hline
1.0 & 20/02/2009 & Ismael Rivera (NUIG) & First version of the deliverable \\ \hline
\end{tabular}
\end{small}

\color{black}

\vfill
%{\bf Explanations of abbreviations on front page}\\
%\\
%%Nature \\
%R: Report \\
%P: Prototype \\
%R/P: Report and Prototype \\
%O: Other \\
% \\
%Dissemination level \\
%PU: Public \\
%PP: Restricted to other FP6 participants \\
%RE: Restricted to specified group \\
%CO: Confidential, only for NEPOMUK partners \\

\newpage

%%%%%%%%%%%%%%%%%%%%%
% Executive Summary %
%%%%%%%%%%%%%%%%%%%%%

\clearpage

\section*{Executive Summary}
\doublespacing

The present deliverable is intented to be the developer's manual of the prototype of the semantic catalogue. That said, whoever developing a component which will interact with the catalogue will find in this document a description of its architecture, the functionality provided together with the API (Application Programming Interface), data interchange formats, code errors and exceptions as well as several examples of usage.

\newpage

%%%%%%%%%%%%%%%%%%%%%
% Document Summary %
%%%%%%%%%%%%%%%%%%%%%

\clearpage

\section*{Document Summary}
% double spacing from here on:
\singlespacing

\begin{small}
\begin{tabular}
	%{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{6cm}|p{2cm}|p{2cm}|}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{6cm}|p{2cm}|p{2cm}|}
	\hline
	\textcolor{white}{\textbf{Code}} & {FP7-ICT-2007-1-216048} & {\textbf{Acronym}} & {FAST}\\ \hline
	\textcolor{white}{\textbf{Full title}} & \multicolumn{3}{l|}{Fast and Advanced Storyboard Tools}\\ \hline
	\textcolor{white}{\textbf{URL}} & \multicolumn{3}{l|}{\url{http://fast.morfeo-project.eu}}\\ \hline
	\textcolor{white}{\textbf{Project officer}} & \multicolumn{3}{l|}{Annalisa Bogliolo}\\ \hline
\end{tabular}
\end{small}

\vspace{0.5cm}

\begin{small}
\begin{tabular}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.25cm}|p{1cm}|p{1cm}|p{6.32cm}|}
	\hline
	\textcolor{white}{\textbf{Deliverable}} & {\textbf{Number}} & {\deliverableNumber} & {\textbf{Name}} & {\deliverableTitle}\\ \hline
	\textcolor{white}{\textbf{Work package}} & {\textbf{Number}} & {\workpackageNumber} & {\textbf{Name}} & {\workpackageTitle}\\ \hline
\end{tabular}
\end{small}

\vspace{0.5cm}

\begin{small}
\begin{tabular}
	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
	\hline
	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {27/02/2009} & {\textbf{Submitted}} & {27/02/2009}\\ \hline
	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{Status} & \multicolumn{2}{l|}{final}\\ \hline
	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{This deliverable is the technical documentation for the prototype developed as part of the WP5. It describes the catalogue's architecture, data interchange formats, APIs and examples of how to interact with it.}\\ \hline
	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
%	{} & \multicolumn{4}{l|}{\authorOne}\\ 
%	{} & \multicolumn{4}{l|}{\authorOne}\\ 
%	{} & \multicolumn{4}{l|}{\authorOne}\\
  \hline
	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {ismael.rivera@deri.org} \\ \cline{4-5}
	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495086} \\ \hline
	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, semantic catalogue, gadget catalogue, RDF store}\\ \hline
\end{tabular}
\end{small}

\newpage

%%%%%%%%%%%%%%%%%%%%%
% TABLE OF CONTENTS %
%%%%%%%%%%%%%%%%%%%%%
\doublespacing
\setcounter{tocdepth}{2}
\tableofcontents
\cleardoublepage
\pagenumbering{arabic}

%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGINNING OF SECTIONS %
%%%%%%%%%%%%%%%%%%%%%%%%%
\rofoot{\small{Page \thepage\ of \pageref{LastPage}}}

\clearpage
\section{Introduction}
\label{sec:introduction}
This section starts establishing the goal and scope of the present document, shows how it is structured and details the relation to others documents and work packages.

\subsection{Goal and Scope}
This is an introductory manual for developers who want to adopt and use the FAST semantic catalogue. It explains the main functionalities implemented so far, an overview of its architecture and a detailed \emph{Application Programming Interface} or API of the complete set of the operations offered through a REST service.

\subsection{Structure of the document}
The deliverable presents both the external and internal architecture in Section \ref{sec:architecture}, then in Section \ref{sec:api} it is detailed the Catalogue API, query formats, interchange formats, error codes and so on, and the document ends by stating some conclusions and future directions of the semantic catalogue in Section \ref{sec:conclusions}.

\subsection{Overall approach for WP5 and relation to other Work Packages and Deliverables}
The main objective of the Work Package 5, as specified in \cite{DoW}, includes the development of the FAST semantic catalogue, considered the back-end of the FAST platform, which will store and index any kind of resources such as screen-flows, screens, back-end Web services descriptions, locate appropriate gadgets, resources and Web services for users, storing and indexing user profiles, and resolve mediation problems between gadgets, back-end services and the different ontologies stored.

The deliverables stated in \cite{DoW} for this WP are:
\begin{description}
	\item[User manual of the FAST Catalogue] This is the user manual of the catalogue, but is out of scope in this iteration of the project.
	\item[Prototype semantic catalogue for screen-flow gadgets and back-end services] This deliverable is a software prototype of the semantic catalogue for experimentation purposes. Apart from the software, it will be created technical documentation and manuals for developers as the present document.
\end{description}

The semantic catalogue relies strongly on the FAST ontology defined in \cite{moeller2009fast_ontology} since the ontology defines how every element within FAST will be internally stored by the catalogue, and takes into consideration specific requirements detailed in \cite{fast_requirements} where the functionalities of the complete FAST tool are defined.

\clearpage
\section{Architecture}
\label{sec:architecture}
This section details the internal catalogue's architecture and which external components will interact with it. Basically three layers can be distinguished: Presentation, Business Logic and Persistence layer (see Figure \ref{fig:architecture}).

\begin{figure}[htb]
\label{fig:architecture}
\begin{center}
	\includegraphics[width=12cm]{images/catalogue_architecture}
	\caption{Catalogue's Architecture}
	\label{fig:CatalogueInternalArchitecture}
\end{center}
\end{figure}

The Presentation layer will be the public interface of the catalogue. The main purpose of this layer is to provide its functionality to other components within the Gadget Visual Studio. The service is offered in a REST (REpresentational State Transfer) style, making it easy to consume than other complex APIs. REST is an architectural style, not a toolkit or a standard. Even though, it makes use of standards like HTTP, URI, Resource Representations (XML, HTML, JSON, JPEG, etc.) or MIME types (text/xml, text/html, application/json, image/jpeg, etc.). Another characteristic of the REST style is its stateless assumption. The catalogue adopts this strategy; therefore every request needs a complete set of information in order to prepare the response.

As part of the Presentation layer, a SPARQL endpoint is offered using the SPARQL protocol service as defined in the SPROT \cite{sprot} specification. The SPARQL endpoint is mostly offered to enable other developers to query directly the catalogue knowledge base using SPARQL queries. This feature is supported by the Sesame RESTful HTTP interface for SPARQL Protocol for RDF.

The Business Logic layer contains all the FAST domain-specific processing. It provides functions to interact with all the elements of the domain model specified in \cite{moeller2009fast_ontology}. It processes the input from the Presentation layer, creating specific objects modelling the business logic, and vice versa. In addition it interacts with the Persistence layer in order to persist the model.

The Persistence layer provides an API which allows the Business layer to work with a standard set of objects that read and save their state to the triple store. For this reason, on top of this layer it is used RDF2Go, an abstraction over triple (and quad) stores. The RDF2Go API allows interacting with the semantic representation of the model (triples) in a generic manner, and also brings the flexibility of choosing different triple stores to persist them. The selected implementation for the triple stored is Sesame 2. Moreover, this framework is completely extensible and configurable regarding to storage mechanisms, inferencers, RDF file formats, query result formats and query languages. For the actual catalogue prototype, the storage mechanism used is the native file storage system of Sesame, the inferencer used is a subset of the RDFS entailment rules \cite{rdfsrules} following a forward-chaining policy.

\clearpage
\section{RESTful Catalogue API}
\label{sec:api}
This section provides a high-level overview of the Catalogue API. It describes the calls or operations supported, specific parameters and responses for each operation, supported interchange formats and some examples to facilitate the understanding of the API.

\subsection{JSON Interchange Format}
The format supported by the Catalogue API for all the request and responses is JSON\footnote{http://www.json.org/}. JSON is a lightweight data interchange format whose simplicity has resulted in widespread use among web developers, easy to read and write and able to using any programming language because its structures map directly to data structures used in most programming languages.

Every HTTP request and response is encoded using the MIME type \texttt{application/json} and the charset \texttt{UTF-8}.

\subsubsection{Example of a screen}
\begin{verbatim}
{
  "uri": null,
  "label": {"en-GB": "Screen - DERI"},
  "description": {"en-GB": "Returns people working in DERI"},
  "creationDate": "2009-02-07T09:59:52+0000",
  "creator": "http://www.ismaelrivera.es/",
    "domainContext": {
      "tags": [],
      "user": null
    },
  "homepage": "http://www.deri.ie/",
  "icon": "http://www.deri.ie/icon.jpg",
  "screenshot": "http://www.deri.ie/screenshot.jpg",
  "rights": "http://creativecommons.org/",
  "version": "1.0",
  "preconditions": [],
  "postconditions": ["?person rdf:type foaf:Person . 
                      ?person foaf:workplaceHomepage http://www.deri.ie/"]
}
\end{verbatim}

\subsubsection{Internationalisation I18n}
In order to offers an adaptable solution to various languages and regions without major engineering changes, internationalisation is considered from an early stage in the catalogue development. Underlying representation technologies used to develop the catalogue (RDF and JSON) allow to implement this feature. This feature is implemented by the addition of a language tag to every 'string' desired. The specification of this language tag is composed by the language code and the country code, following the ISO 639\footnote{http://ftp.ics.uci.edu/pub/ietf/http/related/iso639.txt} for languages and the ISO 3166\footnote{http://userpage.chemie.fu-berlin.de/diverse/doc/ISO\_3166.html} for countries. The following example illustrates how to use it properly in both formats.

\begin{verbatim}
{
  "label": {
    "en-GB": "Simple Example",
    "es-ES": "Ejemplo Sencillo",
    "de-DE": "Einfaches Beispiel"
  },
  "description": {
    "en-GB": "This is a simple example of a screen",
    "es-ES": "Esto es un ejemplo sencillo de pantalla",
    "de-DE": "Dies ist ein einfaches Beispiel für einen Bildschirm"
  },
}
\end{verbatim}

\begin{verbatim}
<rdf:Description rdf:about="http://www.morfeoproject.eu/fast/fco#ExampleScreen">
  <rdf:type rdf:resource="http://www.morfeoproject.eu/fast/fco#Screen"/>
  <rdfs:label xml:lang="en-GB">Simple Example</rdfs:label>
  <rdfs:label xml:lang="es-ES">Ejemplo Sencillo</rdfs:label>
  <rdfs:label xml:lang="de-DE">Einfaches Beispiel</rdfs:label>
  <description
    xmlns="http://purl.org/dc/terms/"
    xml:lang="en-GB">This is a simple example of a screen</description>
  <description
    xmlns="http://purl.org/dc/terms/"
    xml:lang="es-ES">Esto es un ejemplo sencillo de pantalla</description>
  <description
    xmlns="http://purl.org/dc/terms/"
    xml:lang="de-DE">Dies ist ein einfaches Beispiel für einen Bildschirm</description>
</rdf:Description>
\end{verbatim}

By now it is allowed using internationalisation for the label and the description of a screen, but more attributes will be benefited by this feature.

\subsection{API Calls}
This section contains detailed descriptions of the interface provided by the catalogue via REST services, detailing request parameters, response elements, any special errors and examples of requests and responses. The URL format is also specified for each operation, where 'catalogueURL' has to be replaced for the real URL the service is installed (e.g. forge.morfeo-project.org/catalogue).

\subsubsection{CRUD operations for Screens}
A screen is a resource which can be created, retrieved, modified or deleted since the catalogue provides CRUD support for screens. These features are provided following the REST philosophy, using the HTTP methods. Two concepts have to be defined: \emph{collection}, as a set of screens which access URI is http://catalogueURL/screens/ where catalogueURL is the specific URL where the catalogue server is installed, and \emph{member} of the collection, in other words, the screen itself, which access URI is http://catalogueURL/screens/<id> where the <id> has to be replaced by the identifier of a screen. The details of the operations and which HTTP verb has to be used can be found in the Table \ref{tab:crud_screens}.

\begin{table}[htb]
\caption{CRUD operations for Screens}
\label{tab:crud_screens}
\begin{center}
\begin{tabular}{|p{2.5cm}|p{2.5cm}|p{3.5cm}|p{5cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{Resource} &
                         \textcolor{white}{HTTP method} &
                         \textcolor{white}{HTTP body} &
                         \textcolor{white}{Description}\\ \hline
Collection URI & GET & N/A & \textbf{List} the members of the collection.\\ \hline
Collection URI & PUT & N/A & Not used.\\ \hline
Collection URI & POST & JSON representation of the screen & \textbf{Create} a new entry in the collection where the URI is assigned automatically by the collection. The URI created is returned by this operation.\\ \hline
Collection URI & DELETE & N/A & Not used.\\ \hline
Member URI & GET & N/A & \textbf{Retrieve} the addressed member of the collection.\\ \hline
Member URI & PUT & JSON representation of the screen & \textbf{Update} the addressed member of the collection or create it with a defined URI.\\ \hline
Member URI & POST & N/A & Not used. \\ \hline
Member URI & DELETE & N/A & \textbf{Delete} the addressed member of the collection.\\ \hline
\end{tabular}
\end{center}
\end{table}

Here are shown various examples of request and responses using the catalogue server. To create a new screen, a POST request is send to the catalogue server, including the JSON representation of the screen as the body of the request:

\begin{verbatim}
{
  "uri": null,
  "label": {"en-GB": "Screen - DERI"},
  "description": {"en-GB": "Returns people working in DERI"},
  "creationDate": "2009-02-07T09:59:52+0000",
  "creator": "http://www.ismaelrivera.es/",
    "domainContext": {
      "tags": [],
      "user": null
    },
  "homepage": "http://www.deri.ie/",
  "icon": "http://www.deri.ie/icon.jpg",
  "screenshot": "http://www.deri.ie/screenshot.jpg",
  "rights": "http://creativecommons.org/",
  "version": "1.0",
  "preconditions": [],
  "postconditions": ["?person rdf:type foaf:Person . 
                      ?person foaf:workplaceHomepage http://www.deri.ie/"]
}
\end{verbatim}

The URI will be not taken into account since the catalogue will generate a unique URI for each resource created. So, the response for this operation is:

\begin{verbatim}
{
  "creationDate": "2009-02-07T09:59:52+0000",
  "creator": "http://www.ismaelrivera.es/",
  "description": {"en-GB": "Show information about a person working in DERI"},
  "domainContext": {
    "tags": [],
    "user": "null"
  },
  "homepage": "http://www.deri.ie/",
  "icon": "http://www.deri.ie/icon.jpg",
  "label": {"en-GB": "Screen - DERI"},
  "postconditions": [],
  "preconditions": ["?person rdf:type foaf:Person . 
                     ?person foaf:workplaceHomepage http://www.deri.ie/"],
  "rights": "http://creativecommons.org/",
  "screenshot": "http://www.deri.ie/screenshot.jpg",
  "uri": "http://www.morfeoproject.eu/fast/fco#Screen1234552194269",
  "version": "1.0"
}
\end{verbatim}

To obtain all the screens stored in the catalogue a HTTP GET request is sent to the Collection URI and the response may be something like this:
\begin{verbatim}
[
  {
    "creationDate": "2009-02-07T09:59:52+0000",
    "creator": "http://www.ismaelrivera.es/",
    "description": {"en-gb": "Returns people working in DERI"},
    "domainContext": {
      "tags": [],
      "user": null
    },
    "homepage": "http://www.deri.ie/",
    "icon": "http://www.deri.ie/icon.jpg",
    "label": {"en-gb": "Screen - DERI"},
    "postconditions": ["?person rdf:type foaf:Person . 
                        ?person foaf:workplaceHomepage http://www.deri.ie/"],
    "preconditions": [],
    "rights": "http://creativecommons.org/",
    "screenshot": "http://www.deri.ie/screenshot.jpg",
    "uri": "http://www.morfeoproject.eu/fast/fco#Screen1234552164472",
    "version": "1.0"
  },
  {
    "creationDate": "2009-02-07T09:59:52+0000",
    "creator": "http://www.ismaelrivera.es/",
    "description": {"en-gb": "Show information about a person working in DERI"},
    "domainContext": {
      "tags": [],
      "user": null
    },
    "homepage": "http://www.deri.ie/",
    "icon": "http://www.deri.ie/icon.jpg",
    "label": {"en-gb": "Screen - DERI"},
    "postconditions": [],
    "preconditions": ["?person rdf:type foaf:Person . 
                       ?person foaf:workplaceHomepage http://www.deri.ie/"],
    "rights": "http://creativecommons.org/",
    "screenshot": "http://www.deri.ie/screenshot.jpg",
    "uri": "http://www.morfeoproject.eu/fast/fco#Screen1234552194269",
    "version": "1.0"
  },
  {
    "creationDate": "2009-02-07T09:59:52+0000",
    "creator": "http://www.ismaelrivera.es/",
    "description": {"en-gb": "Show information about a person"},
    "domainContext": {
      "tags": [],
      "user": null
    },
    "homepage": "http://www.deri.ie/",
    "icon": "http://www.deri.ie/icon.jpg",
    "label": {"en-gb": "Screen - DERI"},
    "postconditions": [],
    "preconditions": ["?person rdf:type foaf:Person"],
    "rights": "http://creativecommons.org/",
    "screenshot": "http://www.deri.ie/screenshot.jpg",
    "uri": "http://www.morfeoproject.eu/fast/fco#Screen1234552373865",
    "version": "1.0"
  }
]
\end{verbatim}

\subsubsection{CRUD operations for Screen-Flows}
A screenflow is a resource which can be created, retrieved, modified or deleted since the catalogue provides CRUD support for screenflows. These features are provided following the REST philosophy, using the HTTP methods. As in the previous section the concepts \emph{collection} and \emph{member} are used including their URLs. Table \ref{tab:crud_screenflows} details the operations offered for screen-flows.

\begin{table}[htb]
\caption{CRUD operations for Screen-Flows}
\label{tab:crud_screenflows}
\begin{center}
\begin{tabular}{|p{2.5cm}|p{2.5cm}|p{3.5cm}|p{5cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{Resource} &
                         \textcolor{white}{HTTP method} &
                         \textcolor{white}{HTTP body} &
                         \textcolor{white}{Description}\\ \hline
Collection URI & GET & N/A & \textbf{List} the members of the collection.\\ \hline
Collection URI & PUT & N/A & Not used.\\ \hline
Collection URI & POST & JSON representation of the screen-flow & \textbf{Create} a new entry in the collection where the URI is assigned automatically by the collection. The URI created is returned by this operation.\\ \hline
Collection URI & DELETE & N/A & Not used.\\ \hline
Member URI & GET & N/A & \textbf{Retrieve} the addressed member of the collection.\\ \hline
Member URI & PUT & JSON representation of the screen-flow & \textbf{Update} the addressed member of the collection or create it with a defined URI.\\ \hline
Member URI & POST & N/A & Not used. \\ \hline
Member URI & DELETE & N/A & \textbf{Delete} the addressed member of the collection.\\ \hline
\end{tabular}
\end{center}
\end{table}

The following is an example of the request and response while creating a screen-flow:
\begin{verbatim}
{
  "uri": null,
  "label": {"en-GB": "My First ScreenFlow"},
  "description": {"en-GB": "A example of screenflow"},
  "creationDate": "2009-02-05T13:11:41+0000",
  "creator": "http://www.ismaelrivera.es/",
  "domainContext": {
    "tags": ["people"],
    "user": null
  },
  "homepage": "http://www.screenflow.com/",
  "icon": "http://www.screenflow.com/icon.jpg",
  "screenshot": "http://www.screenflow.com/screenshot.jpg",
  "rights": "http://creativecommons.org/",
  "version": "1.0",
  "preconditions": [],
  "postconditions": [],
  "screens": [
    "http://www.morfeoproject.eu/fast/fco#Screen1233750841383",
    "http://www.morfeoproject.eu/fast/fco#Screen1233750828836",
    "http://www.morfeoproject.eu/fast/fco#Screen1233750820914"
  ]
}
\end{verbatim}

\begin{verbatim}
{
  "creationDate": "2009-02-05T13:11:41+0000",
  "creator": "http://www.ismaelrivera.es/",
  "description": {"en-GB": "A example of screenflow"},
  "domainContext": {
    "tags": ["people"],
    "user": null
  },
  "homepage": "http://www.screenflow.com/",
  "icon": "http://www.screenflow.com/icon.jpg",
  "label": {"en-GB": "My First ScreenFlow"},
  "postconditions": [],
  "preconditions": [],
  "rights": "http://creativecommons.org/",
  "screens": [
    "http://www.morfeoproject.eu/fast/fco#Screen1233750841383",
    "http://www.morfeoproject.eu/fast/fco#Screen1233750828836",
    "http://www.morfeoproject.eu/fast/fco#Screen1233750820914"
  ],
  "screenshot": "http://www.screenflow.com/screenshot.jpg",
  "uri": "http://www.morfeoproject.eu/fast/fco#ScreenFlow1233752391898",
  "version": "1.0"
}
\end{verbatim}

\subsubsection{Find}
The find operation search inside the catalogue for any screen which could be somehow related to the gadget the user is creating. It provides a recommended set of screens depending on the domain context, the canvas, etc. The specific URL to access this operation is http://catalogueURL/find using HTTP POST method. From now on, the method 'find' only considers the domain context, being a list of tags and the pre/postconditions of the screens. It will try to satisfy all the unsatisfied preconditions of a given list of screens also known as canvas. The results will be all the screens stored in the catalogue which fulfil some of these preconditions, and are tagged with the tags specified in the domain context. The request parameters are shown in Table \ref{tab:find_request} and the response parameters are shown in Table \ref{tab:find_response}.

\begin{table}[htb]
\caption{Find Request Parameters}
\label{tab:find_request}
\begin{center}
\begin{tabular}{|p{2cm}|p{9.5cm}|p{2cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
                         \textcolor{white}{Description} &
                         \textcolor{white}{Type}\\ \hline
Canvas & The canvas is composed by a list of screens. Only the URI is needed. & Optional \\ \hline
Domain Context & The domain context contains a list of tags and a user. & Optional \\ \hline
Elements & It is a list of resources, previously stored in the catalogue. For example, the list of screens recommended last execution. Only accepts screens. & Optional \\ \hline
\end{tabular}
\end{center}
\end{table}

\begin{table}[htb]
\caption{Find Response Parameters}
\label{tab:find_response}
\begin{center}
\begin{tabular}{|p{2cm}|p{9.5cm}|p{2cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
                         \textcolor{white}{Description} &
                         \textcolor{white}{Type}\\ \hline
N/A & A list of recommended screens is returned. & Required \\ \hline
\end{tabular}
\end{center}
\end{table}

Following is shown an example of usage of this operation considering a canvas with the screen \texttt{http://www.morfeoproject.eu/fast/fco\#Screen1234552373865}.

\begin{verbatim}
{
  canvas: [
    {
      uri: "http://www.morfeoproject.eu/fast/fco#Screen1234552373865"
    }
  ],
  domainContext: {
    tags: [],
	  user: null
  },
  elements: []
}
\end{verbatim}

After the execution of the recommendation algorithm, the response given by the catalogue is:

\begin{verbatim}
[{
  "creationDate": "2009-02-07T09:59:52+0000",
  "creator": "http://www.ismaelrivera.es/",
  "description": {"en-gb": "Returns people working in DERI"},
  "domainContext": {
    "tags": [],
    "user": null
  },
  "homepage": "http://www.deri.ie/",
  "icon": "http://www.deri.ie/icon.jpg",
  "label": {"en-gb": "Screen - DERI"},
  "postconditions": ["?person rdf:type foaf:Person . 
                      ?person foaf:workplaceHomepage http://www.deri.ie/"],
  "preconditions": [],
  "rights": "http://creativecommons.org/",
  "screenshot": "http://www.deri.ie/screenshot.jpg",
  "uri": "http://www.morfeoproject.eu/fast/fco#Screen1234552164472",
  "version": "1.0"
}]
\end{verbatim}

\subsubsection{Check}
This operation verifies the state of certain list of screens depending on a specific criterion. The only criterion accepted in this stage of the prototype is 'reachability'. A screen will be reachable if it all its preconditions are satisfied by postconditions of the screens the canvas contain. The specific URL to access this operation is http://catalogueURL/check using HTTP POST method.

\begin{table}[htb]
\caption{Check Request Parameters}
\label{tab:check_request}
\begin{center}
\begin{tabular}{|p{2cm}|p{9.5cm}|p{2cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
                         \textcolor{white}{Description} &
                         \textcolor{white}{Type}\\ \hline
Canvas & The canvas is composed by a list of screens. Only the URI is needed. & Optional\\ \hline
Domain Context & The domain context contains a list of tags and a user. & Optional\\ \hline
Elements & It is a list of resources, previously stored in the catalogue. For example, the list of screens recommended last execution. Only accepts screens. & Optional\\ \hline
Criterion & The criterion specifies what it has to be check. The only possible supported now is 'reachability' in order to check if the preconditions of a screen are satisfied. & Required\\ \hline
\end{tabular}
\end{center}
\end{table}

\begin{table}[htb]
\caption{Check Response Parameters}
\label{tab:check_response}
\begin{center}
\begin{tabular}{|p{2cm}|p{9.5cm}|p{2cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
                         \textcolor{white}{Description} &
                         \textcolor{white}{Type}\\ \hline
Canvas & The canvas indicating which screens satisfy the critetion. More aditional information may be included, for example, a satisfaction attribute for every precondition for reachability. & Optional\\ \hline
Domain Context & The domain context contains a list of tags and a user. & Optional\\ \hline
Elements & The elements indicating which satisfy the critetion. More aditional information may be included as in the Canvas. & Optional\\ \hline
\end{tabular}
\end{center}
\end{table}

The scenario for the following example is composed by a screen in the canvas which precondition is a foaf:Person which foaf:workplaceHomepage is http://www.deri.ie/, and a screen in the elements list which precondition is a foaf:Person. The state to check in this case is the reachability and satisfaction of the screens and preconditions.

\begin{verbatim}
{
  canvas: [
    {
      uri: "http://www.morfeoproject.eu/fast/fco#Screen1234552194269"
    }
  ],
  domainContext: [],
  elements: [
    {
      uri: "http://www.morfeoproject.eu/fast/fco#Screen1234552373865"
    }  
  ],
  criterion: "reachability"
}
\end{verbatim}

As you can see in the response, no screen is reachable at the moment, and no precondition is satisfied.

\begin{verbatim}
{
  "canvas": [{
    "preconditions": [{
      "expression": "?person rdf:type foaf:Person . 
                     ?person foaf:workplaceHomepage http://www.deri.ie/",
      "satisfied": false
    }],
    "reachability": false,
    "uri": "http://www.morfeoproject.eu/fast/fco#Screen1234552194269"
  }],
  "elements": [{
    "preconditions": [{
      "expression": "?person rdf:type foaf:Person",
      "satisfied": false
    }],
    "reachability": false,
    "uri": "http://www.morfeoproject.eu/fast/fco#Screen1234552373865"
  }]
}
\end{verbatim}

\subsubsection{GetMetadata}
Although the information about a resource can be obtained by its specific retrieval operation supported by the CRUD interface, sometimes is needed to retrieve information which would imply a request per resource, hence this operation allow to get the metadata of a list of resources in only one request. This prototype supports this operation for screenflows and screens. The specific URL to access this operation is http://catalogueURL/getmetadata using HTTP POST method. Table \ref{tab:getmetadata_request} shows the parameters needed for the invocation and Table \ref{tab:getmetadata_response} details the different parameters may contain a response.

\begin{table}[htb]
\caption{GetMetadata Request Parameters}
\label{tab:getmetadata_request}
\begin{center}
\begin{tabular}{|p{2cm}|p{9.5cm}|p{2cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
                         \textcolor{white}{Description} &
                         \textcolor{white}{Type}\\ \hline
N/A & A list of URIs. & Required\\ \hline
\end{tabular}
\end{center}
\end{table}

\begin{table}[htb]
\caption{GetMetadata Response Parameters}
\label{tab:getmetadata_response}
\begin{center}
\begin{tabular}{|p{2cm}|p{9.5cm}|p{2cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
                         \textcolor{white}{Description} &
                         \textcolor{white}{Type}\\ \hline
Screenflows & A set of screenflows with all the metadata associated to them.. & Optional\\ \hline
Screens & A set of screens with all the metadata associated to them. & Optional\\ \hline
\end{tabular}
\end{center}
\end{table}

The following example send two URIs of screens and a URI of a screen-flow to the GetMetadata operation:
\begin{verbatim}
[
  "http://www.morfeoproject.eu/fast/fco#Screen1234552164472",
  "http://www.morfeoproject.eu/fast/fco#Screen1234552194269",
  "http://www.morfeoproject.eu/fast/fco#ScreenFlow1233752391898"
]
\end{verbatim}

The response obtained are two lists, one containing all the metadata regarding to the screen-flows and another one with the information about the screens:
\begin{verbatim}
{
  "screenflows": [{
    "creationDate": "2009-02-05T13:11:41+0000",
    "creator": "http://www.ismaelrivera.es/",
    "description": {"en-gb": "A example of screenflow"},
    "domainContext": {
      "tags": ["people"],
      "user": null
    },
    "homepage": "http://www.screenflow.com/",
    "icon": "http://www.screenflow.com/icon.jpg",
    "label": {"en-gb": "My First ScreenFlow"},
    "postconditions": [],
    "preconditions": [],
    "rights": "http://creativecommons.org/",
    "screens": [
      "http://www.morfeoproject.eu/fast/fco#Screen1233750820914",
      "http://www.morfeoproject.eu/fast/fco#Screen1233750828836",
      "http://www.morfeoproject.eu/fast/fco#Screen1233750841383"
    ],
    "screenshot": "http://www.screenflow.com/screenshot.jpg",
    "uri": "http://www.morfeoproject.eu/fast/fco#ScreenFlow1233752391898",
    "version": "1.0"
  }],
  "screens": [
    {
      "creationDate": "2009-02-07T09:59:52+0000",
      "creator": "http://www.ismaelrivera.es/",
      "description": {"en-gb": "Returns people working in DERI"},
      "domainContext": {
        "tags": [],
        "user": null
      },
      "homepage": "http://www.deri.ie/",
      "icon": "http://www.deri.ie/icon.jpg",
      "label": {"en-gb": "Screen - DERI"},
      "postconditions": ["?person rdf:type foaf:Person . 
                          ?person foaf:workplaceHomepage http://www.deri.ie/"],
      "preconditions": [],
      "rights": "http://creativecommons.org/",
      "screenshot": "http://www.deri.ie/screenshot.jpg",
      "uri": "http://www.morfeoproject.eu/fast/fco#Screen1234552164472",
      "version": "1.0"
    },
    {
      "creationDate": "2009-02-07T09:59:52+0000",
      "creator": "http://www.ismaelrivera.es/",
      "description": {"en-gb": "Show information about a person working in DERI"},
      "domainContext": {
        "tags": [],
        "user": null
      },
      "homepage": "http://www.deri.ie/",
      "icon": "http://www.deri.ie/icon.jpg",
      "label": {"en-gb": "Screen - DERI"},
      "postconditions": [],
      "preconditions": ["?person rdf:type foaf:Person . 
                         ?person foaf:workplaceHomepage http://www.deri.ie/"],
      "rights": "http://creativecommons.org/",
      "screenshot": "http://www.deri.ie/screenshot.jpg",
      "uri": "http://www.morfeoproject.eu/fast/fco#Screen1234552194269",
      "version": "1.0"
    }
  ]
}
\end{verbatim}

\subsubsection{FindAndCheck}
Not supported yet.

\subsection{API Error Codes}
There are two types of error codes, client and server.
\begin{itemize}
	\item Client error codes are generally caused by the client and might be an invalid domain or an invalid request parameter. These errors are accompanied by a 4xx HTTP response code. For example: ResourceNotFound.
	\item Server error codes are generally caused by a server-side issue and should be reported. These errors are accompanied by a 5xx HTTP response code. For example: ServerUnavailable.
\end{itemize}

The following table lists all the error codes.

\begin{table}[htb]
\caption{API Error Codes}
\begin{center}
\begin{tabular}{|p{3.5cm}|p{7cm}|p{3cm}|}
\hline
\rowcolor{fast@lightgrey}\textcolor{white}{Error} &
                         \textcolor{white}{Description} &
                         \textcolor{white}{HTTP Status Code}\\ \hline
ResourceNotFound & The resource <resourceURI> has not been found. & 404 Not Found \\ \hline
AccessFailure & Access to the resource <resourceName> is denied. & 403 Forbidden \\ \hline
InternalError & Request could not be executed due to an internal service error. & 500 Internal Server Error \\ \hline
InvalidAction & The action <actionName> is not valid for this web service.  & 404 Bad Request \\ \hline
InvalidHttpRequest & The HTTP request is invalid. Reason: <reason>. & 400 Bad Request \\ \hline
InvalidLiteral & Illegal literal in the filter expression. & 400 Bad Request \\ \hline
InvalidParameterValue & The specified parameter value is not valid. & 400 Bad Request \\ \hline
InvalidURI & The URI <requestURI> is not valid. & 400 Bad Request \\ \hline
MissingAction & No action was supplied with this request. & 400 Bad Request \\ \hline
MissingParameter & The request must contain the specified missing parameter. & 400 Bad Request \\ \hline
NotYetImplemented & Feature <feature> is not yet available. & 401 Unauthorized \\ \hline
ServiceUnavailable & The service is currently unavailable. Please try again later. &  \\ \hline
UnsupportedHttpVerb & The requested HTTP verb is not supported: <verb>. & 400 Bad Request \\ \hline
URITooLong & The URI exceeded the maximum limit of <maxLength>. & 400 Bad Request \\ \hline
\end{tabular}
\end{center}
\end{table}

\clearpage
\section{Conclusions and Future Work}
\label{sec:conclusions}
As a result of the work done for the back-end catalogue, any one who read this document should be able to understand the capabilities implemented and how to use them. Summarising, this prototype allows the storage of screens and the composition of screen-flows. Future work will involve the storage of the internal components of a screen, such as connectors, filters, forms and back-end services. However, the most interesting work will be done in the optimization of the resources recommendation feature, in other words, it will be investigate new mechanisms to improve the results of the find such as weighting of answers, planning, etc. Last of all, more complicated queries will be taken into account, improving the matching algorithm between pre and postconditions.

\clearpage
\bibliographystyle{apalike}
\bibliography{fast_bib}

\clearpage
\doublespacing
\section*{Appendix A (Lists of Tables and Figures)}
\addcontentsline{toc}{section}{Appendix A (Lists of Tables and Figures)}

\listoftables

\listoffigures

\end{document}

### 2010-01-25 trunk/deliverables/D5_1_2/fast_d5_1_2.tex
draft for deliverable 5.1.2 (M24)
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@471 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-01-25 trunk/deliverables/D5_2_1_appendix/D5_2_1.tex
@@
 
-\newcommand\deliverableNumber{D5.2.1}
-\newcommand\deliverableTitle{Prototype Semantic Catalogue for Screen-Flow Gadgets and Back-End Services}
-\newcommand\deliverableTitleShort{Prototype Semantic Catalogue}
+\newcommand\deliverableNumber{D5.1.2}
+\newcommand\deliverableTitle{User Manual of the FAST Catalogue}
+\newcommand\deliverableTitleShort{Catalogue User Manual}
 \newcommand\workpackageNumber{5}
@@
 \newcommand\authorFour{Author 4}
+\newtheorem{example}{\emph{Example}}
 
@@
 		\vfill
-		{\large \sffamily Date: 20/02/2009}\\
+		{\large \sffamily Date: 25/01/2010}\\
 		\vspace{1cm}
@@
                          \textcolor{white}{\textbf{Change description}}\\ \hline
-1.0 & 20/02/2009 & Ismael Rivera (NUIG) & First version of the deliverable \\ \hline
+1.0 & 25/01/2010 & Ismael Rivera (NUIG) & First version of the deliverable \\ \hline
 \end{tabular}
@@
 	\hline
-	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {27/02/2009} & {\textbf{Submitted}} & {27/02/2009}\\ \hline
+	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {27/02/2010} & {\textbf{Submitted}} & {27/02/2010}\\ \hline
 	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{Status} & \multicolumn{2}{l|}{final}\\ \hline
@@
   \hline
-	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {ismael.rivera@deri.org} \\ \cline{4-5}
-	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495086} \\ \hline
+	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{NUIG} & \textbf{email} & {ismael.rivera@deri.org} \\ \cline{4-5}
+	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495338} \\ \hline
 	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, semantic catalogue, gadget catalogue, RDF store}\\ \hline
@@
 \clearpage
-\section{Introduction}
+\section{Introduction} % (fold)
 \label{sec:introduction}
@@
 
-\subsection{Goal and Scope}
+\subsection{Goal and Scope} % (fold)
+\label{sub:goal_and_scope}
+
 This is an introductory manual for developers who want to adopt and use the FAST semantic catalogue.
@@
 
-\subsection{Structure of the document}
-The deliverable presents both the external and internal architecture in Section \ref{sec:architecture}, then in Section \ref{sec:api} it is detailed the Catalogue API, query formats, interchange formats, error codes and so on, and the document ends by stating some conclusions and future directions of the semantic catalogue in Section \ref{sec:conclusions}.
-
-\subsection{Overall approach for WP5 and relation to other Work Packages and Deliverables}
+% subsection goal_and_scope (end)
+
+\subsection{Structure of the document} % (fold)
+\label{sub:structure_of_the_document}
+
+The deliverable presents both the external and internal architecture in Section \ref{sec:architecture_overview}, then in Section \ref{sec:restful_catalogue_api} it is detailed the Catalogue API, query formats, interchange formats, error codes and so on.
+
+% subsection goal_and_scope (end)
+
+\subsection{Overall approach for WP5 and relation to other Work Packages and Deliverables} % (fold)
+\label{sub:overall_approach_for_wp5}
+
 The main objective of the Work Package 5, as specified in \cite{DoW}, includes the development of the FAST semantic catalogue, considered the back-end of the FAST platform, which will store and index any kind of resources such as screen-flows, screens, back-end Web services descriptions, locate appropriate gadgets, resources and Web services for users, storing and indexing user profiles, and resolve mediation problems between gadgets, back-end services and the different ontologies stored.
@@
 \begin{description}
-	\item[User manual of the FAST Catalogue] This is the user manual of the catalogue, but is out of scope in this iteration of the project.
+	\item[User manual of the FAST Catalogue] This is the user manual of the catalogue.
+Since the semantic catalogue will not provide a user graphical interface by itself: it relies on the GVS; this document will be oriented to developers, explaining technical aspects of the catalogue, its API and several examples to ease its use.
 	\item[Prototype semantic catalogue for screen-flow gadgets and back-end services] This deliverable is a software prototype of the semantic catalogue for experimentation purposes.
-Apart from the software, it will be created technical documentation and manuals for developers as the present document.
 \end{description}
@@
 
+% subsection overall_approach_for_wp5 (end)
+
+\subsection{Changes from D5.2.1} % (fold)
+\label{sub:changes}
+
+This should be the first version of this deliverable since in M12 the deliverable 5.1 was not scheduled.
+Nevertheless, last year an extra document was written as an appendix to the FAST Catalogue in order to detail and clarify technical decisions and to have a technical reference manual.
+Hence, the deliverable D5.2.1 will evolve to become the D5.1.2, so it will be no longer needed by itself.
+
+There are several new sections and major changes which need to be mentioned:
+
+\begin{itemize}
+	\item Section \ref{sec:installation_guide}.
+Installation Guide describes a set of steps to be followed in order to install and execute the FAST Catalogue.
+Also some configuration instructions are specified.
+	\item Section \ref{sub:content_negotiation}. Content Negotiation states how the FAST Catalogue has adopted several of the content negotiation principles.
+	\item Section \ref{sec:restful_catalogue_api}.
+RESTful Catalogue API has been intensively extended and modified in order to  reflect all the new functionality and how every request or response is constructed.
+	\item Several appendixes have been added with the JSON structure of every building block the Catalogue can deal with.
+\end{itemize}
+
+% subsection changes (end)
+
+% section introduction (end)
+
 \clearpage
-\section{Architecture}
-\label{sec:architecture}
-This section details the internal catalogue's architecture and which external components will interact with it.
+\section{Installation Guide} % (fold)
+\label{sec:installation_guide}
+
+This section  provides instructions on how to manually install and configure the FAST Catalogue.
+The first part of this guide presents some requirements to be considered before the installation, then it gives broad general instructions, and the last part contains a more detailed installation notes for specific configurations.
+
+\subsection{System Requirements} % (fold)
+\label{sub:system_requirements}
+
+In addition to the software itself, a standard FAST Catalogue installation has the following requirements:
+\begin{itemize}
+  \item Java\footnote{http://java.sun.com/} (version 6 or later) is required to run the software.
+  \item A Java Servlet Container that supports Java Servlet API 2.4 and Java Server Pages (JSP) 2.0, or newer.
+We recommend using a recent, stable version of Apache Tomcat\footnote{http://tomcat.apache.org/}.
+At the time of writing, this is either version 5.5.x or 6.x.
+\end{itemize}
+
+In addition, there are various optional dependencies which are required if you want to use certain advanced features (see Section \ref{sub:installation_instructions}).
+
+% subsection system_requirements (end)
+
+\subsection{Obtaining the FAST catalogue} % (fold)
+\label{sub:obtaining_the_fast_catalogue}
+
+It is recommended to have a Subversion client installed before you download the code (although you can theoretically download files without Subversion, this would mean tediously downloading each individual file manually).
+The recommended software is the official Subversion client, available from the Subversion project page\footnote{http://subversion.tigris.org/}.
+Note that this client uses a command-line interface, which the instructions below use.
+Alternatively, you can get subversioning software with a graphical user interface such as TortoiseSVN.
+
+To download from the latest release (recommended), enter the following command from the command-line in the directory you wish to download to:
+\begin{verbatim}
+svn checkout https://svn.forge.morfeo-project.org/fast-fp7project/trunk/catalogue_service
+\end{verbatim}
+
+% subsection obtaining_the_fast_catalogue (end)
+
+\subsection{Installation Instructions} % (fold)
+\label{sub:installation_instructions}
+
+Once you have got the source code, first step should be addressed is its compilation.
+You may manually compile the source yourself, however the FAST Catalogue comes along with a script to facilitate this task.
+This script has been made using a Java build tool called Ant \footnote{http://ant.apache.org/}.
+Hence, we recommended to get and install such a tool in order to follow the instructions below.
+
+Now you can compile, package and run the application via:
+\begin{verbatim}
+ant clean
+ant compile
+ant dist
+\end{verbatim}
+
+Or using a task which gather the previous three tasks:
+\begin{verbatim}
+ant main
+\end{verbatim}
+
+After executing with no errors this task, a WAR file should have been created in the /dist directory.
+A WAR file, which stands for  Web Application Archive, is just a JAR file used to distribute a collection of JavaServer Pages, servlets, Java classes, XML files, tag libraries and static Web pages (HTML and related files) that together constitute a Web application, in this case, the FAST Catalogue.
+
+Then, last step is to deply the WAR file into the servlet container. If the server chosen was Apache Tomcat, the procedure for deploying a Web application is:
+\begin{enumerate}
+	\item Stop Tomcat.
+	\item Delete existing deployment.
+If you have previously deployed "foo.war" in TOMCAT\_HOME/webapps, then it has been unpacked into webapps/foo/...
+You must delete this directory and all its contents.
+On Unix, this can be done with
+rm -r \$TOMCAT\_HOME/webapps/foo
+	\item Copy WAR file to TOMCAT\_HOME/webapps/.
+	\item Start Tomcat.
+\end{enumerate}
+
+Now you have successfully built the FAST catalogue, however we highly recommend you to read Section \ref{sub:configuration} and tune your FAST Catalogue instance.
+
+% subsection installation_instructions (end)
+
+\subsection{Configuration} % (fold)
+\label{sub:configuration}
+
+The following section covers the configuration options you may have to set up before start using the Catalogue.
+The FAST Catalogue relies on a Sesame repository for the persistent storage.
+Sesame repositories can be local or remote.
+Remote does not mean, it has to be installed in a different machine, but in a servlet container such as Apache Tomcat.
+The advantage of remote repositories is the possibility of accessing it via a SPARQL endpoint.
+These configuration aspect can be found in the file \emph{repository.properties} in the root directory where the Catalogue has been deployed.
+
+Independently the type of repository, the parameter \emph{serverURL} needs to take the value of the URL where the FAST Catalogue instance has been deployed.
+If you are testing the application in your machine, it may be something like \url{http://localhost:8080/FASTCatalogue}.
+
+Local repositories just need the parameter \emph{storageDir} pointing at a local directory, which has write permission in order to allow Sesame to create files in that directory or to delete any file in the directory.
+For instance:
+
+\begin{verbatim}
+storageDir=D:\\FAST\\catalogue\\repository
+\end{verbatim}
+
+Nevertheless, setting up a remote repository requires a few more steps.
+First the OpenRDF Sesame Server needs to be installed.
+The installation guide of this server is detailed in the Chapter 5 and 6 of the User Guide for Sesame 2.2 \cite{sesame2.2}.
+Once the Sesame server is deployed, and a repository has been created, the parameters sesameServer and repositoryID will take the values of the Sesame server URL and the repository identifier.
+For instance:
+
+\begin{verbatim}
+sesameServer=http://localhost:8880/openrdf-sesame
+repositoryID=c2
+\end{verbatim}
+
+% subsection installation_instructions (end)
+
+\subsection{Hello World!} % (fold)
+\label{sub:hello_world}
+
+It is recommended to check if the FAST Catalogue has been successfully installed.
+To do that, a simple HTTP request, using a Web browser, can be sent to the URL the Catalogue is deployed, and specifying a type of resource such as Screen.
+The URL would be similar to \url{http://localhost:8080/FASTCatalogue/screens}.
+Depending on the variant chosen, the response may be different.
+For JSON, the response should be \verb|[]|, but in HTML the message \verb|No screens found.| should be shown.
+
+% subsection hello_world (end)
+
+% section installation_guide (end)
+
+\clearpage
+\section{Architecture Overview} % (fold)
+\label{sec:architecture_overview}
+
+This section presents a technical overview of the internal catalogue's architecture and which external components will interact with it.
 Basically three layers can be distinguished: Presentation, Business Logic and Persistence layer (see Figure \ref{fig:architecture}).
@@
 
+% section architecture_overview (end)
+
 \clearpage
-\section{RESTful Catalogue API}
-\label{sec:api}
+\section{RESTful Catalogue API} % (fold)
+\label{sec:restful_catalogue_api}
+
 This section provides a high-level overview of the Catalogue API.
@@
 
-\subsection{JSON Interchange Format}
+\subsection{JSON Interchange Format} % (fold)
+\label{sec:json_interchange_format}
+
 The format supported by the Catalogue API for all the request and responses is JSON\footnote{http://www.json.org/}.
@@
 Every HTTP request and response is encoded using the MIME type \texttt{application/json} and the charset \texttt{UTF-8}.
+
+% subsection json_interchange_format (end)
 
@@
 
-\subsubsection{Internationalisation I18n}
+\subsubsection{Internationalisation I18n} % (fold)
+
 In order to offers an adaptable solution to various languages and regions without major engineering changes, internationalisation is considered from an early stage in the catalogue development.
@@
 
-\subsection{API Calls}
+\subsection{Content Negotiation} % (fold)
+\label{sub:content_negotiation}
+
+Before any major details about how content negotiation is addressed by the catalogue, a few concepts need to be understood.
+First of all, say a URI is not a file name.
+It is common to see a URI as the location for a file, but a URI or Universal Resource Identifier is not a URL: it is not a file name or location, it is an identifier (or a reference) to a resource.
+Then by dereferencing a URI the actual resource can be retrieved.
+Let's see a request to retrieve the resource "Screen 123" is sent to the server.
+The resource is just a piece of information, however the server will answer an HTML document with a text describing the screen or an image representing it.
+On the Web, these equivalent representations of a resource are called variants, and content negotiation is the mechanism used to determine which of the representations of most appropiate for a given request.
+
+\begin{figure}[htb]
+\label{fig:ContentNegotiation}
+\begin{center}
+	\includegraphics[width=10cm]{images/content_negotiation}
+	\caption{Content Negotiation}
+\end{center}
+\end{figure}
+
+In summary, the basic idea of content negotiation, as stated in the HTTP 1.1 specification \cite{http1.1}, is to serve the best variant for a resource, taking into account what variants are available, what variants the server may prefer to serve, what the client can accept, and with which preferences: in HTTP, this is done by the client which may send, in its request, Accept headers (Accept, Accept-Language and Accept-Encoding), to communicate its capabilities and preferences in Format, Language and Encoding, respectively.
+
+In fact, what the catalogue really does is "format negotiation" since the alternate representationos are just based on the selection of the media type, through the Accept header, but does not consider different languages or encoding types.
+The formats supported are JSON, RDF+XML and HTML.
+
+Lastly, content negotiation needs to identify which player is going to take the lead on it.
+There are two kinds of content negotiation which are possible in HTTP: server-driven and agent-driven negotiation.
+The approach followed by the catalogue is agent-driven negotiation, hence selecting a specific representation for a resource is responsibility of the user agent.
+If none is specified, by default, the server will choose the JSON representation.
+
+% subsection content_negotiation (end)
+
+\subsection{API Calls} % (fold)
+\label{sub:api_calls}
+
 This section contains detailed descriptions of the interface provided by the catalogue via REST services, detailing request parameters, response elements, any special errors and examples of requests and responses.
 The URL format is also specified for each operation, where 'catalogueURL' has to be replaced for the real URL the service is installed (e.g.
-forge.morfeo-project.org/catalogue).
-
-\subsubsection{CRUD operations for Screens}
+http://demo.fast.morfeo-project.org/catalogue).
+
+\subsubsection{CRUD operations for Screens} % (fold)
+
 A screen is a resource which can be created, retrieved, modified or deleted since the catalogue provides CRUD support for screens.
@@
 
-\subsubsection{CRUD operations for Screen-Flows}
+\subsubsection{CRUD operations for Screen-Flows} % (fold)
+
 A screenflow is a resource which can be created, retrieved, modified or deleted since the catalogue provides CRUD support for screenflows.
@@
 
-\subsubsection{Find}
+\subsubsection{Find} % (fold)
+
 The find operation search inside the catalogue for any screen which could be somehow related to the gadget the user is creating.
@@
 
-\subsubsection{Check}
+\subsubsection{Check} % (fold)
+
 This operation verifies the state of certain list of screens depending on a specific criterion.
@@
 
-\subsubsection{GetMetadata}
+\subsubsection{GetMetadata} % (fold)
+
 Although the information about a resource can be obtained by its specific retrieval operation supported by the CRUD interface, sometimes is needed to retrieve information which would imply a request per resource, hence this operation allow to get the metadata of a list of resources in only one request.
@@
 
-\subsubsection{FindAndCheck}
-Not supported yet.
-
-\subsection{API Error Codes}
+\subsubsection{Screen Find \& Check} % (fold)
+
+Request
+\begin{verbatim}
+{
+  "canvas": [
+    { "uri": "http://purl.oclc.org/fast/ontology/gadget#Screen1257860786707" }
+  ],
+  "elements": [
+  ],
+  "domainContext": {
+    "tags": [],
+    "user": null,
+  },
+  "criterion": "reachability"
+}
+\end{verbatim}
+
+Response
+\begin{verbatim}
+{
+  "canvas": [{
+    "preconditions": [[{
+      "id": "sc",
+      "label": {"en-gb": "A search criteria"},
+      "pattern": "?F http://www.w3.org/1999/02/22-rdf-syntax-ns#type http://aws.amazon.com/AWSECommerceService#SearchCriteria",
+      "satisfied": false,
+      "positive": true
+    }]],
+    "reachability": false,
+    "uri": "http://purl.oclc.org/fast/ontology/gadget#Screen1257860786707"
+  }],
+  "elements": [{
+    "preconditions": [],
+    "reachability": true,
+    "uri": "http://purl.oclc.org/fast/ontology/gadget#Screen1257860530032"
+  }]
+}
+\end{verbatim}
+
+% subsection api_calls (end)
+
+\subsection{Planning}
+\label{sub:planning}
+
+URL: <CATALOGUE\_URL>/planner
+Request
+\begin{verbatim}
+{
+  "goal": "http://purl.oclc.org/fast/ontology/gadget#Screen1257860720035",
+  "canvas": []
+}
+\end{verbatim}
+
+\begin{verbatim}
+Response
+[
+  [
+    "http://purl.oclc.org/fast/ontology/gadget#Screen1257866875435",
+    "http://purl.oclc.org/fast/ontology/gadget#Screen1257865645621"
+  ],
+  [
+    "http://purl.oclc.org/fast/ontology/gadget#Screen1257860530032",
+    "http://purl.oclc.org/fast/ontology/gadget#Screen1257860786707",
+    "http://purl.oclc.org/fast/ontology/gadget#Screen1257860772723"
+  ]
+]
+\end{verbatim}
+
+% subsection planning (end)
+
+\subsection{API Error Codes} % (fold)
 There are two types of error codes, client and server.
@@
 \clearpage
-\section{Conclusions and Future Work}
-\label{sec:conclusions}
-As a result of the work done for the back-end catalogue, any one who read this document should be able to understand the capabilities implemented and how to use them.
-Summarising, this prototype allows the storage of screens and the composition of screen-flows.
-Future work will involve the storage of the internal components of a screen, such as connectors, filters, forms and back-end services.
-However, the most interesting work will be done in the optimization of the resources recommendation feature, in other words, it will be investigate new mechanisms to improve the results of the find such as weighting of answers, planning, etc.
-Last of all, more complicated queries will be taken into account, improving the matching algorithm between pre and postconditions.
-
-\clearpage
 \bibliographystyle{apalike}
-\bibliography{fast_bib}
+\bibliography{fast_d5_1_2}
 
@@
 \doublespacing
-\section*{Appendix A (Lists of Tables and Figures)}
-\addcontentsline{toc}{section}{Appendix A (Lists of Tables and Figures)}
+\section*{Appendix A. Lists of Tables and Figures}
+\addcontentsline{toc}{section}{Appendix A. Lists of Tables and Figures}
 
@@
 
+\clearpage
+\section*{Appendix B. Generic Building Block JSON Structure}
+\addcontentsline{toc}{section}{Appendix B. Generic Building Block JSON Structure}
+
+\begin{verbatim}
+{
+  "code": "http://url.com/.../code.js",
+  "creationDate": "2009-04-20T17:00:00+0100",
+  "creator": "http://www.fabio.es/",
+  "description": {"en-gb": "This is a description of the building block"},
+  "tags": {
+    "tags": [
+      "ebay",
+      "amazon",
+    ],
+    "user": <String> // not defined yet
+  },
+  "homepage": "http://www.homepage.ie/",
+  "icon": "http://url.com/images/icon.png",
+  "id": "Id of the resource",
+  "label": {"en-gb": "Label or title of the building block"},
+  "name": "Amazon Item to Ebay Filter",
+  "rights": "http://creativecommons.org/",
+  "screenshot": "http://www.deri.ie/eBayList-screenshot.jpg",
+  "type": "type of the building block", // screenflow, screen, form, operator y resource
+  "uri": "http://localhost:8080/catalogue/screens/525",
+  "version": "1.0"
+}
+\end{verbatim}
+
+\clearpage
+\section*{Appendix C. Screenflow JSON Structure}
+\addcontentsline{toc}{section}{Appendix C. Screenflow JSON Structure}
+
+\begin{verbatim}
+{
+  <GENERIC BUILDING BLOCK JSON>,
+  "contains": [
+    "http://purl.oclc.org/fast/ontology/gadget#Screen1257860772723",
+    "http://purl.oclc.org/fast/ontology/gadget#Screen1257860720035",
+  ],
+}
+\end{verbatim}
+
+\clearpage
+\section*{Appendix D. Screen JSON Structure}
+\addcontentsline{toc}{section}{Appendix D. Screen JSON Structure}
+
+\begin{verbatim}
+{
+  <GENERIC BUILDING BLOCK JSON>,
+  "postconditions": [[{
+    "id": "url",
+    "label": {"en-gb": "Purchase URL"},
+    "pattern": "?P
+                http://www.w3.org/1999/02/22-rdf-syntax-ns#type
+                http://aws.amazon.com/AWSECommerceService#PurchaseURL",
+    "positive": true
+  }]],
+  "preconditions": [[{
+    "id": "cart",
+    "label": {"en-gb": "A shopping cart"},
+    "pattern": "?C
+                http://www.w3.org/1999/02/22-rdf-syntax-ns#type
+                http://aws.amazon.com/AWSECommerceService#ShoppingCart",
+    "positive": true
+  }]],
+  "code": "URL of the code",
+  "definition": {
+    "buildingblocks: [
+      { // Form
+        "id": "form1", // unique for the container screen
+        "uri": "http://purl.oclc.org/fast/ontology/gadget#Form670238"
+      },
+      { // Operators
+        "id": "op1",
+        "uri": "http://purl.oclc.org/fast/ontology/gadget#Operator213487"
+      },
+      {  // Backend services
+        "id": "bs1",
+        "uri": "http://purl.oclc.org/fast/ontology/gadget#Backendservice38399"
+      },
+      {...}
+    ]
+    "pipes": [
+      {
+        "from": {
+          "buildingblock": "bs1",
+          "condition": "cA"
+        },
+        "to": {
+          "buildingblock": "op1",
+          "condition": "cA", // can be the same id, unique for the building block
+          "action": "filter"
+        }
+      },
+    "triggers": [
+      {
+        "from": {
+          "buildingblock": "form1",
+          "name": "refresh"
+        },
+        "to": {
+          "buildingblock": "bs1",
+          "action": "list"
+        }
+      }
+    ]
+  }
+}
+\end{verbatim}
+
+\clearpage
+\section*{Appendix E. Screen Component JSON Structure}
+\addcontentsline{toc}{section}{Appendix E. Screen Component JSON Structure}
+
+\begin{verbatim}
+{
+  <GENERIC BUILDING BLOCK JSON>,
+  "actions": [{
+    "name": "filter",
+    "preconditions": [[{
+      "id": "item",
+      "label": {"en-gb": "Ebay List"},
+      "pattern": "?Item
+                  http://www.w3.org/1999/02/22-rdf-syntax-ns#type
+                  http://aws.amazon.com/AWSECommerceService#Item",
+      "positive": true
+    }]],
+    "uses": []
+  }],
+  "libraries": [],
+  "postconditions": [[{
+    "id": "filterEbay",
+    "label": {"en-gb": "Ebay List"},
+    "pattern": "?eFilter
+                http://www.w3.org/1999/02/22-rdf-syntax-ns#type
+                http://developer.ebay.com/.../FindItemsAdvanced.html#Request",
+    "positive": true
+  }]],
+  "triggers": ["itemAmazon"]
+}
+\end{verbatim}
+
 \end{document}

### 2010-01-26 trunk/deliverables/D2_2_Ontology/fast_d2_2_2_ontology.tex
done with soft skills part of evaluation section
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@475 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \usepackage{amssymb}
+\usepackage{paralist}
+
+% This is useful for switching in-line comments on and off!
+\usepackage{ifthen}
+\newboolean{showcomments}
+\setboolean{showcomments}{true}
+\ifthenelse{\boolean{showcomments}}
+{
+	\definecolor{Red}{rgb}{1,0,0}
+	\newcommand{\note}[1]{\begin{scriptsize}{{\textsf{\textbf{{\textcolor{Red}{[[#1]]}}}}}}\end{scriptsize}}
+	\definecolor{Orange}{rgb}{1,0.5,0}
+	\newcommand{\todo}[1]{\textsf{\textbf{\textcolor{Orange}{[[TODO: #1]]}}}}
+    \definecolor{Green}{rgb}{0,1,0}
+    \newcommand{\commentreply}[1]{\begin{scriptsize}{\textsf{\textbf{\textcolor{Green}{[[#1]]}}}}\end{scriptsize}}
+} {
+	\newcommand{\note}[1]{\textbf{}}
+	\newcommand{\todo}[1]{}
+    \newcommand{\commentreply}[1]{}
+}
+
 \usepackage{listings}
@@
 	\item In the integration section (Sect.~\ref{sub:integration}) of the ontology definition, the Common Tag vocabulary has been added. It is briefly introduced, and the integrated terms are specified.
+	\item The namespace abbreviation for the FAST gadget ontology changed from \texttt{fast} to \texttt{fgo}, in order to distinguish from other FAST ontologies which may be added in the future.
 \end{itemize}
@@
 \textbf{Properties:} \emph{containsScreen}, \emph{hasLabel}, \emph{hasIcon}, \emph{hasPreCondition}, \emph{hasPostCondition}, \emph{hasTag}, \emph{hasProfile}\\ \hline
-\textcolor{white}{\textbf{Sources of Knowledge}} & FAST Architecture deliverable~\cite{solero2009fast_architecture}, FAST Requirements Specification~\cite{villoslada2009fast_requirements}, FAST State-of-the-art Document~\cite{urmetzer2010fast_state_of_the_art}, EzWeb documentation~\cite{lizcano2008ezweb} \\ \hline
+\textcolor{white}{\textbf{Sources of Knowledge}} & FAST Architecture deliverable~\cite{solero2009fast_architecture}, FAST Requirements Specification~\cite{villoslada2010fast_requirements}, FAST State-of-the-art Document~\cite{urmetzer2010fast_state_of_the_art}, EzWeb documentation~\cite{lizcano2008ezweb} \\ \hline
 \end{longtable}
@@
 
-This section contains a definition of all classes defined in the \texttt{fast} namespace. For classes integrated from other ontologies, see the integration document~\ref{ssub:integration_document}.
+This section contains a definition of all classes defined in the \texttt{fgo} namespace. For classes integrated from other ontologies, see the integration document~\ref{ssub:integration_document}.
 
@@
 
-This section contains a definition of all properties defined in the \texttt{fast} namespace.
+This section contains a definition of all properties defined in the \texttt{fgo} namespace.
 For properties integrated from other ontologies, see the integration document~\ref{ssub:integration_document}.
@@
 As an alternative or supplement to DBpedia identifiers, it is possible to use structured tagging approaches such as MOAT\footnote{\url{http://moat-project.org/}}.
+	\todo{update with reference to common tag!}
 	\item \emph{User Interests:} In the same way as domain contexts, a user's interests can be specified using DBpedia identifiers or structured tagging.
@@
 
-\subsection{Usability as a General Design Principles} % (fold)
+\subsection{Usability as a General Design Principle} % (fold)
 \label{sub:general_design_decisions}
@@
 
-\subsubsection{Avoid of Over-engineering} % (fold)
+\subsubsection{Avoid Over-engineering} % (fold)
 \label{ssub:avoid_of_over_engineering}
 
+The concrete interpretation of this criterion depends to a large degree on the intended use case and user community of the ontology.
+For a large-scale undertaking spanning many domains, and where reasoning is in the centre of attention --- CyC e.g.
+represents this type of ontology  --- a certain amount of complexity is necessary.
+However, in order to appeal to adopters who are familiar with formal knowledge representation, simplicity can be an important factor.
+Good examples of such ontologies are FOAF and SIOC, which would arguably not have had such an impact in the wider Web community if they tried to model their respective domains in a much more complex way, making use of the full range of expressivity offered by e.g.
+the RDFS and OWL ontology languages.
+Indirect support for this argument is provided by \cite{wang2006owl_survey}, who show that most ontologies found on the Web are on the lower end of the expressivity spectrum, indicating that practitioners either don't need or don't understand the full range of possible expressiveness.
+
+At the moment, the use case for the FAST gadget ontology is the tool-internal representation of gadgets and their components, which means that simplicity for users in the wider community is only of secondary importance.
+However, the ontology is still kept relatively basic, both in terms of the number of terms (currently 28 classes and 29 properties\todo{Recount!}), and in the complexity of the class definitions themselves, which are mostly restricted to a straightforward class subsumption hierarchy, with the exception of the sparse use of the boolean class expression \texttt{owl:unionOf}.
+
+As another measure to avoid over-engineering, it was ensured that the logical complexity of the ontology falls within the DL fragment\footnote{We use the Pellet reasoner: \url{http://www.mindswap.org/2003/pellet/demo} to measure complexity.} of OWL.
+While this does not necessarily ensure that an ontology is easy to comprehend by users, it can help to achieve this goal.
+
 % subsubsection avoid_of_over_engineering (end)
 
-\subsubsection{Be a good Web Citizen} % (fold)
+\subsubsection{Be a Good Web Citizen} % (fold)
 \label{ssub:be_a_good_web_citizen}
+
+While ontologies have gained much attention as a key component in the emerging Semantic Web, not all ontologies are meant for use on the Web.
+However, if ontologies are supposed to be deployed on the Web and used in a Web context, they should also be a good Web citizen, meaning they should follow certain rules and best practices, in particular those defined for the Web of linked data~\cite{bernersLee2006linkedData}.
+Of those rules, the first two --- all things are identified with a URI and URIs should be HTTP URIs --- are followed by definition if the ontology in question is formalised in RDF.
+The third one, however --- that information should be served on the Web for each URI --- has often been neglected in the past.
+Terms in ontologies were identified with URIs which were not dereferencable, i.e., the ontology was defined in a namespace URI, but the ontology document was served at an arbitrary other URI (or even not served anywhere at all).
+This has two negative effects on adoption:
+\begin{inparaenum}[(i)]
+	\item automatic agents encountering instance data for an ontology cannot request the ontology from its namespace URI to learn the precise formal definitions of the terms used, and
+	\item human users encountering terms from an ontology cannot look up definitions and documentation for them in a Web browser. 
+\end{inparaenum}
+
+The FAST gadget ontology fulfils this criterion by following the best practices recommendations outlined in \cite{berrueta2008publishing_rdf_vocabularies}, thereby ensuring that the ontology is available within a stable namespace and URI and can be accessed both by machine agents to retrieve its formal definition, as well as by human agents, who will retrieve a human-understandable HTML documentation from the same URI.
+Furthermore, the fourth rule of linked data (links should established between datasets) is covered through the integration of external ontologies, as discussed above in Sec.~\ref{ssub:reuse_of_existing_ontologies}.
 
@@
 
+A lack of good documentation for an ontology will make it hard for potential adopters to comprehend and make use of it, whereas on the other hand, good documentation can be the deciding factor in which ontology gains more uptake.
+On the most basic level, documentation can be provided by the use of annotation properties within the formal definition of the ontology itself (e.g., \texttt{rdfs:comment}).
+However, in practice this should be extended with human-readable documentation, ideally at the available online at the namespace URI of the ontology itself (see above in Sect.~\ref{ssub:be_a_good_web_citizen}).
+The human-readable documentation can be as simple as a list of terms and explanations generated from the ontology source, e.g., with a tool such as \emph{VocDoc}\footnote{\url{http://kantenwerk.org/vocdoc/}}, which was developed specifically to generate documentation for the FAST gadget ontology both in an online format and a print format.
+Much more useful than a simple list of terms, however, are concrete instance examples which illustrate the actual use of the various ontology terms in practice.
+
+Our ontology follows the documentation guideline, both through its online documentation, as well as through this deliverable.
+Both versions of the documentation provide both a lookup list for terms and their definitions, as well as concrete examples of term usage in instances.
+
 % subsubsection be_well_documented (end)
@@
 
+A further guideline to ease and encourage adoption of an ontology is the provision of tool support to users, which will help them to generate instance data.
+Examples for such tools can be online forms such as FOAF-a-matic\footnote{\url{http://www.ldodds.com/foaf/foaf-a-matic}} for FOAF or plugins for popular content management systems to generate SIOC data\footnote{\url{http://sioc-project.org/exporters}}.
+In FAST, instance data for the FAST gadget ontology is never intended to be generated by hand, but rather through the use of the FAST catalogue.
+Additional tool support is therefore not currently necessary.
+
 % subsubsection provide_tool_support (end)
@@
 
-Apart from reusing existing terms, another basic design principle which helps to improve accessibility and understanding of the ontology is modularisation.
-While all terms will (for the time being) reside in the same namespace, classes and properties will nevertheless be grouped according to the specific sub-task which they belong to.
-E.g., terms will be grouped into defining \emph{resources}, \emph{pre- and post-conditions}, \emph{general annotation} or \emph{users and user profiles}.
+Apart from the guidelines proposed by \cite{moeller2009ontology_soft_skills}, another basic design principle which helps to improve accessibility and understanding of the ontology is modularisation.
+While all terms reside in the same namespace, classes and properties are nevertheless be grouped according to the specific sub-tasks which they belong to.
+E.g., terms are grouped into defining \emph{resources}, \emph{pre- and post-conditions}, \emph{general annotation} or \emph{users and user profiles}.
+\todo{Implement this in the online documentation through VocDoc.}
 % subsubsection modularisation (end)
@@
 A solution for representing pre- and post-conditions of screens and screen flows --- a central aspect of FAST technology --- has been proposed, which will now have to be implemented and tested both within the FAST catalogue and the FAST GVS.
+
+The FAST gadget ontology has initially been developed using the original OWL ontology language (or OWL 1).
+Since the start of the project, OWL 2\footnote{\url{http://www.w3.org/TR/owl-overview/}} has been released and is now a W3C recommendation.
+In order to avoid any potential migration issues, we have so far refrained from adopting OWL 2, but may do so in the future.
 
@@
 
-The \emph{Domain Analysis} section of this deliverable implements an ontology requirement specification document for the FAST gadget ontology (which is not to be confused with the FAST requirements deliverable \cite{villoslada2009fast_requirements}!).
+The \emph{Domain Analysis} section of this deliverable implements an ontology requirement specification document for the FAST gadget ontology (which is not to be confused with the FAST requirements deliverable \cite{villoslada2010fast_requirements}!).
 % subsubsection specification (end)

### 2010-01-26 trunk/deliverables/D5_1_2/fast_d5_1_2.tex
deliverable ready for review
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@476 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-\subsection{Overall approach for WP5 and relation to other Work Packages and Deliverables} % (fold)
-\label{sub:overall_approach_for_wp5}
-
-The main objective of the Work Package 5, as specified in \cite{DoW}, includes the development of the FAST semantic catalogue, considered the back-end of the FAST platform, which will store and index any kind of resources such as screen-flows, screens, back-end Web services descriptions, locate appropriate gadgets, resources and Web services for users, storing and indexing user profiles, and resolve mediation problems between gadgets, back-end services and the different ontologies stored.
-
-The deliverables stated in \cite{DoW} for this WP are:
-\begin{description}
-	\item[User manual of the FAST Catalogue] This is the user manual of the catalogue.
-Since the semantic catalogue will not provide a user graphical interface by itself: it relies on the GVS; this document will be oriented to developers, explaining technical aspects of the catalogue, its API and several examples to ease its use.
-	\item[Prototype semantic catalogue for screen-flow gadgets and back-end services] This deliverable is a software prototype of the semantic catalogue for experimentation purposes.
-\end{description}
-
-The semantic catalogue relies strongly on the FAST ontology defined in \cite{moeller2009fast_ontology} since the ontology defines how every element within FAST will be internally stored by the catalogue, and takes into consideration specific requirements detailed in \cite{fast_requirements} where the functionalities of the complete FAST tool are defined.
-
-% subsection overall_approach_for_wp5 (end)
-
-\subsection{Changes from D5.2.1} % (fold)
-\label{sub:changes}
+\subsection{Changes from previous version} % (fold)
+\label{sub:changes_from_previous_version}
 
@@
 RESTful Catalogue API has been intensively extended and modified in order to  reflect all the new functionality and how every request or response is constructed.
+Some of the most important features added in this iteration are:
+	\begin{itemize}
+		\item support for form elements, operators and back-end services (see Section ...),
+		\item convenience operations for recommendation and checking satisfaction and reachability for form elements, operators and back-end services (see Section ...),
+		\item and, last but not least, support for Screen planning (see Section \ref{sub:planning}).
+	\end{itemize}
 	\item Several appendixes have been added with the JSON structure of every building block the Catalogue can deal with.
@@
 
-% subsection changes (end)
+% subsection changes_from_previous_version (end)
 
@@
 
-The format supported by the Catalogue API for all the request and responses is JSON\footnote{http://www.json.org/}.
+The format supported, and in which they must be constructed, by every HTTP request is JSON\footnote{http://www.json.org/}.
 JSON is a lightweight data interchange format whose simplicity has resulted in widespread use among web developers, easy to read and write and able to using any programming language because its structures map directly to data structures used in most programming languages.
 
-Every HTTP request and response is encoded using the MIME type \texttt{application/json} and the charset \texttt{UTF-8}.
-
-% subsection json_interchange_format (end)
-
-\subsubsection{Example of a screen}
-\begin{verbatim}
-{
-  "uri": null,
-  "label": {"en-GB": "Screen - DERI"},
-  "description": {"en-GB": "Returns people working in DERI"},
-  "creationDate": "2009-02-07T09:59:52+0000",
-  "creator": "http://www.ismaelrivera.es/",
+Every HTTP request should be encoded using the MIME type \texttt{application/json} and the charset \texttt{UTF-8}.
+
+The following is an example of a Find \& Check request:
+
+\begin{verbatim}
+{
+  "canvas": [
+    { "uri": "http://localhost:8080/catalogue/screens/238" },
+    { "uri": "http://localhost:8080/catalogue/screens/323" }
+  ],
+  "elements": [
+    { "uri": "http://localhost:8080/catalogue/screens/636" }
+  ],
   "domainContext": {
-      "tags": [],
-      "user": null
-    },
-  "homepage": "http://www.deri.ie/",
-  "icon": "http://www.deri.ie/icon.jpg",
-  "screenshot": "http://www.deri.ie/screenshot.jpg",
-  "rights": "http://creativecommons.org/",
-  "version": "1.0",
-  "preconditions": [],
-  "postconditions": ["?person rdf:type foaf:Person . 
-                      ?person foaf:workplaceHomepage http://www.deri.ie/"]
+    "tags": [
+      {
+        "label": { "en-GB": "Amazon" },
+        "means": "http://dbpedia.org/page/Amazon.com"
+      }
+    ],
+    "user": "irivera",
+  },
+  "criterion": "reachability"
 }
@@
 In order to offers an adaptable solution to various languages and regions without major engineering changes, internationalisation is considered from an early stage in the catalogue development.
-Underlying representation technologies used to develop the catalogue (RDF and JSON) allow to implement this feature.
+Underlying representation technologies used to develop the catalogue (RDF/XML and JSON) allow to implement this feature.
 This feature is implemented by the addition of a language tag to every 'string' desired.
@@
 
-By now it is allowed using internationalisation for the label and the description of a screen, but more attributes will be benefited by this feature.
+Until now, using internationalisation is allowed for the label and the description of a screen, but in the near future, more attributes will benefit from this feature.
+
+% subsection json_interchange_format (end)
 
@@
 
-\subsubsection{CRUD operations for Screens} % (fold)
-
-A screen is a resource which can be created, retrieved, modified or deleted since the catalogue provides CRUD support for screens.
-These features are provided following the REST philosophy, using the HTTP methods.
-Two concepts have to be defined: \emph{collection}, as a set of screens which access URI is http://catalogueURL/screens/ where catalogueURL is the specific URL where the catalogue server is installed, and \emph{member} of the collection, in other words, the screen itself, which access URI is http://catalogueURL/screens/<id> where the <id> has to be replaced by the identifier of a screen.
-The details of the operations and which HTTP verb has to be used can be found in the Table \ref{tab:crud_screens}.
+\subsubsection{CRUD operations} % (fold)
+
+Any building block within the Catalogue is a resource in terms of REST philosophy, hence it can be created, retrieved, modified or deleted, using a certaing URL and a specific HTTP method for every operation.
+Two concepts have to be defined: \emph{collection}, as a set of resources which access URI is http://catalogueURL/<type>/ where catalogueURL is the specific URL where the Catalogue server is installed and <type> is the plural of the name of the building block (e.g.
+screens, services), and \emph{member} of the collection, in other words, the building block itself, which access URI is http://catalogueURL/<type>/<id> where the <id> has to be replaced by the identifier of a specific building block.
+The details of the operations and which HTTP verb has to be used can be found in the Table \ref{tab:crud_operations}.
 
 \begin{table}[htb]
-\caption{CRUD operations for Screens}
-\label{tab:crud_screens}
+\caption{CRUD operations}
+\label{tab:crud_operations}
 \begin{center}
@@
 Collection URI & PUT & N/A & Not used.\\ \hline
-Collection URI & POST & JSON representation of the screen & \textbf{Create} a new entry in the collection where the URI is assigned automatically by the collection.
+Collection URI & POST & JSON representation of the building block & \textbf{Create} a new entry in the collection where the URI is assigned automatically by the collection.
 The URI created is returned by this operation.\\ \hline
@@
 Member URI & GET & N/A & \textbf{Retrieve} the addressed member of the collection.\\ \hline
-Member URI & PUT & JSON representation of the screen & \textbf{Update} the addressed member of the collection or create it with a defined URI.\\ \hline
+Member URI & PUT & JSON representation of the building block & \textbf{Update} the addressed member of the collection or create it with a defined URI.\\ \hline
 Member URI & POST & N/A & Not used. \\ \hline
@@
 
-Here are shown various examples of request and responses using the catalogue server.
+The JSON structure for create any building bock can be found in the appendixes.
+All of them are composed of a common structure which can be found in Section \ref{appendix_b}, and some specific information depending on the type.
+Screen-flows are defined in \ref{appendix_c}, screens in \ref{appendix_d} and forms, operators and back-end services share the same structure detailed in \ref{appendix_e}.
+However, a few examples of request and responses are shown in order to clarify how the request are constructed and sent, and how the responses look like.
+
 To create a new screen, a POST request is send to the catalogue server, including the JSON representation of the screen as the body of the request:
@@
 {
-  "uri": null,
-  "label": {"en-GB": "Screen - DERI"},
-  "description": {"en-GB": "Returns people working in DERI"},
+  "code": "http://demo.fast.morfeo-project.org/code/amazonList.js",
+  "creationDate": "2010-01-26T17:01:13+0000",
+  "creator": "irivera",
+  "description": {"en-gb": "Please fill the description..."},
+  "homepage": "http://fast.morfeo-project.eu/",
+  "icon": "http://demo.fast.morfeo-project.org/images/amazonList.png",
+  "id": "28",
+  "label": {"en-gb": "Amazon Product List"},
+  "name": "ProductList",
+  "postconditions": [
+    [
+      {
+        "id": "item",
+        "label": { "en-gb": "An item" },
+        "pattern": "?I 
+                    http://www.w3.org/1999/02/22-rdf-syntax-ns#type
+                    http://aws.amazon.com/AWSECommerceService#Item",
+        "positive": "true"
+      }
+    ]
+  ],
+  "preconditions": [
+    [
+      {
+        "id": "filter",
+        "label": { "en-gb": "A search criteria" },
+        "pattern": "?F
+                    http://www.w3.org/1999/02/22-rdf-syntax-ns#type
+                    http://aws.amazon.com/AWSECommerceService#SearchCriteria",
+        "positive": "true"
+      }
+    ]
+  ],
+  "rights": "http://creativecommons.org/",
+  "screenshot": "http://demo.fast.morfeo-project.org/images/amazonProductList.png",
+  "tags": [{"label": {"en-gb": "amazon"}}],
+  "version": "0.1"
+}
+\end{verbatim}
+
+The URI of the screen is created using the id specified in the JSON request. So, the response for this operation is:
+
+\begin{verbatim}
+{
+  "code": "http://demo.fast.morfeo-project.org/code/amazonList.js",
+  "creationDate": "2010-01-26T17:01:13+0000",
+  "creator": "irivera",
+  "description": {"en-gb": "Please fill the description..."},
+  "homepage": "http://fast.morfeo-project.eu/",
+  "icon": "http://demo.fast.morfeo-project.org/images/amazonList.png",
+  "id": "28",
+  "label": {"en-gb": "Prueba1"},
+  "name": "ProductList",
+  "postconditions": [
+    [
+      {
+        "id": "item",
+        "label": { "en-gb": "An item" },
+        "pattern": "?I
+                    http://www.w3.org/1999/02/22-rdf-syntax-ns#type
+                    http://aws.amazon.com/AWSECommerceService#Item",
+        "positive": "true"
+      }
+    ]
+  ],
+  "preconditions": [
+    [
+      {
+        "id": "filter",
+        "label": { "en-gb": "A search criteria" },
+        "pattern": "?F
+                    http://www.w3.org/1999/02/22-rdf-syntax-ns#type
+                    http://aws.amazon.com/AWSECommerceService#SearchCriteria",
+        "positive": "true"
+      }
+    ]
+  ],
+  "rights": "http://creativecommons.org/",
+  "screenshot": "http://demo.fast.morfeo-project.org/images/amazonProductList.png",
+  "tags": [{"label": {"en-gb": "amazon"}}],
+  "type": "screen",
+  "uri": "http://localhost:8080/catalogue/screens/28",
+  "version": "0.1"
+}
+\end{verbatim}
+
+To obtain all the screens stored in the catalogue a HTTP GET request is sent to the Collection URI and the response may be something like this:
+\begin{verbatim}
+[
+  {
+    "code": "http://demo.fast.morfeo-project.org/code/amazonList.js",
+    "creationDate": "2010-01-26T17:01:13+0000",
+    "creator": "irivera",
+    "description": {"en-gb": "Please fill the description..."},
+    "homepage": "http://fast.morfeo-project.eu/",
+    "icon": "http://demo.fast.morfeo-project.org/images/amazonList.png",
+    "id": "28",
+    "label": {"en-gb": "Prueba1"},
+    "name": "ProductList",
+    "postconditions": [
+      [
+        {
+          "id": "item",
+          "label": { "en-gb": "An item" },
+          "pattern": "?I
+                      http://www.w3.org/1999/02/22-rdf-syntax-ns#type
+                      http://aws.amazon.com/AWSECommerceService#Item",
+          "positive": "true"
+        }
+      ]
+    ],
+    "preconditions": [
+      [
+        {
+          "id": "filter",
+          "label": { "en-gb": "A search criteria" },
+          "pattern": "?F
+                      http://www.w3.org/1999/02/22-rdf-syntax-ns#type
+                      http://aws.amazon.com/AWSECommerceService#SearchCriteria",
+          "positive": "true"
+        }
+      ]
+    ],
+    "rights": "http://creativecommons.org/",
+    "screenshot": "http://demo.fast.morfeo-project.org/images/amazonProductList.png",
+    "tags": [{"label": {"en-gb": "amazon"}}],
+    "type": "screen",
+    "uri": "http://localhost:8080/catalogue/screens/28",
+    "version": "0.1"
+  },
+  {
     "creationDate": "2009-02-07T09:59:52+0000",
-  "creator": "http://www.ismaelrivera.es/",
-    "domainContext": {
-      "tags": [],
-      "user": null
-    },
-  "homepage": "http://www.deri.ie/",
-  "icon": "http://www.deri.ie/icon.jpg",
-  "screenshot": "http://www.deri.ie/screenshot.jpg",
-  "rights": "http://creativecommons.org/",
-  "version": "1.0",
-  "preconditions": [],
-  "postconditions": ["?person rdf:type foaf:Person . 
-                      ?person foaf:workplaceHomepage http://www.deri.ie/"]
-}
-\end{verbatim}
-
-The URI will be not taken into account since the catalogue will generate a unique URI for each resource created. So, the response for this operation is:
-
-\begin{verbatim}
+    "creator": "fabio",
+    ...,
+	  "uri": "http://localhost:8080/catalogue/screens/14",
+    "version": "1.0"
+  },
   {
     "creationDate": "2009-02-07T09:59:52+0000",
-  "creator": "http://www.ismaelrivera.es/",
-  "description": {"en-GB": "Show information about a person working in DERI"},
-  "domainContext": {
-    "tags": [],
-    "user": "null"
-  },
-  "homepage": "http://www.deri.ie/",
-  "icon": "http://www.deri.ie/icon.jpg",
-  "label": {"en-GB": "Screen - DERI"},
-  "postconditions": [],
-  "preconditions": ["?person rdf:type foaf:Person . 
-                     ?person foaf:workplaceHomepage http://www.deri.ie/"],
-  "rights": "http://creativecommons.org/",
-  "screenshot": "http://www.deri.ie/screenshot.jpg",
-  "uri": "http://www.morfeoproject.eu/fast/fco#Screen1234552194269",
+    "creator": "javier",
+    ...,
+	  "uri": "http://localhost:8080/catalogue/screens/564",
     "version": "1.0"
   }
-\end{verbatim}
-
-To obtain all the screens stored in the catalogue a HTTP GET request is sent to the Collection URI and the response may be something like this:
-\begin{verbatim}
-[
-  {
-    "creationDate": "2009-02-07T09:59:52+0000",
-    "creator": "http://www.ismaelrivera.es/",
-    "description": {"en-gb": "Returns people working in DERI"},
-    "domainContext": {
-      "tags": [],
-      "user": null
-    },
-    "homepage": "http://www.deri.ie/",
-    "icon": "http://www.deri.ie/icon.jpg",
-    "label": {"en-gb": "Screen - DERI"},
-    "postconditions": ["?person rdf:type foaf:Person . 
-                        ?person foaf:workplaceHomepage http://www.deri.ie/"],
-    "preconditions": [],
-    "rights": "http://creativecommons.org/",
-    "screenshot": "http://www.deri.ie/screenshot.jpg",
-    "uri": "http://www.morfeoproject.eu/fast/fco#Screen1234552164472",
-    "version": "1.0"
-  },
-  {
-    "creationDate": "2009-02-07T09:59:52+0000",
-    "creator": "http://www.ismaelrivera.es/",
-    "description": {"en-gb": "Show information about a person working in DERI"},
-    "domainContext": {
-      "tags": [],
-      "user": null
-    },
-    "homepage": "http://www.deri.ie/",
-    "icon": "http://www.deri.ie/icon.jpg",
-    "label": {"en-gb": "Screen - DERI"},
-    "postconditions": [],
-    "preconditions": ["?person rdf:type foaf:Person . 
-                       ?person foaf:workplaceHomepage http://www.deri.ie/"],
-    "rights": "http://creativecommons.org/",
-    "screenshot": "http://www.deri.ie/screenshot.jpg",
-    "uri": "http://www.morfeoproject.eu/fast/fco#Screen1234552194269",
-    "version": "1.0"
-  },
-  {
-    "creationDate": "2009-02-07T09:59:52+0000",
-    "creator": "http://www.ismaelrivera.es/",
-    "description": {"en-gb": "Show information about a person"},
-    "domainContext": {
-      "tags": [],
-      "user": null
-    },
-    "homepage": "http://www.deri.ie/",
-    "icon": "http://www.deri.ie/icon.jpg",
-    "label": {"en-gb": "Screen - DERI"},
-    "postconditions": [],
-    "preconditions": ["?person rdf:type foaf:Person"],
-    "rights": "http://creativecommons.org/",
-    "screenshot": "http://www.deri.ie/screenshot.jpg",
-    "uri": "http://www.morfeoproject.eu/fast/fco#Screen1234552373865",
-    "version": "1.0"
-  }
 ]
@@
 
-\subsubsection{CRUD operations for Screen-Flows} % (fold)
-
-A screenflow is a resource which can be created, retrieved, modified or deleted since the catalogue provides CRUD support for screenflows.
-These features are provided following the REST philosophy, using the HTTP methods.
-As in the previous section the concepts \emph{collection} and \emph{member} are used including their URLs.
-Table \ref{tab:crud_screenflows} details the operations offered for screen-flows.
-
-\begin{table}[htb]
-\caption{CRUD operations for Screen-Flows}
-\label{tab:crud_screenflows}
-\begin{center}
-\begin{tabular}{|p{2.5cm}|p{2.5cm}|p{3.5cm}|p{5cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Resource} &
-                         \textcolor{white}{HTTP method} &
-                         \textcolor{white}{HTTP body} &
-                         \textcolor{white}{Description}\\ \hline
-Collection URI & GET & N/A & \textbf{List} the members of the collection.\\ \hline
-Collection URI & PUT & N/A & Not used.\\ \hline
-Collection URI & POST & JSON representation of the screen-flow & \textbf{Create} a new entry in the collection where the URI is assigned automatically by the collection.
-The URI created is returned by this operation.\\ \hline
-Collection URI & DELETE & N/A & Not used.\\ \hline
-Member URI & GET & N/A & \textbf{Retrieve} the addressed member of the collection.\\ \hline
-Member URI & PUT & JSON representation of the screen-flow & \textbf{Update} the addressed member of the collection or create it with a defined URI.\\ \hline
-Member URI & POST & N/A & Not used. \\ \hline
-Member URI & DELETE & N/A & \textbf{Delete} the addressed member of the collection.\\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-The following is an example of the request and response while creating a screen-flow:
-\begin{verbatim}
-{
-  "uri": null,
-  "label": {"en-GB": "My First ScreenFlow"},
-  "description": {"en-GB": "A example of screenflow"},
-  "creationDate": "2009-02-05T13:11:41+0000",
-  "creator": "http://www.ismaelrivera.es/",
-  "domainContext": {
-    "tags": ["people"],
-    "user": null
-  },
-  "homepage": "http://www.screenflow.com/",
-  "icon": "http://www.screenflow.com/icon.jpg",
-  "screenshot": "http://www.screenflow.com/screenshot.jpg",
-  "rights": "http://creativecommons.org/",
-  "version": "1.0",
-  "preconditions": [],
-  "postconditions": [],
-  "screens": [
-    "http://www.morfeoproject.eu/fast/fco#Screen1233750841383",
-    "http://www.morfeoproject.eu/fast/fco#Screen1233750828836",
-    "http://www.morfeoproject.eu/fast/fco#Screen1233750820914"
-  ]
-}
-\end{verbatim}
-
-\begin{verbatim}
-{
-  "creationDate": "2009-02-05T13:11:41+0000",
-  "creator": "http://www.ismaelrivera.es/",
-  "description": {"en-GB": "A example of screenflow"},
-  "domainContext": {
-    "tags": ["people"],
-    "user": null
-  },
-  "homepage": "http://www.screenflow.com/",
-  "icon": "http://www.screenflow.com/icon.jpg",
-  "label": {"en-GB": "My First ScreenFlow"},
-  "postconditions": [],
-  "preconditions": [],
-  "rights": "http://creativecommons.org/",
-  "screens": [
-    "http://www.morfeoproject.eu/fast/fco#Screen1233750841383",
-    "http://www.morfeoproject.eu/fast/fco#Screen1233750828836",
-    "http://www.morfeoproject.eu/fast/fco#Screen1233750820914"
-  ],
-  "screenshot": "http://www.screenflow.com/screenshot.jpg",
-  "uri": "http://www.morfeoproject.eu/fast/fco#ScreenFlow1233752391898",
-  "version": "1.0"
-}
-\end{verbatim}
-
-\subsubsection{Find} % (fold)
-
-The find operation search inside the catalogue for any screen which could be somehow related to the gadget the user is creating.
-It provides a recommended set of screens depending on the domain context, the canvas, etc.
-The specific URL to access this operation is http://catalogueURL/find using HTTP POST method.
-From now on, the method 'find' only considers the domain context, being a list of tags and the pre/postconditions of the screens.
+\subsubsection{Screen Find} % (fold)
+\label{ssub:screen_find}
+
+The find operation searches inside the catalogue for any screen which could be somehow related to the gadget the user is creating.
+It provides a recommended set of screens depending on the domain context, the canvas, and so on.
+
+The specific URL to access this operation is http://catalogueURL/screens/find using HTTP POST method.
+From now on, the method 'find' only considers the domain context and the pre/postconditions of the screens.
 It will try to satisfy all the unsatisfied preconditions of a given list of screens also known as canvas.
 The results will be all the screens stored in the catalogue which fulfil some of these preconditions, and are tagged with the tags specified in the domain context.
-The request parameters are shown in Table \ref{tab:find_request} and the response parameters are shown in Table \ref{tab:find_response}.
+The request parameters are shown in Table \ref{tab:screen_find_request} and the response parameters are shown in Table \ref{tab:screen_find_response}.
 
 \begin{table}[htb]
-\caption{Find Request Parameters}
-\label{tab:find_request}
+\caption{Screen Find Request Parameters}
+\label{tab:screen_find_request}
 \begin{center}
@@
                          \textcolor{white}{Type}\\ \hline
-Canvas & The canvas is composed by a list of screens. Only the URI is needed. & Optional \\ \hline
-Domain Context & The domain context contains a list of tags and a user. & Optional \\ \hline
-Elements & It is a list of resources, previously stored in the catalogue. For example, the list of screens recommended last execution. Only accepts screens. & Optional \\ \hline
+canvas & The canvas is composed by a list of screens. Only the URI is needed. & Optional \\ \hline
+domainContext & The domain context contains a list of tags and a user. & Optional \\ \hline
+elements & It is a list of resources, previously stored in the catalogue. For example, the list of screens recommended last execution. Only accepts screens. & Optional \\ \hline
 \end{tabular}
@@
 \begin{table}[htb]
-\caption{Find Response Parameters}
-\label{tab:find_response}
+\caption{Screen Find Response Parameters}
+\label{tab:screen_find_response}
 \begin{center}
@@
                          \textcolor{white}{Type}\\ \hline
-N/A & A list of recommended screens is returned. & Required \\ \hline
+N/A & A list of recommended screens URIs is returned. & Required \\ \hline
 \end{tabular}
@@
 
-Following is shown an example of usage of this operation considering a canvas with the screen \texttt{http://www.morfeoproject.eu/fast/fco\#Screen1234552373865}.
+Following is shown an example of usage of this operation considering a canvas with the screen \texttt{http://localhost:8080/catalogue/screens/654}.
 
@@
   canvas: [
-    {
-      uri: "http://www.morfeoproject.eu/fast/fco#Screen1234552373865"
-    }
+    { uri: "http://localhost:8080/catalogue/screens/654" }
   ],
@@
 \begin{verbatim}
-[{
-  "creationDate": "2009-02-07T09:59:52+0000",
-  "creator": "http://www.ismaelrivera.es/",
-  "description": {"en-gb": "Returns people working in DERI"},
-  "domainContext": {
-    "tags": [],
-    "user": null
-  },
-  "homepage": "http://www.deri.ie/",
-  "icon": "http://www.deri.ie/icon.jpg",
-  "label": {"en-gb": "Screen - DERI"},
-  "postconditions": ["?person rdf:type foaf:Person . 
-                      ?person foaf:workplaceHomepage http://www.deri.ie/"],
-  "preconditions": [],
-  "rights": "http://creativecommons.org/",
-  "screenshot": "http://www.deri.ie/screenshot.jpg",
-  "uri": "http://www.morfeoproject.eu/fast/fco#Screen1234552164472",
-  "version": "1.0"
-}]
-\end{verbatim}
-
-\subsubsection{Check} % (fold)
+[
+  "http://localhost:8080/catalogue/screens/371",
+  "http://localhost:8080/catalogue/screens/24",
+  "http://localhost:8080/catalogue/screens/253",
+  "http://localhost:8080/catalogue/screens/12"
+]
+\end{verbatim}
+
+\subsubsection{Screen Check} % (fold)
+\label{ssub:screen_check}
 
@@
 A screen will be reachable if it all its preconditions are satisfied by postconditions of the screens the canvas contain.
-The specific URL to access this operation is http://catalogueURL/check using HTTP POST method.
+The specific URL to access this operation is http://catalogueURL/screens/check using HTTP POST method.
 
 \begin{table}[htb]
-\caption{Check Request Parameters}
-\label{tab:check_request}
+\caption{Screen Check Request Parameters}
+\label{tab:screen_check_request}
 \begin{center}
@@
                          \textcolor{white}{Type}\\ \hline
-Canvas & The canvas is composed by a list of screens. Only the URI is needed. & Optional\\ \hline
-Domain Context & The domain context contains a list of tags and a user. & Optional\\ \hline
-Elements & It is a list of resources, previously stored in the catalogue. For example, the list of screens recommended last execution. Only accepts screens. & Optional\\ \hline
-Criterion & The criterion specifies what it has to be check. The only possible supported now is 'reachability' in order to check if the preconditions of a screen are satisfied. & Required\\ \hline
+canvas & The canvas is composed by a list of screens. Only the URI is needed. & Optional\\ \hline
+domainContext & The domain context contains a list of tags and a user. & Optional\\ \hline
+elements & It is a list of resources, previously stored in the catalogue. For example, the list of screens recommended last execution. Only accepts screens. & Optional\\ \hline
+criterion & The criterion specifies what it has to be check. The only possible supported now is 'reachability' in order to check if the preconditions of a screen are satisfied. & Required\\ \hline
 \end{tabular}
@@
 \begin{table}[htb]
-\caption{Check Response Parameters}
-\label{tab:check_response}
+\caption{Screen Check Response Parameters}
+\label{tab:screen_check_response}
 \begin{center}
@@
 
-The scenario for the following example is composed by a screen in the canvas which precondition is a foaf:Person which foaf:workplaceHomepage is http://www.deri.ie/, and a screen in the elements list which precondition is a foaf:Person.
+The scenario for the following example is composed by a screen in the canvas which precondition is a foaf:Person which foaf:workplaceHomepage is \url{http://www.deri.ie/}, and a screen in the elements list which precondition is a foaf:Person.
 The state to check in this case is the reachability and satisfaction of the screens and preconditions.
@@
 {
-  canvas: [
-    {
-      uri: "http://www.morfeoproject.eu/fast/fco#Screen1234552194269"
-    }
-  ],
-  domainContext: [],
-  elements: [
-    {
-      uri: "http://www.morfeoproject.eu/fast/fco#Screen1234552373865"
-    }  
-  ],
+  "canvas": [
+    { "uri": "http://localhost:8080/catalogue/screens/238" }
+  ],
+  "elements": [
+    { "uri": "http://localhost:8080/catalogue/screens/636" }
+  ],
+  "domainContext": {
+    "tags": [
+      {
+        "label": { "en-GB": "Amazon" },
+        "means": "http://dbpedia.org/page/Amazon.com"
+      }
+    ],
+    "user": "irivera",
+  },
   criterion: "reachability"
@@
 
-As you can see in the response, no screen is reachable at the moment, and no precondition is satisfied.
+The response shows that there are not reachable screens at the moment, and preconditions are not satisfied.
 
@@
     "preconditions": [{
-      "expression": "?person rdf:type foaf:Person . 
+      "label": {"en-gb": "A person working in DERI"},
+      "pattern": "?person rdf:type foaf:Person . 
                   ?person foaf:workplaceHomepage http://www.deri.ie/",
+      "positive": true,
       "satisfied": false
@@
     "reachability": false,
-    "uri": "http://www.morfeoproject.eu/fast/fco#Screen1234552194269"
+    "uri": "http://localhost:8080/catalogue/screens/238"
   }],
@@
     "preconditions": [{
-      "expression": "?person rdf:type foaf:Person",
+      "label": {"en-gb": "A person"},
+      "pattern": "?person rdf:type foaf:Person",
       "satisfied": false
@@
     "reachability": false,
-    "uri": "http://www.morfeoproject.eu/fast/fco#Screen1234552373865"
+    "uri": "http://localhost:8080/catalogue/screens/636"
   }]
@@
 
+\subsubsection{Screen Find \& Check} % (fold)
+\label{ssub:screen_findcheck}
+
+For convenience and to minimise the number of request, the Find operation may be combined with the Check operation.
+This operation will look for new screens based on the given information, and will perform the Check to both the screens in the canvas and the screens in the \emph{elements} attribute.
+
+The format of the request is the same as the one used in the Check operation:
+
+\begin{verbatim}
+{
+  "canvas": [
+    { "uri": "http://localhost:8080/catalogue/screens/238" }
+  ],
+  "elements": [
+    { "uri": "http://localhost:8080/catalogue/screens/636" }
+  ],
+  "domainContext": {
+    "tags": [
+      {
+        "label": { "en-GB": "Amazon" },
+        "means": "http://dbpedia.org/page/Amazon.com"
+      }
+    ],
+    "user": "irivera",
+  },
+  criterion: "reachability"
+}
+\end{verbatim}
+
+The response for the above request is:
+
+\begin{verbatim}
+{
+  "canvas": [{
+    "preconditions": [{
+      "label": {"en-gb": "A person working in DERI"},
+      "pattern": "?person rdf:type foaf:Person . 
+                  ?person foaf:workplaceHomepage http://www.deri.ie/",
+      "positive": true,
+      "satisfied": false
+    }],
+    "reachability": false,
+    "uri": "http://localhost:8080/catalogue/screens/238"
+  }],
+  "elements": [
+    {
+      "preconditions": [{
+        "label": {"en-gb": "A person"},
+        "pattern": "?person rdf:type foaf:Person",
+        "satisfied": false
+      }],
+      "reachability": false,
+      "uri": "http://localhost:8080/catalogue/screens/636"
+    },
+    {
+      "preconditions": [],
+      "reachability": true,
+      "uri": "http://localhost:8080/catalogue/screens/132"
+    }
+  ]
+}
+\end{verbatim}
+
 \subsubsection{GetMetadata} % (fold)
@@
 Although the information about a resource can be obtained by its specific retrieval operation supported by the CRUD interface, sometimes is needed to retrieve information which would imply a request per resource, hence this operation allow to get the metadata of a list of resources in only one request.
-This prototype supports this operation for screenflows and screens.
+This prototype supports this operation for screen-flows and screens.
 The specific URL to access this operation is http://catalogueURL/getmetadata using HTTP POST method.
@@
                          \textcolor{white}{Type}\\ \hline
-Screenflows & A set of screenflows with all the metadata associated to them.. & Optional\\ \hline
-Screens & A set of screens with all the metadata associated to them. & Optional\\ \hline
+screen-flows & A set of screen-flows with all the metadata associated to them.. & Optional\\ \hline
+screens & A set of screens with all the metadata associated to them. & Optional\\ \hline
+forms & A set of forms with all the metadata associated to them. & Optional\\ \hline
+operators & A set of operators with all the metadata associated to them. & Optional\\ \hline
+backendservices & A set of back-end services with all the metadata associated to them. & Optional\\ \hline
 \end{tabular}
@@
 [
-  "http://www.morfeoproject.eu/fast/fco#Screen1234552164472",
-  "http://www.morfeoproject.eu/fast/fco#Screen1234552194269",
-  "http://www.morfeoproject.eu/fast/fco#ScreenFlow1233752391898"
+  "http://localhost:8080/catalogue/screens/12",
+  "http://localhost:8080/catalogue/screens/48",
+  "http://localhost:8080/catalogue/forms/9",
 ]
@@
 {
-  "screenflows": [{
-    "creationDate": "2009-02-05T13:11:41+0000",
-    "creator": "http://www.ismaelrivera.es/",
-    "description": {"en-gb": "A example of screenflow"},
-    "domainContext": {
-      "tags": ["people"],
-      "user": null
-    },
-    "homepage": "http://www.screenflow.com/",
-    "icon": "http://www.screenflow.com/icon.jpg",
-    "label": {"en-gb": "My First ScreenFlow"},
+  "screenflows": [],
+  "screens": [
+    {
+      "creationDate": "2009-02-07T09:59:52+0000",
+      "creator": "irivera",
+      ...,
+      "uri": "http://localhost:8080/catalogue/screens/12",
+      "version": "1.0"
+    },
+    {
+      "creationDate": "2009-02-07T09:59:52+0000",
+      "creator": "irivera",
+      ...,
+      "uri": "http://localhost:8080/catalogue/screens/12",
+      "version": "1.0"
+    }
+  ],
+  "forms": [
+    {
+      "actions": [
+        {
+          "name": "init",
+          "preconditions": [],
+          "uses": []
+        },
+        {
+          "name": "showTable",
+          "preconditions": [{
+            "id": "list",
+            "label": {"en-gb": "A product list"},
+            "pattern": "?PList
+                        http://www.w3.org/1999/02/22-rdf-syntax-ns#type
+                        http://aws.amazon.com/AWSECommerceService#ProductList",
+            "positive": true
+          }],
+          "uses": []
+        }
+      ],
+      ...,
+      "uri": "http://localhost:8080/FASTCatalogue/forms/9",
+      "version": "1.0"
+    },
+  ],
+  "operators": [],
+  "backendservices": []
+}
+\end{verbatim}
+
+\subsubsection{Screen Component Find \& Check} % (fold)
+\label{ssub:screen_component_findcheck}
+
+Similarly to the Screen Find \& Check described in Section \ref{ssub:screen_findcheck}, this operation search for screen components (forms, operators and back-end services) through the catalogue in order to satisfy a given request, and it will attach information about satisfaction and reachability as well.
+
+A request is formed of the following parameters:
+
+\begin{description}
+	\item[preconditions] Preconditions of the screen.
+	\item[postconditions] Postcondition of the screen.
+	\item[canvas] The canvas is composed by a list of screens. Only the URI is needed.
+	\item[forms] It is a list of forms, such as the list of forms recommended last execution.
+	\item[operators] It is a list of operators, such as the list of operators recommended last execution.
+	\item[backendservices] It is a list of back-end services, such as the list of back-end services recommended last execution.
+	\item[domainContext] The domain context contains a list of tags and a user.
+	\item[pipes] List of pipes for the connection between screen components.
+	\item[selectedItem] URI of a screen component from the canvas.
+\end{description}
+
+The screen components recommended are based on the given domain context and the pre/postconditions of the any of the resources in the canvas or the conditions of the parameters preconditions and postconditions.
+Hence, any screen component from the catalogue which has any of these conditions may be suitable for the screen, so it is included in the response, in the list corresponding to the type of screen component.
+
+An example of request is the following:
+
+\begin{verbatim}
+{
+  "preconditions": [{
+    "id": "Searchcriteria_1",
+    "label": {"en-gb": "A search criteria"},
+    "pattern": "?F
+                http://www.w3.org/1999/02/22-rdf-syntax-ns#type
+                http://aws.amazon.com/AWSECommerceService#SearchCriteria",
+    "positive": true,
+  }],
   "postconditions": [],
-    "preconditions": [],
-    "rights": "http://creativecommons.org/",
-    "screens": [
-      "http://www.morfeoproject.eu/fast/fco#Screen1233750820914",
-      "http://www.morfeoproject.eu/fast/fco#Screen1233750828836",
-      "http://www.morfeoproject.eu/fast/fco#Screen1233750841383"
-    ],
-    "screenshot": "http://www.screenflow.com/screenshot.jpg",
-    "uri": "http://www.morfeoproject.eu/fast/fco#ScreenFlow1233752391898",
-    "version": "1.0"
-  }],
-  "screens": [
-    {
-      "creationDate": "2009-02-07T09:59:52+0000",
-      "creator": "http://www.ismaelrivera.es/",
-      "description": {"en-gb": "Returns people working in DERI"},
+  "canvas": [
+    { "uri": "http://localhost:8080/catalogue/forms/1" },
+    { "uri": "http://localhost:8080/catalogue/services/1" }
+  ],
+  "forms": [],
+  "operators": [],
+  "backendservices": [],
   "domainContext": {
@@
   },
-      "homepage": "http://www.deri.ie/",
-      "icon": "http://www.deri.ie/icon.jpg",
-      "label": {"en-gb": "Screen - DERI"},
-      "postconditions": ["?person rdf:type foaf:Person . 
-                          ?person foaf:workplaceHomepage http://www.deri.ie/"],
+  "pipes": [
+    {
+      "from": {
+        "buildingblock": "http://localhost:8080/catalogue/services/1",
+        "condition": "list"
+      },
+      "to": {
+        "action": "showTable",
+        "buildingblock": "http://localhost:8080/catalogue/forms/1",
+        "condition": "list"
+      }
+    }
+  ],
+  "selectedItem": "http://localhost:8080/catalogue/forms/1"
+}
+\end{verbatim}
+
+And the response according to the above request is:
+
+\begin{verbatim}
+{
+  "backendservices": [],
+  "canvas": [
+    {
+      "actions": [{
+        "name": "search",
+        "preconditions": [{
+          "id": "filter",
+          "label": {"en-gb": "A search criteria"},
+          "pattern": "?F 
+                      http://www.w3.org/1999/02/22-rdf-syntax-ns#type
+                      http://aws.amazon.com/AWSECommerceService#SearchCriteria",
+          "positive": true,
+          "satisfied": false
+        }],
+        "satisfied": false,
+        "uses": []
+      }],
+      "reachability": false,
+      "uri": "http://localhost:8080/catalogue/services/1"
+    },
+    {
+      "actions": [
+        {
+          "name": "init",
           "preconditions": [],
-      "rights": "http://creativecommons.org/",
-      "screenshot": "http://www.deri.ie/screenshot.jpg",
-      "uri": "http://www.morfeoproject.eu/fast/fco#Screen1234552164472",
-      "version": "1.0"
-    },
-    {
-      "creationDate": "2009-02-07T09:59:52+0000",
-      "creator": "http://www.ismaelrivera.es/",
-      "description": {"en-gb": "Show information about a person working in DERI"},
-      "domainContext": {
-        "tags": [],
-        "user": null
-      },
-      "homepage": "http://www.deri.ie/",
-      "icon": "http://www.deri.ie/icon.jpg",
-      "label": {"en-gb": "Screen - DERI"},
-      "postconditions": [],
-      "preconditions": ["?person rdf:type foaf:Person . 
-                         ?person foaf:workplaceHomepage http://www.deri.ie/"],
-      "rights": "http://creativecommons.org/",
-      "screenshot": "http://www.deri.ie/screenshot.jpg",
-      "uri": "http://www.morfeoproject.eu/fast/fco#Screen1234552194269",
-      "version": "1.0"
-    }
-  ]
-}
-\end{verbatim}
-
-\subsubsection{Screen Find \& Check} % (fold)
-
-Request
-\begin{verbatim}
-{
-  "canvas": [
-    { "uri": "http://purl.oclc.org/fast/ontology/gadget#Screen1257860786707" }
-  ],
-  "elements": [
-  ],
-  "domainContext": {
-    "tags": [],
-    "user": null,
-  },
-  "criterion": "reachability"
-}
-\end{verbatim}
-
-Response
-\begin{verbatim}
-{
-  "canvas": [{
-    "preconditions": [[{
-      "id": "sc",
-      "label": {"en-gb": "A search criteria"},
-      "pattern": "?F http://www.w3.org/1999/02/22-rdf-syntax-ns#type http://aws.amazon.com/AWSECommerceService#SearchCriteria",
+          "satisfied": true,
+          "uses": []
+        },
+        {
+          "name": "showTable",
+          "preconditions": [{
+            "id": "list",
+            "label": {"en-gb": "A product list"},
+            "pattern": "?PList
+                        http://www.w3.org/1999/02/22-rdf-syntax-ns#type
+                        http://aws.amazon.com/AWSECommerceService#ProductList",
+            "positive": true,
+            "satisfied": false
+          }],
           "satisfied": false,
-      "positive": true
-    }]],
-    "reachability": false,
-    "uri": "http://purl.oclc.org/fast/ontology/gadget#Screen1257860786707"
-  }],
-  "elements": [{
-    "preconditions": [],
+          "uses": []
+        }
+      ],
       "reachability": true,
-    "uri": "http://purl.oclc.org/fast/ontology/gadget#Screen1257860530032"
-  }]
-}
-\end{verbatim}
+      "uri": "http://localhost:8080/catalogue/forms/1"
+    }
+  ],
+  "connections": [
+    {
+      "from": {
+        "buildingblock": null,
+        "condition": "Searchcriteria_1"
+      },
+      "to": {
+        "action": "search",
+        "buildingblock": "http://localhost:8080/catalogue/services/1",
+        "condition": "filter"
+      }
+    },
+  ],
+  "forms": [],
+  "operators": [],
+  "pipes": [
+    {
+      "from": {
+        "buildingblock": "http://localhost:8080/catalogue/services/1",
+        "condition": "list"
+      },
+      "satisfied": true,
+      "to": {
+        "action": "showTable",
+        "buildingblock": "http://localhost:8080/catalogue/forms/1",
+        "condition": "list"
+      }
+    }
+  ],
+  "postconditions": []
+}
+\end{verbatim}
+
+For a screen component to be reachable, its preconditions need to be connected by a pipe to any other screen component which is already reachable, or to any of the screen preconditions.
+Moreover, the pipe needs to be satisfied, in other words, conditions in both edges are compatible.
+
+The response includes an attribute called "connections". This attribute will be retrieved when "selectedItem" is given, and it will contain a list of potential pipes to connect to that item.
+
+% subsubsection screen_component_findcheck (end)
 
@@
 
-URL: <CATALOGUE\_URL>/planner
+While creating a screen-flow, the user need to find proper screens which will satisfy her necessity.
+Once the user has selected one or several screens to be included in her screen-flow, she needs to find screens which satisfy the preconditions of these screens, in order to create an executable screen-flow.
+The operation Find, detailed in Section \ref{ssub:screen_find}, helps the user to find these screens, however the response is just a set of screens which somehow may help to accomplish her goal, but not in a single and straight-forward step.
+
+This functionality offers to the user the possibility to receive a whole set of screens, or plan, which will make a given screen, or goal, reachable.
+These plans take into account the screen which are already in the canvas, in order to minimise the number of screens.
+
+The specific URL to access this operation is \url{http://catalogueURL/planner}, and to execute the operation a POST request needs to be done. The request body is composed of:
+
+\begin{description}
+	\item[goal] Required. URI of the goal screen.
+	\item[canvas] Required. List of screens URIs.
+	\item[page] Optional. Number of the page to be retrieved.
+	\item[per\_page] Optional. Number of plans to be included per page in the response.
+\end{description}
+
+Let's see the following example:
+
 Request
@@
 {
-  "goal": "http://purl.oclc.org/fast/ontology/gadget#Screen1257860720035",
-  "canvas": []
-}
-\end{verbatim}
-
-\begin{verbatim}
+  "goal": "http://localhost:8080/catalogue/screens/35",
+  "canvas": [
+    { "uri": "http://localhost:8080/catalogue/screens/56" },
+    { "uri": "http://localhost:8080/catalogue/screens/85" }
+  ],
+  "page": 1,
+  "per_page": 10 
+}
+\end{verbatim}
+
 Response
+\begin{verbatim}
 [
   [
-    "http://purl.oclc.org/fast/ontology/gadget#Screen1257866875435",
-    "http://purl.oclc.org/fast/ontology/gadget#Screen1257865645621"
+    "http://localhost:8080/catalogue/screens/13",
+    "http://localhost:8080/catalogue/screens/23"
   ],
   [
-    "http://purl.oclc.org/fast/ontology/gadget#Screen1257860530032",
-    "http://purl.oclc.org/fast/ontology/gadget#Screen1257860786707",
-    "http://purl.oclc.org/fast/ontology/gadget#Screen1257860772723"
+    "http://localhost:8080/catalogue/screens/79",
+    "http://localhost:8080/catalogue/screens/37",
+    "http://localhost:8080/catalogue/screens/14"
+  ],
+  [
+    "http://localhost:8080/catalogue/screens/16",
+    "http://localhost:8080/catalogue/screens/44",
+    "http://localhost:8080/catalogue/screens/32"
   ]
@@
 \end{verbatim}
+
+The response is ranked based on adding the minimum number of new screens into the canvas, and giving preference to the screens the user has previously selected (canvas).
+
+In the above example, let's assume the first plan is using the two screens from the canvas to accomplish the goal, and the second plan is able to make the goal reachable by itself.
+That said, a screen-flow with the goal and the second plan could be executable, and it would be composed of just four screens, however, taking the first plan, the resulting screen-flow will be composed of five screens.
+It could make sense to place the second plan in first position, but the plan with just two screens reuse the screens from the canvas which has been added by the user.
 
@@
 \section*{Appendix A. Lists of Tables and Figures}
+\label{appendix_a}
 \addcontentsline{toc}{section}{Appendix A. Lists of Tables and Figures}
@@
 \section*{Appendix B. Generic Building Block JSON Structure}
+\label{appendix_b}
 \addcontentsline{toc}{section}{Appendix B. Generic Building Block JSON Structure}
@@
 \clearpage
-\section*{Appendix C. Screenflow JSON Structure}
-\addcontentsline{toc}{section}{Appendix C. Screenflow JSON Structure}
+\section*{Appendix C. Screen-flow JSON Structure}
+\label{appendix_c}
+\addcontentsline{toc}{section}{Appendix C. Screen-flow JSON Structure}
 
@@
 \section*{Appendix D. Screen JSON Structure}
+\label{appendix_d}
 \addcontentsline{toc}{section}{Appendix D. Screen JSON Structure}
@@
   "postconditions": [[{
-    "id": "url",
+    "id": <identifier:String>,
     "label": {"en-gb": "Purchase URL"},
@@
 \section*{Appendix E. Screen Component JSON Structure}
+\label{appendix_e}
 \addcontentsline{toc}{section}{Appendix E. Screen Component JSON Structure}

### 2010-01-28 trunk/deliverables/D2_2_Ontology/fast_d2_2_2_ontology.tex
finished with templates, classes, instances part, started with basic annotations
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@482 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 	\item Sect.~\ref{sub:goal_and_scope} has been revised to better define the scope of this deliverable with respect to other deliverables.
-	\item The section on our adopted development methodology (previously Sect.~1.3) has been moved to the end of the document (Sect.~\ref{sub:methontology_a_methodology_for_ontology_development}), but otherwise did not change.
+	\item The list of classes and properties has been moved to the appendix section at the back of the document (App.~\ref{sec:ontology_terms}), because it was felt that it would otherwise hinder the flow of the text.
+	\item In response to the M12 review, a UML overview of the ontology in App.~\ref{sec:ontology_overview} now complements the list of terms.
+	\item The section on our adopted development methodology (previously Sect.~1.3) has been moved to the end of the document (App.~\ref{sub:methontology_a_methodology_for_ontology_development}), but otherwise did not change.
 	\item An evaluation section has been added (Sect.~\ref{sec:evaluation}).
@@
 \textbf{Properties:} \emph{containsScreen}, \emph{hasLabel}, \emph{hasIcon}, \emph{hasPreCondition}, \emph{hasPostCondition}, \emph{hasTag}, \emph{hasProfile}\\ \hline
-\textcolor{white}{\textbf{Sources of Knowledge}} & FAST Architecture deliverable~\cite{solero2009fast_architecture}, FAST Requirements Specification~\cite{villoslada2010fast_requirements}, FAST State-of-the-art Document~\cite{urmetzer2010fast_state_of_the_art}, EzWeb documentation~\cite{lizcano2008ezweb} \\ \hline
+\textcolor{white}{\textbf{Sources of Knowledge}} & FAST Architecture deliverable~\cite{urena2010fast_architecture}, FAST Requirements Specification~\cite{villoslada2010fast_requirements}, FAST State-of-the-art Document~\cite{urmetzer2010fast_state_of_the_art}, EzWeb documentation~\cite{lizcano2008ezweb} \\ \hline
 \end{longtable}
@@
 \texttt{hasDescription} & DC & \texttt{dcterms:description} & --- \\ \hline
-\texttt{hasCreationDate} & DC & \texttt{created} & Dates should be formatted according to ISO 8601. \\ \hline
-\texttt{hasRights} & DC & \texttt{rights} & DC terms specify a number of classes to model the rights of resources.
+\texttt{hasCreationDate} & DC & \texttt{dcterms:created} & Dates should be formatted according to ISO 8601. \\ \hline
+\texttt{hasRights} & DC & \texttt{dcterms:rights} & DC terms specify a number of classes to model the rights of resources.
 \texttt{dcterms:rights} links a resource to a \texttt{dcterms:RightsStatement}.
@@
 \\ \hline
-\texttt{hasDomainContext} & DC & \texttt{subject} & DC specifies that subject should be non-literals. To allow simple tagging, we should represent tags as resources (TODO). \\ \hline
-\texttt{hasImage} & FOAF & \texttt{depiction} & The inverse \texttt{foaf:depicts} could also be used. \\ \hline
-\texttt{hasName} & FOAF & \texttt{name} & Apart from \texttt{foaf:name}, which is used for the full name, more fine-grained properties such as \texttt{foaf:firstName}, \texttt{foaf:surname} or \texttt{foaf:title} and \texttt{foaf:nick} can be used as well.
+\texttt{hasDomainContext} & DC & \texttt{dcterms:subject} & DC specifies that subject should be non-literals. To allow simple tagging, we should represent tags as resources (TODO). \\ \hline
+\texttt{hasImage} & FOAF & \texttt{foaf:depiction} & The inverse \texttt{foaf:depicts} could also be used. \\ \hline
+\texttt{hasIcon} & FOAF & sub-property of \texttt{foaf:depiction} & Icons are a specific kind of image, so we can relate this property to the \texttt{foaf:depiciton}. \\ \hline
+\texttt{hasScreenshot} & FOAF & sub-property of \texttt{foaf:depiction} & Screenshots are a specific kind of image, so we can relate this property to the \texttt{foaf:depiciton}. \\ \hline
+\texttt{hasName} & FOAF & \texttt{foaf:name} & Apart from \texttt{foaf:name}, which is used for the full name, more fine-grained properties such as \texttt{foaf:firstName}, \texttt{foaf:surname} or \texttt{foaf:title} and \texttt{foaf:nick} can be used as well.
 \\ \hline
-\texttt{hasEmail} & FOAF & \texttt{mbox} & While \texttt{foaf:mbox} will be used for the plain e-mail address, \texttt{foaf:mbox\_sha1sum} will be used for an obfuscated version of the address.
+\texttt{hasEmail} & FOAF & \texttt{foaf:mbox} & While \texttt{foaf:mbox} will be used for the plain e-mail address, \texttt{foaf:mbox\_sha1sum} will be used for an obfuscated version of the address.
 \\ \hline
@@
 
-In the following sections we will define the FAST gadget ontology.
+While the previous section discussed the various steps preceding the definition of the actual ontology according to our chosen development methodology, this section focusses on concrete usage of the FAST gadget ontology in an RDFS/OWL environment.
+
+\subsection{Classes, Instances and Templates} % (fold)
+\label{sub:classes_instances_and_templates}
+
+In FAST, resources (gadgets, screenflows, screens and their components) exist in different stages throughout their lifecycle.
+Moving top-down from the high-level conceptual description to gadgets in use at runtime, the following differences can be made in terms of the level of instantiation:
+
+\begin{itemize}
+	\item \textbf{Abstract Class Level}: At this level, resources are defined on a categorical level, differentiating between the various kinds of things that are relevant in FAST, as fundamental classes such as screens, forms or services (see Sect.~\ref{ssub:glossary_of_terms}).
+These are implemented technically as classes in OWL.
+	\item \textbf{Screenflow-independent Component Definition}: When a user of the FAST GVS is building a new screenflow or screen, they do so by combining existing building blocks and configuring them.
+These building blocks are not the fundamental classes of FAST, but instead they are specific instances of those classes.
+In other words, users do not combine the idea of a screen with the idea of a service, but instead they combine the specific ``Log In Screen'' with the ``Amazon Web Service''.
+Technically, these concrete building blocks are implemented as instances of the fundamental classes, specified by asserting particular settings such as label or icons, pre- and post-conditions, etc.
+	\item \textbf{Instantiation within a Screenflow at Design-time}: Once a user selects a specific building block in the GVS and adds it to their new screen or screenflow project, they are in principle again instantiating from a conceptual to a concrete level: from the ``Log In Screen'' as such to a particular log in screen as used in that particular screenflow.
+Also in this stage, resources are implemented technically as instances.
+	\item \textbf{Runtime}: At the final level of instantiation, the screenflow has been packaged as a gadget, deployed to a gadget platform and is now being used by an end-user.
+At this level, all gadget components exist only in the form of Web standards compliant code (HTML, JavaScript, CSS).
+Any metadata still necessary or definitions for screen pre- and post-conditions are for performance now defined in the JavaScript native data format JSON, rather than in RDF.
+Therefore, at this level, the resources that make up the gadget are out of the scope of this deliverable.
+\end{itemize}
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics[width=\linewidth]{images/classes_templates_instances.pdf}
+    \caption{Different levels of representation of FAST resources}
+    \label{fig:classes_templates}
+  \end{center}
+\end{figure}
+
+When implementing these four stages, as illustrated in Fig.~\ref{fig:classes_templates}, to OWL DL semantics, one will be faced with an immediate problem.
+Conceptually, once would like to say that $Screen$ is a class, i.e., the set of all screens.
+As argues above, $LogInScreen$ would then be an instance of $Screen$ ($LogInScreen \subset Screen$), while a particular $LogInScreen_x$ in a particular screenflow would again be an instance of $LogInScreen$ ($LogInScreen_x \subset LogInScreen$).
+However, unless we want to move into OWL Full semantics, classes cannot be used as instances at the same time, so we cannot use this modelling approach.
+Furthermore, when we instantiate a resource such as a screen within a particular screenflow, we want to imply that the new instance ($LogInScreen_x$) has the same features such as label, icon or conditions, as $LogInScreen$.
+In other words, what we really require is a something like a \emph{copy} of the original $LogInScreen$.
+
+Neither OWL nor RDFS semantics offer functionality to represent a relationship such as the copy of a resource formally.
+For this reason, we introduce the concept of \emph{templates for RDF instances} in FAST.
+One resource (e.g., $LogInScreen$) will play the role of a \emph{template}, while the other resource (e.g., $LogInScreen_x$) will be called a \emph{copy}.
+Conceptually, copying a resource means defining a named graph that contains all statements about this resource considered to be relevant, and creating a new graph with near-identical statements\footnote{The graphs are not completely identical: the identifiers for the resource and potentially other, contained resources will have changed.}.
+In practice, this procedure can be implemented using a SPARQL CONSTRUCT query, where the WHERE clause matches the graph pattern of the template, and the CONSTRUCT query will build the copy from it.
+The corresponding data on all three levels is shown in List.~\ref{list:templating}, using TriG syntax~\cite{bizer2004trig}.
+
+\singlespacing
+\lstset{
+	caption={Templating example --- different levels of representation of a FAST screen}, 
+	label=list:templating,
+	language=turtle
+}
+\begin{figure}[ht]
+	\lstinputlisting{exampleCode/templating_example.ttl}
+\end{figure}
+\doublespacing
+
+It should be pointed out that the term \emph{RDF template} has been used before, but with a different meaning.
+\cite{davis2003rdf_template} use the term in analogy to XSLT, but acting on an RDF graph rather than an XML tree.
+\cite{kawamoto2006kawawiki} present a semantic wiki, where templates help non-technical users to create semantic descriptions.
+% subsection classes_instances_and_templates (end)
+
+\subsection{Basic Annotation} % (fold)
+\label{sub:basic_annotation}
+
+In this section, we will present the use of some basic resource annotation properties in FAST.
+In general, since screenflows, screens and their components are sub-classes of \htmlref{\texttt{fgo:Resource}}{subs:Resource}, all properties defined to have a domain of \htmlref{\texttt{fgo:Resource}}{subs:Resource} should be used.
+In addition, a number of terms from external ontologies should also be used to annotate resources.
+Note that all of these annotations are valid both for the screenflow-independent component definitions (templates) and the design time instances (copies).
+A detailed description of all terms from the FAST gadget ontology namespace can be found in Sect.~\ref{sec:ontology_terms} of this document.
+
+\singlespacing
+\lstset{
+	caption={Example of basic annotation of a FAST resource}, 
+	label=list:basic_annotation,
+	language=turtle
+}
+\begin{figure}[ht]
+	\lstinputlisting{exampleCode/screen_example.ttl}
+\end{figure}
+\doublespacing
+
+% subsection basic_annotation (end)
 
@@
 
-\subsection{Classes} % (fold)
-\label{sub:classes}
-
-This section contains a definition of all classes defined in the \texttt{fgo} namespace. For classes integrated from other ontologies, see the integration document~\ref{ssub:integration_document}.
-
-\singlespacing
-\begin{small}
-\input{classes}
-\end{small}
-\doublespacing
-
-%subsection classes (end)
-
-\subsection{Properties} % (fold)
-\label{sub:properties}
-
-This section contains a definition of all properties defined in the \texttt{fgo} namespace.
-For properties integrated from other ontologies, see the integration document~\ref{ssub:integration_document}.
-
-\singlespacing
-\begin{small}
-\input{properties}
-\end{small}
-\doublespacing
-
-% subsection properties (end)
-
 \subsection{Extension towards Specific Domains} % (fold)
@@
 \section{Ontology Overview}
+\label{sec:ontology_overview}
 % \addcontentsline{toc}{section}{Ontology Overview}
@@
 \phantomsection
+\section{Ontology Terms}
+\label{sec:ontology_terms}
+\subsection{Classes} % (fold)
+\label{sub:classes}
+
+This section contains a definition of all classes defined in the \texttt{fgo} namespace. For classes integrated from other ontologies, see the integration document~\ref{ssub:integration_document}.
+
+\singlespacing
+\begin{small}
+\input{classes}
+\end{small}
+\doublespacing
+
+%subsection classes (end)
+
+\subsection{Properties} % (fold)
+\label{sub:properties}
+
+This section contains a definition of all properties defined in the \texttt{fgo} namespace.
+For properties integrated from other ontologies, see the integration document~\ref{ssub:integration_document}.
+
+\singlespacing
+\begin{small}
+\input{properties}
+\end{small}
+\doublespacing
+
+% subsection properties (end)
+
+\clearpage
+\phantomsection
 \section{Ontology Code}

### 2010-01-28 trunk/deliverables/D2_2_Ontology/fast_d2_2_2_ontology.tex
version for review
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@483 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \usepackage{array}
+\usepackage{booktabs}
 \usepackage{amssymb}
@@
 
+\subsection{FAST Users} % (fold)
+\label{sub:fast_users}
+
+In modelling resources in FAST, the kind of users that are relevant are the users of the FAST tools themselves (i.e., the GVS), as opposed to end users of the gadgets.
+The following List.~\ref{list:user} shows how an individual user is modelled, using terminology from the FOAF and SIOC vocabularies.
+
+\singlespacing
+\lstset{
+	caption={Basic description of a user in FAST}, 
+	label=list:user,
+	language=turtle
+}
+\begin{figure}[ht]
+	\lstinputlisting{exampleCode/user_example_01.ttl}
+\end{figure}
+\doublespacing
+
+Once a user has been defined in this way, they can e.g.
+be used to specify the creators of individual screen and screenflow resources, as shown in List.~\ref{list:creator}.
+This kind of creation metadata is being added to other metadata as shown in Sect.~\ref{sub:basic_annotation}.
+
+\singlespacing
+\lstset{
+	caption={Modelling the creator of a resource}, 
+	label=list:creator,
+	language=turtle
+}
+\begin{figure}[ht]
+	\lstinputlisting{exampleCode/user_example_02.ttl}
+\end{figure}
+\doublespacing
+
+% subsection fast_users (end)
+
 \subsection{Defining Pre- and Post-conditions} % (fold)
@@
 
-\begin{table}[h]
+\begin{table}[ht]
 	\caption{Different semantics for \texttt{isPositive}}
@@
 	\begin{center}
-\begin{tabular}{|l|c|c|}
-\hline
-& pre-condition & post-condition \\ \hline
-\emph{isPositive ``YES''} & condition must be fulfilled & condition instantiated\\ \hline
-\emph{isPositive ``NO''} & $\neg$condition must be fulfilled & instantiated condition removed \\ \hline
+
+		\begin{tabular}{lcc}
+			\toprule
+			& \textbf{pre-condition} & \textbf{post-condition} \\ 
+			\midrule
+			\emph{isPositive ``YES''} & condition must be fulfilled & condition instantiated\\ 
+			\emph{isPositive ``NO''} & $\neg$condition must be fulfilled & instantiated condition removed \\ 
+			\bottomrule
 		\end{tabular}
@@
 
-\subsection{Instance Examples} % (fold)
-\label{sub:instance_examples}
-
-\subsubsection{Complete Screen with Opaque Code} % (fold)
-\label{ssub:complete_screen}
-
-\singlespacing
-\lstset{
-	caption={Complete Screen with Opaque Code}, 
-	label=list:screen_01
-}
-% \setlength\parindent{0in}
-% \begin{minipage}[t]{\textwidth}
-\lstinputlisting{exampleCode/01_screenExample.ttl}
-% \end{minipage}
-% \setlength\parindent{0.21in}
-\doublespacing
-
-% subsubsection complete_screen (end)
-
-\subsubsection{Complete Screen with Declarative Definition} % (fold)
-\label{ssub:complete_screen_with_declarative_definition}
-
-\singlespacing
-\lstset{
-	caption={Complete Screen with Declarative Definition}, 
-	label=list:screen_01
-}
-% \setlength\parindent{0in}
-% \begin{minipage}[t]{\textwidth}
-\lstinputlisting{exampleCode/02_screenExample.ttl}
-% \end{minipage}
-% \setlength\parindent{0.21in}
-\doublespacing
-% subsubsection complete_screen_with_declarative_definition (end)
-
-% subsection instance_examples (end)
+% 
+% \subsection{Instance Examples} % (fold)
+% \label{sub:instance_examples}
+% 
+% \subsubsection{Complete Screen with Opaque Code} % (fold)
+% \label{ssub:complete_screen}
+% 
+% \singlespacing
+% \lstset{
+% 	caption={Complete Screen with Opaque Code}, 
+% 	label=list:screen_01
+% }
+% % \setlength\parindent{0in}
+% % \begin{minipage}[t]{\textwidth}
+% \lstinputlisting{exampleCode/01_screenExample.ttl}
+% % \end{minipage}
+% % \setlength\parindent{0.21in}
+% \doublespacing
+% 
+% % subsubsection complete_screen (end)
+% 
+% \subsubsection{Complete Screen with Declarative Definition} % (fold)
+% \label{ssub:complete_screen_with_declarative_definition}
+% 
+% \singlespacing
+% \lstset{
+% 	caption={Complete Screen with Declarative Definition}, 
+% 	label=list:screen_02
+% }
+% % \setlength\parindent{0in}
+% % \begin{minipage}[t]{\textwidth}
+% \lstinputlisting{exampleCode/02_screenExample.ttl}
+% % \end{minipage}
+% % \setlength\parindent{0.21in}
+% \doublespacing
+% % subsubsection complete_screen_with_declarative_definition (end)
+% 
+% % subsection instance_examples (end)
 
@@
 
-A solution for representing pre- and post-conditions of screens and screen flows --- a central aspect of FAST technology --- has been proposed, which will now have to be implemented and tested both within the FAST catalogue and the FAST GVS.
+A solution for representing pre- and post-conditions of screens and screen flows --- a central aspect of FAST technology --- has been presented, as it has been implemented and tested both within the FAST catalogue and the FAST GVS.
 
@@
 
-Finally, to contribute to a lasting legacy of the FAST project, the gadget ontology will serve as input to standardisation efforts for the definition of semantic gadgets, which will go beyond the scope of the project itself.
+% Finally, to contribute to a lasting legacy of the FAST project, the gadget ontology will serve as input to standardisation efforts for the definition of semantic gadgets, which will go beyond the scope of the project itself.
 % section conclusions (end)
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \begin{center}

### 2010-02-01 trunk/GVS/media/js/gvs/documents/ScreenCanvasCache.js
Screen loading finished
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@489 d8b0c59b-6544-0410-94d9-83a5658a058e

@@ -112,9 +112,26 @@ var ScreenCanvasCache = Class.create( /** @lends ScreenCanvasCache.prototype */ 
 
     /**
      * Returns the list of postconditions
+     * @type Array
      */
     getPostconditions: function() {
         return this._postconditions;
+    },
+
+    /**
+     * Returns the list of pipes
+     * @type Array
+     */
+    getPipes: function() {
+        return this._pipes;
+    },
+
+    /**
+     * Returns the list of triggers
+     * @type Array
+     */
+    getTriggers: function() {
+        return this._triggers;
     },
 
     /**

### 2010-02-01 trunk/GVS/media/js/gvs/documents/ScreenDocument.js

@@ -833,7 +833,7 @@ var ScreenDocument = Class.create(PaletteDocument,
       */
      _findInstance: function(/** String */ instanceId) {
         return this._canvasInstances.values().detect(function(element) {
-                    return element.getId() == instanceId
+                    return element.getId() == instanceId;
          });
      },
 
@@ -1019,7 +1019,11 @@ var ScreenDocument = Class.create(PaletteDocument,
         }.bind(this));
     },
 
-    _createConditions: function(/** Array */ conditionList, /** DropZone */ area) {
+    /**
+     * Create the pre and post conditions
+     * @private
+     */
+     _createConditions: function(/** Array */ conditionList, /** DropZone */ area) {
         conditionList.each(function(condition) {
             var description = new BuildingBlockDescription(condition);
             var instance = new PrePostInstance(description, this._inferenceEngine, false);
@@ -1035,6 +1039,67 @@ var ScreenDocument = Class.create(PaletteDocument,
     },
 
     /**
+     * Creates the pipes programmatically
+     * @private
+     */
+    _createPipes: function(/** Array */ pipes) {
+        pipes.each(function(pipeData) {
+            var fromInstance;
+            var fromTerminal;
+            if (pipeData.from.buildingblock != "") {
+                fromInstance = this._findInstance(pipeData.from.buildingblock);
+                fromTerminal = fromInstance.getTerminal(pipeData.from.condition);
+            } else {
+                fromInstance = this._description.getPre(pipeData.from.condition);
+                if (!fromInstance) {
+                    fromInstance = this._description.getPost(pipeData.from.condition);
+                }
+                fromTerminal = fromInstance.getTerminal();
+            }
+            var toInstance;
+            var toTerminal;
+            if (pipeData.to.buildingblock != "") {
+                toInstance = this._findInstance(pipeData.to.buildingblock);
+                toTerminal = toInstance.getTerminal(pipeData.to.condition);
+            } else {
+                toInstance = this._description.getPre(pipeData.to.condition);
+                if (!toInstance) {
+                    toInstance = this._description.getPost(pipeData.to.condition);
+                }
+                toTerminal = toInstance.getTerminal();
+            }
+            var pipe = this._pipeFactory.getPipe(fromTerminal, toTerminal);
+            this._description.addPipe(pipe);
+        }.bind(this));
+    },
+
+    /**
+     * Creates the triggers
+     * @private
+     */
+    _createTriggers: function(/** Array */ triggers) {
[…]

### 2010-02-01 trunk/GVS/media/js/gvs/instances/OperatorInstance.js

@@ -13,10 +13,10 @@ var OperatorInstance = Class.create(ComponentInstance,
         
         /**
          * Terminals for screen design
-         * @type Array
+         * @type Hash
          * @private 
          */
-        this._terminals = new Array();
+        this._terminals = new Hash();
         
     },
 
@@ -30,20 +30,7 @@ var OperatorInstance = Class.create(ComponentInstance,
     getTitle: function() {
         return this._buildingBlockDescription.label['en-gb']; 
     },
-    
-    /**
-     * Transform the instance into JSON-like
-     * string
-     * @type String
-     */
-    toJSON: function() {
-        var json = {
-           
-        }
-        return Object.toJSON(json);
-    },
-    
-    
+     
     
     /**
      * @override
@@ -74,7 +61,7 @@ var OperatorInstance = Class.create(ComponentInstance,
                 }     
                 var node = this._view.getConditionNode(pre.id);
                 var terminal = new Terminal(node, options, this, pre.id, action.name);
-                this._terminals.push(terminal);                
+                this._terminals.set(pre.id, terminal);
             }.bind(this));   
         }.bind(this));
         
@@ -99,8 +86,16 @@ var OperatorInstance = Class.create(ComponentInstance,
                 var node = this._view.getConditionNode(post.id);
                 var terminal = new Terminal(node, options, this, post.id);
                 terminal.onWireHandler(handler);
-                this._terminals.push(terminal);
+                this._terminals.set(post.id, terminal);
             }.bind(this)); 
+    },
+
+    /**
+     * Gets a terminal from an id
+     * @type Terminal
+     */
+    getTerminal: function(/** String */ id) {
+        return this._terminals.get(id);
     },
 
     /**
@@ -123,7 +118,7 @@ var OperatorInstance = Class.create(ComponentInstance,
     destroy: function($super) {
         $super();
         if (this._terminals) {
-            this._terminals.each(function(terminal){
+            this._terminals.values().each(function(terminal){
                 terminal.destroy();    
             });
         }
@@ -145,7 +140,7 @@ var OperatorInstance = Class.create(ComponentInstance,
      */
     onUpdate: function(/** Number */ x, /** Number */ y) {
         if (this._terminals) {
-            this._terminals.each(function(terminal){
+            this._terminals.values().each(function(terminal){
                 terminal.updatePosition();    
             });
         }

### 2010-02-01 trunk/GVS/media/js/gvs/piping/Pipe.js

@@ -107,6 +107,7 @@ var Pipe = Class.create(
             });
         }
         this._wire.setOptions(options);
+        this._wire.redraw();
     },
 
     /**

### 2010-02-01 trunk/GVS/media/js/gvs/GVS.js
Incorpored new features to the "demo" version
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@490 d8b0c59b-6544-0410-94d9-83a5658a058e

@@ -241,6 +241,18 @@ var GVSSingleton = function(){
                                         'shortcut': 'Shift+N'
                                     }),
                                     'group': 0
+                                },
+                                'newScreen': {
+                                    'type': 'Action',
+                                    'action': new MenuAction({
+                                        'label': 'Screen',
+                                        'weight': 10,
+                                        'handler': function(){
+                                            this.action("newScreen")
+                                        }.bind(this),
+                                        'shortcut': 'Alt+N'
+                                    }),
+                                    'group': 0
                                 }
                             }
                         },
@@ -259,6 +271,18 @@ var GVSSingleton = function(){
                                             this.action("browseScreenflows");
                                         }.bind(this),
                                         'shortcut': 'Shift+O'
+                                    }),
+                                    'group': 0
+                                },
+                                'browseScreens': {
+                                    'type': 'Action',
+                                    'action': new MenuAction({
+                                        'label': 'Screens',
+                                        'weight': 10,
+                                        'handler': function(){
+                                            this.action("browseScreens")
+                                        }.bind(this),
+                                        'shortcut': 'Alt+N'
                                     }),
                                     'group': 0
                                 }
@@ -318,30 +342,6 @@ var GVSSingleton = function(){
             };
             if (!GlobalOptions.isPublicDemo) {
                 // Include the new screen feature
-                this._menuConfig.file.children['new'].children.newScreen = {
-                    'type': 'Action',
-                    'action': new MenuAction({
-                        'label': 'Screen',
-                        'weight': 10,
-                        'handler': function(){
-                            this.action("newScreen")
-                        }.bind(this),
-                        'shortcut': 'Alt+N'
-                    }),
-                    'group': 0
-                };
-                this._menuConfig.file.children['browse'].children.browseScreens = {
-                    'type': 'Action',
-                    'action': new MenuAction({
-                        'label': 'Screens',
-                        'weight': 10,
-                        'handler': function(){
-                            this.action("browseScreens")
-                        }.bind(this),
-                        'shortcut': 'Alt+N'
-                    }),
-                    'group': 0
-                };
             }
         }
     });

### 2010-02-01 trunk/GVS/media/js/gvs/documents/WelcomeDocument.js

@@ -53,7 +53,7 @@ var WelcomeDocument = Class.create(AbstractDocument,
             label: "Create a Screen from Scratch",
             onClick: function() { gvs.action("newScreen"); }}
         );
-        var openScreenButton = new dijit.form.Button({
+        var browseScreensButton = new dijit.form.Button({
             label: "Browse Screens...",
             onClick: function() { gvs.action("browseScreens"); }}
         );
@@ -63,11 +63,12 @@ var WelcomeDocument = Class.create(AbstractDocument,
         buttonsContainer.appendChild(newScreenflowButton.domNode);
         buttonsContainer.appendChild(new Element("br"));
         buttonsContainer.appendChild(openScreenflowButton.domNode);
+        buttonsContainer.appendChild(new Element("div").update("Screens"));
+        buttonsContainer.appendChild(newScreenButton.domNode);
+        buttonsContainer.appendChild(new Element("br"));
+        buttonsContainer.appendChild(browseScreensButton.domNode);
         if (!GlobalOptions.isPublicDemo) {
-            buttonsContainer.appendChild(new Element("div").update("Screens"));
-            buttonsContainer.appendChild(newScreenButton.domNode);
-            buttonsContainer.appendChild(new Element("br"));
-            buttonsContainer.appendChild(openScreenButton.domNode);
+            
         }
         content.appendChild(buttonsContainer);
     }

### 2010-02-01 trunk/deliverables/D4_3_2/fast_d4_3_2.tex
Fixes error with a URL
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@444 d8b0c59b-6544-0410-94d9-83a5658a058e
Added screen dumps and some editing of the corresponding text. Fine for me now. Albert
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@491 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 Therefore, these precondition ports might be used to determine values to parameters like query keywords coming from a building block inside the screen the service is placed or even an external screen as a screen precondition.
+
+To define the preconditions and postconditions of a new service wrapper the FAST service wrapper tool provides a form that allows the editing of these entries, cf.
+Figure \ref{fig:construct_pre_post_conditions}\footnote{This screen dump is a combination of two different FAST tool screen dumps that show how the final tool will look like.
+We add a real screen dump ASAP.}.
+Note, we allow to edit example values for the preconditions.
+These example values may be used to test the service wrapper.
+
+\begin{figure}
+  \begin{center} \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithPortDefinitions.jpg}
+    \caption{Configuring pre- and postconditions of a service wrapper (photoshoped screen dump)}
+    \label{fig:construct_pre_post_conditions}
+  \end{center}
+\end{figure}
+
 The service wrapper tool composes the request using a template string which will contain placeholders for precondition values.
@@
 
-Figure ~\ref{fig:complex_gadget_architecture} shows the screen to construct the service requests.
+Figure ~\ref{fig:construct_service_request} shows the screen to construct the service requests.
 
@@
   \begin{center}
-    %\includegraphics[width=\linewidth]{images/ComplexGadgetArchitecture.pdf}
-    \caption{...}
+    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithRequestExample.jpg}
+    \caption{Constructing the service request URL and its parameters (photoshoped screen dump)}
     \label{fig:construct_service_request}
@@
 
-In the upper part of Figure ~\ref{fig:complex_gadget_architecture}, the user establishes the name of the Resource Adapter and which parameters or input ports the wrapper is composed of.
-For each input port, the user provides a name, a type, and an example value.
-The example value is just to simulate the behaviour of the wrapper.
-While constructing the request template, these placeholders have to be surrounded by <>.
-As you see in the above example, the request contains <query\_keywords> in the position its value needs to be placed.
+In the top input field of Figure ~\ref{fig:construct_service_request}, the user may drop an example http request taken e.g.
+form the service documentation e.g.
+from \cite{eBayShoppingAPIs}.
+The tool analyses the example request and in the middle of the screen a form for editing the request parameters is provided.
+In our example, the user has connected the \textit{QueryKeywords} parameter with the precondition \textit{search\_key} by adding a corresponding reference to the value field of that parameter.
+In addition, we have retrieved an access key which has been entered as value for the \textit{appid} parameter.
+
+Below the request paramter editing form of Figure ~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed service by sending it to the specified service address (via a server relay).
+Then the placeholders for preconditions are replaced with example values and the resulting http request is shown below the parameter form.
+In addition, the request is send and the response is shown on the bottom of that page.
+This gives the user a fast feedback whether the constructed request works as desired.
 
@@
 However, every building block accepts any (complex) concept as a pre/postcondition, so the wrapper tool needs to be adapted to permit any concept as input or precondition, extracting the required text to construct the request.
+This is planned for the third year.
 
@@
 This response message could be serialize in any format, though the most common formats used nowadays are XML or JSON among others.
-To continue the example it was started in the previous section, the response of the Ebay Shopping service will be in XML as specified in the request, which is the format supported by the wrapper tool.
+To continue the example started in the previous section, the response of the Ebay Shopping service will be in XML as specified in the request, which is the format supported by the wrapper tool.
 
@@
 
-Figure ~\ref{fig:response_service_execution} shows the response obtained after the execution of the service request.
+Figure ~\ref{fig:response_service_execution} shows the data tranformation tab of the wrapper tool.
 
@@
   \begin{center}
-    %\includegraphics[width=\linewidth]{images/ComplexGadgetArchitecture.pdf}
-    \caption{...}
+      \includegraphics[angle=90,width=0.8\linewidth]{images/ServiceWrapperToolGVSWithTransformationRules.jpg}
+    \caption{Interactive, rule based transforming of an XML response to FAST facts (photoshoped screen dump)}
     \label{fig:response_service_execution}
@@
 
-Once the service response, in XML format, has been retrieved, it is displayed both in plain text (debugging/test purposes) and as an interactive object.
-To construct this interactive object, the XML document has to been parsed into a DOM, and a simplified tree representation of that DOM (see lower left part in Figure ~\ref{fig:response_service_execution}) is built up.
-This tree representation of the XML data is will be used as an input to construct transformation rules.
+Once the service response, in XML format, has been retrieved, the transformation tab shows it as an interactive object tree on the left side of Figure ~\ref{fig:response_service_execution}.
+To construct this interactive object, the XML document has been parsed into a DOM, and a simplified tree representation of that DOM is built up.
+This tree representation of the XML data is used as an input to construct transformation rules.
 
 A transformation rule is used to analyse the XML data and to generate domain-specific facts from concepts from the ontologies used by the pre/postconditions of the different building blocks.
-A transformation rule is composed of three elements.
-First, from indicates the origin of the rule, which will be the tagname of a DOM element from the XML document.
+A transformation rule is composed of three elements, cf the middle part of Figure ~\ref{fig:response_service_execution}.
+First, the \textit{from} field indicates the XML elements to be translated by rule.
+These XML elementsare identified by the tagname of a DOM element from the XML document.
 Second, the type of the rule will be set, taking one of the following values: \emph{createObject}, \emph{fillAttributes} or \emph{dummy}.
@@
 The type of that new fact is provided in the third compartment.
-In the example is being explained, the root rule searches for XML elements with tagname \emph{FindItemsResponse} and for each such element a \emph{List} fact is created.
-The resulting facts are shown in a facts tree in the lower right of Figure ~\ref{fig:response_service_execution}.
+In the example being explained, the root rule searches for XML elements with tagname \emph{FindItemsResponse} and for each such element a \emph{List} fact is created.
+The resulting facts are shown in a facts tree in the right of Figure ~\ref{fig:response_service_execution}.
 	\item[action \emph{fillAttributes}] does not create a new fact but it fills the value of the attribute provided as third part of such rules.
@@
 The value of that \emph{Title} element is then transfered to the \emph{productName} attribute of the corresponding \emph{Product} fact.
+Actually, our \textit{from} fields allows also to refer to parts of an XML attribute e.g.
+to \textit{words} 1 through 3.
+It is also possible to combine constant text and elements of multiple XML tree elements.
 	\item[action \emph{dummy}] does not create or modify any facts but such rules are just used to narrow the search space for their sub-rules.
@@
 
-Since FAST is storyboard oriented, the service wrapper designer follows the storyboard idea as well.
+Since FAST is storyboard oriented, the service wrapper tool follows the storyboard idea as well.
 Any time, a change to a transformation rule is done, the transformation process is triggered and the resulted facts tree is directly shown.
@@
 Still, the selected approach suffices for most practical and real world cases.
-However, there are some transformations that these rules cannot perform.
-For example, the XML data for a person may provide two different tags for the first and the last name of a person.
+As a more complex example, the XML data for a person may provide two different tags for the first and the last name of a person.
 Contrarily, a person fact which conforms to a certain ontology for that domain may provide only one \emph{fullname} attribute that shall be filled by a concatenation of the first and the last name.
-To cover such cases directly requires rather complex extensions of our transformation rules.
-To avoid this, intermediate fact formats can be used which would allow generating person facts with separated first and last name attributes.
-Then, additional filters may be used allowing the combination of the first and last name in a single \emph{fullname} attribute.
-However, this can be the role of the operators, since some generic operators can act as filters for aggregation and conversions of facts from multiple sources.
+To achieve this, the \textit{from} field of that tranforamtion rule might look like: \texttt{lastname"', "`firstname}.
+We are also able to do some navigation in the XML tree to follow XRef elements.
+For example the attribute \texttt{grandmother} could be filled using \texttt{mother.mother} in the \textit{from} field.
+
+However, we there are some transformations that these rules cannot perform.
+For example, we do not support any mathematical operations.
+Thus, transforming e.g.
+Fahrenheit into Celsius temperatures is not supported.
+To cover such  cases, intermediate fact formats can be used which would allow generating facts to be further processed by additional filters.
+Such additional filters may be realized using (hand coded) operators, since some generic operators can act as filters for aggregation and conversions of facts from multiple sources.
 Then, service wrappers in combination with these filter operators will allow covering these complex cases.
@@
 Wrapping WSDL-Described Web Services as Moby Services (using SAWSDL)
-http://biomoby.open-bio.org/CVS_CONTENT/moby-live/Java/docs/sawsdlServlet.html
+\url{http://biomoby.open-bio.org/CVS_CONTENT/moby-live/Java/docs/sawsdlServlet.html}
 

### 2010-02-01 trunk/GVS/media/js/gvs/dialogs/FormDialog.js
Controlling not to open external screens with the GVS
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@492 d8b0c59b-6544-0410-94d9-83a5658a058e

@@ -104,6 +104,7 @@ var FormDialog = Class.create( /** @lends FormDialog.prototype */ {
 
     /**
      * This function adds a button with an onclick handler
+     * @type dijit.form.Button
      * @private
      */
     _addButton: function (/** String */ label, /** Function */ handler) {
@@ -114,6 +115,7 @@ var FormDialog = Class.create( /** @lends FormDialog.prototype */ {
 
         
         this._buttonNode.appendChild(button.domNode);
+        return button;
     },
 
     /**

### 2010-02-01 trunk/GVS/media/js/gvs/dialogs/GalleryDialog.js

@@ -30,7 +30,9 @@ var GalleryDialog = Class.create(FormDialog, /** @lends GalleryDialog.prototype 
         // Assigning the passed parameters, or defaults
         this._properties.set('showTitleRow', (properties.showTitleRow || false));
         this._properties.set('elementsPerPage', (properties.elementsPerPage || 10));
-        this._properties.set('onDblClick', (properties.onDblClick || null));
+        this._properties.set('onDblClick', (properties.onDblClick || function(){}));
+        // Disable all the events if the selected row is not valid
+        this._properties.set('disableIfNotValid', (properties.disableIfNotValid || false));
 
         /**
          * Table fields
@@ -45,6 +47,14 @@ var GalleryDialog = Class.create(FormDialog, /** @lends GalleryDialog.prototype 
          * @private
          */
         this._buttons = null;
+
+        /**
+         * List of buttons that will be disabled if the selected row
+         * is not valid
+         * @type Array
+         * @private
+         */
+        this._disabledButtons = new Array();
         
         /**
          * List of rows
@@ -146,22 +156,38 @@ var GalleryDialog = Class.create(FormDialog, /** @lends GalleryDialog.prototype 
 
                 var element = event.findElement(".row");
                 element.addClassName("selected");
+
+                if (row.isValid) {
+                    this._disabledButtons.each(function(button) {
+                        button.attr('disabled', false);
+                    });
+                } else {
+                    this._disabledButtons.each(function(button) {
+                        button.attr('disabled', true);
+                    });
+                }
                 
                 this._selectedRow = row;
             }.bind(this));
             if (this._properties.get('onDblClick')) {
-                rowNode.observe('dblclick', function() {
-                    this._properties.get('onDblClick')(row.key);
-                }.bind(this));
+                if (!this._properties.get('disableIfNotValid') ||
+                    row.isValid) {
+                        rowNode.observe('dblclick', function() {
+                            this._properties.get('onDblClick')(row.key);
+                        }.bind(this));
+                }
             }
             content.appendChild(rowNode);
         }.bind(this));
         if (loadAll && this._rows.size() > 0) {
             this._removeButtons();
             this._buttons.each(function(button){
-                this._addButton(button.value, function() {
+                var buttonWidget = this._addButton(button.value, function() {
                     button.handler(this._selectedRow.key);
                 }.bind(this));
+                if (button.disableIfNotValid) {
[…]

### 2010-02-01 trunk/GVS/media/js/gvs/dialogs/ManageScreensDialog.js

@@ -6,7 +6,8 @@ var ManageScreensDialog = Class.create(GalleryDialog /** @lends ManageScreensDia
      * @extends GalleryDialog
      */ 
     initialize: function($super) {  
-        $super("Screen browser", {'onDblClick': this._openScreen.bind(this)});
+        $super("Screen browser", {'onDblClick': this._openScreen.bind(this),
+                                  'disableIfNotValid': true });
 
         /**
          * List of screens
@@ -66,7 +67,8 @@ var ManageScreensDialog = Class.create(GalleryDialog /** @lends ManageScreensDia
 
         this._setButtons([{
                 'value': 'Open screen',
-                'handler': this._openScreen.bind(this)
+                'handler': this._openScreen.bind(this),
+                'disableIfNotValid': true
             }, {
                 'value': 'Share/Unshare screen',
                 'handler': this._shareScreen.bind(this)
@@ -105,7 +107,8 @@ var ManageScreensDialog = Class.create(GalleryDialog /** @lends ManageScreensDia
                                     data.description['en-gb'],
                                 '<span class=' + (screen.uri ? '"shared"': '"unshared"') +
                                     '>&nbsp;</span>'
-                             ]
+                             ],
+                             'isValid': data.definition ? true : false
                         });
         }.bind(this));
     },

### 2010-02-01 trunk/GVS/media/js/gvs/documents/ScreenDocument.js

@@ -1111,6 +1111,7 @@ var ScreenDocument = Class.create(PaletteDocument,
             this._createConditions(this._canvasCache.getPostconditions(), this._areas.get('post'));
             this._createPipes(this._canvasCache.getPipes());
             this._createTriggers(this._canvasCache.getTriggers());
+            this._refreshReachability();
         }
     }
 });

### 2010-02-01 trunk/deliverables/D4_3_2/fast_d4_3_2.tex
Added some related work.
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@493 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
+The rule based approach of the service wrapper tool has been inspired by triple graph grammars, cf.
+\cite{conf/wg/Schurr94,JSZ97c}.
+While general triple graph grammars allow to relate complex graph structures with each other, in our case the target of a rule is always a single object or attribute.
+This facilitates the whole mechanism, reasonably.
+
 Write about some related work such as:

### 2010-02-01 trunk/deliverables/D4_3_2/fast_d4_3_2.tex
adds more content and integrates Albert's contribution
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@494 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
+% This is useful for switching in-line comments on and off!
+\usepackage{ifthen}
+\newboolean{showcomments}
+\setboolean{showcomments}{true}
+\ifthenelse{\boolean{showcomments}}
+{
+	\definecolor{Red}{rgb}{1,0,0}
+	\newcommand{\note}[1]{\begin{scriptsize}{{\textsf{\textbf{{\textcolor{Red}{[[#1]]}}}}}}\end{scriptsize}}
+	\definecolor{Orange}{rgb}{1,0.5,0}
+	\newcommand{\todo}[1]{\textsf{\textbf{\textcolor{Orange}{[[TODO: #1]]}}}}
+    \definecolor{Green}{rgb}{0,1,0}
+    \newcommand{\commentreply}[1]{\begin{scriptsize}{\textsf{\textbf{\textcolor{Green}{[[#1]]}}}}\end{scriptsize}}
+} {
+	\newcommand{\note}[1]{\textbf{}}
+	\newcommand{\todo}[1]{}
+    \newcommand{\commentreply}[1]{}
+}
+
 \newcommand\deliverableNumber{D4.3.2}
@@
 \newcommand\authorOne{Ismael Rivera, NUIG}
-\newcommand\authorTwo{Author 2}
+\newcommand\authorTwo{Albert Z\"undorf, UniKassel}
 \newcommand\authorThree{Author 3}
@@
 		{\large \sffamily \authorOne}\\
-%		{\large \sffamily \authorTwo}\\
+		{\large \sffamily \authorTwo}\\
 		\vspace{0.5cm}
@@
 							\textcolor{white}{\textbf{Change description}}\\ \hline
-%0.1 & 15.12.2008 & Knud M\"oller (NUIG) & template modelled \\ \hline
-%0.2 & 15.02.2009 & Knud M\"oller (NUIG) & content moved from wiki after internal review \\ \hline
 1.0 & 26.02.2010 & Ismael Rivera (NUIG) & Final version ready for external review \\ \hline
@@
 These back-end services cannot be directly used; hence they need to be encapsulated in what in the FAST platform is called Resource Adapters.
-The application of semantics to the back-end, through the corresponding Resource Adapters, and front-end building blocks assures a powerful instrument in the task of building new gadgets, improving the search, and enhancing the connection among the different building blocks which compose a gadget.
+The application of semantics to the back-end, through the corresponding resource adapters, and front-end building blocks assures a powerful instrument in the task of building new gadgets, improving the search, and enhancing the connection among the different building blocks which compose a gadget.
 
@@
 	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
-	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D4.3.2 is...}\\ \hline
+	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{This deliverable presents techniques to exploit different kind of web services such as RESTful, SOAP-based and WSMO services.
+It exposes how these services are commonly defined, and what FAST need to allow these web services to be use within a gadget.
+In this deliverable is also presented a tool which apply some of these techniques allowing users to create the necessary wrappers for these services.}\\ \hline
 	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
-%	{} & \multicolumn{4}{l|}{}\\ 
+	{} & \multicolumn{4}{l|}{\authorTwo}\\ 
 %	{} & \multicolumn{4}{l|}{}\\ 
@@
 	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {ismael.rivera@deri.org} \\ \cline{4-5}
-	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 ******} \\ \hline
-	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, web services, WSDL, REST, SOAP, WSMO}\\ \hline
+	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495338} \\ \hline
+	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, web service, WSDL, REST, SOAP, WSMO, Resource Adapter}\\ \hline
 \end{tabular}
@@
 This deliverable is structured as follows.
-Section ~\ref{sec:introduction} states the goal, scope and structure of the document.
-Section ~\ref{sec:web_services_wrapping} describes what a service wrapper is, how the platform is able to interact with web services, how they these wrappers are build and then discovered to be reused in any gadget.
-Finally, Section ~\ref{sec:restful_web_services_wrapper_tool} defines the service wrapper tool built to facilitate FAST users to create Resource Adapters of RESTful web services.
+Section~\ref{sec:introduction} states the goal, scope and structure of the document.
+Section~\ref{sec:resource_adapters} describes what a resource adapter is, how the platform is able to interact with web services, how they these wrappers or adapters are build and then discovered to be reused in any gadget.
+Then, Section~\ref{sec:restful_web_services_wrapper_tool} defines the service wrapper tool built to facilitate FAST users to create resource adapters of RESTful web services.
+Finally, Section~\ref{sec:related_work}...\todo{Finish it!}
 
@@
 \clearpage
-\section{Web Services Wrapping} % (fold)
-\label{sec:web_services_wrapping}
+\section{Resource Adapters} % (fold)
+\label{sec:resource_adapters}
 
@@
 
-Therefore, a complex gadget in terms of FAST is aimed to provide a functional access from a graphical user interface to a set of services (SOAP or REST-based Web services) and data sources (Atom/RSS feeds).
+The following sections explain what a resource adapter is, how a user, in this case a resource developer, can build these resource adapters, and how another user, a screen developer for instance, is able to discover and connect them while creating a screen.
+Hence, these three phases are called: building, discovery and connection.
+
+\subsection{Definition} % (fold)
+\label{sub:definition}
+
+A complex gadget in terms of FAST is aimed to provide a functional access from a graphical user interface to a set of services (SOAP or REST-based Web services) and data sources (Atom/RSS feeds).
 These services and data sources need to be modelled and encapsulated within the platform in order to allow the discovery and use of them by other components (i.e.
@@
 This is called Resource Adapter in the FAST platform.
-Figure ~\ref{fig:complex_gadget_architecture} illustrates their position and interaction with the rest of the building blocks within a gadget.
+Figure~\ref{fig:complex_gadget_architecture} illustrates the role they play and which other building blocks they interact with within a gadget.
 
-\begin{figure}
+\begin{figure}[!htb]
   \begin{center}
@@
 
-Once a brief background has been shown, the following sections explain how a user, in this case a resource developer, can build Resource Adapters, and how another user, a screen developer for instance, will discover and connect them while creating a screen.
-Hence, these three phases are called: building, discovery and connection.
+Basically, a resource adapter, from a design point of view, is composed of:
+
+\begin{enumerate}
+	\item \textbf{metadata} such as creator, label, description, version, and so on,
+	\item a set of \textbf{actions} or operations,
+	\item and a set of \textbf{postconditions} or outputs of the resource.
+\end{enumerate}
+
+Figure~\ref{fig:resource_adapter_overview} shows the information above in a graphical manner.
+An action (represented by a white square) is the minimum unit for the functionality a resource may have.
+Imagine a user clicking on a button to get a list of items for sale, or a checkout action performed in order to buy one of the items from the list.
+As an analogy, an action is what a operation is in a SOAP-based web service.
+Moreover, as these operations require some input, the actions may need a certain precondition (green circles) to be satisfied.
+And at last, the execution of these action may cause a postcondition (red circle) be satisfied, or following the same example, an output is returned as result of the execution of an operation.
+
+\begin{figure}[!htb]
+  \begin{center}
+    \includegraphics[width=12cm]{images/resource_adapter.pdf}
+    \caption{Resource Adapter Overview}
+    \label{fig:resource_adapter_overview}
+  \end{center}
+\end{figure}
+
+% subsection definition (end)
 
@@
 
+This section describes different approaches for describing and implementing web services, and how a resource adapter could be generated for these services.
+
 \subsubsection{REST-based Web Services} % (fold)
@@
 
-REST is a term to describe an architecture style, not a standard, of networked systems.
+REST \cite{Fielding2000} is a term to describe an architecture style, not a standard, of networked systems.
 The acronym REST stands for Representational State Transfer.
-REST-based or RESTful web services \cite{Fielding2000} are created identifying all of the conceptual entities or resources which want to be exposed as services.
+REST-based or RESTful web services are created identifying all of the conceptual entities or resources which want to be exposed as services.
 Those entities or resources should be nouns, not verbs (orders, tickets, etc.).
@@
 
-REST is lightweight (not a lot of extra xml markup, human readable results), but unlike SOAP-based web services, which have a standard vocabulary, and commonly used, to describe the web service interface through WSDL, RESTful web services are currently not formally described most of the times.
+RESTful web services are more lightweight than SOAP-based web services, avoiding the verbose XML-based messages, but unlike the former, which have a standard vocabulary, and commonly used, to describe the web service interface through WSDL, RESTful web services are currently not formally described most of the times.
 For a service consumer to understand the context and content of the data that must be sent to and received from the service, both the service consumer and service producer must have an out-of-band agreement.
 This takes the form of documentation, sample code, and an API that the service provider publishes for developers to use.
-For example, the many web-based services available from Google, Yahoo, Flickr, Amazon, and so on have accompanying artefacts describing how to consume the services.
-This style of documenting REST-based web services is fine for use by developers, but it averts tools from programmatically consuming such services and generating artefacts specific to programming languages, as a web service described using WSDL allows.
+For example, the many web-based services available from Google, Yahoo, Flickr, Amazon, and so on have accompanying artifacts describing how to consume the services.
+This style of documenting REST-based web services is fine for use by developers, but it averts tools from programmatically consuming such services and generating executable stubs for a given programming language, as a web service described using WSDL allows.
 Nevertheless, Web Application Description Language (WADL) attempts to resolve some of these issues by providing a means to describe services in terms of schemas, HTTP methods, and the request or response structures exchanged, but this language is not widely adopted yet by RESTful web service developers.
@@
 There is no need of a formal document such as WADL defining the web service, for this reason, building service wrappers for these services involve the correct understanding of the service by a human being and a tool to facilitate this task.
-This tool is explained in detail in Section ~\ref{sec:restful_web_services_wrapper_tool}.
+This tool is explained in detail in Section~\ref{sec:restful_web_services_wrapper_tool}.
 
@@
 
-SOAP-based web Services or "Heavyweight Web Services" use Extensible Markup Language (XML) \cite{XML1.1} messages that follow the Simple Object Access Protocol (SOAP) standard \cite{SOAP1.2} and have been popular with traditional enterprise, usually relying on HTTP for message negotiation and transmission.
+SOAP-based web Services or "Heavyweight Web Services" use Extensible Markup Language (XML) \cite{XML1.1} messages that follow the Simple Object Access Protocol (SOAP) standard \cite{SOAP1.2} and have been popular within traditional enterprise, usually relying on HTTP for message negotiation and transmission.
 In such services, there is often a machine-readable description of the operations offered by the service written in a Web Services Description Language (WSDL) document.
-Hence, the advantage of using WSDL is that it can be programmatically processed.
+The advantage of specifying the service in a using WSDL is that it can be programmatically processed.
 
@@
 From a specific WSDL definition, a set of methods or operations can be easily extracted which encloses a set of inputs and outputs.
-These operations would be transformed into actions, and the inputs and outputs would be used to define Resource Adapter pre/postconditions.
-With the use of this information along with the XML Schema defining the types, a Resource Adapter can be semantically defined.
+These operations would be transformed into \emph{actions}, and the inputs and outputs would be used to define resource adapter \emph{pre/postconditions}.
+But there is a critical problem following this approach.
+The definition of the inputs, outputs, operations and their types are done in a XML Schema document.
+An XML schema describes, in a syntactic manner, the structure of an XML document, but there is no straight-forward relation with the concepts used to describe pre/postconditions in the building blocks.
+In order to illustrate the concepts of WSDL, and the gap between them and semantically-defined building blocks, a shorten version of a real Amazon E-commerce web service will be presented.
+Concretely, the choosen web service performs an item search through the Amazon catalog.
+It expects as input an ItemSearch element, which may contain an actor, an artist or an author, and the item availability.
+The output of the service would be a set of items.
+The WSDL including for this service is given below.
 
-Moreover, there are several frameworks to facilitate the programmatic use of these formal definitions of web services.
-These frameworks shall be use to transform a WSDL definition into an executable Resource Adapter ready to use inside a gadget.
+\singlespacing
+\begin{small}
+\verbatiminput{examples/amazon.wsdl}
+\end{small}
+\doublespacing
+
+As stated before, one of the advantages of describing web services in a formal document is that they can be programmatically consumed.
+In fact, there are many frameworks out there to facilitate such a task, parsing WSDL files, creating stubs to consume the web service and so on.
+However, in order to create a resource adapter from the WSDL defined above, the types defined in the XML Schema need to be translated into concepts from an ontology.
+
+This problem could be addressed in two different ways.
+First one, as a manual process mapping these types and their attributes to concepts and properties, as it was proposed for RESTful services.
+Second approach is taking advantage of SAWSDL, which keeps the process in the programmatically direction.
+
+SAWSDL \cite{sawsdl} stands for Semantic Annotations for WSDL and XML Schema, and it defines how to add semantic annotations to various parts of a WSDL document such as input and output message structures, interfaces and operations.
+These semantic annotations enable the possibility to develop resource adapters from WSDL documents.
+They reference a concept in an ontology or a mapping document, and are either independent of the ontology expression languages and mapping languages.
+
+Basically, there are three main concepts about SAWSDL: \emph{modelReference} to specify the association between a WSDL or XML Schema component and a concept from the ontology, and \emph{liftingSchemaMapping} and \emph{loweringSchemaMapping} to define the mapping mechanisms between semantic data and XML.
+These lowering and lifting schema mappings are not explained in this deliverable, however in Appendix A of the SAWSDL specification \cite{sawsdl} they are described in detail, and illustrated with examples.
+
+To illustrate the usage of these attributes, the following code shows a fragment of the previous example using SAWSDL to semantically annotate the service with concepts from an imaginary ontology defining the Amazon E-commerce web service.
+
+\singlespacing
+\begin{small}
+\verbatiminput{examples/amazon_sawsdl.wsdl}
+\end{small}
+\doublespacing
+
+Hence, SAWSDL enriches the web services descriptions made in WSDL with references to concepts from a semantic model and enables the possibility to create resource adapters from these definitions without a strong user participation in the process.
+If the semantic metadata is missing, a manual annotation process could be considered, where the user would need to relate concepts from an ontology with the WSDL or XML Schema elements, and create the mapping between them manually as well.
 
@@
 The capability of a web service defines its functionality in terms of pre/postconditions, assumptions and effects.
-A web service capability is defined by specifying the following elements: non-functional properties, imported ontologies, used mediators, shared variables, precondition, postcondition, assumption, and effect.
+Apart from that, a web service capability is defined by specifying the following elements: non-functional properties, imported ontologies, used mediators, and shared variables.
+The creation of the resource adapters for these web services will just take into consideration the set of pre/postconditions.
 
-Basically, the web service will offers to a client a postcondition when some conditions are met in the information space (precondition).
-Effects and assumptions can stay out of this study, since they will not be taken into account by any of the building blocks of a gadget.
+In the rest of the section, the Amazon E-commerce web service will be used to build a WSMO web service with the same functionality \cite{wsmo_amazon}.
+In fact, just the small subset of the service, used in previous sections, will be tranformed.
 
-NEED TO BE FINISHED
+First of all, an ontology needs to be extracted from the XML Schema types of the WSDL definition.
+For the shake of simplificity, this example does a one-to-one mapping between the Schema types and the concepts, though a real sceneario may consider an different approach to optimise the modelling.
+
+The concept describing an item search request for the Amazon web service is:
+
+\singlespacing
+\begin{small}
+\begin{verbatim}
+concept itemSearchRequest
+    nonFunctionalProperties
+      dc#description hasValue "A request to search for an item"
+    endNonFunctionalProperties
+    author ofType _string
+    title ofType (1) _string
+    rating ofType ratings#audienceRating
+    minPrice ofType (1) _decimal
+    maxPrice ofType (1) _decimal
+\end{verbatim}
+\end{small}
+\doublespacing
+
+The concept \emph{itemContainer} describes a list of items returned in the item search response, and its code is presented below:
+
+\singlespacing
+\begin{small}
+\begin{verbatim}
+concept itemContainer
+    nonFunctionalProperties
+      dc#description hasValue "Contains a list of items"
+    endNonFunctionalProperties
+    items ofType item
+\end{verbatim}
+\end{small}
+\doublespacing
+
+Next step would be to describe the namespace and non-functional properties declarations for the WSMO Amazon web service, but as it was said, this information is not relevant for the example.
+Hence, the simplified version of the Amazon web service preconditions would look like:
+
+\singlespacing
+\begin{small}
+\begin{verbatim}
+precondition
+  nonFunctionalProperties
+     dc#description hasValue "The Amazon service handles requests 
+     for searching for items by keywords"
+  endNonFunctionalProperties
+  definedBy
+    // A search by author, title, minPrice or maxPrice
+    ?request memberOf am#itemSearchRequest
+\end{verbatim}
+\end{small}
+\doublespacing
+
+It is a fairly simple precondition, which the only condition required is the existence of a \emph{itemSearchRequest} object.
+
+However, postconditions are a bit more difficult to construct.
+They take the form of implication rules, so given a particular precondition, one particular postcondition will take effect.
+To follow the same example, these would be the postconditions for the Amazon search items functionality, including just the criterion author, title, minimum and maximum price as defined in the ontology.
+
+\singlespacing
+\begin{small}
+\begin{verbatim}
+ postcondition
+  nonFunctionalProperties
+    dc#description hasValue "The service returns a list of searched
+      items"
+  endNonFunctionalProperties
+  // The result for a search request by author is a container of products
+  // such that all items in the container have the requested author
+  (?request[
+      author hasValue ?author
+    ] memberOf am#searchItems implies 
+    exists ?container 
+      (?container memberOf am#itemContainer and 
+        forall ?item 
+          (?container[
+              items hasValue ?item
+            ] implies 
+            ?item[
+              author hasValue ?author
+            ]
+          )
+      )
+  ) and
+
+  // The result for a search request by title is a container of products
+  // such that all items in the container have the requested title
+  (?request[
+      title hasValue ?title
+    ] memberOf am#searchItems implies 
+    exists ?container 
+      (?container memberOf am#itemContainer and 
+        forall ?item 
+          (?container[
+              items hasValue ?item
+            ] implies 
+            ?item[
+              title hasValue ?title
+            ]
+          )
+      )
+  ) and
+
+  // The result for a search request by minPrice is a container of products
+  // such that all items in the container have a price >= requested minPrice
+  (?request[
+      minPrice hasValue ?minPrice
+    ] memberOf am#searchItems implies 
+    exists ?container 
+      (?container memberOf am#itemContainer and 
+      forall{?item,?price} 
+        (?container[
+            items hasValue ?item
+          ] and 
+          ?item[
+            price hasValue ?price
+          ] implies 
+          ?price >= ?minPrice
+        )
+      )
+  ) and
+
+  // The result for a search request by maxPrice is a container of products
+  // such that all items in the container have a price <= requested maxPrice
+  (?request[
+      maxPrice hasValue ?maxPrice
+    ] memberOf am#searchItems implies 
+    exists ?container 
+      (?container memberOf am#itemContainer and 
+      forall {?item,?price} 
+        (?container[
+            items hasValue ?item
+          ] and 
+          ?item[
+            price hasValue ?price
+          ] implies 
+          ?price =< ?maxPrice
+        )
+      )
+  ).
+\end{verbatim}
+\end{small}
+\doublespacing
+
+The manner pre/postconditions are defined in WSMO web services are similar to the way they are defined in the Resource Adapters.
+Section 4.4 in \cite{moller2010fast_ontology} gives a detailed explanation on this topic.
+Basically, and using previous examples, the preconditions would be transformed as follows:
+
+\begin{verbatim}
+?request memberOf am#itemSearchRequest => ?request rdf:type am:itemSearchRequest
+\end{verbatim}
+
+And a simple postcondition would be transformed as follows:
+
+\begin{verbatim}
+exists ?container 
+      (?container memberOf am#itemContainer) => ?container rdf:type am:itemContainer
+\end{verbatim}
 
@@
 
-\subsection{Discovering} % (fold)
-\label{sub:discovering}
+\subsection{Discovery} % (fold)
+\label{sub:discovery}
 
-In the building phase, every Resource Adapter is created sharing a common structure for screen components, such as forms, operators and resources.
+In the building phase, every resource adapter is created sharing a common structure for screen components, such as forms, operators and resources.
 Hence, it will have a set of actions which will contain a set of preconditions to satisfy in order to be executed, and after its execution, it may produce any of the conditions of their postconditions.
@@
 
-In these bases, the mechanism is relatively similar to the screen recommender, used to find the best screens to create a screen flow.
+On these basis, the mechanism is relatively similar to the screen recommender, used to find the best screens to create a screen flow.
 Once more, it is based on the pre/postconditions of a set of building blocks.
@@
 % subsection discovery (end)
+
+\subsection{Connection} % (fold)
+\label{sub:connection}
+
+A resource adapter is aimed to perform some action on a web service.
+In order to perform a given action, a precondition may need to be fulfilled, and as a result of the execution a postcondition may be generated.
+These pre/postconditions are intended to be used by pre/postconditions of other building blocks within a gadget.
+This connection must be explicitly defined, because it will indicate the real data-flow between different building blocks.
+In FAST this is commonly called Piping, since a pipe in computer science is just that, a one-way communication channel for interprocess communication; therefore a pipe is a logical connection made between a precondition and a postcondition of different building blocks.
+Figure~\ref{fig:Piping} shows an example of a search form for the Amazon service.
+This is a screen, composed by a form and a resource adapter.
+The form allows the user to introduce a term for the search, producing a postcondition which will be redirect to a resource adapter precondition.
+The execution of the web services produces a list of results, which will be forward to the postcondition of the screen.
+
+\begin{figure}[!htb]
+  \begin{center}
+    \includegraphics[width=12cm]{images/piping.pdf}
+    \caption{Piping}
+    \label{fig:Piping}
+  \end{center}
+\end{figure}
+
+% subsection connection (end)
 
@@
 
-In FAST, the wrapper tool will be in charge of the Resource Adapters' building phase.
-At the moment, just supports RESTful web services.
+In FAST, the wrapper tool will be in charge of the resource adapters' building phase.
+At the moment, it just supports RESTful web services.
 The building phase for this type of web services is done in two steps: a first step will be in charge of the construction of a service request and a second step will analyse the response got from the execution of the service, allowing the extraction of facts contained in that response and mapping them to domain-specific concepts from the ontologies used within FAST by any building block.
@@
 
-To achieve this, service wrappers will be handled as pseudo-screens with preconditions and postconditions (inputs and outputs).
+To achieve this, service wrappers will be handled as any buis with preconditions and postconditions (inputs and outputs).
+Therefore, these precondition ports might be used to determine values to parameters like query keywords coming from a building block inside the screen the service is placed or even an external screen as a screen precondition.
+The service wrapper tool composes the request using a template string which will contain placeholders for precondition values.
+Before sending the request to the service, the placeholders are replaced with their corresponding values and the resulting URL is then ready to be sent the service request.
+
+To achieve this, service wrappers will be handled as any other building block with preconditions and postconditions (inputs and outputs).
 Therefore, these precondition ports might be used to determine values to parameters like query keywords coming from a building block inside the screen the service is placed or even an external screen as a screen precondition.
 
-To define the preconditions and postconditions of a new service wrapper the FAST service wrapper tool provides a form that allows the editing of these entries, cf.
-Figure \ref{fig:construct_pre_post_conditions}\footnote{This screen dump is a combination of two different FAST tool screen dumps that show how the final tool will look like.
-We add a real screen dump ASAP.}.
-Note, we allow to edit example values for the preconditions.
-These example values may be used to test the service wrapper.
+To define the preconditions and postconditions of a new service wrapper, the FAST service wrapper tool provides a form allowing the edition of these entries, cf.
+Figure~\ref{fig:construct_pre_post_conditions}\footnote{Note: These example values may be used to test the service wrapper, but they can be freely edited.}
 
-\begin{figure}
-  \begin{center} \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithPortDefinitions.jpg}
-    \caption{Configuring pre- and postconditions of a service wrapper (photoshoped screen dump)}
+\begin{figure}[!htb]
+  \begin{center}
+    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithRequestExample.jpg}
+    \caption{Configuring pre/postconditions of a service wrapper}
     \label{fig:construct_pre_post_conditions}
@@
 
-Figure ~\ref{fig:construct_service_request} shows the screen to construct the service requests.
+Figure~\ref{fig:construct_service_request} shows the screen to construct the service requests.
+In the top input field of Figure~\ref{fig:construct_service_request}, the user may drop an example HTTP request taken e.g.
+form the service documentation (e.g.
+\cite{eBayShoppingAPIs} for this example).
+The tool analyses the example request and, in the middle of the screen, a form for editing the request parameters is provided.
+This example shows how a user has connected the \textit{QueryKeywords} parameter with the precondition \textit{search\_key} by adding a corresponding reference to the value field of that parameter.
+In addition, an access key has been retrieved and has been entered as value for the \textit{appid} parameter.
 
-\begin{figure}
+\begin{figure}[!htb]
   \begin{center}
     \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithRequestExample.jpg}
-    \caption{Constructing the service request URL and its parameters (photoshoped screen dump)}
+    \caption{Constructing the service request URL and its parameters}
     \label{fig:construct_service_request}
@@
 
-In the top input field of Figure ~\ref{fig:construct_service_request}, the user may drop an example http request taken e.g.
-form the service documentation e.g.
-from \cite{eBayShoppingAPIs}.
-The tool analyses the example request and in the middle of the screen a form for editing the request parameters is provided.
-In our example, the user has connected the \textit{QueryKeywords} parameter with the precondition \textit{search\_key} by adding a corresponding reference to the value field of that parameter.
-In addition, we have retrieved an access key which has been entered as value for the \textit{appid} parameter.
-
-Below the request paramter editing form of Figure ~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed service by sending it to the specified service address (via a server relay).
-Then the placeholders for preconditions are replaced with example values and the resulting http request is shown below the parameter form.
-In addition, the request is send and the response is shown on the bottom of that page.
-This gives the user a fast feedback whether the constructed request works as desired.
-
-\subsubsection{Limitations} % (fold)
-\label{ssub:limitations}
-
-It is worth pointing out that currently the wrapper tool is able to construct input ports for the wrappers using just basic types.
-However, every building block accepts any (complex) concept as a pre/postcondition, so the wrapper tool needs to be adapted to permit any concept as input or precondition, extracting the required text to construct the request.
-This is planned for the third year.
-
-% subsubsection limitations (end)
+Below the request paramter editing form of Figure~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed service by sending it to the specified service address (via a server relay).
+Then the placeholders for preconditions are replaced with example values and the resulting HTTP request is shown below the parameter form.
+In addition, the request is sent and the response is shown on the bottom of that page, which helps the users to figure out quickly whether the constructed request works as desired.
 
@@
 This response message could be serialize in any format, though the most common formats used nowadays are XML or JSON among others.
-To continue the example started in the previous section, the response of the Ebay Shopping service will be in XML as specified in the request, which is the format supported by the wrapper tool.
+To continue the example it was started in the previous section, the response of the Ebay Shopping service will be in XML as specified in the request, which is the format supported by the wrapper tool.
 
-\subsubsection{Translation XML into Facts} % (fold)
-\label{ssub:translation_xml_into_facts}
-
-Figure ~\ref{fig:response_service_execution} shows the data tranformation tab of the wrapper tool.
+Figure~\ref{fig:response_service_execution} shows the transformation data tab of the wrapper tool.
 
@@
     \includegraphics[angle=90,width=0.8\linewidth]{images/ServiceWrapperToolGVSWithTransformationRules.jpg}
-    \caption{Interactive, rule based transforming of an XML response to FAST facts (photoshoped screen dump)}
+    \caption{Interactive rule based transformation of an XML response to FAST facts}
     \label{fig:response_service_execution}
@@
 
-Once the service response, in XML format, has been retrieved, the transformation tab shows it as an interactive object tree on the left side of Figure ~\ref{fig:response_service_execution}.
+Once the service response, in XML format, has been retrieved, the transformation tab shows it as an interactive object tree on the left side of Figure~\ref{fig:response_service_execution}.
 To construct this interactive object, the XML document has been parsed into a DOM, and a simplified tree representation of that DOM is built up.
@@
 A transformation rule is used to analyse the XML data and to generate domain-specific facts from concepts from the ontologies used by the pre/postconditions of the different building blocks.
-A transformation rule is composed of three elements, cf the middle part of Figure ~\ref{fig:response_service_execution}.
+A transformation rule is composed of three elements, cf the middle part of Figure~\ref{fig:response_service_execution}.
 First, the \textit{from} field indicates the XML elements to be translated by rule.
@@
 A detailed explanation of the type of actions to be trigger from the transformation rules is:
+
 \begin{description}
@@
 In the example being explained, the root rule searches for XML elements with tagname \emph{FindItemsResponse} and for each such element a \emph{List} fact is created.
-The resulting facts are shown in a facts tree in the right of Figure ~\ref{fig:response_service_execution}.
+The resulting facts are shown in a facts tree in the right of Figure~\ref{fig:response_service_execution}.
 	\item[action \emph{fillAttributes}] does not create a new fact but it fills the value of the attribute provided as third part of such rules.
@@
 In addition, we may provide some analysis tool, which will help to guarantee that the facts generated by the transformation rules conform to the fact types defined in the corresponding FAST ontology.
-This helps to ensure that the facts generated by the designed pseudo screen will be compatible for precondition ports of subsequent filter steps and or screens.
+This helps to ensure that the facts generated will be compatible for precondition ports of subsequent filter steps.
 
@@
 
-Once the wrapping of a service has been defined and tested in the service wrapper tool, it shall generate an implementation of the desired Resource Adapter in XML, HTML, and JavaScript, ready to be deployed and executed inside a gadget.
+Once the wrapping of a service has been defined and tested in the service wrapper tool, it shall generate an implementation of the desired Resource Adapter in JSON, HTML, and JavaScript, ready to be deployed and executed inside a screen.
 This service wrapper implementation is compliant with the formats required by the Gadget Visual Storyboard Tool (GVS) and it shall be stored inside the FAST catalogue, in order to be found and used by any user.
@@
 The rule based approach of the service wrapper tool has been inspired by triple graph grammars, cf.
-\cite{conf/wg/Schurr94,JSZ97c}.
+\cite{Schurr94,JSZ97c}.
 While general triple graph grammars allow to relate complex graph structures with each other, in our case the target of a rule is always a single object or attribute.
@@
 \url{http://biomoby.open-bio.org/CVS_CONTENT/moby-live/Java/docs/sawsdlServlet.html}
+
+Perhaps, it's worth to mention the service wrappers Yahoo Pipes has, which are a fixed set developed by them, with no possibility for a user to create new ones, as we do.
+(other sites doing the same??)
 

### 2010-02-03 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_part1.tex
Inserted Kassel contribution
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@504 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-\subsection{Service Wrapping Tool}
-\textbf{\textit{Kassel will describe the service wrapping tool they've developed, that creates mapping instances between ontologies. 
-A description of the front-end and the creation of the mapping rules will be given here.}}
+\subsection{Facts Transformation Tool}
+
+For services the Kassel members of the FAST team have developed a service wrapper tool, cf.
+deliverable D4.3.2.
+One part of the service wrapper tool deals with a rule based approach for the transformation of server responses into FAST facts.
+This problem is closely related to the problem of ontology mediation.
+Thus, we plan to use these techniques for the development of a Facts Transformation Tool that allows to create data transformation operators, i.e.
+operators that mediate between two ontologies.
+
+Figure \ref{fig:FactsTransformationToolPortDefinitions} shows a photoshoped preview of the planned data transformation tool.
+The data transformation tool has two tabs.
+The \textit{General} tab allows to define the input and output ports for the desired data transformation operator.
+These ports shall be  typed according to the participating semantic elements to be derived from the FAST catalogue.
+This typing will provide the structure of the input and output data of the desired data transformation operator.
+
+\begin{figure}
+  \begin{center}
+\includegraphics[width=0.8\linewidth]{images/FactsTransformationToolGVSWithPortDefinitions.jpg}
+    \caption{Definition input ports and output ports for data transformation operators (photoshoped screen dump)}
+    \label{fig:FactsTransformationToolPortDefinitions}
+  \end{center}
+\end{figure}
+
+Figure \ref{fig:FactsTransformationToolPortDefinitions} shows the planned editor for transformation rules.
+On the left of Figure \ref{fig:FactsTransformationToolPortDefinitions} there is a tree of example data that is used to validate the transformation rules during editing.
+Thus, the Facts Transformation Tool will execute the transformation rules after any editing in order to give direct feedback on their functionality and to enable the user to validate the achieved transformation.
+The transformation results are shown on the right of Figure \ref{fig:FactsTransformationToolPortDefinitions}.
+
+The middle of Figure \ref{fig:FactsTransformationToolPortDefinitions} shows some possible transformation rules for our example.
+A \textit{createObject} rule may be used to create output fact objects.
+The \textit{Target} field of such a rule specifies the type of the created objects, in our example \textit{List}.
+The top rule fires for each element of the input data with the type specified in the \textit{From} field of the transformation rule.
+Thus, in our example, the input data is searched for objects of type \textit{List} firing for the root of our data.
+
+The second rule of our example creates \textit{Customer} objects.
+This rule is a subrule of the root rule and thus the created objects are added to the \textit{List} object created by the parent rule.
+Similarly, the search for triggering objects of type \textit{Person} is restricted to the subtree below the from element that has triggered the parent rule, i.e.
+to our input \textit{List}.
+
+The third rule is used to fill the attribute \textit{fullName} of the \textit{Customer} objects.
+This rule is an example for a many to one transformation.
+The result is a string concatenation of the \textit{lastName} of the \textit{Person} that has triggered the parent rule plus a text constant plus the \textit{firstName} of the current \textit{Person}.
+The resulting string is transferred to the \textit{fullName} attribute of the corresponding \textit{Customer}.
+
+Note, simple one to one transformation rules may be derived by the ontology mediation approach described above and such rules may be added to the transformation tool, automatically.
+The more complex rules may then be added interactively.
+During editing, the semantics provided by the FAST catalogue may be used to support the user with completion proposals and consistency checking.
+In addition, the direct execution of the rules on example data will help the user to create appropriate rules and to validate their effects.
+
+Note, so far our transformation rules are restricted to simple cases.
+So far we have focused on simple cases allowing a simple transformation approach that does not require to sophisticated skills from our users.
+However, as discussed on the Amazon to GoodRelations example, there exist more complex cases.
+In the third year of the project we will investigate how such more complex cases may be covered, too, and whether it is still possible to keep the rules simple and understandable.
+
+\begin{figure}
+  \begin{center}
+\includegraphics[angle=90,width=0.8\linewidth]{images/FactsTransformationToolGVSWithTransformationRules.jpg}
+    \caption{Interactive, rule based transforming of Person facts (photoshoped screen dump)}
+    \label{fig:FactsTransformationToolTransformationRules}
+  \end{center}
+\end{figure}
 
@@
 
-\textbf{\textit{Kassel will describe the creation of the matching operator in the catalogue and the storage of the mapping rules.}}
+Once the data transformation operator has been defined, it may be added to the FAST catalogue in order to allow its use for the definition of piping configurations of new screens within the FAST CGS tool.
+For this purpose a simple semantics description will be send to FAST catalogue defining the semantics of the operator and the semantics of its input and output port.
+
+Currently, the Service Wrapper Tool uses its own proprietary repository for trees of transformation rules.
+First of all this has technical reasons since the development of the transformation tool is based on an reference architecture and libraries for web based interactive systems provided by Kassel University, cf.
+\cite{AsDH2009CVSM}.
+This repository has the special advantage that it provides support for teams of developers, thus the service wrapper tool allows to merge the contributions of multiple developers.
+However, on the long run, we may store the transformation rules together with the service wrapper and transformation operator definitions in the FAST catalogue.
 
@@
 \label{sec:runtime}
-At runtime, the matching operator's rules are tranformed into JSON data used by the JavaScript gadget.
-Thus, when data come from the two disparate resources, the gadget will automatically transform data instances of the ontology belonging to the first resource, into instances of the ontology belonging to the second resource.
+For deployment, the gadgets implementation is generated in Javascript.
+In addition, a small FAST runtime library (again Javascript) is added to the gadget.
+This runtime library provides notification mechanisms and manages the fact base of the gadget.
+For data transformation we will generate Javascript code that connects itself to the fact base and any time a fact is send to the input port of the operator, it will do the data transformation and send the result to the output port.
+The code generated from the data transformation rules follows a recursive pattern where each rule becomes a Javascript function with parameters for the source and target fact subtrees to be transformed.
+The function body then uses code to search the source tree for sub elements that match the \textit{From} part of the rule and then code for the desired operation is generated.
+For debugging purposes the user may examine the code generated for an operator in the \textit{CodeGen Viewer} tab of the Facts Transformation tool.
 
-\textbf{\textit{Kassel will describe the way the mapping rules are transformed into JSON and how they are used within the JavaScript gadget.}}
+Our code generation for service wrappers and transformation operators uses a template based approach.
+Thus, the \textit{CodeGen Viewer} tab will also allow to examine the templates used for code generation and to adapt the templates in order to meet changing APIs or to meet new requirements.
+Thereby, the code generation is not hard wired into our tool but the code generation may be adapted by (skilled) end users on the fly.

### 2010-02-04 trunk/deliverables/D5_1_2/fast_d5_1_2.tex
added Semantic Web Technologies Evaluation appendix
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@512 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {27/02/2010} & {\textbf{Submitted}} & {27/02/2010}\\ \hline
-	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{Status} & \multicolumn{2}{l|}{final}\\ \hline
+	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
 	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
@@
 
+\clearpage
+\listoftables
+
+\clearpage
+\listoffigures
+
 %%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-The deliverable presents both the external and internal architecture in Section \ref{sec:architecture_overview}, then in Section \ref{sec:restful_catalogue_api} it is detailed the Catalogue API, query formats, interchange formats, error codes and so on.
+The deliverable presents both the external and internal architecture in Section~\ref{sec:architecture_overview}, then in Section~\ref{sec:restful_catalogue_api} it is detailed the Catalogue API, query formats, interchange formats, error codes and so on.
 
@@
 \begin{itemize}
-	\item Section \ref{sec:installation_guide}.
+	\item Section~\ref{sec:installation_guide}.
 Installation Guide describes a set of steps to be followed in order to install and execute the FAST Catalogue.
 Also some configuration instructions are specified.
-	\item Section \ref{sub:content_negotiation}. Content Negotiation states how the FAST Catalogue has adopted several of the content negotiation principles.
-	\item Section \ref{sec:restful_catalogue_api}.
+	\item Section~\ref{sub:content_negotiation}. Content Negotiation states how the FAST Catalogue has adopted several of the content negotiation principles.
+	\item Section~\ref{sec:restful_catalogue_api}.
 RESTful Catalogue API has been intensively extended and modified in order to  reflect all the new functionality and how every request or response is constructed.
@@
 		\item convenience operations for recommendation and checking satisfaction and reachability for form elements, operators and back-end services (see Section ...),
-		\item and, last but not least, support for Screen planning (see Section \ref{sub:planning}).
+		\item and, last but not least, support for Screen planning (see Section~\ref{sub:planning}).
 	\end{itemize}
 	\item Several appendixes have been added with the JSON structure of every building block the Catalogue can deal with.
+	\item Appendix A. Last year an evaluation to choose the best technology for the Catalogue was done. It was not included in any formal deliverable, so this year it has been documented as an appendix.
 \end{itemize}
@@
 
-In addition, there are various optional dependencies which are required if you want to use certain advanced features (see Section \ref{sub:installation_instructions}).
+In addition, there are various optional dependencies which are required if you want to use certain advanced features (see Section~\ref{sub:installation_instructions}).
 
@@
 ant clean
+ant prepare
 ant compile
+\end{verbatim}
+
+Or using a task which gather the previous three tasks and creates WAR file ready to distribute or deploy:
+\begin{verbatim}
 ant dist
-\end{verbatim}
-
-Or using a task which gather the previous three tasks:
-\begin{verbatim}
-ant main
 \end{verbatim}
@@
 
-Now you have successfully built the FAST catalogue, however we highly recommend you to read Section \ref{sub:configuration} and tune your FAST Catalogue instance.
+Now you have successfully built the FAST catalogue, however we highly recommend you to read Section~\ref{sub:configuration} and tune your FAST Catalogue instance.
 
@@
 This section presents a technical overview of the internal catalogue's architecture and which external components will interact with it.
-Basically three layers can be distinguished: Presentation, Business Logic and Persistence layer (see Figure \ref{fig:architecture}).
+Basically three layers can be distinguished: Presentation, Business Logic and Persistence layer (see Figure~\ref{fig:catalogue_architecture}).
 
 \begin{figure}[htb]
-\label{fig:architecture}
+\label{fig:catalogue_architecture}
 \begin{center}
@@
 	\caption{Catalogue's Architecture}
-	\label{fig:CatalogueInternalArchitecture}
 \end{center}
@@
 
+\singlespacing
 \begin{verbatim}
@@
 \end{verbatim}
+\doublespacing
 
@@
 \end{verbatim}
-
+\doublespacing
+
+\singlespacing
 \begin{verbatim}
@@
 \begin{figure}[htb]
-\label{fig:ContentNegotiation}
+\label{fig:content_negotiation}
 \begin{center}
-	\includegraphics[width=10cm]{images/content_negotiation}
+	\includegraphics[width=12cm]{images/content_negotiation}
 	\caption{Content Negotiation}
@@
 
-In summary, the basic idea of content negotiation, as stated in the HTTP 1.1 specification \cite{http1.1}, is to serve the best variant for a resource, taking into account what variants are available, what variants the server may prefer to serve, what the client can accept, and with which preferences: in HTTP, this is done by the client which may send, in its request, Accept headers (Accept, Accept-Language and Accept-Encoding), to communicate its capabilities and preferences in Format, Language and Encoding, respectively.
+In summary, the basic idea of content negotiation, as stated in the HTTP 1.1 specification ~\cite{http1.1}, is to serve the best variant for a resource, taking into account what variants are available, what variants the server may prefer to serve, what the client can accept, and with which preferences: in HTTP, this is done by the client which may send, in its request, Accept headers (Accept, Accept-Language and Accept-Encoding), to communicate its capabilities and preferences in Format, Language and Encoding, respectively.
 
@@
 screens, services), and \emph{member} of the collection, in other words, the building block itself, which access URI is http://catalogueURL/<type>/<id> where the <id> has to be replaced by the identifier of a specific building block.
-The details of the operations and which HTTP verb has to be used can be found in the Table \ref{tab:crud_operations}.
+The details of the operations and which HTTP verb has to be used can be found in the Table~\ref{tab:crud_operations}.
 
@@
 The JSON structure for create any building bock can be found in the appendixes.
-All of them are composed of a common structure which can be found in Section \ref{appendix_b}, and some specific information depending on the type.
-Screen-flows are defined in \ref{appendix_c}, screens in \ref{appendix_d} and forms, operators and back-end services share the same structure detailed in \ref{appendix_e}.
+All of them are composed of a common structure which can be found in Section Appendix B, and some specific information depending on the type.
+Screen-flows are defined in Appendix C, screens in Appendix D and forms, operators and back-end services share the same structure detailed in Appendix E.
 However, a few examples of request and responses are shown in order to clarify how the request are constructed and sent, and how the responses look like.
@@
 \end{verbatim}
+\doublespacing
 
 To obtain all the screens stored in the catalogue a HTTP GET request is sent to the Collection URI and the response may be something like this:
+
+\singlespacing
 \begin{verbatim}
@@
 The results will be all the screens stored in the catalogue which fulfil some of these preconditions, and are tagged with the tags specified in the domain context.
-The request parameters are shown in Table \ref{tab:screen_find_request} and the response parameters are shown in Table \ref{tab:screen_find_response}.
+The request parameters are shown in Table~\ref{tab:screen_find_request} and the response parameters are shown in Table~\ref{tab:screen_find_response}.
 
@@
 The specific URL to access this operation is http://catalogueURL/getmetadata using HTTP POST method.
-Table \ref{tab:getmetadata_request} shows the parameters needed for the invocation and Table \ref{tab:getmetadata_response} details the different parameters may contain a response.
+Table~\ref{tab:getmetadata_request} shows the parameters needed for the invocation and Table~\ref{tab:getmetadata_response} details the different parameters may contain a response.
 
@@
 The following example send two URIs of screens and a URI of a screen-flow to the GetMetadata operation:
+
+\singlespacing
 \begin{verbatim}
@@
 \end{verbatim}
+\doublespacing
 
 The response obtained are two lists, one containing all the metadata regarding to the screen-flows and another one with the information about the screens:
+
+\singlespacing
 \begin{verbatim}
@@
 
-Similarly to the Screen Find \& Check described in Section \ref{ssub:screen_findcheck}, this operation search for screen components (forms, operators and back-end services) through the catalogue in order to satisfy a given request, and it will attach information about satisfaction and reachability as well.
+Similarly to the Screen Find \& Check described in Section~\ref{ssub:screen_findcheck}, this operation search for screen components (forms, operators and back-end services) through the catalogue in order to satisfy a given request, and it will attach information about satisfaction and reachability as well.
 
@@
 Once the user has selected one or several screens to be included in her screen-flow, she needs to find screens which satisfy the preconditions of these screens, in order to create an executable screen-flow.
-The operation Find, detailed in Section \ref{ssub:screen_find}, helps the user to find these screens, however the response is just a set of screens which somehow may help to accomplish her goal, but not in a single and straight-forward step.
+The operation Find, detailed in Section~\ref{ssub:screen_find}, helps the user to find these screens, however the response is just a set of screens which somehow may help to accomplish her goal, but not in a single and straight-forward step.
 
@@
 Request
+\singlespacing
 \begin{verbatim}
@@
 \end{verbatim}
+\doublespacing
 
 Response
+\singlespacing
 \begin{verbatim}
@@
 \clearpage
-\doublespacing
-\section*{Appendix A. Lists of Tables and Figures}
+\section*{Appendix A. Semantic Web Technologies Evaluation}
 \label{appendix_a}
-\addcontentsline{toc}{section}{Appendix A. Lists of Tables and Figures}
-
-\listoftables
-
-\listoffigures
+\addcontentsline{toc}{section}{Appendix A. Semantic Web Technologies Evaluation}
+
+In \cite{urmetzer2010fast_state_of_the_art} were presented different technologies used in the Semantic Web community regarding Web services and catalogues or repositories.
+These are potential technologies to serve as the basis for the Catalogue implementation.
+Said that, an evaluation was done in order to choose the most accurate technology with the most advantages for it.
+Of particular importance in this evaluation is real-time performance, because the GVS (Gadget Visual Storyboard) actually relies on the Catalogue in parts of its user interface.
+
+Basically, two general solutions were considered.
+One was to build the Catalogue, and the internal components or building block of the gadgets, based on a dedicated SWS platform.
+The other was to build a light-weight solution from scratch based on RDF and RDFS.
+
+For both solutions, a few ideas about their potential benefits and disadvantages were in mind beforehand.
+SWS platforms were very tempting to use, because the modelling approach for screens in FAST is quite similar to SWS, in how they define their inputs and outputs.
+E.g., both FAST and WSMO have pre- and post-conditions, OWL-S has inputs/outputs.
+Also, SWS platforms come with ready-made implementations for service composition, which could have been used for automatically combining screens to screen-flows.
+On the downside, it was expected that SWS platforms a bit too slow for some FAST requirements, as the need of real-time discovery.
+On the other hand, building a light-weight implementation directly on RDFS promised to be faster.
+Moreover, there is a broader, more mature and probably more active tool support.
+However, going in this direction would imply to invest a lot more work in modelling and implementation efforts, since all the features that SWS would bring straight away are could not be reused.
+
+To perform the evaluation of the different approaches, an abstract random ontology of 40 concepts was created.
+About 70\% of those were randomly assigned to be sub-concepts of other concepts.
+The abstract ontology was instantiated in WSMO, OWLS and RDFS.
+Then, 3 sample sets of random screens of three different sizes (100, 1000 and 10000) were created, such that each screen has 0..3 random pre- and post-conditions.
+For each set in each technology, every screen was taken to perform a match with any other screen, and the average response time was measured.
+Figure~\ref{fig:comparison_chart} shows the results of the evaluation (RDFS in blue, WSMO in yellow and OWL-S in green).
+Be aware that the scale of the time is logarithmic.
+
+\begin{figure}[htb]
+\label{fig:comparison_chart}
+\begin{center}
+	\includegraphics[width=14cm]{images/comparison_chart}
+	\caption{Semantic Web Technologies Performance Comparison}
+\end{center}
+\end{figure}
+
+It can be seen that RDFS performs rather well, even in the 10.000 size set, whereas the other two are pretty far behind.
+E.g., average response times for WSMO is > 1.5 sec already for the smallest 100 size set.
+Hence, based on this evaluation the approach to follow for the implementation was RDFS, since WSMO and OWL-S simply could not give the performance required for the real-time discovery.
 

### 2010-02-05 trunk/deliverables/D4_3_2/fast_d4_3_2.tex
UPDATE: done some changes regarding Alex review
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@518 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-These back-end services cannot be directly used; hence they need to be encapsulated in what in the FAST platform is called Resource Adapters.
-The application of semantics to the back-end, through the corresponding resource adapters, and front-end building blocks assures a powerful instrument in the task of building new gadgets, improving the search, and enhancing the connection among the different building blocks which compose a gadget.
+These back-end services cannot be directly used; hence they need to be encapsulated or wrapped into the so-called Resource Adapters within the FAST platform.
+The application of semantics to the back-end, through the corresponding resource adapters, and front-end building blocks ensures a powerful instrument in the task of building new gadgets, improving the search, and enhancing the connection among the different building blocks which compose a gadget.
 
-Therefore, the focus of this deliverable is to define how these wrappers will be constructed to allow the FAST platform exploiting web services \cite{GustavoAlonso2003} such as RESTful web services, SOAP-based web services and semantic web services through WSMO, and define mechanisms to connect them within the gadgets.
+Therefore, the focus of this deliverable is to define how these wrappers will be constructed to allow the FAST platform to exploit web services \cite{GustavoAlonso2003} such as RESTful web services, SOAP-based web services and semantic web services through WSMO, and define mechanisms to connect them within the gadgets.
 
@@
 
-The objective of this deliverable is the analysis and development of mechanisms to facilitate the connection between screen-flow gadgets and underlying Web services, based on front-end user requirements and on the semantic descriptions of the service wrappers, and to develop these service wrappers from of the Web services' APIs and formal descriptions.
+The objective of this deliverable is the analysis and development of mechanisms to facilitate the connection between gadgets and underlying Web services, based on front-end user requirements and on the semantic descriptions of the service wrappers, and the development of these service wrappers from the Web services' APIs and formal descriptions.
 
@@
 Section~\ref{sec:introduction} states the goal, scope and structure of the document.
-Section~\ref{sec:resource_adapters} describes what a resource adapter is, how the platform is able to interact with web services, how they these wrappers or adapters are build and then discovered to be reused in any gadget.
+Section~\ref{sec:resource_adapters} describes what a resource adapter is, how the platform is able to interact with web services, how these wrappers or adapters are build and then discovered to be reused in any gadget.
 Then, Section~\ref{sec:restful_web_services_wrapper_tool} defines the service wrapper tool built to facilitate FAST users to create resource adapters of RESTful web services.
@@
 Web Services, Databases, CORBA or RPC interfaces...).
-These services have to be used by the gadgets, but there are several complications to be resolved in order to communicate the front-end (i.e.
-the gadget user interface and logic) with those services, such as the disparity of interfaces and invocation mechanisms.
+These services have to be used by the gadgets, but there are several complications to be resolved in order for the front-end (i.e.
+the gadget user interface and logic) to communicate with those services, such as the disparity of interfaces and invocation mechanisms.
 Another issue is the complexity for an end-user to interact with those services.
-Hence, this interaction will be shift to a user-interaction paradigm through a user-friendly graphical interface (service front-end), allowing both humans and machines to interact with the services through a uniform fashion.
+Hence, this interaction will be shifted to a user-interaction paradigm through a user-friendly graphical interface (service front-end), allowing both humans and machines to interact with the services through a uniform fashion.
 
@@
 A complex gadget in terms of FAST is aimed to provide a functional access from a graphical user interface to a set of services (SOAP or REST-based Web services) and data sources (Atom/RSS feeds).
-These services and data sources need to be modelled and encapsulated within the platform in order to allow the discovery and use of them by other components (i.e.
+These services and data sources need to be modelled and encapsulated within the platform in order to allow being discovered and used by other components (i.e.
 forms and operators).
-This is called Resource Adapter in the FAST platform.
+These service wrappers are called Resource Adapters in the FAST platform.
 Figure~\ref{fig:complex_gadget_architecture} illustrates the role they play and which other building blocks they interact with within a gadget.
@@
 	\item \textbf{metadata} such as creator, label, description, version, and so on,
-	\item a set of \textbf{actions} or operations,
+	\item a set of \textbf{actions} or operations, with their corresponding preconditions,
 	\item and a set of \textbf{postconditions} or outputs of the resource.
@@
 Figure~\ref{fig:resource_adapter_overview} shows the information above in a graphical manner.
-An action (represented by a white square) is the minimum unit for the functionality a resource may have.
+An action (represented by a white square) is the minimal unit of functionality a resource may have.
 Imagine a user clicking on a button to get a list of items for sale, or a checkout action performed in order to buy one of the items from the list.
@@
 
-REST \cite{Fielding2000} is a term to describe an architecture style, not a standard, of networked systems.
+REST \cite{Fielding2000} is a term used to describe an architecture style, not a standard, of networked systems.
 The acronym REST stands for Representational State Transfer.
-REST-based or RESTful web services are created identifying all of the conceptual entities or resources which want to be exposed as services.
+REST-based or RESTful web services are created identifying all the conceptual entities or resources which need to be exposed as services.
 Those entities or resources should be nouns, not verbs (orders, tickets, etc.).
@@
 
-RESTful web services are more lightweight than SOAP-based web services, avoiding the verbose XML-based messages, but unlike the former, which have a standard vocabulary, and commonly used, to describe the web service interface through WSDL, RESTful web services are currently not formally described most of the times.
+RESTful web services are more light-weight than SOAP-based web services, avoiding the verbose XML-based messages, but unlike the former, which have a standard and commonly used vocabulary to describe the web service interface through WSDL, RESTful web services are currently not formally described most of the times.
 For a service consumer to understand the context and content of the data that must be sent to and received from the service, both the service consumer and service producer must have an out-of-band agreement.
 This takes the form of documentation, sample code, and an API that the service provider publishes for developers to use.
-For example, the many web-based services available from Google, Yahoo, Flickr, Amazon, and so on have accompanying artifacts describing how to consume the services.
-This style of documenting REST-based web services is fine for use by developers, but it averts tools from programmatically consuming such services and generating executable stubs for a given programming language, as a web service described using WSDL allows.
+For example, the most web-based services available from Google, Yahoo, Flickr, Amazon, and so on have accompanying artifacts describing how to consume the services.
+This style of documenting REST-based web services is fine for developers' use, but it averts tools from programmatically consuming such services and generating executable stubs for a given programming language, as a web service described using WSDL allows.
 Nevertheless, Web Application Description Language (WADL) attempts to resolve some of these issues by providing a means to describe services in terms of schemas, HTTP methods, and the request or response structures exchanged, but this language is not widely adopted yet by RESTful web service developers.
@@
 In such services, there is often a machine-readable description of the operations offered by the service written in a Web Services Description Language (WSDL) document.
-The advantage of specifying the service in a using WSDL is that it can be programmatically processed.
+The advantage of specifying the service using WSDL is that it can be programmatically processed.
 
@@
 First one, as a manual process mapping these types and their attributes to concepts and properties, as it was proposed for RESTful services.
-Second approach is taking advantage of SAWSDL, which keeps the process in the programmatically direction.
+Second approach is taking advantage of SAWSDL, which would help ensure a programatical (mapping) process.
 
@@
 These semantic annotations enable the possibility to develop resource adapters from WSDL documents.
-They reference a concept in an ontology or a mapping document, and are either independent of the ontology expression languages and mapping languages.
+They reference a concept in an ontology or a mapping document, and are either independent of the ontology expression languages or mapping languages.
 
-Basically, there are three main concepts about SAWSDL: \emph{modelReference} to specify the association between a WSDL or XML Schema component and a concept from the ontology, and \emph{liftingSchemaMapping} and \emph{loweringSchemaMapping} to define the mapping mechanisms between semantic data and XML.
+Basically, there are three main attributes used in SAWSDL: \emph{modelReference} to specify the association between a WSDL or XML Schema component and a concept from the ontology, and \emph{liftingSchemaMapping} and \emph{loweringSchemaMapping} to define the mapping mechanisms between semantic data and XML.
 These lowering and lifting schema mappings are not explained in this deliverable, however in Appendix A of the SAWSDL specification \cite{sawsdl} they are described in detail, and illustrated with examples.
@@
 Hence, SAWSDL enriches the web services descriptions made in WSDL with references to concepts from a semantic model and enables the possibility to create resource adapters from these definitions without a strong user participation in the process.
-If the semantic metadata is missing, a manual annotation process could be considered, where the user would need to relate concepts from an ontology with the WSDL or XML Schema elements, and create the mapping between them manually as well.
+If the semantic metadata is missing, a manual annotation process could be considered, where the user would need to relate concepts from an ontology with the WSDL or XML Schema elements.
 
@@
 Semantic web services bring a number of advantages in the creation of the resources adapters over classic web services.
-Formal and semantic descriptions of web services allow mechanisms to raise its exploitation in a more automatic way.
+Formal and semantic descriptions of web services allow the creation of mechanisms to raise their exploitation in a more automatic way.
 In this section, some general notions of the Web Service Modelling Ontology (WSMO) will be explained and how FAST can make use of it.
@@
 
-Capabilities and interfaces are the two types of Web Service description in WSMO. The capabilities describe the different functions of WSMO, while the interfaces specify:
+Capabilities and interfaces are the two types of web service description in WSMO. The capabilities describe the different functions of the web service, while the interfaces specify:
 \begin{enumerate}
 	\item How to communicate with a web service in order to avail of its functionality. This is called Choreography.
-	\item How the functionality of a web service is enabled by interacting with other Web Services. This is called Orchestration.
+	\item How the functionality of a web service is enabled by interacting with other web services. This is called Orchestration.
 \end{enumerate}
@@
 
-First of all, an ontology needs to be extracted from the XML Schema types of the WSDL definition.
-For the shake of simplificity, this example does a one-to-one mapping between the Schema types and the concepts, though a real sceneario may consider an different approach to optimise the modelling.
+First of all, an ontology is needed describing the Amazon domain.
+For the sake of simplicity, it will be created doing a one-to-one mapping between the XML Schema types and the concepts, though a real sceneario may consider a different approach to optimise the modelling.
 
@@
 
-It is a fairly simple precondition, which the only condition required is the existence of a \emph{itemSearchRequest} object.
+It is a fairly simple precondition, for which the only condition required is the existence of a \emph{itemSearchRequest} object.
 
@@
 Section 4.4 in \cite{moller2010fast_ontology} gives a detailed explanation on this topic.
-Basically, and using previous examples, the preconditions would be transformed as follows:
+Basically, using previous examples, the preconditions would be transformed as follows:
 
@@
 
+In the previous example, there is not a formal expression saying that \emph{?container} must exists, but it is assumed.
+
 % subsection building (end)
@@
 In the building phase, every resource adapter is created sharing a common structure for screen components, such as forms, operators and resources.
-Hence, it will have a set of actions which will contain a set of preconditions to satisfy in order to be executed, and after its execution, it may produce any of the conditions of their postconditions.
+Hence, it will have a set of actions which will contain a set of preconditions to be satisfied in order to be executed, and after its execution, it may produce any of the conditions of their postconditions.
 
-For this reason, a screen developer during the screen development phase is able to reuse these screen components which are already stored in the catalogue, establishing which the screen's pre/postconditions are, and then the system will suggest several screen components which may satisfy these pre/postconditions.
-In a first step, the system will suggest screen components which have any of the preconditions which can be connected to any of the screen preconditions (the connection will be represented by an internal pipe), and the same would happen for the postconditions.
+For this reason, a screen developer during the development phase is able to reuse these screen components which are already stored in the catalogue, establishing which the screen's pre/postconditions are, and then the system will suggest several screen components which may satisfy these pre/postconditions.
+In a first step, the system will suggest screen components having any of the required preconditions, which can be connected to any of the screen preconditions (the connection will be represented by an internal pipe), and the same would happen for the postconditions.
 Once a screen component is inserted into the screen, the system will take its definition into consideration to recommend new screen components which can be connected to the ones the screen is composed of.
@@
 
-A resource adapter is aimed to perform some action on a web service.
+The aim of a resource adapter is to perform some action on a web service.
 In order to perform a given action, a precondition may need to be fulfilled, and as a result of the execution a postcondition may be generated.
@@
 This is a screen, composed by a form and a resource adapter.
-The form allows the user to introduce a term for the search, producing a postcondition which will be redirect to a resource adapter precondition.
-The execution of the web services produces a list of results, which will be forward to the postcondition of the screen.
+The form allows the user to introduce a search term, producing a postcondition which will be redirected to a resource adapter precondition.
+The execution of the web service produces a list of results, which will be forwarded to the postcondition of the screen.
 
@@
 At the moment, it just supports RESTful web services.
-The building phase for this type of web services is done in two steps: a first step will be in charge of the construction of a service request and a second step will analyse the response got from the execution of the service, allowing the extraction of facts contained in that response and mapping them to domain-specific concepts from the ontologies used within FAST by any building block.
+The building phase for this type of web services wrapper is done in two steps: a first step will be in charge of the construction of a service request and a second step will analyse the response got from the execution of the service, allowing the extraction of facts contained in that response and mapping them to domain-specific concepts from the ontologies used within FAST by any building block.
 
@@
 
-To achieve this, service wrappers will be handled as any buis with preconditions and postconditions (inputs and outputs).
-Therefore, these precondition ports might be used to determine values to parameters like query keywords coming from a building block inside the screen the service is placed or even an external screen as a screen precondition.
-The service wrapper tool composes the request using a template string which will contain placeholders for precondition values.
-Before sending the request to the service, the placeholders are replaced with their corresponding values and the resulting URL is then ready to be sent the service request.
-
 To achieve this, service wrappers will be handled as any other building block with preconditions and postconditions (inputs and outputs).
-Therefore, these precondition ports might be used to determine values to parameters like query keywords coming from a building block inside the screen the service is placed or even an external screen as a screen precondition.
+Therefore, these precondition ports might be used to determine values for parameters like query keywords, coming from a building block inside the screen in which the service is placed or even an external screen, as a screen precondition.
 
@@
   \begin{center}
-    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithRequestExample.jpg}
+    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithPortDefinitions.jpg}
     \caption{Configuring pre/postconditions of a service wrapper}
@@
 The service wrapper tool composes the request using a template string which will contain placeholders for precondition values.
-Before sending the request to the service, the placeholders are replaced with their corresponding values and the resulting URL is then ready to be sent the service request.
+Before sending the request to the service, the placeholders are filled with their corresponding values and the resulting URL is then ready to be sent as the service request.
 
@@
 
-Below the request paramter editing form of Figure~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed service by sending it to the specified service address (via a server relay).
+Below the request parameter editing form of Figure~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed service by sending it to the specified service address (via a server relay).
 Then the placeholders for preconditions are replaced with example values and the resulting HTTP request is shown below the parameter form.
@@
 Once the service request is constructed and sent to the service provider, it will send back a response.
-This response message could be serialize in any format, though the most common formats used nowadays are XML or JSON among others.
-To continue the example it was started in the previous section, the response of the Ebay Shopping service will be in XML as specified in the request, which is the format supported by the wrapper tool.
+This response message could be serialized in any format, though the most common formats used nowadays are XML or JSON among others.
+To continue the example that was started in the previous section, the response of the Ebay Shopping service will be in XML as specified in the request, which is the format supported by the wrapper tool.
 
@@
 A transformation rule is used to analyse the XML data and to generate domain-specific facts from concepts from the ontologies used by the pre/postconditions of the different building blocks.
-A transformation rule is composed of three elements, cf the middle part of Figure~\ref{fig:response_service_execution}.
+A transformation rule is composed of three elements (see the middle part of Figure~\ref{fig:response_service_execution}).
 First, the \textit{from} field indicates the XML elements to be translated by rule.
-These XML elementsare identified by the tagname of a DOM element from the XML document.
+These XML elements are identified by the tagname of a DOM element from the XML document.
 Second, the type of the rule will be set, taking one of the following values: \emph{createObject}, \emph{fillAttributes} or \emph{dummy}.
 And third, the target of the rule specifies a certain concept or attribute, to be created or filled.
-A detailed explanation of the type of actions to be trigger from the transformation rules is:
+A detailed explanation of the type of actions to be triggered from the transformation rules is:
 
@@
 The value of that \emph{Title} element is then transfered to the \emph{productName} attribute of the corresponding \emph{Product} fact.
-Actually, our \textit{from} fields allows also to refer to parts of an XML attribute e.g.
+Actually, our \textit{from} fields allow also to refer to parts of an XML attribute e.g.
 to \textit{words} 1 through 3.
@@
 
-However, we there are some transformations that these rules cannot perform.
+However, there are some transformations that these rules cannot perform.
 For example, we do not support any mathematical operations.
@@
 Such additional filters may be realized using (hand coded) operators, since some generic operators can act as filters for aggregation and conversions of facts from multiple sources.
-Then, service wrappers in combination with these filter operators will allow covering these complex cases.
+Then, service wrappers in combination with these filter operators will allow covering some of these complex cases.
 
@@
 \cite{Schurr94,JSZ97c}.
-While general triple graph grammars allow to relate complex graph structures with each other, in our case the target of a rule is always a single object or attribute.
+While general triple graph grammars allow to relate complex structures to each other, in our case the target of a rule is always a single object or attribute.
 This facilitates the whole mechanism, reasonably.

### 2010-02-09 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_mediation.tex
Corrected based on review of DERI contribution
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@528 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 ontologies, and the differences between them need to be determined and overcome to allow for a seamless data exchange between these parties.
-This is a difficult ask, with which web-based agents must cope.
+This is a difficult task, with which web-based agents must cope.
 
@@
     \textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.4.2 is the second iteration of the FAST ontology matching approach.
-The firsty iteration contained a general overview of the literature regarding ontology mediation, paving the way for futher analysis of the scenarios arising in FAST that require ontology mediation.
+The first iteration contained a general overview of the literature regarding ontology mediation, paving the way for futher analysis of the scenarios arising in FAST that require ontology mediation.
 Based on that general overview, this current interation presents the approach to implement ontology matching within FAST.
@@
 FAST is a visual programming platform allowing business users to build enterprise-class mashups, employing various underlying services and generating new ones.
-Resources in FAST are described semantically using different ontologies and vocabularies, and can therefore be combined to what we can call ``intelligent gadgets''.
+Resources in FAST are described semantically using different ontologies and vocabularies, and can therefore be combined to what we can call ``intelligent gadgets'', that rely on semantically enriched building blocks.
 In the services used within the FAST environment, the ontologies involved in the creation of gadgets and building blocks come from different parties.
@@
 
-\paragraph{Overall Approach for WPX and relation to other Work Packages and Deliverables.}
+\paragraph{Overall Approach for WP2 and relation to other Work Packages and Deliverables.}
 Ontology matching will play an important role for the integration of different underlying resources in the user-built gadgets. 
@@
 
+\pagebreak
 \part{Implementing Ontology Matching within the FAST Platform}
@@
 
+\pagebreak
 \part{D2.4.1: Theoretical Foundations for Ontology Mediation}
 \label{part:d2.4.1}
+
 \include{fast_d2_4_2_part2}

### 2010-02-09 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_part1.tex
Corrected based on review of DERI contribution
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@528 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
     \item SMOA distance with a threshold of \texttt{0.5} and
-    \item Wordnet distance using a threshold of \texttt{0.5}.
+    \item Wordnet distance using a threshold of \texttt{0.5}\footnote{Thresholds were selected based on suggestions in the Alignment API documentation}.
 \end{inparaenum}
@@
 By using string distances and giving certain thresholds (Levenshtein and SMOA), we can see that the results are much less precise, but have a better recall, since this allows for entities having similar names to be discovered, at the expense of having quite a few incorrect results (lower precision); the thresholds allow for low-scored cases to be eliminated, although this results in the exclusion of some correct correspondences.
-The last column (JWNL) contains the results of the Wordnet-enabled method, which shows quite an improvement (precision of 0.67 and a recall of 0.75), due to the lexical analysis, which performs a much more relevant comparison of strings, giving a high number of correct results.
+The last column (JWNL) contains the results of the Wordnet-enabled method, which shows quite an improvement (precision of 0.67 and a recall of 0.75\footnote{The scores are orientative, giving a general picture of how the process approaches 100\% efficiency}), due to the lexical analysis, which performs a much more relevant comparison of strings, giving a high number of correct results.
 The precision of the JWNL alignment shows only a tiny drop below the recall value, meaning that the number of incorrect correspondences discovered is small, and the main source of error is from the number of correspondences not discovered.
@@
 We can deduce that the results provided are satisfactory, even though the methods used were simple, string-based ones, and the process was completely automated without any user input.
-We are therefore confident that through some user assistance or an initial input alignment the tool can achieve 100\% correct results.
+We are therefore confident that through some user assistance or an initial input alignment the tool can achieve 100\% correct results, which is our aim for the tool to be used in practice.
 

### 2010-02-09 trunk/deliverables/D4_3_2/fast_d4_3_2.tex
UPDATE: related work section done
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@530 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
+%\clearpage
+%\listoftables
+
+\clearpage
+\listoffigures
+
 %%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-This deliverable is structured as follows.
-Section~\ref{sec:introduction} states the goal, scope and structure of the document.
-Section~\ref{sec:resource_adapters} describes what a resource adapter is, how the platform is able to interact with web services, how these wrappers or adapters are build and then discovered to be reused in any gadget.
-Then, Section~\ref{sec:restful_web_services_wrapper_tool} defines the service wrapper tool built to facilitate FAST users to create resource adapters of RESTful web services.
-Finally, Section~\ref{sec:related_work}...\todo{Finish it!}
+This deliverable is structured as follows: Section~\ref{sec:introduction} states the goal, scope and structure of the document.
+Before going to the core of the document, a brief introduction to some related work is done in Section~\ref{sec:related_work}.
+Then, Section~\ref{sec:resource_adapters} describes what a resource adapter is, how the platform is able to interact with web services, how these wrappers or adapters are build and then discovered to be reused in any gadget.
+Finishing the deliverable, Section~\ref{sec:restful_web_services_wrapper_tool} defines the service wrapper tool built to facilitate FAST users to create resource adapters of RESTful web services.
 
 % section introduction (end)
+
+\clearpage
+\section{Related Work} % (fold)
+\label{sec:related_work}
+
+This section focuses on different approaches dealing with web services integration and reutilisation.
+That is not a novel idea, since one of the basic pillars of web services is to allow any application to access and execute third-party web services over a network, and there are already loads of approaches about this issue.
+However, most of these approaches require a high-level technical profile to accomplish this integration in a programmatically manner.
+The ones studied in this deliverable are those which presents and allow reusing data sources and web services using a graphical interface or wizard.
+
+\cite{urmetzer2010fast_state_of_the_art} presented several tools to create widgets, which allow the interaction with some kind of data sources and services.
+Yahoo!
+Pipes provides a set \emph{modules} to access different kind of data sources, such as RSS feeds, a given web page (HTML code), Flickr images, Google base or the Yahoo!
+search engine, among others.
+These modules are developed by the widget platform itself, they do not allow users to build their own modules, and it lacks of a way to interact with REST or SOAP web service.
+Another tool to reference is Apatar.
+Apatar is more enterprise-oriented, allowing the interaction with databases (MySQL, PostgreSql, Oracle), and powerful enterprise systems such as Salesforces.com or SugarCRM.
+It also provides access to data sources such as XML or text files.
+But again, there is no way to use web services.
+Finally, Presto is another solution enterprise-oriented.
+It allows the common data sources such as RSS and ATOM feeds or XML-based sources.
+However, Presto brings connectors to use services from HP SOA Systinet and any Oracle information technologies including Oracle 9i/10g/11g, Oracle Fusion SOAs, and Oracle Applications.
+
+Summarising up, none of the solutions in the market really facilitate the users to build up their own widgets using any kind of web service in a graphical manner.
+Their approaches are either data-oriented (RSS feeds, databases, raw text), or including a small set of web services in the case of Presto.
+FAST will bring the possibility of integrate REST or SOAP-based web service inside a gadget (by creating service wrappers), and graphical tool will be provided to facilitate the process of building these service wrappers.
+In a similar way, and in the context of bioinformatics web services, \cite{gordon2008} explains how to publish web services to be reused in their tool as Moby (semantic) services.
+They have developed the Daggoo\footnote{http://www.daggoo.net/} prototype which will create these service wrappers to be used in Moby.
+
+% section related_work (end)
 
@@
 The definition of the inputs, outputs, operations and their types are done in a XML Schema document.
-An XML schema describes, in a syntactic manner, the structure of an XML document, but there is no straight-forward relation with the concepts used to describe pre/postconditions in the building blocks.
+An XML schema describes, in a syntactic manner, the structure of an XML document, but there is no straight-forward relation with the concepts used to describe pre and postconditions in the building blocks.
 In order to illustrate the concepts of WSDL, and the gap between them and semantically-defined building blocks, a shorten version of a real Amazon E-commerce web service will be presented.
@@
 
-For this reason, a screen developer during the development phase is able to reuse these screen components which are already stored in the catalogue, establishing which the screen's pre/postconditions are, and then the system will suggest several screen components which may satisfy these pre/postconditions.
+For this reason, a screen developer during the development phase is able to reuse these screen components which are already stored in the catalogue, establishing which the screen's pre and postconditions are, and then the system will suggest several screen components which may satisfy these pre or postconditions.
 In a first step, the system will suggest screen components having any of the required preconditions, which can be connected to any of the screen preconditions (the connection will be represented by an internal pipe), and the same would happen for the postconditions.
@@
 The service wrapper tool composes the request using a template string which will contain placeholders for precondition values.
+An example of a template with one placeholder (<search\_query>), to perform a search on google could be like:
+
+\begin{verbatim}
+http://www.google.com/search?lang=es&q=<search_query>
+\end{verbatim}
+
 Before sending the request to the service, the placeholders are filled with their corresponding values and the resulting URL is then ready to be sent as the service request.
+Following the previous example, the placeholder \emph{search\_query} is filled with a specific search term:
+
+\begin{verbatim}
+http://www.google.com/search?lang=es&q=six+nations+2010
+\end{verbatim}
 
@@
 This tree representation of the XML data is used as an input to construct transformation rules.
+
+The rule based approach of the service wrapper tool has been inspired by triple graph grammars \cite{Schurr94,JSZ97c}.
+While general triple graph grammars allow to relate complex structures to each other, in the approach taken for FAST, the target of a rule is always a single object or attribute making the process as easy as possible for the user.
 
@@
 \clearpage
-\section{Related Work} % (fold)
-\label{sec:related_work}
-
-The rule based approach of the service wrapper tool has been inspired by triple graph grammars, cf.
-\cite{Schurr94,JSZ97c}.
-While general triple graph grammars allow to relate complex structures to each other, in our case the target of a rule is always a single object or attribute.
-This facilitates the whole mechanism, reasonably.
-
-Write about some related work such as:
-
-Wrapping WSDL-Described Web Services as Moby Services (using SAWSDL)
-\url{http://biomoby.open-bio.org/CVS_CONTENT/moby-live/Java/docs/sawsdlServlet.html}
-
-Perhaps, it's worth to mention the service wrappers Yahoo Pipes has, which are a fixed set developed by them, with no possibility for a user to create new ones, as we do.
-(other sites doing the same??)
-
-% section related_work (end)
-
-\clearpage
 \bibliographystyle{apalike}
@@
 
-\clearpage
-\doublespacing
-\section*{Appendix A (Lists of Tables and Figures)}
-\addcontentsline{toc}{section}{Appendix A (Lists of Tables and Figures)}
-
-\listoftables
-
-\listoffigures
-
 \end{document}

### 2010-02-15 trunk/deliverables/D5_1_2/fast_d5_1_2.tex
UPDATE: text about linked data and changes to building blocks structures
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@554 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 Also some configuration instructions are specified.
-	\item Section~\ref{sub:content_negotiation}. Content Negotiation states how the FAST Catalogue has adopted several of the content negotiation principles.
+	\item Section~\ref{sub:linked_data}. Content Negotiation states how the FAST Catalogue has adopted several of the content negotiation principles.
 	\item Section~\ref{sec:restful_catalogue_api}.
@@
 
+Note that it is also possible to get the compiled WAR file directly from the SVN server. This file can be located at \url{https://svn.forge.morfeo-project.org/fast-fp7project/distribution/catalogue}.
+
 Then, last step is to deply the WAR file into the servlet container. If the server chosen was Apache Tomcat, the procedure for deploying a Web application is:
@@
 Independently the type of repository, the parameter \emph{serverURL} needs to take the value of the URL where the FAST Catalogue instance has been deployed.
-If you are testing the application in your machine, it may be something like \url{http://localhost:8080/FASTCatalogue}.
+If you are testing the application in your machine, it may be something like \url{http://catalogueURL/FASTCatalogue}.
 
@@
 To do that, a simple HTTP request, using a Web browser, can be sent to the URL the Catalogue is deployed, and specifying a type of resource such as Screen.
-The URL would be similar to \url{http://localhost:8080/FASTCatalogue/screens}.
+The URL would be similar to \url{http://catalogueURL/FASTCatalogue/screens}.
 Depending on the variant chosen, the response may be different.
@@
   "canvas": [
-    { "uri": "http://localhost:8080/catalogue/screens/238" },
-    { "uri": "http://localhost:8080/catalogue/screens/323" }
+    { "uri": "http://catalogueURL/catalogue/screens/238" },
+    { "uri": "http://catalogueURL/catalogue/screens/323" }
   ],
   "elements": [
-    { "uri": "http://localhost:8080/catalogue/screens/636" }
+    { "uri": "http://catalogueURL/catalogue/screens/636" }
   ],
@@
 
-\subsection{Content Negotiation} % (fold)
-\label{sub:content_negotiation}
+\subsection{Linked Data} % (fold)
+\label{sub:linked_data}
+
+Apart from serving as the back-end system for the Gadget Visual Storyboard or GVS, the catalogue publishes all FAST building blocks as linked data following the four rules or principles dictated in \cite{bernersLee2006linkedData}.
+Hence, building blocks are identified by URIs, concretely HTTP URIs so anyone can look them up.
+The data obtained for these building blocks is served in different standard formats such as JSON (GVS communication), RDF/XML, Turtle, or even HTML+RDFa as a human-readable version.
+And as the forth rule states, these building blocks links to other resources on the web.
+Moreover, principles and best practices from \cite{berrueta2008} and \cite{sauermann2008cool_uris} have been followed for the publication and identification of the building blocks.
+The mechanism to allow its publication/identification in different formats is called Content Negotiation.
 
@@
 
-In summary, the basic idea of content negotiation, as stated in the HTTP 1.1 specification ~\cite{http1.1}, is to serve the best variant for a resource, taking into account what variants are available, what variants the server may prefer to serve, what the client can accept, and with which preferences: in HTTP, this is done by the client which may send, in its request, Accept headers (Accept, Accept-Language and Accept-Encoding), to communicate its capabilities and preferences in Format, Language and Encoding, respectively.
-
-In fact, what the catalogue really does is "format negotiation" since the alternate representationos are just based on the selection of the media type, through the Accept header, but does not consider different languages or encoding types.
-The formats supported are JSON, RDF+XML and HTML.
+In summary, the basic idea of content negotiation, as stated in the \cite{http1.1}, is to serve the best variant for a resource, taking into account what variants are available, what variants the server may prefer to serve, what the client can accept, and with which preferences: in HTTP, this is done by the client which may send, in its request, Accept headers (Accept, Accept-Language and Accept-Encoding), to communicate its capabilities and preferences in Format, Language and Encoding, respectively.
+
+In fact, what the catalogue really does is "format negotiation" since the alternate representationos are just based on the selection of the media type, through the \emph{Accept} header, but does not consider different languages or encoding types.
+The formats supported are JSON, RDF/XML, Turtle and HTML+RDFa.
+Even though the \emph{Accept} header is desired, these different representations can be retrieved by adding the format at the end of the resource URI.
+For instance, to get the RDF representation of the screen 457, performing a GET HTTP request to <catalogue\_url>/screens/457.rdf will serve the information in RDF/XML.
 
@@
 
-To create a new screen, a POST request is send to the catalogue server, including the JSON representation of the screen as the body of the request:
+To create a new screen, a POST request is send to the catalogue server, concretely to http://catalogueURL/screens, including the JSON representation of the screen in the body of the request:
 
@@
   "type": "screen",
-  "uri": "http://localhost:8080/catalogue/screens/28",
+  "uri": "http://catalogueURL/catalogue/screens/28",
   "version": "0.1"
@@
     "type": "screen",
-    "uri": "http://localhost:8080/catalogue/screens/28",
+    "uri": "http://catalogueURL/catalogue/screens/28",
     "version": "0.1"
@@
     ...,
-	  "uri": "http://localhost:8080/catalogue/screens/14",
+	  "uri": "http://catalogueURL/catalogue/screens/14",
     "version": "1.0"
@@
     ...,
-	  "uri": "http://localhost:8080/catalogue/screens/564",
+	  "uri": "http://catalogueURL/catalogue/screens/564",
     "version": "1.0"
@@
 
-Following is shown an example of usage of this operation considering a canvas with the screen \texttt{http://localhost:8080/catalogue/screens/654}.
+Following is shown an example of usage of this operation considering a canvas with the screen \texttt{http://catalogueURL/catalogue/screens/654}.
 
@@
   canvas: [
-    { uri: "http://localhost:8080/catalogue/screens/654" }
+    { uri: "http://catalogueURL/catalogue/screens/654" }
   ],
@@
 [
-  "http://localhost:8080/catalogue/screens/371",
-  "http://localhost:8080/catalogue/screens/24",
-  "http://localhost:8080/catalogue/screens/253",
-  "http://localhost:8080/catalogue/screens/12"
+  "http://catalogueURL/catalogue/screens/371",
+  "http://catalogueURL/catalogue/screens/24",
+  "http://catalogueURL/catalogue/screens/253",
+  "http://catalogueURL/catalogue/screens/12"
 ]
@@
   "canvas": [
-    { "uri": "http://localhost:8080/catalogue/screens/238" }
+    { "uri": "http://catalogueURL/catalogue/screens/238" }
   ],
   "elements": [
-    { "uri": "http://localhost:8080/catalogue/screens/636" }
+    { "uri": "http://catalogueURL/catalogue/screens/636" }
   ],
@@
     "reachability": false,
-    "uri": "http://localhost:8080/catalogue/screens/238"
+    "uri": "http://catalogueURL/catalogue/screens/238"
   }],
@@
     "reachability": false,
-    "uri": "http://localhost:8080/catalogue/screens/636"
+    "uri": "http://catalogueURL/catalogue/screens/636"
   }]
@@
       "reachability": false,
-      "uri": "http://localhost:8080/catalogue/screens/636"
+      "uri": "http://catalogueURL/catalogue/screens/636"
     },
@@
       "reachability": true,
-      "uri": "http://localhost:8080/catalogue/screens/132"
+      "uri": "http://catalogueURL/catalogue/screens/132"
     }
@@
 [
-  "http://localhost:8080/catalogue/screens/12",
-  "http://localhost:8080/catalogue/screens/48",
-  "http://localhost:8080/catalogue/forms/9",
+  "http://catalogueURL/catalogue/screens/12",
+  "http://catalogueURL/catalogue/screens/48",
+  "http://catalogueURL/catalogue/forms/9",
 ]
@@
       ...,
-      "uri": "http://localhost:8080/catalogue/screens/12",
+      "uri": "http://catalogueURL/catalogue/screens/12",
       "version": "1.0"
@@
       ...,
-      "uri": "http://localhost:8080/FASTCatalogue/forms/9",
+      "uri": "http://catalogueURL/FASTCatalogue/forms/9",
       "version": "1.0"
@@
 	\item[backendservices] It is a list of back-end services, such as the list of back-end services recommended last execution.
+	\item[search] Boolean indicating if forms/operators/services are required to be looked for. If \emph{false} the operation performed is just a check. If missed, it will act as \emph{true}.
 	\item[domainContext] The domain context contains a list of tags and a user.
@@
   "canvas": [
-    { "uri": "http://localhost:8080/catalogue/forms/1" },
-    { "uri": "http://localhost:8080/catalogue/services/1" }
+    { "uri": "http://catalogueURL/catalogue/forms/1" },
+    { "uri": "http://catalogueURL/catalogue/services/1" }
   ],
@@
   "backendservices": [],
+  "search": true,
   "domainContext": {
@@
       "from": {
-        "buildingblock": "http://localhost:8080/catalogue/services/1",
+        "buildingblock": "http://catalogueURL/catalogue/services/1",
         "condition": "list"
@@
         "action": "showTable",
-        "buildingblock": "http://localhost:8080/catalogue/forms/1",
+        "buildingblock": "http://catalogueURL/catalogue/forms/1",
         "condition": "list"
@@
   ],
-  "selectedItem": "http://localhost:8080/catalogue/forms/1"
+  "selectedItem": "http://catalogueURL/catalogue/forms/1"
 }
@@
       "reachability": false,
-      "uri": "http://localhost:8080/catalogue/services/1"
+      "uri": "http://catalogueURL/catalogue/services/1"
     },
@@
       "reachability": true,
-      "uri": "http://localhost:8080/catalogue/forms/1"
+      "uri": "http://catalogueURL/catalogue/forms/1"
     }
@@
         "action": "search",
-        "buildingblock": "http://localhost:8080/catalogue/services/1",
+        "buildingblock": "http://catalogueURL/catalogue/services/1",
         "condition": "filter"
@@
 {
-  "goal": "http://localhost:8080/catalogue/screens/35",
+  "goal": "http://catalogueURL/catalogue/screens/35",
   "canvas": [
-    { "uri": "http://localhost:8080/catalogue/screens/56" },
-    { "uri": "http://localhost:8080/catalogue/screens/85" }
+    { "uri": "http://catalogueURL/catalogue/screens/56" },
+    { "uri": "http://catalogueURL/catalogue/screens/85" }
   ],
@@
   [
-    "http://localhost:8080/catalogue/screens/13",
-    "http://localhost:8080/catalogue/screens/23"
+    "http://catalogueURL/catalogue/screens/13",
+    "http://catalogueURL/catalogue/screens/23"
   ],
   [
-    "http://localhost:8080/catalogue/screens/79",
-    "http://localhost:8080/catalogue/screens/37",
-    "http://localhost:8080/catalogue/screens/14"
+    "http://catalogueURL/catalogue/screens/79",
+    "http://catalogueURL/catalogue/screens/37",
+    "http://catalogueURL/catalogue/screens/14"
   ],
   [
-    "http://localhost:8080/catalogue/screens/16",
-    "http://localhost:8080/catalogue/screens/44",
-    "http://localhost:8080/catalogue/screens/32"
+    "http://catalogueURL/catalogue/screens/16",
+    "http://catalogueURL/catalogue/screens/44",
+    "http://catalogueURL/catalogue/screens/32"
   ]
@@
   "creationDate": "2009-04-20T17:00:00+0100",
-  "creator": "http://www.fabio.es/",
+  "creator": "username",
   "description": {"en-gb": "This is a description of the building block"},
-  "tags": {
-    "tags": [
-      "ebay",
-      "amazon",
-    ],
-    "user": <String> // not defined yet
-  },
-  "homepage": "http://www.homepage.ie/",
-  "icon": "http://url.com/images/icon.png",
-  "id": "Id of the resource",
+    "tags":[
+      {
+        "label":{"en-gb":"Amazon"},
+        "means":"http://dbpedia.org/page/Amazon.com"
+      },
+      {
+        "label":{"en-gb":"eBay"},
+        "means":"http://dbpedia.org/page/Ebay"
+      }
+    ],
+  "homepage": "http://url.com/.../homepage",
+  "icon": "http://url.com/images/.../icon.png",
+  "id": "Id of the building block", // i.e. 3545
   "label": {"en-gb": "Label or title of the building block"},
-  "name": "Amazon Item to Ebay Filter",
+  "name": "User-friendly name of the building block",
   "rights": "http://creativecommons.org/",
-  "screenshot": "http://www.deri.ie/eBayList-screenshot.jpg",
-  "type": "type of the building block", // screenflow, screen, form, operator y resource
-  "uri": "http://localhost:8080/catalogue/screens/525",
+  "screenshot": "http://url.com/.../screenshot.jpg",
+  "type": "type of the building block", // screenflow, screen, form, operator or resource
+  "uri": "http://url.com/screens/525",
   "version": "1.0"
@@
   "contains": [
-    "http://purl.oclc.org/fast/ontology/gadget#Screen1257860772723",
-    "http://purl.oclc.org/fast/ontology/gadget#Screen1257860720035",
+    "http://url.com/screens/72723",
+    "http://url.com/screens/20035",
   ],
@@
     }]],
-    "uses": []
+    "uses": [{
+      "id":"cart",
+      "uri":"http://aws.amazon.com/AWSECommerceService#ShoppingCart"
+    }]
   }],
-  "libraries": [],
+  "libraries": [{
+    "language":"JavaScript",
+    "source":"http://url.com/libcode.js"
+  }],
   "postconditions": [[{

### 2010-02-16 trunk/deliverables/D5_1_1/fast_d5_1_1.tex
UPDATE: change 5.1.2 to 5.1.1 and some minor changes to the document
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@558 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-02-15 trunk/deliverables/D5_1_2/fast_d5_1_2.tex
@@
 
-\newcommand\deliverableNumber{D5.1.2}
+\newcommand\deliverableNumber{D5.1.1}
 \newcommand\deliverableTitle{User Manual of the FAST Catalogue}
@@
 \newcommand\authorOne{Ismael Rivera, NUIG}
-\newcommand\authorTwo{Author 2}
+\newcommand\authorTwo{Ciprian Palaghita, Cyntelix}
 \newcommand\authorThree{Author 3}
@@
 		{\large \sffamily \authorOne}\\
+		\vspace{0.2cm}
+		{\large \sffamily \authorTwo}\\
 		\vspace{0.5cm}
 		\vfill
-		{\large \sffamily Date: 25/01/2010}\\
+		{\large \sffamily Date: 19/01/2010}\\
 		\vspace{1cm}
@@
                          \textcolor{white}{\textbf{Change description}}\\ \hline
-1.0 & 25/01/2010 & Ismael Rivera (NUIG) & First version of the deliverable \\ \hline
+1.0 & 19/01/2010 & Ismael Rivera (NUIG) & First version of the deliverable \\
+ & & Ciprian Palaghita (Cyntelix) & \\ \hline
 \end{tabular}
@@
 
-This should be the first version of this deliverable since in M12 the deliverable 5.1 was not scheduled.
-Nevertheless, last year an extra document was written as an appendix to the FAST Catalogue in order to detail and clarify technical decisions and to have a technical reference manual.
-Hence, the deliverable D5.2.1 will evolve to become the D5.1.2, so it will be no longer needed by itself.
+This is the first version of this deliverable since in M12 the deliverable 5.1 was not scheduled.
+Nevertheless, last year an extra document was written as an appendix to the FAST Catalogue in order to detail and clarify technical decisions and to have a technical reference manual, which the content was taken from as the initial draft the present deliverable.
+Hence, the deliverable D5.2.1 will evolve to become the D5.1.1, so it will be no longer needed by itself.
 
@@
 \begin{center}
-\begin{tabular}{|p{2cm}|p{9.5cm}|p{2cm}|}
+\begin{tabular}{|p{2.5cm}|p{9cm}|p{2cm}|}
 \hline
@@
                          \textcolor{white}{Type}\\ \hline
-Canvas & The canvas indicating which screens satisfy the critetion.
+canvas & The canvas indicating which screens satisfy the critetion.
 More aditional information may be included, for example, a satisfaction attribute for every precondition for reachability.
 & Optional\\ \hline
-Domain Context & The domain context contains a list of tags and a user. & Optional\\ \hline
-Elements & The elements indicating which satisfy the critetion. More aditional information may be included as in the Canvas. & Optional\\ \hline
+domainContext & The domain context contains a list of tags and a user. & Optional\\ \hline
+elements & The elements indicating which satisfy the critetion. More aditional information may be included as in the Canvas. & Optional\\ \hline
 \end{tabular}
@@
 \bibliographystyle{apalike}
-\bibliography{fast_d5_1_2}
+\bibliography{fast_d5_1_1}
 

### 2010-02-16 trunk/deliverables/D5_1_1/fast_d5_1_1.tex
changes to linked data section, internationalisation example
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@560 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \usepackage{amssymb}
+
+% This helps to get hyphenation in typewriter font (\texttt)
+\usepackage[htt]{hyphenat}
+
 \parindent0pt
@@
   "label": {
-    "en-GB": "Simple Example",
+    "en-GB": "Product Search",
     "es-ES": "Ejemplo Sencillo",
-    "de-DE": "Einfaches Beispiel"
+    "de-DE": "Produktsuche"
   },
   "description": {
-    "en-GB": "This is a simple example of a screen",
+    "en-GB": "On this screen, the user can enter a search term.",
     "es-ES": "Esto es un ejemplo sencillo de pantalla",
-    "de-DE": "Dies ist ein einfaches Beispiel für einen Bildschirm"
+    "de-DE": "Auf diesem Screen kann der Benutzer einen Suchbegriff eingeben."
   },
@@
 
-Apart from serving as the back-end system for the Gadget Visual Storyboard or GVS, the catalogue publishes all FAST building blocks as linked data following the four rules or principles dictated in \cite{bernersLee2006linkedData}.
-Hence, building blocks are identified by URIs, concretely HTTP URIs so anyone can look them up.
-The data obtained for these building blocks is served in different standard formats such as JSON (GVS communication), RDF/XML, Turtle, or even HTML+RDFa as a human-readable version.
-And as the forth rule states, these building blocks links to other resources on the web.
-Moreover, principles and best practices from \cite{berrueta2008} and \cite{sauermann2008cool_uris} have been followed for the publication and identification of the building blocks.
-The mechanism to allow its publication/identification in different formats is called Content Negotiation.
-
-Before any major details about how content negotiation is addressed by the catalogue, a few concepts need to be understood.
-First of all, say a URI is not a file name.
-It is common to see a URI as the location for a file, but a URI or Universal Resource Identifier is not a URL: it is not a file name or location, it is an identifier (or a reference) to a resource.
-Then by dereferencing a URI the actual resource can be retrieved.
-Let's see a request to retrieve the resource "Screen 123" is sent to the server.
-The resource is just a piece of information, however the server will answer an HTML document with a text describing the screen or an image representing it.
-On the Web, these equivalent representations of a resource are called variants, and content negotiation is the mechanism used to determine which of the representations of most appropiate for a given request.
+Apart from serving as the back-end system for the Gadget Visual Storyboard (GVS), the catalogue publishes all FAST building blocks as linked data, following the principles as originally defined in \cite{bernersLee2006linkedData}, in order to make them available to arbitrary third party applications.
+Each building block is identified by an HTTP URI and hosted through the catalogue so that it can be dereferenced through the same URI.
+For each building block, data is available in representations in different standard formats such as JSON (for communication with the GVS), RDF/XML, Turtle, or even HTML+RDFa as a human-readable version.
+Following the principles and best practices proposed in \cite{berrueta2008} and \cite{sauermann2008cool_uris}, these representations are served based on the request issued by the requesting agent, using a technique called \emph{content negotiation}.
+As required by the forth rule of linked data, individual building blocks link to other data on the Web, thereby preventing so-called isolated ``data islands''.
+
+In order to understand how content negotiation is implemented in the catalogue, a few concepts need to be understood.
+The URI of a particular building block must be understood as the identifier of the building block as such, as opposed to a particular representation (JSON, HTML, etc.) of it.
+In fact, each such representation has its own URI, under which it can be retrieved.
+Using the terminology established in \cite{w3c2004webArchitecture}, the building block is a \emph{non-information resource}, whereas each representation is an \emph{information resource}.
+In the process of content negotiation, the requesting agent will first request the URI of the building block as such (e.g., \texttt{http://catalogue.com/screens/235}), as well as the required representation format (e.g., JSON).
+The catalogue will then inspect the request and redirect the requester to the appropriate representation (e.g., \texttt{http://catalogue.com/screens/235.json}), as illustrated in Fig.~\ref{fig:content_negotiation}.
+On the Web, different representations for the same resource are called \emph{variants}, and content negotiation is the mechanism used to determine which of the representations of most appropriate for a given request.
 
@@
 
-In summary, the basic idea of content negotiation, as stated in the \cite{http1.1}, is to serve the best variant for a resource, taking into account what variants are available, what variants the server may prefer to serve, what the client can accept, and with which preferences: in HTTP, this is done by the client which may send, in its request, Accept headers (Accept, Accept-Language and Accept-Encoding), to communicate its capabilities and preferences in Format, Language and Encoding, respectively.
-
-In fact, what the catalogue really does is "format negotiation" since the alternate representationos are just based on the selection of the media type, through the \emph{Accept} header, but does not consider different languages or encoding types.
+In summary, the basic idea of content negotiation, as stated in the \cite{http1.1}, is to serve the best variant for a resource, taking into account what variants are available, what variants the server may prefer to serve, what the client can accept, and with which preferences.
+In HTTP, this is done by the client which may send, in its request, accept headers (\texttt{Accept}, \texttt{Accept-Language} and \texttt{Accept-Encoding}), to communicate its capabilities and preferences in format, language and encoding, respectively.
+
+In fact, what the catalogue really does is ``format negotiation'' since the alternate representations are just based on the selection of the media type, through the accept header, but does not consider different languages or encoding types.
 The formats supported are JSON, RDF/XML, Turtle and HTML+RDFa.
-Even though the \emph{Accept} header is desired, these different representations can be retrieved by adding the format at the end of the resource URI.
-For instance, to get the RDF representation of the screen 457, performing a GET HTTP request to <catalogue\_url>/screens/457.rdf will serve the information in RDF/XML.
+Even though the accept header is desired, the different representations can also be retrieved directly by dereferencing their own URI.
+The convention used in the catalogue for deriving a representation URI from a building block URI is to simply attach a matching suffix, such as \texttt{.rdf}, \texttt{.json} or \texttt{.html}.
 
@@
 
-To create a new screen, a POST request is send to the catalogue server, concretely to http://catalogueURL/screens, including the JSON representation of the screen in the body of the request:
+To create a new screen, a POST request is sent to the catalogue server, concretely to \texttt{http://cata\-logueURL/screens}, including the JSON representation of the screen in the body of the request:
 

### 2010-02-16 trunk/deliverables/D5_1_1/fast_d5_1_1.tex
UPDATE: minor changes
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@562 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \newcommand\workpackageTitle{Semantic catalogue of screen-flow resources and back-end Web Services}
-\newcommand\authorOne{Ismael Rivera, NUIG}
-\newcommand\authorTwo{Ciprian Palaghita, Cyntelix}
+\newcommand\authorOne{Ciprian Palaghita, Cyntelix}
+\newcommand\authorTwo{Ismael Rivera, NUIG}
 \newcommand\authorThree{Author 3}
@@
 		\vfill
-		{\large \sffamily Date: 19/01/2010}\\
+		{\large \sffamily Date: 19/02/2010}\\
 		\vspace{1cm}
@@
                          \textcolor{white}{\textbf{Change description}}\\ \hline
-1.0 & 19/01/2010 & Ismael Rivera (NUIG) & First version of the deliverable \\
- & & Ciprian Palaghita (Cyntelix) & \\ \hline
+1.0 & 19/02/2010 & Ciprian Palaghita (Cyntelix) & First version of the deliverable \\
+ & & Ismael Rivera (NUIG) & \\ \hline
 \end{tabular}
@@
 \begin{tabular}
-	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
+	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.2cm}|p{1.6cm}|p{3.37cm}|}
 	\hline
@@
 	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
-%	{} & \multicolumn{4}{l|}{\authorOne}\\ 
+	{} & \multicolumn{4}{l|}{\authorTwo}\\ 
 %	{} & \multicolumn{4}{l|}{\authorOne}\\ 
@@
   \hline
-	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{NUIG} & \textbf{email} & {ismael.rivera@deri.org} \\ \cline{4-5}
-	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495338} \\ \hline
+	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{Cyntelix} & \textbf{email} & {cpalaghita@cyntelix.com} \\ \cline{4-5}
+	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 858494258} \\ \hline
 	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, semantic catalogue, gadget catalogue, RDF store}\\ \hline

### 2010-02-16 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_part1.tex
Complete - taskD2.4.2 reacted on reviewer comments
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@563 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-For services the Kassel members of the FAST team have developed a service wrapper tool, cf.
+To support manual ontology mapping, we plan to develop a rule based, interactive Facts Transformation Tool.
+This tool utilizes techniques of our service wrapper tool, cf.
 deliverable D4.3.2.
@@
 This problem is closely related to the problem of ontology mediation.
-Thus, we plan to use these techniques for the development of a Facts Transformation Tool that allows to create data transformation operators, i.e.
+Thus, these techniques will also allow to create data transformation operators, i.e.
 operators that mediate between two ontologies.
 
-Figure \ref{fig:FactsTransformationToolPortDefinitions} shows a photoshoped preview of the planned data transformation tool.
-The data transformation tool has two tabs.
+Figure \ref{fig:FactsTransformationToolPortDefinitions} shows a photoshoped preview of the planned Facts Transformation Tool.
+The Facts Transformation Tool has two tabs.
 The \textit{General} tab allows to define the input and output ports for the desired data transformation operator.
@@
 
+Figure \ref{fig:FactsTransformationToolTransformationRules} shows the planned editor for transformation rules.
+On the left of Figure \ref{fig:FactsTransformationToolTransformationRules} there is a tree of example data that is used to validate the transformation rules during editing.
+Thus, the Facts Transformation Tool will execute the transformation rules after any editing in order to give direct feedback on their functionality and to enable the user to validate the achieved transformation.
+The transformation results are shown on the right of Figure \ref{fig:FactsTransformationToolTransformationRules}.
+
 \begin{figure}
   \begin{center}
-\includegraphics[width=0.8\linewidth]{images/FactsTransformationToolGVSWithPortDefinitions.jpg}
-    \caption{Definition input ports and output ports for data transformation operators (photoshoped screen dump)}
+\includegraphics[width=0.8\linewidth]{images/FactsTransformationToolGVSWithPortDefinitionsNoGlitch2.png}
+    \caption{Definition of operator pre- / postconditions (photoshoped screen dump)}
     \label{fig:FactsTransformationToolPortDefinitions}
@@
 
-Figure \ref{fig:FactsTransformationToolPortDefinitions} shows the planned editor for transformation rules.
-On the left of Figure \ref{fig:FactsTransformationToolPortDefinitions} there is a tree of example data that is used to validate the transformation rules during editing.
-Thus, the Facts Transformation Tool will execute the transformation rules after any editing in order to give direct feedback on their functionality and to enable the user to validate the achieved transformation.
-The transformation results are shown on the right of Figure \ref{fig:FactsTransformationToolPortDefinitions}.
-
-The middle of Figure \ref{fig:FactsTransformationToolPortDefinitions} shows some possible transformation rules for our example.
+The middle of Figure \ref{fig:FactsTransformationToolTransformationRules} shows some possible transformation rules for our example.
 A \textit{createObject} rule may be used to create output fact objects.
@@
 
-Note, simple one to one transformation rules may be derived by the ontology mediation approach described above and such rules may be added to the transformation tool, automatically.
+Note that simple one-to-one transformation rules should be derived with the help of the ontology mediation approach described above.
+Such simple rules should be added to the transformation tool, automatically.
 The more complex rules may then be added interactively.
-During editing, the semantics provided by the FAST catalogue may be used to support the user with completion proposals and consistency checking.
+During editing, the semantics provided by the FAST catalogue will be used to support the user with completion proposals and consistency checking.
 In addition, the direct execution of the rules on example data will help the user to create appropriate rules and to validate their effects.
 
-Note, so far our transformation rules are restricted to simple cases.
+So far our transformation rules are restricted to simple cases.
 So far we have focused on simple cases allowing a simple transformation approach that does not require to sophisticated skills from our users.
 However, as discussed on the Amazon to GoodRelations example, there exist more complex cases.
-In the third year of the project we will investigate how such more complex cases may be covered, too, and whether it is still possible to keep the rules simple and understandable.
+In the third year of the project we will investigate how such more complex cases can be covered, too, and whether it is still possible to keep the rules simple and understandable.
 
@@
 
-Once the data transformation operator has been defined, it may be added to the FAST catalogue in order to allow its use for the definition of piping configurations of new screens within the FAST CGS tool.
+Once the data transformation operator has been defined, it will be added to the FAST catalogue in order to allow its use for the definition of piping configurations of new screens within the FAST CGS tool.
 For this purpose a simple semantics description will be send to FAST catalogue defining the semantics of the operator and the semantics of its input and output port.
 
-Currently, the Service Wrapper Tool uses its own proprietary repository for trees of transformation rules.
-First of all this has technical reasons since the development of the transformation tool is based on an reference architecture and libraries for web based interactive systems provided by Kassel University, cf.
+Currently, the Service Wrapper Tool uses its own repository for trees of transformation rules.
+First of all this has technical reasons: The development of the transformation tool is based on an reference architecture and libraries for web based interactive systems provided by Kassel University, cf.
 \cite{AsDH2009CVSM}.
 This repository has the special advantage that it provides support for teams of developers, thus the service wrapper tool allows to merge the contributions of multiple developers.
-However, on the long run, we may store the transformation rules together with the service wrapper and transformation operator definitions in the FAST catalogue.
+However, on the long run, we shall store the transformation rules together with the service wrapper and transformation operator definitions in the FAST catalogue.
 
@@
 
-At the gadget building phase, whenever two screens are combined, the catalogue will provide the possible matching operators for the two screens.
-This way an operator can be placed between two screens that would otherwise be incompatible, making interoperation between the two possible.
+At the gadget building phase, whenever preconditions and postconditions of screens, resources or operators show a semantic mismatch, the catalogue will offer the possible mediation operators that would solve the incompatibility problem.
+This way an mediation operator can be placed between two pre- postconditions that would otherwise be incompatible, making interoperation between the two possible.
+In seldom cases, the catalogue may fail to retrieve an appropriate mediation operator.
+In such cases building of a new mediation operator with the help of the Facts Transformation Tool will be triggered in order to meet the users need.
 

### 2010-02-17 trunk/deliverables/D5_1_1/fast_d5_1_1.tex
changes to Internationalisation section
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@564 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \begin{small}
-\begin{tabular}{|l|l|l|p{7.5cm}|}
+\begin{tabular}{|l|l|l|p{6.6cm}|}
 \hline
@@
     "en-GB": "Product Search",
-    "es-ES": "Ejemplo Sencillo",
+    "es-ES": "Búsqueda de productos",
     "de-DE": "Produktsuche"
@@
     "en-GB": "On this screen, the user can enter a search term.",
-    "es-ES": "Esto es un ejemplo sencillo de pantalla",
+    "es-ES": "En esta ventana, el usuario puede introducir el criterio de búsqueda",
     "de-DE": "Auf diesem Screen kann der Benutzer einen Suchbegriff eingeben."
-  },
-}
-\end{verbatim}
-\doublespacing
-
-\singlespacing
-\begin{verbatim}
-<rdf:Description rdf:about="http://www.morfeoproject.eu/fast/fco#ExampleScreen">
-  <rdf:type rdf:resource="http://www.morfeoproject.eu/fast/fco#Screen"/>
-  <rdfs:label xml:lang="en-GB">Simple Example</rdfs:label>
-  <rdfs:label xml:lang="es-ES">Ejemplo Sencillo</rdfs:label>
-  <rdfs:label xml:lang="de-DE">Einfaches Beispiel</rdfs:label>
-  <description
-    xmlns="http://purl.org/dc/terms/"
-    xml:lang="en-GB">This is a simple example of a screen</description>
-  <description
-    xmlns="http://purl.org/dc/terms/"
-    xml:lang="es-ES">Esto es un ejemplo sencillo de pantalla</description>
-  <description
-    xmlns="http://purl.org/dc/terms/"
-    xml:lang="de-DE">Dies ist ein einfaches Beispiel für einen Bildschirm</description>
+  }
+}
+\end{verbatim}
+\doublespacing
+
+\singlespacing
+\begin{verbatim}
+<rdf:RDF
+    xmlns:rdf="http://www.w3.org/1999/02/22-rdf-syntax-ns#"
+    xmlns:foaf="http://xmlns.com/foaf/0.1/"
+    xmlns:dc="http://purl.org/dc/terms/"
+    xmlns:fgo="http://purl.oclc.org/fast/ontology/gadget#"
+    xmlns:xsd="http://www.w3.org/2001/XMLSchema#"
+    xmlns:rdfs="http://www.w3.org/2000/01/rdf-schema#"
+    xmlns:ctag="http://commontag.org/ns#">
+  <rdf:Description rdf:about="http://demo.fast.morfeo-project.org/FASTCatalogue/screens/1">
+    <rdf:type rdf:resource="http://purl.oclc.org/fast/ontology/gadget#Screen"/>
+    <rdfs:label xml:lang="en-gb">Product Search</rdfs:label>
+    <rdfs:label xml:lang="en-es">Búsqueda de productos</rdfs:label>
+    <rdfs:label xml:lang="en-de">Produktsuche</rdfs:label>
+    <dc:description xml:lang="en-gb">On this screen, the user can enter a search
+    term.</dc:description>
+    <dc:description xml:lang="en-gb">En esta ventana, el usuario puede introducir
+    el criterio de búsqueda.</dc:description>
+    <dc:description xml:lang="en-gb">Auf diesem Screen kann der Benutzer einen
+    Suchbegriff eingeben.</dc:description>
   </rdf:Description>
-\end{verbatim}
-\doublespacing
-
-Until now, using internationalisation is allowed for the label and the description of a screen, but in the near future, more attributes will benefit from this feature.
+</rdf:RDF>
+\end{verbatim}
+\doublespacing
+
+Internationalisation is being used by the attributes \emph{label} and \emph{description} of any building block, \emph{label} of the pre/postconditions and \emph{label} of the tags.
 

### 2010-02-17 trunk/deliverables/D5_2_2/fast_d5_2_2.tex
final version of 5.2.2
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@565 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-01-26 trunk/deliverables/D5_1_2/fast_d5_1_2.tex
@@
 \usepackage{amssymb}
+
+% This helps to get hyphenation in typewriter font (\texttt)
+\usepackage[htt]{hyphenat}
+
 \parindent0pt
 
-\newcommand\deliverableNumber{D5.1.2}
-\newcommand\deliverableTitle{User Manual of the FAST Catalogue}
-\newcommand\deliverableTitleShort{Catalogue User Manual}
+\newcommand\deliverableNumber{D5.2.2}
+\newcommand\deliverableTitle{Prototype Semantic Catalogue for Screen-Flow Gadgets and Back-End Services}
+\newcommand\deliverableTitleShort{FAST Catalogue}
 \newcommand\workpackageNumber{5}
@@
 		\vfill
-		{\large \sffamily Date: 25/01/2010}\\
+		{\large \sffamily Date: 19/02/2010}\\
 		\vspace{1cm}
@@
                          \textcolor{white}{\textbf{Change description}}\\ \hline
-1.0 & 25/01/2010 & Ismael Rivera (NUIG) & First version of the deliverable \\ \hline
+1.0 & 20/02/2009 & Ismael Rivera (NUIG) & First version of the deliverable \\ \hline
+2.0 & 19/02/2010 & Ismael Rivera (NUIG) & Second version of the deliverable \\ \hline
 \end{tabular}
@@
 
-The present deliverable is intented to be the developer's manual of the prototype of the semantic catalogue.
-That said, whoever developing a component which will interact with the catalogue will find in this document a description of its architecture, the functionality provided together with the API (Application Programming Interface), data interchange formats, code errors and exceptions as well as several examples of usage.
+The present deliverable is an appendix to the software deliverable of the semantic catalogue prototype.
+It includes references to other deliverables where the reader can find further information about the catalogue.
 
@@
 \begin{tabular}
-	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
+	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.2cm}|p{1.6cm}|p{3.37cm}|}
 	\hline
 	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {27/02/2010} & {\textbf{Submitted}} & {27/02/2010}\\ \hline
-	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{Status} & \multicolumn{2}{l|}{final}\\ \hline
+	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
 	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
-	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{This deliverable is the technical documentation for the prototype developed as part of the WP5.
-It describes the catalogue's architecture, data interchange formats, APIs and examples of how to interact with it.}\\ \hline
+	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{This deliverable is the software prototype of the semantic catalogue.
+This document is just an appendix to it.}\\ \hline
 	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
-%	{} & \multicolumn{4}{l|}{\authorOne}\\ 
+%	{} & \multicolumn{4}{l|}{\authorTwo}\\ 
 %	{} & \multicolumn{4}{l|}{\authorOne}\\ 
@@
   \hline
-	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{NUIG} & \textbf{email} & {ismael.rivera@deri.org} \\ \cline{4-5}
+	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {ismael.rivera@deri.org} \\ \cline{4-5}
 	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495338} \\ \hline
@@
 \label{sec:introduction}
-This section starts establishing the goal and scope of the present document, shows how it is structured and details the relation to others documents and work packages.
+This section starts establishing the goal and scope of the present deliverable, and shows references to other deliverables for further information.
 
@@
 
-This is an introductory manual for developers who want to adopt and use the FAST semantic catalogue.
-It explains the main functionalities implemented so far, an overview of its architecture and a detailed \emph{Application Programming Interface} or API of the complete set of the operations offered through a REST service.
+The semantic catalogue will store and index screen-flow resources and back-end Web services, locate appropriate Web services for users, endow gadgets with functionality based on Service descriptions and resolve mediation problems between gadgets and their back-end services.
 
-% subsection goal_and_scope (end)
-
-\subsection{Structure of the document} % (fold)
-\label{sub:structure_of_the_document}
-
-The deliverable presents both the external and internal architecture in Section \ref{sec:architecture_overview}, then in Section \ref{sec:restful_catalogue_api} it is detailed the Catalogue API, query formats, interchange formats, error codes and so on.
+This document serves as an appendix to the semantic catalogue deliverable.
 
@@
 
-This should be the first version of this deliverable since in M12 the deliverable 5.1 was not scheduled.
-Nevertheless, last year an extra document was written as an appendix to the FAST Catalogue in order to detail and clarify technical decisions and to have a technical reference manual.
-Hence, the deliverable D5.2.1 will evolve to become the D5.1.2, so it will be no longer needed by itself.
-
-There are several new sections and major changes which need to be mentioned:
-
-\begin{itemize}
-	\item Section \ref{sec:installation_guide}.
-Installation Guide describes a set of steps to be followed in order to install and execute the FAST Catalogue.
-Also some configuration instructions are specified.
-	\item Section \ref{sub:content_negotiation}. Content Negotiation states how the FAST Catalogue has adopted several of the content negotiation principles.
-	\item Section \ref{sec:restful_catalogue_api}.
-RESTful Catalogue API has been intensively extended and modified in order to  reflect all the new functionality and how every request or response is constructed.
-Some of the most important features added in this iteration are:
-	\begin{itemize}
-		\item support for form elements, operators and back-end services (see Section ...),
-		\item convenience operations for recommendation and checking satisfaction and reachability for form elements, operators and back-end services (see Section ...),
-		\item and, last but not least, support for Screen planning (see Section \ref{sub:planning}).
-	\end{itemize}
-	\item Several appendixes have been added with the JSON structure of every building block the Catalogue can deal with.
-\end{itemize}
+This document is not intended to provide any further information regarding the semantic catalogue prototype.
+For information regarding where the code of the catalogue can be found, installation instructions, system requirements, a overview of the catalogue architecture and a detail API, please read \cite{palaguita2010_usermanual}.
 
@@
 \clearpage
-\section{Installation Guide} % (fold)
-\label{sec:installation_guide}
-
-This section  provides instructions on how to manually install and configure the FAST Catalogue.
-The first part of this guide presents some requirements to be considered before the installation, then it gives broad general instructions, and the last part contains a more detailed installation notes for specific configurations.
-
-\subsection{System Requirements} % (fold)
-\label{sub:system_requirements}
-
-In addition to the software itself, a standard FAST Catalogue installation has the following requirements:
-\begin{itemize}
-  \item Java\footnote{http://java.sun.com/} (version 6 or later) is required to run the software.
-  \item A Java Servlet Container that supports Java Servlet API 2.4 and Java Server Pages (JSP) 2.0, or newer.
-We recommend using a recent, stable version of Apache Tomcat\footnote{http://tomcat.apache.org/}.
-At the time of writing, this is either version 5.5.x or 6.x.
-\end{itemize}
-
-In addition, there are various optional dependencies which are required if you want to use certain advanced features (see Section \ref{sub:installation_instructions}).
-
-% subsection system_requirements (end)
-
-\subsection{Obtaining the FAST catalogue} % (fold)
-\label{sub:obtaining_the_fast_catalogue}
-
-It is recommended to have a Subversion client installed before you download the code (although you can theoretically download files without Subversion, this would mean tediously downloading each individual file manually).
-The recommended software is the official Subversion client, available from the Subversion project page\footnote{http://subversion.tigris.org/}.
-Note that this client uses a command-line interface, which the instructions below use.
-Alternatively, you can get subversioning software with a graphical user interface such as TortoiseSVN.
-
-To download from the latest release (recommended), enter the following command from the command-line in the directory you wish to download to:
-\begin{verbatim}
-svn checkout https://svn.forge.morfeo-project.org/fast-fp7project/trunk/catalogue_service
-\end{verbatim}
-
-% subsection obtaining_the_fast_catalogue (end)
-
-\subsection{Installation Instructions} % (fold)
-\label{sub:installation_instructions}
-
-Once you have got the source code, first step should be addressed is its compilation.
-You may manually compile the source yourself, however the FAST Catalogue comes along with a script to facilitate this task.
-This script has been made using a Java build tool called Ant \footnote{http://ant.apache.org/}.
-Hence, we recommended to get and install such a tool in order to follow the instructions below.
-
-Now you can compile, package and run the application via:
-\begin{verbatim}
-ant clean
-ant compile
-ant dist
-\end{verbatim}
-
-Or using a task which gather the previous three tasks:
-\begin{verbatim}
-ant main
-\end{verbatim}
-
-After executing with no errors this task, a WAR file should have been created in the /dist directory.
-A WAR file, which stands for  Web Application Archive, is just a JAR file used to distribute a collection of JavaServer Pages, servlets, Java classes, XML files, tag libraries and static Web pages (HTML and related files) that together constitute a Web application, in this case, the FAST Catalogue.
-
-Then, last step is to deply the WAR file into the servlet container. If the server chosen was Apache Tomcat, the procedure for deploying a Web application is:
-\begin{enumerate}
-	\item Stop Tomcat.
-	\item Delete existing deployment.
-If you have previously deployed "foo.war" in TOMCAT\_HOME/webapps, then it has been unpacked into webapps/foo/...
-You must delete this directory and all its contents.
-On Unix, this can be done with
-rm -r \$TOMCAT\_HOME/webapps/foo
-	\item Copy WAR file to TOMCAT\_HOME/webapps/.
-	\item Start Tomcat.
-\end{enumerate}
-
-Now you have successfully built the FAST catalogue, however we highly recommend you to read Section \ref{sub:configuration} and tune your FAST Catalogue instance.
-
-% subsection installation_instructions (end)
-
-\subsection{Configuration} % (fold)
-\label{sub:configuration}
-
-The following section covers the configuration options you may have to set up before start using the Catalogue.
-The FAST Catalogue relies on a Sesame repository for the persistent storage.
-Sesame repositories can be local or remote.
-Remote does not mean, it has to be installed in a different machine, but in a servlet container such as Apache Tomcat.
-The advantage of remote repositories is the possibility of accessing it via a SPARQL endpoint.
-These configuration aspect can be found in the file \emph{repository.properties} in the root directory where the Catalogue has been deployed.
-
-Independently the type of repository, the parameter \emph{serverURL} needs to take the value of the URL where the FAST Catalogue instance has been deployed.
-If you are testing the application in your machine, it may be something like \url{http://localhost:8080/FASTCatalogue}.
-
-Local repositories just need the parameter \emph{storageDir} pointing at a local directory, which has write permission in order to allow Sesame to create files in that directory or to delete any file in the directory.
-For instance:
-
-\begin{verbatim}
-storageDir=D:\\FAST\\catalogue\\repository
-\end{verbatim}
-
-Nevertheless, setting up a remote repository requires a few more steps.
-First the OpenRDF Sesame Server needs to be installed.
-The installation guide of this server is detailed in the Chapter 5 and 6 of the User Guide for Sesame 2.2 \cite{sesame2.2}.
-Once the Sesame server is deployed, and a repository has been created, the parameters sesameServer and repositoryID will take the values of the Sesame server URL and the repository identifier.
-For instance:
-
-\begin{verbatim}
-sesameServer=http://localhost:8880/openrdf-sesame
-repositoryID=c2
-\end{verbatim}
-
-% subsection installation_instructions (end)
-
-\subsection{Hello World!} % (fold)
-\label{sub:hello_world}
-
-It is recommended to check if the FAST Catalogue has been successfully installed.
-To do that, a simple HTTP request, using a Web browser, can be sent to the URL the Catalogue is deployed, and specifying a type of resource such as Screen.
-The URL would be similar to \url{http://localhost:8080/FASTCatalogue/screens}.
-Depending on the variant chosen, the response may be different.
-For JSON, the response should be \verb|[]|, but in HTML the message \verb|No screens found.| should be shown.
-
-% subsection hello_world (end)
-
-% section installation_guide (end)
-
-\clearpage
-\section{Architecture Overview} % (fold)
-\label{sec:architecture_overview}
-
-This section presents a technical overview of the internal catalogue's architecture and which external components will interact with it.
-Basically three layers can be distinguished: Presentation, Business Logic and Persistence layer (see Figure \ref{fig:architecture}).
-
-\begin{figure}[htb]
-\label{fig:architecture}
-\begin{center}
-	\includegraphics[width=12cm]{images/catalogue_architecture}
-	\caption{Catalogue's Architecture}
-	\label{fig:CatalogueInternalArchitecture}
-\end{center}
-\end{figure}
-
-The Presentation layer will be the public interface of the catalogue.
-The main purpose of this layer is to provide its functionality to other components within the Gadget Visual Studio.
-The service is offered in a REST (REpresentational State Transfer) style, making it easy to consume than other complex APIs.
-REST is an architectural style, not a toolkit or a standard.
-Even though, it makes use of standards like HTTP, URI, Resource Representations (XML, HTML, JSON, JPEG, etc.) or MIME types (text/xml, text/html, application/json, image/jpeg, etc.).
-Another characteristic of the REST style is its stateless assumption.
-The catalogue adopts this strategy; therefore every request needs a complete set of information in order to prepare the response.
-
-As part of the Presentation layer, a SPARQL endpoint is offered using the SPARQL protocol service as defined in the SPROT \cite{sprot} specification.
-The SPARQL endpoint is mostly offered to enable other developers to query directly the catalogue knowledge base using SPARQL queries.
-This feature is supported by the Sesame RESTful HTTP interface for SPARQL Protocol for RDF.
-
-The Business Logic layer contains all the FAST domain-specific processing.
-It provides functions to interact with all the elements of the domain model specified in \cite{moeller2009fast_ontology}.
-It processes the input from the Presentation layer, creating specific objects modelling the business logic, and vice versa.
-In addition it interacts with the Persistence layer in order to persist the model.
-
-The Persistence layer provides an API which allows the Business layer to work with a standard set of objects that read and save their state to the triple store.
-For this reason, on top of this layer it is used RDF2Go, an abstraction over triple (and quad) stores.
-The RDF2Go API allows interacting with the semantic representation of the model (triples) in a generic manner, and also brings the flexibility of choosing different triple stores to persist them.
-The selected implementation for the triple stored is Sesame 2.
-Moreover, this framework is completely extensible and configurable regarding to storage mechanisms, inferencers, RDF file formats, query result formats and query languages.
-For the actual catalogue prototype, the storage mechanism used is the native file storage system of Sesame, the inferencer used is a subset of the RDFS entailment rules \cite{rdfsrules} following a forward-chaining policy.
-
-% section architecture_overview (end)
-
-\clearpage
-\section{RESTful Catalogue API} % (fold)
-\label{sec:restful_catalogue_api}
-
-This section provides a high-level overview of the Catalogue API.
-It describes the calls or operations supported, specific parameters and responses for each operation, supported interchange formats and some examples to facilitate the understanding of the API.
-
-\subsection{JSON Interchange Format} % (fold)
-\label{sec:json_interchange_format}
-
-The format supported, and in which they must be constructed, by every HTTP request is JSON\footnote{http://www.json.org/}.
-JSON is a lightweight data interchange format whose simplicity has resulted in widespread use among web developers, easy to read and write and able to using any programming language because its structures map directly to data structures used in most programming languages.
-
-Every HTTP request should be encoded using the MIME type \texttt{application/json} and the charset \texttt{UTF-8}.
-
-The following is an example of a Find \& Check request:
-
-\begin{verbatim}
-{
-  "canvas": [
-    { "uri": "http://localhost:8080/catalogue/screens/238" },
-    { "uri": "http://localhost:8080/catalogue/screens/323" }
-  ],
-  "elements": [
-    { "uri": "http://localhost:8080/catalogue/screens/636" }
-  ],
-  "domainContext": {
-    "tags": [
-      {
-        "label": { "en-GB": "Amazon" },
-        "means": "http://dbpedia.org/page/Amazon.com"
-      }
-    ],
-    "user": "irivera",
-  },
-  "criterion": "reachability"
-}
-\end{verbatim}
-
-\subsubsection{Internationalisation I18n} % (fold)
-
-In order to offers an adaptable solution to various languages and regions without major engineering changes, internationalisation is considered from an early stage in the catalogue development.
-Underlying representation technologies used to develop the catalogue (RDF/XML and JSON) allow to implement this feature.
-This feature is implemented by the addition of a language tag to every 'string' desired.
-The specification of this language tag is composed by the language code and the country code, following the ISO 639\footnote{http://ftp.ics.uci.edu/pub/ietf/http/related/iso639.txt} for languages and the ISO 3166\footnote{http://userpage.chemie.fu-berlin.de/diverse/doc/ISO\_3166.html} for countries.
-The following example illustrates how to use it properly in both formats.
-
-\begin{verbatim}
-{
-  "label": {
-    "en-GB": "Simple Example",
-    "es-ES": "Ejemplo Sencillo",
-    "de-DE": "Einfaches Beispiel"
-  },
-  "description": {
-    "en-GB": "This is a simple example of a screen",
-    "es-ES": "Esto es un ejemplo sencillo de pantalla",
-    "de-DE": "Dies ist ein einfaches Beispiel für einen Bildschirm"
-  },
-}
-\end{verbatim}
-
-\begin{verbatim}
-<rdf:Description rdf:about="http://www.morfeoproject.eu/fast/fco#ExampleScreen">
-  <rdf:type rdf:resource="http://www.morfeoproject.eu/fast/fco#Screen"/>
-  <rdfs:label xml:lang="en-GB">Simple Example</rdfs:label>
-  <rdfs:label xml:lang="es-ES">Ejemplo Sencillo</rdfs:label>
-  <rdfs:label xml:lang="de-DE">Einfaches Beispiel</rdfs:label>
-  <description
-    xmlns="http://purl.org/dc/terms/"
-    xml:lang="en-GB">This is a simple example of a screen</description>
-  <description
-    xmlns="http://purl.org/dc/terms/"
-    xml:lang="es-ES">Esto es un ejemplo sencillo de pantalla</description>
-  <description
-    xmlns="http://purl.org/dc/terms/"
-    xml:lang="de-DE">Dies ist ein einfaches Beispiel für einen Bildschirm</description>
-</rdf:Description>
-\end{verbatim}
-
-Until now, using internationalisation is allowed for the label and the description of a screen, but in the near future, more attributes will benefit from this feature.
-
-% subsection json_interchange_format (end)
-
-\subsection{Content Negotiation} % (fold)
-\label{sub:content_negotiation}
-
-Before any major details about how content negotiation is addressed by the catalogue, a few concepts need to be understood.
-First of all, say a URI is not a file name.
-It is common to see a URI as the location for a file, but a URI or Universal Resource Identifier is not a URL: it is not a file name or location, it is an identifier (or a reference) to a resource.
-Then by dereferencing a URI the actual resource can be retrieved.
-Let's see a request to retrieve the resource "Screen 123" is sent to the server.
-The resource is just a piece of information, however the server will answer an HTML document with a text describing the screen or an image representing it.
-On the Web, these equivalent representations of a resource are called variants, and content negotiation is the mechanism used to determine which of the representations of most appropiate for a given request.
-
-\begin{figure}[htb]
-\label{fig:ContentNegotiation}
-\begin{center}
-	\includegraphics[width=10cm]{images/content_negotiation}
-	\caption{Content Negotiation}
-\end{center}
-\end{figure}
-
-In summary, the basic idea of content negotiation, as stated in the HTTP 1.1 specification \cite{http1.1}, is to serve the best variant for a resource, taking into account what variants are available, what variants the server may prefer to serve, what the client can accept, and with which preferences: in HTTP, this is done by the client which may send, in its request, Accept headers (Accept, Accept-Language and Accept-Encoding), to communicate its capabilities and preferences in Format, Language and Encoding, respectively.
-
-In fact, what the catalogue really does is "format negotiation" since the alternate representationos are just based on the selection of the media type, through the Accept header, but does not consider different languages or encoding types.
-The formats supported are JSON, RDF+XML and HTML.
-
-Lastly, content negotiation needs to identify which player is going to take the lead on it.
-There are two kinds of content negotiation which are possible in HTTP: server-driven and agent-driven negotiation.
-The approach followed by the catalogue is agent-driven negotiation, hence selecting a specific representation for a resource is responsibility of the user agent.
-If none is specified, by default, the server will choose the JSON representation.
-
-% subsection content_negotiation (end)
-
-\subsection{API Calls} % (fold)
-\label{sub:api_calls}
-
-This section contains detailed descriptions of the interface provided by the catalogue via REST services, detailing request parameters, response elements, any special errors and examples of requests and responses.
-The URL format is also specified for each operation, where 'catalogueURL' has to be replaced for the real URL the service is installed (e.g.
-http://demo.fast.morfeo-project.org/catalogue).
-
-\subsubsection{CRUD operations} % (fold)
-
-Any building block within the Catalogue is a resource in terms of REST philosophy, hence it can be created, retrieved, modified or deleted, using a certaing URL and a specific HTTP method for every operation.
-Two concepts have to be defined: \emph{collection}, as a set of resources which access URI is http://catalogueURL/<type>/ where catalogueURL is the specific URL where the Catalogue server is installed and <type> is the plural of the name of the building block (e.g.
-screens, services), and \emph{member} of the collection, in other words, the building block itself, which access URI is http://catalogueURL/<type>/<id> where the <id> has to be replaced by the identifier of a specific building block.
-The details of the operations and which HTTP verb has to be used can be found in the Table \ref{tab:crud_operations}.
-
-\begin{table}[htb]
-\caption{CRUD operations}
-\label{tab:crud_operations}
-\begin{center}
-\begin{tabular}{|p{2.5cm}|p{2.5cm}|p{3.5cm}|p{5cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Resource} &
-                         \textcolor{white}{HTTP method} &
-                         \textcolor{white}{HTTP body} &
-                         \textcolor{white}{Description}\\ \hline
-Collection URI & GET & N/A & \textbf{List} the members of the collection.\\ \hline
-Collection URI & PUT & N/A & Not used.\\ \hline
-Collection URI & POST & JSON representation of the building block & \textbf{Create} a new entry in the collection where the URI is assigned automatically by the collection.
-The URI created is returned by this operation.\\ \hline
-Collection URI & DELETE & N/A & Not used.\\ \hline
-Member URI & GET & N/A & \textbf{Retrieve} the addressed member of the collection.\\ \hline
-Member URI & PUT & JSON representation of the building block & \textbf{Update} the addressed member of the collection or create it with a defined URI.\\ \hline
-Member URI & POST & N/A & Not used. \\ \hline
-Member URI & DELETE & N/A & \textbf{Delete} the addressed member of the collection.\\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-The JSON structure for create any building bock can be found in the appendixes.
-All of them are composed of a common structure which can be found in Section \ref{appendix_b}, and some specific information depending on the type.
-Screen-flows are defined in \ref{appendix_c}, screens in \ref{appendix_d} and forms, operators and back-end services share the same structure detailed in \ref{appendix_e}.
-However, a few examples of request and responses are shown in order to clarify how the request are constructed and sent, and how the responses look like.
-
-To create a new screen, a POST request is send to the catalogue server, including the JSON representation of the screen as the body of the request:
-
-\begin{verbatim}
-{
-  "code": "http://demo.fast.morfeo-project.org/code/amazonList.js",
-  "creationDate": "2010-01-26T17:01:13+0000",
-  "creator": "irivera",
-  "description": {"en-gb": "Please fill the description..."},
-  "homepage": "http://fast.morfeo-project.eu/",
-  "icon": "http://demo.fast.morfeo-project.org/images/amazonList.png",
-  "id": "28",
-  "label": {"en-gb": "Amazon Product List"},
-  "name": "ProductList",
-  "postconditions": [
-    [
-      {
-        "id": "item",
-        "label": { "en-gb": "An item" },
-        "pattern": "?I 
-                    http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                    http://aws.amazon.com/AWSECommerceService#Item",
-        "positive": "true"
-      }
-    ]
-  ],
-  "preconditions": [
-    [
-      {
-        "id": "filter",
-        "label": { "en-gb": "A search criteria" },
-        "pattern": "?F
-                    http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                    http://aws.amazon.com/AWSECommerceService#SearchCriteria",
-        "positive": "true"
-      }
-    ]
-  ],
-  "rights": "http://creativecommons.org/",
-  "screenshot": "http://demo.fast.morfeo-project.org/images/amazonProductList.png",
-  "tags": [{"label": {"en-gb": "amazon"}}],
-  "version": "0.1"
-}
-\end{verbatim}
-
-The URI of the screen is created using the id specified in the JSON request. So, the response for this operation is:
-
-\begin{verbatim}
-{
-  "code": "http://demo.fast.morfeo-project.org/code/amazonList.js",
-  "creationDate": "2010-01-26T17:01:13+0000",
-  "creator": "irivera",
-  "description": {"en-gb": "Please fill the description..."},
-  "homepage": "http://fast.morfeo-project.eu/",
-  "icon": "http://demo.fast.morfeo-project.org/images/amazonList.png",
-  "id": "28",
-  "label": {"en-gb": "Prueba1"},
-  "name": "ProductList",
-  "postconditions": [
-    [
-      {
-        "id": "item",
-        "label": { "en-gb": "An item" },
-        "pattern": "?I
-                    http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                    http://aws.amazon.com/AWSECommerceService#Item",
-        "positive": "true"
-      }
-    ]
-  ],
-  "preconditions": [
-    [
-      {
-        "id": "filter",
-        "label": { "en-gb": "A search criteria" },
-        "pattern": "?F
-                    http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                    http://aws.amazon.com/AWSECommerceService#SearchCriteria",
-        "positive": "true"
-      }
-    ]
-  ],
-  "rights": "http://creativecommons.org/",
-  "screenshot": "http://demo.fast.morfeo-project.org/images/amazonProductList.png",
-  "tags": [{"label": {"en-gb": "amazon"}}],
-  "type": "screen",
-  "uri": "http://localhost:8080/catalogue/screens/28",
-  "version": "0.1"
-}
-\end{verbatim}
-
-To obtain all the screens stored in the catalogue a HTTP GET request is sent to the Collection URI and the response may be something like this:
-\begin{verbatim}
-[
-  {
-    "code": "http://demo.fast.morfeo-project.org/code/amazonList.js",
-    "creationDate": "2010-01-26T17:01:13+0000",
-    "creator": "irivera",
-    "description": {"en-gb": "Please fill the description..."},
-    "homepage": "http://fast.morfeo-project.eu/",
-    "icon": "http://demo.fast.morfeo-project.org/images/amazonList.png",
-    "id": "28",
-    "label": {"en-gb": "Prueba1"},
-    "name": "ProductList",
-    "postconditions": [
-      [
-        {
-          "id": "item",
-          "label": { "en-gb": "An item" },
-          "pattern": "?I
-                      http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                      http://aws.amazon.com/AWSECommerceService#Item",
-          "positive": "true"
-        }
-      ]
-    ],
-    "preconditions": [
-      [
-        {
-          "id": "filter",
-          "label": { "en-gb": "A search criteria" },
-          "pattern": "?F
-                      http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                      http://aws.amazon.com/AWSECommerceService#SearchCriteria",
-          "positive": "true"
-        }
-      ]
-    ],
-    "rights": "http://creativecommons.org/",
-    "screenshot": "http://demo.fast.morfeo-project.org/images/amazonProductList.png",
-    "tags": [{"label": {"en-gb": "amazon"}}],
-    "type": "screen",
-    "uri": "http://localhost:8080/catalogue/screens/28",
-    "version": "0.1"
-  },
-  {
-    "creationDate": "2009-02-07T09:59:52+0000",
-    "creator": "fabio",
-    ...,
-	  "uri": "http://localhost:8080/catalogue/screens/14",
-    "version": "1.0"
-  },
-  {
-    "creationDate": "2009-02-07T09:59:52+0000",
-    "creator": "javier",
-    ...,
-	  "uri": "http://localhost:8080/catalogue/screens/564",
-    "version": "1.0"
-  }
-]
-\end{verbatim}
-
-\subsubsection{Screen Find} % (fold)
-\label{ssub:screen_find}
-
-The find operation searches inside the catalogue for any screen which could be somehow related to the gadget the user is creating.
-It provides a recommended set of screens depending on the domain context, the canvas, and so on.
-
-The specific URL to access this operation is http://catalogueURL/screens/find using HTTP POST method.
-From now on, the method 'find' only considers the domain context and the pre/postconditions of the screens.
-It will try to satisfy all the unsatisfied preconditions of a given list of screens also known as canvas.
-The results will be all the screens stored in the catalogue which fulfil some of these preconditions, and are tagged with the tags specified in the domain context.
-The request parameters are shown in Table \ref{tab:screen_find_request} and the response parameters are shown in Table \ref{tab:screen_find_response}.
-
-\begin{table}[htb]
-\caption{Screen Find Request Parameters}
-\label{tab:screen_find_request}
-\begin{center}
-\begin{tabular}{|p{2cm}|p{9.5cm}|p{2cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
-                         \textcolor{white}{Description} &
-                         \textcolor{white}{Type}\\ \hline
-canvas & The canvas is composed by a list of screens. Only the URI is needed. & Optional \\ \hline
-domainContext & The domain context contains a list of tags and a user. & Optional \\ \hline
-elements & It is a list of resources, previously stored in the catalogue. For example, the list of screens recommended last execution. Only accepts screens. & Optional \\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{table}[htb]
-\caption{Screen Find Response Parameters}
-\label{tab:screen_find_response}
-\begin{center}
-\begin{tabular}{|p{2cm}|p{9.5cm}|p{2cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
-                         \textcolor{white}{Description} &
-                         \textcolor{white}{Type}\\ \hline
-N/A & A list of recommended screens URIs is returned. & Required \\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-Following is shown an example of usage of this operation considering a canvas with the screen \texttt{http://localhost:8080/catalogue/screens/654}.
-
-\begin{verbatim}
-{
-  canvas: [
-    { uri: "http://localhost:8080/catalogue/screens/654" }
-  ],
-  domainContext: {
-    tags: [],
-	  user: null
-  },
-  elements: []
-}
-\end{verbatim}
-
-After the execution of the recommendation algorithm, the response given by the catalogue is:
-
-\begin{verbatim}
-[
-  "http://localhost:8080/catalogue/screens/371",
-  "http://localhost:8080/catalogue/screens/24",
-  "http://localhost:8080/catalogue/screens/253",
-  "http://localhost:8080/catalogue/screens/12"
-]
-\end{verbatim}
-
-\subsubsection{Screen Check} % (fold)
-\label{ssub:screen_check}
-
-This operation verifies the state of certain list of screens depending on a specific criterion.
-The only criterion accepted in this stage of the prototype is 'reachability'.
-A screen will be reachable if it all its preconditions are satisfied by postconditions of the screens the canvas contain.
-The specific URL to access this operation is http://catalogueURL/screens/check using HTTP POST method.
-
-\begin{table}[htb]
-\caption{Screen Check Request Parameters}
-\label{tab:screen_check_request}
-\begin{center}
-\begin{tabular}{|p{2cm}|p{9.5cm}|p{2cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
-                         \textcolor{white}{Description} &
-                         \textcolor{white}{Type}\\ \hline
-canvas & The canvas is composed by a list of screens. Only the URI is needed. & Optional\\ \hline
-domainContext & The domain context contains a list of tags and a user. & Optional\\ \hline
-elements & It is a list of resources, previously stored in the catalogue. For example, the list of screens recommended last execution. Only accepts screens. & Optional\\ \hline
-criterion & The criterion specifies what it has to be check. The only possible supported now is 'reachability' in order to check if the preconditions of a screen are satisfied. & Required\\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{table}[htb]
-\caption{Screen Check Response Parameters}
-\label{tab:screen_check_response}
-\begin{center}
-\begin{tabular}{|p{2cm}|p{9.5cm}|p{2cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
-                         \textcolor{white}{Description} &
-                         \textcolor{white}{Type}\\ \hline
-Canvas & The canvas indicating which screens satisfy the critetion.
-More aditional information may be included, for example, a satisfaction attribute for every precondition for reachability.
-& Optional\\ \hline
-Domain Context & The domain context contains a list of tags and a user. & Optional\\ \hline
-Elements & The elements indicating which satisfy the critetion. More aditional information may be included as in the Canvas. & Optional\\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-The scenario for the following example is composed by a screen in the canvas which precondition is a foaf:Person which foaf:workplaceHomepage is \url{http://www.deri.ie/}, and a screen in the elements list which precondition is a foaf:Person.
-The state to check in this case is the reachability and satisfaction of the screens and preconditions.
-
-\begin{verbatim}
-{
-  "canvas": [
-    { "uri": "http://localhost:8080/catalogue/screens/238" }
-  ],
-  "elements": [
-    { "uri": "http://localhost:8080/catalogue/screens/636" }
-  ],
-  "domainContext": {
-    "tags": [
-      {
-        "label": { "en-GB": "Amazon" },
-        "means": "http://dbpedia.org/page/Amazon.com"
-      }
-    ],
-    "user": "irivera",
-  },
-  criterion: "reachability"
-}
-\end{verbatim}
-
-The response shows that there are not reachable screens at the moment, and preconditions are not satisfied.
-
-\begin{verbatim}
-{
-  "canvas": [{
-    "preconditions": [{
-      "label": {"en-gb": "A person working in DERI"},
-      "pattern": "?person rdf:type foaf:Person . 
-                  ?person foaf:workplaceHomepage http://www.deri.ie/",
-      "positive": true,
-      "satisfied": false
-    }],
-    "reachability": false,
-    "uri": "http://localhost:8080/catalogue/screens/238"
-  }],
-  "elements": [{
-    "preconditions": [{
-      "label": {"en-gb": "A person"},
-      "pattern": "?person rdf:type foaf:Person",
-      "satisfied": false
-    }],
-    "reachability": false,
-    "uri": "http://localhost:8080/catalogue/screens/636"
-  }]
-}
-\end{verbatim}
-
-\subsubsection{Screen Find \& Check} % (fold)
-\label{ssub:screen_findcheck}
-
-For convenience and to minimise the number of request, the Find operation may be combined with the Check operation.
-This operation will look for new screens based on the given information, and will perform the Check to both the screens in the canvas and the screens in the \emph{elements} attribute.
-
-The format of the request is the same as the one used in the Check operation:
-
-\begin{verbatim}
-{
-  "canvas": [
-    { "uri": "http://localhost:8080/catalogue/screens/238" }
-  ],
-  "elements": [
-    { "uri": "http://localhost:8080/catalogue/screens/636" }
-  ],
-  "domainContext": {
-    "tags": [
-      {
-        "label": { "en-GB": "Amazon" },
-        "means": "http://dbpedia.org/page/Amazon.com"
-      }
-    ],
-    "user": "irivera",
-  },
-  criterion: "reachability"
-}
-\end{verbatim}
-
-The response for the above request is:
-
-\begin{verbatim}
-{
-  "canvas": [{
-    "preconditions": [{
-      "label": {"en-gb": "A person working in DERI"},
-      "pattern": "?person rdf:type foaf:Person . 
-                  ?person foaf:workplaceHomepage http://www.deri.ie/",
-      "positive": true,
-      "satisfied": false
-    }],
-    "reachability": false,
-    "uri": "http://localhost:8080/catalogue/screens/238"
-  }],
-  "elements": [
-    {
-      "preconditions": [{
-        "label": {"en-gb": "A person"},
-        "pattern": "?person rdf:type foaf:Person",
-        "satisfied": false
-      }],
-      "reachability": false,
-      "uri": "http://localhost:8080/catalogue/screens/636"
-    },
-    {
-      "preconditions": [],
-      "reachability": true,
-      "uri": "http://localhost:8080/catalogue/screens/132"
-    }
-  ]
-}
-\end{verbatim}
-
-\subsubsection{GetMetadata} % (fold)
-
-Although the information about a resource can be obtained by its specific retrieval operation supported by the CRUD interface, sometimes is needed to retrieve information which would imply a request per resource, hence this operation allow to get the metadata of a list of resources in only one request.
-This prototype supports this operation for screen-flows and screens.
-The specific URL to access this operation is http://catalogueURL/getmetadata using HTTP POST method.
-Table \ref{tab:getmetadata_request} shows the parameters needed for the invocation and Table \ref{tab:getmetadata_response} details the different parameters may contain a response.
-
-\begin{table}[htb]
-\caption{GetMetadata Request Parameters}
-\label{tab:getmetadata_request}
-\begin{center}
-\begin{tabular}{|p{2cm}|p{9.5cm}|p{2cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
-                         \textcolor{white}{Description} &
-                         \textcolor{white}{Type}\\ \hline
-N/A & A list of URIs. & Required\\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{table}[htb]
-\caption{GetMetadata Response Parameters}
-\label{tab:getmetadata_response}
-\begin{center}
-\begin{tabular}{|p{2cm}|p{9.5cm}|p{2cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
-                         \textcolor{white}{Description} &
-                         \textcolor{white}{Type}\\ \hline
-screen-flows & A set of screen-flows with all the metadata associated to them.. & Optional\\ \hline
-screens & A set of screens with all the metadata associated to them. & Optional\\ \hline
-forms & A set of forms with all the metadata associated to them. & Optional\\ \hline
-operators & A set of operators with all the metadata associated to them. & Optional\\ \hline
-backendservices & A set of back-end services with all the metadata associated to them. & Optional\\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-The following example send two URIs of screens and a URI of a screen-flow to the GetMetadata operation:
-\begin{verbatim}
-[
-  "http://localhost:8080/catalogue/screens/12",
-  "http://localhost:8080/catalogue/screens/48",
-  "http://localhost:8080/catalogue/forms/9",
-]
-\end{verbatim}
-
-The response obtained are two lists, one containing all the metadata regarding to the screen-flows and another one with the information about the screens:
-\begin{verbatim}
-{
-  "screenflows": [],
-  "screens": [
-    {
-      "creationDate": "2009-02-07T09:59:52+0000",
-      "creator": "irivera",
-      ...,
-      "uri": "http://localhost:8080/catalogue/screens/12",
-      "version": "1.0"
-    },
-    {
-      "creationDate": "2009-02-07T09:59:52+0000",
-      "creator": "irivera",
-      ...,
-      "uri": "http://localhost:8080/catalogue/screens/12",
-      "version": "1.0"
-    }
-  ],
-  "forms": [
-    {
-      "actions": [
-        {
-          "name": "init",
-          "preconditions": [],
-          "uses": []
-        },
-        {
-          "name": "showTable",
-          "preconditions": [{
-            "id": "list",
-            "label": {"en-gb": "A product list"},
-            "pattern": "?PList
-                        http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                        http://aws.amazon.com/AWSECommerceService#ProductList",
-            "positive": true
-          }],
-          "uses": []
-        }
-      ],
-      ...,
-      "uri": "http://localhost:8080/FASTCatalogue/forms/9",
-      "version": "1.0"
-    },
-  ],
-  "operators": [],
-  "backendservices": []
-}
-\end{verbatim}
-
-\subsubsection{Screen Component Find \& Check} % (fold)
-\label{ssub:screen_component_findcheck}
-
-Similarly to the Screen Find \& Check described in Section \ref{ssub:screen_findcheck}, this operation search for screen components (forms, operators and back-end services) through the catalogue in order to satisfy a given request, and it will attach information about satisfaction and reachability as well.
-
-A request is formed of the following parameters:
-
-\begin{description}
-	\item[preconditions] Preconditions of the screen.
-	\item[postconditions] Postcondition of the screen.
-	\item[canvas] The canvas is composed by a list of screens. Only the URI is needed.
-	\item[forms] It is a list of forms, such as the list of forms recommended last execution.
-	\item[operators] It is a list of operators, such as the list of operators recommended last execution.
-	\item[backendservices] It is a list of back-end services, such as the list of back-end services recommended last execution.
-	\item[domainContext] The domain context contains a list of tags and a user.
-	\item[pipes] List of pipes for the connection between screen components.
-	\item[selectedItem] URI of a screen component from the canvas.
-\end{description}
-
-The screen components recommended are based on the given domain context and the pre/postconditions of the any of the resources in the canvas or the conditions of the parameters preconditions and postconditions.
-Hence, any screen component from the catalogue which has any of these conditions may be suitable for the screen, so it is included in the response, in the list corresponding to the type of screen component.
-
-An example of request is the following:
-
-\begin{verbatim}
-{
-  "preconditions": [{
-    "id": "Searchcriteria_1",
-    "label": {"en-gb": "A search criteria"},
-    "pattern": "?F
-                http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                http://aws.amazon.com/AWSECommerceService#SearchCriteria",
-    "positive": true,
-  }],
-  "postconditions": [],
-  "canvas": [
-    { "uri": "http://localhost:8080/catalogue/forms/1" },
-    { "uri": "http://localhost:8080/catalogue/services/1" }
-  ],
-  "forms": [],
-  "operators": [],
-  "backendservices": [],
-  "domainContext": {
-    "tags": [],
-    "user": null
-  },
-  "pipes": [
-    {
-      "from": {
-        "buildingblock": "http://localhost:8080/catalogue/services/1",
-        "condition": "list"
-      },
-      "to": {
-        "action": "showTable",
-        "buildingblock": "http://localhost:8080/catalogue/forms/1",
-        "condition": "list"
-      }
-    }
-  ],
-  "selectedItem": "http://localhost:8080/catalogue/forms/1"
-}
-\end{verbatim}
-
-And the response according to the above request is:
-
-\begin{verbatim}
-{
-  "backendservices": [],
-  "canvas": [
-    {
-      "actions": [{
-        "name": "search",
-        "preconditions": [{
-          "id": "filter",
-          "label": {"en-gb": "A search criteria"},
-          "pattern": "?F 
-                      http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                      http://aws.amazon.com/AWSECommerceService#SearchCriteria",
-          "positive": true,
-          "satisfied": false
-        }],
-        "satisfied": false,
-        "uses": []
-      }],
-      "reachability": false,
-      "uri": "http://localhost:8080/catalogue/services/1"
-    },
-    {
-      "actions": [
-        {
-          "name": "init",
-          "preconditions": [],
-          "satisfied": true,
-          "uses": []
-        },
-        {
-          "name": "showTable",
-          "preconditions": [{
-            "id": "list",
-            "label": {"en-gb": "A product list"},
-            "pattern": "?PList
-                        http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                        http://aws.amazon.com/AWSECommerceService#ProductList",
-            "positive": true,
-            "satisfied": false
-          }],
-          "satisfied": false,
-          "uses": []
-        }
-      ],
-      "reachability": true,
-      "uri": "http://localhost:8080/catalogue/forms/1"
-    }
-  ],
-  "connections": [
-    {
-      "from": {
-        "buildingblock": null,
-        "condition": "Searchcriteria_1"
-      },
-      "to": {
-        "action": "search",
-        "buildingblock": "http://localhost:8080/catalogue/services/1",
-        "condition": "filter"
-      }
-    },
-  ],
-  "forms": [],
-  "operators": [],
-  "pipes": [
-    {
-      "from": {
-        "buildingblock": "http://localhost:8080/catalogue/services/1",
-        "condition": "list"
-      },
-      "satisfied": true,
-      "to": {
-        "action": "showTable",
-        "buildingblock": "http://localhost:8080/catalogue/forms/1",
-        "condition": "list"
-      }
-    }
-  ],
-  "postconditions": []
-}
-\end{verbatim}
-
-For a screen component to be reachable, its preconditions need to be connected by a pipe to any other screen component which is already reachable, or to any of the screen preconditions.
-Moreover, the pipe needs to be satisfied, in other words, conditions in both edges are compatible.
-
-The response includes an attribute called "connections". This attribute will be retrieved when "selectedItem" is given, and it will contain a list of potential pipes to connect to that item.
-
-% subsubsection screen_component_findcheck (end)
-
-% subsection api_calls (end)
-
-\subsection{Planning}
-\label{sub:planning}
-
-While creating a screen-flow, the user need to find proper screens which will satisfy her necessity.
-Once the user has selected one or several screens to be included in her screen-flow, she needs to find screens which satisfy the preconditions of these screens, in order to create an executable screen-flow.
-The operation Find, detailed in Section \ref{ssub:screen_find}, helps the user to find these screens, however the response is just a set of screens which somehow may help to accomplish her goal, but not in a single and straight-forward step.
-
-This functionality offers to the user the possibility to receive a whole set of screens, or plan, which will make a given screen, or goal, reachable.
-These plans take into account the screen which are already in the canvas, in order to minimise the number of screens.
-
-The specific URL to access this operation is \url{http://catalogueURL/planner}, and to execute the operation a POST request needs to be done. The request body is composed of:
-
-\begin{description}
-	\item[goal] Required. URI of the goal screen.
-	\item[canvas] Required. List of screens URIs.
-	\item[page] Optional. Number of the page to be retrieved.
-	\item[per\_page] Optional. Number of plans to be included per page in the response.
-\end{description}
-
-Let's see the following example:
-
-Request
-\begin{verbatim}
-{
-  "goal": "http://localhost:8080/catalogue/screens/35",
-  "canvas": [
-    { "uri": "http://localhost:8080/catalogue/screens/56" },
-    { "uri": "http://localhost:8080/catalogue/screens/85" }
-  ],
-  "page": 1,
-  "per_page": 10 
-}
-\end{verbatim}
-
-Response
-\begin{verbatim}
-[
-  [
-    "http://localhost:8080/catalogue/screens/13",
-    "http://localhost:8080/catalogue/screens/23"
-  ],
-  [
-    "http://localhost:8080/catalogue/screens/79",
-    "http://localhost:8080/catalogue/screens/37",
-    "http://localhost:8080/catalogue/screens/14"
-  ],
-  [
-    "http://localhost:8080/catalogue/screens/16",
-    "http://localhost:8080/catalogue/screens/44",
-    "http://localhost:8080/catalogue/screens/32"
-  ]
-]
-\end{verbatim}
-
-The response is ranked based on adding the minimum number of new screens into the canvas, and giving preference to the screens the user has previously selected (canvas).
-
-In the above example, let's assume the first plan is using the two screens from the canvas to accomplish the goal, and the second plan is able to make the goal reachable by itself.
-That said, a screen-flow with the goal and the second plan could be executable, and it would be composed of just four screens, however, taking the first plan, the resulting screen-flow will be composed of five screens.
-It could make sense to place the second plan in first position, but the plan with just two screens reuse the screens from the canvas which has been added by the user.
-
-% subsection planning (end)
-
-\subsection{API Error Codes} % (fold)
-There are two types of error codes, client and server.
-\begin{itemize}
-	\item Client error codes are generally caused by the client and might be an invalid domain or an invalid request parameter.
-These errors are accompanied by a 4xx HTTP response code.
-For example: ResourceNotFound.
-	\item Server error codes are generally caused by a server-side issue and should be reported. These errors are accompanied by a 5xx HTTP response code. For example: ServerUnavailable.
-\end{itemize}
-
-The following table lists all the error codes.
-
-\begin{table}[htb]
-\caption{API Error Codes}
-\begin{center}
-\begin{tabular}{|p{3.5cm}|p{7cm}|p{3cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Error} &
-                         \textcolor{white}{Description} &
-                         \textcolor{white}{HTTP Status Code}\\ \hline
-ResourceNotFound & The resource <resourceURI> has not been found. & 404 Not Found \\ \hline
-AccessFailure & Access to the resource <resourceName> is denied. & 403 Forbidden \\ \hline
-InternalError & Request could not be executed due to an internal service error. & 500 Internal Server Error \\ \hline
-InvalidAction & The action <actionName> is not valid for this web service.  & 404 Bad Request \\ \hline
-InvalidHttpRequest & The HTTP request is invalid. Reason: <reason>. & 400 Bad Request \\ \hline
-InvalidLiteral & Illegal literal in the filter expression. & 400 Bad Request \\ \hline
-InvalidParameterValue & The specified parameter value is not valid. & 400 Bad Request \\ \hline
-InvalidURI & The URI <requestURI> is not valid. & 400 Bad Request \\ \hline
-MissingAction & No action was supplied with this request. & 400 Bad Request \\ \hline
-MissingParameter & The request must contain the specified missing parameter. & 400 Bad Request \\ \hline
-NotYetImplemented & Feature <feature> is not yet available. & 401 Unauthorized \\ \hline
-ServiceUnavailable & The service is currently unavailable. Please try again later. &  \\ \hline
-UnsupportedHttpVerb & The requested HTTP verb is not supported: <verb>. & 400 Bad Request \\ \hline
-URITooLong & The URI exceeded the maximum limit of <maxLength>. & 400 Bad Request \\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\clearpage
 \bibliographystyle{apalike}
-\bibliography{fast_d5_1_2}
-
-\clearpage
-\doublespacing
-\section*{Appendix A. Lists of Tables and Figures}
-\label{appendix_a}
-\addcontentsline{toc}{section}{Appendix A. Lists of Tables and Figures}
-
-\listoftables
-
-\listoffigures
-
-\clearpage
-\section*{Appendix B. Generic Building Block JSON Structure}
-\label{appendix_b}
-\addcontentsline{toc}{section}{Appendix B. Generic Building Block JSON Structure}
-
-\begin{verbatim}
-{
-  "code": "http://url.com/.../code.js",
-  "creationDate": "2009-04-20T17:00:00+0100",
-  "creator": "http://www.fabio.es/",
-  "description": {"en-gb": "This is a description of the building block"},
-  "tags": {
-    "tags": [
-      "ebay",
-      "amazon",
-    ],
-    "user": <String> // not defined yet
-  },
-  "homepage": "http://www.homepage.ie/",
-  "icon": "http://url.com/images/icon.png",
-  "id": "Id of the resource",
-  "label": {"en-gb": "Label or title of the building block"},
-  "name": "Amazon Item to Ebay Filter",
-  "rights": "http://creativecommons.org/",
-  "screenshot": "http://www.deri.ie/eBayList-screenshot.jpg",
-  "type": "type of the building block", // screenflow, screen, form, operator y resource
-  "uri": "http://localhost:8080/catalogue/screens/525",
-  "version": "1.0"
-}
-\end{verbatim}
-
-\clearpage
-\section*{Appendix C. Screen-flow JSON Structure}
-\label{appendix_c}
-\addcontentsline{toc}{section}{Appendix C. Screen-flow JSON Structure}
-
-\begin{verbatim}
-{
-  <GENERIC BUILDING BLOCK JSON>,
-  "contains": [
-    "http://purl.oclc.org/fast/ontology/gadget#Screen1257860772723",
-    "http://purl.oclc.org/fast/ontology/gadget#Screen1257860720035",
-  ],
-}
-\end{verbatim}
-
-\clearpage
-\section*{Appendix D. Screen JSON Structure}
-\label{appendix_d}
-\addcontentsline{toc}{section}{Appendix D. Screen JSON Structure}
-
-\begin{verbatim}
-{
-  <GENERIC BUILDING BLOCK JSON>,
-  "postconditions": [[{
-    "id": <identifier:String>,
-    "label": {"en-gb": "Purchase URL"},
-    "pattern": "?P
-                http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                http://aws.amazon.com/AWSECommerceService#PurchaseURL",
-    "positive": true
-  }]],
-  "preconditions": [[{
-    "id": "cart",
-    "label": {"en-gb": "A shopping cart"},
-    "pattern": "?C
-                http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                http://aws.amazon.com/AWSECommerceService#ShoppingCart",
-    "positive": true
-  }]],
-  "code": "URL of the code",
-  "definition": {
-    "buildingblocks: [
-      { // Form
-        "id": "form1", // unique for the container screen
-        "uri": "http://purl.oclc.org/fast/ontology/gadget#Form670238"
-      },
-      { // Operators
-        "id": "op1",
-        "uri": "http://purl.oclc.org/fast/ontology/gadget#Operator213487"
-      },
-      {  // Backend services
-        "id": "bs1",
-        "uri": "http://purl.oclc.org/fast/ontology/gadget#Backendservice38399"
-      },
-      {...}
-    ]
-    "pipes": [
-      {
-        "from": {
-          "buildingblock": "bs1",
-          "condition": "cA"
-        },
-        "to": {
-          "buildingblock": "op1",
-          "condition": "cA", // can be the same id, unique for the building block
-          "action": "filter"
-        }
-      },
-    "triggers": [
-      {
-        "from": {
-          "buildingblock": "form1",
-          "name": "refresh"
-        },
-        "to": {
-          "buildingblock": "bs1",
-          "action": "list"
-        }
-      }
-    ]
-  }
-}
-\end{verbatim}
-
-\clearpage
-\section*{Appendix E. Screen Component JSON Structure}
-\label{appendix_e}
-\addcontentsline{toc}{section}{Appendix E. Screen Component JSON Structure}
-
-\begin{verbatim}
-{
-  <GENERIC BUILDING BLOCK JSON>,
-  "actions": [{
-    "name": "filter",
-    "preconditions": [[{
-      "id": "item",
-      "label": {"en-gb": "Ebay List"},
-      "pattern": "?Item
-                  http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                  http://aws.amazon.com/AWSECommerceService#Item",
-      "positive": true
-    }]],
-    "uses": []
-  }],
-  "libraries": [],
-  "postconditions": [[{
-    "id": "filterEbay",
-    "label": {"en-gb": "Ebay List"},
-    "pattern": "?eFilter
-                http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                http://developer.ebay.com/.../FindItemsAdvanced.html#Request",
-    "positive": true
-  }]],
-  "triggers": ["itemAmazon"]
-}
-\end{verbatim}
+\bibliography{fast_d5_2_2}
 

### 2010-02-17 trunk/deliverables/D5_2_2/fast_d5_2_2.tex
minor changes
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@566 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \label{sec:introduction}
-This section starts establishing the goal and scope of the present deliverable, and shows references to other deliverables for further information.
 
-\subsection{Goal and Scope} % (fold)
-\label{sub:goal_and_scope}
-
-The semantic catalogue will store and index screen-flow resources and back-end Web services, locate appropriate Web services for users, endow gadgets with functionality based on Service descriptions and resolve mediation problems between gadgets and their back-end services.
-
-This document serves as an appendix to the semantic catalogue deliverable.
-
-% subsection goal_and_scope (end)
-
-\subsection{Changes from previous version} % (fold)
-\label{sub:changes_from_previous_version}
-
-This document is not intended to provide any further information regarding the semantic catalogue prototype.
-For information regarding where the code of the catalogue can be found, installation instructions, system requirements, a overview of the catalogue architecture and a detail API, please read \cite{palaguita2010_usermanual}.
+Because D5.2 is a pure software deliverable, the sole role of this document is to function as a starting point for further information about the catalogue.
+Nevertheless, the deliverable D5.1 is aimed to fulfil this purpose, so for any further information such as where the code of the catalogue can be found, installation instructions, system requirements, a overview of the catalogue architecture or a detailed API, please read \cite{palaguita2010_usermanual}.
 

### 2010-02-19 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_mediation.tex
Version 1.0 (after integrating corrections)
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@570 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \newcommand\workpackageTitle{Definition of Conceptual Model}
-\newcommand\authorOne{Oszkar Ambrus, NUIG}
-\newcommand\authorTwo{Author 2}
+\newcommand\authorOne{Oszkar Ambrus (NUIG)}
+\newcommand\authorTwo{Albert Zuendorf (UniKassel)}
 \newcommand\authorThree{Author 3}
@@
         \vspace{2cm}
-        {\large \sffamily \authorOne}\\
+        {\large \sffamily \authorOne} \\
+        \vspace{0.5cm}
+        {\large \sffamily \authorTwo} \\
         \vspace{0.5cm}
@@
                             \textcolor{white}{\textbf{Change description}}\\ \hline
-0.1 & 07.01.2010 & Oszkar Ambrus (NUIG) & initial draft \\ \hline
+0.1 & 09.05.2009 & \authorOne & Initial content \\ \hline
+0.4 & 30.11.2009 & \authorOne & Sections 2 and 3 \\ \hline
+0.5 & 07.01.2010 & \authorOne & New structure, Section 4 and 5 \\ \hline
+0.7 & 03.02.2010 & \authorTwo & Section 3.5, 3.6, updated Section 5  \\ \hline
+0.8 & 03.02.2010 & \authorOne & Integration and preliminary finalization  \\ \hline
+0.9 & 03.02.2010 & NUIG, UniKassel & Corrections  \\ \hline
+1.0 & 03.02.2010 & \authorOne & Integration and finalization  \\ \hline
 %0.2 & 19.02.2009 & Oszkar Ambrus (NUIG) & content moved from wiki after internal review \\ \hline
@@
     \hline
-    \textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2009} & {\textbf{Submitted}} & {27/02/2009}\\ \hline
+    \textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {26/02/2010} & {\textbf{Submitted}} & {22/02/2010}\\ \hline
     \textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
@@
 We present the different roles within the platform, tasks and challenges related to them, and the way ontology matching is integrated within each role.}\\ \hline
-    \textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
+    \textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne,  \authorTwo}\\
 %   {} & \multicolumn{4}{l|}{}\\ 
@@
 \label{sec:intro}
-\paragraph{Goal and Scope.}
+\subsection{Goal and Scope.}
 FAST is a visual programming platform allowing business users to build enterprise-class mashups, employing various underlying services and generating new ones.
@@
 In the services used within the FAST environment, the ontologies involved in the creation of gadgets and building blocks come from different parties.
-These ontologies, though different, will sometimes cover the same domain, making ontology matching necessary, as the means to reconcile differences in the various conceptualisations.
+These ontologies, though different, will sometimes cover the same domain, making ontology matching necessary as the means to reconcile differences in the various conceptualisations.
 
@@
 
-\paragraph{Structure of the Document.}
+\subsubsection{Changes to Previous Version}
+
+The first iteration, dealing with the theoretical foundations of ontology matching, is now contained in Part \ref{part:d2.4.1} and the whole of Part \ref{part:d2.4.2} is newly added content, based on the comments of the last review.
+As such, it tackles design and development issues for integrating ontology matching within the FAST platform: it presents the overall strategy and deals with specific problems belonging to each phase of ontology matching in FAST.
+
+\subsection{Structure of the Document.}
 Part \ref{part:d2.4.2} describes the approach of integrating ontology matching into the FAST environment.
@@
 
-\paragraph{Overall Approach for WP2 and relation to other Work Packages and Deliverables.}
+\subsection{Overall Approach for WP2 and relation to other Work Packages and Deliverables.}
 Ontology matching will play an important role for the integration of different underlying resources in the user-built gadgets. 
 We present our initial research in the implementation of ontology matching in this second iteration of D2.4, after having laid the theoretical ground work of ontology matching in the first iteration.
-The final version of the deliverable (i.e., D2.4.3) will finalize the implementation of ontology mediation within the FAST platform.
+The final version of the deliverable (i.e., D2.4.3) will complete the implementation of ontology mediation within the FAST platform.
+
+Parts of this work rely on the results of D4.3.2, where a service wrapper tool is described, that will be modified to be used in matching ontologies.
 
@@
 In this second iteration we established the approach to integrate ontology matching into FAST, and we have shown how this matching is done at the different phases of the gadget life cycle and the approach taken at every level.
-Firstly, have shown the implementation details as well as test scenarios, we have presented a tool for finding the alignments between ontologies as well as a powerful language to represent them, we have described a tool to compose mapping rules in a consistent manner with other FAST components and the means to save them in the Semantic Catalogue.
+Firstly, we have shown the implementation details as well as test scenarios, we have presented a tool for finding the alignments between ontologies as well as a powerful language to represent them, we have described a tool to compose mapping rules in a consistent manner with other FAST components and the means to save them in the Semantic Catalogue.
 Secondly, we described the way in which mapping operators are used in the GVS to combine discrepant screens.

### 2010-02-19 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_part1.tex
Version 1.0 (after integrating corrections)
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@570 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \subsubsection{Automatic Mapping}
+\label{subsubsec:automaticmapping}
 
@@
 \begin{figure}
-  \begin{center}
+    \centering
     \includegraphics[width=0.8\linewidth]{images/FactsTransformationToolGVSWithPortDefinitionsNoGlitch2.png}
@@
     \label{fig:FactsTransformationToolPortDefinitions}
-  \end{center}
 \end{figure}
@@
 
-Note that simple one-to-one transformation rules should be derived with the help of the ontology mediation approach described above.
+Note that simple one-to-one transformation rules should be derived with the help of the ontology mediation approach described in Section \ref{subsubsec:automaticmapping}.
 Such simple rules should be added to the transformation tool, automatically.
@@
 So far our transformation rules are restricted to simple cases.
-So far we have focused on simple cases allowing a simple transformation approach that does not require to sophisticated skills from our users.
+Thus, we have focused on simple cases allowing a simple transformation approach that does not require sophisticated skills from our users.
 However, as discussed on the Amazon to GoodRelations example, there exist more complex cases.
@@
 Once the data transformation operator has been defined, it will be added to the FAST catalogue in order to allow its use for the definition of piping configurations of new screens within the FAST CGS tool.
-For this purpose a simple semantics description will be send to FAST catalogue defining the semantics of the operator and the semantics of its input and output port.
+For this purpose a simple semantic description will be sent to FAST catalogue defining the semantics of the operator and the semantics of its input and output ports.
 
 Currently, the Service Wrapper Tool uses its own repository for trees of transformation rules.
-First of all this has technical reasons: The development of the transformation tool is based on an reference architecture and libraries for web based interactive systems provided by Kassel University, cf.
+First of all this has technical reasons: the development of the transformation tool is based on a reference architecture and libraries for web based interactive systems provided by Kassel University, cf.
 \cite{AsDH2009CVSM}.
@@
 
-At the gadget building phase, whenever preconditions and postconditions of screens, resources or operators show a semantic mismatch, the catalogue will offer the possible mediation operators that would solve the incompatibility problem.
-This way an mediation operator can be placed between two pre- postconditions that would otherwise be incompatible, making interoperation between the two possible.
+At the gadget building phase, whenever preconditions and postconditions of screens, resources or operators show a semantic mismatch, the catalogue will offer the possible matching operators that would solve the incompatibility problem.
+This way a matching operator can be placed between two pre- postconditions that would otherwise be incompatible, making interoperation between the two possible.
 In seldom cases, the catalogue may fail to retrieve an appropriate mediation operator.
-In such cases building of a new mediation operator with the help of the Facts Transformation Tool will be triggered in order to meet the users need.
+In such cases, the possibility of building a new matching operator with the help of the Facts Transformation Tool will be presented, in order to meet the user's need.
 
@@
 This runtime library provides notification mechanisms and manages the fact base of the gadget.
-For data transformation we will generate Javascript code that connects itself to the fact base and any time a fact is send to the input port of the operator, it will do the data transformation and send the result to the output port.
+For data transformation we will generate Javascript code that connects itself to the fact base and any time a fact is sent to the input port of the operator, it will do the data transformation and send the result to the output port.
 The code generated from the data transformation rules follows a recursive pattern where each rule becomes a Javascript function with parameters for the source and target fact subtrees to be transformed.
@@
 Thus, the \textit{CodeGen Viewer} tab will also allow to examine the templates used for code generation and to adapt the templates in order to meet changing APIs or to meet new requirements.
-Thereby, the code generation is not hard wired into our tool but the code generation may be adapted by (skilled) end users on the fly.
+Thereby, the code generation is not hard wired into our tool but may be adapted by (skilled) end users on the fly.

### 2010-02-19 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_mediation.tex
Minor corrections
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@571 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
+\pagebreak
+\part{Conclusion}
+
+\pagebreak
 \section{Summary}
@@
 
-In the first iteration (described in Part \ref{part:d2.4.1} we laid the theoretical foundations of ontology matching.
+In the first iteration (described in Part \ref{part:d2.4.1}) we laid the theoretical foundations of ontology matching.
 After presenting the main problem, and the types of mismatches between ontologies, we have discussed - based on the most prominent research results in these areas - the three main approaches to ontology mediation, that is \textit{ontology mapping} (representing  correspondences between entities), \textit{ontology alignment} (the process of finding the correspondences between ontologies) and \textit{ontology merging }(the process of creating a new ontology based
@@
 
-In this second iteration we established the approach to integrate ontology matching into FAST, and we have shown how this matching is done at the different phases of the gadget life cycle and the approach taken at every level.
+In this second iteration (Part \ref{part:d2.4.2}) we established the approach to integrate ontology matching into FAST, and we have shown how this matching is done at the different phases of the gadget life cycle and the approach taken at every level.
 Firstly, we have shown the implementation details as well as test scenarios, we have presented a tool for finding the alignments between ontologies as well as a powerful language to represent them, we have described a tool to compose mapping rules in a consistent manner with other FAST components and the means to save them in the Semantic Catalogue.
@@
 \doublespacing
+\appendix
 \section*{Appendix A (Lists of Tables and Figures)}

### 2010-02-19 trunk/catalogue_service/src/eu/morfeoproject/fast/catalogue/Catalogue.java
change FormElement to Form
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@573 d8b0c59b-6544-0410-94d9-83a5658a058e

@@ -47,7 +47,7 @@ import eu.morfeoproject.fast.model.BackendService;
 import eu.morfeoproject.fast.model.CTag;
 import eu.morfeoproject.fast.model.Condition;
 import eu.morfeoproject.fast.model.FastModelFactory;
-import eu.morfeoproject.fast.model.FormElement;
+import eu.morfeoproject.fast.model.Form;
 import eu.morfeoproject.fast.model.Library;
 import eu.morfeoproject.fast.model.Operator;
 import eu.morfeoproject.fast.model.Pipe;
@@ -275,7 +275,7 @@ public class Catalogue {
 	}
 	
 	public boolean containsScreenComponent(URI uri) {
-		return containsFormElement(uri)
+		return containsForm(uri)
 			|| containsOperator(uri)
 			|| containsBackendService(uri);
 	}
@@ -284,12 +284,12 @@ public class Catalogue {
 		return containsScreenComponent(sc.getUri());
 	}
 	
-	public boolean containsFormElement(URI feUri) {
-		return tripleStore.isResource(feUri, FGO.FormElement);
+	public boolean containsForm(URI feUri) {
+		return tripleStore.isResource(feUri, FGO.Form);
 	}
 	
-	public boolean containsFormElement(FormElement fe) {
-		return containsFormElement(fe.getUri());
+	public boolean containsForm(Form fe) {
+		return containsForm(fe.getUri());
 	}
 
 	public boolean containsOperator(URI opUri) {
@@ -325,7 +325,7 @@ public class Catalogue {
     		int limit,
     		Set<String> domainContext) throws ClassCastException, ModelRuntimeException {
     	HashSet<URI> results = new HashSet<URI>();
-    	results.addAll(findScreenComponents(container, conditions, toExclude, offset, limit, domainContext, FGO.FormElement));
+    	results.addAll(findScreenComponents(container, conditions, toExclude, offset, limit, domainContext, FGO.Form));
     	results.addAll(findScreenComponents(container, conditions, toExclude, offset, limit, domainContext, FGO.Operator));
     	results.addAll(findScreenComponents(container, conditions, toExclude, offset, limit, domainContext, FGO.BackendService));
     	return results;
@@ -785,7 +785,7 @@ public class Catalogue {
 			return createResource(serverURL, "screenflows", ofClass, id);
 		} else if (ofClass.equals(FGO.Screen)) {
[…]

### 2010-02-19 trunk/catalogue_service/src/eu/morfeoproject/fast/model/FastModelFactory.java

@@ -14,8 +14,8 @@ public class FastModelFactory {
 		return new Screen(null);
 	}
 	
-	public static FormElement createFormElement() {
-		return new FormElement(null);
+	public static Form createForm() {
+		return new Form(null);
 	}
 	
 	public static Operator createOperator() {

### 2010-02-19 trunk/catalogue_service/src/eu/morfeoproject/fast/model/Form.java

@@ -6,9 +6,9 @@ import org.ontoware.rdf2go.vocabulary.RDF;
 
 import eu.morfeoproject.fast.vocabulary.FGO;
 
-public class FormElement extends ScreenComponent {
+public class Form extends ScreenComponent {
 	
-	protected FormElement(URI uri) {
+	protected Form(URI uri) {
 		setUri(uri);
 	}
 	
@@ -16,7 +16,7 @@ public class FormElement extends ScreenComponent {
 	public Model createModel() {
 		Model model = super.createModel();
 		
-		model.addStatement(getUri(), RDF.type, FGO.FormElement);
+		model.addStatement(getUri(), RDF.type, FGO.Form);
 
 		return model;
 	}

### 2010-02-19 trunk/catalogue_service/src/eu/morfeoproject/fast/model/Resource.java

@@ -244,7 +244,7 @@ public abstract class Resource {
 			json.put("type", "screenflow");
 		else if (this instanceof Screen)
 			json.put("type", "screen");
-		else if (this instanceof FormElement)
+		else if (this instanceof Form)
 			json.put("type", "form");
 		else if (this instanceof Operator)
 			json.put("type", "operator");

### 2010-02-19 trunk/catalogue_service/src/eu/morfeoproject/fast/server/FormServlet.java

@@ -24,7 +24,7 @@ import eu.morfeoproject.fast.catalogue.InvalidResourceTypeException;
 import eu.morfeoproject.fast.catalogue.NotFoundException;
 import eu.morfeoproject.fast.catalogue.OntologyInvalidException;
 import eu.morfeoproject.fast.catalogue.ResourceException;
-import eu.morfeoproject.fast.model.FormElement;
+import eu.morfeoproject.fast.model.Form;
 import eu.morfeoproject.fast.model.templates.BuildingBlockTemplate;
 import eu.morfeoproject.fast.model.templates.CollectionTemplate;
 import eu.morfeoproject.fast.model.templates.TemplateManager;
@@ -32,17 +32,17 @@ import eu.morfeoproject.fast.util.Accept;
 import freemarker.template.TemplateException;
 
 /**
- * Servlet implementation class FormElementServlet
+ * Servlet implementation class FormServlet
  */
-public class FormElementServlet extends GenericServlet {
+public class FormServlet extends GenericServlet {
 	private static final long serialVersionUID = 1L;
 
-	static Logger logger = LoggerFactory.getLogger(FormElementServlet.class);
+	static Logger logger = LoggerFactory.getLogger(FormServlet.class);
     
     /**
      * @see HttpServlet#HttpServlet()
      */
-    public FormElementServlet() {
+    public FormServlet() {
         super();
     }
 
@@ -78,7 +78,7 @@ public class FormElementServlet extends GenericServlet {
 						Model model = RDF2Go.getModelFactory().createModel();
 						try {
 							model.open();
-							for (FormElement f : CatalogueAccessPoint.getCatalogue().listFormElements()) {
+							for (Form f : CatalogueAccessPoint.getCatalogue().listForms()) {
 								Model feModel = f.createModel();
 								for (String ns : feModel.getNamespaces().keySet())
 									model.setNamespace(ns, feModel.getNamespace(ns));
@@ -99,11 +99,11 @@ public class FormElementServlet extends GenericServlet {
 							response.setCharacterEncoding(TemplateManager.getDefaultEncoding());
 						if (TemplateManager.getLocale() != null)
 							response.setLocale(TemplateManager.getLocale());
-						CollectionTemplate.process(CatalogueAccessPoint.getCatalogue().listFormElements(), writer);
+						CollectionTemplate.process(CatalogueAccessPoint.getCatalogue().listForms(), writer);
[…]

### 2010-02-19 trunk/catalogue_service/src/eu/morfeoproject/fast/server/GenericServlet.java

@@ -24,7 +24,7 @@ import eu.morfeoproject.fast.model.BackendService;
 import eu.morfeoproject.fast.model.CTag;
 import eu.morfeoproject.fast.model.Condition;
 import eu.morfeoproject.fast.model.FastModelFactory;
-import eu.morfeoproject.fast.model.FormElement;
+import eu.morfeoproject.fast.model.Form;
 import eu.morfeoproject.fast.model.Library;
 import eu.morfeoproject.fast.model.Operator;
 import eu.morfeoproject.fast.model.Pipe;
@@ -238,9 +238,9 @@ public abstract class GenericServlet extends HttpServlet {
 		return post;
 	}
 	
-	protected FormElement parseFormElement(JSONObject jsonFe, URI uri) throws JSONException, IOException {
-		FormElement fe = FastModelFactory.createFormElement();
-		parseScreenComponent(fe, jsonFe, uri);
+	protected Form parseForm(JSONObject jsonForm, URI uri) throws JSONException, IOException {
+		Form fe = FastModelFactory.createForm();
+		parseScreenComponent(fe, jsonForm, uri);
 		return fe;
 	}
 

### 2010-02-19 trunk/catalogue_service/src/eu/morfeoproject/fast/server/MetadataServlet.java

@@ -63,8 +63,8 @@ public class MetadataServlet extends HttpServlet {
 					arrayScreenflows.put(CatalogueAccessPoint.getCatalogue().getScreenFlow(uri).toJSON());
 				else if (CatalogueAccessPoint.getCatalogue().containsScreen(uri))
 					arrayScreens.put(CatalogueAccessPoint.getCatalogue().getScreen(uri).toJSON());
-				else if (CatalogueAccessPoint.getCatalogue().containsFormElement(uri))
-					arrayForms.put(CatalogueAccessPoint.getCatalogue().getFormElement(uri).toJSON());
+				else if (CatalogueAccessPoint.getCatalogue().containsForm(uri))
+					arrayForms.put(CatalogueAccessPoint.getCatalogue().getForm(uri).toJSON());
 				else if (CatalogueAccessPoint.getCatalogue().containsOperator(uri))
 					arrayOperators.put(CatalogueAccessPoint.getCatalogue().getOperator(uri).toJSON());
 				else if (CatalogueAccessPoint.getCatalogue().containsBackendService(uri))

### 2010-02-19 trunk/catalogue_service/src/eu/morfeoproject/fast/server/ScreenComponentFindCheckServlet.java

@@ -152,7 +152,7 @@ public class ScreenComponentFindCheckServlet extends GenericServlet {
 			
 			// add results of 'find' to the list of forms
 			if (search) {
-				Set<URI> formResults = CatalogueAccessPoint.getCatalogue().findScreenComponents(null, conList, all, 0, -1, tags, FGO.FormElement);
+				Set<URI> formResults = CatalogueAccessPoint.getCatalogue().findScreenComponents(null, conList, all, 0, -1, tags, FGO.Form);
 				for (URI uri : formResults)
 					forms.add(CatalogueAccessPoint.getCatalogue().getScreenComponent(uri));
 				// add results of 'find' to the list of operators

### 2010-02-19 trunk/catalogue_service/src/eu/morfeoproject/fast/vocabulary/FGO.java

@@ -4,7 +4,7 @@ import org.ontoware.rdf2go.model.node.URI;
 import org.ontoware.rdf2go.model.node.impl.URIImpl;
 
 /**
- * Vocabulary File. Created by org.ontoware.rdf2go.util.VocabularyWriter on Fri Jan 22 13:22:40 GMT 2010
+ * Vocabulary File. Created by org.ontoware.rdf2go.util.VocabularyWriter on Fri Feb 19 19:25:18 GMT 2010
  * input file: src/eu/morfeoproject/fast/catalogue/ontologies/fgo.rdf
  * namespace: http://purl.oclc.org/fast/ontology/gadget#
  */
@@ -71,10 +71,11 @@ public interface FGO {
     public static final URI Resource = new URIImpl("http://purl.oclc.org/fast/ontology/gadget#Resource", false);
 
     /**
-     * Label: Form Element@en 
-     * Comment: Form elements are UI elements in a particular screen.@en 
+     * Label: Postcondition@en 
+     * Comment: A postcondition is a result condition within a screenflow. It can be seen
+    as an output of the screenflow.@en 
      */
-    public static final URI FormElement = new URIImpl("http://purl.oclc.org/fast/ontology/gadget#FormElement", false);
+    public static final URI Postcondition = new URIImpl("http://purl.oclc.org/fast/ontology/gadget#Postcondition", false);
 
     /**
      * Label: Action@en 
@@ -83,13 +84,6 @@ public interface FGO {
 	be performed when data from a service is received).@en 
      */
     public static final URI Library = new URIImpl("http://purl.oclc.org/fast/ontology/gadget#Library", false);
-
-    /**
-     * Label: Postcondition@en 
-     * Comment: A postcondition is a result condition within a screenflow. It can be seen
-    as an output of the screenflow.@en 
-     */
-    public static final URI Postcondition = new URIImpl("http://purl.oclc.org/fast/ontology/gadget#Postcondition", false);
 
     /**
      * Label: Screen Flow@en 
@@ -216,6 +210,12 @@ public interface FGO {
 	for the condition be true as well.@en 
      */
     public static final URI Condition = new URIImpl("http://purl.oclc.org/fast/ontology/gadget#Condition", false);
+
+    /**
+     * Label: Form Element@en 
+     * Comment: Form elements are UI elements in a particular screen.@en 
+     */
[…]

### 2010-02-19 trunk/catalogue_service/src/eu/morfeoproject/fast/server/BackendServiceServlet.java
fix bug in content negotiation
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@574 d8b0c59b-6544-0410-94d9-83a5658a058e

@@ -69,6 +69,8 @@ public class BackendServiceServlet extends GenericServlet {
 			if (id == null) {
 				// List the members of the collection
 				logger.info("Retrieving all services");
+				// Override format regarding the given extension
+				format = MediaType.forExtension(extension);
 				try {
 					if (format.equals(MediaType.APPLICATION_RDF_XML) ||
 							format.equals(MediaType.APPLICATION_TURTLE)) {

### 2010-02-19 trunk/catalogue_service/src/eu/morfeoproject/fast/server/OperatorServlet.java

@@ -69,6 +69,8 @@ public class OperatorServlet extends GenericServlet {
 			if (id == null) {
 				// List the members of the collection
 				logger.info("Retrieving all operators");
+				// Override format regarding the given extension
+				format = MediaType.forExtension(extension);
 				try {
 					if (format.equals(MediaType.APPLICATION_RDF_XML) ||
 							format.equals(MediaType.APPLICATION_TURTLE)) {

### 2010-02-19 trunk/catalogue_service/src/eu/morfeoproject/fast/server/PostconditionServlet.java

@@ -68,6 +68,8 @@ public class PostconditionServlet extends GenericServlet {
 			if (id == null) {
 				// List the members of the collection
 				logger.info("Retrieving all postconditions");
+				// Override format regarding the given extension
+				format = MediaType.forExtension(extension);
 				try {
 					if (format.equals(MediaType.APPLICATION_RDF_XML) ||
 							format.equals(MediaType.APPLICATION_TURTLE)) {

### 2010-02-19 trunk/catalogue_service/src/eu/morfeoproject/fast/server/PreconditionServlet.java

@@ -68,6 +68,8 @@ public class PreconditionServlet extends GenericServlet {
 			if (id == null) {
 				// List the members of the collection
 				logger.info("Retrieving all preconditions");
+				// Override format regarding the given extension
+				format = MediaType.forExtension(extension);
 				try {
 					if (format.equals(MediaType.APPLICATION_RDF_XML) ||
 							format.equals(MediaType.APPLICATION_TURTLE)) {

### 2010-02-19 trunk/catalogue_service/src/eu/morfeoproject/fast/server/ScreenServlet.java

@@ -70,6 +70,8 @@ public class ScreenServlet extends GenericServlet {
 			if (id == null) {
 				// List the members of the collection
 				logger.info("Retrieving all screens");
+				// Override format regarding the given extension
+				format = MediaType.forExtension(extension);
 				try {
 					if (format.equals(MediaType.APPLICATION_RDF_XML) ||
 							format.equals(MediaType.APPLICATION_TURTLE)) {

### 2010-02-19 trunk/catalogue_service/src/eu/morfeoproject/fast/server/ScreenflowServlet.java

@@ -69,6 +69,8 @@ public class ScreenflowServlet extends GenericServlet {
 			if (id == null) {
 				// List the members of the collection
 				logger.info("Retrieving all screenflows");
+				// Override format regarding the given extension
+				format = MediaType.forExtension(extension);
 				try {
 					if (format.equals(MediaType.APPLICATION_RDF_XML) ||
 							format.equals(MediaType.APPLICATION_TURTLE)) {

### 2010-02-19 trunk/deliverables/D5_1_1/fast_d5_1_1.tex
change any reference from "form element" to "form"
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@576 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 	\begin{itemize}
-		\item support for form elements, operators and back-end services (see Section ...),
-		\item convenience operations for recommendation and checking satisfaction and reachability for form elements, operators and back-end services (see Section ...),
+		\item support for forms, operators and back-end services (see Section ...),
+		\item convenience operations for recommendation and checking satisfaction and reachability for forms, operators and back-end services (see Section ...),
 		\item and, last but not least, support for Screen planning (see Section~\ref{sub:planning}).

### 2010-02-19 trunk/catalogue_service/src/eu/morfeoproject/fast/server/MediaType.java
minor change
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@580 d8b0c59b-6544-0410-94d9-83a5658a058e

@@ -16,7 +16,7 @@ public class MediaType {
 	/* turtle: textual serialization of RDF */
 	public static final String APPLICATION_TURTLE = "text/turtle";
 	
-	/* n3: extension¹ of turtle language expressing a superset of RDF */
+	/* n3: extension of turtle language expressing a superset of RDF */
 	public static final String TEXT_N3 = "text/rdf+n3";
 
 	public static String forExtension(String extension) {

### 2010-02-20 trunk/deliverables/D2_2_Ontology/classes.tex
common tag added
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@463 d8b0c59b-6544-0410-94d9-83a5658a058e
greatly extending conceptualisation part with UML diagrams, also updating to reflect changes in implementations
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@582 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
-\subsubsection{Class: \texttt{fgo:Resource}}
-\label{subs:Resource}
+\subsubsection*{Class: \texttt{fgo:Action}}
+\label{subs:Action}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & Resource \\ \hline
-\textcolor{white}{\textbf{description}} & Anything that is part of a gadget (or the gadget itself). Tentatively 
-    anything that can be `touched' and moved around in the FAST IDE. \\ \hline
-\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fgo:ScreenFlow}}{subs:ScreenFlow}, \htmlref{\texttt{fgo:Screen}}{subs:Screen}, \htmlref{\texttt{fgo:FlowControlElement}}{subs:FlowControlElement}, \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent}, \htmlref{\texttt{fgo:Condition}}{subs:Condition}, \htmlref{\texttt{fgo:Fact}}{subs:Fact}, \htmlref{\texttt{fgo:WithPreConditions}}{subs:WithPreConditions}, \htmlref{\texttt{fgo:WithPostConditions}}{subs:WithPostConditions}, \htmlref{\texttt{fgo:WithConditions}}{subs:WithConditions}, \htmlref{\texttt{fgo:Precondition}}{subs:Precondition}, \htmlref{\texttt{fgo:Postcondition}}{subs:Postcondition}, \htmlref{\texttt{fgo:Pipe}}{subs:Pipe}, \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:contains}}{subs:contains}, \htmlref{\texttt{fgo:hasIcon}}{subs:hasIcon}, \htmlref{\texttt{fgo:hasScreenshot}}{subs:hasScreenshot}, \htmlref{\texttt{fgo:hasTag}}{subs:hasTag}, \htmlref{\texttt{fgo:hasVersion}}{subs:hasVersion}, \htmlref{\texttt{fgo:hasId}}{subs:hasId}, \htmlref{\texttt{fgo:hasName}}{subs:hasName}, \htmlref{\texttt{fgo:hasType}}{subs:hasType}, \htmlref{\texttt{fgo:hasDefinition}}{subs:hasDefinition} \\ \hline
+\textcolor{white}{\textbf{label}} & Action \\ \hline
+\textcolor{white}{\textbf{description}} & An action represents a specific routine which will be performed when a certain
+	condition is fulfilled within a certain screen component (i.e., the action `showTable' will
+	be performed when data from a service is received). \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasUse}}{subs:hasUse} \\ \hline
+\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasAction}}{subs:hasAction} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fgo:ScreenFlow}}
-\label{subs:ScreenFlow}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Screen Flow \\ \hline
-\textcolor{white}{\textbf{description}} & The complete gadget, a set of screens. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\end{tabular}
-\subsubsection{Class: \texttt{fgo:Screen}}
-\label{subs:Screen}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Screen \\ \hline
-\textcolor{white}{\textbf{description}} & An individual screen. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasCode}}{subs:hasCode} \\ \hline
-\end{tabular}
-\subsubsection{Class: \texttt{fgo:FlowControlElement}}
-\label{subs:FlowControlElement}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Flow Control Element \\ \hline
-\textcolor{white}{\textbf{description}} & Any kind of component which can restrict the default flow of screens 
-    in a gadget. \\ \hline
-\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fgo:ScreenFlowStart}}{subs:ScreenFlowStart}, \htmlref{\texttt{fgo:ScreenFlowEnd}}{subs:ScreenFlowEnd}, \htmlref{\texttt{fgo:Connector}}{subs:Connector} \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\end{tabular}
-\subsubsection{Class: \texttt{fgo:ScreenFlowStart}}
-\label{subs:ScreenFlowStart}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Screen Flow Start \\ \hline
-\textcolor{white}{\textbf{description}} & The entry point to a wigdet; the first screen. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:FlowControlElement}}{subs:FlowControlElement} \\ \hline
-\end{tabular}
-\subsubsection{Class: \texttt{fgo:ScreenFlowEnd}}
-\label{subs:ScreenFlowEnd}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Screen Flow End \\ \hline
-\textcolor{white}{\textbf{description}} & A screen that ends the workflow of the gadget. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:FlowControlElement}}{subs:FlowControlElement} \\ \hline
-\end{tabular}
-\subsubsection{Class: \texttt{fgo:Connector}}
-\label{subs:Connector}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Connector \\ \hline
-\textcolor{white}{\textbf{description}} & An explicit connection between two screens. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:FlowControlElement}}{subs:FlowControlElement} \\ \hline
-\end{tabular}
-\subsubsection{Class: \texttt{fgo:ScreenComponent}}
-\label{subs:ScreenComponent}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Screen Component \\ \hline
-\textcolor{white}{\textbf{description}} & A screen component is any resource which is part of a particular screen. \\ \hline
-\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fgo:Operator}}{subs:Operator}, \htmlref{\texttt{fgo:FormElement}}{subs:FormElement}, \htmlref{\texttt{fgo:BackendService}}{subs:BackendService} \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasAction}}{subs:hasAction}, \htmlref{\texttt{fgo:hasTrigger}}{subs:hasTrigger}, \htmlref{\texttt{fgo:hasLibrary}}{subs:hasLibrary} \\ \hline
-\end{tabular}
-\subsubsection{Class: \texttt{fgo:Operator}}
-\label{subs:Operator}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Operator \\ \hline
-\textcolor{white}{\textbf{description}} & Any kind of component that is used to connect backend services to 
-    form elements. Examples are simple pipes, aggregators or various kinds of 
-    filters. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
-\end{tabular}
-\subsubsection{Class: \texttt{fgo:FormElement}}
-\label{subs:FormElement}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Form Element \\ \hline
-\textcolor{white}{\textbf{description}} & Form elements are UI elements in a particular screen. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
-\end{tabular}
-\subsubsection{Class: \texttt{fgo:BackendService}}
+\subsubsection*{Class: \texttt{fgo:BackendService}}
 \label{subs:BackendService}
@@
 \end{tabular}
-\subsubsection{Class: \texttt{fgo:Condition}}
+\subsubsection*{Class: \texttt{fgo:BackendServiceDefinition}}
+\label{subs:BackendServiceDefinition}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Backend Service Definition \\ \hline
+\end{tabular}
+\subsubsection*{Class: \texttt{fgo:Condition}}
 \label{subs:Condition}
@@
 \textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasPreCondition}}{subs:hasPreCondition}, \htmlref{\texttt{fgo:hasPostCondition}}{subs:hasPostCondition}, \htmlref{\texttt{fgo:hasCondition}}{subs:hasCondition} \\ \hline
+\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasCondition}}{subs:hasCondition}, \htmlref{\texttt{fgo:hasPostCondition}}{subs:hasPostCondition}, \htmlref{\texttt{fgo:hasPreCondition}}{subs:hasPreCondition} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fgo:Fact}}
+\subsubsection*{Class: \texttt{fgo:Connector}}
+\label{subs:Connector}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Connector \\ \hline
+\textcolor{white}{\textbf{description}} & An explicit connection between two screens. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:FlowControlElement}}{subs:FlowControlElement} \\ \hline
+\end{tabular}
+\subsubsection*{Class: \texttt{fgo:Definition}}
+\label{subs:Definition}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Resource definition \\ \hline
+\textcolor{white}{\textbf{description}} & Structural and behaviour definition of a resource. \\ \hline
+\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fgo:ScreenDefinition}}{subs:ScreenDefinition} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:contains}}{subs:contains} \\ \hline
+\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasDefinition}}{subs:hasDefinition} \\ \hline
+\end{tabular}
+\subsubsection*{Class: \texttt{fgo:Fact}}
 \label{subs:Fact}
@@
 \end{tabular}
-\subsubsection{Class: \texttt{fgo:WithPreConditions}}
-\label{subs:WithPreConditions}
+\subsubsection*{Class: \texttt{fgo:FlowControlElement}}
+\label{subs:FlowControlElement}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & With-precondition \\ \hline
-\textcolor{white}{\textbf{description}} & Those kinds of resource which can have pre-conditions 
-    (i.e., screens and screen flows). \\ \hline
+\textcolor{white}{\textbf{label}} & Flow Control Element \\ \hline
+\textcolor{white}{\textbf{description}} & Any kind of component which can restrict the default flow of screens 
+    in a gadget. \\ \hline
+\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fgo:Connector}}{subs:Connector}, \htmlref{\texttt{fgo:ScreenFlowEnd}}{subs:ScreenFlowEnd}, \htmlref{\texttt{fgo:ScreenFlowStart}}{subs:ScreenFlowStart} \\ \hline
 \textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasPreCondition}}{subs:hasPreCondition} \\ \hline
-\textcolor{white}{\textbf{unionOf}} & \htmlref{\texttt{fgo:ScreenFlow}}{subs:ScreenFlow}, \htmlref{\texttt{fgo:Screen}}{subs:Screen}, \htmlref{\texttt{fgo:Action}}{subs:Action} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fgo:WithPostConditions}}
-\label{subs:WithPostConditions}
+\subsubsection*{Class: \texttt{fgo:FormDefinition}}
+\label{subs:FormDefinition}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & With-precondition \\ \hline
-\textcolor{white}{\textbf{description}} & Those kinds of resource which can have pre-conditions 
-    (i.e., screens and screen flows). \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasPostCondition}}{subs:hasPostCondition} \\ \hline
-\textcolor{white}{\textbf{unionOf}} & \htmlref{\texttt{fgo:ScreenFlow}}{subs:ScreenFlow}, \htmlref{\texttt{fgo:Screen}}{subs:Screen}, \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
+\textcolor{white}{\textbf{label}} & Form Definition \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fgo:WithConditions}}
-\label{subs:WithConditions}
+\subsubsection*{Class: \texttt{fgo:FormElement}}
+\label{subs:FormElement}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & With-condition \\ \hline
-\textcolor{white}{\textbf{description}} & Those kinds of resource which can have both pre- or post-conditions. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{unionOf}} & \htmlref{\texttt{fgo:WithPreConditions}}{subs:WithPreConditions}, \htmlref{\texttt{fgo:WithPostConditions}}{subs:WithPostConditions} \\ \hline
+\textcolor{white}{\textbf{label}} & Form Element \\ \hline
+\textcolor{white}{\textbf{description}} & Form elements are UI elements in a particular screen. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fgo:Action}}
-\label{subs:Action}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Action \\ \hline
-\textcolor{white}{\textbf{description}} & An action represents a specific routine which will be performed when a certain
-	condition is fulfilled within a certain screen component (i.e., the action `showTable' will
-	be performed when data from a service is received). \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasUse}}{subs:hasUse} \\ \hline
-\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasAction}}{subs:hasAction} \\ \hline
-\end{tabular}
-\subsubsection{Class: \texttt{fgo:Library}}
+\subsubsection*{Class: \texttt{fgo:Library}}
 \label{subs:Library}
@@
 \textcolor{white}{\textbf{label}} & Action \\ \hline
-\textcolor{white}{\textbf{description}} & An action represents a specific routine which will be performed when a certain
-	condition is fulfilled within a certain screen component (i.e., the action `showTable' will
-	be performed when data from a service is received). \\ \hline
+\textcolor{white}{\textbf{description}} & Libraries are references to external code libraries required for the execution of a particular building block at runtime. \\ \hline
 \textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasLanguage}}{subs:hasLanguage}, \htmlref{\texttt{fgo:hasSource}}{subs:hasSource} \\ \hline
@@
 \end{tabular}
-\subsubsection{Class: \texttt{fgo:Precondition}}
+\subsubsection*{Class: \texttt{fgo:Operator}}
+\label{subs:Operator}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Operator \\ \hline
+\textcolor{white}{\textbf{description}} & Any kind of component that is used to connect backend services to 
+    form elements. Examples are simple pipes, aggregators or various kinds of 
+    filters. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
+\end{tabular}
+\subsubsection*{Class: \texttt{fgo:OperatorDefinition}}
+\label{subs:OperatorDefinition}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Operator Definition \\ \hline
+\end{tabular}
+\subsubsection*{Class: \texttt{fgo:Pipe}}
+\label{subs:Pipe}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & pipe or connector \\ \hline
+\textcolor{white}{\textbf{description}} & Define a pipe or connector between two resources. The connection is made
+	specifying the conditions which will be connected. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasIdActionTo}}{subs:hasIdActionTo}, \htmlref{\texttt{fgo:hasIdBBFrom}}{subs:hasIdBBFrom}, \htmlref{\texttt{fgo:hasIdBBTo}}{subs:hasIdBBTo}, \htmlref{\texttt{fgo:hasIdConditionFrom}}{subs:hasIdConditionFrom}, \htmlref{\texttt{fgo:hasIdConditionTo}}{subs:hasIdConditionTo} \\ \hline
+\end{tabular}
+\subsubsection*{Class: \texttt{fgo:Postcondition}}
+\label{subs:Postcondition}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Postcondition \\ \hline
+\textcolor{white}{\textbf{description}} & A postcondition is a result condition within a screenflow. It can be seen
+    as an output of the screenflow. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasCondition}}{subs:hasCondition} \\ \hline
+\end{tabular}
+\subsubsection*{Class: \texttt{fgo:Precondition}}
 \label{subs:Precondition}
@@
 \end{tabular}
-\subsubsection{Class: \texttt{fgo:Postcondition}}
-\label{subs:Postcondition}
+\subsubsection*{Class: \texttt{fgo:Resource}}
+\label{subs:Resource}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & Postcondition \\ \hline
-\textcolor{white}{\textbf{description}} & A postcondition is a result condition within a screenflow. It can be seen
-    as an output of the screenflow. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasCondition}}{subs:hasCondition} \\ \hline
+\textcolor{white}{\textbf{label}} & Resource \\ \hline
+\textcolor{white}{\textbf{description}} & Anything that is part of a gadget (or the gadget itself). Tentatively 
+    anything that can be `touched' and moved around in the FAST IDE. \\ \hline
+\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition}, \htmlref{\texttt{fgo:Fact}}{subs:Fact}, \htmlref{\texttt{fgo:FlowControlElement}}{subs:FlowControlElement}, \htmlref{\texttt{fgo:Pipe}}{subs:Pipe}, \htmlref{\texttt{fgo:Postcondition}}{subs:Postcondition}, \htmlref{\texttt{fgo:Precondition}}{subs:Precondition}, \htmlref{\texttt{fgo:Screen}}{subs:Screen}, \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent}, \htmlref{\texttt{fgo:ScreenFlow}}{subs:ScreenFlow}, \htmlref{\texttt{fgo:Trigger}}{subs:Trigger}, \htmlref{\texttt{fgo:WithConditions}}{subs:WithConditions}, \htmlref{\texttt{fgo:WithPostConditions}}{subs:WithPostConditions}, \htmlref{\texttt{fgo:WithPreConditions}}{subs:WithPreConditions} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:contains}}{subs:contains}, \htmlref{\texttt{fgo:hasDefinition}}{subs:hasDefinition}, \htmlref{\texttt{fgo:hasIcon}}{subs:hasIcon}, \htmlref{\texttt{fgo:hasId}}{subs:hasId}, \htmlref{\texttt{fgo:hasName}}{subs:hasName}, \htmlref{\texttt{fgo:hasScreenshot}}{subs:hasScreenshot}, \htmlref{\texttt{fgo:hasTag}}{subs:hasTag}, \htmlref{\texttt{fgo:hasType}}{subs:hasType}, \htmlref{\texttt{fgo:hasVersion}}{subs:hasVersion} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fgo:Definition}}
-\label{subs:Definition}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Resource definition \\ \hline
-\textcolor{white}{\textbf{description}} & Structural and behaviour definition of a resource. \\ \hline
-\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fgo:ScreenDefinition}}{subs:ScreenDefinition} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:contains}}{subs:contains} \\ \hline
-\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasDefinition}}{subs:hasDefinition} \\ \hline
-\end{tabular}
-\subsubsection{Class: \texttt{fgo:ResourceReference}}
+\subsubsection*{Class: \texttt{fgo:ResourceReference}}
 \label{subs:ResourceReference}
@@
 \end{tabular}
-\subsubsection{Class: \texttt{fgo:ScreenDefinition}}
+\subsubsection*{Class: \texttt{fgo:Screen}}
+\label{subs:Screen}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Screen \\ \hline
+\textcolor{white}{\textbf{description}} & An individual screen. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasCode}}{subs:hasCode} \\ \hline
+\end{tabular}
+\subsubsection*{Class: \texttt{fgo:ScreenComponent}}
+\label{subs:ScreenComponent}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Screen Component \\ \hline
+\textcolor{white}{\textbf{description}} & A screen component is any resource which is part of a particular screen. \\ \hline
+\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fgo:BackendService}}{subs:BackendService}, \htmlref{\texttt{fgo:FormElement}}{subs:FormElement}, \htmlref{\texttt{fgo:Operator}}{subs:Operator} \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasAction}}{subs:hasAction}, \htmlref{\texttt{fgo:hasLibrary}}{subs:hasLibrary}, \htmlref{\texttt{fgo:hasTrigger}}{subs:hasTrigger} \\ \hline
+\end{tabular}
+\subsubsection*{Class: \texttt{fgo:ScreenDefinition}}
 \label{subs:ScreenDefinition}
@@
 \end{tabular}
-\subsubsection{Class: \texttt{fgo:Pipe}}
-\label{subs:Pipe}
+\subsubsection*{Class: \texttt{fgo:ScreenFlow}}
+\label{subs:ScreenFlow}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & pipe or connector \\ \hline
-\textcolor{white}{\textbf{description}} & Define a pipe or connector between two resources. The connection is made
-	specifying the conditions which will be connected. \\ \hline
+\textcolor{white}{\textbf{label}} & Screen Flow \\ \hline
+\textcolor{white}{\textbf{description}} & The complete gadget, a set of screens. \\ \hline
 \textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasIdBBFrom}}{subs:hasIdBBFrom}, \htmlref{\texttt{fgo:hasIdConditionFrom}}{subs:hasIdConditionFrom}, \htmlref{\texttt{fgo:hasIdBBTo}}{subs:hasIdBBTo}, \htmlref{\texttt{fgo:hasIdConditionTo}}{subs:hasIdConditionTo}, \htmlref{\texttt{fgo:hasIdActionTo}}{subs:hasIdActionTo} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fgo:Trigger}}
+\subsubsection*{Class: \texttt{fgo:ScreenFlowEnd}}
+\label{subs:ScreenFlowEnd}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Screen Flow End \\ \hline
+\textcolor{white}{\textbf{description}} & A screen that ends the workflow of the gadget. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:FlowControlElement}}{subs:FlowControlElement} \\ \hline
+\end{tabular}
+\subsubsection*{Class: \texttt{fgo:ScreenFlowStart}}
+\label{subs:ScreenFlowStart}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Screen Flow Start \\ \hline
+\textcolor{white}{\textbf{description}} & The entry point to a wigdet; the first screen. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:FlowControlElement}}{subs:FlowControlElement} \\ \hline
+\end{tabular}
+\subsubsection*{Class: \texttt{fgo:Trigger}}
 \label{subs:Trigger}
@@
 \textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasIdBBFrom}}{subs:hasIdBBFrom}, \htmlref{\texttt{fgo:hasIdBBTo}}{subs:hasIdBBTo}, \htmlref{\texttt{fgo:hasIdActionTo}}{subs:hasIdActionTo}, \htmlref{\texttt{fgo:hasNameFrom}}{subs:hasNameFrom} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasIdActionTo}}{subs:hasIdActionTo}, \htmlref{\texttt{fgo:hasIdBBFrom}}{subs:hasIdBBFrom}, \htmlref{\texttt{fgo:hasIdBBTo}}{subs:hasIdBBTo}, \htmlref{\texttt{fgo:hasNameFrom}}{subs:hasNameFrom} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fgo:FormDefinition}}
-\label{subs:FormDefinition}
+\subsubsection*{Class: \texttt{fgo:WithConditions}}
+\label{subs:WithConditions}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & Form Definition \\ \hline
+\textcolor{white}{\textbf{label}} & With-condition \\ \hline
+\textcolor{white}{\textbf{description}} & Those kinds of resource which can have both pre- or post-conditions. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{unionOf}} & \htmlref{\texttt{fgo:WithPostConditions}}{subs:WithPostConditions}, \htmlref{\texttt{fgo:WithPreConditions}}{subs:WithPreConditions} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fgo:OperatorDefinition}}
-\label{subs:OperatorDefinition}
+\subsubsection*{Class: \texttt{fgo:WithPostConditions}}
+\label{subs:WithPostConditions}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & Operator Definition \\ \hline
+\textcolor{white}{\textbf{label}} & With-post-condition \\ \hline
+\textcolor{white}{\textbf{description}} & Those kinds of resources which can have post-conditions. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasPostCondition}}{subs:hasPostCondition} \\ \hline
+\textcolor{white}{\textbf{unionOf}} & \htmlref{\texttt{fgo:Screen}}{subs:Screen}, \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent}, \htmlref{\texttt{fgo:ScreenFlow}}{subs:ScreenFlow} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fgo:BackendServiceDefinition}}
-\label{subs:BackendServiceDefinition}
+\subsubsection*{Class: \texttt{fgo:WithPreConditions}}
+\label{subs:WithPreConditions}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & Backend Service Definition \\ \hline
+\textcolor{white}{\textbf{label}} & With-pre-condition \\ \hline
+\textcolor{white}{\textbf{description}} & Those kinds of resources which can have pre-conditions. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasPreCondition}}{subs:hasPreCondition} \\ \hline
+\textcolor{white}{\textbf{unionOf}} & \htmlref{\texttt{fgo:Action}}{subs:Action}, \htmlref{\texttt{fgo:Screen}}{subs:Screen}, \htmlref{\texttt{fgo:ScreenFlow}}{subs:ScreenFlow} \\ \hline
 \end{tabular}

### 2010-02-20 trunk/deliverables/D2_2_Ontology/fast_d2_2_2_ontology.tex
greatly extending conceptualisation part with UML diagrams, also updating to reflect changes in implementations
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@582 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
-\documentclass{fast_latex}
+\documentclass[twoside]{fast_latex}
 % yay! We can use any kind of funky diacritic:
@@
 \usepackage{paralist}
+\usepackage{courier}
 
@@
 	extendedchars=true,
-	basicstyle=\footnotesize \sffamily, 
+	% basicstyle=\footnotesize \sffamily, 
+	basicstyle={\footnotesize\ttfamily},
 	stringstyle=\bfseries,
@@
 	breaklines=true,
-	% inputencoding=utf8,
+	inputencoding=utf8,
 }
@@
 \newcommand\workpackageTitle{Definition of Conceptual Model}
-\newcommand\authorOne{Knud M\"oller, NUIG}
+\newcommand\authorOne{Knud Möller, NUIG}
 \newcommand\authorTwo{Ismael Rivera, NUIG}
-\newcommand\authorThree{Author 3}
-\newcommand\authorFour{Author 4}
+\newcommand\authorThree{Marcos Reyes Ureña, TID}
+\newcommand\authorFour{Ciprian Alexandru Palaghita, Cyntelix}
 \newtheorem{example}{\emph{Example}}
@@
 		{\large \sffamily \authorTwo}\\
+		{\large \sffamily \authorThree}\\
+		{\large \sffamily \authorFour}\\
 		\vspace{0.5cm}
@@
 %0.2 & 15.02.2009 & Knud M\"oller (NUIG) & content moved from wiki after internal review \\ \hline
-1.0 & 27.02.2009 & Knud M\"oller (NUIG) & final version (D2.2.1) ready for external review \\ \hline
-2.0 & 27.02.2010 & Knud M\"oller (NUIG) & final version (D2.2.2) ready for external review \\ \hline
+1.0 & 27.02.2009 & \authorOne & final version (D2.2.1) ready for external review \\ \hline
+2.0 & 27.02.2010 & \authorOne & final version (D2.2.2) ready for external review \\ \hline
 \end{tabular}
@@
 The document comprises a discussion of the ontology development methodology used, a number of intermediate representations which document the development process, and the final product in the form of an OWL ontology.}\\ \hline
-	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne, \authorTwo}\\
+	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{p{10.85cm}|}{\authorOne, \authorTwo, \authorThree, \authorFour}\\
 %	{} & \multicolumn{4}{l|}{}\\ 
@@
 
-In this section, we will apply a number of the phases proposed in Methontology to the process of developing the FAST gadget ontology.
+In this section, we will apply a number of the phases proposed in Methontology (see Sect.
+\ref{sub:methontology_a_methodology_for_ontology_development}) to the process of developing the FAST gadget ontology.
 The names of the following sections reflect the names of the phases in Methontology.
@@
 I.e., if a term is listed as a \emph{class} here, this has no direct implications on the concrete implementation of this term.
-For example, while \emph{label} is a type of thing that is important in FAST, it will not necessarily be represented as an \texttt{rdfs:Class} in the implementation step.
-It could just as likely be represented as a property.
-
-\emph{Components/Resources}
-
-\begin{itemize}
-	\item \texttt{Component} (or \texttt{Resource}) - Anything that is part of a gadget (or the gadget itself). Tentatively anything that can be ``touched'' and moved around in the FAST IDE.
-	\begin{itemize}
-		\item \texttt{Screen\_Flow} - A set of screens from which a gadget for a given target platform can be generated.
-		\item \texttt{Screen} - An individual screen.
-		\item \texttt{Backend\_Service} - A Web service which provides data and/or functionality to a screen.
-A backend service will often be external to FAST, and will probably have to be wrapped by the screen.
-		\item \texttt{Flow\_Control\_Element} - Any kind of component which can restrict the default flow of screens in a gadget.
-		\begin{itemize}
-			\item \texttt{Screen\_Flow\_Start} - The entry point to a wigdet; the first screen.
-			\item \texttt{Screen\_Flow\_End} - A screen that ends the workflow of the gadget.
-			\item \texttt{Connector} - An explicit connection between two screens.
-		\end{itemize}
-		\item \texttt{Operator} - Any kind of component that is used to connect backend services to form elements.
-Examples are simple pipes, aggregators or various kinds of filters.
-Operators concern the inner mechanics of a screen and will be covered in years 2 and 3 of the project.
-		\item \texttt{Form\_Element} - Form elements are UI elements in a particular screen.
-	\end{itemize}
-\end{itemize}
-
-\emph{Pre- and Post-conditions}
-
-\begin{itemize}
-	\item \texttt{Condition} - The pre- or post-condition of either a screen or a screenflow.
-In the latter case, each target platform will use these conditions in its own way, or may also ignore them.
+For example, while \emph{label} is a type of thing that is important in FAST, it will not necessarily be represented as an \texttt{owl:Class} in the implementation step, but could just as well be represented as a property.
+
+\singlespacing
+\begin{small}
+\begin{longtable}{|p{4.25cm}|p{10cm}|}
+\caption{\label{tab:glossary_of_terms_classes}FAST Glossary of Terms, Classes}\\
+\hline
+\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Name}} & 
+	\textcolor{white}{\textbf{Description}} \\ \hline
+\endfirsthead
+\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Name}} & 
+	\textcolor{white}{\textbf{Description}} \\ \hline
+\endhead
+\multicolumn{2}{|c|}{\textbf{Gadgets, Components and Building Blocks}} \\ \hline
+\texttt{Gadget} & A FAST gadget is the wrapped and deployed instance of a screen flow.
+It is the end result of the FAST workflow.
+As such, it is an important part of the conceptual model of FAST, but will not be modelled as an entity in the FAST ontology.
+\\ \hline
+\texttt{Building\_Block} & Anything that is part of a gadget (or the gadget itself).
+Tentatively anything that can be ``touched'' and moved around in the FAST IDE, from the most complex units such as \texttt{ScreenFlow}s down to atomic \texttt{FormElement}s like a button or a label in a form.
+\\ \hline
+\texttt{Screen\_Flow} & A set of screens from which a gadget for a given target platform can be generated. \\ \hline
+\texttt{Screen} & An individual screen; the basic unit of user interaction in FAST. \\ \hline
+\texttt{Screen\_Component} & Screens are made up of screen components, which fundamentally include service \texttt{Resources}, \texttt{Operators} and \texttt{Forms}. \\ \hline
+\texttt{Resource} & A service resource in FAST is a wrapper around a Web service (the \texttt{BackendService}), which makes the service available to the platform, e.g., by mapping its definition to FAST facts and actions.
+Note that this is a re-definition of the term ``resource'', as compared to \cite{moeller2009fast_ontology}.
+\\ \hline
+\texttt{Backend\_Service} & A Web service which provides data and/or functionality to a screen.
+The actual backend service is external to FAST, and only available through a wrapper (the service \texttt{Resource}).
+\\ \hline
+\texttt{Operators} & Operators are intended to transform and/or modify data within a screen, usually for preparing data coming from service resources for the use in the screen's interface.
+Operators cover different kinds of data manipulations, from simple aggregation to mediating data with incompatible schemas.
+\\ \hline
+\texttt{Form} & A form is the visual aspect of a screen: its user interface. Each form is made up of individual form elements. \\ \hline
+\texttt{Form\_Element} & Form elements are UI elements in a particular \texttt{Screen}, such as buttons, lists or labels. \\ \hline
+\texttt{Pipe} & Pipes are used to explicitly define the flow of data within a screen, e.g., from service resource to operator to a specific form element. \\ \hline
+\texttt{Action} & Actions are representations of some specific functionality of a building block in FAST.
+Examples are methods of a Web service (e.g., \texttt{getItem}) or functionality to update or change the contents of a \texttt{Form}.
+\\ \hline
+\texttt{Trigger} & Triggers are the flip-side of actions.
+Certain events in a building blocks can cause a trigger to be fired.
+Other building blocks within the same screen, which are listening to it, will react with an \texttt{Action}.
+\\ \hline
+% \texttt{Flow\_Control\_Element} & A special kind of building block which can restrict the default flow of screens in a gadget. \\ \hline
+% \texttt{Screen\_Flow\_Start} & The entry point to a gadget; the first screen. \\ \hline
+% \texttt{Screen\_Flow\_End} & A screen that ends the workflow of the gadget. \\ \hline
+% \texttt{Connector} & An explicit connection between two screens, as opposed to the usual dynamic flow of screens defined by pre- and post-conditions. \\ \hline
+\multicolumn{2}{|c|}{\textbf{Pre- and Post-conditions}} \\ \hline
+\texttt{Condition} & The pre- or post-condition of either a certain kinds of building blocks.
+If the building block is a \texttt{Screen\_Flow}, each target platform will use these conditions in its own way, or may also ignore them.
 E.g., in EzWeb pre- and post-conditions correspond to the concepts of \emph{slot} and \emph{event}.
-	\item \texttt{Pattern} - A set of facts which formally defines a condition. A pattern may contain one or more variables.
-	\item \texttt{Fact} - ``the basic information unit of a FAST gadget''~\cite{solero2009fast_architecture} In terms of RDF, a fact will probably be one statement consisting of \emph{(S, P, O)}.
-\end{itemize}
-
-\emph{Annotation}
-
-\begin{itemize}
-	\item \texttt{String} - A catch-all class of objects that serves for representing short labels, longer descriptions, dates, patterns, etc.
+\\ \hline
+\texttt{Pattern} & A set of facts which formally defines a condition. A pattern may contain one or more variables. \\ \hline
+\texttt{Fact} & The ``basic information unit of a FAST gadget''~\cite{urena2010fast_architecture} In terms of RDF, a fact will probably be one statement consisting of \emph{(S, P, O)}. \\ \hline
+\multicolumn{2}{|c|}{\textbf{Annotation}} \\ \hline
+\texttt{String} & A catch-all class of objects that serves for representing short labels, longer descriptions, dates, patterns, etc.
 If meant for human consumption, strings can have multiple representations for different languages.
-	\item \texttt{Image} - Images are any kind of graphical object, such as icons, screenshots, etc.
-	\item \texttt{Tag} - A short (usually one word) description of a resource.
-\end{itemize}
-
-\emph{Users}
-
-\begin{itemize}
-	\item \texttt{User} - A human user of the FAST IDE.
-	\item \texttt{User\_Profile} - The settings and data known about a particular user.
-\end{itemize}
+\\ \hline
+\texttt{Image} & Images are any kind of graphical object, such as icons, screenshots, etc. \\ \hline
+\texttt{Tag} & A short (usually one word) description of a resource. \\ \hline
+\multicolumn{2}{|c|}{\textbf{User and User Profiles}} \\ \hline
+\texttt{User} & A human user of the FAST IDE. \\ \hline
+\texttt{User\_Profile} & The settings and data known about a particular user. \\ \hline
+\end{longtable}
+\end{small}
+\doublespacing
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics[scale=.66]{images/uml_composition.pdf}
+    \caption{Important composition relationships within the FAST conceptual model}
+    \label{fig:uml_composition}
+  \end{center}
+\end{figure}
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics[scale=.66]{images/uml_screen_component.pdf}
+    \caption{Aggregation of screen components}
+    \label{fig:uml_screen_component}
+  \end{center}
+\end{figure}
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics[scale=.66]{images/uml_definition.pdf}
+    \caption{Building blocks have either code or are defined declaratively}
+    \label{fig:uml_definition}
+  \end{center}
+\end{figure}
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics[scale=.66]{images/uml_condition.pdf}
+    \caption{Building blocks with pre- and post-conditions}
+    \label{fig:uml_condition}
+  \end{center}
+\end{figure}
 
@@
 \endhead
-\multicolumn{2}{|c|}{\textbf{Components/Resources}} \\ \hline
+\multicolumn{2}{|c|}{\textbf{Components/Building Blocks}} \\ \hline
 \texttt{contains} & Many kinds of components in FAST can contain other components: screenflows contain screens, screens contain forms or form elements, etc. \\ \hline
@@
 
-\begin{figure}
+\begin{figure}[ht]
   \begin{center}
@@
   \begin{center}
-    \includegraphics[width=\linewidth]{images/sioc_overview}
+    \includegraphics[width=0.75\linewidth]{images/sioc_overview}
     \caption{Overview of the SIOC ontology}
@@
 In practice, this procedure can be implemented using a SPARQL CONSTRUCT query, where the WHERE clause matches the graph pattern of the template, and the CONSTRUCT query will build the copy from it.
-The corresponding data on all three levels is shown in List.~\ref{list:templating}, using TriG syntax~\cite{bizer2004trig}.
+The corresponding data on all three levels is shown in List.~\ref{list:templating}, using TriG syntax~\cite{bizer2004trig} (an extension of Turtle syntax for named graphs).
 
@@
 
-For the FAST Gadget Ontology, most of the evaluation process will take place as part of the work carried out in WP6 (Experimentation and Evaluation) in years two and three of the project.
+For the FAST Gadget Ontology, most of the evaluation process will take place as part of the general implementation efforts in WPs 3--5, which will either directly or indirectly make use of the ontology and therefore function as means to test its feasibility, as well as providing new input to its evolution.
+Additionally, the work carried out in WP6 (Experimentation and Evaluation) in years two and three of the project is providing valuable input to the evaluation of the ontology.
 % subsubsection evaluation (end)
@@
 In Methontology, documentation is an activity which is automatically carried out throughout the development process, in the form of the various documents which form the output of the individual phases.
-Since this deliverable contains the current versions of all those documents, it is at the same time the ontology itself as well as its documentation.
+Since this deliverable contains the current versions of all those documents, it represents a snapshot of the most recent documentation available at the time of its writing.
 % subsubsection documentation (end)
@@
   \begin{center}
-    \includegraphics[width=.82\linewidth]{images/fgo_uml_diagram}
+    \includegraphics[width=.80\linewidth]{images/fgo_uml_diagram}
     \caption{FAST Gadget Ontology Class Hierarchy}
@@
 \singlespacing
-\begin{small}
-\verbatiminput{fast_gadget.owl}
-\end{small}
+% \begin{small}
+% \verbatiminput{fast_gadget.owl}
+% \end{small}
+\lstset{
+	frame=none,
+	caption={},
+	nolol=true,
+	basicstyle={\footnotesize\ttfamily},
+	label=list:ontology,
+	language=turtle
+}
+\lstinputlisting{fast_gadget.owl}
 

### 2010-02-20 trunk/deliverables/D2_2_Ontology/properties.tex
common tag added
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@463 d8b0c59b-6544-0410-94d9-83a5658a058e
greatly extending conceptualisation part with UML diagrams, also updating to reflect changes in implementations
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@582 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
-\subsubsection{Property: \texttt{fgo:contains}}
+\subsubsection*{Property: \texttt{fgo:contains}}
 \label{subs:contains}
@@
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource}, \htmlref{\texttt{fgo:Definition}}{subs:Definition} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Definition}}{subs:Definition}, \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
 \textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:ResourceReference}}{subs:ResourceReference} \\ \hline
 \end{tabular}
-\subsubsection{Property: \texttt{fgo:hasPreCondition}}
+\subsubsection*{Property: \texttt{fgo:hasAction}}
+\label{subs:hasAction}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has action \\ \hline
+\textcolor{white}{\textbf{description}} & This property indicates which actions are asociated and can be perfomed within 
+	a screen component. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Action}}{subs:Action} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasCode}}
+\label{subs:hasCode}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has code \\ \hline
+\textcolor{white}{\textbf{description}} & URL of the executable code of a particular screen. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Screen}}{subs:Screen} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{foaf:Document} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasCondition}}
+\label{subs:hasCondition}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has condition \\ \hline
+\textcolor{white}{\textbf{description}} & This property links a precondition or postcondition to a certain condition. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Postcondition}}{subs:Postcondition}, \htmlref{\texttt{fgo:Precondition}}{subs:Precondition} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition}, \texttt{rdf:Bag} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasDefinition}}
+\label{subs:hasDefinition}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has definition \\ \hline
+\textcolor{white}{\textbf{description}} & The structure and behaviour of a resource. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Definition}}{subs:Definition} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasIcon}}
+\label{subs:hasIcon}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has icon \\ \hline
+\textcolor{white}{\textbf{description}} & A small graphical representation of any FAST component or sub-component. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{foaf:Image} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasId}}
+\label{subs:hasId}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has id \\ \hline
+\textcolor{white}{\textbf{description}} & The internal ID of the resource within the container resource. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource}, \htmlref{\texttt{fgo:ResourceReference}}{subs:ResourceReference} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasIdActionTo}}
+\label{subs:hasIdActionTo}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has id action to \\ \hline
+\textcolor{white}{\textbf{description}} & action id of the `to' point of the pipe or trigger. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe}, \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasIdBBFrom}}
+\label{subs:hasIdBBFrom}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has id building block from \\ \hline
+\textcolor{white}{\textbf{description}} & building block id of the `from' point of the pipe or trigger. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe}, \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasIdBBTo}}
+\label{subs:hasIdBBTo}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has id building block to \\ \hline
+\textcolor{white}{\textbf{description}} & building block id of the `to' point of the pipe or trigger. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe}, \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasIdConditionFrom}}
+\label{subs:hasIdConditionFrom}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has id condition from \\ \hline
+\textcolor{white}{\textbf{description}} & condition id of the `from' point of the pipe. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasIdConditionTo}}
+\label{subs:hasIdConditionTo}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has id condition to \\ \hline
+\textcolor{white}{\textbf{description}} & condition id of the `to' point of the pipe. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasLanguage}}
+\label{subs:hasLanguage}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has language string \\ \hline
+\textcolor{white}{\textbf{description}} & This property is the language the library is written in. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Library}}{subs:Library} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasLibrary}}
+\label{subs:hasLibrary}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has library \\ \hline
+\textcolor{white}{\textbf{description}} & This property indicates which libraries are used by a screen component. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Library}}{subs:Library} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasName}}
+\label{subs:hasName}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has name \\ \hline
+\textcolor{white}{\textbf{description}} & The name the user gives to the resource. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasNameFrom}}
+\label{subs:hasNameFrom}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has name from \\ \hline
+\textcolor{white}{\textbf{description}} & name of the trigger. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasPattern}}
+\label{subs:hasPattern}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has pattern \\ \hline
+\textcolor{white}{\textbf{description}} & This property links a screen or screenflow to its pre-condition, 
+    i.e., the facts that need to be fulfilled in order for this screen or screenflow 
+    to be reachable. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Fact}}{subs:Fact} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{rdfs:Resource} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasPatternString}}
+\label{subs:hasPatternString}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has pattern string \\ \hline
+\textcolor{white}{\textbf{description}} & This property is the textual representation of a condition. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Fact}}{subs:Fact} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasPostCondition}}
+\label{subs:hasPostCondition}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has post-condition \\ \hline
+\textcolor{white}{\textbf{description}} & This property links certain type of resources to its post-condition, 
+    i.e., the facts that are produced once the screen or screenflow has been 
+    executed. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:WithPostConditions}}{subs:WithPostConditions} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition}, \texttt{rdf:Bag} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasPreCondition}}
 \label{subs:hasPreCondition}
@@
 \end{tabular}
-\subsubsection{Property: \texttt{fgo:hasPostCondition}}
-\label{subs:hasPostCondition}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has post-condition \\ \hline
-\textcolor{white}{\textbf{description}} & This property links certain type of resources to its post-condition, 
-    i.e., the facts that are produced once the screen or screenflow has been 
-    executed. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:WithPostConditions}}{subs:WithPostConditions} \\ \hline
-\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition}, \texttt{rdf:Bag} \\ \hline
-\end{tabular}
-\subsubsection{Property: \texttt{fgo:hasPattern}}
-\label{subs:hasPattern}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has pattern \\ \hline
-\textcolor{white}{\textbf{description}} & This property links a screen or screenflow to its pre-condition, 
-    i.e., the facts that need to be fulfilled in order for this screen or screenflow 
-    to be reachable. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Fact}}{subs:Fact} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{rdfs:Resource} \\ \hline
-\end{tabular}
-\subsubsection{Property: \texttt{fgo:hasIcon}}
-\label{subs:hasIcon}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has icon \\ \hline
-\textcolor{white}{\textbf{description}} & A small graphical representation of any FAST component or sub-component. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{foaf:Image} \\ \hline
-\end{tabular}
-\subsubsection{Property: \texttt{fgo:hasScreenshot}}
+\subsubsection*{Property: \texttt{fgo:hasScreenshot}}
 \label{subs:hasScreenshot}
@@
 \end{tabular}
-\subsubsection{Property: \texttt{fgo:hasCode}}
-\label{subs:hasCode}
+\subsubsection*{Property: \texttt{fgo:hasSource}}
+\label{subs:hasSource}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & has code \\ \hline
-\textcolor{white}{\textbf{description}} & URL of the executable code of a particular screen. \\ \hline
+\textcolor{white}{\textbf{label}} & has source \\ \hline
+\textcolor{white}{\textbf{description}} & URL of the source code of a particular library. \\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Screen}}{subs:Screen} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Library}}{subs:Library} \\ \hline
 \textcolor{white}{\textbf{range}} & \texttt{foaf:Document} \\ \hline
 \end{tabular}
-\subsubsection{Property: \texttt{fgo:hasCondition}}
-\label{subs:hasCondition}
+\subsubsection*{Property: \texttt{fgo:hasTag}}
+\label{subs:hasTag}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & has condition \\ \hline
-\textcolor{white}{\textbf{description}} & This property links a precondition or postcondition to a certain condition. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Precondition}}{subs:Precondition}, \htmlref{\texttt{fgo:Postcondition}}{subs:Postcondition} \\ \hline
-\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition}, \texttt{rdf:Bag} \\ \hline
+\textcolor{white}{\textbf{label}} & has tag \\ \hline
+\textcolor{white}{\textbf{description}} & A tag is used to annotate keywords of a particular resource. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
 \end{tabular}
-\subsubsection{Property: \texttt{fgo:hasAction}}
-\label{subs:hasAction}
+\subsubsection*{Property: \texttt{fgo:hasTrigger}}
+\label{subs:hasTrigger}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & has action \\ \hline
-\textcolor{white}{\textbf{description}} & This property indicates which actions are asociated and can be perfomed within 
-	a screen component. \\ \hline
+\textcolor{white}{\textbf{label}} & has trigger string \\ \hline
+\textcolor{white}{\textbf{description}} & This property represents the events fired within a building block. e.g. a 
+	button to clear the shopping cart will fired a "clear cart" trigger. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasType}}
+\label{subs:hasType}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has type \\ \hline
+\textcolor{white}{\textbf{description}} & The type of the resource. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasUri}}
+\label{subs:hasUri}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has uri \\ \hline
+\textcolor{white}{\textbf{description}} & URI of a particular resource pointing from a reference within
+	another resource. \\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
-\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Action}}{subs:Action} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:ResourceReference}}{subs:ResourceReference} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{foaf:Document} \\ \hline
 \end{tabular}
-\subsubsection{Property: \texttt{fgo:hasUse}}
+\subsubsection*{Property: \texttt{fgo:hasUse}}
 \label{subs:hasUse}
@@
 \end{tabular}
-\subsubsection{Property: \texttt{fgo:hasLibrary}}
-\label{subs:hasLibrary}
+\subsubsection*{Property: \texttt{fgo:hasVersion}}
+\label{subs:hasVersion}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & has library \\ \hline
-\textcolor{white}{\textbf{description}} & This property indicates which libraries are used by a screen component. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
-\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Library}}{subs:Library} \\ \hline
+\textcolor{white}{\textbf{label}} & has version \\ \hline
+\textcolor{white}{\textbf{description}} & The version of a particular screen or screenflow, such as 1.0, 0.2beta, etc. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
 \end{tabular}
-\subsubsection{Property: \texttt{fgo:hasSource}}
-\label{subs:hasSource}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has source \\ \hline
-\textcolor{white}{\textbf{description}} & URL of the source code of a particular library. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Library}}{subs:Library} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{foaf:Document} \\ \hline
-\end{tabular}
-\subsubsection{Property: \texttt{fgo:hasUri}}
-\label{subs:hasUri}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has uri \\ \hline
-\textcolor{white}{\textbf{description}} & URI of a particular resource pointing from a reference within
-	another resource. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:ResourceReference}}{subs:ResourceReference} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{foaf:Document} \\ \hline
-\end{tabular}
-\subsubsection{Property: \texttt{fgo:hasDefinition}}
-\label{subs:hasDefinition}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has definition \\ \hline
-\textcolor{white}{\textbf{description}} & The structure and behaviour of a resource. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Definition}}{subs:Definition} \\ \hline
-\end{tabular}
-\subsubsection{Property: \texttt{fgo:integratesTerm}}
+\subsubsection*{Property: \texttt{fgo:integratesTerm}}
 \label{subs:integratesTerm}
@@
 \textcolor{white}{\textbf{description}} & A way to explicitly say that an ontology uses terms from another namespace. 
-    Maybe a bit redundant, but why not. \\ \hline
+    Maybe a bit redundant, but why not. TODO: Must be moved to a different namespace! \\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
 \end{tabular}
-\subsubsection{Property: \texttt{fgo:hasPatternString}}
-\label{subs:hasPatternString}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has pattern string \\ \hline
-\textcolor{white}{\textbf{description}} & This property is the textual representation of a condition. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Fact}}{subs:Fact} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection{Property: \texttt{fgo:isPositive}}
+\subsubsection*{Property: \texttt{fgo:isPositive}}
 \label{subs:isPositive}
@@
 \end{tabular}
-\subsubsection{Property: \texttt{fgo:hasTag}}
-\label{subs:hasTag}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has tag \\ \hline
-\textcolor{white}{\textbf{description}} & A tag is used to annotate keywords of a particular resource. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection{Property: \texttt{fgo:hasVersion}}
-\label{subs:hasVersion}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has version \\ \hline
-\textcolor{white}{\textbf{description}} & The version of a particular screen or screenflow, such as 1.0, 0.2beta, etc. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection{Property: \texttt{fgo:hasTrigger}}
-\label{subs:hasTrigger}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has trigger string \\ \hline
-\textcolor{white}{\textbf{description}} & This property represents the events fired within a building block. e.g. a 
-	button to clear the shopping cart will fired a "clear cart" trigger. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection{Property: \texttt{fgo:hasLanguage}}
-\label{subs:hasLanguage}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has language string \\ \hline
-\textcolor{white}{\textbf{description}} & This property is the language the library is written in. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Library}}{subs:Library} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection{Property: \texttt{fgo:hasId}}
-\label{subs:hasId}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has id \\ \hline
-\textcolor{white}{\textbf{description}} & The internal ID of the resource within the container resource. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource}, \htmlref{\texttt{fgo:ResourceReference}}{subs:ResourceReference} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection{Property: \texttt{fgo:hasName}}
-\label{subs:hasName}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has name \\ \hline
-\textcolor{white}{\textbf{description}} & The name the user gives to the resource. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection{Property: \texttt{fgo:hasType}}
-\label{subs:hasType}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has type \\ \hline
-\textcolor{white}{\textbf{description}} & The type of the resource. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection{Property: \texttt{fgo:hasIdBBFrom}}
-\label{subs:hasIdBBFrom}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has id building block from \\ \hline
-\textcolor{white}{\textbf{description}} & building block id of the `from' point of the pipe or trigger. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe}, \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection{Property: \texttt{fgo:hasIdConditionFrom}}
-\label{subs:hasIdConditionFrom}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has id condition from \\ \hline
-\textcolor{white}{\textbf{description}} & condition id of the `from' point of the pipe. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection{Property: \texttt{fgo:hasIdBBTo}}
-\label{subs:hasIdBBTo}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has id building block to \\ \hline
-\textcolor{white}{\textbf{description}} & building block id of the `to' point of the pipe or trigger. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe}, \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection{Property: \texttt{fgo:hasIdConditionTo}}
-\label{subs:hasIdConditionTo}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has id condition to \\ \hline
-\textcolor{white}{\textbf{description}} & condition id of the `to' point of the pipe. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection{Property: \texttt{fgo:hasIdActionTo}}
-\label{subs:hasIdActionTo}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has id action to \\ \hline
-\textcolor{white}{\textbf{description}} & action id of the `to' point of the pipe or trigger. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe}, \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection{Property: \texttt{fgo:hasNameFrom}}
-\label{subs:hasNameFrom}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has name from \\ \hline
-\textcolor{white}{\textbf{description}} & name of the trigger. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}

### 2010-02-20 trunk/deliverables/D2_2_Ontology/fast_d2_2_2_ontology.tex
greatly extending conceptualisation part with UML diagrams, also updating to reflect changes in implementations
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@583 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 The glossary is seeded from the specification document (see previous section), but groups the terms according to relatedness.
-Also, new terms are added to this document, whereas the specification document stays the same.
+Also, new terms are added to this document, whereas the specification document mostly remains in its original state.
 
@@
 For example, while \emph{label} is a type of thing that is important in FAST, it will not necessarily be represented as an \texttt{owl:Class} in the implementation step, but could just as well be represented as a property.
+Table~\ref{tab:glossary_of_terms_classes} lists all classes considered in the conceptual model of FAST, loosely grouped according to their relevance for specific aspects of the platform.
 
@@
 \multicolumn{2}{|c|}{\textbf{Gadgets, Components and Building Blocks}} \\ \hline
-\texttt{Gadget} & A FAST gadget is the wrapped and deployed instance of a screen flow.
+\texttt{Gadget} & A FAST gadget is the wrapped and deployed instance of a \texttt{Screen\_Flow}.
 It is the end result of the FAST workflow.
-As such, it is an important part of the conceptual model of FAST, but will not be modelled as an entity in the FAST ontology.
+As such, it is an important part of the conceptual model of FAST.
+However, it will not be modelled as an entity in the FAST ontology, since it does not does need to be handled by either GVS or catalogue.
 \\ \hline
 \texttt{Building\_Block} & Anything that is part of a gadget (or the gadget itself).
-Tentatively anything that can be ``touched'' and moved around in the FAST IDE, from the most complex units such as \texttt{ScreenFlow}s down to atomic \texttt{FormElement}s like a button or a label in a form.
+Tentatively anything that can be ``touched'' and moved around in the FAST IDE, from the most complex units such as \texttt{Screen\_Flow}s down to atomic \texttt{Form\_Element}s like a button or a label in a form.
 \\ \hline
@@
 \\ \hline
+\multicolumn{2}{|c|}{\textbf{Pre- and Post-conditions}} \\ \hline
+\texttt{Condition} & The pre- or post-condition of either a certain kinds of building blocks.
+If the building block is a \texttt{Screen\_Flow}, each target platform will use these conditions in its own way, or may also ignore them.
+E.g., in EzWeb pre- and post-conditions correspond to the concepts of \emph{slot} and \emph{event}.
+\\ \hline
+\texttt{Pattern} & A set of facts which formally defines a condition. A pattern may contain one or more variables. \\ \hline
+\texttt{Fact} & The ``basic information unit of a FAST gadget''~\cite{urena2010fast_architecture} In terms of RDF, a fact will probably be one statement consisting of \emph{(S, P, O)}. \\ \hline
+\multicolumn{2}{|c|}{\textbf{Implementation}} \\ \hline
+\texttt{Code} & In case the implementation of a building block is hard-coded, the \texttt{Code} represents the source code that defines it. \\ \hline
+\texttt{Library} & For hard-coded building blocks, certain programming libraries may be necessary. The are represented as a \texttt{Library}. \\ \hline
+\texttt{Definition} & For building blocks which do not rely on a hard-coded implementation, but which are instead defined declaratively, the \texttt{Definition} represents this. \\ \hline
 % \texttt{Flow\_Control\_Element} & A special kind of building block which can restrict the default flow of screens in a gadget. \\ \hline
@@
 % \texttt{Connector} & An explicit connection between two screens, as opposed to the usual dynamic flow of screens defined by pre- and post-conditions. \\ \hline
-\multicolumn{2}{|c|}{\textbf{Pre- and Post-conditions}} \\ \hline
-\texttt{Condition} & The pre- or post-condition of either a certain kinds of building blocks.
-If the building block is a \texttt{Screen\_Flow}, each target platform will use these conditions in its own way, or may also ignore them.
-E.g., in EzWeb pre- and post-conditions correspond to the concepts of \emph{slot} and \emph{event}.
-\\ \hline
-\texttt{Pattern} & A set of facts which formally defines a condition. A pattern may contain one or more variables. \\ \hline
-\texttt{Fact} & The ``basic information unit of a FAST gadget''~\cite{urena2010fast_architecture} In terms of RDF, a fact will probably be one statement consisting of \emph{(S, P, O)}. \\ \hline
 \multicolumn{2}{|c|}{\textbf{Annotation}} \\ \hline
@@
 
+Regarding the layering from the actual gadget down to the embedded Web services, the central classes in the conceptual model of FAST can be grouped into three levels: the gadget and screen flow level at the top, the level of individual screens in the middle, and the level of Web services at the bottom.
+This layering is reflected in Fig.~\ref{fig:uml_composition}, which illustrates the central compositional relationships of the model.
+As the figure shows, a \emph{gadget} in FAST is mainly composed of a \emph{screen flow}, which it wraps.
+The \emph{screen flow} is composed of one or more \emph{screens}, which is turn comprises the three \emph{screen components} of the visual front-end (the \emph{form}), the back end service (the \emph{resource}) and the data \emph{operator}.
+Finally, \emph{forms} are assembled from \emph{form elements}, while \emph{resources} wrap \emph{backend services}.
+
 \begin{figure}[ht]
@@
 
+Figure~\ref{fig:uml_screen_component} looks closer at the features of individual screens.
+These are forms, resources and operators, which are all specialisations of the general \emph{screen component} concept.
+Each such screen component may or may not have an action or trigger associated with them, which declaratively define their behaviour in terms of their own functionality or functionality they can trigger in other building blocks.
+
 \begin{figure}[ht]
@@
 
+Regarding their implementation, different kinds of building blocks in FAST can either be defined hard-coded as a piece of source code, or they can be defined declaratively in the terms of the FAST ontology (using pipes, operators, etc.).
+In the former case, a building block such as a screen would have been implemented and added to the FAST platform by an engineer, whereas in the latter case, ordinary users of the FAST platform would have assembled the building block using the tools available in the GVS\footnote{Currently, only screens can be assembled in the GVS.
+Forms may or may not be editable in the GVS in the future.
+Other kinds of building blocks are not planned to be editable.}.
+This principle is reflected in Fig.~\ref{fig:uml_definition}, showing how certain kinds of building blocks aggregate either a \texttt{Definition} or \texttt{Code}, whereas other kinds of building blocks are always hard-coded.
+
 \begin{figure}[ht]
@@
 \end{figure}
+
+Figure~\ref{fig:uml_condition} illustrates how different building blocks relate to the concepts of pre- and post-conditions.
+While all building blocks can essentially have such conditions, only screens and screen flows directly aggregate both kinds of conditions.
+However, in the case of the three different kinds of screen components (form, resource and operator), the relation to a pre-condition is only established through their actions, which represent their basic functionality.
+For all building blocks, both pre- and post-conditions are entirely optional.
 
@@
 In this section we will investigate out a number of established ontologies which cover part of the requirements laid out in the specification document, as well as the glossary of terms from the conceptualisation stage.
-In particular, we will look at the Friend of a Friend (FOAF) ontology, Semantically Interlinked Online Communities (SIOC) and the Dublin Core metadata specification.
-The outcome of this process will be a living \emph{integration document} at \ref{ssub:integration_document}.
+In particular, we will look at the Friend of a Friend (FOAF) ontology, Semantically Interlinked Online Communities (SIOC), the Dublin Core metadata specification and Common Tag.
+Looking at the integrated vocabularies and ontologies, it becomes apparent that we integrate terms for those parts of the conceptual model which are not unique for FAST, i.e., users and user profiles, basic annotation needs or the tagging of building blocks.
+Aspects of the model which are unique to FAST, however, had been modelled from scratch.
+\todo{idea: move these last sentences to the conclusion.}
+The outcome of the integration process is the living \emph{integration document} at \ref{ssub:integration_document}.
 
@@
 \begin{small}
-\begin{longtable}{|p{2.6cm}|p{1.5cm}|p{3cm}|p{7cm}|}
+\begin{longtable}{|p{3cm}|p{1.5cm}|p{3.5cm}|p{6.1cm}|}
 \caption{\label{tab:integration_document}FAST Ontology integration document}\\
@@
 \texttt{hasDescription} & DC & \texttt{dcterms:description} & --- \\ \hline
-\texttt{hasCreationDate} & DC & \texttt{dcterms:created} & Dates should be formatted according to ISO 8601. \\ \hline
+\texttt{hasCreationDate} & DC & \texttt{dcterms:created} & Dates should be formatted according to ISO 8601~\cite{w3c_iso8601_1997}. \\ \hline
 \texttt{hasRights} & DC & \texttt{dcterms:rights} & DC terms specify a number of classes to model the rights of resources.
@@
 \\ \hline
-\texttt{hasDomainContext} & DC & \texttt{dcterms:subject} & DC specifies that subject should be non-literals. To allow simple tagging, we should represent tags as resources (TODO). \\ \hline
+\texttt{hasDomainContext} & DC & \texttt{dcterms:subject} & DC specifies that subject shoulds be non-literals.
+To allow simple tagging, we should represent tags as resources (\todo{common tag}).
+\\ \hline
 \texttt{hasImage} & FOAF & \texttt{foaf:depiction} & The inverse \texttt{foaf:depicts} could also be used. \\ \hline
@@
 Conceptually, once would like to say that $Screen$ is a class, i.e., the set of all screens.
-As argues above, $LogInScreen$ would then be an instance of $Screen$ ($LogInScreen \subset Screen$), while a particular $LogInScreen_x$ in a particular screenflow would again be an instance of $LogInScreen$ ($LogInScreen_x \subset LogInScreen$).
+As argued above, $LogInScreen$ would then be an instance of $Screen$ ($LogInScreen \subset Screen$), while a particular $LogInScreen_x$ in a particular screenflow would again be an instance of $LogInScreen$ ($LogInScreen_x \subset LogInScreen$).
 However, unless we want to move into OWL Full semantics, classes cannot be used as instances at the same time, so we cannot use this modelling approach.
@@
 
+The FAST ontology is continuously being evaluated through its direct and indirect employment in the implementation work packages WP3-5, as well as through an ongoing dialogue between the ontology designers and the platform implementers.
+The use case-based evaluation performed in WP6 provides additional input to this process.
+This kind of evaluation has lead to the gradual change, extension and adjustment of the documents produced in the different stages in the ontology development process, such as the glossary of terms in the conceptualisation phase, the integration document and the formal OWL ontology document.
+In addition to this kind of continuous evaluation we have also looked into other kinds of metrics ontology evaluation, as discussed in the following paragraphs.
+
 \subsection{Usability as a General Design Principle} % (fold)

### 2010-02-20 trunk/deliverables/D2_2_Ontology/fast_d2_2_2_ontology.tex
rearranged conceptualisation section
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@584 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 In the conceptualisation phase, we first produce a \emph{glossary of terms} which lists all classes and properties of our ontology.
-The glossary is seeded from the specification document (see previous section), but groups the terms according to relatedness.
-Also, new terms are added to this document, whereas the specification document mostly remains in its original state.
+After introducing the general terms, the conceptual model is then refined by relating the different terms with regards to type hierarchy, composition and aggregation.
 
@@
 \label{ssub:glossary_of_terms}
+
+The glossary of terms is seeded from the specification document (see previous section), but groups the terms according to relatedness.
+Also, the evolution of the ontology is mainly reflected in this document, as new terms being introduced based on the employment and ongoing evaluation of the ontology are added here, whereas the specification document remains in its original state.
 
@@
 
-Regarding the layering from the actual gadget down to the embedded Web services, the central classes in the conceptual model of FAST can be grouped into three levels: the gadget and screen flow level at the top, the level of individual screens in the middle, and the level of Web services at the bottom.
-This layering is reflected in Fig.~\ref{fig:uml_composition}, which illustrates the central compositional relationships of the model.
-As the figure shows, a \emph{gadget} in FAST is mainly composed of a \emph{screen flow}, which it wraps.
-The \emph{screen flow} is composed of one or more \emph{screens}, which is turn comprises the three \emph{screen components} of the visual front-end (the \emph{form}), the back end service (the \emph{resource}) and the data \emph{operator}.
-Finally, \emph{forms} are assembled from \emph{form elements}, while \emph{resources} wrap \emph{backend services}.
-
-\begin{figure}[ht]
-  \begin{center}
-    \includegraphics[scale=.66]{images/uml_composition.pdf}
-    \caption{Important composition relationships within the FAST conceptual model}
-    \label{fig:uml_composition}
-  \end{center}
-\end{figure}
-
-Figure~\ref{fig:uml_screen_component} looks closer at the features of individual screens.
-These are forms, resources and operators, which are all specialisations of the general \emph{screen component} concept.
-Each such screen component may or may not have an action or trigger associated with them, which declaratively define their behaviour in terms of their own functionality or functionality they can trigger in other building blocks.
-
-\begin{figure}[ht]
-  \begin{center}
-    \includegraphics[scale=.66]{images/uml_screen_component.pdf}
-    \caption{Aggregation of screen components}
-    \label{fig:uml_screen_component}
-  \end{center}
-\end{figure}
-
-Regarding their implementation, different kinds of building blocks in FAST can either be defined hard-coded as a piece of source code, or they can be defined declaratively in the terms of the FAST ontology (using pipes, operators, etc.).
-In the former case, a building block such as a screen would have been implemented and added to the FAST platform by an engineer, whereas in the latter case, ordinary users of the FAST platform would have assembled the building block using the tools available in the GVS\footnote{Currently, only screens can be assembled in the GVS.
-Forms may or may not be editable in the GVS in the future.
-Other kinds of building blocks are not planned to be editable.}.
-This principle is reflected in Fig.~\ref{fig:uml_definition}, showing how certain kinds of building blocks aggregate either a \texttt{Definition} or \texttt{Code}, whereas other kinds of building blocks are always hard-coded.
-
-\begin{figure}[ht]
-  \begin{center}
-    \includegraphics[scale=.66]{images/uml_definition.pdf}
-    \caption{Building blocks have either code or are defined declaratively}
-    \label{fig:uml_definition}
-  \end{center}
-\end{figure}
-
-Figure~\ref{fig:uml_condition} illustrates how different building blocks relate to the concepts of pre- and post-conditions.
-While all building blocks can essentially have such conditions, only screens and screen flows directly aggregate both kinds of conditions.
-However, in the case of the three different kinds of screen components (form, resource and operator), the relation to a pre-condition is only established through their actions, which represent their basic functionality.
-For all building blocks, both pre- and post-conditions are entirely optional.
-
-\begin{figure}[ht]
-  \begin{center}
-    \includegraphics[scale=.66]{images/uml_condition.pdf}
-    \caption{Building blocks with pre- and post-conditions}
-    \label{fig:uml_condition}
-  \end{center}
-\end{figure}
-
 % paragraph classes (end)
@@
 
-All properties are listed in Tab.~\ref{tab:glossary_of_terms_properties} (many of the properties covered in this table will also have an inverse).
+All relevant properties of the FAST conceptual model are listed in Tab.~\ref{tab:glossary_of_terms_properties}.
+Many of the properties covered in this table will also have an inverse, which are not listed.
 
@@
 \doublespacing
+
+\subsubsection{Term Relations within the Conceptual Model} % (fold)
+\label{ssub:term_relations_within_the_conceptual_model}
+
+Regarding the layering from the actual gadget down to the embedded Web services, the central classes in the conceptual model of FAST can be grouped into three levels: the gadget and screen flow level at the top, the level of individual screens in the middle, and the level of Web services at the bottom.
+This layering is reflected in Fig.~\ref{fig:uml_composition}, which illustrates the central compositional relationships of the model.
+As the figure shows, a \emph{gadget} in FAST is mainly composed of a \emph{screen flow}, which it wraps.
+The \emph{screen flow} is composed of one or more \emph{screens}, which is turn comprises the three \emph{screen components} of the visual front-end (the \emph{form}), the back end service (the \emph{resource}) and the data \emph{operator}.
+Finally, \emph{forms} are assembled from \emph{form elements}, while \emph{resources} wrap \emph{backend services}.
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics[scale=.66]{images/uml_composition.pdf}
+    \caption{Important composition relationships within the FAST conceptual model}
+    \label{fig:uml_composition}
+  \end{center}
+\end{figure}
+
+Figure~\ref{fig:uml_screen_component} looks closer at the features of individual screens.
+These are forms, resources and operators, which are all specialisations of the general \emph{screen component} concept.
+Each such screen component may or may not have an action or trigger associated with them, which declaratively define their behaviour in terms of their own functionality or functionality they can trigger in other building blocks.
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics[scale=.66]{images/uml_screen_component.pdf}
+    \caption{Aggregation of screen components}
+    \label{fig:uml_screen_component}
+  \end{center}
+\end{figure}
+
+Regarding their implementation, different kinds of building blocks in FAST can either be defined hard-coded as a piece of source code, or they can be defined declaratively in the terms of the FAST ontology (using pipes, operators, etc.).
+In the former case, a building block such as a screen would have been implemented and added to the FAST platform by an engineer, whereas in the latter case, ordinary users of the FAST platform would have assembled the building block using the tools available in the GVS\footnote{Currently, only screens can be assembled in the GVS.
+Forms may or may not be editable in the GVS in the future.
+Other kinds of building blocks are not planned to be editable.}.
+This principle is reflected in Fig.~\ref{fig:uml_definition}, showing how certain kinds of building blocks aggregate either a \texttt{Definition} or \texttt{Code}, whereas other kinds of building blocks are always hard-coded.
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics[scale=.66]{images/uml_definition.pdf}
+    \caption{Building blocks have either code or are defined declaratively}
+    \label{fig:uml_definition}
+  \end{center}
+\end{figure}
+
+Figure~\ref{fig:uml_condition} illustrates how different building blocks relate to the concepts of pre- and post-conditions.
+While all building blocks can essentially have such conditions, only screens and screen flows directly aggregate both kinds of conditions.
+However, in the case of the three different kinds of screen components (form, resource and operator), the relation to a pre-condition is only established through their actions, which represent their basic functionality.
+For all building blocks, both pre- and post-conditions are entirely optional.
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics[scale=.66]{images/uml_condition.pdf}
+    \caption{Building blocks with pre- and post-conditions}
+    \label{fig:uml_condition}
+  \end{center}
+\end{figure}
+
+% subsubsection term_relations_within_the_conceptual_model (end)
 

### 2010-02-22 trunk/deliverables/D2_2_Ontology/fast_d2_2_2_ontology.tex
cleared todos, rearranged document (overview UML no longer at the end), extended conclusion
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@585 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \newboolean{showcomments}
-\setboolean{showcomments}{true}
+\setboolean{showcomments}{false}
 \ifthenelse{\boolean{showcomments}}
@@
 In the conceptualisation phase, we first produce a \emph{glossary of terms} which lists all classes and properties of our ontology.
-After introducing the general terms, the conceptual model is then refined by relating the different terms with regards to type hierarchy, composition and aggregation.
-
-\subsubsection{Glossary of Terms} % (fold)
-\label{ssub:glossary_of_terms}
-
 The glossary of terms is seeded from the specification document (see previous section), but groups the terms according to relatedness.
 Also, the evolution of the ontology is mainly reflected in this document, as new terms being introduced based on the employment and ongoing evaluation of the ontology are added here, whereas the specification document remains in its original state.
-
-\paragraph{Classes} % (fold)
-\label{par:classes}
+After introducing the general terms, the conceptual model is then refined by relating the different terms with regards to type hierarchy, composition and aggregation.
+
+\subsubsection{Glossary of Terms: Classes} % (fold)
+\label{subs:classes}
 
 ``Classes'' here should be read as ``types of things'' in general.
-I.e., if a term is listed as a \emph{class} here, this has no direct implications on the concrete implementation of this term.
+I.e., if a term is listed as a \emph{class} here, this has no direct implications on the concrete implementation of the term.
 For example, while \emph{label} is a type of thing that is important in FAST, it will not necessarily be represented as an \texttt{owl:Class} in the implementation step, but could just as well be represented as a property.
@@
 \texttt{Screen} & An individual screen; the basic unit of user interaction in FAST. \\ \hline
-\texttt{Screen\_Component} & Screens are made up of screen components, which fundamentally include service \texttt{Resources}, \texttt{Operators} and \texttt{Forms}. \\ \hline
+\texttt{Screen\_Component} & Screens are made up of screen components, which fundamentally include service \texttt{Resource}s, \texttt{Operator}s and \texttt{Form}s. \\ \hline
 \texttt{Resource} & A service resource in FAST is a wrapper around a Web service (the \texttt{BackendService}), which makes the service available to the platform, e.g., by mapping its definition to FAST facts and actions.
-Note that this is a re-definition of the term ``resource'', as compared to \cite{moeller2009fast_ontology}.
+Note that this meaning of the term ``resource'' is different from the previous version of the FAST ontology in \cite{moeller2009fast_ontology}.
 \\ \hline
@@
 \texttt{Trigger} & Triggers are the flip-side of actions.
-Certain events in a building blocks can cause a trigger to be fired.
+Certain events in a building block can cause a trigger to be fired.
 Other building blocks within the same screen, which are listening to it, will react with an \texttt{Action}.
@@
 \multicolumn{2}{|c|}{\textbf{Pre- and Post-conditions}} \\ \hline
-\texttt{Condition} & The pre- or post-condition of either a certain kinds of building blocks.
+\texttt{Condition} & The pre- or post-condition of a certain kinds of building block.
 If the building block is a \texttt{Screen\_Flow}, each target platform will use these conditions in its own way, or may also ignore them.
@@
 \\ \hline
+\texttt{Pre-condition} & The pre-condition of a building block comprises of a set of \texttt{Fact}s that need to be fulfilled or available in order for the building block to be activated. \\ \hline
+\texttt{Post-condition} & The post-condition of a building block is a set of \texttt{Fact}s it can produce, and which will be thus become available to other building blocks. \\ \hline
 \texttt{Pattern} & A set of facts which formally defines a condition. A pattern may contain one or more variables. \\ \hline
-\texttt{Fact} & The ``basic information unit of a FAST gadget''~\cite{urena2010fast_architecture} In terms of RDF, a fact will probably be one statement consisting of \emph{(S, P, O)}. \\ \hline
+\texttt{Fact} & The ``basic information unit of a FAST gadget''~\cite{urena2010fast_architecture}.
+In terms of RDF, a fact is a statement consisting of a subject, predicate and object \emph{(S, P, O)}.
+\\ \hline
+\texttt{Ontology} & \texttt{Fact}s in FAST are represented as RDF triples, using terminology from a particular vocabulary or ontology.
+A building block can explicitly specify which ontology or ontologies it uses to specify its pre- and post-conditions (e.g., in the case of a \texttt{Screen}), or which ontology or ontologies it can otherwise handle (e.g., in terms of a mediation \texttt{Operator}).
+\\ \hline
 \multicolumn{2}{|c|}{\textbf{Implementation}} \\ \hline
 \texttt{Code} & In case the implementation of a building block is hard-coded, the \texttt{Code} represents the source code that defines it. \\ \hline
-\texttt{Library} & For hard-coded building blocks, certain programming libraries may be necessary. The are represented as a \texttt{Library}. \\ \hline
-\texttt{Definition} & For building blocks which do not rely on a hard-coded implementation, but which are instead defined declaratively, the \texttt{Definition} represents this. \\ \hline
+\texttt{Library} & For hard-coded building blocks, certain programming libraries may be necessary. The are represented as instances of \texttt{Library}. \\ \hline
+\texttt{Definition} & For building blocks which do not rely on a hard-coded implementation, but which are instead defined declaratively, the definition is represented by this class. \\ \hline
 % \texttt{Flow\_Control\_Element} & A special kind of building block which can restrict the default flow of screens in a gadget. \\ \hline
@@
 % \texttt{Connector} & An explicit connection between two screens, as opposed to the usual dynamic flow of screens defined by pre- and post-conditions. \\ \hline
+\multicolumn{2}{|c|}{\textbf{User and User Profiles}} \\ \hline
+\texttt{User} & A human user involved in the lifecycle of a FAST gadget.
+The kind of \texttt{User} in FAST that is most relevant to the gadget ontology is a user of the FAST IDE, since these users need to be represented and handled by the FAST platform directly (as opposed to gadget end users).
+\\ \hline
+\texttt{User\_Profile} & The settings and data known about a particular user of the FAST platform. \\ \hline
+\texttt{User\_Role} & Throughout the lifecycle of a FAST gadget different kinds of users are involved.
+These users are distinguished by playing a particular \texttt{User\_Role}.
+The particular user roles identified in \cite{villoslada2010fast_requirements} are \emph{end users} of a gadget, \emph{key users}, \emph{consultants} (i.e., gadget developers), \emph{screen developers} and \emph{resource developers}.
+Any particular \texttt{User} can theoretically play several roles at the same time.
+\\ \hline
 \multicolumn{2}{|c|}{\textbf{Annotation}} \\ \hline
@@
 \\ \hline
-\texttt{Image} & Images are any kind of graphical object, such as icons, screenshots, etc. \\ \hline
-\texttt{Tag} & A short (usually one word) description of a resource. \\ \hline
-\multicolumn{2}{|c|}{\textbf{User and User Profiles}} \\ \hline
-\texttt{User} & A human user of the FAST IDE. \\ \hline
-\texttt{User\_Profile} & The settings and data known about a particular user. \\ \hline
+\texttt{Image} & Images are any kind of graphical object, such as icons, screenshots, etc., and are used in the UI of the FAST platform to represent building blocks to users. \\ \hline
+\texttt{Tag} & A short (usually one word) description of a building block.
+Rather than being only simple literal string, a \texttt{Tag} in FAST is a complex object consisting of a string representation as well as a unique identifier, which unambiguously specifies the tags meaning.
+\\ \hline
 \end{longtable}
@@
 
-\paragraph{Properties} % (fold)
-\label{par:properties}
+\subsubsection{Glossary of Terms: Properties} % (fold)
+\label{subs:properties}
 
 All relevant properties of the FAST conceptual model are listed in Tab.~\ref{tab:glossary_of_terms_properties}.
-Many of the properties covered in this table will also have an inverse, which are not listed.
+Not all properties which will eventually be defined in the formal ontology are listed in the table: in many cases, these concrete properties simply reflect particular instances of generic composition or aggregation relations, which become apparent in the diagrams in Sect.~\ref{ssub:term_relations_within_the_conceptual_model}.
+Usually, these concrete properties will follow a general naming scheme of \texttt{hasCLASS\_NAME}, e.g., \texttt{hasDefinition} or \texttt{hasPreCondition}.
 
@@
 \multicolumn{2}{|c|}{\textbf{Components/Building Blocks}} \\ \hline
-\texttt{contains} & Many kinds of components in FAST can contain other components: screenflows contain screens, screens contain forms or form elements, etc. \\ \hline
-\texttt{hasPreCondition} & This property links a screen or screenflow to its pre-condition, i.e., the facts that need to be fulfilled in order for this screen or screenflow to be reachable. \\ \hline
+\texttt{contains} & Many kinds of building blocks in FAST can contain other building blocks: screenflows contain screens, screens contain forms or operators, forms contain form elements, etc.
+In many cases, specific properties for particular containment relations could be specified in the formal ontology.
+\\ \hline
+\texttt{hasPreCondition} & This property links a building block to its pre-condition, i.e., the facts that need to be fulfilled in order for this screen or screenflow to be reachable. \\ \hline
 \texttt{hasPostCondition} & This property links a screen or screenflow to its post-condition, i.e., the facts that are produced once the screen or screenflow has been executed. \\ \hline
@@
 In particular, we will look at the Friend of a Friend (FOAF) ontology, Semantically Interlinked Online Communities (SIOC), the Dublin Core metadata specification and Common Tag.
-Looking at the integrated vocabularies and ontologies, it becomes apparent that we integrate terms for those parts of the conceptual model which are not unique for FAST, i.e., users and user profiles, basic annotation needs or the tagging of building blocks.
-Aspects of the model which are unique to FAST, however, had been modelled from scratch.
-\todo{idea: move these last sentences to the conclusion.}
 The outcome of the integration process is the living \emph{integration document} at \ref{ssub:integration_document}.
@@
 While the previous section discussed the various steps preceding the definition of the actual ontology according to our chosen development methodology, this section focusses on concrete usage of the FAST gadget ontology in an RDFS/OWL environment.
+As a point of reference, an overview of all concrete classes will be given in Sect.~\ref{sub:ontology_overview}, followed by a number of sections detailing concrete aspects of the usage of the FAST gadget ontology.
+
+\subsection{Ontology Overview} % (fold)
+\label{sub:ontology_overview}
+
+This section gives a general, birds-eye-view of the ontology in terms of its classes and properties.
+Fig.~\ref{fig:ontology_uml} shows all classes in their subsumption hierarchy as a UML diagram, including the properties defining those classes as their domain.
+A more detailed list of classes and properties, complete with human-readable comments and documentation can be found in the back of the document in App.~\ref{sec:ontology_terms}, while the complete ontology code in OWL is given in App.~\ref{sec:ontology_code}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\linewidth]{images/fgo_uml_diagram}
+    \caption{FAST Gadget Ontology Class Hierarchy}
+    \label{fig:ontology_uml}
+  \end{center}
+\end{figure}
+
+% subsection ontology_overview (end)
 
@@
 \begin{itemize}
-	\item \textbf{Abstract Class Level}: At this level, resources are defined on a categorical level, differentiating between the various kinds of things that are relevant in FAST, as fundamental classes such as screens, forms or services (see Sect.~\ref{ssub:glossary_of_terms}).
+	\item \textbf{Abstract Class Level}: At this level, resources are defined on a categorical level, differentiating between the various kinds of things that are relevant in FAST, as fundamental classes such as screens, forms or resources (see Sect.~\ref{sub:conceptualisation}).
 These are implemented technically as classes in OWL.
@@
 We have specified what the individual components of a gadget are, and how they interrelate.
-In developing our ontology, we have followed  well-known methodology for ontology development.
-During the course of this process, a number of living documents --- the requirements specification document, the glossary of terms, the integration document and finally the implementation document --- were produced.
-Those documents serve as a tool for development and as documentation at the same time, and will be extended and updated in the coming two years of the project and beyond.
-The driving force behind these changes will be testing and evaluation of the ontology within the FAST project, and possibly in other contexts as well.
+In developing our ontology, we have followed Methontology, a well-known methodology for ontology development, which we have extended with additional UML diagrams to define the conceptual model of the FAST platform.
+During the course of this process, a number of living documents --- the requirements specification document, the glossary of terms and conceptual model, the integration document and finally the implementation document --- were produced.
+Looking at the integration document, it becomes apparent that we integrate terms for those parts of the conceptual model which are not unique for FAST, i.e., users and user profiles, basic annotation needs or the tagging of building blocks.
+Aspects of the model which are unique to FAST, however, had been modelled from scratch.
+The various documents eventually leading to the FAST gadget ontology serve as tools for development and as documentation at the same time, and are being extended and updated throughout the duration of the project and beyond.
+The driving force behind these changes is the testing and evaluation of the ontology within the FAST project, and possibly in other contexts as well.
 
@@
 
-\clearpage
-\phantomsection
-\section{Ontology Overview}
-\label{sec:ontology_overview}
-% \addcontentsline{toc}{section}{Ontology Overview}
-
-\begin{figure}[ht]
-  \begin{center}
-    \includegraphics[width=.80\linewidth]{images/fgo_uml_diagram}
-    \caption{FAST Gadget Ontology Class Hierarchy}
-    \label{fig:ontology_uml}
-  \end{center}
-\end{figure}
+% \clearpage
+% \phantomsection
+% \section{Ontology Overview}
+% \label{sec:ontology_overview}
+% % \addcontentsline{toc}{section}{Ontology Overview}
+% 
+% \begin{figure}[ht]
+%   \begin{center}
+%     \includegraphics[width=.80\linewidth]{images/fgo_uml_diagram}
+%     \caption{FAST Gadget Ontology Class Hierarchy}
+%     \label{fig:ontology_uml}
+%   \end{center}
+% \end{figure}
 
@@
 \section{Ontology Code}
+\label{sec:ontology_code}
 % \addcontentsline{toc}{section}{Ontology Code}

### 2010-02-23 trunk/deliverables/D2_2_Ontology/classes.tex
review-ready version ready
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@589 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2009-12-11 trunk/deliverables/D2_2_Ontology/classes.tex
@@
-\subsubsection{Class: \texttt{fast:Resource}}
-\label{subs:Resource}
+\subsubsection*{Class: \texttt{fgo:Action}}
+\label{subs:Action}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & Resource \\ \hline
-\textcolor{white}{\textbf{description}} & Anything that is part of a gadget (or the gadget itself). Tentatively 
-    anything that can be `touched' and moved around in the FAST IDE. \\ \hline
-\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fast:Screen\_Flow}}{subs:ScreenFlow}, \htmlref{\texttt{fast:Screen}}{subs:Screen}, \htmlref{\texttt{fast:Backend\_Service}}{subs:BackendService}, \htmlref{\texttt{fast:Flow\_Control\_Element}}{subs:FlowControlElement}, \htmlref{\texttt{fast:Operator}}{subs:Operator}, \htmlref{\texttt{fast:Form\_Element}}{subs:FormElement}, \htmlref{\texttt{fast:Condition}}{subs:Condition}, \htmlref{\texttt{fast:WithConditions}}{subs:WithConditions} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fast:contains}}{subs:contains}, \htmlref{\texttt{fast:hasIcon}}{subs:hasIcon}, \htmlref{\texttt{fast:hasScreenshot}}{subs:hasScreenshot} \\ \hline
-\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fast:contains}}{subs:contains} \\ \hline
+\textcolor{white}{\textbf{label}} & Action \\ \hline
+\textcolor{white}{\textbf{description}} & An action represents a specific routine which will be performed when a certain
+condition is fulfilled within a certain screen component. Examples are methods of a Web service (e.g., getItem) or functionality to update or change the contents of a form. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasUse}}{subs:hasUse} \\ \hline
+\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasAction}}{subs:hasAction} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fast:Screen\_Flow}}
-\label{subs:ScreenFlow}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Screen Flow \\ \hline
-\textcolor{white}{\textbf{description}} & The complete gadget, a set of screens. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
-\end{tabular}
-\subsubsection{Class: \texttt{fast:Screen}}
-\label{subs:Screen}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Screen \\ \hline
-\textcolor{white}{\textbf{description}} & An individual screen. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
-\end{tabular}
-\subsubsection{Class: \texttt{fast:Backend\_Service}}
+\subsubsection*{Class: \texttt{fgo:BackendService}}
 \label{subs:BackendService}
@@
 \textcolor{white}{\textbf{description}} & A Web service which provides data and/or functionality to a screen.
-    A backend service will often be external to FAST, and will probably have to be 
-    wrapped by the screen. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
+The actual backend service is external to FAST, and only available through a wrapper (the service Resource).
+\\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fast:Flow\_Control\_Element}}
-\label{subs:FlowControlElement}
+\subsubsection*{Class: \texttt{fgo:BuildingBlock}}
+\label{subs:BuildingBlock}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & Flow Control Element \\ \hline
-\textcolor{white}{\textbf{description}} & Any kind of component which can restrict the default flow of screens 
-    in a gadget. \\ \hline
-\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fast:Screen\_Flow\_Start}}{subs:ScreenFlowStart}, \htmlref{\texttt{fast:Screen\_Flow\_End}}{subs:ScreenFlowEnd}, \htmlref{\texttt{fast:Connector}}{subs:Connector} \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{label}} & BuildingBlock \\ \hline
+\textcolor{white}{\textbf{description}} & Anything that is part of a gadget.
+Tentatively anything that can be `touched' and moved around in the FAST IDE, from the most complex units such as screen flows, down to atomic form elements like a button or a label in a form.
+\\ \hline
+\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fgo:Action}}{subs:Action}, \htmlref{\texttt{fgo:BackendService}}{subs:BackendService}, \htmlref{\texttt{fgo:Condition}}{subs:Condition}, \htmlref{\texttt{fgo:Fact}}{subs:Fact}, \htmlref{\texttt{fgo:FormElement}}{subs:FormElement}, \htmlref{\texttt{fgo:Library}}{subs:Library}, \htmlref{\texttt{fgo:Pipe}}{subs:Pipe}, \htmlref{\texttt{fgo:Screen}}{subs:Screen}, \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent}, \htmlref{\texttt{fgo:ScreenFlow}}{subs:ScreenFlow}, \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:contains}}{subs:contains}, \htmlref{\texttt{fgo:hasIcon}}{subs:hasIcon}, \htmlref{\texttt{fgo:hasScreenshot}}{subs:hasScreenshot} \\ \hline
+\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:contains}}{subs:contains} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fast:Screen\_Flow\_Start}}
-\label{subs:ScreenFlowStart}
+\subsubsection*{Class: \texttt{fgo:Condition}}
+\label{subs:Condition}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & Screen Flow Start \\ \hline
-\textcolor{white}{\textbf{description}} & The entry point to a wigdet; the first screen. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Flow\_Control\_Element}}{subs:FlowControlElement} \\ \hline
+\textcolor{white}{\textbf{label}} & Condition \\ \hline
+\textcolor{white}{\textbf{description}} & The pre- or post-condition of a certain kinds of building block.
+If the building block is a screen flow, each target platform will use these conditions in its own way, or may also ignore them.
+E.g., in EzWeb pre- and post-conditions correspond to the concepts of slot and event.
+A condition can be seen as a RDF bag of facts, where every fact has to be true
+for the condition be true as well. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasPostCondition}}{subs:hasPostCondition}, \htmlref{\texttt{fgo:hasPreCondition}}{subs:hasPreCondition} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fast:Screen\_Flow\_End}}
-\label{subs:ScreenFlowEnd}
+\subsubsection*{Class: \texttt{fgo:Fact}}
+\label{subs:Fact}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & Screen Flow End \\ \hline
-\textcolor{white}{\textbf{description}} & A screen that ends the workflow of the gadget. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Flow\_Control\_Element}}{subs:FlowControlElement} \\ \hline
+\textcolor{white}{\textbf{label}} & Fact \\ \hline
+\textcolor{white}{\textbf{description}} & The `basic information unit of a FAST gadget'. In terms of RDF, a fact is a statement consisting of a subject, predicate and object (S, P, O). \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasPattern}}{subs:hasPattern}, \htmlref{\texttt{fgo:hasPatternString}}{subs:hasPatternString}, \htmlref{\texttt{fgo:isPositive}}{subs:isPositive} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fast:Connector}}
-\label{subs:Connector}
+\subsubsection*{Class: \texttt{fgo:Form}}
+\label{subs:Form}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & Connector \\ \hline
-\textcolor{white}{\textbf{description}} & An explicit connection between two screens. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Flow\_Control\_Element}}{subs:FlowControlElement} \\ \hline
+\textcolor{white}{\textbf{label}} & Form \\ \hline
+\textcolor{white}{\textbf{description}} & A form is the visual aspect of a screen: its user interface. Each form is made up of individual form elements. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasFormElement}}{subs:hasFormElement} \\ \hline
+\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasForm}}{subs:hasForm} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fast:Operator}}
+\subsubsection*{Class: \texttt{fgo:FormElement}}
+\label{subs:FormElement}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Form Element \\ \hline
+\textcolor{white}{\textbf{description}} & Form elements are UI elements in a particular screen, such as buttons, lists or labels. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasFormElement}}{subs:hasFormElement} \\ \hline
+\end{tabular}
+\subsubsection*{Class: \texttt{fgo:Library}}
+\label{subs:Library}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Library \\ \hline
+\textcolor{white}{\textbf{description}} & Libraries are references to external code libraries required for the execution of a particular building block at runtime. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasLanguage}}{subs:hasLanguage} \\ \hline
+\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasLibrary}}{subs:hasLibrary} \\ \hline
+\end{tabular}
+\subsubsection*{Class: \texttt{fgo:Operator}}
 \label{subs:Operator}
@@
 \textcolor{white}{\textbf{label}} & Operator \\ \hline
-\textcolor{white}{\textbf{description}} & Any kind of component that is used to connect backend services to 
-    form elements. Examples are simple pipes, aggregators or various kinds of 
-    filters. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{description}} & Operators are intended to transform and/or modify data within a screen, usually for preparing data coming from service resources for the use in the screen's interface.
+Operators cover different kinds of data manipulations, from simple aggregation to mediating data with incompatible schemas.
+\\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
+\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasOperator}}{subs:hasOperator} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fast:Form\_Element}}
-\label{subs:FormElement}
+\subsubsection*{Class: \texttt{fgo:Pipe}}
+\label{subs:Pipe}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & Form Element \\ \hline
-\textcolor{white}{\textbf{description}} & Form elements are UI elements in a particular screen. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{label}} & pipe or connector \\ \hline
+\textcolor{white}{\textbf{description}} & Pipes are used to explicitly define the flow of data within a screen, e.g., from service resource to operator to a specific form element. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasIdActionTo}}{subs:hasIdActionTo}, \htmlref{\texttt{fgo:hasIdBBFrom}}{subs:hasIdBBFrom}, \htmlref{\texttt{fgo:hasIdBBTo}}{subs:hasIdBBTo}, \htmlref{\texttt{fgo:hasIdConditionFrom}}{subs:hasIdConditionFrom}, \htmlref{\texttt{fgo:hasIdConditionTo}}{subs:hasIdConditionTo} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fast:Condition}}
-\label{subs:Condition}
+\subsubsection*{Class: \texttt{fgo:Resource}}
+\label{subs:Resource}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & Condition \\ \hline
-\textcolor{white}{\textbf{description}} & The pre- or post-condition of either a screen or a screenflow. In 
-    the latter case, each target platform will use these conditions in its own way, 
-    or may also ignore them. E.g., in EzWeb pre- and post-conditions correspond to 
-    the concepts of slot and event. \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fast:hasPattern}}{subs:hasPattern}, \htmlref{\texttt{fast:isPositive}}{subs:isPositive} \\ \hline
-\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fast:hasPreCondition}}{subs:hasPreCondition}, \htmlref{\texttt{fast:hasPostCondition}}{subs:hasPostCondition} \\ \hline
+\textcolor{white}{\textbf{label}} & Resource \\ \hline
+\textcolor{white}{\textbf{description}} & A service resource in FAST is a wrapper around a Web service (the backend service), which makes the service available to the platform, e.g., by mapping its definition to FAST facts and actions.
+\\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
+\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasResource}}{subs:hasResource} \\ \hline
 \end{tabular}
-\subsubsection{Class: \texttt{fast:WithConditions}}
-\label{subs:WithConditions}
+\subsubsection*{Class: \texttt{fgo:Screen}}
+\label{subs:Screen}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & With-condition \\ \hline
-\textcolor{white}{\textbf{description}} & Those kinds of resource which can have pre- or post-conditions 
-    (i.e., screens and screen flows). \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fast:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fast:hasPreCondition}}{subs:hasPreCondition}, \htmlref{\texttt{fast:hasPostCondition}}{subs:hasPostCondition} \\ \hline
-\textcolor{white}{\textbf{unionOf}} & \htmlref{\texttt{fast:Screen}}{subs:Screen}, \htmlref{\texttt{fast:Screen\_Flow}}{subs:ScreenFlow} \\ \hline
+\textcolor{white}{\textbf{label}} & Screen \\ \hline
+\textcolor{white}{\textbf{description}} & An individual screen; the basic unit of user interaction in FAST.
+A screen is the interface through which a user gets access to data and functionality of a backend service.
+\\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasForm}}{subs:hasForm}, \htmlref{\texttt{fgo:hasOperator}}{subs:hasOperator}, \htmlref{\texttt{fgo:hasResource}}{subs:hasResource} \\ \hline
 \end{tabular}
+\subsubsection*{Class: \texttt{fgo:ScreenComponent}}
+\label{subs:ScreenComponent}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Screen Component \\ \hline
+\textcolor{white}{\textbf{description}} & Screens are made up of screen components, which fundamentally include service resources, operators and forms. \\ \hline
+\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fgo:Form}}{subs:Form}, \htmlref{\texttt{fgo:Operator}}{subs:Operator}, \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasAction}}{subs:hasAction}, \htmlref{\texttt{fgo:hasTrigger}}{subs:hasTrigger} \\ \hline
+\end{tabular}
+\subsubsection*{Class: \texttt{fgo:ScreenFlow}}
+\label{subs:ScreenFlow}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Screen Flow \\ \hline
+\textcolor{white}{\textbf{description}} & A set of screens from which a gadget for a given target platform can be generated. \\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\end{tabular}
+\subsubsection*{Class: \texttt{fgo:Trigger}}
+\label{subs:Trigger}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & Trigger \\ \hline
+\textcolor{white}{\textbf{description}} & Triggers are the flip-side of actions.
+Certain events in a building block can cause a trigger to be fired.
+Other building blocks within the same screen, which are listening to it, will react with an action.
+\\ \hline
+\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasIdActionTo}}{subs:hasIdActionTo}, \htmlref{\texttt{fgo:hasIdBBFrom}}{subs:hasIdBBFrom}, \htmlref{\texttt{fgo:hasIdBBTo}}{subs:hasIdBBTo}, \htmlref{\texttt{fgo:hasNameFrom}}{subs:hasNameFrom} \\ \hline
+\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasTrigger}}{subs:hasTrigger} \\ \hline
+\end{tabular}
+\subsubsection*{Class: \texttt{fgo:WithCode}}
+\label{subs:WithCode}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & With Code \\ \hline
+\textcolor{white}{\textbf{description}} & Any kind of building block that can be defined as a whole through code. \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasCode}}{subs:hasCode}, \htmlref{\texttt{fgo:hasLibrary}}{subs:hasLibrary} \\ \hline
+\textcolor{white}{\textbf{unionOf}} & \htmlref{\texttt{fgo:Form}}{subs:Form}, \htmlref{\texttt{fgo:Operator}}{subs:Operator}, \htmlref{\texttt{fgo:Resource}}{subs:Resource}, \htmlref{\texttt{fgo:Screen}}{subs:Screen} \\ \hline
+\end{tabular}
+\subsubsection*{Class: \texttt{fgo:WithDefinition}}
+\label{subs:WithDefinition}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & With Definition \\ \hline
+\textcolor{white}{\textbf{description}} & Any kind of building block that can be defined declaratively in the GVS. \\ \hline
+\textcolor{white}{\textbf{unionOf}} & \htmlref{\texttt{fgo:Form}}{subs:Form}, \htmlref{\texttt{fgo:Screen}}{subs:Screen} \\ \hline
+\end{tabular}
+\subsubsection*{Class: \texttt{fgo:WithPostConditions}}
+\label{subs:WithPostConditions}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & With Post-condition \\ \hline
+\textcolor{white}{\textbf{description}} & Those kinds of building blocks which can have post-conditions. \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasPostCondition}}{subs:hasPostCondition} \\ \hline
+\textcolor{white}{\textbf{unionOf}} & \htmlref{\texttt{fgo:Screen}}{subs:Screen}, \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent}, \htmlref{\texttt{fgo:ScreenFlow}}{subs:ScreenFlow} \\ \hline
+\end{tabular}
+\subsubsection*{Class: \texttt{fgo:WithPreConditions}}
+\label{subs:WithPreConditions}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & With Pre-condition \\ \hline
+\textcolor{white}{\textbf{description}} & Those kinds of building blocks which can have pre-conditions. \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasPreCondition}}{subs:hasPreCondition} \\ \hline
+\textcolor{white}{\textbf{unionOf}} & \htmlref{\texttt{fgo:Action}}{subs:Action}, \htmlref{\texttt{fgo:Screen}}{subs:Screen}, \htmlref{\texttt{fgo:ScreenFlow}}{subs:ScreenFlow} \\ \hline
+\end{tabular}

### 2010-02-23 trunk/deliverables/D2_2_Ontology/fast_d2_2_2_ontology.tex
review-ready version ready
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@589 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \usepackage{courier}
+
+% This helps to get hyphenation in typewriter font (\texttt)
+\usepackage[htt]{hyphenat}
 
@@
 After briefly addressing the topic of related work in Sect.~\ref{sec:related_work} --- which is mainly covered in a different deliverable from this work package, \cite{urmetzer2010fast_state_of_the_art} --- the structure of the remainder of this document follows directly from our adopted methodology: we will perform a domain analysis in Sect.~\ref{sec:domain_analysis}, covering the stages of specification, conceptualisation and integration of existing vocabularies.
-This is then followed by the implementation step in Sect.~\ref{sec:ontology}, which covers the ontology in its final form (at this moment of the project).
+This is then followed by the implementation step in Sect.~\ref{sec:ontology}, which covers the ontology in its form at this moment of the project.
 Finishing the deliverable, we present our conclusions and outlook on years two and three of the project in Sect.~\ref{sec:conclusions}.
@@
 \label{sub:changes_from_previous_version}
+
+\todo{update this section, refer to change requests in last review}
 
@@
 While both are not strictly speaking ontologies, they nevertheless provide a very good insight in how to conceptualise the domain.
-Additionally, we consider work done in the area of Semantic Web Services, firstly because the gadgets designed in FAST will connect to Semantic and conventional Web services, and secondly because there are a number of similarities between the way SWS are often formalised and the way we envision the interaction of the different components of a gadget in the FAST IDE.
+Additionally, we consider work done in the area of Semantic Web Services, firstly because the gadgets designed in FAST will eventually connect to Semantic and conventional Web services, and secondly because there are a number of similarities between the way SWS are often formalised and the way we envision the interaction of the different components of a gadget in the FAST IDE.
 
@@
 \textcolor{white}{\textbf{Name}} & \textbf{FAST Gadget Ontology} \\ \hline
-\textcolor{white}{\textbf{Domain}} & Intelligent Gadgets \\ \hline
+\textcolor{white}{\textbf{Domain}} & Intelligent and complex gadgets \\ \hline
 \textcolor{white}{\textbf{Purpose}} & The FAST gadget ontology conceptualises the domain of intelligent gadgets as defined in the FAST development platform.
@@
 In FAST, a description of each gadget component and resource is available to the IDE.
-Form this description, the IDE can construct its interface, determine which components can be connected to which other components, or make suggestions to the user in order to aid in the gadget development process.
+From this description, the IDE can construct its interface, determine which components can be connected to which other components, or make suggestions to the user in order to aid in the gadget development process.
 
@@
 \textcolor{white}{\textbf{Level of Formality}} & formal (OWL ontology) \\ \hline
-\textcolor{white}{\textbf{Scope}} & \textbf{Concepts:} \emph{Component}, \emph{Resource}, \emph{Screen}, \emph{Backend Service}, \emph{Flow Control Element}, \emph{Operator}, \emph{Screenflow Start}, \emph{Screenflow End}, \emph{Connector}, \emph{Form Element}, \emph{Pre-condition}, \emph{Post-condition}, \emph{Query}, \emph{Label}, \emph{Icon}, \emph{User}, \emph{User Profile}, \emph{Tag}
+\textcolor{white}{\textbf{Scope}} & \textbf{Concepts:} \emph{Component}, \emph{Resource}, \emph{Screen}, \emph{Screen flow}, \emph{Backend Service}, \emph{Flow Control Element}, \emph{Operator}, \emph{Screenflow Start}, \emph{Screenflow End}, \emph{Connector}, \emph{Form Element}, \emph{Pre-condition}, \emph{Post-condition}, \emph{Query}, \emph{Label}, \emph{Icon}, \emph{User}, \emph{User Profile}, \emph{Tag}
 
@@
 \\ \hline
-\texttt{Building\_Block} & Anything that is part of a gadget (or the gadget itself).
-Tentatively anything that can be ``touched'' and moved around in the FAST IDE, from the most complex units such as \texttt{Screen\_Flow}s down to atomic \texttt{Form\_Element}s like a button or a label in a form.
-\\ \hline
-\texttt{Screen\_Flow} & A set of screens from which a gadget for a given target platform can be generated. \\ \hline
-\texttt{Screen} & An individual screen; the basic unit of user interaction in FAST. \\ \hline
+\texttt{Building\_Block} & Anything that is part of a gadget.
+Tentatively anything that can be ``touched'' and moved around in the FAST IDE, from the most complex units such as \texttt{Screen\_Flow}s, down to atomic \texttt{Form\_Element}s like a button or a label in a form.
+\\ \hline
+\texttt{Screen\_Flow} & A set of screens from which a gadget in a given format (which can cover one or more target platforms) can be generated. \\ \hline
+\texttt{Screen} & An individual screen; the basic unit of user interaction in FAST. A screen is the interface through which a user gets access to data and functionality of a backend service. \\ \hline
 \texttt{Screen\_Component} & Screens are made up of screen components, which fundamentally include service \texttt{Resource}s, \texttt{Operator}s and \texttt{Form}s. \\ \hline
@@
 Regarding their implementation, different kinds of building blocks in FAST can either be defined hard-coded as a piece of source code, or they can be defined declaratively in the terms of the FAST ontology (using pipes, operators, etc.).
-In the former case, a building block such as a screen would have been implemented and added to the FAST platform by an engineer, whereas in the latter case, ordinary users of the FAST platform would have assembled the building block using the tools available in the GVS\footnote{Currently, only screens can be assembled in the GVS.
+In the former case, a building block such as a screen would have been implemented and added to the FAST platform by an engineer, whereas in the latter case, ordinary users of the FAST platform would have assembled the building block using the tools available in the GVS\footnote{At this moment, only screens can be assembled in the GVS.
 Forms may or may not be editable in the GVS in the future.
-Other kinds of building blocks are not planned to be editable.}.
+Other kinds of building blocks are not currently planned to be editable.}.
 This principle is reflected in Fig.~\ref{fig:uml_definition}, showing how certain kinds of building blocks aggregate either a \texttt{Definition} or \texttt{Code}, whereas other kinds of building blocks are always hard-coded.
+It should be noted that, in theory, there is no reason why not any kind of building block could be defined either way.
+However, the concrete set of building blocks which can be defined declaratively or in code is changing dynamically according to the current state and plans of the FAST project.
+The situation depicted in Fig.~\ref{fig:uml_definition} --- \texttt{Screen} and any kind of \texttt{Screen\_Component} can be defined in code, whereas only \texttt{Screen} and \texttt{Form} can be defined declaratively --- reflects the status at M24 of the project.
 
@@
 
-In this section we will investigate out a number of established ontologies which cover part of the requirements laid out in the specification document, as well as the glossary of terms from the conceptualisation stage.
-In particular, we will look at the Friend of a Friend (FOAF) ontology, Semantically Interlinked Online Communities (SIOC), the Dublin Core metadata specification and Common Tag.
+In this section we will investigate a number of established ontologies which cover part of the requirements laid out in the specification document, as well as the glossary of terms from the conceptualisation stage.
+In particular, we will look at the Dublin Core metadata specification, the Friend of a Friend (FOAF) ontology, Semantically Interlinked Online Communities (SIOC)  and Common Tag.
 The outcome of the integration process is the living \emph{integration document} at \ref{ssub:integration_document}.
 
+\subsubsection{Dublin Core} % (fold)
+\label{ssub:dublin_core}
+
+Dublin Core (DC) is a set of metadata elements for describing online resources in order to support indexing, searching and finding them.
+Typical examples of terms from DC are \emph{title}, \emph{creator}, \emph{subject}, \emph{description}, \emph{date} or \emph{rights}.
+In principle, DC can be applied to a wide range of resources, but is typically used in the context of describing media (textual documents, video, sound or image).
+The initial work on DC was done at an invitational workshop in Dublin, Ohio, US (hence the name) in 1995.
+Since then, several revisions and extensions have been applied to the vocabulary.
+
+While it is possible to express DC in plain HTML, the more widely used language is probably RDF.
+The original 15 terms were all properties defined in the DC elements namespace (\url{http://purl.org/dc/elements/1.1/}, short \texttt{dc}).
+The properties were not formally defined with respect to their range and domain, or whether they are object or datatype properties.
+However, the assumption was that the values for each property would always be literals, as opposed to complex values represented by a URI.
+More recently, the DC elements namespace has been declared legacy and is now superseded by the DC terms namespace (\url{http://purl.org/dc/terms/}, short \texttt{dcterms})~\cite{dcterms2008}, which includes more precise re-definitions of all 15 original terms, as well as a number of new terms (resulting in a total of 55 terms).
+This new iteration of DC now provides the facility to use structured values for properties and specifies what type these values will have.
+E.g., the property \texttt{dcterms:creator} now supersedes the original \texttt{dc:creator} and specifies that its value has the type \texttt{dcterms:Agent}.
+
+In order to reconcile with the simplicity of the original, flat DC elements and the newer, more structured DC terms, DC also includes the so-called ``DumbDown'' principle, which basically says that structured values of DC terms should always carry a literal description as well, in order to cater for ``dumb'' agents processing the data.
+These literals should be expressed using properties such as \emph{rdfs:label} or \emph{rdf:value}.
+
+Because of the open nature of the RDF model and the open-world assumption usually applied for RDFS and OWL, it is possible to integrate DC with other ontologies such as FOAF.
+E.g., it is possible and good practice to make a statement saying that the \texttt{dcterms:creator} of a document is an instance of \texttt{foaf:Person}.
+By inference following from the DC terms specification, this person would then also be a \texttt{dcterms:Agent}.
+This scenario is illustrated in Fig.~\ref{fig:dcterms_example}.
+A similar, but slightly out-dated example is given in \url{http://dublincore.org/documents/dcq-rdf-xml/}.
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\linewidth]{images/dcterms_example.pdf}
+    \caption{Example of Dublin Core data}
+    \label{fig:dcterms_example}
+  \end{center}
+\end{figure}
+
+Looking at the requirements specification document, DC can be integrated into the FAST gadget ontology in many ways.
+A lot of the metadata needs for screens, screenflows and their components are identical to those of the media resources originally intended by DC.
+For this reason, many properties such as \texttt{title}, \texttt{creator}, \texttt{rights} or various \texttt{date} properties can be used directly (more details in the integration document).
+% subsubsection dublin_core (end)
 \subsubsection{FOAF} % (fold)
@@
 % subsubsection sioc (end)
-
-\subsubsection{Dublin Core} % (fold)
-\label{ssub:dublin_core}
-
-Dublin Core (DC) is a set of metadata elements for describing online resources in order to support indexing, searching and finding them.
-Typical examples of terms from DC are \emph{title}, \emph{creator}, \emph{subject}, \emph{description}, \emph{date} or \emph{rights}.
-In principle, DC can be applied to a wide range of resources, but is typically used in the context of describing media (textual documents, video, sound or image).
-The initial work on DC was done at an invitational workshop in Dublin, Ohio, US (hence the name) in 1995.
-Since then, several revisions and extensions have been applied to the vocabulary.
-
-While it is possible to express DC in plain HTML, the more widely used language is probably RDF.
-The original 15 terms were all properties defined in the DC elements namespace (\url{http://purl.org/dc/elements/1.1/}, short \texttt{dc}).
-The properties were not formally defined with respect to their range and domain, or whether they are object or datatype properties.
-However, the assumption was that the values for each property would always be literals, as opposed to complex values represented by a URI.
-More recently, the DC elements namespace has been declared legacy and is now superseded by the DC terms namespace (\url{http://purl.org/dc/terms/}, short \texttt{dcterms})~\cite{dcterms2008}, which includes more precise re-definitions of all 15 original terms, as well as a number of new terms (resulting in a total of 55 terms).
-This new iteration of DC now provides the facility to use structured values for properties and specifies what type these values will have.
-E.g., the property \texttt{dcterms:creator} now supersedes the original \texttt{dc:creator} and specifies that its value has the type \texttt{dcterms:Agent}.
-
-In order to reconcile with the simplicity of the original, flat DC elements and the newer, more structured DC terms, DC also includes the so-called ``DumbDown'' principle, which basically says that structured values of DC terms should always carry a literal description as well, in order to cater for ``dumb'' agents processing the data.
-These literals should be expressed using properties such as \emph{rdfs:label} or \emph{rdf:value}.
-
-Because of the open nature of the RDF model and the open-world assumption usually applied for RDFS and OWL, it is possible to integrate DC with other ontologies such as FOAF.
-E.g., it is possible and good practice to make a statement saying that the \texttt{dcterms:creator} of a document is an instance of \texttt{foaf:Person}.
-By inference following from the DC terms specification, this person would then also be a \texttt{dcterms:Agent}.
-This scenario is illustrated in Fig.~\ref{fig:dcterms_example}.
-A similar, but slightly out-dated example is given in \url{http://dublincore.org/documents/dcq-rdf-xml/}.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\linewidth]{images/dcterms_example.pdf}
-    \caption{Example of Dublin Core data}
-    \label{fig:dcterms_example}
-  \end{center}
-\end{figure}
-
-Looking at the requirements specification document, DC can be integrated into the FAST gadget ontology in many ways.
-A lot of the metadata needs for screens, screenflows and their components are identical to those of the media resources originally intended by DC.
-For this reason, many properties such as \texttt{title}, \texttt{creator}, \texttt{rights} or various \texttt{date} properties can be used directly (more details in the integration document).
-% subsubsection dublin_core (end)
 
@@
 \\ \hline
-\texttt{hasDomainContext} & DC & \texttt{dcterms:subject} & DC specifies that subject shoulds be non-literals.
+% \texttt{hasDomainContext} & DC & \texttt{dcterms:subject} & DC specifies that subject shoulds be non-literals.
 To allow simple tagging, we should represent tags as resources (\todo{common tag}).
@@
 \\ \hline
+\texttt{hasVersion} & DOAP & \texttt{daop:revision} & DOAP (Description of a Project) is a widely used vocabulary for basic annotation of software and other projects. \\ \hline
 \end{longtable}
@@
 
+\todo{add graph figures to all examples}
+
 While the previous section discussed the various steps preceding the definition of the actual ontology according to our chosen development methodology, this section focusses on concrete usage of the FAST gadget ontology in an RDFS/OWL environment.
@@
 Fig.~\ref{fig:ontology_uml} shows all classes in their subsumption hierarchy as a UML diagram, including the properties defining those classes as their domain.
+All \texttt{rdfs:subClassOf}/\texttt{rdfs:superClassOf} relations are modelled with the subsumption arrow, while \texttt{owl:union\_of} relations are modelled using the filled composition diamond.
 A more detailed list of classes and properties, complete with human-readable comments and documentation can be found in the back of the document in App.~\ref{sec:ontology_terms}, while the complete ontology code in OWL is given in App.~\ref{sec:ontology_code}.
@@
     \includegraphics[width=\linewidth]{images/fgo_uml_diagram}
-    \caption{FAST Gadget Ontology Class Hierarchy}
+    \caption{FAST Gadget Class Hierarchy}
     \label{fig:ontology_uml}
@@
 These are implemented technically as classes in OWL.
-	\item \textbf{Screenflow-independent Component Definition}: When a user of the FAST GVS is building a new screenflow or screen, they do so by combining existing building blocks and configuring them.
+	\item \textbf{Prototypical Component Definition}: When a user of the FAST GVS is building a new screenflow or screen, they do so by combining existing building blocks and configuring them.
 These building blocks are not the fundamental classes of FAST, but instead they are specific instances of those classes.
@@
 At this level, all gadget components exist only in the form of Web standards compliant code (HTML, JavaScript, CSS).
-Any metadata still necessary or definitions for screen pre- and post-conditions are for performance now defined in the JavaScript native data format JSON, rather than in RDF.
+Any metadata still necessary or definitions for screen pre- and post-conditions are for performance now defined in the JavaScript-native data format JSON, rather than in RDF.
 Therefore, at this level, the resources that make up the gadget are out of the scope of this deliverable.
@@
 When implementing these four stages, as illustrated in Fig.~\ref{fig:classes_templates}, to OWL DL semantics, one will be faced with an immediate problem.
-Conceptually, once would like to say that $Screen$ is a class, i.e., the set of all screens.
+Conceptually, one would like to say that $Screen$ is a class, i.e., the set of all screens.
 As argued above, $LogInScreen$ would then be an instance of $Screen$ ($LogInScreen \subset Screen$), while a particular $LogInScreen_x$ in a particular screenflow would again be an instance of $LogInScreen$ ($LogInScreen_x \subset LogInScreen$).
@@
 One resource (e.g., $LogInScreen$) will play the role of a \emph{template}, while the other resource (e.g., $LogInScreen_x$) will be called a \emph{copy}.
-Conceptually, copying a resource means defining a named graph that contains all statements about this resource considered to be relevant, and creating a new graph with near-identical statements\footnote{The graphs are not completely identical: the identifiers for the resource and potentially other, contained resources will have changed.}.
+Conceptually, copying a resource means defining a named graph that contains all statements about this resource considered to be relevant, and creating a new graph with near-identical statements\footnote{The graphs are not completely identical: the identifiers for the resource and potentially other, contained resources will have changed.} \todo{explain this better}.
 In practice, this procedure can be implemented using a SPARQL CONSTRUCT query, where the WHERE clause matches the graph pattern of the template, and the CONSTRUCT query will build the copy from it.
@@
 In addition, a number of terms from external ontologies should also be used to annotate resources.
-Note that all of these annotations are valid both for the screenflow-independent component definitions (templates) and the design time instances (copies).
+Note that all of these annotations are valid both for the prototypical component definitions (templates) and the design time instances (copies).
 A detailed description of all terms from the FAST gadget ontology namespace can be found in Sect.~\ref{sec:ontology_terms} of this document.
@@
 \label{sub:defining_pre_and_post_conditions}
+
+\todo{change examples to good relations}
 
@@
 Since graph patterns with variable cannot directly be expressed in RDF (short of using blank nodes, which cannot have labels), the graph patterns are expressed as RDF literals and linked \texttt{fast:Condition} instances using the \texttt{fast:hasPattern} property.
-Each such condition will be linked to a screen or screenflow using either the \texttt{fast:hasPreCondition} or \texttt{fast:hasPostCondition} property.
+Each such condition will be linked to a screen or screenflow using either the \texttt{fast:has\-PreCondition} or \texttt{fast:has\-Post\-Condition} property.
 
@@
 	\todo{update with reference to common tag!}
+	\todo{give concrete example with code, maybe Good Relations?}
 	\item \emph{User Interests:} In the same way as domain contexts, a user's interests can be specified using DBpedia identifiers or structured tagging.
@@
 
-It should be added that for any given screenflow, the concrete definition of both domain context and could be derived automatically from any semantic description which the backend services might have.
+It should be added that for any given screenflow, the concrete definition of both domain context could be derived automatically from any semantic description which the backend services might have.
 While this might happen directly, without any transformation or mapping, this approach would be impractical in practice.
@@
 
-In the following, we will look at each of those guidelines and discuss in how far the FAST gadget ontology fulfils them.
+In the following, we will look at each of those guidelines and discuss how far the FAST gadget ontology fulfils them.
 
@@
 The concrete interpretation of this criterion depends to a large degree on the intended use case and user community of the ontology.
-For a large-scale undertaking spanning many domains, and where reasoning is in the centre of attention --- CyC e.g.
+For a large-scale undertaking spanning many domains, and where reasoning is in the centre of attention --- CyC\footnote{\url{http://cyc.com/cyc/opencyc}} e.g.
 represents this type of ontology  --- a certain amount of complexity is necessary.
@@
 At the moment, the use case for the FAST gadget ontology is the tool-internal representation of gadgets and their components, which means that simplicity for users in the wider community is only of secondary importance.
-However, the ontology is still kept relatively basic, both in terms of the number of terms (currently 28 classes and 29 properties\todo{Recount!}), and in the complexity of the class definitions themselves, which are mostly restricted to a straightforward class subsumption hierarchy, with the exception of the sparse use of the boolean class expression \texttt{owl:unionOf}.
+However, the ontology is still kept relatively basic, both in terms of the number of terms (currently 19 classes and 22 properties\todo{Recount!}), and in the complexity of the class definitions themselves, which are mostly restricted to a straightforward class subsumption hierarchy, with the exception of the sparse use of the boolean class expression \texttt{owl:unionOf}.
 
@@
 
-The FAST gadget ontology fulfils this criterion by following the best practices recommendations outlined in \cite{berrueta2008publishing_rdf_vocabularies}, thereby ensuring that the ontology is available within a stable namespace and URI and can be accessed both by machine agents to retrieve its formal definition, as well as by human agents, who will retrieve a human-understandable HTML documentation from the same URI.
-Furthermore, the fourth rule of linked data (links should established between datasets) is covered through the integration of external ontologies, as discussed above in Sec.~\ref{ssub:reuse_of_existing_ontologies}.
+The FAST gadget ontology fulfils the previous criterion by following the best practices recommendations outlined in \cite{berrueta2008publishing_rdf_vocabularies}, thereby ensuring that the ontology is available within a stable namespace and URI and can be accessed both by machine agents to retrieve its formal definition, as well as by human agents, who will retrieve a human-understandable HTML documentation from the same URI.
+Furthermore, the fourth rule of linked data (links should be established between datasets) is covered through the integration of external ontologies, as discussed above in Sec.~\ref{ssub:reuse_of_existing_ontologies}.
 
@@
 Our ontology follows the documentation guideline, both through its online documentation, as well as through this deliverable.
-Both versions of the documentation provide both a lookup list for terms and their definitions, as well as concrete examples of term usage in instances.
+The two versions of the documentation provide both a lookup list for terms and their definitions, as well as concrete examples of term usage in instances.
 
@@
 
-A further guideline to ease and encourage adoption of an ontology is the provision of tool support to users, which will help them to generate instance data.
-Examples for such tools can be online forms such as FOAF-a-matic\footnote{\url{http://www.ldodds.com/foaf/foaf-a-matic}} for FOAF or plugins for popular content management systems to generate SIOC data\footnote{\url{http://sioc-project.org/exporters}}.
+A further guideline to ease and encourage adoption of an ontology is the provision of tool support to users, which will help them generate instance data.
+Examples of such tools can be online forms such as FOAF-a-matic\footnote{\url{http://www.ldodds.com/foaf/foaf-a-matic}} for FOAF or plugins for popular content management systems to generate SIOC data\footnote{\url{http://sioc-project.org/exporters}}.
 In FAST, instance data for the FAST gadget ontology is never intended to be generated by hand, but rather through the use of the FAST catalogue.
@@
 The various documents eventually leading to the FAST gadget ontology serve as tools for development and as documentation at the same time, and are being extended and updated throughout the duration of the project and beyond.
-The driving force behind these changes is the testing and evaluation of the ontology within the FAST project, and possibly in other contexts as well.
+The driving force behind these changes is the testing and evaluating of the ontology within the FAST project, and possibly in other contexts as well.
 
@@
 In order to avoid any potential migration issues, we have so far refrained from adopting OWL 2, but may do so in the future.
+
+\todo{Extend conclusion!}
 
@@
 We decided to use the \emph{Methontology} methodology, which was first introduced in \cite{gomez_perez1996conceptualise_ontologies} and \cite{fernandez1997methontology}.
-Methontology provides a good balance of formalisation and streamlining of the design process on the one hand, and a looseness in its requirements on how strictly one needs to follow the individual phases and steps.
+Methontology provides a good balance of formalisation and streamlining of the design process on the one hand, and on the other hand a looseness in its requirements on how strictly one needs to follow the individual phases and steps.
 The authors specifically advocate an ``evolving prototype life cycle'', which allows the ontology to ``grow depending on its needs'' \cite{fernandez1997methontology}.
@@
 
-Knowledge Acquisition in the context of Methontology is a loose collective term to all activities which are aimed at gathering background knowledge to clearly determine purpose and scope of the ontology.
+Knowledge Acquisition in the context of Methontology is a loose collective term for all activities which are aimed at gathering background knowledge to clearly determine purpose and scope of the ontology.
 As such, it feeds directly into the specification phase, but can just as well be relevant at later stages of the design and development process.
@@
 This can either mean connecting the emerging ontology to upper or meta ontologies (in the case that there are appropriate super classes or properties in those ontologies which can act as connection joints), or the reuse of terms from other ontologies (in the case that other ontologies have already defined classes or properties which fall in the scope of the emerging ontology).
-A typical example for a meta-ontology is OpenCyc, while the FOAF vocabulary \cite{brickley2004foaf} is a example of an often-reused ontology to describe people.
+A typical example for a meta-ontology is OpenCyc, while the FOAF vocabulary \cite{brickley2004foaf} is an example of an often-reused ontology to describe people.
 As the output of this phase, an \emph{integration document} is suggested, which gives detailed information about which terms were taken from which external ontology.
@@
 Below is the complete code of the FAST gadget ontology in Turtle syntax. The latest version of this file can always be accessed at \url{http://purl.oclc.org/fast/ontology/gadget}.
+\todo{check the line lengths for nice layout in this document}
 
@@
 }
-\lstinputlisting{fast_gadget.owl}
+\lstinputlisting{clean.ttl}
 

### 2010-02-23 trunk/deliverables/D2_2_Ontology/properties.tex
review-ready version ready
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@589 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Definition}}{subs:Definition}, \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:ResourceReference}}{subs:ResourceReference} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
 \end{tabular}
@@
 \textcolor{white}{\textbf{label}} & has action \\ \hline
-\textcolor{white}{\textbf{description}} & This property indicates which actions are asociated and can be perfomed within 
+\textcolor{white}{\textbf{description}} & This property indicates which actions are associated and can be perfomed within 
 a screen component. \\ \hline
@@
 \textcolor{white}{\textbf{label}} & has code \\ \hline
-\textcolor{white}{\textbf{description}} & URL of the executable code of a particular screen. \\ \hline
+\textcolor{white}{\textbf{description}} & URL of the executable code of a particular building block (currently screens or screen components). \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:WithCode}}{subs:WithCode} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{foaf:Document} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasForm}}
+\label{subs:hasForm}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has form \\ \hline
+\textcolor{white}{\textbf{description}} & If a screen is defined declaratively, this property links it to its form (i.e., its visual user interface). \\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
 \textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Screen}}{subs:Screen} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{foaf:Document} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Form}}{subs:Form} \\ \hline
 \end{tabular}
-\subsubsection*{Property: \texttt{fgo:hasCondition}}
-\label{subs:hasCondition}
+\subsubsection*{Property: \texttt{fgo:hasFormElement}}
+\label{subs:hasFormElement}
 \begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
 \hline
-\textcolor{white}{\textbf{label}} & has condition \\ \hline
-\textcolor{white}{\textbf{description}} & This property links a precondition or postcondition to a certain condition. \\ \hline
+\textcolor{white}{\textbf{label}} & has form element \\ \hline
+\textcolor{white}{\textbf{description}} & If a form is defined declaratively, its elements are linked to it with this property. \\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Postcondition}}{subs:Postcondition}, \htmlref{\texttt{fgo:Precondition}}{subs:Precondition} \\ \hline
-\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition}, \texttt{rdf:Bag} \\ \hline
-\end{tabular}
-\subsubsection*{Property: \texttt{fgo:hasDefinition}}
-\label{subs:hasDefinition}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has definition \\ \hline
-\textcolor{white}{\textbf{description}} & The structure and behaviour of a resource. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Definition}}{subs:Definition} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Form}}{subs:Form} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:FormElement}}{subs:FormElement} \\ \hline
 \end{tabular}
@@
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
 \textcolor{white}{\textbf{range}} & \texttt{foaf:Image} \\ \hline
-\end{tabular}
-\subsubsection*{Property: \texttt{fgo:hasId}}
-\label{subs:hasId}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has id \\ \hline
-\textcolor{white}{\textbf{description}} & The internal ID of the resource within the container resource. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource}, \htmlref{\texttt{fgo:ResourceReference}}{subs:ResourceReference} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
 \end{tabular}
@@
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:WithCode}}{subs:WithCode} \\ \hline
 \textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Library}}{subs:Library} \\ \hline
-\end{tabular}
-\subsubsection*{Property: \texttt{fgo:hasName}}
-\label{subs:hasName}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has name \\ \hline
-\textcolor{white}{\textbf{description}} & The name the user gives to the resource. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
 \end{tabular}
@@
 \end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasOperator}}
+\label{subs:hasOperator}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has operator \\ \hline
+\textcolor{white}{\textbf{description}} & If a screen is defined declaratively, this property links it to an operator it might contain. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Screen}}{subs:Screen} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Operator}}{subs:Operator} \\ \hline
+\end{tabular}
 \subsubsection*{Property: \texttt{fgo:hasPattern}}
@@
 \textcolor{white}{\textbf{label}} & has pattern \\ \hline
-\textcolor{white}{\textbf{description}} & This property links a screen or screenflow to its pre-condition, 
-    i.e., the facts that need to be fulfilled in order for this screen or screenflow 
-    to be reachable. \\ \hline
+\textcolor{white}{\textbf{description}} & This property links the abstract representation of a fact (i.e., an RDF triple) in FAST to a graph resource containing the actual triple. \\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
@@
 \textcolor{white}{\textbf{label}} & has pattern string \\ \hline
-\textcolor{white}{\textbf{description}} & This property is the textual representation of a condition. \\ \hline
+\textcolor{white}{\textbf{description}} & This property links to the textual representation of a fact (i.e., an RDF triple), e.g., in N3. \\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
@@
 \textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:WithPostConditions}}{subs:WithPostConditions} \\ \hline
-\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition}, \texttt{rdf:Bag} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition} \\ \hline
 \end{tabular}
@@
 \textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:WithPreConditions}}{subs:WithPreConditions} \\ \hline
-\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition}, \texttt{rdf:Bag} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasResource}}
+\label{subs:hasResource}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has resource \\ \hline
+\textcolor{white}{\textbf{description}} & If a screen is defined declaratively, this property links it to its service resource (i.e., its backend). \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Screen}}{subs:Screen} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
 \end{tabular}
@@
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
 \textcolor{white}{\textbf{range}} & \texttt{foaf:Image} \\ \hline
-\end{tabular}
-\subsubsection*{Property: \texttt{fgo:hasSource}}
-\label{subs:hasSource}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has source \\ \hline
-\textcolor{white}{\textbf{description}} & URL of the source code of a particular library. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Library}}{subs:Library} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{foaf:Document} \\ \hline
-\end{tabular}
-\subsubsection*{Property: \texttt{fgo:hasTag}}
-\label{subs:hasTag}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has tag \\ \hline
-\textcolor{white}{\textbf{description}} & A tag is used to annotate keywords of a particular resource. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
 \end{tabular}
@@
 \textcolor{white}{\textbf{label}} & has trigger string \\ \hline
-\textcolor{white}{\textbf{description}} & This property represents the events fired within a building block. e.g. a 
-	button to clear the shopping cart will fired a "clear cart" trigger. \\ \hline
+\textcolor{white}{\textbf{description}} & This property links a buildiong block to a Trigger. \\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
 \textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection*{Property: \texttt{fgo:hasType}}
-\label{subs:hasType}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has type \\ \hline
-\textcolor{white}{\textbf{description}} & The type of the resource. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection*{Property: \texttt{fgo:hasUri}}
-\label{subs:hasUri}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has uri \\ \hline
-\textcolor{white}{\textbf{description}} & URI of a particular resource pointing from a reference within
-	another resource. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:ResourceReference}}{subs:ResourceReference} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{foaf:Document} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
 \end{tabular}
@@
 \textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Action}}{subs:Action} \\ \hline
-\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:ResourceReference}}{subs:ResourceReference} \\ \hline
-\end{tabular}
-\subsubsection*{Property: \texttt{fgo:hasVersion}}
-\label{subs:hasVersion}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has version \\ \hline
-\textcolor{white}{\textbf{description}} & The version of a particular screen or screenflow, such as 1.0, 0.2beta, etc. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Resource}}{subs:Resource} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection*{Property: \texttt{fgo:integratesTerm}}
-\label{subs:integratesTerm}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & integratesTerm \\ \hline
-\textcolor{white}{\textbf{description}} & A way to explicitly say that an ontology uses terms from another namespace. 
-    Maybe a bit redundant, but why not. TODO: Must be moved to a different namespace! \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{rdfs:Resource} \\ \hline
 \end{tabular}
@@
 \textcolor{white}{\textbf{description}} & Facts can be set to a specific scope: design time, execution time,
-	or both of them. This will define when they have to be taken into account by the
+or both of them. This property will define when they have to be taken into account by the
 inference engine or reasoner. \\ \hline

### 2010-02-23 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_mediation.tex
version numbers changed
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@591 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
                             \textcolor{white}{\textbf{Change description}}\\ \hline
-0.1 & 09.05.2009 & \authorOne & Initial content \\ \hline
-0.4 & 30.11.2009 & \authorOne & Sections 2 and 3 \\ \hline
-0.5 & 07.01.2010 & \authorOne & New structure, Section 4 and 5 \\ \hline
-0.7 & 03.02.2010 & \authorTwo & Section 3.5, 3.6, updated Section 5  \\ \hline
-0.8 & 03.02.2010 & \authorOne & Integration and preliminary finalization  \\ \hline
-0.9 & 03.02.2010 & NUIG, UniKassel & Corrections  \\ \hline
-1.0 & 03.02.2010 & \authorOne & Integration and finalization  \\ \hline
+% 0.1 & 09.05.2009 & \authorOne & Initial content \\ \hline
+% 0.4 & 30.11.2009 & \authorOne & Sections 2 and 3 \\ \hline
+% 0.5 & 07.01.2010 & \authorOne & New structure, Section 4 and 5 \\ \hline
+% 0.7 & 03.02.2010 & \authorTwo & Section 3.5, 3.6, updated Section 5  \\ \hline
+% 0.8 & 03.02.2010 & \authorOne & Integration and preliminary finalization  \\ \hline
+% 0.9 & 03.02.2010 & NUIG, UniKassel & Corrections  \\ \hline
+% 1.0 & 03.02.2010 & \authorOne & Integration and finalization  \\ \hline
 %0.2 & 19.02.2009 & Oszkar Ambrus (NUIG) & content moved from wiki after internal review \\ \hline
 %0.1 & 07.01.2009 & Oszkar Ambrus (NUIG) & final version ready for external review \\ \hline
+1.0 & 27.02.2009 & \authorOne & final version (D2.4.1) ready for external review \\ \hline
+2.0 & 27.02.2010 & \authorOne & final version (D2.4.2) ready for external review \\ \hline
 \end{tabular}

### 2010-02-23 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_mediation.tex
Final corrections: submitted to wiki
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@592 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \newcommand\workpackageTitle{Definition of Conceptual Model}
-\newcommand\authorOne{Oszkar Ambrus (NUIG)}
-\newcommand\authorTwo{Albert Zuendorf (UniKassel)}
-\newcommand\authorThree{Author 3}
-\newcommand\authorFour{Author 4}
+\newcommand\authorOne{Oszk\'{a}r Ambrus (NUIG)}
+\newcommand\authorTwo{Albert Z\"{u}ndorf (UniKassel)}
+\newcommand\authorThree{J\"{o}rn Friedrich Dreyer (UniKassel)}
+\newcommand\authorFour{Sebasti\'{a}n Ortega (UPM)}
 \newtheorem{example}{\emph{Example}}
@@
         {\large \sffamily \authorOne} \\
-        \vspace{0.5cm}
         {\large \sffamily \authorTwo} \\
+        {\large \sffamily \authorThree} \\
+        {\large \sffamily \authorFour} \\
         \vspace{0.5cm}
         \vfill
-        {\large \sffamily Date: 20/02/2010}\\
+        {\large \sffamily Date: 27/02/2010}\\
         \vspace{1cm}
@@
     \hline
-    \textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {26/02/2010} & {\textbf{Submitted}} & {22/02/2010}\\ \hline
+    \textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2010} & {\textbf{Submitted}} & {27/02/2010}\\ \hline
     \textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
@@
 We present the different roles within the platform, tasks and challenges related to them, and the way ontology matching is integrated within each role.}\\ \hline
-    \textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne,  \authorTwo}\\
+    \textcolor{white}{\textbf{Authors}} & \multicolumn{4}{p{10.85cm}|}{\authorOne,  \authorTwo, \authorThree, \authorFour}\\
 %   {} & \multicolumn{4}{l|}{}\\ 
@@
 
-The first iteration, dealing with the theoretical foundations of ontology matching, is now contained in Part \ref{part:d2.4.1} and the whole of Part \ref{part:d2.4.2} is newly added content, based on the comments of the last review.
+The first iteration, dealing with the theoretical foundations of ontology matching, is now contained in the Appendix, and the whole body of this document is newly added content, based on the comments of the last review.
 As such, it tackles design and development issues for integrating ontology matching within the FAST platform: it presents the overall strategy and deals with specific problems belonging to each phase of ontology matching in FAST.
@@
 \subsection{Structure of the Document.}
-Part \ref{part:d2.4.2} describes the approach of integrating ontology matching into the FAST environment.
+This document describes the approach of integrating ontology matching into the FAST environment.
 Sect.~\ref{sec:overview} gives the big picture of ontology matching in fast, which the remaining three sections expand.
@@
 %Describe the rest.
-Part \ref{part:d2.4.1} contains the first iteration of this document, D2.4.1, laying down the theoretical foundations of ontology matching.
-We discuss ontologies in general in Sect.~\ref{sec:ontologies}, present the ontology mediation problem in Sect.~\ref{sec:mediationproblem} , discuss the types  of mismatches between ontologies in Sect.~\ref{sec:ontologymismatches} and detail the main approaches to ontology mediation in Sect.~\ref{sec:mediationapproaches}.
+The Appendix contains the first iteration of this document, D2.4.1, laying down the theoretical foundations of ontology matching.
+We discuss ontologies in general in App.~\ref{sec:ontologies}, present the ontology mediation problem in App.~\ref{sec:mediationproblem} , discuss the types  of mismatches between ontologies in App.~\ref{sec:ontologymismatches} and detail the main approaches to ontology mediation in App.~\ref{sec:mediationapproaches}.
 
@@
 
-\pagebreak
-\part{Implementing Ontology Matching within the FAST Platform}
-\label{part:d2.4.2}
+%\pagebreak
+%\part{Implementing Ontology Matching within the FAST Platform}
+%\label{part:d2.4.2}
 
@@
 
-\pagebreak
-\part{D2.4.1: Theoretical Foundations for Ontology Mediation}
-\label{part:d2.4.1}
-
-\include{fast_d2_4_2_part2}
-
-\pagebreak
-\part{Conclusion}
-
-\pagebreak
 \section{Summary}
@@
 
-In the first iteration (described in Part \ref{part:d2.4.1}) we laid the theoretical foundations of ontology matching.
+In the first iteration (described in the Appendix) we laid the theoretical foundations of ontology matching.
 After presenting the main problem, and the types of mismatches between ontologies, we have discussed - based on the most prominent research results in these areas - the three main approaches to ontology mediation, that is \textit{ontology mapping} (representing  correspondences between entities), \textit{ontology alignment} (the process of finding the correspondences between ontologies) and \textit{ontology merging }(the process of creating a new ontology based
@@
 
-In this second iteration (Part \ref{part:d2.4.2}) we established the approach to integrate ontology matching into FAST, and we have shown how this matching is done at the different phases of the gadget life cycle and the approach taken at every level.
+In this second iteration (described in the main body of this document) we established the approach to integrate ontology matching into FAST, and we have shown how this matching is done at the different phases of the gadget life cycle and the approach taken at every level.
 Firstly, we have shown the implementation details as well as test scenarios, we have presented a tool for finding the alignments between ontologies as well as a powerful language to represent them, we have described a tool to compose mapping rules in a consistent manner with other FAST components and the means to save them in the Semantic Catalogue.
@@
 
+%\pagebreak
+%\part{D2.4.1: Theoretical Foundations for Ontology Mediation}
+%\label{part:d2.4.1}
+
+\clearpage
+\phantomsection
+\appendix
+
+\include{fast_d2_4_2_part2}
+
+\clearpage
+\doublespacing
+\appendix
+\section{Lists of Tables and Figures}
+%\addcontentsline{toc}{section}{Appendix A (Lists of Tables and Figures)}
+
+\listoftables
+\listoffigures
+
 \clearpage
@@
 
-\clearpage
-\doublespacing
-\appendix
-\section*{Appendix A (Lists of Tables and Figures)}
-\addcontentsline{toc}{section}{Appendix A (Lists of Tables and Figures)}
-
-\listoftables
-
-\listoffigures
-
 \end{document}

### 2010-02-24 trunk/deliverables/D2_2_Ontology/classes.tex
making more improvements, added piping example figure
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@597 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fgo:Action}}{subs:Action}, \htmlref{\texttt{fgo:BackendService}}{subs:BackendService}, \htmlref{\texttt{fgo:Condition}}{subs:Condition}, \htmlref{\texttt{fgo:Fact}}{subs:Fact}, \htmlref{\texttt{fgo:FormElement}}{subs:FormElement}, \htmlref{\texttt{fgo:Library}}{subs:Library}, \htmlref{\texttt{fgo:Pipe}}{subs:Pipe}, \htmlref{\texttt{fgo:Screen}}{subs:Screen}, \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent}, \htmlref{\texttt{fgo:ScreenFlow}}{subs:ScreenFlow}, \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:contains}}{subs:contains}, \htmlref{\texttt{fgo:hasIcon}}{subs:hasIcon}, \htmlref{\texttt{fgo:hasScreenshot}}{subs:hasScreenshot} \\ \hline
-\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:contains}}{subs:contains} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:contains}}{subs:contains}, \htmlref{\texttt{fgo:hasCopy}}{subs:hasCopy}, \htmlref{\texttt{fgo:hasIcon}}{subs:hasIcon}, \htmlref{\texttt{fgo:hasScreenshot}}{subs:hasScreenshot}, \htmlref{\texttt{fgo:hasTemplate}}{subs:hasTemplate} \\ \hline
+\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:contains}}{subs:contains}, \htmlref{\texttt{fgo:hasCopy}}{subs:hasCopy}, \htmlref{\texttt{fgo:hasTemplate}}{subs:hasTemplate} \\ \hline
 \end{tabular}

### 2010-02-24 trunk/deliverables/D2_2_Ontology/fast_d2_2_2_ontology.tex
making more improvements, added piping example figure
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@597 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \newboolean{showcomments}
-\setboolean{showcomments}{false}
+\setboolean{showcomments}{true}
 \ifthenelse{\boolean{showcomments}}
@@
 At this level, all gadget components exist only in the form of Web standards compliant code (HTML, JavaScript, CSS).
-Any metadata still necessary or definitions for screen pre- and post-conditions are for performance now defined in the JavaScript-native data format JSON, rather than in RDF.
-Therefore, at this level, the resources that make up the gadget are out of the scope of this deliverable.
+For performance reasons, the FAST catalogue transforms any metadata still necessary, as well as definitions for screen pre- and post-conditions, from their original RDF representations into the JavaScript Object Notation (JSON)\footnote{\url{http://www.json.org/}}.
+At this level, the resources that make up the gadget have moved out of the scope of the FAST platform and therefore out of the scope of this deliverable.
 \end{itemize}
@@
 One resource (e.g., $LogInScreen$) will play the role of a \emph{template}, while the other resource (e.g., $LogInScreen_x$) will be called a \emph{copy}.
-Conceptually, copying a resource means defining a named graph that contains all statements about this resource considered to be relevant, and creating a new graph with near-identical statements\footnote{The graphs are not completely identical: the identifiers for the resource and potentially other, contained resources will have changed.} \todo{explain this better}.
-In practice, this procedure can be implemented using a SPARQL CONSTRUCT query, where the WHERE clause matches the graph pattern of the template, and the CONSTRUCT query will build the copy from it.
+Conceptually, copying a resource involves the following steps:
+\begin{inparaenum}[(i)]
+	\item Defining a named graph $G_1$ that contains all statements about this resource considered to be relevant.
+	\item Creating a second graph $G_2$ with copies of all statements in $G_1$.
+	\item During this copying process, the identifiers for the main resource and certain other resources in the graph will be changed from $I_{template}$ to $I_{copy}$.
+\end{inparaenum}	
+In practice, this procedure can be implemented using a SPARQL CONSTRUCT query, where the WHERE clause defines the graph pattern of the template, and the CONSTRUCT query will build the copy from it.
 The corresponding data on all three levels is shown in List.~\ref{list:templating}, using TriG syntax~\cite{bizer2004trig} (an extension of Turtle syntax for named graphs).
@@
 \lstset{
-	caption={Templating example --- different levels of representation of a FAST screen}, 
+	caption={Templating example --- different levels of representation of a FAST screen (TriG)}, 
 	label=list:templating,
@@
 \doublespacing
+
+Figure \ref{fig:templating_example_graph} illustrates the same example graphically, using different colours to represent the two different graphs: the red graph in the background represents the original template resource, whereas the green graph in the front in represents the resource copy as it would be used in a concrete screen flow.
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics[width=.75\linewidth]{images/templating_example.pdf}
+    \caption{Templating example --- different levels of representation of a FAST screen (Diagramme)}
+    \label{fig:templating_example_graph}
+  \end{center}
+\end{figure}
 
@@
 
-Once a user has been defined in this way, they can e.g.
-be used to specify the creators of individual screen and screenflow resources, as shown in List.~\ref{list:creator}.
+Once a user has been defined in this way, they can for example be used to specify the creators of individual screen and screenflow resources, as shown in List.~\ref{list:creator}.
 This kind of creation metadata is being added to other metadata as shown in Sect.~\ref{sub:basic_annotation}.
@@
 % % subsection instance_examples (end)
+
+\subsection{Piping within Screens} % (fold)
+\label{sub:piping_within_screens}
+
+\todo{needs text and code}
+
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics[width=\linewidth]{images/declarativeDefinitionExample.pdf}
+    \caption{Explicit piping within a screen from a post- to a pre-condition}
+    \label{fig:piping}
+  \end{center}
+\end{figure}
+
+% subsection piping_within_screens (end)
 
@@
 	\item \textbf{be a good Web citizen} by following the principles of linked data,
-	\item \textbf{be well documented} to make it easy for implementers to understand and use the ontology correctly and
+	\item \textbf{document well} to make it easy for implementers to understand and use the ontology correctly and
 	\item \textbf{provide tool support} to ease the work-load involved in generating instance data.
@@
 
-\subsubsection{Be Well Documented} % (fold)
+\subsubsection{Document Well} % (fold)
 \label{ssub:be_well_documented}
@@
 We decided to use the \emph{Methontology} methodology, which was first introduced in \cite{gomez_perez1996conceptualise_ontologies} and \cite{fernandez1997methontology}.
-Methontology provides a good balance of formalisation and streamlining of the design process on the one hand, and on the other hand a looseness in its requirements on how strictly one needs to follow the individual phases and steps.
+Methontology provides a good balance of formalisation and streamlining of the design process on the one hand, while on the other hand its requirements on how strictly one needs to follow the individual phases and steps are intentionally left relatively loose.
 The authors specifically advocate an ``evolving prototype life cycle'', which allows the ontology to ``grow depending on its needs'' \cite{fernandez1997methontology}.

### 2010-02-24 trunk/deliverables/D2_2_Ontology/properties.tex
making more improvements, added piping example figure
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@597 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \textcolor{white}{\textbf{range}} & \texttt{foaf:Document} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasCopy}}
+\label{subs:hasCopy}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has copy \\ \hline
+\textcolor{white}{\textbf{description}} & Links the template of a resource as used in the palette or catalogue of available resources to its copy, as used in a particular screen flow. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
 \end{tabular}
@@
 \end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasTemplate}}
+\label{subs:hasTemplate}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has template \\ \hline
+\textcolor{white}{\textbf{description}} & Links the copy of a resource as used in a particular screen flow to its template, as used in the palette or catalogue of available resources. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\end{tabular}
 \subsubsection*{Property: \texttt{fgo:hasTrigger}}

### 2010-02-24 trunk/deliverables/D2_2_Ontology/fast_d2_2_2_ontology.tex
extended text for templating section
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@598 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-Figure \ref{fig:templating_example_graph} illustrates the same example graphically, using different colours to represent the two different graphs: the red graph in the background represents the original template resource, whereas the green graph in the front in represents the resource copy as it would be used in a concrete screen flow.
+Figure \ref{fig:templating_example_graph} illustrates the same example graphically, using different colours to represent the two different graphs: the red graph $G_1$ in the background represents the original template resource, whereas the green graph $G_2$ in the front in represents the resource copy as it would be used in a concrete screen flow.
+\texttt{catalogue:LoginScreen} $\subset I_{template}$ has been changed to \texttt{catalogue:LoginScreen\_74382} $\subset I_{copy}$.
 

### 2010-02-25 trunk/deliverables/D2_2_Ontology/classes.tex
pretty much done with the very last revisions. more figures, better text, ontology code revised.
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@602 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasUse}}{subs:hasUse} \\ \hline
 \textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasAction}}{subs:hasAction} \\ \hline
@@
 \\ \hline
-\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fgo:Action}}{subs:Action}, \htmlref{\texttt{fgo:BackendService}}{subs:BackendService}, \htmlref{\texttt{fgo:Condition}}{subs:Condition}, \htmlref{\texttt{fgo:Fact}}{subs:Fact}, \htmlref{\texttt{fgo:FormElement}}{subs:FormElement}, \htmlref{\texttt{fgo:Library}}{subs:Library}, \htmlref{\texttt{fgo:Pipe}}{subs:Pipe}, \htmlref{\texttt{fgo:Screen}}{subs:Screen}, \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent}, \htmlref{\texttt{fgo:ScreenFlow}}{subs:ScreenFlow}, \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:contains}}{subs:contains}, \htmlref{\texttt{fgo:hasCopy}}{subs:hasCopy}, \htmlref{\texttt{fgo:hasIcon}}{subs:hasIcon}, \htmlref{\texttt{fgo:hasScreenshot}}{subs:hasScreenshot}, \htmlref{\texttt{fgo:hasTemplate}}{subs:hasTemplate} \\ \hline
+\textcolor{white}{\textbf{super\_class\_of}} & \htmlref{\texttt{fgo:Action}}{subs:Action}, \htmlref{\texttt{fgo:BackendService}}{subs:BackendService}, \htmlref{\texttt{fgo:Condition}}{subs:Condition}, \htmlref{\texttt{fgo:FormElement}}{subs:FormElement}, \htmlref{\texttt{fgo:Library}}{subs:Library}, \htmlref{\texttt{fgo:Pipe}}{subs:Pipe}, \htmlref{\texttt{fgo:Screen}}{subs:Screen}, \htmlref{\texttt{fgo:ScreenComponent}}{subs:ScreenComponent}, \htmlref{\texttt{fgo:ScreenFlow}}{subs:ScreenFlow}, \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:contains}}{subs:contains}, \htmlref{\texttt{fgo:hasCopy}}{subs:hasCopy}, \htmlref{\texttt{fgo:hasIcon}}{subs:hasIcon}, \htmlref{\texttt{fgo:hasScreenshot}}{subs:hasScreenshot}, \htmlref{\texttt{fgo:hasTemplate}}{subs:hasTemplate}, \htmlref{\texttt{fgo:uses}}{subs:uses} \\ \hline
 \textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:contains}}{subs:contains}, \htmlref{\texttt{fgo:hasCopy}}{subs:hasCopy}, \htmlref{\texttt{fgo:hasTemplate}}{subs:hasTemplate} \\ \hline
@@
 E.g., in EzWeb pre- and post-conditions correspond to the concepts of slot and event.
-A condition can be seen as a RDF bag of facts, where every fact has to be true
-for the condition be true as well. \\ \hline
+A condition is made up of individual facts, where each fact is represented internally as an RDF triple, usually involving a variable of blank node. \\ \hline
 \textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
-\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasPostCondition}}{subs:hasPostCondition}, \htmlref{\texttt{fgo:hasPreCondition}}{subs:hasPreCondition} \\ \hline
-\end{tabular}
-\subsubsection*{Class: \texttt{fgo:Fact}}
-\label{subs:Fact}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & Fact \\ \hline
-\textcolor{white}{\textbf{description}} & The `basic information unit of a FAST gadget'. In terms of RDF, a fact is a statement consisting of a subject, predicate and object (S, P, O). \\ \hline
-\textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasPattern}}{subs:hasPattern}, \htmlref{\texttt{fgo:hasPatternString}}{subs:hasPatternString}, \htmlref{\texttt{fgo:isPositive}}{subs:isPositive} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasPattern}}{subs:hasPattern}, \htmlref{\texttt{fgo:isPositive}}{subs:isPositive} \\ \hline
+\textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:from}}{subs:from}, \htmlref{\texttt{fgo:hasPostCondition}}{subs:hasPostCondition}, \htmlref{\texttt{fgo:hasPreCondition}}{subs:hasPreCondition}, \htmlref{\texttt{fgo:to}}{subs:to} \\ \hline
 \end{tabular}
@@
 \textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasIdActionTo}}{subs:hasIdActionTo}, \htmlref{\texttt{fgo:hasIdBBFrom}}{subs:hasIdBBFrom}, \htmlref{\texttt{fgo:hasIdBBTo}}{subs:hasIdBBTo}, \htmlref{\texttt{fgo:hasIdConditionFrom}}{subs:hasIdConditionFrom}, \htmlref{\texttt{fgo:hasIdConditionTo}}{subs:hasIdConditionTo} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:from}}{subs:from}, \htmlref{\texttt{fgo:to}}{subs:to} \\ \hline
 \end{tabular}
@@
 \textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasIdActionTo}}{subs:hasIdActionTo}, \htmlref{\texttt{fgo:hasIdBBFrom}}{subs:hasIdBBFrom}, \htmlref{\texttt{fgo:hasIdBBTo}}{subs:hasIdBBTo}, \htmlref{\texttt{fgo:hasNameFrom}}{subs:hasNameFrom} \\ \hline
 \textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:hasTrigger}}{subs:hasTrigger} \\ \hline

### 2010-02-25 trunk/deliverables/D2_2_Ontology/fast_d2_2_2_ontology.tex
pretty much done with the very last revisions. more figures, better text, ontology code revised.
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@602 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \tableofcontents
+
+\clearpage
+\doublespacing
+\phantomsection
+\section*{Lists of Tables and Figures}
+% \addcontentsline{toc}{section}{Lists of Tables, Figures and Listings}
+
+\listoftables
+\listoffigures
+\lstlistoflistings
+
 \cleardoublepage
-\pagenumbering{arabic}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
+\pagenumbering{arabic}
 \clearpage
@@
 FOAF provides the property \texttt{depiction} to express this general relation.
-However, \texttt{hasIcon} (our property for saying that something is the icon of a thing) can be considered a specialisation of this, and so we will say that \texttt{fast:hasIcon} is a specialisation of \texttt{foaf:depiction}.
-In terms of RDFS, we will say that \texttt{<fast:hasIcon> <rdfs:subPropertyOf> <foaf:depiction>}.
+However, \texttt{hasIcon} (our property for saying that something is the icon of a thing) can be considered a specialisation of this, and so we will say that \texttt{fgo:hasIcon} is a specialisation of \texttt{foaf:depiction}.
+In terms of RDFS, we will say that \texttt{<fgo:hasIcon> <rdfs:subPropertyOf> <foaf:depiction>}.
 % subsubsection foaf (end)
@@
   \begin{center}
-    \includegraphics[width=0.75\linewidth]{images/sioc_overview}
+    \includegraphics[width=0.8\linewidth]{images/sioc_overview}
     \caption{Overview of the SIOC ontology}
@@
 The basic idea of conceptual tagging is to address the problem of ambiguity in the meaning of tags (e.g., does ``jaguar'' mean the car, the animal or the operating system?) as it occurs in free-text tagging.
-Rather than relying on methods such as clustering, applied to a whole corpus of tagged resources, conceptual tagging intends to disambiguate a tag right from the moment it is used, by linking it to an unambiguous concept from vocabularies such as DBpedia (representing Wikipedia as linked data) or Freebase.
+Rather than relying on methods such as clustering, applied to a whole corpus of tagged resources, conceptual tagging intends to disambiguate a tag right from the moment it is used, by linking it to an unambiguous concept from vocabularies such as Dbpedia~\cite{Auer07dbpedia} (representing Wikipedia as linked data) or Freebase\footnote{\url{http://www.freebase.com/}}.
 E.g., in order to disambiguate ``cocoa'' as meaning the API, not the drink or plant, it would be represented as a resource (rather than a plain literal) and linked to the DBpedia resource \url{http://dbpedia.org/resource/Cocoa_%28API%29}.
@@
 \\ \hline
-\texttt{hasVersion} & DOAP & \texttt{daop:revision} & DOAP (Description of a Project) is a widely used vocabulary for basic annotation of software and other projects. \\ \hline
+\texttt{hasVersion} & DOAP & \texttt{doap:revision} & DOAP (Description of a Project) is a widely used vocabulary for basic annotation of software and other projects. \\ \hline
 \end{longtable}
@@
 \label{sec:ontology}
-
-\todo{add graph figures to all examples}
 
@@
     \includegraphics[width=.75\linewidth]{images/templating_example.pdf}
-    \caption{Templating example --- different levels of representation of a FAST screen (Diagramme)}
+    \caption{Templating example --- different levels of representation of a FAST screen (diagramme)}
     \label{fig:templating_example_graph}
@@
 
-\todo{change examples to good relations}
-
-The facts which define the pre- and post-conditions of screens and screenflows in FAST will be modelled as individual RDF triples.
-In effect, this means that conditions, which are sets of facts, can be modelled as graph patterns using SPARQL notation \cite{sparql2008spec}.
+The facts which define the pre- and post-conditions of screens and screenflows in FAST are modelled as individual RDF triples.
+In effect, this means that conditions, which are sets of facts, are modelled as graph patterns using SPARQL notation \cite{sparql2008spec}.
 E.g., a simple pre-condition such as \emph{``There has to be a user''} could be expressed as in List.~\ref{ex:simple_precondition}.
@@
 In logical terms, it means something along the lines of \emph{``there exists a sioc:User''}.
-For the pre-condition to be fulfilled, it needs to be executed against the RDF graph in question (most likely the facts on the current canvas).
+In order to determine if this  pre-condition can be fulfilled, it needs to be executed against the RDF graph in question.
+This graph could for example consist of the set of post-conditions of all screens currently available on the canvas.
 
@@
 
-Post-conditions will be expressed in a similar fashion.
-At design time, post-condition patterns will have variable in the same way as pre-condition patterns.
-When a screen is added to the canvas, the pattern needs to be materialised.
+Post-conditions are expressed in a similar fashion.
+In the prototypical phase (i.e., as part of templates), post-condition patterns have variables in the same way as pre-condition patterns.
+When a building block is copied to the canvas, the pattern needs to be materialised.
 To do this, each variable will be replaced with a randomly generated URI or blank node.
-At runtime, variables will be replaced with actual values once the screen for which the post-condition holds has been executed.
+At runtime, variables are then replaced with actual values once the screen for which the post-condition holds has been executed.
 As an example, consider the post-condition of a login screen.
@@
 \lstset{
-	caption={Simple post-condition}, 
+	caption={Post-condition at the prototypical stage}, 
 	label=ex:postcondition,
@@
 
-Once added to the canvas in the GVS (``instantiated at design time''), each variable would be replaced with a blank node identifier, as shown in example~\ref{ex:postcondition_canvas}.
-The pre-condition in example~\ref{ex:simple_precondition} could now be executed successfully as a SPARQL query against the canvas graph.
+Once added to the canvas in the GVS (``instantiated at design time''), each variable is then replaced with a blank node identifier, as shown in example~\ref{ex:postcondition_canvas}.
+This is necessary because, while variables are available in SPARQL, they cannot be represented directly in RDF.
+The pre-condition in example~\ref{ex:simple_precondition} could now be matched successfully as a SPARQL query against the canvas graph.
 
@@
 \lstset{
-	caption={Post-condition on the FAST GVS canvas}, 
+	caption={Post-condition as instantiated during design time (on the FAST GVS canvas)}, 
 	label=ex:postcondition_canvas,
@@
 E.g., this could result in a graph such as in example~\ref{ex:postcondition_runtime}.
+However, in the current implementation of FAST, gadgets do not use an RDF representation of facts at runtime, but instead a translation of the facts into the JSON format.
 
@@
 
-Since graph patterns with variable cannot directly be expressed in RDF (short of using blank nodes, which cannot have labels), the graph patterns are expressed as RDF literals and linked \texttt{fast:Condition} instances using the \texttt{fast:hasPattern} property.
-Each such condition will be linked to a screen or screenflow using either the \texttt{fast:has\-PreCondition} or \texttt{fast:has\-Post\-Condition} property.
-
-Since SPARQL has no direct support for negation, and since there is a need to allow screens to remove facts from the canvas graph as well as adding them, both pre- and post-conditions can be modelled to be either \emph{positive} or \emph{negative} using the \texttt{fast:isPositive} property.
+Regarding the representation of condition graph patterns and their linking to building blocks in the FAST backend, two parallel approaches are being followed. Each graph pattern is 
+\begin{inparaenum}[(i)]
+	\item represented as a named graph which is linked from an \texttt{fgo:Condition} using the \texttt{fgo:hasPattern} property, and
+	\item as a string, linked from the condition using the \texttt{fgo:hasPatternString} property.
+This second approach is followed for convenient rendering in user interfaces and to allow optional support the use of a triple store without support for named graphs.
+\end{inparaenum}
+Each such condition will be linked to a screen or screenflow using either the \texttt{fgo:has\-PreCondition} or \texttt{fgo:has\-Post\-Condition} property.
+A concrete example of this principle is given in List.~\ref{list:post_condition}.
+
+\singlespacing
+\lstset{
+	caption={Linking screens to conditions}, 
+	label=list:post_condition,
+	language=turtle
+}
+\begin{figure}[ht]
+	\lstinputlisting{exampleCode/post_condition_example.ttl}
+\end{figure}
+\doublespacing
+
+Since SPARQL has no direct support for negation, and since there is a need to allow screens to remove facts from the canvas graph as well as adding them, both pre- and post-conditions can be modelled to be either \emph{positive} or \emph{negative} using the \texttt{fgo:isPositive} property.
 For pre-conditions, the interpretation of \emph{positive} is that the condition must hold, while \emph{negative} means that the negation of the condition must hold.
@@
 
-\subsection{Piping within Screens} % (fold)
+\subsection{Conditions and Piping within Screens} % (fold)
 \label{sub:piping_within_screens}
 
-\todo{needs text and code}
+In year two of the FAST project, work has been started to extend the flexibility of the GVS to also include the creation of new screens from building blocks, as opposed to only assembling pre-made screens into screen flows.
+The gadget ontology supports this by modelling screen internals such as \emph{forms}, \emph{service resources} and \emph{operators}.
+Each of those building blocks can have their own pre- and post-conditions (either directly or through their \emph{actions}).
+However, while conditions on the screen/screen flow level are matched automatically and the flow of data is establish dynamically, on the screen-internal level this is defined explicitly through \emph{pipes}.
+Pipes are themselves components of a particular screen and connect matching post- and pre-conditions of other components in the same screen.
 
@@
     \includegraphics[width=\linewidth]{images/declarativeDefinitionExample.pdf}
-    \caption{Explicit piping within a screen from a post- to a pre-condition}
+    \caption{Explicit piping within a screen from a post- to a pre-condition (diagramme)}
     \label{fig:piping}
@@
 
+The concept of piping on this level of the FAST platform is illustrated graphically in Fig.~\ref{fig:piping}, and in code in List.~\ref{list:piping}\footnote{The example does not show a complete screen definition, but only the aspects which are relevant for this section.}.
+The screen is composed of (possibly among other things) a backend service resource, which wraps the Amazon search service, as well as a frontend Web form, which shows a list of Amazon products to the user.
+Additionally, the screen contains an instance of \texttt{fgo:Pipe}, which links the resource's post-condition (a product list) to the matching pre-condition of the form's ``showTable'' action.
+The desired functionality of this arrangement is that, whenever the service has returned a list of products in response to a search, this list will be piped to the form, where it will be displayed to the gadget's user.
+
+\singlespacing
+\lstset{
+	caption={Explicit piping within a screen from a post- to a pre-condition (code)}, 
+	label=list:piping,
+	language=turtle
+}
+\begin{figure}[ht]
+	\lstinputlisting{exampleCode/piping_example.ttl}
+\end{figure}
+\doublespacing
+
 % subsection piping_within_screens (end)
@@
 
-At this point in the development of FAST, there are three main entry points at which external ontologies come into play:
+While the FAST gadget ontology provides the framework and scaffolding for the formal description of complex gadgets, any  tailoring of a gadget or building block for a specific domain requires the integration of external domain ontologies.
+These ontologies can either be existing domain ontologies (e.g., the Good Relations ontology~\cite{Hepp:2008jr}\footnote{\url{http://www.heppnetz.de/projects/goodrelations/}} for e-Commerce) or ontologies which were purpose-built for FAST.
+There are three main entry points at which external ontologies come into play:
 
 \begin{itemize}
-	\item \emph{Domain Contexts:} By providing a domain context for a screen (or any other FAST resource), it is possible to express what domain this resource is relevant for.
+	\item \emph{Domain Contexts:} By providing a domain context for a screen (or any other FAST building block), it is possible to express what domain this resource is relevant for.
 Examples for domain contexts are ``medicine'', ``EU projects'', ``catering'', ``human resources'', ``books'', etc.
-As defined in the Integration Document in Sect.~\ref{ssub:integration_document}, domain contexts are given using the \emph{dcterms:subject} property.
-One option for finding domain identifiers to be used with this property is the use of DBpedia~\cite{Auer07dbpedia} identifiers such as \url{http://dbpedia.org/resource/Medicine} (DBpedia is a large RDF corpus automatically extracted from the info boxes of Wikipedia entries).
-This is now a widely used and recommended practice in the Semantic Web community (similar corpora could also be used).
-As an alternative or supplement to DBpedia identifiers, it is possible to use structured tagging approaches such as MOAT\footnote{\url{http://moat-project.org/}}.
-	\todo{update with reference to common tag!}
-	\todo{give concrete example with code, maybe Good Relations?}
-	\item \emph{User Interests:} In the same way as domain contexts, a user's interests can be specified using DBpedia identifiers or structured tagging.
-Both can then be used in combination to suggest resources to the user.
+As defined in the Integration Document in Sect.~\ref{ssub:integration_document}, we integrate the \texttt{ctag:tagged} property from the Common Tag vocabulary to define domain contexts as complex tags.
+As discussed in Sect.~\ref{ssub:common_tag}, one option for finding domain identifiers to be used with this property is the use of DBpedia identifiers such as \url{http://dbpedia.org/resource/Medicine} (DBpedia is a large RDF corpus automatically extracted from the info boxes of Wikipedia entries).
+This is now a widely used and recommended practice in the Semantic Web community.
+Similar corpora such as FreeBase could also be used.
+A concrete example of defining the domain context of the login screen to the Amazon service is given in List.~\ref{list:domain_context}, showing the use of \texttt{ctag:means} to disambiguate the meaning of the tag ``Amazon.com''.
+	\item \emph{User Interests:} In the same way as domain contexts, a user's interests can be specified using DBpedia identifiers and structured tagging.
+Matching domain contexts to used interests, the GVS (through the catalogue) can then suggest relevant resources to the user.
 	\item \emph{Pre- and Post-conditions:} In specifying pre- and post-conditions, terminology for an open-ended number of domains is necessary.
@@
 \end{itemize}
+
+\singlespacing
+\lstset{
+	caption={Defining a screen's domain context using the Common Tag vocabulary}, 
+	label=list:domain_context,
+	language=turtle
+}
+\begin{figure}[ht]
+	\lstinputlisting{exampleCode/domain_context_example.ttl}
+\end{figure}
+\doublespacing
 
@@
 \clearpage
-\doublespacing
-\phantomsection
-\section*{Lists of Tables and Figures}
-\addcontentsline{toc}{section}{Lists of Tables and Figures}
-
-\listoftables
-\listoffigures
-
-\clearpage
 \phantomsection

### 2010-02-25 trunk/deliverables/D2_2_Ontology/properties.tex
pretty much done with the very last revisions. more figures, better text, ontology code revised.
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@602 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:from}}
+\label{subs:from}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & from \\ \hline
+\textcolor{white}{\textbf{description}} & This property defines the starting point of a pipe within a screen. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition} \\ \hline
 \end{tabular}
@@
 \end{tabular}
-\subsubsection*{Property: \texttt{fgo:hasIdActionTo}}
-\label{subs:hasIdActionTo}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has id action to \\ \hline
-\textcolor{white}{\textbf{description}} & action id of the `to' point of the pipe or trigger. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe}, \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection*{Property: \texttt{fgo:hasIdBBFrom}}
-\label{subs:hasIdBBFrom}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has id building block from \\ \hline
-\textcolor{white}{\textbf{description}} & building block id of the `from' point of the pipe or trigger. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe}, \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection*{Property: \texttt{fgo:hasIdBBTo}}
-\label{subs:hasIdBBTo}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has id building block to \\ \hline
-\textcolor{white}{\textbf{description}} & building block id of the `to' point of the pipe or trigger. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe}, \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection*{Property: \texttt{fgo:hasIdConditionFrom}}
-\label{subs:hasIdConditionFrom}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has id condition from \\ \hline
-\textcolor{white}{\textbf{description}} & condition id of the `from' point of the pipe. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
-\subsubsection*{Property: \texttt{fgo:hasIdConditionTo}}
-\label{subs:hasIdConditionTo}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has id condition to \\ \hline
-\textcolor{white}{\textbf{description}} & condition id of the `to' point of the pipe. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
 \subsubsection*{Property: \texttt{fgo:hasLanguage}}
@@
 \end{tabular}
-\subsubsection*{Property: \texttt{fgo:hasNameFrom}}
-\label{subs:hasNameFrom}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has name from \\ \hline
-\textcolor{white}{\textbf{description}} & name of the trigger. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Trigger}}{subs:Trigger} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
-\end{tabular}
 \subsubsection*{Property: \texttt{fgo:hasOperator}}
@@
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Fact}}{subs:Fact} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition} \\ \hline
 \textcolor{white}{\textbf{range}} & \texttt{rdfs:Resource} \\ \hline
@@
 \textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Fact}}{subs:Fact} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:string} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:Condition} \\ \hline
 \end{tabular}
@@
 \end{tabular}
-\subsubsection*{Property: \texttt{fgo:hasUse}}
-\label{subs:hasUse}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has use string \\ \hline
-\textcolor{white}{\textbf{description}} & This property indicates concepts used within a building block, without being a 
-pre/postcondition. \\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Action}}{subs:Action} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{rdfs:Resource} \\ \hline
-\end{tabular}
 \subsubsection*{Property: \texttt{fgo:isPositive}}
@@
 \textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Fact}}{subs:Fact} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition} \\ \hline
 \textcolor{white}{\textbf{range}} & \texttt{xsd:boolean} \\ \hline
 \end{tabular}
+\subsubsection*{Property: \texttt{fgo:to}}
+\label{subs:to}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & to \\ \hline
+\textcolor{white}{\textbf{description}} & This property defines the end point of a pipe within a screen. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Pipe}}{subs:Pipe} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition} \\ \hline
+\end{tabular}
+\subsubsection*{Property: \texttt{fgo:uses}}
+\label{subs:uses}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & uses \\ \hline
+\textcolor{white}{\textbf{description}} & This property indicates concepts used within a building block, without being a 
+pre- or post-condition. \\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{rdfs:Resource} \\ \hline
+\end{tabular}

### 2010-02-26 trunk/deliverables/D2_2_Ontology/classes.tex
the final version that will be submitted to the EC. Mainly revised ontology code, introduction and conclusion
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@603 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \textcolor{white}{\textbf{sub\_class\_of}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
-\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasPattern}}{subs:hasPattern}, \htmlref{\texttt{fgo:isPositive}}{subs:isPositive} \\ \hline
+\textcolor{white}{\textbf{in\_domain\_of}} & \htmlref{\texttt{fgo:hasPattern}}{subs:hasPattern}, \htmlref{\texttt{fgo:hasPatternString}}{subs:hasPatternString}, \htmlref{\texttt{fgo:isPositive}}{subs:isPositive} \\ \hline
 \textcolor{white}{\textbf{in\_range\_of}} & \htmlref{\texttt{fgo:from}}{subs:from}, \htmlref{\texttt{fgo:hasPostCondition}}{subs:hasPostCondition}, \htmlref{\texttt{fgo:hasPreCondition}}{subs:hasPreCondition}, \htmlref{\texttt{fgo:to}}{subs:to} \\ \hline

### 2010-02-26 trunk/deliverables/D2_2_Ontology/fast_d2_2_2_ontology.tex
the final version that will be submitted to the EC. Mainly revised ontology code, introduction and conclusion
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@603 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \lstdefinelanguage{turtle} 
-{morekeywords={@prefix, a}, 
+{morekeywords={@prefix, a, true, false}, 
 sensitive=false, 
@@
 
-This deliverable defines the ontology and conceptual model for the semantic characterisation of complex gadgets in the FAST project.
+This deliverable defines the abstract conceptual model and concrete ontology for the semantic characterisation of complex gadgets in the FAST project.
 As such, it feeds directly into the various implementation efforts within the project.
-In the document, we present the ontology development methodology (Methontology) and design principles we have applied.
-Following Methontology, our design and development process results in a number of different living documents, which both function as a development tool, as well as documentation for the ontology itself.
+Throughout the document, we apply the \emph{Methontology} ontology development methodology.
+Our design and development process is therefore defined by a number of different living documents, which function both as a development tool, as well as documentation for the ontology itself.
 The major part of the deliverable is made up of these documents, resulting in the final \emph{implementation document} which formally defines the ontology in OWL DL semantics.
+Throughout the document, we provide extensive code examples and visual representations of the different aspects of both model and ontology.
 
@@
 The ontology provides a formal description of the conceptual model of FAST in general, and of the screen/screenflow architecture in particular.
-The document comprises a discussion of the ontology development methodology used, a number of intermediate representations which document the development process, and the final product in the form of an OWL ontology.}\\ \hline
+The document a number of successive intermediate representations which document the development process, and the final product in the form of an OWL ontology.
+Numerous code examples and visual representations of the ontology terms are provided.}\\ \hline
 	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{p{10.85cm}|}{\authorOne, \authorTwo, \authorThree, \authorFour}\\
@@
 The ontology formally defines the parts which a complex gadget is made of, how the parts interrelate, what kind of metadata is available for each part, how users of the system are represented, etc.
+It should be pointed out that the conceptual model and ontology are not meant to cover every conceivable aspect of the development of complex gadgets, not even every conceivable aspect of the FAST platform.
+Rather, the goal is the following: FAST is about the \emph{development of complex gadgets for end users}, and so the conceptual model and ontology provide the terminology and conceptual framework that is needed for specifically this task.
+In other words, while this document provides the theoretical underpinning for the implementation-oriented work in the project, the implementation-oriented work on the other hand defines the scope and requirements for the model.
 
@@
 
-This deliverable is structured as follows: as part of the introduction, we will present the methodology we chose to adopt for the development of the FAST gadget ontology, by discussing  each of its proposed stages in detail.
-Following, we will conclude the introduction by elaborating more on some of the general design decisions we are following.
-After briefly addressing the topic of related work in Sect.~\ref{sec:related_work} --- which is mainly covered in a different deliverable from this work package, \cite{urmetzer2010fast_state_of_the_art} --- the structure of the remainder of this document follows directly from our adopted methodology: we will perform a domain analysis in Sect.~\ref{sec:domain_analysis}, covering the stages of specification, conceptualisation and integration of existing vocabularies.
-This is then followed by the implementation step in Sect.~\ref{sec:ontology}, which covers the ontology in its form at this moment of the project.
-Finishing the deliverable, we present our conclusions and outlook on years two and three of the project in Sect.~\ref{sec:conclusions}.
+This deliverable is structured as follows: 
+we will start off by outlining the main changes that this document has undergone since its previous version, so as to enable the reader to quickly identify new material.
+After briefly addressing the topic of related work in Sect.~\ref{sec:related_work} --- which is mainly covered in a different deliverable from this work package, \cite{urmetzer2010fast_state_of_the_art} --- the structure of the following sections follows directly from our adopted ontology development methodology: we will perform a domain analysis in Sect.~\ref{sec:domain_analysis}, covering the stages of specification, conceptualisation and integration of existing vocabularies.
+This is then followed by the implementation step in Sect.~\ref{sec:ontology}, which covers the concrete ontology in its form at this moment of the project.
+After a section discussing an evaluation of the FAST gadget ontology, the deliverable will be concluded with a summary of our work in Sect.~\ref{sec:conclusions}.\ref{sub:methontology_a_methodology_for_ontology_development}
+A number of appendices have been added to the document: App.~\ref{sec:methodology} presents the development methodology adopted by us, App.~\ref{sec:ontology_terms} provides documentation for the ontology in the form of a detailed list of terms, while App.~\ref{sec:ontology_code} contains the complete code of the ontology.
 
@@
 \label{sub:changes_from_previous_version}
-
-\todo{update this section, refer to change requests in last review}
 
@@
 	\item The list of classes and properties has been moved to the appendix section at the back of the document (App.~\ref{sec:ontology_terms}), because it was felt that it would otherwise hinder the flow of the text.
-	\item In response to the M12 review, a UML overview of the ontology in App.~\ref{sec:ontology_overview} now complements the list of terms.
-	\item The section on our adopted development methodology (previously Sect.~1.3) has been moved to the end of the document (App.~\ref{sub:methontology_a_methodology_for_ontology_development}), but otherwise did not change.
-	\item An evaluation section has been added (Sect.~\ref{sec:evaluation}).
-	\item The section on design principles (previously Sect.~1.4, now Sect.~\ref{sub:general_design_decisions}) has been moved to the general evaluation section and extended to reflect research done by the authors within the past year.
-	\item In the integration section (Sect.~\ref{sub:integration}) of the ontology definition, the Common Tag vocabulary has been added. It is briefly introduced, and the integrated terms are specified.
-	\item The namespace abbreviation for the FAST gadget ontology changed from \texttt{fast} to \texttt{fgo}, in order to distinguish from other FAST ontologies which may be added in the future.
+	\item In response to the M12 review, numerous graphical representations both of the underlying conceptual model as well as the concrete gadget ontology have been added to the document, using UML notation.
+As an example, Sect.~\ref{sub:ontology_overview} now provides a complete overview of the ontology with all classes and properties.
+	\item Also in response to the previous review, conceptual model and ontology have been extended both in breadth and depth.
+	\item As an addition to the glossary of terms in the conceptualisation phase, a section discussing and illustrating the relations between the different classes has been provided.
+	\item In order to reflect the extension of the scope of FAST towards screen creation (as opposed to screen flow creation), the conceptual model and ontology have been extended accordingly.
+	\item A mechanism for RDF templating has been devised and is discussed in Sect.~\ref{sub:classes_instances_and_templates}.
+	\item As the result of the ongoing dialogue between ontology developers and application implementers, many classes and properties have been added, changed or removed.
+	\item The section on our adopted development methodology (previously Sect.~1.3) has been moved to the end of the document (App.~\ref{sub:methontology_a_methodology_for_ontology_development}), but otherwise did not change significantly.
+	\item An evaluation section has been added (Sect.~\ref{sec:evaluation}), based on research done by the authors within the past year.
+	\item The section on design principles (previously Sect.~1.4, now Sect.~\ref{sub:general_design_decisions}) has been moved to the general evaluation section.
+	\item In the integration section (Sect.~\ref{sub:integration}) of the ontology definition, the Common Tag vocabulary for complex tagging has been added.
+It is briefly introduced, and the integrated terms are specified.
+	\item On a technical note, the namespace abbreviation for the FAST gadget ontology changed from \texttt{fast} to \texttt{fgo}, in order to distinguish from other FAST ontologies which may be added in the future.
 \end{itemize}
@@
 
+\clearpage
 \section{Related Work} % (fold)
@@
 
-\section{Domain Analysis} % (fold)
+\clearpage
+\section{Domain Analysis and Conceptualisation} % (fold)
 \label{sec:domain_analysis}
@@
 
-\subsection{Conceptualisation} % (fold)
+\subsection{Conceptual Model} % (fold)
 \label{sub:conceptualisation}
@@
 
+\clearpage
 \singlespacing
@@
 
-\begin{figure}[ht]
+\begin{figure}
   \begin{center}
@@
 
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\linewidth]{images/dcterms_example.pdf}
+    \caption{Example of Dublin Core data}
+    \label{fig:dcterms_example}
+  \end{center}
+\end{figure}
+
 Because of the open nature of the RDF model and the open-world assumption usually applied for RDFS and OWL, it is possible to integrate DC with other ontologies such as FOAF.
@@
 
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\linewidth]{images/dcterms_example.pdf}
-    \caption{Example of Dublin Core data}
-    \label{fig:dcterms_example}
-  \end{center}
-\end{figure}
-
 Looking at the requirements specification document, DC can be integrated into the FAST gadget ontology in many ways.
@@
 % subsubsection dublin_core (end)
+
 \subsubsection{FOAF} % (fold)
@@
 }
-\begin{figure}[ht]
+\begin{figure}
 	\lstinputlisting{exampleCode/common_tag_example.ttl}
@@
 %
+\clearpage
 \section{The FAST Gadget Ontology} % (fold)
@@
 \lstset{
-	caption={Templating example --- different levels of representation of a FAST screen (TriG)}, 
+	caption={Templating example --- different levels of representation of a FAST screen (code)}, 
 	label=list:templating,
@@
 }
-\begin{figure}[ht]
+\begin{figure}
 	\lstinputlisting{exampleCode/templating_example.ttl}
@@
 }
-\begin{figure}[ht]
+\begin{figure}
 	\lstinputlisting{exampleCode/screen_example.ttl}
@@
 }
-\begin{figure}[ht]
+\begin{figure}
 	\lstinputlisting{exampleCode/user_example_01.ttl}
@@
 }
-\begin{figure}[ht]
+\begin{figure}
 	\lstinputlisting{exampleCode/user_example_02.ttl}
@@
 }
-\begin{figure}[ht]
+\begin{figure}
 \begin{lstlisting}
@@
 }
-\begin{figure}[ht]
+\begin{figure}
 	\lstinputlisting{exampleCode/post_condition_example.ttl}
@@
 			\midrule
-			\emph{isPositive ``YES''} & condition must be fulfilled & condition instantiated\\ 
-			\emph{isPositive ``NO''} & $\neg$condition must be fulfilled & instantiated condition removed \\ 
+			\emph{isPositive: true} & condition must be fulfilled & condition instantiated\\ 
+			\emph{isPositive: false} & $\neg$condition must be fulfilled & instantiated condition removed \\ 
 			\bottomrule
@@
 }
-\begin{figure}[ht]
+\begin{figure}
 	\lstinputlisting{exampleCode/piping_example.ttl}
@@
 }
-\begin{figure}[ht]
+\begin{figure}
 	\lstinputlisting{exampleCode/domain_context_example.ttl}
@@
 
+\clearpage
 \section{Evaluation} % (fold)
@@
 At the moment, the use case for the FAST gadget ontology is the tool-internal representation of gadgets and their components, which means that simplicity for users in the wider community is only of secondary importance.
-However, the ontology is still kept relatively basic, both in terms of the number of terms (currently 19 classes and 22 properties\todo{Recount!}), and in the complexity of the class definitions themselves, which are mostly restricted to a straightforward class subsumption hierarchy, with the exception of the sparse use of the boolean class expression \texttt{owl:unionOf}.
+However, the ontology is still kept relatively basic, both in terms of the number of terms (currently 18 classes and 22 properties), and in the complexity of the class definitions themselves, which are mostly restricted to a straightforward class subsumption hierarchy, with the exception of the sparse use of the boolean class expression \texttt{owl:unionOf}.
 
@@
 E.g., terms are grouped into defining \emph{resources}, \emph{pre- and post-conditions}, \emph{general annotation} or \emph{users and user profiles}.
-\todo{Implement this in the online documentation through VocDoc.}
+% \todo{Implement this in the online documentation through VocDoc.}
 % subsubsection modularisation (end)
@@
 
+\clearpage
 \section{Conclusions and Outlook} % (fold)
@@
 Aspects of the model which are unique to FAST, however, had been modelled from scratch.
+
 The various documents eventually leading to the FAST gadget ontology serve as tools for development and as documentation at the same time, and are being extended and updated throughout the duration of the project and beyond.
-The driving force behind these changes is the testing and evaluating of the ontology within the FAST project, and possibly in other contexts as well.
-
-A solution for representing pre- and post-conditions of screens and screen flows --- a central aspect of FAST technology --- has been presented, as it has been implemented and tested both within the FAST catalogue and the FAST GVS.
+The driving force behind these changes is the dialogue between ontology and application developers, as well as the testing and evaluating of the ontology within the FAST project.
+
+Apart from the conceptual model and ontology as such, we have introduced the notion of RDF graph templates as a means to overcome the strict separation of classes and instances in OWL.
+Within FAST, we use these templates to distinguish between prototypical building blocks and instances of such prototypes within a concrete screen flow.
+Also, a solution for representing pre- and post-conditions of building blocks such as screens flows, screens and screen components --- a central contribution of FAST technology --- has been presented, implemented and tested both within the FAST catalogue and the FAST GVS.
 
@@
 
-\todo{Extend conclusion!}
+The latest version of the ontology and its documentation can always be accessed at \url{http://purl.oclc.org/fast/ontology/gadget}.
 
@@
 \section{Development Methodology}
+\label{sec:methodology}
 % \addcontentsline{toc}{section}{Development Methodology}
@@
 For a more detailed discussion, we point the reader to the original papers proposing this methodology.
-The phases are \emph{specification}, \emph{knowledge acquisistion}, \emph{conceptualisation}, \emph{integration}, \emph{implementation}, \emph{evaluation} and \emph{documentation}.
+The phases are \emph{specification}, \emph{knowledge acquisition}, \emph{conceptualisation}, \emph{integration}, \emph{implementation}, \emph{evaluation} and \emph{documentation}.
 It is important to note that the phases do not have to be visited slavishly in the order presented here.
@@
 % 
-% \begin{figure}[ht]
+% \begin{figure}
 %   \begin{center}
@@
 Below is the complete code of the FAST gadget ontology in Turtle syntax. The latest version of this file can always be accessed at \url{http://purl.oclc.org/fast/ontology/gadget}.
-\todo{check the line lengths for nice layout in this document}
 

### 2010-02-26 trunk/deliverables/D2_2_Ontology/properties.tex
the final version that will be submitted to the EC. Mainly revised ontology code, introduction and conclusion
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@603 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \textcolor{white}{\textbf{label}} & contains \\ \hline
-\textcolor{white}{\textbf{description}} & Many kinds of components in FAST can contain other components: 
-screenflows contain screens, screens contain forms or form elements, etc.
-It points to a resource reference to give a unique `id' to each component
-within the resource. \\ \hline
+\textcolor{white}{\textbf{description}} & Many kinds of components in FAST can contain other components: screenflows contain screens, screens contain forms, operators and resources, forms contain form elements, etc.
+\\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
@@
 \textcolor{white}{\textbf{label}} & has action \\ \hline
-\textcolor{white}{\textbf{description}} & This property indicates which actions are associated and can be perfomed within 
-a screen component. \\ \hline
+\textcolor{white}{\textbf{description}} & This property indicates which actions are associated and can be perfomed within a screen component. \\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
@@
 \textcolor{white}{\textbf{label}} & has code \\ \hline
-\textcolor{white}{\textbf{description}} & URL of the executable code of a particular building block (currently screens or screen components). \\ \hline
+\textcolor{white}{\textbf{description}} & This property links to the executable code of a particular building block (currently screens or screen components). \\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
@@
 \textcolor{white}{\textbf{label}} & has copy \\ \hline
-\textcolor{white}{\textbf{description}} & Links the template of a resource as used in the palette or catalogue of available resources to its copy, as used in a particular screen flow. \\ \hline
+\textcolor{white}{\textbf{description}} & Links the template of a building block as used in the palette or catalogue of available building blocks to its copy, as used in a particular screen flow.
+\\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
@@
 \textcolor{white}{\textbf{label}} & has language string \\ \hline
-\textcolor{white}{\textbf{description}} & This property is the language the library is written in. \\ \hline
+\textcolor{white}{\textbf{description}} & This property defines the programming language a particular programming library uses. \\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
@@
 \textcolor{white}{\textbf{label}} & has library \\ \hline
-\textcolor{white}{\textbf{description}} & This property indicates which libraries are used by a screen component. \\ \hline
+\textcolor{white}{\textbf{description}} & This property indicates which programming libraries are used by the code of a screen component. \\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
@@
 \textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Fact}}{subs:Fact} \\ \hline
-\textcolor{white}{\textbf{range}} & \texttt{xsd:Condition} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:Condition}}{subs:Condition} \\ \hline
+\textcolor{white}{\textbf{range}} & \texttt{xsd:String} \\ \hline
 \end{tabular}
@@
 \textcolor{white}{\textbf{label}} & has post-condition \\ \hline
-\textcolor{white}{\textbf{description}} & This property links certain type of resources to its post-condition, 
-i.e., the facts that are produced once the screen or screenflow has been 
-executed. \\ \hline
+\textcolor{white}{\textbf{description}} & This property links certain type of building blocks to their post-condition, i.e., the facts that are produced once the building block has been executed.
+\\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
@@
 \textcolor{white}{\textbf{label}} & has pre-condition \\ \hline
-\textcolor{white}{\textbf{description}} & This property links certain type of resources to its pre-condition, i.e., 
-the facts that need to be fulfilled in order for this screen or screenflow to be 
-reachable. \\ \hline
+\textcolor{white}{\textbf{description}} & This property links certain type of building blocks to their pre-condition, i.e., the facts that need to be fulfilled in order for the building block to be reachable.
+\\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
@@
 \textcolor{white}{\textbf{label}} & has screenshot \\ \hline
-\textcolor{white}{\textbf{description}} & An image which shows a particular screen or screenflow in action, to 
-aid users in deciding which screen or screenflow to choose out of many. \\ \hline
+\textcolor{white}{\textbf{description}} & An image which shows a particular screen or screenflow in action, to aid users in deciding which screen or screenflow to choose out of many. \\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
@@
 \textcolor{white}{\textbf{label}} & has template \\ \hline
-\textcolor{white}{\textbf{description}} & Links the copy of a resource as used in a particular screen flow to its template, as used in the palette or catalogue of available resources. \\ \hline
+\textcolor{white}{\textbf{description}} & Links the copy of a building block as used in a particular screen flow to its template, as used in the palette or catalogue of available building blocks.
+\\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
@@
 \textcolor{white}{\textbf{label}} & has trigger string \\ \hline
-\textcolor{white}{\textbf{description}} & This property links a buildiong block to a Trigger. \\ \hline
+\textcolor{white}{\textbf{description}} & This property links a buildiong block to a trigger. \\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
@@
 \textcolor{white}{\textbf{label}} & is positive \\ \hline
-\textcolor{white}{\textbf{description}} & Facts can be set to a specific scope: design time, execution time,
-or both of them. This property will define when they have to be taken into account by the
-inference engine or reasoner. \\ \hline
+\textcolor{white}{\textbf{description}} & Facts can be set to a specific scope: design time, execution time, or both of them.
+This property defines how they have to be taken into account by the inference engine or reasoner.
+\\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:DatatypeProperty} \\ \hline
@@
 \textcolor{white}{\textbf{label}} & uses \\ \hline
-\textcolor{white}{\textbf{description}} & This property indicates concepts used within a building block, without being a 
-pre- or post-condition. \\ \hline
+\textcolor{white}{\textbf{description}} & This property indicates concepts used within a building block, without being a pre- or post-condition. \\ \hline
 \textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline

### 2010-04-13 trunk/deliverables/D2_2_Ontology/fast_d2_2_3_ontology.tex
post deliverable submission start on third version
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@779 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-02-26 trunk/deliverables/D2_2_Ontology/fast_d2_2_2_ontology.tex
@@
 % explicit hyphenations
-\hyphenation{RDF-Re-po-si-to-ry}
-\hyphenation{name-space}
-\hyphenation{So-cket-A-dap-ter}
+% \hyphenation{RDF-Re-po-si-to-ry}
+% \hyphenation{name-space}
+% \hyphenation{So-cket-A-dap-ter}
 
@@
 
-\pagenumbering{roman}
+% \pagenumbering{roman}
 
@@
 %\lofoot{\small\textcolor{fast@lightgrey}{Task Force Ontologies}}
-\cofoot{\small{FAST 216048 --- \deliverableTitleShort}}
-\rofoot{\small{\thepage}}
+%\rofoot{\small{\thepage}}
+\cofoot{\small{FAST $\bullet$ 216048 $\bullet$ \deliverableTitleShort{} $\bullet$ Page \thepage\ of \pageref{LastPage}}}
 
@@
 The ontology provides a formal description of the conceptual model of FAST in general, and of the screen/screenflow architecture in particular.
-The document a number of successive intermediate representations which document the development process, and the final product in the form of an OWL ontology.
+The deliverable contains a number of successive intermediate representations which document the development process, and the final product in the form of an OWL ontology.
 Numerous code examples and visual representations of the ontology terms are provided.}\\ \hline
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\pagenumbering{arabic}
+% \pagenumbering{arabic}
 \clearpage
-\rofoot{\small{Page \thepage\ of \pageref{LastPage}}}
+% \rofoot{\small{Page \thepage\ of \pageref{LastPage}}}
 
@@
 There are small differences in meaning between this term and the term \emph{widget}, which is also widely used.
-However, for the purpose of this document, it should be noted that whenever we use the term \emph{gadget}, we also refer to what \emph{widget} at the same time.
+However, for the purpose of this document, it should be noted that whenever we use the term \emph{gadget}, we might also have used the \emph{widget} at the same time.
 
@@
 For performance reasons, the FAST catalogue transforms any metadata still necessary, as well as definitions for screen pre- and post-conditions, from their original RDF representations into the JavaScript Object Notation (JSON)\footnote{\url{http://www.json.org/}}.
+The JSON structures used at this level are described in \cite{palaghita2010catalogue_manual}.
 At this level, the resources that make up the gadget have moved out of the scope of the FAST platform and therefore out of the scope of this deliverable.
@@
 For this reason, we introduce the concept of \emph{templates for RDF instances} in FAST.
+Templates are similar to the notion of prototypes in programming languages such as JavaScript.
 One resource (e.g., $LogInScreen$) will play the role of a \emph{template}, while the other resource (e.g., $LogInScreen_x$) will be called a \emph{copy}.

### 2010-06-13 trunk/papers/icsoc2010/paper.tex
first draft
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@841 d8b0c59b-6544-0410-94d9-83a5658a058e

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\usepackage{url}
\urldef{\mailsa}\path|{ismael.rivera, knud.moeller}@deri.org| 
\urldef{\mailsb}\path|zuendorf@cs.uni-kassel.de|
%\urldef{\mailsc}\path|erika.siebert-cole, peter.strasser, lncs}@springer.com|    
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\usepackage{amssymb}

\newenvironment{listing}
{\begin{list}{}{\setlength{\leftmargin}{1em}}\item\scriptsize\bfseries}
{\end{list}}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{An improved platform for Web service wrapping, discovery and consumption}

% a short form should be given in case it is too long for the running head
\titlerunning{Enhancing the discovery and consumption of Web services}

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{Ismael Rivera\and Knud M\"oller\and Albert Z\"undorf}
%
\authorrunning{Ismael Rivera\and Knud M\"oller\and Albert Z\"undorf}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{DERI, National University of Ireland, Galway\\
University of Kassel, Germany\\
\mailsa\\
\mailsb\\
%\mailsc\\
%\url{http://www.springer.com/lncs}
}

%
% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem"
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".
%

\toctitle{Lecture Notes in Computer Science}
\tocauthor{Authors' Instructions}
\maketitle

\begin{abstract}
B2B systems integration was revolutionized by the introduction of web services. The way enterprise systems communicated was dramatically transformed, and by adopting these technologies, the relationships with providers and customers were strengthened. The main advantages seen by companies which adopted the technology were an increase of operational efficiencies, and reduction of costs. In this scenario, high qualified software developers tackle with the integration, by means of the understanding of WSDL documents or human-readable specifications. However this model fails when it tries to target the long tail of enterprise software demand, and as a result, the end-users. Discovery and consumption of web services is far from a straight-forward task for an end-user, meaning potential users, willing to create their task-specific applications, have been ruled out. This paper presents an approach to facilitate the discovery and consumption of web services by end-users on the Internet, breaking the gap between business web services and the latter. The process includes: (a) a way to generate ready-to-use web services wrappers, and (b) a catalogue which users can browse to search for web services fitting their needs. These web services wrappers are described by a model created, although highly influenced by well-known web services modeling standards. Furthermore, two prototypes have been developed to serve as a proof of concept.
\keywords{web services, linked data, end-users, discovery, consumption}
\end{abstract}

\section{Introduction}

Business-to-business or B2B integration is easier said than done. Integration is a big challenge, in most cases requires a huge amount of effort, and deal with many factors and tecnologies for protocols, architectures, security standards, among others. Solutions based on open standards have reduced the complexity of integrating different business applications between different companies and partners. Web services and other XML standards such as RosettaNet, ebXML or OAG have been a boon to the world of B2B. By web services standards we are refering to the following open standards: Web Services Description Language (WSDL - to describe), Universal Description, Discovery and Integration (UDDI - to advertise and syndicate), Simple Object Access Protocol (SOAP - to communicate) and Web Services Flow Language (WSFL - to define work flows). Therefore, a common scenario is a company using WSDL to describe its public and private web service, publishing them either using a public or private repository using UDDI, where these web services using SOAP-based messages to achieve dynamic integration between different disparate applications.

Dealing with Business-to-consumer (B2C) integration, the adoption of these standards have offered some advantages as well, however the interaction and consumption of web services requires a programming skills and understanding of many technical details about the technology, which is an obstacle for an end-user with limited knowledge on the matter. Regarding discovery, most of the publishing and discovery platforms are syntactic-base, leading to poor precision and recall, making difficult to navigate through a large variety of web services \cite{pilioura_acm2009}, and as a result not empowering the end-users to tackle easily with this task. Other solutions like Semantic Web Service promised many advantages in discovery, composition and consumption of web services, independently of the provider's platform, the location, the service implementation, or the data format. However, their adoption was not as expected [TODO: any reference?], the further requirements needed and extra work have not being seen to be worth by enterprises for the benefits they offer. In the same direction as B2C, goverments are currently opening their data to the public as linked data. So far, the efforts are being focused on data, and the users have to deal with the development of applications to use that data. It is easily to predict that sooner or later they will move to open their services for administration tasks and single-window applications. Hence, they will be benefict by the research done around these concepts and technologies.

The motivation of our work has been strongly influenced by the end-users' needs. We are targeting non-skilled users in term of programming and software development knowledge, intending to empower them with tools to discover and consume web services in a straight-forward manner. The web services and their specifications are not published right away in the platform, since we were just reimplementing UDDI. However, their definition and behaviour are wrapped in what we call a ``web service wrapper''. The resulting products of the wrapping process are two artifacts: a specific definition of the web service to be used by this tool, and a ready-to-consume (inside a web browser) piece of code with the proper functions to invoke the web service.

The remainder of this paper is organised as follows. Section 2 covers the state of the art in web services publication, discovery and wrapping/consumption. Section 3 describes our platform for publishing web services and empowering end-users to discover services. Next, in Section 4 we present how we tackle with the web services wrapping, available possibilies and limitations. Section 5 gives the reader a clearer use case and how this work is integrated into the FAST platrofmr, and finally, we present our main conclusions and highlight future lines of research in Section 6.

\section{Related work}
\label{sec:related_work}

Web services have been around for a long time since they first appeared. One of their most important claims has been (syntactic) interoperability between third-party systems and applications based on different platforms / programming languages. System integration, within a company or between systems from different enterprises became easier with the adoption of the web services standard technologies, being the Web Service Description or WSDL the most extended standard used for the definition of web services. In fact, there are many integrated development environments (IDEs) that can deal with WSDL, and create a plain vanilla proxy class for the desired platform / programming language, getting a more ``developer-friendly'' API built around proxy classes, instead of having to know all request parameter names and possible values for the different calls to the Web service. Even though, in an integration of two or more systems scenario, developers must read and understand their specifications in order to accomplish the task, in a programmatically-fashion way. Another advantage of the use of WSDL is validation. Having the request messages (methods and parameters) and response messages (return parameters / structures) built upon XML schemas, a Web service engine can validate the message conforming to a schema. However, as it has been said before, the use of these languages and tools requires a qualified developer. In a inter-company system integration scenario these technologies are of high value, but when a company opens its services to the public, it is clear that the common end-user has been left off-side.

There are several tools out there situated a step forward to facilitate the interaction with data sources and services from around the Web. Yahoo! Pipes provides a set of modules to access different kind of data sources, such as RSS feeds, a given web page (HTML code), Flickr images, Google base or the Yahoo! search engine, among others. These modules are developed by the widget platform itself, they do not allow users to build their own modules, and it lacks of a way to interact with REST or SOAP web service. Another tool worth to mention is Apatar, an enterprise-oriented data integration and ETL tool. It allows the interaction with databases (MySQL, PostgreSql, Oracle), powerful enterprise systems such as Salesforce CRM, SugarCRM and Goldmine CRM, and access to data sources such as XML or text files. It enables non-developers to design and perform data transformations in a visual way through its graphical interface; however it misses support for Web services. Yet another tool, in the same direction as Apatar, could be JackBe Presto, defined as an enterprise-ready mashup solution. Presto permits to integrate and extract data from a great variety of data sources such as RSS and ATOM feeds or XML-based sources, and empower the user to create connectors to use services from HP SOA Systinet and any Oracle information technologies including Oracle 9i/10g/11g, Oracle Fusion SOAs, and Oracle Applications.

Summarising up, none of the solutions in the market really facilitate end-users to build up their own applications or widgets allowing the interaction with Web services created by third-party providers. The solutions found are mainly data-oriented (RSS feeds, databases, raw text). Several tools permits some sort of (Web) service integration, but are meant to be used within an entreprise level by savvy business users or developers. [TODO: rewrite this paragraph] This paper presents an approach to leverage the possibility of integrate REST or SOAP-based web service inside a widget or any application for a web browser (by creating service wrappers), and graphical tool will be provided to facilitate the process of building these service wrappers.

In the context of publishing and discovering Web services, a service provider have well-known and widely used technologies to accomplish the task of publication, such as the Universal Description, Discovery and Integration (UDDI) or WS-Discovery. The UDDI service registry specification \cite{uddi2004} is currently one of the core standards in the Web service technology stack and an integral part of every major SOA vendor's technology strategy and offers to requesters the ability to discover services via the Internet. In a few words, UDDI serves as a centralised repository of WSDL documents. A similar concept is iServe \cite{pedrinaci_ores2010}. This platform aims to publish Web services as what they called Lined Services Linked Services --linked data describing services, storing Web service definitions in a semantic-annotation fashion so semantic Web technologies for Web service discovery and processing can make use of them. However, the platform does not deal with the step from the definition to the consumption of the services.

A different approach for Web service discovery is the Web Services Dynamic Discovery (WS-Discovery) specification. The core of this approach is a multicast discovery protocol. Service providers and consumers  listen each other for new services specifications within a network, so there is no need of a centralised registry. As a drawback, WS-Discovery do not support Internet-scale discovery, therefore make if useless for our purposes.

[TODO: rewrite/extend this] The rule based approach of the service wrapper tool has been inspired by triple graph grammars, cf. \cite{schurr94}. While general triple graph grammars allow to relate complex graph structures with each other, in our case the target of a rule is always a single object or attribute. This facilitates the whole mechanism, reasonably.

\section{Wrapping web services (Kassel)}
\label{sec:wrapping_web_services}

%This section talks about the service wrapper tool (Kassel prototype)
%
%What it offers (afaik now it supports RESTful services, if support for SOAP-based services is also implemented, talk about it, otherwise explain how you expect to address it in the future work section
%
%How to create a wrapper (test-and-construct approach)
%
%Place this component within the FAST architecture
%
%Talk about formats: JSON and XML??

In our approach, wrapping web services is done in two steps: a first step is in charge of the construction of a service request and a second step analyzes the response got from the execution of the service, allowing the analysis of response data and the mapping to domain-specific concepts. From this input, we generate JavaScript code to be embedded into web gadgets. This JavaScript code takes query parameters, constructs the service request, analysis the service response and returns result data in the desired format.

\subsection{Constructing service requests} % (fold)
\label{sub:constructing_service_requests}

As a first approach, the interaction with services is limited to simple GET requests as supported e.g. by REST services. Support for other services like SOAP services is planned. A service request is thus assembled using a certain URL and a set of parameters. As an example, the Ebay Shopping web service will be studied. The following URL is invoked to retrieve a list of items corresponding to the search keywords \emph{USB}.

\url{http://open.api.sandbox.ebay.com/shopping?appid=KasselUn-efea-4b93-9505-5dc2ef1ceecd&version=517&callname=FindItems&ItemSort=EndTime&QueryKeywords=USB&responseencoding=XML}

As you may see, the URL invoked is http://open.api.sandbox.ebay.com/shopping and the parameters used in the example are:
\begin{description}
	\item[appid] this is the application ID obtained to use the API.
	\item[version] the API version.
	\item[callname] in this case FindItems to search through all items in Ebay.
	\item[itemsort] sorting method for the list of items.
	\item[querykeywords] list of keywords.
	\item[responseencoding] format of the response message obtained by the invocation of the request.
\end{description}

If needed, a detailed specification of the Ebay Shopping API can be found at \cite{eBayShoppingAPIs}.

The above URL for searching items in Ebay is followed by the query parameters, which take the form \textit{argument=value}. In this example, the only relevant parameter the end user may want to provide is \emph{querykeywords}. Thus the end user may use some input field to provide a query keyword and pass this value to our service wrapper operation. Other parameters can be set to a default value. However, in order to develop a more generic wrapper to a service, more or even all parameters might be passed by the user.

To define the desired input parameters and the type of the expected result of a new service wrapper the FAST service wrapper tool provides a web based form that allows the editing of these entries, cf. Figure \ref{fig:construct_pre_post_conditions}. Note, we allow to edit example values for the input parameters. These example values may be used to test the service wrapper.

%\begin{figure}
%  \begin{center} \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithPortDefinitions.jpg}
%    \caption{Configuring parameters of a service wrapper (photoshoped screen dump)}
%    \label{fig:construct_pre_post_conditions}
%  \end{center}
%\end{figure}

The service wrapper tool composes the request using a template string which will contain placeholders for input values. Before sending the request to the service, the placeholders are replaced with their corresponding values and the resulting URL is then ready to be sent as the service request.

Figure ~\ref{fig:construct_service_request} shows the web based tool to construct the service requests.

%\begin{figure}
%  \begin{center}
%    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithRequestExample.jpg}
%    \caption{Constructing the service request URL and its parameters (photoshoped screen dump)}
%    \label{fig:construct_service_request}
%  \end{center}
%\end{figure}

In the top input field of Figure ~\ref{fig:construct_service_request}, the user may drop an example http request taken e.g. form the service documentation e.g. from \cite{eBayShoppingAPIs}. The tool analyses the example request and in the middle of the screen a form for editing the request parameters is provided. In our example, the user has connected the \textit{QueryKeywords} parameter with the input parameter \textit{search\_key} by adding a corresponding reference to the value field of that parameter. In addition, we have retrieved an access key which has been entered as value for the \textit{appid} parameter.

Below the request parameter editing form of Figure ~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed URL by sending it to the specified service address (via a server relay). Then the placeholders for input parameters are replaced by their example values and the resulting http request is shown below the parameter form. In addition, the request is send and the response is shown on the bottom of that page. This gives the user a fast feedback whether the constructed request works as desired.

\subsubsection{Limitations} % (fold)
\label{ssub:limitations}

It is worth pointing out that currently the wrapper tool is able to construct input ports for the wrappers using just basic types. To construct URL requests from more complex input parameters, as e.g. a person object, we are currently developing access operations that will allow to access the values of fields of complex objects. For example, the expression \emph{customer.address} may refer to the \emph{address} field of a \emph{customer} parameter.

% subsubsection limitations (end)

% subsection constructing_service_requests (end)

\subsection{Interpreting service responses} % (fold)
\label{sub:interpreting_service_responses}

Once the service request is constructed and sent to the service provider, it will send back a response. This response message could be serialized in any format, though the most common formats used nowadays are XML or JSON among others. To continue the example started in the previous section, the response of the Ebay Shopping service will be in XML as specified in the request.

\subsubsection{Translation XML into Facts} % (fold)
\label{ssub:translation_xml_into_facts}

Figure ~\ref{fig:response_service_execution} shows the data transformation tab of the wrapper tool.

%\begin{figure}
%  \begin{center}
%      \includegraphics[angle=90,width=0.8\linewidth]{images/ServiceWrapperToolGVSWithTransformationRules.jpg}
%    \caption{Interactive, rule based transforming of an XML response to domain objects (photoshoped screen %dump)}
%    \label{fig:response_service_execution}
%  \end{center}
%\end{figure}

Once the service response, in XML format, has been retrieved, the transformation tab shows it as an interactive object tree on the left side of Figure ~\ref{fig:response_service_execution}. To construct this interactive object tree, the XML document has been parsed into a DOM, and a simplified tree representation of that DOM is built up.

A transformation rule is used to analyze the XML data and to generate domain-specific objects from concepts from the ontologies used by the parameters of the different building blocks. A transformation rule is composed of three elements, cf. the middle part of Figure ~\ref{fig:response_service_execution}. First, the \textit{from} field indicates the XML elements to be translated by the rule. These XML elements are identified by the tag name of a DOM element from the XML document. Second, the type of the rule will be set, taking one of the following values: \emph{createObject}, \emph{fillAttributes} or \emph{dummy}. And third, the target of the rule specifies a certain concept or attribute, to be created or filled. A detailed explanation of the type of actions to be trigger from the transformation rules is:
\begin{description}
	\item[action \emph{createObject}] specifies the creation of a new domain object. The type of that new object is provided in the third compartment. In the example being explained, the root rule searches for XML elements with tag name \emph{FindItemsResponse} and for each such element a \emph{List} object is created. The resulting objects are shown in a facts tree in the right of Figure ~\ref{fig:response_service_execution}.
	\item[action \emph{fillAttributes}] does not create a new object but it fills the value of the attribute provided as third part of such rules. In our example, the third transformation rule searches for XML elements with tag name \emph{Title}. Note, the rule is a sub-rule of the second rule, which generates \emph{Product} objects. Thus, the sub-rule searches for \emph{Title} tags only in the subtree of the XML data that has been identified by an application of the parent rule before. For example, the \emph{Item} rule may just have been applied to the first \emph{Item} element of the XML data. Then, the \emph{Title} rule is applied only to the first \emph{Item} sub-tree of the XML data and thus it will find only one \emph{Title} element in that sub-tree (not visible in Figure ~\ref{fig:response_service_execution} ). The value of that \emph{Title} element is then transfered to the \emph{productName} attribute of the corresponding \emph{Product} object. Actually, our \textit{from} fields allows also to refer to parts of an XML attribute e.g. to \textit{words} 1 through 3. It is also possible to combine constant text and elements of multiple XML tree elements. 
	\item[action \emph{dummy}] does not create or modify any objects but such rules are just used to narrow the search space for their sub-rules. For example, in Amazon product data, the XML data for an item contains sections for \emph{minimum price}, \emph{maximum price}, and \emph{average price}. Each such section contains the \emph{plain price} and the \emph{formatted price}. Thus, in the Amazon case, a rule that searches for \emph{formatted price} elements within an \emph{Item} element would retrieve three matches. Using a dummy rule, we may first search for \emph{minimum price} elements and then search for \emph{formatted price} elements within that sub-tree.
\end{description}

Our tool follows an interactive paradigm. Therefore, any time a change to a transformation rule is done, the transformation process is triggered and the resulting facts tree is directly shown. This process helps the user to deal with the slightly complex semantics of the transformation rules avoiding errors or mistakes. In addition, out tool is ontology-driven, therefore, the service wrapper designer shall retrieve the domain-specific types from a FAST ontology server together with the structure of each type, i.e. together with a description of the attributes of each object. Thus, the transformation rule editor is able to provide selection boxes for the target element of the rules. For a \emph{createObject} rule, this selection box shows the object types available for that domain. For the \emph{fillAttributes} rules, the selection box shows the attributes of the object type chosen in the parent rule. In addition, we may provide some analysis tool, which will help to guarantee that the objects generated by the transformation rules conform to the object types defined in the corresponding FAST ontology. This helps to ensure that the objects generated by the designed service wrapper will be compatible for input parameters of subsequent filter steps and or gadgets.

% subsubsection translation_xml_into_facts (end)

% subsection interpreting_service_responses (end)

\subsection{Generating a Resource Adapter} % (fold)
\label{sub:generating_a_resource_adapter}

Once the wrapping of a service has been defined and tested in the service wrapper tool, we generate an implementation of the desired Resource Adapter in XML, HTML, and JavaScript, ready to be deployed and executed inside a web gadget.

% subsection generating_a_resource_adapter (end)

\subsection{Limitations} % (fold)
\label{sub:limitations}

The rule driven approach presented above is somewhat limited. It is deliberately restricted to such a simple rule mechanism in order to keep things simple enough for end-users. Still, the selected approach suffices for many practical and real world cases. As a more complex example, the XML data for a person may provide two different tags for the first and the last name of a person. Contrarily, a person fact which conforms to a certain ontology for that domain may provide only one \emph{fullname} attribute that shall be filled by a concatenation of the first and the last name. To achieve this, the \textit{from} field of that tranforamtion rule might look like: \texttt{lastname"', "`firstname}. We are also able to do some navigation in the XML tree to follow XRef elements. For example the attribute \texttt{grandmother} could be filled using \texttt{mother.mother} in the \textit{from} field.

However, there are some transformations that these rules cannot perform. For example, we do not support any mathematical operations. Thus, transforming e.g. Fahrenheit into Celsius temperatures is not supported. To cover such  cases, intermediate object formats can be used which would allow generating objects to be further processed by additional filters. Such additional filters may be realized using (hand coded) operators, since some generic operators can act as filters for aggregation and conversions of objects from multiple sources. Then, service wrappers in combination with these filter operators will allow covering these complex cases.

% subsection limitations (end)

% section restful_web_services_wrapper_tool (end)

\section{Discovery: the marketplace}
\label{sec:discovery}

The area of web service publication and discovery has been subject for a lot of research since the very beginning the concept was coined. The current state of the art provides several solutions and strategies which providers and consumers may take advantage of. However, as explained in previous sections, a large number of them suffer from limitated syntactic-based descriptions and a simple keyword-based search, and the gap between discovery and consumption is an obstacle for their usage by a non-technical common user. In order to ligthen the issue, this paper presents yet another publishing platform, permiting any enterprise or individual to publish their public domain web services, with an enhanced semantic search based on the definition of the services, and serving``web service wrappers'' for easily consumption in any web application.

\subsection{Overview}
\label{ssec:overview}

The main difference of this platform with regards to the state of the art is that it is clearly targeting a different kind of user, and covers many deficiencies others tools does not tackle with. As a briefly overview, the platform being presented:

\begin{itemize}
	\item offers web services as Linked Data;
	\item serves web services wrappers ready-to-consume in any web application;
	\item follows well-known best practices for publishing data on the web;
	\item supports managing its resources via its RESTful API;
	\item supports content negotiation so the different clients may retrieved the information in JSON, RDF/XML, RDF/N3 or a human-readable HTML version;
	\item provices advanced search capabilites, based on the service formal definition and inferences extracted from it;
	\item a SPARQL endpoint to access directly to the data allowing advanced queries;
\end{itemize}

This platform was intented and it is being used as part of the FAST storyboard tool. Others components communicates with it via its RESTful API, but a
That said, The process of publishing a web service requires two steps: (i) generation of the wrapper (see Sec. \ref{sec:wrapping_web_services}, and (ii) creation of the resource in the publishing platform. The creation of the resources is done via the public RESTful API, and the language selected for the input for any operation is JSON. The following is an example of part of a request sent to create a new wrapper into the platform:

\begin{listing}
\begin{verbatim}
{
  "code": "http://demo.fast.morfeo-project.org/code/amazonService.js",
  "creationDate": "2010-01-26T17:01:13+0000",
  "description": {"en-gb": "Amazon web service"},
  "label": {"en-gb": "Amazon web service"},
  "preconditions": [
    [
      {
        "id": "item",
        "label": { "en-gb": "An item" },
        "pattern": "?I http://www.w3.org/1999/02/22-rdf-syntax-ns#type http://aws.amazon.com/AWSECommerceService#Item",
        "positive": "true"
      }
    ]
  ],
  "postconditions": [
    ...
  ],
  ...
\end{verbatim}
\label{lis:json_request}
\end{listing}

The architecture, as presented in the Figure \ref{fig:}, comprises a RDF store used for persistence, a business layer dealing with the model and reasoning, and public façade providing the core functionality as a RESTful API, and a SPARQL endpoint accessing directly the RDF store. This presentation layer is aimed to interact with the FAST Gadget Visual Storyboard, or any other third-party application.

The RESTful API and the SPARQL endpoint are part of the presentation layer. The SPARQL endpoint is offered using the SPARQL protocol service as defined in the SPROT specification \cite{sprot} and is aimed to enable third-party developers to query directly the knowledge base using SPARQL queries. This feature is supported by the Sesame RESTful HTTP interface for SPARQL Protocol for RDF.

The business logic layer contains all the domain-specific processing and reasoning. It provides functions to interact with all the elements of the domain model specified in \cite{sec:model}, acting as a mediator between the public RESTful API and the persistence layer.

The persistence layer provides an API to work with a standard set of objects representing the model. The interaction with the underlying RDF store is made via RDF2Go, an abstraction over triple (and quad) stores, which allows to program against rdf2go interfaces and choose any the RDF store implementation. This allows having a completely extensible and configurable framework for storage mechanisms, inferencers, RDF file formats, query result formats and query languages...

How to publish (RESTful app, anyone can create a wrapper and publish it)

How to browse: will allow users to go through the given domains (ontologies) and concepts, and will explain how to find the services they want (operation Find of the catalogue) => how the recommendation/reasoning works

It offers the wrapped web services as linked data and has an open sparql endpoint, allowing other developers to create applications on top of it...

\subsection{Serving Linked Data} % (fold)
\label{sub:linked_data}

The recently success of the web of Data has influenced the way information is now published on the web, and has been widely adopted by the academic community, large companies such as The New York Times and BBC, and national governments such as US and UK made public commitments toward open data. The main aim of the linking data is to find other related data based upon previously known data, in terms of the connections or links between them. We apply this concept in order to provide metadata about the web services as linked data, following the principles as originally defined in \cite{bizer_ijswis2009}, in order to make them available to arbitrary third-party applications. Each web service is identified by an HTTP URI and hosted in the catalogue so that it can be dereferenced through the same URI. For each building block, data is available in representations in different standard formats such as JSON (for communication with the GVS), RDF/XML, Turtle, or even HTML+RDFa as a human-readable version. The principles and best practices proposed in \cite{berrueta2008} and \cite{sauermann2008cool_uris} are also taken into account, these representations are served based on the request issued by the requesting agent, using a technique called \emph{content negotiation}. As required by the forth rule of linked data, individual building blocks link to other data on the web, thereby preventing so-called isolated ``data islands''.

In order to understand how content negotiation is implemented in the catalogue, a few concepts need to be understood. The URI of a particular building block must be understood as the identifier of the building block as such, as opposed to a particular representation (JSON, HTML, etc.) of it. In fact, each such representation has its own URI, under which it can be retrieved. Using the terminology established in \cite{w3c2004webArchitecture}, the building block is a \emph{non-information resource}, whereas each representation is an \emph{information resource}. In the process of content negotiation, the requesting agent will first request the URI of the building block as such (e.g., \texttt{http://catalogue.com/screens/235}), as well as the required representation format (e.g., JSON). The catalogue will then inspect the request and redirect the requester to the appropriate representation (e.g., \texttt{http://catalogue.com/screens/235.json}), as illustrated in Fig.~\ref{fig:content_negotiation}.
On the web, different representations for the same resource are called \emph{variants}, and content negotiation is the mechanism used to determine which of the representations of most appropriate for a given request.

%\begin{figure}[htb]
%\label{fig:content_negotiation}
%\begin{center}
%	\includegraphics[width=12cm]{images/content_negotiation}
%	\caption{Content Negotiation}
%\end{center}
%\end{figure}

In summary, the basic idea of content negotiation, as stated in the \cite{http1.1}, is to serve the best variant for a resource, taking into account what variants are available, what variants the server may prefer to serve, what the client can accept, and with which preferences. In HTTP, this is done by the client which may send, in its request, accept headers (\texttt{Accept}, \texttt{Accept-Language} and \texttt{Accept-Encoding}), to communicate its capabilities and preferences in format, language and encoding, respectively.

In fact, what the catalogue really does is ``format negotiation'' since the alternate representations are just based on the selection of the media type, through the accept header, but does not consider different languages or encoding types. The formats supported are JSON, RDF/XML, Turtle and HTML+RDFa. Even though the accept header is desired, the different representations can also be retrieved directly by dereferencing their own URI. The convention used in the catalogue for deriving a representation URI from a building block URI is to simply attach a matching suffix, such as \texttt{.rdf}, \texttt{.json} or \texttt{.html}.

Lastly, content negotiation needs to identify which player is going to take the lead on it. There are two kinds of content negotiation which are possible in HTTP: server-driven and agent-driven negotiation. The approach followed by the catalogue is agent-driven negotiation, hence selecting a specific representation for a resource is responsibility of the user agent. If none is specified, by default, the server will choose the JSON representation.

\subsection{Simplified Model/Ontology (Knud)}
The model used to define the web service wrappers within this application has been influenced by WSDL, the main web service definition language, and some semantic-enriched approaches such as WSMO and OWL-S. There is no need to reinvent the wheel,

A WSDL operation is mapped to an action. Its inputs and outputs are called pre-postconditions (a WSMO concept).

This section could talk about a the part of the ontology describing the back-end services, how we remove the unnecessary information from WSDL and similar (not need anymore because the code has been generated), however we reuse some of the concepts from these technologies to define them (preconditions, postconditions, actions, and so on) allowing discovery and composition.

%\begin{figure}[ht]
%	\centering
%	\includegraphics[width=8.31cm]{uddi.png}
%	\caption{UDDI overview}
%	\label{fig:uddi}
%\end{figure}

%\section{Wrapping web services (Kassel)}
%\label{sec:wrapping_web_services}

%This section talks about the service wrapper tool (Kassel prototype)

%What it offers

%How to create a wrapper (test-and-construct approach)

%Place this component within the FAST architecture

%Talk about formats: JSON and XML??

%\subsection{Code-generation phase}
%How the executable code is build:

% - mapping from operations and parameters to actions and preconditions

% - mapping from operation results to postconditions

%Now, the calls are done through the FAST api, which acts as a proxy <- imply they need to construct the widget using the FAST GVS which doesn't give an open solution (proposal in future work)

\section{Use case}
maybe we should provide some kind of use case to show the benefits of the approach and the possibilities this sort of tools?

\section{Conclusions and future work}

Future work:

Talk about wrapping WSDL-based services

Talk about WSMO-lite services

Use JSONRequest to replace the AJAX XMLHttpRequest and its problems (cross-domain calls), so there's no need for a Proxy.

\subsubsection*{Acknowledgments.} This work is supported in part by the European Commission
under the first call of its Seventh Framework Program (FAST STREP Project, grant
INFSO-ICT-216048) and in part by Science Foundation Ireland under Grant No.
SFI/08/CE/I1380 (L\'ion-2).

\bibliographystyle{apalike}
\bibliography{paper}

\section*{Checklist of Items to be Sent to Volume Editors}
Here is a checklist of everything the volume editor requires from you:

\begin{itemize}
\settowidth{\leftmargin}{{\Large$\square$}}\advance\leftmargin\labelsep
\itemsep8pt\relax
\renewcommand\labelitemi{{\lower1.5pt\hbox{\Large$\square$}}}

\item The final \LaTeX{} source files
\item A final PDF file
\item A copyright form, signed by one author on behalf of all of the
authors of the paper.
\item A readme giving the name and email address of the
corresponding author.
\end{itemize}
\end{document}

### 2010-06-13 trunk/papers/icsoc2010/paper.tex
changes in introduction
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@842 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 Section 2 covers the state of the art in web services publication, discovery and wrapping/consumption.
-Section 3 describes our platform for publishing web services and empowering end-users to discover services.
-Next, in Section 4 we present how we tackle with the web services wrapping, available possibilies and limitations.
+In Section 4 we present how we tackle with the web services wrapping, available possibilies and limitations.
+Next, section 3 describes our platform for publishing web services and empowering end-users to discover services.
 Section 5 gives the reader a clearer use case and how this work is integrated into the FAST platrofmr, and finally, we present our main conclusions and highlight future lines of research in Section 6.

### 2010-06-14 trunk/papers/icsoc2010/paper.tex
split it into different files
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@844 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-\begin{abstract}
-B2B systems integration was revolutionized by the introduction of web services.
-The way enterprise systems communicated was dramatically transformed, and by adopting these technologies, the relationships with providers and customers were strengthened.
-The main advantages seen by companies which adopted the technology were an increase of operational efficiencies, and reduction of costs.
-In this scenario, high qualified software developers tackle with the integration, by means of the understanding of WSDL documents or human-readable specifications.
-However this model fails when it tries to target the long tail of enterprise software demand, and as a result, the end-users.
-Discovery and consumption of web services is far from a straight-forward task for an end-user, meaning potential users, willing to create their task-specific applications, have been ruled out.
-This paper presents an approach to facilitate the discovery and consumption of web services by end-users on the Internet, breaking the gap between business web services and the latter.
-The process includes: (a) a way to generate ready-to-use web services wrappers, and (b) a catalogue which users can browse to search for web services fitting their needs.
-These web services wrappers are described by a model created, although highly influenced by well-known web services modeling standards.
-Furthermore, two prototypes have been developed to serve as a proof of concept.
-\keywords{web services, linked data, end-users, discovery, consumption}
-\end{abstract}
+\input{paper_abstract}
 
-\section{Introduction}
+\input{paper_sec_introduction}
 
-Business-to-business or B2B integration is easier said than done.
-Integration is a big challenge, in most cases requires a huge amount of effort, and deal with many factors and tecnologies for protocols, architectures, security standards, among others.
-Solutions based on open standards have reduced the complexity of integrating different business applications between different companies and partners.
-Web services and other XML standards such as RosettaNet, ebXML or OAG have been a boon to the world of B2B.
-By web services standards we are refering to the following open standards: Web Services Description Language (WSDL - to describe), Universal Description, Discovery and Integration (UDDI - to advertise and syndicate), Simple Object Access Protocol (SOAP - to communicate) and Web Services Flow Language (WSFL - to define work flows).
-Therefore, a common scenario is a company using WSDL to describe its public and private web service, publishing them either using a public or private repository using UDDI, where these web services using SOAP-based messages to achieve dynamic integration between different disparate applications.
+\input{paper_sec_related_work}
 
-Dealing with Business-to-consumer (B2C) integration, the adoption of these standards have offered some advantages as well, however the interaction and consumption of web services requires a programming skills and understanding of many technical details about the technology, which is an obstacle for an end-user with limited knowledge on the matter.
-Regarding discovery, most of the publishing and discovery platforms are syntactic-base, leading to poor precision and recall, making difficult to navigate through a large variety of web services \cite{pilioura_acm2009}, and as a result not empowering the end-users to tackle easily with this task.
-Other solutions like Semantic Web Service promised many advantages in discovery, composition and consumption of web services, independently of the provider's platform, the location, the service implementation, or the data format.
-However, their adoption was not as expected [TODO: any reference?], the further requirements needed and extra work have not being seen to be worth by enterprises for the benefits they offer.
-In the same direction as B2C, goverments are currently opening their data to the public as linked data.
-So far, the efforts are being focused on data, and the users have to deal with the development of applications to use that data.
-It is easily to predict that sooner or later they will move to open their services for administration tasks and single-window applications.
-Hence, they will be benefict by the research done around these concepts and technologies.
+\input{paper_sec_wrapping}
 
-The motivation of our work has been strongly influenced by the end-users' needs.
-We are targeting non-skilled users in term of programming and software development knowledge, intending to empower them with tools to discover and consume web services in a straight-forward manner.
-The web services and their specifications are not published right away in the platform, since we were just reimplementing UDDI.
-However, their definition and behaviour are wrapped in what we call a ``web service wrapper''.
-The resulting products of the wrapping process are two artifacts: a specific definition of the web service to be used by this tool, and a ready-to-consume (inside a web browser) piece of code with the proper functions to invoke the web service.
+\input{paper_sec_catalogue}
 
-The remainder of this paper is organised as follows.
-Section 2 covers the state of the art in web services publication, discovery and wrapping/consumption.
-In Section 4 we present how we tackle with the web services wrapping, available possibilies and limitations.
-Next, section 3 describes our platform for publishing web services and empowering end-users to discover services.
-Section 5 gives the reader a clearer use case and how this work is integrated into the FAST platrofmr, and finally, we present our main conclusions and highlight future lines of research in Section 6.
+\input{paper_sec_use_case}
 
-\section{Related work}
-\label{sec:related_work}
+\input{paper_sec_conclusions}
 
-Web services have been around for a long time since they first appeared.
-One of their most important claims has been (syntactic) interoperability between third-party systems and applications based on different platforms / programming languages.
-System integration, within a company or between systems from different enterprises became easier with the adoption of the web services standard technologies, being the Web Service Description or WSDL the most extended standard used for the definition of web services.
-In fact, there are many integrated development environments (IDEs) that can deal with WSDL, and create a plain vanilla proxy class for the desired platform / programming language, getting a more ``developer-friendly'' API built around proxy classes, instead of having to know all request parameter names and possible values for the different calls to the Web service.
-Even though, in an integration of two or more systems scenario, developers must read and understand their specifications in order to accomplish the task, in a programmatically-fashion way.
-Another advantage of the use of WSDL is validation.
-Having the request messages (methods and parameters) and response messages (return parameters / structures) built upon XML schemas, a Web service engine can validate the message conforming to a schema.
-However, as it has been said before, the use of these languages and tools requires a qualified developer.
-In a inter-company system integration scenario these technologies are of high value, but when a company opens its services to the public, it is clear that the common end-user has been left off-side.
-
-There are several tools out there situated a step forward to facilitate the interaction with data sources and services from around the Web.
-Yahoo!
-Pipes provides a set of modules to access different kind of data sources, such as RSS feeds, a given web page (HTML code), Flickr images, Google base or the Yahoo!
-search engine, among others.
-These modules are developed by the widget platform itself, they do not allow users to build their own modules, and it lacks of a way to interact with REST or SOAP web service.
-Another tool worth to mention is Apatar, an enterprise-oriented data integration and ETL tool.
-It allows the interaction with databases (MySQL, PostgreSql, Oracle), powerful enterprise systems such as Salesforce CRM, SugarCRM and Goldmine CRM, and access to data sources such as XML or text files.
-It enables non-developers to design and perform data transformations in a visual way through its graphical interface; however it misses support for Web services.
-Yet another tool, in the same direction as Apatar, could be JackBe Presto, defined as an enterprise-ready mashup solution.
-Presto permits to integrate and extract data from a great variety of data sources such as RSS and ATOM feeds or XML-based sources, and empower the user to create connectors to use services from HP SOA Systinet and any Oracle information technologies including Oracle 9i/10g/11g, Oracle Fusion SOAs, and Oracle Applications.
-
-Summarising up, none of the solutions in the market really facilitate end-users to build up their own applications or widgets allowing the interaction with Web services created by third-party providers.
-The solutions found are mainly data-oriented (RSS feeds, databases, raw text).
-Several tools permits some sort of (Web) service integration, but are meant to be used within an entreprise level by savvy business users or developers.
-[TODO: rewrite this paragraph] This paper presents an approach to leverage the possibility of integrate REST or SOAP-based web service inside a widget or any application for a web browser (by creating service wrappers), and graphical tool will be provided to facilitate the process of building these service wrappers.
-
-In the context of publishing and discovering Web services, a service provider have well-known and widely used technologies to accomplish the task of publication, such as the Universal Description, Discovery and Integration (UDDI) or WS-Discovery.
-The UDDI service registry specification \cite{uddi2004} is currently one of the core standards in the Web service technology stack and an integral part of every major SOA vendor's technology strategy and offers to requesters the ability to discover services via the Internet.
-In a few words, UDDI serves as a centralised repository of WSDL documents.
-A similar concept is iServe \cite{pedrinaci_ores2010}.
-This platform aims to publish Web services as what they called Lined Services Linked Services --linked data describing services, storing Web service definitions in a semantic-annotation fashion so semantic Web technologies for Web service discovery and processing can make use of them.
-However, the platform does not deal with the step from the definition to the consumption of the services.
-
-A different approach for Web service discovery is the Web Services Dynamic Discovery (WS-Discovery) specification.
-The core of this approach is a multicast discovery protocol.
-Service providers and consumers  listen each other for new services specifications within a network, so there is no need of a centralised registry.
-As a drawback, WS-Discovery do not support Internet-scale discovery, therefore make if useless for our purposes.
-
-[TODO: rewrite/extend this] The rule based approach of the service wrapper tool has been inspired by triple graph grammars, cf.
-\cite{schurr94}.
-While general triple graph grammars allow to relate complex graph structures with each other, in our case the target of a rule is always a single object or attribute.
-This facilitates the whole mechanism, reasonably.
-
-\section{Wrapping web services (Kassel)}
-\label{sec:wrapping_web_services}
-
-%This section talks about the service wrapper tool (Kassel prototype)
-%
-%What it offers (afaik now it supports RESTful services, if support for SOAP-based services is also implemented, talk about it, otherwise explain how you expect to address it in the future work section
-%
-%How to create a wrapper (test-and-construct approach)
-%
-%Place this component within the FAST architecture
-%
-%Talk about formats: JSON and XML??
-
-In our approach, wrapping web services is done in two steps: a first step is in charge of the construction of a service request and a second step analyzes the response got from the execution of the service, allowing the analysis of response data and the mapping to domain-specific concepts.
-From this input, we generate JavaScript code to be embedded into web gadgets.
-This JavaScript code takes query parameters, constructs the service request, analysis the service response and returns result data in the desired format.
-
-\subsection{Constructing service requests} % (fold)
-\label{sub:constructing_service_requests}
-
-As a first approach, the interaction with services is limited to simple GET requests as supported e.g.
-by REST services.
-Support for other services like SOAP services is planned.
-A service request is thus assembled using a certain URL and a set of parameters.
-As an example, the Ebay Shopping web service will be studied.
-The following URL is invoked to retrieve a list of items corresponding to the search keywords \emph{USB}.
-
-\url{http://open.api.sandbox.ebay.com/shopping?appid=KasselUn-efea-4b93-9505-5dc2ef1ceecd&version=517&callname=FindItems&ItemSort=EndTime&QueryKeywords=USB&responseencoding=XML}
-
-As you may see, the URL invoked is http://open.api.sandbox.ebay.com/shopping and the parameters used in the example are:
-\begin{description}
-	\item[appid] this is the application ID obtained to use the API.
-	\item[version] the API version.
-	\item[callname] in this case FindItems to search through all items in Ebay.
-	\item[itemsort] sorting method for the list of items.
-	\item[querykeywords] list of keywords.
-	\item[responseencoding] format of the response message obtained by the invocation of the request.
-\end{description}
-
-If needed, a detailed specification of the Ebay Shopping API can be found at \cite{eBayShoppingAPIs}.
-
-The above URL for searching items in Ebay is followed by the query parameters, which take the form \textit{argument=value}.
-In this example, the only relevant parameter the end user may want to provide is \emph{querykeywords}.
-Thus the end user may use some input field to provide a query keyword and pass this value to our service wrapper operation.
-Other parameters can be set to a default value.
-However, in order to develop a more generic wrapper to a service, more or even all parameters might be passed by the user.
-
-To define the desired input parameters and the type of the expected result of a new service wrapper the FAST service wrapper tool provides a web based form that allows the editing of these entries, cf.
-Figure \ref{fig:construct_pre_post_conditions}.
-Note, we allow to edit example values for the input parameters.
-These example values may be used to test the service wrapper.
-
-%\begin{figure}
-%  \begin{center} \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithPortDefinitions.jpg}
-%    \caption{Configuring parameters of a service wrapper (photoshoped screen dump)}
-%    \label{fig:construct_pre_post_conditions}
-%  \end{center}
-%\end{figure}
-
-The service wrapper tool composes the request using a template string which will contain placeholders for input values.
-Before sending the request to the service, the placeholders are replaced with their corresponding values and the resulting URL is then ready to be sent as the service request.
-
-Figure ~\ref{fig:construct_service_request} shows the web based tool to construct the service requests.
-
-%\begin{figure}
-%  \begin{center}
-%    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithRequestExample.jpg}
-%    \caption{Constructing the service request URL and its parameters (photoshoped screen dump)}
-%    \label{fig:construct_service_request}
-%  \end{center}
-%\end{figure}
-
-In the top input field of Figure ~\ref{fig:construct_service_request}, the user may drop an example http request taken e.g.
-form the service documentation e.g.
-from \cite{eBayShoppingAPIs}.
-The tool analyses the example request and in the middle of the screen a form for editing the request parameters is provided.
-In our example, the user has connected the \textit{QueryKeywords} parameter with the input parameter \textit{search\_key} by adding a corresponding reference to the value field of that parameter.
-In addition, we have retrieved an access key which has been entered as value for the \textit{appid} parameter.
-
-Below the request parameter editing form of Figure ~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed URL by sending it to the specified service address (via a server relay).
-Then the placeholders for input parameters are replaced by their example values and the resulting http request is shown below the parameter form.
-In addition, the request is send and the response is shown on the bottom of that page.
-This gives the user a fast feedback whether the constructed request works as desired.
-
-\subsubsection{Limitations} % (fold)
-\label{ssub:limitations}
-
-It is worth pointing out that currently the wrapper tool is able to construct input ports for the wrappers using just basic types.
-To construct URL requests from more complex input parameters, as e.g.
-a person object, we are currently developing access operations that will allow to access the values of fields of complex objects.
-For example, the expression \emph{customer.address} may refer to the \emph{address} field of a \emph{customer} parameter.
-
-% subsubsection limitations (end)
-
-% subsection constructing_service_requests (end)
-
-\subsection{Interpreting service responses} % (fold)
-\label{sub:interpreting_service_responses}
-
-Once the service request is constructed and sent to the service provider, it will send back a response.
-This response message could be serialized in any format, though the most common formats used nowadays are XML or JSON among others.
-To continue the example started in the previous section, the response of the Ebay Shopping service will be in XML as specified in the request.
-
-\subsubsection{Translation XML into Facts} % (fold)
-\label{ssub:translation_xml_into_facts}
-
-Figure ~\ref{fig:response_service_execution} shows the data transformation tab of the wrapper tool.
-
-%\begin{figure}
-%  \begin{center}
-%      \includegraphics[angle=90,width=0.8\linewidth]{images/ServiceWrapperToolGVSWithTransformationRules.jpg}
-%    \caption{Interactive, rule based transforming of an XML response to domain objects (photoshoped screen %dump)}
-%    \label{fig:response_service_execution}
-%  \end{center}
-%\end{figure}
-
-Once the service response, in XML format, has been retrieved, the transformation tab shows it as an interactive object tree on the left side of Figure ~\ref{fig:response_service_execution}.
-To construct this interactive object tree, the XML document has been parsed into a DOM, and a simplified tree representation of that DOM is built up.
-
-A transformation rule is used to analyze the XML data and to generate domain-specific objects from concepts from the ontologies used by the parameters of the different building blocks.
-A transformation rule is composed of three elements, cf.
-the middle part of Figure ~\ref{fig:response_service_execution}.
-First, the \textit{from} field indicates the XML elements to be translated by the rule.
-These XML elements are identified by the tag name of a DOM element from the XML document.
-Second, the type of the rule will be set, taking one of the following values: \emph{createObject}, \emph{fillAttributes} or \emph{dummy}.
-And third, the target of the rule specifies a certain concept or attribute, to be created or filled.
-A detailed explanation of the type of actions to be trigger from the transformation rules is:
-\begin{description}
-	\item[action \emph{createObject}] specifies the creation of a new domain object.
-The type of that new object is provided in the third compartment.
-In the example being explained, the root rule searches for XML elements with tag name \emph{FindItemsResponse} and for each such element a \emph{List} object is created.
-The resulting objects are shown in a facts tree in the right of Figure ~\ref{fig:response_service_execution}.
-	\item[action \emph{fillAttributes}] does not create a new object but it fills the value of the attribute provided as third part of such rules.
-In our example, the third transformation rule searches for XML elements with tag name \emph{Title}.
-Note, the rule is a sub-rule of the second rule, which generates \emph{Product} objects.
-Thus, the sub-rule searches for \emph{Title} tags only in the subtree of the XML data that has been identified by an application of the parent rule before.
-For example, the \emph{Item} rule may just have been applied to the first \emph{Item} element of the XML data.
-Then, the \emph{Title} rule is applied only to the first \emph{Item} sub-tree of the XML data and thus it will find only one \emph{Title} element in that sub-tree (not visible in Figure ~\ref{fig:response_service_execution} ).
-The value of that \emph{Title} element is then transfered to the \emph{productName} attribute of the corresponding \emph{Product} object.
-Actually, our \textit{from} fields allows also to refer to parts of an XML attribute e.g.
-to \textit{words} 1 through 3.
-It is also possible to combine constant text and elements of multiple XML tree elements.
-	\item[action \emph{dummy}] does not create or modify any objects but such rules are just used to narrow the search space for their sub-rules.
-For example, in Amazon product data, the XML data for an item contains sections for \emph{minimum price}, \emph{maximum price}, and \emph{average price}.
-Each such section contains the \emph{plain price} and the \emph{formatted price}.
-Thus, in the Amazon case, a rule that searches for \emph{formatted price} elements within an \emph{Item} element would retrieve three matches.
-Using a dummy rule, we may first search for \emph{minimum price} elements and then search for \emph{formatted price} elements within that sub-tree.
-\end{description}
-
-Our tool follows an interactive paradigm.
-Therefore, any time a change to a transformation rule is done, the transformation process is triggered and the resulting facts tree is directly shown.
-This process helps the user to deal with the slightly complex semantics of the transformation rules avoiding errors or mistakes.
-In addition, out tool is ontology-driven, therefore, the service wrapper designer shall retrieve the domain-specific types from a FAST ontology server together with the structure of each type, i.e.
-together with a description of the attributes of each object.
-Thus, the transformation rule editor is able to provide selection boxes for the target element of the rules.
-For a \emph{createObject} rule, this selection box shows the object types available for that domain.
-For the \emph{fillAttributes} rules, the selection box shows the attributes of the object type chosen in the parent rule.
-In addition, we may provide some analysis tool, which will help to guarantee that the objects generated by the transformation rules conform to the object types defined in the corresponding FAST ontology.
-This helps to ensure that the objects generated by the designed service wrapper will be compatible for input parameters of subsequent filter steps and or gadgets.
-
-% subsubsection translation_xml_into_facts (end)
-
-% subsection interpreting_service_responses (end)
-
-\subsection{Generating a Resource Adapter} % (fold)
-\label{sub:generating_a_resource_adapter}
-
-Once the wrapping of a service has been defined and tested in the service wrapper tool, we generate an implementation of the desired Resource Adapter in XML, HTML, and JavaScript, ready to be deployed and executed inside a web gadget.
-
-% subsection generating_a_resource_adapter (end)
-
-\subsection{Limitations} % (fold)
-\label{sub:limitations}
-
-The rule driven approach presented above is somewhat limited.
-It is deliberately restricted to such a simple rule mechanism in order to keep things simple enough for end-users.
-Still, the selected approach suffices for many practical and real world cases.
-As a more complex example, the XML data for a person may provide two different tags for the first and the last name of a person.
-Contrarily, a person fact which conforms to a certain ontology for that domain may provide only one \emph{fullname} attribute that shall be filled by a concatenation of the first and the last name.
-To achieve this, the \textit{from} field of that tranforamtion rule might look like: \texttt{lastname"', "`firstname}.
-We are also able to do some navigation in the XML tree to follow XRef elements.
-For example the attribute \texttt{grandmother} could be filled using \texttt{mother.mother} in the \textit{from} field.
-
-However, there are some transformations that these rules cannot perform.
-For example, we do not support any mathematical operations.
-Thus, transforming e.g.
-Fahrenheit into Celsius temperatures is not supported.
-To cover such  cases, intermediate object formats can be used which would allow generating objects to be further processed by additional filters.
-Such additional filters may be realized using (hand coded) operators, since some generic operators can act as filters for aggregation and conversions of objects from multiple sources.
-Then, service wrappers in combination with these filter operators will allow covering these complex cases.
-
-% subsection limitations (end)
-
-% section restful_web_services_wrapper_tool (end)
-
-\section{Discovery: the marketplace}
-\label{sec:discovery}
-
-The area of web service publication and discovery has been subject for a lot of research since the very beginning the concept was coined.
-The current state of the art provides several solutions and strategies which providers and consumers may take advantage of.
-However, as explained in previous sections, a large number of them suffer from limitated syntactic-based descriptions and a simple keyword-based search, and the gap between discovery and consumption is an obstacle for their usage by a non-technical common user.
-In order to ligthen the issue, this paper presents yet another publishing platform, permiting any enterprise or individual to publish their public domain web services, with an enhanced semantic search based on the definition of the services, and serving``web service wrappers'' for easily consumption in any web application.
-
-\subsection{Overview}
-\label{ssec:overview}
-
-The main difference of this platform with regards to the state of the art is that it is clearly targeting a different kind of user, and covers many deficiencies others tools does not tackle with.
-As a briefly overview, the platform being presented:
-
-\begin{itemize}
-	\item offers web services as Linked Data;
-	\item serves web services wrappers ready-to-consume in any web application;
-	\item follows well-known best practices for publishing data on the web;
-	\item supports managing its resources via its RESTful API;
-	\item supports content negotiation so the different clients may retrieved the information in JSON, RDF/XML, RDF/N3 or a human-readable HTML version;
-	\item provices advanced search capabilites, based on the service formal definition and inferences extracted from it;
-	\item a SPARQL endpoint to access directly to the data allowing advanced queries;
-\end{itemize}
-
-This platform was intented and it is being used as part of the FAST storyboard tool. Others components communicates with it via its RESTful API, but a
-That said, The process of publishing a web service requires two steps: (i) generation of the wrapper (see Sec.
-\ref{sec:wrapping_web_services}, and (ii) creation of the resource in the publishing platform.
-The creation of the resources is done via the public RESTful API, and the language selected for the input for any operation is JSON.
-The following is an example of part of a request sent to create a new wrapper into the platform:
-
-\begin{listing}
-\begin{verbatim}
-{
-  "code": "http://demo.fast.morfeo-project.org/code/amazonService.js",
-  "creationDate": "2010-01-26T17:01:13+0000",
-  "description": {"en-gb": "Amazon web service"},
-  "label": {"en-gb": "Amazon web service"},
-  "preconditions": [
-    [
-      {
-        "id": "item",
-        "label": { "en-gb": "An item" },
-        "pattern": "?I http://www.w3.org/1999/02/22-rdf-syntax-ns#type http://aws.amazon.com/AWSECommerceService#Item",
-        "positive": "true"
-      }
-    ]
-  ],
-  "postconditions": [
-    ...
-  ],
-  ...
-\end{verbatim}
-\label{lis:json_request}
-\end{listing}
-
-The architecture, as presented in the Figure \ref{fig:}, comprises a RDF store used for persistence, a business layer dealing with the model and reasoning, and public façade providing the core functionality as a RESTful API, and a SPARQL endpoint accessing directly the RDF store.
-This presentation layer is aimed to interact with the FAST Gadget Visual Storyboard, or any other third-party application.
-
-The RESTful API and the SPARQL endpoint are part of the presentation layer.
-The SPARQL endpoint is offered using the SPARQL protocol service as defined in the SPROT specification \cite{sprot} and is aimed to enable third-party developers to query directly the knowledge base using SPARQL queries.
-This feature is supported by the Sesame RESTful HTTP interface for SPARQL Protocol for RDF.
-
-The business logic layer contains all the domain-specific processing and reasoning.
-It provides functions to interact with all the elements of the domain model specified in \cite{sec:model}, acting as a mediator between the public RESTful API and the persistence layer.
-
-The persistence layer provides an API to work with a standard set of objects representing the model.
-The interaction with the underlying RDF store is made via RDF2Go, an abstraction over triple (and quad) stores, which allows to program against rdf2go interfaces and choose any the RDF store implementation.
-This allows having a completely extensible and configurable framework for storage mechanisms, inferencers, RDF file formats, query result formats and query languages...
-
-How to publish (RESTful app, anyone can create a wrapper and publish it)
-
-How to browse: will allow users to go through the given domains (ontologies) and concepts, and will explain how to find the services they want (operation Find of the catalogue) => how the recommendation/reasoning works
-
-It offers the wrapped web services as linked data and has an open sparql endpoint, allowing other developers to create applications on top of it...
-
-\subsection{Serving Linked Data} % (fold)
-\label{sub:linked_data}
-
-The recently success of the web of Data has influenced the way information is now published on the web, and has been widely adopted by the academic community, large companies such as The New York Times and BBC, and national governments such as US and UK made public commitments toward open data.
-The main aim of the linking data is to find other related data based upon previously known data, in terms of the connections or links between them.
-We apply this concept in order to provide metadata about the web services as linked data, following the principles as originally defined in \cite{bizer_ijswis2009}, in order to make them available to arbitrary third-party applications.
-Each web service is identified by an HTTP URI and hosted in the catalogue so that it can be dereferenced through the same URI.
-For each building block, data is available in representations in different standard formats such as JSON (for communication with the GVS), RDF/XML, Turtle, or even HTML+RDFa as a human-readable version.
-The principles and best practices proposed in \cite{berrueta2008} and \cite{sauermann2008cool_uris} are also taken into account, these representations are served based on the request issued by the requesting agent, using a technique called \emph{content negotiation}.
-As required by the forth rule of linked data, individual building blocks link to other data on the web, thereby preventing so-called isolated ``data islands''.
-
-In order to understand how content negotiation is implemented in the catalogue, a few concepts need to be understood.
-The URI of a particular building block must be understood as the identifier of the building block as such, as opposed to a particular representation (JSON, HTML, etc.) of it.
-In fact, each such representation has its own URI, under which it can be retrieved.
-Using the terminology established in \cite{w3c2004webArchitecture}, the building block is a \emph{non-information resource}, whereas each representation is an \emph{information resource}.
-In the process of content negotiation, the requesting agent will first request the URI of the building block as such (e.g., \texttt{http://catalogue.com/screens/235}), as well as the required representation format (e.g., JSON).
-The catalogue will then inspect the request and redirect the requester to the appropriate representation (e.g., \texttt{http://catalogue.com/screens/235.json}), as illustrated in Fig.~\ref{fig:content_negotiation}.
-On the web, different representations for the same resource are called \emph{variants}, and content negotiation is the mechanism used to determine which of the representations of most appropriate for a given request.
-
-%\begin{figure}[htb]
-%\label{fig:content_negotiation}
-%\begin{center}
-%	\includegraphics[width=12cm]{images/content_negotiation}
-%	\caption{Content Negotiation}
-%\end{center}
-%\end{figure}
-
-In summary, the basic idea of content negotiation, as stated in the \cite{http1.1}, is to serve the best variant for a resource, taking into account what variants are available, what variants the server may prefer to serve, what the client can accept, and with which preferences.
-In HTTP, this is done by the client which may send, in its request, accept headers (\texttt{Accept}, \texttt{Accept-Language} and \texttt{Accept-Encoding}), to communicate its capabilities and preferences in format, language and encoding, respectively.
-
-In fact, what the catalogue really does is ``format negotiation'' since the alternate representations are just based on the selection of the media type, through the accept header, but does not consider different languages or encoding types.
-The formats supported are JSON, RDF/XML, Turtle and HTML+RDFa.
-Even though the accept header is desired, the different representations can also be retrieved directly by dereferencing their own URI.
-The convention used in the catalogue for deriving a representation URI from a building block URI is to simply attach a matching suffix, such as \texttt{.rdf}, \texttt{.json} or \texttt{.html}.
-
-Lastly, content negotiation needs to identify which player is going to take the lead on it.
-There are two kinds of content negotiation which are possible in HTTP: server-driven and agent-driven negotiation.
-The approach followed by the catalogue is agent-driven negotiation, hence selecting a specific representation for a resource is responsibility of the user agent.
-If none is specified, by default, the server will choose the JSON representation.
-
-\subsection{Simplified Model/Ontology (Knud)}
-The model used to define the web service wrappers within this application has been influenced by WSDL, the main web service definition language, and some semantic-enriched approaches such as WSMO and OWL-S.
-There is no need to reinvent the wheel,
-
-A WSDL operation is mapped to an action. Its inputs and outputs are called pre-postconditions (a WSMO concept).
-
-This section could talk about a the part of the ontology describing the back-end services, how we remove the unnecessary information from WSDL and similar (not need anymore because the code has been generated), however we reuse some of the concepts from these technologies to define them (preconditions, postconditions, actions, and so on) allowing discovery and composition.
-
-%\begin{figure}[ht]
-%	\centering
-%	\includegraphics[width=8.31cm]{uddi.png}
-%	\caption{UDDI overview}
-%	\label{fig:uddi}
-%\end{figure}
-
-%\section{Wrapping web services (Kassel)}
-%\label{sec:wrapping_web_services}
-
-%This section talks about the service wrapper tool (Kassel prototype)
-
-%What it offers
-
-%How to create a wrapper (test-and-construct approach)
-
-%Place this component within the FAST architecture
-
-%Talk about formats: JSON and XML??
-
-%\subsection{Code-generation phase}
-%How the executable code is build:
-
-% - mapping from operations and parameters to actions and preconditions
-
-% - mapping from operation results to postconditions
-
-%Now, the calls are done through the FAST api, which acts as a proxy <- imply they need to construct the widget using the FAST GVS which doesn't give an open solution (proposal in future work)
-
-\section{Use case}
-maybe we should provide some kind of use case to show the benefits of the approach and the possibilities this sort of tools?
-
-\section{Conclusions and future work}
-
-Future work:
-
-Talk about wrapping WSDL-based services
-
-Talk about WSMO-lite services
-
-Use JSONRequest to replace the AJAX XMLHttpRequest and its problems (cross-domain calls), so there's no need for a Proxy.
-
-\subsubsection*{Acknowledgments.} This work is supported in part by the European Commission
-under the first call of its Seventh Framework Program (FAST STREP Project, grant
-INFSO-ICT-216048) and in part by Science Foundation Ireland under Grant No.
+\subsubsection*{Acknowledgments.} This work is supported in part by the European Commission under the first call of its Seventh Framework Program (FAST STREP Project, grant INFSO-ICT-216048) and in part by Science Foundation Ireland under Grant No.
 SFI/08/CE/I1380 (L\'ion-2).

### 2010-06-14 trunk/papers/icsoc2010/paper_abstract.tex
split it into different files
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@844 d8b0c59b-6544-0410-94d9-83a5658a058e

\begin{abstract}
B2B systems integration was revolutionized by the introduction of web services. The way enterprise systems communicated was dramatically transformed, and by adopting these technologies, the relationships with providers and customers were strengthened. The main advantages seen by companies which adopted the technology were an increase of operational efficiencies, and reduction of costs. In this scenario, high qualified software developers tackle with the integration, by means of the understanding of WSDL documents or human-readable specifications. However this model fails when it tries to target the long tail of enterprise software demand, and as a result, the end-users. Discovery and consumption of web services is far from a straight-forward task for an end-user, meaning potential users, willing to create their task-specific applications, have been ruled out. This paper presents an approach to facilitate the discovery and consumption of web services by end-users on the Internet, breaking the gap between business web services and the latter. The process includes: (a) a way to generate ready-to-use web services wrappers, and (b) a catalogue which users can browse to search for web services fitting their needs. These web services wrappers are described by a model created, although highly influenced by well-known web services modeling standards. Furthermore, two prototypes have been developed to serve as a proof of concept.
\keywords{web services, linked data, end-users, discovery, consumption}
\end{abstract}

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_catalogue.tex
split it into different files
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@844 d8b0c59b-6544-0410-94d9-83a5658a058e

\section{Discovery: the marketplace}
\label{sec:discovery}

The area of web service publication and discovery has been subject for a lot of research since the very beginning the concept was coined. The current state of the art provides several solutions and strategies which providers and consumers may take advantage of. However, as explained in previous sections, a large number of them suffer from limitated syntactic-based descriptions and a simple keyword-based search, and the gap between discovery and consumption is an obstacle for their usage by a non-technical common user. In order to ligthen the issue, this paper presents yet another publishing platform, permiting any enterprise or individual to publish their public domain web services, with an enhanced semantic search based on the definition of the services, and serving``web service wrappers'' for easily consumption in any web application.

\subsection{Overview}
\label{ssec:overview}

The main difference of this platform with regards to the state of the art is that it is clearly targeting a different kind of user, and covers many deficiencies others tools does not tackle with. As a briefly overview, the platform being presented:

\begin{itemize}
	\item offers web services as Linked Data;
	\item serves web services wrappers ready-to-consume in any web application;
	\item follows well-known best practices for publishing data on the web;
	\item supports managing its resources via its RESTful API;
	\item supports content negotiation so the different clients may retrieved the information in JSON, RDF/XML, RDF/N3 or a human-readable HTML version;
	\item provices advanced search capabilites, based on the service formal definition and inferences extracted from it;
	\item a SPARQL endpoint to access directly to the data allowing advanced queries;
\end{itemize}

This platform was intented and it is being used as part of the FAST storyboard tool. Others components communicates with it via its RESTful API, but a
That said, The process of publishing a web service requires two steps: (i) generation of the wrapper (see Sec. \ref{sec:wrapping_web_services}, and (ii) creation of the resource in the publishing platform. The creation of the resources is done via the public RESTful API, and the language selected for the input for any operation is JSON. The following is an example of part of a request sent to create a new wrapper into the platform:

\begin{listing}
\begin{verbatim}
{
  "code": "http://demo.fast.morfeo-project.org/code/amazonService.js",
  "creationDate": "2010-01-26T17:01:13+0000",
  "description": {"en-gb": "Amazon web service"},
  "label": {"en-gb": "Amazon web service"},
  "preconditions": [
    [
      {
        "id": "item",
        "label": { "en-gb": "An item" },
        "pattern": "?item 
                    http://www.w3.org/1999/02/22-rdf-syntax-ns#type
                    http://aws.amazon.com/AWSECommerceService#Item",
        "positive": "true"
      }
    ]
  ],
  "postconditions": [
    ...
  ],
  ...
\end{verbatim}
\label{lis:json_request}
\end{listing}

The architecture, as presented in the Figure \ref{fig:}, comprises a RDF store used for persistence, a business layer dealing with the model and reasoning, and public façade providing the core functionality as a RESTful API, and a SPARQL endpoint accessing directly the RDF store. This presentation layer is aimed to interact with the FAST Gadget Visual Storyboard, or any other third-party application.

The RESTful API and the SPARQL endpoint are part of the presentation layer. The SPARQL endpoint is offered using the SPARQL protocol service as defined in the SPROT specification \cite{sprot} and is aimed to enable third-party developers to query directly the knowledge base using SPARQL queries. This feature is supported by the Sesame RESTful HTTP interface for SPARQL Protocol for RDF.

The business logic layer contains all the domain-specific processing and reasoning. It provides functions to interact with all the elements of the domain model specified in \cite{sec:model}, acting as a mediator between the public RESTful API and the persistence layer.

The persistence layer provides an API to work with a standard set of objects representing the model. The interaction with the underlying RDF store is made via RDF2Go, an abstraction over triple (and quad) stores, which allows to program against rdf2go interfaces and choose any the RDF store implementation. This allows having a completely extensible and configurable framework for storage mechanisms, inferencers, RDF file formats, query result formats and query languages...

How to publish (RESTful app, anyone can create a wrapper and publish it)

How to browse: will allow users to go through the given domains (ontologies) and concepts, and will explain how to find the services they want (operation Find of the catalogue) => how the recommendation/reasoning works

It offers the wrapped web services as linked data and has an open sparql endpoint, allowing other developers to create applications on top of it...

\input{paper_ssec_ontology}

\input{paper_ssec_linked_data}

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_conclusions.tex
split it into different files
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@844 d8b0c59b-6544-0410-94d9-83a5658a058e

\section{Conclusions and future work}

Future work:

Talk about wrapping WSDL-based services

Talk about WSMO-lite services

Use JSONRequest to replace the AJAX XMLHttpRequest and its problems (cross-domain calls), so there's no need for a Proxy.

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_introduction.tex
split it into different files
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@844 d8b0c59b-6544-0410-94d9-83a5658a058e

\section{Introduction}

Business-to-business or B2B integration is easier said than done. Integration is a big challenge, in most cases requires a huge amount of effort, and deal with many factors and tecnologies for protocols, architectures, security standards, among others. Solutions based on open standards have reduced the complexity of integrating different business applications between different companies and partners. Web services and other XML standards such as RosettaNet, ebXML or OAG have been a boon to the world of B2B. By web services standards we are refering to the following open standards: Web Services Description Language (WSDL - to describe), Universal Description, Discovery and Integration (UDDI - to advertise and syndicate), Simple Object Access Protocol (SOAP - to communicate) and Web Services Flow Language (WSFL - to define work flows). Therefore, a common scenario is a company using WSDL to describe its public and private web service, publishing them either using a public or private repository using UDDI, where these web services using SOAP-based messages to achieve dynamic integration between different disparate applications.

Dealing with Business-to-consumer (B2C) integration, the adoption of these standards have offered some advantages as well, however the interaction and consumption of web services requires a programming skills and understanding of many technical details about the technology, which is an obstacle for an end-user with limited knowledge on the matter. Regarding discovery, most of the publishing and discovery platforms are syntactic-base, leading to poor precision and recall, making difficult to navigate through a large variety of web services \cite{pilioura_acm2009}, and as a result not empowering the end-users to tackle easily with this task. Other solutions like Semantic Web Service promised many advantages in discovery, composition and consumption of web services, independently of the provider's platform, the location, the service implementation, or the data format. However, their adoption was not as expected [TODO: any reference?], the further requirements needed and extra work have not being seen to be worth by enterprises for the benefits they offer. In the same direction as B2C, goverments are currently opening their data to the public as linked data. So far, the efforts are being focused on data, and the users have to deal with the development of applications to use that data. It is easily to predict that sooner or later they will move to open their services for administration tasks and single-window applications. Hence, they will be benefict by the research done around these concepts and technologies.

The motivation of our work has been strongly influenced by the end-users' needs. We are targeting non-skilled users in term of programming and software development knowledge, intending to empower them with tools to discover and consume web services in a straight-forward manner. The web services and their specifications are not published right away in the platform, since we were just reimplementing UDDI. However, their definition and behaviour are wrapped in what we call a ``web service wrapper''. The resulting products of the wrapping process are two artifacts: a specific definition of the web service to be used by this tool, and a ready-to-consume (inside a web browser) piece of code with the proper functions to invoke the web service.

The remainder of this paper is organised as follows. Section 2 covers the state of the art in web services publication, discovery and wrapping/consumption. In Section 4 we present how we tackle with the web services wrapping, available possibilies and limitations. Next, section 3 describes our platform for publishing web services and empowering end-users to discover services. Section 5 gives the reader a clearer use case and how this work is integrated into the FAST platrofmr, and finally, we present our main conclusions and highlight future lines of research in Section 6.

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_related_work.tex
split it into different files
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@844 d8b0c59b-6544-0410-94d9-83a5658a058e

\section{Related work}
\label{sec:related_work}

Web services have been around for a long time since they first appeared. One of their most important claims has been (syntactic) interoperability between third-party systems and applications based on different platforms / programming languages. System integration, within a company or between systems from different enterprises became easier with the adoption of the web services standard technologies, being the Web Service Description or WSDL the most extended standard used for the definition of web services. In fact, there are many integrated development environments (IDEs) that can deal with WSDL, and create a plain vanilla proxy class for the desired platform / programming language, getting a more ``developer-friendly'' API built around proxy classes, instead of having to know all request parameter names and possible values for the different calls to the Web service. Even though, in an integration of two or more systems scenario, developers must read and understand their specifications in order to accomplish the task, in a programmatically-fashion way. Another advantage of the use of WSDL is validation. Having the request messages (methods and parameters) and response messages (return parameters / structures) built upon XML schemas, a Web service engine can validate the message conforming to a schema. However, as it has been said before, the use of these languages and tools requires a qualified developer. In a inter-company system integration scenario these technologies are of high value, but when a company opens its services to the public, it is clear that the common end-user has been left off-side.

There are several tools out there situated a step forward to facilitate the interaction with data sources and services from around the Web. Yahoo! Pipes provides a set of modules to access different kind of data sources, such as RSS feeds, a given web page (HTML code), Flickr images, Google base or the Yahoo! search engine, among others. These modules are developed by the widget platform itself, they do not allow users to build their own modules, and it lacks of a way to interact with REST or SOAP web service. Another tool worth to mention is Apatar, an enterprise-oriented data integration and ETL tool. It allows the interaction with databases (MySQL, PostgreSql, Oracle), powerful enterprise systems such as Salesforce CRM, SugarCRM and Goldmine CRM, and access to data sources such as XML or text files. It enables non-developers to design and perform data transformations in a visual way through its graphical interface; however it misses support for Web services. Yet another tool, in the same direction as Apatar, could be JackBe Presto, defined as an enterprise-ready mashup solution. Presto permits to integrate and extract data from a great variety of data sources such as RSS and ATOM feeds or XML-based sources, and empower the user to create connectors to use services from HP SOA Systinet and any Oracle information technologies including Oracle 9i/10g/11g, Oracle Fusion SOAs, and Oracle Applications.

Summarising up, none of the solutions in the market really facilitate end-users to build up their own applications or widgets allowing the interaction with Web services created by third-party providers. The solutions found are mainly data-oriented (RSS feeds, databases, raw text). Several tools permits some sort of (Web) service integration, but are meant to be used within an entreprise level by savvy business users or developers. [TODO: rewrite this paragraph] This paper presents an approach to leverage the possibility of integrate REST or SOAP-based web service inside a widget or any application for a web browser (by creating service wrappers), and graphical tool will be provided to facilitate the process of building these service wrappers.

In the context of publishing and discovering Web services, a service provider have well-known and widely used technologies to accomplish the task of publication, such as the Universal Description, Discovery and Integration (UDDI) or WS-Discovery. The UDDI service registry specification \cite{uddi2004} is currently one of the core standards in the Web service technology stack and an integral part of every major SOA vendor's technology strategy and offers to requesters the ability to discover services via the Internet. In a few words, UDDI serves as a centralised repository of WSDL documents. A similar concept is iServe \cite{pedrinaci_ores2010}. This platform aims to publish Web services as what they called Lined Services Linked Services --linked data describing services, storing Web service definitions in a semantic-annotation fashion so semantic Web technologies for Web service discovery and processing can make use of them. However, the platform does not deal with the step from the definition to the consumption of the services.

A different approach for Web service discovery is the Web Services Dynamic Discovery (WS-Discovery) specification. The core of this approach is a multicast discovery protocol. Service providers and consumers  listen each other for new services specifications within a network, so there is no need of a centralised registry. As a drawback, WS-Discovery do not support Internet-scale discovery, therefore make if useless for our purposes.

[TODO: rewrite/extend this] The rule based approach of the service wrapper tool has been inspired by triple graph grammars, cf. \cite{schurr94}. While general triple graph grammars allow to relate complex graph structures with each other, in our case the target of a rule is always a single object or attribute. This facilitates the whole mechanism, reasonably.

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_use_case.tex
split it into different files
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@844 d8b0c59b-6544-0410-94d9-83a5658a058e

\section{Use case}
\label{sec:use_case}

The two artifacts presented in this paper may be deployed and used separately by third-party applications. The intention was to show the advances made regarding web service discovery and consumption. However, they were developed in order to be part of the FAST platform. FAST constitutes a novel approach to application composition and business process definition from a top-down user-centric perspective. Therefore, it is aimed to allow users to create their own applications by combining different components, or building blocks, such as graphical forms and back-end services in terms of their inputs and outputs. These two components, as described in Section 3 and 4, are fully integrated and present a high cohesion the rest of the components of the FAST architecture (see Picture \ref{pic:fast_architecture}.

[TODO: create a simple picture of the FAST architecture, highlighting these two components.]

For a better understanding of the use case, it is worth to describe a few concepts related to FAST. A Gadget is the output product, aimed to be used in a specific mashup platform. Inside FAST, the platform-independent gadget is called Screenflow, which comprises a set of domain-specific Screens connected through their pre and postconditions. A screen is the most complex building block fully functional by itself. It is composed by a form conveying the graphical interface, and a set of operators and resources binding to the back-end services.

Then, the web service wrapper tool, which is used to create the formal definitions of back-end services, transforming them into building blocks to be stored and reused within the FAST platform. It is integrated into the graphical IDE, and for the storage interacts with the catalogue, where the formal service descriptions and the executable code encapsulating the service behaviour is stored.

The publishing platform presented, called ``Resource Catalogue'' within the FAST platform, covers several important purposes:
\begin{description}
	\item[Publication and search of gadgets] The output of the FAST IDE are gadgets, which are self-contained and ready to use for users. They need to be made available for download, as well as being discover according to their formal description.
	\item[Publication and search of building blocks] Every single component or building block definition is stored in the catalogue and it provides powerful ways of searching, recommendation and planning. 
	\item[Storage and indexing of user profiles and histories] Each user of the FAST platform will have a representation in the catalogue, thus making it possible for the IDE to make informed guesses about what they want in a particular situation, based on their previous behaviour.
	\item[Facilitate ontology mediation] Because FAST allows the integration of third-party external services, it is not always guaranteed that all the data coming from a service is directly compatible with the FAST platform. In such cases, mediation between the different ontologies is necessary. The catalogue provide some grade of mediation functionality.
\end{description}

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_wrapping.tex
split it into different files
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@844 d8b0c59b-6544-0410-94d9-83a5658a058e

\section{Wrapping web services (Kassel)}
\label{sec:wrapping_web_services}

%This section talks about the service wrapper tool (Kassel prototype)
%
%What it offers (afaik now it supports RESTful services, if support for SOAP-based services is also implemented, talk about it, otherwise explain how you expect to address it in the future work section
%
%How to create a wrapper (test-and-construct approach)
%
%Place this component within the FAST architecture
%
%Talk about formats: JSON and XML??

In our approach, wrapping web services is done in two steps: a first step is in charge of the construction of a service request and a second step analyzes the response got from the execution of the service, allowing the analysis of response data and the mapping to domain-specific concepts. From this input, we generate JavaScript code to be embedded into web gadgets. This JavaScript code takes query parameters, constructs the service request, analysis the service response and returns result data in the desired format.

\subsection{Constructing service requests} % (fold)
\label{sub:constructing_service_requests}

As a first approach, the interaction with services is limited to simple GET requests as supported e.g. by REST services. Support for other services like SOAP services is planned. A service request is thus assembled using a certain URL and a set of parameters. As an example, the Ebay Shopping web service will be studied. The following URL is invoked to retrieve a list of items corresponding to the search keywords \emph{USB}.

\url{http://open.api.sandbox.ebay.com/shopping?appid=KasselUn-efea-4b93-9505-5dc2ef1ceecd&version=517&callname=FindItems&ItemSort=EndTime&QueryKeywords=USB&responseencoding=XML}

As you may see, the URL invoked is http://open.api.sandbox.ebay.com/shopping and the parameters used in the example are:
\begin{description}
	\item[appid] this is the application ID obtained to use the API.
	\item[version] the API version.
	\item[callname] in this case FindItems to search through all items in Ebay.
	\item[itemsort] sorting method for the list of items.
	\item[querykeywords] list of keywords.
	\item[responseencoding] format of the response message obtained by the invocation of the request.
\end{description}

If needed, a detailed specification of the Ebay Shopping API can be found at \cite{eBayShoppingAPIs}.

The above URL for searching items in Ebay is followed by the query parameters, which take the form \textit{argument=value}. In this example, the only relevant parameter the end user may want to provide is \emph{querykeywords}. Thus the end user may use some input field to provide a query keyword and pass this value to our service wrapper operation. Other parameters can be set to a default value. However, in order to develop a more generic wrapper to a service, more or even all parameters might be passed by the user.

To define the desired input parameters and the type of the expected result of a new service wrapper the FAST service wrapper tool provides a web based form that allows the editing of these entries, cf. Figure \ref{fig:construct_pre_post_conditions}. Note, we allow to edit example values for the input parameters. These example values may be used to test the service wrapper.

%\begin{figure}
%  \begin{center} \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithPortDefinitions.jpg}
%    \caption{Configuring parameters of a service wrapper (photoshoped screen dump)}
%    \label{fig:construct_pre_post_conditions}
%  \end{center}
%\end{figure}

The service wrapper tool composes the request using a template string which will contain placeholders for input values. Before sending the request to the service, the placeholders are replaced with their corresponding values and the resulting URL is then ready to be sent as the service request.

Figure ~\ref{fig:construct_service_request} shows the web based tool to construct the service requests.

%\begin{figure}
%  \begin{center}
%    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithRequestExample.jpg}
%    \caption{Constructing the service request URL and its parameters (photoshoped screen dump)}
%    \label{fig:construct_service_request}
%  \end{center}
%\end{figure}

In the top input field of Figure ~\ref{fig:construct_service_request}, the user may drop an example http request taken e.g. form the service documentation e.g. from \cite{eBayShoppingAPIs}. The tool analyses the example request and in the middle of the screen a form for editing the request parameters is provided. In our example, the user has connected the \textit{QueryKeywords} parameter with the input parameter \textit{search\_key} by adding a corresponding reference to the value field of that parameter. In addition, we have retrieved an access key which has been entered as value for the \textit{appid} parameter.

Below the request parameter editing form of Figure ~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed URL by sending it to the specified service address (via a server relay). Then the placeholders for input parameters are replaced by their example values and the resulting http request is shown below the parameter form. In addition, the request is send and the response is shown on the bottom of that page. This gives the user a fast feedback whether the constructed request works as desired.

\subsubsection{Limitations} % (fold)
\label{ssub:limitations}

It is worth pointing out that currently the wrapper tool is able to construct input ports for the wrappers using just basic types. To construct URL requests from more complex input parameters, as e.g. a person object, we are currently developing access operations that will allow to access the values of fields of complex objects. For example, the expression \emph{customer.address} may refer to the \emph{address} field of a \emph{customer} parameter.

% subsubsection limitations (end)

% subsection constructing_service_requests (end)

\subsection{Interpreting service responses} % (fold)
\label{sub:interpreting_service_responses}

Once the service request is constructed and sent to the service provider, it will send back a response. This response message could be serialized in any format, though the most common formats used nowadays are XML or JSON among others. To continue the example started in the previous section, the response of the Ebay Shopping service will be in XML as specified in the request.

\subsubsection{Translation XML into Facts} % (fold)
\label{ssub:translation_xml_into_facts}

Figure ~\ref{fig:response_service_execution} shows the data transformation tab of the wrapper tool.

%\begin{figure}
%  \begin{center}
%      \includegraphics[angle=90,width=0.8\linewidth]{images/ServiceWrapperToolGVSWithTransformationRules.jpg}
%    \caption{Interactive, rule based transforming of an XML response to domain objects (photoshoped screen %dump)}
%    \label{fig:response_service_execution}
%  \end{center}
%\end{figure}

Once the service response, in XML format, has been retrieved, the transformation tab shows it as an interactive object tree on the left side of Figure ~\ref{fig:response_service_execution}. To construct this interactive object tree, the XML document has been parsed into a DOM, and a simplified tree representation of that DOM is built up.

A transformation rule is used to analyze the XML data and to generate domain-specific objects from concepts from the ontologies used by the parameters of the different building blocks. A transformation rule is composed of three elements, cf. the middle part of Figure ~\ref{fig:response_service_execution}. First, the \textit{from} field indicates the XML elements to be translated by the rule. These XML elements are identified by the tag name of a DOM element from the XML document. Second, the type of the rule will be set, taking one of the following values: \emph{createObject}, \emph{fillAttributes} or \emph{dummy}. And third, the target of the rule specifies a certain concept or attribute, to be created or filled. A detailed explanation of the type of actions to be trigger from the transformation rules is:
\begin{description}
	\item[action \emph{createObject}] specifies the creation of a new domain object. The type of that new object is provided in the third compartment. In the example being explained, the root rule searches for XML elements with tag name \emph{FindItemsResponse} and for each such element a \emph{List} object is created. The resulting objects are shown in a facts tree in the right of Figure ~\ref{fig:response_service_execution}.
	\item[action \emph{fillAttributes}] does not create a new object but it fills the value of the attribute provided as third part of such rules. In our example, the third transformation rule searches for XML elements with tag name \emph{Title}. Note, the rule is a sub-rule of the second rule, which generates \emph{Product} objects. Thus, the sub-rule searches for \emph{Title} tags only in the subtree of the XML data that has been identified by an application of the parent rule before. For example, the \emph{Item} rule may just have been applied to the first \emph{Item} element of the XML data. Then, the \emph{Title} rule is applied only to the first \emph{Item} sub-tree of the XML data and thus it will find only one \emph{Title} element in that sub-tree (not visible in Figure ~\ref{fig:response_service_execution} ). The value of that \emph{Title} element is then transfered to the \emph{productName} attribute of the corresponding \emph{Product} object. Actually, our \textit{from} fields allows also to refer to parts of an XML attribute e.g. to \textit{words} 1 through 3. It is also possible to combine constant text and elements of multiple XML tree elements. 
	\item[action \emph{dummy}] does not create or modify any objects but such rules are just used to narrow the search space for their sub-rules. For example, in Amazon product data, the XML data for an item contains sections for \emph{minimum price}, \emph{maximum price}, and \emph{average price}. Each such section contains the \emph{plain price} and the \emph{formatted price}. Thus, in the Amazon case, a rule that searches for \emph{formatted price} elements within an \emph{Item} element would retrieve three matches. Using a dummy rule, we may first search for \emph{minimum price} elements and then search for \emph{formatted price} elements within that sub-tree.
\end{description}

Our tool follows an interactive paradigm. Therefore, any time a change to a transformation rule is done, the transformation process is triggered and the resulting facts tree is directly shown. This process helps the user to deal with the slightly complex semantics of the transformation rules avoiding errors or mistakes. In addition, out tool is ontology-driven, therefore, the service wrapper designer shall retrieve the domain-specific types from a FAST ontology server together with the structure of each type, i.e. together with a description of the attributes of each object. Thus, the transformation rule editor is able to provide selection boxes for the target element of the rules. For a \emph{createObject} rule, this selection box shows the object types available for that domain. For the \emph{fillAttributes} rules, the selection box shows the attributes of the object type chosen in the parent rule. In addition, we may provide some analysis tool, which will help to guarantee that the objects generated by the transformation rules conform to the object types defined in the corresponding FAST ontology. This helps to ensure that the objects generated by the designed service wrapper will be compatible for input parameters of subsequent filter steps and or gadgets.

% subsubsection translation_xml_into_facts (end)

% subsection interpreting_service_responses (end)

\subsection{Generating a Resource Adapter} % (fold)
\label{sub:generating_a_resource_adapter}

Once the wrapping of a service has been defined and tested in the service wrapper tool, we generate an implementation of the desired Resource Adapter in XML, HTML, and JavaScript, ready to be deployed and executed inside a web gadget.

% subsection generating_a_resource_adapter (end)

\subsection{Limitations} % (fold)
\label{sub:limitations}

The rule driven approach presented above is somewhat limited. It is deliberately restricted to such a simple rule mechanism in order to keep things simple enough for end-users. Still, the selected approach suffices for many practical and real world cases. As a more complex example, the XML data for a person may provide two different tags for the first and the last name of a person. Contrarily, a person fact which conforms to a certain ontology for that domain may provide only one \emph{fullname} attribute that shall be filled by a concatenation of the first and the last name. To achieve this, the \textit{from} field of that tranforamtion rule might look like: \texttt{lastname"', "`firstname}. We are also able to do some navigation in the XML tree to follow XRef elements. For example the attribute \texttt{grandmother} could be filled using \texttt{mother.mother} in the \textit{from} field.

However, there are some transformations that these rules cannot perform. For example, we do not support any mathematical operations. Thus, transforming e.g. Fahrenheit into Celsius temperatures is not supported. To cover such  cases, intermediate object formats can be used which would allow generating objects to be further processed by additional filters. Such additional filters may be realized using (hand coded) operators, since some generic operators can act as filters for aggregation and conversions of objects from multiple sources. Then, service wrappers in combination with these filter operators will allow covering these complex cases.

% subsection limitations (end)

% section restful_web_services_wrapper_tool (end)

### 2010-06-14 trunk/papers/icsoc2010/paper_ssec_linked_data.tex
split it into different files
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@844 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-02-16 trunk/deliverables/D5_1_1/fast_d5_1_1.tex
@@
-\documentclass{fast_latex}
-
-\usepackage{inputenc}
-\usepackage{lastpage}
-\usepackage{setspace}
-\usepackage{graphicx}
-\usepackage[pdfborder={0 0 0}]{hyperref}		% turn on when latex is used (not miktec)
-\usepackage{url} % LEO: urls \url{}
-\usepackage{verbatim} % code and comment
-\usepackage{longtable}
-\usepackage{xspace}   % whitespace after a macro if no punctuation after the macro
-\usepackage{multirow}
-\usepackage{colortbl}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{amssymb}
-
-% This helps to get hyphenation in typewriter font (\texttt)
-\usepackage[htt]{hyphenat}
-
-\parindent0pt
-
-\newcommand\deliverableNumber{D5.1.1}
-\newcommand\deliverableTitle{User Manual of the FAST Catalogue}
-\newcommand\deliverableTitleShort{Catalogue User Manual}
-\newcommand\workpackageNumber{5}
-\newcommand\workpackageTitle{Semantic catalogue of screen-flow resources and back-end Web Services}
-\newcommand\authorOne{Ciprian Palaghita, Cyntelix}
-\newcommand\authorTwo{Ismael Rivera, NUIG}
-\newcommand\authorThree{Author 3}
-\newcommand\authorFour{Author 4}
-\newtheorem{example}{\emph{Example}}
-
-\begin{document}
-% explicit hyphenations
-\hyphenation{RDF-Re-po-si-to-ry}
-\hyphenation{name-space}
-
-%\fontfamily{tahoma}\selectfont
-\def\note#1{\marginpar{\footnotesize#1}} % use this to show the notes in the document
-%\def\note#1{} % use this to hide the notes
-
-\newenvironment{definition}{}{}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% TITLE PAGES 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\thispagestyle{empty}
-
-\pagenumbering{roman}
-
-\begin{flushright}
-	\includegraphics[width=3cm]{images/FP7_logo}
-\end{flushright}
-
-\vspace{1cm}
-
-%\begin{minipage}[p]{15cm}
-	\begin{center}
-		\includegraphics{images/FAST_logo}\\
-		\vspace{1cm}
-		{\LARGE{\sffamily \emph{FAST AND ADVANCED STORYBOARD TOOLS}}}\\
-		\vspace{0.5cm}
-		{\LARGE \sffamily \emph{FP7-ICT-2007-1-216048}}\\
-		\vspace{0.5cm}
-		{\LARGE \sffamily \emph{http://fast.morfeo-project.eu}}\\
-		\vspace{4cm}
-		{\LARGE \sffamily \textbf{Deliverable \deliverableNumber}}\\
-		\vspace{0.5cm}
-		{\LARGE \sffamily \textbf{\deliverableTitle}}\\
-		\vspace{2cm}
-		{\large \sffamily \authorOne}\\
-		\vspace{0.2cm}
-		{\large \sffamily \authorTwo}\\
-		\vspace{0.5cm}
-		\vfill
-		{\large \sffamily Date: 19/02/2010}\\
-		\vspace{1cm}
-		{\sffamily FAST is partially funded by the E.C. (grant code: FP7-ICT-2007-1-216048).}
-
-	\end{center}
-%\end{minipage}
-
-\clearpage
-%%%%%%%%%%%%%%
-% NEXT PAGES %
-%%%%%%%%%%%%%%
-\pagestyle{scrheadings}
-
-\lohead{\includegraphics[width=4cm]{images/FAST_logo_transparent}}
-%\cohead{\small\textcolor{fast@lightgrey}{\deliverableTitle}}
-\rohead{\small{\today}}
-%\lofoot{\small\textcolor{fast@lightgrey}{Task Force Ontologies}}
-\cofoot{\small{FAST 216048 --- \deliverableTitleShort}}
-\rofoot{\small{\thepage}}
-
-\newpage
-
-\clearpage
-
-\section*{Version History}
-
-\begin{small}
-\begin{tabular}{|l|l|l|p{7.5cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{\textbf{Rev. No.}} &
-                         \textcolor{white}{\textbf{Date}} &
-                         \textcolor{white}{\textbf{Author (Partner)}} &
-                         \textcolor{white}{\textbf{Change description}}\\ \hline
-1.0 & 19/02/2010 & Ciprian Palaghita (Cyntelix) & First version of the deliverable \\
- & & Ismael Rivera (NUIG) & \\ \hline
-\end{tabular}
-\end{small}
-
-\color{black}
-
-\vfill
-%{\bf Explanations of abbreviations on front page}\\
-%\\
-%%Nature \\
-%R: Report \\
-%P: Prototype \\
-%R/P: Report and Prototype \\
-%O: Other \\
-% \\
-%Dissemination level \\
-%PU: Public \\
-%PP: Restricted to other FP6 participants \\
-%RE: Restricted to specified group \\
-%CO: Confidential, only for NEPOMUK partners \\
-
-\newpage
-
-%%%%%%%%%%%%%%%%%%%%%
-% Executive Summary %
-%%%%%%%%%%%%%%%%%%%%%
-
-\clearpage
-
-\section*{Executive Summary}
-\doublespacing
-
-The present deliverable is intented to be the developer's manual of the prototype of the semantic catalogue.
-That said, whoever developing a component which will interact with the catalogue will find in this document a description of its architecture, the functionality provided together with the API (Application Programming Interface), data interchange formats, code errors and exceptions as well as several examples of usage.
-
-\newpage
-
-%%%%%%%%%%%%%%%%%%%%%
-% Document Summary %
-%%%%%%%%%%%%%%%%%%%%%
-
-\clearpage
-
-\section*{Document Summary}
-% double spacing from here on:
-\singlespacing
-
-\begin{small}
-\begin{tabular}
-	%{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{6cm}|p{2cm}|p{2cm}|}
-	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{6cm}|p{2cm}|p{2cm}|}
-	\hline
-	\textcolor{white}{\textbf{Code}} & {FP7-ICT-2007-1-216048} & {\textbf{Acronym}} & {FAST}\\ \hline
-	\textcolor{white}{\textbf{Full title}} & \multicolumn{3}{l|}{Fast and Advanced Storyboard Tools}\\ \hline
-	\textcolor{white}{\textbf{URL}} & \multicolumn{3}{l|}{\url{http://fast.morfeo-project.eu}}\\ \hline
-	\textcolor{white}{\textbf{Project officer}} & \multicolumn{3}{l|}{Annalisa Bogliolo}\\ \hline
-\end{tabular}
-\end{small}
-
-\vspace{0.5cm}
-
-\begin{small}
-\begin{tabular}
-	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.25cm}|p{1cm}|p{1cm}|p{6.32cm}|}
-	\hline
-	\textcolor{white}{\textbf{Deliverable}} & {\textbf{Number}} & {\deliverableNumber} & {\textbf{Name}} & {\deliverableTitle}\\ \hline
-	\textcolor{white}{\textbf{Work package}} & {\textbf{Number}} & {\workpackageNumber} & {\textbf{Name}} & {\workpackageTitle}\\ \hline
-\end{tabular}
-\end{small}
-
-\vspace{0.5cm}
-
-\begin{small}
-\begin{tabular}
-	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.2cm}|p{1.6cm}|p{3.37cm}|}
-	\hline
-	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {27/02/2010} & {\textbf{Submitted}} & {27/02/2010}\\ \hline
-	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
-	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
-	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{This deliverable is the technical documentation for the prototype developed as part of the WP5.
-It describes the catalogue's architecture, data interchange formats, APIs and examples of how to interact with it.}\\ \hline
-	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
-	{} & \multicolumn{4}{l|}{\authorTwo}\\ 
-%	{} & \multicolumn{4}{l|}{\authorOne}\\ 
-%	{} & \multicolumn{4}{l|}{\authorOne}\\
-  \hline
-	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{Cyntelix} & \textbf{email} & {cpalaghita@cyntelix.com} \\ \cline{4-5}
-	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 858494258} \\ \hline
-	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, semantic catalogue, gadget catalogue, RDF store}\\ \hline
-\end{tabular}
-\end{small}
-
-\newpage
-
-%%%%%%%%%%%%%%%%%%%%%
-% TABLE OF CONTENTS %
-%%%%%%%%%%%%%%%%%%%%%
-\doublespacing
-\setcounter{tocdepth}{2}
-\tableofcontents
-\cleardoublepage
-\pagenumbering{arabic}
-
-\clearpage
-\listoftables
-
-\clearpage
-\listoffigures
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-% BEGINNING OF SECTIONS %
-%%%%%%%%%%%%%%%%%%%%%%%%%
-\rofoot{\small{Page \thepage\ of \pageref{LastPage}}}
-
-\clearpage
-\section{Introduction} % (fold)
-\label{sec:introduction}
-This section starts establishing the goal and scope of the present document, shows how it is structured and details the relation to others documents and work packages.
-
-\subsection{Goal and Scope} % (fold)
-\label{sub:goal_and_scope}
-
-This is an introductory manual for developers who want to adopt and use the FAST semantic catalogue.
-It explains the main functionalities implemented so far, an overview of its architecture and a detailed \emph{Application Programming Interface} or API of the complete set of the operations offered through a REST service.
-
-% subsection goal_and_scope (end)
-
-\subsection{Structure of the document} % (fold)
-\label{sub:structure_of_the_document}
-
-The deliverable presents both the external and internal architecture in Section~\ref{sec:architecture_overview}, then in Section~\ref{sec:restful_catalogue_api} it is detailed the Catalogue API, query formats, interchange formats, error codes and so on.
-
-% subsection goal_and_scope (end)
-
-\subsection{Changes from previous version} % (fold)
-\label{sub:changes_from_previous_version}
-
-This is the first version of this deliverable since in M12 the deliverable 5.1 was not scheduled.
-Nevertheless, last year an extra document was written as an appendix to the FAST Catalogue in order to detail and clarify technical decisions and to have a technical reference manual, which the content was taken from as the initial draft the present deliverable.
-Hence, the deliverable D5.2.1 will evolve to become the D5.1.1, so it will be no longer needed by itself.
-
-There are several new sections and major changes which need to be mentioned:
-
-\begin{itemize}
-	\item Section~\ref{sec:installation_guide}.
-Installation Guide describes a set of steps to be followed in order to install and execute the FAST Catalogue.
-Also some configuration instructions are specified.
-	\item Section~\ref{sub:linked_data}. Content Negotiation states how the FAST Catalogue has adopted several of the content negotiation principles.
-	\item Section~\ref{sec:restful_catalogue_api}.
-RESTful Catalogue API has been intensively extended and modified in order to  reflect all the new functionality and how every request or response is constructed.
-Some of the most important features added in this iteration are:
-	\begin{itemize}
-		\item support for form elements, operators and back-end services (see Section ...),
-		\item convenience operations for recommendation and checking satisfaction and reachability for form elements, operators and back-end services (see Section ...),
-		\item and, last but not least, support for Screen planning (see Section~\ref{sub:planning}).
-	\end{itemize}
-	\item Several appendixes have been added with the JSON structure of every building block the Catalogue can deal with.
-	\item Appendix A. Last year an evaluation to choose the best technology for the Catalogue was done. It was not included in any formal deliverable, so this year it has been documented as an appendix.
-\end{itemize}
-
-% subsection changes_from_previous_version (end)
-
-% section introduction (end)
-
-\clearpage
-\section{Installation Guide} % (fold)
-\label{sec:installation_guide}
-
-This section  provides instructions on how to manually install and configure the FAST Catalogue.
-The first part of this guide presents some requirements to be considered before the installation, then it gives broad general instructions, and the last part contains a more detailed installation notes for specific configurations.
-
-\subsection{System Requirements} % (fold)
-\label{sub:system_requirements}
-
-In addition to the software itself, a standard FAST Catalogue installation has the following requirements:
-\begin{itemize}
-  \item Java\footnote{http://java.sun.com/} (version 6 or later) is required to run the software.
-  \item A Java Servlet Container that supports Java Servlet API 2.4 and Java Server Pages (JSP) 2.0, or newer.
-We recommend using a recent, stable version of Apache Tomcat\footnote{http://tomcat.apache.org/}.
-At the time of writing, this is either version 5.5.x or 6.x.
-\end{itemize}
-
-In addition, there are various optional dependencies which are required if you want to use certain advanced features (see Section~\ref{sub:installation_instructions}).
-
-% subsection system_requirements (end)
-
-\subsection{Obtaining the FAST catalogue} % (fold)
-\label{sub:obtaining_the_fast_catalogue}
-
-It is recommended to have a Subversion client installed before you download the code (although you can theoretically download files without Subversion, this would mean tediously downloading each individual file manually).
-The recommended software is the official Subversion client, available from the Subversion project page\footnote{http://subversion.tigris.org/}.
-Note that this client uses a command-line interface, which the instructions below use.
-Alternatively, you can get subversioning software with a graphical user interface such as TortoiseSVN.
-
-To download from the latest release (recommended), enter the following command from the command-line in the directory you wish to download to:
-\begin{verbatim}
-svn checkout https://svn.forge.morfeo-project.org/fast-fp7project/trunk/catalogue_service
-\end{verbatim}
-
-% subsection obtaining_the_fast_catalogue (end)
-
-\subsection{Installation Instructions} % (fold)
-\label{sub:installation_instructions}
-
-Once you have got the source code, first step should be addressed is its compilation.
-You may manually compile the source yourself, however the FAST Catalogue comes along with a script to facilitate this task.
-This script has been made using a Java build tool called Ant \footnote{http://ant.apache.org/}.
-Hence, we recommended to get and install such a tool in order to follow the instructions below.
-
-Now you can compile, package and run the application via:
-\begin{verbatim}
-ant clean
-ant prepare
-ant compile
-\end{verbatim}
-
-Or using a task which gather the previous three tasks and creates WAR file ready to distribute or deploy:
-\begin{verbatim}
-ant dist
-\end{verbatim}
-
-After executing with no errors this task, a WAR file should have been created in the /dist directory.
-A WAR file, which stands for  Web Application Archive, is just a JAR file used to distribute a collection of JavaServer Pages, servlets, Java classes, XML files, tag libraries and static Web pages (HTML and related files) that together constitute a Web application, in this case, the FAST Catalogue.
-
-Note that it is also possible to get the compiled WAR file directly from the SVN server. This file can be located at \url{https://svn.forge.morfeo-project.org/fast-fp7project/distribution/catalogue}.
-
-Then, last step is to deply the WAR file into the servlet container. If the server chosen was Apache Tomcat, the procedure for deploying a Web application is:
-\begin{enumerate}
-	\item Stop Tomcat.
-	\item Delete existing deployment.
-If you have previously deployed "foo.war" in TOMCAT\_HOME/webapps, then it has been unpacked into webapps/foo/...
-You must delete this directory and all its contents.
-On Unix, this can be done with
-rm -r \$TOMCAT\_HOME/webapps/foo
-	\item Copy WAR file to TOMCAT\_HOME/webapps/.
-	\item Start Tomcat.
-\end{enumerate}
-
-Now you have successfully built the FAST catalogue, however we highly recommend you to read Section~\ref{sub:configuration} and tune your FAST Catalogue instance.
-
-% subsection installation_instructions (end)
-
-\subsection{Configuration} % (fold)
-\label{sub:configuration}
-
-The following section covers the configuration options you may have to set up before start using the Catalogue.
-The FAST Catalogue relies on a Sesame repository for the persistent storage.
-Sesame repositories can be local or remote.
-Remote does not mean, it has to be installed in a different machine, but in a servlet container such as Apache Tomcat.
-The advantage of remote repositories is the possibility of accessing it via a SPARQL endpoint.
-These configuration aspect can be found in the file \emph{repository.properties} in the root directory where the Catalogue has been deployed.
-
-Independently the type of repository, the parameter \emph{serverURL} needs to take the value of the URL where the FAST Catalogue instance has been deployed.
-If you are testing the application in your machine, it may be something like \url{http://catalogueURL/FASTCatalogue}.
-
-Local repositories just need the parameter \emph{storageDir} pointing at a local directory, which has write permission in order to allow Sesame to create files in that directory or to delete any file in the directory.
-For instance:
-
-\begin{verbatim}
-storageDir=D:\\FAST\\catalogue\\repository
-\end{verbatim}
-
-Nevertheless, setting up a remote repository requires a few more steps.
-First the OpenRDF Sesame Server needs to be installed.
-The installation guide of this server is detailed in the Chapter 5 and 6 of the User Guide for Sesame 2.2 \cite{sesame2.2}.
-Once the Sesame server is deployed, and a repository has been created, the parameters sesameServer and repositoryID will take the values of the Sesame server URL and the repository identifier.
-For instance:
-
-\begin{verbatim}
-sesameServer=http://localhost:8880/openrdf-sesame
-repositoryID=c2
-\end{verbatim}
-
-% subsection installation_instructions (end)
-
-\subsection{Hello World!} % (fold)
-\label{sub:hello_world}
-
-It is recommended to check if the FAST Catalogue has been successfully installed.
-To do that, a simple HTTP request, using a Web browser, can be sent to the URL the Catalogue is deployed, and specifying a type of resource such as Screen.
-The URL would be similar to \url{http://catalogueURL/FASTCatalogue/screens}.
-Depending on the variant chosen, the response may be different.
-For JSON, the response should be \verb|[]|, but in HTML the message \verb|No screens found.| should be shown.
-
-% subsection hello_world (end)
-
-% section installation_guide (end)
-
-\clearpage
-\section{Architecture Overview} % (fold)
-\label{sec:architecture_overview}
-
-This section presents a technical overview of the internal catalogue's architecture and which external components will interact with it.
-Basically three layers can be distinguished: Presentation, Business Logic and Persistence layer (see Figure~\ref{fig:catalogue_architecture}).
-
-\begin{figure}[htb]
-\label{fig:catalogue_architecture}
-\begin{center}
-	\includegraphics[width=12cm]{images/catalogue_architecture}
-	\caption{Catalogue's Architecture}
-\end{center}
-\end{figure}
-
-The Presentation layer will be the public interface of the catalogue.
-The main purpose of this layer is to provide its functionality to other components within the Gadget Visual Studio.
-The service is offered in a REST (REpresentational State Transfer) style, making it easy to consume than other complex APIs.
-REST is an architectural style, not a toolkit or a standard.
-Even though, it makes use of standards like HTTP, URI, Resource Representations (XML, HTML, JSON, JPEG, etc.) or MIME types (text/xml, text/html, application/json, image/jpeg, etc.).
-Another characteristic of the REST style is its stateless assumption.
-The catalogue adopts this strategy; therefore every request needs a complete set of information in order to prepare the response.
-
-As part of the Presentation layer, a SPARQL endpoint is offered using the SPARQL protocol service as defined in the SPROT \cite{sprot} specification.
-The SPARQL endpoint is mostly offered to enable other developers to query directly the catalogue knowledge base using SPARQL queries.
-This feature is supported by the Sesame RESTful HTTP interface for SPARQL Protocol for RDF.
-
-The Business Logic layer contains all the FAST domain-specific processing.
-It provides functions to interact with all the elements of the domain model specified in \cite{moeller2009fast_ontology}.
-It processes the input from the Presentation layer, creating specific objects modelling the business logic, and vice versa.
-In addition it interacts with the Persistence layer in order to persist the model.
-
-The Persistence layer provides an API which allows the Business layer to work with a standard set of objects that read and save their state to the triple store.
-For this reason, on top of this layer it is used RDF2Go, an abstraction over triple (and quad) stores.
-The RDF2Go API allows interacting with the semantic representation of the model (triples) in a generic manner, and also brings the flexibility of choosing different triple stores to persist them.
-The selected implementation for the triple stored is Sesame 2.
-Moreover, this framework is completely extensible and configurable regarding to storage mechanisms, inferencers, RDF file formats, query result formats and query languages.
-For the actual catalogue prototype, the storage mechanism used is the native file storage system of Sesame, the inferencer used is a subset of the RDFS entailment rules \cite{rdfsrules} following a forward-chaining policy.
-
-% section architecture_overview (end)
-
-\clearpage
-\section{RESTful Catalogue API} % (fold)
-\label{sec:restful_catalogue_api}
-
-This section provides a high-level overview of the Catalogue API.
-It describes the calls or operations supported, specific parameters and responses for each operation, supported interchange formats and some examples to facilitate the understanding of the API.
-
-\subsection{JSON Interchange Format} % (fold)
-\label{sec:json_interchange_format}
-
-The format supported, and in which they must be constructed, by every HTTP request is JSON\footnote{http://www.json.org/}.
-JSON is a lightweight data interchange format whose simplicity has resulted in widespread use among web developers, easy to read and write and able to using any programming language because its structures map directly to data structures used in most programming languages.
-
-Every HTTP request should be encoded using the MIME type \texttt{application/json} and the charset \texttt{UTF-8}.
-
-The following is an example of a Find \& Check request:
-
-\singlespacing
-\begin{verbatim}
-{
-  "canvas": [
-    { "uri": "http://catalogueURL/catalogue/screens/238" },
-    { "uri": "http://catalogueURL/catalogue/screens/323" }
-  ],
-  "elements": [
-    { "uri": "http://catalogueURL/catalogue/screens/636" }
-  ],
-  "domainContext": {
-    "tags": [
-      {
-        "label": { "en-GB": "Amazon" },
-        "means": "http://dbpedia.org/page/Amazon.com"
-      }
-    ],
-    "user": "irivera",
-  },
-  "criterion": "reachability"
-}
-\end{verbatim}
-\doublespacing
-
-\subsubsection{Internationalisation I18n} % (fold)
-
-In order to offers an adaptable solution to various languages and regions without major engineering changes, internationalisation is considered from an early stage in the catalogue development.
-Underlying representation technologies used to develop the catalogue (RDF/XML and JSON) allow to implement this feature.
-This feature is implemented by the addition of a language tag to every 'string' desired.
-The specification of this language tag is composed by the language code and the country code, following the ISO 639\footnote{http://ftp.ics.uci.edu/pub/ietf/http/related/iso639.txt} for languages and the ISO 3166\footnote{http://userpage.chemie.fu-berlin.de/diverse/doc/ISO\_3166.html} for countries.
-The following example illustrates how to use it properly in both formats.
-
-\singlespacing
-\begin{verbatim}
-{
-  "label": {
-    "en-GB": "Product Search",
-    "es-ES": "Ejemplo Sencillo",
-    "de-DE": "Produktsuche"
-  },
-  "description": {
-    "en-GB": "On this screen, the user can enter a search term.",
-    "es-ES": "Esto es un ejemplo sencillo de pantalla",
-    "de-DE": "Auf diesem Screen kann der Benutzer einen Suchbegriff eingeben."
-  },
-}
-\end{verbatim}
-\doublespacing
-
-\singlespacing
-\begin{verbatim}
-<rdf:Description rdf:about="http://www.morfeoproject.eu/fast/fco#ExampleScreen">
-  <rdf:type rdf:resource="http://www.morfeoproject.eu/fast/fco#Screen"/>
-  <rdfs:label xml:lang="en-GB">Simple Example</rdfs:label>
-  <rdfs:label xml:lang="es-ES">Ejemplo Sencillo</rdfs:label>
-  <rdfs:label xml:lang="de-DE">Einfaches Beispiel</rdfs:label>
-  <description
-    xmlns="http://purl.org/dc/terms/"
-    xml:lang="en-GB">This is a simple example of a screen</description>
-  <description
-    xmlns="http://purl.org/dc/terms/"
-    xml:lang="es-ES">Esto es un ejemplo sencillo de pantalla</description>
-  <description
-    xmlns="http://purl.org/dc/terms/"
-    xml:lang="de-DE">Dies ist ein einfaches Beispiel für einen Bildschirm</description>
-</rdf:Description>
-\end{verbatim}
-\doublespacing
-
-Until now, using internationalisation is allowed for the label and the description of a screen, but in the near future, more attributes will benefit from this feature.
-
-% subsection json_interchange_format (end)
-
-\subsection{Linked Data} % (fold)
+\subsection{Serving Linked Data} % (fold)
 \label{sub:linked_data}
 
-Apart from serving as the back-end system for the Gadget Visual Storyboard (GVS), the catalogue publishes all FAST building blocks as linked data, following the principles as originally defined in \cite{bernersLee2006linkedData}, in order to make them available to arbitrary third party applications.
-Each building block is identified by an HTTP URI and hosted through the catalogue so that it can be dereferenced through the same URI.
+The recently success of the web of Data has influenced the way information is now published on the web, and has been widely adopted by the academic community, large companies such as The New York Times and BBC, and national governments such as US and UK made public commitments toward open data.
+The main aim of the linking data is to find other related data based upon previously known data, in terms of the connections or links between them.
+We apply this concept in order to provide metadata about the web services as linked data, following the principles as originally defined in \cite{bizer_ijswis2009}, in order to make them available to arbitrary third-party applications.
+Each web service is identified by an HTTP URI and hosted in the catalogue so that it can be dereferenced through the same URI.
 For each building block, data is available in representations in different standard formats such as JSON (for communication with the GVS), RDF/XML, Turtle, or even HTML+RDFa as a human-readable version.
-Following the principles and best practices proposed in \cite{berrueta2008} and \cite{sauermann2008cool_uris}, these representations are served based on the request issued by the requesting agent, using a technique called \emph{content negotiation}.
-As required by the forth rule of linked data, individual building blocks link to other data on the Web, thereby preventing so-called isolated ``data islands''.
+The principles and best practices proposed in \cite{berrueta2008} and \cite{sauermann2008cool_uris} are also taken into account, these representations are served based on the request issued by the requesting agent, using a technique called \emph{content negotiation}.
+As required by the forth rule of linked data, individual building blocks link to other data on the web, thereby preventing so-called isolated ``data islands''.
 
@@
 The catalogue will then inspect the request and redirect the requester to the appropriate representation (e.g., \texttt{http://catalogue.com/screens/235.json}), as illustrated in Fig.~\ref{fig:content_negotiation}.
-On the Web, different representations for the same resource are called \emph{variants}, and content negotiation is the mechanism used to determine which of the representations of most appropriate for a given request.
+On the web, different representations for the same resource are called \emph{variants}, and content negotiation is the mechanism used to determine which of the representations of most appropriate for a given request.
 
-\begin{figure}[htb]
-\label{fig:content_negotiation}
-\begin{center}
-	\includegraphics[width=12cm]{images/content_negotiation}
-	\caption{Content Negotiation}
-\end{center}
-\end{figure}
+%\begin{figure}[htb]
+%\label{fig:content_negotiation}
+%\begin{center}
+%	\includegraphics[width=12cm]{images/content_negotiation}
+%	\caption{Content Negotiation}
+%\end{center}
+%\end{figure}
 
@@
 If none is specified, by default, the server will choose the JSON representation.
-
-% subsection content_negotiation (end)
-
-\subsection{API Calls} % (fold)
-\label{sub:api_calls}
-
-This section contains detailed descriptions of the interface provided by the catalogue via REST services, detailing request parameters, response elements, any special errors and examples of requests and responses.
-The URL format is also specified for each operation, where 'catalogueURL' has to be replaced for the real URL the service is installed (e.g.
-http://demo.fast.morfeo-project.org/catalogue).
-
-\subsubsection{CRUD operations} % (fold)
-
-Any building block within the Catalogue is a resource in terms of REST philosophy, hence it can be created, retrieved, modified or deleted, using a certaing URL and a specific HTTP method for every operation.
-Two concepts have to be defined: \emph{collection}, as a set of resources which access URI is http://catalogueURL/<type>/ where catalogueURL is the specific URL where the Catalogue server is installed and <type> is the plural of the name of the building block (e.g.
-screens, services), and \emph{member} of the collection, in other words, the building block itself, which access URI is http://catalogueURL/<type>/<id> where the <id> has to be replaced by the identifier of a specific building block.
-The details of the operations and which HTTP verb has to be used can be found in the Table~\ref{tab:crud_operations}.
-
-\begin{table}[htb]
-\caption{CRUD operations}
-\label{tab:crud_operations}
-\begin{center}
-\begin{tabular}{|p{2.5cm}|p{2.5cm}|p{3.5cm}|p{5cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Resource} &
-                         \textcolor{white}{HTTP method} &
-                         \textcolor{white}{HTTP body} &
-                         \textcolor{white}{Description}\\ \hline
-Collection URI & GET & N/A & \textbf{List} the members of the collection.\\ \hline
-Collection URI & PUT & N/A & Not used.\\ \hline
-Collection URI & POST & JSON representation of the building block & \textbf{Create} a new entry in the collection where the URI is assigned automatically by the collection.
-The URI created is returned by this operation.\\ \hline
-Collection URI & DELETE & N/A & Not used.\\ \hline
-Member URI & GET & N/A & \textbf{Retrieve} the addressed member of the collection.\\ \hline
-Member URI & PUT & JSON representation of the building block & \textbf{Update} the addressed member of the collection or create it with a defined URI.\\ \hline
-Member URI & POST & N/A & Not used. \\ \hline
-Member URI & DELETE & N/A & \textbf{Delete} the addressed member of the collection.\\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-The JSON structure for create any building bock can be found in the appendixes.
-All of them are composed of a common structure which can be found in Section Appendix B, and some specific information depending on the type.
-Screen-flows are defined in Appendix C, screens in Appendix D and forms, operators and back-end services share the same structure detailed in Appendix E.
-However, a few examples of request and responses are shown in order to clarify how the request are constructed and sent, and how the responses look like.
-
-To create a new screen, a POST request is sent to the catalogue server, concretely to \texttt{http://cata\-logueURL/screens}, including the JSON representation of the screen in the body of the request:
-
-\singlespacing
-\begin{verbatim}
-{
-  "code": "http://demo.fast.morfeo-project.org/code/amazonList.js",
-  "creationDate": "2010-01-26T17:01:13+0000",
-  "creator": "irivera",
-  "description": {"en-gb": "Please fill the description..."},
-  "homepage": "http://fast.morfeo-project.eu/",
-  "icon": "http://demo.fast.morfeo-project.org/images/amazonList.png",
-  "id": "28",
-  "label": {"en-gb": "Amazon Product List"},
-  "name": "ProductList",
-  "postconditions": [
-    [
-      {
-        "id": "item",
-        "label": { "en-gb": "An item" },
-        "pattern": "?I 
-                    http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                    http://aws.amazon.com/AWSECommerceService#Item",
-        "positive": "true"
-      }
-    ]
-  ],
-  "preconditions": [
-    [
-      {
-        "id": "filter",
-        "label": { "en-gb": "A search criteria" },
-        "pattern": "?F
-                    http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                    http://aws.amazon.com/AWSECommerceService#SearchCriteria",
-        "positive": "true"
-      }
-    ]
-  ],
-  "rights": "http://creativecommons.org/",
-  "screenshot": "http://demo.fast.morfeo-project.org/images/amazonProductList.png",
-  "tags": [{"label": {"en-gb": "amazon"}}],
-  "version": "0.1"
-}
-\end{verbatim}
-\doublespacing
-
-The URI of the screen is created using the id specified in the JSON request. So, the response for this operation is:
-
-\singlespacing
-\begin{verbatim}
-{
-  "code": "http://demo.fast.morfeo-project.org/code/amazonList.js",
-  "creationDate": "2010-01-26T17:01:13+0000",
-  "creator": "irivera",
-  "description": {"en-gb": "Please fill the description..."},
-  "homepage": "http://fast.morfeo-project.eu/",
-  "icon": "http://demo.fast.morfeo-project.org/images/amazonList.png",
-  "id": "28",
-  "label": {"en-gb": "Prueba1"},
-  "name": "ProductList",
-  "postconditions": [
-    [
-      {
-        "id": "item",
-        "label": { "en-gb": "An item" },
-        "pattern": "?I
-                    http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                    http://aws.amazon.com/AWSECommerceService#Item",
-        "positive": "true"
-      }
-    ]
-  ],
-  "preconditions": [
-    [
-      {
-        "id": "filter",
-        "label": { "en-gb": "A search criteria" },
-        "pattern": "?F
-                    http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                    http://aws.amazon.com/AWSECommerceService#SearchCriteria",
-        "positive": "true"
-      }
-    ]
-  ],
-  "rights": "http://creativecommons.org/",
-  "screenshot": "http://demo.fast.morfeo-project.org/images/amazonProductList.png",
-  "tags": [{"label": {"en-gb": "amazon"}}],
-  "type": "screen",
-  "uri": "http://catalogueURL/catalogue/screens/28",
-  "version": "0.1"
-}
-\end{verbatim}
-\doublespacing
-
-To obtain all the screens stored in the catalogue a HTTP GET request is sent to the Collection URI and the response may be something like this:
-
-\singlespacing
-\begin{verbatim}
-[
-  {
-    "code": "http://demo.fast.morfeo-project.org/code/amazonList.js",
-    "creationDate": "2010-01-26T17:01:13+0000",
-    "creator": "irivera",
-    "description": {"en-gb": "Please fill the description..."},
-    "homepage": "http://fast.morfeo-project.eu/",
-    "icon": "http://demo.fast.morfeo-project.org/images/amazonList.png",
-    "id": "28",
-    "label": {"en-gb": "Prueba1"},
-    "name": "ProductList",
-    "postconditions": [
-      [
-        {
-          "id": "item",
-          "label": { "en-gb": "An item" },
-          "pattern": "?I
-                      http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                      http://aws.amazon.com/AWSECommerceService#Item",
-          "positive": "true"
-        }
-      ]
-    ],
-    "preconditions": [
-      [
-        {
-          "id": "filter",
-          "label": { "en-gb": "A search criteria" },
-          "pattern": "?F
-                      http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                      http://aws.amazon.com/AWSECommerceService#SearchCriteria",
-          "positive": "true"
-        }
-      ]
-    ],
-    "rights": "http://creativecommons.org/",
-    "screenshot": "http://demo.fast.morfeo-project.org/images/amazonProductList.png",
-    "tags": [{"label": {"en-gb": "amazon"}}],
-    "type": "screen",
-    "uri": "http://catalogueURL/catalogue/screens/28",
-    "version": "0.1"
-  },
-  {
-    "creationDate": "2009-02-07T09:59:52+0000",
-    "creator": "fabio",
-    ...,
-	  "uri": "http://catalogueURL/catalogue/screens/14",
-    "version": "1.0"
-  },
-  {
-    "creationDate": "2009-02-07T09:59:52+0000",
-    "creator": "javier",
-    ...,
-	  "uri": "http://catalogueURL/catalogue/screens/564",
-    "version": "1.0"
-  }
-]
-\end{verbatim}
-\doublespacing
-
-\subsubsection{Screen Find} % (fold)
-\label{ssub:screen_find}
-
-The find operation searches inside the catalogue for any screen which could be somehow related to the gadget the user is creating.
-It provides a recommended set of screens depending on the domain context, the canvas, and so on.
-
-The specific URL to access this operation is http://catalogueURL/screens/find using HTTP POST method.
-From now on, the method 'find' only considers the domain context and the pre/postconditions of the screens.
-It will try to satisfy all the unsatisfied preconditions of a given list of screens also known as canvas.
-The results will be all the screens stored in the catalogue which fulfil some of these preconditions, and are tagged with the tags specified in the domain context.
-The request parameters are shown in Table~\ref{tab:screen_find_request} and the response parameters are shown in Table~\ref{tab:screen_find_response}.
-
-\begin{table}[htb]
-\caption{Screen Find Request Parameters}
-\label{tab:screen_find_request}
-\begin{center}
-\begin{tabular}{|p{2.5cm}|p{9cm}|p{2cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
-                         \textcolor{white}{Description} &
-                         \textcolor{white}{Type}\\ \hline
-canvas & The canvas is composed by a list of screens. Only the URI is needed. & Optional \\ \hline
-domainContext & The domain context contains a list of tags and a user. & Optional \\ \hline
-elements & It is a list of resources, previously stored in the catalogue. For example, the list of screens recommended last execution. Only accepts screens. & Optional \\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{table}[htb]
-\caption{Screen Find Response Parameters}
-\label{tab:screen_find_response}
-\begin{center}
-\begin{tabular}{|p{2.5cm}|p{9cm}|p{2cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
-                         \textcolor{white}{Description} &
-                         \textcolor{white}{Type}\\ \hline
-N/A & A list of recommended screens URIs is returned. & Required \\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-Following is shown an example of usage of this operation considering a canvas with the screen \texttt{http://catalogueURL/catalogue/screens/654}.
-
-\singlespacing
-\begin{verbatim}
-{
-  canvas: [
-    { uri: "http://catalogueURL/catalogue/screens/654" }
-  ],
-  domainContext: {
-    tags: [],
-	  user: null
-  },
-  elements: []
-}
-\end{verbatim}
-\doublespacing
-
-After the execution of the recommendation algorithm, the response given by the catalogue is:
-
-\singlespacing
-\begin{verbatim}
-[
-  "http://catalogueURL/catalogue/screens/371",
-  "http://catalogueURL/catalogue/screens/24",
-  "http://catalogueURL/catalogue/screens/253",
-  "http://catalogueURL/catalogue/screens/12"
-]
-\end{verbatim}
-\doublespacing
-
-\subsubsection{Screen Check} % (fold)
-\label{ssub:screen_check}
-
-This operation verifies the state of certain list of screens depending on a specific criterion.
-The only criterion accepted in this stage of the prototype is 'reachability'.
-A screen will be reachable if it all its preconditions are satisfied by postconditions of the screens the canvas contain.
-The specific URL to access this operation is http://catalogueURL/screens/check using HTTP POST method.
-
-\begin{table}[htb]
-\caption{Screen Check Request Parameters}
-\label{tab:screen_check_request}
-\begin{center}
-\begin{tabular}{|p{2.5cm}|p{9cm}|p{2cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
-                         \textcolor{white}{Description} &
-                         \textcolor{white}{Type}\\ \hline
-canvas & The canvas is composed by a list of screens. Only the URI is needed. & Optional\\ \hline
-domainContext & The domain context contains a list of tags and a user. & Optional\\ \hline
-elements & It is a list of resources, previously stored in the catalogue. For example, the list of screens recommended last execution. Only accepts screens. & Optional\\ \hline
-criterion & The criterion specifies what it has to be check. The only possible supported now is 'reachability' in order to check if the preconditions of a screen are satisfied. & Required\\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{table}[htb]
-\caption{Screen Check Response Parameters}
-\label{tab:screen_check_response}
-\begin{center}
-\begin{tabular}{|p{2.5cm}|p{9cm}|p{2cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
-                         \textcolor{white}{Description} &
-                         \textcolor{white}{Type}\\ \hline
-canvas & The canvas indicating which screens satisfy the critetion.
-More aditional information may be included, for example, a satisfaction attribute for every precondition for reachability.
-& Optional\\ \hline
-domainContext & The domain context contains a list of tags and a user. & Optional\\ \hline
-elements & The elements indicating which satisfy the critetion. More aditional information may be included as in the Canvas. & Optional\\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-The scenario for the following example is composed by a screen in the canvas which precondition is a foaf:Person which foaf:workplaceHomepage is \url{http://www.deri.ie/}, and a screen in the elements list which precondition is a foaf:Person.
-The state to check in this case is the reachability and satisfaction of the screens and preconditions.
-
-\singlespacing
-\begin{verbatim}
-{
-  "canvas": [
-    { "uri": "http://catalogueURL/catalogue/screens/238" }
-  ],
-  "elements": [
-    { "uri": "http://catalogueURL/catalogue/screens/636" }
-  ],
-  "domainContext": {
-    "tags": [
-      {
-        "label": { "en-GB": "Amazon" },
-        "means": "http://dbpedia.org/page/Amazon.com"
-      }
-    ],
-    "user": "irivera",
-  },
-  criterion: "reachability"
-}
-\end{verbatim}
-\doublespacing
-
-The response shows that there are not reachable screens at the moment, and preconditions are not satisfied.
-
-\singlespacing
-\begin{verbatim}
-{
-  "canvas": [{
-    "preconditions": [{
-      "label": {"en-gb": "A person working in DERI"},
-      "pattern": "?person rdf:type foaf:Person . 
-                  ?person foaf:workplaceHomepage http://www.deri.ie/",
-      "positive": true,
-      "satisfied": false
-    }],
-    "reachability": false,
-    "uri": "http://catalogueURL/catalogue/screens/238"
-  }],
-  "elements": [{
-    "preconditions": [{
-      "label": {"en-gb": "A person"},
-      "pattern": "?person rdf:type foaf:Person",
-      "satisfied": false
-    }],
-    "reachability": false,
-    "uri": "http://catalogueURL/catalogue/screens/636"
-  }]
-}
-\end{verbatim}
-\doublespacing
-
-\subsubsection{Screen Find \& Check} % (fold)
-\label{ssub:screen_findcheck}
-
-For convenience and to minimise the number of request, the Find operation may be combined with the Check operation.
-This operation will look for new screens based on the given information, and will perform the Check to both the screens in the canvas and the screens in the \emph{elements} attribute.
-
-The format of the request is the same as the one used in the Check operation:
-
-\singlespacing
-\begin{verbatim}
-{
-  "canvas": [
-    { "uri": "http://catalogueURL/catalogue/screens/238" }
-  ],
-  "elements": [
-    { "uri": "http://catalogueURL/catalogue/screens/636" }
-  ],
-  "domainContext": {
-    "tags": [
-      {
-        "label": { "en-GB": "Amazon" },
-        "means": "http://dbpedia.org/page/Amazon.com"
-      }
-    ],
-    "user": "irivera",
-  },
-  criterion: "reachability"
-}
-\end{verbatim}
-\doublespacing
-
-The response for the above request is:
-
-\singlespacing
-\begin{verbatim}
-{
-  "canvas": [{
-    "preconditions": [{
-      "label": {"en-gb": "A person working in DERI"},
-      "pattern": "?person rdf:type foaf:Person . 
-                  ?person foaf:workplaceHomepage http://www.deri.ie/",
-      "positive": true,
-      "satisfied": false
-    }],
-    "reachability": false,
-    "uri": "http://catalogueURL/catalogue/screens/238"
-  }],
-  "elements": [
-    {
-      "preconditions": [{
-        "label": {"en-gb": "A person"},
-        "pattern": "?person rdf:type foaf:Person",
-        "satisfied": false
-      }],
-      "reachability": false,
-      "uri": "http://catalogueURL/catalogue/screens/636"
-    },
-    {
-      "preconditions": [],
-      "reachability": true,
-      "uri": "http://catalogueURL/catalogue/screens/132"
-    }
-  ]
-}
-\end{verbatim}
-\doublespacing
-
-\subsubsection{GetMetadata} % (fold)
-
-Although the information about a resource can be obtained by its specific retrieval operation supported by the CRUD interface, sometimes is needed to retrieve information which would imply a request per resource, hence this operation allow to get the metadata of a list of resources in only one request.
-This prototype supports this operation for screen-flows and screens.
-The specific URL to access this operation is http://catalogueURL/getmetadata using HTTP POST method.
-Table~\ref{tab:getmetadata_request} shows the parameters needed for the invocation and Table~\ref{tab:getmetadata_response} details the different parameters may contain a response.
-
-\begin{table}[htb]
-\caption{GetMetadata Request Parameters}
-\label{tab:getmetadata_request}
-\begin{center}
-\begin{tabular}{|p{2.5cm}|p{9cm}|p{2cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
-                         \textcolor{white}{Description} &
-                         \textcolor{white}{Type}\\ \hline
-N/A & A list of URIs. & Required\\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{table}[htb]
-\caption{GetMetadata Response Parameters}
-\label{tab:getmetadata_response}
-\begin{center}
-\begin{tabular}{|p{2.5cm}|p{9cm}|p{2cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
-                         \textcolor{white}{Description} &
-                         \textcolor{white}{Type}\\ \hline
-screen-flows & A set of screen-flows with all the metadata associated to them.. & Optional\\ \hline
-screens & A set of screens with all the metadata associated to them. & Optional\\ \hline
-forms & A set of forms with all the metadata associated to them. & Optional\\ \hline
-operators & A set of operators with all the metadata associated to them. & Optional\\ \hline
-backendservices & A set of back-end services with all the metadata associated to them. & Optional\\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-The following example send two URIs of screens and a URI of a screen-flow to the GetMetadata operation:
-
-\singlespacing
-\begin{verbatim}
-[
-  "http://catalogueURL/catalogue/screens/12",
-  "http://catalogueURL/catalogue/screens/48",
-  "http://catalogueURL/catalogue/forms/9",
-]
-\end{verbatim}
-\doublespacing
-
-The response obtained are two lists, one containing all the metadata regarding to the screen-flows and another one with the information about the screens:
-
-\singlespacing
-\begin{verbatim}
-{
-  "screenflows": [],
-  "screens": [
-    {
-      "creationDate": "2009-02-07T09:59:52+0000",
-      "creator": "irivera",
-      ...,
-      "uri": "http://catalogueURL/catalogue/screens/12",
-      "version": "1.0"
-    },
-    {
-      "creationDate": "2009-02-07T09:59:52+0000",
-      "creator": "irivera",
-      ...,
-      "uri": "http://catalogueURL/catalogue/screens/12",
-      "version": "1.0"
-    }
-  ],
-  "forms": [
-    {
-      "actions": [
-        {
-          "name": "init",
-          "preconditions": [],
-          "uses": []
-        },
-        {
-          "name": "showTable",
-          "preconditions": [{
-            "id": "list",
-            "label": {"en-gb": "A product list"},
-            "pattern": "?PList
-                        http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                        http://aws.amazon.com/AWSECommerceService#ProductList",
-            "positive": true
-          }],
-          "uses": []
-        }
-      ],
-      ...,
-      "uri": "http://catalogueURL/FASTCatalogue/forms/9",
-      "version": "1.0"
-    },
-  ],
-  "operators": [],
-  "backendservices": []
-}
-\end{verbatim}
-\doublespacing
-
-\subsubsection{Screen Component Find \& Check} % (fold)
-\label{ssub:screen_component_findcheck}
-
-Similarly to the Screen Find \& Check described in Section~\ref{ssub:screen_findcheck}, this operation search for screen components (forms, operators and back-end services) through the catalogue in order to satisfy a given request, and it will attach information about satisfaction and reachability as well.
-
-A request is formed of the following parameters:
-
-\begin{description}
-	\item[preconditions] Preconditions of the screen.
-	\item[postconditions] Postcondition of the screen.
-	\item[canvas] The canvas is composed by a list of screens. Only the URI is needed.
-	\item[forms] It is a list of forms, such as the list of forms recommended last execution.
-	\item[operators] It is a list of operators, such as the list of operators recommended last execution.
-	\item[backendservices] It is a list of back-end services, such as the list of back-end services recommended last execution.
-	\item[search] Boolean indicating if forms/operators/services are required to be looked for. If \emph{false} the operation performed is just a check. If missed, it will act as \emph{true}.
-	\item[domainContext] The domain context contains a list of tags and a user.
-	\item[pipes] List of pipes for the connection between screen components.
-	\item[selectedItem] URI of a screen component from the canvas.
-\end{description}
-
-The screen components recommended are based on the given domain context and the pre/postconditions of the any of the resources in the canvas or the conditions of the parameters preconditions and postconditions.
-Hence, any screen component from the catalogue which has any of these conditions may be suitable for the screen, so it is included in the response, in the list corresponding to the type of screen component.
-
-An example of request is the following:
-
-\singlespacing
-\begin{verbatim}
-{
-  "preconditions": [{
-    "id": "Searchcriteria_1",
-    "label": {"en-gb": "A search criteria"},
-    "pattern": "?F
-                http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                http://aws.amazon.com/AWSECommerceService#SearchCriteria",
-    "positive": true,
-  }],
-  "postconditions": [],
-  "canvas": [
-    { "uri": "http://catalogueURL/catalogue/forms/1" },
-    { "uri": "http://catalogueURL/catalogue/services/1" }
-  ],
-  "forms": [],
-  "operators": [],
-  "backendservices": [],
-  "search": true,
-  "domainContext": {
-    "tags": [],
-    "user": null
-  },
-  "pipes": [
-    {
-      "from": {
-        "buildingblock": "http://catalogueURL/catalogue/services/1",
-        "condition": "list"
-      },
-      "to": {
-        "action": "showTable",
-        "buildingblock": "http://catalogueURL/catalogue/forms/1",
-        "condition": "list"
-      }
-    }
-  ],
-  "selectedItem": "http://catalogueURL/catalogue/forms/1"
-}
-\end{verbatim}
-\doublespacing
-
-And the response according to the above request is:
-
-\singlespacing
-\begin{verbatim}
-{
-  "backendservices": [],
-  "canvas": [
-    {
-      "actions": [{
-        "name": "search",
-        "preconditions": [{
-          "id": "filter",
-          "label": {"en-gb": "A search criteria"},
-          "pattern": "?F 
-                      http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                      http://aws.amazon.com/AWSECommerceService#SearchCriteria",
-          "positive": true,
-          "satisfied": false
-        }],
-        "satisfied": false,
-        "uses": []
-      }],
-      "reachability": false,
-      "uri": "http://catalogueURL/catalogue/services/1"
-    },
-    {
-      "actions": [
-        {
-          "name": "init",
-          "preconditions": [],
-          "satisfied": true,
-          "uses": []
-        },
-        {
-          "name": "showTable",
-          "preconditions": [{
-            "id": "list",
-            "label": {"en-gb": "A product list"},
-            "pattern": "?PList
-                        http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                        http://aws.amazon.com/AWSECommerceService#ProductList",
-            "positive": true,
-            "satisfied": false
-          }],
-          "satisfied": false,
-          "uses": []
-        }
-      ],
-      "reachability": true,
-      "uri": "http://catalogueURL/catalogue/forms/1"
-    }
-  ],
-  "connections": [
-    {
-      "from": {
-        "buildingblock": null,
-        "condition": "Searchcriteria_1"
-      },
-      "to": {
-        "action": "search",
-        "buildingblock": "http://catalogueURL/catalogue/services/1",
-        "condition": "filter"
-      }
-    },
-  ],
-  "forms": [],
-  "operators": [],
-  "pipes": [
-    {
-      "from": {
-        "buildingblock": "http://catalogueURL/catalogue/services/1",
-        "condition": "list"
-      },
-      "satisfied": true,
-      "to": {
-        "action": "showTable",
-        "buildingblock": "http://catalogueURL/catalogue/forms/1",
-        "condition": "list"
-      }
-    }
-  ],
-  "postconditions": []
-}
-\end{verbatim}
-\doublespacing
-
-For a screen component to be reachable, its preconditions need to be connected by a pipe to any other screen component which is already reachable, or to any of the screen preconditions.
-Moreover, the pipe needs to be satisfied, in other words, conditions in both edges are compatible.
-
-The response includes an attribute called "connections". This attribute will be retrieved when "selectedItem" is given, and it will contain a list of potential pipes to connect to that item.
-
-% subsubsection screen_component_findcheck (end)
-
-% subsection api_calls (end)
-
-\subsection{Planning}
-\label{sub:planning}
-
-While creating a screen-flow, the user need to find proper screens which will satisfy her necessity.
-Once the user has selected one or several screens to be included in her screen-flow, she needs to find screens which satisfy the preconditions of these screens, in order to create an executable screen-flow.
-The operation Find, detailed in Section~\ref{ssub:screen_find}, helps the user to find these screens, however the response is just a set of screens which somehow may help to accomplish her goal, but not in a single and straight-forward step.
-
-This functionality offers to the user the possibility to receive a whole set of screens, or plan, which will make a given screen, or goal, reachable.
-These plans take into account the screen which are already in the canvas, in order to minimise the number of screens.
-
-The specific URL to access this operation is \url{http://catalogueURL/planner}, and to execute the operation a POST request needs to be done. The request body is composed of:
-
-\begin{description}
-	\item[goal] Required. URI of the goal screen.
-	\item[canvas] Required. List of screens URIs.
-	\item[page] Optional. Number of the page to be retrieved.
-	\item[per\_page] Optional. Number of plans to be included per page in the response.
-\end{description}
-
-Let's see the following example:
-
-Request
-\singlespacing
-\begin{verbatim}
-{
-  "goal": "http://catalogueURL/catalogue/screens/35",
-  "canvas": [
-    { "uri": "http://catalogueURL/catalogue/screens/56" },
-    { "uri": "http://catalogueURL/catalogue/screens/85" }
-  ],
-  "page": 1,
-  "per_page": 10 
-}
-\end{verbatim}
-\doublespacing
-
-Response
-\singlespacing
-\begin{verbatim}
-[
-  [
-    "http://catalogueURL/catalogue/screens/13",
-    "http://catalogueURL/catalogue/screens/23"
-  ],
-  [
-    "http://catalogueURL/catalogue/screens/79",
-    "http://catalogueURL/catalogue/screens/37",
-    "http://catalogueURL/catalogue/screens/14"
-  ],
-  [
-    "http://catalogueURL/catalogue/screens/16",
-    "http://catalogueURL/catalogue/screens/44",
-    "http://catalogueURL/catalogue/screens/32"
-  ]
-]
-\end{verbatim}
-\doublespacing
-
-The response is ranked based on adding the minimum number of new screens into the canvas, and giving preference to the screens the user has previously selected (canvas).
-
-In the above example, let's assume the first plan is using the two screens from the canvas to accomplish the goal, and the second plan is able to make the goal reachable by itself.
-That said, a screen-flow with the goal and the second plan could be executable, and it would be composed of just four screens, however, taking the first plan, the resulting screen-flow will be composed of five screens.
-It could make sense to place the second plan in first position, but the plan with just two screens reuse the screens from the canvas which has been added by the user.
-
-% subsection planning (end)
-
-\subsection{API Error Codes} % (fold)
-There are two types of error codes, client and server.
-\begin{itemize}
-	\item Client error codes are generally caused by the client and might be an invalid domain or an invalid request parameter.
-These errors are accompanied by a 4xx HTTP response code.
-For example: ResourceNotFound.
-	\item Server error codes are generally caused by a server-side issue and should be reported. These errors are accompanied by a 5xx HTTP response code. For example: ServerUnavailable.
-\end{itemize}
-
-The following table lists all the error codes.
-
-\begin{table}[htb]
-\caption{API Error Codes}
-\begin{center}
-\begin{tabular}{|p{3.5cm}|p{7cm}|p{3cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Error} &
-                         \textcolor{white}{Description} &
-                         \textcolor{white}{HTTP Status Code}\\ \hline
-ResourceNotFound & The resource <resourceURI> has not been found. & 404 Not Found \\ \hline
-AccessFailure & Access to the resource <resourceName> is denied. & 403 Forbidden \\ \hline
-InternalError & Request could not be executed due to an internal service error. & 500 Internal Server Error \\ \hline
-InvalidAction & The action <actionName> is not valid for this web service.  & 404 Bad Request \\ \hline
-InvalidHttpRequest & The HTTP request is invalid. Reason: <reason>. & 400 Bad Request \\ \hline
-InvalidLiteral & Illegal literal in the filter expression. & 400 Bad Request \\ \hline
-InvalidParameterValue & The specified parameter value is not valid. & 400 Bad Request \\ \hline
-InvalidURI & The URI <requestURI> is not valid. & 400 Bad Request \\ \hline
-MissingAction & No action was supplied with this request. & 400 Bad Request \\ \hline
-MissingParameter & The request must contain the specified missing parameter. & 400 Bad Request \\ \hline
-NotYetImplemented & Feature <feature> is not yet available. & 401 Unauthorized \\ \hline
-ServiceUnavailable & The service is currently unavailable. Please try again later. &  \\ \hline
-UnsupportedHttpVerb & The requested HTTP verb is not supported: <verb>. & 400 Bad Request \\ \hline
-URITooLong & The URI exceeded the maximum limit of <maxLength>. & 400 Bad Request \\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\clearpage
-\bibliographystyle{apalike}
-\bibliography{fast_d5_1_1}
-
-\clearpage
-\section*{Appendix A. Semantic Web Technologies Evaluation}
-\label{appendix_a}
-\addcontentsline{toc}{section}{Appendix A. Semantic Web Technologies Evaluation}
-
-In \cite{urmetzer2010fast_state_of_the_art} were presented different technologies used in the Semantic Web community regarding Web services and catalogues or repositories.
-These are potential technologies to serve as the basis for the Catalogue implementation.
-Said that, an evaluation was done in order to choose the most accurate technology with the most advantages for it.
-Of particular importance in this evaluation is real-time performance, because the GVS (Gadget Visual Storyboard) actually relies on the Catalogue in parts of its user interface.
-
-Basically, two general solutions were considered.
-One was to build the Catalogue, and the internal components or building block of the gadgets, based on a dedicated SWS platform.
-The other was to build a light-weight solution from scratch based on RDF and RDFS.
-
-For both solutions, a few ideas about their potential benefits and disadvantages were in mind beforehand.
-SWS platforms were very tempting to use, because the modelling approach for screens in FAST is quite similar to SWS, in how they define their inputs and outputs.
-E.g., both FAST and WSMO have pre- and post-conditions, OWL-S has inputs/outputs.
-Also, SWS platforms come with ready-made implementations for service composition, which could have been used for automatically combining screens to screen-flows.
-On the downside, it was expected that SWS platforms a bit too slow for some FAST requirements, as the need of real-time discovery.
-On the other hand, building a light-weight implementation directly on RDFS promised to be faster.
-Moreover, there is a broader, more mature and probably more active tool support.
-However, going in this direction would imply to invest a lot more work in modelling and implementation efforts, since all the features that SWS would bring straight away are could not be reused.
-
-To perform the evaluation of the different approaches, an abstract random ontology of 40 concepts was created.
-About 70\% of those were randomly assigned to be sub-concepts of other concepts.
-The abstract ontology was instantiated in WSMO, OWLS and RDFS.
-Then, 3 sample sets of random screens of three different sizes (100, 1000 and 10000) were created, such that each screen has 0..3 random pre- and post-conditions.
-For each set in each technology, every screen was taken to perform a match with any other screen, and the average response time was measured.
-Figure~\ref{fig:comparison_chart} shows the results of the evaluation (RDFS in blue, WSMO in yellow and OWL-S in green).
-Be aware that the scale of the time is logarithmic.
-
-\begin{figure}[htb]
-\label{fig:comparison_chart}
-\begin{center}
-	\includegraphics[width=14cm]{images/comparison_chart}
-	\caption{Semantic Web Technologies Performance Comparison}
-\end{center}
-\end{figure}
-
-It can be seen that RDFS performs rather well, even in the 10.000 size set, whereas the other two are pretty far behind.
-E.g., average response times for WSMO is > 1.5 sec already for the smallest 100 size set.
-Hence, based on this evaluation the approach to follow for the implementation was RDFS, since WSMO and OWL-S simply could not give the performance required for the real-time discovery.
-
-\clearpage
-\section*{Appendix B. Generic Building Block JSON Structure}
-\label{appendix_b}
-\addcontentsline{toc}{section}{Appendix B. Generic Building Block JSON Structure}
-
-\singlespacing
-\begin{verbatim}
-{
-  "code": "http://url.com/.../code.js",
-  "creationDate": "2009-04-20T17:00:00+0100",
-  "creator": "username",
-  "description": {"en-gb": "This is a description of the building block"},
-    "tags":[
-      {
-        "label":{"en-gb":"Amazon"},
-        "means":"http://dbpedia.org/page/Amazon.com"
-      },
-      {
-        "label":{"en-gb":"eBay"},
-        "means":"http://dbpedia.org/page/Ebay"
-      }
-    ],
-  "homepage": "http://url.com/.../homepage",
-  "icon": "http://url.com/images/.../icon.png",
-  "id": "Id of the building block", // i.e. 3545
-  "label": {"en-gb": "Label or title of the building block"},
-  "name": "User-friendly name of the building block",
-  "rights": "http://creativecommons.org/",
-  "screenshot": "http://url.com/.../screenshot.jpg",
-  "type": "type of the building block", // screenflow, screen, form, operator or resource
-  "uri": "http://url.com/screens/525",
-  "version": "1.0"
-}
-\end{verbatim}
-\doublespacing
-
-\clearpage
-\section*{Appendix C. Screen-flow JSON Structure}
-\label{appendix_c}
-\addcontentsline{toc}{section}{Appendix C. Screen-flow JSON Structure}
-
-\singlespacing
-\begin{verbatim}
-{
-  <GENERIC BUILDING BLOCK JSON>,
-  "contains": [
-    "http://url.com/screens/72723",
-    "http://url.com/screens/20035",
-  ],
-}
-\end{verbatim}
-\doublespacing
-
-\clearpage
-\section*{Appendix D. Screen JSON Structure}
-\label{appendix_d}
-\addcontentsline{toc}{section}{Appendix D. Screen JSON Structure}
-
-\singlespacing
-\begin{verbatim}
-{
-  <GENERIC BUILDING BLOCK JSON>,
-  "postconditions": [[{
-    "id": <identifier:String>,
-    "label": {"en-gb": "Purchase URL"},
-    "pattern": "?P
-                http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                http://aws.amazon.com/AWSECommerceService#PurchaseURL",
-    "positive": true
-  }]],
-  "preconditions": [[{
-    "id": "cart",
-    "label": {"en-gb": "A shopping cart"},
-    "pattern": "?C
-                http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                http://aws.amazon.com/AWSECommerceService#ShoppingCart",
-    "positive": true
-  }]],
-  "code": "URL of the code",
-  "definition": {
-    "buildingblocks: [
-      { // Form
-        "id": "form1", // unique for the container screen
-        "uri": "http://purl.oclc.org/fast/ontology/gadget#Form670238"
-      },
-      { // Operators
-        "id": "op1",
-        "uri": "http://purl.oclc.org/fast/ontology/gadget#Operator213487"
-      },
-      {  // Backend services
-        "id": "bs1",
-        "uri": "http://purl.oclc.org/fast/ontology/gadget#Backendservice38399"
-      },
-      {...}
-    ]
-    "pipes": [
-      {
-        "from": {
-          "buildingblock": "bs1",
-          "condition": "cA"
-        },
-        "to": {
-          "buildingblock": "op1",
-          "condition": "cA", // can be the same id, unique for the building block
-          "action": "filter"
-        }
-      },
-    "triggers": [
-      {
-        "from": {
-          "buildingblock": "form1",
-          "name": "refresh"
-        },
-        "to": {
-          "buildingblock": "bs1",
-          "action": "list"
-        }
-      }
-    ]
-  }
-}
-\end{verbatim}
-\doublespacing
-
-\clearpage
-\section*{Appendix E. Screen Component JSON Structure}
-\label{appendix_e}
-\addcontentsline{toc}{section}{Appendix E. Screen Component JSON Structure}
-
-\singlespacing
-\begin{verbatim}
-{
-  <GENERIC BUILDING BLOCK JSON>,
-  "actions": [{
-    "name": "filter",
-    "preconditions": [[{
-      "id": "item",
-      "label": {"en-gb": "Ebay List"},
-      "pattern": "?Item
-                  http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                  http://aws.amazon.com/AWSECommerceService#Item",
-      "positive": true
-    }]],
-    "uses": [{
-      "id":"cart",
-      "uri":"http://aws.amazon.com/AWSECommerceService#ShoppingCart"
-    }]
-  }],
-  "libraries": [{
-    "language":"JavaScript",
-    "source":"http://url.com/libcode.js"
-  }],
-  "postconditions": [[{
-    "id": "filterEbay",
-    "label": {"en-gb": "Ebay List"},
-    "pattern": "?eFilter
-                http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                http://developer.ebay.com/.../FindItemsAdvanced.html#Request",
-    "positive": true
-  }]],
-  "triggers": ["itemAmazon"]
-}
-\end{verbatim}
-\doublespacing
-
-\end{document}

### 2010-06-14 trunk/papers/icsoc2010/paper_ssec_ontology.tex
split it into different files
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@844 d8b0c59b-6544-0410-94d9-83a5658a058e

\subsection{Simplified Model/Ontology (Knud)}
The model used to define the web service wrappers within this application has been influenced by WSDL, the main web service definition language, and some semantic-enriched approaches such as WSMO and OWL-S. There is no need to reinvent the wheel,

A WSDL operation is mapped to an action. Its inputs and outputs are called pre-postconditions (a WSMO concept).

This section could talk about a the part of the ontology describing the back-end services, how we remove the unnecessary information from WSDL and similar (not need anymore because the code has been generated), however we reuse some of the concepts from these technologies to define them (preconditions, postconditions, actions, and so on) allowing discovery and composition.

%\begin{figure}[ht]
%	\centering
%	\includegraphics[width=8.31cm]{uddi.png}
%	\caption{UDDI overview}
%	\label{fig:uddi}
%\end{figure}

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_introduction.tex
reference added
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@845 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-Dealing with Business-to-consumer (B2C) integration, the adoption of these standards have offered some advantages as well, however the interaction and consumption of web services requires a programming skills and understanding of many technical details about the technology, which is an obstacle for an end-user with limited knowledge on the matter.
+Dealing with Business-to-consumer (B2C) integration, the adoption of these standards have offered some advantages as well, however the interaction and consumption of web services requires a programming skills and understanding of many technical details about the technology, which is an obstacle for an end-user with limited knowledge on the matter (in the sense as defined in \cite{fuchsloch2010}).
 Regarding discovery, most of the publishing and discovery platforms are syntactic-base, leading to poor precision and recall, making difficult to navigate through a large variety of web services \cite{pilioura_acm2009}, and as a result not empowering the end-users to tackle easily with this task.

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_conclusions.tex
draft for conclusions section
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@846 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-Future work:
+It has been demostrated that adopting web services standards lead enterprises to increase operational efficiency, reduce costs and strength the relations with partners.
+Many of these standards have been widely adopted, such as WSDL, XML and UDDI, and many companies offer access to their information and operational system through web services.
+However, when they tackle with end-users, the process for publication and consumption is not well defined.
+They common way of publishing is directly by providing some sort of definition in their websites, such a WSDL document and further details in a human readable HTML page.
+The discovery is made through a search engine or aggregator, mainly in a syntatic-based way, and for their consumption, in most cases, high programming skills are required.
+The work presented in this paper empowers the end-user with a platform to easily discover services based on functional behaviour (pre/postconditions) and other metadata, being able to download a so-called resource adapter allowing the consumption of web services using languages to execute within a web browser, and a tool to transform, in an interactive manner, formal definitions of web service into these resource adapter, ready to being publish into the platform.
 
-Talk about wrapping WSDL-based services
-
-Talk about WSMO-lite services
-
-Use JSONRequest to replace the AJAX XMLHttpRequest and its problems (cross-domain calls), so there's no need for a Proxy.
+The current wrapping tool version permits to create resource adapters for RESTful web services.
+Next step is to tackle with SOAP-based web services which are described in WSDL documents, hence we will cover the two most used paradigms for leveraging web services.
+However, we consider to include semantically enriched WSDL documents using SAWSDL, and to support WSMO-lite services.
+Another limitation of the platform is inherent from web clients technologies and the cross-domain policy for security.
+This is solved within FAST using platform dependent API calls.
+Hence, the code generated makes use of the FAST API, which will be transformed depending on the mashup platform the user intends to deploy the gadget.
+To avoid these issues, and to be able to provide platform-independent code right away, we are researching into JSONRequest proposed by \cite{crockford2006} and other solutions.

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_use_case.tex
minor change
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@847 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 Therefore, it is aimed to allow users to create their own applications by combining different components, or building blocks, such as graphical forms and back-end services in terms of their inputs and outputs.
-These two components, as described in Section 3 and 4, are fully integrated and present a high cohesion the rest of the components of the FAST architecture (see Picture \ref{pic:fast_architecture}.
+These two components, as described in Section 3 and 4, are fully integrated and present a high cohesion the rest of the components of the FAST architecture (see Picture~\ref{pic:fast_architecture}.
 

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_introduction.tex
introduction updated
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@848 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-06-14 trunk/papers/icsoc2010/paper_sec_introduction.tex
@@
+%!TEX root = paper.tex
+
 \section{Introduction}
 
-Business-to-business or B2B integration is easier said than done.
-Integration is a big challenge, in most cases requires a huge amount of effort, and deal with many factors and tecnologies for protocols, architectures, security standards, among others.
+Business-to-business (B2B) integration is easier said than done.
+Integration is a big challenge, in most cases requires a huge amount of effort, and deal with many factors and technologies for protocols, architectures, security standards, among others.
 Solutions based on open standards have reduced the complexity of integrating different business applications between different companies and partners.
 Web services and other XML standards such as RosettaNet, ebXML or OAG have been a boon to the world of B2B.
-By web services standards we are refering to the following open standards: Web Services Description Language (WSDL - to describe), Universal Description, Discovery and Integration (UDDI - to advertise and syndicate), Simple Object Access Protocol (SOAP - to communicate) and Web Services Flow Language (WSFL - to define work flows).
+By web services standards we are referring to the following open standards: Web Services Description Language (WSDL - to describe), Universal Description, Discovery and Integration (UDDI - to advertise and syndicate), Simple Object Access Protocol (SOAP - to communicate) and Web Services Flow Language (WSFL - to define work flows).
 Therefore, a common scenario is a company using WSDL to describe its public and private web service, publishing them either using a public or private repository using UDDI, where these web services using SOAP-based messages to achieve dynamic integration between different disparate applications.
 
-Dealing with Business-to-consumer (B2C) integration, the adoption of these standards have offered some advantages as well, however the interaction and consumption of web services requires a programming skills and understanding of many technical details about the technology, which is an obstacle for an end-user with limited knowledge on the matter.
-Regarding discovery, most of the publishing and discovery platforms are syntactic-base, leading to poor precision and recall, making difficult to navigate through a large variety of web services \cite{pilioura_acm2009}, and as a result not empowering the end-users to tackle easily with this task.
-Other solutions like Semantic Web Service promised many advantages in discovery, composition and consumption of web services, independently of the provider's platform, the location, the service implementation, or the data format.
-However, their adoption was not as expected [TODO: any reference?], the further requirements needed and extra work have not being seen to be worth by enterprises for the benefits they offer.
-In the same direction as B2C, goverments are currently opening their data to the public as linked data.
-So far, the efforts are being focused on data, and the users have to deal with the development of applications to use that data.
-It is easily to predict that sooner or later they will move to open their services for administration tasks and single-window applications.
-Hence, they will be benefict by the research done around these concepts and technologies.
+Dealing with Business-to-consumer (B2C) integration, the adoption of these standards have offered some advantages as well, however the interaction and consumption of web services requires programming skills and understanding of many technical details about the technology, which is an obstacle for an end-user with limited knowledge on the matter (in the sense as defined in \cite{fuchsloch2010}).
+Regarding discovery, most of the publishing and discovery platforms are syntax-based, leading to poor precision and recall [TODO: of what?], making difficult to navigate through a large variety of web services \cite{pilioura_acm2009},
+therefore preventing end-users from performing these tasks. 
+Other solutions like Semantic Web Services promised many advantages in discovery, composition and consumption of web services, independently of the provider's platform, location,  service implementation or data format.
+However, as of yet they have not been widely adopted, 
+possibly because the perceived potential benefits of SWS did not justify the additional investments required for businesses. 
+In a similar vein as B2C, governments are currently opening their data to the public as linked data. 
+So far, the efforts are restricted to data alone, so that users left to find applications of that data. 
+A natural further development in the public sector would be that their services for administration tasks and single-window applications will be opened as well.
+Hence, governments will be benefit from the research done around these concepts and technologies.
 
 The motivation of our work has been strongly influenced by the end-users' needs.
-We are targeting non-skilled users in term of programming and software development knowledge, intending to empower them with tools to discover and consume web services in a straight-forward manner.
-The web services and their specifications are not published right away in the platform, since we were just reimplementing UDDI.
+We are targeting users which are non-skilled in term of programming and software development, intending to empower them with tools to discover and consume web services in a straight-forward manner.
+The web services and their specifications are not published right away in the platform, since this would simply mean to reimplement UDDI.
 However, their definition and behaviour are wrapped in what we call a ``web service wrapper''.
@@
 Section 2 covers the state of the art in web services publication, discovery and wrapping/consumption.
-In Section 4 we present how we tackle with the web services wrapping, available possibilies and limitations.
+In Section 4 we present how we tackle with the web services wrapping, available possibilities and limitations.
 Next, section 3 describes our platform for publishing web services and empowering end-users to discover services.
-Section 5 gives the reader a clearer use case and how this work is integrated into the FAST platrofmr, and finally, we present our main conclusions and highlight future lines of research in Section 6.
+Section 5 gives the reader a clearer use case and how this work is integrated into the FAST platform, and finally, we present our main conclusions and highlight future lines of research in Section 6.

### 2010-06-14 trunk/papers/icsoc2010/paper_abstract.tex
abstract revised
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@849 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
+%!TEX root = paper.tex
+
 \begin{abstract}
-B2B systems integration was revolutionized by the introduction of web services.
+B2B systems integration was revolutionised by the introduction of web services.
 The way enterprise systems communicated was dramatically transformed, and by adopting these technologies, the relationships with providers and customers were strengthened.
 The main advantages seen by companies which adopted the technology were an increase of operational efficiencies, and reduction of costs.
-In this scenario, high qualified software developers tackle with the integration, by means of the understanding of WSDL documents or human-readable specifications.
+In this scenario, highly qualified software developers are responsible for the integration of services with other systems, by means of the analysing of WSDL service description or human-readable specifications.
 However this model fails when it tries to target the long tail of enterprise software demand, and as a result, the end-users.
 Discovery and consumption of web services is far from a straight-forward task for an end-user, meaning potential users, willing to create their task-specific applications, have been ruled out.
-This paper presents an approach to facilitate the discovery and consumption of web services by end-users on the Internet, breaking the gap between business web services and the latter.
+This paper presents an approach to facilitate the discovery and consumption of web services by end-users on the Internet, closing the gap between business web services and the latter.
 The process includes: (a) a way to generate ready-to-use web services wrappers, and (b) a catalogue which users can browse to search for web services fitting their needs.
-These web services wrappers are described by a model created, although highly influenced by well-known web services modeling standards.
+These web services wrappers are described by a model created, although highly influenced by well-known web services modelling standards.
 Furthermore, two prototypes have been developed to serve as a proof of concept.

### 2010-06-14 trunk/papers/icsoc2010/paper.tex
changed title
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@850 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 % first the title is needed
-\title{An improved platform for Web service wrapping, discovery and consumption}
+\title{Web Service Wrapping, Discovery and Consumption --- More Power to the End-User}
 
 % a short form should be given in case it is too long for the running head
-\titlerunning{Enhancing the discovery and consumption of Web services}
+\titlerunning{Web Service Wrapping, Discovery and Consumption --- More Power to the End-User}
 

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_introduction.tex
introduction updated
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@851 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \section{Introduction}
+\label{sec:introduction}
 
@@
 Web services and other XML standards such as RosettaNet, ebXML or OAG have been a boon to the world of B2B.
-By web services standards we are referring to the following open standards: Web Services Description Language (WSDL - to describe), Universal Description, Discovery and Integration (UDDI - to advertise and syndicate), Simple Object Access Protocol (SOAP - to communicate) and Web Services Flow Language (WSFL - to define work flows).
+By web services standards we are referring to the following open standards: Web Services Description Language (WSDL --- to describe), Universal Description, Discovery and Integration (UDDI --- to advertise and syndicate), Simple Object Access Protocol (SOAP --- to communicate) and Web Services Flow Language (WSFL --- to define work flows).
 Therefore, a common scenario is a company using WSDL to describe its public and private web service, publishing them either using a public or private repository using UDDI, where these web services using SOAP-based messages to achieve dynamic integration between different disparate applications.
 
-Dealing with Business-to-consumer (B2C) integration, the adoption of these standards have offered some advantages as well, however the interaction and consumption of web services requires programming skills and understanding of many technical details about the technology, which is an obstacle for an end-user with limited knowledge on the matter (in the sense as defined in \cite{fuchsloch2010}).
-Regarding discovery, most of the publishing and discovery platforms are syntax-based, leading to poor precision and recall [TODO: of what?], making difficult to navigate through a large variety of web services \cite{pilioura_acm2009},
+Dealing with Business-to-consumer (B2C) integration, the adoption of these standards have offered some advantages as well, however the interaction and consumption of web services requires programming skills and understanding of many technical details about the technology, which is an obstacle for an end-user with limited knowledge on the matter (in the sense as defined in~\cite{fuchsloch2010}).
+Regarding discovery, most of the publishing and discovery platforms are syntax-based, leading to poor precision and recall in finding the correct service, making difficult to navigate through a large variety of web services~\cite{pilioura_acm2009},
 therefore preventing end-users from performing these tasks. 

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_related_work.tex
content reduced
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@853 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 System integration, within a company or between systems from different enterprises became easier with the adoption of the web services standard technologies, being the Web Service Description or WSDL the most extended standard used for the definition of web services.
-In fact, there are many integrated development environments (IDEs) that can deal with WSDL, and create a plain vanilla proxy class for the desired platform / programming language, getting a more ``developer-friendly'' API built around proxy classes, instead of having to know all request parameter names and possible values for the different calls to the Web service.
-Even though, in an integration of two or more systems scenario, developers must read and understand their specifications in order to accomplish the task, in a programmatically-fashion way.
-Another advantage of the use of WSDL is validation.
-Having the request messages (methods and parameters) and response messages (return parameters / structures) built upon XML schemas, a Web service engine can validate the message conforming to a schema.
-However, as it has been said before, the use of these languages and tools requires a qualified developer.
-In a inter-company system integration scenario these technologies are of high value, but when a company opens its services to the public, it is clear that the common end-user has been left off-side.
-
-There are several tools out there situated a step forward to facilitate the interaction with data sources and services from around the Web.
+Many integrated development environments (IDEs) can deal with WSDL to facilitate the integration task, however it is still required that developers read and understand their specifications, and implement part of the logic in a programmatically-fashion way.
+A step forward, there are several tools which facilitate the interaction with data sources and services on the Web.
 Yahoo!
-Pipes provides a set of modules to access different kind of data sources, such as RSS feeds, a given web page (HTML code), Flickr images, Google base or the Yahoo!
-search engine, among others.
-These modules are developed by the widget platform itself, they do not allow users to build their own modules, and it lacks of a way to interact with REST or SOAP web service.
-Another tool worth to mention is Apatar, an enterprise-oriented data integration and ETL tool.
-It allows the interaction with databases (MySQL, PostgreSql, Oracle), powerful enterprise systems such as Salesforce CRM, SugarCRM and Goldmine CRM, and access to data sources such as XML or text files.
-It enables non-developers to design and perform data transformations in a visual way through its graphical interface; however it misses support for Web services.
-Yet another tool, in the same direction as Apatar, could be JackBe Presto, defined as an enterprise-ready mashup solution.
-Presto permits to integrate and extract data from a great variety of data sources such as RSS and ATOM feeds or XML-based sources, and empower the user to create connectors to use services from HP SOA Systinet and any Oracle information technologies including Oracle 9i/10g/11g, Oracle Fusion SOAs, and Oracle Applications.
-
-Summarising up, none of the solutions in the market really facilitate end-users to build up their own applications or widgets allowing the interaction with Web services created by third-party providers.
+Pipes, Apatar, JackBe Presto, Microsoft Popfly and NetVibes, among others provide a set of modules to access different kind of data sources, such as RSS feeds, a given web page (HTML code), Flickr images, Google base or the Yahoo!
+search engine, databases (MySQL, PostgreSql, Oracle), and powerful enterprise systems such as Salesforce CRM, SugarCRM and Goldmine CRM.
+However, none of the solutions in the market really facilitate end-users to build up their own applications or widgets allowing the interaction with Web services created by third-party providers.
 The solutions found are mainly data-oriented (RSS feeds, databases, raw text).
@@
 As a drawback, WS-Discovery do not support Internet-scale discovery, therefore make if useless for our purposes.
-
-[TODO: rewrite/extend this] The rule based approach of the service wrapper tool has been inspired by triple graph grammars, cf.
-\cite{schurr94}.
-While general triple graph grammars allow to relate complex graph structures with each other, in our case the target of a rule is always a single object or attribute.
-This facilitates the whole mechanism, reasonably.

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_wrapping.tex
figures added
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@855 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \label{sec:wrapping_web_services}
-
-%This section talks about the service wrapper tool (Kassel prototype)
-%
-%What it offers (afaik now it supports RESTful services, if support for SOAP-based services is also implemented, talk about it, otherwise explain how you expect to address it in the future work section
-%
-%How to create a wrapper (test-and-construct approach)
-%
-%Place this component within the FAST architecture
-%
-%Talk about formats: JSON and XML??
 
@@
 As an example, the Ebay Shopping web service will be studied.
-The following URL is invoked to retrieve a list of items corresponding to the search keywords \emph{USB}.
+The following URL is invoked to retrieve a list of items corresponding to the search keywords \emph{USB}:
 
-\url{http://open.api.sandbox.ebay.com/shopping?appid=KasselUn-efea-4b93-9505-5dc2ef1ceecd&version=517&callname=FindItems&ItemSort=EndTime&QueryKeywords=USB&responseencoding=XML}
+\begin{listing}
+\begin{verbatim}
+http://open.api.sandbox.ebay.com/shopping?appid=KasselUn-efea-4b93-9505-5dc2ef1ceecd
+  &version=517&callname=FindItems&ItemSort=EndTime&QueryKeywords=USB&responseencoding=XML
+\end{verbatim}
+\end{listing}
 
-As you may see, the URL invoked is http://open.api.sandbox.ebay.com/shopping and the parameters used in the example are:
+As you may see, the URL invoked is \url{http://open.api.sandbox.ebay.com/shopping} and the parameters used in the example are:
 \begin{description}
@@
 
-%\begin{figure}
-%  \begin{center} \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithPortDefinitions.jpg}
-%    \caption{Configuring parameters of a service wrapper (photoshoped screen dump)}
-%    \label{fig:construct_pre_post_conditions}
-%  \end{center}
-%\end{figure}
+\begin{figure}
+  \begin{center} \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithPortDefinitions.png}
+    \caption{Configuring parameters of a service wrapper}
+    \label{fig:construct_pre_post_conditions}
+  \end{center}
+\end{figure}
 
@@
 
-Figure ~\ref{fig:construct_service_request} shows the web based tool to construct the service requests.
+Figure~\ref{fig:construct_service_request} shows the web based tool to construct the service requests.
 
-%\begin{figure}
-%  \begin{center}
-%    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithRequestExample.jpg}
-%    \caption{Constructing the service request URL and its parameters (photoshoped screen dump)}
-%    \label{fig:construct_service_request}
-%  \end{center}
-%\end{figure}
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithRequestExample.png}
+    \caption{Constructing the service request URL and its parameters}
+    \label{fig:construct_service_request}
+  \end{center}
+\end{figure}
 
@@
 
-%\begin{figure}
-%  \begin{center}
-%      \includegraphics[angle=90,width=0.8\linewidth]{images/ServiceWrapperToolGVSWithTransformationRules.jpg}
-%    \caption{Interactive, rule based transforming of an XML response to domain objects (photoshoped screen %dump)}
-%    \label{fig:response_service_execution}
-%  \end{center}
-%\end{figure}
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithTransformationRules.png}
+    \caption{Interactive, rule based transforming of an XML response to domain objects}
+    \label{fig:response_service_execution}
+  \end{center}
+\end{figure}
 

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_introduction.tex
bib changed
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@857 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 Next, section 3 describes our platform for publishing web services and empowering end-users to discover services.
-Section 5 gives the reader a clearer use case and how this work is integrated into the FAST platform, and finally, we present our main conclusions and highlight future lines of research in Section 6.
+Section~\ref{sec:use_case} gives the reader a clearer use case and how this work is integrated into the FAST platform, and finally, we present our main conclusions and highlight future lines of research in Section 6.

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_use_case.tex
bib changed
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@857 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
+%!TEX root = paper.tex
+
 \section{Use case}
@@
 
-The two artifacts presented in this paper may be deployed and used separately by third-party applications.
-The intention was to show the advances made regarding web service discovery and consumption.
-However, they were developed in order to be part of the FAST platform.
+The intention of this paper was to show the advances made regarding web service discovery and consumption through the two artifacts developed as part of our research.
+While these artifacts may be deployed and used separately by third-party applications, they were originally developed to form the backbone of the FAST platform.
 FAST constitutes a novel approach to application composition and business process definition from a top-down user-centric perspective.

### 2010-06-14 trunk/papers/icsoc2010/paper_ssec_linked_data.tex
reduced content
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@858 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-In order to understand how content negotiation is implemented in the catalogue, a few concepts need to be understood.
-The URI of a particular building block must be understood as the identifier of the building block as such, as opposed to a particular representation (JSON, HTML, etc.) of it.
-In fact, each such representation has its own URI, under which it can be retrieved.
-Using the terminology established in \cite{w3c2004webArchitecture}, the building block is a \emph{non-information resource}, whereas each representation is an \emph{information resource}.
-In the process of content negotiation, the requesting agent will first request the URI of the building block as such (e.g., \texttt{http://catalogue.com/screens/235}), as well as the required representation format (e.g., JSON).
-The catalogue will then inspect the request and redirect the requester to the appropriate representation (e.g., \texttt{http://catalogue.com/screens/235.json}), as illustrated in Fig.~\ref{fig:content_negotiation}.
-On the web, different representations for the same resource are called \emph{variants}, and content negotiation is the mechanism used to determine which of the representations of most appropriate for a given request.
-
-%\begin{figure}[htb]
-%\label{fig:content_negotiation}
-%\begin{center}
-%	\includegraphics[width=12cm]{images/content_negotiation}
-%	\caption{Content Negotiation}
-%\end{center}
-%\end{figure}
-
-In summary, the basic idea of content negotiation, as stated in the \cite{http1.1}, is to serve the best variant for a resource, taking into account what variants are available, what variants the server may prefer to serve, what the client can accept, and with which preferences.
+To allow a third-party application to retrieve the content in the format required, we support content negotiation as stated in the \cite{http1.1}, serving the best variant for a resource, taking into account what variants are available, what variants the server may prefer to serve, what the client can accept, and with which preferences.
 In HTTP, this is done by the client which may send, in its request, accept headers (\texttt{Accept}, \texttt{Accept-Language} and \texttt{Accept-Encoding}), to communicate its capabilities and preferences in format, language and encoding, respectively.
-
-In fact, what the catalogue really does is ``format negotiation'' since the alternate representations are just based on the selection of the media type, through the accept header, but does not consider different languages or encoding types.
-The formats supported are JSON, RDF/XML, Turtle and HTML+RDFa.
-Even though the accept header is desired, the different representations can also be retrieved directly by dereferencing their own URI.
-The convention used in the catalogue for deriving a representation URI from a building block URI is to simply attach a matching suffix, such as \texttt{.rdf}, \texttt{.json} or \texttt{.html}.
-
-Lastly, content negotiation needs to identify which player is going to take the lead on it.
-There are two kinds of content negotiation which are possible in HTTP: server-driven and agent-driven negotiation.
-The approach followed by the catalogue is agent-driven negotiation, hence selecting a specific representation for a resource is responsibility of the user agent.
-If none is specified, by default, the server will choose the JSON representation.
+Concretely, the approach followed is agent-driven negotiation, where the user agent selects the specific representation for a resource.

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_catalogue.tex
some major changes
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@861 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
-\section{Discovery: the marketplace}
+\section{Discovery: web services publishing platform}
 \label{sec:discovery}
@@
 \begin{itemize}
+  \item allows functional discovery through web services pre and postconditions;
+	\item serves web services wrappers ready-to-consume in any web application;
+	\item provices advanced search capabilites, based on the service formal definition and inferences extracted from it;
+	\item supports managing its resources via its RESTful API;
+	\item a SPARQL endpoint to access directly to the data allowing advanced queries;
 	\item offers web services as Linked Data;
-	\item serves web services wrappers ready-to-consume in any web application;
 	\item follows well-known best practices for publishing data on the web;
-	\item supports managing its resources via its RESTful API;
 	\item supports content negotiation so the different clients may retrieved the information in JSON, RDF/XML, RDF/N3 or a human-readable HTML version;
-	\item provices advanced search capabilites, based on the service formal definition and inferences extracted from it;
-	\item a SPARQL endpoint to access directly to the data allowing advanced queries;
 \end{itemize}
 
-This platform was intented and it is being used as part of the FAST storyboard tool. Others components communicates with it via its RESTful API, but a
-That said, The process of publishing a web service requires two steps: (i) generation of the wrapper (see Sec.
+This platform was intented and it is being used as part of the FAST storyboard tool, and it communicates with others components through its RESTful API, using JSON for the input of the requests.
+This paper will focus on the functionality regarding web services publication and discovery, and the process to publish any third-party web service into the platform.
+This process requires two steps: (i) generation of the wrapper (see Sec.
 \ref{sec:wrapping_web_services}, and (ii) creation of the resource in the publishing platform.
-The creation of the resources is done via the public RESTful API, and the language selected for the input for any operation is JSON.
-The following is an example of part of a request sent to create a new wrapper into the platform:
+The following is an example of part of a request sent to create a new wrapper into the platform (for the sake of clarity, some parts of the request have been ommitted, but the main idea is shown):
 
@@
 {
-  "code": "http://demo.fast.morfeo-project.org/code/amazonService.js",
+  "code": "http://demo.fast.morfeo-project.org/code/amazonSearch.js",
   "creationDate": "2010-01-26T17:01:13+0000",
-  "description": {"en-gb": "Amazon web service"},
-  "label": {"en-gb": "Amazon web service"},
-  "preconditions": [
-    [
-      {
+  "description": {"en-gb": "Amazon web service - Search"},
+  "label": {"en-gb": "Amazon search"},
+  "actions": [{
+    "name": "filter",
+    "preconditions": [[{
       "id": "item",
-        "label": { "en-gb": "An item" },
+      "label": {"en-gb": "Ebay List"},
       "pattern": "?item
@@
                   http://aws.amazon.com/AWSECommerceService#Item",
-        "positive": "true"
-      }
-    ]
-  ],
+      "positive": true
+  }]],
   "postconditions": [
@@
 
-The architecture, as presented in the Figure \ref{fig:}, comprises a RDF store used for persistence, a business layer dealing with the model and reasoning, and public façade providing the core functionality as a RESTful API, and a SPARQL endpoint accessing directly the RDF store.
+The enriched search capabilities are supported by the definition of pre and postcondition.
+It allows to define the inputs and outputs of the services using concepts from any ontology, and find web services which match and hence is able to communicate with other web services or building block.
+The concepts of pre/postconditions were highly influenced by WSMO \cite{roman2005}, make it simpler and implemented in RDFS for a better performance.
+The model used to define the web service wrappers and the ontology created are detailed in Section~\ref{ssec:ontology}.
+
+\subsection{Platform architecture}
+\label{ssec:architecture}
+
+The architecture comprises a RDF store used for persistence, a business layer dealing with the model and reasoning, and public façade providing the core functionality as a RESTful API, and a SPARQL endpoint accessing directly the RDF store.
 This presentation layer is aimed to interact with the FAST Gadget Visual Storyboard, or any other third-party application.
@@
 The business logic layer contains all the domain-specific processing and reasoning.
-It provides functions to interact with all the elements of the domain model specified in \cite{sec:model}, acting as a mediator between the public RESTful API and the persistence layer.
+It provides functions to interact with all the elements of the domain model specified in \cite{ssec:ontology}, acting as a mediator between the public RESTful API and the persistence layer.
 
-The persistence layer provides an API to work with a standard set of objects representing the model.
+The persistence layer provides an API allowing to work with a standard set of objects representing the model.
 The interaction with the underlying RDF store is made via RDF2Go, an abstraction over triple (and quad) stores, which allows to program against rdf2go interfaces and choose any the RDF store implementation.
-This allows having a completely extensible and configurable framework for storage mechanisms, inferencers, RDF file formats, query result formats and query languages...
-
-How to publish (RESTful app, anyone can create a wrapper and publish it)
-
-How to browse: will allow users to go through the given domains (ontologies) and concepts, and will explain how to find the services they want (operation Find of the catalogue) => how the recommendation/reasoning works
-
-It offers the wrapped web services as linked data and has an open sparql endpoint, allowing other developers to create applications on top of it...
+This allows having a completely extensible and configurable framework for storage mechanisms, inferencers, RDF file formats, query result formats and query languages.
 

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_use_case.tex
checklist at end removed
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@862 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 The intention of this paper was to show the advances made regarding web service discovery and consumption through the two artifacts developed as part of our research.
-While these artifacts may be deployed and used separately by third-party applications, they were originally developed to form the backbone of the FAST platform.
+While these artifacts may be deployed and used separately by third-party applications, they were originally developed to form the backbone of the FAST platform~\cite{hoyer2009fast}.
 FAST constitutes a novel approach to application composition and business process definition from a top-down user-centric perspective.

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_wrapping.tex
changed ebay reference to footnote
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@863 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-In the top input field of Figure ~\ref{fig:construct_service_request}, the user may drop an example http request taken e.g.
+In the top input field of Figure ~\ref{fig:construct_service_request}, the user may drop an example HTTP request taken e.g.
 form the service documentation e.g.
-from \cite{eBayShoppingAPIs}.
+from its website\footnote{http://developer.ebay.com/products/shopping/}.
 The tool analyses the example request and in the middle of the screen a form for editing the request parameters is provided.

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_use_case.tex
architecture overview added
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@865 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 FAST constitutes a novel approach to application composition and business process definition from a top-down user-centric perspective.
-Therefore, it is aimed to allow users to create their own applications by combining different components, or building blocks, such as graphical forms and back-end services in terms of their inputs and outputs.
-These two components, as described in Section 3 and 4, are fully integrated and present a high cohesion the rest of the components of the FAST architecture (see Picture~\ref{pic:fast_architecture}.
+Therefore, it is aimed at allowing users (mainly business users at the management level, without programming experience) to create their own situational applications by visually combining different components, or building blocks, such as graphical forms and back-end services in terms of their inputs and outputs.
+These two components, as described in Sects.
+3 and 4, are fully integrated and present a high cohesion with the rest of the components of the FAST architecture (see Fig.~\ref{fig:fast_architecture}).
 
-[TODO: create a simple picture of the FAST architecture, highlighting these two components.]
+\begin{figure}[ht]
+  \begin{center}
+    \includegraphics[width=.9\linewidth]{images/fast_architecture.png}
+    \caption{Overview of the FAST architecture}
+    \label{fig:fast_architecture}
+  \end{center}
+\end{figure}
 
-For a better understanding of the use case, it is worth to describe a few concepts related to FAST.
-A Gadget is the output product, aimed to be used in a specific mashup platform.
-Inside FAST, the platform-independent gadget is called Screenflow, which comprises a set of domain-specific Screens connected through their pre and postconditions.
+In order to facilitate a better understanding of the use case, we will describe a number of concepts related to FAST in this section.
+A \emph{gadget} is the end product of the platform, aimed to be used in a specific mashup platform.
+Within FAST, the platform-independent gadget is called \emph{screenflow}, which comprises a set of domain-specific \emph{screens} connected through their \emph{pre-} and \emph{post-conditions}.
 A screen is the most complex building block fully functional by itself.
-It is composed by a form conveying the graphical interface, and a set of operators and resources binding to the back-end services.
+It is composed by a \emph{form} conveying the graphical interface, and a set of \emph{operators} and \emph{back-end services}, wrapped into so-called \emph{service resources}.
 

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_catalogue.tex
ontology references removed
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@867 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
+%!TEX root = paper.tex
+
 \section{Discovery: web services publishing platform}
@@
 
-The architecture comprises a RDF store used for persistence, a business layer dealing with the model and reasoning, and public façade providing the core functionality as a RESTful API, and a SPARQL endpoint accessing directly the RDF store.
+The architecture comprises a RDF store used for persistence, a business layer dealing with the model and reasoning, and public facade providing the core functionality as a RESTful API, and a SPARQL endpoint accessing directly the RDF store.
 This presentation layer is aimed to interact with the FAST Gadget Visual Storyboard, or any other third-party application.
@@
 
-The business logic layer contains all the domain-specific processing and reasoning.
-It provides functions to interact with all the elements of the domain model specified in \cite{ssec:ontology}, acting as a mediator between the public RESTful API and the persistence layer.
+The business logic layer contains all the domain-specific processing and reasoning. It provides functions to interact with all the elements of the domain model,
+% specified in \ref{ssec:ontology}
+acting as a mediator between the public RESTful API and the persistence layer.
 
@@
 
-\input{paper_ssec_ontology}
+% \input{paper_ssec_ontology}
 

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_use_case.tex
ontology references removed
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@867 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-Then, the web service wrapper tool, which is used to create the formal definitions of back-end services, transforming them into building blocks to be stored and reused within the FAST platform.
-It is integrated into the graphical IDE, and for the storage interacts with the catalogue, where the formal service descriptions and the executable code encapsulating the service behaviour is stored.
+The web service wrapper tool is then used to create the formal definitions of back-end services, transforming them into building blocks to be stored and reused within the FAST platform.
+The tool is integrated directly into the graphical IDE, and for the storage interacts with the catalogue, where the formal service descriptions and the executable code encapsulating the service behaviour is stored.
 
-The publishing platform presented, called ``Resource Catalogue'' within the FAST platform, covers several important purposes:
+The publishing platform presented, called ``Metadata Catalogue'' within the FAST platform, covers several important purposes:
 \begin{description}

### 2010-06-14 trunk/papers/icsoc2010/paper_ssec_ontology.tex
ontology references removed
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@867 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
+%!TEX root = paper.tex
+
 \subsection{Simplified Model/Ontology (Knud)}
+\label{ssec:ontology}
 The model used to define the web service wrappers within this application has been influenced by WSDL, the main web service definition language, and some semantic-enriched approaches such as WSMO and OWL-S.

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_conclusions.tex
conclusion revised
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@868 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-06-14 trunk/papers/icsoc2010/paper_sec_conclusions.tex
@@
+%!TEX root = paper.tex
+
 \section{Conclusions and future work}
+\label{sec:conclusions}
 
-Future work:
+It has been demonstrated that adopting web services standards lead enterprises to increase operational efficiency, reduce costs and strengthen the relations with partners.
+Many of these standards have been widely adopted, such as WSDL, XML and UDDI, and many companies offer access to their information and operational systems through web services.
+However, when dealing with end-users, the process for publication and consumption is not well defined. 
+The common way of publishing services is by providing some sort of definition on their websites, such a WSDL document and further details in a human-readable HTML page. 
+Discovery is performed through a search engine or aggregator, mainly in a syntax-based way. For consumption, in most cases, consultants with good programming skills are required. 
+In a move towards improving this situation, the work presented in this paper empowers the end-user with a platform (the \emph{catalogue}) to easily discover services based on functional behaviour (pre-/post-conditions) and other metadata, being able to download a so-called resource adapter allowing the consumption of web services using standard languages to execute within a web browser, and a tool to transform, in an interactive manner, formal definitions of web services into these resource adapters, ready for being publisher into the catalogue.
 
-Talk about wrapping WSDL-based services
+The current version of the wrapping tool permits to create resource adapters for RESTful web services. 
+As a next step, we will tackle SOAP-based web services described in WSDL documents.
+Once this is done, we will cover the two major paradigms for defining web service interfaces.
+However, we also consider to include semantically enriched WSDL documents using SAWSDL, and to support other SWS approaches such as WSMO-lite services.
 
-Talk about WSMO-lite services
-
-Use JSONRequest to replace the AJAX XMLHttpRequest and its problems (cross-domain calls), so there's no need for a Proxy.
+Another limitation of the platform is inherent in web client technologies and the cross-domain policy for security.
+This is solved within FAST using platform-dependent API calls.
+Hence, the code generated makes use of the FAST API, which will be transformed depending on the mashup platform the user intends to deploy the gadget.
+To avoid these issues, and to be able to provide platform-independent code right away, we are planning to deploy the JSONRequest approach as proposed by \cite{crockford2006} and other solutions.

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_use_case.tex
conclusion revised
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@868 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \begin{description}
-	\item[Publication and search of gadgets] The output of the FAST IDE are gadgets, which are self-contained and ready to use for users.
-They need to be made available for download, as well as being discover according to their formal description.
-	\item[Publication and search of building blocks] Every single component or building block definition is stored in the catalogue and it provides powerful ways of searching, recommendation and planning.
-	\item[Storage and indexing of user profiles and histories] Each user of the FAST platform will have a representation in the catalogue, thus making it possible for the IDE to make informed guesses about what they want in a particular situation, based on their previous behaviour.
+	\item[Publication and search of gadgets] The output of the FAST IDE are gadgets, which are self-contained and ready to use.
+They need to be made available for download, as well as being discovered according to their formal description.
+	\item[Publication and search of building blocks] Every single component or building block definition is stored in the catalogue, which facilitates searching, recommendation and planning (the suggestion of combinations of building blocks to fulfil certain requirements).
+	\item[Storage and indexing of user profiles and histories] Each user of the FAST platform has a representation in the catalogue, thus making it possible for the IDE to make informed guesses about what they want in a particular situation, based on their previous behaviour.
 	\item[Facilitate ontology mediation] Because FAST allows the integration of third-party external services, it is not always guaranteed that all the data coming from a service is directly compatible with the FAST platform.
 In such cases, mediation between the different ontologies is necessary.
-The catalogue provide some grade of mediation functionality.
+The catalogue provide some grade of mediation functionality (which is not covered in this paper).
 \end{description}

### 2010-06-14 trunk/papers/icsoc2010/paper.tex
checklist at end removed
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@862 d8b0c59b-6544-0410-94d9-83a5658a058e
changed short title
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@869 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 % a short form should be given in case it is too long for the running head
-\titlerunning{Web Service Wrapping, Discovery and Consumption --- More Power to the End-User}
+\titlerunning{Web Service Wrapping, Discovery and Consumption}
 
@@
 
-\section*{Checklist of Items to be Sent to Volume Editors}
-Here is a checklist of everything the volume editor requires from you:
-
-\begin{itemize}
-\settowidth{\leftmargin}{{\Large$\square$}}\advance\leftmargin\labelsep
-\itemsep8pt\relax
-\renewcommand\labelitemi{{\lower1.5pt\hbox{\Large$\square$}}}
-
-\item The final \LaTeX{} source files
-\item A final PDF file
-\item A copyright form, signed by one author on behalf of all of the
-authors of the paper.
-\item A readme giving the name and email address of the
-corresponding author.
-\end{itemize}
+% \section*{Checklist of Items to be Sent to Volume Editors}
+% Here is a checklist of everything the volume editor requires from you:
+% 
+% 
+% \begin{itemize}
+% \settowidth{\leftmargin}{{\Large$\square$}}\advance\leftmargin\labelsep
+% \itemsep8pt\relax
+% \renewcommand\labelitemi{{\lower1.5pt\hbox{\Large$\square$}}}
+% 
+% \item The final \LaTeX{} source files
+% \item A final PDF file
+% \item A copyright form, signed by one author on behalf of all of the
+% authors of the paper.
+% \item A readme giving the name and email address of the
+% corresponding author.
+% \end{itemize}
 \end{document}

### 2010-06-14 trunk/papers/icsoc2010/paper_abstract.tex
minor change
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@866 d8b0c59b-6544-0410-94d9-83a5658a058e
fixed some spelling mistkes
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@871 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 The main advantages seen by companies which adopted the technology were an increase of operational efficiencies, and reduction of costs.
-In this scenario, highly qualified software developers are responsible for the integration of services with other systems, by means of the analysing of WSDL service description or human-readable specifications.
+In this scenario, highly qualified software developers are responsible for the integration of services with other systems, by means of the analysis of WSDL service description or human-readable specifications.
 However this model fails when it tries to target the long tail of enterprise software demand, and as a result, the end-users.
@@
 Furthermore, two prototypes have been developed to serve as a proof of concept.
-\keywords{web services, linked data, end-users, discovery, consumption}
+\keywords{web services, end-users, discovery, consumption, linked data}
 \end{abstract}

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_related_work.tex
some sentences rewritten
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@872 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 Several tools permits some sort of (Web) service integration, but are meant to be used within an entreprise level by savvy business users or developers.
-[TODO: rewrite this paragraph] This paper presents an approach to leverage the possibility of integrate REST or SOAP-based web service inside a widget or any application for a web browser (by creating service wrappers), and graphical tool will be provided to facilitate the process of building these service wrappers.
+The solution presented in this paper leverage the possibility of integrating REST or SOAP-based web service inside visual applications running in a browser (e.g.
+widgets) by providing a platform to create, publish and discover web services wrappers.
 
@@
 A similar concept is iServe \cite{pedrinaci_ores2010}.
-This platform aims to publish Web services as what they called Lined Services Linked Services --linked data describing services, storing Web service definitions in a semantic-annotation fashion so semantic Web technologies for Web service discovery and processing can make use of them.
+That platform aims to publish Web services as what they called Linked Services Linked Services ---linked data describing services---, storing web service definitions in a semantic-annotation fashion so other semantic web technologies for web service discovery and processing may take advantage of it.
 However, the platform does not deal with the step from the definition to the consumption of the services.
 
-A different approach for Web service discovery is the Web Services Dynamic Discovery (WS-Discovery) specification.
+A different approach for web service discovery is the Web Services Dynamic Discovery (WS-Discovery) specification.
 The core of this approach is a multicast discovery protocol.

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_wrapping.tex
removed "Kassel"
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@870 d8b0c59b-6544-0410-94d9-83a5658a058e
some corrections
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@873 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
-\section{Wrapping web services (Kassel)}
+\section{Wrapping web services}
 \label{sec:wrapping_web_services}
@@
 
-If needed, a detailed specification of the Ebay Shopping API can be found at \cite{eBayShoppingAPIs}.
-
 The above URL for searching items in Ebay is followed by the query parameters, which take the form \textit{argument=value}.
@@
 
-Figure~\ref{fig:construct_service_request} shows the web based tool to construct the service requests.
+An example of service request construction can be found in the Figure~\ref{fig:construct_service_request}.
+In the top input field the user may drop an example HTTP request taken e.g.
+form the service documentation e.g.
+from its website\footnote{http://developer.ebay.com/products/shopping/}.
+The tool analyses the example request and in the middle of the screen a form for editing the request parameters is provided.
+In our example, the user has connected the \textit{QueryKeywords} parameter with the input parameter \textit{search\_key} by adding a corresponding reference to the value field of that parameter.
+In addition, we have retrieved an access key which has been entered as value for the \textit{appid} parameter.
 
@@
 \end{figure}
-
-In the top input field of Figure ~\ref{fig:construct_service_request}, the user may drop an example HTTP request taken e.g.
-form the service documentation e.g.
-from its website\footnote{http://developer.ebay.com/products/shopping/}.
-The tool analyses the example request and in the middle of the screen a form for editing the request parameters is provided.
-In our example, the user has connected the \textit{QueryKeywords} parameter with the input parameter \textit{search\_key} by adding a corresponding reference to the value field of that parameter.
-In addition, we have retrieved an access key which has been entered as value for the \textit{appid} parameter.
 

### 2010-06-14 trunk/papers/icsoc2010/paper_sec_catalogue.tex
some corrections
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@874 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-\section{Discovery: web services publishing platform}
+\section{Publishing and discovery platform}
 \label{sec:discovery}
@@
 However, as explained in previous sections, a large number of them suffer from limitated syntactic-based descriptions and a simple keyword-based search, and the gap between discovery and consumption is an obstacle for their usage by a non-technical common user.
-In order to ligthen the issue, this paper presents yet another publishing platform, permiting any enterprise or individual to publish their public domain web services, with an enhanced semantic search based on the definition of the services, and serving``web service wrappers'' for easily consumption in any web application.
+In order to ligthen the issue, this paper presents yet another publishing platform, permiting any enterprise or individual to publish their public domain web services, with an enhanced semantic search based on the definition of the services, and serving``web service wrappers'' for an easy consumption in any web application.
 

### 2010-06-14 trunk/papers/icsoc2010/paper_ssec_ontology.tex
half way to be done...
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@875 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-\subsection{Simplified Model/Ontology (Knud)}
+\subsection{Conceptual Model}
 \label{ssec:ontology}
-The model used to define the web service wrappers within this application has been influenced by WSDL, the main web service definition language, and some semantic-enriched approaches such as WSMO and OWL-S.
-There is no need to reinvent the wheel,
-
-A WSDL operation is mapped to an action. Its inputs and outputs are called pre-postconditions (a WSMO concept).
-
-This section could talk about a the part of the ontology describing the back-end services, how we remove the unnecessary information from WSDL and similar (not need anymore because the code has been generated), however we reuse some of the concepts from these technologies to define them (preconditions, postconditions, actions, and so on) allowing discovery and composition.
-
-%\begin{figure}[ht]
-%	\centering
-%	\includegraphics[width=8.31cm]{uddi.png}
-%	\caption{UDDI overview}
-%	\label{fig:uddi}
-%\end{figure}
+The conceptual model used to define the web service wrappers within this application has been influenced by WSDL, the main web service definition language, and some semantic-enriched approaches such as WSMO and OWL-S.
+It belongs to a more complex conceptual model for FAST, which can be grouped into three levels: the gadget and screenflow level at the top, the level of individual screens in the middle, and the level of web services at the bottom.
+In a few words, a screenflow is a runnable set of screens, and a screen is a executable set of building blocks with a specific purpose.
+This section only describes in detail the former level, defining the different components a screen comprises.
+These three different kinds of screen components are form, operator and resource, and share the same structure.
+Any screen component has a set of actions, which need a set of preconditions (inputs) to be fulfilled in order to execute the action (operation in WSDL), and provides a set of postcondition (ouputs).
+These pre- and post-conditions are defined as individual RDF triples.
+This means that conditions, which are sets of facts, can be modelled as graph patterns using SPARQL notation \cite{sparql2008spec}, where a statement like ``there exists a sioc:User'' will be transformed into a simple pattern like ``a variable ?user is of type sioc:User''.

### 2010-06-15 trunk/papers/icsoc2010/paper_abstract.tex
last changes everywhere
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@876 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \begin{abstract}
-B2B systems integration was revolutionised by the introduction of web services.
-The way enterprise systems communicated was dramatically transformed, and by adopting these technologies, the relationships with providers and customers were strengthened.
-The main advantages seen by companies which adopted the technology were an increase of operational efficiencies, and reduction of costs.
-In this scenario, highly qualified software developers are responsible for the integration of services with other systems, by means of the analysis of WSDL service description or human-readable specifications.
-However this model fails when it tries to target the long tail of enterprise software demand, and as a result, the end-users.
-Discovery and consumption of web services is far from a straight-forward task for an end-user, meaning potential users, willing to create their task-specific applications, have been ruled out.
-This paper presents an approach to facilitate the discovery and consumption of web services by end-users on the Internet, closing the gap between business web services and the latter.
-The process includes: (a) a way to generate ready-to-use web services wrappers, and (b) a catalogue which users can browse to search for web services fitting their needs.
-These web services wrappers are described by a model created, although highly influenced by well-known web services modelling standards.
-Furthermore, two prototypes have been developed to serve as a proof of concept.
+B2B systems integration was revolutionised by the introduction of web services,
+transforming the way enterprise systems communicate and strengthening the  relationships with providers and customers.
+The main advantage seen by companies was an increase of operational efficiencies, and reduction of costs.
+In this scenario, highly qualified software developers are responsible for the integration of services with other systems, through the analysis of formal service description or human-readable specifications.
+However, this model fails when targeting the long tail of enterprise software demand, the end-users.
+Discovery and consumption of web services is far from a straight-forward task for an end-user.
+This means that potential users trying to create task-specific applications are left out.
+This paper presents an approach to facilitate the discovery and consumption of business web services by end-users, closing the gap between the two.
+The approach includes: (a) a method to generate ready-to-use web service wrappers, and (b) a catalogue which users can browse to search for web services fitting their needs.
+% These web services wrappers are described by a model created, although highly influenced by well-known web services modelling standards. 
 \keywords{web services, end-users, discovery, consumption, linked data}

### 2010-06-15 trunk/papers/icsoc2010/paper_sec_catalogue.tex
last changes everywhere
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@876 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 The current state of the art provides several solutions and strategies which providers and consumers may take advantage of.
-However, as explained in previous sections, a large number of them suffer from limitated syntactic-based descriptions and a simple keyword-based search, and the gap between discovery and consumption is an obstacle for their usage by a non-technical common user.
-In order to ligthen the issue, this paper presents yet another publishing platform, permiting any enterprise or individual to publish their public domain web services, with an enhanced semantic search based on the definition of the services, and serving``web service wrappers'' for an easy consumption in any web application.
+However, as explained in previous sections, a large number of them suffer from limited syntactic-based descriptions and a simple keyword-based search, and the gap between discovery and consumption is an obstacle for their usage by a non-technical common user.
+In order to lighten the issue, this paper presents yet another publishing platform, permitting any enterprise or individual to publish their public domain web services, with an enhanced semantic search based on the definition of the services, and serving ``web service wrappers'' for an easy consumption in any web application.
 
@@
 	\item serves web services wrappers ready-to-consume in any web application;
-	\item provices advanced search capabilites, based on the service formal definition and inferences extracted from it;
+	\item provides advanced search capabilities, based on the service formal definition and inferences extracted from it;
 	\item supports managing its resources via its RESTful API;
@@
 
-This platform was intented and it is being used as part of the FAST storyboard tool, and it communicates with others components through its RESTful API, using JSON for the input of the requests.
+This platform was intended and it is being used as part of the storyboard tool of the FAST platform~\cite{hoyer2009fast}, and it communicates with other components through its RESTful API, using JSON for the input of the requests.
 This paper will focus on the functionality regarding web services publication and discovery, and the process to publish any third-party web service into the platform.
@@
 \ref{sec:wrapping_web_services}, and (ii) creation of the resource in the publishing platform.
-The following is an example of part of a request sent to create a new wrapper into the platform (for the sake of clarity, some parts of the request have been ommitted, but the main idea is shown):
+The following is an example of part of a request sent to create a new wrapper into the platform (for the sake of clarity, some parts of the request have been omitted, but the main idea is shown):
 
@@
 
-The business logic layer contains all the domain-specific processing and reasoning. It provides functions to interact with all the elements of the domain model,
-% specified in \ref{ssec:ontology}
+The business logic layer contains all the domain-specific processing and reasoning. It provides functions to interact with all the elements of the domain model specified in \ref{ssec:ontology},
 acting as a mediator between the public RESTful API and the persistence layer.
@@
 
-% \input{paper_ssec_ontology}
+\input{paper_ssec_ontology}
 

### 2010-06-15 trunk/papers/icsoc2010/paper_sec_introduction.tex
last changes everywhere
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@876 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-06-14 trunk/papers/icsoc2010/paper_sec_introduction.tex
@@
 Next, section 3 describes our platform for publishing web services and empowering end-users to discover services.
-Section 5 gives the reader a clearer use case and how this work is integrated into the FAST platform, and finally, we present our main conclusions and highlight future lines of research in Section 6.
+Section~\ref{sec:use_case} gives the reader a clearer use case and how this work is integrated into the FAST platform, and finally, we present our main conclusions and highlight future lines of research in Sect.~\ref{sec:conclusions}.

### 2010-06-15 trunk/papers/icsoc2010/paper_sec_use_case.tex
last changes everywhere
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@876 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 The intention of this paper was to show the advances made regarding web service discovery and consumption through the two artifacts developed as part of our research.
-While these artifacts may be deployed and used separately by third-party applications, they were originally developed to form the backbone of the FAST platform~\cite{hoyer2009fast}.
+While these artifacts may be deployed and used separately by third-party applications, they were originally developed to form the backbone of the FAST platform.
 FAST constitutes a novel approach to application composition and business process definition from a top-down user-centric perspective.

### 2010-06-15 trunk/papers/icsoc2010/paper_sec_wrapping.tex
some corrections
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@874 d8b0c59b-6544-0410-94d9-83a5658a058e
last changes everywhere
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@876 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
+%!TEX root = paper.tex
+
 \section{Wrapping web services}
@@
 
-To define the desired input parameters and the type of the expected result of a new service wrapper the FAST service wrapper tool provides a web based form that allows the editing of these entries, cf.
+To define the desired input parameters and the type of the expected result of a new service wrapper the service wrapper tool provides a web based form that allows the editing of these entries, cf.
 Figure \ref{fig:construct_pre_post_conditions}.
@@
 Once the service request is constructed and sent to the service provider, it will send back a response.
-This response message could be serialized in any format, though the most common formats used nowadays are XML or JSON among others.
+This response message could be serialized in any format, though the most common formats used nowadays are XML or JSON.
 To continue the example started in the previous section, the response of the Ebay Shopping service will be in XML as specified in the request.
@@
 This process helps the user to deal with the slightly complex semantics of the transformation rules avoiding errors or mistakes.
-In addition, out tool is ontology-driven, therefore, the service wrapper designer shall retrieve the domain-specific types from a FAST ontology server together with the structure of each type, i.e.
+In addition, our tool is ontology-driven, therefore the service wrapper designer retrieves the domain-specific types from a backend server (such as the catalogue discussed in Sect.~\ref{sec:discovery}) together with the structure of each type, i.e.
 together with a description of the attributes of each object.
@@
 For the \emph{fillAttributes} rules, the selection box shows the attributes of the object type chosen in the parent rule.
-In addition, we may provide some analysis tool, which will help to guarantee that the objects generated by the transformation rules conform to the object types defined in the corresponding FAST ontology.
+In addition, we may provide some analysis tool, which will help to guarantee that the objects generated by the transformation rules conform to the object types defined in the corresponding conceptual model (see Sect.~\ref{ssec:ontology}).
 This helps to ensure that the objects generated by the designed service wrapper will be compatible for input parameters of subsequent filter steps and or gadgets.

### 2010-06-15 trunk/papers/icsoc2010/paper_sec_introduction.tex
some changes
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@877 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-06-14 trunk/papers/icsoc2010/paper_sec_introduction.tex
@@
 However, as of yet they have not been widely adopted, 
-possibly because the perceived potential benefits of SWS did not justify the additional investments required for businesses. 
+possibly because the perceived potential benefits of SWS did not justify the additional investments required for businesses~\cite{shi2007}.
 In a similar vein as B2C, governments are currently opening their data to the public as linked data. 
@@
 Section 2 covers the state of the art in web services publication, discovery and wrapping/consumption.
-In Section 4 we present how we tackle with the web services wrapping, available possibilities and limitations.
-Next, section 3 describes our platform for publishing web services and empowering end-users to discover services.
-Section 5 gives the reader a clearer use case and how this work is integrated into the FAST platform, and finally, we present our main conclusions and highlight future lines of research in Section 6.
+In Section~\ref{sec:wrapping_web_services} we present how we tackle with the web services wrapping, available possibilities and limitations.
+Next, Section~\ref{sec:discovery} describes our platform for publishing web services and empowering end-users to discover services.
+Section~\ref{sec:use_case} gives the reader a clearer use case and how this work is integrated into the FAST platform, and finally, we present our main conclusions and highlight future lines of research in Section~\ref{sec:conclusions}.

### 2010-06-15 trunk/papers/icsoc2010/paper_sec_related_work.tex
some changes
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@877 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 In the context of publishing and discovering Web services, a service provider have well-known and widely used technologies to accomplish the task of publication, such as the Universal Description, Discovery and Integration (UDDI) or WS-Discovery.
-The UDDI service registry specification \cite{uddi2004} is currently one of the core standards in the Web service technology stack and an integral part of every major SOA vendor's technology strategy and offers to requesters the ability to discover services via the Internet.
+The UDDI service registry specification~\cite{uddi2004} is currently one of the core standards in the Web service technology stack and an integral part of every major SOA vendor's technology strategy and offers to requesters the ability to discover services via the Internet.
 In a few words, UDDI serves as a centralised repository of WSDL documents.
-A similar concept is iServe \cite{pedrinaci_ores2010}.
+A similar concept is iServe~\cite{pedrinaci_ores2010}.
 That platform aims to publish Web services as what they called Linked Services Linked Services ---linked data describing services---, storing web service definitions in a semantic-annotation fashion so other semantic web technologies for web service discovery and processing may take advantage of it.
@@
 
-A different approach for web service discovery is the Web Services Dynamic Discovery (WS-Discovery) specification.
+A different approach for web service discovery is the Web Services Dynamic Discovery (WS-Discovery) specification~\cite{beatty2005}.
 The core of this approach is a multicast discovery protocol.

### 2010-06-15 trunk/papers/icsoc2010/paper_abstract.tex
abstract revised
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@878 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 transforming the way enterprise systems communicate and strengthening the  relationships with providers and customers.
-The main advantage seen by companies was an increase of operational efficiencies, and reduction of costs.
-In this scenario, highly qualified software developers are responsible for the integration of services with other systems, through the analysis of formal service description or human-readable specifications.
+The main advantage seen by companies was an increase of operational efficiencies, and a reduction of costs.
+In this scenario, highly qualified software developers are responsible for the integration of services with other systems, through the analysis of formal service descriptions or human-readable specifications.
 However, this model fails when targeting the long tail of enterprise software demand, the end-users. 
-Discovery and consumption of web services is far from a straight-forward task for an end-user.
-This means that potential users trying to create task-specific applications are left out.
-This paper presents an approach to facilitate the discovery and consumption of business web services by end-users, closing the gap between the two.
-The approach includes: (a) a method to generate ready-to-use web service wrappers, and (b) a catalogue which users can browse to search for web services fitting their needs.
+Discovery and consumption of web services are difficult tasks for end-users. 
+This means that potential long tail of end-users creating task-specific applications from existing services is as of yet completely untapped. 
+This paper presents an approach to facilitate the discovery and consumption of business web services by end-users, closing the gap between the two. The approach includes: 
+\begin{inparaenum}[(a)]
+	\item a method to generate ready-to-use web service wrappers, and 
+	\item a catalogue which users can browse to search for web services fitting their needs. 
+\end{inparaenum}
 % These web services wrappers are described by a model created, although highly influenced by well-known web services modelling standards. 

### 2010-06-15 trunk/papers/icsoc2010/paper_sec_use_case.tex
new architecture picture
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@880 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 Therefore, it is aimed at allowing users (mainly business users at the management level, without programming experience) to create their own situational applications by visually combining different components, or building blocks, such as graphical forms and back-end services in terms of their inputs and outputs.
-These two components, as described in Sects.
-3 and 4, are fully integrated and present a high cohesion with the rest of the components of the FAST architecture (see Fig.~\ref{fig:fast_architecture}).
+The work covered in this paper are the components highlighted and surronded by a dashed frame in the Fig.~\ref{fig:fast_architecture}.
+These components have public interfaces to communicate with any application developed by a third-party developer or company, and as described in Sects.
+3 and 4, are fully integrated and present a high cohesion with the rest of the components of the FAST architecture.
 
@@
   \begin{center}
-    \includegraphics[width=.9\linewidth]{images/fast_architecture.png}
+    \includegraphics[width=.9\linewidth]{images/fast_architecture.pdf}
     \caption{Overview of the FAST architecture}
@@
 
-The publishing platform presented, called ``Metadata Catalogue'' within the FAST platform, covers several important purposes:
+The publishing platform presented, also called ``Catalogue'' within the FAST platform, covers several important purposes:
 \begin{description}

### 2010-06-15 trunk/papers/icsoc2010/paper_ssec_linked_data.tex
small change
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@881 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-The recently success of the web of Data has influenced the way information is now published on the web, and has been widely adopted by the academic community, large companies such as The New York Times and BBC, and national governments such as US and UK made public commitments toward open data.
+The recently success of the Web of Data has influenced the way information is now published on the web, and has been widely adopted by the academic community, large companies such as The New York Times and BBC, and national governments such as US and UK made public commitments toward open data.
 The main aim of the linking data is to find other related data based upon previously known data, in terms of the connections or links between them.
 We apply this concept in order to provide metadata about the web services as linked data, following the principles as originally defined in \cite{bizer_ijswis2009}, in order to make them available to arbitrary third-party applications.
-Each web service is identified by an HTTP URI and hosted in the catalogue so that it can be dereferenced through the same URI.
+Each web service is identified by an HTTP URI and hosted in the publishing platform so that it can be dereferenced through the same URI.
 For each building block, data is available in representations in different standard formats such as JSON (for communication with the GVS), RDF/XML, Turtle, or even HTML+RDFa as a human-readable version.

### 2010-06-15 trunk/papers/icsoc2010/paper_sec_catalogue.tex
minor change
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@882 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 The persistence layer provides an API allowing to work with a standard set of objects representing the model.
-The interaction with the underlying RDF store is made via RDF2Go, an abstraction over triple (and quad) stores, which allows to program against rdf2go interfaces and choose any the RDF store implementation.
+The interaction with the underlying RDF store is made via RDF2Go, an abstraction over triple (and quad) stores, which allows to program against rdf2go interfaces and choose any RDF store implementation.
 This allows having a completely extensible and configurable framework for storage mechanisms, inferencers, RDF file formats, query result formats and query languages.

### 2010-06-15 trunk/papers/icsoc2010/paper_sec_wrapping.tex
wrapping section updated
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@885 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-In our approach, wrapping web services is done in two steps: a first step is in charge of the construction of a service request and a second step analyzes the response got from the execution of the service, allowing the analysis of response data and the mapping to domain-specific concepts.
+In our approach, wrapping web services is done in two steps: a first step is in charge of the construction of a service request and a second step analyses the response received from the execution of the service, allowing the analysis of response data and the mapping to domain-specific concepts.
 From this input, we generate JavaScript code to be embedded into web gadgets.
-This JavaScript code takes query parameters, constructs the service request, analysis the service response and returns result data in the desired format.
+This JavaScript code takes query parameters, constructs the service request, analyses the service response and returns result data in the desired format.
 
@@
 
-To define the desired input parameters and the type of the expected result of a new service wrapper the service wrapper tool provides a web based form that allows the editing of these entries, cf.
-Figure \ref{fig:construct_pre_post_conditions}.
+To define the desired input parameters and the type of the expected result of a new service wrapper the service wrapper tool provides a web based form that allows the editing of these entries, see Fig.~\ref{fig:construct_pre_post_conditions}.
 Note, we allow to edit example values for the input parameters.
@@
 Then the placeholders for input parameters are replaced by their example values and the resulting http request is shown below the parameter form.
-In addition, the request is send and the response is shown on the bottom of that page.
+In addition, the request is sent and the response is shown on the bottom of that page.
 This gives the user a fast feedback whether the constructed request works as desired.
@@
 
-Once the service request is constructed and sent to the service provider, it will send back a response.
-This response message could be serialized in any format, though the most common formats used nowadays are XML or JSON.
-To continue the example started in the previous section, the response of the Ebay Shopping service will be in XML as specified in the request.
+The service request is constructed and sent to the service provider, who will then send back a response.
+This response message could be serialised in any format, though the most common formats used nowadays are XML or JSON.
+To continue the example started in the previous section, we assume the response of the Ebay Shopping service will be in XML as specified in the request.
 
-\subsubsection{Translation XML into Facts} % (fold)
+\subsubsection{Translation of XML into Facts} % (fold)
 \label{ssub:translation_xml_into_facts}
-
-Figure ~\ref{fig:response_service_execution} shows the data transformation tab of the wrapper tool.
 
@@
 
-Once the service response, in XML format, has been retrieved, the transformation tab shows it as an interactive object tree on the left side of Figure ~\ref{fig:response_service_execution}.
+Once the service response, in XML format, has been retrieved, the transformation tab of the wrapper tool shows it as an interactive object tree, as seen on the left side of Fig.~\ref{fig:response_service_execution}.
 To construct this interactive object tree, the XML document has been parsed into a DOM, and a simplified tree representation of that DOM is built up.
 
-A transformation rule is used to analyze the XML data and to generate domain-specific objects from concepts from the ontologies used by the parameters of the different building blocks.
-A transformation rule is composed of three elements, cf.
-the middle part of Figure ~\ref{fig:response_service_execution}.
+A transformation rule is used to analyse the XML data and to generate domain-specific objects from concepts from the ontologies used by the parameters of the different building blocks.
+A transformation rule is composed of three elements, as seen in the middle part of Fig.~\ref{fig:response_service_execution}.
 First, the \textit{from} field indicates the XML elements to be translated by the rule.
@@
 And third, the target of the rule specifies a certain concept or attribute, to be created or filled.
-A detailed explanation of the type of actions to be trigger from the transformation rules is:
+A detailed explanation of the type of actions to be trigger from the transformation rules is as follows:
 \begin{description}
@@
 Contrarily, a person fact which conforms to a certain ontology for that domain may provide only one \emph{fullname} attribute that shall be filled by a concatenation of the first and the last name.
-To achieve this, the \textit{from} field of that tranforamtion rule might look like: \texttt{lastname"', "`firstname}.
+To achieve this, the \textit{from} field of that transformation rule might look like: \texttt{lastname"', "`firstname}.
 We are also able to do some navigation in the XML tree to follow XRef elements.
@@
 To cover such  cases, intermediate object formats can be used which would allow generating objects to be further processed by additional filters.
-Such additional filters may be realized using (hand coded) operators, since some generic operators can act as filters for aggregation and conversions of objects from multiple sources.
+Such additional filters may be realised using (hand coded) operators, since some generic operators can act as filters for aggregation and conversions of objects from multiple sources.
 Then, service wrappers in combination with these filter operators will allow covering these complex cases.

### 2010-06-15 trunk/papers/icsoc2010/paper_sec_catalogue.tex
updating catalogue section
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@886 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 The current state of the art provides several solutions and strategies which providers and consumers may take advantage of.
-However, as explained in previous sections, a large number of them suffer from limited syntactic-based descriptions and a simple keyword-based search, and the gap between discovery and consumption is an obstacle for their usage by a non-technical common user.
-In order to lighten the issue, this paper presents yet another publishing platform, permitting any enterprise or individual to publish their public domain web services, with an enhanced semantic search based on the definition of the services, and serving ``web service wrappers'' for an easy consumption in any web application.
+However, as explained in previous sections, a large number of them suffer from limited syntax-based descriptions and simple keyword-based search, and the gap between discovery and consumption is an obstacle for their usage by non-technical end-users.
+In order to improve this issue, this paper presents a novel publishing platform, permitting any enterprise or individual to publish their public domain web services, with an enhanced semantic search based on the definition of the services, and serving web service wrappers for easy consumption in web applications.
 
@@
 
-The main difference of this platform with regards to the state of the art is that it is clearly targeting a different kind of user, and covers many deficiencies others tools does not tackle with.
-As a briefly overview, the platform being presented:
+The main difference of this platform with regards to the state of the art is that it is targeting a different kind of user, and covers many deficiencies others tools do not tackle.
+As a brief overview, the platform being presented:
 
 \begin{itemize}
-  \item allows functional discovery through web services pre and postconditions;
-	\item serves web services wrappers ready-to-consume in any web application;
-	\item provides advanced search capabilities, based on the service formal definition and inferences extracted from it;
+  \item allows functional discovery through web service pre- and post-conditions;
+	\item serves web services wrappers ready to consume in web applications;
+	\item provides advanced search capabilities, based on the formal service definition and inferences extracted from it;
 	\item supports managing its resources via its RESTful API;
-	\item a SPARQL endpoint to access directly to the data allowing advanced queries;
-	\item offers web services as Linked Data;
+	\item offers a SPARQL endpoint, giving direct access to the data through complex queries;
+	\item offers web service descriptions as linked data;
 	\item follows well-known best practices for publishing data on the web;
-	\item supports content negotiation so the different clients may retrieved the information in JSON, RDF/XML, RDF/N3 or a human-readable HTML version;
+	\item supports content negotiation so that different clients may retrieve information in their preferred format, choosing from JSON, RDF/XML, RDF/N3 or a human-readable HTML version.
 \end{itemize}
@@
 This platform was intended and it is being used as part of the storyboard tool of the FAST platform~\cite{hoyer2009fast}, and it communicates with other components through its RESTful API, using JSON for the input of the requests.
-This paper will focus on the functionality regarding web services publication and discovery, and the process to publish any third-party web service into the platform.
-This process requires two steps: (i) generation of the wrapper (see Sec.
-\ref{sec:wrapping_web_services}, and (ii) creation of the resource in the publishing platform.
+This paper will focus on the functionality regarding web service publication and discovery, and the process to publish any third-party web service into the platform.
+This process requires two steps:
+\begin{inparaenum}[(i)]
+	\item generation of the wrapper (see Sect. \ref{sec:wrapping_web_services}), and 
+	\item creation of the resource in the publishing platform. 
+\end{inparaenum}
 The following is an example of part of a request sent to create a new wrapper into the platform (for the sake of clarity, some parts of the request have been omitted, but the main idea is shown):
@@
 
-The enriched search capabilities are supported by the definition of pre and postcondition.
-It allows to define the inputs and outputs of the services using concepts from any ontology, and find web services which match and hence is able to communicate with other web services or building block.
-The concepts of pre/postconditions were highly influenced by WSMO \cite{roman2005}, make it simpler and implemented in RDFS for a better performance.
-The model used to define the web service wrappers and the ontology created are detailed in Section~\ref{ssec:ontology}.
+The enriched search capabilities are supported by the definition of pre- and post-conditions.
+They allows to define the inputs and outputs of the services and other building blocks using concepts from any ontology, and in this way to find web services or other building blocks which match each other and can therefore be integrated.
+The concepts of pre-/post-conditions were strongly influenced by WSMO~\cite{roman2005}, simplified and implemented in RDFS for better live performance.
+The model used to define the web service wrappers and the ontology created are described in more detail in Sect.~\ref{ssec:ontology}.
 

### 2010-06-15 trunk/papers/icsoc2010/paper_sec_catalogue.tex
updating catalogue section
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@887 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-The architecture comprises a RDF store used for persistence, a business layer dealing with the model and reasoning, and public facade providing the core functionality as a RESTful API, and a SPARQL endpoint accessing directly the RDF store.
-This presentation layer is aimed to interact with the FAST Gadget Visual Storyboard, or any other third-party application.
+The architecture comprises an RDF store used for persistence, a business layer dealing with the model and reasoning, and a public facade providing the core functionality as a RESTful API, as well as a SPARQL endpoint accessing the RDF store directly.
+This presentation layer is aimed to interact with the FAST Gadget Visual Storyboard (see Sect.~\ref{sec:use_case}), or any other third-party application.
 
 The RESTful API and the SPARQL endpoint are part of the presentation layer.
-The SPARQL endpoint is offered using the SPARQL protocol service as defined in the SPROT specification \cite{sprot} and is aimed to enable third-party developers to query directly the knowledge base using SPARQL queries.
+The SPARQL endpoint is offered using the SPARQL protocol service as defined in the SPROT specification~\cite{sprot} and is aimed to enable third-party developers to directly query the knowledge base using SPARQL queries.
 This feature is supported by the Sesame RESTful HTTP interface for SPARQL Protocol for RDF.
@@
 The persistence layer provides an API allowing to work with a standard set of objects representing the model.
-The interaction with the underlying RDF store is made via RDF2Go, an abstraction over triple (and quad) stores, which allows to program against rdf2go interfaces and choose any RDF store implementation.
+The interaction with the underlying RDF store is made via RDF2Go, an abstraction over triple (and quad) stores, which allows to program against RDF2Go interfaces and choose any RDF store implementation.
 This allows having a completely extensible and configurable framework for storage mechanisms, inferencers, RDF file formats, query result formats and query languages.

### 2010-06-15 trunk/papers/icsoc2010/paper_ssec_linked_data.tex
updating catalogue section
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@887 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
+%!TEX root = paper.tex
+
 \subsection{Serving Linked Data} % (fold)
@@
 
-The recently success of the Web of Data has influenced the way information is now published on the web, and has been widely adopted by the academic community, large companies such as The New York Times and BBC, and national governments such as US and UK made public commitments toward open data.
+The recently success of the Web of Data has influenced the way information is now published on the web, and has been widely adopted by the academic community.
+Also large companies such as The New York Times and BBC, and national governments such as US and UK made public commitments toward open data.
 The main aim of the linking data is to find other related data based upon previously known data, in terms of the connections or links between them.
-We apply this concept in order to provide metadata about the web services as linked data, following the principles as originally defined in \cite{bizer_ijswis2009}, in order to make them available to arbitrary third-party applications.
+We apply this concept in order to provide metadata about the web services as linked data, following the principles as defined in~\cite{bizer_ijswis2009}, in order to make them available to arbitrary third-party applications.
 Each web service is identified by an HTTP URI and hosted in the publishing platform so that it can be dereferenced through the same URI.
-For each building block, data is available in representations in different standard formats such as JSON (for communication with the GVS), RDF/XML, Turtle, or even HTML+RDFa as a human-readable version.
-The principles and best practices proposed in \cite{berrueta2008} and \cite{sauermann2008cool_uris} are also taken into account, these representations are served based on the request issued by the requesting agent, using a technique called \emph{content negotiation}.
-As required by the forth rule of linked data, individual building blocks link to other data on the web, thereby preventing so-called isolated ``data islands''.
+For each building block, data is available in representations in different standard formats such as JSON (for communication with the applications such as the widget designer shown in Fig.~\ref{fig:fast_architecture}), RDF/XML, Turtle, or even HTML+RDFa as a human-readable version.
+The principles and best practices proposed in \cite{berrueta2008} and \cite{sauermann2008cool_uris} are also taken into account, so that these representations are served based on the request issued by the requesting agent, using a technique called \emph{content negotiation}.
+As suggested by linked data principles, individual building blocks link to other data on the web, thereby preventing so-called isolated ``data islands''.
 

### 2010-09-20 trunk/papers/webist2011/authors.tex
adapted to WEBIST format
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@980 d8b0c59b-6544-0410-94d9-83a5658a058e

\author{{\textmd {\textbf }}\\
  \fontsize{11}{13}\selectfont
  Ismael Rivera, Knud M\"oller\\
  \fontsize{9}{11}\selectfont
  \textit{DERI, National University of Ireland, Galway}\\
  \fontsize{9}{11}\selectfont
  \textit{ismael.rivera@deri.org, knud.moeller@deri.org}\\[12pt]
  %
  \fontsize{11}{13}\selectfont
  Albert Z\"undorf\\
  \fontsize{9}{11}\selectfont
  \textit{University of Kassel, Germany}\\
  \fontsize{9}{11}\selectfont
  \textit{zuendorf@cs.uni-kassel.de}\\[12pt]
  %
  % repeat previous group of commands here for further authors
}

### 2010-09-20 trunk/papers/webist2011/paper_sec_introduction.tex
adapted to WEBIST format
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@980 d8b0c59b-6544-0410-94d9-83a5658a058e

%!TEX root = paper.tex

\section{Introduction}
\label{sec:introduction}

Business-to-business (B2B) integration is still a significant challenge, often requiring extensive efforts in terms of different aspects and technologies for protocols, architectures or security. Solutions based on open standards have reduced the complexity of integrating different business applications between different companies and partners. 
In this sense, Web services and XML standards such as RosettaNet, ebXML or OAG have been a boon to the world of B2B. By web services standards we are referring to the following open standards: Web Services Description Language (WSDL --- to describe), Universal Description, Discovery and Integration (UDDI --- to advertise and syndicate), Simple Object Access Protocol (SOAP --- to communicate) and Web Services Flow Language (WSFL --- to define work flows). 
A common scenario  in their usage is a company using WSDL to describe its web services, publishing them through a repository using UDDI, while these web services use SOAP-based messages to achieve dynamic integration between different, disparate applications.

While the adoption of these standards has offered some advantages in business-to-consumer (B2C) integration as well,  the interaction and consumption of web services still requires programming skills and a deep understanding of the  technology, which poses an obstacle for an end-user with limited knowledge on the matter (in the sense as defined in~\cite{fuchsloch2010}). 
Regarding the discovery of the right service for the right task --- a crucial requirement for the dynamic use of WS --- most publishing platforms are syntax-based, %leading to poor precision and recall in finding the correct service, 
making it difficult to navigate through a large number of web services~\cite{pilioura_acm2009}, 
therefore preventing end-users from performing these tasks. 
Solutions such as Semantic Web Services (SWS) promised many advantages in discovery, composition and consumption of web services, independently of the provider's platform, location, service implementation or data format. 
However, as of yet they have not been widely adopted, 
possibly because the perceived potential benefits of SWS did not justify the additional investments required for businesses~\cite{shi2007}.
In a similar vein as B2C, governments are currently opening their data to the public as linked data. 
So far, the efforts are restricted to data only, so that users are left alone in finding applications of that data. 
A natural further development in the public sector would be that their services for administration tasks and single-window applications will be opened as well. Hence, governments will benefit from the research done around these concepts and technologies.

The motivation of our work has been strongly influenced by the end-users' needs. We are targeting users which are non-skilled in term of programming and software development, empowering them with a platform to discover and consume web services in a straight-forward manner.
The web services and their specifications are not published right away in the platform, since this would simply mean to reimplement UDDI.
However, their definition and behaviour are wrapped in what we call a ``web service wrapper''. The products resulting from the wrapping process are two artifacts: a specific definition of the web service to be used by this tool, and a ready-to-consume (inside a web browser) piece of code with the proper functions to invoke the web service.

The remainder of this paper is organised as follows. Section~\ref{sec:related_work} covers the state of the art in web services publication, discovery and wrapping/consumption. In Section~\ref{sec:wrapping_web_services} we present our approach for web services wrapping, available possibilities and limitations. Next, Sect.~\ref{sec:discovery} describes our platform for publishing web services and empowering end-users to discover services. Section~\ref{sec:use_case} gives the reader a clearer use case and how this work is integrated into the FAST platform, and finally, we present our main conclusions and highlight future lines of research in Section~\ref{sec:conclusions}.

### 2010-09-20 trunk/papers/webist2011/paper_sec_related_work.tex
adapted to WEBIST format
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@980 d8b0c59b-6544-0410-94d9-83a5658a058e

%!TEX root = paper.tex

\section{Related work}
\label{sec:related_work}

Web services have been around for a long time since they first appeared.
One of the most important claims about their benefits has been (syntactic) interoperability between third-party systems and applications based on different platforms / programming languages. 
System integration, within a company or between systems from different enterprises, became easier with the adoption of web service standard technologies, in particular the Web Service Description Language (WSDL), the most extensive standard used for the definition of web services. 
Many integrated development environments (IDEs) can deal with WSDL to facilitate the integration task, however it is still required that developers read and understand their specifications, and implement part of the logic programmatically. 
As a step forward, there are several tools which facilitate the interaction with data sources and services on the Web. Yahoo! Pipes, Apatar, JackBe Presto, Microsoft Popfly (now defunct) and NetVibes, among others provide a set of modules to access different kind of data sources, such as RSS feeds, a given web page (HTML code), Flickr images, Google base or the Yahoo! search engine, databases (MySQL, PostgreSql, Oracle), and powerful enterprise systems such as Salesforce CRM, SugarCRM and Goldmine CRM. 
However, none of the solutions in the market really facilitates end-users to build  their own applications or widgets allowing the interaction with web services created by third-party providers. The solutions found are mainly data-oriented (RSS feeds, databases, raw text). Several tools permits some sort of (web) service integration, but are meant to be used within an enterprise level by savvy business users or developers. The solution presented in this paper leverage the possibility of integrating REST or SOAP-based web service inside visual applications running in a browser (e.g. widgets) by providing a platform to create, publish and discover web services wrappers.

In the context of publishing and discovering Web services, service providers have well-known and widely used technologies to accomplish the task of publication, such as the Universal Description, Discovery and Integration (UDDI) or WS-Discovery. The UDDI service registry specification~\cite{uddi2004} is currently one of the core standards in the Web service technology stack and an integral part of every major SOA vendor's technology strategy and offers to requesters the ability to discover services via the Internet. 
In short, UDDI serves as a centralised repository of WSDL documents. A similar concept is iServe~\cite{pedrinaci_ores2010}. This platform aims to publish web services as what they called Linked Services --- linked data describing services ---, storing web service definitions as semantic annotations, so that other semantic web-aware applications may take advantage of it. However, the platform does not deal with the step from the definition to the consumption of the services.

A different approach for web service discovery is the Web Services Dynamic Discovery (WS-Discovery) specification~\cite{beatty2005}. The core of this approach is a multicast discovery protocol. Service providers and consumers listen to each other for new services specifications within a network, so there is no need of a centralised registry. As a drawback, WS-Discovery do not support Internet-scale discovery, making it useless for our purposes.

### 2010-09-20 trunk/papers/webist2011/paper_ssec_ontology.tex
adapted to WEBIST format
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@980 d8b0c59b-6544-0410-94d9-83a5658a058e

%!TEX root = paper.tex

\subsection{Conceptual Model}
\label{ssec:ontology}
The conceptual model used to define the web service wrappers within this application has been influenced by WSDL, the main web service definition language, and semantic approaches such as WSMO and OWL-S. It belongs to a more complex conceptual model for FAST, which can be grouped into three levels: the gadget and screenflow level at the top, the level of individual screens in the middle, and the level of web services at the bottom. 
In short, a screenflow is an executable (in a Web browser) set of screens, while a screen is a set of building blocks with a specific purpose. These building blocks --- \emph{forms} for the UI, \emph{operators} for data manipulation and \emph{back-end service} --- all share the same structure. Every building block has a set of actions (operations in WSDL), each of which needs a set of pre-conditions (inputs) to be fulfilled in order to be executed, and provides a set of post-conditions (outputs) to other building blocks. These pre- and post-conditions are defined as individual RDF triples. This means that conditions can be modelled as graph patterns using SPARQL notation \cite{sparql2008spec}, where e.g. the post-condition of a login service such as ``there exists a user'' will be transformed into a simple pattern like ``a variable \texttt{?user} is of type \texttt{sioc:User}'', or ``\texttt{?user a sioc:User}''. Using this mechanism, extended by RDFS entailment rules, services and other building blocks can be matched automatically. The publishing and discovery platform (catalogue) can employ this functionality to support the discovery of web services based on the current user needs (expressed in the same way as pre-conditions). For further details on the conceptual model, see its definition in \cite{moeller2010fast_ontology}.

### 2010-09-20 trunk/papers/webist2011/abstract.tex
added abstract
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@981 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-06-15 trunk/papers/icsoc2010/paper_abstract.tex
@@
 
-\begin{abstract}
 B2B systems integration was revolutionised by the introduction of web services,
@@
 \end{inparaenum}
-% These web services wrappers are described by a model created, although highly influenced by well-known web services modelling standards. 
-\keywords{web services, end-users, discovery, consumption, linked data}
-\end{abstract}

### 2010-09-27 trunk/papers/webist2011/paper_ssec_ontology.tex
%!TEX root = paper.tex
\subsection{Conceptual Model}
\label{ssec:ontology}
The conceptual model used to define the web service wrappers within this application has been influenced by WSDL, the main web service definition language, and semantic approaches such as WSMO and OWL-S. It belongs to a more complex conceptual model for FAST, which can be grouped into three levels: the gadget and screenflow level at the top, the level of individual screens in the middle, and the level of web services at the bottom. 
In short, a screenflow is an executable (in a Web browser) set of screens, while a scree

@@
 In short, a screenflow is an executable (in a Web browser) set of screens, while a screen is a set of building blocks with a specific purpose.
-These building blocks --- \emph{forms} for the UI, \emph{operators} for data manipulation and \emph{back-end service} --- all share the same structure.
+These building blocks --- \emph{forms} for the UI, \emph{operators} for data manipulation and \emph{back-end services} --- all share the same structure.
 Every building block has a set of actions (operations in WSDL), each of which needs a set of pre-conditions (inputs) to be fulfilled in order to be executed, and provides a set of post-conditions (outputs) to other building blocks.

### 2010-09-28 trunk/papers/webist2011/paper_ssec_discovery.tex
added discovery section
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@983 d8b0c59b-6544-0410-94d9-83a5658a058e

%!TEX root = paper.tex

\subsection{Discovery mechanisms} % (fold)
\label{ssec:discovery}

The platform presented in this paper is intended to support web service discovery and recommendation capabilities.
As explained in detail in Section~\ref{ssec:ontology}, there are different types of building block which can be modelled
using the platform, leading to a necessity for different kind of discovery or recommendation mechanisms, depending
on the level of composition to deal with: screen-flow or screen. In the screen-flow composition, the smallest
functional unit is the screen, hance a screen-flow is composed by interconnecting a number of screens between 
each other. In fact, this composition can be applied to any sort of building block since they all share the
same structure: a set of pre- and post-conditions which make possible the communication with other building blocks.
That said, technically the algorithms presented in this section could be used for the discovery and composition of
any type of building block stored in the platform.

At the screen-flow design level, the composition is made using a set of screens, and the pre- and post-conditions 
of the screen-flow itself, used as entry and exit points. A screen is only reachable when its pre-conditions are satisfied
as well, meaning that the post-conditions of another screen, or the pre-conditions of the screen-flow are compatible 
with them.

The main goal of the discovery process is to help the user to make the screen-flow executable, in other words, to make
all the screens within a screen-flow reachable. To do that, the platform offers two mechanisms to find and recommend
screens, previously stored in the building block base (the catalogue), which will make the screen-flow executable: a
simple discovery based on pre- and postconditions, and a multi-step discovery or planning mechanism. Like any other
search engine or recommender system, the results obtained in the discovery process need to rank before being presented
to the user. The ranking algorithm is covered in Section~\ref{sssec:ranking}.

\subsubsection{Simple discovery based on pre- and post-conditions}
\label{sssec:simple_discovery}

At a particular state of a composition, ie. at screen-flow design, several of the building blocks used to compose it, 
screens in the case of a screen-flow, might be unreachable, therefore there must be pre-conditions not satisfied. The
platform will assist the process by recommending building blocks which will satisfy these pre-conditions. In a nutshell,
it collects the pre-conditions of all the unreachable building blocks, used as a graph pattern (see Section~\ref{ssec:ontology})
in order to be matched against the post-conditions of any building block. In the following scenario, there are two 
screens: \emph{s1} and \emph{s2}. \emph{s1} has as a pre-condition: ``there exists a search criteria'', and as a 
post-condition: ``there exists a item''. \emph{s2} has just a pre-condition stating: ``there exists a search criteria''.
The same example formally defined:

\begin{listing}
\begin{verbatim}
:G1 { :s1 a fgo:Screen .      
      :s1 fgo:hasPrecondition c1 .
      :s1 fgo:hasPostcondition c2 .
      :c1 fgo:hasPattern GC1 .
      :c2 fgo:hasPattern GC2 .
      :s2 a fgo:Screen .      
      :s2 fgo:hasPrecondition c3 .
      :c3 fgo:hasPattern GC3 }
:GC1 { _:x a amazon:SearchCriteria }
:GC2 { _:x a amazon:Item }
:GC3 { _:x a amazon:SearchCriteria }
\end{verbatim}
\label{lis:discovery_example}
\end{listing}

The algorithm will construct a SPARQL query to retrieve building blocks satisfying the pre-conditions \emph{c1} and \emph{c3}. 
It would look something like:

\begin{listing}
\begin{verbatim}
PREFIX rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#>
PREFIX fgo: <http://purl.oclc.org/fast/ontology/gadget#>
PREFIX amazon:<http://aws.amazon.com/AWSECommerceService#>

SELECT DISTINCT ?bb 
WHERE { 
  ?bb rdf:type fgo:Screen . 
  {
    {
      ?bb fgo:hasPostCondition ?c .
      ?c fgo:hasPattern ?p .
      GRAPH ?p { ?x rdf:type amazon:SearchCriteria } 
    }
    UNION
    {
      ?bb fgo:hasPostCondition ?c .
      ?c fgo:hasPattern ?p .
      GRAPH ?p { ?x rdf:type amazon:Item } 
    }
  }
  FILTER (?bb != <http://fast.org/screens/S1>) 
  FILTER (?bb != <http://fast.org/screens/S2>) 
}
\end{verbatim}
\label{lis:sparql_find}
\end{listing}

However, you might have realised that the pre-condition \emph{c3} would be satisfied whether the screen \emph{s1} were 
reachable, hence it will removed those pre-conditions that would be satisfied by the current elements being used in
the composition. The removal of unnecessary pre-conditions improves the performance of the algorithm, and it leads to find
better results for the user, since the focus is given to make satisfy the pre-conditions which could not be satisfied
by any of the elements of the current composition.

\subsubsection{Enhanced discovery: search tree planning}
\label{sssec:planning}

In artiﬁcial intelligence, the term \emph{planning} originally meant a search for a sequence of logical operators or actions
that transform an initial world state into a desired goal state. Presently, it also includes many decision-theoretic ideas,
imperfect state information, and game-theoretic equilibria.

This paper applies the concept of planning to the design of building blocks. Back to the screen-flow design, the goal is to
make the screen-flow executable, the initial state is a certain screen which is not yet reachable, and the plans are sets of 
screens which are reachable by themselves, and accomplish the goal. The algorithm has been influenced by the ideas behing the 
heuristic search. For a certain state, i.e. the initial state which contains the screen to make reachable, a large tree of 
possible continuations is considered, in fact any screen would fit. Those screens which post-conditions do not satisfy in any
form the unsatisfied pre-conditions are discarded, reducing the branches of the tree. A branch stops growing when a screens is
already reachable (i.e. it has no pre-conditions) becoming a leaf of the tree. Once there are not screens added in a certain
step, the algorithm stops and discards the incomplete branches, in other words, those branches where the leaf is not reachable.

It is worth pointing out some of the tree structure is pre-computed beforehand. Any time a building block is inserted into the
catalogue, the algorithm is executed following two approaches: \emph{forward search} and \emph{backward search}. The forward
search approach finds the building blocks whose pre-conditions will be satisfied by the post-conditions of the new building 
block while the backward search finds the building blocks whose post-conditions will satisfy the pre-conditions of the recently
created building block. This data is stored and used for the tree or plan builder algorithm, without needing to chec the
compatibility of pre- and post-conditions for every single building block at every request, increasing highly the performance.

\subsubsection{Results ranking}
\label{sssec:ranking}

In the previous sections there have been presented two ways of building block discovery for the composition or design of new
building blocks. The ensure compatibility based on the functional specification of the building blocks, however both are leaving
out the measurement of quality of the results. In other words, it can be seen as a search engine, returning a set of results
for a certain input, and as any other single search engine, a ranking mechanism is needed.

The simple discovery algorithm, in order to rank the results, applies the following rules:
\begin{enumerate}
 \item it gives a higher position to those building blocks which satisfy the highest number of pre-conditions;
 \item it prioritise building blocks created by the same user who is querying;
 \item it adjust the rank by using the ratings given to the building blocks, and their popularity in terms of usage statistics;
 \item it weights the results according to non-functional features such as availability. This is only applied for what we call
``web service wrapper'', and it is calculated periodically by invoking the wrapped web services generating an up-time rate.
\end{enumerate}

For the planning case, the objective is not only to produce a plan but also to satisfy user-specified preferences, or what is
known as \emph{preference-based planning}. The ranking algorithm:
\begin{enumerate}
 \item minimizes the size of the plans, after removing the elements of the plan which are already in the canvas, so it gives
priority to the elements the user has already inserted,
 \item adjust the rank by using the rules 2, 3 and 4 used in the ranking algorithm of simple discovery.
\end{enumerate}

### 2010-10-23 trunk/papers/icsoc2010/paper_sec_wrapping.tex
some changes addressing other kinds of services
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1023 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-As a first approach, the interaction with services is limited to simple GET requests as supported e.g.
+In this paper we illustrate the interaction with services on simple GET requests as supported e.g.
 by REST services.
-Support for other services like SOAP services is planned.
-A service request is thus assembled using a certain URL and a set of parameters.
-As an example, the Ebay Shopping web service will be studied.
-The following URL is invoked to retrieve a list of items corresponding to the search keywords \emph{USB}:
+Support for other services like POST based services is under construction.
+In case of a simple GET request, a service request is assembled using a certain URL and a set of parameters.
+(A POST request would have a text block to be posted.
+In addition, other header data may be needed to transfer e.g.
+session ids.) As an example, the Ebay Shopping web service will be studied.
+As documented on the corresponding web site, the following URL is invoked to retrieve a list of items corresponding to the search keywords \emph{USB}:
 

### 2010-10-25 trunk/papers/webist2011/paper_sec_catalogue.tex
%!TEX root = paper.tex
\subsection{Conceptual Model}
\label{ssec:ontology}
The conceptual model used to define the web service wrappers within this application has been influenced by WSDL, the main web service definition language, and semantic approaches such as WSMO and OWL-S. It belongs to a more complex conceptual model for FAST, which can be grouped into three levels: the gadget and screenflow level at the top, the level of individual screens in the middle, and the level of web services at the bottom. 
In short, a screenflow is an executable (in a Web browser) set of screens, while a scree

--- 2010-06-15 trunk/papers/icsoc2010/paper_sec_catalogue.tex
@@
 
-The area of web service publication and discovery has been subject for a lot of research since the very beginning the concept was coined.
-The current state of the art provides several solutions and strategies which providers and consumers may take advantage of.
-However, as explained in previous sections, a large number of them suffer from limited syntax-based descriptions and simple keyword-based search, and the gap between discovery and consumption is an obstacle for their usage by non-technical end-users.
-In order to improve this issue, this paper presents a novel publishing platform, permitting any enterprise or individual to publish their public domain web services, with an enhanced semantic search based on the definition of the services, and serving web service wrappers for easy consumption in web applications.
+The area of web service publication and discovery has been subject for a lot of research since the very beginning the 
+concept was coined. The current state of the art provides several solutions and strategies which providers and 
+consumers may take advantage of. However, as explained in previous sections, a large number of them suffer from 
+limited syntax-based descriptions and simple keyword-based search, while other more complex approach did not success
+to be widely spread regarding the few benefits offered compared to their drawbacks. Moreover, the gap between discovery 
+and consumption is an obstacle for their usage by non-technical end-users. In order to improve this issue, 
+this paper presents a novel publishing platform, permitting any enterprise or individual to publish their public 
+domain web services, with an enhanced semantic search based on the definition of the services, and serving web 
+service wrappers for easy consumption in web applications.
 
+% Overview subsection
 \subsection{Overview}
@@
 \begin{inparaenum}[(i)]
-	\item generation of the wrapper (see Sect. \ref{sec:wrapping_web_services}), and 
+	\item generation of the wrapper (see Section \ref{sec:wrapping_web_services}), and 
 	\item creation of the resource in the publishing platform. 
@@
 {
-  "code": "http://demo.fast.morfeo-project.org/code/amazonSearch.js",
+  "code": "http://fast.morfeo-project.org/amazon-search.js",
   "creationDate": "2010-01-26T17:01:13+0000",
@@
         http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                  http://aws.amazon.com/AWSECommerceService#Item",
-      "positive": true
+        http://aws.amazon.com/AWSECommerceService#Item"
   }]],
@@
 The enriched search capabilities are supported by the definition of pre- and post-conditions.
-They allows to define the inputs and outputs of the services and other building blocks using concepts from any ontology, and in this way to find web services or other building blocks which match each other and can therefore be integrated.
+They allow to define the inputs and outputs of the services and other building blocks using concepts from any ontology, and in this way to find web services or other building blocks which match each other and can therefore be integrated.
 The concepts of pre-/post-conditions were strongly influenced by WSMO~\cite{roman2005}, simplified and implemented in RDFS for better live performance.
-The model used to define the web service wrappers and the ontology created are described in more detail in Sect.~\ref{ssec:ontology}.
+The model used to define the web service wrappers and the ontology created are described in more detail in Section~\ref{ssec:ontology}.
 
+% Architecture subsection
 \subsection{Platform architecture}
@@
 The architecture comprises an RDF store used for persistence, a business layer dealing with the model and reasoning, and a public facade providing the core functionality as a RESTful API, as well as a SPARQL endpoint accessing the RDF store directly.
-This presentation layer is aimed to interact with the FAST Gadget Visual Storyboard (see Sect.~\ref{sec:use_case}), or any other third-party application.
+This presentation layer is aimed to interact with the FAST Gadget Visual Storyboard (see Section~\ref{sec:use_case}), or any other third-party application.
 
@@
 
+% Conceptual model subsection
 \input{paper_ssec_ontology}
 
+% Discovery and recommendations subsection
+\input{paper_ssec_discovery}
+
+% Conceptual model subsection
 \input{paper_ssec_linked_data}

### 2010-10-25 trunk/papers/webist2011/paper_ssec_discovery.tex
added some changes changes
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1033 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 The platform presented in this paper is intended to support web service discovery and recommendation capabilities.
-As explained in detail in Section~\ref{ssec:ontology}, there are different types of building block which can be modelled
-using the platform, leading to a necessity for different kind of discovery or recommendation mechanisms, depending
+As explained in detail in Section~\ref{ssec:ontology}, there are different types of building blocks which can be modelled
+using the platform, leading to a necessity for different kinds of discovery or recommendation mechanisms, depending
 on the level of composition to deal with: screen-flow or screen. In the screen-flow composition, the smallest
-functional unit is the screen, hance a screen-flow is composed by interconnecting a number of screens between 
-each other. In fact, this composition can be applied to any sort of building block since they all share the
-same structure: a set of pre- and post-conditions which make possible the communication with other building blocks.
+functional unit is the screen, meaning that a screen-flow is composed by interconnecting a number of screens between 
+each other. This composition can be applied to any sort of building block since they all share the
+same structure: a set of pre- and post-conditions, which make possible the communication with other building blocks.
 That said, technically the algorithms presented in this section could be used for the discovery and composition of
@@
 At the screen-flow design level, the composition is made using a set of screens, and the pre- and post-conditions 
-of the screen-flow itself, used as entry and exit points. A screen is only reachable when its pre-conditions are satisfied
-as well, meaning that the post-conditions of another screen, or the pre-conditions of the screen-flow are compatible 
-with them.
+of the screen-flow itself used as entry and exit points of the screen-flow. A screen is only reachable when its 
+pre-conditions are satisfied as well, meaning that the post-conditions of another screen, or the pre-conditions of 
+the screen-flow are compatible with them.
 
 The main goal of the discovery process is to help the user to make the screen-flow executable, in other words, to make
-all the screens within a screen-flow reachable. To do that, the platform offers two mechanisms to find and recommend
+all the screens within a screen-flow reachable. The platform offers two mechanisms to find and recommend
 screens, previously stored in the building block base (the catalogue), which will make the screen-flow executable: a
-simple discovery based on pre- and postconditions, and a multi-step discovery or planning mechanism. Like any other
-search engine or recommender system, the results obtained in the discovery process need to rank before being presented
+simple discovery based on pre- and postconditions, and a multi-step discovery or planning algorithm. Like any other
+search engine or recommender system, the results are ranked before being presented
 to the user. The ranking algorithm is covered in Section~\ref{sssec:ranking}.
@@
 post-condition: ``there exists a item''. \emph{s2} has just a pre-condition stating: ``there exists a search criteria''.
-The same example formally defined:
+Here you can see the same example but formally defined:
 
@@
 The algorithm will construct a SPARQL query to retrieve building blocks satisfying the pre-conditions \emph{c1} and \emph{c3}. 
-It would look something like:
+The query, although simplified for the sake of clarity, would look something like:
 
@@
 
-However, you might have realised that the pre-condition \emph{c3} would be satisfied whether the screen \emph{s1} were 
-reachable, hence it will removed those pre-conditions that would be satisfied by the current elements being used in
-the composition. The removal of unnecessary pre-conditions improves the performance of the algorithm, and it leads to find
-better results for the user, since the focus is given to make satisfy the pre-conditions which could not be satisfied
-by any of the elements of the current composition.
+However, in the example presented, it is straight-forward to spot that the pre-condition \emph{c3} would be
+satisfied whether the screen \emph{s1} was reachable, being unnecessary to query the recommender algorithm 
+to retrieve screens satisfying this pre-condition, hence these pre-conditions are removed for the construction
+of the query reducing the complexity of the problem, hence improving the overall performance of the algorithm.
+Moreover, it leads to find better results for the user, since the focus is given to make satisfy the 
+pre-conditions which could not be satisfied by any of the elements of the current composition.
 
@@
 
-This paper applies the concept of planning to the design of building blocks. Back to the screen-flow design, the goal is to
-make the screen-flow executable, the initial state is a certain screen which is not yet reachable, and the plans are sets of 
-screens which are reachable by themselves, and accomplish the goal. The algorithm has been influenced by the ideas behing the 
-heuristic search. For a certain state, i.e. the initial state which contains the screen to make reachable, a large tree of 
-possible continuations is considered, in fact any screen would fit. Those screens which post-conditions do not satisfy in any
-form the unsatisfied pre-conditions are discarded, reducing the branches of the tree. A branch stops growing when a screens is
-already reachable (i.e. it has no pre-conditions) becoming a leaf of the tree. Once there are not screens added in a certain
-step, the algorithm stops and discards the incomplete branches, in other words, those branches where the leaf is not reachable.
+This paper applies the concept of planning to the design of building blocks. Back to the screen-flow design, the goal
+would be to make the screen-flow executable; the initial state would be a certain screen which is not yet reachable, 
+and the plans would be sets of screens which are reachable by themselves, and accomplish the goal. 
+The algorithm has been influenced by the ideas behing the heuristic search. For a certain state, i.e. the initial 
+state which contains the screen to make reachable, a large tree of possible continuations is considered, in 
+fact any screen would fit. Those screens which post-conditions do not satisfy in any form the unsatisfied 
+pre-conditions are discarded, reducing the branches of the tree. A branch stops growing when a screens is
+already reachable (i.e. it has no pre-conditions) becoming a leaf of the tree. Once there are not screens 
+added in a certain step, the algorithm stops and discards the incomplete branches, in other words, 
+those branches where the leaf is not reachable.
 
-It is worth pointing out some of the tree structure is pre-computed beforehand. Any time a building block is inserted into the
-catalogue, the algorithm is executed following two approaches: \emph{forward search} and \emph{backward search}. The forward
-search approach finds the building blocks whose pre-conditions will be satisfied by the post-conditions of the new building 
-block while the backward search finds the building blocks whose post-conditions will satisfy the pre-conditions of the recently
-created building block. This data is stored and used for the tree or plan builder algorithm, without needing to chec the
-compatibility of pre- and post-conditions for every single building block at every request, increasing highly the performance.
+It is worth pointing out some of the tree structure is pre-computed beforehand to speed up the querying process at runtime. 
+Any time a building block is inserted into the catalogue, the algorithm is executed following two approaches: 
+\emph{forward search} and \emph{backward search}. The forward search approach finds the building blocks whose 
+pre-conditions will be satisfied by the post-conditions of the new building block while the backward search 
+finds the building blocks whose post-conditions will satisfy the pre-conditions of the recently
+created building block. This data is stored and used for the tree or plan builder algorithm, without
+needing to check the compatibility of pre- and post-conditions for every single building block at every 
+request.
 
@@
 
-In the previous sections there have been presented two ways of building block discovery for the composition or design of new
-building blocks. The ensure compatibility based on the functional specification of the building blocks, however both are leaving
-out the measurement of quality of the results. In other words, it can be seen as a search engine, returning a set of results
-for a certain input, and as any other single search engine, a ranking mechanism is needed.
+Previous sections presented different mechanisms to search or discover building blocks which may satisfy the user needs.
+The results ensure compatibility based on the functional specification of the building blocks, however they are not
+ranked, and their quality is not measured. This section explains the ranking techniques applied for the different
+discovery mechanisms.
 
-The simple discovery algorithm, in order to rank the results, applies the following rules:
+The ranking algorithm for~\ref{sssec:simple_discovery} applies the following rules:
 \begin{enumerate}
@@
 priority to the elements the user has already inserted,
- \item adjust the rank by using the rules 2, 3 and 4 used in the ranking algorithm of simple discovery.
+ \item adjust the rank by using the rules 2, 3 and 4 used from the ranking algorithm of simple discovery based on 
+pre- and post-conditions.
 \end{enumerate}

### 2010-10-25 trunk/papers/webist2011/paper_sec_wrapping.tex
changes made by Albert
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1024 d8b0c59b-6544-0410-94d9-83a5658a058e
added some changes changes
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1034 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-10-23 trunk/papers/icsoc2010/paper_sec_wrapping.tex
@@
 \begin{verbatim}
-http://open.api.sandbox.ebay.com/shopping?appid=KasselUn-efea-4b93-9505-5dc2ef1ceecd
-  &version=517&callname=FindItems&ItemSort=EndTime&QueryKeywords=USB&responseencoding=XML
+http://open.api.sandbox.ebay.com/shopping?
+  appid=KasselUn-efea-4b93-9505-5dc2ef1ceecd&
+  version=517&callname=FindItems&
+  ItemSort=EndTime&QueryKeywords=USB&
+  responseencoding=XML
 \end{verbatim}
@@
 This process helps the user to deal with the slightly complex semantics of the transformation rules avoiding errors or mistakes.
-In addition, our tool is ontology-driven, therefore the service wrapper designer retrieves the domain-specific types from a backend server (such as the catalogue discussed in Sect.~\ref{sec:discovery}) together with the structure of each type, i.e.
+In addition, our tool is ontology-driven, therefore the service wrapper designer retrieves the domain-specific types from a backend server (such as the catalogue discussed in Section~\ref{sec:discovery}) together with the structure of each type, i.e.
 together with a description of the attributes of each object.
@@
 For the \emph{fillAttributes} rules, the selection box shows the attributes of the object type chosen in the parent rule.
-In addition, we may provide some analysis tool, which will help to guarantee that the objects generated by the transformation rules conform to the object types defined in the corresponding conceptual model (see Sect.~\ref{ssec:ontology}).
+In addition, we may provide some analysis tool, which will help to guarantee that the objects generated by the transformation rules conform to the object types defined in the corresponding conceptual model (see Section~\ref{ssec:ontology}).
 This helps to ensure that the objects generated by the designed service wrapper will be compatible for input parameters of subsequent filter steps and or gadgets.

### 2011-01-16 FAST D4_3_3/fast_d4_3_3.tex
internal draft
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1124 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-02-01 trunk/deliverables/D4_3_2/fast_d4_3_2.tex
@@
 
-% This is useful for switching in-line comments on and off!
-\usepackage{ifthen}
-\newboolean{showcomments}
-\setboolean{showcomments}{true}
-\ifthenelse{\boolean{showcomments}}
-{
-	\definecolor{Red}{rgb}{1,0,0}
-	\newcommand{\note}[1]{\begin{scriptsize}{{\textsf{\textbf{{\textcolor{Red}{[[#1]]}}}}}}\end{scriptsize}}
-	\definecolor{Orange}{rgb}{1,0.5,0}
-	\newcommand{\todo}[1]{\textsf{\textbf{\textcolor{Orange}{[[TODO: #1]]}}}}
-    \definecolor{Green}{rgb}{0,1,0}
-    \newcommand{\commentreply}[1]{\begin{scriptsize}{\textsf{\textbf{\textcolor{Green}{[[#1]]}}}}\end{scriptsize}}
-} {
-	\newcommand{\note}[1]{\textbf{}}
-	\newcommand{\todo}[1]{}
-    \newcommand{\commentreply}[1]{}
-}
-
-\newcommand\deliverableNumber{D4.3.2}
+\newcommand\deliverableNumber{D4.3.3}
 \newcommand\deliverableTitle{Mechanisms for Gadget-Service Connections and Gadget Functionality}
@@
 \newcommand\authorOne{Ismael Rivera, NUIG}
-\newcommand\authorTwo{Albert Z\"undorf, UniKassel}
+\newcommand\authorTwo{Albert Zündorf, UniKas}
 \newcommand\authorThree{Author 3}
@@
 		{\large \sffamily \authorOne}\\
-		{\large \sffamily \authorTwo}\\
+%		{\large \sffamily \authorTwo}\\
 		\vspace{0.5cm}
@@
 							\textcolor{white}{\textbf{Change description}}\\ \hline
-1.0 & 26.02.2010 & Ismael Rivera (NUIG) & Final version ready for external review \\ \hline
+%0.1 & 15.12.2008 & Knud M\"oller (NUIG) & template modelled \\ \hline
+%0.2 & 15.02.2009 & Knud M\"oller (NUIG) & content moved from wiki after internal review \\ \hline
+1.0 & 26.02.2010 & Ismael Rivera (NUIG), Albert Zündorf (UniKas) & Final version ready for external review \\ \hline
 \end{tabular}
@@
 These back-end services cannot be directly used; hence they need to be encapsulated in what in the FAST platform is called Resource Adapters.
-The application of semantics to the back-end, through the corresponding resource adapters, and front-end building blocks assures a powerful instrument in the task of building new gadgets, improving the search, and enhancing the connection among the different building blocks which compose a gadget.
+The application of semantics to the back-end, through the corresponding Resource Adapters, and front-end building blocks assures a powerful instrument in the task of building new gadgets, improving the search, and enhancing the connection among the different building blocks which compose a gadget.
 
@@
 	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
-	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{This deliverable presents techniques to exploit different kind of web services such as RESTful, SOAP-based and WSMO services.
-It exposes how these services are commonly defined, and what FAST need to allow these web services to be use within a gadget.
-In this deliverable is also presented a tool which apply some of these techniques allowing users to create the necessary wrappers for these services.}\\ \hline
+	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D4.3.3 is...}\\ \hline
 	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
-	{} & \multicolumn{4}{l|}{\authorTwo}\\ 
+%	{} & \multicolumn{4}{l|}{}\\ 
 %	{} & \multicolumn{4}{l|}{}\\ 
@@
 	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {ismael.rivera@deri.org} \\ \cline{4-5}
-	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495338} \\ \hline
-	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, web service, WSDL, REST, SOAP, WSMO, Resource Adapter}\\ \hline
+	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 ******} \\ \hline
+	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, web services, WSDL, REST, SOAP, WSMO}\\ \hline
 \end{tabular}
@@
 This deliverable is structured as follows.
-Section~\ref{sec:introduction} states the goal, scope and structure of the document.
-Section~\ref{sec:resource_adapters} describes what a resource adapter is, how the platform is able to interact with web services, how they these wrappers or adapters are build and then discovered to be reused in any gadget.
-Then, Section~\ref{sec:restful_web_services_wrapper_tool} defines the service wrapper tool built to facilitate FAST users to create resource adapters of RESTful web services.
-Finally, Section~\ref{sec:related_work}...\todo{Finish it!}
+Section ~\ref{sec:introduction} states the goal, scope and structure of the document.
+Section ~\ref{sec:web_services_wrapping} describes what a service wrapper is, how the platform is able to interact with web services, how these wrappers are build and then discovered to be reused in any gadget.
+Finally, Section ~\ref{sec:restful_web_services_wrapper_tool} defines the service wrapper tool built to facilitate FAST users to create Resource Adapters of RESTful web services.
 
@@
 \clearpage
-\section{Resource Adapters} % (fold)
-\label{sec:resource_adapters}
+\section{Web Services Wrapping} % (fold)
+\label{sec:web_services_wrapping}
 
@@
 
-The following sections explain what a resource adapter is, how a user, in this case a resource developer, can build these resource adapters, and how another user, a screen developer for instance, is able to discover and connect them while creating a screen.
-Hence, these three phases are called: building, discovery and connection.
-
-\subsection{Definition} % (fold)
-\label{sub:definition}
-
-A complex gadget in terms of FAST is aimed to provide a functional access from a graphical user interface to a set of services (SOAP or REST-based Web services) and data sources (Atom/RSS feeds).
+Therefore, a complex gadget in terms of FAST is aimed to provide a functional access from a graphical user interface to a set of services (SOAP or REST-based Web services) and data sources (Atom/RSS feeds).
 These services and data sources need to be modelled and encapsulated within the platform in order to allow the discovery and use of them by other components (i.e.
@@
 This is called Resource Adapter in the FAST platform.
-Figure~\ref{fig:complex_gadget_architecture} illustrates the role they play and which other building blocks they interact with within a gadget.
+Figure ~\ref{fig:complex_gadget_architecture} illustrates their position and interaction with the rest of the building blocks within a gadget.
 
-\begin{figure}[!htb]
+\begin{figure}
   \begin{center}
@@
 
-Basically, a resource adapter, from a design point of view, is composed of:
-
-\begin{enumerate}
-	\item \textbf{metadata} such as creator, label, description, version, and so on,
-	\item a set of \textbf{actions} or operations,
-	\item and a set of \textbf{postconditions} or outputs of the resource.
-\end{enumerate}
-
-Figure~\ref{fig:resource_adapter_overview} shows the information above in a graphical manner.
-An action (represented by a white square) is the minimum unit for the functionality a resource may have.
-Imagine a user clicking on a button to get a list of items for sale, or a checkout action performed in order to buy one of the items from the list.
-As an analogy, an action is what a operation is in a SOAP-based web service.
-Moreover, as these operations require some input, the actions may need a certain precondition (green circles) to be satisfied.
-And at last, the execution of these action may cause a postcondition (red circle) be satisfied, or following the same example, an output is returned as result of the execution of an operation.
-
-\begin{figure}[!htb]
-  \begin{center}
-    \includegraphics[width=12cm]{images/resource_adapter.pdf}
-    \caption{Resource Adapter Overview}
-    \label{fig:resource_adapter_overview}
-  \end{center}
-\end{figure}
-
-% subsection definition (end)
+Once a brief background has been shown, the following sections explain how a user, in this case a resource developer, can build Resource Adapters, and how another user, a screen developer for instance, will discover and connect them while creating a screen.
+Hence, these three phases are called: building, discovery and connection.
 
@@
 
-This section describes different approaches for describing and implementing web services, and how a resource adapter could be generated for these services.
-
 \subsubsection{REST-based Web Services} % (fold)
@@
 
-REST \cite{Fielding2000} is a term to describe an architecture style, not a standard, of networked systems.
+REST is a term to describe an architecture style, not a standard, of networked systems.
 The acronym REST stands for Representational State Transfer.
-REST-based or RESTful web services are created identifying all of the conceptual entities or resources which want to be exposed as services.
+REST-based or RESTful web services \cite{Fielding2000} are created identifying all of the conceptual entities or resources which want to be exposed as services.
 Those entities or resources should be nouns, not verbs (orders, tickets, etc.).
@@
 
-RESTful web services are more lightweight than SOAP-based web services, avoiding the verbose XML-based messages, but unlike the former, which have a standard vocabulary, and commonly used, to describe the web service interface through WSDL, RESTful web services are currently not formally described most of the times.
+REST is lightweight (not a lot of extra xml markup, human readable results), but unlike SOAP-based web services, which have a standard vocabulary, and commonly used, to describe the web service interface through WSDL, RESTful web services are currently not formally described most of the times.
 For a service consumer to understand the context and content of the data that must be sent to and received from the service, both the service consumer and service producer must have an out-of-band agreement.
 This takes the form of documentation, sample code, and an API that the service provider publishes for developers to use.
-For example, the many web-based services available from Google, Yahoo, Flickr, Amazon, and so on have accompanying artifacts describing how to consume the services.
-This style of documenting REST-based web services is fine for use by developers, but it averts tools from programmatically consuming such services and generating executable stubs for a given programming language, as a web service described using WSDL allows.
+For example, the many web-based services available from Google, Yahoo, Flickr, Amazon, and so on have accompanying artefacts describing how to consume the services.
+This style of documenting REST-based web services is fine for use by developers, but it averts tools from programmatically consuming such services and generating artefacts specific to programming languages, as a web service described using WSDL allows.
 Nevertheless, Web Application Description Language (WADL) attempts to resolve some of these issues by providing a means to describe services in terms of schemas, HTTP methods, and the request or response structures exchanged, but this language is not widely adopted yet by RESTful web service developers.
@@
 There is no need of a formal document such as WADL defining the web service, for this reason, building service wrappers for these services involve the correct understanding of the service by a human being and a tool to facilitate this task.
-This tool is explained in detail in Section~\ref{sec:restful_web_services_wrapper_tool}.
+This tool is explained in detail in Section ~\ref{sec:restful_web_services_wrapper_tool}.
 
@@
 
-SOAP-based web Services or "Heavyweight Web Services" use Extensible Markup Language (XML) \cite{XML1.1} messages that follow the Simple Object Access Protocol (SOAP) standard \cite{SOAP1.2} and have been popular within traditional enterprise, usually relying on HTTP for message negotiation and transmission.
+SOAP-based web Services or "Heavyweight Web Services" use Extensible Markup Language (XML) \cite{XML1.1} messages that follow the Simple Object Access Protocol (SOAP) standard \cite{SOAP1.2} and have been popular with traditional enterprise, usually relying on HTTP for message negotiation and transmission.
 In such services, there is often a machine-readable description of the operations offered by the service written in a Web Services Description Language (WSDL) document.
-The advantage of specifying the service in a using WSDL is that it can be programmatically processed.
+Hence, the advantage of using WSDL is that it can be programmatically processed.
 
@@
 From a specific WSDL definition, a set of methods or operations can be easily extracted which encloses a set of inputs and outputs.
-These operations would be transformed into \emph{actions}, and the inputs and outputs would be used to define resource adapter \emph{pre/postconditions}.
-But there is a critical problem following this approach.
-The definition of the inputs, outputs, operations and their types are done in a XML Schema document.
-An XML schema describes, in a syntactic manner, the structure of an XML document, but there is no straight-forward relation with the concepts used to describe pre/postconditions in the building blocks.
-In order to illustrate the concepts of WSDL, and the gap between them and semantically-defined building blocks, a shorten version of a real Amazon E-commerce web service will be presented.
-Concretely, the choosen web service performs an item search through the Amazon catalog.
-It expects as input an ItemSearch element, which may contain an actor, an artist or an author, and the item availability.
-The output of the service would be a set of items.
-The WSDL including for this service is given below.
+These operations would be transformed into actions, and the inputs and outputs would be used to define Resource Adapter pre/postconditions.
+With the use of this information along with the XML Schema defining the types, a Resource Adapter can be semantically defined.
 
-\singlespacing
-\begin{small}
-\verbatiminput{examples/amazon.wsdl}
-\end{small}
-\doublespacing
-
-As stated before, one of the advantages of describing web services in a formal document is that they can be programmatically consumed.
-In fact, there are many frameworks out there to facilitate such a task, parsing WSDL files, creating stubs to consume the web service and so on.
-However, in order to create a resource adapter from the WSDL defined above, the types defined in the XML Schema need to be translated into concepts from an ontology.
-
-This problem could be addressed in two different ways.
-First one, as a manual process mapping these types and their attributes to concepts and properties, as it was proposed for RESTful services.
-Second approach is taking advantage of SAWSDL, which keeps the process in the programmatically direction.
-
-SAWSDL \cite{sawsdl} stands for Semantic Annotations for WSDL and XML Schema, and it defines how to add semantic annotations to various parts of a WSDL document such as input and output message structures, interfaces and operations.
-These semantic annotations enable the possibility to develop resource adapters from WSDL documents.
-They reference a concept in an ontology or a mapping document, and are either independent of the ontology expression languages and mapping languages.
-
-Basically, there are three main concepts about SAWSDL: \emph{modelReference} to specify the association between a WSDL or XML Schema component and a concept from the ontology, and \emph{liftingSchemaMapping} and \emph{loweringSchemaMapping} to define the mapping mechanisms between semantic data and XML.
-These lowering and lifting schema mappings are not explained in this deliverable, however in Appendix A of the SAWSDL specification \cite{sawsdl} they are described in detail, and illustrated with examples.
-
-To illustrate the usage of these attributes, the following code shows a fragment of the previous example using SAWSDL to semantically annotate the service with concepts from an imaginary ontology defining the Amazon E-commerce web service.
-
-\singlespacing
-\begin{small}
-\verbatiminput{examples/amazon_sawsdl.wsdl}
-\end{small}
-\doublespacing
-
-Hence, SAWSDL enriches the web services descriptions made in WSDL with references to concepts from a semantic model and enables the possibility to create resource adapters from these definitions without a strong user participation in the process.
-If the semantic metadata is missing, a manual annotation process could be considered, where the user would need to relate concepts from an ontology with the WSDL or XML Schema elements, and create the mapping between them manually as well.
+To integrate WSDL defined services into the FAST platform, we again propose an interactivce approach: We extend the tool for REST service wrapping with the ability of loading WSDL definitions of a service and this WSDL definition is then exploited for guidance of the user through the Resource Adapter construction, cf.
+Section ~\ref{sec:restful_web_services_wrapper_tool}.
 
@@
 The capability of a web service defines its functionality in terms of pre/postconditions, assumptions and effects.
-Apart from that, a web service capability is defined by specifying the following elements: non-functional properties, imported ontologies, used mediators, and shared variables.
-The creation of the resource adapters for these web services will just take into consideration the set of pre/postconditions.
+A web service capability is defined by specifying the following elements: non-functional properties, imported ontologies, used mediators, shared variables, precondition, postcondition, assumption, and effect.
 
-In the rest of the section, the Amazon E-commerce web service will be used to build a WSMO web service with the same functionality \cite{wsmo_amazon}.
-In fact, just the small subset of the service, used in previous sections, will be tranformed.
+Basically, the web service will offers to a client a postcondition when some conditions are met in the information space (precondition).
+Effects and assumptions can stay out of this study, since they will not be taken into account by any of the building blocks of a gadget.
 
-First of all, an ontology needs to be extracted from the XML Schema types of the WSDL definition.
-For the shake of simplificity, this example does a one-to-one mapping between the Schema types and the concepts, though a real sceneario may consider an different approach to optimise the modelling.
-
-The concept describing an item search request for the Amazon web service is:
-
-\singlespacing
-\begin{small}
-\begin{verbatim}
-concept itemSearchRequest
-    nonFunctionalProperties
-      dc#description hasValue "A request to search for an item"
-    endNonFunctionalProperties
-    author ofType _string
-    title ofType (1) _string
-    rating ofType ratings#audienceRating
-    minPrice ofType (1) _decimal
-    maxPrice ofType (1) _decimal
-\end{verbatim}
-\end{small}
-\doublespacing
-
-The concept \emph{itemContainer} describes a list of items returned in the item search response, and its code is presented below:
-
-\singlespacing
-\begin{small}
-\begin{verbatim}
-concept itemContainer
-    nonFunctionalProperties
-      dc#description hasValue "Contains a list of items"
-    endNonFunctionalProperties
-    items ofType item
-\end{verbatim}
-\end{small}
-\doublespacing
-
-Next step would be to describe the namespace and non-functional properties declarations for the WSMO Amazon web service, but as it was said, this information is not relevant for the example.
-Hence, the simplified version of the Amazon web service preconditions would look like:
-
-\singlespacing
-\begin{small}
-\begin{verbatim}
-precondition
-  nonFunctionalProperties
-     dc#description hasValue "The Amazon service handles requests 
-     for searching for items by keywords"
-  endNonFunctionalProperties
-  definedBy
-    // A search by author, title, minPrice or maxPrice
-    ?request memberOf am#itemSearchRequest
-\end{verbatim}
-\end{small}
-\doublespacing
-
-It is a fairly simple precondition, which the only condition required is the existence of a \emph{itemSearchRequest} object.
-
-However, postconditions are a bit more difficult to construct.
-They take the form of implication rules, so given a particular precondition, one particular postcondition will take effect.
-To follow the same example, these would be the postconditions for the Amazon search items functionality, including just the criterion author, title, minimum and maximum price as defined in the ontology.
-
-\singlespacing
-\begin{small}
-\begin{verbatim}
- postcondition
-  nonFunctionalProperties
-    dc#description hasValue "The service returns a list of searched
-      items"
-  endNonFunctionalProperties
-  // The result for a search request by author is a container of products
-  // such that all items in the container have the requested author
-  (?request[
-      author hasValue ?author
-    ] memberOf am#searchItems implies 
-    exists ?container 
-      (?container memberOf am#itemContainer and 
-        forall ?item 
-          (?container[
-              items hasValue ?item
-            ] implies 
-            ?item[
-              author hasValue ?author
-            ]
-          )
-      )
-  ) and
-
-  // The result for a search request by title is a container of products
-  // such that all items in the container have the requested title
-  (?request[
-      title hasValue ?title
-    ] memberOf am#searchItems implies 
-    exists ?container 
-      (?container memberOf am#itemContainer and 
-        forall ?item 
-          (?container[
-              items hasValue ?item
-            ] implies 
-            ?item[
-              title hasValue ?title
-            ]
-          )
-      )
-  ) and
-
-  // The result for a search request by minPrice is a container of products
-  // such that all items in the container have a price >= requested minPrice
-  (?request[
-      minPrice hasValue ?minPrice
-    ] memberOf am#searchItems implies 
-    exists ?container 
-      (?container memberOf am#itemContainer and 
-      forall{?item,?price} 
-        (?container[
-            items hasValue ?item
-          ] and 
-          ?item[
-            price hasValue ?price
-          ] implies 
-          ?price >= ?minPrice
-        )
-      )
-  ) and
-
-  // The result for a search request by maxPrice is a container of products
-  // such that all items in the container have a price <= requested maxPrice
-  (?request[
-      maxPrice hasValue ?maxPrice
-    ] memberOf am#searchItems implies 
-    exists ?container 
-      (?container memberOf am#itemContainer and 
-      forall {?item,?price} 
-        (?container[
-            items hasValue ?item
-          ] and 
-          ?item[
-            price hasValue ?price
-          ] implies 
-          ?price =< ?maxPrice
-        )
-      )
-  ).
-\end{verbatim}
-\end{small}
-\doublespacing
-
-The manner pre/postconditions are defined in WSMO web services are similar to the way they are defined in the Resource Adapters.
-Section 4.4 in \cite{moller2010fast_ontology} gives a detailed explanation on this topic.
-Basically, and using previous examples, the preconditions would be transformed as follows:
-
-\begin{verbatim}
-?request memberOf am#itemSearchRequest => ?request rdf:type am:itemSearchRequest
-\end{verbatim}
-
-And a simple postcondition would be transformed as follows:
-
-\begin{verbatim}
-exists ?container 
-      (?container memberOf am#itemContainer) => ?container rdf:type am:itemContainer
-\end{verbatim}
+Like WSDL defined services, we incorporate WSMO services into the FAST platform in an interactive storyboard driven approach.
+The FAST tool described in Section ~\ref{sec:restful_web_services_wrapper_tool} reads the WSMO specification of a service and exploits this information again for guidance for the contstruction of an Resource Adapter.
 
@@
 
-\subsection{Discovery} % (fold)
-\label{sub:discovery}
+\subsection{Discovering} % (fold)
+\label{sub:discovering}
 
-In the building phase, every resource adapter is created sharing a common structure for screen components, such as forms, operators and resources.
+In the building phase, every Resource Adapter is created sharing a common structure for screen components, such as forms, operators and resources.
 Hence, it will have a set of actions which will contain a set of preconditions to satisfy in order to be executed, and after its execution, it may produce any of the conditions of their postconditions.
 
-For this reason, a screen developer during the screen development phase is able to reuse these screen components which are already stored in the catalogue, establishing which the screen's pre/postconditions are, and then the system will suggest several screen components which may satisfy these pre/postconditions.
+For this reason, a screen developer during the screen development phase is able to reuse these screen components which are already stored in the catalog, establishing which the screen's pre/postconditions are, and then the system will suggest several screen components which may satisfy these pre/postconditions.
 In a first step, the system will suggest screen components which have any of the preconditions which can be connected to any of the screen preconditions (the connection will be represented by an internal pipe), and the same would happen for the postconditions.
@@
 
-On these basis, the mechanism is relatively similar to the screen recommender, used to find the best screens to create a screen flow.
+In these bases, the mechanism is relatively similar to the screen recommender, used to find the best screens to create a screen flow.
 Once more, it is based on the pre/postconditions of a set of building blocks.
@@
 
-\subsection{Connection} % (fold)
-\label{sub:connection}
-
-A resource adapter is aimed to perform some action on a web service.
-In order to perform a given action, a precondition may need to be fulfilled, and as a result of the execution a postcondition may be generated.
-These pre/postconditions are intended to be used by pre/postconditions of other building blocks within a gadget.
-This connection must be explicitly defined, because it will indicate the real data-flow between different building blocks.
-In FAST this is commonly called Piping, since a pipe in computer science is just that, a one-way communication channel for interprocess communication; therefore a pipe is a logical connection made between a precondition and a postcondition of different building blocks.
-Figure~\ref{fig:Piping} shows an example of a search form for the Amazon service.
-This is a screen, composed by a form and a resource adapter.
-The form allows the user to introduce a term for the search, producing a postcondition which will be redirect to a resource adapter precondition.
-The execution of the web services produces a list of results, which will be forward to the postcondition of the screen.
-
-\begin{figure}[!htb]
-  \begin{center}
-    \includegraphics[width=12cm]{images/piping.pdf}
-    \caption{Piping}
-    \label{fig:Piping}
-  \end{center}
-\end{figure}
-
-% subsection connection (end)
-
 % section web_services_wrapping (end)
@@
 \clearpage
-\section{RESTful Web Services Wrapper Tool} % (fold)
+\section{Web Service Wrapper Tool} % (fold)
 \label{sec:restful_web_services_wrapper_tool}
 
-In FAST, the wrapper tool will be in charge of the resource adapters' building phase.
-At the moment, it just supports RESTful web services.
-The building phase for this type of web services is done in two steps: a first step will be in charge of the construction of a service request and a second step will analyse the response got from the execution of the service, allowing the extraction of facts contained in that response and mapping them to domain-specific concepts from the ontologies used within FAST by any building block.
+In FAST, the wrapper tool will be in charge of the Resource Adapters' building phase.
+The building of Resource Adapters is done in three steps: in a first step, we define which pre- and postconditions are used to pass parameters between FAST building blocks and the Resource adapter.
+The second step is in charge of the construction of a service request and the third step analyses the response got from the execution of the service, allowing the extraction of facts contained in that response and mapping them to domain-specific concepts from the ontologies used within FAST by any building block.
+
+\subsection{Defining pre- and postconditions} % (fold)
+\label{sub:resource_adapter_parameters}
+
+We will demonstrate the use of the FAST Service Wrapper Tool on the example of a simple Formula1 service.
+In the first tab of the FAST Service Wrapper Tool, the user defines the name and the pre- and postconditions of the desired Resource Adapter, cf.
+Fig.
+\ref{fig:construct_pre_post_conditions}.
+For the precondition, we provide a name, a type, and an example value.
+The type of a pre- or postcondition is chosen from a drop down list of available concepts.
+This list of available concepts is retrieved from the FAST ontology catalog.
+The example value is used for test runs during subsequent construction steps.
+We keep track of example values provided for different domain concepts and propose them, when the domain concept is re-used e.g.
+in another Resource Adapter.
+
+\begin{figure}
+  \begin{center} \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithPrePostConds.png}
+    \caption{Configuring pre- and postconditions of a Resource Adapter}
+    \label{fig:construct_pre_post_conditions}
+  \end{center}
+\end{figure}
 
@@
 
-As a first approach, the interaction with these services will be limited to retrieve information to feed the gadgets using simply GET requests.
-A service request would be assembled using a certain URL and a set of parameters.
-As an example, Ebay Shopping web service will be studied.
-The following URL is invoked to retrieve a list of items corresponding to certain search keywords:
+In our simple case, the service request consists just of a URL that is easily retrieved from API documentation page of our Formula1 service.
+This URL consists of the service address \url{http:// ergast.com/api} followed by the year we are querying for followed by the category \url{drivers} we are interested in.
+The user may grab such an URL from the API documentation page of the targeted web service and drop it into the Request tab of the FAST Service Wrapper Tool, cf.
+Figure ~\ref{fig:construct_service_request}.
+The FAST Service Wrapper Tool will than analyze the dropped URL and split it into its parts and provide a convenient form allowing to edit e.g.
+all the parameters of the URL.
+In our example the URL has no parameters so the parameter form shows only the base URL.
 
-\url{http://open.api.sandbox.ebay.com/shopping?appid=KasselUn-efea-4b93-9505-5dc2ef1ceecd&version=517&callname=FindItems&ItemSort=EndTime&QueryKeywords=USB&responseencoding=XML}
+For WSDL or WSMO services, the FAST Service Wrapper Tool provides additional support for the construction of the request.
+The user may point the FAST Service Wrapper Tool to the WSDL or WSMO definition of the targeted service.
+The FAST Service Wrapper Tool then reads this definition and analysis it.
+With the help of this information, the FAST Service Wrapper Tool provides interactive guidance for selecting the desired operation from a service and then the FAST Service Wrapper Tool retrieves the operation descriptions and fills the request parameter form, adequately.
 
-As you may see, the URL invoked is http://open.api.sandbox.ebay.com/shopping and the parameters used in the example are:
-\begin{description}
-	\item[appid] this is the application ID obtained to use the API.
-	\item[version] the API version.
-	\item[callname] in this case FindItems to search through all items in Ebay.
-	\item[itemsort] sorting method for the list of items.
-	\item[querykeywords] list of keywords.
-	\item[responseencoding] format of the response message obtained by the invocation of the request.
-\end{description}
+Now we have to connect the request parameters to the parameter values passed via the preconditions to the Resource Adapter.
+In our case, we replace the year part of the URL with a placeholder for the precondition season \url{<season>}, see Request URL input field at the top of Figure ~\ref{fig:construct_service_request}.
 
-If needed, a detailed specification of the Ebay Shopping API can be found at \cite{eBayShoppingAPIs}.
+In our example, we use a simple GET based web service.
+The FAST Service Wrapper tool also supports the other kinds of URL operations as POST, PUT, etc.
+For POST requests, one usually passes parameters within a special post body.
+For unstructured POST bodies, we provide a text area allowing the user to fill in placeholders for references to precondition values.
+In case of structured POST bodies as in the case of WSDL and WSMO services the POST body uses a structured text format based e.g.
+on JSON or XML.
+In that case, the body text is again passed and the parameters are included in the request parameter form part.
+Then the user may easily use the request parameter form part to fill in precondition placeholders.
 
-The above URL for searching items in Ebay is followed by the query parameters, which take the form \textit{argument=value}, where the arguments and values are URL encoded, and are separated by an ampersand (\&).
-For instance, the only relevant parameter the user would need to specify is \emph{querykeywords}, thus somehow the service has to receive an input value for it, while the other parameters can be set to a default value.
-However, in order to develop a generic wrapper for the any service, all parameters might be set by the user.
+Figure ~\ref{fig:construct_service_request} shows the screen to construct the service requests.
 
-To achieve this, service wrappers will be handled as any buis with preconditions and postconditions (inputs and outputs).
-Therefore, these precondition ports might be used to determine values to parameters like query keywords coming from a building block inside the screen the service is placed or even an external screen as a screen precondition.
-The service wrapper tool composes the request using a template string which will contain placeholders for precondition values.
-Before sending the request to the service, the placeholders are replaced with their corresponding values and the resulting URL is then ready to be sent the service request.
-
-To achieve this, service wrappers will be handled as any other building block with preconditions and postconditions (inputs and outputs).
-Therefore, these precondition ports might be used to determine values to parameters like query keywords coming from a building block inside the screen the service is placed or even an external screen as a screen precondition.
-
-To define the preconditions and postconditions of a new service wrapper, the FAST service wrapper tool provides a form allowing the edition of these entries, cf.
-Figure~\ref{fig:construct_pre_post_conditions}\footnote{Note: These example values may be used to test the service wrapper, but they can be freely edited.}
-
-\begin{figure}[!htb]
+\begin{figure}
   \begin{center}
-    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithRequestExample.jpg}
-    \caption{Configuring pre/postconditions of a service wrapper}
-    \label{fig:construct_pre_post_conditions}
-  \end{center}
-\end{figure}
-
-The service wrapper tool composes the request using a template string which will contain placeholders for precondition values.
-Before sending the request to the service, the placeholders are replaced with their corresponding values and the resulting URL is then ready to be sent the service request.
-
-Figure~\ref{fig:construct_service_request} shows the screen to construct the service requests.
-In the top input field of Figure~\ref{fig:construct_service_request}, the user may drop an example HTTP request taken e.g.
-form the service documentation (e.g.
-\cite{eBayShoppingAPIs} for this example).
-The tool analyses the example request and, in the middle of the screen, a form for editing the request parameters is provided.
-This example shows how a user has connected the \textit{QueryKeywords} parameter with the precondition \textit{search\_key} by adding a corresponding reference to the value field of that parameter.
-In addition, an access key has been retrieved and has been entered as value for the \textit{appid} parameter.
-
-\begin{figure}[!htb]
-  \begin{center}
-    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithRequestExample.jpg}
-    \caption{Constructing the service request URL and its parameters}
+    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSRequestTemplate.png}
+    \caption{Constructing the service request and its parameters}
     \label{fig:construct_service_request}
@@
 
-Below the request paramter editing form of Figure~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed service by sending it to the specified service address (via a server relay).
-Then the placeholders for preconditions are replaced with example values and the resulting HTTP request is shown below the parameter form.
-In addition, the request is sent and the response is shown on the bottom of that page, which helps the users to figure out quickly whether the constructed request works as desired.
+Below the request paramter editing form of Figure ~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed request by sending it to the specified service address (via a server relay).
+Then the placeholders for preconditions are replaced with example values and the resulting request is shown below the parameter form.
+In addition, the request is send and the response is shown on the bottom of that page.
+This gives the user a fast feedback whether the constructed request works as desired.
+
+\subsubsection{Limitations} % (fold)
+\label{ssub:limitations}
+
+It is worth pointing out that currently the wrapper tool is able to construct input ports for the wrappers using just basic types.
+However, every building block accepts any (complex) concept as a pre/postcondition, so the wrapper tool needs to be adapted to permit any concept as input or precondition, extracting the required text to construct the request.
+This should be easy to achieve before the project review.
+
+% subsubsection limitations (end)
 
@@
 This response message could be serialize in any format, though the most common formats used nowadays are XML or JSON among others.
-To continue the example it was started in the previous section, the response of the Ebay Shopping service will be in XML as specified in the request, which is the format supported by the wrapper tool.
+To continue the example started in the previous section, the response of the Formula1 service will is in XML.
+However, the FAST wrapper tool supports JSON, similarly.
 
-Figure~\ref{fig:response_service_execution} shows the transformation data tab of the wrapper tool.
+\subsubsection{Translation XML/JSON into Facts} % (fold)
+\label{ssub:translation_xml_into_facts}
+
+Figure ~\ref{fig:response_service_execution} shows the data tranformation tab of the wrapper tool.
 
@@
   \begin{center}
-    \includegraphics[angle=90,width=0.8\linewidth]{images/ServiceWrapperToolGVSWithTransformationRules.jpg}
-    \caption{Interactive rule based transformation of an XML response to FAST facts}
+      \includegraphics[angle=90,width=0.8\linewidth]{images/ServiceWrapperToolGVSWithTransformationRulesF1.png}
+    \caption{Interactive, rule based transforming of an XML response to FAST facts (photoshoped screen dump)}
     \label{fig:response_service_execution}
@@
 
-Once the service response, in XML format, has been retrieved, the transformation tab shows it as an interactive object tree on the left side of Figure~\ref{fig:response_service_execution}.
-To construct this interactive object, the XML document has been parsed into a DOM, and a simplified tree representation of that DOM is built up.
-This tree representation of the XML data is used as an input to construct transformation rules.
+Once the service response, in XML or JSON format, has been retrieved, the transformation tab shows it as an interactive object tree on the left side of Figure ~\ref{fig:response_service_execution}.
+To construct this interactive object tree, the XML or JSON document is parsed into a DOM, and a simplified tree representation of that DOM is built up.
+This tree representation of the XML/JSON data is used as an input to construct transformation rules.
 
-A transformation rule is used to analyse the XML data and to generate domain-specific facts from concepts from the ontologies used by the pre/postconditions of the different building blocks.
-A transformation rule is composed of three elements, cf the middle part of Figure~\ref{fig:response_service_execution}.
+A transformation rule is used to analyse the XML/JSON data and to generate domain-specific facts from concepts from the ontologies used by the pre/postconditions of the different building blocks.
+A transformation rule is composed of three elements, cf the middle part of Figure ~\ref{fig:response_service_execution}.
 First, the \textit{from} field indicates the XML elements to be translated by rule.
-These XML elementsare identified by the tagname of a DOM element from the XML document.
+These XML elements are identified by the tagname of a DOM element from the XML document.
 Second, the type of the rule will be set, taking one of the following values: \emph{createObject}, \emph{fillAttributes} or \emph{dummy}.
@@
 A detailed explanation of the type of actions to be trigger from the transformation rules is:
-
 \begin{description}
@@
 The type of that new fact is provided in the third compartment.
-In the example being explained, the root rule searches for XML elements with tagname \emph{FindItemsResponse} and for each such element a \emph{List} fact is created.
-The resulting facts are shown in a facts tree in the right of Figure~\ref{fig:response_service_execution}.
+In the example being explained, the root rule searches for XML elements with tagname \emph{DriverTable} and for each such element a \emph{List} fact is created.
+The resulting facts are shown in a facts tree in the right of Figure ~\ref{fig:response_service_execution}.
 	\item[action \emph{fillAttributes}] does not create a new fact but it fills the value of the attribute provided as third part of such rules.
-In our example, the third transformation rule searches for XML elements with tagname \emph{Title}.
-Note, the rule is a sub-rule of the second rule, which generates \emph{Product} facts.
-Thus, the sub-rule searches for \emph{Title} tags only in the subtree of the XML data that has been identified by an application of the parent rule before.
-For example, the \emph{Item} rule may just have been applied to the first \emph{Item} element of the XML data.
-Then, the \emph{Title} rule is applied only to the first \emph{Item} sub-tree of the XML data and thus it will find only one \emph{Title} element in that sub-tree (not visible in Figure ~\ref{fig:response_service_execution} ).
-The value of that \emph{Title} element is then transfered to the \emph{productName} attribute of the corresponding \emph{Product} fact.
+In our example, the third transformation rule searches for XML elements with tagname \emph{FamilyName}.
+Note, the rule is a sub-rule of the second rule, which generates \emph{Person} facts.
+Thus, the sub-rule searches for \emph{FamilyName} tags only in the subtree of the XML data that has been identified by an application of the parent rule before.
+For example, the \emph{Driver} rule may just have been applied to the first \emph{Driver} element of the XML data.
+Then, the \emph{FamilyName} rule is applied only to the first \emph{Driver} sub-tree of the XML data and thus it will find only one \emph{FamilyName} element in that sub-tree.
+The value of that \emph{FamilyName} element is then transfered to the \emph{name} attribute of the corresponding \emph{Person} fact.
 Actually, our \textit{from} fields allows also to refer to parts of an XML attribute e.g.
@@
 It is also possible to combine constant text and elements of multiple XML tree elements.
+A \textit{from} field may also refer to multiple tags of the source tree, e.g.
+to combine a full name from a first name and a family name field.
 	\item[action \emph{dummy}] does not create or modify any facts but such rules are just used to narrow the search space for their sub-rules.
@@
 
-Since FAST is storyboard oriented, the service wrapper tool follows the storyboard idea as well.
-Any time, a change to a transformation rule is done, the transformation process is triggered and the resulted facts tree is directly shown.
+Since FAST is storyboard oriented, the FAST Service Wrapper Tool follows the storyboard idea as well.
+Any time, a change to a transformation rule is done, the transformation process is triggered and the resulting facts tree is directly shown.
 This process helps the user to deal with the slightly complex semantics of the transformation rules avoiding errors or mistakes.
-In addition, FAST is semantic-driven, therefore, the service wrapper designer shall retrieve the domain-specific types from a FAST ontology server together with the structure of each type, i.e.
+In addition, FAST is semantic-driven, therefore, the service wrapper designer retrieves the domain-specific types from the FAST ontology catalog server together with the structure of each type, i.e.
 together with a description of the attributes of each fact.
@@
 In addition, we may provide some analysis tool, which will help to guarantee that the facts generated by the transformation rules conform to the fact types defined in the corresponding FAST ontology.
-This helps to ensure that the facts generated will be compatible for precondition ports of subsequent filter steps.
+This helps to ensure that the facts generated by the designed Resource Adapter will be compatible for preconditions of subsequent filter steps and or screens.
 
@@
 
-Once the wrapping of a service has been defined and tested in the service wrapper tool, it shall generate an implementation of the desired Resource Adapter in JSON, HTML, and JavaScript, ready to be deployed and executed inside a screen.
-This service wrapper implementation is compliant with the formats required by the Gadget Visual Storyboard Tool (GVS) and it shall be stored inside the FAST catalogue, in order to be found and used by any user.
+Once the wrapping of a service has been defined and tested in the FAST Service Wrapper Tool, it generates an implementation of the desired Resource Adapter in XML, HTML, and JavaScript, ready to be deployed and executed inside a gadget.
+In addition, a description of the new building block is added to the FAST catalog.
+Thereby, the new Resource Adapter becomes available for the construction of new screens and screen flows.
 
@@
 The rule based approach of the service wrapper tool has been inspired by triple graph grammars, cf.
-\cite{Schurr94,JSZ97c}.
+\cite{conf/wg/Schurr94,JSZ97c}.
 While general triple graph grammars allow to relate complex graph structures with each other, in our case the target of a rule is always a single object or attribute.
@@
 \url{http://biomoby.open-bio.org/CVS_CONTENT/moby-live/Java/docs/sawsdlServlet.html}
-
-Perhaps, it's worth to mention the service wrappers Yahoo Pipes has, which are a fixed set developed by them, with no possibility for a user to create new ones, as we do.
-(other sites doing the same??)
 

### 2011-01-28 trunk/papers/webist2011/paper_sec_catalogue.tex
power shortening and improving
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1208 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-06-15 trunk/papers/icsoc2010/paper_sec_catalogue.tex
@@
 
-\section{Publishing and discovery platform}
+\section{Publishing and Discovery Platform}
 \label{sec:discovery}
 
-The area of web service publication and discovery has been subject for a lot of research since the very beginning the concept was coined.
-The current state of the art provides several solutions and strategies which providers and consumers may take advantage of.
-However, as explained in previous sections, a large number of them suffer from limited syntax-based descriptions and simple keyword-based search, and the gap between discovery and consumption is an obstacle for their usage by non-technical end-users.
-In order to improve this issue, this paper presents a novel publishing platform, permitting any enterprise or individual to publish their public domain web services, with an enhanced semantic search based on the definition of the services, and serving web service wrappers for easy consumption in web applications.
+As explained in previous sections, current solutions and strategies for web service publication and discovery suffer from 
+limited syntax-based descriptions and simple keyword-based search, while other more complex approach failed because the added complexity did not offer sufficient benefits. 
+This paper therefore presents a novel publishing platform (the \emph{catalogue}), permitting any enterprise or individual to publish their public web services, providing enhanced semantic search service wrappers for easy consumption in web applications.
 
+% Overview subsection
 \subsection{Overview}
@@
 
-The main difference of this platform with regards to the state of the art is that it is targeting a different kind of user, and covers many deficiencies others tools do not tackle.
-As a brief overview, the platform being presented:
+One of the main difference of this platform with regards to the state of the art is that it is targeting a different kind of user. As a brief overview, the platform being presented:
 
@@
 
-This platform was intended and it is being used as part of the storyboard tool of the FAST platform~\cite{hoyer2009fast}, and it communicates with other components through its RESTful API, using JSON for the input of the requests.
-This paper will focus on the functionality regarding web service publication and discovery, and the process to publish any third-party web service into the platform.
-This process requires two steps:
-\begin{inparaenum}[(i)]
-	\item generation of the wrapper (see Sect. \ref{sec:wrapping_web_services}), and 
-	\item creation of the resource in the publishing platform. 
-\end{inparaenum}
-The following is an example of part of a request sent to create a new wrapper into the platform (for the sake of clarity, some parts of the request have been omitted, but the main idea is shown):
-
-\begin{listing}
-\begin{verbatim}
-{
-  "code": "http://demo.fast.morfeo-project.org/code/amazonSearch.js",
-  "creationDate": "2010-01-26T17:01:13+0000",
-  "description": {"en-gb": "Amazon web service - Search"},
-  "label": {"en-gb": "Amazon search"},
-  "actions": [{
-    "name": "filter",
-    "preconditions": [[{
-      "id": "item",
-      "label": {"en-gb": "Ebay List"},
-      "pattern": "?item
-                  http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                  http://aws.amazon.com/AWSECommerceService#Item",
-      "positive": true
-  }]],
-  "postconditions": [
-    ...
-  ],
-  ...
-\end{verbatim}
-\label{lis:json_request}
-\end{listing}
+% This paper will focus on the functionality regarding web service publication and discovery, and the process to publish any third-party web service into the platform. This process requires two steps:
+% \begin{inparaenum}[(i)]
+% 	\item generation of the wrapper (see Section \ref{sec:wrapping_web_services}), and 
+% 	\item creation of the resource in the publishing platform. 
+% \end{inparaenum}
+% The following is an example of part of a request sent to create a new wrapper into the platform (for the sake of clarity, some parts of the request have been omitted, but the main idea is shown):
+% 
+% \begin{listing}
+% \begin{verbatim}
+% {
+%   "code": "http://fast.morfeo-project.org/amazon-search.js",
+%   "creationDate": "2010-01-26T17:01:13+0000",
+%   "description": {"en-gb": "Amazon web service - Search"},
+%   "label": {"en-gb": "Amazon search"},
+%   "actions": [{
+%     "name": "filter",
+%     "preconditions": [[{
+%       "id": "item",
+%       "label": {"en-gb": "Ebay List"},
+%       "pattern": "?item
+%         http://www.w3.org/1999/02/22-rdf-syntax-ns#type
+%         http://aws.amazon.com/AWSECommerceService#Item"
+%   }]],
+%   "postconditions": [
+%     ...
+%   ],
+%   ...
+% \end{verbatim}
+% \label{lis:json_request}
+% \end{listing}
 
 The enriched search capabilities are supported by the definition of pre- and post-conditions.
-They allows to define the inputs and outputs of the services and other building blocks using concepts from any ontology, and in this way to find web services or other building blocks which match each other and can therefore be integrated.
+They allow to define the inputs and outputs of the services and other building blocks using concepts from any ontology, and in this way to find web services or other building blocks which can be integrated.
 The concepts of pre-/post-conditions were strongly influenced by WSMO~\cite{roman2005}, simplified and implemented in RDFS for better live performance.
-The model used to define the web service wrappers and the ontology created are described in more detail in Sect.~\ref{ssec:ontology}.
 
-\subsection{Platform architecture}
-\label{ssec:architecture}
+% % Architecture subsection
+% \subsection{Platform architecture}
+% \label{ssec:architecture}
 
-The architecture comprises a RDF store used for persistence, a business layer dealing with the model and reasoning, and public facade providing the core functionality as a RESTful API, and a SPARQL endpoint accessing directly the RDF store.
-This presentation layer is aimed to interact with the FAST Gadget Visual Storyboard, or any other third-party application.
+The catalogue architecture comprises an RDF store used for persistence, a business layer dealing with the model and reasoning, and a public facade providing the core functionality as a RESTful API, as well as a SPARQL endpoint accessing the RDF store directly.
+This presentation layer is aimed to interact with the FAST Gadget Visual Storyboard (see Fig.~\ref{fig:fast_architecture}), or any other third-party application.
 
-The RESTful API and the SPARQL endpoint are part of the presentation layer.
-The SPARQL endpoint is offered using the SPARQL protocol service as defined in the SPROT specification \cite{sprot} and is aimed to enable third-party developers to query directly the knowledge base using SPARQL queries.
+% The RESTful API and the SPARQL endpoint are part of the presentation layer.
+The SPARQL endpoint is offered using the SPARQL protocol service as defined in the SPROT specification~\cite{sprot} and is aimed to enable third-party developers to directly query the knowledge base using SPARQL queries.
 This feature is supported by the Sesame RESTful HTTP interface for SPARQL Protocol for RDF.
-
-The business logic layer contains all the domain-specific processing and reasoning. It provides functions to interact with all the elements of the domain model specified in \ref{ssec:ontology},
-acting as a mediator between the public RESTful API and the persistence layer.
-
-The persistence layer provides an API allowing to work with a standard set of objects representing the model.
-The interaction with the underlying RDF store is made via RDF2Go, an abstraction over triple (and quad) stores, which allows to program against rdf2go interfaces and choose any RDF store implementation.
+% 
+% The business logic layer contains all the domain-specific processing and reasoning. It provides functions to interact with all the elements of the domain model specified in \ref{ssec:ontology},
+% acting as a mediator between the public RESTful API and the persistence layer.
+% 
+% The persistence layer provides an API allowing to work with a standard set of objects representing the model.
+The interaction with the underlying RDF store is made via RDF2Go, an abstraction over triple (and quad) stores, which allows to program against RDF2Go interfaces and choose any RDF store implementation.
 This allows having a completely extensible and configurable framework for storage mechanisms, inferencers, RDF file formats, query result formats and query languages.
 
+% Conceptual model subsection
 \input{paper_ssec_ontology}
 
+% Discovery and recommendations subsection
+\input{paper_ssec_discovery}
+
+% Conceptual model subsection
 \input{paper_ssec_linked_data}

### 2011-01-28 trunk/papers/webist2011/paper_sec_introduction.tex
%!TEX root = paper.tex
\subsection{Conceptual Model}
\label{ssec:ontology}
The conceptual model used to define the web service wrappers within this application has been influenced by WSDL, the main web service definition language, and semantic approaches such as WSMO and OWL-S. It belongs to a more complex conceptual model for FAST, which can be grouped into three levels: the gadget and screenflow level at the top, the level of individual screens in the middle, and the level of web services at the bottom. 
In short, a screenflow is an executable (in a Web browser) set of screens, while a scree

@@
 Business-to-business (B2B) integration is still a significant challenge, often requiring extensive efforts in terms of different aspects and technologies for protocols, architectures or security.
-Solutions based on open standards have reduced the complexity of integrating different business applications between different companies and partners.
-In this sense, Web services and XML standards such as RosettaNet, ebXML or OAG have been a boon to the world of B2B.
+Solutions based on open standards
+such as RosettaNet, ebXML, the Web Service Description Language (WSDL), Universal Description, Discovery and Integration (UDDI) or the Simple Object Access Protocol (SOAP)
+have reduced the complexity of integrating different business applications between different companies and partners. 
+% In this sense, Web services and XML standards such as RosettaNet, ebXML or OAG have been a boon to the world of B2B.
 By web services standards we are referring to the following open standards: Web Services Description Language (WSDL --- to describe), Universal Description, Discovery and Integration (UDDI --- to advertise and syndicate), Simple Object Access Protocol (SOAP --- to communicate) and Web Services Flow Language (WSFL --- to define work flows).
-A common scenario  in their usage is a company using WSDL to describe its web services, publishing them through a repository using UDDI, while these web services use SOAP-based messages to achieve dynamic integration between different, disparate applications.
+% A common scenario  in their usage is a company using WSDL to describe its web services, publishing them through a repository using UDDI, while these web services use SOAP-based messages to achieve dynamic integration between different, disparate applications.
 
-While the adoption of these standards has offered some advantages in business-to-consumer (B2C) integration as well,  the interaction and consumption of web services still requires programming skills and a deep understanding of the  technology, which poses an obstacle for an end-user with limited knowledge on the matter (in the sense as defined in~\cite{fuchsloch2010}).
+While the adoption of these standards has offered some advantages in business-to-consumer (B2C) integration as well,  the interaction and consumption of web services still requires programming skills and a deep understanding of the  technology, which poses an obstacle for an end-user with limited knowledge on the matter.% (in the sense as defined in~\cite{fuchsloch2010}).
 Regarding the discovery of the right service for the right task --- a crucial requirement for the dynamic use of WS --- most publishing platforms are syntax-based, %leading to poor precision and recall in finding the correct service,
@@
 We are targeting users which are non-skilled in term of programming and software development, empowering them with a platform to discover and consume web services in a straight-forward manner.
-The web services and their specifications are not published right away in the platform, since this would simply mean to reimplement UDDI.
+The web services and their specifications are not published right away in the platform, since this would simply mean to re-implement UDDI.
 However, their definition and behaviour are wrapped in what we call a ``web service wrapper''.
@@
 
-The remainder of this paper is organised as follows.
+% The remainder of this paper is organised as follows.
 Section~\ref{sec:related_work} covers the state of the art in web services publication, discovery and wrapping/consumption.
 In Section~\ref{sec:wrapping_web_services} we present our approach for web services wrapping, available possibilities and limitations.
-Next, Sect.~\ref{sec:discovery} describes our platform for publishing web services and empowering end-users to discover services.
+Next, Section~\ref{sec:discovery} describes our platform for publishing web services and empowering end-users to discover services.
 Section~\ref{sec:use_case} gives the reader a clearer use case and how this work is integrated into the FAST platform, and finally, we present our main conclusions and highlight future lines of research in Section~\ref{sec:conclusions}.

### 2011-01-28 trunk/papers/webist2011/paper_sec_related_work.tex
power shortening and improving
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1208 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-\section{Related work}
+\section{Related Work}
 \label{sec:related_work}
@@
 
-A different approach for web service discovery is the Web Services Dynamic Discovery (WS-Discovery) specification~\cite{beatty2005}.
+% A different approach for web service discovery is the Web Services Dynamic Discovery (WS-Discovery) specification~\cite{beatty2005}.
 The core of this approach is a multicast discovery protocol.

### 2011-01-28 trunk/papers/webist2011/paper_sec_use_case.tex
adapted to WEBIST format
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@980 d8b0c59b-6544-0410-94d9-83a5658a058e
power shortening and improving
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1208 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-06-14 trunk/papers/icsoc2010/paper_sec_use_case.tex
@@
 
-\section{Use case}
+\section{Context: The FAST Platform}
 \label{sec:use_case}
 
-The intention of this paper was to show the advances made regarding web service discovery and consumption through the two artifacts developed as part of our research.
+The intention of this paper is to demonstrate the advances made regarding web service discovery and consumption through two artifacts developed as part of our research: the \emph{publishing and discovery platform} and the \emph{service wrapper tool}.
 While these artifacts may be deployed and used separately by third-party applications, they were originally developed to form the backbone of the FAST platform~\cite{hoyer2009fast}.
-FAST constitutes a novel approach to application composition and business process definition from a top-down user-centric perspective.
-Therefore, it is aimed at allowing users (mainly business users at the management level, without programming experience) to create their own situational applications by visually combining different components, or building blocks, such as graphical forms and back-end services in terms of their inputs and outputs.
-These two components, as described in Sects.
-3 and 4, are fully integrated and present a high cohesion with the rest of the components of the FAST architecture (see Fig.~\ref{fig:fast_architecture}).
+FAST constitutes a novel approach to application composition and business process definition from a user-centric perspective.
+It is aimed at allowing users without previous programming experience to create their own situational applications by visually combining different \emph{building blocks}, such as graphical forms and back-end services, based on their inputs and outputs (or \emph{pre-} and \emph{post-conditions}).
+The work covered in this paper are the components highlighted and surrounded by a dashed frame in the Fig.~\ref{fig:fast_architecture}.
+Communication within the platform is mostly done via a RESTful API, using JSON as an exchange syntax.
+%These components have public interfaces to communicate with any application developed by a third-party developer or company, and as described in Sects.
+3 and 4, are fully integrated and present a high cohesion with the rest of the components of the FAST architecture.
 
@@
   \begin{center}
-    \includegraphics[width=.9\linewidth]{images/fast_architecture.png}
+    \includegraphics[width=\linewidth]{images/fast_architecture.pdf}
     \caption{Overview of the FAST architecture}
@@
 
-In order to facilitate a better understanding of the use case, we will describe a number of concepts related to FAST in this section.
-A \emph{gadget} is the end product of the platform, aimed to be used in a specific mashup platform.
+In order to facilitate a better understanding of these components, we describe a number of concepts related to FAST in this section. 
+A \emph{gadget} is the end product of the platform, ready to be run in any ordinary web browser, usually through a mashup platform. 
 Within FAST, the platform-independent gadget is called \emph{screenflow}, which comprises a set of domain-specific \emph{screens} connected through their \emph{pre-} and \emph{post-conditions}.
 A screen is the most complex building block fully functional by itself.
+Visually, it is similar to a tab in a tabbed application.
 It is composed by a \emph{form} conveying the graphical interface, and a set of \emph{operators} and \emph{back-end services}, wrapped into so-called \emph{service resources}.
 
-The web service wrapper tool is then used to create the formal definitions of back-end services, transforming them into building blocks to be stored and reused within the FAST platform.
-The tool is integrated directly into the graphical IDE, and for the storage interacts with the catalogue, where the formal service descriptions and the executable code encapsulating the service behaviour is stored.
-
-The publishing platform presented, called ``Metadata Catalogue'' within the FAST platform, covers several important purposes:
-\begin{description}
-	\item[Publication and search of gadgets] The output of the FAST IDE are gadgets, which are self-contained and ready to use.
-They need to be made available for download, as well as being discovered according to their formal description.
-	\item[Publication and search of building blocks] Every single component or building block definition is stored in the catalogue, which facilitates searching, recommendation and planning (the suggestion of combinations of building blocks to fulfil certain requirements).
-	\item[Storage and indexing of user profiles and histories] Each user of the FAST platform has a representation in the catalogue, thus making it possible for the IDE to make informed guesses about what they want in a particular situation, based on their previous behaviour.
-	\item[Facilitate ontology mediation] Because FAST allows the integration of third-party external services, it is not always guaranteed that all the data coming from a service is directly compatible with the FAST platform.
-In such cases, mediation between the different ontologies is necessary.
-The catalogue provide some grade of mediation functionality (which is not covered in this paper).
-\end{description}
+The service wrapper tool is used to create wrappers for third-party web services, transforming them into building blocks to be stored and reused within the FAST platform.
+The publishing platform presented, also called \emph{catalogue} within the FAST platform, covers several important purposes, such as storage, indexing, publication and search of gadgets, gadget building blocks and user profiles, as well as some level of ontology mediation to facilitate the integration of services from different parties (first steps in this direction are outlined in \cite{Ambrus:2009it}).

### 2011-01-28 trunk/papers/webist2011/paper_sec_wrapping.tex
power shortening and improving
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1208 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-\section{Wrapping web services}
+\section{Wrapping Web Services}
 \label{sec:wrapping_web_services}
@@
 
-\subsection{Constructing service requests} % (fold)
+\subsection{Constructing Service Requests} % (fold)
 \label{sub:constructing_service_requests}
@@
 In addition, other header data may be needed to transfer e.g.
-session ids.) As an example, the Ebay Shopping web service will be studied.
+session ids).
+As an example, the Ebay Shopping web service will be studied.
 As documented on the corresponding web site, the following URL is invoked to retrieve a list of items corresponding to the search keywords \emph{USB}:
@@
 	\item[version] the API version.
-	\item[callname] in this case FindItems to search through all items in Ebay.
-	\item[itemsort] sorting method for the list of items.
-	\item[querykeywords] list of keywords.
+	\item[callname] in this case \emph{FindItems} to search through all items in Ebay.
+	\item[ItemSort] sorting method for the list of items.
+	\item[QueryKeywords] list of keywords.
 	\item[responseencoding] format of the response message obtained by the invocation of the request.
@@
 
-To define the desired input parameters and the type of the expected result of a new service wrapper the service wrapper tool provides a web based form that allows the editing of these entries, see Fig.~\ref{fig:construct_pre_post_conditions}.
+To define the desired input parameters and the type of the expected result of a new service wrapper the service wrapper tool provides a web-based form that allows the editing of these entries, see Fig.~\ref{fig:construct_pre_post_conditions}.
 Note, we allow to edit example values for the input parameters.
@@
 
-The service wrapper tool composes the request using a template string which will contain placeholders for input values.
+% The service wrapper tool composes the request using a template string which will contain placeholders for input values.
 Before sending the request to the service, the placeholders are replaced with their corresponding values and the resulting URL is then ready to be sent as the service request.
@@
 An example of service request construction can be found in the Figure~\ref{fig:construct_service_request}.
-In the top input field the user may drop an example HTTP request taken e.g.
-form the service documentation e.g.
-from its website\footnote{http://developer.ebay.com/products/shopping/}.
+In the top input field the user may drop an example HTTP request, taken e.g.
+from the service's documentation\footnote{\url{http://developer.ebay.com/products/shopping/}}.
 The tool analyses the example request and in the middle of the screen a form for editing the request parameters is provided.
@@
 
-Below the request parameter editing form of Figure ~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed URL by sending it to the specified service address (via a server relay).
-Then the placeholders for input parameters are replaced by their example values and the resulting http request is shown below the parameter form.
-In addition, the request is sent and the response is shown on the bottom of that page.
-This gives the user a fast feedback whether the constructed request works as desired.
+Below the request parameter editing form of Figure~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed URL by sending it to the specified service address (via a server relay).
+The service's response is shown on the bottom of the page. This gives the user a fast feedback whether the constructed request works as desired.
 
@@
 
-\subsection{Interpreting service responses} % (fold)
+\subsection{Interpreting Service Responses} % (fold)
 \label{sub:interpreting_service_responses}
 
-The service request is constructed and sent to the service provider, who will then send back a response.
-This response message could be serialised in any format, though the most common formats used nowadays are XML or JSON.
-To continue the example started in the previous section, we assume the response of the Ebay Shopping service will be in XML as specified in the request.
+A web service can choose to send a response message in any desired structured format. In our example, we assume an XML tree, but a JSON structure is equally possible.
 
-\subsubsection{Translation of XML into Facts} % (fold)
+\subsubsection{Translation of the Response Tree into Facts} % (fold)
 \label{ssub:translation_xml_into_facts}
@@
 
-Once the service response, in XML format, has been retrieved, the transformation tab of the wrapper tool shows it as an interactive object tree, as seen on the left side of Fig.~\ref{fig:response_service_execution}.
-To construct this interactive object tree, the XML document has been parsed into a DOM, and a simplified tree representation of that DOM is built up.
+Once the service response has been retrieved, the transformation tab of the wrapper tool shows it as an interactive object tree, as seen on the left side of Fig.~\ref{fig:response_service_execution}.
 
-A transformation rule is used to analyse the XML data and to generate domain-specific objects from concepts from the ontologies used by the parameters of the different building blocks.
+A transformation rule is used to analyse the response data and to generate domain-specific objects from concepts from the ontologies used by the parameters of the different building blocks.
 A transformation rule is composed of three elements, as seen in the middle part of Fig.~\ref{fig:response_service_execution}.
-First, the \textit{from} field indicates the XML elements to be translated by the rule.
-These XML elements are identified by the tag name of a DOM element from the XML document.
+First, the \textit{from} field indicates the source elements to be translated by the rule.
 Second, the type of the rule will be set, taking one of the following values: \emph{createObject}, \emph{fillAttributes} or \emph{dummy}.
@@
 A detailed explanation of the type of actions to be trigger from the transformation rules is as follows:
-\begin{description}
-	\item[action \emph{createObject}] specifies the creation of a new domain object.
+
+\paragraph{createObject} % (fold)
+\label{par:createobject}
+
+specifies the creation of a new domain object.
 The type of that new object is provided in the third compartment.
@@
 The resulting objects are shown in a facts tree in the right of Figure ~\ref{fig:response_service_execution}.
-	\item[action \emph{fillAttributes}] does not create a new object but it fills the value of the attribute provided as third part of such rules.
+
+% paragraph createobject (end)
+
+\paragraph{fillAttributes} % (fold)
+\label{par:fillattributes}
+
+does not create a new object but it fills the value of the attribute provided as third part of such rules.
 In our example, the third transformation rule searches for XML elements with tag name \emph{Title}.
@@
 The value of that \emph{Title} element is then transfered to the \emph{productName} attribute of the corresponding \emph{Product} object.
-Actually, our \textit{from} fields allows also to refer to parts of an XML attribute e.g.
-to \textit{words} 1 through 3.
-It is also possible to combine constant text and elements of multiple XML tree elements.
-	\item[action \emph{dummy}] does not create or modify any objects but such rules are just used to narrow the search space for their sub-rules.
+
+% paragraph fillattributes (end)
+
+\paragraph{dummy} % (fold)
+\label{par:dummy}
+
+does not create or modify any objects but such rules are just used to narrow the search space for their sub-rules.
 For example, in Amazon product data, the XML data for an item contains sections for \emph{minimum price}, \emph{maximum price}, and \emph{average price}.
@@
 Using a dummy rule, we may first search for \emph{minimum price} elements and then search for \emph{formatted price} elements within that sub-tree.
-\end{description}
+
+% paragraph dummy (end)
 
@@
 For the \emph{fillAttributes} rules, the selection box shows the attributes of the object type chosen in the parent rule.
-In addition, we may provide some analysis tool, which will help to guarantee that the objects generated by the transformation rules conform to the object types defined in the corresponding conceptual model (see Section~\ref{ssec:ontology}).
+% In addition, we may provide some analysis tool, which will help to guarantee that the objects generated by the transformation rules conform to the object types defined in the corresponding conceptual model (see Section~\ref{ssec:ontology}).
 This helps to ensure that the objects generated by the designed service wrapper will be compatible for input parameters of subsequent filter steps and or gadgets.

### 2011-01-28 trunk/papers/webist2011/paper_ssec_discovery.tex
power shortening and improving
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1208 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-\subsection{Discovery mechanisms} % (fold)
+\subsection{Discovery Mechanisms} % (fold)
 \label{ssec:discovery}
 
-The platform presented in this paper is intended to support web service discovery and recommendation capabilities.
-As explained in detail in Section~\ref{ssec:ontology}, there are different types of building blocks which can be modelled
-using the platform, leading to a necessity for different kinds of discovery or recommendation mechanisms, depending
-on the level of composition to deal with: screen-flow or screen. In the screen-flow composition, the smallest
-functional unit is the screen, meaning that a screen-flow is composed by interconnecting a number of screens between 
-each other. This composition can be applied to any sort of building block since they all share the
-same structure: a set of pre- and post-conditions, which make possible the communication with other building blocks.
-That said, technically the algorithms presented in this section could be used for the discovery and composition of
-any type of building block stored in the platform.
+% Due to the diversity of building blocks available, there is a necessity for different kinds of discovery or recommendation mechanisms, depending on the level of composition: screen-flow or screen.
 
-At the screen-flow design level, the composition is made using a set of screens, and the pre- and post-conditions 
-of the screen-flow itself used as entry and exit points of the screen-flow. A screen is only reachable when its 
-pre-conditions are satisfied as well, meaning that the post-conditions of another screen, or the pre-conditions of 
-the screen-flow are compatible with them.
+The main goal of the discovery process is to aid the user in finding suitable building blocks to complement the ones they are already using.
+E.g., on the screen-flow level, this would mean to suggest additional screens to make existing screens within a screen-flow reachable, and the screen-flow therefore executable.
+The platform offers two mechanisms to find and recommend screens (or other building blocks) stored in the catalogue: a simple discovery based on pre- and postconditions, and a multi-step discovery or planning algorithm.
+Before being presented to the user, the results are being ranked, as discussed in Sect.~\ref{sssec:ranking}.
 
-The main goal of the discovery process is to help the user to make the screen-flow executable, in other words, to make
-all the screens within a screen-flow reachable. The platform offers two mechanisms to find and recommend
-screens, previously stored in the building block base (the catalogue), which will make the screen-flow executable: a
-simple discovery based on pre- and postconditions, and a multi-step discovery or planning algorithm. Like any other
-search engine or recommender system, the results are ranked before being presented
-to the user. The ranking algorithm is covered in Section~\ref{sssec:ranking}.
-
-\subsubsection{Simple discovery based on pre- and post-conditions}
+\subsubsection{Simple Discovery Based on Pre- and Post-Conditions}
 \label{sssec:simple_discovery}
 
-At a particular state of a composition, ie. at screen-flow design, several of the building blocks used to compose it, 
-screens in the case of a screen-flow, might be unreachable, therefore there must be pre-conditions not satisfied. The
-platform will assist the process by recommending building blocks which will satisfy these pre-conditions. In a nutshell,
-it collects the pre-conditions of all the unreachable building blocks, used as a graph pattern (see Section~\ref{ssec:ontology})
-in order to be matched against the post-conditions of any building block. In the following scenario, there are two 
-screens: \emph{s1} and \emph{s2}. \emph{s1} has as a pre-condition: ``there exists a search criteria'', and as a 
-post-condition: ``there exists a item''. \emph{s2} has just a pre-condition stating: ``there exists a search criteria''.
-Here you can see the same example but formally defined:
+In this simple approach, the platform will assist the process by recommending all building blocks which will satisfy currently unfulfilled pre-conditions. 
+The pre-conditions of all the unreachable building blocks are collected as a graph pattern, which is then matched against the post-conditions of all available building blocks. 
+In the following scenario, there are two 
+screens: \emph{s1} and \emph{s2}. \emph{s1} has as a pre-condition: \emph{``there exists a search criteria''}, and as a 
+post-condition: \emph{``there exists a item''}. \emph{s2} has just a pre-condition stating: \emph{``there exists a search criteria''}.
 

### 2011-01-28 trunk/papers/webist2011/paper_ssec_ontology.tex
power shortening and improving
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1208 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-09-20 trunk/papers/webist2011/paper_ssec_ontology.tex
@@
 \label{ssec:ontology}
-The conceptual model used to define the web service wrappers within this application has been influenced by WSDL, the main web service definition language, and semantic approaches such as WSMO and OWL-S.
-It belongs to a more complex conceptual model for FAST, which can be grouped into three levels: the gadget and screenflow level at the top, the level of individual screens in the middle, and the level of web services at the bottom.
-In short, a screenflow is an executable (in a Web browser) set of screens, while a screen is a set of building blocks with a specific purpose.
-These building blocks --- \emph{forms} for the UI, \emph{operators} for data manipulation and \emph{back-end service} --- all share the same structure.
-Every building block has a set of actions (operations in WSDL), each of which needs a set of pre-conditions (inputs) to be fulfilled in order to be executed, and provides a set of post-conditions (outputs) to other building blocks.
-These pre- and post-conditions are defined as individual RDF triples.
-This means that conditions can be modelled as graph patterns using SPARQL notation \cite{sparql2008spec}, where e.g.
-the post-condition of a login service such as ``there exists a user'' will be transformed into a simple pattern like ``a variable \texttt{?user} is of type \texttt{sioc:User}'', or ``\texttt{?user a sioc:User}''.
+The conceptual model used to define the web service wrappers within this application has been influenced by both WSDL and semantic approaches such as WSMO and OWL-S. 
+It is part of a more complex conceptual model for FAST~\cite{moeller2010fast_ontology}, which can be grouped into three levels: the gadget and screenflow level at the top, the level of individual screens in the middle, and the level of web services at the bottom.
+All these building blocks and their sub-parts
+% --- \emph{forms} for the UI, \emph{operators} for data manipulation and \emph{back-end services} --- 
+share the same structure: a set of actions (like operations in WSDL), each of which needs a set of pre-conditions (inputs) to be fulfilled in order to be executed, and provides a set of post-conditions (outputs) to other building blocks.
+These pre- and post-conditions are defined as RDF graph patterns.
+E.g., the post-condition of a login service such as ``there exists a user'' will be expressed as a simple pattern such as ``\texttt{?user a sioc:User}''\footnote{We use Trig (\url{http://www4.wiwiss.fu-berlin.de/bizer/TriG/}) and SPARQL (\url{http://www.w3.org/TR/rdf-sparql-query/}) notation for RDF graphs throughout this paper.}.
 Using this mechanism, extended by RDFS entailment rules, services and other building blocks can be matched automatically.
-The publishing and discovery platform (catalogue) can employ this functionality to support the discovery of web services based on the current user needs (expressed in the same way as pre-conditions).
-For further details on the conceptual model, see its definition in \cite{moeller2010fast_ontology}.
+The publishing and discovery platform can employ this functionality to support the internal discovery of web services based on the current user needs (expressed in the same way as pre-conditions).

### 2011-01-28 trunk/papers/webist2011/abstract.tex
super-mega shortening done. Down to 6 pages from 10 1/2
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1211 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-B2B systems integration was revolutionised by the introduction of web services,
-transforming the way enterprise systems communicate and strengthening the  relationships with providers and customers.
-The main advantage seen by companies was an increase of operational efficiencies, and a reduction of costs.
-In this scenario, highly qualified software developers are responsible for the integration of services with other systems, through the analysis of formal service descriptions or human-readable specifications.
+In B2B systems integration and web services, many companies see the advantage of increased operational efficiencies and a reduction of costs. 
+In this scenario, highly qualified software developers are responsible for the integration of services with other systems.
 However, this model fails when targeting the long tail of enterprise software demand, the end-users. 
@@
 \begin{inparaenum}[(a)]
-	\item a method to generate ready-to-use web service wrappers, and 
-	\item a catalogue which users can browse to search for web services fitting their needs. 
+	\item a catalogue which users can browse to search for web services fitting their needs, and
+	\item a method to generate ready-to-use web service wrappers to use in the catalogue.
 \end{inparaenum}

### 2011-01-28 trunk/papers/webist2011/paper_sec_conclusions.tex
super-mega shortening done. Down to 6 pages from 10 1/2
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1211 d8b0c59b-6544-0410-94d9-83a5658a058e

%!TEX root = paper.tex

\section{Conclusions and future work}
\label{sec:conclusions}

It has been argued that adopting web services standards helps enterprises to increase operational efficiency, reduce costs and strengthen the relations with partners. 
A range of WS standards help in this regard.
% Many of these standards have been widely adopted, such as WSDL, XML and UDDI, and many companies offer access to their information and operational systems through web services. 
However, when dealing with end-users, the process for publication and consumption is not well supported. 
% The common way of publishing services is by providing some sort of definition on their websites, such a WSDL document and further details in a human-readable HTML page. 
% Discovery is performed through a search engine or aggregator, mainly in a syntax-based way. For consumption, in most cases, consultants with good programming skills are required. 
In a move towards improving this situation, the work presented in this paper empowers the end-user with a platform (the \emph{catalogue}) to easily select services based on functional behaviour (pre-/post-conditions) and other metadata, being able to download a so-called resource adapter allowing the consumption of web services using standard languages to execute within a web browser, and a tool to transform, in an interactive manner, formal definitions of web services into these resource adapters, ready for being published into the catalogue.

The current version of the wrapping tool permits to create resource adapters for RESTful web services.
SOAP-based services are also experimentally supported. As potential future work, we consider to include semantically enriched WSDL documents using SAWSDL, and to support other SWS approaches such as WSMO-lite services --- however, real-life examples of these are still hard to find.

% Another limitation of the platform is inherent in web client technologies and the cross-domain policy for security. This is solved within FAST using platform-dependent API calls. Hence, the code generated makes use of the FAST API, which will be transformed depending on the mashup platform the user intends to deploy the gadget. To avoid these issues, and to be able to provide platform-independent code right away, we are planning to deploy the JSONRequest approach as proposed by \cite{crockford2006} and other solutions.

### 2011-01-28 trunk/papers/webist2011/paper_sec_introduction.tex
super-mega shortening done. Down to 6 pages from 10 1/2
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1211 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 Business-to-business (B2B) integration is still a significant challenge, often requiring extensive efforts in terms of different aspects and technologies for protocols, architectures or security.
-Solutions based on open standards
+While the adoption of open standards 
 such as RosettaNet, ebXML, the Web Service Description Language (WSDL), Universal Description, Discovery and Integration (UDDI) or the Simple Object Access Protocol (SOAP)
-have reduced the complexity of integrating different business applications between different companies and partners. 
+have reduced the complexity of integrating business applications between different companies and partners,
 % In this sense, Web services and XML standards such as RosettaNet, ebXML or OAG have been a boon to the world of B2B.
@@
 % A common scenario  in their usage is a company using WSDL to describe its web services, publishing them through a repository using UDDI, while these web services use SOAP-based messages to achieve dynamic integration between different, disparate applications.
-
-While the adoption of these standards has offered some advantages in business-to-consumer (B2C) integration as well,  the interaction and consumption of web services still requires programming skills and a deep understanding of the  technology, which poses an obstacle for an end-user with limited knowledge on the matter.% (in the sense as defined in~\cite{fuchsloch2010}).
-Regarding the discovery of the right service for the right task --- a crucial requirement for the dynamic use of WS --- most publishing platforms are syntax-based, %leading to poor precision and recall in finding the correct service,
+and offered some advantages in business-to-consumer (B2C) integration as well, the interaction and consumption of web services still requires programming skills and a deep understanding of the  technology, which poses an obstacle for an end-user with limited knowledge of the matter.
+% (in the sense as defined in~\cite{fuchsloch2010}).
+Regarding the selection of the right service for the right task --- a crucial requirement for the dynamic use of web services --- most publishing platforms are syntax-based, %leading to poor precision and recall in finding the correct service,
 making it difficult to navigate through a large number of web services~\cite{pilioura_acm2009}, 
-therefore preventing end-users from performing these tasks. 
-Solutions such as Semantic Web Services (SWS) promised many advantages in discovery, composition and consumption of web services, independently of the provider's platform, location, service implementation or data format.
+preventing end-users from performing these tasks. 
+Solutions such as Semantic Web Services (SWS) promised many advantages in this respect.
+% discovery, composition and consumption of web services, independently of the provider's platform, location, service implementation or data format. 
 However, as of yet they have not been widely adopted, 
-possibly because the perceived potential benefits of SWS did not justify the additional investments required for businesses~\cite{shi2007}.
-In a similar vein as B2C, governments are currently opening their data to the public as linked data. 
-So far, the efforts are restricted to data only, so that users are left alone in finding applications of that data. 
-A natural further development in the public sector would be that their services for administration tasks and single-window applications will be opened as well.
+possibly because the perceived potential benefits did not justify the additional investments~\cite{shi2007}.
+% In a similar vein as B2C, governments are currently opening their data to the public as linked data. 
+% So far, the efforts are restricted to data only, so that users are left alone in finding applications of that data. 
+% A natural further development in the public sector would be that their services for administration tasks and single-window applications will be opened as well.
 Hence, governments will benefit from the research done around these concepts and technologies.
@@
 The motivation of our work has been strongly influenced by the end-users' needs.
-We are targeting users which are non-skilled in term of programming and software development, empowering them with a platform to discover and consume web services in a straight-forward manner.
-The web services and their specifications are not published right away in the platform, since this would simply mean to re-implement UDDI.
-However, their definition and behaviour are wrapped in what we call a ``web service wrapper''.
-The products resulting from the wrapping process are two artifacts: a specific definition of the web service to be used by this tool, and a ready-to-consume (inside a web browser) piece of code with the proper functions to invoke the web service.
+We are targeting users which are non-skilled in programming and software development, empowering them with a platform to select and consume web services in a straight-forward manner.
+Rather than publishing services directly in our platform, we leave existing third-party services untouched and instead integrate them through service wrappers.
+The products resulting from the wrapping process are two artifacts: a specific definition of the web service to be used by this tool, and a ready-to-use piece of code with the proper functions to invoke the web service.
 

### 2011-01-28 trunk/papers/webist2011/paper_sec_related_work.tex
super-mega shortening done. Down to 6 pages from 10 1/2
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1211 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-Web services have been around for a long time since they first appeared.
+Web services have been around for a long time.
 One of the most important claims about their benefits has been (syntactic) interoperability between third-party systems and applications based on different platforms / programming languages. 
-System integration, within a company or between systems from different enterprises, became easier with the adoption of web service standard technologies, in particular the Web Service Description Language (WSDL), the most extensive standard used for the definition of web services.
-Many integrated development environments (IDEs) can deal with WSDL to facilitate the integration task, however it is still required that developers read and understand their specifications, and implement part of the logic programmatically.
+System integration, within a company or between systems from different enterprises, became easier with the adoption of web service standard technologies such as WSDL. 
+While many integrated development environments (IDEs) can deal with WSDL to facilitate the integration task, trained developers are still required for this.
+% however it is still required that developers read and understand their specifications, and implement part of the logic programmatically. 
 As a step forward, there are several tools which facilitate the interaction with data sources and services on the Web.
 Yahoo!
-Pipes, Apatar, JackBe Presto, Microsoft Popfly (now defunct) and NetVibes, among others provide a set of modules to access different kind of data sources, such as RSS feeds, a given web page (HTML code), Flickr images, Google base or the Yahoo!
-search engine, databases (MySQL, PostgreSql, Oracle), and powerful enterprise systems such as Salesforce CRM, SugarCRM and Goldmine CRM.
-However, none of the solutions in the market really facilitates end-users to build  their own applications or widgets allowing the interaction with web services created by third-party providers.
+Pipes, Apatar, JackBe Presto and NetVibes, among others provide a set of modules to access different kind of data sources, such as RSS feeds, a given web page (HTML code), Flickr images, databases, and even powerful enterprise systems such as Salesforce CRM or Goldmine CRM.
+However, none of these solutions facilitate end-users to build their own applications allowing the interaction with web services created by third-party providers.
 The solutions found are mainly data-oriented (RSS feeds, databases, raw text).
 Several tools permits some sort of (web) service integration, but are meant to be used within an enterprise level by savvy business users or developers.
-The solution presented in this paper leverage the possibility of integrating REST or SOAP-based web service inside visual applications running in a browser (e.g.
-widgets) by providing a platform to create, publish and discover web services wrappers.
+The solution presented in this paper leverages the possibility of integrating RESTful or SOAP-based web services inside browser-based applications (i.e., widgets or gadgets) by providing a platform to create, publish and select web services wrappers.
 
-In the context of publishing and discovering Web services, service providers have well-known and widely used technologies to accomplish the task of publication, such as the Universal Description, Discovery and Integration (UDDI) or WS-Discovery.
-The UDDI service registry specification~\cite{uddi2004} is currently one of the core standards in the Web service technology stack and an integral part of every major SOA vendor's technology strategy and offers to requesters the ability to discover services via the Internet.
-In short, UDDI serves as a centralised repository of WSDL documents.
+In the context of publishing and discovering Web services, service providers have well-known and widely used technologies to accomplish the task of publication, such as the Universal Description, Discovery and Integration (UDDI)~\cite{uddi2004}.
+UDDI serves as a centralised repository of WSDL documents.
 A similar concept is iServe~\cite{pedrinaci_ores2010}.
-This platform aims to publish web services as what they called Linked Services --- linked data describing services ---, storing web service definitions as semantic annotations, so that other semantic web-aware applications may take advantage of it.
-However, the platform does not deal with the step from the definition to the consumption of the services.
+This platform aims to publish web services as what they called Linked Services --- linked data describing services ---, storing web service definitions as semantic annotations, so that other semantically aware applications may take advantage of it.
+However, the platform does not handle the step from definition to consumption of the services.
 

### 2011-01-28 trunk/papers/webist2011/paper_sec_use_case.tex
super-mega shortening done. Down to 6 pages from 10 1/2
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1211 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 While these artifacts may be deployed and used separately by third-party applications, they were originally developed to form the backbone of the FAST platform~\cite{hoyer2009fast}.
-FAST constitutes a novel approach to application composition and business process definition from a user-centric perspective.
+FAST constitutes a novel approach to application composition from a user-centric perspective.
 It is aimed at allowing users without previous programming experience to create their own situational applications by visually combining different \emph{building blocks}, such as graphical forms and back-end services, based on their inputs and outputs (or \emph{pre-} and \emph{post-conditions}).
-The work covered in this paper are the components highlighted and surrounded by a dashed frame in the Fig.~\ref{fig:fast_architecture}.
+The work covered in this paper are the components highlighted in Fig.~\ref{fig:fast_architecture} by a dashed line.
 Communication within the platform is mostly done via a RESTful API, using JSON as an exchange syntax.
@@
 
-In order to facilitate a better understanding of these components, we describe a number of concepts related to FAST in this section. 
+In order to provide a better understanding of these components, we describe a number of concepts related to FAST in this section. 
 A \emph{gadget} is the end product of the platform, ready to be run in any ordinary web browser, usually through a mashup platform. 
-Within FAST, the platform-independent gadget is called \emph{screenflow}, which comprises a set of domain-specific \emph{screens} connected through their \emph{pre-} and \emph{post-conditions}.
-A screen is the most complex building block fully functional by itself.
-Visually, it is similar to a tab in a tabbed application.
+An undeployed gadget is also called \emph{screenflow}, which comprises a set of \emph{screens} connected through their \emph{pre-} and \emph{post-conditions}.
+A screen is the most complex building block fully functional by itself,
+visually similar to a tab in a tabbed application.
 It is composed by a \emph{form} conveying the graphical interface, and a set of \emph{operators} and \emph{back-end services}, wrapped into so-called \emph{service resources}.
 
+The publishing platform, also called \emph{catalogue}, covers several important purposes, such as storage, indexing, publication and search of gadgets, gadget building blocks and user profiles.
+% , as well as some level of ontology mediation to facilitate the integration of services from different parties (first steps in this direction are outlined in \cite{Ambrus:2009it}).
 The service wrapper tool is used to create wrappers for third-party web services, transforming them into building blocks to be stored and reused within the FAST platform.
-The publishing platform presented, also called \emph{catalogue} within the FAST platform, covers several important purposes, such as storage, indexing, publication and search of gadgets, gadget building blocks and user profiles, as well as some level of ontology mediation to facilitate the integration of services from different parties (first steps in this direction are outlined in \cite{Ambrus:2009it}).

### 2011-01-28 trunk/papers/webist2011/paper_sec_wrapping.tex
super-mega shortening done. Down to 6 pages from 10 1/2
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1211 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-In our approach, wrapping web services is done in two steps: a first step is in charge of the construction of a service request and a second step analyses the response received from the execution of the service, allowing the analysis of response data and the mapping to domain-specific concepts.
-From this input, we generate JavaScript code to be embedded into web gadgets.
-This JavaScript code takes query parameters, constructs the service request, analyses the service response and returns result data in the desired format.
+Before third-party services can be used by our platform, they need to be provided with a service wrapper. In our approach, this is done in two steps: 
+\begin{inparaenum}[(i)]
+	\item constructing an exemplary service request, and 
+	\item analysing the response received from the execution of the service, allowing the mapping of the response data to domain-specific concepts to be used as pre- and post-conditions.
+\end{inparaenum}
+From this input, the tool then generates the actual wrapper --- JavaScript code to be embedded into web gadgets.
 
@@
 
-In this paper we illustrate the interaction with services on simple GET requests as supported e.g.
-by REST services.
-Support for other services like POST based services is under construction.
-In case of a simple GET request, a service request is assembled using a certain URL and a set of parameters.
-(A POST request would have a text block to be posted.
-In addition, other header data may be needed to transfer e.g.
-session ids).
-As an example, the Ebay Shopping web service will be studied.
-As documented on the corresponding web site, the following URL is invoked to retrieve a list of items corresponding to the search keywords \emph{USB}:
+In this section we illustrate the interaction with RESTful services on simple GET requests\footnote{Other methods such as POST or SOAP-based services are also supported.}. 
+In this case, a service request is assembled using its URL and a set of parameters. As an example, we will look at the eBay Shopping web service. 
+To define the desired pre-condition and the type of post-condition of a new service wrapper, the service wrapper tool provides a form (not shown here).
+For the current example, we assume that the user has defined a pre-condition \emph{search\_key}.
 
-\begin{listing}
-\begin{verbatim}
+We are invoking the service to retrieve a list of items corresponding to the search keyword ``USB'':
+
+\begin{lstlisting}
 http://open.api.sandbox.ebay.com/shopping?
-  appid=KasselUn-efea-4b93-9505-5dc2ef1ceecd&
-  version=517&callname=FindItems&
-  ItemSort=EndTime&QueryKeywords=USB&
-  responseencoding=XML
-\end{verbatim}
-\end{listing}
+  appid=KasselUn-efea-xxx&version=517&
+  callname=FindItems&ItemSort=EndTime&
+  QueryKeywords=USB&ResponseEncoding=XML
+\end{lstlisting}
 
-As you may see, the URL invoked is \url{http://open.api.sandbox.ebay.com/shopping} and the parameters used in the example are:
-\begin{description}
-	\item[appid] this is the application ID obtained to use the API.
-	\item[version] the API version.
-	\item[callname] in this case \emph{FindItems} to search through all items in Ebay.
-	\item[ItemSort] sorting method for the list of items.
-	\item[QueryKeywords] list of keywords.
-	\item[responseencoding] format of the response message obtained by the invocation of the request.
-\end{description}
+% The parameters used in the example are:
+% \textbf{appid} (the application ID obtained to use the API),
+% \textbf{version} (the API version),
+% \textbf{callname} (in this case \emph{FindItems} to search through all items in eBay),
+% \textbf{ItemSort} (sorting method for the list of items),
+% \textbf{QueryKeywords} (list of keywords), and
+% \textbf{ResponseEncoding} (format of the response message obtained by the invocation of the request).
 
-The above URL for searching items in Ebay is followed by the query parameters, which take the form \textit{argument=value}.
-In this example, the only relevant parameter the end user may want to provide is \emph{querykeywords}.
-Thus the end user may use some input field to provide a query keyword and pass this value to our service wrapper operation.
-Other parameters can be set to a default value.
-However, in order to develop a more generic wrapper to a service, more or even all parameters might be passed by the user.
+% Note that we allow to edit example values for the input parameters, which can be used to test the service wrapper.
 
-To define the desired input parameters and the type of the expected result of a new service wrapper the service wrapper tool provides a web-based form that allows the editing of these entries, see Fig.~\ref{fig:construct_pre_post_conditions}.
-Note, we allow to edit example values for the input parameters.
-These example values may be used to test the service wrapper.
-
-\begin{figure}
-  \begin{center} \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithPortDefinitions.png}
-    \caption{Configuring parameters of a service wrapper}
-    \label{fig:construct_pre_post_conditions}
-  \end{center}
-\end{figure}
+% \begin{figure}
+%   \begin{center} \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithPortDefinitions.png}
+%     \caption{Configuring parameters of a service wrapper}
+%     \label{fig:construct_pre_post_conditions}
+%   \end{center}
+% \end{figure}
 
@@
 
-An example of service request construction can be found in the Figure~\ref{fig:construct_service_request}.
-In the top input field the user may drop an example HTTP request, taken e.g.
-from the service's documentation\footnote{\url{http://developer.ebay.com/products/shopping/}}.
-The tool analyses the example request and in the middle of the screen a form for editing the request parameters is provided.
-In our example, the user has connected the \textit{QueryKeywords} parameter with the input parameter \textit{search\_key} by adding a corresponding reference to the value field of that parameter.
-In addition, we have retrieved an access key which has been entered as value for the \textit{appid} parameter.
+The various parameters in the example are defined by the service provider.
+However, the most relevant one for our example is \emph{QueryKeywords}, which communicates to the services what we are looking for.
+An example of service request construction is shown in the Fig.~\ref{fig:construct_service_request}.
+In the ``Request'' input field the user enters an example HTTP request, e.g.
+the one above.
+The tool analyses the request and provides a form for editing the request parameters.
+In our example, the user has connected the service's \textit{QueryKeywords} parameter with the wrapper's \textit{search\_key} pre-condition.
+Requests constructed in this way can then be sent off to the service, whose response is shown in the bottom text area.
 
@@
     \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithRequestExample.png}
-    \caption{Constructing the service request URL and its parameters}
+    \caption{Constructing the service request}
     \label{fig:construct_service_request}
@@
 
-Below the request parameter editing form of Figure~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed URL by sending it to the specified service address (via a server relay).
-The service's response is shown on the bottom of the page. This gives the user a fast feedback whether the constructed request works as desired.
-
-\subsubsection{Limitations} % (fold)
-\label{ssub:limitations}
-
-It is worth pointing out that currently the wrapper tool is able to construct input ports for the wrappers using just basic types.
+% \subsubsection{Limitations} % (fold)
+% \label{ssub:limitations}
+% 
+% It is worth pointing out that currently the wrapper tool is able to construct input ports for the wrappers using just basic types.
 To construct URL requests from more complex input parameters, as e.g.
@@
 For example, the expression \emph{customer.address} may refer to the \emph{address} field of a \emph{customer} parameter.
-
-% subsubsection limitations (end)
+% 
+% % subsubsection limitations (end)
 
@@
 
-A web service can choose to send a response message in any desired structured format. In our example, we assume an XML tree, but a JSON structure is equally possible.
+% A web service can choose to send a response message in any desired structured format. In our example, we assume an XML tree, but a JSON structure is equally possible.
 
-\subsubsection{Translation of the Response Tree into Facts} % (fold)
-\label{ssub:translation_xml_into_facts}
+% \subsubsection{Translation of the Response Tree into Facts} % (fold)
+% \label{ssub:translation_xml_into_facts}
+
+Once the service response has been retrieved, the transformation tab of the wrapper tool shows it as an interactive object tree, as seen in Fig.~\ref{fig:response_service_execution}.
 
@@
     \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithTransformationRules.png}
-    \caption{Interactive, rule based transforming of an XML response to domain objects}
+    \caption{Rule-based transformation of service response}
     \label{fig:response_service_execution}
@@
 
-Once the service response has been retrieved, the transformation tab of the wrapper tool shows it as an interactive object tree, as seen on the left side of Fig.~\ref{fig:response_service_execution}.
-
-A transformation rule is used to analyse the response data and to generate domain-specific objects from concepts from the ontologies used by the parameters of the different building blocks.
-A transformation rule is composed of three elements, as seen in the middle part of Fig.~\ref{fig:response_service_execution}.
-First, the \textit{from} field indicates the source elements to be translated by the rule.
-Second, the type of the rule will be set, taking one of the following values: \emph{createObject}, \emph{fillAttributes} or \emph{dummy}.
-And third, the target of the rule specifies a certain concept or attribute, to be created or filled.
-A detailed explanation of the type of actions to be trigger from the transformation rules is as follows:
+A transformation rule is used to analyse the response data and generate the wrapper's post-conditions.
+Such a rule is composed of three elements, as seen in the middle part of Fig.~\ref{fig:response_service_execution}.
+The \textit{from} field indicates the source elements to be translated by the rule.
+The second part defines the type of rule (see below).
+Thirdly, the target of the rule specifies a certain concept or attribute, to be created or filled.
+Rules can be chained together in a rule tree.
+A detailed explanation of the different rule types is as follows:
 
@@
 
-specifies the creation of a new domain object.
-The type of that new object is provided in the third compartment.
-In the example being explained, the root rule searches for XML elements with tag name \emph{FindItemsResponse} and for each such element a \emph{List} object is created.
-The resulting objects are shown in a facts tree in the right of Figure ~\ref{fig:response_service_execution}.
+specifies the creation of a new post-condition of the type specified in the third part of the rule.
+In the example in Fig.~\ref{fig:construct_service_request}, the root rule could search for source elements with name \emph{FindItemsResponse} and create a post-condition of type \emph{List} for each such element.
+The resulting objects are shown in a facts tree in the right of the figure.
 
@@
 
-does not create a new object but it fills the value of the attribute provided as third part of such rules.
-In our example, the third transformation rule searches for XML elements with tag name \emph{Title}.
-Note, the rule is a sub-rule of the second rule, which generates \emph{Product} objects.
+only ever appear as sub-rules in the rule tree.
+They match source elements defined in the first part of the rule.
+The third part defines the attribute of the super-rule that will be filled (e.g., \emph{fullName} in Fig.~\ref{fig:response_service_execution}).
+% In Fig.~\ref{fig:response_service_execution}, the third transformation rule matches elements with the name \emph{Title}. The 
+% Note that this rule is a sub-rule of the second rule, which generates \emph{Product} objects.
 Thus, the sub-rule searches for \emph{Title} tags only in the subtree of the XML data that has been identified by an application of the parent rule before.
@@
 
-does not create or modify any objects but such rules are just used to narrow the search space for their sub-rules.
-For example, in Amazon product data, the XML data for an item contains sections for \emph{minimum price}, \emph{maximum price}, and \emph{average price}.
-Each such section contains the \emph{plain price} and the \emph{formatted price}.
-Thus, in the Amazon case, a rule that searches for \emph{formatted price} elements within an \emph{Item} element would retrieve three matches.
-Using a dummy rule, we may first search for \emph{minimum price} elements and then search for \emph{formatted price} elements within that sub-tree.
+does not create or modify any objects but instead narrows the search space for their sub-rules, by selecting certain elements in the source tree and ignoring others.
+
+% E.g., in Amazon product data, each item contains elements for \emph{minimum price}, \emph{maximum price}, and \emph{average price}.
+Each such element in turn contains a \emph{plain price} and \emph{formatted price}.
+In this case, a rule that searches for \emph{formatted price} elements within an \emph{Item} element would retrieve three matches.
+Using a dummy rule, we can first search for \emph{minimum price} elements and then search for \emph{formatted price} elements within that sub-tree.
 
@@
 Our tool follows an interactive paradigm.
-Therefore, any time a change to a transformation rule is done, the transformation process is triggered and the resulting facts tree is directly shown.
-This process helps the user to deal with the slightly complex semantics of the transformation rules avoiding errors or mistakes.
-In addition, our tool is ontology-driven, therefore the service wrapper designer retrieves the domain-specific types from a backend server (such as the catalogue discussed in Section~\ref{sec:discovery}) together with the structure of each type, i.e.
+Any time a change to a transformation rule is done, the transformation process is triggered and the resulting facts tree is shown.
+This aids the user to deal with the complexity of the transformation rules, preventing errors or mistakes.
+In addition, the tool is ontology-driven, meaning that possible post-condition types and their attributes are selected from domain ontologies loaded in the system.
+% therefore the service wrapper designer retrieves the domain-specific types from a backend server (such as the catalogue discussed in Section~\ref{sec:discovery}) together with the structure of each type, i.e.
 together with a description of the attributes of each object.
@@
 
-\subsection{Generating a Resource Adapter} % (fold)
+\subsection{Generating the Service Wrapper} % (fold)
 \label{sub:generating_a_resource_adapter}
 
-Once the wrapping of a service has been defined and tested in the service wrapper tool, we generate an implementation of the desired Resource Adapter in XML, HTML, and JavaScript, ready to be deployed and executed inside a web gadget.
+Once the wrapping of a service has been defined and tested in the tool, we generate an implementation of the desired specification in XML, HTML, and JavaScript, ready to be deployed and executed inside a web gadget.
 
@@
 
-\subsection{Limitations} % (fold)
-\label{sub:limitations}
-
-The rule driven approach presented above is somewhat limited.
+% \subsection{Limitations} % (fold)
+% \label{sub:limitations}
+% 
+% The rule driven approach presented above is somewhat limited.
 It is deliberately restricted to such a simple rule mechanism in order to keep things simple enough for end-users.
@@
 For example the attribute \texttt{grandmother} could be filled using \texttt{mother.mother} in the \textit{from} field.
-
-However, there are some transformations that these rules cannot perform.
+% 
+% However, there are some transformations that these rules cannot perform.
 For example, we do not support any mathematical operations.
@@
 Then, service wrappers in combination with these filter operators will allow covering these complex cases.
-
+% 
 % subsection limitations (end)

### 2011-01-28 trunk/papers/webist2011/paper_ssec_discovery.tex
super-mega shortening done. Down to 6 pages from 10 1/2
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1211 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-\begin{listing}
-\begin{verbatim}
+\lstset{
+	language=sparql
+}
+\begin{lstlisting}
 :G1 { :s1 a fgo:Screen .      
@@
 :GC3 { _:x a amazon:SearchCriteria }
-\end{verbatim}
-\label{lis:discovery_example}
-\end{listing}
+\end{lstlisting}
 
@@
 
-\begin{listing}
-\begin{verbatim}
-PREFIX rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#>
-PREFIX fgo: <http://purl.oclc.org/fast/ontology/gadget#>
-PREFIX amazon:<http://aws.amazon.com/AWSECommerceService#>
-
+\begin{lstlisting}
 SELECT DISTINCT ?bb 
 WHERE { 
-  ?bb rdf:type fgo:Screen . 
-  {
-    {
+  ?bb a fgo:Screen . 
+  { { 
       ?bb fgo:hasPostCondition ?c .
       ?c fgo:hasPattern ?p .
-      GRAPH ?p { ?x rdf:type amazon:SearchCriteria } 
-    }
-    UNION
-    {
+      GRAPH ?p { ?x a amazon:SearchCriteria } 
+    } UNION {
       ?bb fgo:hasPostCondition ?c .
       ?c fgo:hasPattern ?p .
-      GRAPH ?p { ?x rdf:type amazon:Item } 
-    }
-  }
+      GRAPH ?p { ?x a amazon:Item } 
+  } }
   FILTER (?bb != <http://fast.org/screens/S1>) 
-  FILTER (?bb != <http://fast.org/screens/S2>) 
-}
-\end{verbatim}
-\label{lis:sparql_find}
-\end{listing}
+  FILTER (?bb != <http://fast.org/screens/S2>) }
+\end{lstlisting}
 
-However, in the example presented, it is straight-forward to spot that the pre-condition \emph{c3} would be
-satisfied whether the screen \emph{s1} was reachable, being unnecessary to query the recommender algorithm 
-to retrieve screens satisfying this pre-condition, hence these pre-conditions are removed for the construction
-of the query reducing the complexity of the problem, hence improving the overall performance of the algorithm.
-Moreover, it leads to find better results for the user, since the focus is given to make satisfy the 
-pre-conditions which could not be satisfied by any of the elements of the current composition.
+% However, in the example presented, it is straight-forward to spot that the pre-condition \emph{c3} would be satisfied whether the screen \emph{s1} was reachable, being unnecessary to query the recommender algorithm to retrieve screens satisfying this pre-condition, hence these pre-conditions are removed for the construction of the query reducing the complexity of the problem, hence improving the overall performance of the algorithm.
+Moreover, it leads to find better results for the user, since the focus is given to make satisfy the pre-conditions which could not be satisfied by any of the elements of the current composition.
 
-\subsubsection{Enhanced discovery: search tree planning}
+\subsubsection{Enhanced Discovery: Search Tree Planning}
 \label{sssec:planning}
 
-In artiﬁcial intelligence, the term \emph{planning} originally meant a search for a sequence of logical operators or actions
-that transform an initial world state into a desired goal state. Presently, it also includes many decision-theoretic ideas,
-imperfect state information, and game-theoretic equilibria.
+In artificial intelligence, the term \emph{planning} originally meant to search for a sequence of logical operators or actions that transform an initial state into a desired goal state.% Presently, it also includes many decision-theoretic ideas, imperfect state information, and game-theoretic equilibria.
 
-This paper applies the concept of planning to the design of building blocks. Back to the screen-flow design, the goal
-would be to make the screen-flow executable; the initial state would be a certain screen which is not yet reachable, 
-and the plans would be sets of screens which are reachable by themselves, and accomplish the goal. 
-The algorithm has been influenced by the ideas behing the heuristic search. For a certain state, i.e. the initial 
-state which contains the screen to make reachable, a large tree of possible continuations is considered, in 
-fact any screen would fit. Those screens which post-conditions do not satisfy in any form the unsatisfied 
-pre-conditions are discarded, reducing the branches of the tree. A branch stops growing when a screens is
-already reachable (i.e. it has no pre-conditions) becoming a leaf of the tree. Once there are not screens 
-added in a certain step, the algorithm stops and discards the incomplete branches, in other words, 
-those branches where the leaf is not reachable.
+In contrast to the simple approach, the planning approach finds sets of building blocks to fulfil the pre-conditions of a given building block (e.g., a screen). For a certain state, i.e., the initial 
+state which contains the pre-condition to fulfil, a large search tree of possible continuations is considered.
+Those building blocks cannot satisfy the unsatisfied pre-conditions are discarded, reducing the branches of the tree.
+A branch stops growing when a building block is reachable (i.e., it has no unfulfilled pre-conditions).
+Once there are no screens
+added in a certain step, the algorithm stops and discards all incomplete branches.
 
-It is worth pointing out some of the tree structure is pre-computed beforehand to speed up the querying process at runtime. 
+It should be pointed out that some of the tree structure is pre-computed to speed up the querying process at runtime. 
 Any time a building block is inserted into the catalogue, the algorithm is executed following two approaches: 
@@
 finds the building blocks whose post-conditions will satisfy the pre-conditions of the recently
-created building block. This data is stored and used for the tree or plan builder algorithm, without
-needing to check the compatibility of pre- and post-conditions for every single building block at every 
-request.
+created building block.
 
@@
 
-Previous sections presented different mechanisms to search or discover building blocks which may satisfy the user needs.
-The results ensure compatibility based on the functional specification of the building blocks, however they are not
-ranked, and their quality is not measured. This section explains the ranking techniques applied for the different
-discovery mechanisms.
+This section explains the ranking techniques applied for the different discovery mechanisms.
 
-The ranking algorithm for~\ref{sssec:simple_discovery} applies the following rules:
-\begin{enumerate}
+The ranking algorithm for the simple approach applies the following rules:
+\begin{inparaenum}[(1)]
 	\item it gives a higher position to those building blocks which satisfy the highest number of pre-conditions;
- \item it prioritise building blocks created by the same user who is querying;
- \item it adjust the rank by using the ratings given to the building blocks, and their popularity in terms of usage statistics;
- \item it weights the results according to non-functional features such as availability. This is only applied for what we call
-``web service wrapper'', and it is calculated periodically by invoking the wrapped web services generating an up-time rate.
-\end{enumerate}
+	\item it prioritises building blocks created by the same user who is querying;
+	\item it adjusts the rank by using the ratings given to the building blocks, and their popularity in terms of usage statistics;
+	\item it weights the results according to non-functional features such as availability.
+This is only applied for what we call ``web service wrapper'', and it is calculated periodically by invoking the wrapped web services.
+\end{inparaenum}
 
@@
 known as \emph{preference-based planning}. The ranking algorithm:
-\begin{enumerate}
- \item minimizes the size of the plans, after removing the elements of the plan which are already in the canvas, so it gives
+\begin{inparaenum}[(1)]
+ \item minimises the size of the plans, after removing the elements of the plan which are already in the canvas, so it gives
 priority to the elements the user has already inserted,
@@
 pre- and post-conditions.
-\end{enumerate}
+\end{inparaenum}

### 2011-01-28 trunk/papers/webist2011/paper_ssec_linked_data.tex
super-mega shortening done. Down to 6 pages from 10 1/2
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1211 d8b0c59b-6544-0410-94d9-83a5658a058e

%!TEX root = paper.tex

\subsection{Serving Linked Data} % (fold)
\label{sub:linked_data}

The idea of a Web of data has recently seen a remarkable uptake, e.g. highlighted by large players such as the New York Times, the BBC or an increasing number of national governments (most notably the US and UK governments).
We apply this concept in order to provide metadata about the web services as linked data, following the principles as defined in~\cite{bizer_ijswis2009}, in order to make them available to arbitrary third-party applications. Each web service is identified by an HTTP URI and hosted in the publishing platform so that it can be dereferenced through the same URI. For each building block, data is available in representations in different standard formats such as JSON (for communication with the applications such as the widget designer shown in Fig.~\ref{fig:fast_architecture}), RDF/XML, Turtle, or even HTML+RDFa as a human-readable version. To do this, we employ content-negotiation, such as proposed as a best practice in \cite{sauermann2008cool_uris}.
%The principles and best practices proposed in \cite{berrueta2008} and  are also taken into account, so that these representations are served based on the request issued by the requesting agent, using a technique called \emph{content negotiation}. 
%As suggested by linked data principles, individual building blocks link to other data on the web, thereby preventing so-called isolated ``data islands''.

% To allow a third-party application to retrieve the content in the format required, we support content negotiation as defined in \cite{http1.1} and proposed as best practice in \cite{sauermann2008cool_uris}, serving the best variant for a resource, taking into account what variants are available, what variants the server may prefer to serve, what the client can accept, and with which preferences.
% In HTTP, this is done by the client which may send, in its request, accept headers (\texttt{Accept}, \texttt{Accept-Language} and \texttt{Accept-Encoding}), to communicate its capabilities and preferences in format, language and encoding, respectively. Concretely, the approach followed is agent-driven negotiation, where the user agent selects the specific representation for a resource.

### 2011-01-31 trunk/deliverables/D2_2_Ontology/fast_d2_2_3_ontology.tex
template now in sty instead of cls file
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1216 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
-\documentclass[twoside]{fast_latex}
+% \documentclass[twoside]{article}
+\documentclass{article}
+
+\usepackage{fast_latex}
 % yay! We can use any kind of funky diacritic:
@@
 
-\newcommand\deliverableNumber{D2.2.2}
+\newcommand\deliverableNumber{D2.2.3}
 \newcommand\deliverableTitle{Ontology and Conceptual Model for the Semantic Characterisation of Complex Gadgets}
@@
 		\vfill
-		{\large \sffamily Date: 27/02/2010}\\
+		{\large \sffamily Date: 27/02/2011}\\
 		\vspace{1cm}
@@
 \cofoot{\small{FAST $\bullet$ 216048 $\bullet$ \deliverableTitleShort{} $\bullet$ Page \thepage\ of \pageref{LastPage}}}
-
-\newpage
-%
-%\section*{Authors}
-%\authorOne \\
-%%\authorTwo \\
-%%\authorThree\\
-%%\authorFour\\
-%
-%
-%\vfill
-%%\section*{Project Co-ordinator}
-%%\hspace*{-2,5cm}\begin{minipage}[p]{14cm}
-%%Dr. Ansgar Bernardi \\
-%%German Research Center for Artificial Intelligence (DFKI) GmbH \\
-%%Trippstadter Strasse 122 \\
-%%D 67663 Kaiserslautern \\
-%%Germany \\
-%%Email: bernardi@dfki.uni-kl.de, phone: +49 631 205 3582, fax: +49 631 205 4910 \\
-%%\end{minipage}
-%
-%
-%\section*{Partners}
-%Telef\'onica Investigaci\'on y Desarrollo (Telef\'onica I+D), Spain \\
-%National University of Ireland, Galway (NUIG), Republic of Ireland \\
-%Kassel University, Germany \\
-%Cyntelix Corporation, Republic of Ireland \\
-%SAP AG, Germany \\
-%Universidad Polit\'ecnica de Madrid (UPM), Spain
-%
-%
-%\vspace{0.3cm}
-%\begin{footnotesize}
-%Copyright: FAST Consortium 2009\\
-%\end{footnotesize}
-
-%\clearpage
 
@@
 2.0 & 27.02.2010 & \authorOne & final version (D2.2.2) ready for external review \\ \hline
+3.0 & 27.02.2011 & \authorOne & final version (D2.2.3) ready for external review \\ \hline
 \end{tabular}
@@
 	\hline
-	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2010} & {\textbf{Submitted}} & {27/02/2010}\\ \hline
+	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2011} & {\textbf{Submitted}} & {27/02/2011}\\ \hline
 	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
@@
 
+\todo{Rewrite this whole section with focus on prototypes, prototype theory, prototype-based programming, etc.}
+
+\cite{Taivalsaari96classes_vs_prototypes}
+
 In FAST, resources (gadgets, screenflows, screens and their components) exist in different stages throughout their lifecycle.
@@
     \includegraphics[width=\linewidth]{images/classes_templates_instances.pdf}
-    \caption{Different levels of representation of FAST resources}
+    \caption{Different levels of representation of FAST building blocks}
     \label{fig:classes_templates}
@@
 \phantomsection
+\section{The VocDoc Ontology Documentation Tool}
+\label{sec:vocdoc}
+
+\todo{include documentation for VocDoc}
+
+\clearpage
+\phantomsection
 \bibliographystyle{apalike}

### 2011-01-31 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/client/gui/RequestGUI.java
Fix: Position of the WSDL Button
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1219 d8b0c59b-6544-0410-94d9-83a5658a058e

@@ -4,6 +4,7 @@ import com.google.gwt.event.dom.client.ClickEvent;
 import com.google.gwt.event.dom.client.ClickHandler;
 import com.google.gwt.user.client.ui.Button;
 import com.google.gwt.user.client.ui.FlexTable;
+import com.google.gwt.user.client.ui.FlowPanel;
 import com.google.gwt.user.client.ui.HasHorizontalAlignment;
 import com.google.gwt.user.client.ui.Label;
 import com.google.gwt.user.client.ui.TextArea;
@@ -87,9 +88,13 @@ public class RequestGUI
 		designer.requestButton = new Button("Send Request", designer.requestHandler);
 		designer.requestButton.setStyleName("fastButton");
 		designer.requestButton.addStyleName("sc-FixedWidthButton");
-		requestTable.setWidget(numRows, 1, designer.requestButton);
+		
+		//the flow panel holds the "send request" and "WSDL" button 
+		FlowPanel flowPanel = new FlowPanel();
+		flowPanel.add(designer.requestButton);
+		flowPanel.add(wsdlButton);
+		requestTable.setWidget(numRows, 1, flowPanel);
 		requestTable.setCellSpacing(15);
-		requestTable.setWidget(numRows, 2, wsdlButton);
 		
 		numRows++;
 

### 2011-01-31 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/server/RequestServiceImpl.java

@@ -65,6 +65,9 @@ public class RequestServiceImpl extends RemoteServiceServlet implements RequestS
 		return responseBody;
 	}
 	
+	/**
+	 * TODO dk or tg: parse the header!
+	 * */
 	@Override
 	public String sendHttpRequest_POST(String url, HashMap<String, String> headers, String body)
 	{
@@ -72,7 +75,7 @@ public class RequestServiceImpl extends RemoteServiceServlet implements RequestS
 		HttpClient httpclient = new HttpClient();
 	    PostMethod method = new PostMethod(url);
 	    
-	    //needed for bis responses, but then we should cut the body not containing answerers
+	    //needed for big responses, but then we should cut the body not containing answerers
 //	    BufferedReader br = null;
 	    
 	    //parse given body and create parameters
@@ -157,7 +160,6 @@ public class RequestServiceImpl extends RemoteServiceServlet implements RequestS
 			boolean localDemo = true;
 			if (localDemo)
 			{
-				
 				path = "../../GVS/static/";
 			}
 			else if ( classLocation.contains("ServiceDesignerWep"))

### 2011-01-31 trunk/catalogue_service/src/eu/morfeoproject/fast/catalogue/Catalogue.java
fixes problem with "from" and "to" URIs in pipes and triggers, when they reference the screen they belong to
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1220 d8b0c59b-6544-0410-94d9-83a5658a058e

@@ -1144,6 +1144,7 @@ public class Catalogue {
 			screen.setCreationDate(new Date());
 		// persists the screen
 		saveScreen(screen);
+		printStatements();
 		// create plans for the screen
 		if (planner != null) planner.add(screen);
 		scRecommender.rebuild(); //FIXME do not rebuild every time a new screen is added!
@@ -1275,13 +1276,15 @@ public class Catalogue {
 			}
 			for (Pipe pipe : screen.getPipes()) {
 				URI pUri = tripleStore.createURI(copyUri+"/pipes/"+UUID.randomUUID().toString());
-				Pipe newPipe = pipe.clone(copyUri, pUri);
+				Screen copy = BuildingBlockFactory.createScreen(copyUri);
+				Pipe newPipe = pipe.clone(copy, pUri);
 				tripleStore.addModel(newPipe.toRDF2GoModel(), clonesGraph);
 				tripleStore.addStatement(clonesGraph, copyUri, FGO.contains, pUri);
 			}
 			for (Trigger trigger : screen.getTriggers()) {
 				URI tUri = tripleStore.createURI(copyUri+"/triggers/"+UUID.randomUUID().toString());
-				Trigger newTrigger = trigger.clone(copyUri, tUri);
+				Screen copy = BuildingBlockFactory.createScreen(copyUri);
+				Trigger newTrigger = trigger.clone(copy, tUri);
 				tripleStore.addModel(newTrigger.toRDF2GoModel(), clonesGraph);
 				tripleStore.addStatement(clonesGraph, copyUri, FGO.contains, tUri);
 			}

### 2011-01-31 trunk/catalogue_service/src/eu/morfeoproject/fast/catalogue/builder/BuildingBlockJSONBuilder.java

@@ -314,7 +314,7 @@ public class BuildingBlockJSONBuilder {
 	}
 	
 	private static Pipe parsePipe(Screen screen, JSONObject jsonPipe) throws JSONException, IOException {
-		Pipe pipe = BuildingBlockFactory.createPipe(screen == null ? null : screen.getUri());
+		Pipe pipe = BuildingBlockFactory.createPipe(screen);
 		JSONObject pipeFrom = jsonPipe.getJSONObject("from");
 		JSONObject pipeTo = jsonPipe.getJSONObject("to");
 		String bbFrom = pipeFrom.isNull("buildingblock") || pipeFrom.getString("buildingblock").equals("") ? null : pipeFrom.getString("buildingblock");
@@ -331,7 +331,7 @@ public class BuildingBlockJSONBuilder {
 	}
 	
 	private static Trigger parseTrigger(Screen screen, JSONObject jsonTrigger) throws JSONException, IOException {
-		Trigger trigger = new Trigger(screen == null ? null : screen.getUri());
+		Trigger trigger = new Trigger(screen);
 		JSONObject from = jsonTrigger.getJSONObject("from");
 		JSONObject to = jsonTrigger.getJSONObject("to");
 		String bbFrom = from.isNull("buildingblock") || from.getString("buildingblock").equals("") ? null : from.getString("buildingblock");

### 2011-01-31 trunk/catalogue_service/src/eu/morfeoproject/fast/catalogue/builder/BuildingBlockRDF2GoBuilder.java

@@ -367,7 +367,7 @@ public class BuildingBlockRDF2GoBuilder {
 	}
 	
 	private static Pipe retrievePipe(Screen screen, URI pipeURI, Model model) {
-		Pipe pipe = BuildingBlockFactory.createPipe(screen.getUri(), pipeURI);
+		Pipe pipe = BuildingBlockFactory.createPipe(screen, pipeURI);
 		ClosableIterator<Statement> cIt = model.findStatements(pipeURI, Variable.ANY, Variable.ANY);
 		for ( ; cIt.hasNext(); ) {
 			Statement stmt = cIt.next();
@@ -405,7 +405,7 @@ public class BuildingBlockRDF2GoBuilder {
 	}
 
 	private static Trigger retrieveTrigger(Screen screen, URI triggerURI, Model model) {
-		Trigger trigger = BuildingBlockFactory.createTrigger(screen.getUri(), triggerURI);
+		Trigger trigger = BuildingBlockFactory.createTrigger(screen, triggerURI);
 		ClosableIterator<Statement> cIt = model.findStatements(triggerURI, Variable.ANY, Variable.ANY);
 		for ( ; cIt.hasNext(); ) {
 			Statement stmt = cIt.next();

### 2011-01-31 trunk/catalogue_service/src/eu/morfeoproject/fast/catalogue/model/BuildingBlock.java

@@ -192,7 +192,7 @@ public abstract class BuildingBlock {
 		JSONObject json = new JSONObject();
 
 		json.put("uri", getUri() == null ? JSONObject.NULL : getUri().toString());
-		json.put("template", getPrototype() == null ? JSONObject.NULL : getPrototype());
+		json.put("prototype", getPrototype() == null ? JSONObject.NULL : getPrototype());
 		if (this.getLabels() == null) json.put("label", JSONObject.NULL);
 		else {
 			JSONObject jsonLabels = new JSONObject();

### 2011-01-31 trunk/catalogue_service/src/eu/morfeoproject/fast/catalogue/model/Pipe.java

@@ -13,20 +13,20 @@ import eu.morfeoproject.fast.catalogue.vocabulary.FGO;
 public class Pipe implements Comparable<Pipe> {
 
 	private URI uri;
-	private URI screenURI;
+	private Screen screen;
 	private URI bbFrom;
 	private String conditionFrom;
 	private URI bbTo;
 	private String actionTo;
 	private String conditionTo;
 
-	public Pipe(URI screenURI) {
+	public Pipe(Screen screen) {
 		super();
-		this.screenURI = screenURI;
+		this.screen = screen;
 	}
 	
-	public Pipe(URI screenURI, URI uri) {
-		this(screenURI);
+	public Pipe(Screen screen, URI uri) {
+		this(screen);
 		this.uri = uri;
 	}
 	
@@ -38,12 +38,12 @@ public class Pipe implements Comparable<Pipe> {
 		this.uri = uri;
 	}
 
-	public URI getScreenUri() {
-		return screenURI;
+	public Screen getScreen() {
+		return screen;
 	}
 
-	public void setScreenUri(URI screenURI) {
-		this.screenURI = screenURI;
+	public void setScreen(Screen screen) {
+		this.screen = screen;
 	}
 
 	public URI getBBFrom() {
@@ -106,7 +106,13 @@ public class Pipe implements Comparable<Pipe> {
 		if (this.uri != null && other.getUri() != null) {
 			comparison = this.uri.compareTo(other.getUri());
 			if (comparison != EQUAL) return comparison;
-		}	
+		}
+//		if (this.screen == null && other.getScreen() != null) return AFTER;
+//		if (this.screen != null && other.getScreen() == null) return BEFORE;
+//		if (this.screen != null && other.getScreen() != null) {
+//			comparison = this.screen.compareTo(other.getScreen());
+//			if (comparison != EQUAL) return comparison;
+//		}
 		if (this.bbFrom == null && other.getBBFrom() != null) return AFTER;
 		if (this.bbFrom != null && other.getBBFrom() == null) return BEFORE;
 		if (this.bbFrom != null && other.getBBFrom() != null) {
@@ -166,10 +172,10 @@ public class Pipe implements Comparable<Pipe> {
 //		}
 		
 		URI from = model.createURI(bbFrom == null ? 
-				screenURI + "/preconditions/" + conditionFrom : 
[…]

### 2011-01-31 trunk/catalogue_service/src/eu/morfeoproject/fast/catalogue/model/Trigger.java

@@ -13,19 +13,19 @@ import eu.morfeoproject.fast.catalogue.vocabulary.FGO;
 public class Trigger implements Comparable<Trigger> {
 
 	private URI uri;
-	private URI screenURI;
+	private Screen screen;
 	private URI bbFrom;
 	private String nameFrom;
 	private URI bbTo;
 	private String actionTo;
 
-	public Trigger(URI screenURI) {
+	public Trigger(Screen screen) {
 		super();
-		this.screenURI = screenURI;
+		this.screen = screen;
 	}
 	
-	public Trigger(URI screenURI, URI uri) {
-		this(screenURI);
+	public Trigger(Screen screen, URI uri) {
+		this(screen);
 		this.uri = uri;
 	}
 	
@@ -37,12 +37,12 @@ public class Trigger implements Comparable<Trigger> {
 		this.uri = uri;
 	}
 
-	public URI getScreenUri() {
-		return screenURI;
+	public Screen getScreen() {
+		return screen;
 	}
 
-	public void setScreenUri(URI screenURI) {
-		this.screenURI = screenURI;
+	public void setScreen(Screen screen) {
+		this.screen = screen;
 	}
 
 	public URI getBBFrom() {
@@ -98,6 +98,12 @@ public class Trigger implements Comparable<Trigger> {
 			comparison = this.uri.compareTo(other.getUri());
 			if (comparison != EQUAL) return comparison;
 		}	
+//		if (this.screen == null && other.getScreen() != null) return AFTER;
+//		if (this.screen != null && other.getScreen() == null) return BEFORE;
+//		if (this.screen != null && other.getScreen() != null) {
+//			comparison = this.screen.compareTo(other.getScreen());
+//			if (comparison != EQUAL) return comparison;
+//		}
 		if (this.bbFrom == null && other.getBBFrom() != null) return AFTER;
 		if (this.bbFrom != null && other.getBBFrom() == null) return BEFORE;
 		if (this.bbFrom != null && other.getBBFrom() != null) {
@@ -144,8 +150,8 @@ public class Trigger implements Comparable<Trigger> {
 		model.open();
 		model.setNamespace("fgo", FGO.NS_FGO.toString());
 		
-		URI from = model.createURI((bbFrom == null ? screenURI : bbFrom) + "/triggers/" + nameFrom);
[…]

### 2011-01-31 trunk/catalogue_service/src/eu/morfeoproject/fast/catalogue/model/factory/BuildingBlockFactory.java

@@ -88,20 +88,20 @@ public class BuildingBlockFactory {
 		return new Condition(uri);
 	}
 	
-	public static Pipe createPipe(URI screenUri) {
-		return new Pipe(screenUri);
+	public static Pipe createPipe(Screen screen) {
+		return new Pipe(screen);
 	}
 	
-	public static Pipe createPipe(URI screenUri, URI uri) {
-		return new Pipe(screenUri, uri);
+	public static Pipe createPipe(Screen screen, URI uri) {
+		return new Pipe(screen, uri);
 	}
 	
-	public static Trigger createTrigger(URI screenUri) {
-		return new Trigger(screenUri);
+	public static Trigger createTrigger(Screen screen) {
+		return new Trigger(screen);
 	}
 	
-	public static Trigger createTrigger(URI screenUri, URI uri) {
-		return new Trigger(screenUri, uri);
+	public static Trigger createTrigger(Screen screen, URI uri) {
+		return new Trigger(screen, uri);
 	}
 	
 	public static Action createAction() {

### 2011-01-31 trunk/papers/webist2011/paper_sec_wrapping.tex
rotated Figure 3 for larger size.
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1221 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-\subsection{Constructing Service Requests} % (fold)
-\label{sub:constructing_service_requests}
+%\subsection{Constructing Service Requests} % (fold)
+%\label{sub:constructing_service_requests}
 
-In this section we illustrate the interaction with RESTful services on simple GET requests\footnote{Other methods such as POST or SOAP-based services are also supported.}. 
+We now illustrate the interaction with RESTful services on simple GET requests\footnote{Other methods such as POST or SOAP-based services are also supported.}. 
 In this case, a service request is assembled using its URL and a set of parameters. As an example, we will look at the eBay Shopping web service. 
@@
 For the current example, we assume that the user has defined a pre-condition \emph{search\_key}.
-
 We are invoking the service to retrieve a list of items corresponding to the search keyword ``USB'':
@@
 
-\subsection{Interpreting Service Responses} % (fold)
-\label{sub:interpreting_service_responses}
+%\subsection{Interpreting Service Responses} % (fold)
+%\label{sub:interpreting_service_responses}
 
@@
   \begin{center}     
-    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithTransformationRules.png}
+    \includegraphics[height=\linewidth, angle=90]{images/ServiceWrapperToolGVSWithTransformationRules.png}
     \caption{Rule-based transformation of service response}
@@
 
-\subsection{Generating the Service Wrapper} % (fold)
-\label{sub:generating_a_resource_adapter}
+%\subsection{Generating the Service Wrapper} % (fold)
+%\label{sub:generating_a_resource_adapter}
 

### 2011-01-31 trunk/catalogue_service/src/eu/morfeoproject/fast/catalogue/Catalogue.java
fixes problem when a condition is selected in a find&check (screen design) request
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1223 d8b0c59b-6544-0410-94d9-83a5658a058e

@@ -1144,7 +1144,6 @@ public class Catalogue {
 			screen.setCreationDate(new Date());
 		// persists the screen
 		saveScreen(screen);
-		printStatements();
 		// create plans for the screen
 		if (planner != null) planner.add(screen);
 		scRecommender.rebuild(); //FIXME do not rebuild every time a new screen is added!

### 2011-01-31 trunk/catalogue_service/src/eu/morfeoproject/fast/catalogue/services/ScreenComponentFindCheckServlet.java

@@ -99,11 +99,11 @@ public class ScreenComponentFindCheckServlet extends GenericServlet {
 				}
 			} 
 			// flag to search or not for new components
-			boolean search = input.has("search") ? input.getBoolean("search") : true;
-			boolean iserve = input.has("iserve") ? input.getBoolean("iserve") : true;
+			boolean search = input.optBoolean("search", true);
+			boolean iserve = input.optBoolean("iserve", true);
 			// pagination variables
-			int offset = input.has("offset") ? input.getInt("offset") : 0;
-			int limit = input.has("limit") ? input.getInt("limit") : -1;
+			int offset = input.optInt("offset", 0);
+			int limit = input.optInt("limit", -1);
 			
 			// do the real work
 			//-----------------------------
@@ -384,7 +384,7 @@ public class ScreenComponentFindCheckServlet extends GenericServlet {
 			}
 		} else {
 			Condition condition = (Condition) selectedItem;
-			if (preconditions.contains(condition)) {
+			if (getConditionById(preconditions, condition.getId()) != null) {
 				// look for pipes from the precondition to <somewhere>
 				for (ScreenComponent sc : canvas) {
 					for (Action action : sc.getActions()) {
@@ -401,7 +401,7 @@ public class ScreenComponentFindCheckServlet extends GenericServlet {
 						}
 					}
 				}
-			} else if (postconditions.contains(condition)) {
+			} else if (getConditionById(postconditions, condition.getId()) != null) {
 				// look for pipes from <somewhere> to the postcondition
 				for (ScreenComponent sc : canvas) {
 					for (Condition con : sc.getPostconditions()) {

### 2011-01-31 trunk/papers/webist2011/paper_ssec_discovery.tex
squeezing things to get down to 6 pages - done
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1224 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
   ?bb a fgo:Screen . 
-  { { 
-      ?bb fgo:hasPostCondition ?c .
+  { { ?bb fgo:hasPostCondition ?c .
       ?c fgo:hasPattern ?p .
@@
       ?c fgo:hasPattern ?p .
-      GRAPH ?p { ?x a amazon:Item } 
-  } }
+      GRAPH ?p { ?x a amazon:Item } } }
   FILTER (?bb != <http://fast.org/screens/S1>) 
@@
 
-In artificial intelligence, the term \emph{planning} originally meant to search for a sequence of logical operators or actions that transform an initial state into a desired goal state.% Presently, it also includes many decision-theoretic ideas, imperfect state information, and game-theoretic equilibria.
+% In artificial intelligence, the term \emph{planning} originally meant to search for a sequence of logical operators or actions that transform an initial state into a desired goal state.% Presently, it also includes many decision-theoretic ideas, imperfect state information, and game-theoretic equilibria.
 

### 2011-01-31 trunk/catalogue_service/src/eu/morfeoproject/fast/catalogue/model/Action.java
fixes problem with JSON representation of "uses"
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1225 d8b0c59b-6544-0410-94d9-83a5658a058e

@@ -81,7 +81,14 @@ public class Action {
 		}
 		json.put("preconditions", preArray);
 		// uses
-		json.put("uses", getUses());
+		JSONArray usesArray = new JSONArray();
+		for (String id : getUses().keySet()) {
+			JSONObject useObject = new JSONObject();
+			useObject.put("id", id);
+			useObject.put("uri", getUses().get(id));
+			usesArray.put(useObject);
+		}
+		json.put("uses", usesArray);
 		return json;
 	}
 

### 2011-01-31 trunk/papers/webist2011/paper_sec_wrapping.tex
figure 3 back to landscape mode, but better quality
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1226 d8b0c59b-6544-0410-94d9-83a5658a058e
finally fig. 3 in good quality!!
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1227 d8b0c59b-6544-0410-94d9-83a5658a058e
after using some magic spells, the paper is back to good quality images and six pages
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1228 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2011-01-28 trunk/papers/webist2011/paper_sec_wrapping.tex
@@
 
-In this section we illustrate the interaction with RESTful services on simple GET requests\footnote{Other methods such as POST or SOAP-based services are also supported.}. 
+We now illustrate the interaction with RESTful services on simple GET requests\footnote{Other methods such as POST or SOAP-based services are also supported.}. 
 In this case, a service request is assembled using its URL and a set of parameters. As an example, we will look at the eBay Shopping web service. 
@@
 
-\begin{figure}
+\begin{figure}[hb]
   \begin{center}     
-    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithTransformationRules.png}
+    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithTransformationRules-HQ.png}
     \caption{Rule-based transformation of service response}

### 2011-01-31 trunk/papers/webist2011/paper_sec_conclusions.tex
added some more hints for future work
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1230 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-The current version of the wrapping tool permits to create resource adapters for RESTful web services.
-SOAP-based services are also experimentally supported.
+As potential future work, we consider to include semantically enriched WSDL documents using SAWSDL, and to support other SWS approaches such as WSMO-lite services, making the web service wrapping a semi-automatic (or automatic) process.
+As regards to the discovery, once the catalogue begins to be populated, applying techniques such as frequent itemset mining or collaborative filtering could be an interesting way to offer better recommendations to the users.
+
+% The current version of the wrapping tool permits to create resource adapters for RESTful web services.
+% SOAP-based services are also experimentally supported.
 As potential future work, we consider to include semantically enriched WSDL documents using SAWSDL, and to support other SWS approaches such as WSMO-lite services --- however, real-life examples of these are still hard to find.

### 2011-02-02 trunk/deliverables/D2_2_Ontology/fast_d2_2_3_ontology.tex
prototyping instead of templating
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1245 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \usepackage{amssymb}
+\usepackage{amsmath}
 \usepackage{paralist}
@@
 
-\subsection{Classes, Instances and Templates} % (fold)
+\subsection{Levels of Representation} % (fold)
 \label{sub:classes_instances_and_templates}
-
-\todo{Rewrite this whole section with focus on prototypes, prototype theory, prototype-based programming, etc.}
-
-\cite{Taivalsaari96classes_vs_prototypes}
 
@@
 
-When implementing these four stages, as illustrated in Fig.~\ref{fig:classes_templates}, to OWL DL semantics, one will be faced with an immediate problem.
-Conceptually, one would like to say that $Screen$ is a class, i.e., the set of all screens.
+When implementing these four stages, as illustrated in Fig.~\ref{fig:classes_templates}, in OWL DL semantics, we are faced with an immediate problem.
+Conceptually, we would like to say that $Screen$ is a class, i.e., the set of all screens.
 As argued above, $LogInScreen$ would then be an instance of $Screen$ ($LogInScreen \subset Screen$), while a particular $LogInScreen_x$ in a particular screenflow would again be an instance of $LogInScreen$ ($LogInScreen_x \subset LogInScreen$).
@@
 Furthermore, when we instantiate a resource such as a screen within a particular screenflow, we want to imply that the new instance ($LogInScreen_x$) has the same features such as label, icon or conditions, as $LogInScreen$.
-In other words, what we really require is a something like a \emph{copy} of the original $LogInScreen$.
-
-Neither OWL nor RDFS semantics offer functionality to represent a relationship such as the copy of a resource formally.
+In other words, what we really require is a something like a \emph{copy} or a \emph{clone} of the original $LogInScreen$.
+
+\subsubsection{Prototypes} % (fold)
+\label{ssub:prototypes}
+
+% subsubsection prototypes (end)
+
+Neither OWL nor RDFS semantics offer functionality to represent a relationship such as the clone of a resource formally.
+However, we can find inspiration for this kind of semantics in other areas, such as programming languages, and more particularly in object-oriented programming languages (OO).
+In OO, there is a general dichotomy between \emph{class-based} languages and \emph{prototype-based} languages.
+Both paradigms address the issue of defining similarity between objects, but do so in a different way.
+The class-based paradigm requires an a-priori classification and categorisation of the world (the class hierarchy), into which individual objects (the instances) are then grouped.
+All instances of a class inherit the attributes and functionality of the class.
+Changing this behaviour usually involves changing the class.
+Prototype-based languages, on the other hand, do not assume prior categorisations.
+Instead, one starts with specific examples and then generalises from there.
+These languages do not show a class vs.
+instance distinction.
+Instead, all objects have the same status, and new objects can be \emph{cloned} from other objects, which function as \emph{prototypical objects} in this situation.
+The new object (or clone) will inherit all attributes and functionality from the prototype.
+Changing behaviour can be done on each object individually.
+
+\todo{Rewrite this whole section with focus on prototypes, prototype theory, prototype-based programming, etc.}
+
+\cite{Taivalsaari96classes_vs_prototypes}
+
+\begin{itemize}
+	\item in a prototypical programming language, creating an object from a prototype means to clone it --- copy all its attributes\footnote{Whether or not this is true copy or merely a pointer is implementation-dependant.
+This is more or less transparent for the user.}
+	\item in RDF, what would ``all its attributes'' mean?
+	\item this is the old question of what constitutes an object in RDF
+	\item should we use the minimal self-contained graph \cite{Tummarello:2008kl}?
+	\item maybe some other named graph? But the RDF model doesn't feature named graphs
+	\item let's just assume some unspecified ``object graph'' for now --- the exact definition of the object graph is application dependent.
+In FAST, it is actually always known which triples constitute the object graph of a building block.
+	\item so, now that we can say that the prototype object, identified by its URI (we therefore call it $P_{URI}$), is defined by its object graph, we can define the relation between the prototype and its clone ($C_{URI}$) as follows:
+\end{itemize}
+
+\begin{equation}
+S:= \begin{cases}
+i = 0..M,\quad j =0..N  & \text{for }M >N \\
+i=0..M,\quad j=0..M & \text{for }N > M
+\end{cases}
+\end{equation}
+
 For this reason, we introduce the concept of \emph{templates for RDF instances} in FAST.
@@
 In practice, this procedure can be implemented using a SPARQL CONSTRUCT query, where the WHERE clause defines the graph pattern of the template, and the CONSTRUCT query will build the copy from it.
-The corresponding data on all three levels is shown in List.~\ref{list:templating}, using TriG syntax~\cite{bizer2004trig} (an extension of Turtle syntax for named graphs).
+The corresponding data on all three levels is shown in List.~\ref{list:prototyping}, using TriG syntax~\cite{bizer2004trig} (an extension of Turtle syntax for named graphs).
 
@@
 \lstset{
-	caption={Templating example --- different levels of representation of a FAST screen (code)}, 
-	label=list:templating,
+	caption={Prototyping example --- different levels of representation of a FAST screen (code)}, 
+	label=list:prototyping,
 	language=turtle
@@
 \begin{figure}
-	\lstinputlisting{exampleCode/templating_example.ttl}
-\end{figure}
-\doublespacing
-
-Figure \ref{fig:templating_example_graph} illustrates the same example graphically, using different colours to represent the two different graphs: the red graph $G_1$ in the background represents the original template resource, whereas the green graph $G_2$ in the front in represents the resource copy as it would be used in a concrete screen flow.
+	\lstinputlisting{exampleCode/prototypeExample.ttl}
+\end{figure}
+\doublespacing
+
+Figure \ref{fig:prototype_example_graph} illustrates the same example graphically, using different colours to represent the two different graphs: the red graph $G_1$ in the background represents the original template resource, whereas the green graph $G_2$ in the front in represents the resource copy as it would be used in a concrete screen flow.
 \texttt{catalogue:LoginScreen} $\subset I_{template}$ has been changed to \texttt{catalogue:LoginScreen\_74382} $\subset I_{copy}$.
@@
   \begin{center}
-    \includegraphics[width=.75\linewidth]{images/templating_example.pdf}
-    \caption{Templating example --- different levels of representation of a FAST screen (diagramme)}
-    \label{fig:templating_example_graph}
+    \includegraphics[width=.75\linewidth]{images/prototypeExample.pdf}
+    \caption{Prototyping example --- different levels of representation of a FAST screen (diagram)}
+    \label{fig:prototype_example_graph}
   \end{center}

### 2011-02-03 trunk/deliverables/D2_2_Ontology/fast_d2_2_3_ontology.tex
in the middle of writing about prototypes
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1250 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
+\lstdefinelanguage{javascript}
+{
+morekeywords={var},
+sensitive=false,
+morecomment=[l]{//},
+morestring=[b]",
+}
+
 \lstset{
@@
 \newcommand\authorFour{Ciprian Alexandru Palaghita, Cyntelix}
+
 \newtheorem{example}{\emph{Example}}
+\newtheorem{theorem}{Theorem}
+\newtheorem{definition}[theorem]{Definition}
 
@@
 
-% subsubsection prototypes (end)
-
 Neither OWL nor RDFS semantics offer functionality to represent a relationship such as the clone of a resource formally.
@@
 In OO, there is a general dichotomy between \emph{class-based} languages and \emph{prototype-based} languages.
+Java, C++ and Smalltalk are examples of the former, while JavaScript\footnote{The name of the standard is ECMAScript, but we will use JavaScript in this deliverable, since it is the most prominent dialect.} (as the most prominent) and Self (as the oldest) are examples of the latter.
 Both paradigms address the issue of defining similarity between objects, but do so in a different way.
 The class-based paradigm requires an a-priori classification and categorisation of the world (the class hierarchy), into which individual objects (the instances) are then grouped.
-All instances of a class inherit the attributes and functionality of the class.
+All instances of a class inherit its attributes and functionality.
 Changing this behaviour usually involves changing the class.
-Prototype-based languages, on the other hand, do not assume prior categorisations.
-Instead, one starts with specific examples and then generalises from there.
+Prototype-based languages, on the other hand, do not assume prior categorisation.
 These languages do not show a class vs.
 instance distinction.
-Instead, all objects have the same status, and new objects can be \emph{cloned} from other objects, which function as \emph{prototypical objects} in this situation.
+Instead, all objects have the same status, and new objects can be \emph{cloned} from other objects, which function as \emph{prototypical objects} in this situation, thereby starting with a specific example and generalising from there.
 The new object (or clone) will inherit all attributes and functionality from the prototype.
 Changing behaviour can be done on each object individually.
-
-\todo{Rewrite this whole section with focus on prototypes, prototype theory, prototype-based programming, etc.}
-
-\cite{Taivalsaari96classes_vs_prototypes}
-
-\begin{itemize}
-	\item in a prototypical programming language, creating an object from a prototype means to clone it --- copy all its attributes\footnote{Whether or not this is true copy or merely a pointer is implementation-dependant.
-This is more or less transparent for the user.}
-	\item in RDF, what would ``all its attributes'' mean?
-	\item this is the old question of what constitutes an object in RDF
-	\item should we use the minimal self-contained graph \cite{Tummarello:2008kl}?
-	\item maybe some other named graph? But the RDF model doesn't feature named graphs
-	\item let's just assume some unspecified ``object graph'' for now --- the exact definition of the object graph is application dependent.
-In FAST, it is actually always known which triples constitute the object graph of a building block.
-	\item so, now that we can say that the prototype object, identified by its URI (we therefore call it $P_{URI}$), is defined by its object graph, we can define the relation between the prototype and its clone ($C_{URI}$) as follows:
-\end{itemize}
+As an aside, the difference between the two paradigms can be traced all the way back to classical philosophy.
+In this sense, the class-based paradigm has its roots in Plato (ideal forms) and Aristotle (classification of things), while the origin of the prototype-based paradigm only appeared much later as a criticism of the classical model, e.g.
+by Wittgenstein (the idea of \emph{family resemblance}) and Eleanor Rosch, who introduced prototype theory within cognitive science.
+While most language designers, let alone programmers, are probably unaware of this heritage~\cite{Taivalsaari96classes_vs_prototypes}, it underlines the importance of this distinction to see how far back it goes.
+
+% subsubsection prototypes (end)
+
+\subsubsection{Prototype-based Semantics for RDF} % (fold)
+\label{ssub:prototype_based_semantics_for_rdf}
+
+While of course RDFS and OWL semantics are not identical to class-based OO semantics (the open and closed world assumption being one case in point), they are quite similar in their distinction between classes and instances, and the a-priori categorisation of the world.
+Along these lines, what we need for FAST is a similar RDF equivalent for prototype-based OO semantics.
+To achieve this, let's first look again at what creating a new object from a prototype means in a prototypical programming language: essentially, this process means to \emph{clone the prototype}, i.e., to copy all its properties and attributes to the new object\footnote{Whether or not this is true copy or merely a pointer is implementation-dependant.
+However, this is usually transparent for the user.}.
+Following our login screen example, a snippet of JavaScript code of what this can look like in practice is shown in List.~\ref{list:js_prototyping}.
+The \texttt{loginScreen} object is created with a number of properties, such as \texttt{label} and \texttt{icon}.
+A second \texttt{loginScreen\_74382} object is then created using the \texttt{create} function, using \texttt{loginScreen} as its prototype.
+\texttt{loginScreen\_74382} inherits all properties from its prototype, but can also overwrite or add individual ones.
+
+\singlespacing
+\lstset{
+	caption={Creation of a new object from a prototype in JavaScript}, 
+	label=list:js_prototyping,
+	language=javascript
+}
+\begin{figure}
+	\lstinputlisting{exampleCode/prototype.js}
+\end{figure}
+\doublespacing
+
+\paragraph{Determining the Prototype's Sub-graph} % (fold)
+\label{par:determining_the_correct_sub_graph}
+
+In an attempt to define prototype-based RDF semantics, we first have to specify what the RDF analogue to an object from OO is.
+Tentatively, we could say that the resource URI is analogous to the object.
+However, this will not help us when we want to create a new resource $C_{URI}$ from a prototypical resource $P_{URI}$: a URI merely \emph{identifies} a particular resource, without defining which statements belong to it.
+We will not know which properties should be cloned, or rather, which new RDF statements should hold for the clone.
+An RDF graph is simply a set of RDF triples, without any internal boundaries reflecting which of these triples make up a given real-world object (such as a login screen).
+Exactly which triples these are is difficult to determine.
+A number of approximations are possible.
+For the remainder of this discussion, let $o$ be a real-world object, $URI_o$ the URI that identifies it, and $G_o$ the RDF graph containing all triples that are considered part of $o$.
+Also, let $s$ be an arbitrary RDF statement, and let $Subj(s)$, $Pred(s)$ and $Obj(s)$ be the subject, predicate and object of $s$, respectively.
+The simplest approximation of which triples comprise a real-world object $o$ would be to select all those triples in which $URI_o$ is either subject or object\footnote{Note: there is potential for confusion between the term \emph{real-world object} (a tree, a person, a login screen) and the \emph{object} of an RDF triple.
+The two are two entirely different things!}:
+\begin{align}
+	\label{equation:simple_object_graph}
+	G_o := \{\;s\;|\;URI_o=Subj(s) \textnormal{ or } URI_o=Obj(s)\}
+\end{align}
+
+However, this definition can easily 
+\begin{inparaenum}[(i)]
+	\item contain false positives (triples that point to $URI_o$, but are not considered part of $o$), 
+	\item lack important triples (e.g., in the case where the object in a triple $<URI, Pred, Obj>$ is itself a complex object), or 
+	\item have dangling blank nodes. 
+\end{inparaenum}
+At least the dangling blank node problem can be addressed by extending the Def.~\ref{equation:simple_object_graph} to the \emph{minimal self-contained graph} (MSG)~\cite{Tummarello:2008kl}, which basically selects the sub-graph so that there are no blank node as leafs.
+Unfortunately, the MSG still does not necessarily give us the desired sub-graph, since problems (i) and (ii) are not addressed.
+At the end of the day, these issues are highly context- and interpretation-dependant, which is why there cannot be a deterministic way to know which triples in a graph belong to a particular real-world object.
+In the general case, we will therefore simply leave the definition of $G_o$ undefined, to be solved individually for each application and use case.
+A SPARQL graph pattern would be well suited for defining any particular occurrence of $G_o$.
+For the FAST use case of defining the object graph of a prototypical building block $p$, the definition of $G_p$ is as given in List.~\ref{list:prototype_graph_definition} (here showing the definition for a screen; other types of building blocks are slightly different).
+
+\singlespacing
+\lstset{
+	caption={SPARQL query to select the object graph of a screen in FAST}, 
+	label=list:prototype_graph_definition,
+	language=sparql
+}
+\begin{figure}
+	\lstinputlisting{exampleCode/screen_object_graph.sparql}
+\end{figure}
+\doublespacing
+
+% paragraph determining_the_correct_sub_graph (end)
+
+\paragraph{Deriving the Clone Graph} % (fold)
+\label{par:deriving_the_}
+
+Now that we know the (application-dependant) graph $G_p$ of the prototypical object $p$, we can define what it means to create a new graph $P_c$ for its clone $c$.
+In prose, the clone graph will contain all statements of the prototypical graph, but all mentions of $URI_p$ will be replaced with $URI_c$.
+Formally, we say that this mapping of $G_p$ into $G_c$ is the prototype mapping $pm$, defined as follows:
 
 \begin{equation}
-S:= \begin{cases}
-i = 0..M,\quad j =0..N  & \text{for }M >N \\
-i=0..M,\quad j=0..M & \text{for }N > M
+pm: G_p \mapsto G_c := \begin{cases}
+<URI_c, Pred, Obj >, \textnormal{ if } URI_p = Obj(s) \\
+<Subj, URI_c, Obj >, \textnormal{ if } URI_p = Pred(s) \\
+<Subj, Pred, URI_c >, \textnormal{ if } URI_p = Subj(s) \\
+<Subj, Pred, Obj >, \textnormal{ if } URI_p \notin s
 \end{cases}
@@
 \cite{kawamoto2006kawawiki} present a semantic wiki, where templates help non-technical users to create semantic descriptions.
+
+% paragraph deriving_the_ (end)
+
+% subsubsection prototype_based_semantics_for_rdf (end)
+
 % subsection classes_instances_and_templates (end)

### 2011-02-07 trunk/deliverables/D5_1_2/fast_d5_1_2.tex
first version of the deliverable for M36
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1252 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-02-04 trunk/deliverables/D5_1_2/fast_d5_1_2.tex
@@
 \usepackage{amssymb}
+\usepackage{amsmath}
+
+% This helps to get hyphenation in typewriter font (\texttt)
+\usepackage[htt]{hyphenat}
+
 \parindent0pt
@@
 \newcommand\workpackageTitle{Semantic catalogue of screen-flow resources and back-end Web Services}
-\newcommand\authorOne{Ismael Rivera, NUIG}
-\newcommand\authorTwo{Author 2}
+\newcommand\authorOne{Ciprian Palaghita, Cyntelix}
+\newcommand\authorTwo{Ismael Rivera, NUIG}
 \newcommand\authorThree{Author 3}
@@
 		{\large \sffamily \authorOne}\\
+		\vspace{0.2cm}
+		{\large \sffamily \authorTwo}\\
 		\vspace{0.5cm}
 		\vfill
-		{\large \sffamily Date: 25/01/2010}\\
+		{\large \sffamily Date: 17/02/2011}\\
 		\vspace{1cm}
@@
 %\lofoot{\small\textcolor{fast@lightgrey}{Task Force Ontologies}}
-\cofoot{\small{FAST 216048 --- \deliverableTitleShort}}
-\rofoot{\small{\thepage}}
+% \cofoot{\small{FAST 216048 --- \deliverableTitleShort}}
+% \rofoot{\small{\thepage}}
+
+\cofoot{\small{FAST $\bullet$ 216048 $\bullet$ \deliverableTitleShort{} $\bullet$ Page \thepage\ of \pageref{LastPage}}}
 
@@
                          \textcolor{white}{\textbf{Change description}}\\ \hline
-1.0 & 25/01/2010 & Ismael Rivera (NUIG) & First version of the deliverable \\ \hline
+1.0 & 19/02/2010 & Ciprian Palaghita (Cyntelix) & First version of the deliverable \\
+ & & Ismael Rivera (NUIG) & \\ \hline
+2.0 & 17/02/2011 & Ciprian Palaghita (Cyntelix) & Second version of the deliverable \\
+ & & Ismael Rivera (NUIG) & \\ \hline
 \end{tabular}
@@
 \begin{tabular}
-	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.28cm}|p{1.6cm}|p{3.29cm}|}
+	{| >{\columncolor{fast@lightgrey}}p{3.25cm}|p{1.4cm}|p{3.2cm}|p{1.6cm}|p{3.37cm}|}
 	\hline
-	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {27/02/2010} & {\textbf{Submitted}} & {27/02/2010}\\ \hline
+	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2011} & {\textbf{Submitted}} & {25/02/2011}\\ \hline
 	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
@@
 \cleardoublepage
-\pagenumbering{arabic}
+% \pagenumbering{arabic}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%
-\rofoot{\small{Page \thepage\ of \pageref{LastPage}}}
+% \rofoot{\small{Page \thepage\ of \pageref{LastPage}}}
 
@@
 
-The deliverable presents both the external and internal architecture in Section~\ref{sec:architecture_overview}, then in Section~\ref{sec:restful_catalogue_api} it is detailed the Catalogue API, query formats, interchange formats, error codes and so on.
+The deliverable presents both the external and internal architecture in Section~\ref{sec:architecture_overview}, then in Section~\ref{sec:catalogue_api} it is detailed the Catalogue API, query formats, interchange formats, error codes and so on.
 
@@
 
-This should be the first version of this deliverable since in M12 the deliverable 5.1 was not scheduled.
-Nevertheless, last year an extra document was written as an appendix to the FAST Catalogue in order to detail and clarify technical decisions and to have a technical reference manual.
-Hence, the deliverable D5.2.1 will evolve to become the D5.1.2, so it will be no longer needed by itself.
-
-There are several new sections and major changes which need to be mentioned:
-
+changes from M24 version:
 \begin{itemize}
-	\item Section~\ref{sec:installation_guide}.
-Installation Guide describes a set of steps to be followed in order to install and execute the FAST Catalogue.
-Also some configuration instructions are specified.
-	\item Section~\ref{sub:content_negotiation}. Content Negotiation states how the FAST Catalogue has adopted several of the content negotiation principles.
-	\item Section~\ref{sec:restful_catalogue_api}.
-RESTful Catalogue API has been intensively extended and modified in order to  reflect all the new functionality and how every request or response is constructed.
-Some of the most important features added in this iteration are:
-	\begin{itemize}
-		\item support for form elements, operators and back-end services (see Section ...),
-		\item convenience operations for recommendation and checking satisfaction and reachability for form elements, operators and back-end services (see Section ...),
-		\item and, last but not least, support for Screen planning (see Section~\ref{sub:planning}).
-	\end{itemize}
-	\item Several appendixes have been added with the JSON structure of every building block the Catalogue can deal with.
-	\item Appendix A. Last year an evaluation to choose the best technology for the Catalogue was done. It was not included in any formal deliverable, so this year it has been documented as an appendix.
+ \item 'find' and 'check' operations for screens (screen-flow design) has been removed. the same functionaly was offered by 'find\&check' doing just 1 request.
+ \item API calls (both find\&check, planner, metadata) updated to reflect the changes in the prototype
+ \item Adds operation for building block cloning.
+ \item added appendix explaining how the recommendation/planning is done
+ \item architecture have been modified reflecting the changes in the catalogue's architecture, and including new components such as the machine learning library (Mahout) and a library to query iServe.
+ \item uses Sindice in the 'ontology importer' mechanism
 \end{itemize}
@@
 A WAR file, which stands for  Web Application Archive, is just a JAR file used to distribute a collection of JavaServer Pages, servlets, Java classes, XML files, tag libraries and static Web pages (HTML and related files) that together constitute a Web application, in this case, the FAST Catalogue.
+
+Note that it is also possible to get the compiled WAR file directly from the SVN server. This file can be located at \url{https://svn.forge.morfeo-project.org/fast-fp7project/distribution/catalogue}.
 
@@
 Independently the type of repository, the parameter \emph{serverURL} needs to take the value of the URL where the FAST Catalogue instance has been deployed.
-If you are testing the application in your machine, it may be something like \url{http://localhost:8080/FASTCatalogue}.
+If you are testing the application in your machine, it may be something like \verb|http://catalogueURL/FASTCatalogue|.
 
@@
 \begin{verbatim}
-storageDir=D:\\FAST\\catalogue\\repository
+storageDir=/opt/fast-catalogue/repository
 \end{verbatim}
@@
 sesameServer=http://localhost:8880/openrdf-sesame
-repositoryID=c2
+repositoryID=repository-ex1
 \end{verbatim}
@@
 It is recommended to check if the FAST Catalogue has been successfully installed.
-To do that, a simple HTTP request, using a Web browser, can be sent to the URL the Catalogue is deployed, and specifying a type of resource such as Screen.
-The URL would be similar to \url{http://localhost:8080/FASTCatalogue/screens}.
+To do that, a simple HTTP request, using a Web browser, can be sent to the URL the Catalogue is deployed, and specifying a type of building block such as Screen.
+The URL would be similar to \verb|http://catalogueURL/FASTCatalogue/screens|.
 Depending on the variant chosen, the response may be different.
@@
 
-This section presents a technical overview of the internal catalogue's architecture and which external components will interact with it.
-Basically three layers can be distinguished: Presentation, Business Logic and Persistence layer (see Figure~\ref{fig:catalogue_architecture}).
+This sections presents a technical overview of the Catalogue's architecture (see Figure~\ref{fig:catalogue_architecture}), and how the interaction with external applications or APIs is done.
+The Catalogue platform has a modular design, which are spread in basically three layers, depending on the nature of each component:
+\begin{itemize}
+ \item The top layer represents the public API for the Catalogue.
+It puts together all the available information and services to offer the maximum powerful and flexibility for other IDEs or applications.
+In FAST, here is where the Gadget Visual Storyboard
+ \item The core layer implements the object model, and provides the business logic for the different processes involved in the Catalogue, integrating third party libraries where needed.
+ \item The pumbling layer offers an abstraction to RDF model logic and triple stores.
+\end{itemize}
 
@@
 
-The Presentation layer will be the public interface of the catalogue.
-The main purpose of this layer is to provide its functionality to other components within the Gadget Visual Studio.
-The service is offered in a REST (REpresentational State Transfer) style, making it easy to consume than other complex APIs.
+\subsection{Presentation layer / API}
+
+The Presentation layer will be the public interface or API of the Catalogue.
+The main purpose of this layer is to provide its functionality to the Gadget Visual Storyboard, or any other third party IDE or application.
+The API is offered in a REST (REpresentational State Transfer) style, making it easy to consume than other complex APIs.
 REST is an architectural style, not a toolkit or a standard.
-Even though, it makes use of standards like HTTP, URI, Resource Representations (XML, HTML, JSON, JPEG, etc.) or MIME types (text/xml, text/html, application/json, image/jpeg, etc.).
+Even though, it makes use of standards like HTTP, URI, Resource Representations or MIME types.
 Another characteristic of the REST style is its stateless assumption.
-The catalogue adopts this strategy; therefore every request needs a complete set of information in order to prepare the response.
-
-As part of the Presentation layer, a SPARQL endpoint is offered using the SPARQL protocol service as defined in the SPROT \cite{sprot} specification.
-The SPARQL endpoint is mostly offered to enable other developers to query directly the catalogue knowledge base using SPARQL queries.
+The Catalogue adopts this strategy; therefore every request needs a complete set of information in order to prepare the response.
+The API provided is explained in detail in Section~\ref{sec:catalogue_api}.
+
+As part of the Presentation layer, a SPARQL endpoint has been implemented using the SPARQL protocol service as defined in the SPROT \cite{sprot} specification.
+The SPARQL endpoint is mostly offered to enable other developers to query directly the Catalogue knowledge base using SPARQL queries.
 This feature is supported by the Sesame RESTful HTTP interface for SPARQL Protocol for RDF.
 
-The Business Logic layer contains all the FAST domain-specific processing.
-It provides functions to interact with all the elements of the domain model specified in \cite{moeller2009fast_ontology}.
-It processes the input from the Presentation layer, creating specific objects modelling the business logic, and vice versa.
-In addition it interacts with the Persistence layer in order to persist the model.
-
-The Persistence layer provides an API which allows the Business layer to work with a standard set of objects that read and save their state to the triple store.
-For this reason, on top of this layer it is used RDF2Go, an abstraction over triple (and quad) stores.
+\subsection{Core layer}
+
+The Core or Business Logic layer contains all the FAST domain-specific processing.
+The resources and relations between each other, which define the domain model, are extracted from the FAST ontology~\cite{moeller2011fast_ontology}.
+These layer provides support to the functions defined in the API, and interact with the persistence layer to retrieve and store any information of the data model.
+
+This layer provides several mechanisms to search and browse these different resources.
+To know more about these capabilities, please read Sections~\ref{ssub:screen_findcheck},~\ref{ssub:screen_component_findcheck} and~\ref{sub:planning} where the API calls are defined, and Appendix F for more details about the implementation.
+
+At the bottom part of the Gadget Architecture~\cite{reyes2011gadget_architecture} are the back-end services.
+These are RESTful services, SOAP-based web services, or similar.
+In FAST, these web services are wrapped, extracting their definition and encapsulation their behaviour.
+To creation of these service wrappers is on demand, and the input for this process is a particular definition of a web service.
+However, the Catalogue does not crawl the Web looking for Web services.
+Instead, and to enrich the collection of available web services, the Catalogue integrates and queries iServe\footnote{http://iserve.kmi.open.ac.uk/}, a platform for publishing Semantic Web Services as linked data, no matter their original format.
+Therefore, thousands of web services defined or annotated in a semantic way, are also served from the Catalogue, in a uniform manner.
+
+The Catalogue is a semantic application.
+As a result, it has to handle with ontologies.
+Apart from the core ontologies being used across the whole application (see~\cite{moeller2011fast_ontology}), the building blocks may reference classes and properties from diverse vocabularies and ontologies.
+To have a better understanding of them, the Catalogue imports them inside its repository any time it detects a building block is using an unknown class, property, or in other words, an unknown ontology.
+While importing an ontology, the Catalogue tries to get its definition from the URI of a given class or property (as stated in~\cite{berrueta2008}).
+However, sometimes this may be not succesful, and the Catalogue will try to get its definition querying Sindice\footnote{http://www.sindice.com/}, a semantic search engine, which cralws the web and keeps provenance metadata for the ontologies.
+
+\subsection{Plumbing layer}
+
+The Plumbing layer abstract the interaction with triple stores, RDF models and physical persistence to disk, database, etc.
+It provides the different mechanisms to allow the Catalogue the storage of the domain objects into a triple store in its underlying representation of RDF.
+
+There are many triple stores solutions, any of them with its own advantages and disadvantages.
+To make the Catalogue independent of the solution adopted, it makes uses of RDF2Go, an abstraction over triple (and quad) stores.
 The RDF2Go API allows interacting with the semantic representation of the model (triples) in a generic manner, and also brings the flexibility of choosing different triple stores to persist them.
-The selected implementation for the triple stored is Sesame 2.
-Moreover, this framework is completely extensible and configurable regarding to storage mechanisms, inferencers, RDF file formats, query result formats and query languages.
-For the actual catalogue prototype, the storage mechanism used is the native file storage system of Sesame, the inferencer used is a subset of the RDFS entailment rules \cite{rdfsrules} following a forward-chaining policy.
+
+The current implementation of the Catalogue prototype at M36 uses Sesame 2 as triple store.
+This solution has a wide community behind, and it has attractive advantages such as being completely extensible and configurable regarding to storage mechanisms, inferencers, RDF file formats, query result formats and query languages.
 
@@
 \clearpage
-\section{RESTful Catalogue API} % (fold)
-\label{sec:restful_catalogue_api}
+\section{Catalogue API} % (fold)
+\label{sec:catalogue_api}
 
@@
   "canvas": [
-    { "uri": "http://localhost:8080/catalogue/screens/238" },
-    { "uri": "http://localhost:8080/catalogue/screens/323" }
-  ],
-  "elements": [
-    { "uri": "http://localhost:8080/catalogue/screens/636" }
+    { "uri": "http://catalogueURL/screens/238" },
+    { "uri": "http://catalogueURL/screens/323" }
+  ],
+  "palette": [
+    { "uri": "http://catalogueURL/screens/636" }
   ],
@@
     ],
-    "user": "irivera",
+    "user": "http://ismaelrivera.es/#me",
   },
@@
   "label": {
-    "en-GB": "Simple Example",
-    "es-ES": "Ejemplo Sencillo",
-    "de-DE": "Einfaches Beispiel"
+    "en-GB": "Product Search",
+    "es-ES": "Búsqueda de productos",
+    "de-DE": "Produktsuche"
   },
   "description": {
-    "en-GB": "This is a simple example of a screen",
-    "es-ES": "Esto es un ejemplo sencillo de pantalla",
-    "de-DE": "Dies ist ein einfaches Beispiel für einen Bildschirm"
-  },
-}
-\end{verbatim}
-\doublespacing
-
-\singlespacing
-\begin{verbatim}
-<rdf:Description rdf:about="http://www.morfeoproject.eu/fast/fco#ExampleScreen">
-  <rdf:type rdf:resource="http://www.morfeoproject.eu/fast/fco#Screen"/>
-  <rdfs:label xml:lang="en-GB">Simple Example</rdfs:label>
-  <rdfs:label xml:lang="es-ES">Ejemplo Sencillo</rdfs:label>
-  <rdfs:label xml:lang="de-DE">Einfaches Beispiel</rdfs:label>
-  <description
-    xmlns="http://purl.org/dc/terms/"
-    xml:lang="en-GB">This is a simple example of a screen</description>
-  <description
-    xmlns="http://purl.org/dc/terms/"
-    xml:lang="es-ES">Esto es un ejemplo sencillo de pantalla</description>
-  <description
-    xmlns="http://purl.org/dc/terms/"
-    xml:lang="de-DE">Dies ist ein einfaches Beispiel für einen Bildschirm</description>
+    "en-GB": "On this screen, the user can enter a search term.",
+    "es-ES": "En esta ventana, el usuario puede introducir el criterio de búsqueda",
+    "de-DE": "Auf diesem Screen kann der Benutzer einen Suchbegriff eingeben."
+  }
+}
+\end{verbatim}
+\doublespacing
+
+\singlespacing
+\begin{verbatim}
+<rdf:RDF
+    xmlns:rdf="http://www.w3.org/1999/02/22-rdf-syntax-ns#"
+    xmlns:foaf="http://xmlns.com/foaf/0.1/"
+    xmlns:dc="http://purl.org/dc/terms/"
+    xmlns:fgo="http://purl.oclc.org/fast/ontology/gadget#"
+    xmlns:xsd="http://www.w3.org/2001/XMLSchema#"
+    xmlns:rdfs="http://www.w3.org/2000/01/rdf-schema#"
+    xmlns:ctag="http://commontag.org/ns#">
+  <rdf:Description rdf:about="http://demo.fast.morfeo-project.org/FASTCatalogue/screens/1">
+    <rdf:type rdf:resource="http://purl.oclc.org/fast/ontology/gadget#Screen"/>
+    <rdfs:label xml:lang="en-gb">Product Search</rdfs:label>
+    <rdfs:label xml:lang="en-es">Búsqueda de productos</rdfs:label>
+    <rdfs:label xml:lang="en-de">Produktsuche</rdfs:label>
+    <dc:description xml:lang="en-gb">On this screen, the user can enter a search
+    term.</dc:description>
+    <dc:description xml:lang="en-gb">En esta ventana, el usuario puede introducir
+    el criterio de búsqueda.</dc:description>
+    <dc:description xml:lang="en-gb">Auf diesem Screen kann der Benutzer einen
+    Suchbegriff eingeben.</dc:description>
   </rdf:Description>
-\end{verbatim}
-\doublespacing
-
-Until now, using internationalisation is allowed for the label and the description of a screen, but in the near future, more attributes will benefit from this feature.
+</rdf:RDF>
+\end{verbatim}
+\doublespacing
+
+Internationalisation is being used by the attributes \emph{label} and \emph{description} of any building block, \emph{label} of the pre-/post-conditions and \emph{label} of the tags.
 
@@
 
-\subsection{Content Negotiation} % (fold)
-\label{sub:content_negotiation}
-
-Before any major details about how content negotiation is addressed by the catalogue, a few concepts need to be understood.
-First of all, say a URI is not a file name.
-It is common to see a URI as the location for a file, but a URI or Universal Resource Identifier is not a URL: it is not a file name or location, it is an identifier (or a reference) to a resource.
-Then by dereferencing a URI the actual resource can be retrieved.
-Let's see a request to retrieve the resource "Screen 123" is sent to the server.
-The resource is just a piece of information, however the server will answer an HTML document with a text describing the screen or an image representing it.
-On the Web, these equivalent representations of a resource are called variants, and content negotiation is the mechanism used to determine which of the representations of most appropiate for a given request.
+\subsection{Linked Data} % (fold)
+\label{sub:linked_data}
+
+Apart from serving as the back-end system for the Gadget Visual Storyboard (GVS), the catalogue publishes all FAST building blocks as linked data, following the principles as originally defined in \cite{bernersLee2006linkedData}, in order to make them available to arbitrary third party applications.
+Each building block is identified by an HTTP URI and hosted through the catalogue so that it can be dereferenced through the same URI.
+For each building block, data is available in representations in different standard formats such as JSON (for communication with the GVS), RDF/XML, Turtle, or even HTML+RDFa as a human-readable version.
+Following the principles and best practices proposed in \cite{berrueta2008} and \cite{sauermann2008cool_uris}, these representations are served based on the request issued by the requesting agent, using a technique called \emph{content negotiation}.
+As required by the forth rule of linked data, individual building blocks link to other data on the Web, thereby preventing so-called isolated ``data islands''.
+
+In order to understand how content negotiation is implemented in the catalogue, a few concepts need to be understood.
+The URI of a particular building block must be understood as the identifier of the building block as such, as opposed to a particular representation (JSON, HTML, etc.) of it.
+In fact, each such representation has its own URI, under which it can be retrieved.
+Using the terminology established in \cite{w3c2004webArchitecture}, the building block is a \emph{non-information resource}, whereas each representation is an \emph{information resource}.
+In the process of content negotiation, the requesting agent will first request the URI of the building block as such (e.g., \verb|http://catalogueURL/screens/235|), as well as the required representation format (e.g., JSON).
+The catalogue will then inspect the request and redirect the requester to the appropriate representation (e.g., \verb|http://catalogueURL/screens/235.json|), as illustrated in Fig.~\ref{fig:content_negotiation}.
+On the Web, different representations for the same resource are called \emph{variants}, and content negotiation is the mechanism used to determine which of the representations of most appropriate for a given request.
 
@@
 
-In summary, the basic idea of content negotiation, as stated in the HTTP 1.1 specification ~\cite{http1.1}, is to serve the best variant for a resource, taking into account what variants are available, what variants the server may prefer to serve, what the client can accept, and with which preferences: in HTTP, this is done by the client which may send, in its request, Accept headers (Accept, Accept-Language and Accept-Encoding), to communicate its capabilities and preferences in Format, Language and Encoding, respectively.
-
-In fact, what the catalogue really does is "format negotiation" since the alternate representationos are just based on the selection of the media type, through the Accept header, but does not consider different languages or encoding types.
-The formats supported are JSON, RDF+XML and HTML.
+In summary, the basic idea of content negotiation, as stated in the \cite{http1.1}, is to serve the best variant for a resource, taking into account what variants are available, what variants the server may prefer to serve, what the client can accept, and with which preferences.
+In HTTP, this is done by the client which may send, in its request, accept headers (\texttt{Accept}, \texttt{Accept-Language} and \texttt{Accept-Encoding}), to communicate its capabilities and preferences in format, language and encoding, respectively.
+
+In fact, what the catalogue really does is ``format negotiation'' since the alternate representations are just based on the selection of the media type, through the accept header, but does not consider different languages or encoding types.
+The formats supported are JSON, RDF/XML, Turtle and HTML+RDFa.
+Even though the accept header is desired, the different representations can also be retrieved directly by dereferencing their own URI.
+The convention used in the catalogue for deriving a representation URI from a building block URI is to simply attach a matching suffix, such as \texttt{.rdf}, \texttt{.json} or \texttt{.html}.
 
@@
 
-This section contains detailed descriptions of the interface provided by the catalogue via REST services, detailing request parameters, response elements, any special errors and examples of requests and responses.
-The URL format is also specified for each operation, where 'catalogueURL' has to be replaced for the real URL the service is installed (e.g.
+The Catalogue API is a programmatic interface into many of the Catalogue's features.
+It includes not only control over creation, modification and deletion of building blocks, but it has been extended to support advanced search and recommendation capabilities.
+
+This section contains detailed descriptions of the operations permitted by the API, detailing request parameters, response elements, any special errors and examples of requests and responses.
+The URL format is also specified for each operation, where \emph{catalogueURL} has to be replaced for the real URL the service is installed (e.g.
 http://demo.fast.morfeo-project.org/catalogue).
 
-\subsubsection{CRUD operations} % (fold)
-
-Any building block within the Catalogue is a resource in terms of REST philosophy, hence it can be created, retrieved, modified or deleted, using a certaing URL and a specific HTTP method for every operation.
-Two concepts have to be defined: \emph{collection}, as a set of resources which access URI is http://catalogueURL/<type>/ where catalogueURL is the specific URL where the Catalogue server is installed and <type> is the plural of the name of the building block (e.g.
-screens, services), and \emph{member} of the collection, in other words, the building block itself, which access URI is http://catalogueURL/<type>/<id> where the <id> has to be replaced by the identifier of a specific building block.
+\subsection{Managing Building Blocks} % (fold)
+\label{sub:managing_building_blocks}
+
+Any building block within the Catalogue is a resource in terms of REST philosophy, therefore it can be created, retrieved, modified or deleted, using a certaing URL and a specific HTTP method for every operation.
+Two concepts should be clarified: \emph{collection}, as a set of resources which access URI is \verb|http://catalogueURL/[type]| where \emph{catalogueURL} is  where the Catalogue server is installed and \verb|[type]| is the plural of the name of the building block (e.g.
+screens, operators), and \emph{member} of the collection, in other words, the building block itself, which access URI is \verb|http://catalogueURL/[type]/[id]| where \verb|[id]| has to be replaced by the identifier of a specific building block.
 The details of the operations and which HTTP verb has to be used can be found in the Table~\ref{tab:crud_operations}.
+These methods are to what is called \emph{Prototype} of building blocks (see \cite{moeller2011fast_ontology}).
 
@@
 
-To create a new screen, a POST request is send to the catalogue server, including the JSON representation of the screen as the body of the request:
+To create a new screen, a POST request is sent to the catalogue server, concretely to \texttt{http://cata\-logueURL/screens}, including the JSON representation of the screen in the body of the request:
 
@@
   "creationDate": "2010-01-26T17:01:13+0000",
-  "creator": "irivera",
+  "creator": "http://ismaelrivera.es/#me",
   "description": {"en-gb": "Please fill the description..."},
@@
   "label": {"en-gb": "Amazon Product List"},
-  "name": "ProductList",
   "postconditions": [
-    [
     {
@@
       "label": { "en-gb": "An item" },
-        "pattern": "?I 
+      "pattern": "?item 
                   http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                    http://aws.amazon.com/AWSECommerceService#Item",
+                  http://fast.morfeo-project.org/ontologies/amazon#Item",
       "positive": "true"
     }
-    ]
   ],
   "preconditions": [
-    [
     {
       "id": "filter",
-        "label": { "en-gb": "A search criteria" },
-        "pattern": "?F
+      "label": { "en-gb": "A search request" },
+      "pattern": "?search
                   http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                    http://aws.amazon.com/AWSECommerceService#SearchCriteria",
+                  http://fast.morfeo-project.org/ontologies/amazon#SearchRequest",
       "positive": "true"
     }
-    ]
   ],
@@
   "tags": [{"label": {"en-gb": "amazon"}}],
-  "version": "0.1"
+  "version": "1.0"
 }
@@
   "label": {"en-gb": "Prueba1"},
-  "name": "ProductList",
   "postconditions": [
-    [
       {
@@
         "label": { "en-gb": "An item" },
-        "pattern": "?I
+        "pattern": "?item
                     http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                    http://aws.amazon.com/AWSECommerceService#Item",
+                    http://fast.morfeo-project.org/ontologies/amazon#Item",
         "positive": "true"
@@
         "id": "filter",
-        "label": { "en-gb": "A search criteria" },
-        "pattern": "?F
+        "label": { "en-gb": "A search request" },
+        "pattern": "?search
                     http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                    http://aws.amazon.com/AWSECommerceService#SearchCriteria",
+                    http://fast.morfeo-project.org/ontologies/amazon#SearchRequest",
         "positive": "true"
@@
   "type": "screen",
-  "uri": "http://localhost:8080/catalogue/screens/28",
+  "uri": "http://catalogueURL/screens/28",
   "version": "0.1"
@@
     "creationDate": "2010-01-26T17:01:13+0000",
-    "creator": "irivera",
+    "creator": "http://ismaelrivera.es/#me",
     "description": {"en-gb": "Please fill the description..."},
@@
     "label": {"en-gb": "Prueba1"},
-    "name": "ProductList",
     "postconditions": [
@@
           "label": { "en-gb": "An item" },
-          "pattern": "?I
+          "pattern": "?item
                       http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                      http://aws.amazon.com/AWSECommerceService#Item",
+                      http://fast.morfeo-project.org/ontologies/amazon#Item",
           "positive": "true"
@@
           "id": "filter",
-          "label": { "en-gb": "A search criteria" },
-          "pattern": "?F
+          "label": { "en-gb": "A search request" },
+          "pattern": "?search
                       http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                      http://aws.amazon.com/AWSECommerceService#SearchCriteria",
+                      http://fast.morfeo-project.org/ontologies/amazon#SearchRequest",
           "positive": "true"
@@
     "type": "screen",
-    "uri": "http://localhost:8080/catalogue/screens/28",
+    "uri": "http://catalogueURL/screens/28",
     "version": "0.1"
@@
   {
-    "creationDate": "2009-02-07T09:59:52+0000",
+    "creationDate": "2011-02-07T09:59:52+0000",
     "creator": "fabio",
     ...,
-	  "uri": "http://localhost:8080/catalogue/screens/14",
+    "uri": "http://catalogueURL/screens/14",
     "version": "1.0"
@@
   {
-    "creationDate": "2009-02-07T09:59:52+0000",
+    "creationDate": "2011-02-07T09:59:52+0000",
     "creator": "javier",
     ...,
-	  "uri": "http://localhost:8080/catalogue/screens/564",
+    "uri": "http://catalogueURL/screens/564",
     "version": "1.0"
@@
 
-\subsubsection{Screen Find} % (fold)
-\label{ssub:screen_find}
-
-The find operation searches inside the catalogue for any screen which could be somehow related to the gadget the user is creating.
-It provides a recommended set of screens depending on the domain context, the canvas, and so on.
-
-The specific URL to access this operation is http://catalogueURL/screens/find using HTTP POST method.
-From now on, the method 'find' only considers the domain context and the pre/postconditions of the screens.
-It will try to satisfy all the unsatisfied preconditions of a given list of screens also known as canvas.
-The results will be all the screens stored in the catalogue which fulfil some of these preconditions, and are tagged with the tags specified in the domain context.
-The request parameters are shown in Table~\ref{tab:screen_find_request} and the response parameters are shown in Table~\ref{tab:screen_find_response}.
+\subsubsection{Prototype cloning} % (fold)
+\label{ssub:prototype_cloning}
+A building block can be cloned.
+Making a clone of a prototype is as simple as sending a HTTP POST request to \verb|http://catalogueURL/cloning| indicating the URI of the building block to be clone, as in this example:
+
+\singlespacing
+\begin{verbatim}
+{
+  "uri": "http://catalogueURL/forms/7"
+}
+\end{verbatim}
+\doublespacing
+
+Example of response:
+
+\singlespacing
+\begin{verbatim}
+{
+  "clone": "http://catalogueURL/forms/clones/745"
+  "uri": "http://catalogueURL/forms/7"
+}
+\end{verbatim}
+\doublespacing
+
+As clones are building blocks as well, the operations explained in Section~\ref{sub:managing_building_blocks} are valid to retrieve and delete them.
+
+\subsubsection{GetMetadata} % (fold)
+
+Retrieves all the metadata about a list of building blocks or classes (concepts) stored in the catalogue. All building blocks are supported, no matter if it is a prototype or a clone.
+
+The URL to access this operation is \verb|http://catalogueURL/getmetadata| using HTTP POST method.
+Table~\ref{tab:getmetadata_request} shows the parameters needed for the invocation and Table~\ref{tab:getmetadata_response} details the different parameters of a response.
 
 \begin{table}[htb]
-\caption{Screen Find Request Parameters}
-\label{tab:screen_find_request}
+\caption{GetMetadata Request Parameters}
+\label{tab:getmetadata_request}
 \begin{center}
-\begin{tabular}{|p{2cm}|p{9.5cm}|p{2cm}|}
+\begin{tabular}{|p{2.5cm}|p{9cm}|p{2cm}|}
 \hline
@@
                          \textcolor{white}{Type}\\ \hline
-canvas & The canvas is composed by a list of screens. Only the URI is needed. & Optional \\ \hline
-domainContext & The domain context contains a list of tags and a user. & Optional \\ \hline
-elements & It is a list of resources, previously stored in the catalogue. For example, the list of screens recommended last execution. Only accepts screens. & Optional \\ \hline
+N/A & A list of URIs. & Required\\ \hline
 \end{tabular}
@@
 \begin{table}[htb]
-\caption{Screen Find Response Parameters}
-\label{tab:screen_find_response}
+\caption{GetMetadata Response Parameters}
+\label{tab:getmetadata_response}
 \begin{center}
-\begin{tabular}{|p{2cm}|p{9.5cm}|p{2cm}|}
+\begin{tabular}{|p{2.5cm}|p{11.5cm}|}
+\hline
+\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
+                         \textcolor{white}{Description}\\ \hline
+screen-flows & A list of screen-flows with all the metadata associated to them.\\ \hline
+screens & A list of screens with all the metadata associated to them.\\ \hline
+forms & A list of forms with all the metadata associated to them.\\ \hline
+operators & A list of operators with all the metadata associated to them.\\ \hline
+backendservices & A list of back-end services with all the metadata associated to them.\\ \hline
+classes & A list of classes or concepts, with metadata such as label, properties, subclasses, etc.\\ \hline
+\end{tabular}
+\end{center}
+\end{table}
+
+Example request:
+
+\singlespacing
+\begin{verbatim}
+[
+  "http://catalogueURL/screens/48",
+  "http://catalogueURL/forms/clones/339",
+  "http://fast.morfeo-project.org/ontologies/amazon#Item",
+]
+\end{verbatim}
+\doublespacing
+
+Example response:
+
+\singlespacing
+\begin{verbatim}
+{
+  "screenflows": [],
+  "screens": [
+    {
+      "creationDate": "2011-02-07T09:59:52+0000",
+      "creator": "http://ismaelrivera.es/#me",
+      ...,
+      "uri": "http://catalogueURL/screens/48",
+      "version": "1.0"
+    }
+  ],
+  "forms": [
+    {
+      "actions": [
+        {
+          "name": "init",
+          "preconditions": [],
+          "uses": []
+        },
+        {
+          "name": "showTable",
+          "preconditions": [{
+            "id": "list",
+            "label": {"en-gb": "A product list"},
+            "pattern": "?urlList
+                        http://www.w3.org/1999/02/22-rdf-syntax-ns#type
+                        http://fast.morfeo-project.org/ontologies/amazon#ProductList",
+            "positive": true
+          }],
+          "uses": []
+        }
+      ],
+      ...,
+      "uri": "http://catalogueURL/FASTCatalogue/forms/clones/249",
+      "version": "1.0"
+    },
+  ],
+  "operators": [],
+  "backendservices": []
+  "classes": [
+    {
+      "attributes": [
+        {
+          "type": "http://xmlns.com/foaf/0.1/Document",
+          "uri": "http://fast.morfeo-project.org/ontologies/amazon#smallImage"
+        },
+        {
+          "type": "http://xmlns.com/foaf/0.1/Document",
+          "uri": "http://fast.morfeo-project.org/ontologies/amazon#mediumImage"
+        },
+        {
+          "type": "http://xmlns.com/foaf/0.1/Document",
+          "uri": "http://fast.morfeo-project.org/ontologies/amazon#largeImage"
+        },
+        {
+          "type": "http://www.w3.org/2001/XMLSchema#double",
+          "uri": "http://fast.morfeo-project.org/ontologies/amazon#price"
+        }
+      ],
+      "label": {"en": "An item handled by the Amazon Service"},
+      "tags": [{
+        "label": {"en": "amazon"},
+        "means": "http://www.amazon.com"
+      }],
+      "uri": "http://fast.morfeo-project.org/ontologies/amazon#Item"
+    }
+  ]
+}
+\end{verbatim}
+\doublespacing
+
+\subsubsection{Screen Find \& Check} % (fold)
+\label{ssub:screen_findcheck}
+
+Searches screens and check the compatibility/reachability of a given screen-flow.
+
+\begin{table}[htb]
+\caption{Screen Find \& Check Request Parameters}
+\label{tab:screen_check_request}
+\begin{center}
+\begin{tabular}{|p{2.5cm}|p{9cm}|p{2cm}|}
 \hline
@@
                          \textcolor{white}{Type}\\ \hline
-N/A & A list of recommended screens URIs is returned. & Required \\ \hline
+canvas & The canvas is composed by a list of screens (clones). Only the URI is needed. & Required\\ \hline
+palette & It is a list of screens (prototypes), not in the canvas but which must be check based on the specified criterion. & Required\\ \hline
+domainContext & The domain context contains a list of tags and the user who makes the request. & Required\\ \hline
+criterion & The criterion specifies what it has to be check. The only possible supported now is 'reachability' in order to check if the pre-conditions of a screen are satisfied. & Required\\ \hline
 \end{tabular}
@@
 
-Following is shown an example of usage of this operation considering a canvas with the screen \texttt{http://localhost:8080/catalogue/screens/654}.
-
-\singlespacing
-\begin{verbatim}
-{
-  canvas: [
-    { uri: "http://localhost:8080/catalogue/screens/654" }
-  ],
-  domainContext: {
-    tags: [],
-	  user: null
-  },
-  elements: []
-}
-\end{verbatim}
-\doublespacing
-
-After the execution of the recommendation algorithm, the response given by the catalogue is:
-
-\singlespacing
-\begin{verbatim}
-[
-  "http://localhost:8080/catalogue/screens/371",
-  "http://localhost:8080/catalogue/screens/24",
-  "http://localhost:8080/catalogue/screens/253",
-  "http://localhost:8080/catalogue/screens/12"
-]
-\end{verbatim}
-\doublespacing
-
-\subsubsection{Screen Check} % (fold)
-\label{ssub:screen_check}
-
-This operation verifies the state of certain list of screens depending on a specific criterion.
-The only criterion accepted in this stage of the prototype is 'reachability'.
-A screen will be reachable if it all its preconditions are satisfied by postconditions of the screens the canvas contain.
-The specific URL to access this operation is http://catalogueURL/screens/check using HTTP POST method.
-
 \begin{table}[htb]
-\caption{Screen Check Request Parameters}
-\label{tab:screen_check_request}
+\caption{Screen Find \& Check Response Parameters}
+\label{tab:screen_check_response}
 \begin{center}
-\begin{tabular}{|p{2cm}|p{9.5cm}|p{2cm}|}
+\begin{tabular}{|p{2.5cm}|p{11.5cm}|}
 \hline
 \rowcolor{fast@lightgrey}\textcolor{white}{Name} &
-                         \textcolor{white}{Description} &
-                         \textcolor{white}{Type}\\ \hline
-canvas & The canvas is composed by a list of screens. Only the URI is needed. & Optional\\ \hline
-domainContext & The domain context contains a list of tags and a user. & Optional\\ \hline
-elements & It is a list of resources, previously stored in the catalogue. For example, the list of screens recommended last execution. Only accepts screens. & Optional\\ \hline
-criterion & The criterion specifies what it has to be check. The only possible supported now is 'reachability' in order to check if the preconditions of a screen are satisfied. & Required\\ \hline
+                         \textcolor{white}{Description}\\ \hline
+canvas & List of screens in the canvas, which extra information indicating satisfaction of the criterion, for the screens themselves and their pre-conditions.\\ \hline
+palette & A list of screens obtained by the search, attaching information about the satisfaction of the criterion.\\ \hline
+domainContext & It contains the same value as in the request.\\ \hline
+limit & Integer. It is used to limit the number of results returned. Default value is -1 (no limit).\\ \hline
+offset & Integer. It is used to jump or skip a number of results to be returned. Default value is 0.\\ \hline
 \end{tabular}
@@
 
-\begin{table}[htb]
-\caption{Screen Check Response Parameters}
-\label{tab:screen_check_response}
-\begin{center}
-\begin{tabular}{|p{2cm}|p{9.5cm}|p{2cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
-                         \textcolor{white}{Description} &
-                         \textcolor{white}{Type}\\ \hline
-Canvas & The canvas indicating which screens satisfy the critetion.
-More aditional information may be included, for example, a satisfaction attribute for every precondition for reachability.
-& Optional\\ \hline
-Domain Context & The domain context contains a list of tags and a user. & Optional\\ \hline
-Elements & The elements indicating which satisfy the critetion. More aditional information may be included as in the Canvas. & Optional\\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-The scenario for the following example is composed by a screen in the canvas which precondition is a foaf:Person which foaf:workplaceHomepage is \url{http://www.deri.ie/}, and a screen in the elements list which precondition is a foaf:Person.
-The state to check in this case is the reachability and satisfaction of the screens and preconditions.
+Example of the request:
 
@@
   "canvas": [
-    { "uri": "http://localhost:8080/catalogue/screens/238" }
-  ],
-  "elements": [
-    { "uri": "http://localhost:8080/catalogue/screens/636" }
+    { "uri": "http://catalogueURL/screens/clones/338" }
+  ],
+  "palette": [
+    { "uri": "http://catalogueURL/screens/636" }
   ],
@@
 
-The response shows that there are not reachable screens at the moment, and preconditions are not satisfied.
+The response for the above request is:
 
@@
     "reachability": false,
-    "uri": "http://localhost:8080/catalogue/screens/238"
+    "uri": "http://catalogueURL/screens/clones/338"
   }],
-  "elements": [{
+  "palette": [
+    {
       "preconditions": [{
@@
       "reachability": false,
-    "uri": "http://localhost:8080/catalogue/screens/636"
-  }]
-}
-\end{verbatim}
-\doublespacing
-
-\subsubsection{Screen Find \& Check} % (fold)
-\label{ssub:screen_findcheck}
-
-For convenience and to minimise the number of request, the Find operation may be combined with the Check operation.
-This operation will look for new screens based on the given information, and will perform the Check to both the screens in the canvas and the screens in the \emph{elements} attribute.
-
-The format of the request is the same as the one used in the Check operation:
-
-\singlespacing
-\begin{verbatim}
-{
-  "canvas": [
-    { "uri": "http://localhost:8080/catalogue/screens/238" }
-  ],
-  "elements": [
-    { "uri": "http://localhost:8080/catalogue/screens/636" }
-  ],
-  "domainContext": {
-    "tags": [
-      {
-        "label": { "en-GB": "Amazon" },
-        "means": "http://dbpedia.org/page/Amazon.com"
-      }
-    ],
-    "user": "irivera",
-  },
-  criterion: "reachability"
-}
-\end{verbatim}
-\doublespacing
-
-The response for the above request is:
-
-\singlespacing
-\begin{verbatim}
-{
-  "canvas": [{
-    "preconditions": [{
-      "label": {"en-gb": "A person working in DERI"},
-      "pattern": "?person rdf:type foaf:Person . 
-                  ?person foaf:workplaceHomepage http://www.deri.ie/",
-      "positive": true,
-      "satisfied": false
-    }],
-    "reachability": false,
-    "uri": "http://localhost:8080/catalogue/screens/238"
-  }],
-  "elements": [
-    {
-      "preconditions": [{
-        "label": {"en-gb": "A person"},
-        "pattern": "?person rdf:type foaf:Person",
-        "satisfied": false
-      }],
-      "reachability": false,
-      "uri": "http://localhost:8080/catalogue/screens/636"
+      "uri": "http://catalogueURL/screens/636"
     },
@@
       "reachability": true,
-      "uri": "http://localhost:8080/catalogue/screens/132"
+      "uri": "http://catalogueURL/screens/132"
     }
@@
 
-\subsubsection{GetMetadata} % (fold)
-
-Although the information about a resource can be obtained by its specific retrieval operation supported by the CRUD interface, sometimes is needed to retrieve information which would imply a request per resource, hence this operation allow to get the metadata of a list of resources in only one request.
-This prototype supports this operation for screen-flows and screens.
-The specific URL to access this operation is http://catalogueURL/getmetadata using HTTP POST method.
-Table~\ref{tab:getmetadata_request} shows the parameters needed for the invocation and Table~\ref{tab:getmetadata_response} details the different parameters may contain a response.
-
-\begin{table}[htb]
-\caption{GetMetadata Request Parameters}
-\label{tab:getmetadata_request}
-\begin{center}
-\begin{tabular}{|p{2cm}|p{9.5cm}|p{2cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
-                         \textcolor{white}{Description} &
-                         \textcolor{white}{Type}\\ \hline
-N/A & A list of URIs. & Required\\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\begin{table}[htb]
-\caption{GetMetadata Response Parameters}
-\label{tab:getmetadata_response}
-\begin{center}
-\begin{tabular}{|p{2cm}|p{9.5cm}|p{2cm}|}
-\hline
-\rowcolor{fast@lightgrey}\textcolor{white}{Name} &
-                         \textcolor{white}{Description} &
-                         \textcolor{white}{Type}\\ \hline
-screen-flows & A set of screen-flows with all the metadata associated to them.. & Optional\\ \hline
-screens & A set of screens with all the metadata associated to them. & Optional\\ \hline
-forms & A set of forms with all the metadata associated to them. & Optional\\ \hline
-operators & A set of operators with all the metadata associated to them. & Optional\\ \hline
-backendservices & A set of back-end services with all the metadata associated to them. & Optional\\ \hline
-\end{tabular}
-\end{center}
-\end{table}
-
-The following example send two URIs of screens and a URI of a screen-flow to the GetMetadata operation:
-
-\singlespacing
-\begin{verbatim}
-[
-  "http://localhost:8080/catalogue/screens/12",
-  "http://localhost:8080/catalogue/screens/48",
-  "http://localhost:8080/catalogue/forms/9",
-]
-\end{verbatim}
-\doublespacing
-
-The response obtained are two lists, one containing all the metadata regarding to the screen-flows and another one with the information about the screens:
-
-\singlespacing
-\begin{verbatim}
-{
-  "screenflows": [],
-  "screens": [
-    {
-      "creationDate": "2009-02-07T09:59:52+0000",
-      "creator": "irivera",
-      ...,
-      "uri": "http://localhost:8080/catalogue/screens/12",
-      "version": "1.0"
-    },
-    {
-      "creationDate": "2009-02-07T09:59:52+0000",
-      "creator": "irivera",
-      ...,
-      "uri": "http://localhost:8080/catalogue/screens/12",
-      "version": "1.0"
-    }
-  ],
-  "forms": [
-    {
-      "actions": [
-        {
-          "name": "init",
-          "preconditions": [],
-          "uses": []
-        },
-        {
-          "name": "showTable",
-          "preconditions": [{
-            "id": "list",
-            "label": {"en-gb": "A product list"},
-            "pattern": "?PList
-                        http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                        http://aws.amazon.com/AWSECommerceService#ProductList",
-            "positive": true
-          }],
-          "uses": []
-        }
-      ],
-      ...,
-      "uri": "http://localhost:8080/FASTCatalogue/forms/9",
-      "version": "1.0"
-    },
-  ],
-  "operators": [],
-  "backendservices": []
-}
-\end{verbatim}
-\doublespacing
+A more detailed explanation regarding the search, recommendation, and ranking algorithms used by this operation can be found in Appendix F.
 
@@
 
-Similarly to the Screen Find \& Check described in Section~\ref{ssub:screen_findcheck}, this operation search for screen components (forms, operators and back-end services) through the catalogue in order to satisfy a given request, and it will attach information about satisfaction and reachability as well.
+Similarly to the Screen Find \& Check described in Section~\ref{ssub:screen_findcheck}, this operation search for screen components (forms, operators and back-end services) through the catalogue in order to satisfy a given request, and attaches information regarding building block satisfaction and reachability as well.
 
@@
 	\item[postconditions] Postcondition of the screen.
-	\item[canvas] The canvas is composed by a list of screens. Only the URI is needed.
-	\item[forms] It is a list of forms, such as the list of forms recommended last execution.
-	\item[operators] It is a list of operators, such as the list of operators recommended last execution.
-	\item[backendservices] It is a list of back-end services, such as the list of back-end services recommended last execution.
-	\item[domainContext] The domain context contains a list of tags and a user.
-	\item[pipes] List of pipes for the connection between screen components.
-	\item[selectedItem] URI of a screen component from the canvas.
+	\item[canvas] The canvas is composed by a list of clones of forms, operators and back-end services. Only the URI is needed.
+	\item[forms] It is a list of forms (prototypes), not in the canvas, but related to the request (may be the list of forms retrieved in previous searches).
+	\item[operators] It is a list of operators (prototypes), not in the canvas, but related to the request (may be the list of operators retrieved in previous searches).
+	\item[backendservices] It is a list of back-end services (prototypes), not in the canvas, but related to the request (may be the list of back-end services retrieved in previous searches).
+	\item[domainContext] The domain context contains a list of tags and the user who makes the request.
+	\item[pipes] List of pipes for the connection between screen components and conditions. If the building block in a pipe is the screen itself, the value must be set to \emph{null}.
+	\item[selectedItem] URI of a screen component from the canvas, or id of a pre-/post-condition.
+	\item[search] Boolean indicating if forms/operators/services are required to be looked for. If \emph{false} the operation performed is just a check. If omitted, default value is \emph{true}.
+	\item[iserve] Boolean indicating if a search has to be perfomed in the iServe service. If omitted, default value is \emph{false}. If \emph{search} is false, this parameter is discarded.
+	\item[limit] Integer. It is used to limit the number of results returned. Default value is -1 (no limit).
+	\item[offset] Integer. It is used to jump or skip a number of results to be returned. Default value is 0.
 \end{description}
 
-The screen components recommended are based on the given domain context and the pre/postconditions of the any of the resources in the canvas or the conditions of the parameters preconditions and postconditions.
-Hence, any screen component from the catalogue which has any of these conditions may be suitable for the screen, so it is included in the response, in the list corresponding to the type of screen component.
-
-An example of request is the following:
+Example request:
 
@@
     "id": "Searchcriteria_1",
-    "label": {"en-gb": "A search criteria"},
-    "pattern": "?F
+    "label": {"en-gb": "A search request"},
+    "pattern": "?search
                 http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                http://aws.amazon.com/AWSECommerceService#SearchCriteria",
+                http://fast.morfeo-project.org/ontologies/amazon#SearchRequest",
     "positive": true,
@@
   "canvas": [
-    { "uri": "http://localhost:8080/catalogue/forms/1" },
-    { "uri": "http://localhost:8080/catalogue/services/1" }
+    { "uri": "http://catalogueURL/forms/clones/112" },
+    { "uri": "http://catalogueURL/services/clones/312" }
   ],
@@
   "backendservices": [],
+  "search": true,
   "domainContext": {
     "tags": [],
-    "user": null
+    "user": "http://ismaelrivera.es#me"
   },
@@
       "from": {
-        "buildingblock": "http://localhost:8080/catalogue/services/1",
+        "buildingblock": "http://catalogueURL/services/clones/312",
         "condition": "list"
@@
         "action": "showTable",
-        "buildingblock": "http://localhost:8080/catalogue/forms/1",
+        "buildingblock": "http://catalogueURL/forms/clones/112",
         "condition": "list"
@@
   ],
-  "selectedItem": "http://localhost:8080/catalogue/forms/1"
-}
-\end{verbatim}
-\doublespacing
-
-And the response according to the above request is:
+  "selectedItem": "http://catalogueURL/forms/clones/112"
+  "search": true,
+  "iserve": true
+}
+\end{verbatim}
+\doublespacing
+
+Example response:
 
@@
           "id": "filter",
-          "label": {"en-gb": "A search criteria"},
-          "pattern": "?F 
+          "label": {"en-gb": "A search request"},
+          "pattern": "?search 
                       http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                      http://aws.amazon.com/AWSECommerceService#SearchCriteria",
+                      http://fast.morfeo-project.org/ontologies/amazon#SearchRequest",
           "positive": true,
@@
       "reachability": false,
-      "uri": "http://localhost:8080/catalogue/services/1"
+      "uri": "http://catalogueURL/services/clones/312"
     },
@@
             "label": {"en-gb": "A product list"},
-            "pattern": "?PList
+            "pattern": "?urlList
                         http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                        http://aws.amazon.com/AWSECommerceService#ProductList",
+                        http://fast.morfeo-project.org/ontologies/amazon#ProductList",
             "positive": true,
@@
       "reachability": true,
-      "uri": "http://localhost:8080/catalogue/forms/1"
+      "uri": "http://catalogueURL/forms/clones/112"
     }
@@
         "action": "search",
-        "buildingblock": "http://localhost:8080/catalogue/services/1",
+        "buildingblock": "http://catalogueURL/services/clones/312",
         "condition": "filter"
@@
   ],
-  "postconditions": []
-}
-\end{verbatim}
-\doublespacing
-
-For a screen component to be reachable, its preconditions need to be connected by a pipe to any other screen component which is already reachable, or to any of the screen preconditions.
+  "postconditions": [],
+  "iserve": [
+    {
+      "address": "http://ws.audioscrobbler.com/2.0/?method=user.getfriends
+                  &user={http://xmlns.com/foaf/0.1/Person}&api_key={APIKey}",
+      "classes": ["http://xmlns.com/foaf/0.1/Person"],
+      "doc": "http://iserve.kmi.open.ac.uk/resource/documents/[id]/hrest.html",
+      "label": "LastFm Friends",
+      "service": "http://iserve.kmi.open.ac.uk/resource/services/[id]#LastFmFriends"
+    },
+    ...
+  ]
+}
+\end{verbatim}
+\doublespacing
+
+For a screen component to be reachable, its pre-conditions need to be connected by a pipe to any other screen component which is already reachable, or to any of the screen pre-conditions.
 Moreover, the pipe needs to be satisfied, in other words, conditions in both edges are compatible.
@@
 The response includes an attribute called "connections". This attribute will be retrieved when "selectedItem" is given, and it will contain a list of potential pipes to connect to that item.
+
+A more detailed explanation regarding the search, recommendation, and ranking algorithms used by this operation can be found in Appendix F.
 
@@
 
-While creating a screen-flow, the user need to find proper screens which will satisfy her necessity.
-Once the user has selected one or several screens to be included in her screen-flow, she needs to find screens which satisfy the preconditions of these screens, in order to create an executable screen-flow.
-The operation Find, detailed in Section~\ref{ssub:screen_find}, helps the user to find these screens, however the response is just a set of screens which somehow may help to accomplish her goal, but not in a single and straight-forward step.
-
-This functionality offers to the user the possibility to receive a whole set of screens, or plan, which will make a given screen, or goal, reachable.
-These plans take into account the screen which are already in the canvas, in order to minimise the number of screens.
-
-The specific URL to access this operation is \url{http://catalogueURL/planner}, and to execute the operation a POST request needs to be done. The request body is composed of:
+Search for plans (sets of screens) which accomplish a goal. A goal is a screen or set of screens which must be made reachable.
+
+The URL to access this operation is \verb|http://catalogueURL/planner|. This operation is invoked sending an HTTP POST request with the following parameters:
 
 \begin{description}
-	\item[goal] Required. URI of the goal screen.
-	\item[canvas] Required. List of screens URIs.
+	\item[goal] Required. URI or list of URIs (clones) conforming the goal.
+	\item[canvas] Required. List of screens URIs (clones).
 	\item[page] Optional. Number of the page to be retrieved.
@@
 
-Let's see the following example:
-
-Request
-\singlespacing
-\begin{verbatim}
-{
-  "goal": "http://localhost:8080/catalogue/screens/35",
+Example request (single goal):
+
+\singlespacing
+\begin{verbatim}
+{
+  "goal": "http://catalogueURL/screens/clones/135",
   "canvas": [
-    { "uri": "http://localhost:8080/catalogue/screens/56" },
-    { "uri": "http://localhost:8080/catalogue/screens/85" }
+    { "uri": "http://catalogueURL/screens/clones/135" },
+    { "uri": "http://catalogueURL/screens/clones/656" },
+    { "uri": "http://catalogueURL/screens/clones/385" }
   ],
@@
 
-Response
+Example request (multi goal):
+
+\singlespacing
+\begin{verbatim}
+{
+  "goal": [
+    "http://catalogueURL/screens/clones/135",
+    "http://catalogueURL/screens/clones/537"
+  ]
+  "canvas": [
+    { "uri": "http://catalogueURL/screens/clones/135" },
+    { "uri": "http://catalogueURL/screens/clones/537" },
+    { "uri": "http://catalogueURL/screens/clones/656" },
+    { "uri": "http://catalogueURL/screens/clones/385" }
+  ],
+  "page": 1,
+  "per_page": 10 
+}
+\end{verbatim}
+\doublespacing
+
+Example response:
+
 \singlespacing
@@
   [
-    "http://localhost:8080/catalogue/screens/13",
-    "http://localhost:8080/catalogue/screens/23"
+    "http://catalogueURL/screens/clones/113",
+    "http://catalogueURL/screens/clones/423"
   ],
   [
-    "http://localhost:8080/catalogue/screens/79",
-    "http://localhost:8080/catalogue/screens/37",
-    "http://localhost:8080/catalogue/screens/14"
+    "http://catalogueURL/screens/clones/669",
+    "http://catalogueURL/screens/clones/37",
+    "http://catalogueURL/screens/clones/114"
   ],
   [
-    "http://localhost:8080/catalogue/screens/16",
-    "http://localhost:8080/catalogue/screens/44",
-    "http://localhost:8080/catalogue/screens/32"
+    "http://catalogueURL/screens/clones/711",
+    "http://catalogueURL/screens/clones/464",
+    "http://catalogueURL/screens/clones/322"
   ]
@@
 
-The response is ranked based on adding the minimum number of new screens into the canvas, and giving preference to the screens the user has previously selected (canvas).
-
-In the above example, let's assume the first plan is using the two screens from the canvas to accomplish the goal, and the second plan is able to make the goal reachable by itself.
-That said, a screen-flow with the goal and the second plan could be executable, and it would be composed of just four screens, however, taking the first plan, the resulting screen-flow will be composed of five screens.
-It could make sense to place the second plan in first position, but the plan with just two screens reuse the screens from the canvas which has been added by the user.
+A more detailed explanation regarding the algorithms used for the planning can be found in Appendix F.
 
@@
 
-The following table lists all the error codes.
+The possible error codes are listed in Table~\ref{table:error_codes}.
 
 \begin{table}[htb]
+\label{table:error_codes}
 \caption{API Error Codes}
@@
 InternalError & Request could not be executed due to an internal service error. & 500 Internal Server Error \\ \hline
-InvalidAction & The action <actionName> is not valid for this web service.  & 404 Bad Request \\ \hline
 InvalidHttpRequest & The HTTP request is invalid. Reason: <reason>. & 400 Bad Request \\ \hline
-InvalidLiteral & Illegal literal in the filter expression. & 400 Bad Request \\ \hline
-InvalidParameterValue & The specified parameter value is not valid. & 400 Bad Request \\ \hline
 InvalidURI & The URI <requestURI> is not valid. & 400 Bad Request \\ \hline
-MissingAction & No action was supplied with this request. & 400 Bad Request \\ \hline
 MissingParameter & The request must contain the specified missing parameter. & 400 Bad Request \\ \hline
@@
   "code": "http://url.com/.../code.js",
-  "creationDate": "2009-04-20T17:00:00+0100",
-  "creator": "http://www.fabio.es/",
+  "creationDate": "2011-04-20T17:00:00+0100",
+  "creator": "username",
   "description": {"en-gb": "This is a description of the building block"},
-  "tags": {
-    "tags": [
-      "ebay",
-      "amazon",
-    ],
-    "user": <String> // not defined yet
-  },
-  "homepage": "http://www.homepage.ie/",
-  "icon": "http://url.com/images/icon.png",
-  "id": "Id of the resource",
+    "tags":[
+      {
+        "label":{"en-gb":"Amazon"},
+        "means":"http://dbpedia.org/page/Amazon.com"
+      },
+      {
+        "label":{"en-gb":"eBay"},
+        "means":"http://dbpedia.org/page/Ebay"
+      }
+    ],
+  "homepage": "http://url.com/.../homepage",
+  "icon": "http://url.com/images/.../icon.png",
+  "id": "Id of the building block", // i.e. 3545
   "label": {"en-gb": "Label or title of the building block"},
-  "name": "Amazon Item to Ebay Filter",
   "rights": "http://creativecommons.org/",
-  "screenshot": "http://www.deri.ie/eBayList-screenshot.jpg",
-  "type": "type of the building block", // screenflow, screen, form, operator y resource
-  "uri": "http://localhost:8080/catalogue/screens/525",
+  "screenshot": "http://url.com/.../screenshot.jpg",
+  "type": "type of the building block", // screenflow, screen, form, operator or service
+  "uri": "http://url.com/screens/525",
   "version": "1.0"
@@
   "contains": [
-    "http://purl.oclc.org/fast/ontology/gadget#Screen1257860772723",
-    "http://purl.oclc.org/fast/ontology/gadget#Screen1257860720035",
+    "http://url.com/screens/clones/72723",
+    "http://url.com/screens/clones/20035",
   ],
@@
   "postconditions": [[{
-    "id": <identifier:String>,
+    "id": "Id of the building block", // i.e. 3545
     "label": {"en-gb": "Purchase URL"},
-    "pattern": "?P
+    "pattern": "?url
                 http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                http://aws.amazon.com/AWSECommerceService#PurchaseURL",
+                http://fast.morfeo-project.org/ontologies/amazon#PurchaseURL",
     "positive": true
@@
     "label": {"en-gb": "A shopping cart"},
-    "pattern": "?C
+    "pattern": "?cart
                 http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                http://aws.amazon.com/AWSECommerceService#ShoppingCart",
+                http://fast.morfeo-project.org/ontologies/amazon#ShoppingCart",
     "positive": true
@@
     "buildingblocks: [
-      { // Form
-        "id": "form1", // unique for the container screen
-        "uri": "http://purl.oclc.org/fast/ontology/gadget#Form670238"
-      },
-      { // Operators
-        "id": "op1",
-        "uri": "http://purl.oclc.org/fast/ontology/gadget#Operator213487"
-      },
-      {  // Backend services
-        "id": "bs1",
-        "uri": "http://purl.oclc.org/fast/ontology/gadget#Backendservice38399"
-      },
-      {...}
+      {
+        "uri": "http://catalogueURL/forms/clones/670238"
+        "uri": "http://catalogueURL/operators/clones/213487"
+        "uri": "http://catalogueURL/services/clones/38399"
+      },
     ]
@@
         "from": {
-          "buildingblock": "bs1",
+          "buildingblock": "http://catalogueURL/services/clones/38399",
           "condition": "cA"
@@
         "to": {
-          "buildingblock": "op1",
-          "condition": "cA", // can be the same id, unique for the building block
+          "buildingblock": "http://catalogueURL/operators/clones/213487",
           "action": "filter"
+          "condition": "cA",
         }
@@
         "from": {
-          "buildingblock": "form1",
+          "buildingblock": "http://catalogueURL/forms/clones/670238",
           "name": "refresh"
@@
         "to": {
-          "buildingblock": "bs1",
+          "buildingblock": "http://catalogueURL/services/clones/38399",
           "action": "list"
@@
       "label": {"en-gb": "Ebay List"},
-      "pattern": "?Item
+      "pattern": "?item
                   http://www.w3.org/1999/02/22-rdf-syntax-ns#type
-                  http://aws.amazon.com/AWSECommerceService#Item",
+                  http://fast.morfeo-project.org/ontologies/amazon#Item",
       "positive": true
     }]],
-    "uses": []
+    "uses": [{
+      "id":"cart",
+      "uri":"http://fast.morfeo-project.org/ontologies/amazon#ShoppingCart"
+    }]
   }],
-  "libraries": [],
+  "libraries": [{
+    "language":"JavaScript",
+    "source":"http://url.com/libcode.js"
+  }],
   "postconditions": [[{
@@
 
+\clearpage
+\section*{Appendix F. Algorithms for building block recommendation}
+\label{appendix_f}
+\addcontentsline{toc}{section}{Appendix F. Algorithms for building block recommendation}
+
+The Catalogue serves as a metadata repository for the different kind of building blocks used by the Gadgets in the FAST platform, or so-called building blocks.
+The number of building blocks stored in the catalogue is arbitrary, and a simple browsing or keyword-based search through the whole catalogue would be not usable if the number is relatively high.
+
+There are two scenarios where the user may be assisted: screen-flow design, and screen design.
+At screen-flow design, a user needs to find the proper screens to fulfil her goal.
+At screen design, the scenario is a bit more complex, where different types of building blocks are involved, making screen creation process more challenging.
+Therefore, the Catalogue includes several mechanisms in order to help in these situations.
+
+\subsection*{F.1 Pre-/post-conditions matching search}
+
+The building blocks are connected (implicit or explicitaly) to each other via their pre- and post-conditions.
+These connections are guaranteed by a matching method which determine when a pre-condition is satisfied by a post-condition.
+
+At screen-flow or screen design, the user needs to find new building blocks which may satisfy elements from the canvas.
+There is a high probability that building blocks which pre-/post-conditions are compatible might be used to accomplish her goal.
+The goal for the algorithm is to satisfy a set of pre-conditions, which are still not satisfied from the elements in the canvas.
+Therefore, this algorithm searchs the Catalogue for building blocks which post-conditions satisfy in a certain way some or all of these pre-conditions.
+
+\subsection*{F.2 Planning}
+
+One of the limitations of the pre-/post-conditions matching algorithm is that the output is limited to satisfy a given set of conditions, but the building blocks suggested may include new pre-conditions to be satisfied, where the user has to add building blocks iteratively until the goal is accomplished (screen-flow or screen is executable).
+
+A set of screens can be seen as a graph (see Figure~\ref{fig:screens_graph}), where the vertices (or nodes) are the screens, and the edges represent whether a screen satisfied or is satisfied but another screen.
+There is a special node connected to all the screens with no pre-conditions (those are reachable by default).
+For a given screen, a plan is the path from itself to the special node.
+These paths or plans are then offered to the user as a set of screens which makes a specific screen reachable.
+
+\begin{figure}[htb]
+\label{fig:screens_graph}
+\begin{center}
+	\includegraphics{images/screens_graph}
+	\caption{Graph of compatible screens by pre-/post-conditions}
+\end{center}
+\end{figure}
+
+If the number of vertices and edges is high, the number of different paths can be as well high.
+This cause problems such as a huge number of different paths between two vertices, or paths which are incredible long, in other words, the number of results or their quality may not be appropiate to be presented to the user, which may expect a few results and with the minimum number of screens as possible.
+This is a common problem, and there exist many are many algorithms which solve it, such as Dijkstra or the A* search algorithm which are complete and will always find a solution if one exists.
+
+Our approach has been to implement the A* algorithm, as it calculates a heuristic estimate of the distance to the goal for each node, apart from the path-cost function also included in Dijkstra.
+In our case, the cost for all edges is the same, and the heuristic function is used to prioritize the screens which are being used already by a given canvas.
+This way, we try to minimise the number of screens to be used.
+
+\subsection*{F.3 Association Rule Mining}
+
+Association Rule Mining is a well known method for the discovery of relations between variables in large databases.
+Given a set of transactions, find rules that will predict the occurrence of an item based on the occurrences of other items in the transaction.
+For example, the rule \verb|{onions, potatoes} = {burger}| found in the sales data of a supermarket would indicate that if a customer buys onions and potatoes together, he or she is likely to also buy burger.
+
+This method is divided in two steps: frequent itemset mining and rule generation.
+There are many algorithms for frequent itemset mining as Apriori, Eclat and FP-Growth.
+This is still a computationally expensive task, and when a mined dataset has a considerable size it is resource intensive \cite{Han:2000:MFP:335191.335372}.
+However, FP-Growth has demostrated to execute faster than the rest, which lead the open source project called Mahout\footnote{http://mahout.apache.org/} to create its parallel implementation of the algorithm \cite{Li:2008:PPF:1454008.1454027}.
+
+We are not reinventing the wheel doing our own implementation of any algorithm for frequent pattern mining.
+We opted for integrating the Catalogue with Mahout to make uses of its power, and its scalability advantages.
+
+This algorithm is used both in screen-flow design and screen-design. The approach is similar in both cases, finding building blocks which co-ocurrence is high.
+
+As an example, given the following set of screen-flows:
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|p{3.5cm}|p{10cm}|}
+\hline
+\rowcolor{fast@lightgrey}\textcolor{white}{Screen-flow} &
+                         \textcolor{white}{Screens} \\ \hline
+screenflow-A & screen-1, screen-2 \\ \hline
+screenflow-B & screen-1, screen-3, screen-4 \\ \hline
+screenflow-C & screen-1, screen-4 \\ \hline
+screenflow-D & screen-4, screen-5, screen-7, screen-8 \\ \hline
+screenflow-E & screen-3, screen-4, screen-9 \\ \hline
+screenflow-F & screen-5, screen-7, screen-2 \\ \hline
+screenflow-G & screen-10, screen-11, screen-5, screen-7 \\ \hline
+screenflow-H & screen-1, screen-2, screen-3 \\ \hline
+\end{tabular}
+\end{center}
+\end{table}
+
+The algorithm will search for the screens with the highest co-ocurrence for a given input, i.e.
+a screen-flow during its design phase.
+For example, for a screen-flow which contains the screen-5, the algorithm will recommend screen-7 as its co-ocurrence is high.
+
+However, the algorithm used for the operations Find \& Check at screen-flow and screen design, are a combination of the pre-/post-conditions matching and frequent itemset mining algorithms.
+The output items of the frequent itemset mining algorithm are ranked depending on their co-ocurrence, adding the output of the pre-/post-conditions matching search algorithm as if their co-ocurrence was one.
+
 \end{document}

### 2011-02-07 trunk/GVS/fast/templates/js_list.py
Basic support for iserve services coming from the catalogue
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1253 d8b0c59b-6544-0410-94d9-83a5658a058e

@@ -139,6 +139,8 @@ file_list = [
     "gvs/dialogs/StandaloneEmbeddingDialog.js",
     "gvs/dialogs/ButtonArrayDialog.js",
     "gvs/dialogs/ManageBuildingBlocksDialog.js",
+    "gvs/dialogs/iServeDialog.js",
+
 
     # Plans
     "gvs/plans/PlanPanel.js",

### 2011-02-07 trunk/GVS/media/js/gvs/catalogue/inferenceEngine/ScreenInferenceEngine.js

@@ -44,8 +44,11 @@ var ScreenInferenceEngine = Class.create( /** @lends ScreenInferenceEngine.proto
             body.search = false;
         } else {
             body.search = true;
-            // TODO: Control this
-            body.iserve = false;
+            if (GVS.getUser().getiServe()) {
+                body.iserve = true;
+            } else {
+                body.iserve = false;
+            }
         }
         body = Object.extend(body, elements);
         return Object.toJSON(body);

### 2011-02-07 trunk/GVS/media/js/gvs/dialogs/PreferencesDialog.js

@@ -42,9 +42,12 @@ var PreferencesDialog = Class.create(ConfirmDialog /** @lends PreferencesDialog.
                     'value': user.getEzWebURL(),
                     'regExp': FormDialog.URL_VALIDATION,
                     'message': FormDialog.INVALID_URL_MESSAGE},
-            {'type':'checkbox', 'label': 'Magic Catalogue Enabled',
+            {'type':'checkbox', 'label': 'Magic Catalogue:',
                     'name': 'catalogueMagic',
-                    'checked': user.getCatalogueMagic()}
+                    'checked': user.getCatalogueMagic()},
+            {'type':'checkbox', 'label': 'Search iServe services: <br /> (It may be slow)',
+                    'name': 'iServe',
+                    'checked': user.getiServe()}
         ];
 
         this._setContent(formData);
@@ -79,6 +82,7 @@ var PreferencesDialog = Class.create(ConfirmDialog /** @lends PreferencesDialog.
         this._getForm().email.value = user.getEmail();
         this._getForm().ezWebURL.value = user.getEzWebURL();
         this._getForm().catalogueMagic.checked = user.getCatalogueMagic() ? 'checked' : 'unchecked';
+        this._getForm().iServe.checked = user.getiServe() ? 'checked' : 'unchecked';
         $super();
     }
 });

### 2011-02-07 trunk/GVS/media/js/gvs/dialogs/iServeDialog.js

@@ -1,0 +1,89 @@
+var iServeDialog = Class.create(ConfirmDialog /** @lends iServeDialog.prototype */, {
+    /**
+     * Shows iServe services
+     * @constructs
+     * @extends ConfirmDialog
+     */
+    initialize: function($super, iServeList) {
+        $super("iServe External Services", ConfirmDialog.OK);
+
+        /**
+         * List of iServe services
+         * @private
+         * @type Array
+         */
+        this._iServeList = iServeList;
+    },
+
+    // **************** PUBLIC METHODS **************** //
+
+
+    // **************** PRIVATE METHODS **************** //
+
+
+    /**
+     * initDialogInterface
+     * This function creates the dom structure
+     * @private
+     * @override
+     */
+    _initDialogInterface: function () {
+        var content = new Element("div");
+
+        var header = new Element("h2").update("About this services");
+        content.appendChild(header);
+
+        var p1 = new Element("p").update(
+            'This services are extracted from ' +
+            '<a href="http://iserve.kmi.open.ac.uk/">iServe</a>, ' +
+            'which is a platform to discover services through their semantic ' +
+            'information.<br /> All of them are related to the building block you ' +
+            'are creating, so hopefully some of them will be useful for you.'
+        );
+        content.appendChild(p1);
+
+        var p2 = new Element("p").update(
+            'The proposed services are not GVS-compliant, so you ' +
+            'will have to wrap them. You can use the GVS to do so. '
+        );
+        content.appendChild(p2);
+
+        var table = new Element("table", {"class": "propertiesTable"});
+        table.setStyle();
+        content.appendChild(table);
+
+        var tHead = new Element("tr", {"class": "tableHeader"});
+        table.appendChild(tHead);
+
+
+        var serviceTitle = new Element("td").update("Service Name");
+        tHead.appendChild(serviceTitle);
+
+        var serviceUri = new Element("td").update("Service URL");
+        tHead.appendChild(serviceUri);
+
+        var info = new Element("td").update("Get more info");
+        tHead.appendChild(info);
+
+        this._iServeList.each(function(service) {
+            var row = new Element("tr");
+            table.appendChild(row);
+
+            var name = service.label || "-";
+            var serviceName = new Element("td").update(name);
+            row.appendChild(serviceName);
+
+            var url = new Element("td").update(service.address);
+            row.appendChild(url);
+
+            var moreinfo = new Element("td").update(
[…]

### 2011-02-07 trunk/GVS/media/js/gvs/documents/PaletteDocument.js

@@ -380,6 +380,7 @@ var PaletteDocument = Class.create(AbstractDocument, /** @lends PaletteDocument.
             this._selectedElement.getView().setSelected(true);
         } else {
             this._selectedElement = null;
+            Utils.hideMessage();
         }
 
         this._updateToolbar(this._selectedElement);

### 2011-02-07 trunk/GVS/media/js/gvs/documents/ScreenDocument.js

@@ -970,6 +970,22 @@ var ScreenDocument = Class.create(PaletteDocument,
         if (GVS.getUser().getCatalogueMagic()) {
             this._findCheckCallback(reachabilityData);
         }
+        if (reachabilityData.iserve && reachabilityData.iserve.length > 0) {
+
+            var text = new Element("span");
+            text.update("Don't find a service in the catalogue? <br /> Some " +
+                        "external services could fit your needs. ");
+            var link = new Element("a", {"href": "javascript:"}).update("Show them");
+            link.observe("click", function() {
+                this._showiServe(reachabilityData.iserve);
+            }.bind(this));
+            text.appendChild(link);
+            Utils.showMessage(text);
+        } else {
+            Utils.hideMessage();
+        }
+
+
         this._updatePanes();
     },
 
@@ -1330,6 +1346,15 @@ var ScreenDocument = Class.create(PaletteDocument,
             }
             GVS.getDocumentController().openExternalTool("Screen Debugger", uri);
         }
+    },
+
+    /**
+     * Show iserve services
+     * @private
+     */
+    _showiServe: function(iServeList) {
+        var dialog = new iServeDialog(iServeList);
+        dialog.show();
     }
 });
 

### 2011-02-07 trunk/GVS/media/js/gvs/user/User.js

@@ -45,6 +45,13 @@ var User = Class.create( /** @lends User.prototype */ {
          */
         this._catalogueMagic = false;
 
+        /**
+         * Do you want iserve integration?
+         * @private
+         * @type Boolean
+         */
+        this._iServe = false;
+
         //Bring the user from the server
         this._getUser();
     },
@@ -115,6 +122,20 @@ var User = Class.create( /** @lends User.prototype */ {
     },
 
     /**
+     * @type Boolean
+     */
+    getiServe: function() {
+        return this._iServe;
+    },
+
+    /**
+     * @param Boolean
+     */
+    setiServe: function(iServe) {
+        this._iServe = iServe;
+    },
+
+    /**
      * This function updates the user object
      * based on the data passed
      */
@@ -123,7 +144,8 @@ var User = Class.create( /** @lends User.prototype */ {
         this._firstName = userData.firstName;
         this._lastName = userData.lastName;
         this._email = userData.email;
-        this._catalogueMagic = (userData.catalogueMagic != undefined);
+        this._catalogueMagic = (userData.catalogueMagic !== undefined);
+        this._iServe = (userData.iServe !== undefined);
 
         this._ezWebURL = userData.ezWebURL;
 

### 2011-02-07 trunk/GVS/media/js/gvs/utils/Utils.js

@@ -109,6 +109,10 @@ Object.extend(Utils, {
         }
     },
 
+    hideMessage: function() {
+        $("messages").setStyle({"opacity": "0"});
+    },
+
     showErrorMessage: function(/** String */ text, /** Object(optional) */ options) {
         this.showMessage(text, Object.extend(options || {}, {'error':true}));
     },

### 2011-02-08 trunk/deliverables/D2_2_Ontology/fast_d2_2_3_ontology.tex
slowly getting closer to finishing the damn prototyping section
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1256 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 In the general case, we will therefore simply leave the definition of $G_o$ undefined, to be solved individually for each application and use case.
-A SPARQL graph pattern would be well suited for defining any particular occurrence of $G_o$.
+
+In many cases, a SPARQL graph pattern would be well suited for defining $G_o$.
+Depending on the form of the query and the optimisations available on the query processor, evaluating this graph pattern directly might not scale, so that a programmatic solution should be preferred.
+However, even in those cases defining $G_o$ with a SPARQL graph pattern is still useful for its explanatory benefits.
 For the FAST use case of defining the object graph of a prototypical building block $p$, the definition of $G_p$ is as given in List.~\ref{list:prototype_graph_definition} (here showing the definition for a screen; other types of building blocks are slightly different).
+In essence, the object graph we would like to define for a screen contains its basic annotations (label, icon, etc.), its pre- and post-conditions, its tags and its pipes and triggers.
 
@@
 \lstset{
-	caption={SPARQL query to select the object graph of a screen in FAST}, 
+	caption={SPARQL graph pattern to select the object graph of a screen in FAST}, 
 	label=list:prototype_graph_definition,
@@
 
+% The SPARQL thing is much harder than I thought. Here is an example of the general problem:
+% I have a large RDF graph with many person instances, as well as other data.
+I want to define a graph pattern in SPARQL to select all relevant triples for a person P.
+I define this patterns as follows in prose:
+% - all triples with P as the subject
+% - for each triples (P foo:address A) I also want all triples with A as the subject
+% The first question is: how do I efficiently express that as a SPARQL graph pattern?
+% The other question is: how do I build the CONSTRUCT clause to output this from a SPARQL-based filter?
+% Maybe the answer is that I shouldn't use SPARQL at all?
+
+% SELECT DISTINCT *
+% WHERE {
+% <http://data.semanticweb.org/person/knud-moeller> ?p ?pub .
+% ?pub a swrc:InProceedings .
+% ?pub ?p_pred ?p_obj .
+% }
+
 % paragraph determining_the_correct_sub_graph (end)
@@
 
-Now that we know the (application-dependant) graph $G_p$ of the prototypical object $p$, we can define what it means to create a new graph $P_c$ for its clone $c$.
+Now that we know the (application-dependant) graph $G_p$ of the prototypical object $p$, we can define what it means to create a new graph $G_c$ for its clone $c$.
 In prose, the clone graph will contain all statements of the prototypical graph, but all mentions of $URI_p$ will be replaced with $URI_c$.
@@
 
+\subsubsection{Discussion} % (fold)
+\label{ssub:discussion}
+
+This section proposed an approach for prototype-based semantics for RDF.
+This was done on the backdrop of different levels of building block representations within FAST, but could have much wider use.
+Business function modelling and other use cases have been mentioned in private discussions.
+E.g., there might be a class \texttt{BusinessFunction} with a concrete instance \texttt{ShippingBF}.
+If we want to say that there are further specialisations of \texttt{ShippingBF}, such as a \texttt{FedexBF} and a \texttt{TNTBF}, the usual class-based modelling paradigm won't help us, but prototype-based modelling will \cite{bhiri2010businessFunctionModelling}.
+
+% subsubsection discussion (end)
+
 % subsection classes_instances_and_templates (end)
@@
 In addition, a number of terms from external ontologies should also be used to annotate resources.
-Note that all of these annotations are valid both for the prototypical component definitions (templates) and the design time instances (copies).
+Note that all of these annotations are valid both for the prototypical component definitions (prototypes) and the design time instances (clones).
 A detailed description of all terms from the FAST gadget ontology namespace can be found in Sect.~\ref{sec:ontology_terms} of this document.

### 2011-02-08 trunk/deliverables/D2_2_Ontology/fast_d2_2_3_ontology.tex
prototype part done, in principle
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1257 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 For the FAST use case of defining the object graph of a prototypical building block $p$, the definition of $G_p$ is as given in List.~\ref{list:prototype_graph_definition} (here showing the definition for a screen; other types of building blocks are slightly different).
-In essence, the object graph we would like to define for a screen contains its basic annotations (label, icon, etc.), its pre- and post-conditions, its tags and its pipes and triggers.
+In essence, the object graph we would like to define for a screen contains its basic annotations (label, icon, etc.), its pre- and post-conditions and its pipes and triggers.
 
@@
 \begin{equation}
+	\label{equation:prototype_mapping}
 pm: G_p \mapsto G_c := \begin{cases}
@@
 
-For this reason, we introduce the concept of \emph{templates for RDF instances} in FAST.
-Templates are similar to the notion of prototypes in programming languages such as JavaScript.
-One resource (e.g., $LogInScreen$) will play the role of a \emph{template}, while the other resource (e.g., $LogInScreen_x$) will be called a \emph{copy}.
-Conceptually, copying a resource involves the following steps:
-\begin{inparaenum}[(i)]
-	\item Defining a named graph $G_1$ that contains all statements about this resource considered to be relevant.
-	\item Creating a second graph $G_2$ with copies of all statements in $G_1$.
-	\item During this copying process, the identifiers for the main resource and certain other resources in the graph will be changed from $I_{template}$ to $I_{copy}$.
-\end{inparaenum}	
-In practice, this procedure can be implemented using a SPARQL CONSTRUCT query, where the WHERE clause defines the graph pattern of the template, and the CONSTRUCT query will build the copy from it.
-The corresponding data on all three levels is shown in List.~\ref{list:prototyping}, using TriG syntax~\cite{bizer2004trig} (an extension of Turtle syntax for named graphs).
+However, there are cases in which this mapping will not be sufficient.
+More specifically, when the prototype's object graph contains resources which themselves need to be cloned, their URIs need to be replaced in the same fashion as shown in Def.~\ref{equation:prototype_mapping}.
+An example of this in FAST is the cloning of screens: the URIs of conditions, triggers and pipes are based on their screen, and therefore need to change along with it.
+As discussed for the definition of the object graph above, selecting which sub-resources need to be cloned is application-dependant.
+While we do not propose any particular mechanism for selecting sub-resources, also here a SPARQL graph pattern is a possibility, as shown in List.~\ref{list:sub_resources}.
+Here, the replacement algorithm would have to be performed for each \texttt{?sub\_res}.
+
+\singlespacing
+\lstset{
+	caption={SPARQL SELECT to select sub-resources of a screen in FAST}, 
+	label=list:sub_resources,
+	language=sparql
+}
+\begin{figure}
+	\lstinputlisting{exampleCode/sub_resources.sparql}
+\end{figure}
+\doublespacing
+
+Finally, coming back to the levels of representation in FAST, List.~\ref{list:prototyping} shows the class level, prototypical instance and cloned instance of a screen, using TriG syntax~\cite{bizer2004trig} (an extension of Turtle syntax for named graphs).
+Using the terminology established so far, $URI_p$ is \texttt{catalogue:LogInScreen} and $URI_c$ is \texttt{catalogue:LogInScreen\_74382}, while the corresponding object graphs $G_p$ and $G_c$ are \texttt{catalogue:prototype\_graph} and \texttt{catalogue:clone\_graph}, respectively.
+Prototype and clone refer to each other using the \texttt{fgo:has-Prototype} and \texttt{fgo:hasClone} properties.
 
@@
 
-Figure \ref{fig:prototype_example_graph} illustrates the same example graphically, using different colours to represent the two different graphs: the red graph $G_1$ in the background represents the original template resource, whereas the green graph $G_2$ in the front in represents the resource copy as it would be used in a concrete screen flow.
-\texttt{catalogue:LoginScreen} $\subset I_{template}$ has been changed to \texttt{catalogue:LoginScreen\_74382} $\subset I_{copy}$.
+Figure \ref{fig:prototype_example_graph} illustrates the same example graphically, using different colours to represent the two different object graphs: the red graph $G_p$ in the background represents the graph of the prototypical instance, whereas the green graph $G_c$ in the front in represents the new clone as it would be used in a concrete screen flow.
+% \texttt{catalogue:LoginScreen} $\subset I_{template}$ has been changed to \texttt{catalogue:LoginScreen\_74382} $\subset I_{copy}$.
 
@@
 
-It should be pointed out that the term \emph{RDF template} has been used before, but with a different meaning.
-\cite{davis2003rdf_template} use the term in analogy to XSLT, but acting on an RDF graph rather than an XML tree.
-\cite{kawamoto2006kawawiki} present a semantic wiki, where templates help non-technical users to create semantic descriptions.
-
 % paragraph deriving_the_ (end)
@@
 This section proposed an approach for prototype-based semantics for RDF.
-This was done on the backdrop of different levels of building block representations within FAST, but could have much wider use.
+This was done against the backdrop of different levels of building block representations within FAST, but could have much wider use.
 Business function modelling and other use cases have been mentioned in private discussions.
@@
 If we want to say that there are further specialisations of \texttt{ShippingBF}, such as a \texttt{FedexBF} and a \texttt{TNTBF}, the usual class-based modelling paradigm won't help us, but prototype-based modelling will \cite{bhiri2010businessFunctionModelling}.
+
+Additionally, it should  be pointed out that in practice FAST does not currently make use of the prototyping mechanism to the fullest extent.
+Certain situations \emph{require} the creation of new objects from prototypes (e.g., the use of several operators of the same kind in a single screen during the screen design phase), while other situations simply benefit from the approach because it leads to a more coherent URI minting scheme (e.g., the cloning of conditions, pipes and triggers for screens).
+However, in yet other situations no new objects are generated, since there is no compelling reason to do so (e.g., screen components such as forms, operators and service resources are not considered part of a screen's object graph, and are therefore not cloned with it).
+With this in mind, FAST should be considered a first ``testing in the field'' of our proposal for prototype-based semantics, rather than a fully-fledged implementation.
+
+Finally, in previous versions of this deliverable we used the terms \emph{template} and \emph{copy} instead of \emph{prototype} and \emph{clone}.
+The change to the new terminology was done because it is much more established and well-defined.
+Also, the term \emph{RDF template} had been used in prior work in the community, but with a different meaning.
+\cite{davis2003rdf_template} use the term in analogy to XSLT, but acting on an RDF graph rather than an XML tree.
+\cite{kawamoto2006kawawiki} present a semantic wiki, where templates help non-technical users to create semantic descriptions.
 
@@
 Post-conditions are expressed in a similar fashion.
-In the prototypical phase (i.e., as part of templates), post-condition patterns have variables in the same way as pre-condition patterns.
-When a building block is copied to the canvas, the pattern needs to be materialised.
+As part of prototypical instances, post-condition patterns have variables in the same way as pre-condition patterns.
+When a building block is cloned to the canvas, the pattern needs to be materialised.
 To do this, each variable will be replaced with a randomly generated URI or blank node.

### 2011-02-12 trunk/deliverables/D5_1_2/fast_d5_1_2.tex
changes suggested by Alex are done, and I have added new sections. not yet finished!
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1265 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-\usepackage{inputenc}
+\usepackage[utf8]{inputenc}
 \usepackage{lastpage}
@@
                          \textcolor{white}{\textbf{Change description}}\\ \hline
-1.0 & 19/02/2010 & Ciprian Palaghita (Cyntelix) & First version of the deliverable \\
+1.0 & 19/02/2010 & Ciprian Palaghita (Cyntelix) & Release for M24 \\
  & & Ismael Rivera (NUIG) & \\ \hline
-2.0 & 17/02/2011 & Ciprian Palaghita (Cyntelix) & Second version of the deliverable \\
+2.0 & 17/02/2011 & Ciprian Palaghita (Cyntelix) & Release for M36 \\
  & & Ismael Rivera (NUIG) & \\ \hline
@@
 
-changes from M24 version:
+This sections enumerates and gives a brief overview of the changes made in the deliverable with regards to the version 1.0 released for the milestone M24.
+
 \begin{itemize}
- \item 'find' and 'check' operations for screens (screen-flow design) has been removed. the same functionaly was offered by 'find\&check' doing just 1 request.
- \item API calls (both find\&check, planner, metadata) updated to reflect the changes in the prototype
- \item Adds operation for building block cloning.
- \item added appendix explaining how the recommendation/planning is done
- \item architecture have been modified reflecting the changes in the catalogue's architecture, and including new components such as the machine learning library (Mahout) and a library to query iServe.
- \item uses Sindice in the 'ontology importer' mechanism
+ \item The installation guide has been modified to include some minor changes in the configuration, and a better example for the ``Hello World!''.
+ \item Two operations have been removed: 'Find' and 'Check' for screens, used at screen-flow design, since the same functionaly was offered by the operation Find\&Check, so no functionality is duplicated or overlapped.
+However, using this unique operation does not mean that every request does a \emph{find} and a \emph{check}, these actions are optional and can be specified in the request.
+ \item There have been some substantial changes in the API methods Find\&Check and Planner and Metadata reflect the changes made in the Catalogue prototype.
+ \item There have been created new operations in the API to support the prototype-based semantics for RDF as defined in D2.2.3~\cite{moeller2011fast_ontology}.
+ \item The Catalogue Architecture has been updated to reflect the new functionality and other changes.
+ \item There have been implemented new mechanisms for managing and importing ontologies. These mechanism are explained in~\ref{ssec:logic_tier}.
+ \item A new appendix has been added in order to explain the decisions and logic behind the main methods of the API, how the building block recommendation works, algorithm supporting the building block planning, matching and inferencing in pre-/post-conditions, and the integration with iServe\footnote{http://iserve.kmi.open.ac.uk/}, enhancing the web service discovery.
+ \item There are two new resources, and its CRUD operations: concepts (classes) and samples (instances).
 \end{itemize}
@@
 The following section covers the configuration options you may have to set up before start using the Catalogue.
+The configuration is done in the file \emph{repository.properties} in the root directory where the Catalogue has been deployed.
+
+The resources (building blocks) created in the Catalogue have an unique identifier or URI.
+This URI is constructed relatively to the Catalogue URL which is configured by the parameter \emph{serverURL}.
+If you are testing the application in your machine, it may look like \verb|http://localhost:8080/FASTCatalogue|.
+
 The FAST Catalogue relies on a Sesame repository for the persistent storage.
-Sesame repositories can be local or remote.
-Remote does not mean, it has to be installed in a different machine, but in a servlet container such as Apache Tomcat.
-The advantage of remote repositories is the possibility of accessing it via a SPARQL endpoint.
-These configuration aspect can be found in the file \emph{repository.properties} in the root directory where the Catalogue has been deployed.
-
-Independently the type of repository, the parameter \emph{serverURL} needs to take the value of the URL where the FAST Catalogue instance has been deployed.
-If you are testing the application in your machine, it may be something like \verb|http://catalogueURL/FASTCatalogue|.
-
-Local repositories just need the parameter \emph{storageDir} pointing at a local directory, which has write permission in order to allow Sesame to create files in that directory or to delete any file in the directory.
-For instance:
+Sesame repositories can be \emph{local} or \emph{remote}.
+Local repositories just need the parameter \emph{storageDir} pointing at a local directory, with write permission in order to allow Sesame to create and delete the necessary binary files to store and retrieve the data.
 
@@
 
-Nevertheless, setting up a remote repository requires a few more steps.
+Setting up a remote repository requires a few more steps.
+\emph{Remote} means the repository is installed in a web server and accessed via HTTP, independently if it is in the same machine or in a different dedicated machine.
 First the OpenRDF Sesame Server needs to be installed.
@@
 Once the Sesame server is deployed, and a repository has been created, the parameters sesameServer and repositoryID will take the values of the Sesame server URL and the repository identifier.
-For instance:
-
-\begin{verbatim}
-sesameServer=http://localhost:8880/openrdf-sesame
-repositoryID=repository-ex1
-\end{verbatim}
+
+\begin{verbatim}
+sesameServer=http://sesameServerURL/openrdf-sesame
+repositoryID=repository-id
+\end{verbatim}
+
+Note: If you want to provide the data of in the repository via a SPARQL endpoint, the easiest way it is to configure the Catalogue using a remote repository, as in this case Sesame provides an HTTP accessible SPARQL endpoint by default.
 
@@
 It is recommended to check if the FAST Catalogue has been successfully installed.
-To do that, a simple HTTP request, using a Web browser, can be sent to the URL the Catalogue is deployed, and specifying a type of building block such as Screen.
-The URL would be similar to \verb|http://catalogueURL/FASTCatalogue/screens|.
-Depending on the variant chosen, the response may be different.
-For JSON, the response should be \verb|[]|, but in HTML the message \verb|No screens found.| should be shown.
+The simpliest manner is to send a HTTP GET request to retrieve concepts (classes) and attributes (properties) of the predefined ontologies to \verb|http://catalogueURL/FASTCatalogue/concepts| (e.g.
+typing the address in a Web browser).
+The response is a list of all the concepts and their associated metadata, as in the following example:
+
+\singlespacing
+\begin{verbatim}
+[
+  {
+    "attributes": [],
+    "label": {"en": "A request to search for an item"},
+    "tags": [{
+      "label": {"en": "amazon"},
+      "means": "http://www.amazon.com"
+    }],
+    "uri": "http://fast.morfeo-project.org/ontologies/amazon#SearchRequest"
+  },
+  {
+    "attributes": [
+      {
+        "type": "http://xmlns.com/foaf/0.1/Document",
+        "uri": "http://fast.morfeo-project.org/ontologies/amazon#smallImage"
+      },
+      {
+        "type": "http://www.w3.org/2001/XMLSchema#double",
+        "uri": "http://fast.morfeo-project.org/ontologies/amazon#price"
+      }
+    ],
+    "label": {"en": "An item handled by the Amazon Service"},
+    "tags": [{
+      "label": {"en": "amazon"},
+      "means": "http://www.amazon.com"
+    }],
+    "uri": "http://fast.morfeo-project.org/ontologies/amazon#Item"
+  },
+  ...
+]
+\end{verbatim}
+\doublespacing
+
+By default, the Catalogue does not include any building block, therefore any GET request for any type of building block, as for instance to \verb|http://catalogueURL/FASTCatalogue/screens| should return an empty list \verb|[]| (if JSON requested) or \verb|No screens found| (if HTML requested).
 
@@
 \begin{itemize}
- \item The top layer represents the public API for the Catalogue.
-It puts together all the available information and services to offer the maximum powerful and flexibility for other IDEs or applications.
-In FAST, here is where the Gadget Visual Storyboard
- \item The core layer implements the object model, and provides the business logic for the different processes involved in the Catalogue, integrating third party libraries where needed.
- \item The pumbling layer offers an abstraction to RDF model logic and triple stores.
+ \item \emph{Presentation tier} is the top layer which represents the public API for the Catalogue.
+It brings together all the relaying functionality in a simple but flexible manner to be used by other IDEs or applications.
+In FAST, this is the unifying interface with other components such as the Gadget Visual Storyboard or the Service Wrapper.
+ \item \emph{Logic tier} is the core layer which implements the business logic for the different processes involved in the Catalogue, and communicates with third party libraries or APIs.
+ \item \emph{Data tier} is the bottom layer. It offers storage capabilities in triple stores, relational databases and binary files, and acts as an abstraction to RDF model logic.
 \end{itemize}
@@
 \begin{center}
-	\includegraphics[width=12cm]{images/catalogue_architecture}
+	\includegraphics[width=13cm]{images/catalogue_architecture}
 	\caption{Catalogue's Architecture}
@@
 
-\subsection{Presentation layer / API}
-
-The Presentation layer will be the public interface or API of the Catalogue.
+\subsection{Presentation tier / Public API}
+
+The Presentation tier is the public interface or API of the Catalogue.
 The main purpose of this layer is to provide its functionality to the Gadget Visual Storyboard, or any other third party IDE or application.
 The API is offered in a REST (REpresentational State Transfer) style, making it easy to consume than other complex APIs.
-REST is an architectural style, not a toolkit or a standard.
-Even though, it makes use of standards like HTTP, URI, Resource Representations or MIME types.
+REST is an architectural style, not a toolkit or a standard , even though, it makes use of standards like HTTP, URI, Resource Representations or MIME types.
 Another characteristic of the REST style is its stateless assumption.
@@
 
-As part of the Presentation layer, a SPARQL endpoint has been implemented using the SPARQL protocol service as defined in the SPROT \cite{sprot} specification.
+As part of the Presentation tier, a SPARQL endpoint has been implemented using the SPARQL protocol service as defined in the SPROT \cite{sprot} specification.
 The SPARQL endpoint is mostly offered to enable other developers to query directly the Catalogue knowledge base using SPARQL queries.
@@
 
-\subsection{Core layer}
-
-The Core or Business Logic layer contains all the FAST domain-specific processing.
+\subsection{Logic tier}
+\label{ssec:logic_tier}
+
+The Logic tier or business logic layer contains all the FAST domain-specific processing.
 The resources and relations between each other, which define the domain model, are extracted from the FAST ontology~\cite{moeller2011fast_ontology}.
-These layer provides support to the functions defined in the API, and interact with the persistence layer to retrieve and store any information of the data model.
+This layer provides support to the functions defined in the API, and interact with the data tier to retrieve and store any information of the data model.
 
@@
 
-At the bottom part of the Gadget Architecture~\cite{reyes2011gadget_architecture} are the back-end services.
+At the bottom level of the Gadget Architecture~\cite{reyes2011gadget_architecture} are the back-end services.
 These are RESTful services, SOAP-based web services, or similar.
-In FAST, these web services are wrapped, extracting their definition and encapsulation their behaviour.
-To creation of these service wrappers is on demand, and the input for this process is a particular definition of a web service.
+In FAST, these web services are wrapped, extracting their definition and encapsulating their behaviour.
+The creation of these service wrappers is on demand, and the input for this process is a particular definition of a web service.
 However, the Catalogue does not crawl the Web looking for Web services.
-Instead, and to enrich the collection of available web services, the Catalogue integrates and queries iServe\footnote{http://iserve.kmi.open.ac.uk/}, a platform for publishing Semantic Web Services as linked data, no matter their original format.
+Instead, and to enrich the collection of available web services, the Catalogue integrates and queries iServe, a platform for publishing Semantic Web Services as linked data, no matter their original format.
 Therefore, thousands of web services defined or annotated in a semantic way, are also served from the Catalogue, in a uniform manner.
@@
 The Catalogue is a semantic application.
-As a result, it has to handle with ontologies.
-Apart from the core ontologies being used across the whole application (see~\cite{moeller2011fast_ontology}), the building blocks may reference classes and properties from diverse vocabularies and ontologies.
-To have a better understanding of them, the Catalogue imports them inside its repository any time it detects a building block is using an unknown class, property, or in other words, an unknown ontology.
-While importing an ontology, the Catalogue tries to get its definition from the URI of a given class or property (as stated in~\cite{berrueta2008}).
-However, sometimes this may be not succesful, and the Catalogue will try to get its definition querying Sindice\footnote{http://www.sindice.com/}, a semantic search engine, which cralws the web and keeps provenance metadata for the ontologies.
-
-\subsection{Plumbing layer}
-
-The Plumbing layer abstract the interaction with triple stores, RDF models and physical persistence to disk, database, etc.
-It provides the different mechanisms to allow the Catalogue the storage of the domain objects into a triple store in its underlying representation of RDF.
-
-There are many triple stores solutions, any of them with its own advantages and disadvantages.
+As a consequence, it must handle and work with ontologies and vocabularies.
+The business model itself is described by in the FAST ontology, and the Catalogue goes along with a few established ontologies integrated in its core (see~\cite{moeller2011fast_ontology}).
+However, the definition of the pre-/post-conditions of the building blocks are open to any ontology or vocabulary, and they may reference classes and properties from any of them.
+To have a better understanding of these classes, properties, and so on, the Catalogue needs to import into its own knowledge base the ontology in which they are described along with the relations with other classes and properties.
+In order to adquire these vocabularies, it follows the basic recipe dictated in~\cite{berrueta2008} which says that deferencing a vocabulary URI should provide a machine-processable (RDF) content back.
+Nonetheless, this is just a good practice and not every ontology published on the Internet follows it.
+Hence, going a step further, whenever an ontology cannot be retrieved by deferencing its URI, the Catalogue queries Sindice\footnote{http://www.sindice.com/}, a semantic search engine, which crawls and contains thousands of vocabularies and metadata as their provenance, which the Catalogue can take advantage of.
+
+\subsection{Data tier}
+
+The Data tier abstracts the interaction with triple stores, RDF models and physical persistence to disk, database, etc.
+It provides different mechanisms to allow the Catalogue the storage of domain objects into a triple store in its underlying representation of RDF.
+
+There are many triple stores solutions, all of them with their own advantages and disadvantages.
 To make the Catalogue independent of the solution adopted, it makes uses of RDF2Go, an abstraction over triple (and quad) stores.
@@
 
-The format supported, and in which they must be constructed, by every HTTP request is JSON\footnote{http://www.json.org/}.
-JSON is a lightweight data interchange format whose simplicity has resulted in widespread use among web developers, easy to read and write and able to using any programming language because its structures map directly to data structures used in most programming languages.
+The body of the requests to the API must be in JSON\footnote{http://www.json.org/}.
+JSON is a lightweight data interchange format which simplicity has resulted in widespread use among web developers, easy to read and write and able to process and parse using any programming language because its structures map directly to data structures used in most programming languages.
 
@@
   "canvas": [
-    { "uri": "http://catalogueURL/screens/238" },
-    { "uri": "http://catalogueURL/screens/323" }
+    { "uri": "http://catalogueURL/screens/clones/2338" },
+    { "uri": "http://catalogueURL/screens/clones/1323" }
   ],
@@
     xmlns:ctag="http://commontag.org/ns#">
-  <rdf:Description rdf:about="http://demo.fast.morfeo-project.org/FASTCatalogue/screens/1">
+  <rdf:Description rdf:about="http://catalogueURL/screens/1">
     <rdf:type rdf:resource="http://purl.oclc.org/fast/ontology/gadget#Screen"/>
@@
 In order to understand how content negotiation is implemented in the catalogue, a few concepts need to be understood.
-The URI of a particular building block must be understood as the identifier of the building block as such, as opposed to a particular representation (JSON, HTML, etc.) of it.
+The URI of a particular building block must be understood as the identifier of the building block such, as opposed to a particular representation (JSON, HTML, etc.) of it.
 In fact, each such representation has its own URI, under which it can be retrieved.
@@
 The catalogue will then inspect the request and redirect the requester to the appropriate representation (e.g., \verb|http://catalogueURL/screens/235.json|), as illustrated in Fig.~\ref{fig:content_negotiation}.
-On the Web, different representations for the same resource are called \emph{variants}, and content negotiation is the mechanism used to determine which of the representations of most appropriate for a given request.
+On the Web, different representations for the same resource are called \emph{variants}, and content negotiation is the mechanism used to determine which of the representations is most appropriate for a given request.
 
@@
 Even though the accept header is desired, the different representations can also be retrieved directly by dereferencing their own URI.
-The convention used in the catalogue for deriving a representation URI from a building block URI is to simply attach a matching suffix, such as \texttt{.rdf}, \texttt{.json} or \texttt{.html}.
+The convention used in the catalogue for deriving a representation URI from a building block URI is to simply attach a matching suffix, such as \texttt{.rdf}, \texttt{.ttl}, \texttt{.json} or \texttt{.html} for RDF/XML, Turle, JSON or HTML respectively.
 
@@
 
-\subsection{API Calls} % (fold)
+\subsection{API calls} % (fold)
 \label{sub:api_calls}
@@
 
-\subsection{Managing Building Blocks} % (fold)
+\subsection{Managing building blocks} % (fold)
 \label{sub:managing_building_blocks}
@@
 Any building block within the Catalogue is a resource in terms of REST philosophy, therefore it can be created, retrieved, modified or deleted, using a certaing URL and a specific HTTP method for every operation.
-Two concepts should be clarified: \emph{collection}, as a set of resources which access URI is \verb|http://catalogueURL/[type]| where \emph{catalogueURL} is  where the Catalogue server is installed and \verb|[type]| is the plural of the name of the building block (e.g.
-screens, operators), and \emph{member} of the collection, in other words, the building block itself, which access URI is \verb|http://catalogueURL/[type]/[id]| where \verb|[id]| has to be replaced by the identifier of a specific building block.
+Two concepts should be clarified: \emph{collection}, as a set of resources are accessed at the URL \verb|http://catalogueURL/[type]| where \emph{catalogueURL} is  where the Catalogue server is installed and \verb|[type]| is the plural of the name of the building block (e.g.
+screens, operators), and \emph{member} of the collection, in other words, the building block itself, being accessed at the URL \verb|http://catalogueURL/[type]/[id]| where \verb|[id]| has to be replaced by the identifier of a specific building block.
 The details of the operations and which HTTP verb has to be used can be found in the Table~\ref{tab:crud_operations}.
-These methods are to what is called \emph{Prototype} of building blocks (see \cite{moeller2011fast_ontology}).
-
-\begin{table}[htb]
+
+\begin{table}[htb!]
 \caption{CRUD operations}
@@
 
-The JSON structure for create any building bock can be found in the appendixes.
+The JSON structure for creating any building block can be found in the appendixes.
 All of them are composed of a common structure which can be found in Section Appendix B, and some specific information depending on the type.
 Screen-flows are defined in Appendix C, screens in Appendix D and forms, operators and back-end services share the same structure detailed in Appendix E.
-However, a few examples of request and responses are shown in order to clarify how the request are constructed and sent, and how the responses look like.
+However, a few examples of request and responses are shown in order to clarify how the requests are constructed and sent, and how the responses look like.
 
@@
   "tags": [{"label": {"en-gb": "amazon"}}],
-  "type": "screen",
   "uri": "http://catalogueURL/screens/28",
@@
     "tags": [{"label": {"en-gb": "amazon"}}],
-    "type": "screen",
     "uri": "http://catalogueURL/screens/28",
@@
 
+\subsection{Managing concepts (classes) and attributes (properties)} % (fold)
+\label{sub:managing_concepts_attributes}
+
+This method allows the creation of new classes and properties which do not belong to any existing ontology or vocabulary.
+The URI of this new class has the Catalogue URL as prefix, in order to be dereferenceable, followed by a domain, which groups these classes, and a name (http://catalogueURL/concepts/[domain]/[name]).
+
+The URL for this operation is \verb|http://catalogueURL/concepts|.
+
+The concepts are another type of resource, hence the common CRUD operations (GET to retrieve, POST to create, PUT to update and DELETE to remove) are supported for them.
+
+Example of concept:
+
+\singlespacing
+\begin{verbatim}
+{          
+  "description": {"en-gb": "Represents a school with its coordinates."}, 
+  "label": {"en-gb": "School"},
+  "tags":[{
+    "label": {"en-gb": "school"}
+   }],
+  "subclassOf": "http://www.w3.org/2003/01/geo/wgs84_pos#SpatialThing",
+  "attributes": [
+    {
+      "uri": "http://www.w3.org/2003/01/geo/wgs84_pos#lat",
+      "type": "http://www.w3.org/2001/XMLSchema#double"
+    },
+    {
+      "uri": "http://www.w3.org/2003/01/geo/wgs84_pos#long",
+      "type": "http://www.w3.org/2001/XMLSchema#double"
+    }
+  "name": "school",
+  "domain": "institutions"
+}
+\end{verbatim}
+\doublespacing
+
+\subsection{Managing samples/instances of concepts} % (fold)
+\label{sub:managing_samples}
+
+It is allowed to create instances of classes, or so-called samples of concepts.
+The samples are another type of resource, hence the common CRUD operations (GET to retrieve, POST to create, PUT to update and DELETE to remove) are supported for them.
+
+Example of sample:
+
+\singlespacing
+\begin{verbatim}
+{
+  "concept": "http://catalogueURL/concepts/institutions/school",
+  "attributes": [
+    {
+      "uri": "http://www.w3.org/2003/01/geo/wgs84_pos#lat",
+      "value": "60.025"
+    },
+    {
+      "uri": "http://www.w3.org/2003/01/geo/wgs84_pos#long",
+      "value": "34.224"
+    }
+}
+\end{verbatim}
+\doublespacing
+
 \subsubsection{Prototype cloning} % (fold)
 \label{ssub:prototype_cloning}
-A building block can be cloned.
-Making a clone of a prototype is as simple as sending a HTTP POST request to \verb|http://catalogueURL/cloning| indicating the URI of the building block to be clone, as in this example:
+
+The resources created by the requests~\ref{ssub:managing_building_blocks} are prototypes.
+To create a clone of a prototype is as simple as sending a HTTP POST request to \verb|http://catalogueURL/cloning| indicating the URI of the building block to be cloned.
+It is no the aim of this deliverable to explain the logic behind the prototypes and clones.
+For a better understanding on this topic,~\cite{moeller2011fast_ontology} contains an extend section about prototype-based semantics for RDF,
+
+Request example:
 
@@
 
-\subsubsection{GetMetadata} % (fold)
-
-Retrieves all the metadata about a list of building blocks or classes (concepts) stored in the catalogue. All building blocks are supported, no matter if it is a prototype or a clone.
-
-The URL to access this operation is \verb|http://catalogueURL/getmetadata| using HTTP POST method.
-Table~\ref{tab:getmetadata_request} shows the parameters needed for the invocation and Table~\ref{tab:getmetadata_response} details the different parameters of a response.
-
-\begin{table}[htb]
-\caption{GetMetadata Request Parameters}
-\label{tab:getmetadata_request}
+\subsubsection{Metadata} % (fold)
+\label{ssub:metadata}
+
+Retrieves all the metadata about a list of building blocks or classes (concepts) stored in the catalogue.
+This functionality is already provided by~\ref{sub:managing_building_blocks} if one specific building block, or the entire collection, is to be retrieved.
+However, this operation provides a way to get the metadata of an arbitrary number of resources, specified by their URIs, with no disctintion of their type.
+
+The URL to access this operation is \verb|http://catalogueURL/metadata|.
+The operation is invoked sending a HTTP POST request, Table~\ref{tab:metadata_request} shows the parameters needed for the invocation and Table~\ref{tab:metadata_response} details the different parameters of the response.
+
+\begin{table}[htb!]
+\caption{Metadata Request Parameters}
+\label{tab:metadata_request}
 \begin{center}
@@
 \begin{table}[htb]
-\caption{GetMetadata Response Parameters}
-\label{tab:getmetadata_response}
+\caption{Metadata Response Parameters}
+\label{tab:metadata_response}
 \begin{center}
@@
 
-Searches screens and check the compatibility/reachability of a given screen-flow.
-
-\begin{table}[htb]
+Searches for new screens and checks the satisfaction/reachability of screens in a given screen-flow.
+The response will contain all the elements pass on the request, plus information regarding the conditions satisfaction, and therefore whether the screens are reachable.
+Moreover, if screens are not reachable (their pre-conditions are not satisfied), it searchs for new screens which make these screens to be reachable.
+
+The URL for this operation is \verb|http://catalogueURL/screens/findcheck|.
+
+\begin{table}[htb!]
 \caption{Screen Find \& Check Request Parameters}
@@
                          \textcolor{white}{Type}\\ \hline
-canvas & The canvas is composed by a list of screens (clones). Only the URI is needed. & Required\\ \hline
-palette & It is a list of screens (prototypes), not in the canvas but which must be check based on the specified criterion. & Required\\ \hline
-domainContext & The domain context contains a list of tags and the user who makes the request. & Required\\ \hline
-criterion & The criterion specifies what it has to be check. The only possible supported now is 'reachability' in order to check if the pre-conditions of a screen are satisfied. & Required\\ \hline
+canvas & The canvas is subdivided in three lists:
+\begin{description}
+ \item[screens] List of the screens of the screen-flow. Only the URI is needed.
+ \item[preconditions] List of the pre-conditions of the screen-flow.
+ \item[postconditions] List of the post-conditions of the screen-flow.
+ \end{description} & Required\\ \hline
+palette & It is a list of screens (prototypes), which are not in the canvas but need to be checked based on the specified criterion. & Optional\\ \hline
+domainContext & The domain context contains a list of tags and the user who makes the request. & Optional\\ \hline
+criterion & The criterion specifies what it has to be check.
+At the moment, only 'reachability' (checks conditions satisfaction and screen reachability) is supported, but this parameter may support other values.
+& Required\\ \hline
 \end{tabular}
@@
 
-\begin{table}[htb]
+\begin{table}[htb!]
 \caption{Screen Find \& Check Response Parameters}
@@
                          \textcolor{white}{Description}\\ \hline
-canvas & List of screens in the canvas, which extra information indicating satisfaction of the criterion, for the screens themselves and their pre-conditions.\\ \hline
+canvas & List of screens in the canvas, with extra information indicating satisfaction of the criterion, for the screens themselves and their pre-conditions.\\ \hline
 palette & A list of screens obtained by the search, attaching information about the satisfaction of the criterion.\\ \hline
@@
 
+The URL for this operation is \verb|http://catalogueURL/components/findcheck|.
+
 A request is formed of the following parameters:
@@
 	\item[selectedItem] URI of a screen component from the canvas, or id of a pre-/post-condition.
-	\item[search] Boolean indicating if forms/operators/services are required to be looked for. If \emph{false} the operation performed is just a check. If omitted, default value is \emph{true}.
-	\item[iserve] Boolean indicating if a search has to be perfomed in the iServe service. If omitted, default value is \emph{false}. If \emph{search} is false, this parameter is discarded.
+	\item[actions] List of actions to be executed in the operation.
+Three actions can be specified: \emph{find}, \emph{check} and \emph{iserve}.
+The action \emph{find} indicates that it must search for forms/operators/services.
+The action \emph{check} checks whether the components/conditions/pipes are satisfied/reachable.
+\emph{iserve} is set to look for web services descriptions in the iServe platform.
 	\item[limit] Integer. It is used to limit the number of results returned. Default value is -1 (no limit).
@@
   "backendservices": [],
-  "search": true,
   "domainContext": {
@@
   ],
-  "selectedItem": "http://catalogueURL/forms/clones/112"
-  "search": true,
-  "iserve": true
+  "selectedItem": "http://catalogueURL/forms/clones/112",
+  "actions": ["find", "check", "iserve"]
 }
@@
 
-\subsection{API Error Codes} % (fold)
+\subsection{API error codes} % (fold)
 There are two types of error codes, client and server.
@@
 
-\begin{table}[htb]
+\begin{table}[htb!]
 \label{table:error_codes}
@@
 \clearpage
-\section*{Appendix A. Semantic Web Technologies Evaluation}
+\section*{Appendix A. Semantic Web technologies evaluation}
 \label{appendix_a}
@@
 Also, SWS platforms come with ready-made implementations for service composition, which could have been used for automatically combining screens to screen-flows.
-On the downside, it was expected that SWS platforms a bit too slow for some FAST requirements, as the need of real-time discovery.
+On the downside, it was concluded that SWS platforms a bit too slow for some FAST requirements, as the need of real-time discovery.
 On the other hand, building a light-weight implementation directly on RDFS promised to be faster.
@@
 \clearpage
-\section*{Appendix B. Generic Building Block JSON Structure}
+\section*{Appendix B. Generic building block JSON structure}
 \label{appendix_b}
@@
   "screenshot": "http://url.com/.../screenshot.jpg",
-  "type": "type of the building block", // screenflow, screen, form, operator or service
   "uri": "http://url.com/screens/525",
@@
 \clearpage
-\section*{Appendix D. Screen JSON Structure}
+\section*{Appendix D. Screen JSON structure}
 \label{appendix_d}
@@
 \clearpage
-\section*{Appendix E. Screen Component JSON Structure}
+\section*{Appendix E. Screen Component JSON structure}
 \label{appendix_e}
@@
 
-The Catalogue serves as a metadata repository for the different kind of building blocks used by the Gadgets in the FAST platform, or so-called building blocks.
+The Catalogue serves as a metadata repository for the different kind of resources used by the Gadgets in the FAST platform, or so-called building blocks.
 The number of building blocks stored in the catalogue is arbitrary, and a simple browsing or keyword-based search through the whole catalogue would be not usable if the number is relatively high.
@@
 
-At screen-flow or screen design, the user needs to find new building blocks which may satisfy elements from the canvas.
-There is a high probability that building blocks which pre-/post-conditions are compatible might be used to accomplish her goal.
-The goal for the algorithm is to satisfy a set of pre-conditions, which are still not satisfied from the elements in the canvas.
-Therefore, this algorithm searchs the Catalogue for building blocks which post-conditions satisfy in a certain way some or all of these pre-conditions.
+At screen-flow or screen design level, the user needs to find new building blocks which may satisfy elements from the canvas.
+There is a high probability that building blocks which have compatible pre-/post-conditions might be used to accomplish her goal.
+The goal for the algorithm is to satisfy a set of pre-conditions, which are still not satisfied, of the elements in the canvas.
+Therefore, this algorithm searches the Catalogue for building blocks which have post-conditions that satisfy in a certain way some or all of these pre-conditions.
+
+\textcolor{red}{TODO: write about inferencing/reasoning done in the Catalogue}
 
@@
 
-A set of screens can be seen as a graph (see Figure~\ref{fig:screens_graph}), where the vertices (or nodes) are the screens, and the edges represent whether a screen satisfied or is satisfied but another screen.
+A set of screens can be seen as a graph (see Figure~\ref{fig:screens_graph}), where the vertices (or nodes) are the screens, and the edges represent whether a screen satisfies or is satisfied by another screen (the arrow points towards the screen which gets satisfied).
 There is a special node connected to all the screens with no pre-conditions (those are reachable by default).
@@
 If the number of vertices and edges is high, the number of different paths can be as well high.
-This cause problems such as a huge number of different paths between two vertices, or paths which are incredible long, in other words, the number of results or their quality may not be appropiate to be presented to the user, which may expect a few results and with the minimum number of screens as possible.
-This is a common problem, and there exist many are many algorithms which solve it, such as Dijkstra or the A* search algorithm which are complete and will always find a solution if one exists.
+This causes problems such as a huge number of different paths between two vertices, or paths which are incredible long, in other words, the number of results or their quality may not be appropiate to be presented to the user, which may expect a few results and with the minimum number of screens as possible.
+This is a common problem, and there  are many algorithms which solve it, such as Dijkstra or the A* search algorithm which are complete and will always find a solution if one exists.
 
@@
 
-However, the algorithm used for the operations Find \& Check at screen-flow and screen design, are a combination of the pre-/post-conditions matching and frequent itemset mining algorithms.
+However, the algorithm used for the operations Find \& Check at screen-flow and screen design, is a combination of the pre-/post-conditions matching and frequent itemset mining algorithms.
 The output items of the frequent itemset mining algorithm are ranked depending on their co-ocurrence, adding the output of the pre-/post-conditions matching search algorithm as if their co-ocurrence was one.
 
+\subsection*{F.4 iServe platform integration}
+
+\textcolor{red}{TODO: write this section}
+
 \end{document}

### 2011-02-15 trunk/deliverables/D2_2_Ontology/fast_d2_2_3_ontology.tex
fleshing out the vocdoc section
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1268 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 After briefly addressing the topic of related work in Sect.~\ref{sec:related_work} --- which is mainly covered in a different deliverable from this work package, \cite{urmetzer2010fast_state_of_the_art} --- the structure of the following sections follows directly from our adopted ontology development methodology: we will perform a domain analysis in Sect.~\ref{sec:domain_analysis}, covering the stages of specification, conceptualisation and integration of existing vocabularies.
-This is then followed by the implementation step in Sect.~\ref{sec:ontology}, which covers the concrete ontology in its form at this moment of the project.
-After a section discussing an evaluation of the FAST gadget ontology, the deliverable will be concluded with a summary of our work in Sect.~\ref{sec:conclusions}.\ref{sub:methontology_a_methodology_for_ontology_development}
+This is then followed by the implementation step in Sect.~\ref{sec:ontology}, which covers the concrete ontology in its form at this moment of the project, as well as some basic considerations such as an early proposal for prototypical semantics for RDF.
+After a section discussing an evaluation of the FAST gadget ontology in Sect.~\ref{sec:evaluation}, the deliverable will be concluded with a summary of our work in Sect.~\ref{sec:conclusions}.\ref{sub:methontology_a_methodology_for_ontology_development}
 A number of appendices have been added to the document: App.~\ref{sec:methodology} presents the development methodology adopted by us, App.~\ref{sec:ontology_terms} provides documentation for the ontology in the form of a detailed list of terms, while App.~\ref{sec:ontology_code} contains the complete code of the ontology.
@@
 
-The following changes have been made with respect to the previous version of this deliverable, D2.2.1.~\cite{moeller2009fast_ontology}:
+The following changes have been made with respect to the previous version of this deliverable, D2.2.2.~\cite{moeller2010fast_ontology}.
+A history of previous changes is kept in App.~\ref{sec:old_changes}.
 
 \begin{itemize}
-	\item Sect.~\ref{sub:goal_and_scope} has been revised to better define the scope of this deliverable with respect to other deliverables.
-	\item The list of classes and properties has been moved to the appendix section at the back of the document (App.~\ref{sec:ontology_terms}), because it was felt that it would otherwise hinder the flow of the text.
-	\item In response to the M12 review, numerous graphical representations both of the underlying conceptual model as well as the concrete gadget ontology have been added to the document, using UML notation.
-As an example, Sect.~\ref{sub:ontology_overview} now provides a complete overview of the ontology with all classes and properties.
-	\item Also in response to the previous review, conceptual model and ontology have been extended both in breadth and depth.
-	\item As an addition to the glossary of terms in the conceptualisation phase, a section discussing and illustrating the relations between the different classes has been provided.
-	\item In order to reflect the extension of the scope of FAST towards screen creation (as opposed to screen flow creation), the conceptual model and ontology have been extended accordingly.
-	\item A mechanism for RDF templating has been devised and is discussed in Sect.~\ref{sub:classes_instances_and_templates}.
-	\item As the result of the ongoing dialogue between ontology developers and application implementers, many classes and properties have been added, changed or removed.
-	\item The section on our adopted development methodology (previously Sect.~1.3) has been moved to the end of the document (App.~\ref{sub:methontology_a_methodology_for_ontology_development}), but otherwise did not change significantly.
-	\item An evaluation section has been added (Sect.~\ref{sec:evaluation}), based on research done by the authors within the past year.
-	\item The section on design principles (previously Sect.~1.4, now Sect.~\ref{sub:general_design_decisions}) has been moved to the general evaluation section.
-	\item In the integration section (Sect.~\ref{sub:integration}) of the ontology definition, the Common Tag vocabulary for complex tagging has been added.
-It is briefly introduced, and the integrated terms are specified.
-	\item On a technical note, the namespace abbreviation for the FAST gadget ontology changed from \texttt{fast} to \texttt{fgo}, in order to distinguish from other FAST ontologies which may be added in the future.
+	\item Section~\ref{sub:classes_instances_and_templates} has been significantly extended with a more in-depth discussion of prototype theory, and a first proposal for prototype-based semantics for RDF, as they are employed in FAST.
+	\item A section for the \emph{VocDoc} tool, which was developed in part for generating online and PDF documentation for the FAST ontology, has been added as App.~\ref{sec:vocdoc}.
+	\item The ontology itself has remained more or less stable since D2.2.2.~\cite{moeller2010fast_ontology}.
+However, a handful of small changes and corrections have been performed, e.g., to reflect the terminology changes in the prototype section (\emph{prototype} instead of \emph{template}, \emph{clone} instead of \emph{copy}).
+	\item The document as a whole has been revised, and minor changes (mostly, spelling, grammar and small updates to reflect latest developments) have been throughout the deliverable.
 \end{itemize}
@@
 
+\doublespacing
 \clearpage
@@
 
-\todo{include documentation for VocDoc}
-
+\emph{VocDoc} is a tool for the auto-generation of documentation for RDF vocabularies and ontologies.
+VocDoc was originally developed for use within the FAST project, but has since been used in other contexts as well, such as for generating the online documentation of the Semantic Web Conference Ontology\footnote{\url{http://data.semanticweb.org/ns/swc/ontology}}.
+As input, VocDoc parses the original RDF source files of the ontology, and then produces documentation in different output formats, such as HTML for online purposes, and La\TeX{} for print purposes.
+
+\subsection{Configuring VocDoc Processing} % (fold)
+\label{sub:configuration}
+
+In principle, VocDoc will parse any input RDF and produce output accordingly, scanning the graph for T-Box statements from the \texttt{rdf}\footnote{\url{http://www.w3.org/1999/02/22-rdf-syntax-ns\#}},
+\texttt{rdfs}\footnote{\url{http://www.w3.org/2000/01/rdf-schema\#}} and 
+\texttt{owl}\footnote{\url{http://www.w3.org/2002/07/owl\#}} 
+namespaces such as class and property definitions, class hierarchies or domain and range specifications.
+However, the output of VocDoc can be further manipulated and improved by using specific annotation statements in the input graph, providing a configuration file and using command line parameters.
+
+\subsubsection{Annotation Properties} % (fold)
+\label{ssub:annotation_properties}
+
+VocDoc looks for a number of special annotation properties to structure and enhance its output.
+All of these properties must be used on the main \texttt{owl:Ontology} instance in the input graph.
+Examples for these annotations can be found in App.~\ref{sec:ontology_code}.
+
+\paragraph{Title and Logo} % (fold)
+\label{par:title_and_logo}
+
+\texttt{rdfs:label} is used to specify the name of the vocabulary, which will appear as the ontology title in the output document.
+\texttt{foaf:logo} is used to specify the URL of a logo image, which will also be added to the output.
+
+% paragraph title_and_logo (end)
+
+\paragraph{Dates} % (fold)
+\label{par:dates}
+
+The Dublin Core vocabulary is used to specify relevant dates in the ontology development process.
+Specifically, \texttt{dcterms:created} indicates the date when the vocabulary was first created, while \texttt{dcterms:modified} indicates the date when the vocabulary was last changed (i.e., this would always be the date when the current version of the ontology was created).
+
+% paragraph dates (end)
+
+\paragraph{Authors} % (fold)
+\label{par:authors}
+
+\texttt{dcterms:creator}  is used to specify the main authors of the vocabulary or ontology.
+There can be several creators, but each one must be an instance of \texttt{foaf:Person}, complete with a \texttt{foaf:name}, \texttt{foaf:surname} and \texttt{foaf:homepage}.
+Optionally, it is also possible to instantiate a \texttt{foaf:Organization} with name and homepage and say that the person is \texttt{foaf:member} of it.
+Additional  contributors to the vocabulary can be added can be added using \texttt{dcterms:contributor}.
+Details for contributors are defined in the same way as authors.
+However, it is also possible to have any kind of \texttt{foaf:Agent} as the value.
+This way, it is possible to have a group as a contributor, e.g., ``members of the FAST project''.
+
+% paragraph authors (end)
+
+\paragraph{Licensing Information} % (fold)
+\label{par:licensing_information}
+
+Human-readable licensing information can be added using the property \texttt{dc:license} to point to a resource which contains human-readable licensing information.
+The license resource itself should be modelled with an \texttt{rdfs:label} (e.g.
+``License Statement'') and an \texttt{rdfs:comment} containing the actual license text as it should appear in the output document.
+\texttt{rdfs:seeAlso} can be used to point to a detailed version of the license (e.g., \texttt{http://creative-commons.org/licenses/by-nc/3.0/}).
+
+% paragraph licensing_information (end)
+
+\paragraph{General Documentation} % (fold)
+\label{par:general_documentation}
+
+An arbitrary number of sections containing further human-readable documentation can be added.
+Each section is defined as an individual RDF resource (of arbitrary type), using \texttt{rdfs:label} (for the section's title) and \texttt{rdfs:comment} (for the section's content).
+These sections are linked to the ontology in the desired order using the \texttt{vs:userDocs} property.
+The object of the \texttt{userDocs} statement must be a list construct with the sections in the desired order.
+
+% paragraph general_documentation (end)
+
+% subsubsection annotation_properties (end)
+
+% subsection configuration (end)
+
+\begin{itemize}
+	\item document configuration file
+	\item will take any RDF, but the use of specific annotations will improve the usefulness of the output 
+	\item document RDF annotation vocabulary 
+	\item describe HTML output (example screenshots)
+	\item describe latex output (part of this deliverable was produced with VocDoc)
+\end{itemize}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\linewidth]{images/vocdoc_preamble}
+    \caption{Screenshot showing the HTML output of VocDoc}
+    \label{fig:vocdoc_preamble}
+  \end{center}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\linewidth]{images/vocdoc_overview}
+    \caption{Screenshot showing the HTML output of VocDoc}
+    \label{fig:vocdoc_overview}
+  \end{center}
+\end{figure}
+
+\clearpage
+\phantomsection
+\section{Past Changes} % (fold)
+\label{sec:old_changes}
+
+For keeping a record of past changes, this section simply lists the changes that were made between the previous two versions of this deliverable, D2.2.1~\cite{moeller2009fast_ontology} and D2.2.2~\cite{moeller2010fast_ontology}.
+
+\begin{itemize}
+	\item Sect.~\ref{sub:goal_and_scope} has been revised to better define the scope of this deliverable with respect to other deliverables.
+	\item The list of classes and properties has been moved to the appendix section at the back of the document (App.~\ref{sec:ontology_terms}), because it was felt that it would otherwise hinder the flow of the text.
+	\item In response to the M12 review, numerous graphical representations both of the underlying conceptual model as well as the concrete gadget ontology have been added to the document, using UML notation.
+As an example, Sect.~\ref{sub:ontology_overview} now provides a complete overview of the ontology with all classes and properties.
+	\item Also in response to the previous review, conceptual model and ontology have been extended both in breadth and depth.
+	\item As an addition to the glossary of terms in the conceptualisation phase, a section discussing and illustrating the relations between the different classes has been provided.
+	\item In order to reflect the extension of the scope of FAST towards screen creation (as opposed to screen flow creation), the conceptual model and ontology have been extended accordingly.
+	\item A mechanism for RDF templating has been devised and is discussed in Sect.~\ref{sub:classes_instances_and_templates}.
+	\item As the result of the ongoing dialogue between ontology developers and application implementers, many classes and properties have been added, changed or removed.
+	\item The section on our adopted development methodology (previously Sect.~1.3) has been moved to the end of the document (App.~\ref{sub:methontology_a_methodology_for_ontology_development}), but otherwise did not change significantly.
+	\item An evaluation section has been added (Sect.~\ref{sec:evaluation}), based on research done by the authors within the past year.
+	\item The section on design principles (previously Sect.~1.4, now Sect.~\ref{sub:general_design_decisions}) has been moved to the general evaluation section.
+	\item In the integration section (Sect.~\ref{sub:integration}) of the ontology definition, the Common Tag vocabulary for complex tagging has been added.
+It is briefly introduced, and the integrated terms are specified.
+	\item On a technical note, the namespace abbreviation for the FAST gadget ontology changed from \texttt{fast} to \texttt{fgo}, in order to distinguish from other FAST ontologies which may be added in the future.
+\end{itemize}
+
+% section changes_from_previous_version (end)
+
+\singlespacing
 \clearpage

### 2011-02-15 trunk/deliverables/D5_1_2/fast_d5_1_2.tex
- architecture picture updated - includes iServe appendix - includes explanation about reasoning/inferencing - some other minor corrections
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1269 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 However, using this unique operation does not mean that every request does a \emph{find} and a \emph{check}, these actions are optional and can be specified in the request.
- \item There have been some substantial changes in the API methods Find\&Check and Planner and Metadata reflect the changes made in the Catalogue prototype.
+ \item There have been some substantial changes in the API methods, Find\&Check, Planner and Metadata reflect these changes made in the Catalogue prototype.
  \item There have been created new operations in the API to support the prototype-based semantics for RDF as defined in D2.2.3~\cite{moeller2011fast_ontology}.
@@
  \item A new appendix has been added in order to explain the decisions and logic behind the main methods of the API, how the building block recommendation works, algorithm supporting the building block planning, matching and inferencing in pre-/post-conditions, and the integration with iServe\footnote{http://iserve.kmi.open.ac.uk/}, enhancing the web service discovery.
- \item There are two new resources, and its CRUD operations: concepts (classes) and samples (instances).
+ \item There are two new resources, and their CRUD operations: concepts (classes) and samples (instances).
 \end{itemize}
@@
 
-Once you have got the source code, first step should be addressed is its compilation.
+Once you have got the source code, first step that should be addressed is its compilation.
 You may manually compile the source yourself, however the FAST Catalogue comes along with a script to facilitate this task.
@@
 This URI is constructed relatively to the Catalogue URL which is configured by the parameter \emph{serverURL}.
-If you are testing the application in your machine, it may look like \verb|http://localhost:8080/FASTCatalogue|.
+If you are testing the application on your machine, it may look like \verb|http://localhost:8080/FASTCatalogue|.
 
@@
 Setting up a remote repository requires a few more steps.
-\emph{Remote} means the repository is installed in a web server and accessed via HTTP, independently if it is in the same machine or in a different dedicated machine.
+\emph{Remote} means the repository is installed in a web server and accessed via HTTP, independently if it is on the same machine or on a different dedicated machine.
 First the OpenRDF Sesame Server needs to be installed.
@@
 
-Note: If you want to provide the data of in the repository via a SPARQL endpoint, the easiest way it is to configure the Catalogue using a remote repository, as in this case Sesame provides an HTTP accessible SPARQL endpoint by default.
+Note: If you want to provide the data in the repository via a SPARQL endpoint, for flexible data access to third-party applications, the easiest way it is to configure the Catalogue using a remote repository, as in this case Sesame provides an HTTP accessible SPARQL endpoint by default.
 
@@
 
-This sections presents a technical overview of the Catalogue's architecture (see Figure~\ref{fig:catalogue_architecture}), and how the interaction with external applications or APIs is done.
+This sections presents a technical overview of the Catalogue's architecture (see Figure~\ref{fig:catalogue_architecture}), and how the interaction with external applications or API is done.
 The Catalogue platform has a modular design, which are spread in basically three layers, depending on the nature of each component:
@@
  \item \emph{Presentation tier} is the top layer which represents the public API for the Catalogue.
-It brings together all the relaying functionality in a simple but flexible manner to be used by other IDEs or applications.
+It brings together all the underlying functionality in a simple but flexible manner to be used by other IDEs or applications.
 In FAST, this is the unifying interface with other components such as the Gadget Visual Storyboard or the Service Wrapper.
@@
 The API is offered in a REST (REpresentational State Transfer) style, making it easy to consume than other complex APIs.
-REST is an architectural style, not a toolkit or a standard , even though, it makes use of standards like HTTP, URI, Resource Representations or MIME types.
+REST is an architectural style, not a toolkit or a standard, even though, it makes use of standards like HTTP, URI, Resource Representations or MIME types.
 Another characteristic of the REST style is its stateless assumption.
@@
 As a consequence, it must handle and work with ontologies and vocabularies.
-The business model itself is described by in the FAST ontology, and the Catalogue goes along with a few established ontologies integrated in its core (see~\cite{moeller2011fast_ontology}).
+The business model itself is described in the FAST ontology, which is included in the Catalogue along with a few established ontologies integrated in its core (see~\cite{moeller2011fast_ontology}).
 However, the definition of the pre-/post-conditions of the building blocks are open to any ontology or vocabulary, and they may reference classes and properties from any of them.
 To have a better understanding of these classes, properties, and so on, the Catalogue needs to import into its own knowledge base the ontology in which they are described along with the relations with other classes and properties.
-In order to adquire these vocabularies, it follows the basic recipe dictated in~\cite{berrueta2008} which says that deferencing a vocabulary URI should provide a machine-processable (RDF) content back.
+In order to acquire these vocabularies, it follows the basic recipe dictated in~\cite{berrueta2008} which says that deferencing a vocabulary URI should provide a machine-processable (RDF) content back.
 Nonetheless, this is just a good practice and not every ontology published on the Internet follows it.
@@
 The body of the requests to the API must be in JSON\footnote{http://www.json.org/}.
-JSON is a lightweight data interchange format which simplicity has resulted in widespread use among web developers, easy to read and write and able to process and parse using any programming language because its structures map directly to data structures used in most programming languages.
+JSON is a lightweight data interchange format which simplicity has resulted in widespread use among web developers, being easy to read, write, be processed and parsed using any programming language because its structures map directly to data structures used in most programming languages.
 
@@
 
-The JSON structure for creating any building block can be found in the appendixes.
+The JSON structure for creating any building block can be found in the appendices.
 All of them are composed of a common structure which can be found in Section Appendix B, and some specific information depending on the type.
@@
 
-The resources created by the requests~\ref{ssub:managing_building_blocks} are prototypes.
+The resources created by the requests described in Section~\ref{sub:managing_building_blocks} are prototypes.
 To create a clone of a prototype is as simple as sending a HTTP POST request to \verb|http://catalogueURL/cloning| indicating the URI of the building block to be cloned.
-It is no the aim of this deliverable to explain the logic behind the prototypes and clones.
-For a better understanding on this topic,~\cite{moeller2011fast_ontology} contains an extend section about prototype-based semantics for RDF,
+It is not the aim of this deliverable to explain the logic behind the prototypes and clones.
+For a better understanding on this topic,~\cite{moeller2011fast_ontology} contains an extend section about prototype-based semantics for RDF.
 
@@
 
-As clones are building blocks as well, the operations explained in Section~\ref{sub:managing_building_blocks} are valid to retrieve and delete them.
+As clones are building blocks as well, the operations explained in Section~\ref{sub:managing_building_blocks} are valid to retrieve and delete them, being \verb|http://catalogueURL/[type]/clones/[id]| the URI to send the HTTP request.
 
@@
 Searches for new screens and checks the satisfaction/reachability of screens in a given screen-flow.
-The response will contain all the elements pass on the request, plus information regarding the conditions satisfaction, and therefore whether the screens are reachable.
-Moreover, if screens are not reachable (their pre-conditions are not satisfied), it searchs for new screens which make these screens to be reachable.
+The response will contain all the elements pass on passed in the request, plus information regarding the conditions' satisfaction, and therefore, whether the screens are reachable.
+Moreover, if screens are not reachable (their pre-conditions are not satisfied), it searchs for new screens which would make these screens reachable.
 
@@
                          \textcolor{white}{Type}\\ \hline
-canvas & The canvas is subdivided in three lists:
+canvas & The canvas is divided in three lists:
 \begin{description}
@@
 domainContext & The domain context contains a list of tags and the user who makes the request. & Optional\\ \hline
-criterion & The criterion specifies what it has to be check.
+criterion & The criterion specifies what has to be checked.
 At the moment, only 'reachability' (checks conditions satisfaction and screen reachability) is supported, but this parameter may support other values.
@@
 
-Similarly to the Screen Find \& Check described in Section~\ref{ssub:screen_findcheck}, this operation search for screen components (forms, operators and back-end services) through the catalogue in order to satisfy a given request, and attaches information regarding building block satisfaction and reachability as well.
+Similarly to the Screen Find \& Check described in Section~\ref{ssub:screen_findcheck}, this operation searches for screen components (forms, operators and back-end services) through the catalogue in order to satisfy a given request, and attaches information regarding building block satisfaction and reachability as well.
 
@@
 Also, SWS platforms come with ready-made implementations for service composition, which could have been used for automatically combining screens to screen-flows.
-On the downside, it was concluded that SWS platforms a bit too slow for some FAST requirements, as the need of real-time discovery.
+On the downside, it was concluded that SWS platforms are a bit too slow for some FAST requirements, as the need of real-time discovery.
 On the other hand, building a light-weight implementation directly on RDFS promised to be faster.
@@
 
-The building blocks are connected (implicit or explicitaly) to each other via their pre- and post-conditions.
+The building blocks are connected (implicitly or explicitaly) to each other via their pre- and post-conditions.
 These connections are guaranteed by a matching method which determine when a pre-condition is satisfied by a post-condition.
@@
 The goal for the algorithm is to satisfy a set of pre-conditions, which are still not satisfied, of the elements in the canvas.
-Therefore, this algorithm searches the Catalogue for building blocks which have post-conditions that satisfy in a certain way some or all of these pre-conditions.
-
-\textcolor{red}{TODO: write about inferencing/reasoning done in the Catalogue}
+Therefore, this algorithm searches the Catalogue for building blocks which have post-conditions that satisfy, in a certain way, some or all of these pre-conditions.
+
+The matching algorithm used to determine the compatibility of two conditions is a SPARQL query which compares whether two condition patterns are compatible.
+How the pre- and post-conditions are defined is explained in~\cite{moeller2011fast_ontology}, but as an example:
+
+\singlespacing
+\begin{verbatim}
+?user a sioc:User ;
+      foaf:accountName ?account_name .
+\end{verbatim}
+\doublespacing
+
+It describes a hypothetical condition requiring the existence of a \verb|sioc:User| which must have a certain \verb|foaf:accountName|.
+
+Defining the condition patterns in this manner allows the usage of inference or reasoning mechanisms, which would relax the restrictions to connect building blocks between each other.
+The Catalogue supports RDFS Entailment Rules, as defined in~\cite{patrickHayes2004rdfSemantics}.
+
+As an example, enabling the inferencing capabilities would allow building blocks pre-conditions to be satisfied by a more specialised condition, in terms of classes and subclasses.
+In the following scenario:
+
+\singlespacing
+\begin{verbatim}
+sioc:User a rdfs:Class .
+ex:Professor rdfs:subClassOf sioc:User .
+\end{verbatim}
+\doublespacing
+
+A screen A which pre-condition pattern is \verb|?user a sioc:User| would be satisfied by a screen B which post-condition pattern is \verb|?professor a ex:Professor|, as the pre-condition requires the existence of a more generic class than the one defined in the post-condition.
 
@@
 
-\textcolor{red}{TODO: write this section}
+iServe is a platform for publishing Semantic Web Services as linked data~\cite{oro23093}.
+In a nutshell, it crawls the Web for web services descriptions and imports them into the platform, using the Minimal Service Model (MSM), which provides a common vocabulary for service annotations of different formalisms such as OWL-S, SAWSDL and WSMO.
+To facilitate the consumption and manipulation of the data, iServe provides three interfaces: a Web-based application (iServe Browser), a RESTful API and a SPARQL endpoint.
+
+In FAST, the definition of the building blocks, and therefore of the web services have many similarities to the MSM.
+Without going into much detail, both have operations (or actions) which have inputs and outputs.
+These inputs and outputs are linked to classes in an ontology.
+In FAST, an input/output is defined by a set of condition or triple patterns, which says the class of the condition and adds some restrictions to it (values for certain properties, etc.).
+iServe references the classes or types of the inputs and outputs using the property \emph{modelReference} from the SAWSDL vocabulary.
+
+The Catalogue queries the iServe platform in the Find \& Check for screen components operation.
+The goal of this operation is to find forms, operators or web services wrappers which may fit into a given screen (canvas).
+The integration is done sending SPARQL queries to its SPARQL endpoint\footnote{http://iserve.kmi.open.ac.uk/data/execute-query}, looking for web services which operations, inputs, or outputs reference (sawsdl:modelReference) the classes used by a set of conditions (e.g.
+unsatisfied conditions in the canvas).
 

### 2011-02-17 trunk/deliverables/D2_2_Ontology/fast_d2_2_3_ontology.tex
ready for review
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1272 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \usepackage{courier}
+\usepackage{subfig}
 
@@
 	\item The document as a whole has been revised, and minor changes (mostly, spelling, grammar and small updates to reflect latest developments) have been throughout the deliverable.
+	\item The online documentation of the FAST Gadget ontology at \url{http://purl.oclc.org/fast/ontology/gadget} has been updated to reflect the changes in this version of the deliverable.
 \end{itemize}
@@
     \includegraphics[width=\linewidth]{images/fgo_uml_diagram}
-    \caption{FAST Gadget Class Hierarchy}
+    \caption{FAST Gadget Ontology Class Hierarchy}
     \label{fig:ontology_uml}
@@
 With this in mind, FAST should be considered a first ``testing in the field'' of our proposal for prototype-based semantics, rather than a fully-fledged implementation.
+
+Furthermore, it should be mentioned that our proposal makes no assumption on how the cloning process is actually implemented, i.e., whether any cloned triples are actually materialised into the database, or if the cloning process should be handled as inference rules on top of the RDF layer.
+In the FAST catalogue, we are currently following the former approach.
 
@@
 }
-\lstinputlisting{clean.ttl}
+\lstinputlisting{fgo2011-02-11.ttl}
 
@@
 VocDoc was originally developed for use within the FAST project, but has since been used in other contexts as well, such as for generating the online documentation of the Semantic Web Conference Ontology\footnote{\url{http://data.semanticweb.org/ns/swc/ontology}}.
-As input, VocDoc parses the original RDF source files of the ontology, and then produces documentation in different output formats, such as HTML for online purposes, and La\TeX{} for print purposes.
+As input, VocDoc parses the original RDF source files of the ontology, and then produces documentation in different output formats, such as HTML for online purposes, DOT\footnote{\url{http://www.graphviz.org/doc/info/lang.html}} for a visual overview and La\TeX{} for print purposes.
+
+VocDoc is available at \url{http://kantenwerk.org/vocdoc}.
+
+\subsection{VocDoc Output} % (fold)
+\label{sub:vocdoc_output}
+
+At the moment, VocDoc will either produce La\TeX{}, DOT or HTML output, depending on the parameters specified when running the tool (see Sect.~\ref{ssub:command_line_parameters}).
+
+The La\TeX{} output is currently intended for inclusion in other documents.
+It consists of a list ontology terms, divided into classes and properties, with a detailed description for each.
+These descriptions include relations to other ontology terms (hierarchical or other), as well as various other attributes available through OWL and RDFS semantics.
+Related terms are linked to each other via hyperlinks.
+App.~\ref{sec:ontology_terms} of this deliverable was created entirely via VocDoc and serves as an example of this kind of output.
+
+The DOT output of VocDoc will produce a UML class hierarchy of the input ontology's classes in the DOT language. The ontology overview in Fig.~\ref{fig:ontology_uml} was produced in this fashion.
+
+The HTML option of VocDoc produces a stand-alone HTML document for publishing on the Web.
+It includes the same interlinked term list as the La\TeX{} output.
+In addition, and if the appropriate annotations are present in the input graph (see Sect.~\ref{sub:configuration}), the output document will also include a preamble with metadata such as authors and change date, licensing information, an overview with terms group by topic, and extended additional documentation.
+
+\begin{figure}
+  \begin{center}
+	\subfloat[Preamble of the HTML output]{
+	    \includegraphics[width=0.485\linewidth]{images/vocdoc_preamble}
+	    \label{fig:vocdoc_preamble}
+	}
+	\subfloat[Overview section of the HTML output]{
+	    \includegraphics[width=0.485\linewidth]{images/vocdoc_overview}
+	    \label{fig:vocdoc_overview}
+	}
+	\caption{Screenshots showing the HTML output of VocDoc}
+	\label{fig:vocdoc_html}
+  \end{center}
+\end{figure}
+
+% \begin{figure}
+%   \begin{center}
+%     \includegraphics[width=\linewidth]{images/vocdoc_overview}
+%     \caption{Screenshot showing the HTML output of VocDoc}
+%     \label{fig:vocdoc_overview}
+%   \end{center}
+% \end{figure}
+
+% subsection vocdoc_output (end)
 
@@
 
-In principle, VocDoc will parse any input RDF and produce output accordingly, scanning the graph for T-Box statements from the \texttt{rdf}\footnote{\url{http://www.w3.org/1999/02/22-rdf-syntax-ns\#}},
+In principle, VocDoc will parse any input RDF and produce output accordingly, scanning the graph for T-Box terms from the \texttt{rdf}\footnote{\url{http://www.w3.org/1999/02/22-rdf-syntax-ns\#}}, 
 \texttt{rdfs}\footnote{\url{http://www.w3.org/2000/01/rdf-schema\#}} and 
@@
 namespaces such as class and property definitions, class hierarchies or domain and range specifications.
-However, the output of VocDoc can be further manipulated and improved by using specific annotation statements in the input graph, providing a configuration file and using command line parameters.
-
-\subsubsection{Annotation Properties} % (fold)
+However, the output of VocDoc can be further manipulated and improved by using specific annotation statements in the input graph, providing a configuration file and using command line parameters, as explained in the sections below.
+
+\subsubsection{T-Box Definitions and Annotations} % (fold)
+\label{ssub:recognised_t_box_terms}
+
+VocDoc will scan the input graph for instances of the following classes from the \texttt{owl} namespace:
+
+\begin{itemize}
+	\item \textbf{Classes}: \texttt{owl:Class} and \texttt{owl:DeprecatedClass}
+	\item \textbf{Properties}: \texttt{owl:ObjectProperty}, \texttt{owl:DatatypeProperty}, \texttt{owl:FunctionalProp-erty}, \texttt{owl:InverseFunctionalProperty}, \texttt{owl:TransitiveProperty}, \texttt{owl:Symme-tricProperty} and \texttt{owl:DeprecatedProperty}
+	\item \textbf{Individuals}: instances of any of the classes defined in this ontology
+\end{itemize}
+
+For each class, the tool will furthermore find the class hierarchy (via \texttt{rdfs:subClassOf}), as well as class definitions based on OWL constructs such as \texttt{owl:unionOf}, \texttt{owl:intersectionOf} and \texttt{owl:oneOf}.
+Additionally, it is checked if the class is in the \texttt{rdfs:domain} or \texttt{rdfs:range} of any property defined in the same ontology.
+
+Each property is checked for its domain and range \texttt{rdfs:range} and \texttt{rdfs:domain}, placed in a term hierarchy using \texttt{rdfs:subPropertyOf} and \texttt{rdfs:superPropertyOf} and checked for \texttt{owl:inverseOf} statements.
+
+All terms in the ontology (classes and properties) can also have further annotations, such as \texttt{rdfs:label}, \texttt{rdfs:comment} (the documentation for an individual term), \texttt{dc:subject} (for grouping ontology terms in different thematic categories) and \texttt{vs:term\_status}, which can have values such as ``testing'' or ``stable''.
+
+% subsubsection recognised_t_box_terms (end)
+
+\subsubsection{Ontology Annotations} % (fold)
 \label{ssub:annotation_properties}
@@
 
+\subsubsection{Configuration File} % (fold)
+\label{ssub:configuration_file}
+
+VocDoc requires the presence of a configuration file, which is used to provide the tool with additional data required for processing the input ontology.
+The configuration file consists of the following simple key-value pairs (\texttt{key=value}):
+
+\begin{itemize}
+	\item \texttt{ontology\_namespace}: The namespace URI in which the ontology resides. This is used to disambiguate between different instances of \texttt{owl:Ontology}. This parameter is mandatory.
+	\item \texttt{doc\_root}: If HTML output is produced, this URI indicates where the ontology documentation and source will eventually be located online. This parameter is mandatory.
+	\item \texttt{alphabetise}: A boolean which sets if term lists should be outputted in alphabetical order.
+	\item \texttt{GA\_code}: If HTML output is produced, VocDoc can append special code needed to make Google Analytics\footnote{\url{http://www.google.com/analytics/}} work with the documentation.
+The code is the application code provided by Google for the user.
+\end{itemize}
+
+% subsubsection configuration_file (end)
+
 % subsection configuration (end)
 
+\subsection{Running VocDoc} % (fold)
+\label{sub:running_vocdoc}
+
+VocDoc is implemented in the Ruby scripting language as a command line tool, using the Redland RDF API\footnote{\url{http://librdf.org/}} for parsing and querying the input RDF graph.
+The general syntax for running as shown below, where \texttt{source\_file} indicates the input RDF file (in Turtle syntax) and \texttt{target\_folder} indicates the folder in which to write the output La\TeX{} or HTML files.
+
+\begin{verbatim}
+ruby vocdoc.rb [parameters] source_file target_folder
+\end{verbatim}
+
+\subsubsection{Command-line Parameters} % (fold)
+\label{ssub:command_line_parameters}
+
+The following command-line parameters are available:
+
+\singlespacing
+\begin{verbatim}
+-h, --help          Displays help message
+-v, --version       Display the version, then exit
+-q, --quiet         Output as little as possible, overrides verbose
+-V, --verbose       Verbose output
+--latex             output latex source into target folder, divided
+                      into "classes.tex" and "properties.tex". This 
+                      is the default.
+--html              output html source ("ontology.html") into 
+                      target folder
+\end{verbatim}
+\doublespacing
+
+% subsubsection command_line_parameters (end)
+
+\subsubsection{Requirements} % (fold)
+\label{ssub:requirements}
+
+The following requirements must be fulfilled in order to run VocDoc:
+
 \begin{itemize}
-	\item document configuration file
-	\item will take any RDF, but the use of specific annotations will improve the usefulness of the output 
-	\item document RDF annotation vocabulary 
-	\item describe HTML output (example screenshots)
-	\item describe latex output (part of this deliverable was produced with VocDoc)
+	\item Ruby must be installed.
+	\item The Redland RDF library and Ruby language bindings must be installed.
+	\item A configuration file \texttt{owl\_doc.conf} must be present in the folder where VocDoc is run.
 \end{itemize}
 
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\linewidth]{images/vocdoc_preamble}
-    \caption{Screenshot showing the HTML output of VocDoc}
-    \label{fig:vocdoc_preamble}
-  \end{center}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\linewidth]{images/vocdoc_overview}
-    \caption{Screenshot showing the HTML output of VocDoc}
-    \label{fig:vocdoc_overview}
-  \end{center}
-\end{figure}
+% subsubsection requirements (end)
+
+% subsection running_vocdoc (end)
 

### 2011-02-21 trunk/deliverables/D4_3/fast_d4_3_2.tex
incorporating Albert's changes
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1218 d8b0c59b-6544-0410-94d9-83a5658a058e
Added subsection 3.6 discussing other approaches to service wrapping.
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1274 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2011-01-16 FAST D4_3_3/fast_d4_3_3.tex
@@
-\documentclass{fast_latex}
+\documentclass{article}
+
+\usepackage{fast_latex}
 % yay! We can use any kind of funky diacritic:
@@
 %0.2 & 15.02.2009 & Knud M\"oller (NUIG) & content moved from wiki after internal review \\ \hline
-1.0 & 26.02.2010 & Ismael Rivera (NUIG), Albert Zündorf (UniKas) & Final version ready for external review \\ \hline
+1.0 & 26.02.2011 & Ismael Rivera (NUIG), & Final version ready for external review \\ 
+ & & Albert Zündorf (UniKas) & \\ \hline
 \end{tabular}
@@
 
+\subsection{Other approaches} % (fold)
+\label{sub:other_approaches}
+
+There is a number of related work that we have consider to use for service wrapping instead of building a Service Wrapper
+Tool ourself. However, due to the special needs of the Gadget Development Process it turned out that we were not able 
+to use an existing approach directly but we were only able to grab 
+ideas from existing approaches and come up with our own implementation. 
+As discussed, service wrapping requires three steps.
+First we have to define the connection between the desired service wrapper and the other building blocks of FAST.
+This is achieved by the definition of pre- and postconditions in our tool.
+Since this is very FAST specific, finding an existing tool for this purpose did not work.
+However, this is a simple step anyhow requiring only
+limited implementation efforts.
+
+Second, we have to construct the request and to send it to the target service. There is already a number of REST clients 
+for this purpose. There are even web based REST clients that may be used in a browser. We could easily have used such 
+a standard REST client. However, we considered the user interfaces of these REST clients as too complicated for our
+end users. In addition, we wanted an easy and simple connection to the pre- and postconditions of the service wrapper. 
+This required the integration of a template expansion mechanism into the REST client which was most easy to achieve 
+with a new interface build by ourself. Again the effort for this step was limited and justified by the smooth integration into 
+the overall tool.
+
+Third, we needed a data mediation step. This might have been achieved by e.g. XSLT tools. XSLT tools work on 
+XML input and provide a template based generation of textual output. The problem with XSLT is a quite 
+complex syntax which is not adequate for our end users. It would also have been difficult to provide our end users 
+with direct feedback, as required by FAST's  storyboard philosophy. Similarly, interactive help e.g. by selection boxes for target 
+types and target attribute name would not have been possible, easily.
+Instead of XSLT we might also have used model based transformation languages like ATL (Atlas Transformation Language) or TGGs (Triple Graph Grammars).
+While these
+approaches work on a higher level of abstraction compared to XSLT, we still consider them too complicated for our end users. 
+In addition, it would have been rather difficult to achieve an integrated interactive support as we have achieved in our 
+own implementation. However, our translation rules are heavily influenced by TGGs which was a tremendous help 
+for coming up with a simple and easy to use yet powerful approach.
+
+% subsection Other approaches (end)
+
 % section restful_web_services_wrapper_tool (end)
@@
 \addcontentsline{toc}{section}{References}
-\bibliography{fast_d4_3_2}
+\bibliography{fast_d4_3}
 

### 2011-02-21 trunk/deliverables/D2_2_Ontology/fast_d2_2_3_ontology.tex
improved ontology UML diagram
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1275 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 However, a handful of small changes and corrections have been performed, e.g., to reflect the terminology changes in the prototype section (\emph{prototype} instead of \emph{template}, \emph{clone} instead of \emph{copy}).
+	\item Based on comments from the previous review, Fig.~\ref{fig:ontology_uml} has been corrected and improved with a legend.
 	\item The document as a whole has been revised, and minor changes (mostly, spelling, grammar and small updates to reflect latest developments) have been throughout the deliverable.
@@
 As the figure shows, a \emph{gadget} in FAST is mainly composed of a \emph{screen flow}, which it wraps.
-The \emph{screen flow} is composed of one or more \emph{screens}, which is turn comprises the three \emph{screen components} of the visual front-end (the \emph{form}), the back end service (the \emph{resource}) and the data \emph{operator}.
+The \emph{screen flow} is composed of one or more \emph{screens}, which in turn comprises the three \emph{screen components} of the visual front-end (the \emph{form}), the back end service (the \emph{resource}) and the data \emph{operator}.
 Finally, \emph{forms} are assembled from \emph{form elements}, while \emph{resources} wrap \emph{backend services}.

### 2011-02-21 trunk/deliverables/D4_3/fast_d4_3_2.tex
Ismael review: I fixed some mistakes, and make a few comments for improvement
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1277 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
-\documentclass{article}
+\documentclass{fast_latex}
 
-\usepackage{fast_latex}
 % yay! We can use any kind of funky diacritic:
@@
 
-\newcommand\deliverableNumber{D4.3.3}
+\newcommand\deliverableNumber{D4.3.2}
 \newcommand\deliverableTitle{Mechanisms for Gadget-Service Connections and Gadget Functionality}
@@
 \newcommand\workpackageTitle{Visual composition of screen-flow resources and interoperability with back-end Web Services}
-\newcommand\authorOne{Ismael Rivera, NUIG}
-\newcommand\authorTwo{Albert Zündorf, UniKas}
-\newcommand\authorThree{Author 3}
+\newcommand\authorOne{Ismael Rivera (NUIG)}
+\newcommand\authorTwo{Albert Zündorf (UniKas)}
+\newcommand\authorThree{Knud M\"oller (NUIG)}
 \newcommand\authorFour{Author 4}
@@
 		{\large \sffamily \authorTwo}\\
+		{\large \sffamily \authorThree}\\
 		\vspace{0.5cm}
 		\vfill
-		{\large \sffamily Date: 26/02/2010}\\
+		{\large \sffamily Date: 27/02/2010}\\
 		\vspace{1cm}
@@
 %0.2 & 15.02.2009 & Knud M\"oller (NUIG) & content moved from wiki after internal review \\ \hline
-1.0 & 26.02.2011 & Ismael Rivera (NUIG), & Final version ready for external review \\ 
+1.0 & 26.02.2010 & Ismael Rivera (NUIG), & Final version ready for external review \\ 
  & & Albert Zündorf (UniKas) & \\ \hline
+2.0 & 27.02.2011 & Ismael Rivera (NUIG), & Final version ready for external review \\ 
+ & & Albert Zündorf (UniKas) & \\
+ & & Knud M\"oller (NUIG) & \\ \hline
 \end{tabular}
@@
 	\hline
-	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2009} & {\textbf{Submitted}} & {27/02/2009}\\ \hline
+	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2011} & {\textbf{Submitted}} & {27/02/2011}\\ \hline
 	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
 	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
-	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D4.3.3 is...}\\ \hline
-	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne}\\
+	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D4.3.2 is...}\\ \hline
+	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne, \authorTwo, \authorThree}\\
 %	{} & \multicolumn{4}{l|}{}\\ 
@@
 	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {ismael.rivera@deri.org} \\ \cline{4-5}
-	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 ******} \\ \hline
+	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 ?????} \\ \hline
 	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, web services, WSDL, REST, SOAP, WSMO}\\ \hline
@@
 This deliverable is structured as follows.
-Section ~\ref{sec:introduction} states the goal, scope and structure of the document.
-Section ~\ref{sec:web_services_wrapping} describes what a service wrapper is, how the platform is able to interact with web services, how these wrappers are build and then discovered to be reused in any gadget.
-Finally, Section ~\ref{sec:restful_web_services_wrapper_tool} defines the service wrapper tool built to facilitate FAST users to create Resource Adapters of RESTful web services.
+Section~\ref{sec:introduction} states the goal, scope and structure of the document.
+Section~\ref{sec:web_services_wrapping} describes what a service wrapper is, how the platform is able to interact with web services, how these wrappers are build and then discovered to be reused in any gadget.
+Finally, Section~\ref{sec:restful_web_services_wrapper_tool} defines the service wrapper tool built to facilitate FAST users to create Resource Adapters of RESTful web services.
+
+% subsection structure_of_the_document (end)
+
+\subsection{Changes from Previous Version} % (fold)
+\label{sub:changes_from_previous_version}
+
+\textcolor{red}{To be done...}
+
+% subsection changes_from_previous_version (end)
 
@@
 This is called Resource Adapter in the FAST platform.
-Figure ~\ref{fig:complex_gadget_architecture} illustrates their position and interaction with the rest of the building blocks within a gadget.
+Figure~\ref{fig:complex_gadget_architecture} illustrates their position and interaction with the rest of the building blocks within a gadget.
 
@@
 From a specific WSDL definition, a set of methods or operations can be easily extracted which encloses a set of inputs and outputs.
-These operations would be transformed into actions, and the inputs and outputs would be used to define Resource Adapter pre/postconditions.
+These operations would be transformed into actions, and the inputs and outputs would be used to define Resource Adapter pre-/post-conditions.
 With the use of this information along with the XML Schema defining the types, a Resource Adapter can be semantically defined.
@@
 To integrate WSDL defined services into the FAST platform, we again propose an interactivce approach: We extend the tool for REST service wrapping with the ability of loading WSDL definitions of a service and this WSDL definition is then exploited for guidance of the user through the Resource Adapter construction, cf.
-Section ~\ref{sec:restful_web_services_wrapper_tool}.
+Section~\ref{sec:restful_web_services_wrapper_tool}.
 
@@
 A web service in WSMO defines one and only one capability.
-The capability of a web service defines its functionality in terms of pre/postconditions, assumptions and effects.
-A web service capability is defined by specifying the following elements: non-functional properties, imported ontologies, used mediators, shared variables, precondition, postcondition, assumption, and effect.
+The capability of a web service defines its functionality in terms of pre-/post-conditions, assumptions and effects.
+A web service capability is defined by specifying the following elements: non-functional properties, imported ontologies, used mediators, shared variables, pre-condition, post-condition, assumption, and effect.
 
-Basically, the web service will offers to a client a postcondition when some conditions are met in the information space (precondition).
+Basically, the web service will offers to a client a post-condition when some conditions are met in the information space (pre-condition).
 Effects and assumptions can stay out of this study, since they will not be taken into account by any of the building blocks of a gadget.
@@
 Like WSDL defined services, we incorporate WSMO services into the FAST platform in an interactive storyboard driven approach.
-The FAST tool described in Section ~\ref{sec:restful_web_services_wrapper_tool} reads the WSMO specification of a service and exploits this information again for guidance for the contstruction of an Resource Adapter.
+The FAST tool described in Section~\ref{sec:restful_web_services_wrapper_tool} reads the WSMO specification of a service and exploits this information again for guidance for the construction of a Resource Adapter.
 
@@
 In the building phase, every Resource Adapter is created sharing a common structure for screen components, such as forms, operators and resources.
-Hence, it will have a set of actions which will contain a set of preconditions to satisfy in order to be executed, and after its execution, it may produce any of the conditions of their postconditions.
+Hence, it will have a set of actions which will contain a set of pre-conditions to satisfy in order to be executed, and after its execution, it may produce any of the conditions of their post-conditions.
 
-For this reason, a screen developer during the screen development phase is able to reuse these screen components which are already stored in the catalog, establishing which the screen's pre/postconditions are, and then the system will suggest several screen components which may satisfy these pre/postconditions.
-In a first step, the system will suggest screen components which have any of the preconditions which can be connected to any of the screen preconditions (the connection will be represented by an internal pipe), and the same would happen for the postconditions.
-Once a screen component is inserted into the screen, the system will take its definition into consideration to recommend new screen components which can be connected to the ones the screen is composed of.
+For this reason, a screen developer during the screen development phase is able to reuse these screen components which are already stored in the Catalogue, establishing which the screen's pre-/post-conditions are, and then the system will suggest several screen components which may satisfy these pre-/post-conditions.
+In a first step, the system will suggest screen components which have any of the pre-conditions which can be connected to any of the screen pre-conditions (the connection will be represented by an internal pipe), and the same would happen for the post-conditions.
+Once a screen component is inserted into the screen, the system will take its definition (screen components, pipes, etc.) into consideration to recommend new screen components which can be connected to the ones the screen is composed of.
 
-In these bases, the mechanism is relatively similar to the screen recommender, used to find the best screens to create a screen flow.
-Once more, it is based on the pre/postconditions of a set of building blocks.
+For a further explanation on how the recommendation algorithms are implemented, please see~\cite{rivera2011catalogue_user_manual}.
 
@@
 \clearpage
-\section{Web Service Wrapper Tool} % (fold)
+\section{Service Wrapper Tool} % (fold)
 \label{sec:restful_web_services_wrapper_tool}
 
-In FAST, the wrapper tool will be in charge of the Resource Adapters' building phase.
-The building of Resource Adapters is done in three steps: in a first step, we define which pre- and postconditions are used to pass parameters between FAST building blocks and the Resource adapter.
-The second step is in charge of the construction of a service request and the third step analyses the response got from the execution of the service, allowing the extraction of facts contained in that response and mapping them to domain-specific concepts from the ontologies used within FAST by any building block.
+In FAST, the Service Wrapper Tool will be in charge of the Resource Adapters' building phase.
+The building of resource adapters is done in three steps: in a first step, we define which pre- and post-conditions are used to pass parameters between FAST building blocks and the resource adapter.
+The second step is in charge of the construction of a service request and the third step analyses the response got from the execution of the service, allowing the extraction of facts contained in that response and mapping them to domain-specific classes or concepts from the ontologies used within FAST by the building blocks' conditions.
 
-\subsection{Defining pre- and postconditions} % (fold)
+\subsection{Defining pre- and post-conditions} % (fold)
 \label{sub:resource_adapter_parameters}
 
-We will demonstrate the use of the FAST Service Wrapper Tool on the example of a simple Formula1 service.
-In the first tab of the FAST Service Wrapper Tool, the user defines the name and the pre- and postconditions of the desired Resource Adapter, cf.
+We will demonstrate the use of the Service Wrapper Tool on the example of a simple \emph{Formula1} service.
+In the first tab of the Service Wrapper Tool, the user defines the name and the pre- and post-conditions of the desired Resource Adapter, cf.
 Fig.
 \ref{fig:construct_pre_post_conditions}.
-For the precondition, we provide a name, a type, and an example value.
-The type of a pre- or postcondition is chosen from a drop down list of available concepts.
-This list of available concepts is retrieved from the FAST ontology catalog.
-The example value is used for test runs during subsequent construction steps.
+For the pre-condition, we provide a name, a type, and an example value.
+The type of a pre- or post-condition is chosen from a drop down list of available concepts.
+The list of available concepts, together with its properties and some example data (instances) is retrieved from the Catalogue.
+The example value is used for test requests during subsequent construction steps.
 We keep track of example values provided for different domain concepts and propose them, when the domain concept is re-used e.g.
@@
   \begin{center} \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSWithPrePostConds.png}
-    \caption{Configuring pre- and postconditions of a Resource Adapter}
+    \caption{Configuring pre- and post-conditions of a Resource Adapter}
     \label{fig:construct_pre_post_conditions}
@@
 In our simple case, the service request consists just of a URL that is easily retrieved from API documentation page of our Formula1 service.
-This URL consists of the service address \url{http:// ergast.com/api} followed by the year we are querying for followed by the category \url{drivers} we are interested in.
-The user may grab such an URL from the API documentation page of the targeted web service and drop it into the Request tab of the FAST Service Wrapper Tool, cf.
-Figure ~\ref{fig:construct_service_request}.
-The FAST Service Wrapper Tool will than analyze the dropped URL and split it into its parts and provide a convenient form allowing to edit e.g.
+This URL consists of the service address \url{http://ergast.com/api} followed by the year we are querying for followed by the category \verb|drivers| we are interested in.
+The user may grab such an URL from the API documentation page of the targeted web service and drop it into the Request tab of the Service Wrapper Tool, cf.
+Figure~\ref{fig:construct_service_request}.
+The Service Wrapper Tool will than analyze the dropped URL and split it into its parts and provide a convenient form allowing to edit e.g.
 all the parameters of the URL.
@@
 
-For WSDL or WSMO services, the FAST Service Wrapper Tool provides additional support for the construction of the request.
-The user may point the FAST Service Wrapper Tool to the WSDL or WSMO definition of the targeted service.
-The FAST Service Wrapper Tool then reads this definition and analysis it.
-With the help of this information, the FAST Service Wrapper Tool provides interactive guidance for selecting the desired operation from a service and then the FAST Service Wrapper Tool retrieves the operation descriptions and fills the request parameter form, adequately.
+For WSDL or WSMO services, the Service Wrapper Tool provides additional support for the construction of the request.
+The user may point to the WSDL or WSMO definition of the targeted service.
+The Service Wrapper Tool then reads this definition and analyses it.
+With the help of this information, the Service Wrapper Tool provides interactive guidance for selecting the desired operation from a service and then it retrieves the operation descriptions and fills the request parameter form adequately.
 
-Now we have to connect the request parameters to the parameter values passed via the preconditions to the Resource Adapter.
-In our case, we replace the year part of the URL with a placeholder for the precondition season \url{<season>}, see Request URL input field at the top of Figure ~\ref{fig:construct_service_request}.
+Now we have to connect the request parameters to the parameter values passed via the pre-conditions to the Resource Adapter.
+In our case, we replace the year part of the URL with a placeholder for the pre-condition season \verb|<season>|, see Request URL input field at the top of Figure~\ref{fig:construct_service_request}.
 
-In our example, we use a simple GET based web service.
-The FAST Service Wrapper tool also supports the other kinds of URL operations as POST, PUT, etc.
+In our example, we use a simple HTTP GET request.
+The Service Wrapper Tool also supports the other kinds of HTTP methods as POST, PUT or DELETE.
 For POST requests, one usually passes parameters within a special post body.
-For unstructured POST bodies, we provide a text area allowing the user to fill in placeholders for references to precondition values.
+For unstructured POST bodies, we provide a text area allowing the user to fill in placeholders for references to pre-condition values.
 In case of structured POST bodies as in the case of WSDL and WSMO services the POST body uses a structured text format based e.g.
@@
 In that case, the body text is again passed and the parameters are included in the request parameter form part.
-Then the user may easily use the request parameter form part to fill in precondition placeholders.
+Then the user may easily use the request parameter form part to fill in pre-condition placeholders.
 
-Figure ~\ref{fig:construct_service_request} shows the screen to construct the service requests.
+\textcolor{red}{WSDL is usually used with SOAP and XML, JSON wouldn't be common in this case.
+And from my understanding, the user has to hand-write the POST body?
+therefore the SOAP message?
+that's doable, but pretty hard and tedious.
+- Ismael}
+
+\textcolor{red}{Is PUT and DELETE supported? if so, how? - Ismael}
+
+Figure~\ref{fig:construct_service_request} shows the screen to construct the service requests.
 
@@
 
-Below the request paramter editing form of Figure ~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed request by sending it to the specified service address (via a server relay).
-Then the placeholders for preconditions are replaced with example values and the resulting request is shown below the parameter form.
-In addition, the request is send and the response is shown on the bottom of that page.
+Below the request parameter editing form of Figure~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed request by sending it to the specified service address (via a server relay).
+Then the placeholders for pre-conditions are replaced with example values and the resulting request is shown below the parameter form.
+In addition, the request is sent and the response is shown on the bottom of that page.
 This gives the user a fast feedback whether the constructed request works as desired.
@@
 
-It is worth pointing out that currently the wrapper tool is able to construct input ports for the wrappers using just basic types.
-However, every building block accepts any (complex) concept as a pre/postcondition, so the wrapper tool needs to be adapted to permit any concept as input or precondition, extracting the required text to construct the request.
+It is worth pointing out that currently the Service Wrapper Tool is able to construct input ports for the wrappers using just basic types.
+However, every building block accepts any (complex) concept as a pre-/post-condition, so the wrapper tool needs to be adapted to permit any concept as input or pre-condition, extracting the required text to construct the request.
 This should be easy to achieve before the project review.
+
+\textcolor{red}{I guess this paragraph will be rewritten...because my understanding is that we do support constructing wrappers with ``complex'' pre-/post-conditions and not just basic types.
+- Ismael}
 
@@
 Once the service request is constructed and sent to the service provider, it will send back a response.
-This response message could be serialize in any format, though the most common formats used nowadays are XML or JSON among others.
-To continue the example started in the previous section, the response of the Formula1 service will is in XML.
-However, the FAST wrapper tool supports JSON, similarly.
+This response message could be serialize in any format, though the most common formats used nowadays are XML or JSON.
+To continue the example started in the previous section, the response of the Formula1 service will is formatted in XML.
+Th Service Wrapper Tool supports both JSON and XML.
 
@@
 
-Figure ~\ref{fig:response_service_execution} shows the data tranformation tab of the wrapper tool.
+Figure~\ref{fig:response_service_execution} shows the data tranformation tab of the wrapper tool.
 
@@
       \includegraphics[angle=90,width=0.8\linewidth]{images/ServiceWrapperToolGVSWithTransformationRulesF1.png}
-    \caption{Interactive, rule based transforming of an XML response to FAST facts (photoshoped screen dump)}
+    \caption{Interactive, rule based transforming of an XML response to FAST facts}
     \label{fig:response_service_execution}
@@
 
-Once the service response, in XML or JSON format, has been retrieved, the transformation tab shows it as an interactive object tree on the left side of Figure ~\ref{fig:response_service_execution}.
+Once the service response has been retrieved, it is shown in the transformation tab as an interactive object tree, as it can be seeing on the left side of Figure~\ref{fig:response_service_execution}.
 To construct this interactive object tree, the XML or JSON document is parsed into a DOM, and a simplified tree representation of that DOM is built up.
@@
 
-A transformation rule is used to analyse the XML/JSON data and to generate domain-specific facts from concepts from the ontologies used by the pre/postconditions of the different building blocks.
-A transformation rule is composed of three elements, cf the middle part of Figure ~\ref{fig:response_service_execution}.
+A transformation rule is used to analyse the XML/JSON data and to generate domain-specific facts from concepts from the ontologies used by the pre-/post-conditions of the different building blocks.
+A transformation rule is composed of three elements, as the middle part of Figure~\ref{fig:response_service_execution} shows.
 First, the \textit{from} field indicates the XML elements to be translated by rule.
@@
 And third, the target of the rule specifies a certain concept or attribute, to be created or filled.
-A detailed explanation of the type of actions to be trigger from the transformation rules is:
+A detailed explanation of the type of actions to be triggered from the transformation rules is:
 \begin{description}
@@
 In the example being explained, the root rule searches for XML elements with tagname \emph{DriverTable} and for each such element a \emph{List} fact is created.
-The resulting facts are shown in a facts tree in the right of Figure ~\ref{fig:response_service_execution}.
+The resulting facts are shown in a facts tree in the right of Figure~\ref{fig:response_service_execution}.
 	\item[action \emph{fillAttributes}] does not create a new fact but it fills the value of the attribute provided as third part of such rules.
@@
 
-Since FAST is storyboard oriented, the FAST Service Wrapper Tool follows the storyboard idea as well.
+Since FAST is storyboard oriented, the Service Wrapper Tool follows the storyboard idea as well.
 Any time, a change to a transformation rule is done, the transformation process is triggered and the resulting facts tree is directly shown.
 This process helps the user to deal with the slightly complex semantics of the transformation rules avoiding errors or mistakes.
-In addition, FAST is semantic-driven, therefore, the service wrapper designer retrieves the domain-specific types from the FAST ontology catalog server together with the structure of each type, i.e.
+In addition, FAST is semantic-driven, therefore, the service wrapper designer retrieves the domain-specific types from the Catalogue server together with the structure of each type, i.e.
 together with a description of the attributes of each fact.
@@
 In addition, we may provide some analysis tool, which will help to guarantee that the facts generated by the transformation rules conform to the fact types defined in the corresponding FAST ontology.
-This helps to ensure that the facts generated by the designed Resource Adapter will be compatible for preconditions of subsequent filter steps and or screens.
+This helps to ensure that the facts generated by the designed Resource Adapter will be compatible for pre-conditions of subsequent filter steps and or screens.
+
+\textcolor{red}{We may provide some analysis tool? is it implemented? otherwise I don't see it's worth to point this out, or maybe just in a ``future work'' section - Ismael}
 
@@
 
-Once the wrapping of a service has been defined and tested in the FAST Service Wrapper Tool, it generates an implementation of the desired Resource Adapter in XML, HTML, and JavaScript, ready to be deployed and executed inside a gadget.
-In addition, a description of the new building block is added to the FAST catalog.
-Thereby, the new Resource Adapter becomes available for the construction of new screens and screen flows.
+Once the wrapping of a service has been defined and tested in the Service Wrapper Tool, it generates an implementation of the desired Resource Adapter in XML, HTML, and JavaScript, ready to be deployed and executed inside a gadget.
+In addition, a description of the new building block is added to the FAST Catalogue.
+Thereby, the new Resource Adapter becomes available for the construction of new screens and screen-flows.
+
+\textcolor{red}{XML, HTML and Javascript?
+I don't get why XML and HTML is generated.
+The service wrapper should be just a javascript encapsulation of the web service, isn't it?
+HTML files should be for forms and XML for specific platform configuration files for deployment?
+- Ismael}
 
@@
 The rule driven approach presented above is somewhat limited.
-It is deliberately restricted to such a simple rule mechanism in order to keep things simple enough for end-users.
-Still, the selected approach suffices for most practical and real world cases.
+It is deliberately restricted to such a simple rule mechanism in order to keep things simple enough for end users.
+However, the selected approach suffices for most practical and real world cases.
 As a more complex example, the XML data for a person may provide two different tags for the first and the last name of a person.
@@
 For example the attribute \texttt{grandmother} could be filled using \texttt{mother.mother} in the \textit{from} field.
+\textcolor{red}{This paragraph doesn't talk about limitations, these are features! - Ismael}
 
@@
 To cover such cases, intermediate fact formats can be used which would allow generating facts to be further processed by additional filters.
-Such additional filters may be realized using (hand coded) operators, since some generic operators can act as filters for aggregation and conversions of facts from multiple sources.
+Such additional filters may be realized using hard-coded operators, since some generic operators can act as filters for aggregation and conversions of facts from multiple sources.
 Then, service wrappers in combination with these filter operators will allow covering these complex cases.
@@
 Tool ourself. However, due to the special needs of the Gadget Development Process it turned out that we were not able 
-to use an existing approach directly but we were only able to grab 
-ideas from existing approaches and come up with our own implementation. 
+to use an existing approach directly but we were only able to grab ideas from existing approaches and come up with our own implementation. 
 As discussed, service wrapping requires three steps.
 First we have to define the connection between the desired service wrapper and the other building blocks of FAST.
-This is achieved by the definition of pre- and postconditions in our tool.
+This is achieved by the definition of pre- and post-conditions in our tool.
 Since this is very FAST specific, finding an existing tool for this purpose did not work.
-However, this is a simple step anyhow requiring only
-limited implementation efforts.
+However, this is a simple step anyhow requiring only limited implementation efforts.
 
@@
 a standard REST client. However, we considered the user interfaces of these REST clients as too complicated for our
-end users. In addition, we wanted an easy and simple connection to the pre- and postconditions of the service wrapper. 
+end users. In addition, we wanted an easy and simple connection to the pre- and post-conditions of the service wrapper. 
 This required the integration of a template expansion mechanism into the REST client which was most easy to achieve 
-with a new interface build by ourself. Again the effort for this step was limited and justified by the smooth integration into 
-the overall tool.
+with a new interface build by ourself. Again the effort for this step was limited and justified by the smooth integration into the overall tool.
 
@@
 XML input and provide a template based generation of textual output. The problem with XSLT is a quite 
-complex syntax which is not adequate for our end users. It would also have been difficult to provide our end users 
-with direct feedback, as required by FAST's  storyboard philosophy. Similarly, interactive help e.g. by selection boxes for target 
-types and target attribute name would not have been possible, easily.
+complex syntax which is not adequate for our end users.\textcolor{red}{Well, no more than writing a SOAP message :) - Ismael} It would also have been difficult to provide our end users with direct feedback, as required by FAST's storyboard philosophy.
+Similarly, interactive help e.g.
+by selection boxes for target types and target attribute name would not have been possible, easily.
 Instead of XSLT we might also have used model based transformation languages like ATL (Atlas Transformation Language) or TGGs (Triple Graph Grammars).
-While these
-approaches work on a higher level of abstraction compared to XSLT, we still consider them too complicated for our end users. 
+While these approaches work on a higher level of abstraction compared to XSLT, we still consider them too complicated for our end users.
 In addition, it would have been rather difficult to achieve an integrated interactive support as we have achieved in our 
@@
 for coming up with a simple and easy to use yet powerful approach.
+
+\textcolor{red}{I couldn't really understand how a wrapper for a WSDL is constructed.
+The wrapper tool uses any library (Axis, etc) to parse the WSDL definition and extract the operations with their inputs and outputs?
+If so, it should be mention.
+Or should be the user who reads the WSDL document, checks the endpoint URL for the operation that wants to wrap, and then construct the wrapper as in the RESTful case, writting the SOAP message for the request body?
+- Ismael}
 
@@
 
+\textcolor{red}{is this section done? - Ismael}
+
 % section related_work (end)

### 2011-02-21 trunk/catalogue_service/src/eu/morfeoproject/fast/catalogue/model/Screen.java
'code' and 'definition' are returned if both are defined, there is no preference for any (before if 'code' was found, the 'definition' was not returned)
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1278 d8b0c59b-6544-0410-94d9-83a5658a058e

@@ -102,33 +102,43 @@ public class Screen extends WithConditions {
 		URI screenUri = getUri();
 		
 		model.addStatement(screenUri, RDF.type, FGO.Screen);
+		
+		/* 
+		 * 'code' and 'definition' define the behaviour of the building block.
+		 * only one of them should be declared, but it's up to the application
+		 * to decide whether code or the definition will be used, here both
+		 * elements are supported and retrieved.
+		 */
+		
+		//--- CODE ---//
 		if (getCode() != null) {
 			model.addStatement(screenUri, FGO.hasCode, this.getCode());
-		} else {
-			// building blocks
-			for (URI uri : getBuildingBlocks()) {
-				model.addStatement(screenUri, FGO.contains, uri);
+		}
+		
+		//--- DEFINITION ---//
+		// building blocks
+		for (URI uri : getBuildingBlocks()) {
+			model.addStatement(screenUri, FGO.contains, uri);
+		}
+		// pipes
+		for (Pipe pipe : getPipes()) {
+			URI pUri = pipe.getUri();
+			if (pUri == null) {
+				pUri = model.createURI(screenUri+"/pipes/"+UUID.randomUUID().toString());
+				pipe.setUri(pUri);
 			}
-			// pipes
-			for (Pipe pipe : getPipes()) {
-				URI pUri = pipe.getUri();
-				if (pUri == null) {
-					pUri = model.createURI(screenUri+"/pipes/"+UUID.randomUUID().toString());
-					pipe.setUri(pUri);
-				}
-				model.addStatement(screenUri, FGO.contains, pUri);
-				model.addModel(pipe.toRDF2GoModel());
+			model.addStatement(screenUri, FGO.contains, pUri);
+			model.addModel(pipe.toRDF2GoModel());
+		}
+		// triggers
+		for (Trigger trigger : getTriggers()) {
+			URI tUri = trigger.getUri();
+			if (tUri == null) {
+				tUri = model.createURI(screenUri+"/triggers/"+UUID.randomUUID().toString());
+				trigger.setUri(tUri);
 			}
-			// triggers
-			for (Trigger trigger : getTriggers()) {
-				URI tUri = trigger.getUri();
-				if (tUri == null) {
-					tUri = model.createURI(screenUri+"/triggers/"+UUID.randomUUID().toString());
-					trigger.setUri(tUri);
-				}
-				model.addStatement(screenUri, FGO.contains, tUri);
[…]

### 2011-02-22 trunk/deliverables/D2_2_Ontology/fast_d2_2_3_ontology.tex
final version
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1279 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 A number of appendices have been added to the document: App.~\ref{sec:methodology} presents the development methodology adopted by us, App.~\ref{sec:ontology_terms} provides documentation for the ontology in the form of a detailed list of terms, while App.~\ref{sec:ontology_code} contains the complete code of the ontology.
+Appendix~\ref{sec:vocdoc} describes and documents the VocDoc tool, which was developed to automatically create ontology documentation from the original ontology code.
 
@@
 	\item Based on comments from the previous review, Fig.~\ref{fig:ontology_uml} has been corrected and improved with a legend.
-	\item The document as a whole has been revised, and minor changes (mostly, spelling, grammar and small updates to reflect latest developments) have been throughout the deliverable.
-	\item The online documentation of the FAST Gadget ontology at \url{http://purl.oclc.org/fast/ontology/gadget} has been updated to reflect the changes in this version of the deliverable.
+	\item Also in response to reviewer comments, some concrete examples for instances of the various classes in the FAST ontology have been added to Sect.~\ref{ssub:term_relations_within_the_conceptual_model}.
+	\item The document as a whole has been revised, and minor changes (mostly, spelling, grammar and small updates to reflect latest developments) have been made throughout the deliverable.
+	\item The online documentation of the FAST Gadget ontology at \url{http://purl.oclc.org/fast/ontology/gadget} has been updated to reflect the changes in this version of the ontology.
 \end{itemize}
@@
 \textbf{Properties:} \emph{containsScreen}, \emph{hasLabel}, \emph{hasIcon}, \emph{hasPreCondition}, \emph{hasPostCondition}, \emph{hasTag}, \emph{hasProfile}\\ \hline
-\textcolor{white}{\textbf{Sources of Knowledge}} & FAST Architecture deliverable~\cite{urena2010fast_architecture}, FAST Requirements Specification~\cite{villoslada2010fast_requirements}, FAST State-of-the-art Document~\cite{urmetzer2010fast_state_of_the_art}, EzWeb documentation~\cite{lizcano2008ezweb} \\ \hline
+\textcolor{white}{\textbf{Sources of Knowledge}} & FAST Architecture deliverable~\cite{lopez2011fast_architecture}, FAST Requirements Specification~\cite{villoslada2010fast_requirements}, FAST State-of-the-art Document~\cite{urmetzer2010fast_state_of_the_art}, EzWeb documentation~\cite{lizcano2008ezweb} \\ \hline
 \end{longtable}
@@
 \texttt{Pattern} & A set of facts which formally defines a condition. A pattern may contain one or more variables. \\ \hline
-\texttt{Fact} & The ``basic information unit of a FAST gadget''~\cite{urena2010fast_architecture}.
+\texttt{Fact} & The ``basic information unit of a FAST gadget''~\cite{lopez2011fast_architecture}.
 In terms of RDF, a fact is a statement consisting of a subject, predicate and object \emph{(S, P, O)}.
@@
 Finally, \emph{forms} are assembled from \emph{form elements}, while \emph{resources} wrap \emph{backend services}.
+To illustrate, Fig.~\ref{fig:screen} shows the picture of a screen, as it is actually rendered in a web browser.
 
@@
 
+\begin{figure}
+  \begin{center}
+    \includegraphics[scale=.5]{images/screen.png}
+    \caption{A FAST screen in a web browser}
+    \label{fig:screen}
+  \end{center}
+\end{figure}
+
 Figure~\ref{fig:uml_screen_component} looks closer at the features of individual screens.
@@
 Each such screen component may or may not have an action or trigger associated with them, which declaratively define their behaviour in terms of their own functionality or functionality they can trigger in other building blocks.
+Some concrete examples of various different screen components are given in Fig.~\ref{fig:building_blocks}, which shows a screenshot from the screen composition tool in the FAST GVS \cite{lopez2011fast_architecture}.
 
@@
     \label{fig:uml_screen_component}
+  \end{center}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[scale=.66]{images/building_blocks.pdf}
+    \caption{Examples of FAST building blocks during screen composition}
+    \label{fig:building_blocks}
   \end{center}
@@
 
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=\linewidth]{images/declarativeDefinitionExample.pdf}
-    \caption{Explicit piping within a screen from a post- to a pre-condition (diagramme)}
-    \label{fig:piping}
-  \end{center}
-\end{figure}
-
 The concept of piping on this level of the FAST platform is illustrated graphically in Fig.~\ref{fig:piping}, and in code in List.~\ref{list:piping}\footnote{The example does not show a complete screen definition, but only the aspects which are relevant for this section.}.
@@
 The desired functionality of this arrangement is that, whenever the service has returned a list of products in response to a search, this list will be piped to the form, where it will be displayed to the gadget's user.
+
+\begin{figure}[h]
+  \begin{center}
+    \includegraphics[width=\linewidth]{images/declarativeDefinitionExample.pdf}
+    \caption{Explicit piping within a screen from a post- to a pre-condition (diagram)}
+    \label{fig:piping}
+  \end{center}
+\end{figure}
 

### 2011-02-22 trunk/deliverables/D2_2_Ontology/properties.tex
final version
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1279 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 \end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasClone}}
+\label{subs:hasClone}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has clone \\ \hline
+\textcolor{white}{\textbf{description}} & Links the prototype of a building block as used in the palette or catalogue of available building blocks to its clone, as used in a particular screen flow or screen.
+\\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\textcolor{white}{\textbf{inverse\_of}} & \htmlref{\texttt{fgo:hasPrototype}}{subs:hasPrototype} \\ \hline
+\end{tabular}
 \subsubsection*{Property: \texttt{fgo:hasCode}}
@@
 \textcolor{white}{\textbf{range}} & \texttt{foaf:Document} \\ \hline
-\end{tabular}
-\subsubsection*{Property: \texttt{fgo:hasCopy}}
-\label{subs:hasCopy}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has copy \\ \hline
-\textcolor{white}{\textbf{description}} & Links the template of a building block as used in the palette or catalogue of available building blocks to its copy, as used in a particular screen flow.
-\\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
-\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
 \end{tabular}
@@
 \end{tabular}
+\subsubsection*{Property: \texttt{fgo:hasPrototype}}
+\label{subs:hasPrototype}
+\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
+\hline
+\textcolor{white}{\textbf{label}} & has prototype \\ \hline
+\textcolor{white}{\textbf{description}} & Links the clone of a building block as used in a particular screen flow or screen to its prototype, as used in the palette or catalogue of available building blocks.
+\\ \hline
+\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
+\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
+\textcolor{white}{\textbf{inverse\_of}} & \htmlref{\texttt{fgo:hasClone}}{subs:hasClone} \\ \hline
+\end{tabular}
 \subsubsection*{Property: \texttt{fgo:hasResource}}
@@
 \textcolor{white}{\textbf{range}} & \texttt{foaf:Image} \\ \hline
-\end{tabular}
-\subsubsection*{Property: \texttt{fgo:hasTemplate}}
-\label{subs:hasTemplate}
-\begin{tabular}{| >{\columncolor{fast@lightgrey}}p{2.5cm}|p{12cm}|}
-\hline
-\textcolor{white}{\textbf{label}} & has template \\ \hline
-\textcolor{white}{\textbf{description}} & Links the copy of a building block as used in a particular screen flow to its template, as used in the palette or catalogue of available building blocks.
-\\ \hline
-\textcolor{white}{\textbf{type}} & \texttt{owl:ObjectProperty} \\ \hline
-\textcolor{white}{\textbf{domain}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
-\textcolor{white}{\textbf{range}} & \htmlref{\texttt{fgo:BuildingBlock}}{subs:BuildingBlock} \\ \hline
 \end{tabular}

### 2011-02-22 trunk/deliverables/D2_2_Ontology/fast_d2_2_3_ontology.tex
final version
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1280 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 Not all properties which will eventually be defined in the formal ontology are listed in the table: in many cases, these concrete properties simply reflect particular instances of generic composition or aggregation relations, which become apparent in the diagrams in Sect.~\ref{ssub:term_relations_within_the_conceptual_model}.
-Usually, these concrete properties will follow a general naming scheme of \texttt{hasCLASS\_NAME}, e.g., \texttt{hasDefinition} or \texttt{hasPreCondition}.
+Usually, these concrete properties will follow a general naming scheme of \texttt{hasCLASS\_NAME}, e.g., \texttt{hasPreCondition}.
 
@@
 
-In FAST, resources (gadgets, screenflows, screens and their components) exist in different stages throughout their lifecycle.
+In FAST, all different kinds of entities (gadgets, screenflows, screens and their components) exist in different stages throughout their lifecycle.
 Moving top-down from the high-level conceptual description to gadgets in use at runtime, the following differences can be made in terms of the level of instantiation:

### 2011-02-24 trunk/deliverables/D2_4_Mediation/fast_d2_4_3_part1.tex
new files for version 2.4.3
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1282 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-02-19 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_part1.tex
@@
 
-To support manual ontology mapping, we plan to develop a rule based, interactive Facts Transformation Tool.
+To support manual ontology mapping, we have developed a rule based, interactive Facts Transformation Tool.
 This tool utilizes techniques of our service wrapper tool, cf.
@@
 This problem is closely related to the problem of ontology mediation.
-Thus, these techniques will also allow to create data transformation operators, i.e.
+Thus, these techniques also allow to create data transformation operators, i.e.
 operators that mediate between two ontologies.
 
-Figure \ref{fig:FactsTransformationToolPortDefinitions} shows a photoshoped preview of the planned Facts Transformation Tool.
-The Facts Transformation Tool has two tabs.
-The \textit{General} tab allows to define the input and output ports for the desired data transformation operator.
-These ports shall be  typed according to the participating semantic elements to be derived from the FAST catalogue.
+Figure \ref{fig:FactsTransformationToolPortDefinitions} shows a screen shot of the  Facts Transformation Tool within the GVS environment.
+The Facts Transformation Tool has three tabs.
+The \textit{General} tab allows to define the pre- and postconditions for the desired data transformation operator.
+These conditions are  typed according to the participating semantic elements to be derived from the FAST catalogue.
 This typing will provide the structure of the input and output data of the desired data transformation operator.
 
-Figure \ref{fig:FactsTransformationToolTransformationRules} shows the planned editor for transformation rules.
+Figure \ref{fig:FactsTransformationToolTransformationRules} shows the editor for transformation rules.
 On the left of Figure \ref{fig:FactsTransformationToolTransformationRules} there is a tree of example data that is used to validate the transformation rules during editing.
@@
     \includegraphics[width=0.8\linewidth]{images/FactsTransformationToolGVSWithPortDefinitionsNoGlitch2.png}
-    \caption{Definition of operator pre- / postconditions (photoshoped screen dump)}
+    \caption{Definition of operator pre- / postconditions}
     \label{fig:FactsTransformationToolPortDefinitions}
@@
 
-Note that simple one-to-one transformation rules should be derived with the help of the ontology mediation approach described in Section \ref{subsubsec:automaticmapping}.
-Such simple rules should be added to the transformation tool, automatically.
+Note that simple one-to-one transformation rules may be derived with the help of the ontology mediation approach described in Section \ref{subsubsec:automaticmapping}.
+Such simple rules may be added to the transformation tool, automatically.
+We try to add this functionality within the remaining project time.
 The more complex rules may then be added interactively.
-During editing, the semantics provided by the FAST catalogue will be used to support the user with completion proposals and consistency checking.
-In addition, the direct execution of the rules on example data will help the user to create appropriate rules and to validate their effects.
+During editing, the semantics provided by the FAST catalogue is used to support the user with completion proposals and consistency checking.
+In addition, the direct execution of the rules on example data helps the user to create appropriate rules and to validate their effects.
 
-So far our transformation rules are restricted to simple cases.
-Thus, we have focused on simple cases allowing a simple transformation approach that does not require sophisticated skills from our users.
-However, as discussed on the Amazon to GoodRelations example, there exist more complex cases.
-In the third year of the project we will investigate how such more complex cases can be covered, too, and whether it is still possible to keep the rules simple and understandable.
+So far our transformation rules are restricted to simple cases. We have focused on simple cases allowing a simple transformation approach that does not require sophisticated skills from our users.
 
@@
 
-Once the data transformation operator has been defined, it will be added to the FAST catalogue in order to allow its use for the definition of piping configurations of new screens within the FAST CGS tool.
-For this purpose a simple semantic description will be sent to FAST catalogue defining the semantics of the operator and the semantics of its input and output ports.
+Once the data transformation operator has been defined, it may be added to the FAST catalogue in order to allow its use for the definition of piping configurations of new screens within the FAST CGS tool.
+For this purpose a simple semantic description is sent to the FAST catalogue defining the semantics of the operator and the semantics of its pre- and postconditions.
 
-Currently, the Service Wrapper Tool uses its own repository for trees of transformation rules.
+The Service Wrapper Tool uses its own repository for trees of transformation rules.
 First of all this has technical reasons: the development of the transformation tool is based on a reference architecture and libraries for web based interactive systems provided by Kassel University, cf.
@@
 This repository has the special advantage that it provides support for teams of developers, thus the service wrapper tool allows to merge the contributions of multiple developers.
-However, on the long run, we shall store the transformation rules together with the service wrapper and transformation operator definitions in the FAST catalogue.
 
@@
 
-At the gadget building phase, whenever preconditions and postconditions of screens, resources or operators show a semantic mismatch, the catalogue will offer the possible matching operators that would solve the incompatibility problem.
+At the gadget building phase, whenever preconditions and postconditions of screens, resources or operators show a semantic mismatch, the catalogue offers the possible matching operators that would solve the incompatibility problem.
 This way a matching operator can be placed between two pre- postconditions that would otherwise be incompatible, making interoperation between the two possible.
 In seldom cases, the catalogue may fail to retrieve an appropriate mediation operator.
-In such cases, the possibility of building a new matching operator with the help of the Facts Transformation Tool will be presented, in order to meet the user's need.
+In such cases, the possibility of building a new matching operator with the help of the Facts Transformation Tool is presented, in order to meet the user's need.
 
@@
 This runtime library provides notification mechanisms and manages the fact base of the gadget.
-For data transformation we will generate Javascript code that connects itself to the fact base and any time a fact is sent to the input port of the operator, it will do the data transformation and send the result to the output port.
+For data transformation we generate Javascript code that connects itself to the fact base and any time a fact is sent to the precondition of the operator, it does the data transformation and sends the result to the postcondition.
 The code generated from the data transformation rules follows a recursive pattern where each rule becomes a Javascript function with parameters for the source and target fact subtrees to be transformed.

### 2011-02-24 trunk/deliverables/D2_4_Mediation/fast_d2_4_3_part2.tex
new files for version 2.4.3
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1282 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-01-14 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_part2.tex
@@
 \item \textit{instance-based/schema-based} matching: A schema-based
-matcher uses different approaches to determine correspondence between
+matcher uses different  \linebreak approaches to determine correspondence between
 ontologies based on concepts and relations \cite{debruin2005wsml}. 
@@
 
+\begin{center}
+\tablecaption{Class by attribute mapping pattern and example. {\newline}Taken from \cite{debruin2005wsml}.}
+\label{tab:MappingPattern}
+\tablehead{}
+\begin{supertabular}{|m{0.8\linewidth}|}
+\hline
+\textbf{Name}: Class by Attribute Mapping \\\hline
+\textbf{Problem}: The extension of a class in one ontology
+corresponds to the extension of a class in another ontology, provided
+that all individuals in the extension have a particular attribute
+value. \\\hline
+\textbf{Solution:}
+
+\textit{Solution description}: A mapping is established
+between a class/attribute/attribute value combination in one ontology
+and a class in another ontology.
+
+\textit{Mapping syntax:}
+
+\texttt{mapping ::= classMapping(direction A
+B attributeValueCondition(P o))} \\\hline
+\textbf{Example:}
+
+\texttt{classMapping(Human Female
+attributeValueCondition(hasGender
+}{\textquotedblleft}female{\textquotedblright})) \\\hline
+\end{supertabular}
+\end{center}
+
 The example in Table \ref{tab:MappingPattern} \cite{debruin2005wsml} shows a mapping element from the
@@
 distinguish the gender.
-
-\begin{center}
-\tablecaption{Class by attribute mapping pattern and example. {\newline}Taken from \cite{debruin2005wsml}.}
-\label{tab:MappingPattern}
-\tablehead{}
-\begin{supertabular}{|m{0.8\linewidth}|}
-\hline
-\textbf{Name}: Class by Attribute Mapping \\\hline
-\textbf{Problem}: The extension of a class in one ontology
-corresponds to the extension of a class in another ontology, provided
-that all individuals in the extension have a particular attribute
-value. \\\hline
-\textbf{Solution:}
-
-\textit{Solution description}: A mapping is established
-between a class/attribute/attribute value combination in one ontology
-and a class in another ontology.
-
-\textit{Mapping syntax:}
-
-\texttt{mapping ::= classMapping(direction A
-B attributeValueCondition(P o))} \\\hline
-\textbf{Example:}
-
-\texttt{classMapping(Human Female
-attributeValueCondition(hasGender
-}{\textquotedblleft}female{\textquotedblright})) \\\hline
-\end{supertabular}
-\end{center}
 

### 2011-02-24 trunk/deliverables/D4_3/fast_d4_3_2.tex
Answered some questions.
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1285 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 on JSON or XML.
-In that case, the body text is again passed and the parameters are included in the request parameter form part.
-Then the user may easily use the request parameter form part to fill in pre-condition placeholders.
+In that case, our tool parses the body text and the user is able to adapt the parameters via the request parameter form part.
+This means, the user may easily use the request parameter form part to fill in pre-condition placeholders.
 
@@
 that's doable, but pretty hard and tedious.
-- Ismael}
+- Ismael
+Well as I understood, in case of WSDL the post body is structured (XML?).
+Thus, we parse that body, identify the key value pairs of the parameters and come up with an input form as we do for URL parameters.
+Easily done.
+I have changed the text above slightly to reflect this better.
+}
 
-\textcolor{red}{Is PUT and DELETE supported? if so, how? - Ismael}
+\textcolor{red}{Is PUT and DELETE supported? if so, how? - Ismael
+Well, it will be supported for the review meeting. Actually Dennis has done a lot of work on this and it is pretty complete. 
+After all a post request looks very much like a get request.
+Yes, the data is provided in some body.
+But, the FAST API, that works as relay for us, provides a method, where you pass an url, the request kind (GET, PUT, DELETE, ...) and the parameters.
+Depending on the kind, this method puts the parameters either in the URL or in some body.
+No difference for us.
+}
 

### 2011-02-24 trunk/deliverables/D4_3/fast_d4_3_2.tex
added a few more comments
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1286 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 - Ismael
-Well as I understood, in case of WSDL the post body is structured (XML?).
+\textcolor{blue}{Well as I understood, in case of WSDL the post body is structured (XML?).
 Thus, we parse that body, identify the key value pairs of the parameters and come up with an input form as we do for URL parameters.
 Easily done.
-I have changed the text above slightly to reflect this better.
-}
+I have changed the text above slightly to reflect this better. - Albert}
+\textcolor{red}{yes and no. it uses SOAP, which relies on XML, but it's something different. the SOAP message will have more info that pre-/post-condition needs. - Ismael}
 
-\textcolor{red}{Is PUT and DELETE supported? if so, how? - Ismael
-Well, it will be supported for the review meeting. Actually Dennis has done a lot of work on this and it is pretty complete. 
+\textcolor{red}{Is PUT and DELETE supported? if so, how? - Ismael}
+\textcolor{blue}{Well, it will be supported for the review meeting. Actually Dennis has done a lot of work on this and it is pretty complete. 
 After all a post request looks very much like a get request.
@@
 Depending on the kind, this method puts the parameters either in the URL or in some body.
-No difference for us.
-}
+No difference for us.}
+\textcolor{red}{ok, grand!}
 

### 2011-02-25 trunk/deliverables/D5_1_2/fast_d5_1_2.tex
added some changes to "planning appendix"
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1288 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-One of the limitations of the pre-/post-conditions matching algorithm is that the output is limited to satisfy a given set of conditions, but the building blocks suggested may include new pre-conditions to be satisfied, where the user has to add building blocks iteratively until the goal is accomplished (screen-flow or screen is executable).
-
-A set of screens can be seen as a graph (see Figure~\ref{fig:screens_graph}), where the vertices (or nodes) are the screens, and the edges represent whether a screen satisfies or is satisfied by another screen (the arrow points towards the screen which gets satisfied).
+One of the limitations of the pre-/post-conditions matching algorithm is that its output is limited to a set of building blocks which satisfy a given set of conditions, but they may include new pre-conditions which are not yet satisfied.
+The consequence is that the user has to add building blocks iteratively until her goal is accomplished (screen-flow is executable).
+
+The approach followed by this algorithm is to find sets of screens, or so-called \emph{plans}, which fully accomplish a certain goal in a one-step action.
+
+From the point of view of this algorithm, the screen knowledge base is seen as a graph, as represented in Figure~\ref{fig:screens_graph}), where the vertices (or nodes) are the screens, and the edges represent whether a screen satisfies or is satisfied by another screen (the arrow points towards the screen which satisfies).
 There is a special node connected to all the screens with no pre-conditions (those are reachable by default).
-For a given screen, a plan is the path from itself to the special node.
-These paths or plans are then offered to the user as a set of screens which makes a specific screen reachable.
 
@@
 	\includegraphics{images/screens_graph}
-	\caption{Graph of compatible screens by pre-/post-conditions}
+	\caption{Graph of compatible screens by their pre-/post-conditions}
 \end{center}
@@
 
-If the number of vertices and edges is high, the number of different paths can be as well high.
-This causes problems such as a huge number of different paths between two vertices, or paths which are incredible long, in other words, the number of results or their quality may not be appropiate to be presented to the user, which may expect a few results and with the minimum number of screens as possible.
-This is a common problem, and there  are many algorithms which solve it, such as Dijkstra or the A* search algorithm which are complete and will always find a solution if one exists.
-
-Our approach has been to implement the A* algorithm, as it calculates a heuristic estimate of the distance to the goal for each node, apart from the path-cost function also included in Dijkstra.
-In our case, the cost for all edges is the same, and the heuristic function is used to prioritize the screens which are being used already by a given canvas.
-This way, we try to minimise the number of screens to be used.
+Therefore, the aim of the algorithm is to find the paths or plans from a given screen to the special node or root.
+This is a very well-known problem in graph theory, and therefore there are many graph and tree search algorithms dealing with it, such as Floyd's, Dijkstra, A* or depth-first search, among others.
+
+Floyd's, Dijkstra and A* are focused on finding shortest paths for a given problem.
+In the case of A*, the advantage is that a heuristic function may be define to prioritise a set of screens (i.e.
+screens in the current canvas).
+However, finding the shortest path is not the intention of this algorithm.
+In fact, it is preferred to find a small set of possible solutions, where depth-first search is a good fit as it finds all the paths from a given node to another, avoiding cycles.
+The paths returned by depth-first are then ranked in order to minimise the path lenght, and to maximise the number of co-occurrences of elements in the path and the canvas.
+
+The depth-first search algorithm has not been re-implemented for the catalogue.
+Instead, it relies on Neo4j\footnote{http://neo4j.org/}, a flexible graph database, which apart from providing the algorithm, it allows the storage and access of the graph.
+The advantages of using Neo4j are that it is highly optimized for storing graph structures for maximum performance and scalability, and supports several search algorithms, including as well Dijkstra and A*, which provides a great flexibility to the catalogue.
 

### 2011-02-25 trunk/deliverables/D4_3/fast_d4_3_2.tex
fixed problem with }
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1287 d8b0c59b-6544-0410-94d9-83a5658a058e
first review done
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1293 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
+\textcolor{red}{Check if this section reflects the structure...}
+
 % subsection structure_of_the_document (end)
@@
 
-\subsection{Discovering} % (fold)
-\label{sub:discovering}
+\subsection{Discovery} % (fold)
+\label{sub:discovery}
 
@@
 
-For this reason, a screen developer during the screen development phase is able to reuse these screen components which are already stored in the Catalogue, establishing which the screen's pre-/post-conditions are, and then the system will suggest several screen components which may satisfy these pre-/post-conditions.
-In a first step, the system will suggest screen components which have any of the pre-conditions which can be connected to any of the screen pre-conditions (the connection will be represented by an internal pipe), and the same would happen for the post-conditions.
-Once a screen component is inserted into the screen, the system will take its definition (screen components, pipes, etc.) into consideration to recommend new screen components which can be connected to the ones the screen is composed of.
-
-For a further explanation on how the recommendation algorithms are implemented, please see~\cite{rivera2011catalogue_user_manual}.
+For this reason, a screen developer during the screen development phase is able to reuse these screen components which are already stored in the Catalogue, establishing which the screen's pre-/post-conditions are.
+These pre-/post-conditions and building blocks are then used to find other relevant screen components which may satisfy these pre-/post-conditions, or be a good fit for the current problem.
+A more detailed explanation on this topic can be found in~\cite{rivera2011catalogue_user_manual}.
 
 % subsection discovery (end)
+
+\subsection{Connection} % (fold)
+\label{sub:connection}
+
+The aim of a resource adapter is to perform some action on a web service.
+In order to perform a given action, a precondition may need to be fulfilled, and as a result of the execution a postcondition may be generated.
+These pre/postconditions are intended to be used by pre/postconditions of other building blocks within a gadget.
+This connection must be explicitly defined, because it will indicate the real data-flow between different building blocks.
+In FAST this is commonly called Piping, since a pipe in computer science is just that, a one-way communication channel for interprocess communication; therefore a pipe is a logical connection made between a precondition and a postcondition of different building blocks.
+Figure~\ref{fig:Piping} shows an example of a search form for the Amazon service.
+This is a screen, composed by a form and a resource adapter.
+The form allows the user to introduce a search term, producing a postcondition which will be redirected to a resource adapter precondition.
+The execution of the web service produces a list of results, which will be forwarded to the postcondition of the screen.
+
+\begin{figure}[!htb]
+  \begin{center}
+    \includegraphics[width=12cm]{images/piping.pdf}
+    \caption{Piping}
+    \label{fig:Piping}
+  \end{center}
+\end{figure}
+
+% subsection connection (end)
 
@@
 
-In FAST, the Service Wrapper Tool will be in charge of the Resource Adapters' building phase.
-The building of resource adapters is done in three steps: in a first step, we define which pre- and post-conditions are used to pass parameters between FAST building blocks and the resource adapter.
-The second step is in charge of the construction of a service request and the third step analyses the response got from the execution of the service, allowing the extraction of facts contained in that response and mapping them to domain-specific classes or concepts from the ontologies used within FAST by the building blocks' conditions.
+In FAST, the Service Wrapper Tool deals with the Resource Adapters' building phase.
+The building of resource adapters is done in three steps: in a first step, the pre- and post-conditions used to pass parameters between FAST building blocks and the resource adapter are defined.
+The second step constructs and parses the request to the Web service, and the third step analyses the response obtained from the execution of the service, allowing the extraction of facts contained in that response and mapping them to domain-specific classes or concepts from the ontologies used by the building blocks' conditions.
+
+This deliverable takes as an example a \emph{Formula1} RESTful Web service in order to explain in an easy manner certain aspects of the Service Wrapper Tool.
 
@@
 
-We will demonstrate the use of the Service Wrapper Tool on the example of a simple \emph{Formula1} service.
-In the first tab of the Service Wrapper Tool, the user defines the name and the pre- and post-conditions of the desired Resource Adapter, cf.
-Fig.
-\ref{fig:construct_pre_post_conditions}.
-For the pre-condition, we provide a name, a type, and an example value.
-The type of a pre- or post-condition is chosen from a drop down list of available concepts.
+In the first tab of the Service Wrapper Tool, the user defines the name and the pre- and post-conditions of the desired Resource Adapter, as illustrated in Figure~\ref{fig:construct_pre_post_conditions}.
+For the pre-condition, a name, a type, and an example value are provided.
+The type of a pre- or post-condition is chosen from a drop-down list of available concepts.
 The list of available concepts, together with its properties and some example data (instances) is retrieved from the Catalogue.
-The example value is used for test requests during subsequent construction steps.
-We keep track of example values provided for different domain concepts and propose them, when the domain concept is re-used e.g.
-in another Resource Adapter.
+The type of the properties or attributes of the concepts may be basic types, such as string, integer, date, etc., or other concepts, providing a good flexibility and support for a any kind of input/output for the Web services.
 
@@
 
+These example values (instances) are used for testing requests during subsequent construction steps, and to assist the user with previous values for pre-/post-conditions which have already being used in another resource adapters.
+
 \subsection{Constructing service requests} % (fold)
@@
 
-In our simple case, the service request consists just of a URL that is easily retrieved from API documentation page of our Formula1 service.
-This URL consists of the service address \url{http://ergast.com/api} followed by the year we are querying for followed by the category \verb|drivers| we are interested in.
+Following with the proposed example, the service request consists of a URL template that is easily retrieved from Formula1 API documentation.
+This URL template consists of the service URL or address \url{http://ergast.com/api} followed by the year to query for, and by the category \verb|drivers| we are interested in.
 The user may grab such an URL from the API documentation page of the targeted web service and drop it into the Request tab of the Service Wrapper Tool, cf.
 Figure~\ref{fig:construct_service_request}.
-The Service Wrapper Tool will than analyze the dropped URL and split it into its parts and provide a convenient form allowing to edit e.g.
+The Service Wrapper Tool will then analyse the dropped URL and split it into its parts and provide a convenient form allowing to edit e.g.
 all the parameters of the URL.
@@
 that's doable, but pretty hard and tedious.
-- Ismael
+- Ismael}
 \textcolor{blue}{Well as I understood, in case of WSDL the post body is structured (XML?).
@@
 
-\subsubsection{Limitations} % (fold)
-\label{ssub:limitations}
-
-It is worth pointing out that currently the Service Wrapper Tool is able to construct input ports for the wrappers using just basic types.
-However, every building block accepts any (complex) concept as a pre-/post-condition, so the wrapper tool needs to be adapted to permit any concept as input or pre-condition, extracting the required text to construct the request.
-This should be easy to achieve before the project review.
-
-\textcolor{red}{I guess this paragraph will be rewritten...because my understanding is that we do support constructing wrappers with ``complex'' pre-/post-conditions and not just basic types.
-- Ismael}
-
-% subsubsection limitations (end)
-
 % subsection constructing_service_requests (end)
@@
 This response message could be serialize in any format, though the most common formats used nowadays are XML or JSON.
-To continue the example started in the previous section, the response of the Formula1 service will is formatted in XML.
-Th Service Wrapper Tool supports both JSON and XML.
+To continue the example started in the previous section, the response of the Formula1 service will is formatted in XML, although the Service Wrapper Tool supports both JSON and XML.
 
@@
 
-Figure~\ref{fig:response_service_execution} shows the data tranformation tab of the wrapper tool.
+Once the service response has been retrieved, it is shown in the transformation tab as an interactive object tree, as it can be seeing on the left side of Figure~\ref{fig:response_service_execution}.
+To construct this interactive object tree, the XML or JSON document is parsed into a DOM, and a simplified tree representation of the given DOM is built up.
+This tree representation of the XML/JSON data is used as an input to construct the transformation rules.
 
@@
 
-Once the service response has been retrieved, it is shown in the transformation tab as an interactive object tree, as it can be seeing on the left side of Figure~\ref{fig:response_service_execution}.
-To construct this interactive object tree, the XML or JSON document is parsed into a DOM, and a simplified tree representation of that DOM is built up.
-This tree representation of the XML/JSON data is used as an input to construct transformation rules.
-
 A transformation rule is used to analyse the XML/JSON data and to generate domain-specific facts from concepts from the ontologies used by the pre-/post-conditions of the different building blocks.
@@
 These XML elements are identified by the tagname of a DOM element from the XML document.
-Second, the type of the rule will be set, taking one of the following values: \emph{createObject}, \emph{fillAttributes} or \emph{dummy}.
+Second, the type of the rule, taking one of the following values: \emph{createObject}, \emph{fillAttributes} or \emph{dummy}.
 And third, the target of the rule specifies a certain concept or attribute, to be created or filled.
@@
 	\item[action \emph{createObject}] specifies the creation of a new fact object.
-The type of that new fact is provided in the third compartment.
+The type of that new fact is provided in the third input box.
 In the example being explained, the root rule searches for XML elements with tagname \emph{DriverTable} and for each such element a \emph{List} fact is created.
@@
 In addition, FAST is semantic-driven, therefore, the service wrapper designer retrieves the domain-specific types from the Catalogue server together with the structure of each type, i.e.
-together with a description of the attributes of each fact.
-Thus, the transformation rule editor is able to provide selection boxes for the target element of the rules.
+a list of the properties or attributes of each concept.
+Therefore, the transformation rule editor is able to provide selection boxes for the target element of the rules.
 For a \emph{createObject} rule, this selection box shows the fact types available for that domain.
 For the \emph{fillAttributes} rules, the selection box shows the attributes of the fact type chosen in the parent rule.
-In addition, we may provide some analysis tool, which will help to guarantee that the facts generated by the transformation rules conform to the fact types defined in the corresponding FAST ontology.
-This helps to ensure that the facts generated by the designed Resource Adapter will be compatible for pre-conditions of subsequent filter steps and or screens.
-
-\textcolor{red}{We may provide some analysis tool? is it implemented? otherwise I don't see it's worth to point this out, or maybe just in a ``future work'' section - Ismael}
+This rule ensures that the attributes of the generated fact match the type of the concept's properties defined in the ontology, and guarantees all mandatory attributes have a mapping rule.
 
@@
 
-Once the wrapping of a service has been defined and tested in the Service Wrapper Tool, it generates an implementation of the desired Resource Adapter in XML, HTML, and JavaScript, ready to be deployed and executed inside a gadget.
-In addition, a description of the new building block is added to the FAST Catalogue.
-Thereby, the new Resource Adapter becomes available for the construction of new screens and screen-flows.
-
-\textcolor{red}{XML, HTML and Javascript?
-I don't get why XML and HTML is generated.
-The service wrapper should be just a javascript encapsulation of the web service, isn't it?
-HTML files should be for forms and XML for specific platform configuration files for deployment?
-- Ismael}
+Once the wrapping of a service has been defined and tested in the Service Wrapper Tool, the user can invoke the creation of the Resource Adapter.
+This process triggers two actions: the generation of the Javascript code which encapsulates the calls to the Web service, and the creation of a new building block in the Catalogue, with its proper metadata.
+Thereby, the new Resource Adapter becomes available for the creation of new screens, and ready to be embedded in a gadget.
 
@@
 
+\textcolor{red}{Wait for Albert's answer regarding this (and the Javascript eval()). and then rewrite this section as something we do, not limitations - Ismael}
+
 % subsection limitations (end)
 
-\subsection{Other approaches} % (fold)
-\label{sub:other_approaches}
+% section restful_web_services_wrapper_tool (end)
+
+\clearpage
+\section{Related Work} % (fold)
+\label{sec:related_work}
 
@@
 XML input and provide a template based generation of textual output. The problem with XSLT is a quite 
-complex syntax which is not adequate for our end users.\textcolor{red}{Well, no more than writing a SOAP message :) - Ismael} It would also have been difficult to provide our end users with direct feedback, as required by FAST's storyboard philosophy.
+complex syntax which is not adequate for our end users.
+It would also have been difficult to provide our end users with direct feedback, as required by FAST's storyboard philosophy.
 Similarly, interactive help e.g.
@@
 for coming up with a simple and easy to use yet powerful approach.
-
-\textcolor{red}{I couldn't really understand how a wrapper for a WSDL is constructed.
-The wrapper tool uses any library (Axis, etc) to parse the WSDL definition and extract the operations with their inputs and outputs?
-If so, it should be mention.
-Or should be the user who reads the WSDL document, checks the endpoint URL for the operation that wants to wrap, and then construct the wrapper as in the RESTful case, writting the SOAP message for the request body?
-- Ismael}
-
-% subsection Other approaches (end)
-
-% section restful_web_services_wrapper_tool (end)
-
-\clearpage
-\section{Related Work} % (fold)
-\label{sec:related_work}
-
-The rule based approach of the service wrapper tool has been inspired by triple graph grammars, cf.
-\cite{conf/wg/Schurr94,JSZ97c}.
-While general triple graph grammars allow to relate complex graph structures with each other, in our case the target of a rule is always a single object or attribute.
-This facilitates the whole mechanism, reasonably.
-
-Write about some related work such as:
-
-Wrapping WSDL-Described Web Services as Moby Services (using SAWSDL)
-\url{http://biomoby.open-bio.org/CVS_CONTENT/moby-live/Java/docs/sawsdlServlet.html}
-
-\textcolor{red}{is this section done? - Ismael}
 

### 2011-02-25 trunk/deliverables/D5_1_2/fast_d5_1_2.tex
adds some changes
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1294 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 		\vfill
-		{\large \sffamily Date: 17/02/2011}\\
+		{\large \sffamily Date: 25/02/2011}\\
 		\vspace{1cm}
@@
  & & Ismael Rivera (NUIG) & \\ \hline
-2.0 & 17/02/2011 & Ciprian Palaghita (Cyntelix) & Release for M36 \\
+2.0 & 25/02/2011 & Ciprian Palaghita (Cyntelix) & Release for M36 \\
  & & Ismael Rivera (NUIG) & \\ \hline
@@
 
-From the point of view of this algorithm, the screen knowledge base is seen as a graph, as represented in Figure~\ref{fig:screens_graph}), where the vertices (or nodes) are the screens, and the edges represent whether a screen satisfies or is satisfied by another screen (the arrow points towards the screen which satisfies).
+From the point of view of this algorithm, the screen knowledge base is seen as a graph, as represented in Figure~\ref{fig:screens_graph}, where the vertices (or nodes) are the screens, and the edges represent whether a screen satisfies or is satisfied by another screen (the arrow points towards the screen which satisfies).
 There is a special node connected to all the screens with no pre-conditions (those are reachable by default).
@@
 However, finding the shortest path is not the intention of this algorithm.
-In fact, it is preferred to find a small set of possible solutions, where depth-first search is a good fit as it finds all the paths from a given node to another, avoiding cycles.
-The paths returned by depth-first are then ranked in order to minimise the path lenght, and to maximise the number of co-occurrences of elements in the path and the canvas.
+In fact, it is preferred to find a small set of possible solutions, thus depth-first search is a good fit as it finds all the paths from a given node to another, avoiding cycles.
+Then, the paths returned by depth-first are then ranked in order to minimise the path lenght, and to maximise the number of co-occurrences of elements in the path and the canvas.
 
 The depth-first search algorithm has not been re-implemented for the catalogue.
-Instead, it relies on Neo4j\footnote{http://neo4j.org/}, a flexible graph database, which apart from providing the algorithm, it allows the storage and access of the graph.
+Instead, it relies on Neo4j\footnote{http://neo4j.org/}, a flexible graph database, which apart from providing the aforesaid algorithm, it provides the necessary mechanism to store, access and query the graph.
 The advantages of using Neo4j are that it is highly optimized for storing graph structures for maximum performance and scalability, and supports several search algorithms, including as well Dijkstra and A*, which provides a great flexibility to the catalogue.

### 2011-02-25 trunk/deliverables/D5_2/fast_d5_2_3.tex
added brief appendix for D5.2.3
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1297 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-02-17 trunk/deliverables/D5_2_2/fast_d5_2_2.tex
@@
 
-\newcommand\deliverableNumber{D5.2.2}
+\newcommand\deliverableNumber{D5.2.3}
 \newcommand\deliverableTitle{Prototype Semantic Catalogue for Screen-Flow Gadgets and Back-End Services}
@@
 % explicit hyphenations
-\hyphenation{RDF-Re-po-si-to-ry}
-\hyphenation{name-space}
+%\hyphenation{RDF-Re-po-si-to-ry}
+%\hyphenation{name-space}
 
@@
 		\vfill
-		{\large \sffamily Date: 19/02/2010}\\
+		{\large \sffamily Date: 25/02/2011}\\
 		\vspace{1cm}
@@
 %\lofoot{\small\textcolor{fast@lightgrey}{Task Force Ontologies}}
-\cofoot{\small{FAST 216048 --- \deliverableTitleShort}}
-\rofoot{\small{\thepage}}
+% \cofoot{\small{FAST 216048 --- \deliverableTitleShort}}
+% \rofoot{\small{\thepage}}
+\cofoot{\small{FAST $\bullet$ 216048 $\bullet$ \deliverableTitleShort{} $\bullet$ Page \thepage\ of \pageref{LastPage}}}
 
@@
 2.0 & 19/02/2010 & Ismael Rivera (NUIG) & Second version of the deliverable \\ \hline
+3.0 & 25/02/2011 & Ismael Rivera (NUIG) & Third version of the deliverable \\ \hline
 \end{tabular}
@@
 	\hline
-	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {27/02/2010} & {\textbf{Submitted}} & {27/02/2010}\\ \hline
+	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {27/02/2010} & {\textbf{Submitted}} & {26/02/2010}\\ \hline
 	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
@@
 Because D5.2 is a pure software deliverable, the sole role of this document is to function as a starting point for further information about the catalogue.
-Nevertheless, the deliverable D5.1 is aimed to fulfil this purpose, so for any further information such as where the code of the catalogue can be found, installation instructions, system requirements, a overview of the catalogue architecture or a detailed API, please read \cite{palaguita2010_usermanual}.
+Nevertheless, the deliverable D5.1 is aimed to fulfil this purpose, so for any further information such as where the code of the catalogue can be found, installation instructions, system requirements, a overview of the catalogue architecture or a detailed API, please read \cite{palaguita2011_usermanual}.
 
@@
 \bibliographystyle{apalike}
-\bibliography{fast_d5_2_2}
+\bibliography{fast_d5_2}
 

### 2011-02-25 trunk/deliverables/D4_3/fast_d4_3_2.tex
Added new screen shot for WSDL service and adapted the description.
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1299 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 In our example, we use a simple HTTP GET request.
-The Service Wrapper Tool also supports the other kinds of HTTP methods as POST, PUT or DELETE.
+The Service Wrapper Tool also supports the other kinds of HTTP methods as POST request.
+We might support PUT and DELETE requests in the future, too.
 For POST requests, one usually passes parameters within a special post body.
 For unstructured POST bodies, we provide a text area allowing the user to fill in placeholders for references to pre-condition values.
-In case of structured POST bodies as in the case of WSDL and WSMO services the POST body uses a structured text format based e.g.
-on JSON or XML.
-In that case, our tool parses the body text and the user is able to adapt the parameters via the request parameter form part.
-This means, the user may easily use the request parameter form part to fill in pre-condition placeholders.
+In case of structured POST bodies as in the case of WSDL and WSMO services the POST body uses a structured text format.
+In Figure~\ref{fig:construct_service_request_wsdl} the user has visited the semantic description of a WSDL service for a geoip service.
+This pops up a window with a drop down box with a list of operations offered by that service.
+In our example the user chooses the GetGeoIP operation.
+This allows our tool to retrieve the list of parameters for the GetGEoIP operation and to prefill the post body text area shown in the right center of the screen with parameter settings.
+In our example only one paramter with name IPAddress is required.
+Then the user might connect parameters to preconditions using \verb|<IPAddress>| as the value for the parameter.
+Similar support for WSMO service will be provided in the future.
 
-\textcolor{red}{WSDL is usually used with SOAP and XML, JSON wouldn't be common in this case.
+%\textcolor{red}{WSDL is usually used with SOAP and XML, JSON wouldn't be common in this case.
 And from my understanding, the user has to hand-write the POST body?
@@
 - Ismael}
-\textcolor{blue}{Well as I understood, in case of WSDL the post body is structured (XML?).
-Thus, we parse that body, identify the key value pairs of the parameters and come up with an input form as we do for URL parameters.
+%\textcolor{blue}{Well as I understood, in case of WSDL the post body is structured (XML?).
+Thus, we parse that body, identify %the key value pairs of the parameters and come up with an input form as we do for URL parameters.
 Easily done.
-I have changed the text above slightly to reflect this better. - Albert}
-\textcolor{red}{yes and no. it uses SOAP, which relies on XML, but it's something different. the SOAP message will have more info that pre-/post-condition needs. - Ismael}
-
-\textcolor{red}{Is PUT and DELETE supported? if so, how? - Ismael}
-\textcolor{blue}{Well, it will be supported for the review meeting. Actually Dennis has done a lot of work on this and it is pretty complete. 
-After all a post request looks very much like a get request.
+%I have changed the text above slightly to reflect this better. - Albert}
+%\textcolor{red}{yes and no. it uses SOAP, which relies on XML, but it's something different. the SOAP message will have more info that pre-/post-condition needs. - Ismael}
+%\textcolor{red}{Is PUT and DELETE supported? if so, how? - Ismael}
+%\textcolor{blue}{Well, it will be supported for the review meeting. Actually Dennis has done a lot of work on this and it is pretty complete. 
+%After all a post request looks very much like a get request.
 Yes, the data is provided in some body.
@@
 No difference for us.}
-\textcolor{red}{ok, grand!}
+%\textcolor{red}{ok, grand!}
 
@@
     \label{fig:construct_service_request}
+  \end{center}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSRequestTemplateWSDL.png}
+    \caption{Importing WSDL service descriptions}
+    \label{fig:construct_service_request_wsdl}
   \end{center}

### 2011-02-25 trunk/deliverables/D5_1_2/fast_d5_1_2.tex
fixes a typo
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1300 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-This sections enumerates and gives a brief overview of the changes made in the deliverable with regards to the version 1.0 released for the milestone M24.
+This section enumerates and gives a brief overview of the changes made in the deliverable with regards to the version 1.0 released for the milestone M24.
 

### 2011-02-25 trunk/ServiceScreenDesignerWep/gensrc/fast/facttool/client/DeleteAttributeButton.java
Added license note
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1301 d8b0c59b-6544-0410-94d9-83a5658a058e

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 /*
  * GWT compliant generated by Fujaba - CodeGen2 version 03.12.2009
  */

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/mediation/client/DataTransformationTool.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.mediation.client;
 
 import java.util.Iterator;

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/mediation/client/gui/MediationRuleGUI.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.mediation.client.gui;
 
 import java.util.Iterator;

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/client/FastTool.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.servicescreen.client;
 
 import fast.servicescreen.client.gui.codegen_js.CodeGenViewer;

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/client/RequestServiceAsync.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.servicescreen.client;
 
 import java.util.HashMap;

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/client/ServiceScreenDesignerWep.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.servicescreen.client;
 
 import java.util.Iterator;

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/client/URL_Settings.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.servicescreen.client;
 
 /**

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/client/gui/RequestGUI.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.servicescreen.client.gui;
 
 import com.google.gwt.event.dom.client.ClickEvent;

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/client/gui/RuleGUI.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.servicescreen.client.gui;
 
 import java.util.ArrayList;

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/client/gui/TabWidget.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.servicescreen.client.gui;
 
 import com.google.gwt.user.client.ui.HTML;

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/client/gui/TemplateGUI.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.servicescreen.client.gui;
 
 import java.util.Iterator;

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/client/gui/codegen_js/CodeGenViewer.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.servicescreen.client.gui.codegen_js;
 
 import com.google.gwt.core.client.GWT;

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/client/gui/codegen_js/CodeGenerator_JSON.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.servicescreen.client.gui.codegen_js;
 
 import java.util.HashMap;

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/client/gui/codegen_js/CodeGenerator_reqJSON.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.servicescreen.client.gui.codegen_js;
 
 import java.util.ArrayList;

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/client/gui/parser/ExtendedRuleParser.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.servicescreen.client.gui.parser;
 
 import java.util.LinkedHashMap;

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/client/gui/parser/Kind.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.servicescreen.client.gui.parser;  
 
 /**

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/client/gui/parser/OperationHandler.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.servicescreen.client.gui.parser;
 
 import java.util.ArrayList;

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/client/gui/parser/Token.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.servicescreen.client.gui.parser;
 
 /**

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/client/rpc/WSDLHandler.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.servicescreen.client.rpc;
 
 import java.util.HashMap;

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/server/RequestServiceImpl.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.servicescreen.server;
 
 import java.io.File;

### 2011-02-25 trunk/ServiceScreenDesignerWep/src/fast/servicescreen/server/RequestServlet.java

@@ -1,3 +1,25 @@
+/**
+ * Copyright (c) 2008-2011, FAST Consortium
+ * 
+ * This file is part of FAST Platform.
+ * 
+ * FAST Platform is free software: you can redistribute it and/or modify
+ * it under the terms of the GNU Affero General Public License as published by
+ * the Free Software Foundation, either version 3 of the License, or
+ * (at your option) any later version.
+ * 
+ * FAST Platform is distributed in the hope that it will be useful, but
+ * WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY
+ * or FITNESS FOR A PARTICULAR PURPOSE. See the GNU Affero General Public
+ * License for more details.
+ * 
+ * You should have received a copy of the GNU Affero General Public License
+ * along with FAST Platform. If not, see <http://www.gnu.org/licenses/>.
+ * 
+ * Info about members and contributors of the FAST Consortium
+ * is available at http://fast.morfeo-project.eu
+ *
+ **/
 package fast.servicescreen.server;
 
 import java.io.IOException;

### 2011-02-25 trunk/deliverables/D4_3/fast_d4_3_2.tex
final review done. ready to send
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1302 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2011-02-25 trunk/deliverables/D4_3/fast_d4_3_2.tex
@@
 \newcommand\authorTwo{Albert Zündorf (UniKas)}
-\newcommand\authorThree{Knud M\"oller (NUIG)}
+\newcommand\authorThree{Author 3}
 \newcommand\authorFour{Author 4}
@@
 		{\large \sffamily \authorTwo}\\
-		{\large \sffamily \authorThree}\\
 		\vspace{0.5cm}
 		\vfill
-		{\large \sffamily Date: 27/02/2010}\\
+		{\large \sffamily Date: 25/02/2010}\\
 		\vspace{1cm}
@@
 							\textcolor{white}{\textbf{Change description}}\\ \hline
-%0.1 & 15.12.2008 & Knud M\"oller (NUIG) & template modelled \\ \hline
-%0.2 & 15.02.2009 & Knud M\"oller (NUIG) & content moved from wiki after internal review \\ \hline
 1.0 & 26.02.2010 & Ismael Rivera (NUIG), & Final version ready for external review \\ 
@@
 2.0 & 27.02.2011 & Ismael Rivera (NUIG), & Final version ready for external review \\ 
- & & Albert Zündorf (UniKas) & \\
- & & Knud M\"oller (NUIG) & \\ \hline
+ & & Albert Zündorf (UniKas) & \\ \hline
 \end{tabular}
@@
 	\hline
-	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2011} & {\textbf{Submitted}} & {27/02/2011}\\ \hline
+	\textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2011} & {\textbf{Submitted}} & {25/02/2011}\\ \hline
 	\textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
 	\textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
-	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D4.3.2 is...}\\ \hline
-	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne, \authorTwo, \authorThree}\\
+	\textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{This deliverable presents techniques to exploit different kind of web services such as RESTful, SOAP-based and WSMO services.
+It gives an overview of the characteristics of each type, and how they match into the FAST philosophy, where the Resource Adapter encapsulates they functionality and make possible their execution.
+Then, the Service Wrapper Tool, which creates the resource adapters, is described, and how it addresses the different kind of supported web services.}\\ \hline
+	\textcolor{white}{\textbf{Authors}} & \multicolumn{4}{l|}{\authorOne, \authorTwo}\\
 %	{} & \multicolumn{4}{l|}{}\\ 
@@
 	\textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {ismael.rivera@deri.org} \\ \cline{4-5}
-	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 ?????} \\ \hline
-	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, web services, WSDL, REST, SOAP, WSMO}\\ \hline
+	\textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495338} \\ \hline
+	\textcolor{white}{\textbf{Keywords}} & \multicolumn{4}{p{10.85cm}|}{FAST, web service, WSDL, REST, SOAP, WSMO, Resource Adapter}\\ \hline
 \end{tabular}
@@
 Section~\ref{sec:introduction} states the goal, scope and structure of the document.
-Section~\ref{sec:web_services_wrapping} describes what a service wrapper is, how the platform is able to interact with web services, how these wrappers are build and then discovered to be reused in any gadget.
-Finally, Section~\ref{sec:restful_web_services_wrapper_tool} defines the service wrapper tool built to facilitate FAST users to create Resource Adapters of RESTful web services.
-
-\textcolor{red}{Check if this section reflects the structure...}
+Section~\ref{sec:web_services_wrapping} describes what a service wrapper is, how the platform is able to interact with web services, and how these wrappers are built and discovered to be reused in the gadgets.
+Then, Section~\ref{sec:restful_web_services_wrapper_tool} defines the Service Wrapper Tool, built to facilitate FAST users to create Resource Adapters of different kind of web services.
+Finally, Section~\ref{sec:related_work} presents similar approaches and related work that have been investigated and re-used in a certain level.
 
@@
 
-\textcolor{red}{To be done...}
+This section enumerates and gives a brief overview of the changes made in the deliverable with regards to the version 1.0 released for the milestone M24.
+
+\begin{itemize}
+ \item The related work section has been renamed to ``Similar Approaches and Related Work'', and includes new approaches dealing with the topics covered by the Service Wrapper Tool, and how it re-uses concepts and techniques of some of them.
+ \item ``Resources Adapters'' sections has been renamed to ``Wrapping Services'', and small changes have been made to this section.
+ \item ``RESTful Web Service Wrapper Tool'' has been renamed to ``Service Wrapper Tool'' as the tool has evolved and now it supports the wrapping of WSDL and WSMO services, as well as RESTful services.
+Hence, this section describes how these web services are supported and how to construct them.
+\end{itemize}
 
@@
 Now we have to connect the request parameters to the parameter values passed via the pre-conditions to the Resource Adapter.
-In our case, we replace the year part of the URL with a placeholder for the pre-condition season \verb|<season>|, see Request URL input field at the top of Figure~\ref{fig:construct_service_request}.
+In our case, we replace the year part of the URL with a placeholder for the pre-condition season \verb|<season>|, see Request URL input field at the top of Figure~\ref{fig:construct_service_request}, which shows the screen to construct the service requests.
 
-In our example, we use a simple HTTP GET request.
-The Service Wrapper Tool also supports the other kinds of HTTP methods as POST, PUT or DELETE.
-For POST requests, one usually passes parameters within a special post body.
-For unstructured POST bodies, we provide a text area allowing the user to fill in placeholders for references to pre-condition values.
-In case of structured POST bodies as in the case of WSDL and WSMO services the POST body uses a structured text format based e.g.
-on JSON or XML.
-In that case, our tool parses the body text and the user is able to adapt the parameters via the request parameter form part.
-This means, the user may easily use the request parameter form part to fill in pre-condition placeholders.
-
-\textcolor{red}{WSDL is usually used with SOAP and XML, JSON wouldn't be common in this case.
-And from my understanding, the user has to hand-write the POST body?
-therefore the SOAP message?
-that's doable, but pretty hard and tedious.
-- Ismael}
-\textcolor{blue}{Well as I understood, in case of WSDL the post body is structured (XML?).
-Thus, we parse that body, identify the key value pairs of the parameters and come up with an input form as we do for URL parameters.
-Easily done.
-I have changed the text above slightly to reflect this better. - Albert}
-\textcolor{red}{yes and no. it uses SOAP, which relies on XML, but it's something different. the SOAP message will have more info that pre-/post-condition needs. - Ismael}
-
-\textcolor{red}{Is PUT and DELETE supported? if so, how? - Ismael}
-\textcolor{blue}{Well, it will be supported for the review meeting. Actually Dennis has done a lot of work on this and it is pretty complete. 
-After all a post request looks very much like a get request.
-Yes, the data is provided in some body.
-But, the FAST API, that works as relay for us, provides a method, where you pass an url, the request kind (GET, PUT, DELETE, ...) and the parameters.
-Depending on the kind, this method puts the parameters either in the URL or in some body.
-No difference for us.}
-\textcolor{red}{ok, grand!}
-
-Figure~\ref{fig:construct_service_request} shows the screen to construct the service requests.
+In this example, an HTTP GET request has to be constructed.
+The Service Wrapper Tool supports HTTP POST requests, and its design makes possible the addition of other HTTP verbs as HEAD, PUT or DELETE, among others.
+While in GET request the parameters are set in the URL itself, POST requests may include a body of data sent after the header lines.
+For unstructured POST bodies, a text area is provided allowing the user to fill in placeholders for references to pre-condition values.
+However, in structured POST bodies as in the case of WSDL and WSMO services, the POST body uses a structured text format.
+In Figure~\ref{fig:construct_service_request_wsdl} the user has loaded a WSDL definition of a \emph{geoip} web service into the tool.
+Then, a modal window shows a drop-down list with the different operations offered by the service.
+After the user selects an operation, in this example the GetGeoIP operation, the service wrapper tool retrieves the list of parameters for the GetGEoIP operation and construct the template of the POST body which is inserted in its specific textarea.
+The GetGeoIP operations accepts one parameter: IPAddress.
+Then the user is allow to connect parameters to preconditions using \verb|<IPAddress>| as the value for the parameter.
 
@@
 
-Below the request parameter editing form of Figure~\ref{fig:construct_service_request}, a \textit{Send Request} button allows to validate the constructed request by sending it to the specified service address (via a server relay).
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\linewidth]{images/ServiceWrapperToolGVSRequestTemplateWSDL.png}
+    \caption{Importing WSDL service descriptions}
+    \label{fig:construct_service_request_wsdl}
+  \end{center}
+\end{figure}
+
+Below the request parameter editing form of Figure~\ref{fig:construct_service_request}, a \emph{Send Request} button allows to validate the constructed request by sending it to the specified service address.
 Then the placeholders for pre-conditions are replaced with example values and the resulting request is shown below the parameter form.
-In addition, the request is sent and the response is shown on the bottom of that page.
+In addition, the request is sent and the response is shown at the bottom of interface.
 This gives the user a fast feedback whether the constructed request works as desired.
@@
       \includegraphics[angle=90,width=0.8\linewidth]{images/ServiceWrapperToolGVSWithTransformationRulesF1.png}
-    \caption{Interactive, rule based transforming of an XML response to FAST facts}
+    \caption{Interactive, rule based transformation of a XML response into FAST facts}
     \label{fig:response_service_execution}
@@
 
-A transformation rule is used to analyse the XML/JSON data and to generate domain-specific facts from concepts from the ontologies used by the pre-/post-conditions of the different building blocks.
+A transformation rule approach is used to analyse the XML/JSON data and to generate domain-specific facts from concepts of the ontologies used by the pre-/post-conditions of the different building blocks.
 A transformation rule is composed of three elements, as the middle part of Figure~\ref{fig:response_service_execution} shows.
@@
 
+This rule driven approach is highly intuitive and eases the mapping process from the services responses to the FAST facts.
+It allows mapping basic types attributes, but also the creation of nested facts, such as having a fact \emph{Address} inside a fact \emph{Person}.
+Besides, it supports special operations, as attribute concatenation, in order to mediate with certain problems, for example, the fact Person has an attribute \emph{fullname} which is composed by the \emph{firstname} and \emph{lastname} elements of the response.
+This is done by setting \emph{lastname"', "`firstname} into the \emph{from} field of the transformation rule.
+Another special operation is the access to inner elements in the XML DOM.
+For example, an hypothetical attribute \emph{grandmother} could be filled using the rule \emph{mother.mother}.
+
+However, not every imaginable problem can be solved by a transformation rule.
+Mathematical operations, as converting temperatures from Fahrenheit into Celsius, or similar are not supported.
+However, complex cases as these may be solved by the use or creation of another kind of building block: the operator.
+More information about operators and the Data Mediation Tool can be found in~\cite{ambrus2011_mediation}.
+
 Since FAST is storyboard oriented, the Service Wrapper Tool follows the storyboard idea as well.
@@
 
-\subsection{Limitations} % (fold)
-\label{sub:limitations}
-
-The rule driven approach presented above is somewhat limited.
-It is deliberately restricted to such a simple rule mechanism in order to keep things simple enough for end users.
-However, the selected approach suffices for most practical and real world cases.
-As a more complex example, the XML data for a person may provide two different tags for the first and the last name of a person.
-Contrarily, a person fact which conforms to a certain ontology for that domain may provide only one \emph{fullname} attribute that shall be filled by a concatenation of the first and the last name.
-To achieve this, the \textit{from} field of that tranforamtion rule might look like: \texttt{lastname"', "`firstname}.
-We are also able to do some navigation in the XML tree to follow XRef elements.
-For example the attribute \texttt{grandmother} could be filled using \texttt{mother.mother} in the \textit{from} field.
-\textcolor{red}{This paragraph doesn't talk about limitations, these are features! - Ismael}
-
-However, there are some transformations that these rules cannot perform.
-For example, we do not support any mathematical operations.
-Thus, transforming e.g.
-Fahrenheit into Celsius temperatures is not supported.
-To cover such cases, intermediate fact formats can be used which would allow generating facts to be further processed by additional filters.
-Such additional filters may be realized using hard-coded operators, since some generic operators can act as filters for aggregation and conversions of facts from multiple sources.
-Then, service wrappers in combination with these filter operators will allow covering these complex cases.
-
-\textcolor{red}{Wait for Albert's answer regarding this (and the Javascript eval()). and then rewrite this section as something we do, not limitations - Ismael}
-
-% subsection limitations (end)
-
 % section restful_web_services_wrapper_tool (end)
@@
 \clearpage
-\section{Related Work} % (fold)
+\section{Similar Approaches and Related Work} % (fold)
 \label{sec:related_work}
@@
 \doublespacing
-\section*{Appendix A (Lists of Tables and Figures)}
-\addcontentsline{toc}{section}{Appendix A (Lists of Tables and Figures)}
+\section*{Appendix A (Lists of Figures)}
+\addcontentsline{toc}{section}{Appendix A (Lists of Figures)}
 
-\listoftables
+%\listoftables
 

### 2011-02-26 trunk/deliverables/D2_4_Mediation/fast_d2_4_3_mediation.tex
new files for version 2.4.3
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1282 d8b0c59b-6544-0410-94d9-83a5658a058e
getting final version done
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1304 d8b0c59b-6544-0410-94d9-83a5658a058e

--- 2010-02-23 trunk/deliverables/D2_4_Mediation/fast_d2_4_2_mediation.tex
@@
 \usepackage{setspace}
-\usepackage{graphicx}
 \usepackage[pdfborder={0 0 0}]{hyperref}        % turn on when latex is used (not miktec)
@@
 \usepackage{colortbl}
-\usepackage{longtable}
 \usepackage{array}
@@
 \usepackage{tikz}
-\usepackage{paralist}
 \usetikzlibrary{plotmarks}
@@
 
-\newcommand\deliverableNumber{D2.4.2}
+\newcommand\deliverableNumber{D2.4.3}
 \newcommand\deliverableTitle{Mediation amongst ontologies: Application to the FAST platform}
@@
 \newcommand\workpackageTitle{Definition of Conceptual Model}
-\newcommand\authorOne{Oszk\'{a}r Ambrus (NUIG)}
-\newcommand\authorTwo{Albert Z\"{u}ndorf (UniKassel)}
-\newcommand\authorThree{J\"{o}rn Friedrich Dreyer (UniKassel)}
-\newcommand\authorFour{Sebasti\'{a}n Ortega (UPM)}
+\newcommand\authorOne{Oszkar Ambrus (NUIG)}
+\newcommand\authorTwo{Knud M\"{o}ller (NUIG)}
+\newcommand\authorThree{Albert Z\"{u}ndorf (UniKassel)}
+\newcommand\authorFour{J\"{o}rn Friedrich Dreyer (UniKassel)}
+\newcommand\authorFive{Sebasti\'{a}n Ortega (UPM)}
 \newtheorem{example}{\emph{Example}}
@@
 
-\pagenumbering{roman}
+%\pagenumbering{roman}
+\pagenumbering{arabic}
 
@@
         {\large \sffamily \authorFour} \\
+        {\large \sffamily \authorFive} \\
         \vspace{0.5cm}
         \vfill
-        {\large \sffamily Date: 27/02/2010}\\
+        {\large \sffamily Date: 27/02/2011}\\
         \vspace{1cm}
@@
 %\lofoot{\small\textcolor{fast@lightgrey}{Task Force Ontologies}}
-\cofoot{\small{FAST 216048 --- \deliverableTitleShort}}
-\rofoot{\small{\thepage}}
+\cofoot{\small{FAST $\bullet$ 216048 $\bullet$ \deliverableTitleShort\ $\bullet$ Page \thepage\ of \pageref{LastPage}}}
+%\rofoot{\small{}}
+%\rofoot{\small{\thepage}}
 
@@
 2.0 & 27.02.2010 & \authorOne & final version (D2.4.2) ready for external review \\ \hline
+3.0 & 27.02.2011 & \authorTwo & final version (D2.4.3) ready for external review \\ \hline
 \end{tabular}
@@
     \hline
-    \textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2010} & {\textbf{Submitted}} & {27/02/2010}\\ \hline
+    \textcolor{white}{\textbf{Delivery data}} & {\textbf{Due date}} & {28/02/2011} & {\textbf{Submitted}} & {27/02/2011}\\ \hline
     \textcolor{white}{\textbf{Status}} & \multicolumn{2}{l|}{} & \multicolumn{2}{l|}{final}\\ \hline
     \textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
-    \textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.4.2 is the second iteration of the FAST ontology matching approach.
-The first iteration contained a general overview of the literature regarding ontology mediation, paving the way for futher analysis of the scenarios arising in FAST that require ontology mediation.
-Based on that general overview, this current interation presents the approach to implement ontology matching within FAST.
+    \textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.4.3 is the third iteration of the FAST ontology matching approach.
+The first iteration contained a general overview of the literature regarding ontology mediation, paving the way for further analysis of the scenarios arising in FAST that require ontology mediation.
+Based on that general overview, this current iteration presents the approach to implement ontology matching within FAST.
 We present the different roles within the platform, tasks and challenges related to them, and the way ontology matching is integrated within each role.}\\ \hline
-    \textcolor{white}{\textbf{Authors}} & \multicolumn{4}{p{10.85cm}|}{\authorOne,  \authorTwo, \authorThree, \authorFour}\\
+    \textcolor{white}{\textbf{Authors}} & \multicolumn{4}{p{10.85cm}|}{\authorOne,  \authorTwo, \authorThree, \authorFour, \authorFive}\\
 %   {} & \multicolumn{4}{l|}{}\\ 
@@
     \hline
-    \textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {oszkar.ambrus@deri.org} \\ \cline{4-5}
+    \textcolor{white}{\textbf{Deliverable Owner}} & \multicolumn{2}{l|}{\authorOne} & \textbf{email} & {knud.moeller@deri.org} \\ \cline{4-5}
     \textcolor{white}{\textbf{(Partner)}} & \multicolumn{2}{l|}{} & \textbf{phone} & {+353 91 495112} \\ \hline
@@
 \tableofcontents
+\clearpage
+\section*{Lists of Tables and Figures}
+\listoftables
+\listoffigures
 \cleardoublepage
-\pagenumbering{arabic}
+%\pagenumbering{arabic}
 
@@
 \label{sec:intro}
-\subsection{Goal and Scope.}
-FAST is a visual programming platform allowing business users to build enterprise-class mashups, employing various underlying services and generating new ones.
+\subsection{Goal and Scope}
+FAST is a visual programming platform allowing non-technical end users to build situational mashups, employing various underlying services and generating new ones.
 Resources in FAST are described semantically using different ontologies and vocabularies, and can therefore be combined to what we can call ``intelligent gadgets'', that rely on semantically enriched building blocks.
@@
 
-We aim to integrate ontology matching into the FAST platform to allow for discrepant systems to intercommunicate.
+We aim to integrate a form of ontology matching into the FAST platform to allow for discrepant systems to intercommunicate.
 We try to find a solution to these differences which range from simple naming differences to more complex data heterogeneity, behavioural differences, discrepant paradigms and other disparities.
@@
 
-\subsubsection{Changes to Previous Version}
+\subsubsection{Changes from Previous Version}
 
-The first iteration, dealing with the theoretical foundations of ontology matching, is now contained in the Appendix, and the whole body of this document is newly added content, based on the comments of the last review.
-As such, it tackles design and development issues for integrating ontology matching within the FAST platform: it presents the overall strategy and deals with specific problems belonging to each phase of ontology matching in FAST.
+The following changes have been made with respect to the previous version of this deliverable, D2.4.2.~\cite{ambrus2010mediation}. A history of previous changes is kept in App.~\ref{sec:old_changes}.
 
-\subsection{Structure of the Document.}
+\begin{itemize}
+	\item In response to the previous review, we have extended and improved the text in order to better position ontology matching in FAST throughout the document.
+	\item The previous D2.4.2 introduced the \emph{fact transformation tool}.
+At the time, this tool was an early stand-alone prototype.
+In the meantime, the fact tool has been properly integrated with the FAST tool chain.
+These improvements are reflected in both text and screenshots in Sects.~\ref{sec:fact_tool} and \ref{sec:creating_the_operator}.
+New content has been added to show how the output of the tool is used in screen development.
+	\item The document as a whole has been revised, and minor changes (mostly, spelling, grammar and small updates to reflect latest developments) have been made throughout the deliverable.
+\end{itemize}
+
+\subsection{Structure of the Document}
 This document describes the approach of integrating ontology matching into the FAST environment.
 Sect.~\ref{sec:overview} gives the big picture of ontology matching in fast, which the remaining three sections expand.
-Thus, Sect.~\ref{sec:ontologyengineering} describes a tool to automatically discover alignments, a scenario and example ontologies to demonstrate the capabilites of the tool, and the testing procedure with the results.
-The remaing part of the section deals with a tool to define the discovered mapping rules in a way that is consistent with other similar tasks within the FAST platform, and describes the way in which matching operators are stored within the Semantic Catalogue.
+Thus, Sect.~\ref{sec:ontologyengineering} describes a tool to automatically discover alignments, a scenario and example ontologies to demonstrate the capabilities of the tool, and the testing procedure with the results.
+The remaining part of the section deals with a tool to define the discovered mapping rules in a way that is consistent with other similar tasks within the FAST platform, and describes the way in which matching operators are stored within the Semantic Catalogue.
 Sect.~\ref{sec:gadgetbuilding} shortly describes the way the matching operator is used when composing screenflows and Sect.~\ref{sec:runtime} describes how the matching operator and the defined rules are converted and used in the final, deployed gadget.
@@
 
-\include{fast_d2_4_2_part1}
+\include{fast_d2_4_3_part1}
 
@@
 
-For the last and final iteration of deliverable D2.4 we plan to evaluate real-life scenarios and finalize the implementation of ontology matching for the FAST platform.
+For the last and final iteration of deliverable D2.4 we plan to evaluate real-life scenarios and finalise the implementation of ontology matching for the FAST platform.
 
@@
 
-\include{fast_d2_4_2_part2}
+\include{fast_d2_4_3_part2}
 
 \clearpage
-\doublespacing
-\appendix
-\section{Lists of Tables and Figures}
-%\addcontentsline{toc}{section}{Appendix A (Lists of Tables and Figures)}
+\phantomsection
+\section{Past Changes} % (fold)
+\label{sec:old_changes}
 
-\listoftables
-\listoffigures
+For keeping a record of past changes, this section simply lists the changes that were made between the previous two versions of this deliverable, D2.4.1~\cite{ambrus2009mediation} and D2.4.2.~\cite{ambrus2010mediation}:
+
+The first iteration, dealing with the theoretical foundations of ontology matching, had been moved to App.~\ref{sec:background}.
+The preceding main body of D2.4.2 document is newly added content, based on the comments of the last review (M24).
+As such, it tackles design and development issues for integrating ontology matching within the FAST platform: it presents the overall strategy and deals with specific problems belonging to each phase of ontology matching in FAST.
 

### 2011-02-26 trunk/deliverables/D2_4_Mediation/fast_d2_4_3_part1.tex
New screen dumps
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1284 d8b0c59b-6544-0410-94d9-83a5658a058e
getting final version done
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1304 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
+%!TEX root = fast_d2_4_3_mediation.tex
+
 \lstset{
@@
 
-% \begin{abstract}
-% The FAST environment allows users to graphically compose intelligent, i.e.
-semantically annotated gadgets from predefined building blocks and deploy them on various mashup platforms, thus enabling the interconnection of different systems and services.
-Since different parties involved in the creation of gadgets and building blocks conceptualise knowledge in different ontologies, using ontology matching as a means to reconcile differences in the various conceptualisations is a crucial issue.
-% This paper discusses first steps in our effort to integrate ontology matching in an end-user-oriented environment such as FAST, using an e-commerce-related scenario.
-We evaluate a number of tools and approaches for solving different levels of complexity in ontology matching and
-% %, which we address in this paper.
-We present a scenario of interconnecting e-commerce web services within a FAST gadget, the ontologies created to represent those services and the alignment tool and language used to generate and represent the ontology mappings.
-The testing aims firstly, to find out whether and how state-of-the-art ontology matching makes it possible to mash-up different systems and secondly, to
-% define the direction of integrating ontology matching into the FAST environment.
-% 
-% \keywords{Ontology Matching, User-oriented, Mashups, E-commerce}
-% \end{abstract}
-
 \section{Overview}
 \label{sec:overview}
+
+In this section, we introduce some general terminology and definitions regarding ontology matching, as well as positioning ontology matching in FAST in terms of user roles.
 
@@
 
-FAST uses ontologies to conceptualise the underlying resources used by the different components.
+FAST uses ontologies to conceptualise resources and data used by the different components.
 Ontologies embody the fundamental vehicle for conceptualising data on semantic systems; they describe the context and semantic background of data that should be known to all agents using it~\cite{gruber93towards}.
@@
 This is also true for FAST, where gadget building blocks can originate from different providers, who might use different ontologies to describe them.
-To ensure interoperability, the task of ontology matching is therefore critical in FAST.
+To ensure interoperability, the task of ontology matching is therefore important in FAST.
 
@@
 
-The gadget life cycle in FAST has several phases and roles associated, as detailed in \cite{hoyer-fast}.
+As detailed in \cite{hoyer2009fast}, the gadget life cycle in FAST has several phases and roles associated.
 Here, we list the ones relevant for the ontology matching tasks, in decreasing order of the measure in which knowledge about ontologies is required.
@@
 The \textit{resource developer} then uses these ontologies to annotate resources created in FAST, which resources will be used by the \textit{screen developer}, who will also specify the resource type at the pre- and postconditions of a screen.
+These three roles are the most critical in FAST with respect to ontology matching, and topics relevant to these roles will be discussed in detail in Sect.~\ref{sec:ontologyengineering}.
     \item Ontology matching is needed by the \textit{gadget developer} at the design-time of a screen-flow (gadget).
@@
 %the FAST user interface determines if two resources can be combined or not (see Figure \ref{fig:screens} for a conceptual example of resources that can not be combined).
+Section~\ref{sec:gadgetbuilding} outlines how the gadget developer if affected by ontology matching.
     %\item The \emph{gadget developer} combines screens to screen-flows and gadgets, and only uses ontology matching implicitly.
@@
 Only at run-time the actual mapping of instance data has to be performed.
+This aspect of ontology matching is briefly discussed in Sect.~\ref{sec:runtime}.
 \end{enumerate}
@@
 \label{sec:ontologyengineering}
+
+In the FAST lifecycle, the ontology engineering phase is the most relevant for the topic of this deliverable.
+Relevant roles are the ontology engineer, who performs the main work in this phase, as well as the the resource developer and the screen developer, who are directly affected.
 
@@
 
-\subsection{Facts Transformation Tool}
+\subsection{Fact Transformation Tool}
+\label{sec:fact_tool}
 
@@
     \centering
-    \includegraphics[width=0.8\linewidth]{images/FactsTransformationToolGVSWithPortDefinitionsNoGlitch2.png}
+    \includegraphics[width=0.8\linewidth]{images/GVSMediationToolPreconditions.png}
     \caption{Definition of operator pre- / postconditions}
@@
   \begin{center}
-\includegraphics[angle=90,width=0.8\linewidth]{images/FactsTransformationToolGVSWithTransformationRules.jpg}
-    \caption{Interactive, rule based transforming of Person facts (photoshoped screen dump)}
+\includegraphics[width=\linewidth]{images/FactsTransformationToolGVSWithTransformationRules.png}
+    \caption{Interactive, rule-based transformation of Person facts}
     \label{fig:FactsTransformationToolTransformationRules}
@@
 \subsection{Creating the Operator}
+\label{sec:creating_the_operator}
 
-Once the data transformation operator has been defined, it may be added to the FAST catalogue in order to allow its use for the definition of piping configurations of new screens within the FAST CGS tool.
+Once the data transformation operator has been defined, it may be added to the FAST catalogue in order to allow its use for the definition of piping configurations of new screens within the FAST GVS tool.
 For this purpose a simple semantic description is sent to the FAST catalogue defining the semantics of the operator and the semantics of its pre- and postconditions.
+This is shown in Fig.~\ref{fig:operator_in_screen_builder}, where the newly created operator for mediating between Person and Customer concepts (i.e., transforming instances of one concepts into instances of the other) has been saved and is now available in the GVS screen builder interface.
+
+\begin{figure}[h]
+  \begin{center}
+\includegraphics[width=.6\linewidth]{images/operator_closeup}
+    \caption{A newly created operator in the GVS screen builder}
+    \label{fig:operator_in_screen_builder}
+  \end{center}
+\end{figure}
 

### 2011-02-26 trunk/deliverables/D2_4_Mediation/fast_d2_4_3_part2.tex
getting final version done
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1304 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
-\section{Ontologies}
+\section{Ontology Mediation Background}
+\label{sec:background}
+
+\subsection{Ontologies}
 \label{sec:ontologies}
@@
 
-\subsection{Definitions}
+\subsubsection{Definitions}
 As specified in \cite{gruber2008ontology} {\textquotedblleft}an ontology defines a
@@
 
-\subsection{General Description}
+\subsubsection{General Description}
 An ontology is used to define the vocabulary that agents use to exchange
@@
 
-\section{The Mediation Problem} %%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{The Mediation Problem} %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \label{sec:mediationproblem}
@@
 
-\subsection{Motivating Example}
+\subsubsection{Motivating Example}
 \label{subsec:mediationproblem_motivatingexample}
@@
 
-\subsection{Problem Statement}
+\subsubsection{Problem Statement}
 \label{subsec:mediationproblem_problemstatement}
@@
 
-\subsection{Mediation Examples}
+\subsubsection{Mediation Examples}
 \label{subsec:mediationproblem_examples}
@@
 
-\section{Ontology Mismatches}
+\subsection{Ontology Mismatches}
 \label{sec:ontologymismatches}
@@
 
-\subsection{Language Level Mismatches}
+\subsubsection{Language Level Mismatches}
 \label{subsec:ontologymismatches_languagelevel}
@@
 
-\subsubsection{Syntax}
+\paragraph{Syntax}
 Different languages obviously use different syntax. Therefore the same
@@
 
-\subsubsection{Logical representation}
+\paragraph{Logical representation}
 A slightly more complicated mismatch is when a logical notion can not be
@@
 
-\subsubsection{Semantics of primitives}
+\paragraph{Semantics of primitives}
 This difference is a more subtle one, and occurs at the meta-model of
@@
 
-\subsubsection{Language expressiveness}
+\paragraph{Language expressiveness}
 This mismatch, at the meta-model level of the languages, is the one
@@
 
-\subsection{Ontology Level Mismatches}
+\subsubsection{Ontology Level Mismatches}
 \label{subsec:ontologymismatches_ontologylevel}
@@
 A conceptualisation mismatch is a difference in how the domain is
-modeled (conceptualised), therefore we will have different concepts and
+modelled (conceptualised), therefore we will have different concepts and
 different relations between them. Explication mismatches are
@@
 
-\subsubsection{Conceptualisation Mismatches}
+\paragraph{Conceptualisation Mismatches}
 These are differences in the conceptualisation of the domain, that is
@@
 
-\subsubsection{Explication Mismatches}
+\paragraph{Explication Mismatches}
 The first two types of explication mismatches (\textit{paradigm} and
 \textit{concept description}) result from the \textbf{style of
-modeling.} The next two types (\textit{synonym terms} and
+modelling.} The next two types (\textit{synonym terms} and
 \textit{homonym terms}) can be classified as \textbf{terminological
@@
 
-\section{Ontology Mediation Approaches}
+\subsection{Ontology Mediation Approaches}
 \label{sec:mediationapproaches}
@@
 
-\subsection{Ontology Mapping}
+\subsubsection{Ontology Mapping}
 \label{subsec:mediationapproaches_ontologymapping}
@@
 temporal properties, but may participate in certain activities at given
-timepoints. 
+time points. 
 \end{itemize}
@@
 
-\subsection{Ontology Alignment}
+\subsubsection{Ontology Alignment}
 \label{subsec:mediationapproaches_ontologyalignment}
@@
 
-\subsubsection{The Alignment Process}
+\paragraph{The Alignment Process}
 The alignment process proposed in \cite{debruin2005wsml} relieves the user of
@@
 
-\subsubsection{Classification of Approaches}
+\paragraph{Classification of Approaches}
 An implementation of Match may use multiple matching algorithms or
@@
 
-\subsubsection{Overview of Approaches}
+\paragraph{Overview of Approaches}
 In this section we give a brief overview of some of the main approaches
@@
 iterations the actual mapping between the ontologies is obtained. 
-\subsubsection{Alignment Examples}
+\paragraph{Alignment Examples}
 This section provides a few examples of the matching
@@
 
-\subsection{Ontology Merging}
+\subsubsection{Ontology Merging}
 \label{subsec:mediationapproaches_ontologymerging}
@@
 
-\subsubsection{PROMPT}
+\paragraph{PROMPT}
 \label{subsubsec:mediationapproaches_ontologymerging_PROMPT}
@@
 
-\subsubsection{OntoMerge}
+\paragraph{OntoMerge}
 OntoMerge \cite{dou2002ontology_translation} is an online ontology translation system,

### 2011-02-26 trunk/deliverables/D2_4_Mediation/fast_d2_4_3_mediation.tex
getting final version done
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1305 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 
-In D2.4.1 we reviewed current work and literature, discussing the main areas of the topic, that is the core problem of ontology mediation, the types of mismatches that occur between ontologies, and the three main ontology mediation approaches, namely (1) the representation of correspondences between ontologies, (2) the (semi-)automatic process of discovering these correspondences and (3) creation of a new ontology from the source ontologies.
+In \cite{ambrus2009mediation} we reviewed current work and literature, discussing the main areas of the topic: the core problem of ontology mediation, the types of mismatches that occur between ontologies, and the three main ontology mediation approaches, namely (1) the representation of correspondences between ontologies, (2) the (semi-)automatic process of discovering these correspondences and (3) creation of a new ontology from the source ontologies.
 
-Based on the theoretical foundations laid down in D2.4.1 we present our approach to integrating ontology matching into the FAST platform.
+Based on these theoretical foundations, the second~\cite{ambrus2010mediation} as well as this final iteration of the deliverable present our approach to integrating ontology matching and the definition of mapping rules into the FAST platform.
 We present the roles related to ontology matching in FAST, and the way ontology matching is achieved within each one of those roles.
+We have developed and integrated a tool to allow end users to define mappings of varying complexity and create new operator building blocks to be used in the screen design process.
 
@@
     \textcolor{white}{\textbf{Dissemination Level}} & \multicolumn{4}{l|}{Public $\boxtimes$ / Consortium $\square$}\\ \hline
-    \textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.4.3 is the third iteration of the FAST ontology matching approach.
+    \textcolor{white}{\textbf{Short description of contents}} & \multicolumn{4}{p{10.85cm}|}{D2.4.3 is the third iteration of the FAST ontology matching and mapping approach.
 The first iteration contained a general overview of the literature regarding ontology mediation, paving the way for further analysis of the scenarios arising in FAST that require ontology mediation.
-Based on that general overview, this current iteration presents the approach to implement ontology matching within FAST.
-We present the different roles within the platform, tasks and challenges related to them, and the way ontology matching is integrated within each role.}\\ \hline
+Based on that general overview, the second and third iteration presents the approach to implement ontology matching within FAST.
+We present the different roles within the platform, tasks and challenges related to them, and the way ontology matching is integrated within each role, including the \emph{fact transformation tool}, which has become part of the FAST tool chain.}\\ \hline
     \textcolor{white}{\textbf{Authors}} & \multicolumn{4}{p{10.85cm}|}{\authorOne,  \authorTwo, \authorThree, \authorFour, \authorFive}\\
@@
 \clearpage
-\section*{Lists of Tables and Figures}
+% \section*{Lists of Tables and Figures}
 \listoftables
@@
 
-We aim to integrate a form of ontology matching into the FAST platform to allow for discrepant systems to intercommunicate.
+We aim to integrate a form of ontology matching into the FAST platform to allow for discrepant systems to intercommunicate --- or, more precisely, previously incompatible building blocks in our case.
 We try to find a solution to these differences which range from simple naming differences to more complex data heterogeneity, behavioural differences, discrepant paradigms and other disparities.
 The reconciliation of these differences is called ontology matching (often mentioned in the specialised literature as ontology mediation, or ontology alignment, thus we will sometimes use these terms as well).
+
+While we have explored ways of automatically defining mappings between different conceptualisations, our focus within FAST is less on research into new such methods.
+Instead, we aim to support the end user in defining such mapping rules within the FAST tool chain and integrate the process within the platform.
+To this end, we have developed and integrated a fact transformation tool, which is part of the FAST GVS and presented in this deliverable.
 
@@
 Thus, Sect.~\ref{sec:ontologyengineering} describes a tool to automatically discover alignments, a scenario and example ontologies to demonstrate the capabilities of the tool, and the testing procedure with the results.
-The remaining part of the section deals with a tool to define the discovered mapping rules in a way that is consistent with other similar tasks within the FAST platform, and describes the way in which matching operators are stored within the Semantic Catalogue.
+The remaining part of the section deals with the \emph{fact transformation tool}, developed to define the discovered mapping rules in a way that is consistent with other similar tasks within the FAST platform, and describes the way in which matching operators are stored within the Semantic Catalogue.
 Sect.~\ref{sec:gadgetbuilding} shortly describes the way the matching operator is used when composing screenflows and Sect.~\ref{sec:runtime} describes how the matching operator and the defined rules are converted and used in the final, deployed gadget.
@@
 %Describe the rest.
-The Appendix contains the first iteration of this document, D2.4.1, laying down the theoretical foundations of ontology matching.
+The Appendix contains the main content of first iteration of this document, D2.4.1, laying down the theoretical foundations of ontology matching.
 We discuss ontologies in general in App.~\ref{sec:ontologies}, present the ontology mediation problem in App.~\ref{sec:mediationproblem} , discuss the types  of mismatches between ontologies in App.~\ref{sec:ontologymismatches} and detail the main approaches to ontology mediation in App.~\ref{sec:mediationapproaches}.
@@
 
-%\pagebreak
-%\part{Implementing Ontology Matching within the FAST Platform}
-%\label{part:d2.4.2}
-
 \include{fast_d2_4_3_part1}
-
-\section{Summary}
-\label{sec:summary}
-With the Semantic Web winning more and more ground both in research and industry, the mismatches between ontologies used by the inter-operating systems present an increasing problem.
-Ontology matching deals with identifying these mismatches, and overcoming them.
-
-In the first iteration (described in the Appendix) we laid the theoretical foundations of ontology matching.
-After presenting the main problem, and the types of mismatches between ontologies, we have discussed - based on the most prominent research results in these areas - the three main approaches to ontology mediation, that is \textit{ontology mapping} (representing  correspondences between entities), \textit{ontology alignment} (the process of finding the correspondences between ontologies) and \textit{ontology merging }(the process of creating a new ontology based
-on the source ontologies that need to be reconciled).
-
-In this second iteration (described in the main body of this document) we established the approach to integrate ontology matching into FAST, and we have shown how this matching is done at the different phases of the gadget life cycle and the approach taken at every level.
-Firstly, we have shown the implementation details as well as test scenarios, we have presented a tool for finding the alignments between ontologies as well as a powerful language to represent them, we have described a tool to compose mapping rules in a consistent manner with other FAST components and the means to save them in the Semantic Catalogue.
-Secondly, we described the way in which mapping operators are used in the GVS to combine discrepant screens.
-Lastly, we presented the way the matching operator and mapping rules are deployed and exploited in the JavaScript gadget.
-
-For the last and final iteration of deliverable D2.4 we plan to evaluate real-life scenarios and finalise the implementation of ontology matching for the FAST platform.
 

### 2011-02-26 trunk/deliverables/D2_4_Mediation/fast_d2_4_3_part1.tex
getting final version done
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1305 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 Relevant roles are the ontology engineer, who performs the main work in this phase, as well as the the resource developer and the screen developer, who are directly affected.
+
+In this section, we present a number of tools that can be used to help the ontology engineer in creating mapping rules to mediate between different conceptualisations.
+Some ontologies relevant for an e-commerce use case are introduced, and their automatic mapping is evaluated using one of the tools.
+The evaluation shows that automatic mapping is feasible only in certain cases, while other cases require a manual approach.
+The section concludes with the presentation of a tool developed in FAST to facilitate the use of (either automatically or manually created) mapping rules within the FAST tool chain.
 
@@
 The first row (reference) shows the reference alignment, which, naturally, has both perfect precision and recall.
-We can observe what intuition has predicted, namely that pure string equality (equality) is far too simple and irrelevant, by only taking identical labels.
-By using string distances and giving certain thresholds (Levenshtein and SMOA), we can see that the results are much less precise, but have a better recall, since this allows for entities having similar names to be discovered, at the expense of having quite a few incorrect results (lower precision); the thresholds allow for low-scored cases to be eliminated, although this results in the exclusion of some correct correspondences.
-The last column (JWNL) contains the results of the Wordnet-enabled method, which shows quite an improvement (precision of 0.67 and a recall of 0.75\footnote{The scores are orientative, giving a general picture of how the process approaches 100\% efficiency}), due to the lexical analysis, which performs a much more relevant comparison of strings, giving a high number of correct results.
+We can observe what intuition has predicted, namely that pure string equality on term labels is far too simple and therefore irrelevant.
+By using string distances and giving certain thresholds (``Levenshtein and SMOA''), we can see that the results are much less precise, but have a better recall, since this allows for entities having similar names to be discovered, at the expense of having quite a few incorrect results; the thresholds allow for low-scored cases to be eliminated, although this results in the exclusion of some correct correspondences.
+The last row (``JWNL'') contains the results of the Wordnet-enabled method, which shows quite an improvement (precision of 0.67 and a recall of 0.75),
+% \footnote{The scores are orientative, giving a general picture of how the process approaches 100\% efficiency}
+due to the lexical analysis, which performs a much more relevant comparison of strings, giving a high number of correct results.
 The precision of the JWNL alignment shows only a tiny drop below the recall value, meaning that the number of incorrect correspondences discovered is small, and the main source of error is from the number of correspondences not discovered.
@@
 
-To support manual ontology mapping, we have developed a rule based, interactive Facts Transformation Tool.
-This tool utilizes techniques of our service wrapper tool, cf.
-deliverable D4.3.2.
-One part of the service wrapper tool deals with a rule based approach for the transformation of server responses into FAST facts.
-This problem is closely related to the problem of ontology mediation.
-Thus, these techniques also allow to create data transformation operators, i.e.
-operators that mediate between two ontologies.
-
-Figure \ref{fig:FactsTransformationToolPortDefinitions} shows a screen shot of the  Facts Transformation Tool within the GVS environment.
-The Facts Transformation Tool has three tabs.
-The \textit{General} tab allows to define the pre- and postconditions for the desired data transformation operator.
-These conditions are  typed according to the participating semantic elements to be derived from the FAST catalogue.
-This typing will provide the structure of the input and output data of the desired data transformation operator.
-
-Figure \ref{fig:FactsTransformationToolTransformationRules} shows the editor for transformation rules.
-On the left of Figure \ref{fig:FactsTransformationToolTransformationRules} there is a tree of example data that is used to validate the transformation rules during editing.
-Thus, the Facts Transformation Tool will execute the transformation rules after any editing in order to give direct feedback on their functionality and to enable the user to validate the achieved transformation.
-The transformation results are shown on the right of Figure \ref{fig:FactsTransformationToolTransformationRules}.
+To support the application of previously created mapping rules, we have developed a rule based, interactive \emph{fact transformation tool} (or \emph{fact tool} for short).
+This tool utilises techniques of our service wrapper tool.
+One part of the service wrapper tool~\cite{rivera2011connecting} uses a rule-based approach for the transformation of web service responses into FAST facts (i.e., instances of ontology concepts).
+The application of ontology mediation rules can be modelled in a similar vein.
+In the fact tool, we therefore use analogous rule-based techniques to create data transformation operators, i.e., operators that mediate between two different conceptualisation.
 
@@
 
-The middle of Figure \ref{fig:FactsTransformationToolTransformationRules} shows some possible transformation rules for our example.
-A \textit{createObject} rule may be used to create output fact objects.
-The \textit{Target} field of such a rule specifies the type of the created objects, in our example \textit{List}.
-The top rule fires for each element of the input data with the type specified in the \textit{From} field of the transformation rule.
-Thus, in our example, the input data is searched for objects of type \textit{List} firing for the root of our data.
-
-The second rule of our example creates \textit{Customer} objects.
-This rule is a subrule of the root rule and thus the created objects are added to the \textit{List} object created by the parent rule.
-Similarly, the search for triggering objects of type \textit{Person} is restricted to the subtree below the from element that has triggered the parent rule, i.e.
-to our input \textit{List}.
-
-The third rule is used to fill the attribute \textit{fullName} of the \textit{Customer} objects.
-This rule is an example for a many to one transformation.
-The result is a string concatenation of the \textit{lastName} of the \textit{Person} that has triggered the parent rule plus a text constant plus the \textit{firstName} of the current \textit{Person}.
-The resulting string is transferred to the \textit{fullName} attribute of the corresponding \textit{Customer}.
-
-Note that simple one-to-one transformation rules may be derived with the help of the ontology mediation approach described in Section \ref{subsubsec:automaticmapping}.
-Such simple rules may be added to the transformation tool, automatically.
-We try to add this functionality within the remaining project time.
-The more complex rules may then be added interactively.
-During editing, the semantics provided by the FAST catalogue is used to support the user with completion proposals and consistency checking.
-In addition, the direct execution of the rules on example data helps the user to create appropriate rules and to validate their effects.
-
-So far our transformation rules are restricted to simple cases. We have focused on simple cases allowing a simple transformation approach that does not require sophisticated skills from our users.
+Figure \ref{fig:FactsTransformationToolPortDefinitions} shows a screen shot of the  fact transformation tool within the GVS environment.
+The tool has four tabs\footnote{The \texttt{CodeGen Viewer} tab is explained later in Sect.~\ref{sec:runtime}.}, of which the \textit{General} tab (currently shown in the screen shot) is employed to define the pre- and postconditions for the desired data transformation operator in a first general step.
+Using the terminology introduced previously, this allows to define a basic \emph{level 0} mapping between two concepts (\texttt{PersonList} and \texttt{CustomerList} in the example).
+The fact tool communicates with the FAST catalogue to select concepts and provide example instances (as JSON data structures) that help the user to perform the correct choice.
+% These conditions are typed according to the participating semantic elements to be derived from the FAST catalogue.
+This typing will provide the structure of the input and output data of the desired data transformation operator.
+After this step, the mapping can be refined to a higher level by defining concrete transformation rules.
+These can be edited in the \emph{exTransformation} tab as shown in Fig.~\ref{fig:FactsTransformationToolTransformationRules}.
+On the left of the tab the user is presented with a tree of example data for the input concept, while the right side of the tab shows a similar tree with the derived data for the output concept.
+This way, the user can continuously observe the effect of changes in the rule they are currently editing.
+The fact tool constantly executes the transformation rules after any editing, giving direct feedback on their effect and thus enabling the user to validate the achieved transformation.
 
@@
 
+Examples for the rules themselves are shown in the rule tree in the middle of Fig.~\ref{fig:FactsTransformationToolTransformationRules}.
+Starting with a root rule, additional rules can be added as child-rules, while existing rules can be moved around the tree or deleted.
+Rules can have different types, such as \emph{createObject} or \emph{fillAttributes}.
+A \textit{createObject} rule can be used to create output fact objects.
+The \textit{Target} field of such a rule specifies the type of the created objects (i.e., a concept from the FAST catalogue), in our example \textit{List}.
+The top rule fires for each element of the input data with the type specified in the \textit{From} field of the transformation rule.
+Thus, in our example, the input data is searched for objects of type \textit{List} firing for the root of our data.
+The second rule of our example creates \textit{Customer} objects.
+This rule is a sub-rule of the root rule and thus the created objects are added to the \textit{List} object created by the parent rule.
+Similarly, the search for triggering objects of type \textit{Person} is restricted to the subtree below the from element that has triggered the parent rule, i.e.
+to our input \textit{List}.
+The third rule is used to fill the attribute \textit{fullName} of the \textit{Customer} objects.
+This rule is an example for a many-to-one transformation.
+The result is a string concatenation of the \textit{lastName} of the \textit{Person} that has triggered the parent rule plus a text constant plus the \textit{firstName} of the current \textit{Person}.
+The resulting string is transferred to the \textit{fullName} attribute of the corresponding \textit{Customer}.
+
+Note that simple one-to-one transformation rules may be derived with the help of the ontology mediation approach described in Sect.~\ref{subsubsec:automaticmapping}.
+Such simple rules can be added to the transformation tool, automatically.
+The more complex rules may then be added interactively.
+During editing, the semantics provided by the FAST catalogue are used to support the user with completion proposals and consistency checking.
+In addition, the direct execution of the rules on example data helps the user to create appropriate rules and to validate their effects.
+
+So far our transformation rules are restricted to simple cases. We have focused on simple cases allowing a simple transformation approach that does not require sophisticated skills from our users.
+
 \subsection{Creating the Operator}
@@
 For this purpose a simple semantic description is sent to the FAST catalogue defining the semantics of the operator and the semantics of its pre- and postconditions.
-This is shown in Fig.~\ref{fig:operator_in_screen_builder}, where the newly created operator for mediating between Person and Customer concepts (i.e., transforming instances of one concepts into instances of the other) has been saved and is now available in the GVS screen builder interface.
+This is shown in Fig.~\ref{fig:operator_in_screen_builder}, where the newly created data transformator for mediating between Person and Customer concepts (i.e., transforming instances of one concepts into instances of the other) has been saved and is now available in the GVS screen builder interface as an filter operator.
 
@@
 The Service Wrapper Tool uses its own repository for trees of transformation rules.
-First of all this has technical reasons: the development of the transformation tool is based on a reference architecture and libraries for web based interactive systems provided by Kassel University, cf.
-\cite{AsDH2009CVSM}.
+First of all this has technical reasons: the development of the transformation tool is based on a reference architecture and libraries for web-based interactive systems provided by Kassel University~\cite{AsDH2009CVSM}.
 This repository has the special advantage that it provides support for teams of developers, thus the service wrapper tool allows to merge the contributions of multiple developers.
 
-\section{Gadget Building Phase}
+\section{Gadget Building  and Runtime}
+
+Neither gadget (or screenflow) building or runtime are directly involved with the ontology mapping process.
+However, the steps performed previously affect these phases of the lifecycle as well, as discussed briefly in the sections below.
+
+\subsection{Gadget Building} % (fold)
 \label{sec:gadgetbuilding}
-Every time ontology matching is performed and an alignment is created between two underlying resources, the corresponding alignment rules are stored and an operator is created in the catalogue.
 
-At the gadget building phase, whenever preconditions and postconditions of screens, resources or operators show a semantic mismatch, the catalogue offers the possible matching operators that would solve the incompatibility problem.
-This way a matching operator can be placed between two pre- postconditions that would otherwise be incompatible, making interoperation between the two possible.
+Every time ontology matching is performed and an alignment is created between two underlying resources, the corresponding alignment rules are stored and an operator building block is created in the catalogue, to be available during screen development.
+
+During the gadget building phase, whenever pre- and post-conditions of screens, resources or operators show a semantic mismatch, the catalogue offers the possible matching operators that would solve the incompatibility problem.
+This way a matching operator can be placed between two pre- and postconditions that would otherwise be incompatible, making interoperation between the two possible.
 In seldom cases, the catalogue may fail to retrieve an appropriate mediation operator.
-In such cases, the possibility of building a new matching operator with the help of the Facts Transformation Tool is presented, in order to meet the user's need.
+In such cases, the possibility of building a new mapping operator with the help of the fact transformation tool is presented, allowing the user to assume the role of ontology engineer in special cases.
 
-\section{Runtime} 
+\subsection{Runtime} 
 \label{sec:runtime}
-For deployment, the gadgets implementation is generated in Javascript.
+For deployment, the gadget's implementation is generated as an executable bundle containing a portion of Javascript code.
 In addition, a small FAST runtime library (again Javascript) is added to the gadget.
 This runtime library provides notification mechanisms and manages the fact base of the gadget.
-For data transformation we generate Javascript code that connects itself to the fact base and any time a fact is sent to the precondition of the operator, it does the data transformation and sends the result to the postcondition.
-The code generated from the data transformation rules follows a recursive pattern where each rule becomes a Javascript function with parameters for the source and target fact subtrees to be transformed.
+For data transformation purposes (i.e., ontology mapping) we generate Javascript code that connects itself to the fact base.
+Any time a fact is sent to the pre-condition of the operator, it performs the data transformation and sends the result to the post-condition.
+The code generated from the transformation rules follows a recursive pattern, where each rule is represented as a Javascript function with parameters for the source and target fact subtrees to be transformed.
 The function body then uses code to search the source tree for sub elements that match the \textit{From} part of the rule and then code for the desired operation is generated.
-For debugging purposes the user may examine the code generated for an operator in the \textit{CodeGen Viewer} tab of the Facts Transformation tool.
+For debugging purposes the user can examine the code generated for an operator in the \textit{CodeGen Viewer} tab of the fact tool.
 
-Our code generation for service wrappers and transformation operators uses a template based approach.
+Our code generation for service wrappers and transformation operators uses a template-based approach.
 Thus, the \textit{CodeGen Viewer} tab will also allow to examine the templates used for code generation and to adapt the templates in order to meet changing APIs or to meet new requirements.
 Thereby, the code generation is not hard wired into our tool but may be adapted by (skilled) end users on the fly.
+
+\section{Summary}
+\label{sec:summary}
+With the Semantic Web winning more and more ground both in research and industry, the mismatches between ontologies used by the inter-operating systems present an increasing problem.
+Ontology matching deals with identifying these mismatches, and overcoming them.
+
+The initial work on the ontology matching comprised an overview of the area in general, and was reflected in the first iteration of this deliverable (\cite{ambrus2009mediation}, now in App.~\ref{sec:background}).
+In it, we discussed the theoretical foundations of ontology matching.
+After presenting the main problem and the types of mismatches between ontologies, we presented --- based on a study of the state-of-the-art in these areas --- the three main approaches to ontology mediation, i.e., \textit{ontology mapping} (representing  correspondences between entities), \textit{ontology alignment} (the process of finding the correspondences between ontologies) and \textit{ontology merging} (the process of creating a new ontology based on the source ontologies that need to be reconciled).
+
+The second iteration of the document~\cite{ambrus2010mediation} saw a positioning of ontology matching within the FAST project, both in terms of the FAST lifecycle and its associated user roles.
+It was defined which phases are affected in which way, and which actions by particular user roles are involved.
+We have shown how matching is done at the different phases of the gadget life cycle and the approach taken at every level.
+Additionally, we covered different candidates for tools to support automatic matching algorithms and evaluated the most promising ones on the backdrop of an e-commerce scenario.
+It was found that, while automatic matching is feasible for some, mostly simple cases, the manual definition of mapping rules will still be necessary for many real-life cases.
+Research into new methods for automatic matching of complex cases was deemed outside the scope of this project.
+Instead, the focus was on supporting users of the FAST tool to manually define mapping rules in an intuitive fashion.
+To this end, a first standalone prototype of the \emph{fact transformation tool} was developed and presented briefly.
+
+This final iteration of the deliverable focussed mainly on the improvement of the fact tool and its transformation of a standalone tool into an integral part of the FAST tool chain, thereby supporting users and allowing them to use the same kind of interaction and user interface throughout the platform.
+The fact tool presented in this deliverable and the service wrapping tool presented in \cite{rivera2011connecting} therefore use a similar graphical approach.
+Without leaving the FAST platform, users can now assume the role of ontology engineer and define mapping rules of varying complexity, which can then be saved as operator building blocks, to be used in the screen development process.
+To allow this, the fact tool communicates with the FAST catalogue to guide users in the rule creation process and outputs definitions according to the FAST gadget ontology, as well as code to be used in the deployed gadget.

### 2011-02-27 trunk/deliverables/D5_1_2/fast_d5_1_2.tex
minor changes in "planning" appendix
git-svn-id: https://svn.forge.morfeo-project.org/fast-fp7project@1327 d8b0c59b-6544-0410-94d9-83a5658a058e

@@
 As a consequence, it must handle and work with ontologies and vocabularies.
-The business model itself is described in the FAST ontology, which is included in the Catalogue along with a few established ontologies integrated in its core (see~\cite{moeller2011fast_ontology}).
+The business model itself is described in the FAST ontology, which is included in the catalogue along with a few other established ontologies (see~\cite{moeller2011fast_ontology}).
 However, the definition of the pre-/post-conditions of the building blocks are open to any ontology or vocabulary, and they may reference classes and properties from any of them.
@@
 Therefore, the aim of the algorithm is to find the paths or plans from a given screen to the special node or root.
-This is a very well-known problem in graph theory, and therefore there are many graph and tree search algorithms dealing with it, such as Floyd's, Dijkstra, A* or depth-first search, among others.
+This is a very well-known problem in graph theory, and therefore there are many graph and tree search algorithms dealing with it, such as Floyd's, Dijkstra, A* or breadth-first search, among others.
 
@@
 However, finding the shortest path is not the intention of this algorithm.
-In fact, it is preferred to find a small set of possible solutions, thus depth-first search is a good fit as it finds all the paths from a given node to another, avoiding cycles.
-Then, the paths returned by depth-first are then ranked in order to minimise the path lenght, and to maximise the number of co-occurrences of elements in the path and the canvas.
-
-The depth-first search algorithm has not been re-implemented for the catalogue.
-Instead, it relies on Neo4j\footnote{http://neo4j.org/}, a flexible graph database, which apart from providing the aforesaid algorithm, it provides the necessary mechanism to store, access and query the graph.
-The advantages of using Neo4j are that it is highly optimized for storing graph structures for maximum performance and scalability, and supports several search algorithms, including as well Dijkstra and A*, which provides a great flexibility to the catalogue.
+In fact, it is preferred to find a small set of possible solutions, thus breadth-first search is a good fit as it finds all the paths from a given node to another, avoiding cycles.
+Then, the paths returned by breadth-first are then ranked in order to minimise the path lenght, and to maximise the number of co-occurrences of elements in the path and the canvas.
+
+The graph is updated any time a screen is created, updated or deleted in order to reflect the creation/deletion of nodes and their relationships.
+It is stored in the triple store where the building blocks metadata and other information is stored as well, and an implementation of the breadth-first search algorithm has been created for searching the paths or plans.
+
+Besides, the Catalogue has been integrated with Neo4j\footnote{http://neo4j.org/}, a flexible graph database, which provides the necessary mechanisms to store, access and query the graph.
+Neo4j provides different graph search algorithms, including breadth-first search.
+The advantages of using Neo4j are that it is highly optimized for storing graph structures maximising performance and scalability, and supports several search algorithms, including as well Dijkstra and A*, which provides a great flexibility to the catalogue in case other algorithms may be needed.
 
